### 2015-04-09 rapport/component.tex
rapport_090415

%-------------------------------------
\subsection{Component models}
A \textit{component} is an independent object which represents a functionnality of an application, and which is associated to a set of \textit{interfaces} used to plug an instance of a component with other ones. The object composed of multiple component instances linked together through their interfaces is called a component assembly. As a result, a component assembly is the code of an entire application represented as a set of splitted functionnalities. Such software engineering model is known to increase \textit{code re-use} and \textit{productivity}, as one component can be used in different applications, but also \textit{maintainability} and \textit{scalability}, as functionnalities of an application are clearly splitted in different and independent components. More recently, component models have been studied to write HPC applications such as $L^2C$~\cite{} (Low Level Component Model) and CCA~\cite{} (Common Component Architecture).

In the rest of this paper, the component model $L^2C$ is considered and used by the solution. A component is defined as
\begin{equation}
C(I,F),
\end{equation}
where $I$ is a set of available interfaces for the given component, and $F$ is the set of functions to answer the functionnality of the component. A component instance, as for an object, is the actual creation of the component, and more than one instance can be created. The components of $L^2C$ are limited to a few interfaces without the introduction of overheads. In this paper we are interested in four types of interfaces. The three first one are \textit{use} and \textit{provide} interfaces (Figure~\ref{AB}) to use a functionnality provided by another component instance, and \textit{use-multiple} interface (Figure~\ref{ABC}) to use a vector of functionnalities provided by a set of component instances. A condensed notation of a use-multiple is given in the Figure~\ref{ABCbis}.

\begin{figure}[h!]
\begin{center}
  \subfigure[][$A$ uses $B$ and $A$ uses $C$]{
  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (A) at (0,0) {$A$};
   \node[component] (B) at (2,+0.5) {$B$};
   \node[component] (C) at (2,-0.5) {$C$};

  \path[-]
    ([yshift=1ex]A.east) edge [connection] (B.west)
    ([yshift=-1ex]A.east) edge [connection] (C.west);
  \end{tikzpicture}
  \label{AB}
  }
  \hspace{2cm}
  \subfigure[][$A$ uses multiple $B$ instances]{
  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (A) at (0,0) {$A$};
   \node[component] (B) at (2,+0.5) {$B_1$};
   \node[component] (C) at (2,-0.5) {$B_2$};

  \path[-]
    (A.east) edge [connection] node {} (B.west)
             edge [connection] node {} (C.west);

  \end{tikzpicture}
  \label{ABC}
  }
  \hspace{2cm}
  \subfigure[][$A$ uses multiple $B$ instances (condensed notation)]{
  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (A) at (0,0) {$A$};
   \node[component] (B) at (2,0) {$B$};

  \path[-]
    (A.east) edge [connection] node {1,n} (B.west);

  \end{tikzpicture}
  \label{ABCbis}
  }
  \caption{$L^2C$ use-provide and use-multiple interfaces}
  \label{interfaces}
\end{center}
\end{figure}

An instance $i$ of a component has the same properties than the component it is instanciated from, but it is also asociated to a resource $r$ (a core, a thread etc.) of the set of all available resources $R$.
\begin{equation}
C_i(I,F,r)
\end{equation}
A component assembly represents the application as a set of component instances and the composition of those instances through their interfaces. A component assembly is defined as
\begin{equation}
\alpha (\mathcal{C},L,R),
\end{equation}
where $\mathcal{C}$ is the set of component instances used, $L$ is the set of links between interfaces of component instances, and $R$ is the set of available resources. The component assembly of $L^2C$ offers a way to write an assembly for each MPI process, and to connect some of their components by MPI communicators. This connection is an \textit{MPI interface}. In the rest of this paper we consider a broader interface called \textit{synchronization}, which corresponds to the MPI interface of $L^2C$ but which could also be a \textit{thread synchronization} instead.

Using those four interfaces (use, use-multiple, provide and synchronization), it is possible to write an application with medium-grain components, enough fine to keep efficiency and to stay close to the execution model, and enough coarse to take advantage of components. The efficiency of $L^2C$ has been proved, while increasing the productivity and maintainability of HPC applications, but moreover it has also been shown that the protability of such component-based HPC application is improved~\cite{}.

%-------------------------------------
\subsection{Stencil program parallelization using components}
In this section is studied the parallelization of a numerical simulation using the given definitions of a component, an assembly and four interfaces. First, the different parallelization techniques explained in the Section~\ref{sect:parall} are translated to component-based parallel assemblies, and an approximation of the general case is discussed for the solution presented in this paper. In the rest of this section, the following example of dependencies is used 
\begin{equation}
A<((B<D) \parallel C)<E \ll A'<((B'<D') \parallel C'),
\label{eq:dep}
\end{equation}
 where $A,B,C,D,E,A',B',C',D'$ are numerical computations. %In this example, $\parallel$ is a binary expression which means that a parallel execution of the right and left computations is possible. It is equivalent to $\not>$. 
This example, can be represented as a directed acyclic graph as in the Figure~\ref{example}.

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (A) at (0,0) {$A$};
   \node[component] (B) at (1,+0.5) {$B$};
   \node[component] (C) at (1,-0.5) {$C$};
   \node[component] (D) at (2,+0.5) {$D$};
   \node[component] (E) at (3,0) {$E$};
   \node[component] (AA) at (4,0) {$A'$};
   \node[component] (BB) at (5,+0.5) {$B'$};
   \node[component] (CC) at (5,-0.5) {$C'$};
   \node[component] (DD) at (6,+0.5) {$D'$};

  \path[->]
    (A) edge node {} (B)
        edge node {} (C)
    (B) edge node {} (D)
    (D) edge node {} (E)
    (C) edge node {} (E)
    (E) edge node {} (AA)
    (AA) edge node {} (BB)
        edge node {} (CC)
    (BB) edge node {} (DD);
  \end{tikzpicture}
  \caption{Dependencies example~(\ref{eq:dep}) represented as a DAG.}
  \label{example}
\end{center}
\end{figure}

%---------
%---------
\paragraph{Coarse- and fine-grain data parallelism.} 
As explained in the Section~\ref{sect:parall}, to apply coarse-grain data parallelism on numerical simulations, a domain decomposition (or graph partitioning) has to be done on the mesh and also on the data mapped on it. When this important step is solved, the same program can be applied on each subpart of the data, on the different resources. In addition to this, a set of synchronizations are needed between the resources at each time step to compute the neighborhood $\mathcal{N}$ correctly. When using a component model, the computing part of the application, which is called for each time step, can be seen as a component assembly which is duplicated on different resources, which is possible using $L^2C$ for example. However, it is needed to classify computations in the component assembly to identify when synchronizations are needed (denoted $\ll$ in the Section~\ref{sect:dep}). For this reason, a \textit{sequence} component $SEQ$, which uses one after the other a set of computation components, is defined as 
\begin{equation}
SEQ(\{provide,use-multiple\},\{sequence\}).
\end{equation} 
Thus, the \textit{sequence} function (represented in the Algorithm~(\ref{alg:seq})) is responsible for the loop which calls as much \textit{uses} as the number of components connected to the \textit{use-multiple}. In addition to this, a \textit{synchronized sequence} component $SSEQ$, which uses one after the other a set of computation components, but which also proceeds synchronizations between them, is defined as 
\begin{equation}
SSEQ(\{provide,use-multiple, synchronization\},\{ssequence\}).
\end{equation}
In this case, the \textit{ssequence} function (represented in the Algorithm~(\ref{alg:sseq})) is responsible for a loop of two steps, first a synchronization step, second a use step. Finally, each numerical computation is represented by a component 
\begin{equation}
K(\{provide\},\{compute\}),
\end{equation}
 where the function \textit{compute} (represented in the Algorithm~(\ref{alg:comp})) is responsible for the sapce loop and the numerical computation on each element of the mesh.

\begin{center}
\begin{minipage}[t]{7cm}
  \vspace{0pt} 
 \begin{algorithm}[H]
 \ForAll{component $cp$ to use}{
 use the provide interface of $cp$
 }
 \caption{sequence function}
 \label{alg:seq}
 \end{algorithm}
 \end{minipage}%
 \hspace{0.5cm}
\begin{minipage}[t]{7cm}
  \vspace{0pt}
 \begin{algorithm}[H]
 \ForAll{component $cp$ to use}{
 synchronizations\\
 use the provide interface of $cp$
 }
 \label{alg:sseq}
 \caption{ssequence function}
 \end{algorithm}
 \end{minipage}%
\end{center}

%\begin{minipage}[t]{5cm}
 % \vspace{0pt}
 \begin{center}
 \begin{algorithm}[H]
 \ForAll{elements $d$ in $D$}{
 $w(d)=e(R(d),R(\mathcal{N}(d)))$
 }
 \caption{compute function}
 \label{alg:comp}
 \end{algorithm}
 \end{center}
%\end{minipage}%
Using the three components $SSEQ$, $SEQ$, and $K$, the component assembly of the dependencies~(\ref{eq:dep}) is represented in the Figure~~\ref{exdataparall}. One can notice, in this Figure, that the parallel dependencies ($\parallel$) are lost. However, in a SPMD parallelization, the parallelism is obtained from executing the same program on different part of the data. As a result, a correct sequence of computations is sufficient to expose coarse-grain data parallelism.
%---------
\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
  \node[component] (SSEQ) at (0,0) {$SSEQ$};
  \node (sync) at (4,0) {Synchronizations};
  \node[component] (SEQ1) at (-2.5,-1) {$SEQ$};
  \node[component] (SEQ2) at (2.5,-1) {$SEQ$};
  \node[component] (A) at (-4,-2) {$A$};
  \node[component] (B) at (-3,-2) {$B$};
  \node[component] (C) at (-2,-2) {$C$};
  \node[component] (D) at (-1,-2) {$D$};
  \node[component] (E) at (0,-2) {$E$};
  \node[component] (AA) at (1,-2) {$A'$};
  \node[component] (BB) at (2,-2) {$B'$};
  \node[component] (CC) at (3,-2) {$C'$};
  \node[component] (DD) at (4,-2) {$D'$};

  \path[-]
    (SSEQ) edge [connection] node {} (SEQ1)
           edge [connection] node {} (SEQ2)
           edge [mpiconn] node {} (sync)
    (SEQ1) edge [connection] node {} (A)
           edge [connection] node {} (B)
           edge [connection] node {} (C)
           edge [connection] node {} (D)
           edge [connection] node {} (E)
    (SEQ2) edge [connection] node {} (AA)
           edge [connection] node {} (BB)
           edge [connection] node {} (CC)
           edge [connection] node {} (DD);

  \end{tikzpicture}
\caption{Example assembly for coarse- and fine-grain data parallelism}
\label{exdataparall}
\end{center}
\end{figure}
The overall computation assembly of an SPMD numerical simulation is composed of a duplication of the computation (or dependencies) assembly described above. Each duplication is applied on an available resource as illustrated in the Figure~\ref{overall}. In the rest of this paper, this overall computation assembly of an SPMD application (coarse-grain data parallelism) is represented as a single computation assembly, and the $SSEQ$ component indicate where the synchronizations are needed.

%---------
\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
  \node[smallcp] (SSEQ) at (0,0) {$SSEQ$};
  \node (sync) at (3.5,0) {Synchronizations};
  \node[smallcp] (SEQ1) at (-1.5,-1) {$SEQ$};
  \node[smallcp] (SEQ2) at (1.5,-1) {$SEQ$};
  \node[smallcp] (A) at (-2.5,-2) {$A$};
  \node[smallcp] (B) at (-2,-2) {$B$};
  \node[smallcp] (C) at (-1.5,-2) {$C$};
  \node[smallcp] (D) at (-1,-2) {$D$};
  \node[smallcp] (E) at (-0.5,-2) {$E$};
  \node[smallcp] (AA) at (0.5,-2) {$A'$};
  \node[smallcp] (BB) at (1.1,-2) {$B'$};
  \node[smallcp] (CC) at (1.9,-2) {$C'$};
  \node[smallcp] (DD) at (2.5,-2) {$D'$};
  \draw [resource] (-3,-2.5) rectangle +(6cm, 3cm);
  \node (R1) at (0,-3) {r1};

  \node[smallcp] (SSEQ2) at (7,0) {$SSEQ$};
  \node[smallcp] (SEQ3) at (5.5,-1) {$SEQ$};
  \node[smallcp] (SEQ4) at (8.5,-1) {$SEQ$};
  \node[smallcp] (A2) at (4.5,-2) {$A$};
  \node[smallcp] (B2) at (5,-2) {$B$};
  \node[smallcp] (C2) at (5.5,-2) {$C$};
  \node[smallcp] (D2) at (6,-2) {$D$};
  \node[smallcp] (E2) at (6.5,-2) {$E$};
  \node[smallcp] (AA2) at (7.5,-2) {$A'$};
  \node[smallcp] (BB2) at (8.1,-2) {$B'$};
  \node[smallcp] (CC2) at (8.9,-2) {$C'$};
  \node[smallcp] (DD2) at (9.5,-2) {$D'$};
  \draw [resource] (4,-2.5) rectangle +(6cm, 3cm);
  \node (R2) at (7,-3) {r2};

  \path[-]
    (SSEQ) edge [connection] node {} (SEQ1)
           edge [connection] node {} (SEQ2)
           edge [mpiconn] node {} (sync)
    (SEQ1) edge [connection] node {} (A)
           edge [connection] node {} (B)
           edge [connection] node {} (C)
           edge [connection] node {} (D)
           edge [connection] node {} (E)
    (SEQ2) edge [connection] node {} (AA)
           edge [connection] node {} (BB)
           edge [connection] node {} (CC)
           edge [connection] node {} (DD)
    (SSEQ2) edge [connection] node {} (SEQ3)
           edge [connection] node {} (SEQ4)
           edge [mpiconn] node {} (sync)
    (SEQ3) edge [connection] node {} (A2)
           edge [connection] node {} (B2)
           edge [connection] node {} (C2)
           edge [connection] node {} (D2)
           edge [connection] node {} (E2)
    (SEQ4) edge [connection] node {} (AA2)
           edge [connection] node {} (BB2)
           edge [connection] node {} (CC2)
           edge [connection] node {} (DD2);

  \end{tikzpicture}
\caption{Overall SPMD computation assembly of the example~(\ref{eq:dep}) on two resources.}
\label{overall}
\end{center}
\end{figure}

But the assembly obtained from the instanciation of $SEQ$, $SSEQ$, and $K$ does not only expose coarse-grain data parallelism. Actually, the generic notation of a computation component $K$, which represents a numerical computation, is important to identifiy the different computations of a numerical simulation as different functionnalities of the program, but in addition to this, exposing a computation as a component, also exposes the fine-grain data parallelism level. In fact, a numerical computation is nothing more than a loop (or a nested loop) on the mesh elements and a numerical expression to compute (Algorithm~(\ref{alg:comp})). As a result a computation component is ideal to introduce SIMD and SIMT parallelism.

The general case of assembly to introduce coarse- and fine-grain data parallelism in numerical simulations is represented in the Figure~\ref{dataparall}.

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (SSEQ) at (0,0) {SSEQ};
   \node[component] (SEQ) [right=of SSEQ] {SEQ};
   \node[component] (K) [right=of SEQ] {K};

  \path[-]
    (SSEQ)  edge [connection] node {1,n} (SEQ)
            edge [mpiconn] node {} (-1,0)
    (SEQ)	 edge [connection] node {1,n}	(K);
\end{tikzpicture}
\caption{General assembly description for coarse- and fine-grain data parallelism}
\label{dataparall}
\end{center}
\end{figure}
%---------

%---------
%---------
\paragraph{Data and task parallelism}
As also explained in the Section~\ref{sect:parall}, to apply task parallelism to numerical simulations, it is needed to identify the different tasks of the application and their dependencies. With the introduction of $SSEQ$, $SEQ$, and $K$ components, a part of this work is already done. Actually, each computation component $K$ represents a task. However, to be able to express all dependencies between computations, the $PAR$ component is defined as 
\begin{equation}
PAR(\{provide,use-multiple\},\{parallel\}),
\end{equation}
 and corresponds to the expression $\parallel$ or $\not<$. The function $parallel$ creates a thread for each component to use and join all threads at the end.

\begin{algorithm}[H]
 \ForAll{component $cp$ to use}{
 create thread $t$
 $t$ uses the provide interface of $cp$
 }
 join all threads
 \label{alg:par}
 \caption{parallel function}
 \end{algorithm}

\medskip
In addition to this, a composition of the components $SSEQ$, $SEQ$, $PAR$ and $K$ is needed to be able to represent complex dependencies. For example, the example dependencies~(\ref{eq:dep}) is represented in the Figure~\ref{exallparall}.
%---------
\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
  \node[component] (SSEQ) at (0,0) {$SSEQ$};
  \node[component] (SEQ1) at (-2.5,-1) {$SEQ$};
  \node[component] (SEQ2) at (2.5,-1) {$SEQ$};
  \node[component] (A) at (-3.5,-2) {$A$};
  \node[component] (E) at (-1.5,-2) {$E$};
  \node[component] (PAR1) at (-2.5,-2) {$PAR$};
  \node[component] (AA) at (2,-2) {$A'$};
  \node[component] (PAR2) at (3,-2) {$PAR$};
  \node[component] (SEQ3) at (-3,-3) {$SEQ$};
  \node[component] (C) at (-2,-3) {$C$};
  \node[component] (SEQ4) at (2.5,-3) {$SEQ$};
  \node[component] (CC) at (3.5,-3) {$C$};
  \node[component] (B) at (-3.5,-4) {$B$};
  \node[component] (D) at (-2.5,-4) {$D$};
  \node[component] (BB) at (2,-4) {$B$};
  \node[component] (DD) at (3,-4) {$D$};

  \path[-]
    (SSEQ) edge [connection] node {} (SEQ1)
           edge [connection] node {} (SEQ2)
           edge [mpiconn] node {} (2,0)
    (SEQ1) edge [connection] node {} (A)
           edge [connection] node {} (PAR1)
           edge [connection] node {} (E)
    (SEQ2) edge [connection] node {} (AA)
           edge [connection] node {} (PAR2)
    (PAR1) edge [connection] node {} (SEQ3)
           edge [connection] node {} (C)
    (PAR2) edge [connection] node {} (SEQ4)
           edge [connection] node {} (CC)
    (SEQ3) edge [connection] node {} (B)
           edge [connection] node {} (D)
    (SEQ4) edge [connection] node {} (BB)
           edge [connection] node {} (DD);

  \end{tikzpicture}
\caption{Example assembly for coarse- and fine-grain data parallelism}
\label{exallparall}
\end{center}
\end{figure}

The general case of assembly to introduce data parallelism and task parallelism in numerical simulations is represented in the Figure~\ref{allparall}.

%---------
\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (SSEQ) at (0,0) {SSEQ};
   \node[component] (SEQ) [right=of SSEQ] {SEQ};
   \node[component] (PAR) [right=of SEQ] {PAR};
   \node[component] (K)   [right=of PAR] {K};

  \path[-]
    (SSEQ)  edge  [bend right=70,connection] node  [swap] {1,n} (PAR)
    (SSEQ)  edge  [connection]                node          {1,n} (SEQ)
            edge  [mpiconn] node {} (-1,0)
    (SEQ)	  edge  [connection]                node          {1,n} (PAR)
    (PAR)   edge  [bend right=40,connection]  node  [swap]  {1,n} (SEQ)
    (SEQ)   edge  [bend right=40,connection]  node  [swap]  {1,n} (SSEQ)
    (PAR)   edge  [bend left=50,connection]   node  [swap]  {1,n} (SSEQ)
    (SEQ)   edge  [bend left=60,connection]   node  [swap]  {1,n} (K)
    (PAR)	  edge  [connection]                node          {1,n} (K);
\end{tikzpicture}
\caption{General assembly description for data and task parallelism}
\label{allparall}
\end{center}
\end{figure}
%---------

%---------
%---------
\paragraph{Intermediate expressivity}
The component assembly defined in the Figure~\ref{allparall} is usefull to expose data parallelism, task parallelism, and to have a complete expressivity for dependencies between computations. For the general case of scientific applications, this assembly is needed, however, this assembly has a certain number of drawbacks. First, $SSEQ$, $SEQ$ and $PAR$ do not represent functionnalities of the application but control of functionnalities. As a result, the role of a component is twisted. Second, in the case of a complex application, with complex dependencies, the component assembly is going to be riddled with $SSEQ$, $SEQ$ and $PAR$ components, making difficult the reading of the application. Thus, it decreases the maintainability bring by component models. Finally, this expressivity is as complex as a \textit{task graph}~\cite{}, and as a result, two additional difficulties get out if such an assembly is used in our solution: the automatic detection of complex dependencies from a simple description of the application, and the \textit{scheduling} of such tasks. This last point is particularly critical and difficult. Because of the compositions of $SSEQ$, $SEQ$ and $PAR$ components illustrated in the Figure~\ref{allparall}, and because of the definition of $PAR$ which simply creates a thread for each component to use, it is possible that the work of each thread becomes unbalanced. For example, the Figure~\ref{unbalanced} represents an unbalanced example.

%---------
\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
  \node[component] (PAR1) at (0,0) {$PAR$};
  \node[component] (SEQ3) at (-1,-1) {$SEQ$};
  \node[component] (A) at (1,-1) {$A$};
  \node[component] (B) at (-2,-2) {$B$};
  \node[component] (C) at (-1,-2) {$C$};
  \node[component] (D) at (0,-2) {$D$};

  \path[-]
    (PAR1) edge [connection] node {} (SEQ3)
           edge [connection] node {} (A)
    (SEQ3) edge [connection] node {} (B)
           edge [connection] node {} (C)
           edge [connection] node {} (D);

  \end{tikzpicture}
\caption{Unbalanced assembly example}
\label{unbalanced}
\end{center}
\end{figure}

However, in existing component models for HPC (for exemple in $L^2C$) only the component object is defined, and all components are created, destroyed, and called the same way. Thus, it is not possible to manage a $PAR$ component to dynamically balance the workload on threads. For this reason, it seems that the combination of tasks and components are needed as well as a scheduler. The scheduling of a task graph is an active research domain and it could be an interesting perspective to work on the combination of $L^2C$ and StarPU~\cite{}, for example. However, this work aims to propose a light solution using component contributions, and to illustrate the advantages and drawbacks of such a solution. To only use components in our solution, and to avoid unbalanced and heavy assemblies, we propose an intermediate assembly which is an approximation of the general case of the Figure~\ref{allparall}. We claim, in the rest of this paper, that this approximation is enough expressive and produces efficient applications for the specific case of mesh-based numerical simulations.

The proposed intermediate assembly is represented in the Figure~\ref{approx}. The same basic association of components $SSEQ$, $SEQ$, $PAR$ and $K$ is found, however a limitation of the possible compositions of those components is applied. It is no longer possible to use a $SEQ$ or $SSEQ$ from a $PAR$ component, as well as using a $SSEQ$ from a $SEQ$ component and vice versa. The most important part of those limitations is to avoid too much imbalance. Actually, from a $PAR$ component it is only possible to use one computation $K$ in each created thread. In the specific case of mesh-based numerical simulations, as most numerical computations $c(R,w,D,e)$ (except boundary conditions) are applied on a complete set of mesh elements $D$ and computes a numerical expression $e$, two numerical computations are approximately balanced, or at least the imbalance is limited.

%If solving those problems for general cases seems needed in one hand (and this point is a perspective of this work), when focusing on mesh-based numerical simulations, such an expressivity is not essential on the other hand. To explain this claim, the Figure~\ref{approx} represents an approximation of the assembly of the Figure~\ref{allparall}. In this assembly, the available compositions are limited. Acually, a $SSEQ$ component exclusively uses multiple $SEQ$ components, which exclusively uses multiple $PAR$ or $K$. Finally, a $PAR$ component exclusively uses multiple $K$.

%---------
\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (SSEQ) at (0,0) {SSEQ};
   \node[component] (SEQ) [right=of SSEQ] {SEQ};
   \node[component] (PAR) [right=of SEQ] {PAR};
   \node[component] (K) [right=of PAR] {K};

  \path[->]
             (SSEQ)  edge  [connection]               node           {1,n}   (SEQ)
                    edge [mpiconn] node {} (-1,0)
             (SEQ)   edge [connection]          node       {1,n}   (PAR)
             (SEQ)   edge [bend left=100,connection]   node  [swap]   {1,n}   (K)
             (PAR)   edge [connection]          node       {1,n}   (K);
\end{tikzpicture}
\caption{General assembly description approximation}
\label{approx}
\end{center}
\end{figure}

Using this approximated assembly, the assembly of the example is given in the Figure~\ref{exapprox}. The limitation of such an approximation is clearer in this example: the dependencies $(B<D) \parallel C$ and $(B'<D') \parallel C'$ can only be expressed by $B \parallel C<D$ which is a less precised dependency. But as explained before, this limitation of expressivity also guarantees a limitation of the imbalance which is due to the use of components only, and not task graphs and schedulers. Finally, the proposed approximation still exposes the three level of parallelism introduced in the Section~\ref{}.%When introducing task parallelism scheduling, this limitation can be a big disadvantage if the task (or functionnality) $B$ is short compared to $C$. However, in the case of a mesh-based numerical simulation, and as explained in the Section~\ref{sect:formalism}, most numerical computation $c(R,w,D,e)$ of a stencil program $\mathcal{P}(\mathcal{M},\Delta,\Gamma,\mathcal{T})$ are computed on all elements of $E_i \subset \mathcal{M}$, and computes a single numerical expression $e$. As a result most computations are close in terms of amount of numerical operations. Thus, thanks to the specific domain we are focussing on, the approximated assembly of the Figure~\ref{approx} is satisfactory and it makes the solution lighter by the only use of the component model $L^2C$, and without the need of an external task graph tool and scheduler. Finally, even if this assembly is an approximation of the real dependencies assembly, the three level of parallelism are still available in the solution.

%---------
\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
  \node[component] (SSEQ) at (0,0) {$SSEQ$};
  \node[component] (SEQ1) at (-2.5,-1) {$SEQ$};
  \node[component] (SEQ2) at (2.5,-1) {$SEQ$};
  \node[component] (A) at (-4,-2) {$A$};
  \node[component] (PAR1) at (-3,-2) {$PAR$};
  \node[component] (D) at (-2,-2) {$D$};
  \node[component] (E) at (-1,-2) {$E$};
  \node[component] (AA) at (1.5,-2) {$A'$};
  \node[component] (PAR2) at (2.5,-2) {$PAR$};
  \node[component] (DD) at (3.5,-2) {$D$};
  \node[component] (B) at (-3.5,-3) {$B$};
  \node[component] (C) at (-2.5,-3) {$C$};
  \node[component] (BB) at (2,-3) {$B$};
  \node[component] (CC) at (3,-3) {$C$};

  \path[-]
    (SSEQ) edge [connection] node {} (SEQ1)
            edge [connection] node {} (SEQ2)
            edge [mpiconn] node {} (2,0)
    (SEQ1) edge [connection] node {} (A)
           edge [connection] node {} (PAR1)
           edge [connection] node {} (D)
           edge [connection] node {} (E)
    (SEQ2) edge [connection] node {} (AA)
           edge [connection] node {} (PAR2)
           edge [connection] node {} (DD)
    (PAR1) edge [connection] node {} (B)
           edge [connection] node {} (C)
    (PAR2) edge [connection] node {} (BB)
           edge [connection] node {} (CC);

  \end{tikzpicture}
\caption{Example assembly approximation}
\label{exapprox}
\end{center}
\end{figure}

### 2015-04-09 rapport/formalism.tex
rapport_090415

To numerically solve a set of PDEs, iterative methods are frequently used to approximate the solution more easily by a discretization of the time and the space (which is called a mesh). Thus, the PDEs are transformed to a set of numerical computations applied at each time step and on all elements of the mesh. Among the numerical computations is found a set of numerical schemes, also called stencil computations, and a set of auxiliary numerical computations also needed to perform the simulation. 
%A stencil is a computation applied, at each time iteration, to a complete set of elements of a mesh. 
In this paper is distinguished the concept of stencil shape (often called a stencil) from the concept of stencil computation. A stencil computation involves values of a simulated quantity (pressure, speed etc.) on a given element of the mesh and also on a given neighborhood of the element, denoted $\mathcal{N}$. A stencil shape is the shape of the needed neighborhood $\mathcal{N}$ in a stencil computation. As a result, a mesh-based numerical simulation is composed of a certain number of stencil computations, which is called in this paper \textit{multi-stencils}, involving one or more stencil shapes. In applied mathematics, an explicit numerical scheme is typically a stencil computation, but, on the other hand, an implicit numerical scheme also requires the resolution of a linear system at each time iteration. This last kind of scheme is above the scope of this work, however explicit schemes already represent a huge part of PDE solvers. In this section is defined the formalism of a \textit{stencil program}, a description of how to parallelize such programs and the associated definitions of a \textit{dependency} and a \textit{synchronization}. %Finally, a definition of a \textit{phase}, a \textit{group}, and a \textit{kernel}, from which numerical computations of a stencil program can be classified for any kind of mesh, and from which different levels of parallelism can be detected, are given.

%-------------------------------------
\subsection{Definitions}

A stencil program or application is a program in which stencil expressions (or computations) are computed among other kind of numerical expressions, which do not involve neighborhood information. If the space domain is denoted $\Omega$, the time domain $\tau$, a stencil program is denoted as 
\begin{equation}
\mathcal{P}(\mathcal{M},\Delta,\Gamma,\mathcal{T}),
\end{equation}
where $\mathcal{M}$ is a mesh which discretizes $\Omega$, $\Delta$ is the set of quantities and auxiliary data mapped onto $\mathcal{M}$, $\Gamma$ is the ordered set of numerical computations in the program, and $\mathcal{T}$ is the discretization of the time domain $\tau$. $\mathcal{M}$ can be considered as a set of different sets of elements $\{E_0,...,E_n\}$ and a connectivity between them. For example, $\mathcal{M}$ could be composed of $E_0$ the set of cells, $E_1$ the set of nodes and $E_2$ the set of edges, such that each cell is composed of four edges, and each edge is composed of a source and a destination node. 
%If the operator $A(B)$ means that $A$ is applied on $B$, a 
A stencil program is composed of a set of quantities to simulate (pressure, speed etc.), mapped on the mesh, which is denoted $\Delta(E_k), 0 \leq k \leq n$. For all $t \in \mathcal{T}$ the set of numerical computations are applied on the data, which is simply denoted by $\Delta=\Gamma(\Delta)$.%such that for $\Delta_1, \Delta_2 \subset \Delta$, $\Delta(t) = \Gamma(\Delta_1(t-1),\Delta_2(t))$. 
In a mathematical numerical expression, a quantity to simulate has $d$ dimensions for the space domain and one additional dimension for the time, for example $\mathcal{A}(x,y,z,t)$. However, the implementation of a quantity in $\Delta$ to a data structure $\delta$ does not represent the time dimension in the rest of this paper. Thus, for a given quantity $\mathcal{A}$ at $t$ and $t-1$, eventually two different data structures (arrays) $\delta, \delta'$ are used. %As a result, the expression of a computation can be simplified to $\Delta=\Gamma(\Delta_1)$, which means that in a stencil program, the set of quantities $\Delta$ are computed from a given subset of quantities $\Delta_1$ on which is applied the set of computations.

A numerical computation in $\Gamma$ is denoted 
\begin{equation}
c(R,w,D,e), 
\end{equation}
where $R \subset \Delta, w \in \Delta$ are respectively the set of data read and written during the numerical expression $e$ (a single data is written in a single computation), and $D$ is one of the subsets $E_i \subset \mathcal{M}$. It has to be noticed that during a numerical simulation, at each time iteration, all the elements of a mesh are computed. However, it happens that the computation of the mesh elements is splitted in different computations (for example the computation of the physical border). In this case additional $E_i$ can be specified for the mesh $\mathcal{M}$. %For this reason, the sub-domain concept is introduced. Thus, i
In a computation $c$, $\forall d \in D$, the numerical expression $e(d)$ is applied. If the number of computations in a stencil program $\mathcal{P}(\mathcal{M},\Delta,\Gamma,\mathcal{T})$ is $card(\Gamma)=m$, such that $\bigcup_{i=0}^{m-1}c_i = \Gamma$, then $\bigcup_{i=0}^{m-1}R_i \cup w_i = \Delta$. %Finally, in this work, and because of the choice to not represent the time dimension in a data $\delta \in \Delta$, a computation $c(R,w,D,e)$ always verifies $R \cap w = \emptyset$ otherwize an implicit numerical scheme has to be solve (a relation between a $\mathcal{A}(t)$ and $\mathcal{A}(t)$ itself).

Two different types of computations can be handled by a stencil program: stencil computations or numerical computations which do not involve neighborhood information (as $map$ or $zip$ in functionnal languages). The needed neighborhood $\mathcal{N}$ of a stencil computation is the shape of the stencil. For regular meshes (as for example in finite difference methods), a stencil shape is not difficult to express, however, in the general purpose, the neighborhood is specific to the kind of mesh it is applied on, and could be difficult to express. For example, the neighborhood for unstructured meshes is composed of different concepts which are explained by the operators of Laszlo and Dobkin~\cite{}. This paper gives a general definition of a stencil computation, with a generic vision of a neighborhood $\mathcal{N}$. %independently of a specific neighborhood shape which depends on the mesh used in the simulation. 
Actually, the solution presented in this paper aims to handle any kind of mesh. A stencil computation is denoted by the general notation
\begin{equation}
s(R,w,D,e,\mathcal{N}),
\end{equation}
where $R \subset \Delta, w \in \Delta$ are read and written in the numerical expression $e$, which uses a given neighborhood function denoted $\mathcal{N}$ on a sub-domain $D$. In a stencil computation $s$, $\forall d \in D$, the numerical expression is applied such that $w(d) = e(R(d),R(\mathcal{N}(d))$. Finally, in this work, a stencil computation $s(R,w,D,e,\mathcal{N})$ always verifies $R \cap w = \emptyset$, otherwize an implicit numerical scheme has to be solve (a relation between a a quantity $\mathcal{A}(t)$ and $\mathcal{A}(t)$ itself at the same time iteration).

The second kind of numerical computation is denoted
\begin{equation}
l(R,w,D,e).
\end{equation}
This is a local numerical expression, where $e$ does not involve $\mathcal{N}(d)$, and where $e$ is applied such that $w(d) = e(R(d))$.

%-------------------------------------
\subsection{Stencil program parallelization}
\label{sect:parall}
Mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.

Maybe the more important parallelization technique of such simulations is called \textit{SPMD} (Single Program, Multiple Data) which can also be called \textit{coarse-grain data parallelism}. This concept represents the well-known and broadly-used domain decomposition technique to split the mesh in balanced sub-meshes for available resources. Because of the good locality degree of numerical simulations, only a small number of additional elements, called a \textit{recorvery} or a \textit{ghost} area, are needed to exchange information computed by another resource to correctly compute stencils. As the amount of communications of such parallelization are rather small compared to the amount of computations (small neighborhood), and also by the addition of recovery of communications by computations, this parallelization technique has proved many times its efficiency and scalability on numerical simulations~\cite{}. This parallelization technique can be applied on shared and distributed memory architectures (clusters, multi-core CPUs), and it stays the most efficient parallelization technique if the numerical simulation has to be executed on a distributed memory architecture. For this reason, this parallelization stays a reference in the domain and is still studied for more complex simulations, using unstructured meshes, multi-grids techniques etc.

However, because of energy consumption constraints, most of the \textit{top500}~\footnote{http://www.top500.org} supercomputer centers are not only composed of clusters of multi-core machines, but also of accelerator cards such as GPGPU and many-cores. In addition to this, vectorization units are also available on most CPUs. This introduces another kind of parallelization, linked to the first one, called in this paper \textit{fine-grain data parallelism}, which is composed of \textit{SIMD} (Single Instruction, Multiple Data) and \textit{SIMT} (Single Instruction, Multiple Threads) parallelization techniques. Such techniques typically compute a numerical expression in parallel on a vector, using vectorization units or threads, which perfectly matches the type of computation in the nested loops of numerical simulations.

Finally, the last parallelization technique which can be applied on various architectures (clusters, multi-cores, many-cores, GPGPU), is the \textit{task parallelism} technique. Instead of spliting the data of the simulation (or the vector used in a numerical expression), the different tasks of the simulation are identified and scheduled over the available resources during the execution (at runtime)~\cite{}. As numerical computations of a simulation can be considered as tasks, linked together through data dependencies for example, this parallelization technique can be used for HPC numerical simulations too.

This section discussed the various parallelization techniques which can be used for numerical simulations. As a result, it seems that, by a combination of existing techniques, modern heterogeneous parallel architectures can be fully harnessed by such applications~\cite{}. In this paper is presented a component-based model to expose those three kinds of parallelism in the overall simulation, while hiding them from the user (numericians).

%-------------------------------------
\subsection{Dependencies and synchronizations}
\label{sect:dep}
To define and explain the component-based implicit parallelism model presented in this paper, two additional definitions handle the relations between data read and written into computations. Those relations are responsible for dependencies and synchronizations.

\medskip
\begin{mydef}
\textit{For two computations $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, it is said that $c_2$ is dependant from $c_1$, denoted $c_1<c_2$, if $w_1 \cap R_2 \neq \emptyset$. In this case, $c_1$ has to be computed before $c_2$. The binary relation $c_1<c_2$ represents a \textit{dependency}.}
\end{mydef}

\medskip
Dependencies lead to one computation before another, but when the program is parallelized some dependencies are also responsible for synchronizations between different processes. The opposite definition can also be given,

\medskip
\begin{mydef}
\textit{For two computations $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, it is said that $c_2$ is independant from $c_1$, denoted $c_1 \not <c_2$ or $c_1 \parallel c_2$, if $w_1 \cap R_2 = \emptyset$.}
\end{mydef}

Considering a coarse-grain data parallelized stencil program $\mathcal{P}$,

\medskip
\begin{mydef}
\textit{A depedency between two computations $c_1<c_2$ is a \textit{synchronisation}, denoted $c_1 \ll c_2$, if $c_2=s_2(R_2,w_2,D_2,e_2)$, and $w_1 \cap R_2 \neq \emptyset$.}
\end{mydef}

\medskip
Actually, in a parallel stencil program, synchronizations between processes are always needed for the set of data $R$ of a stencil $s(R,w,D,e)$ because the expression $e$ involves a neighborhood $\mathcal{N}$ which could be handled by another processor or resource.

### 2015-04-10 rapport/csm.tex
rapport_100415

%-------------------------------------
\subsection{Phase, group and kernel}
Using the formalism described in Sections~\ref{sect:formalism} and~\ref{sect:component}, the following definitions are given for the solution.

\medskip
\begin{mydef}
For a stencil program $\mathcal{P}(\mathcal{M},\Delta,\Gamma,\mathcal{T})$, a \emph{phase} $\Phi \subset \Gamma$ is a subset of ordered computations $\{c_0,...,c_{n-1}\} \in \Gamma$ such that $\forall 0 \leq i,j \leq n-1, \nexists c_i,c_j \in \Phi$, which verifies $c_i \ll c_j$.
\end{mydef}

One can notice, however, that in a phase $Phi$ of a stencil program it is possible to have a pair $c_i,c_j \in \Phi$, which verifies $c_i<c_j$. As a result, a phase is equivalent to a sequence of computations $SEQ$. The set of phases of a stencil program are ordered and a synchronization is needed between two different \emph{phases} of a stencil program $\mathcal{P}$. In other terms for two phases $\Phi_1$ and $\Phi_2$ $\exists c_i \in \Phi_1, c_j \in \Phi_2$ such that $c_i \ll c_j$. If $m$ is the number of stencil computations in a phase $\Phi_i$ among $n$ total computations, the synchronizations of $\Phi_i$ are applied on $\bigcup_{j=0}^{m-1}R_j$. Instead of introducing an additional definition for the equivalent of $SSEQ$, and because $sequence(sequence(a,b),sequence(c,d))=sequence(a,b,c,d)$, synchronizations needed by a new phase are commputed inside the Phase itself before computations. A phase component is defined as
\begin{equation}
Phase(\{provide,use-multiple, synchronization\},\{phase\}).
\end{equation}
 The function $phase$ performs the needed synchronizations before to use each component of the sequence

\begin{algorithm}[H]
 synchronizations\\
 \ForAll{component $cp$ to use}{
 use the provide interface of $cp$
 }
 \caption{phase function}
 \end{algorithm}

\begin{mydef}
For a stencil program $\mathcal{P}(\mathcal{M},\Delta,\Gamma,\mathcal{T})$, a \emph{group} $\mathcal{G} \subset \Gamma$ is a subset of unordered computations $\{c_0,...,c_{n-1}\}$ such that $\forall 0 \leq i,j \leq n-1, \nexists c_i,c_j \in \mathcal{G}$, which verifies $c_i<c_j$.
\end{mydef}

Thus, a dependency exists between two different \emph{groups} of a stencil program $\mathcal{P}$, but the computations inside a \emph{group} are unordered, without dependencies. As a result, the Group component is defined as $PAR$
\begin{equation}
Group(\{provide,use-multiple\},\{group\}),
\end{equation}
 where the function $group$ creates a thread for each component to use, and join all threads at the end.

\begin{algorithm}[H]
 \ForAll{component $cp$ to use}{
 create thread $t$
 $t$ uses the provide interface of $cp$
 }
 join all threads
 \label{alg:par}
 \caption{group function}
 \end{algorithm}

\begin{mydef}
For a stencil program $\mathcal{P}(\mathcal{M},\Delta,\Gamma,\mathcal{T})$, a \emph{kernel} $\mathcal{K} \subset \Gamma$ is a subset of unordered computations $\{c_0,...,c_{n-1}\}$ such that $\forall 0 \leq i,j \leq n-1, \nexists c_i,c_j \in \mathcal{K}$, which verifies $c_i<c_j$, and where $\forall i,j<n, D_i=D_j$.
\end{mydef}

\medskip
Thus, a \emph{kernel} could contains $n$ computations $\{c_0,...,c_{n-1}\}$, but handles a single space loop in a coarser computation $c'$ such that $c'(\bigcup_{i=0}^{n-1}R_i,\bigcup_{i=0}^{n-1}w_i,D,\{e_0,...,e_{n-1}\})$. This definition differs from the component $K$ described in this section as a kernel could contains more than one computation if the space domain is the same. This definition of a kernel improves performances as the memory bandwidth is better used. A kernel component if defined as
\begin{equation}
Kernel(\{provide\},\{kernel\}).
\end{equation}
 The function $kernel$ performs the set of numerical computations

\begin{algorithm}[H]
 \ForAll{elements $d$ in $D$}{
 $w(d)=e_1(R(d),R(\mathcal{N}(d)))$\\
 $w(d)=e_2(R(d),R(\mathcal{N}(d)))$\\
 ...
 }
 \caption{kernel function}
 \end{algorithm}

\medskip
The component assembly of a mesh-based numerical simulation using phase, group and kernel components is represented in the Figure~\ref{phgpk}. This assembly is equivalent to the one presented in the Figure~\ref{approx}, but $SSEQ$ and $SEQ$ are fusionned in $phase$, and the component $kernel$ is bit different from $K$.

%---------
\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (phase) at (0,0) {phase};
   \node[component] (group) [right=of phase] {group};
   \node[component] (kernel) [right=of group] {kernel};

  \path[->]
             (phase)  edge  [connection]               node           {1,n}   (group)
                      edge [mpiconn] node {} (-1,0)
             (group)   edge [connection]          node       {1,n}   (kernel)
             (phase)   edge [bend left=100,connection]   node  [swap]   {1,n}   (kernel);
\end{tikzpicture}
\caption{General assembly proposed in our solution}
\label{phgpk}
\end{center}
\end{figure}

%-------------------------------------
\subsection{Overall model}

%---------
\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (phase) at (0,0) {Phase};
   \node[component] (group) [right=of phase] {Group};
   \node[component] (kernel) [right=of group] {Kernel};
   \node[dcomponent] (init) [above=1.5cm of group] {Initializer};
   \node[component] (data) [above=1.5cm of phase] {Data};
   \node[component] (time) [left=of phase] {Time};
   \node[component] (drapp) [above=1.5cm of time] {DrApp};
   \node[dcomponent] (conv) [left=of time] {Conv};
   \node[component] (dr) [above=1.5cm of drapp] {Driver};
   \node[component] (dds) [above=1.5cm of data] {DDS};
   \node[component] (connec) [above=1.5cm of init] {Connector};

  \path[->]
  	(dr)		edge  [connection]	node	{}	(dds)
  	(dr)		edge  [connection]	node	{}	(drapp)
  	(drapp)		edge  [connection]	node	{1,n}	(data)
  	(drapp)		edge  [connection]	node	{}	(time)
  	(data)  	edge  [connection]	node	{}	(dds)
  	(data)  	edge  [connection]	node	{}	(connec)
  	(connec)	edge  [connection]	node	{}	(dds)
  				edge  [mpiconn]		node	{}	(4,3)
  	(data)		edge  [connection]	node	{1,n}	(init)
  	(time)		edge  [connection]	node	{}	(conv)
  	(time)		edge  [connection]	node	{1,n}	(phase)
  	(phase)		edge  [connection]	node	{1,n}	(data)
  	(phase)		edge  [connection]	node	{1,n}	(group)
  	(phase)		edge  [bend right=50,connection]   node  [swap]   {1,n}   (kernel)
  	(group)		edge  [connection]	node	{1,n}	(kernel)
  	(kernel)	edge  [connection]	node	{1,n}	(data);
\end{tikzpicture}
\caption{Overall assembly of the Component Stencil Model.}
\label{phgpk}
\end{center}
\end{figure}

### 2015-04-10 rapport/csm.tex
rapport_100415_with_csm

@@
 If $m$ is the number of stencil computations in a phase $\Phi_i$ among $n$ total computations, the synchronizations of $\Phi_i$ are applied on $\bigcup_{j=0}^{m-1}R_j$.
+
 Instead of introducing an additional definition for the equivalent of $SSEQ$, and because $sequence(sequence(a,b),sequence(c,d))=sequence(a,b,c,d)$, synchronizations needed by a new phase are commputed inside the Phase itself before computations.
@@
 \medskip
-The component assembly of a mesh-based numerical simulation using phase, group and kernel components is represented in the Figure~\ref{phgpk}.
+The component assembly of the computations of a mesh-based numerical simulation using phase, group and kernel components is represented in the Figure~\ref{phgpk}.
 This assembly is equivalent to the one presented in the Figure~\ref{approx}, but $SSEQ$ and $SEQ$ are fusionned in $phase$, and the component $kernel$ is bit different from $K$.
@@
 
+This component assembly of computations can be, by definition, extracted from an ordered list of typed computations $s(R,w,D,e,\mathcal{N})$ or $l(R,w,D,e)$.
+$R$, $w$ and $D$ are used to build Phases, Groups and Kernels, while $e$ and $\mathcal{N}$ are used inside the Kernel components.
+As a result, we can define a function $\lambda$ to build the computation assembly
+\begin{equation}
+\lambda : list_o(c(R,w,D,e)) \rightarrow \alpha ({{Phase},{Group},{Kernel}},L,R)
+\end{equation}
+
 %-------------------------------------
-\subsection{Overall model}
+\subsection{Overall assembly}
+
+\paragraph{The SIPSim model.} The SIPSim model~\cite{} is a model which proposes a systematic way to create an SPMD implicit parallelism solution for mesh-based numerical simulations through four concepts.
+The first concept is a \textit{Distributed Data Structure} (DDS), which represents the distributed mesh.
+Thus, the DDS manages a domain decomposition, or more broadly a graph partitioning problem which balances the number of elements on available resources, and which tries to minimize the number of needed synchronizations between resources.
+It has to be noticed that this particular task could be managed by an external graph partitionner.
+In addition to this, the DDS also has to offer an efficient access to an element of a mesh and to its neighborhood.
+The second concept is the \textit{Distributed Property Map} (DPMap), which is responsible to map a quantity to simulate (a data) onto the distributed mesh.
+The third concept is an \textit{applicator}, which is responsible for applying a set of synchronizations before a set of numerical computations written by the user (also called operations).
+Finally, the last concept contains the \textit{interfaces} needed to write an operation in a sequential programming style while using DPMaps mapped onto the DDS.
+Those interfaces are directly linked to the type of DDS used.
+
+\paragraph{CSM overall assembly.} From the SIPSim model, CSM uses the concept of DDS and DPMap and transforms them in components to build the overall component assembly.
+The implementation of those components are not described in this paper, and it is possible to imagine an adhoc implementation in simple cases, as a Cartesian mesh, or existing external implementations (libraries) in more complex cases.
+This choice is an implementation choice and will be compatible with CSM.
+In addition to those two new components, the concept of \textit{interface} of the SIPSim model has to be kept to write the implementation of the expressions $e$ of the numerical computations on distributed data structures.
+But this concept is not needed as a component.
+It could be provided by the DDS itself or it could be introduced as a header file for example.
+
+The overall assembly produced by the model is presented in the Figure~\ref{csmassembly}.
+First, the component \textit{DDS} and the component \textit{Data}, which corresponds to the DPMap concept of the SIPSim model, are introduced in the assembly with the same functionnality than described above.
+
+Second, one can notice that the assembly described in the previous Figure~\ref{phgpk} is, of course, inserted in the overall assembly.
+However, even if the \textit{Phase} component is still responsible for starting a set of synchronizations, those synchronizations are in fact performed by the \textit{Connector} component.
+Actually, the \textit{Phase} component uses the data concerned by its synchronizations, and then each data uses the \textit{Connector}, which finally performs the synchronizations.
+As a result, the \textit{Connector} component can be implemented with threads, or with MPI, without the modification of \textit{Phase} and \textit{Data} components.
+This is a separation of concerns between the data of the simulation and the parallelization strategy which increases the portability of the application.
+This strategy has been more described in the paper~\cite{}.
+
+The third important point to notice is the introduction of the component \textit{Time} which is responsible for the time iteration steps of the simulation.
+To establish the end of time steps, most of the time a convergence function is computed on data, but it also happens that a fix number of iterations is precised.
+For this reason, the instanciation of the component \textit{Convergence} is optional.
+The \textit{Driver} component starts the simulation and launches the initialization of the DDS, and the \textit{DrApp} component starts the initialization of the data and also the time loop of the simulation.
+Finally, the \textit{Initializer} component is linked to the initialization of data (read a file format, default value, or numerical computation).
+The computation components, specific to each simulation, are \textit{Kernel} and \textit{Convergence}, and to perform numerical expressions $e$ those components have to be connected to the Data of each $R$ and $w$.
 
@@
    \node[component] (kernel) [right=of group] {Kernel};
-   \node[dcomponent] (init) [above=1.5cm of group] {Initializer};
+   \node[component] (init) [above=1.5cm of group] {Initializer};
    \node[component] (data) [above=1.5cm of phase] {Data};
@@
    \node[component] (drapp) [above=1.5cm of time] {DrApp};
-   \node[dcomponent] (conv) [left=of time] {Conv};
+   \node[dcomponent] (conv) [left=of time] {Convergence};
    \node[component] (dr) [above=1.5cm of drapp] {Driver};
@@
   	(group)		edge  [connection]	node	{1,n}	(kernel)
-  	(kernel)	edge  [connection]	node	{1,n}	(data);
+  	(kernel)	edge  [connection]	node	{1,n}	(data)
+  	(conv)	edge  [connection,dashed]	node	{1,n}	(data);
 \end{tikzpicture}
 \caption{Overall assembly of the Component Stencil Model.}
-\label{phgpk}
+\label{csmassembly}
 \end{center}
 \end{figure}
+
+This overall assembly can be compiled from a set of sequential and descriptive information. First an unordered list of data is needed. A data is defined as
+\begin{equation}
+data(init_f,E_i,type)
+\end{equation}
+, where $init_f$ is the function to initialize the data (in the Initializer component), $Ei$ is the set of elements of the mesh on which the data is applied, and $type$ is the actual type of data ($int$, $double$ etc.).
+
+The second needed information is the description of the mesh to use as
+\begin{equation}
+mesh(type,dimension,size)
+\end{equation}
+, where $type$ of is the topology of the mesh (for example Cartesian or k-sided unstructured meshes), $dimension$ is the dimension of the mesh, and $size$ is the size of the mesh.
+
+As a result, we can define a function $\Lambda$ to build the overall simulation assembly of the Figure~\ref{csmassembly} as
+\begin{equation}
+\Lambda : list_o(c(R,w,D,e)) \times list(data(init_f,E_i,type)) \times mesh(type,dimension,size) \rightarrow \alpha (\mathcal{C},L,R)
+\end{equation}
+
+The needed information of the function $Lambda$ are typically produced by numericians who try to solve a set of partial differential equations by existing numerical methods, by a combination of methods, or by their own methods.
+Once the parallel assembly $\alpha$ is created, the engineer or a numerican developper of the lab can write the set of numerical computation components $Kernel$, the set of $Initializer$ components, and eventually the $Convergence$ component.
+All those components are written with an imperative sequential programming style using the interfaces linked to the DDS.
+As a result, multiple levels of separation of concerns are introduced.
+First the separation of concerns between the numerician and the person in charge of the implementation, and second the separation of concerns between the implementation and the parallelization of the simulation, which is generated.
+
+%\paragraph{Evaluation.}
+% for different kind of mesh

### 2015-04-16 rapport/component.tex
rapportv1_160415

@@
 %-------------------------------------
-\subsection{Component models}
-A \textit{component} is an independent object which represents a functionnality of an application, and which is associated to a set of \textit{interfaces} used to plug an instance of a component with other ones.
-The object composed of multiple component instances linked together through their interfaces is called a component assembly.
-As a result, a component assembly is the code of an entire application represented as a set of splitted functionnalities.
-Such software engineering model is known to increase \textit{code re-use} and \textit{productivity}, as one component can be used in different applications, but also \textit{maintainability} and \textit{scalability}, as functionnalities of an application are clearly splitted in different and independent components.
-More recently, component models have been studied to write HPC applications such as $L^2C$~\cite{} (Low Level Component Model) and CCA~\cite{} (Common Component Architecture).
-
-In the rest of this paper, the component model $L^2C$ is considered and used by the solution. A component is defined as
-\begin{equation}
-C(I,F),
-\end{equation}
-where $I$ is a set of available interfaces for the given component, and $F$ is the set of functions to answer the functionnality of the component.
-A component instance, as for an object, is the actual creation of the component, and more than one instance can be created.
-The components of $L^2C$ are limited to a few interfaces without the introduction of overheads.
-In this paper we are interested in four types of interfaces.
-The three first one are \textit{use} and \textit{provide} interfaces (Figure~\ref{AB}) to use a functionnality provided by another component instance, and \textit{use-multiple} interface (Figure~\ref{ABC}) to use a vector of functionnalities provided by a set of component instances.
-A condensed notation of a use-multiple is given in the Figure~\ref{ABCbis}.
-
-\begin{figure}[h!]
-\begin{center}
-  \subfigure[][$A$ uses $B$ and $A$ uses $C$]{
-  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (A) at (0,0) {$A$};
-   \node[component] (B) at (2,+0.5) {$B$};
-   \node[component] (C) at (2,-0.5) {$C$};
-
-  \path[-]
-    ([yshift=1ex]A.east) edge [connection] (B.west)
-    ([yshift=-1ex]A.east) edge [connection] (C.west);
-  \end{tikzpicture}
-  \label{AB}
-  }
-  \hspace{2cm}
-  \subfigure[][$A$ uses multiple $B$ instances]{
-  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (A) at (0,0) {$A$};
-   \node[component] (B) at (2,+0.5) {$B_1$};
-   \node[component] (C) at (2,-0.5) {$B_2$};
-
-  \path[-]
-    (A.east) edge [connection] node {} (B.west)
-             edge [connection] node {} (C.west);
-
-  \end{tikzpicture}
-  \label{ABC}
-  }
-  \hspace{2cm}
-  \subfigure[][$A$ uses multiple $B$ instances (condensed notation)]{
-  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (A) at (0,0) {$A$};
-   \node[component] (B) at (2,0) {$B$};
-
-  \path[-]
-    (A.east) edge [connection] node {1,n} (B.west);
-
-  \end{tikzpicture}
-  \label{ABCbis}
-  }
-  \caption{$L^2C$ use-provide and use-multiple interfaces}
-  \label{interfaces}
-\end{center}
-\end{figure}
-
-An instance $i$ of a component has the same properties than the component it is instanciated from, but it is also asociated to a resource $r$ (a core, a thread etc.) of the set of all available resources $R$.
-\begin{equation}
-C_i(I,F,r)
-\end{equation}
-A component assembly represents the application as a set of component instances and the composition of those instances through their interfaces. A component assembly is defined as
-\begin{equation}
-\alpha (\mathcal{C},L,R),
-\end{equation}
-where $\mathcal{C}$ is the set of component instances used, $L$ is the set of links between interfaces of component instances, and $R$ is the set of available resources.
-The component assembly of $L^2C$ offers a way to write an assembly for each MPI process, and to connect some of their components by MPI communicators.
-This connection is an \textit{MPI interface}.
-In the rest of this paper we consider a broader interface called \textit{synchronization}, which corresponds to the MPI interface of $L^2C$ but which could also be a \textit{thread synchronization} instead.
-
-Using those four interfaces (use, use-multiple, provide and synchronization), it is possible to write an application with medium-grain components, enough fine to keep efficiency and to stay close to the execution model, and enough coarse to take advantage of components.
-The efficiency of $L^2C$ has been proved, while increasing the productivity and maintainability of HPC applications, but moreover it has also been shown that the protability of such component-based HPC application is improved~\cite{}.
+\subsection{A primitive component model}
+Definitions of a primitive component (ports, interfaces), a component instanciation, an assembly + simple semantics on the behavior of a primitive component.
+Cite L2C as an implementation of such a model + introduction of notations (use,provide,use-multiple,sync).
 
 %-------------------------------------
-\subsection{Stencil program parallelization using components}
-In this section is studied the parallelization of a numerical simulation using the given definitions of a component, an assembly and four interfaces.
-First, the different parallelization techniques explained in the Section~\ref{sect:parall} are translated to component-based parallel assemblies, and an approximation of the general case is discussed for the solution presented in this paper.
-In the rest of this section, the following example of dependencies is used
-\begin{equation}
-A<((B<D) \parallel C)<E \ll A'<((B'<D') \parallel C'),
-\label{eq:dep}
-\end{equation}
- where $A,B,C,D,E,A',B',C',D'$ are numerical computations.
-%In this example, $\parallel$ is a binary expression which means that a parallel execution of the right and left computations is possible.
-It is equivalent to $\not>$.
-This example, can be represented as a directed acyclic graph as in the Figure~\ref{example}.
+\subsection{Control components}
+Definition of three control components, and of the kernel component. Explain the behavior.
 
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (A) at (0,0) {$A$};
-   \node[component] (B) at (1,+0.5) {$B$};
-   \node[component] (C) at (1,-0.5) {$C$};
-   \node[component] (D) at (2,+0.5) {$D$};
-   \node[component] (E) at (3,0) {$E$};
-   \node[component] (AA) at (4,0) {$A'$};
-   \node[component] (BB) at (5,+0.5) {$B'$};
-   \node[component] (CC) at (5,-0.5) {$C'$};
-   \node[component] (DD) at (6,+0.5) {$D'$};
-
-  \path[->]
-    (A) edge node {} (B)
-        edge node {} (C)
-    (B) edge node {} (D)
-    (D) edge node {} (E)
-    (C) edge node {} (E)
-    (E) edge node {} (AA)
-    (AA) edge node {} (BB)
-        edge node {} (CC)
-    (BB) edge node {} (DD);
-  \end{tikzpicture}
-  \caption{Dependencies example~(\ref{eq:dep}) represented as a DAG.}
-  \label{example}
-\end{center}
-\end{figure}
-
-%---------
-%---------
-\paragraph{Coarse- and fine-grain data parallelism.} 
-As explained in the Section~\ref{sect:parall}, to apply coarse-grain data parallelism on numerical simulations, a domain decomposition (or graph partitioning) has to be done on the mesh and also on the data mapped on it.
-When this important step is solved, the same program can be applied on each subpart of the data, on the different resources.
-In addition to this, a set of synchronizations are needed between the resources at each time step to compute the neighborhood $\mathcal{N}$ correctly.
-When using a component model, the computing part of the application, which is called for each time step, can be seen as a component assembly which is duplicated on different resources, which is possible using $L^2C$ for example.
-However, it is needed to classify computations in the component assembly to identify when synchronizations are needed (denoted $\ll$ in the Section~\ref{sect:dep}).
-For this reason, a \textit{sequence} component $SEQ$, which uses one after the other a set of computation components, is defined as
-\begin{equation}
-SEQ(\{provide,use-multiple\},\{sequence\}).
-\end{equation} 
-Thus, the \textit{sequence} function (represented in the Algorithm~(\ref{alg:seq})) is responsible for the loop which calls as much \textit{uses} as the number of components connected to the \textit{use-multiple}.
-In addition to this, a \textit{synchronized sequence} component $SSEQ$, which uses one after the other a set of computation components, but which also proceeds synchronizations between them, is defined as
-\begin{equation}
-SSEQ(\{provide,use-multiple, synchronization\},\{ssequence\}).
-\end{equation}
-In this case, the \textit{ssequence} function (represented in the Algorithm~(\ref{alg:sseq})) is responsible for a loop of two steps, first a synchronization step, second a use step.
-Finally, each numerical computation is represented by a component
-\begin{equation}
-K(\{provide\},\{compute\}),
-\end{equation}
- where the function \textit{compute} (represented in the Algorithm~(\ref{alg:comp})) is responsible for the sapce loop and the numerical computation on each element of the mesh.
-
-\begin{center}
-\begin{minipage}[t]{7cm}
-  \vspace{0pt} 
- \begin{algorithm}[H]
- \ForAll{component $cp$ to use}{
- use the provide interface of $cp$
- }
- \caption{sequence function}
- \label{alg:seq}
- \end{algorithm}
- \end{minipage}%
- \hspace{0.5cm}
-\begin{minipage}[t]{7cm}
-  \vspace{0pt}
- \begin{algorithm}[H]
- \ForAll{component $cp$ to use}{
- synchronizations\\
- use the provide interface of $cp$
- }
- \label{alg:sseq}
- \caption{ssequence function}
- \end{algorithm}
- \end{minipage}%
-\end{center}
-
-%\begin{minipage}[t]{5cm}
- % \vspace{0pt}
- \begin{center}
- \begin{algorithm}[H]
- \ForAll{elements $d$ in $D$}{
- $w(d)=e(R(d),R(\mathcal{N}(d)))$
- }
- \caption{compute function}
- \label{alg:comp}
- \end{algorithm}
- \end{center}
-%\end{minipage}%
-Using the three components $SSEQ$, $SEQ$, and $K$, the component assembly of the dependencies~(\ref{eq:dep}) is represented in the Figure~~\ref{exdataparall}.
-One can notice, in this Figure, that the parallel dependencies ($\parallel$) are lost.
-However, in a SPMD parallelization, the parallelism is obtained from executing the same program on different part of the data.
-As a result, a correct sequence of computations is sufficient to expose coarse-grain data parallelism.
-%---------
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-  \node[component] (SSEQ) at (0,0) {$SSEQ$};
-  \node (sync) at (4,0) {Synchronizations};
-  \node[component] (SEQ1) at (-2.5,-1) {$SEQ$};
-  \node[component] (SEQ2) at (2.5,-1) {$SEQ$};
-  \node[component] (A) at (-4,-2) {$A$};
-  \node[component] (B) at (-3,-2) {$B$};
-  \node[component] (C) at (-2,-2) {$C$};
-  \node[component] (D) at (-1,-2) {$D$};
-  \node[component] (E) at (0,-2) {$E$};
-  \node[component] (AA) at (1,-2) {$A'$};
-  \node[component] (BB) at (2,-2) {$B'$};
-  \node[component] (CC) at (3,-2) {$C'$};
-  \node[component] (DD) at (4,-2) {$D'$};
-
-  \path[-]
-    (SSEQ) edge [connection] node {} (SEQ1)
-           edge [connection] node {} (SEQ2)
-           edge [mpiconn] node {} (sync)
-    (SEQ1) edge [connection] node {} (A)
-           edge [connection] node {} (B)
-           edge [connection] node {} (C)
-           edge [connection] node {} (D)
-           edge [connection] node {} (E)
-    (SEQ2) edge [connection] node {} (AA)
-           edge [connection] node {} (BB)
-           edge [connection] node {} (CC)
-           edge [connection] node {} (DD);
-
-  \end{tikzpicture}
-\caption{Example assembly for coarse- and fine-grain data parallelism}
-\label{exdataparall}
-\end{center}
-\end{figure}
-The overall computation assembly of an SPMD numerical simulation is composed of a duplication of the computation (or dependencies) assembly described above.
-Each duplication is applied on an available resource as illustrated in the Figure~\ref{overall}.
-In the rest of this paper, this overall computation assembly of an SPMD application (coarse-grain data parallelism) is represented as a single computation assembly, and the $SSEQ$ component indicate where the synchronizations are needed.
-
-%---------
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-  \node[smallcp] (SSEQ) at (0,0) {$SSEQ$};
-  \node (sync) at (3.5,0) {Synchronizations};
-  \node[smallcp] (SEQ1) at (-1.5,-1) {$SEQ$};
-  \node[smallcp] (SEQ2) at (1.5,-1) {$SEQ$};
-  \node[smallcp] (A) at (-2.5,-2) {$A$};
-  \node[smallcp] (B) at (-2,-2) {$B$};
-  \node[smallcp] (C) at (-1.5,-2) {$C$};
-  \node[smallcp] (D) at (-1,-2) {$D$};
-  \node[smallcp] (E) at (-0.5,-2) {$E$};
-  \node[smallcp] (AA) at (0.5,-2) {$A'$};
-  \node[smallcp] (BB) at (1.1,-2) {$B'$};
-  \node[smallcp] (CC) at (1.9,-2) {$C'$};
-  \node[smallcp] (DD) at (2.5,-2) {$D'$};
-  \draw [resource] (-3,-2.5) rectangle +(6cm, 3cm);
-  \node (R1) at (0,-3) {r1};
-
-  \node[smallcp] (SSEQ2) at (7,0) {$SSEQ$};
-  \node[smallcp] (SEQ3) at (5.5,-1) {$SEQ$};
-  \node[smallcp] (SEQ4) at (8.5,-1) {$SEQ$};
-  \node[smallcp] (A2) at (4.5,-2) {$A$};
-  \node[smallcp] (B2) at (5,-2) {$B$};
-  \node[smallcp] (C2) at (5.5,-2) {$C$};
-  \node[smallcp] (D2) at (6,-2) {$D$};
-  \node[smallcp] (E2) at (6.5,-2) {$E$};
-  \node[smallcp] (AA2) at (7.5,-2) {$A'$};
-  \node[smallcp] (BB2) at (8.1,-2) {$B'$};
-  \node[smallcp] (CC2) at (8.9,-2) {$C'$};
-  \node[smallcp] (DD2) at (9.5,-2) {$D'$};
-  \draw [resource] (4,-2.5) rectangle +(6cm, 3cm);
-  \node (R2) at (7,-3) {r2};
-
-  \path[-]
-    (SSEQ) edge [connection] node {} (SEQ1)
-           edge [connection] node {} (SEQ2)
-           edge [mpiconn] node {} (sync)
-    (SEQ1) edge [connection] node {} (A)
-           edge [connection] node {} (B)
-           edge [connection] node {} (C)
-           edge [connection] node {} (D)
-           edge [connection] node {} (E)
-    (SEQ2) edge [connection] node {} (AA)
-           edge [connection] node {} (BB)
-           edge [connection] node {} (CC)
-           edge [connection] node {} (DD)
-    (SSEQ2) edge [connection] node {} (SEQ3)
-           edge [connection] node {} (SEQ4)
-           edge [mpiconn] node {} (sync)
-    (SEQ3) edge [connection] node {} (A2)
-           edge [connection] node {} (B2)
-           edge [connection] node {} (C2)
-           edge [connection] node {} (D2)
-           edge [connection] node {} (E2)
-    (SEQ4) edge [connection] node {} (AA2)
-           edge [connection] node {} (BB2)
-           edge [connection] node {} (CC2)
-           edge [connection] node {} (DD2);
-
-  \end{tikzpicture}
-\caption{Overall SPMD computation assembly of the example~(\ref{eq:dep}) on two resources.}
-\label{overall}
-\end{center}
-\end{figure}
-
-But the assembly obtained from the instanciation of $SEQ$, $SSEQ$, and $K$ does not only expose coarse-grain data parallelism.
-Actually, the generic notation of a computation component $K$, which represents a numerical computation, is important to identifiy the different computations of a numerical simulation as different functionnalities of the program, but in addition to this, exposing a computation as a component, also exposes the fine-grain data parallelism level.
-In fact, a numerical computation is nothing more than a loop (or a nested loop) on the mesh elements and a numerical expression to compute (Algorithm~(\ref{alg:comp})).
-As a result a computation component is ideal to introduce SIMD and SIMT parallelism.
-
-The general case of assembly to introduce coarse- and fine-grain data parallelism in numerical simulations is represented in the Figure~\ref{dataparall}.
-
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (SSEQ) at (0,0) {SSEQ};
-   \node[component] (SEQ) [right=of SSEQ] {SEQ};
-   \node[component] (K) [right=of SEQ] {K};
-
-  \path[-]
-    (SSEQ)  edge [connection] node {1,n} (SEQ)
-            edge [mpiconn] node {} (-1,0)
-    (SEQ)	 edge [connection] node {1,n}	(K);
-\end{tikzpicture}
-\caption{General assembly description for coarse- and fine-grain data parallelism}
-\label{dataparall}
-\end{center}
-\end{figure}
-%---------
-
-%---------
-%---------
-\paragraph{Data and task parallelism}
-As also explained in the Section~\ref{sect:parall}, to apply task parallelism to numerical simulations, it is needed to identify the different tasks of the application and their dependencies.
-With the introduction of $SSEQ$, $SEQ$, and $K$ components, a part of this work is already done.
-Actually, each computation component $K$ represents a task.
-However, to be able to express all dependencies between computations, the $PAR$ component is defined as
-\begin{equation}
-PAR(\{provide,use-multiple\},\{parallel\}),
-\end{equation}
- and corresponds to the expression $\parallel$ or $\not<$. The function $parallel$ creates a thread for each component to use and join all threads at the end.
-
-\begin{algorithm}[H]
- \ForAll{component $cp$ to use}{
- create thread $t$
- $t$ uses the provide interface of $cp$
- }
- join all threads
- \label{alg:par}
- \caption{parallel function}
- \end{algorithm}
-
-\medskip
-In addition to this, a composition of the components $SSEQ$, $SEQ$, $PAR$ and $K$ is needed to be able to represent complex dependencies.
-For example, the example dependencies~(\ref{eq:dep}) is represented in the Figure~\ref{exallparall}.
-%---------
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-  \node[component] (SSEQ) at (0,0) {$SSEQ$};
-  \node[component] (SEQ1) at (-2.5,-1) {$SEQ$};
-  \node[component] (SEQ2) at (2.5,-1) {$SEQ$};
-  \node[component] (A) at (-3.5,-2) {$A$};
-  \node[component] (E) at (-1.5,-2) {$E$};
-  \node[component] (PAR1) at (-2.5,-2) {$PAR$};
-  \node[component] (AA) at (2,-2) {$A'$};
-  \node[component] (PAR2) at (3,-2) {$PAR$};
-  \node[component] (SEQ3) at (-3,-3) {$SEQ$};
-  \node[component] (C) at (-2,-3) {$C$};
-  \node[component] (SEQ4) at (2.5,-3) {$SEQ$};
-  \node[component] (CC) at (3.5,-3) {$C$};
-  \node[component] (B) at (-3.5,-4) {$B$};
-  \node[component] (D) at (-2.5,-4) {$D$};
-  \node[component] (BB) at (2,-4) {$B$};
-  \node[component] (DD) at (3,-4) {$D$};
-
-  \path[-]
-    (SSEQ) edge [connection] node {} (SEQ1)
-           edge [connection] node {} (SEQ2)
-           edge [mpiconn] node {} (2,0)
-    (SEQ1) edge [connection] node {} (A)
-           edge [connection] node {} (PAR1)
-           edge [connection] node {} (E)
-    (SEQ2) edge [connection] node {} (AA)
-           edge [connection] node {} (PAR2)
-    (PAR1) edge [connection] node {} (SEQ3)
-           edge [connection] node {} (C)
-    (PAR2) edge [connection] node {} (SEQ4)
-           edge [connection] node {} (CC)
-    (SEQ3) edge [connection] node {} (B)
-           edge [connection] node {} (D)
-    (SEQ4) edge [connection] node {} (BB)
-           edge [connection] node {} (DD);
-
-  \end{tikzpicture}
-\caption{Example assembly for coarse- and fine-grain data parallelism}
-\label{exallparall}
-\end{center}
-\end{figure}
-
-The general case of assembly to introduce data parallelism and task parallelism in numerical simulations is represented in the Figure~\ref{allparall}.
-
-%---------
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (SSEQ) at (0,0) {SSEQ};
-   \node[component] (SEQ) [right=of SSEQ] {SEQ};
-   \node[component] (PAR) [right=of SEQ] {PAR};
-   \node[component] (K)   [right=of PAR] {K};
-
-  \path[-]
-    (SSEQ)  edge  [bend right=70,connection] node  [swap] {1,n} (PAR)
-    (SSEQ)  edge  [connection]                node          {1,n} (SEQ)
-            edge  [mpiconn] node {} (-1,0)
-    (SEQ)	  edge  [connection]                node          {1,n} (PAR)
-    (PAR)   edge  [bend right=40,connection]  node  [swap]  {1,n} (SEQ)
-    (SEQ)   edge  [bend right=40,connection]  node  [swap]  {1,n} (SSEQ)
-    (PAR)   edge  [bend left=50,connection]   node  [swap]  {1,n} (SSEQ)
-    (SEQ)   edge  [bend left=60,connection]   node  [swap]  {1,n} (K)
-    (PAR)	  edge  [connection]                node          {1,n} (K);
-\end{tikzpicture}
-\caption{General assembly description for data and task parallelism}
-\label{allparall}
-\end{center}
-\end{figure}
-%---------
-
-%---------
-%---------
-\paragraph{Intermediate expressivity}
-The component assembly defined in the Figure~\ref{allparall} is usefull to expose data parallelism, task parallelism, and to have a complete expressivity for dependencies between computations.
-For the general case of scientific applications, this assembly is needed, however, this assembly has a certain number of drawbacks.
-First, $SSEQ$, $SEQ$ and $PAR$ do not represent functionnalities of the application but control of functionnalities.
-As a result, the role of a component is twisted.
-Second, in the case of a complex application, with complex dependencies, the component assembly is going to be riddled with $SSEQ$, $SEQ$ and $PAR$ components, making difficult the reading of the application.
-Thus, it decreases the maintainability bring by component models.
-Finally, this expressivity is as complex as a \textit{task graph}~\cite{}, and as a result, two additional difficulties get out if such an assembly is used in our solution: the automatic detection of complex dependencies from a simple description of the application, and the \textit{scheduling} of such tasks.
-This last point is particularly critical and difficult.
-Because of the compositions of $SSEQ$, $SEQ$ and $PAR$ components illustrated in the Figure~\ref{allparall}, and because of the definition of $PAR$ which simply creates a thread for each component to use, it is possible that the work of each thread becomes unbalanced.
-For example, the Figure~\ref{unbalanced} represents an unbalanced example.
-
-%---------
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-  \node[component] (PAR1) at (0,0) {$PAR$};
-  \node[component] (SEQ3) at (-1,-1) {$SEQ$};
-  \node[component] (A) at (1,-1) {$A$};
-  \node[component] (B) at (-2,-2) {$B$};
-  \node[component] (C) at (-1,-2) {$C$};
-  \node[component] (D) at (0,-2) {$D$};
-
-  \path[-]
-    (PAR1) edge [connection] node {} (SEQ3)
-           edge [connection] node {} (A)
-    (SEQ3) edge [connection] node {} (B)
-           edge [connection] node {} (C)
-           edge [connection] node {} (D);
-
-  \end{tikzpicture}
-\caption{Unbalanced assembly example}
-\label{unbalanced}
-\end{center}
-\end{figure}
-
-However, in existing component models for HPC (for exemple in $L^2C$) only the component object is defined, and all components are created, destroyed, and called the same way.
-Thus, it is not possible to manage a $PAR$ component to dynamically balance the workload on threads.
-For this reason, it seems that the combination of tasks and components are needed as well as a scheduler.
-The scheduling of a task graph is an active research domain and it could be an interesting perspective to work on the combination of $L^2C$ and StarPU~\cite{}, for example.
-However, this work aims to propose a light solution using component contributions, and to illustrate the advantages and drawbacks of such a solution.
-To only use components in our solution, and to avoid unbalanced and heavy assemblies, we propose an intermediate assembly which is an approximation of the general case of the Figure~\ref{allparall}.
-We claim, in the rest of this paper, that this approximation is enough expressive and produces efficient applications for the specific case of mesh-based numerical simulations.
-
-The proposed intermediate assembly is represented in the Figure~\ref{approx}.
-The same basic association of components $SSEQ$, $SEQ$, $PAR$ and $K$ is found, however a limitation of the possible compositions of those components is applied.
-It is no longer possible to use a $SEQ$ or $SSEQ$ from a $PAR$ component, as well as using a $SSEQ$ from a $SEQ$ component and vice versa.
-The most important part of those limitations is to avoid too much imbalance.
-Actually, from a $PAR$ component it is only possible to use one computation $K$ in each created thread.
-In the specific case of mesh-based numerical simulations, as most numerical computations $c(R,w,D,e)$ (except boundary conditions) are applied on a complete set of mesh elements $D$ and computes a numerical expression $e$, two numerical computations are approximately balanced, or at least the imbalance is limited.
-
-%If solving those problems for general cases seems needed in one hand (and this point is a perspective of this work), when focusing on mesh-based numerical simulations, such an expressivity is not essential on the other hand.
-To explain this claim, the Figure~\ref{approx} represents an approximation of the assembly of the Figure~\ref{allparall}.
-In this assembly, the available compositions are limited.
-Acually, a $SSEQ$ component exclusively uses multiple $SEQ$ components, which exclusively uses multiple $PAR$ or $K$.
-Finally, a $PAR$ component exclusively uses multiple $K$.
-
-%---------
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (SSEQ) at (0,0) {SSEQ};
-   \node[component] (SEQ) [right=of SSEQ] {SEQ};
-   \node[component] (PAR) [right=of SEQ] {PAR};
-   \node[component] (K) [right=of PAR] {K};
-
-  \path[->]
-             (SSEQ)  edge  [connection]               node           {1,n}   (SEQ)
-                    edge [mpiconn] node {} (-1,0)
-             (SEQ)   edge [connection]          node       {1,n}   (PAR)
-             (SEQ)   edge [bend left=100,connection]   node  [swap]   {1,n}   (K)
-             (PAR)   edge [connection]          node       {1,n}   (K);
-\end{tikzpicture}
-\caption{General assembly description approximation}
-\label{approx}
-\end{center}
-\end{figure}
-
-Using this approximated assembly, the assembly of the example is given in the Figure~\ref{exapprox}.
-The limitation of such an approximation is clearer in this example: the dependencies $(B<D) \parallel C$ and $(B'<D') \parallel C'$ can only be expressed by $B \parallel C<D$ which is a less precised dependency.
-But as explained before, this limitation of expressivity also guarantees a limitation of the imbalance which is due to the use of components only, and not task graphs and schedulers.
-Finally, the proposed approximation still exposes the three level of parallelism introduced in the Section~\ref{}.%When introducing task parallelism scheduling, this limitation can be a big disadvantage if the task (or functionnality) $B$ is short compared to $C$.
-However, in the case of a mesh-based numerical simulation, and as explained in the Section~\ref{sect:formalism}, most numerical computation $c(R,w,D,e)$ of a stencil program $\mathcal{P}(\mathcal{M},\Delta,\Gamma,\mathcal{T})$ are computed on all elements of $E_i \subset \mathcal{M}$, and computes a single numerical expression $e$.
-As a result most computations are close in terms of amount of numerical operations.
-Thus, thanks to the specific domain we are focussing on, the approximated assembly of the Figure~\ref{approx} is satisfactory and it makes the solution lighter by the only use of the component model $L^2C$, and without the need of an external task graph tool and scheduler.
-Finally, even if this assembly is an approximation of the real dependencies assembly, the three level of parallelism are still available in the solution.
-
-%---------
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-  \node[component] (SSEQ) at (0,0) {$SSEQ$};
-  \node[component] (SEQ1) at (-2.5,-1) {$SEQ$};
-  \node[component] (SEQ2) at (2.5,-1) {$SEQ$};
-  \node[component] (A) at (-4,-2) {$A$};
-  \node[component] (PAR1) at (-3,-2) {$PAR$};
-  \node[component] (D) at (-2,-2) {$D$};
-  \node[component] (E) at (-1,-2) {$E$};
-  \node[component] (AA) at (1.5,-2) {$A'$};
-  \node[component] (PAR2) at (2.5,-2) {$PAR$};
-  \node[component] (DD) at (3.5,-2) {$D$};
-  \node[component] (B) at (-3.5,-3) {$B$};
-  \node[component] (C) at (-2.5,-3) {$C$};
-  \node[component] (BB) at (2,-3) {$B$};
-  \node[component] (CC) at (3,-3) {$C$};
-
-  \path[-]
-    (SSEQ) edge [connection] node {} (SEQ1)
-            edge [connection] node {} (SEQ2)
-            edge [mpiconn] node {} (2,0)
-    (SEQ1) edge [connection] node {} (A)
-           edge [connection] node {} (PAR1)
-           edge [connection] node {} (D)
-           edge [connection] node {} (E)
-    (SEQ2) edge [connection] node {} (AA)
-           edge [connection] node {} (PAR2)
-           edge [connection] node {} (DD)
-    (PAR1) edge [connection] node {} (B)
-           edge [connection] node {} (C)
-    (PAR2) edge [connection] node {} (BB)
-           edge [connection] node {} (CC);
-
-  \end{tikzpicture}
-\caption{Example assembly approximation}
-\label{exapprox}
-\end{center}
-\end{figure}
+%-------------------------------------
+\subsection{Transformation to a parallel assembly}
+Transformation from a parallel algorithm using the directives: PARALLEL, PARALLEL SECTIONS, SECTION and FORALL, to a component assembly

### 2015-04-16 rapport/formalism.tex
rapportv1_160415

@@
-To numerically solve a set of PDEs, iterative methods are frequently used to approximate the solution more easily by a discretization of the time and the space (which is called a mesh).
-Thus, the PDEs are transformed to a set of numerical computations applied at each time step and on all elements of the mesh.
-Among the numerical computations is found a set of numerical schemes, also called stencil computations, and a set of auxiliary numerical computations also needed to perform the simulation.
-%A stencil is a computation applied, at each time iteration, to a complete set of elements of a mesh. 
-In this paper is distinguished the concept of stencil shape (often called a stencil) from the concept of stencil computation.
-A stencil computation involves values of a simulated quantity (pressure, speed etc.) on a given element of the mesh and also on a given neighborhood of the element, denoted $\mathcal{N}$.
-A stencil shape is the shape of the needed neighborhood $\mathcal{N}$ in a stencil computation.
-As a result, a mesh-based numerical simulation is composed of a certain number of stencil computations, which is called in this paper \textit{multi-stencils}, involving one or more stencil shapes.
-In applied mathematics, an explicit numerical scheme is typically a stencil computation, but, on the other hand, an implicit numerical scheme also requires the resolution of a linear system at each time iteration.
-This last kind of scheme is above the scope of this work, however explicit schemes already represent a huge part of PDE solvers.
-In this section is defined the formalism of a \textit{stencil program}, a description of how to parallelize such programs and the associated definitions of a \textit{dependency} and a \textit{synchronization}.
-%Finally, a definition of a \textit{phase}, a \textit{group}, and a \textit{kernel}, from which numerical computations of a stencil program can be classified for any kind of mesh, and from which different levels of parallelism can be detected, are given.
+To numerically solve a set of PDEs, iterative methods are frequently used to approximate the solution by a step by step phenomena.
+Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied on a mesh (space discretization).
+In other words, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain.
+Among the numerical computations is found a set of numerical schemes, also called \textit{stencil computations}.
+A formal definition of a \textit{stencil program}, and the parallelization of such program, are presented in this section.
 
@@
 
-A stencil program or application is a program in which stencil expressions (or computations) are computed among other kind of numerical expressions, which do not involve neighborhood information.
-If the space domain is denoted $\Omega$, the time domain $\tau$, a stencil program is denoted as
-\begin{equation}
-\mathcal{P}(\mathcal{M},\Delta,\Gamma,\mathcal{T}),
-\end{equation}
-where $\mathcal{M}$ is a mesh which discretizes $\Omega$, $\Delta$ is the set of quantities and auxiliary data mapped onto $\mathcal{M}$, $\Gamma$ is the ordered set of numerical computations in the program, and $\mathcal{T}$ is the discretization of the time domain $\tau$.
-$\mathcal{M}$ can be considered as a set of different sets of elements $\{E_0,...,E_n\}$ and a connectivity between them.
-For example, $\mathcal{M}$ could be composed of $E_0$ the set of cells, $E_1$ the set of nodes and $E_2$ the set of edges, such that each cell is composed of four edges, and each edge is composed of a source and a destination node.
-%If the operator $A(B)$ means that $A$ is applied on $B$, a 
-A stencil program is composed of a set of quantities to simulate (pressure, speed etc.), mapped on the mesh, which is denoted $\Delta(E_k), 0 \leq k \leq n$.
-For all $t \in \mathcal{T}$ the set of numerical computations are applied on the data, which is simply denoted by $\Delta=\Gamma(\Delta)$.%such that for $\Delta_1, \Delta_2 \subset \Delta$, $\Delta(t) = \Gamma(\Delta_1(t-1),\Delta_2(t))$.
-In a mathematical numerical expression, a quantity to simulate has $d$ dimensions for the space domain and one additional dimension for the time, for example $\mathcal{A}(x,y,z,t)$.
-However, the implementation of a quantity in $\Delta$ to a data structure $\delta$ does not represent the time dimension in the rest of this paper.
-Thus, for a given quantity $\mathcal{A}$ at $t$ and $t-1$, eventually two different data structures (arrays) $\delta, \delta'$ are used.
-%As a result, the expression of a computation can be simplified to $\Delta=\Gamma(\Delta_1)$, which means that in a stencil program, the set of quantities $\Delta$ are computed from a given subset of quantities $\Delta_1$ on which is applied the set of computations.
+A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ of a set of PDEs.
 
-A numerical computation in $\Gamma$ is denoted 
-\begin{equation}
-c(R,w,D,e), 
-\end{equation}
-where $R \subset \Delta, w \in \Delta$ are respectively the set of data read and written during the numerical expression $e$ (a single data is written in a single computation), and $D$ is one of the subsets $E_i \subset \mathcal{M}$.
-It has to be noticed that during a numerical simulation, at each time iteration, all the elements of a mesh are computed.
+\medskip
+\begin{mydef}
+\textit{A mesh is a connected undirected graph without bridges.}
+\end{mydef}
+
+$E_i$ denotes a set of elements constructed by a function $elem_i$ which defines a precise association of nodes and edges of a mesh.
+If we denote $nodes$ the function to get nodes of a mesh and $edges$ the same for edges, the funtion $elem_i$ is defined as
+\begin{equation*}
+elem_i : nodes(\mathcal{M}) \times edges(\mathcal{M}) \rightarrow E_i
+\end{equation*}
+For example, the set of cells $E_0$ in a Cartesian 2D mesh could be defined by exactly four nodes and four edges connected as a cycle.
+But we could also define another set of elements $E_1$ as the simple set of nodes etc.
+
+A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) and hybrid if mixing those.
+
+\begin{figure}[!h]\begin{center}
+  \resizebox{8cm}{!}{\includegraphics{./images/maillages.pdf}}
+  \caption{From left to right, Cartesian, curvilinear and unstructured meshes.}
+  \label{fig:mesh}
+\end{center}\end{figure}
+
+The discretization of the continuous time domain $\mathcal{T}$ is denoted $T$. $T$ is responsible for the iteration time steps of the numerical simulation.
+
+In a numerical simulation a set of data, or quantities, are applied onto the mesh and represent the set of values to compute, or to use, for computation.
+We denote the set of data applied on the mesh by $\Delta$, such that $\delta \in \Delta$ is a function which associates each element $el \in E_i$ to a value $v \in V$,
+\begin{equation*}
+\delta : E_i \rightarrow V
+\end{equation*}
+One can notice that in applied mathematics, the signature of $\delta$ would be $\delta : E_i \times T \rightarrow V$, however when programming a numerical simulation it is not wise to store all values of each time iteration.
+
+Finally, in a numerical simulation are performed a set of ordered numerical computations denoted $\Gamma$. A numerical computation in $\Gamma$ is denoted 
+\begin{equation*}
+c(R,w,D,e). 
+\end{equation*}
+In this definition $R \subset \Delta, w \in \Delta$ are respectively the data read and written to compute the numerical expression $e$ (a single data is written in a single computation).
+$D$ is one of the subsets $E_i \subset \mathcal{M}$.
+It has to be noticed that at each time iteration, all the elements of a mesh are computed.
 However, it happens that the computation of the mesh elements is splitted in different computations (for example the computation of the physical border).
 In this case additional $E_i$ can be specified for the mesh $\mathcal{M}$.
-%For this reason, the sub-domain concept is introduced.
-Thus, i
-In a computation $c$, $\forall d \in D$, the numerical expression $e(d)$ is applied.
-If the number of computations in a stencil program $\mathcal{P}(\mathcal{M},\Delta,\Gamma,\mathcal{T})$ is $card(\Gamma)=m$, such that $\bigcup_{i=0}^{m-1}c_i = \Gamma$, then $\bigcup_{i=0}^{m-1}R_i \cup w_i = \Delta$.
-%Finally, in this work, and because of the choice to not represent the time dimension in a data $\delta \in \Delta$, a computation $c(R,w,D,e)$ always verifies $R \cap w = \emptyset$ otherwize an implicit numerical scheme has to be solve (a relation between a $\mathcal{A}(t)$ and $\mathcal{A}(t)$ itself).
+The numerical expression $e$ computes the data $w$ from a set of data $R$.
+\begin{equation*}
+e : R \rightarrow w
+\end{equation*}
 
-Two different types of computations can be handled by a stencil program: stencil computations or numerical computations which do not involve neighborhood information (as $map$ or $zip$ in functionnal languages).
-The needed neighborhood $\mathcal{N}$ of a stencil computation is the shape of the stencil.
-For regular meshes (as for example in finite difference methods), a stencil shape is not difficult to express, however, in the general purpose, the neighborhood is specific to the kind of mesh it is applied on, and could be difficult to express.
-For example, the neighborhood for unstructured meshes is composed of different concepts which are explained by the operators of Laszlo and Dobkin~\cite{}.
-This paper gives a general definition of a stencil computation, with a generic vision of a neighborhood $\mathcal{N}$.
-%independently of a specific neighborhood shape which depends on the mesh used in the simulation.
-Actually, the solution presented in this paper aims to handle any kind of mesh. A stencil computation is denoted by the general notation
+A \textit{multi-stencil program} is defined by the quadruplet
+\begin{equation}
+\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma),
+\end{equation}
+
+If the number of computations in $\Gamma$ is $card(\Gamma)=m$, such that $\bigcup_{i=0}^{m-1}c_i = \Gamma$, then $\bigcup_{i=0}^{m-1}R_i \cup w_i \subseteq \Delta$.
+
+A computation $c \in \Gamma$ can be of two different types.
+The first type is called a \textit{stencil computation} and involves an additional information denoted $\mathcal{N}$ which represents a neighborhood function around an element
+\begin{equation}
+\mathcal{N} : E_i \rightarrow E_k \times E_k \times ...
+\end{equation}
+, where $E_k$ is a set of elements of $\mathcal{M}$ which could be equal to $E_i$ or not.
+In other words, from an element $el \in E_i$ the function $\mathcal{N}$ returns a set of elements of $E_k$ which represents the neighborhood of $el$.
+The function $\mathcal{N}$ is also sometimes called the \textit{stencil shape}, or the \textit{stencil} in applied mathematics.
+
+A \textit{stencil computation} is a quintuplet
 \begin{equation}
@@
 \end{equation}
-where $R \subset \Delta, w \in \Delta$ are read and written in the numerical expression $e$, which uses a given neighborhood function denoted $\mathcal{N}$ on a sub-domain $D$.
-In a stencil computation $s$, $\forall d \in D$, the numerical expression is applied such that $w(d) = e(R(d),R(\mathcal{N}(d))$.
-Finally, in this work, a stencil computation $s(R,w,D,e,\mathcal{N})$ always verifies $R \cap w = \emptyset$, otherwize an implicit numerical scheme has to be solve (a relation between a a quantity $\mathcal{A}(t)$ and $\mathcal{A}(t)$ itself at the same time iteration).
+where $R \subset \Delta, w \in \Delta$ are read and written in the numerical expression $e$, and $D$ is one the subsets $E_i$.
+In a stencil computation $s$, $\forall d \in D$, the stencil numerical expression $e$ is applied such that $w(d) = e(R(d),R(\mathcal{N}(d))$.
+In this work, a stencil computation $s(R,w,D,e,\mathcal{N})$ always verifies $R \cap w = \emptyset$, otherwize an implicit numerical scheme has to be solve which is over the scope of this paper.
 
-The second kind of numerical computation is denoted
+Figure~\ref{} gives an example of a stencil computation $s(R,w,D,e,\mathcal{N})$, on a two dimensional Cartesian mesh, where $R=\{A,B\}$, $w=C$, and where for $(x,y) \in D$
+\begin{equation*}
+e(R(x,y),R(\mathcal{N}(x,y)) = A(x,y)+B(x,y+1)+B(x,y-1)+B(x+1,y)+B(x-1,y).
+\end{equation*}
+
+Finally, the second type of numerical computation is a local computation where $e$ does not involve a neighborhood function $\mathcal{N}$
 \begin{equation}
-l(R,w,D,e).
+l(R,w,D,e)
 \end{equation}
-This is a local numerical expression, where $e$ does not involve $\mathcal{N}(d)$, and where $e$ is applied such that $w(d) = e(R(d))$.
 
 %-------------------------------------
-\subsection{Stencil program parallelization}
+\subsection{Parallelization}
 \label{sect:parall}
-Mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
+% Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
 
-Maybe the more important parallelization technique of such simulations is called \textit{SPMD} (Single Program, Multiple Data) which can also be called \textit{coarse-grain data parallelism}.
+% Maybe the more important parallelization technique of such simulations is called \textit{SPMD} (Single Program, Multiple Data) which can also be called \textit{coarse-grain data parallelism}.
 This concept represents the well-known and broadly-used domain decomposition technique to split the mesh in balanced sub-meshes for available resources.
@@
 
-However, because of energy consumption constraints, most of the \textit{top500}~\footnote{http://www.top500.org} supercomputer centers are not only composed of clusters of multi-core machines, but also of accelerator cards such as GPGPU and many-cores.
+% However, because of energy consumption constraints, most of the \textit{top500}~\footnote{http://www.top500.org} supercomputer centers are not only composed of clusters of multi-core machines, but also of accelerator cards such as GPGPU and many-cores.
 In addition to this, vectorization units are also available on most CPUs.
@@
 
-Finally, the last parallelization technique which can be applied on various architectures (clusters, multi-cores, many-cores, GPGPU), is the \textit{task parallelism} technique.
+% Finally, the last parallelization technique which can be applied on various architectures (clusters, multi-cores, many-cores, GPGPU), is the \textit{task parallelism} technique.
 Instead of spliting the data of the simulation (or the vector used in a numerical expression), the different tasks of the simulation are identified and scheduled over the available resources during the execution (at runtime)~\cite{}.
@@
 
-This section discussed the various parallelization techniques which can be used for numerical simulations.
+% This section discussed the various parallelization techniques which can be used for numerical simulations.
 As a result, it seems that, by a combination of existing techniques, modern heterogeneous parallel architectures can be fully harnessed by such applications~\cite{}.
@@
 
-%-------------------------------------
-\subsection{Dependencies and synchronizations}
-\label{sect:dep}
-To define and explain the component-based implicit parallelism model presented in this paper, two additional definitions handle the relations between data read and written into computations.
+% %-------------------------------------
+% \subsection{Dependencies and synchronizations}
+% \label{sect:dep}
+% To define and explain the component-based implicit parallelism model presented in this paper, two additional definitions handle the relations between data read and written into computations.
 Those relations are responsible for dependencies and synchronizations.
 
-\medskip
-\begin{mydef}
-\textit{For two computations $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, it is said that $c_2$ is dependant from $c_1$, denoted $c_1<c_2$, if $w_1 \cap R_2 \neq \emptyset$.
+% \medskip
+% \begin{mydef}
+% \textit{For two computations $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, it is said that $c_2$ is dependant from $c_1$, denoted $c_1<c_2$, if $w_1 \cap R_2 \neq \emptyset$.
 In this case, $c_1$ has to be computed before $c_2$.
 The binary relation $c_1<c_2$ represents a \textit{dependency}.}
-\end{mydef}
+% \end{mydef}
 
-\medskip
-Dependencies lead to one computation before another, but when the program is parallelized some dependencies are also responsible for synchronizations between different processes.
+% \medskip
+% Dependencies lead to one computation before another, but when the program is parallelized some dependencies are also responsible for synchronizations between different processes.
 The opposite definition can also be given,
 
-\medskip
-\begin{mydef}
-\textit{For two computations $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, it is said that $c_2$ is independant from $c_1$, denoted $c_1 \not <c_2$ or $c_1 \parallel c_2$, if $w_1 \cap R_2 = \emptyset$.}
-\end{mydef}
+% \medskip
+% \begin{mydef}
+% \textit{For two computations $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, it is said that $c_2$ is independant from $c_1$, denoted $c_1 \not <c_2$ or $c_1 \parallel c_2$, if $w_1 \cap R_2 = \emptyset$.}
+% \end{mydef}
 
-Considering a coarse-grain data parallelized stencil program $\mathcal{P}$,
+% Considering a coarse-grain data parallelized multi-stencil program $\mathcal{MSP}$,
 
-\medskip
-\begin{mydef}
-\textit{A depedency between two computations $c_1<c_2$ is a \textit{synchronisation}, denoted $c_1 \ll c_2$, if $c_2=s_2(R_2,w_2,D_2,e_2)$, and $w_1 \cap R_2 \neq \emptyset$.}
-\end{mydef}
+% \medskip
+% \begin{mydef}
+% \textit{A depedency between two computations $c_1<c_2$ is a \textit{synchronisation}, denoted $c_1 \ll c_2$, if $c_2=s_2(R_2,w_2,D_2,e_2)$, and $w_1 \cap R_2 \neq \emptyset$.}
+% \end{mydef}
 
-\medskip
-Actually, in a parallel stencil program, synchronizations between processes are always needed for the set of data $R$ of a stencil $s(R,w,D,e)$ because the expression $e$ involves a neighborhood $\mathcal{N}$ which could be handled by another processor or resource.
+% \medskip
+% Actually, in a parallel stencil program, synchronizations between processes are always needed for the set of data $R$ of a stencil $s(R,w,D,e)$ because the expression $e$ involves a neighborhood $\mathcal{N}$ which could be handled by another processor or resource.

### 2015-04-16 rapport/formalism.tex
rapportv1_160415

@@
 \label{sect:parall}
-% Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
+Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
 
-% Maybe the more important parallelization technique of such simulations is called \textit{SPMD} (Single Program, Multiple Data) which can also be called \textit{coarse-grain data parallelism}.
-This concept represents the well-known and broadly-used domain decomposition technique to split the mesh in balanced sub-meshes for available resources.
-Because of the good locality degree of numerical simulations, only a small number of additional elements, called a \textit{recorvery} or a \textit{ghost} area, are needed to exchange information computed by another resource to correctly compute stencils.
-As the amount of communications of such parallelization are rather small compared to the amount of computations (small neighborhood), and also by the addition of recovery of communications by computations, this parallelization technique has proved many times its efficiency and scalability on numerical simulations~\cite{}.
-This parallelization technique can be applied on shared and distributed memory architectures (clusters, multi-core CPUs), and it stays the most efficient parallelization technique if the numerical simulation has to be executed on a distributed memory architecture.
-For this reason, this parallelization stays a reference in the domain and is still studied for more complex simulations, using unstructured meshes, multi-grids techniques etc.
+The \textit{SPMD} (Single Program, Multiple Data) parallelization technique which is called \textit{coarse-grain data parallelism} in the rest of this paper, is broadly used because of its scalability.
+This parallelization concept is the well-known domain decomposition technique where the mesh is splitted in balanced sub-meshes for available resources, and the same program is applied on each sub-part of the initial mesh.
+Because of the good locality degree of numerical simulations, only a small number of additional elements, called a \textit{recorvery} or a \textit{ghost} area, are needed to get information computed on another resource.
+This ghost area is however unavoidable to correctly compute stencils because of the neighborhood $\mathcal{N}$.
+As the amount of data to synchronize between resources is rather small compared to the amount of computation (small neighborhood), and also by the addition of recovery of synchronizations by computations, this parallelization technique has proved many times its efficiency and scalability on numerical simulations~\cite{}.
+This parallelization technique can be applied on shared and distributed memory architectures (clusters, multi-core CPUs), and it is one of the most efficient parallelization technique if the numerical simulation has to be executed on a distributed memory architecture such as a cluster.
+Figure~\ref{} illustrates this parallelization technique.
+
+More formally, in a SPMD parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $n$ balanced sub-meshes (for $n$ resources) $\{\mathcal{M}_0,\dots,\mathcal{M}_{n-1}\}$.
+This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
+The second step of the parallelization is to identify in the ordered list of computation $\Gamma$ when synchronizations are needed between resources.
+
+\medskip
+\begin{mydef}
+\textit{A \textit{synchronisation} is needed between two computations $c_1$ and $c_2$, denoted $c_1 \ll c_2$, if $c_2=s_2(R_2,w_2,D_2,e_2)$, and $w_1 \cap R_2 \neq \emptyset$.}
+\end{mydef}
+
+Actually, a synchronization can only be needed by the data read in a stencil computation, and only if this data has been modified before, which means that it has been written before.
+A new ordered list of computations $\Gamma'$ is obtained from the insertion of the synchronizations at the good places.
+The final step of this parallelization is to run the same ordered list of computation $\Gamma'$ on each resource.
+For each resource $0 \leq k \leq n-1$ a multi-stencil program defined by
+\begin{equation}
+\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma'),
+\end{equation}
+is runned.
+
+PARALLEL,SYNC definitions + algorithm
 

### 2015-04-20 rapport/formalism.tex
rapport_200415

--- 2015-04-16 rapport/formalism.tex
@@
 
-A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ of a set of PDEs.
+A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ of a set of PDEs and is defined as followed.
 
-\medskip
 \begin{mydef}
-\textit{A mesh is a connected undirected graph without bridges.}
+\textit{A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V$ is the set of vertices and $E$ the set of edges.
+The set of edges $E$ of a mesh $\mathcal{M}=(V,E)$ does not contain bridges.}
 \end{mydef}
 
-$E_i$ denotes a set of elements constructed by a function $elem_i$ which defines a precise association of nodes and edges of a mesh.
-If we denote $nodes$ the function to get nodes of a mesh and $edges$ the same for edges, the funtion $elem_i$ is defined as
-\begin{equation*}
-elem_i : nodes(\mathcal{M}) \times edges(\mathcal{M}) \rightarrow E_i
-\end{equation*}
-For example, the set of cells $E_0$ in a Cartesian 2D mesh could be defined by exactly four nodes and four edges connected as a cycle.
-But we could also define another set of elements $E_1$ as the simple set of nodes etc.
+\begin{mydef}
+$D_i$ is a set of elements of a mesh $\mathcal{M}=(V,E)$, constructed by a function $domain_i$ which defines a precise association between $V$ and $E$, $domain_i : V \times E \rightarrow D_i$.
+\end{mydef}
+For example, the set of cells $D_0$ in a Cartesian 2D mesh could be defined by exactly four vertices and four edges connected as a cycle.
+But we could also define another set of elements $D_1$ as the simple set of vertices $V$ etc.
 
-A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) and hybrid if mixing those.
+A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) and hybrid.
 
@@
 
-The discretization of the continuous time domain $\mathcal{T}$ is denoted $T$. $T$ is responsible for the iteration time steps of the numerical simulation.
+\begin{mydef}
+The discretization of the continuous time domain $\mathcal{T}$ is denoted $T$ such that $\forall t_i, t_{i+1} \in T, \exists \Delta t \in \mathbb{R}$, $t_{i+1} = t_i + \Delta t$.
+Thus, $T$ is responsible for the iteration time steps of the numerical simulation.
+\end{mydef}
 
 In a numerical simulation a set of data, or quantities, are applied onto the mesh and represent the set of values to compute, or to use, for computation.
-We denote the set of data applied on the mesh by $\Delta$, such that $\delta \in \Delta$ is a function which associates each element $el \in E_i$ to a value $v \in V$,
-\begin{equation*}
-\delta : E_i \rightarrow V
-\end{equation*}
+
+\begin{mydef}
+The set of data applied on the mesh is denoted by $\Delta$, such that $\delta \in \Delta$ is a function which associates each element $d \in D_i$ to a value $v \in V$, $\delta : D_i \rightarrow V$.
+\end{mydef}
 One can notice that in applied mathematics, the signature of $\delta$ would be $\delta : E_i \times T \rightarrow V$, however when programming a numerical simulation it is not wise to store all values of each time iteration.
 
-Finally, in a numerical simulation are performed a set of ordered numerical computations denoted $\Gamma$. A numerical computation in $\Gamma$ is denoted 
-\begin{equation*}
-c(R,w,D,e). 
-\end{equation*}
-In this definition $R \subset \Delta, w \in \Delta$ are respectively the data read and written to compute the numerical expression $e$ (a single data is written in a single computation).
-$D$ is one of the subsets $E_i \subset \mathcal{M}$.
+\begin{mydef}
+A numerical expression $\text{exp}$ is a function which represents how to compute, for an element $d \in D_i$, a data $w \in \Delta$ (written data) with a set $R \subset \Delta$ of input data (read data), $\text{exp} : R \times D_i \rightarrow w \times D_i$.
+\end{mydef}
+
+\begin{mydef}
+A computation $c$ of a numerical simulation is defined as $c(R,w,D,\text{exp})$, where $R \subset \Delta, w \in \Delta$ and $\text{exp}$ a numerical expression $\text{exp} : R \rightarrow w$.
+$D$ is one of the subsets $D_i \subset \mathcal{M}$, such that $w : D \rightarrow V$.
+\end{mydef}
 It has to be noticed that at each time iteration, all the elements of a mesh are computed.
 However, it happens that the computation of the mesh elements is splitted in different computations (for example the computation of the physical border).
-In this case additional $E_i$ can be specified for the mesh $\mathcal{M}$.
-The numerical expression $e$ computes the data $w$ from a set of data $R$.
+In this case additional $D_i$ can be specified for the mesh $\mathcal{M}$.
+
+\begin{mydef}
+The set of $n$ ordered computations of a numerical simulation is denoted $\Gamma = [c_i]_{0 \leq i \leq n-1}$, such that $\forall c_i,c_j$ with $i \leq j$, $c_i$ is computed before $c_j$, and $c_j$ can be computed only when $c_i$ is finished.
+\end{mydef}
+
+\begin{mydef}
+Finally, a \textit{multi-stencil program} is defined by the quadruplet $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$.
+\end{mydef}
+%If the number of computations in $\Gamma$ is $card(\Gamma)=n$, such that $\bigcup_{i=0}^{n-1}c_i = \Gamma$, then $\bigcup_{i=0}^{m-1}R_i \cup w_i \subseteq \Delta$.
+
+\begin{mydef}
+The neighborhood $\mathcal{N}$ of an element $d \in D_i$ is a function to define a set of elements in any $D_k \subset \mathcal{M}$, $\mathcal{N} : D_i \rightarrow D_k \times D_k \times \dots$.
+\end{mydef}
+The function $\mathcal{N}$ is also sometimes called the \textit{stencil shape}, or the \textit{stencil} in applied mathematics.
+
+A computation $c \in \Gamma$ can be of two different types. The first type is called a \textit{stencil computation}.
+
+\begin{mydef}
+A \textit{stencil computation} is defined as a quintuplet $s(R,w,D,\text{exp},\mathcal{N})$, where $R \subset \Delta, w \in \Delta$, $D$ is one the subsets $D_i$ and $w : D \rightarrow V$.
+\end{mydef}
+In a stencil computation $s$, $\forall d \in D$, the stencil numerical expression $\text{exp}$ is applied such that $w(d) = \text{exp}(R(d),R(\mathcal{N}(d))$.
+In this work, a stencil computation $s(R,w,D,\text{exp},\mathcal{N})$ always verifies $R \cap w = \emptyset$, otherwize an implicit numerical scheme has to be solve which is over the scope of this paper.
+
+Figure~\ref{} gives an example of a stencil computation $s(R,w,D,\text{exp},\mathcal{N})$, where $\mathcal{M}(V,E)$ is a two dimensional Cartesian mesh.
+A single domain $D$ is defined in this example and is composed of cells formed by a cycle of four vertices $v \in V$ and four edges $e \in E$.
+Furthermore, in this example $R=\{A\}$, $w=B$, and for $(x,y) \in D$ the neighborhood function is
 \begin{equation*}
-e : R \rightarrow w
+\mathcal{N} : (x,y) \rightarrow \{(x,y+1),(x,y-1),(x+1,y),(x-1,y)\}.
+\end{equation*}
+Finally, the numerical expression of this example is 
+\begin{equation*}
+\text{exp}(A(x,y),A(\mathcal{N}(x,y)) = B(x,y) = A(x,y)+(A(x,y+1)+A(x,y-1)+A(x+1,y)+A(x-1,y))/4.
 \end{equation*}
 
-A \textit{multi-stencil program} is defined by the quadruplet
-\begin{equation}
-\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma),
-\end{equation}
-
-If the number of computations in $\Gamma$ is $card(\Gamma)=m$, such that $\bigcup_{i=0}^{m-1}c_i = \Gamma$, then $\bigcup_{i=0}^{m-1}R_i \cup w_i \subseteq \Delta$.
-
-A computation $c \in \Gamma$ can be of two different types.
-The first type is called a \textit{stencil computation} and involves an additional information denoted $\mathcal{N}$ which represents a neighborhood function around an element
-\begin{equation}
-\mathcal{N} : E_i \rightarrow E_k \times E_k \times ...
-\end{equation}
-, where $E_k$ is a set of elements of $\mathcal{M}$ which could be equal to $E_i$ or not.
-In other words, from an element $el \in E_i$ the function $\mathcal{N}$ returns a set of elements of $E_k$ which represents the neighborhood of $el$.
-The function $\mathcal{N}$ is also sometimes called the \textit{stencil shape}, or the \textit{stencil} in applied mathematics.
-
-A \textit{stencil computation} is a quintuplet
-\begin{equation}
-s(R,w,D,e,\mathcal{N}),
-\end{equation}
-where $R \subset \Delta, w \in \Delta$ are read and written in the numerical expression $e$, and $D$ is one the subsets $E_i$.
-In a stencil computation $s$, $\forall d \in D$, the stencil numerical expression $e$ is applied such that $w(d) = e(R(d),R(\mathcal{N}(d))$.
-In this work, a stencil computation $s(R,w,D,e,\mathcal{N})$ always verifies $R \cap w = \emptyset$, otherwize an implicit numerical scheme has to be solve which is over the scope of this paper.
-
-Figure~\ref{} gives an example of a stencil computation $s(R,w,D,e,\mathcal{N})$, on a two dimensional Cartesian mesh, where $R=\{A,B\}$, $w=C$, and where for $(x,y) \in D$
-\begin{equation*}
-e(R(x,y),R(\mathcal{N}(x,y)) = A(x,y)+B(x,y+1)+B(x,y-1)+B(x+1,y)+B(x-1,y).
-\end{equation*}
-
-Finally, the second type of numerical computation is a local computation where $e$ does not involve a neighborhood function $\mathcal{N}$
-\begin{equation}
-l(R,w,D,e)
-\end{equation}
+Finally, the second type of numerical computation is a local computation.
+\begin{mydef}
+A local computation is a quadruplet $l(R,w,D,\text{exp})$, where $e$ does not involve a neighborhood function $\mathcal{N}$.
+\end{mydef}
 
@@
 \label{sect:parall}
-% Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
+Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
 
-% Maybe the more important parallelization technique of such simulations is called \textit{SPMD} (Single Program, Multiple Data) which can also be called \textit{coarse-grain data parallelism}.
-This concept represents the well-known and broadly-used domain decomposition technique to split the mesh in balanced sub-meshes for available resources.
-Because of the good locality degree of numerical simulations, only a small number of additional elements, called a \textit{recorvery} or a \textit{ghost} area, are needed to exchange information computed by another resource to correctly compute stencils.
-As the amount of communications of such parallelization are rather small compared to the amount of computations (small neighborhood), and also by the addition of recovery of communications by computations, this parallelization technique has proved many times its efficiency and scalability on numerical simulations~\cite{}.
-This parallelization technique can be applied on shared and distributed memory architectures (clusters, multi-core CPUs), and it stays the most efficient parallelization technique if the numerical simulation has to be executed on a distributed memory architecture.
-For this reason, this parallelization stays a reference in the domain and is still studied for more complex simulations, using unstructured meshes, multi-grids techniques etc.
+\paragraph{Coarse-grain data parallelism.} In a data parallelization technique, the idea is to split the data on which the program is computed in balanced sub-parts, one for each available resource.
+The same sequential program can afterwards be applied on each sub-part simultaneously, with some additioinal synchronizations between resources to update the data not computed locally and thus to guarantee a correct result.
 
-% However, because of energy consumption constraints, most of the \textit{top500}~\footnote{http://www.top500.org} supercomputer centers are not only composed of clusters of multi-core machines, but also of accelerator cards such as GPGPU and many-cores.
-In addition to this, vectorization units are also available on most CPUs.
-This introduces another kind of parallelization, linked to the first one, called in this paper \textit{fine-grain data parallelism}, which is composed of \textit{SIMD} (Single Instruction, Multiple Data) and \textit{SIMT} (Single Instruction, Multiple Threads) parallelization techniques.
-Such techniques typically compute a numerical expression in parallel on a vector, using vectorization units or threads, which perfectly matches the type of computation in the nested loops of numerical simulations.
+More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\{\mathcal{M}_0,\dots,\mathcal{M}_{n-1}\}$.
+This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
+As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\{\Delta_0,\dots,\Delta_{n-1}\}$.
+The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma'$.
 
-% Finally, the last parallelization technique which can be applied on various architectures (clusters, multi-cores, many-cores, GPGPU), is the \textit{task parallelism} technique.
-Instead of spliting the data of the simulation (or the vector used in a numerical expression), the different tasks of the simulation are identified and scheduled over the available resources during the execution (at runtime)~\cite{}.
-As numerical computations of a simulation can be considered as tasks, linked together through data dependencies for example, this parallelization technique can be used for HPC numerical simulations too.
+\begin{mydef}
+For $n$ the number of computations in $\Gamma$, and $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \ll c_j$, if $c_j=s_j(R_j,w_j,D_j,\text{exp}_j)$, and $w_i \cap R_j \neq \emptyset$.
+Moreover, the data to update is $w_i \cap R_j$.
+\end{mydef}
 
-% This section discussed the various parallelization techniques which can be used for numerical simulations.
-As a result, it seems that, by a combination of existing techniques, modern heterogeneous parallel architectures can be fully harnessed by such applications~\cite{}.
-In this paper is presented a component-based model to expose those three kinds of parallelism in the overall simulation, while hiding them from the user (numericians).
+Actually, a synchronization can only be needed by the data read in a stencil computation, and only if this data has been modified before, which means that it has been written before.
+This synchronization is needed because the neighborhood function $\mathcal{N}$ of the stencil computation involves values computed on different resources.
 
-% %-------------------------------------
-% \subsection{Dependencies and synchronizations}
-% \label{sect:dep}
-% To define and explain the component-based implicit parallelism model presented in this paper, two additional definitions handle the relations between data read and written into computations.
-Those relations are responsible for dependencies and synchronizations.
+\begin{mydef}
+A synchronization between two computations $c_i \ll c_j$ is defined as a specific computation $\text{update}(w_i \cap R_j)$.
+\end{mydef}
 
-% \medskip
-% \begin{mydef}
-% \textit{For two computations $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, it is said that $c_2$ is dependant from $c_1$, denoted $c_1<c_2$, if $w_1 \cap R_2 \neq \emptyset$.
-In this case, $c_1$ has to be computed before $c_2$.
-The binary relation $c_1<c_2$ represents a \textit{dependency}.}
-% \end{mydef}
+We denote an function to insert at a precise position of a given ordered list, a computation $c_{up}$, as
+\begin{equation}
+\text{insert} : [c_k]_{0 \leq k \leq n-1} \times \text{position} \times c_{up} \rightarrow [c_0,\dots,c_{\text{position}-1},c_{up},\dots,c_{n-1}]
+\end{equation}
 
-% \medskip
-% Dependencies lead to one computation before another, but when the program is parallelized some dependencies are also responsible for synchronizations between different processes.
-The opposite definition can also be given,
+The new ordered list of computations $\Gamma'$ is obtained from the insertion of the synchronizations at the good position. Algorithm~\ref{alg:gammaprime} illustrates how $\Gamma'$ is build.
+\begin{center}
+\begin{algorithm}
+$\Gamma' = \Gamma$\\
+$\text{offset} = 0$\\
+\For{$c_j \in \Gamma$}{
+	\uIf{$c_j$ is a stencil}{
+	\uIf{$\Gamma[c_0:c_j]==[]$}{
+	$\text{insert}(\Gamma',0,\text{update}(R_j)$\\
+	$\text{offset}++$\\
+	}
+	\uElse{
+	\For{$c_i \in \Gamma[c_0:c_j]$}{
+	\uIf{$c_i \ll c_j$}{
+	$\text{insert}(\Gamma',j+\text{offset},\text{update}(w_i \cap R_j)$\\
+	$\text{offset}++$
+	}
+	}
+	}
+	}
+	}
+\caption{Creation of $\Gamma'$ from $\Gamma$}
+\label{alg:gammaprime}
+\end{algorithm}
+\end{center}
 
-% \medskip
-% \begin{mydef}
-% \textit{For two computations $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, it is said that $c_2$ is independant from $c_1$, denoted $c_1 \not <c_2$ or $c_1 \parallel c_2$, if $w_1 \cap R_2 = \emptyset$.}
-% \end{mydef}
+ The final step of this parallelization is to run $\Gamma'$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program defined by
+\begin{equation}
+\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma'),
+\end{equation}
+is runned.
 
-% Considering a coarse-grain data parallelized multi-stencil program $\mathcal{MSP}$,
+We denote this parallelization technique a coarse-grain data parallelization in contrast with the same technique applied to the finer level of a single computation.
+In this case, for a computation $c(R,w,D,\text{exp})$, the local domain $D$ is divided for $p$ resources $\{D_0,\dots,D_{p-1}\}$, and each resource $k$ is responsible for a sub-computation $c_k(R,w,D_k,\text{exp})$.
+This finer data parallelization does not require the addition of synchronizations because each computation .
+This finer data parallelization technique is not directly adressed by the work presented in this paper as it will be explained later.
 
-% \medskip
-% \begin{mydef}
-% \textit{A depedency between two computations $c_1<c_2$ is a \textit{synchronisation}, denoted $c_1 \ll c_2$, if $c_2=s_2(R_2,w_2,D_2,e_2)$, and $w_1 \cap R_2 \neq \emptyset$.}
-% \end{mydef}
-
-% \medskip
-% Actually, in a parallel stencil program, synchronizations between processes are always needed for the set of data $R$ of a stencil $s(R,w,D,e)$ because the expression $e$ involves a neighborhood $\mathcal{N}$ which could be handled by another processor or resource.
+\paragraph{Task parallelism.}

### 2015-04-20 rapport/formalism.tex
rapport_200415

@@
 
-Figure~\ref{} gives an example of a stencil computation $s(R,w,D,\text{exp},\mathcal{N})$, where $\mathcal{M}(V,E)$ is a two dimensional Cartesian mesh.
+Figure~\ref{fig:ex} gives an example of a stencil computation $s(R,w,D,\text{exp},\mathcal{N})$, where $\mathcal{M}(V,E)$ is a two dimensional Cartesian mesh.
 A single domain $D$ is defined in this example and is composed of cells formed by a cycle of four vertices $v \in V$ and four edges $e \in E$.
@@
 \end{equation*}
+
+\begin{figure}[!h]\begin{center}
+  \resizebox{8cm}{!}{\includegraphics{./images/example.pdf}}
+  \caption{Example of a stencil computation.}
+  \label{fig:ex}
+\end{center}\end{figure}
 

### 2015-04-30 rapport/formalism.tex
rapport_300415

@@
 \paragraph{Coarse-grain data parallelism.} In a data parallelization technique, the idea is to split the data on which the program is computed in balanced sub-parts, one for each available resource.
-The same sequential program can afterwards be applied on each sub-part simultaneously, with some additioinal synchronizations between resources to update the data not computed locally and thus to guarantee a correct result.
+The same sequential program can afterwards be applied on each sub-part simultaneously, with some additioinal synchronizations between resources to update the data not computed locally, and thus to guarantee a correct result.
 
@@
 As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\{\Delta_0,\dots,\Delta_{n-1}\}$.
-The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma'$.
+The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma_{data}$.
 
 \begin{mydef}
-For $n$ the number of computations in $\Gamma$, and $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \ll c_j$, if $c_j=s_j(R_j,w_j,D_j,\text{exp}_j)$, and $w_i \cap R_j \neq \emptyset$.
-Moreover, the data to update is $w_i \cap R_j$.
+For $n$ the number of computations in $\Gamma$, and $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j=s_j(R_j,w_j,D_j,\text{exp}_j)$, and $w_i \subset R_j$.
+Moreover, the data to update is $w_i \cap R_j = w_i$.
 \end{mydef}
@@
 \begin{mydef}
-A synchronization between two computations $c_i \ll c_j$ is defined as a specific computation $\text{update}(w_i \cap R_j)$.
+A synchronization between two computations $c_i \pprec c_j$ is defined as a specific computation $\text{update}(w_i \cap R_j)=\text{update}(w_i)$.
 \end{mydef}
 
-We denote an function to insert at a precise position of a given ordered list, a computation $c_{up}$, as
+\begin{mydef}
+The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[c_i]_{0 \leq i \leq n-1}$ and $l_2=[w_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[c_0,\dots,c_{n-1},w_0,\dots,w_{m-1}]$.
+\end{mydef}
+
+\begin{mydef}
+From the ordered list of computation $\Gamma$, a new ordered list $\Gamma_{data}$ is obtained from the call $\Gamma_{data} = T_{data}(\Gamma,0)$, where $T_{data}$ is the recursive function defined as
+\begin{equation*}
+T_{data}(\Gamma,i) = 
+\begin{cases} 	[\Gamma[i]] \cdot T_{data}(\Gamma,i+1) & \mbox{if }\forall j\leq i\mbox{, } c_j \not\pprec c_i\\
+				[\text{update}(\{w_k\}),\Gamma[i]] \cdot T_{data}(\Gamma,i+1) & \forall k \leq i\mbox{, }c_k \pprec c_i\\
+				[] & \mbox{if }i=|\Gamma|.
+\end{cases}
+\end{equation*}
+\end{mydef}
+
+The construction of $\Gamma_{data}$ implies that if a synchronization is needed it is always proceeded before the computation wich needs it.
+However, an optimization is possible to combine synchronizations together, and thus to synchronize more than one data at a time.
+If the number of synchronizations is reduced, final performance of the program will be better.
+For this reason, we define a reduction to apply to $\Gamma_{data}$, and which is performed as a post-transformation.
+
+\begin{myprop}
+Denoting $c_k$ and $c_j$, with $k<j$, two computations of the ordered list $\Gamma_{data}$ for which a synchronization is needed.
+If $c_i$, with $i<j$ is the computation such that $c_i \pprec c_j$ and such that $\text{update}(w_i)$ is performed before $c_j$ in $\Gamma_{data}$.
+And if the needed synchronization before $k$ is $\text{update}(w_l)$, where $l<k$.
+Then, if $i<k$, a single synchronization phase is needed before $c_k$ and is $\text{update}(w_l\cup w_i)$.
+\end{myprop}
+
+\begin{proof}
+By definition, the synchronization of $w_i$ can be performed between $c_i$ and $c_j$.
+As $i<k$ the computation $c_i$ is not performed between $k$ and $j$.
+As a result, the synchronization of $w_i$ can be performed between $c_i$ and $c_k$.
+\end{proof}
+
+In other words, for each synchronization in $\Gamma_{data}$, if the computation $c_i$ from which $c_j$ is synchronized occurs before the last synchronization, the two synchronizations can be reduced to one with the union of the data to update.
+
+% We denote an function to insert at a precise position of a given ordered list, a computation $c_{up}$, as
+% \begin{equation}
+% \text{insert} : [c_k]_{0 \leq k \leq n-1} \times \text{position} \times c_{up} \rightarrow [c_0,\dots,c_{\text{position}-1},c_{up},\dots,c_{n-1}]
+% \end{equation}
+
+% The new ordered list of computations $\Gamma'$ is obtained from the insertion of the synchronizations at the good position. Algorithm~\ref{alg:gammaprime} illustrates how $\Gamma'$ is build.
+% \begin{center}
+% \begin{algorithm}
+% $\Gamma' = \Gamma$\\
+% $\text{offset} = 0$\\
+% \For{$c_j \in \Gamma$}{
+% 	\uIf{$c_j$ is a stencil}{
+% 	\uIf{$\Gamma[c_0:c_j]==[]$}{
+% 	$\text{insert}(\Gamma',0,\text{update}(R_j)$\\
+% 	$\text{offset}++$\\
+% 	}
+% 	\uElse{
+% 	\For{$c_i \in \Gamma[c_0:c_j]$}{
+% 	\uIf{$c_i \pprec c_j$}{
+% 	$\text{insert}(\Gamma',j+\text{offset},\text{update}(w_i \cap R_j)$\\
+% 	$\text{offset}++$
+% 	}
+% 	}
+% 	}
+% 	}
+% 	}
+% \caption{Creation of $\Gamma'$ from $\Gamma$}
+% \label{alg:gammaprime}
+% \end{algorithm}
+% \end{center}
+
+ The final step of this parallelization is to run $\Gamma_{data}$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program defined by
 \begin{equation}
-\text{insert} : [c_k]_{0 \leq k \leq n-1} \times \text{position} \times c_{up} \rightarrow [c_0,\dots,c_{\text{position}-1},c_{up},\dots,c_{n-1}]
-\end{equation}
-
-The new ordered list of computations $\Gamma'$ is obtained from the insertion of the synchronizations at the good position. Algorithm~\ref{alg:gammaprime} illustrates how $\Gamma'$ is build.
-\begin{center}
-\begin{algorithm}
-$\Gamma' = \Gamma$\\
-$\text{offset} = 0$\\
-\For{$c_j \in \Gamma$}{
-	\uIf{$c_j$ is a stencil}{
-	\uIf{$\Gamma[c_0:c_j]==[]$}{
-	$\text{insert}(\Gamma',0,\text{update}(R_j)$\\
-	$\text{offset}++$\\
-	}
-	\uElse{
-	\For{$c_i \in \Gamma[c_0:c_j]$}{
-	\uIf{$c_i \ll c_j$}{
-	$\text{insert}(\Gamma',j+\text{offset},\text{update}(w_i \cap R_j)$\\
-	$\text{offset}++$
-	}
-	}
-	}
-	}
-	}
-\caption{Creation of $\Gamma'$ from $\Gamma$}
-\label{alg:gammaprime}
-\end{algorithm}
-\end{center}
-
- The final step of this parallelization is to run $\Gamma'$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program defined by
-\begin{equation}
-\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma'),
+\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data}),
 \end{equation}
@@
 We denote this parallelization technique a coarse-grain data parallelization in contrast with the same technique applied to the finer level of a single computation.
-In this case, for a computation $c(R,w,D,\text{exp})$, the local domain $D$ is divided for $p$ resources $\{D_0,\dots,D_{p-1}\}$, and each resource $k$ is responsible for a sub-computation $c_k(R,w,D_k,\text{exp})$.
-This finer data parallelization does not require the addition of synchronizations because each computation .
-This finer data parallelization technique is not directly adressed by the work presented in this paper as it will be explained later.
+In this case, for a computation $c(R,w,D,\text{exp})$, the local domain $D$ is partitionned for $p$ resources $\{D_0,\dots,D_{p-1}\}$, and each resource $k$ is responsible for a sub-computation $c_k(R,w,D_k,\text{exp}')$.
+%This finer data parallelization is particularly adapted to shared memory architectures (multi-cores, many-cores or GPUs), and in such case it does not require the addition of synchronizations because the computed element is local to the sub-domain $D_k$.
+Actually, if the computation is a stencil, the computation only read neighborhood elements but does not write them which does not require a synchronization, and if the computation is local nothing is read or written in the sub-domain of another resource.
+This finer data parallelization technique is not directly adressed by the work presented in this paper, but is exhibited, as it will be explained later.
 
 \paragraph{Task parallelism.}
+A task parallelization technique is a technique to transform a program as a dependency graph of different tasks.
+A dependency graph exhibits parallel tasks, or on the contrary sequential execution of tasks.
+In a multi-stencil program a task is defined as a computation $c(R,w,D,\text{exp})$.
+As a result, the dependency graph of a multi-stencil program represents the dependencies between local and stencil computations of $\Gamma$.
+
+\begin{mydef}
+For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $i \leq j$, it is said that $c_j$ is dependant from $c_i$, denoted $c_i\prec c_j$, if $w_i \cap R_j \neq \emptyset$.
+In this case, $c_i$ has to be computed before $c_j$.
+The binary relation $c_i\prec c_j$ represents a \textit{dependency}.
+\end{mydef}
+
+\begin{myprop}
+The binary relation $\prec$ is not transitive. 
+\end{myprop}
+
+\begin{proof}
+Considering three computations $c_0(R_0,w_0,D_0,e_0)$, $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, where $w_0\cap R_1 \neq \emptyset$ and $w_1\cap R_2 \neq \emptyset$.
+In this case two dependencies can be extracted $c_0 \prec c_1$ and $c_1 \prec c_2$.
+However, the dependency $c_0 \prec c_2$ is not true as $w_0\cap R_2 = \emptyset$.
+\end{proof}
+
+It has been proved that the binary relation $\prec$ is not transitive.
+This property is due to the fact that $\prec$ is defined as a data dependency between computations.
+A dependency chain, however, creates another type of dependency only due to time.
+
+\begin{mydef}
+For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, a time dependency is denoted $c_i \blacktriangleleft c_j$ and means that $c_i$ has to be computed before $c_j$.
+\end{mydef}
+
+\begin{myprop}
+For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$ such that $c_i \prec c_j$, $c_i \blacktriangleleft c_j$ is verified.
+\end{myprop}
+
+\begin{proof}
+By definition $c_i \prec c_j$ means that $w_i\cap R_j \neq \emptyset$ and that $c_i$ has to computed before $c_j$. %Thus, the relation $c_i \prec c_j$ is stronger than $c_i \blacktriangleleft c_j$.
+\end{proof}
+
+\begin{myprop}
+The binary relation $\blacktriangleleft$ is transitive. 
+\end{myprop}
+
+\begin{proof}
+Considering three computations $c_0(R_0,w_0,D_0,e_0)$, $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, where $c_0 \blacktriangleleft c_1$ and $c_1 \blacktriangleleft c_2$.
+$c_0$ is computed before $c_1$ and $c_1$ is computed before $c_2$, as a result $c_0$ is computed before $c_2$ and the relation $c_0 \blacktriangleleft c_2$ is verified.
+\end{proof}
+
+\begin{mydef}
+A directed acyclic graph (DAG) $G(V,A)$ is a graph where the edges are directed from a source to a destination vertex, and where following the direction of edges, no cycle can be found from a vertex $u$ to itself.
+A directed edge is called an arc, and for two vertices $v,u \in V$ an arc from $u$ to $v$ is denoted $(\overset{\frown}{u,v}) \in A$.
+\end{mydef}
+
+From an ordered list of computations, a directed dependency graph $\Gamma_{task}(V,A)$ can be built finding all pairs of computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $i<j$, such that $c_i \blacktriangleleft c_j$.
+Those pairs of relations can be found from pairs of data dependencies $c_i \prec c_j$.
+
+\begin{mydef}
+From an ordered list of computation $c(R,w,D,\text{exp})$, a directed dependency graph $\Gamma_{task}(V,A)$ is obtained from the call $T_{task}(\Gamma,0)$, where $T_{task}$ is the recursive function
+\begin{equation*}
+T_{task}(\Gamma,i,\Gamma_{task}(V,A)) = 
+\begin{cases} 	(V,A) & \mbox{if }i=|\Gamma|\\
+				T_{task}(\Gamma,i+1,\Gamma_{task}(V\cup c_i, A\cup (\overset{\frown}{c_k,c_i})))\mbox{, }\forall k \leq i \mbox{, } c_k\prec c_i & \mbox{if }i<|\Gamma|
+\end{cases}
+\end{equation*}
+\end{mydef}
+
+This constructive function is possible because the input is an ordered list.
+Actually, if $c_k\prec c_i$ then $k<i$.
+As a result, $c_k$ is already in $V$ when the arc $(\overset{\frown}{c_k,c_i})$ is built.
+
+\begin{myprop}
+The directed graph $\Gamma_{task}$ is an acyclic graph.
+\end{myprop}
+
+\begin{proof}
+$\Gamma_{task}$ is built from $\Gamma$ which is an ordered and sequential list of computations.
+Moreover, each computation of the list $\Gamma$ is associated to a vertex of $V$, even if the same computation is represented more than once in $\Gamma$.
+As a result it is not possible to go back to a previous computation and to create a cycle.
+\end{proof}
+
+Using the function $T_{task}$ to build $\Gamma_{task}$, however, duplication of dependencies may occur because of the transitivity of the relation $\blacktriangleleft$.
+Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,D_k,e_k)$, $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $k<i<j$ and where $c_k \prec c_i$, $c_i \prec c_j$ and $c_k \prec c_j$, then the \textit{pseudo-transitivity} $c_0 \blacktriangleleft c_2$ is verified because of the relations $c_0 \blacktriangleleft c_1$ and $c_1 \blacktriangleleft c_2$.
+However, $c_0 \blacktriangleleft c_2$ is also directly represented by the binary relation $c_0 \prec c_2$.
+As a result, a duplication of the dependency is created in $\Gamma_{task}$.
+In Figure~\ref{fig:duplication} an example of duplication is given for the relation $c_1 \blacktriangleleft c_4$.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (c0) at (0,0) {$c_0$};
+   \node[component] (c1) at (1,0) {$c_1$};
+   \node[component] (c2) at (2,0) {$c_2$};
+   \node[component] (c3) at (3,0) {$c_3$};
+   \node[component] (c4) at (4,0) {$c_4$};
+
+  \path[->]
+    (c0) edge node {} (c1)
+    (c1.east) edge node {} (c2)
+    	 edge [bend right=50] node [swap] {} (c4)
+    (c2) edge node {} (c3)
+    (c3) edge node {} (c4);
+  \end{tikzpicture}
+  \caption{Useless duplication of the dependency $c_1 \blacktriangleleft c_4$}
+  \label{fig:duplication}
+\end{center}
+\end{figure}
+
+Another view of the relations $\prec$ and $\blacktriangleleft$ is that $c_i\blacktriangleleft c_j$ in $\Gamma_{task}$ is a path from $c_i$ to $c_j$ of any size, while $c_i \prec c_j$ is a path of size $1$ from $c_i$ to $c_j$.
+And in any case a path in the graph represents a time dependency between computations.
+Scheduling a graph dependency which do not contains duplication of information is easier.
+For this reason, we define a reduction to apply to $\Gamma_{task}$, and which is performed as a post-transformation.
+In the example of Figure~\ref{fig:duplication} the dependency represented by the path $c_1 \blacktriangleleft c_2 \blacktriangleleft c_3 \blacktriangleleft c_4$ gives more information than the one directly represented by $c_1 \prec c_4$.
+The proposed reduction is based on this property.
+
+\begin{mydef}
+The dependency DAG $\Gamma_{task}(V,A)$ is reduced to the dependency graph $\Gamma_{task}(V,A')$, where $A' \subseteq A$.
+A direct arc $(\overset{\frown}{c_i,c_j})$, $i<j$, of $A$, which means that $c_i \prec c_j$, is included in $A'$ if no longer path $c_i \blacktriangleleft c_j$ exists in $A$ because of transitivity.
+\end{mydef}
+
+%For each resource and from the previous parallelization technique has been obtained a multi-stencil program $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma')$.
+A new stencil program $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma'')$ is built to combine the previous coarse-grain data parallelization technique to a task parallelization technique.
+
+%The set of tasks to perform in a $\mathcal{MSP}_k$ is the set of computations $\Gamma'$.
+From $\Gamma'$ the task graph $\Gamma''$ is built and is a directed acyclic graph (DAG).
+We denote the DAG $\Gamma''(V,A)$, where $V$ is the set of vertices of the DAG, and $A$ is the set of directed edges between vertices.
+Each vertex of the set $V$ of $\Gamma''$ corresponds to a computation in the ordered list $\Gamma'$, and the set $A$ is built from the definition of a dependency.
+
+%In Algorithm~\ref{alg:gammasecond}, $\text{adj}$ denotes the adjacency list of the DAG $\Gamma''$, which associates two vertices together to form a directed edge.
+%In other words, Algorithm~\ref{alg:gammasecond} illustrates the creation of the set of directed edges $A$, thus th creation of $\Gamma''$.
+
+% \begin{center}
+% \begin{algorithm}
+% $adj = \emptyset$\\
+% \For{$c_j \in \Gamma'$}{
+% 	\For{$c_i \in \Gamma'[c_0:c_j]$}{
+% 	\uIf{$c_i<c_j$}{
+% 	adj.append($(c_i,c_j)$)
+% 	}
+% 	}
+% 	}
+% \caption{Creation of $\Gamma''$ from $\Gamma'$}
+% \label{alg:gammasecond}
+% \end{algorithm}
+% \end{center}
+
+%However, the adjacency list obtained from Algorithm~\ref{alg:gammasecond} contains dummy edges...
+
+\paragraph{Hybrid parallelism.}
+It is also possible to combine coarse-grain data and task parallelization techniques to get hybrid parallelism, sometimes more efficient on hybrid architectures.
+A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list with the addition of needed synchronizations.
+On the other hand, the task parallelization technique builds, from an ordered list of computation, a graph dependency $\Gamma_{task}$.
+As a result $T_{data}$ and $T_{task}$ can be composed to build $\Gamma_{hybrid}$ such as $k$ multi-stencil programs
+\begin{equation*}
+\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{hybrid})
+\end{equation*}
+are responsible for an hybrid parallelization of a multi-stencil program. The set of computation $\Gamma_{hybrid}$ is a DAG of dependencies and synchronizations and can be built from the call to 
+\begin{equation*}
+T_{task}(T_{data}(\Gamma,0),0,\Gamma_{hybrid}).
+\end{equation*}

### 2015-05-04 rapport/formalism.tex
rapport_040415

@@
-To numerically solve a set of PDEs, iterative methods are frequently used to approximate the solution by a step by step phenomena.
+To numerically solve a set of PDEs, iterative methods (finite difference, finite volume or finite element methods) are frequently used to approximate the solution through a discretized (step by step) phenomena.
 Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied on a mesh (space discretization).
 In other words, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain.
-Among the numerical computations is found a set of numerical schemes, also called \textit{stencil computations}.
-A formal definition of a \textit{stencil program}, and the parallelization of such program, are presented in this section.
+Among the numerical computations is found a set of numerical schemes, also called \textit{stencil computations}, and a set of local computations also needed to perform the simulation.
+In the following Section, a formal definition of a \textit{stencil program} and its computations is given.
+Then, the different parallelization techniques which can be applied on such program, are presented.
 
@@
 \begin{mydef}
-The discretization of the continuous time domain $\mathcal{T}$ is denoted $T$ such that $\forall t_i, t_{i+1} \in T, \exists \Delta t \in \mathbb{R}$, $t_{i+1} = t_i + \Delta t$.
+The discretization of the continuous time domain $\mathcal{T}$ is denoted $T$ such that $\forall\mbox{ }t_i\mbox{, }t_{i+1} \in T\mbox{, }\exists\mbox{ }\Delta t \in \mathbb{R}$\mbox{, }$t_{i+1} = t_i + \Delta t$.
 Thus, $T$ is responsible for the iteration time steps of the numerical simulation.
@@
 \end{mydef}
-One can notice that in applied mathematics, the signature of $\delta$ would be $\delta : E_i \times T \rightarrow V$, however when programming a numerical simulation it is not wise to store all values of each time iteration.
+One can notice that in applied mathematics, the signature of $\delta$ would be $\delta : D_i \times T \rightarrow V$, however when programming a numerical simulation it is not wise to store all values of each time iteration.
 
@@
 \begin{mydef}
-A computation $c$ of a numerical simulation is defined as $c(R,w,D,\text{exp})$, where $R \subset \Delta, w \in \Delta$ and $\text{exp}$ a numerical expression $\text{exp} : R \rightarrow w$.
+A computation $c$ of a numerical simulation is defined as $c(R,w,D,\text{exp})$, where $R \subset \Delta, w \in \Delta$ and $\text{exp}$ a numerical expression.
 $D$ is one of the subsets $D_i \subset \mathcal{M}$, such that $w : D \rightarrow V$.
@@
 
+As already mentioned, the ordered list $\Gamma$ can be composed of two different types of computations, stencil and local computations, which will be defined in the rest of this Section.
+
 \begin{mydef}
-The neighborhood $\mathcal{N}$ of an element $d \in D_i$ is a function to define a set of elements in any $D_k \subset \mathcal{M}$, $\mathcal{N} : D_i \rightarrow D_k \times D_k \times \dots$.
+The neighborhood $\mathcal{N}$ of an element $d \in D_i$ is a function to obtain a set of elements in any $D_k \subset \mathcal{M}$, $\mathcal{N} : D_i \rightarrow D_k \times D_k \times \dots$.
 \end{mydef}
 The function $\mathcal{N}$ is also sometimes called the \textit{stencil shape}, or the \textit{stencil} in applied mathematics.
-
-A computation $c \in \Gamma$ can be of two different types. The first type is called a \textit{stencil computation}.
+In this paper we distinguish a stencil shape from a \textit{stencil computation} defined as followed:
 
@@
 In this work, a stencil computation $s(R,w,D,\text{exp},\mathcal{N})$ always verifies $R \cap w = \emptyset$, otherwize an implicit numerical scheme has to be solve which is over the scope of this paper.
+As a result, the ordered list $\Gamma$ of a multi-stencil program can be composed of a set of stencil computations applied on one or more stencil shapes.
 
@@
 
+A stencil program and stencil and local computations have been formally defined in this section.
+This formalism is used in the next Section to define two parallelization techniques of a multi-stencil program.
+
 %-------------------------------------
@@
 
-More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\{\mathcal{M}_0,\dots,\mathcal{M}_{n-1}\}$.
+More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
 This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
-As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\{\Delta_0,\dots,\Delta_{n-1}\}$.
+As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\{\Delta_0,\dots,\Delta_{p-1}\}$.
 The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma_{data}$.
@@
 \begin{mydef}
-For $n$ the number of computations in $\Gamma$, and $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j=s_j(R_j,w_j,D_j,\text{exp}_j)$, and $w_i \subset R_j$.
+For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j=s_j(R_j,w_j,D_j,\text{exp}_j)$, and $w_i \subset R_j$.
 Moreover, the data to update is $w_i \cap R_j = w_i$.
@@
 
-Actually, a synchronization can only be needed by the data read in a stencil computation, and only if this data has been modified before, which means that it has been written before.
-This synchronization is needed because the neighborhood function $\mathcal{N}$ of the stencil computation involves values computed on different resources.
+Actually, a synchronization can only be needed by the data read by a stencil computation (not local), and only if this data has been modified before, which means that it has been written before.
+This synchronization is needed because the neighborhood function $\mathcal{N}$ of a stencil computation involves values computed on different resources.
 
@@
 \begin{mydef}
-The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[c_i]_{0 \leq i \leq n-1}$ and $l_2=[w_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[c_0,\dots,c_{n-1},w_0,\dots,w_{m-1}]$.
+An \textit{updated-computation} a computation which needs a data synchronization before it can be performed.
+If $c_i \pprec c_j$, with $i<j$, the computation $c_j$ is transformed to an updated-computation $c^*_j(R_j,w_j,D_j,\text{exp}_j,\text{update}(w_i))$ such that $\text{update}(w_i)$ is performed before the evaluation of $exp_j$.
+The updated-computation can also be denoted as $c^*_j(c_j,\text{update}(w_i))$.
+\end{mydef}
+
+\begin{mydef}
+The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[c_i]_{0 \leq i \leq n-1}$ and $l_2=[c'_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[c_0,\dots,c_{n-1},c'_0,\dots,c'_{m-1}]$.
 \end{mydef}
@@
 \begin{cases} 	[\Gamma[i]] \cdot T_{data}(\Gamma,i+1) & \mbox{if }\forall j\leq i\mbox{, } c_j \not\pprec c_i\\
-				[\text{update}(\{w_k\}),\Gamma[i]] \cdot T_{data}(\Gamma,i+1) & \forall k \leq i\mbox{, }c_k \pprec c_i\\
+				[c^*_i(\Gamma[i],\text{update}(\{w_k\})] \cdot T_{data}(\Gamma,i+1) & \forall k < i\mbox{, }c_k \pprec c_i\\
 				[] & \mbox{if }i=|\Gamma|.
@@
 \begin{myprop}
-Denoting $c_k$ and $c_j$, with $k<j$, two computations of the ordered list $\Gamma_{data}$ for which a synchronization is needed.
-If $c_i$, with $i<j$ is the computation such that $c_i \pprec c_j$ and such that $\text{update}(w_i)$ is performed before $c_j$ in $\Gamma_{data}$.
-And if the needed synchronization before $k$ is $\text{update}(w_l)$, where $l<k$.
-Then, if $i<k$, a single synchronization phase is needed before $c_k$ and is $\text{update}(w_l\cup w_i)$.
+Denoting four computations $c_l$, $c_k$, $c_i$ and $c_j$ of $\Gamma$, with $l<k$, $i<j$, and where $c_l \pprec c_k$ and $c_i \pprec c_j$.
+Then, computations $c_k$ and $c_j$ are tansformed to $c^*_k(c_k,update(w_l)$ and $c^*_j(c_j,update(w_i)$.
+If $i<k$, a single update call can be performed at $k$ such that $c^*_k(c_k,update(w_l\cup w_i)$, and in this case, the computation $c_j$ is not transformed.
 \end{myprop}
@@
 \end{equation}
-is runned.
+is performed.
 
@@
 \begin{mydef}
-For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $i \leq j$, it is said that $c_j$ is dependant from $c_i$, denoted $c_i\prec c_j$, if $w_i \cap R_j \neq \emptyset$.
+For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $i < j$, it is said that $c_j$ is data dependant from $c_i$, denoted $c_i\prec c_j$, if $w_i \cap R_j \neq \emptyset$.
 In this case, $c_i$ has to be computed before $c_j$.
@@
 \end{mydef}
+
+As a result, the relation $\blacktriangleleft$ is more general than $\prec$.
 
@@
 From an ordered list of computations, a directed dependency graph $\Gamma_{task}(V,A)$ can be built finding all pairs of computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $i<j$, such that $c_i \blacktriangleleft c_j$.
-Those pairs of relations can be found from pairs of data dependencies $c_i \prec c_j$.
+Those time dependencies can be found from pairs of data dependencies $c_i \prec c_j$.
 
 \begin{mydef}
-From an ordered list of computation $c(R,w,D,\text{exp})$, a directed dependency graph $\Gamma_{task}(V,A)$ is obtained from the call $T_{task}(\Gamma,0)$, where $T_{task}$ is the recursive function
+For two directed graphs $G(V,A)$ and $G'(V',A')$, the union $(V,A)\cup (V',A')$ is defined as the union of each set $(V\cup V', A \cup A')$.
+\end{mydef}
+
+\begin{mydef}
+From an ordered list $\Gamma$ of computations $c(R,w,D,\text{exp})$, a directed dependency graph $\Gamma_{task}(V,A)$ is obtained from the call $T_{task}(\Gamma,0)$, where $T_{task}$ is the recursive function
 \begin{equation*}
-T_{task}(\Gamma,i,\Gamma_{task}(V,A)) = 
-\begin{cases} 	(V,A) & \mbox{if }i=|\Gamma|\\
-				T_{task}(\Gamma,i+1,\Gamma_{task}(V\cup c_i, A\cup (\overset{\frown}{c_k,c_i})))\mbox{, }\forall k \leq i \mbox{, } c_k\prec c_i & \mbox{if }i<|\Gamma|
+T_{task}(\Gamma,i) = 
+\begin{cases} 	(\{\},\{\}) & \mbox{if }i=|\Gamma|\\
+				(c_i, \{(\overset{\frown}{c_k,c_i})\mbox{, }\forall k < i \mbox{, } c_k\prec c_i \})\cup T_{task}(\Gamma,i+1) & \mbox{if }i<|\Gamma|
 \end{cases}
@@
 Using the function $T_{task}$ to build $\Gamma_{task}$, however, duplication of dependencies may occur because of the transitivity of the relation $\blacktriangleleft$.
-Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,D_k,e_k)$, $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $k<i<j$ and where $c_k \prec c_i$, $c_i \prec c_j$ and $c_k \prec c_j$, then the \textit{pseudo-transitivity} $c_0 \blacktriangleleft c_2$ is verified because of the relations $c_0 \blacktriangleleft c_1$ and $c_1 \blacktriangleleft c_2$.
-However, $c_0 \blacktriangleleft c_2$ is also directly represented by the binary relation $c_0 \prec c_2$.
+Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,D_k,e_k)$, $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $k<i<j$ and where $c_k \prec c_i$, $c_i \prec c_j$ and $c_k \prec c_j$, then the transitivity $c_k \blacktriangleleft c_j$ is verified because of the relations $c_k \blacktriangleleft c_i$ and $c_i \blacktriangleleft c_j$.
+However, $c_k \blacktriangleleft c_j$ is also directly represented by the binary relation $c_k \prec c_j$.
 As a result, a duplication of the dependency is created in $\Gamma_{task}$.
@@
 \begin{mydef}
-The dependency DAG $\Gamma_{task}(V,A)$ is reduced to the dependency graph $\Gamma_{task}(V,A')$, where $A' \subseteq A$.
+The dependency graph $\Gamma_{task}(V,A)$ is reduced to the dependency graph $\Gamma_{task}(V,A')$, where $A' \subseteq A$.
 A direct arc $(\overset{\frown}{c_i,c_j})$, $i<j$, of $A$, which means that $c_i \prec c_j$, is included in $A'$ if no longer path $c_i \blacktriangleleft c_j$ exists in $A$ because of transitivity.
@@
 It is also possible to combine coarse-grain data and task parallelization techniques to get hybrid parallelism, sometimes more efficient on hybrid architectures.
-A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list with the addition of needed synchronizations.
+A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list of computations $c$ and updated-computations $c^*$.
 On the other hand, the task parallelization technique builds, from an ordered list of computation, a graph dependency $\Gamma_{task}$.
@@
 \end{equation*}
-are responsible for an hybrid parallelization of a multi-stencil program. The set of computation $\Gamma_{hybrid}$ is a DAG of dependencies and synchronizations and can be built from the call to 
+are responsible for an hybrid parallelization of a multi-stencil program.
+The set of computation $\Gamma_{hybrid}$ is a dependency graph between computations $c$ and updated-computations $c^*$ and can be built from the call to
 \begin{equation*}
-T_{task}(T_{data}(\Gamma,0),0,\Gamma_{hybrid}).
+T_{task}(T_{data}(\Gamma,0),0).
 \end{equation*}

### 2015-05-25 rapport/component.tex
rapport_25_05

@@
 \subsection{A primitive component model}
-Definitions of a primitive component (ports, interfaces), a component instanciation, an assembly + simple semantics on the behavior of a primitive component.
-Cite L2C as an implementation of such a model + introduction of notations (use,provide,use-multiple,sync).
 
@@
 \subsection{Control components}
-Definition of three control components, and of the kernel component. Explain the behavior.
 
@@
 \subsection{Transformation to a parallel assembly}
-Transformation from a parallel algorithm using the directives: PARALLEL, PARALLEL SECTIONS, SECTION and FORALL, to a component assembly

### 2015-05-25 rapport/formalism.tex
rapport_25_05

--- 2015-04-20 rapport/formalism.tex
@@
-To numerically solve a set of PDEs, iterative methods are frequently used to approximate the solution by a step by step phenomena.
+To numerically solve a set of PDEs, iterative methods (finite difference, finite volume or finite element methods) are frequently used to approximate the solution through a discretized (step by step) phenomena.
 Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied on a mesh (space discretization).
 In other words, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain.
-Among the numerical computations is found a set of numerical schemes, also called \textit{stencil computations}.
-A formal definition of a \textit{stencil program}, and the parallelization of such program, are presented in this section.
+Among the numerical computations is found a set of numerical schemes, also called \textit{stencil computations}, and a set of local computations also needed to perform the simulation.
+In the following Section, a formal definition of a \textit{stencil program} and its computations is given.
+Then, the different parallelization techniques which can be applied on such program, are presented.
 
 %-------------------------------------
-\subsection{Definitions}
+%\subsection{Definitions}
 
@@
 
-%-------------------------------------
-\subsection{Parallelization}
-\label{sect:parall}
-Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
-
-\paragraph{Coarse-grain data parallelism.} In a data parallelization technique, the idea is to split the data on which the program is computed in balanced sub-parts, one for each available resource.
-The same sequential program can afterwards be applied on each sub-part simultaneously, with some additioinal synchronizations between resources to update the data not computed locally and thus to guarantee a correct result.
-
-More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\{\mathcal{M}_0,\dots,\mathcal{M}_{n-1}\}$.
-This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
-As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\{\Delta_0,\dots,\Delta_{n-1}\}$.
-The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma'$.
-
-\begin{mydef}
-For $n$ the number of computations in $\Gamma$, and $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \ll c_j$, if $c_j=s_j(R_j,w_j,D_j,\text{exp}_j)$, and $w_i \cap R_j \neq \emptyset$.
-Moreover, the data to update is $w_i \cap R_j$.
-\end{mydef}
-
-Actually, a synchronization can only be needed by the data read in a stencil computation, and only if this data has been modified before, which means that it has been written before.
-This synchronization is needed because the neighborhood function $\mathcal{N}$ of the stencil computation involves values computed on different resources.
-
-\begin{mydef}
-A synchronization between two computations $c_i \ll c_j$ is defined as a specific computation $\text{update}(w_i \cap R_j)$.
-\end{mydef}
-
-We denote an function to insert at a precise position of a given ordered list, a computation $c_{up}$, as
-\begin{equation}
-\text{insert} : [c_k]_{0 \leq k \leq n-1} \times \text{position} \times c_{up} \rightarrow [c_0,\dots,c_{\text{position}-1},c_{up},\dots,c_{n-1}]
-\end{equation}
-
-The new ordered list of computations $\Gamma'$ is obtained from the insertion of the synchronizations at the good position. Algorithm~\ref{alg:gammaprime} illustrates how $\Gamma'$ is build.
-\begin{center}
-\begin{algorithm}
-$\Gamma' = \Gamma$\\
-$\text{offset} = 0$\\
-\For{$c_j \in \Gamma$}{
-	\uIf{$c_j$ is a stencil}{
-	\uIf{$\Gamma[c_0:c_j]==[]$}{
-	$\text{insert}(\Gamma',0,\text{update}(R_j)$\\
-	$\text{offset}++$\\
-	}
-	\uElse{
-	\For{$c_i \in \Gamma[c_0:c_j]$}{
-	\uIf{$c_i \ll c_j$}{
-	$\text{insert}(\Gamma',j+\text{offset},\text{update}(w_i \cap R_j)$\\
-	$\text{offset}++$
-	}
-	}
-	}
-	}
-	}
-\caption{Creation of $\Gamma'$ from $\Gamma$}
-\label{alg:gammaprime}
-\end{algorithm}
-\end{center}
-
- The final step of this parallelization is to run $\Gamma'$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program defined by
-\begin{equation}
-\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma'),
-\end{equation}
-is runned.
-
-We denote this parallelization technique a coarse-grain data parallelization in contrast with the same technique applied to the finer level of a single computation.
-In this case, for a computation $c(R,w,D,\text{exp})$, the local domain $D$ is divided for $p$ resources $\{D_0,\dots,D_{p-1}\}$, and each resource $k$ is responsible for a sub-computation $c_k(R,w,D_k,\text{exp})$.
-This finer data parallelization does not require the addition of synchronizations because each computation .
-This finer data parallelization technique is not directly adressed by the work presented in this paper as it will be explained later.
-
-\paragraph{Task parallelism.}
+A stencil program and stencil and local computations have been formally defined in this section.
+This formalism is used in the next Section to define two parallelization techniques of a multi-stencil program.

### 2015-05-25 rapport/parallelism.tex
rapport_25_05

Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.

%------------------------------
\subsection{Coarse-grain data parallelism}
In a data parallelization technique, the idea is to split the data on which the program is computed in balanced sub-parts, one for each available resource. The same sequential program can afterwards be applied on each sub-part simultaneously, with some additioinal synchronizations between resources to update the data not computed locally, and thus to guarantee a correct result.

More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$. This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper. As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\{\Delta_0,\dots,\Delta_{p-1}\}$. The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma_{data}$.

\begin{mydef}
For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j=s_j(R_j,w_j,D_j,\text{exp}_j)$, and $w_i \subset R_j$. Moreover, the data to update is $w_i \cap R_j = w_i$.
\end{mydef}

Actually, a synchronization can only be needed by the data read by a stencil computation (not local), and only if this data has been modified before, which means that it has been written before. This synchronization is needed because the neighborhood function $\mathcal{N}$ of a stencil computation involves values computed on different resources.

\begin{mydef}
A synchronization between two computations $c_i \pprec c_j$ is defined as a specific computation $\text{update}(w_i \cap R_j)=\text{update}(w_i)$.
\end{mydef}

\begin{mydef}
An \textit{updated-computation} a computation which needs a data synchronization before it can be performed. If $c_i \pprec c_j$, with $i<j$, the computation $c_j$ is transformed to an updated-computation $c^*_j(R_j,w_j,D_j,\text{exp}_j,\text{update}(w_i))$ such that $\text{update}(w_i)$ is performed before the evaluation of $exp_j$. The updated-computation can also be denoted as $c^*_j(c_j,\text{update}(w_i))$.
\end{mydef}

\begin{mydef}
The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[c_i]_{0 \leq i \leq n-1}$ and $l_2=[c'_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[c_0,\dots,c_{n-1},c'_0,\dots,c'_{m-1}]$.
\end{mydef}

\begin{mydef}
From the ordered list of computation $\Gamma$, a new ordered list $\Gamma_{data}$ is obtained from the call $\Gamma_{data} = T_{data}(\Gamma,0)$, where $T_{data}$ is the recursive function defined as
\begin{equation*}
T_{data}(\Gamma,i) = 
\begin{cases} 	[\Gamma[i]] \cdot T_{data}(\Gamma,i+1) & \mbox{if }\forall j\leq i\mbox{, } c_j \not\pprec c_i\\
				[c^*_i(\Gamma[i],\text{update}(\{w_k\})] \cdot T_{data}(\Gamma,i+1) & \forall k < i\mbox{, }c_k \pprec c_i\\
				[] & \mbox{if }i=|\Gamma|.
\end{cases}
\end{equation*}
\end{mydef}

The construction of $\Gamma_{data}$ implies that if a synchronization is needed it is always proceeded before the computation wich needs it. However, an optimization is possible to combine synchronizations together, and thus to synchronize more than one data at a time. If the number of synchronizations is reduced, final performance of the program will be better. For this reason, we define a reduction to apply to $\Gamma_{data}$, and which is performed as a post-transformation.

\begin{myprop}
Denoting four computations $c_l$, $c_k$, $c_i$ and $c_j$ of $\Gamma$, with $l<k$, $i<j$, and where $c_l \pprec c_k$ and $c_i \pprec c_j$. Then, computations $c_k$ and $c_j$ are tansformed to $c^*_k(c_k,update(w_l)$ and $c^*_j(c_j,update(w_i)$. If $i<k$, a single update call can be performed at $k$ such that $c^*_k(c_k,update(w_l\cup w_i)$, and in this case, the computation $c_j$ is not transformed.
\end{myprop}

\begin{proof}
By definition, the synchronization of $w_i$ can be performed between $c_i$ and $c_j$. As $i<k$ the computation $c_i$ is not performed between $k$ and $j$. As a result, the synchronization of $w_i$ can be performed between $c_i$ and $c_k$.
\end{proof}

In other words, for each synchronization in $\Gamma_{data}$, if the computation $c_i$ from which $c_j$ is synchronized occurs before the last synchronization, the two synchronizations can be reduced to one with the union of the data to update.

 The final step of this parallelization is to run $\Gamma_{data}$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program defined by
\begin{equation}
\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data}),
\end{equation}
is performed.

We denote this parallelization technique a coarse-grain data parallelization in contrast with the same technique applied to the finer level of a single computation. In this case, for a computation $c(R,w,D,\text{exp})$, the local domain $D$ is partitionned for $p$ resources $\{D_0,\dots,D_{p-1}\}$, and each resource $k$ is responsible for a sub-computation $c_k(R,w,D_k,\text{exp}')$.
%This finer data parallelization is particularly adapted to shared memory architectures (multi-cores, many-cores or GPUs), and in such case it does not require the addition of synchronizations because the computed element is local to the sub-domain $D_k$. Actually, if the computation is a stencil, the computation only read neighborhood elements but does not write them which does not require a synchronization, and if the computation is local nothing is read or written in the sub-domain of another resource. 
This finer data parallelization technique is not directly adressed by the work presented in this paper, but is exhibited, as it will be explained later.

%------------------------------
\subsection{Task parallelism}
A task parallelization technique is a technique to transform a program as a dependency graph of different tasks. A dependency graph exhibits parallel tasks, or on the contrary sequential execution of tasks. In a multi-stencil program a task is defined as a computation $c(R,w,D,\text{exp})$. As a result, the dependency graph of a multi-stencil program represents the dependencies between local and stencil computations of $\Gamma$.

\begin{mydef}
For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $i < j$, it is said that $c_j$ is data dependant from $c_i$, denoted $c_i\prec c_j$, if $w_i \cap R_j \neq \emptyset$. In this case, $c_i$ has to be computed before $c_j$. The binary relation $c_i\prec c_j$ represents a \textit{dependency}.
\end{mydef}

\begin{myprop}
The binary relation $\prec$ is not transitive. 
\end{myprop}

\begin{proof}
Considering three computations $c_0(R_0,w_0,D_0,e_0)$, $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, where $w_0\cap R_1 \neq \emptyset$ and $w_1\cap R_2 \neq \emptyset$. In this case two dependencies can be extracted $c_0 \prec c_1$ and $c_1 \prec c_2$. However, the dependency $c_0 \prec c_2$ is not true as $w_0\cap R_2 = \emptyset$.
\end{proof}

It has been proved that the binary relation $\prec$ is not transitive. This property is due to the fact that $\prec$ is defined as a data dependency between computations. A dependency chain, however, creates another type of dependency only due to time.

\begin{mydef}
For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, a time dependency is denoted $c_i \blacktriangleleft c_j$ and means that $c_i$ has to be computed before $c_j$.
\end{mydef}

As a result, the relation $\blacktriangleleft$ is more general than $\prec$.

\begin{myprop}
For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$ such that $c_i \prec c_j$, $c_i \blacktriangleleft c_j$ is verified.
\end{myprop}

\begin{proof}
By definition $c_i \prec c_j$ means that $w_i\cap R_j \neq \emptyset$ and that $c_i$ has to computed before $c_j$. %Thus, the relation $c_i \prec c_j$ is stronger than $c_i \blacktriangleleft c_j$.
\end{proof}

\begin{myprop}
The binary relation $\blacktriangleleft$ is transitive. 
\end{myprop}

\begin{proof}
Considering three computations $c_0(R_0,w_0,D_0,e_0)$, $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, where $c_0 \blacktriangleleft c_1$ and $c_1 \blacktriangleleft c_2$. $c_0$ is computed before $c_1$ and $c_1$ is computed before $c_2$, as a result $c_0$ is computed before $c_2$ and the relation $c_0 \blacktriangleleft c_2$ is verified. 
\end{proof}

\begin{mydef}
A directed acyclic graph (DAG) $G(V,A)$ is a graph where the edges are directed from a source to a destination vertex, and where following the direction of edges, no cycle can be found from a vertex $u$ to itself. A directed edge is called an arc, and for two vertices $v,u \in V$ an arc from $u$ to $v$ is denoted $(\overset{\frown}{u,v}) \in A$.
\end{mydef}

From an ordered list of computations, a directed dependency graph $\Gamma_{task}(V,A)$ can be built finding all pairs of computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $i<j$, such that $c_i \blacktriangleleft c_j$. Those time dependencies can be found from pairs of data dependencies $c_i \prec c_j$.

\begin{mydef}
For two directed graphs $G(V,A)$ and $G'(V',A')$, the union $(V,A)\cup (V',A')$ is defined as the union of each set $(V\cup V', A \cup A')$.
\end{mydef}

\begin{mydef}
From an ordered list $\Gamma$ of computations $c(R,w,D,\text{exp})$, a directed dependency graph $\Gamma_{task}(V,A)$ is obtained from the call $T_{task}(\Gamma,0)$, where $T_{task}$ is the recursive function
\begin{equation*}
T_{task}(\Gamma,i) = 
\begin{cases} 	(\{\},\{\}) & \mbox{if }i=|\Gamma|\\
				(c_i, \{(\overset{\frown}{c_k,c_i})\mbox{, }\forall k < i \mbox{, } c_k\prec c_i \})\cup T_{task}(\Gamma,i+1) & \mbox{if }i<|\Gamma|
\end{cases}
\end{equation*}
\end{mydef}

This constructive function is possible because the input is an ordered list. Actually, if $c_k\prec c_i$ then $k<i$. As a result, $c_k$ is already in $V$ when the arc $(\overset{\frown}{c_k,c_i})$ is built.

\begin{myprop}
The directed graph $\Gamma_{task}$ is an acyclic graph.
\end{myprop}

\begin{proof}
$\Gamma_{task}$ is built from $\Gamma$ which is an ordered and sequential list of computations. Moreover, each computation of the list $\Gamma$ is associated to a vertex of $V$, even if the same computation is represented more than once in $\Gamma$. As a result it is not possible to go back to a previous computation and to create a cycle.
\end{proof}

Using the function $T_{task}$ to build $\Gamma_{task}$, however, duplication of dependencies may occur because of the transitivity of the relation $\blacktriangleleft$. Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,D_k,e_k)$, $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $k<i<j$ and where $c_k \prec c_i$, $c_i \prec c_j$ and $c_k \prec c_j$, then the transitivity $c_k \blacktriangleleft c_j$ is verified because of the relations $c_k \blacktriangleleft c_i$ and $c_i \blacktriangleleft c_j$. However, $c_k \blacktriangleleft c_j$ is also directly represented by the binary relation $c_k \prec c_j$. As a result, a duplication of the dependency is created in $\Gamma_{task}$. In Figure~\ref{fig:duplication} an example of duplication is given for the relation $c_1 \blacktriangleleft c_4$.

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (c0) at (0,0) {$c_0$};
   \node[component] (c1) at (1,0) {$c_1$};
   \node[component] (c2) at (2,0) {$c_2$};
   \node[component] (c3) at (3,0) {$c_3$};
   \node[component] (c4) at (4,0) {$c_4$};

  \path[->]
    (c0) edge node {} (c1)
    (c1.east) edge node {} (c2)
    	 edge [bend right=50] node [swap] {} (c4)
    (c2) edge node {} (c3)
    (c3) edge node {} (c4);
  \end{tikzpicture}
  \caption{Useless duplication of the dependency $c_1 \blacktriangleleft c_4$}
  \label{fig:duplication}
\end{center}
\end{figure}

Another view of the relations $\prec$ and $\blacktriangleleft$ is that $c_i\blacktriangleleft c_j$ in $\Gamma_{task}$ is a path from $c_i$ to $c_j$ of any size, while $c_i \prec c_j$ is a path of size $1$ from $c_i$ to $c_j$. And in any case a path in the graph represents a time dependency between computations. Scheduling a graph dependency which do not contains duplication of information is easier. 
%For this reason, we define a reduction to apply to $\Gamma_{task}$, and which is performed as a post-transformation. In the example of Figure~\ref{fig:duplication} the dependency represented by the path $c_1 \blacktriangleleft c_2 \blacktriangleleft c_3 \blacktriangleleft c_4$ gives more information than the one directly represented by $c_1 \prec c_4$. The proposed reduction is based on this property.

% \begin{mydef}
% The dependency graph $\Gamma_{task}(V,A)$ is reduced to the dependency graph $\Gamma_{task}(V,A')$, where $A' \subseteq A$. A direct arc $(\overset{\frown}{c_i,c_j})$, $i<j$, of $A$, which means that $c_i \prec c_j$, is included in $A'$ if no longer path $c_i \blacktriangleleft c_j$ exists in $A$ because of transitivity.
% \end{mydef}
\begin{mydef}
A DAG is transitive if it contains an arc $(\overset{\frown}{u,v})$ between any two vertices such that there is a path from $u$ to $v$. The transitive closure of a DAG $G=(V,E)$, is the DAG $G_T=(V,E_T)$ for which $E_T$ is the minimal subset of $V \times V$ that includes $E$ and makes $G_T$ transitive.
\end{mydef}

\begin{mydef}
An arc $(\overset{\frown}{u,v})$ of a DAG is redundant if there is a path from $u$ to $v$ in the DAG that does not include the edge. A DAG that does not contain any redundant arc is called minimal. The transitive reduction of a DAG $G$ is its unique minimal sub-graph having the same transitive closure.
\end{mydef}

Thus, the removal of the redundant dependencies of $\Gamma_{task}$ consists in applying a transitive reduction.

%------------------------------
\subsection{Hybrid parallelism}
It is also possible to combine coarse-grain data and task parallelization techniques to get hybrid parallelism, sometimes more efficient on hybrid architectures. A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list of computations $c$ and updated-computations $c^*$. On the other hand, the task parallelization technique builds, from an ordered list of computation, a graph dependency $\Gamma_{task}$. As a result $T_{data}$ and $T_{task}$ can be composed to build $\Gamma_{hybrid}$. Thus, $k$ multi-stencil programs
\begin{equation*}
\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{hybrid})
\end{equation*}
are responsible for an hybrid parallelization of a multi-stencil program. The set of computation $\Gamma_{hybrid}$ is a dependency graph between computations $c$ and updated-computations $c^*$ and can be built from the call to 
\begin{equation*}
T_{task}(T_{data}(\Gamma,0),0).
\end{equation*}

%------------------------------
\subsection{Series-parallel graphs and series-parallel tree decomposition}

### 2015-05-26 rapport/parallelism.tex
rapport_26_05

@@
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (c0) at (0,0) {$c_0$};
-   \node[component] (c1) at (1,0) {$c_1$};
-   \node[component] (c2) at (2,0) {$c_2$};
-   \node[component] (c3) at (3,0) {$c_3$};
-   \node[component] (c4) at (4,0) {$c_4$};
+   \node[] (c0) at (0,0) {$c_0$};
+   \node[] (c1) at (1,0) {$c_1$};
+   \node[] (c2) at (2,0) {$c_2$};
+   \node[] (c3) at (3,0) {$c_3$};
+   \node[] (c4) at (4,0) {$c_4$};
 
@@
 \subsection{Series-parallel graphs and series-parallel tree decomposition}
+In this section we argue that the graphs $\Gamma_{task}$ or $\Gamma_{hybrid}$, previously defined, on which an approximation will be defined, are minimal series-parallel graphs.
+For this reason, the structure of those graphs can be represented as binary trees of parallel and series compositions of sub-graphs, also called \emph{binary decomposition tree}~\cite{Valdes:1979:RSP:800135.804393}.
+First, the needed definitions on series-parallel graphs will be given...
+
+\paragraph{GSP and MSP classes.} 
+A vertex $v$ of a DAG $G$ is a \emph{source} if no edge of $G$ enters $v$.
+Similarly, a vertex $v$ is a \emph{sink} if no edge of $G$ leaves $v$.
+In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of minimal series-parallel DAGs (MSP).
+
+\begin{mydef}Minimal Series Parallel
+\begin{itemize}
+\item The DAG having a single vertex and no edges is MSP.
+\item If $G_1=(V_1,E_1)$ and $G_2=(V_2,E_2)$ are two MSP DAGs, so is either of the DAGs constructed by the following operations:
+\begin{itemize}
+\item Parallel composition: $G_p=(V_1\cup V_2,E_1\cup E_2)$.
+\item Series composition: $G_s=(V_1\cup V_2,E_1\cup E_2\cup (N_1 \times R_2))$, where $N_1$ is the set of sinks of $G_1$ and $R_2$ is the set of sources of $G_2$.
+\end{itemize}
+\end{itemize}
+\end{mydef}
+
+\begin{mydef}
+A DAG is \emph{General Series Parallel} (GSP) if and only if its transitive reduction is a MSP DAG.
+\end{mydef}
+
+A \emph{binary decomposition tree} is a tree having a leaf for each vertex of the MSP DAG it represents, and whose internal nodes are labelled $S$ or $P$ to indicate respectively the series or parallel composition of the MSP sub-DAGs represented by the subtrees rooted at $S$ or $P$.
+Figures~\ref{fig:gsp}, ~\ref{fig:msp} and~\ref{fig:t} respectively give an example of a GSP DAG, its transitive reduction which is MSP, and its tree decomposition.
+
+\begin{figure}[h!]
+\begin{center}
+\subfigure[][\label{fig:gsp}]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (a) at (0,0) {$a$};
+   \node[] (b) at (1,0) {$b$};
+   \node[] (c) at (2,0.5) {$c$};
+   \node[] (d) at (3,0.5) {$d$};
+   \node[] (e) at (2,-0.5) {$e$};
+   \node[] (f) at (3,-0.5) {$f$};
+   \node[] (g) at (4,0) {$g$};
+   \node[] (h) at (1,-1) {$h$};
+   \node[] (i) at (2,-1) {$i$};
+   \node[] (j) at (3,-1) {$j$};
+
+  \path[->]
+    (a) edge node {} (b)
+        edge [bend left=50] node [swap] {} (d)
+        edge node {} (h)
+    (b) edge node {} (c)
+        edge node {} (e)
+    (c) edge node {} (d)
+        edge node {} (f)
+    (e) edge node {} (d)
+        edge node {} (f)
+    (d) edge node {} (g)
+    (f) edge node {} (g)
+    (h) edge node {} (i)
+        edge [bend right=50] node [swap] {} (j)
+    (i) edge node {} (j);
+  \end{tikzpicture}
+}
+\hspace{10pt}
+\subfigure[][\label{fig:msp}]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+  \node[] (a) at (0,0) {$a$};
+   \node[] (b) at (1,0) {$b$};
+   \node[] (c) at (2,0.5) {$c$};
+   \node[] (d) at (3,0.5) {$d$};
+   \node[] (e) at (2,-0.5) {$e$};
+   \node[] (f) at (3,-0.5) {$f$};
+   \node[] (g) at (4,0) {$g$};
+   \node[] (h) at (1,-1) {$h$};
+   \node[] (i) at (2,-1) {$i$};
+   \node[] (j) at (3,-1) {$j$};
+
+  \path[->]
+    (a) edge node {} (b)
+        edge node {} (h)
+    (b) edge node {} (c)
+        edge node {} (e)
+    (c) edge node {} (d)
+        edge node {} (f)
+    (e) edge node {} (d)
+        edge node {} (f)
+    (d) edge node {} (g)
+    (f) edge node {} (g)
+    (h) edge node {} (i)
+    (i) edge node {} (j);
+  \end{tikzpicture}
+}
+\end{center}
+\caption{(a) GSP and (b) MSP DAGs}
+\label{fig:gspmsp}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (S1) at (0,0) {$\mathcal{S}$};
+   \node[] (P1) at (0.5,1) {$\mathcal{P}$};
+   \node[] (a) at (-0.5,1) {$a$};
+   \node[] (S2) at (-1,2) {$\mathcal{S}$}; %+1
+   \node[] (S3) at (2,2) {$\mathcal{S}$}; %-1
+   \node[] (b) at (-2,3) {$b$};
+   \node[] (S4) at (0,3) {$\mathcal{S}$};
+   \node[] (P2) at (-1,4) {$\mathcal{P}$};
+   \node[] (g) at (1,4) {$g$};
+   \node[] (S5) at (-2,5) {$\mathcal{S}$};
+   \node[] (S6) at (0,5) {$\mathcal{S}$};
+   \node[] (c) at (-2.5,6) {$c$};
+   \node[] (d) at (-1.5,6) {$d$};
+   \node[] (e) at (-0.5,6) {$e$};
+   \node[] (f) at (0.5,6) {$f$};
+
+   \node[] (h) at (1,3) {$h$};
+   \node[] (S7) at (3,3) {$\mathcal{S}$};
+   \node[] (i) at (2.5,4) {$i$};
+   \node[] (j) at (3.5,4) {$j$};
+
+  \path[->]
+    (S1) edge node {} (a)
+          edge node {} (P1)
+    (P1) edge node {} (S2)
+          edge node {} (S3)
+    (S2) edge node {} (b)
+        edge node {} (S4)
+    (S4) edge node {} (P2)
+          edge node {} (g)
+    (P2) edge node {} (S5)
+          edge node {} (S6)
+    (S5) edge node {} (c)
+          edge node {} (d)
+    (S6) edge node {} (e)
+          edge node {} (f)
+    (S3) edge node {} (h)
+         edge node {} (S7)
+    (S7) edge node {} (i)
+          edge node {} (j);
+  \end{tikzpicture}
+  \caption{Binary decomposition tree of the MSP of Figure~\ref{fig:msp}}
+  \label{fig:t}
+\end{center}
+\end{figure}
+
+Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have also proposed a linear algorithm to know if a DAG is MSP, and if it is to decompose it to the associated series-parallel tree.
+This algorithm is based on the duality of the class of MSP DAGs, with the class of \emph{Two Terminal Series Parallel} DAGs (TTSP) from which a binary tree decomposition can be performed in linear time.
+Other algorithms have also been proposed, all of them in linear time~\cite{Schoenmakers95anew}.
+
+Finally, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identify a forbidden shape, or subgraph, called $N$ and represented in Figure~\ref{fig:n}.
+
+\begin{mydef}
+A DAG $G$ is GSP if and only if its transitive closure does not contain $N$ as a subgraph.
+\end{mydef}
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (a) at (0,0) {$a$};
+   \node[] (b) at (1,0) {$b$};
+   \node[] (c) at (0,-1) {$c$};
+   \node[] (d) at (1,-1) {$d$};
+
+  \path[->]
+    (a) edge node {} (b)
+    (c) edge node {} (b)
+        edge node {} (d);
+  \end{tikzpicture}
+  \caption{Forbidden $N$ subgraph shape for a DAG to be GSP.}
+  \label{fig:n}
+\end{center}
+\end{figure}
+
+\paragraph{Multi-stencil programs.}
+We are interesting in GSP and MSP classes of graphs to be able to represent computations of a multi-stencil program as a set of sequences or parallel executions.
+Actually, such a representation can directly be dumped to a parallel language, in which \emph{sequence} of instructions and \emph{parallel} execution of instructions are defined.
+Series-parallel trees can also be used as input of scheduling optimizations~\cite{Finta1996323,Wang20082684} to improve task parallelism efficiency.
+Finally, in this paper is presented that such a series-parallel representation can also be dumped to component models by defining specific control components.
+Thus, the proposed DSL inheritates, among others, some software engineering advantages of component models, such as code re-use, productivity and maintainability.
+
+However, even by applying the transitive reduction on the computation DAG $\Gamma_{task}$ (or $\Gamma_{hybrid}$), it is possible to obtain a DAG not MSP.
+Actually, it is possible in a multi-stencil program to have a set of computations such that their dependencies form the forbidden $N$ subgraph.
+For example, if $c_0$, $c_1$, $c_2$ and $c_3$ are computations of a $\mathcal{MSP}$ such that $c_0 \prec c_1$, $c_2 \prec c_1$ and $c_2 \prec c_3$, the \emph{zigzag} relation $c_0 \prec c_1 \succ c_2 \prec c_3$ which form a forbidden subgraph of MSP DAGs is found in $\Gamma_{task}$.
+For this reason, we have made the choice to over-constrain such a case by adding the relation $c_0 \prec c_3$ such that a complete graph is created and can be translated to a series-parallel decomposition as illustrated in Figure~\ref{fig:allover}.
+
+\begin{figure}[h!]
+\begin{center}
+\subfigure[][Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (c0) at (0,0) {$c_0$};
+   \node[] (c1) at (2,0) {$c_1$};
+   \node[] (c2) at (0,-1) {$c_2$};
+   \node[] (c3) at (2,-1) {$c_3$};
+
+  \path[->]
+    (c0) edge node {} (c1)
+          edge [dashed] node [swap] {} (c3)
+    (c2) edge node {} (c1)
+        edge node {} (c3);
+  \end{tikzpicture}
+}
+\hspace{50pt}
+\subfigure[][Series-parallel tree associated to the over-constraint\label{fig:treeover}]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (P1) at (0,0) {$\mathcal{P}$};
+   \node[] (S1) at (-1,1) {$\mathcal{S}$};
+   \node[] (S2) at (1,1) {$\mathcal{S}$};
+   \node[] (c0) at (-1.5,2) {$c_0$};
+   \node[] (c1) at (-0.5,2) {$c_1$};
+   \node[] (c2) at (0.5,2) {$c_2$};
+   \node[] (c3) at (1.5,2) {$c_3$};
+
+  \path[->]
+    (P1) edge [dashed] node [swap] {} (S1)
+          edge [dashed] node [swap] {} (S2)
+    (S1) edge node {} (c0)
+          edge node {} (c1)
+    (S2) edge node {} (c2)
+        edge node {} (c3);
+  \end{tikzpicture}
+}
+\caption{Deletion of forbidden subgraphs.}
+\label{fig:allover}
+\end{center}
+\end{figure}
+
+This approximation on the dependencies is acceptable for multi-stencil programs, first because it rarely appends, second because of the relative homogeneity of computations.
+Actually, all computations, except the ones on the physical border of the domain, are performed on an entire domain of the mesh, and as a computation performs a single quantity at a time ($w_i$ of $c_i$ is a singleton), the amount of arithmetic operations in a computation are quite homogeneous.
+
+After this over-constraint is applied, $\Gamma_{task}$ (or $\Gamma_{hybrid}$) is a MSP DAG. As a result, the binary series-parallel tree decomposition can be computed in linear time.

### 2015-05-27 rapport/component.tex
rapport_27_05

@@
+Instead of dumping the series-parallel tree decomposition of $\Gamma_{task}$ (or $\Gamma_{hybrid}$) to a functionnal high abstraction level language, or to an imperative high abstraction level language, such as OpenMP, we have chosen to dump it to a component model assembly.
+Component models are an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228} where an application is represented as a set of components, each of them responsible for a functionnality, and linked together through their interfaces.
+As a result, a component assembly represents the functional dependencies of the application.
+Each component is an independant black box, and one component could be replaced in the assembly by another one having same interfaces.
+Thus, component models improve code re-use, scalability and maintainability of applications~\cite{Szyperski:2002:CSB:515228,bigot:inria-00388508}.
+Moreover, component models bring a separation of concerns in the application by a clear division of functionnalities in different components.
+
+Recent work on component models have shown a simultaneous answer to performance, maintainability and portability of application~\cite{l2c} which makes component models a serious and interesting candidate for modern high performance computing.
+However, to obtain performance in applications, overheads introduced by separation of concerns in independant components, and their communications, have to be reduced as much as possible at runtime.
+For this reason, only two component models are nowadays considered as HPC-oriented: $L^2C$~\cite{l2c} and CCA~\cite{Allan04acomponent}.
+Both of them propose simple primitive components, and interactions between components are reduced to virtual method calls and MPI communications.
+The resulting component assembly, however, is close to the execution model and some components represent low level concerns which are difficult to implement just like a standard HPC code.
+A DSL hide those difficulties to the end-user while keeping advantages of component models.
+
+Proposing a DSL which is dumped into a component assembly has many advantages:
+\begin{itemize}
+\item advantages of component models are inherited for the DSL itself. Thus, the DSL is easier to maintain, scalable, portable and still efficient;
+\item the DSL conception phase is improved because of code re-use and productivity;
+\item the DSL is light. Its compiler generates a component assembly instead of code.
+\end{itemize}
+
+In this Section, a primitive component model, from which $L^2C$ could be defined, is presented.
+From this primitive model, three \emph{control} components are introduced to finally dump the binary tree decomposition to a component assembly described in the last section.
+
 %-------------------------------------

### 2015-05-27 rapport/parallelism.tex
rapport_27_05

@@
 \paragraph{Multi-stencil programs.}
-We are interesting in GSP and MSP classes of graphs to be able to represent computations of a multi-stencil program as a set of sequences or parallel executions.
+We are interested in GSP and MSP classes of graphs to be able to represent computations of a multi-stencil program as a set of sequences or parallel executions.
 Actually, such a representation can directly be dumped to a parallel language, in which \emph{sequence} of instructions and \emph{parallel} execution of instructions are defined.
@@
 
-After this over-constraint is applied, $\Gamma_{task}$ (or $\Gamma_{hybrid}$) is a MSP DAG. As a result, the binary series-parallel tree decomposition can be computed in linear time.
+After these over-constraints are applied, $\Gamma_{task}$ (or $\Gamma_{hybrid}$) is a MSP DAG.
+As a result, the binary series-parallel tree decomposition can be computed in linear time.
+At this point, the $\mathcal{MSP}$ program could be dumped to any parallel language in which a sequence and a parallel execution is available.
+For example, we could imagine a parallel functionnal language with a binary \emph{sequence} function and a binary \emph{parallel} function such that the tree of Figure~\ref{fig:t} corresponds the following nested call:
+
+\begin{equation}
+\begin{split}
+\text{sequence}(a,\text{parallel}(&\\
+&\text{sequence}(b,\text{sequence}(\text{parallel}(\text{sequence}(c,d),\text{sequence}(e,f)),g)),\\
+&\text{sequence}(h,\text{sequence}(i,j))\\
+&))
+\end{split}
+\end{equation}

### 2015-06-02 rapport/component.tex
rapport_02_06_15

--- 2015-05-25 rapport/component.tex
@@
+Instead of dumping the series-parallel tree decomposition of $\Gamma_{task}$ (or $\Gamma_{hybrid}$) to a functionnal high abstraction level language, or to an imperative high abstraction level language, such as OpenMP, we have chosen to dump it to a component model assembly.
+Component model is an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228} which improves code re-use, scalability and maintainability of applications~\cite{Szyperski:2002:CSB:515228,bigot:inria-00388508}.
+Moreover, component models bring a separation of concerns in the application by a clear division of functionnalities in different components.
+Recent work on component models have shown a simultaneous answer to performance, maintainability and portability of application~\cite{l2c} which makes component models a serious and interesting candidate for modern high performance computing.
+
+Proposing a DSL which is dumped into a component assembly has many advantages:
+\begin{itemize}
+\item advantages of component models are inherited for the DSL itself.
+Thus, the DSL is easier to maintain, scalable, portable and still efficient; %a part portable, le reste devrait etre montrable dans evaluations
+\item the DSL conception phase is improved because of code re-use and productivity; %SEQ,PAR,SSEQ peuvent etre utilises pour tout DSL
+\item the DSL conception phase is facilitated by the separation of concerns in the set of components; %montrer assemblage, et les differentes preoccupations dans les differents composants
+\item the DSL is light as its compiler generates a component assembly instead of code.
+%lié aussi a la separation de preoccupations, ce n'est pas au concepteur de DSL de gérer la generation de code, il utilise des composants qui peuvent etre codes par d'autres
+\end{itemize}
+
 %-------------------------------------
-\subsection{A primitive component model}
+\subsection{Definitions}
+%definir les trucs communs a tous les modeles a composants : composant, port (use, use-multiple, provide), interfaces, assemblages (descriptive, idl), connections valides
+
+%certains modeles propose des concepts plus evolues tels que la hierarchie, la genericite, les connecteurs,
+
+%where an application is represented as a set of components, each of them responsible for a functionnality, and linked together through their interfaces.
+As a result, a component assembly represents the functional dependencies of the application.
+Each component is an independant black box, and one component could be replaced in the assembly by another one having same interfaces.
+Thus, component models
+
+%However, to obtain performance in applications, overheads introduced by separation of concerns of independant components, and their communications, have to be reduced as much as possible at runtime.
+For this reason, only two component models are nowadays considered as HPC-oriented: $L^2C$~\cite{l2c} and CCA~\cite{Allan04acomponent}.
+Both of them propose simple primitive components, and interactions between components are reduced to virtual method calls and MPI communications.
+The resulting component assembly, however, is close to the execution model and some components represent low level concerns which are as difficult to implement as standard HPC codes.
+A DSL hide those difficulties to the end-user while keeping advantages of component models.
 

### 2015-06-02 rapport/msp.tex
rapport_02_06_15

--- 2015-05-27 rapport/parallelism.tex
@@
-Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
-
-%------------------------------
-\subsection{Coarse-grain data parallelism}
-In a data parallelization technique, the idea is to split the data on which the program is computed in balanced sub-parts, one for each available resource.
-The same sequential program can afterwards be applied on each sub-part simultaneously, with some additioinal synchronizations between resources to update the data not computed locally, and thus to guarantee a correct result.
-
-More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
-This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
-As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\{\Delta_0,\dots,\Delta_{p-1}\}$.
-The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma_{data}$.
-
-\begin{mydef}
-For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j=s_j(R_j,w_j,D_j,\text{exp}_j)$, and $w_i \subset R_j$.
-Moreover, the data to update is $w_i \cap R_j = w_i$.
-\end{mydef}
-
-Actually, a synchronization can only be needed by the data read by a stencil computation (not local), and only if this data has been modified before, which means that it has been written before.
-This synchronization is needed because the neighborhood function $\mathcal{N}$ of a stencil computation involves values computed on different resources.
-
-\begin{mydef}
-A synchronization between two computations $c_i \pprec c_j$ is defined as a specific computation $\text{update}(w_i \cap R_j)=\text{update}(w_i)$.
-\end{mydef}
-
-\begin{mydef}
-An \textit{updated-computation} a computation which needs a data synchronization before it can be performed.
-If $c_i \pprec c_j$, with $i<j$, the computation $c_j$ is transformed to an updated-computation $c^*_j(R_j,w_j,D_j,\text{exp}_j,\text{update}(w_i))$ such that $\text{update}(w_i)$ is performed before the evaluation of $exp_j$.
-The updated-computation can also be denoted as $c^*_j(c_j,\text{update}(w_i))$.
-\end{mydef}
-
-\begin{mydef}
-The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[c_i]_{0 \leq i \leq n-1}$ and $l_2=[c'_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[c_0,\dots,c_{n-1},c'_0,\dots,c'_{m-1}]$.
-\end{mydef}
-
-\begin{mydef}
-From the ordered list of computation $\Gamma$, a new ordered list $\Gamma_{data}$ is obtained from the call $\Gamma_{data} = T_{data}(\Gamma,0)$, where $T_{data}$ is the recursive function defined as
-\begin{equation*}
-T_{data}(\Gamma,i) = 
-\begin{cases} 	[\Gamma[i]] \cdot T_{data}(\Gamma,i+1) & \mbox{if }\forall j\leq i\mbox{, } c_j \not\pprec c_i\\
-				[c^*_i(\Gamma[i],\text{update}(\{w_k\})] \cdot T_{data}(\Gamma,i+1) & \forall k < i\mbox{, }c_k \pprec c_i\\
-				[] & \mbox{if }i=|\Gamma|.
-\end{cases}
-\end{equation*}
-\end{mydef}
-
-The construction of $\Gamma_{data}$ implies that if a synchronization is needed it is always proceeded before the computation wich needs it.
-However, an optimization is possible to combine synchronizations together, and thus to synchronize more than one data at a time.
-If the number of synchronizations is reduced, final performance of the program will be better.
-For this reason, we define a reduction to apply to $\Gamma_{data}$, and which is performed as a post-transformation.
-
-\begin{myprop}
-Denoting four computations $c_l$, $c_k$, $c_i$ and $c_j$ of $\Gamma$, with $l<k$, $i<j$, and where $c_l \pprec c_k$ and $c_i \pprec c_j$.
-Then, computations $c_k$ and $c_j$ are tansformed to $c^*_k(c_k,update(w_l)$ and $c^*_j(c_j,update(w_i)$.
-If $i<k$, a single update call can be performed at $k$ such that $c^*_k(c_k,update(w_l\cup w_i)$, and in this case, the computation $c_j$ is not transformed.
-\end{myprop}
-
-\begin{proof}
-By definition, the synchronization of $w_i$ can be performed between $c_i$ and $c_j$.
-As $i<k$ the computation $c_i$ is not performed between $k$ and $j$.
-As a result, the synchronization of $w_i$ can be performed between $c_i$ and $c_k$.
-\end{proof}
-
-In other words, for each synchronization in $\Gamma_{data}$, if the computation $c_i$ from which $c_j$ is synchronized occurs before the last synchronization, the two synchronizations can be reduced to one with the union of the data to update.
-
- The final step of this parallelization is to run $\Gamma_{data}$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program defined by
-\begin{equation}
-\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data}),
-\end{equation}
-is performed.
-
-We denote this parallelization technique a coarse-grain data parallelization in contrast with the same technique applied to the finer level of a single computation.
-In this case, for a computation $c(R,w,D,\text{exp})$, the local domain $D$ is partitionned for $p$ resources $\{D_0,\dots,D_{p-1}\}$, and each resource $k$ is responsible for a sub-computation $c_k(R,w,D_k,\text{exp}')$.
-%This finer data parallelization is particularly adapted to shared memory architectures (multi-cores, many-cores or GPUs), and in such case it does not require the addition of synchronizations because the computed element is local to the sub-domain $D_k$.
-Actually, if the computation is a stencil, the computation only read neighborhood elements but does not write them which does not require a synchronization, and if the computation is local nothing is read or written in the sub-domain of another resource.
-This finer data parallelization technique is not directly adressed by the work presented in this paper, but is exhibited, as it will be explained later.
-
-%------------------------------
-\subsection{Task parallelism}
-A task parallelization technique is a technique to transform a program as a dependency graph of different tasks.
-A dependency graph exhibits parallel tasks, or on the contrary sequential execution of tasks.
-In a multi-stencil program a task is defined as a computation $c(R,w,D,\text{exp})$.
-As a result, the dependency graph of a multi-stencil program represents the dependencies between local and stencil computations of $\Gamma$.
-
-\begin{mydef}
-For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $i < j$, it is said that $c_j$ is data dependant from $c_i$, denoted $c_i\prec c_j$, if $w_i \cap R_j \neq \emptyset$.
-In this case, $c_i$ has to be computed before $c_j$.
-The binary relation $c_i\prec c_j$ represents a \textit{dependency}.
-\end{mydef}
-
-\begin{myprop}
-The binary relation $\prec$ is not transitive. 
-\end{myprop}
-
-\begin{proof}
-Considering three computations $c_0(R_0,w_0,D_0,e_0)$, $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, where $w_0\cap R_1 \neq \emptyset$ and $w_1\cap R_2 \neq \emptyset$.
-In this case two dependencies can be extracted $c_0 \prec c_1$ and $c_1 \prec c_2$.
-However, the dependency $c_0 \prec c_2$ is not true as $w_0\cap R_2 = \emptyset$.
-\end{proof}
-
-It has been proved that the binary relation $\prec$ is not transitive.
-This property is due to the fact that $\prec$ is defined as a data dependency between computations.
-A dependency chain, however, creates another type of dependency only due to time.
-
-\begin{mydef}
-For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, a time dependency is denoted $c_i \blacktriangleleft c_j$ and means that $c_i$ has to be computed before $c_j$.
-\end{mydef}
-
-As a result, the relation $\blacktriangleleft$ is more general than $\prec$.
-
-\begin{myprop}
-For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$ such that $c_i \prec c_j$, $c_i \blacktriangleleft c_j$ is verified.
-\end{myprop}
-
-\begin{proof}
-By definition $c_i \prec c_j$ means that $w_i\cap R_j \neq \emptyset$ and that $c_i$ has to computed before $c_j$. %Thus, the relation $c_i \prec c_j$ is stronger than $c_i \blacktriangleleft c_j$.
-\end{proof}
-
-\begin{myprop}
-The binary relation $\blacktriangleleft$ is transitive. 
-\end{myprop}
-
-\begin{proof}
-Considering three computations $c_0(R_0,w_0,D_0,e_0)$, $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, where $c_0 \blacktriangleleft c_1$ and $c_1 \blacktriangleleft c_2$.
-$c_0$ is computed before $c_1$ and $c_1$ is computed before $c_2$, as a result $c_0$ is computed before $c_2$ and the relation $c_0 \blacktriangleleft c_2$ is verified.
-\end{proof}
-
-\begin{mydef}
-A directed acyclic graph (DAG) $G(V,A)$ is a graph where the edges are directed from a source to a destination vertex, and where following the direction of edges, no cycle can be found from a vertex $u$ to itself.
-A directed edge is called an arc, and for two vertices $v,u \in V$ an arc from $u$ to $v$ is denoted $(\overset{\frown}{u,v}) \in A$.
-\end{mydef}
-
-From an ordered list of computations, a directed dependency graph $\Gamma_{task}(V,A)$ can be built finding all pairs of computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $i<j$, such that $c_i \blacktriangleleft c_j$.
-Those time dependencies can be found from pairs of data dependencies $c_i \prec c_j$.
-
-\begin{mydef}
-For two directed graphs $G(V,A)$ and $G'(V',A')$, the union $(V,A)\cup (V',A')$ is defined as the union of each set $(V\cup V', A \cup A')$.
-\end{mydef}
-
-\begin{mydef}
-From an ordered list $\Gamma$ of computations $c(R,w,D,\text{exp})$, a directed dependency graph $\Gamma_{task}(V,A)$ is obtained from the call $T_{task}(\Gamma,0)$, where $T_{task}$ is the recursive function
-\begin{equation*}
-T_{task}(\Gamma,i) = 
-\begin{cases} 	(\{\},\{\}) & \mbox{if }i=|\Gamma|\\
-				(c_i, \{(\overset{\frown}{c_k,c_i})\mbox{, }\forall k < i \mbox{, } c_k\prec c_i \})\cup T_{task}(\Gamma,i+1) & \mbox{if }i<|\Gamma|
-\end{cases}
-\end{equation*}
-\end{mydef}
-
-This constructive function is possible because the input is an ordered list.
-Actually, if $c_k\prec c_i$ then $k<i$.
-As a result, $c_k$ is already in $V$ when the arc $(\overset{\frown}{c_k,c_i})$ is built.
-
-\begin{myprop}
-The directed graph $\Gamma_{task}$ is an acyclic graph.
-\end{myprop}
-
-\begin{proof}
-$\Gamma_{task}$ is built from $\Gamma$ which is an ordered and sequential list of computations.
-Moreover, each computation of the list $\Gamma$ is associated to a vertex of $V$, even if the same computation is represented more than once in $\Gamma$.
-As a result it is not possible to go back to a previous computation and to create a cycle.
-\end{proof}
-
-Using the function $T_{task}$ to build $\Gamma_{task}$, however, duplication of dependencies may occur because of the transitivity of the relation $\blacktriangleleft$.
-Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,D_k,e_k)$, $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $k<i<j$ and where $c_k \prec c_i$, $c_i \prec c_j$ and $c_k \prec c_j$, then the transitivity $c_k \blacktriangleleft c_j$ is verified because of the relations $c_k \blacktriangleleft c_i$ and $c_i \blacktriangleleft c_j$.
-However, $c_k \blacktriangleleft c_j$ is also directly represented by the binary relation $c_k \prec c_j$.
-As a result, a duplication of the dependency is created in $\Gamma_{task}$.
-In Figure~\ref{fig:duplication} an example of duplication is given for the relation $c_1 \blacktriangleleft c_4$.
-
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (c0) at (0,0) {$c_0$};
-   \node[] (c1) at (1,0) {$c_1$};
-   \node[] (c2) at (2,0) {$c_2$};
-   \node[] (c3) at (3,0) {$c_3$};
-   \node[] (c4) at (4,0) {$c_4$};
-
-  \path[->]
-    (c0) edge node {} (c1)
-    (c1.east) edge node {} (c2)
-    	 edge [bend right=50] node [swap] {} (c4)
-    (c2) edge node {} (c3)
-    (c3) edge node {} (c4);
-  \end{tikzpicture}
-  \caption{Useless duplication of the dependency $c_1 \blacktriangleleft c_4$}
-  \label{fig:duplication}
-\end{center}
-\end{figure}
-
-Another view of the relations $\prec$ and $\blacktriangleleft$ is that $c_i\blacktriangleleft c_j$ in $\Gamma_{task}$ is a path from $c_i$ to $c_j$ of any size, while $c_i \prec c_j$ is a path of size $1$ from $c_i$ to $c_j$.
-And in any case a path in the graph represents a time dependency between computations.
-Scheduling a graph dependency which do not contains duplication of information is easier.
-%For this reason, we define a reduction to apply to $\Gamma_{task}$, and which is performed as a post-transformation.
-In the example of Figure~\ref{fig:duplication} the dependency represented by the path $c_1 \blacktriangleleft c_2 \blacktriangleleft c_3 \blacktriangleleft c_4$ gives more information than the one directly represented by $c_1 \prec c_4$.
-The proposed reduction is based on this property.
-
-% \begin{mydef}
-% The dependency graph $\Gamma_{task}(V,A)$ is reduced to the dependency graph $\Gamma_{task}(V,A')$, where $A' \subseteq A$.
-A direct arc $(\overset{\frown}{c_i,c_j})$, $i<j$, of $A$, which means that $c_i \prec c_j$, is included in $A'$ if no longer path $c_i \blacktriangleleft c_j$ exists in $A$ because of transitivity.
-% \end{mydef}
-\begin{mydef}
-A DAG is transitive if it contains an arc $(\overset{\frown}{u,v})$ between any two vertices such that there is a path from $u$ to $v$.
-The transitive closure of a DAG $G=(V,E)$, is the DAG $G_T=(V,E_T)$ for which $E_T$ is the minimal subset of $V \times V$ that includes $E$ and makes $G_T$ transitive.
-\end{mydef}
-
-\begin{mydef}
-An arc $(\overset{\frown}{u,v})$ of a DAG is redundant if there is a path from $u$ to $v$ in the DAG that does not include the edge.
-A DAG that does not contain any redundant arc is called minimal.
-The transitive reduction of a DAG $G$ is its unique minimal sub-graph having the same transitive closure.
-\end{mydef}
-
-Thus, the removal of the redundant dependencies of $\Gamma_{task}$ consists in applying a transitive reduction.
-
-%------------------------------
-\subsection{Hybrid parallelism}
-It is also possible to combine coarse-grain data and task parallelization techniques to get hybrid parallelism, sometimes more efficient on hybrid architectures.
-A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list of computations $c$ and updated-computations $c^*$.
-On the other hand, the task parallelization technique builds, from an ordered list of computation, a graph dependency $\Gamma_{task}$.
-As a result $T_{data}$ and $T_{task}$ can be composed to build $\Gamma_{hybrid}$.
-Thus, $k$ multi-stencil programs
-\begin{equation*}
-\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{hybrid})
-\end{equation*}
-are responsible for an hybrid parallelization of a multi-stencil program.
-The set of computation $\Gamma_{hybrid}$ is a dependency graph between computations $c$ and updated-computations $c^*$ and can be built from the call to
-\begin{equation*}
-T_{task}(T_{data}(\Gamma,0),0).
-\end{equation*}
-
-%------------------------------
-\subsection{Series-parallel graphs and series-parallel tree decomposition}
 In this section we argue that the graphs $\Gamma_{task}$ or $\Gamma_{hybrid}$, previously defined, on which an approximation will be defined, are minimal series-parallel graphs.
@@
 
-\paragraph{GSP and MSP classes.} 
+%--------------------
+\subsection{GSP and MSP classes}
 A vertex $v$ of a DAG $G$ is a \emph{source} if no edge of $G$ enters $v$.
@@
 \begin{center}
-\subfigure[][\label{fig:gsp}]{
+\subfloat[][\label{fig:gsp}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
 \hspace{10pt}
-\subfigure[][\label{fig:msp}]{
+\subfloat[][\label{fig:msp}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
 
-Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have also proposed a linear algorithm to know if a DAG is MSP, and if it is to decompose it to the associated series-parallel tree.
+Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have also proposed a linear algorithm to know if a DAG is MSP and, if it is, to decompose it to its associated binary decomposition tree.
 This algorithm is based on the duality of the class of MSP DAGs, with the class of \emph{Two Terminal Series Parallel} DAGs (TTSP) from which a binary tree decomposition can be performed in linear time.
-Other algorithms have also been proposed, all of them in linear time~\cite{Schoenmakers95anew}.
-
-Finally, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identify a forbidden shape, or subgraph, called $N$ and represented in Figure~\ref{fig:n}.
+However, in this paper we are interested in DAGs that we already know as MSP.
+Thus, we only give the needed definitions to understand the binary decomposition algorithm, and not the recognition algorithm.
+To let the reader understand this algorithm, which is modified in this work, the needed definitions are given.
+%  Other algorithms have also been proposed, all of them in linear time~\cite{Schoenmakers95anew}.
 
 \begin{mydef}
+The \emph{line digraph} of a digraph $G$ is a digraph $L(G)$ that has:
+\begin{itemize}
+\item a vertex $f(e)$ for each edge $e$ of $G$; and
+\item an edge $(f(e_1),f(e_2))$ for each pair of edges of $G$ of the form $e_1=(u,v)$, $e_2=(v,w)$.
+\end{itemize}
+\end{mydef}
+
+\begin{mydef}Two Terminal Series Parallel
+\begin{itemize}
+\item A digraph consisting of two vertices joined by a single edge is TTSP.
+\item If $G_1$ and $G_2$ are TTSP digraphs, so is the digraph obtained by either of the following operations:
+\begin{itemize}
+\item \emph{Two terminal parallel composition}: identify the sourc of $G_1$ with the source of $G_2$ and the sink of $G_1$ with the sink of $G_2$.
+\item \emph{Two terminal series composition}: identify the sink of $G_1$ with the source of $G_2$.
+\end{itemize}
+\end{itemize}
+\end{mydef}
+
+\begin{myth}
+If the DAG $G$ is a MSP graph, its \emph{inverse line DAG} $L^{-1}(G)$ is TTSP.
+\end{myth}
+
+The binary decomposition tree algorithm is based on the fact that the decomposition can be obtained as a byproduct of a reduction process.
+In order to obtain the decomposition, Valdes \& Al associate a label with each edge of the digraph being reduced.
+Initially the label of each edge is a trivial binary tree consisting of a single node.
+As the reduction process introduces new edges the rules of Figure~\ref{fig:rules} are used to compute the binary trees used to label them.
+The algorithm ends when the TTSP graph is reduced to its minimum, i.e.\ two vertices and a single edge between them.
+
+\begin{figure}[h!]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};
+
+  \path[->]
+    (x) edge node [above] {$T_1$} (y)
+    (y) edge node [above] {$T_2$} (z);
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+\subfloat[]{
+$\Rightarrow$
+}
+\hspace{50pt}
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
+   \node[] (s) at (1.5,0.5) {$\mathcal{S}$};
+   \node[] (t1) at (1,1.5) {$T_1$};
+   \node[] (t2) at (2,1.5) {$T_2$};
+
+  \path[->]
+    (x) edge node {} (y)
+    (s) edge node {} (t1)
+        edge node {} (t2);
+  \end{tikzpicture}
+}
+\\
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
+
+  \path[->]
+    (x) edge [bend left] node [above] {$T_1$} (y);
+        edge node [above] {$T_2$} (y);
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+\subfloat[]{
+$\Rightarrow$
+}
+\hspace{50pt}
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
+   \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
+   \node[] (t1) at (1,1.5) {$T_1$};
+   \node[] (t2) at (2,1.5) {$T_2$};
+
+  \path[->]
+    (x) edge node {} (y)
+    (p) edge node {} (t1)
+        edge node {} (t2);
+  \end{tikzpicture}
+}
+\caption{Reduction rules of the decomposition tree algorithm.}
+\label{fig:rules}
+\end{center}
+\end{figure}
+
+Finally, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identify a forbidden shape, or subgraph, called $N$ and represented in Figure~\ref{fig:n}, such that
+
+\begin{myth}
 A DAG $G$ is GSP if and only if its transitive closure does not contain $N$ as a subgraph.
-\end{mydef}
+\end{myth}
 
@@
 
-\paragraph{Multi-stencil programs.}
-We are interested in GSP and MSP classes of graphs to be able to represent computations of a multi-stencil program as a set of sequences or parallel executions.
+%------------------
+\subsection{Multi-stencil programs.}
+We are interested in GSP and MSP classes of graphs to be able to represent computations of a multi-stencil program as a set of sequences and parallel executions.
 Actually, such a representation can directly be dumped to a parallel language, in which \emph{sequence} of instructions and \emph{parallel} execution of instructions are defined.
-Series-parallel trees can also be used as input of scheduling optimizations~\cite{Finta1996323,Wang20082684} to improve task parallelism efficiency.
+Series-parallel trees can also be used as input of scheduling optimizations~\cite{Finta1996323,Wang20082684} to improve task parallelism efficiency, which opens more perspectives to this work.
 Finally, in this paper is presented that such a series-parallel representation can also be dumped to component models by defining specific control components.
-Thus, the proposed DSL inheritates, among others, some software engineering advantages of component models, such as code re-use, productivity and maintainability.
+Thus, the proposed DSL inheritates software engineering advantages of component models, such as code re-use, productivity and maintainability.
 
@@
 \begin{center}
-\subfigure[][Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
+\subfloat[][Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
 \hspace{50pt}
-\subfigure[][Series-parallel tree associated to the over-constraint\label{fig:treeover}]{
+\subfloat[][Series-parallel tree associated to the over-constraint\label{fig:treeover}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
 As a result, the binary series-parallel tree decomposition can be computed in linear time.
-At this point, the $\mathcal{MSP}$ program could be dumped to any parallel language in which a sequence and a parallel execution is available.
+At this point, the $\mathcal{MSP}$ program could be dumped to any parallel language in which a sequence and a parallel execution are available.
 For example, we could imagine a parallel functionnal language with a binary \emph{sequence} function and a binary \emph{parallel} function such that the tree of Figure~\ref{fig:t} corresponds the following nested call:

### 2015-06-02 rapport/msp.tex
rapport_02_06_15

@@
-In this section we argue that the graphs $\Gamma_{task}$ or $\Gamma_{hybrid}$, previously defined, on which an approximation will be defined, are minimal series-parallel graphs.
+In this section we argue that the graphs $\Gamma_{hybrid}$ or $\Gamma_{task}$, previously defined, on which an approximation will be defined, are minimal series-parallel graphs.
 For this reason, the structure of those graphs can be represented as binary trees of parallel and series compositions of sub-graphs, also called \emph{binary decomposition tree}~\cite{Valdes:1979:RSP:800135.804393}.
@@
 \begin{center}
+%FIRST RULE
 \subfloat[]{
@@
 \hspace{50pt}
+%ARROW
 \subfloat[]{
@@
 \hspace{50pt}
+%SECOND RULE
 \subfloat[]{
@@
   \path[->]
-    (x) edge [bend left] node [above] {$T_1$} (y);
+    (x) edge [bend left] node [above] {$T_1$} (y)
         edge node [above] {$T_2$} (y);
@@
 
-However, even by applying the transitive reduction on the computation DAG $\Gamma_{task}$ (or $\Gamma_{hybrid}$), it is possible to obtain a DAG not MSP.
+Returning back to the parallelization formalism of Section~\ref{}, two important steps have to be performed to make $\Gamma_{hybrid}$ a MSP graph:
+\begin{itemize}
+\item transitive reduction,
+\item deletion of the forbidden $N$ shape.
+\end{itemize}
+
+The transitive reduction has already been applied on $\Gamma_{hybrid}$ by the transitivity of the relation $\blacktriangleleft$.
+However, even by applying this transitive reduction it is possible to obtain a DAG which is not MSP.
 Actually, it is possible in a multi-stencil program to have a set of computations such that their dependencies form the forbidden $N$ subgraph.
-For example, if $c_0$, $c_1$, $c_2$ and $c_3$ are computations of a $\mathcal{MSP}$ such that $c_0 \prec c_1$, $c_2 \prec c_1$ and $c_2 \prec c_3$, the \emph{zigzag} relation $c_0 \prec c_1 \succ c_2 \prec c_3$ which form a forbidden subgraph of MSP DAGs is found in $\Gamma_{task}$.
+For example, if $c_0$, $c_1$, $c_2$ and $c_3$ are computations of a $\mathcal{MSP}$ such that $c_0 \prec c_1$, $c_2 \prec c_1$ and $c_2 \prec c_3$, the \emph{zigzag} relation $c_0 \prec c_1 \succ c_2 \prec c_3$ which form a forbidden subgraph of MSP DAGs is found in $\Gamma_{hybrid}$.
 For this reason, we have made the choice to over-constrain such a case by adding the relation $c_0 \prec c_3$ such that a complete graph is created and can be translated to a series-parallel decomposition as illustrated in Figure~\ref{fig:allover}.
@@
 
-After these over-constraints are applied, $\Gamma_{task}$ (or $\Gamma_{hybrid}$) is a MSP DAG.
-As a result, the binary series-parallel tree decomposition can be computed in linear time.
-At this point, the $\mathcal{MSP}$ program could be dumped to any parallel language in which a sequence and a parallel execution are available.
+After these over-constraints are applied, $\Gamma_{hybrid}$ is a MSP DAG.
+As a result, the binary tree decomposition algorithm of Valdes \& Al can be applied on $\Gamma_{hybrid}$.
+However, because of updated-computations inserted in $\Gamma_{data}$, we have to define additional reduction rules for the algorithm.
+For an updated-computation $c^*_j(c_j,\text{update}(\{w_i\text{, }\forall i\text{ }c_i \prec c_j\}))$, we define $*_j$ as the update needed by $c_j$, $*_j = \text{update}(\{w_i\text{, }\forall i\text{ }c_i \prec c_j\})$.
+As a result, the updated-computation $c^*_j$ is defined by $c^*_j(c_j,*_j)$.
+The initial label of the edges of the graph to reduce are the computations of the multi-stencil program.
+Four new rules are needed to perform the binary tree decomposition of $\Gamma_{hybrid}$ and are defined in Figure~\ref{fig:newrules}.
+
+\begin{figure}[h!]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+%FIRST RULE
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};
+
+  \path[->]
+    (x) edge node [above] {$c_1^*$} (y)
+    (y) edge node [above] {$c_2$} (z);
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+%ARROW
+\subfloat[]{
+$\Rightarrow$
+}
+\hspace{50pt}
+%TRANSFORMATION
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
+   \node[] (s1) at (1.5,0.5) {$\mathcal{S}$};
+   \node[] (s2) at (1,1.5) {$\mathcal{S}$};
+   \node[] (star) at (0.5,2.5) {$*_1$};
+   \node[] (t1) at (1.5,2.5) {$c_1$};
+   \node[] (t2) at (2,1.5) {$c_2$};
+
+  \path[->]
+    (x) edge node {} (y)
+    (s1) edge node {} (s2)
+        edge node {} (t2)
+    (s2) edge node {} (star)
+         edge node {} (t1);
+  \end{tikzpicture}
+}
+\\
+%SECOND RULE
+%FIRST RULE
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};
+
+  \path[->]
+    (x) edge node [above] {$c_1^*$} (y)
+    (y) edge node [above] {$c_2^*$} (z);
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+%ARROW
+\subfloat[]{
+$\Rightarrow$
+}
+\hspace{50pt}
+%TRANSFORMATION
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
+   \node[] (s1) at (1.5,0.5) {$\mathcal{S}$};
+   \node[] (s2) at (0.5,1.5) {$\mathcal{S}$};
+   \node[] (s3) at (2.5,1.5) {$\mathcal{S}$};
+   \node[] (star1) at (0,2.5) {$*_1$};
+   \node[] (t1) at (1,2.5) {$c_1$};
+   \node[] (star2) at (2,2.5) {$*_2$};
+   \node[] (t2) at (3,2.5) {$c_2$};
+
+  \path[->]
+    (x) edge node {} (y)
+    (s1) edge node {} (s2)
+        edge node {} (s3)
+    (s2) edge node {} (star1)
+         edge node {} (t1)
+    (s3) edge node {} (star2)
+         edge node {} (t2);
+  \end{tikzpicture}
+}
+\\
+%THIRD RULE
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
+
+  \path[->]
+    (x) edge [bend left] node [above] {$c_1^*$} (y)
+        edge [bend right] node [below] {$c_2^*$} (y);
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+%ARROW
+\subfloat[]{
+$\Rightarrow$
+}
+\hspace{50pt}
+%TRANSFORMATION
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
+   \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
+   \node[] (s) at (1,1.5) {$\mathcal{S}$};
+   \node[] (star) at (0.5,2.5) {$*_1$};
+   \node[] (t1) at (1.5,2.5) {$c_1$};
+   \node[] (t2) at (2,1.5) {$c_2$};
+
+  \path[->]
+    (x) edge node {} (y)
+    (p) edge node {} (s)
+        edge node {} (t2)
+    (s) edge node {} (star)
+        edge node {} (t1);
+  \end{tikzpicture}
+}
+\\
+%FOURTH RULE
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
+
+  \path[->]
+    (x) edge [bend left] node [above] {$c_1^*$} (y)
+        edge [bend right] node [below] {$c_2^*$} (y);
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+%ARROW
+\subfloat[]{
+$\Rightarrow$
+}
+\hspace{50pt}
+%TRANSFORMATION
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
+   \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
+   \node[] (s2) at (0.5,1.5) {$\mathcal{S}$};
+   \node[] (s3) at (2.5,1.5) {$\mathcal{S}$};
+   \node[] (star1) at (0,2.5) {$*_1$};
+   \node[] (t1) at (1,2.5) {$c_1$};
+   \node[] (star2) at (2,2.5) {$*_2$};
+   \node[] (t2) at (3,2.5) {$c_2$};
+
+  \path[->]
+    (x) edge node {} (y)
+    (p) edge node {} (s2)
+        edge node {} (s3)
+    (s2) edge node {} (star1)
+         edge node {} (t1)
+    (s3) edge node {} (star2)
+         edge node {} (t2);
+  \end{tikzpicture}
+}
+\caption{Reduction rules of the decomposition tree algorithm.}
+\label{fig:newrules}
+\end{center}
+\end{figure}
+
+Using those additional reduction rules, the binary series-parallel tree decomposition can be computed in linear time.
+A complete example of reduction is given in Figure~\ref{fig:example}.
+The initial $\Gamma_{hybrid}$ DAG and its inverse line graph are given first, followed by the set of reductions.
+The label of the final single edge corresponds to the binary tree decomposition of the initial $\Gamma_{hybrid}$ graph.
+
+%====BIG EXAMPLE
+\begin{figure}[ht!]
+\begin{center}
+% INITIAL GRAPH
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (c0) at (0,0) {$c_0$};
+   \node[] (c1) at (1,0) {$c_1^*$};
+   \node[] (c2) at (2,0.5) {$c_2$};
+   \node[] (c3) at (2,-0.5) {$c_3$};
+   \node[] (c4) at (3,0.5) {$c_4^*$};
+   \node[] (c5) at (3,-0.5) {$c_5$};
+   \node[] (c6) at (4,0.5) {$c_6$};
+   \node[] (c7) at (5,0) {$c_7$};
+   \node[] (c8) at (6,0) {$c_8^*$};
+
+  \path[->]
+    (c0) edge node {} (c1)
+    (c1) edge node {} (c2)
+         edge node {} (c3)
+    (c2) edge node {} (c4)
+    (c4) edge node {} (c6)
+    (c3) edge node {} (c5)
+    (c5) edge node {} (c7)
+    (c6) edge node {} (c7)
+    (c7) edge node {} (c8);
+  \end{tikzpicture}
+}
+\\
+% INVERSE LINE GRAPH
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c0d) at (1,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (3,0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (3,-0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (4,0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c7d) at (6,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};
+
+  \path[->]
+    (c0s) edge node [above] {$c_0$} (c0d)
+    (c0d) edge node [above] {$c_1^*$} (c1d)
+    (c1d) edge node [above] {$c_2$} (c2d)
+          edge node [above] {$c_3$} (c3d)
+    (c2d) edge node [above] {$c_4^*$} (c4d)
+    (c4d) edge node [above] {$c_6$} (c6d)
+    (c6d) edge node [above] {$c_7$} (c7d)
+    (c7d) edge node [above] {$c_8$} (c8d)
+    (c3d) edge node [above] {$c_5$} (c6d);
+  \end{tikzpicture}
+}
+\\
+% TRANSFORMATIONS
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (3,0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (3,-0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (4,0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};
+
+   %reduction c0 c1
+   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,1.5) {$c_0$};
+   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
+   \node[] (star1) at (1,2.5) {$*_1$};
+   \node[] (c1) at (2,2.5) {$c_1$};
+
+   %reduction c7 c8
+   \node[] (s3) at (6,0.5) {$\mathcal{S}$};
+   \node[] (c7) at (5.5,1.5) {$c_7$};
+   \node[] (s4) at (6.5,1.5) {$\mathcal{S}$};
+   \node[] (star8) at (6,2.5) {$*_8$};
+   \node[] (c8) at (7,2.5) {$c_8$};
+
+  \path[->]
+    (c0s) edge node {} (c1d)
+    (c1d) edge node [above] {$c_2$} (c2d)
+          edge node [above] {$c_3$} (c3d)
+    (c2d) edge node [above] {$c_4^*$} (c4d)
+    (c4d) edge node [above] {$c_6$} (c6d)
+    (c6d) edge node [above] {} (c8d)
+    (c3d) edge node [above] {$c_5$} (c6d)
+
+    %reduction c0 c1
+    (s1) edge node {} (c0)
+         edge node {} (s2)
+    (s2) edge node {} (star1)
+         edge node {} (c1)
+    %reduction c7 c8
+    (s3) edge node {} (c7)
+         edge node {} (s4)
+    (s4) edge node {} (star8)
+         edge node {} (c8);
+  \end{tikzpicture}
+}
+\\
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (4,0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};
+
+   %reduction c0 c1
+   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,1.5) {$c_0$};
+   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
+   \node[] (star1) at (1,2.5) {$*_1$};
+   \node[] (c1) at (2,2.5) {$c_1$};
+
+   %reduction c7 c8
+   \node[] (s3) at (6,0.5) {$\mathcal{S}$};
+   \node[] (c7) at (5.5,1.5) {$c_7$};
+   \node[] (s4) at (6.5,1.5) {$\mathcal{S}$};
+   \node[] (star8) at (6,2.5) {$*_8$};
+   \node[] (c8) at (7,2.5) {$c_8$};
+
+   %reduction c3 c5
+   \node[] (s5) at (3.5,-2) {$\mathcal{S}$};
+   \node[] (c3) at (3,-1) {$c_3$};
+   \node[] (c5) at (4,-1) {$c_5$};
+
+   %reduction c2 c4
+   \node[] (s6) at (3.5,1) {$\mathcal{S}$};
+   \node[] (c2) at (3,2) {$c_2$};
+   \node[] (s7) at (4,2) {$\mathcal{S}$};
+   \node[] (star4) at (3.5,3) {$*_4$};
+   \node[] (c4) at (4.5,3) {$c_4$};
+
+  \path[->]
+    (c0s) edge node {} (c1d)
+    (c1d) edge [bend left] node [above] {} (c4d)
+    (c4d) edge node [above] {$c_6$} (c6d)
+    (c6d) edge node [above] {} (c8d)
+    (c1d) edge [bend right] node [above] {} (c6d)
+
+    %reduction c0 c1
+    (s1) edge node {} (c0)
+         edge node {} (s2)
+    (s2) edge node {} (star1)
+         edge node {} (c1)
+    %reduction c7 c8
+    (s3) edge node {} (c7)
+         edge node {} (s4)
+    (s4) edge node {} (star8)
+         edge node {} (c8)
+
+    %reduction c3 c5
+    (s5) edge node {} (c3)
+         edge node {} (c5)
+
+    %reduction c7 c8
+    (s6) edge node {} (c2)
+         edge node {} (s7)
+    (s7) edge node {} (star4)
+         edge node {} (c4);
+  \end{tikzpicture}
+}
+\end{center}
+\caption{Big example}
+\label{fig:example}
+\end{figure}
+
+\begin{figure}[ht!]
+\ContinuedFloat
+\begin{center}
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};
+
+   %reduction c0 c1
+   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,1.5) {$c_0$};
+   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
+   \node[] (star1) at (1,2.5) {$*_1$};
+   \node[] (c1) at (2,2.5) {$c_1$};
+
+   %reduction c7 c8
+   \node[] (s3) at (6,0.5) {$\mathcal{S}$};
+   \node[] (c7) at (5.5,1.5) {$c_7$};
+   \node[] (s4) at (6.5,1.5) {$\mathcal{S}$};
+   \node[] (star8) at (6,2.5) {$*_8$};
+   \node[] (c8) at (7,2.5) {$c_8$};
+
+   %reduction c3 c5
+   \node[] (s5) at (3.5,-2) {$\mathcal{S}$};
+   \node[] (c3) at (3,-1) {$c_3$};
+   \node[] (c5) at (4,-1) {$c_5$};
+
+   %reduction c2 c6
+   \node[] (s8) at (3.5,1) {$\mathcal{S}$};
+   \node[] (c6) at (4,2) {$c_6$};
+   \node[] (s6) at (3,2) {$\mathcal{S}$};
+   \node[] (c2) at (2.5,3) {$c_2$};
+   \node[] (s7) at (3.5,3) {$\mathcal{S}$};
+   \node[] (star4) at (3,4) {$*_4$};
+   \node[] (c4) at (4,4) {$c_4$};
+
+  \path[->]
+    (c0s) edge node {} (c1d)
+    (c1d) edge [bend left] node [above] {} (c6d)
+    (c6d) edge node [above] {} (c8d)
+    (c1d) edge [bend right] node [above] {} (c6d)
+
+    %reduction c0 c1
+    (s1) edge node {} (c0)
+         edge node {} (s2)
+    (s2) edge node {} (star1)
+         edge node {} (c1)
+    %reduction c7 c8
+    (s3) edge node {} (c7)
+         edge node {} (s4)
+    (s4) edge node {} (star8)
+         edge node {} (c8)
+
+    %reduction c3 c5
+    (s5) edge node {} (c3)
+         edge node {} (c5)
+
+    %reduction c2 c6
+    (s8) edge node {} (s6)
+         edge node {} (c6)
+    (s6) edge node {} (c2)
+         edge node {} (s7)
+    (s7) edge node {} (star4)
+         edge node {} (c4);
+  \end{tikzpicture}
+}
+\\
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (8,0) {};
+
+   %reduction c0 c1
+   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,1.5) {$c_0$};
+   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
+   \node[] (star1) at (1,2.5) {$*_1$};
+   \node[] (c1) at (2,2.5) {$c_1$};
+
+   %reduction c7 c8
+   \node[] (s3) at (7,0.5) {$\mathcal{S}$};
+   \node[] (c7) at (6.5,1.5) {$c_7$};
+   \node[] (s4) at (7.5,1.5) {$\mathcal{S}$};
+   \node[] (star8) at (7,2.5) {$*_8$};
+   \node[] (c8) at (8,2.5) {$c_8$};
+
+   %reduction c2 c4 c6 c3 c5
+   \node[] (p1) at (4,0.5) {$\mathcal{P}$};
+   \node[] (s5) at (3,1) {$\mathcal{S}$};
+   \node[] (c3) at (2.5,2) {$c_3$};
+   \node[] (c5) at (3.5,2) {$c_5$};
+   \node[] (s8) at (5,1) {$\mathcal{S}$};
+   \node[] (c6) at (5.5,2) {$c_6$};
+   \node[] (s6) at (4.5,2) {$\mathcal{S}$};
+   \node[] (c2) at (4,3) {$c_2$};
+   \node[] (s7) at (5,3) {$\mathcal{S}$};
+   \node[] (star4) at (4.5,4) {$*_4$};
+   \node[] (c4) at (5.5,4) {$c_4$};
+
+  \path[->]
+    (c0s) edge node {} (c1d)
+    (c1d) edge node [above] {} (c6d)
+    (c6d) edge node [above] {} (c8d)
+
+    %reduction c0 c1
+    (s1) edge node {} (c0)
+         edge node {} (s2)
+    (s2) edge node {} (star1)
+         edge node {} (c1)
+    %reduction c7 c8
+    (s3) edge node {} (c7)
+         edge node {} (s4)
+    (s4) edge node {} (star8)
+         edge node {} (c8)
+
+    %reduction c2 c4 c6 c3 c5
+    (p1) edge node {} (s5)
+         edge node {} (s8)
+    (s5) edge node {} (c3)
+         edge node {} (c5)
+    (s8) edge node {} (s6)
+         edge node {} (c6)
+    (s6) edge node {} (c2)
+         edge node {} (s7)
+    (s7) edge node {} (star4)
+         edge node {} (c4);
+  \end{tikzpicture}
+}
+\\
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (8,0) {};
+
+   %all
+   \node[] (s9) at (2.5,1.5) {$\mathcal{S}$};
+   \node[] (s10) at (4,0.5) {$\mathcal{S}$};
+
+   %reduction c0 c1
+   \node[] (s1) at (1,2) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,3) {$c_0$};
+   \node[] (s2) at (1.5,3) {$\mathcal{S}$};
+   \node[] (star1) at (1,4) {$*_1$};
+   \node[] (c1) at (2,4) {$c_1$};
+
+   %reduction c7 c8
+   \node[] (s3) at (7,1) {$\mathcal{S}$};
+   \node[] (c7) at (6.5,2) {$c_7$};
+   \node[] (s4) at (7.5,2) {$\mathcal{S}$};
+   \node[] (star8) at (7,3) {$*_8$};
+   \node[] (c8) at (8,3) {$c_8$};
+
+   %reduction c2 c4 c6 c3 c5
+   \node[] (p1) at (4,2) {$\mathcal{P}$};
+   \node[] (s5) at (3,2.5) {$\mathcal{S}$};
+   \node[] (c3) at (2.5,3.5) {$c_3$};
+   \node[] (c5) at (3.5,3.5) {$c_5$};
+   \node[] (s8) at (5,2.5) {$\mathcal{S}$};
+   \node[] (c6) at (5.5,3.5) {$c_6$};
+   \node[] (s6) at (4.5,3.5) {$\mathcal{S}$};
+   \node[] (c2) at (4,4.5) {$c_2$};
+   \node[] (s7) at (5,4.5) {$\mathcal{S}$};
+   \node[] (star4) at (4.5,5.5) {$*_4$};
+   \node[] (c4) at (5.5,5.5) {$c_4$};
+
+  \path[->]
+    (c0s) edge node {} (c8d)
+
+    %reduction c0 c1
+    (s1) edge node {} (c0)
+         edge node {} (s2)
+    (s2) edge node {} (star1)
+         edge node {} (c1)
+    %reduction c7 c8
+    (s3) edge node {} (c7)
+         edge node {} (s4)
+    (s4) edge node {} (star8)
+         edge node {} (c8)
+
+    %reduction c2 c4 c6 c3 c5
+    (p1) edge node {} (s5)
+         edge node {} (s8)
+    (s5) edge node {} (c3)
+         edge node {} (c5)
+    (s8) edge node {} (s6)
+         edge node {} (c6)
+    (s6) edge node {} (c2)
+         edge node {} (s7)
+    (s7) edge node {} (star4)
+         edge node {} (c4)
+
+    %all
+    (s9) edge node {} (s1)
+         edge node {} (p1)
+    (s10) edge node {} (s9)
+          edge node {} (s3);
+  \end{tikzpicture}
+}
+\end{center}
+\caption{Big example}
+\label{fig:example}
+\end{figure}
+
+At this point, the $\mathcal{MSP}$ program can be dumped to any parallel language in which a sequence and a parallel execution are available.
 For example, we could imagine a parallel functionnal language with a binary \emph{sequence} function and a binary \emph{parallel} function such that the tree of Figure~\ref{fig:t} corresponds the following nested call:

### 2015-06-02 rapport/parallelism.tex
rapport_02_06_15

--- 2015-05-25 rapport/parallelism.tex
@@
 
+%et alors qu'est-ce qu'on fait de Gamma_task ? Il faut le scheduler, donner les travail aux ressources ! On va voir plus tard dans cet article...
+
 %------------------------------
@@
 
-%------------------------------
-\subsection{Series-parallel graphs and series-parallel tree decomposition}
+The rest of this paper deals with the hybrid parallelization of multi-stencil programs described above.
+The first reason for this choice is that this paper does not adress the portability problem, which however will be a perspective.
+Second, by adressing the hybrid parallelism case, we also are able to adress data parallelism and task parallelism, from which hybrid parallelism is built.

### 2015-06-04 rapport/component.tex
rapport_04_06

--- 2015-05-25 rapport/component.tex
@@
-%-------------------------------------
-\subsection{A primitive component model}
+Instead of dumping the series-parallel tree decomposition of $\Gamma_{hybrid}$ (or $\Gamma_{task}$) to a functionnal high abstraction level language, or to an imperative high abstraction level language, such as OpenMP, we have chosen to dump it to a component model assembly.
+Component model is an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228} which improves code re-use, scalability and maintainability of applications~\cite{Szyperski:2002:CSB:515228,bigot:inria-00388508}.
+Moreover, component models bring a separation of concerns in the application by a clear division of functionnalities in different components.
+Recent work on component models have shown a simultaneous answer to performance, maintainability and portability of application~\cite{l2c} which makes component models a serious and interesting candidate for maintainable modern high performance computing.
+
+Proposing a DSL which is dumped to a component assembly has many advantages:
+\begin{itemize}
+\item advantages of component models are inherited for the DSL itself.
+Thus, the DSL is easier to maintain, scalable, portable and still efficient; %a part portable, le reste devrait etre montrable dans evaluations
+\item the DSL conception phase is improved because of code re-use and productivity; %SEQ,PAR,SSEQ peuvent etre utilises pour tout DSL
+\item the DSL conception phase is facilitated by the separation of concerns; %montrer assemblage, et les differentes preoccupations dans les differents composants
+\item the DSL is light as its compilation does not generate codes but a light component assembly.
+%lié aussi a la separation de preoccupations, ce n'est pas au concepteur de DSL de gérer la generation de code, il utilise des composants qui peuvent etre codes par d'autres
+\end{itemize}
+
+Component-based software engineering~\cite{Szyperski:2002:CSB:515228} extends the concept of class by specifying in its interface not only the services it offers, or public methods, but also all its possible interactions with outer world, including the services it requires.
+As a result, each component is an independant entity composed of a set of services its provides, and a set of services it requires (and uses).
+
+A \emph{port} is an entity embbedded in the component which makes possible an \emph{assembly} of components.
+An assembly of components (of instanciated components) is a way to actually produce a complete application by a set of components and their interactions.
+An interface of provided services of a component is associated to a \emph{provide} port, while an interface of required services is associated to a \emph{use} port.
+It is also possible to group more than one required interface into a single port, as a list, called a \emph{use-multiple} port.
+As illustrated in Figures~\ref{fig:control}, a provide port will be represented by a white circle, a use port by a black circle and a use-multiple port by a black circle with a white $m$ in it.
+
+Many component models exist, each of them with its particularities.
+However, components, ports, interfaces and assemblies are common concepts to all component models.
+The rest of this section will only use those concepts to define how the series-parallel tree decomposition, described in the previous Section, can be dumped to a component assembly, whatever the component model is.
 
 %-------------------------------------
-\subsection{Control components}
+\subsection{Control-components}
+As already explained, our aim is to dump the series-parallel tree decomposition of $\Gamma_{hybrid}$ to a component assembly.
+As a result, specific components have to be proposed such that $\mathcal{S}$ and $\mathcal{P}$ nodes of the decomposition can be associated to components.
+For this reason, we propose three specific \emph{control-components}.
+Those components are called control-components because they are used in a different way than usual components.
+Actually, as explained before, a component represents a precise functionnality of the application, while control-components represent a way to control the execution time of other components.
+Each control-component contains a single provide port, this port is associated to a single method which correspond to the execution of the control-component.
+
+\paragraph{Sequence component.} A sequence component, denoted $SEQ$, is a control-component for which the provided service is to garantee that the required services (provided by other components) are used one after the other in a precise order.
+Thus, the sequence component apply synchronous and ordered calls to required services.
+To do so, an ordered use-multiple port is needed.
+
+\paragraph{Synchronized-sequence component.} A synchronized-sequence component, denoted $SSEQ$, is a control-component for which the provided service is to garantee that a set of data synchronizations is performed before a sequence of other required services.
+A $SSEQ$ component is thus composed of two use-multiple ports, one not necessarily ordered to link the $SSEQ$ to data to synchronize (could be more than one), and the second, ordered, to be connect $SSEQ$ to other services, as $SEQ$ does.
+
+\paragraph{Parallel component.} A parallel component, denoted $PAR$, is a control-component for which the provided service is to call simultaneously (asynchronously) a set of required services, and to wait the end of all asynchronous calls before terminating.
+Thus, $PAR$ is composed of a single use-multiple port which does not need to be ordered.
+
+\begin{figure}[h!]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+%FIRST RULE
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (seq) at (0,0) {$SEQ$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (1.5,0) {$m$};
+
+  \path[->]
+    (p) edge node {} (seq)
+    (seq) edge node {} (u);
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+%ARROW
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (seq) at (0,0) {$SSEQ$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u1) at (1.5,0) {$m$};
+   \node[use] (u2) at (0,1) {$m$};
+
+  \path[->]
+    (p) edge node {} (seq)
+    (seq) edge node {} (u1)
+    	  edge node {} (u2);
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+%ARROW
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (seq) at (0,0) {$PAR$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (1.5,0) {$m$};
+
+  \path[->]
+    (p) edge node {} (seq)
+    (seq) edge node {} (u);
+  \end{tikzpicture}
+}
+\caption{$SEQ$, $SSEQ$ and $PAR$ control-components.}
+\label{fig:control}
+\end{center}
+\end{figure}
+
+To be able to dump the series-parallel tree decomposition of $\Gamma_{hybrid}$ to a component assembly, each computation $c_i$ has to be associated to a component.
+Such a component contains a single provide port for which the service is to perform the computation.
+As a computation $c_i(R_i,w_i,D_i,\text{exp}_i)$ manipulates data $R$ and $w$, the component needs a use-multiple port to be linked to data.
+
+\begin{figure}[h!]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (c) at (0,0) {$c_i$};
+   \node[provide] (p) at (-1,0) {};
+   \node[use] (u) at (0,1) {$m$};
+
+  \path[->]
+    (p) edge node {} (c)
+    (c) edge node {} (u);
+  \end{tikzpicture}
+  \caption{Computation component}
+\label{fig:control}
+\end{center}
+\end{figure}
+
+%-------------------------------------
+\subsection{Canonical series-parallel tree decomposition}
+In the series-parallel binary tree decomposition of $\Gamma_{hybrid}$, $\mathcal{S}$ and $\mathcal{P}$ are binary operators.
+On the ther hand, definitions of $SEQ$, $SSEQ$ and $PAR$ are not binary, mainly because of HPC concerns.
+Actually, as less as the number of calls is, the better the performance is.
+In the work of Finta \& Al~\cite{Finta:1996:SUS:235643.235649}, a canonical form of the series-parallel binary tree decomposition has also been studied.
+The canonical form consists in merging two identical and successive compositions ($\mathcal{S}$ or $\mathcal{P}$) as a single one.
+In the case of $\Gamma_{hybrid}$, however, a particular case appear when $\mathcal{S}$ is the source node of a leaf of updates $*_i$.
+In this case, it is forbidden to merge two sequence compositions $\mathcal{S}$.
+The canonical form of the final binary tree decomposition of Figure~\ref{fig:example}, is represented in Figure~\ref{fig:canon}.
+
+\begin{figure}[h!]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (s10) at (4,0.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,2) {$c_0$};
+   \node[] (s2) at (1.5,2) {$\mathcal{S}$};
+   \node[] (star1) at (1,3) {$*_1$};
+   \node[] (c1) at (2,3) {$c_1$};
+   \node[] (c7) at (6.5,2) {$c_7$};
+   \node[] (s4) at (7.5,2) {$\mathcal{S}$};
+   \node[] (star8) at (7,3) {$*_8$};
+   \node[] (c8) at (8,3) {$c_8$};
+   \node[] (p1) at (4,2) {$\mathcal{P}$};
+   \node[] (s5) at (3,2.5) {$\mathcal{S}$};
+   \node[] (c3) at (2.5,3.5) {$c_3$};
+   \node[] (c5) at (3.5,3.5) {$c_5$};
+   \node[] (s8) at (5,2.5) {$\mathcal{S}$};
+   \node[] (c6) at (6,3.5) {$c_6$};
+   \node[] (c2) at (4,3.5) {$c_2$};
+   \node[] (s7) at (5,3.5) {$\mathcal{S}$};
+   \node[] (star4) at (4.5,4.5) {$*_4$};
+   \node[] (c4) at (5.5,4.5) {$c_4$};
+
+  \path[->]
+    (s10) edge node {} (c0)
+    	  edge node {} (s2)
+    	  edge node {} (p1)
+    	  edge node {} (c7)
+    	  edge node {} (s4)
+    (s2) edge node {} (star1)
+         edge node {} (c1)
+    (s4) edge node {} (star8)
+         edge node {} (c8)
+    (p1) edge node {} (s5)
+         edge node {} (s8)
+    (s5) edge node {} (c3)
+         edge node {} (c5)
+    (s8) edge node {} (c2)
+    	 edge node {} (s7)
+         edge node {} (c6)
+    (s7) edge node {} (star4)
+         edge node {} (c4);
+  \end{tikzpicture}
+\caption{Canonical form of the example of Figure~\ref{fig:example}.}
+\label{fig:canon}
+\end{center}
+\end{figure}
 

### 2015-06-04 rapport/msp.tex
rapport_04_06

@@
 
+It has to be indicate that for a DAG $G$, $L^{-1}(G)$ is not unique.
+Thus, in the work of Valdes \& Al, and in this work, $L^{-1}(G)$ will refer to the unique digraph having a single source and a single sink whose line digraph is $L(L^{-1}(G))=G$.
+
 The binary decomposition tree algorithm is based on the fact that the decomposition can be obtained as a byproduct of a reduction process.
@@
     (x) edge [bend left] node [above] {$T_1$} (y)
-        edge node [above] {$T_2$} (y);
+        edge [bend right] node [below] {$T_2$} (y);
   \end{tikzpicture}

### 2015-06-05 rapport/component.tex
rapport_05_06

@@
 
+Every use-multiple port of a control-component can be linked in the assembly to another control-component or to a computation component.
+
 %-------------------------------------
@@
 Actually, as less as the number of calls is, the better the performance is.
-In the work of Finta \& Al~\cite{Finta:1996:SUS:235643.235649}, a canonical form of the series-parallel binary tree decomposition has also been studied.
+In the work of Finta \& Al~\cite{Finta:1996:SUS:235643.235649}, a canonical form of the series-parallel binary tree decomposition has been introduced.
 The canonical form consists in merging two identical and successive compositions ($\mathcal{S}$ or $\mathcal{P}$) as a single one.
@@
 
+From this canonical representation of the multi-stencil program $\mathcal{MSP}$, a direct dump to a component assembly is possible following the transformation rules of Figure~\ref{fig:dump}.
+
+\begin{figure}[h!]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+%SEQ RULE
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (x) at (0,0) {$\mathcal{S}$};
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+%ARROW
+\subfloat[]{
+$\Rightarrow$
+}
+\hspace{50pt}
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (seq) at (0,0) {$SEQ$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (1.5,0) {$m$};
+
+  \path[->]
+    (p) edge node {} (seq)
+    (seq) edge node {} (u);
+\end{tikzpicture}
+}
+%SSEQ RULE
+\\
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (x) at (0,0) {$\mathcal{S}$};
+   \node[] (y) at (0,1) {$*_i$};
+
+  \path[->]
+    (x) edge node {} (y);
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+\subfloat[]{
+$\Rightarrow$
+}
+\hspace{50pt}
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (seq) at (0,0) {$SSEQ$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u1) at (1.5,0) {$m$};
+   \node[use] (u2) at (0,1) {$m$};
+
+  \path[->]
+    (p) edge node {} (seq)
+    (seq) edge node {} (u1)
+    	  edge node {} (u2);
+  \end{tikzpicture}
+}
+\\
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (x) at (0,0) {$\mathcal{P}$};
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+\subfloat[]{
+$\Rightarrow$
+}
+\hspace{50pt}
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (seq) at (0,0) {$PAR$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (1.5,0) {$m$};
+
+  \path[->]
+    (p) edge node {} (seq)
+    (seq) edge node {} (u);
+  \end{tikzpicture}
+}
+\caption{Direct dump rules from the canonical form to the component assembly.}
+\label{fig:dump}
+\end{center}
+\end{figure}
+
+For example, the canonical decomposition tree of Figure~\ref{fig:canon} is dumped into the component assembly of Figure~\ref{fig:assembly}.
+
+\begin{figure}[h!]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   %seq0
+   \node[component] (seq0) at (0,0) {$SEQ$};
+   \node[provide] (seq0p) [left = 1cm of seq0] {};
+   \node[use] (seq0u) [right = 1cm of seq0] {$m$};
+   %c0
+   \node[component] (c0) [right = 1.5cm of seq0u] {$c_0$};
+   \node[use] (c0u) [above = 0.8cm of c0] {$m$};
+   %sseq0
+   \node[component] (sseq0) [below = 1.5cm of c0] {$SSEQ$};
+   \node[use] (sseq0u1) [below = 0.8cm of c0] {$m$};
+   \node[use] (sseq0u2) [right = 1cm of sseq0] {$m$};
+   %c1
+   \node[component] (c1) [right = 1cm of sseq0u2] {$c_1$};
+   \node[use] (c1u) [right = 1cm of c1] {$m$};
+   %par0
+   \node[component] (par0) [below = 1cm of sseq0] {$PAR$};
+   \node[use] (par0u) [right = 1cm of par0] {$m$};
+   %seq2
+   \node[component] (seq2) [right = 1cm of par0u] {$SEQ$};
+   \node[use] (seq2u) [right = 1cm of seq2] {$m$};
+   %c3
+   \node[component] (c3) [right = 1cm of seq2u] {$c_3$};
+   \node[use] (c3u) [above = 0.8cm of c3] {$m$};
+   %c5
+   \node[component] (c5) [below = 1.5cm of c3] {$c_5$};
+   \node[use] (c5u) [below = 0.8cm of c3] {$m$};
+   %seq3
+   \node[component] (seq3) [below = 3cm of seq2] {$SEQ$};
+   \node[use] (seq3u) [right = 1cm of seq3] {$m$};
+   %c2
+   \node[component] (c2) [right = 1cm of seq3u] {$c_2$};
+   \node[use] (c2u) [above = 0.8cm of c2] {$m$};
+   %sseq1
+   \node[component] (sseq1) [below = 1.5cm of c2] {$SSEQ$};
+   \node[use] (sseq1u1) [below = 0.8cm of c2] {$m$};
+   \node[use] (sseq1u2) [right = 1.5cm of sseq1] {$m$};
+   %c4
+   \node[component] (c4) [right = 1cm of sseq1u2] {$c_4$};
+   \node[use] (c4u) [above = 0.8cm of c4] {$m$};
+   %c6
+   \node[component] (c6) [below = 1.5cm of sseq1] {$c_6$};
+   \node[use] (c6u) [above = 0.8cm of c6] {$m$};
+   %c7
+   \node[component] (c7) [below = 5cm of par0] {$c_7$};
+   \node[use] (c7u) [above = 0.8cm of c7] {$m$};
+   %sseq2
+   \node[component] (sseq2) [below = 1.5cm of c7] {$SSEQ$};
+   \node[use] (sseq2u1) [below = 0.8cm of c7] {$m$};
+   \node[use] (sseq2u2) [right = 1.5cm of sseq2] {$m$};
+   %c8
+   \node[component] (c8) [right = 1cm of sseq2u2] {$c_8$};
+   \node[use] (c8u) [above = 0.8cm of c8] {$m$};
+
+   \path[->]
+   %seq0
+    (seq0) edge node {} (seq0u)
+    (seq0p) edge node {} (seq0)
+   %c0
+    (seq0u) edge node {} (c0)
+    (c0) edge node {} (c0u)
+   %seq1
+     (seq0u) edge node {} (sseq0.west)
+     (sseq0) edge node {} (sseq0u1)
+     (sseq0) edge node {} (sseq0u2)
+   %c1
+     (sseq0u2) edge node {} (c1)
+      (c1) edge node {} (c1u)
+   %par0
+    (seq0u) edge node {} (par0.west)
+    (par0) edge node {} (par0u)
+   %seq2
+    (par0u) edge node {} (seq2.west)
+    (seq2) edge node {} (seq2u)
+   %c3
+    (seq2u) edge node {} (c3)
+    (c3) edge node {} (c3u)
+   %c5
+    (seq2u) edge node {} (c5)
+    (c5) edge node {} (c5u)
+   %seq3
+    (par0u) edge node {} (seq3.west)
+    (seq3) edge node {} (seq3u)
+   %c2
+    (seq3u) edge node {} (c2)
+    (c2) edge node {} (c2u)
+   %sseq1
+    (seq3u) edge node {} (sseq1.west)
+    (sseq1) edge node {} (sseq1u1)
+    (sseq1) edge node {} (sseq1u2)
+   %c4
+    (sseq1u2) edge node {} (c4)
+    (c4) edge node {} (c4u)
+   %c6
+    (seq3u) edge node {} (c6)
+    (c6) edge node {} (c6u)
+   %c7
+    (seq0u) edge node {} (c7.west)
+    (c7) edge node {} (c7u)
+   %sseq2
+    (seq0u) edge node {} (sseq2.west)
+    (sseq2) edge node {} (sseq2u1)
+    (sseq2) edge node {} (sseq2u2)
+   %c8
+    (sseq2u2) edge node {} (c8)
+    (c8) edge node {} (c8u);
+\end{tikzpicture}
+\caption{Direct dump rules from the canonical form to the component assembly.}
+\label{fig:assembly}
+\end{center}
+\end{figure}
+
 %-------------------------------------

### 2015-06-05 rapport/component.tex
rapport_05_06

@@
 %-------------------------------------
-\subsection{Transformation to a parallel assembly}
+\subsection{Complete component assembly}
+At this point of the work, the computation part of the parallel $\mathcal{MSP}$ program is solved and dumped to a component assembly.
+Of course, this component assembly does not represent the entire $\mathcal{MSP}$ program.
+Actually, additionnal services are needed.
+For example, computations are performed on quantities, or data, thus those data have to be accessible somewhere through services of components in the assembly.
+Those quantities are mapped onto a distributed mesh, which also means that the mesh has to be accessible by the data.
+In addition to this, and as explained in Section~\ref{sect:dataparal}, the data parallelization needs the management of specific updated-computations $c_i^*$, which means that the complete assembly have to manage an update service, as well as the partitioning of the mesh and the quantities mapped onto it.
+
+However, there is not a unique way to design the rest of the assembly and to make accessible the different services (data, update, mesh etc.).
+The work presented until know can be considered as a generic solution to solve the parallel computation part, but the rest of the assembly is dependant of the implementation choice.
+However, the implementation presented in this work, and detailed in the next Section, follows the SIPSim model~\cite{} which seems to be an interesting solution to dump onto component models.

### 2015-06-05 rapport/parallelism.tex
rapport_05_06

@@
 \subsection{Coarse-grain data parallelism}
+\label{sect:dataparal}
 In a data parallelization technique, the idea is to split the data on which the program is computed in balanced sub-parts, one for each available resource.

### 2015-06-10 rapport/formalism.tex
rapport_10_06

@@
 To numerically solve a set of PDEs, iterative methods (finite difference, finite volume or finite element methods) are frequently used to approximate the solution through a discretized (step by step) phenomena.
-Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied on a mesh (space discretization).
+Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied onto a mesh (space discretization).
 In other words, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain.
-Among the numerical computations is found a set of numerical schemes, also called \textit{stencil computations}, and a set of local computations also needed to perform the simulation.
-In the following Section, a formal definition of a \textit{stencil program} and its computations is given.
+Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}, and a set of auxiliary computations also needed to perform the simulation, and also called \emph{local computations}.
+In the following Section, formal definitions of a \textit{stencil program} and its computations are given.
 Then, the different parallelization techniques which can be applied on such program, are presented.
@@
 %-------------------------------------
-%\subsection{Definitions}
+\subsection{Time, mesh and data}
 
@@
 \begin{mydef}
-$D_i$ is a set of elements of a mesh $\mathcal{M}=(V,E)$, constructed by a function $domain_i$ which defines a precise association between $V$ and $E$, $domain_i : V \times E \rightarrow D_i$.
+$D_i$ is a set of elements of a mesh $\mathcal{M}=(V,E)$, constructed by a function $dom_i$ which defines a precise association between $V$ and $E$, $dom_i : V \times E \rightarrow D_i$.
 \end{mydef}
@@
 \begin{mydef}
-The set of data applied on the mesh is denoted by $\Delta$, such that $\delta \in \Delta$ is a function which associates each element $d \in D_i$ to a value $v \in V$, $\delta : D_i \rightarrow V$.
+The set of data applied on the mesh is denoted by $\Delta$, such that $\delta \in \Delta$ is a function which associates each element $d \in D_i$ (the domain it is applied on) to a value $v \in V$, $\delta : D_i \rightarrow V$.
+In the rest of this paper, the domain of a data $\delta$ can be given by the function $domain(\delta)=D_i$.
 \end{mydef}
@@
 
+%-----------------------
+\subsection{Computations}
+
 \begin{mydef}
-A numerical expression $\text{exp}$ is a function which represents how to compute, for an element $d \in D_i$, a data $w \in \Delta$ (written data) with a set $R \subset \Delta$ of input data (read data), $\text{exp} : R \times D_i \rightarrow w \times D_i$.
+A numerical expression $\text{exp}$ is a function which represents how to compute a value for an element $d \in domain(w)=D_i$, using the set $R \subset \Delta$ of input data (read data), $\text{exp} : R \times D_i \rightarrow w \times D_i$.
 \end{mydef}
@@
 \begin{mydef}
-A computation $c$ of a numerical simulation is defined as $c(R,w,D,\text{exp})$, where $R \subset \Delta, w \in \Delta$ and $\text{exp}$ a numerical expression.
-$D$ is one of the subsets $D_i \subset \mathcal{M}$, such that $w : D \rightarrow V$.
+A computation $c$ of a numerical simulation is defined as $c(R,w,\text{exp})$, where $R \subset \Delta, w \in \Delta$ and $\text{exp}$ a numerical expression.
+%$D$ is one of the subsets $D_i \subset \mathcal{M}$, such that $w : D \rightarrow V$.
 \end{mydef}
 It has to be noticed that at each time iteration, all the elements of a mesh are computed.
-However, it happens that the computation of the mesh elements is splitted in different computations (for example the computation of the physical border).
+However, it happens that computations of the mesh elements are splitted in different domains, as for example the computation of the physical border.
 In this case additional $D_i$ can be specified for the mesh $\mathcal{M}$.
@@
 \begin{mydef}
-The set of $n$ ordered computations of a numerical simulation is denoted $\Gamma = [c_i]_{0 \leq i \leq n-1}$, such that $\forall c_i,c_j$ with $i \leq j$, $c_i$ is computed before $c_j$, and $c_j$ can be computed only when $c_i$ is finished.
+The set of $n$ ordered computations of a numerical simulation is denoted $\Gamma = [c_i]_{0 \leq i \leq n-1}$, such that $\forall c_i,c_j \in \Gamma$, if $i \leq j$, then $c_i$ is computed before $c_j$, and $c_j$ can be computed only when $c_i$ is finished.
 \end{mydef}
@@
 \begin{mydef}
-Finally, a \textit{multi-stencil program} is defined by the quadruplet $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$.
+A \textit{multi-stencil program} is defined by the quadruplet $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$.
 \end{mydef}
@@
 \begin{mydef}
-A \textit{stencil computation} is defined as a quintuplet $s(R,w,D,\text{exp},\mathcal{N})$, where $R \subset \Delta, w \in \Delta$, $D$ is one the subsets $D_i$ and $w : D \rightarrow V$.
+A \textit{stencil computation} is defined as a quadruplet $s(R,w,\text{exp},\mathcal{N})$, where $R \subset \Delta$, and $w : D \rightarrow V \in \Delta$.
 \end{mydef}
 In a stencil computation $s$, $\forall d \in D$, the stencil numerical expression $\text{exp}$ is applied such that $w(d) = \text{exp}(R(d),R(\mathcal{N}(d))$.
-In this work, a stencil computation $s(R,w,D,\text{exp},\mathcal{N})$ always verifies $R \cap w = \emptyset$, otherwize an implicit numerical scheme has to be solve which is over the scope of this paper.
+In this work, a stencil computation $s(R,w,\text{exp},\mathcal{N})$ always verifies $R \cap \{w\} = \emptyset$, otherwize an implicit numerical scheme has to be solve which is over the scope of this paper.
 As a result, the ordered list $\Gamma$ of a multi-stencil program can be composed of a set of stencil computations applied on one or more stencil shapes.
 
-Figure~\ref{fig:ex} gives an example of a stencil computation $s(R,w,D,\text{exp},\mathcal{N})$, where $\mathcal{M}(V,E)$ is a two dimensional Cartesian mesh.
-A single domain $D$ is defined in this example and is composed of cells formed by a cycle of four vertices $v \in V$ and four edges $e \in E$.
+Figure~\ref{fig:ex} gives an example of a stencil computation $s(R,w,\text{exp},\mathcal{N})$, where $\mathcal{M}(V,E)$ is a two dimensional Cartesian mesh.
+A single domain $D=domain(w)$ is defined in this example and is composed of cells formed by a cycle of four vertices $v \in V$ and four edges $e \in E$.
 Furthermore, in this example $R=\{A\}$, $w=B$, and for $(x,y) \in D$ the neighborhood function is
@@
 
-Finally, the second type of numerical computation is a local computation.
+The second type of numerical computation is a local computation.
 \begin{mydef}
-A local computation is a quadruplet $l(R,w,D,\text{exp})$, where $e$ does not involve a neighborhood function $\mathcal{N}$.
+A local computation is a triplet $l(R,w,\text{exp})$, where $exp$ does not involve a neighborhood function $\mathcal{N}$.
 \end{mydef}

### 2015-06-10 rapport/msp.tex
rapport_10_06

@@
 For this reason, the structure of those graphs can be represented as binary trees of parallel and series compositions of sub-graphs, also called \emph{binary decomposition tree}~\cite{Valdes:1979:RSP:800135.804393}.
-First, the needed definitions on series-parallel graphs will be given...
+First, the needed definitions on series-parallel graphs are given, and second the case of the dag $\Gamma_{hybrid}$ is studied.
 
@@
 This algorithm is based on the duality of the class of MSP DAGs, with the class of \emph{Two Terminal Series Parallel} DAGs (TTSP) from which a binary tree decomposition can be performed in linear time.
-However, in this paper we are interested in DAGs that we already know as MSP.
+As it will be explained later, in this paper we are interested in DAGs that we already know to be MSP.
 Thus, we only give the needed definitions to understand the binary decomposition algorithm, and not the recognition algorithm.
@@
 
-It has to be indicate that for a DAG $G$, $L^{-1}(G)$ is not unique.
+It has to be indicated that for a DAG $G$, $L^{-1}(G)$ is not unique.
 Thus, in the work of Valdes \& Al, and in this work, $L^{-1}(G)$ will refer to the unique digraph having a single source and a single sink whose line digraph is $L(L^{-1}(G))=G$.
 
-The binary decomposition tree algorithm is based on the fact that the decomposition can be obtained as a byproduct of a reduction process.
+The binary decomposition tree algorithm is based on the fact that the decomposition can be obtained as a byproduct of a reduction process on $L^{-1}(G)$, which is TTSP.
 In order to obtain the decomposition, Valdes \& Al associate a label with each edge of the digraph being reduced.
@@
 Series-parallel trees can also be used as input of scheduling optimizations~\cite{Finta1996323,Wang20082684} to improve task parallelism efficiency, which opens more perspectives to this work.
-Finally, in this paper is presented that such a series-parallel representation can also be dumped to component models by defining specific control components.
+Finally, in this paper is presented that such a series-parallel representation can also be dumped to component models by defining specific \emph{control components}.
 Thus, the proposed DSL inheritates software engineering advantages of component models, such as code re-use, productivity and maintainability.
 
-Returning back to the parallelization formalism of Section~\ref{}, two important steps have to be performed to make $\Gamma_{hybrid}$ a MSP graph:
+Returning back to the parallelization formalism of Section~\ref{sect:parallelism}, two important steps have to be performed to make $\Gamma_{hybrid}$ a MSP graph:
 \begin{itemize}
@@
    \node[] (c0) at (0,0) {$c_0$};
+   \node[] (bc0) at (-1,0) {};
    \node[] (c1) at (2,0) {$c_1$};
    \node[] (c2) at (0,-1) {$c_2$};
+   \node[] (bc2) at (-1,-1) {};
    \node[] (c3) at (2,-1) {$c_3$};
@@
   \path[->]
+    (bc0) edge [dotted] node {} (c0)
+    (bc2) edge [dotted] node {} (c2)
     (c0) edge node {} (c1)
@@
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (P1) at (0,0) {$\mathcal{P}$};
-   \node[] (S1) at (-1,1) {$\mathcal{S}$};
-   \node[] (S2) at (1,1) {$\mathcal{S}$};
-   \node[] (c0) at (-1.5,2) {$c_0$};
-   \node[] (c1) at (-0.5,2) {$c_1$};
-   \node[] (c2) at (0.5,2) {$c_2$};
-   \node[] (c3) at (1.5,2) {$c_3$};
+   \node[] (S0) at (0,0) {$\mathcal{S}$};
+   \node[] (bS0) at (-1,-1) {};
+   \node[] (P1) at (-1,1) {$\mathcal{P}$};
+   \node[] (aP1) at (-1.5,2) {};
+   \node[] (aaP1) at (-0.5,2) {};
+   \node[] (P2) at (1,1) {$\mathcal{P}$};
+   \node[] (c0) at (-2,3) {$c_0$};
+   \node[] (c1) at (0.5,3) {$c_1$};
+   \node[] (c2) at (0,3) {$c_2$};
+   \node[] (c3) at (2,3) {$c_3$};
 
   \path[->]
-    (P1) edge [dashed] node [swap] {} (S1)
-          edge [dashed] node [swap] {} (S2)
-    (S1) edge node {} (c0)
-          edge node {} (c1)
-    (S2) edge node {} (c2)
+    (bS0) edge [dotted] node {} (S0)
+    (S0) edge node {} (P1)
+         edge node {} (P2)
+    (P1) edge [dotted] node [swap] {} (aP1)
+          edge [dotted] node [swap] {} (aaP1)
+    (aP1) edge node {} (c0)
+    (aaP1) edge node {} (c2)
+    (P2) edge node {} (c1)
         edge node {} (c3);
@@
 
-This approximation on the dependencies is acceptable for multi-stencil programs, first because it rarely appends, second because of the relative homogeneity of computations.
+We consider this approximation acceptable for multi-stencil programs, because of the relative homogeneity of computations.
 Actually, all computations, except the ones on the physical border of the domain, are performed on an entire domain of the mesh, and as a computation performs a single quantity at a time ($w_i$ of $c_i$ is a singleton), the amount of arithmetic operations in a computation are quite homogeneous.
+However, this approximation can also be responsible for a useless wait for computation $c_3$ and less performance.
+For this reason, perspectives of this work open to a new kind of component models.
 
@@
 However, because of updated-computations inserted in $\Gamma_{data}$, we have to define additional reduction rules for the algorithm.
-For an updated-computation $c^*_j(c_j,\text{update}(\{w_i\text{, }\forall i\text{ }c_i \prec c_j\}))$, we define $*_j$ as the update needed by $c_j$, $*_j = \text{update}(\{w_i\text{, }\forall i\text{ }c_i \prec c_j\})$.
+%For an updated-computation $c^*_j(c_j,\text{update}(\{w_i\text{, }\forall i\text{ }c_i \prec c_j\}))$, we define $*_j$ as the update needed by $c_j$, $*_j = \text{update}(\{w_i\text{, }\forall i\text{ }c_i \prec c_j\})$.
 As a result, the updated-computation $c^*_j$ is defined by $c^*_j(c_j,*_j)$.
-The initial label of the edges of the graph to reduce are the computations of the multi-stencil program.
+The initial label of the edges of the graph $L^{-1}(\Gamma_{hybrid})$ to reduce are the computations of the multi-stencil program.
 Four new rules are needed to perform the binary tree decomposition of $\Gamma_{hybrid}$ and are defined in Figure~\ref{fig:newrules}.
+The symetric transforations are not represented.
 
@@
     (x) edge [bend left] node [above] {$c_1^*$} (y)
-        edge [bend right] node [below] {$c_2^*$} (y);
+        edge [bend right] node [below] {$c_2$} (y);
   \end{tikzpicture}
@@
 At this point, the $\mathcal{MSP}$ program can be dumped to any parallel language in which a sequence and a parallel execution are available.
-For example, we could imagine a parallel functionnal language with a binary \emph{sequence} function and a binary \emph{parallel} function such that the tree of Figure~\ref{fig:t} corresponds the following nested call:
+For example, we could imagine a parallel functionnal language with a binary \emph{sequence} function and a binary \emph{parallel} function.% such that the tree of Figure~\ref{fig:t} corresponds the following nested call:
 
-\begin{equation}
-\begin{split}
-\text{sequence}(a,\text{parallel}(&\\
-&\text{sequence}(b,\text{sequence}(\text{parallel}(\text{sequence}(c,d),\text{sequence}(e,f)),g)),\\
-&\text{sequence}(h,\text{sequence}(i,j))\\
-&))
-\end{split}
-\end{equation}
+% \begin{equation}
+% \begin{split}
+% \text{sequence}(a,\text{parallel}(&\\
+% &\text{sequence}(b,\text{sequence}(\text{parallel}(\text{sequence}(c,d),\text{sequence}(e,f)),g)),\\
+% &\text{sequence}(h,\text{sequence}(i,j))\\
+% &))
+% \end{split}
+% \end{equation}

### 2015-06-11 rapport/msp.tex
rapport_11_06

@@
 The initial label of the edges of the graph $L^{-1}(\Gamma_{hybrid})$ to reduce are the computations of the multi-stencil program.
-Four new rules are needed to perform the binary tree decomposition of $\Gamma_{hybrid}$ and are defined in Figure~\ref{fig:newrules}.
+%Four new rules are needed to perform the binary tree decomposition of $\Gamma_{hybrid}$ and are defined in Figure~\ref{fig:newrules}.
 The symetric transforations are not represented.
 
-\begin{figure}[h!]
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-%FIRST RULE
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};
+% \begin{figure}[h!]
+% \captionsetup[subfigure]{labelformat=empty}
+% \begin{center}
+% %FIRST RULE
+% \subfloat[]{
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
+%    \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};
 
-  \path[->]
-    (x) edge node [above] {$c_1^*$} (y)
-    (y) edge node [above] {$c_2$} (z);
-  \end{tikzpicture}
-  }
-\hspace{50pt}
-%ARROW
-\subfloat[]{
-$\Rightarrow$
-}
-\hspace{50pt}
-%TRANSFORMATION
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-   \node[] (s1) at (1.5,0.5) {$\mathcal{S}$};
-   \node[] (s2) at (1,1.5) {$\mathcal{S}$};
-   \node[] (star) at (0.5,2.5) {$*_1$};
-   \node[] (t1) at (1.5,2.5) {$c_1$};
-   \node[] (t2) at (2,1.5) {$c_2$};
+%   \path[->]
+%     (x) edge node [above] {$c_1^*$} (y)
+%     (y) edge node [above] {$c_2$} (z);
+%   \end{tikzpicture}
+%   }
+% \hspace{50pt}
+% %ARROW
+% \subfloat[]{
+% $\Rightarrow$
+% }
+% \hspace{50pt}
+% %TRANSFORMATION
+% \subfloat[]{
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
+%    \node[] (s1) at (1.5,0.5) {$\mathcal{S}$};
+%    \node[] (s2) at (1,1.5) {$\mathcal{S}$};
+%    \node[] (star) at (0.5,2.5) {$*_1$};
+%    \node[] (t1) at (1.5,2.5) {$c_1$};
+%    \node[] (t2) at (2,1.5) {$c_2$};
 
-  \path[->]
-    (x) edge node {} (y)
-    (s1) edge node {} (s2)
-        edge node {} (t2)
-    (s2) edge node {} (star)
-         edge node {} (t1);
-  \end{tikzpicture}
-}
-\\
-%SECOND RULE
-%FIRST RULE
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};
+%   \path[->]
+%     (x) edge node {} (y)
+%     (s1) edge node {} (s2)
+%         edge node {} (t2)
+%     (s2) edge node {} (star)
+%          edge node {} (t1);
+%   \end{tikzpicture}
+% }
+% \\
+% %SECOND RULE
+% %FIRST RULE
+% \subfloat[]{
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
+%    \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};
 
-  \path[->]
-    (x) edge node [above] {$c_1^*$} (y)
-    (y) edge node [above] {$c_2^*$} (z);
-  \end{tikzpicture}
-  }
-\hspace{50pt}
-%ARROW
-\subfloat[]{
-$\Rightarrow$
-}
-\hspace{50pt}
-%TRANSFORMATION
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-   \node[] (s1) at (1.5,0.5) {$\mathcal{S}$};
-   \node[] (s2) at (0.5,1.5) {$\mathcal{S}$};
-   \node[] (s3) at (2.5,1.5) {$\mathcal{S}$};
-   \node[] (star1) at (0,2.5) {$*_1$};
-   \node[] (t1) at (1,2.5) {$c_1$};
-   \node[] (star2) at (2,2.5) {$*_2$};
-   \node[] (t2) at (3,2.5) {$c_2$};
+%   \path[->]
+%     (x) edge node [above] {$c_1^*$} (y)
+%     (y) edge node [above] {$c_2^*$} (z);
+%   \end{tikzpicture}
+%   }
+% \hspace{50pt}
+% %ARROW
+% \subfloat[]{
+% $\Rightarrow$
+% }
+% \hspace{50pt}
+% %TRANSFORMATION
+% \subfloat[]{
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
+%    \node[] (s1) at (1.5,0.5) {$\mathcal{S}$};
+%    \node[] (s2) at (0.5,1.5) {$\mathcal{S}$};
+%    \node[] (s3) at (2.5,1.5) {$\mathcal{S}$};
+%    \node[] (star1) at (0,2.5) {$*_1$};
+%    \node[] (t1) at (1,2.5) {$c_1$};
+%    \node[] (star2) at (2,2.5) {$*_2$};
+%    \node[] (t2) at (3,2.5) {$c_2$};
 
-  \path[->]
-    (x) edge node {} (y)
-    (s1) edge node {} (s2)
-        edge node {} (s3)
-    (s2) edge node {} (star1)
-         edge node {} (t1)
-    (s3) edge node {} (star2)
-         edge node {} (t2);
-  \end{tikzpicture}
-}
-\\
-%THIRD RULE
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
+%   \path[->]
+%     (x) edge node {} (y)
+%     (s1) edge node {} (s2)
+%         edge node {} (s3)
+%     (s2) edge node {} (star1)
+%          edge node {} (t1)
+%     (s3) edge node {} (star2)
+%          edge node {} (t2);
+%   \end{tikzpicture}
+% }
+% \\
+% %THIRD RULE
+% \subfloat[]{
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
 
-  \path[->]
-    (x) edge [bend left] node [above] {$c_1^*$} (y)
-        edge [bend right] node [below] {$c_2$} (y);
-  \end{tikzpicture}
-  }
-\hspace{50pt}
-%ARROW
-\subfloat[]{
-$\Rightarrow$
-}
-\hspace{50pt}
-%TRANSFORMATION
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-   \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
-   \node[] (s) at (1,1.5) {$\mathcal{S}$};
-   \node[] (star) at (0.5,2.5) {$*_1$};
-   \node[] (t1) at (1.5,2.5) {$c_1$};
-   \node[] (t2) at (2,1.5) {$c_2$};
+%   \path[->]
+%     (x) edge [bend left] node [above] {$c_1^*$} (y)
+%         edge [bend right] node [below] {$c_2$} (y);
+%   \end{tikzpicture}
+%   }
+% \hspace{50pt}
+% %ARROW
+% \subfloat[]{
+% $\Rightarrow$
+% }
+% \hspace{50pt}
+% %TRANSFORMATION
+% \subfloat[]{
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
+%    \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
+%    \node[] (s) at (1,1.5) {$\mathcal{S}$};
+%    \node[] (star) at (0.5,2.5) {$*_1$};
+%    \node[] (t1) at (1.5,2.5) {$c_1$};
+%    \node[] (t2) at (2,1.5) {$c_2$};
 
-  \path[->]
-    (x) edge node {} (y)
-    (p) edge node {} (s)
-        edge node {} (t2)
-    (s) edge node {} (star)
-        edge node {} (t1);
-  \end{tikzpicture}
-}
-\\
-%FOURTH RULE
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
+%   \path[->]
+%     (x) edge node {} (y)
+%     (p) edge node {} (s)
+%         edge node {} (t2)
+%     (s) edge node {} (star)
+%         edge node {} (t1);
+%   \end{tikzpicture}
+% }
+% \\
+% %FOURTH RULE
+% \subfloat[]{
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
 
-  \path[->]
-    (x) edge [bend left] node [above] {$c_1^*$} (y)
-        edge [bend right] node [below] {$c_2^*$} (y);
-  \end{tikzpicture}
-  }
-\hspace{50pt}
-%ARROW
-\subfloat[]{
-$\Rightarrow$
-}
-\hspace{50pt}
-%TRANSFORMATION
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-   \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
-   \node[] (s2) at (0.5,1.5) {$\mathcal{S}$};
-   \node[] (s3) at (2.5,1.5) {$\mathcal{S}$};
-   \node[] (star1) at (0,2.5) {$*_1$};
-   \node[] (t1) at (1,2.5) {$c_1$};
-   \node[] (star2) at (2,2.5) {$*_2$};
-   \node[] (t2) at (3,2.5) {$c_2$};
+%   \path[->]
+%     (x) edge [bend left] node [above] {$c_1^*$} (y)
+%         edge [bend right] node [below] {$c_2^*$} (y);
+%   \end{tikzpicture}
+%   }
+% \hspace{50pt}
+% %ARROW
+% \subfloat[]{
+% $\Rightarrow$
+% }
+% \hspace{50pt}
+% %TRANSFORMATION
+% \subfloat[]{
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
+%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
+%    \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
+%    \node[] (s2) at (0.5,1.5) {$\mathcal{S}$};
+%    \node[] (s3) at (2.5,1.5) {$\mathcal{S}$};
+%    \node[] (star1) at (0,2.5) {$*_1$};
+%    \node[] (t1) at (1,2.5) {$c_1$};
+%    \node[] (star2) at (2,2.5) {$*_2$};
+%    \node[] (t2) at (3,2.5) {$c_2$};
 
-  \path[->]
-    (x) edge node {} (y)
-    (p) edge node {} (s2)
-        edge node {} (s3)
-    (s2) edge node {} (star1)
-         edge node {} (t1)
-    (s3) edge node {} (star2)
-         edge node {} (t2);
-  \end{tikzpicture}
-}
-\caption{Reduction rules of the decomposition tree algorithm.}
-\label{fig:newrules}
-\end{center}
-\end{figure}
+%   \path[->]
+%     (x) edge node {} (y)
+%     (p) edge node {} (s2)
+%         edge node {} (s3)
+%     (s2) edge node {} (star1)
+%          edge node {} (t1)
+%     (s3) edge node {} (star2)
+%          edge node {} (t2);
+%   \end{tikzpicture}
+% }
+% \caption{Reduction rules of the decomposition tree algorithm.}
+% \label{fig:newrules}
+% \end{center}
+% \end{figure}
 
-Using those additional reduction rules, the binary series-parallel tree decomposition can be computed in linear time.
+% Using those additional reduction rules, the binary series-parallel tree decomposition can be computed in linear time. 
 A complete example of reduction is given in Figure~\ref{fig:example}.
@@
    \node[] (c0) at (0.5,1.5) {$c_0$};
+   %\node[] (c1) at (1.5,1.5) {$c_1^*$};
    \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
@@
    \node[] (c7) at (5.5,1.5) {$c_7$};
+   %\node[] (c8) at (6.5,1.5) {$c_8^*$};
    \node[] (s4) at (6.5,1.5) {$\mathcal{S}$};
@@
     (s1) edge node {} (c0)
+    %     edge node {} (c1)
         edge node {} (s2)
@@
     (s3) edge node {} (c7)
+    %     edge node {} (c8);
         edge node {} (s4)
@@
    \node[] (c2) at (3,2) {$c_2$};
+   %\node[] (c4) at (4,2) {$c_4^*$};
    \node[] (s7) at (4,2) {$\mathcal{S}$};
@@
     (s1) edge node {} (c0)
+    %     edge node {} (c1)
          edge node {} (s2)
@@
     (s3) edge node {} (c7)
+    %     edge node {} (c8)
          edge node {} (s4)
@@
     (s6) edge node {} (c2)
+    %     edge node {} (c4);
          edge node {} (s7)
@@
    \node[] (c2) at (2.5,3) {$c_2$};
+   %\node[] (c4) at (3.5,3) {$c_4^*$};
    \node[] (s7) at (3.5,3) {$\mathcal{S}$};
@@
    \node[] (c7) at (6.5,1.5) {$c_7$};
+   %\node[] (c8) at (7.5,1.5) {$c_8^*$};
    \node[] (s4) at (7.5,1.5) {$\mathcal{S}$};
@@
    \node[] (c2) at (4,3) {$c_2$};
+   %\node[] (c4) at (5,3) {$c_4^*$};
    \node[] (s7) at (5,3) {$\mathcal{S}$};
@@
     (s6) edge node {} (c2)
+         %edge node {} (c4);
          edge node {} (s7)

### 2015-06-12 rapport/component.tex
rapport_12_06

@@
 \begin{itemize}
+\item as for any DSL, the user manipulates a high-abstraction level language;
 \item advantages of component models are inherited for the DSL itself.
@@
 
-Component-based software engineering~\cite{Szyperski:2002:CSB:515228} extends the concept of class by specifying in its interface not only the services it offers, or public methods, but also all its possible interactions with outer world, including the services it requires.
+Component-based software engineering~\cite{Szyperski:2002:CSB:515228} extends the concept of class by specifying in its interfaces not only the services it offers, or public methods, but also all its possible interactions with outer world, including the services it requires.
 As a result, each component is an independant entity composed of a set of services its provides, and a set of services it requires (and uses).
@@
 A \emph{port} is an entity embbedded in the component which makes possible an \emph{assembly} of components.
-An assembly of components (of instanciated components) is a way to actually produce a complete application by a set of components and their interactions.
-An interface of provided services of a component is associated to a \emph{provide} port, while an interface of required services is associated to a \emph{use} port.
+An assembly of components (of instanciated components) is a way to actually connect components together and to produce a complete application, as a set of components and their interactions.
+An interface of provided services of a component is associated to a \emph{provide} port, while required services are associated to \emph{use} ports.
 It is also possible to group more than one required interface into a single port, as a list, called a \emph{use-multiple} port.
@@
 Those components are called control-components because they are used in a different way than usual components.
-Actually, as explained before, a component represents a precise functionnality of the application, while control-components represent a way to control the execution time of other components.
-Each control-component contains a single provide port, this port is associated to a single method which correspond to the execution of the control-component.
+Actually, as explained before, a component represents a precise functionnality of the application, while control-components represent a way to control the execution of components.
+Each control-component contains a single provide port, this port is associated to a single method which corresponds to the execution of the control-component.
+
+\paragraph{Synchronization component.} A synchronization component, denoted $SYNC$, is a control-component for which the provided service is to garantee that an update of data is performed.
+A $SYNC$ component is thus composed of a single use-multiple port which will be connected to data to update.
 
@@
 To do so, an ordered use-multiple port is needed.
-
-\paragraph{Synchronized-sequence component.} A synchronized-sequence component, denoted $SSEQ$, is a control-component for which the provided service is to garantee that a set of data synchronizations is performed before a sequence of other required services.
-A $SSEQ$ component is thus composed of two use-multiple ports, one not necessarily ordered to link the $SSEQ$ to data to synchronize (could be more than one), and the second, ordered, to be connect $SSEQ$ to other services, as $SEQ$ does.
 
@@
 \captionsetup[subfigure]{labelformat=empty}
+
 \begin{center}
-%FIRST RULE
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (sync) at (1.5,0) {$SYNC$};
+   \node[provide] (p) at (0,0) {};
+   \node[use] (u) at (1.5,1) {$m$};
+
+  \path[->]
+    (p) edge node {} (sync)
+    (sync) edge node {} (u);
+  \end{tikzpicture}
+}
+\end{center}
+\hspace{50pt}
+
+\begin{center}
 \subfloat[]{
@@
   }
+  \end{center}
 \hspace{50pt}
-%ARROW
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (seq) at (0,0) {$SSEQ$};
-   \node[provide] (p) at (-1.5,0) {};
-   \node[use] (u1) at (1.5,0) {$m$};
-   \node[use] (u2) at (0,1) {$m$};
 
-  \path[->]
-    (p) edge node {} (seq)
-    (seq) edge node {} (u1)
-    	  edge node {} (u2);
-  \end{tikzpicture}
-  }
-\hspace{50pt}
-%ARROW
+\begin{center}
 \subfloat[]{
@@
 }
+\end{center}
 \caption{$SEQ$, $SSEQ$ and $PAR$ control-components.}
 \label{fig:control}
-\end{center}
 \end{figure}
 
+\medskip
 To be able to dump the series-parallel tree decomposition of $\Gamma_{hybrid}$ to a component assembly, each computation $c_i$ has to be associated to a component.
 Such a component contains a single provide port for which the service is to perform the computation.
-As a computation $c_i(R_i,w_i,D_i,\text{exp}_i)$ manipulates data $R$ and $w$, the component needs a use-multiple port to be linked to data.
+As a computation $c_i(R_i,w_i,\text{exp}_i)$ manipulates data $R$ and $w$, the component needs a use-multiple port to be linked to data.
 
@@
 
-Every use-multiple port of a control-component can be linked in the assembly to another control-component or to a computation component.
+Use-multiple ports of $SEQ$ and $PAR$ can be equally linked in the assembly other control-components or to computation components.
 
@@
 In the series-parallel binary tree decomposition of $\Gamma_{hybrid}$, $\mathcal{S}$ and $\mathcal{P}$ are binary operators.
-On the ther hand, definitions of $SEQ$, $SSEQ$ and $PAR$ are not binary, mainly because of HPC concerns.
-Actually, as less as the number of calls is, the better the performance is.
+However, those operators could also accept more than two operands, resulting in less functionnal or component calls in the final dump, and thus to better performance.
+Moreover, definitions of $SEQ$ and $PAR$ are not binary (use-multiple).
 In the work of Finta \& Al~\cite{Finta:1996:SUS:235643.235649}, a canonical form of the series-parallel binary tree decomposition has been introduced.
 The canonical form consists in merging two identical and successive compositions ($\mathcal{S}$ or $\mathcal{P}$) as a single one.
-In the case of $\Gamma_{hybrid}$, however, a particular case appear when $\mathcal{S}$ is the source node of a leaf of updates $*_i$.
+%In the case of $\Gamma_{hybrid}$, however, a particular case appear when $\mathcal{S}$ is the source node of a leaf of updates $*_i$.
 In this case, it is forbidden to merge two sequence compositions $\mathcal{S}$.
@@
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (s10) at (4,0.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,2) {$c_0$};
-   \node[] (s2) at (1.5,2) {$\mathcal{S}$};
-   \node[] (star1) at (1,3) {$*_1$};
-   \node[] (c1) at (2,3) {$c_1$};
-   \node[] (c7) at (6.5,2) {$c_7$};
-   \node[] (s4) at (7.5,2) {$\mathcal{S}$};
-   \node[] (star8) at (7,3) {$*_8$};
-   \node[] (c8) at (8,3) {$c_8$};
-   \node[] (p1) at (4,2) {$\mathcal{P}$};
-   \node[] (s5) at (3,2.5) {$\mathcal{S}$};
-   \node[] (c3) at (2.5,3.5) {$c_3$};
-   \node[] (c5) at (3.5,3.5) {$c_5$};
-   \node[] (s8) at (5,2.5) {$\mathcal{S}$};
-   \node[] (c6) at (6,3.5) {$c_6$};
-   \node[] (c2) at (4,3.5) {$c_2$};
-   \node[] (s7) at (5,3.5) {$\mathcal{S}$};
-   \node[] (star4) at (4.5,4.5) {$*_4$};
-   \node[] (c4) at (5.5,4.5) {$c_4$};
+   \node[] (s0) at (0,0) {$\mathcal{S}$};
+   \node[] (c0) at (-3,1) {$c_0$};
+   \node[] (star1) at (-2,1) {$*_1$};
+   \node[] (c1) at (-1,1) {$c_1^*$};
+
+   \node[] (p0) at (0,1) {$\mathcal{P}$};
+   \node[] (s1) at (-1,2) {$\mathcal{S}$};
+   \node[] (p1) at (-2,3) {$\mathcal{P}$};
+   \node[] (c4) at (-1,3) {$c_4^*$};
+   \node[] (c6) at (-0,3) {$c_6$};
+   \node[] (c2) at (-2.5,4) {$c_2$};
+   \node[] (star4) at (-1.5,4) {$*_4$};
+   \node[] (s2) at (1,2) {$\mathcal{S}$};
+   \node[] (c3) at (0.5,3) {$c_3$};
+   \node[] (c5) at (1.5,3) {$c_5$};
+
+   \node[] (c7) at (1,1) {$c_7$};
+   \node[] (star8) at (2,1) {$*_8$};
+   \node[] (c8) at (3,1) {$c_8^*$};
 
   \path[->]
-    (s10) edge node {} (c0)
+    (s0) edge node {} (c0)
+         edge node {} (star1)
+         edge node {} (c1)
+         edge node {} (p0)
+         edge node {} (c7)
+         edge node {} (star8)
+         edge node {} (c8)
+    (p0) edge node {} (s1)
          edge node {} (s2)
-    	  edge node {} (p1)
-    	  edge node {} (c7)
-    	  edge node {} (s4)
-    (s2) edge node {} (star1)
-         edge node {} (c1)
-    (s4) edge node {} (star8)
-         edge node {} (c8)
-    (p1) edge node {} (s5)
-         edge node {} (s8)
-    (s5) edge node {} (c3)
-         edge node {} (c5)
-    (s8) edge node {} (c2)
-    	 edge node {} (s7)
+    (s1) edge node {} (p1)
+         edge node {} (c4)
          edge node {} (c6)
-    (s7) edge node {} (star4)
-         edge node {} (c4);
+    (p1) edge node {} (c2)
+         edge node {} (star4)
+    (s2) edge node {} (c3)
+         edge node {} (c5);
   \end{tikzpicture}
@@
 From this canonical representation of the multi-stencil program $\mathcal{MSP}$, a direct dump to a component assembly is possible following the transformation rules of Figure~\ref{fig:dump}.
+One can note that this dump is possible because in the construction of the series-parallel tree decomposition, a node $*_i$ is by definition a leaf.
+As a result, $*_i$ can be mapped onto the control-component $SYNC$which has a single use-multiple port linked to the data (as the computation component).
 
@@
 }
-%SSEQ RULE
-\\
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (x) at (0,0) {$\mathcal{S}$};
-   \node[] (y) at (0,1) {$*_i$};
-
-  \path[->]
-    (x) edge node {} (y);
-  \end{tikzpicture}
-  }
-\hspace{50pt}
-\subfloat[]{
-$\Rightarrow$
-}
-\hspace{50pt}
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (seq) at (0,0) {$SSEQ$};
-   \node[provide] (p) at (-1.5,0) {};
-   \node[use] (u1) at (1.5,0) {$m$};
-   \node[use] (u2) at (0,1) {$m$};
-
-  \path[->]
-    (p) edge node {} (seq)
-    (seq) edge node {} (u1)
-    	  edge node {} (u2);
-  \end{tikzpicture}
-}
 \\
@@
 }
+%SYNC
+\\
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (y) at (0,0) {$*_i$};
+  \end{tikzpicture}
+  }
+\hspace{50pt}
+\subfloat[]{
+$\Rightarrow$
+}
+\hspace{50pt}
+\subfloat[]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (sync) at (0,0) {$SYNC$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (0,1) {$m$};
+
+  \path[->]
+    (p) edge node {} (sync)
+    (sync) edge node {} (u);
+  \end{tikzpicture}
+}
 \caption{Direct dump rules from the canonical form to the component assembly.}
@@
    %c0
-   \node[component] (c0) [right = 1.5cm of seq0u] {$c_0$};
+   \node[component] (c0) [right = 2.5cm of seq0u] {$c_0$};
    \node[use] (c0u) [above = 0.8cm of c0] {$m$};
-   %sseq0
-   \node[component] (sseq0) [below = 1.5cm of c0] {$SSEQ$};
-   \node[use] (sseq0u1) [below = 0.8cm of c0] {$m$};
-   \node[use] (sseq0u2) [right = 1cm of sseq0] {$m$};
+   \node[] (r0) [right = 0.5cm of c0u] {$R_0$};
+   %sync0
+   \node[component] (sync0) [below = 1.5cm of c0] {$SYNC$};
+   \node[use] (sync0u) [above = 0.8cm of sync0] {$m$};
+   \node[] (star1) [right = 0.5cm of sync0u] {$*_1$};
    %c1
-   \node[component] (c1) [right = 1cm of sseq0u2] {$c_1$};
-   \node[use] (c1u) [right = 1cm of c1] {$m$};
+   \node[component] (c1) [below = 1.5cm of sync0] {$c_1^*$};
+   \node[use] (c1u) [above = 0.8cm of c1] {$m$};
+   \node[] (r1) [right = 0.5cm of c1u] {$R_1$};
    %par0
-   \node[component] (par0) [below = 1cm of sseq0] {$PAR$};
+   \node[component] (par0) [below = 1cm of c1] {$PAR$};
    \node[use] (par0u) [right = 1cm of par0] {$m$};
+  %seq1
+   \node[component] (seq1) [right = 1cm of par0u] {$SEQ$};
+   \node[use] (seq1u) [right = 1cm of seq1] {$m$};
+   %par1
+   \node[component] (par1) [right = 1cm of seq1u] {$PAR$};
+   \node[use] (par1u) [right = 1cm of par1] {$m$};
+  %c2
+   \node[component] (c2) [right = 1.5cm of par1u] {$c_2$};
+   \node[use] (c2u) [above = 0.8cm of c2] {$m$};
+   \node[] (r2) [right = 0.5cm of c2u] {$R_2$};
+  %sync1
+   \node[component] (sync1) [below = 1.5cm of c2] {$SYNC$};
+   \node[use] (sync1u) [above = 0.8cm of sync1] {$m$};
+   \node[] (star4) [right = 0.5cm of sync1u] {$*_4$};
+  %c4
+   \node[component] (c4) [below = 1.5cm of par1] {$c_4^*$};
+   \node[use] (c4u) [above = 0.8cm of c4] {$m$};
+   \node[] (r4) [right = 0.5cm of c4u] {$R_4$};
+  %c6
+   \node[component] (c6) [below = 1.5cm of c4] {$c_6$};
+   \node[use] (c6u) [above = 0.8cm of c6] {$m$};
+   \node[] (r6) [right = 0.5cm of c6u] {$R_6$};
   %seq2
-   \node[component] (seq2) [right = 1cm of par0u] {$SEQ$};
+   \node[component] (seq2) [below = 4.5cm of seq1] {$SEQ$};
    \node[use] (seq2u) [right = 1cm of seq2] {$m$};
@@
    \node[use] (c3u) [above = 0.8cm of c3] {$m$};
+   \node[] (r3) [right = 0.5cm of c3u] {$R_3$};
   %c5
    \node[component] (c5) [below = 1.5cm of c3] {$c_5$};
-   \node[use] (c5u) [below = 0.8cm of c3] {$m$};
-   %seq3
-   \node[component] (seq3) [below = 3cm of seq2] {$SEQ$};
-   \node[use] (seq3u) [right = 1cm of seq3] {$m$};
-   %c2
-   \node[component] (c2) [right = 1cm of seq3u] {$c_2$};
-   \node[use] (c2u) [above = 0.8cm of c2] {$m$};
-   %sseq1
-   \node[component] (sseq1) [below = 1.5cm of c2] {$SSEQ$};
-   \node[use] (sseq1u1) [below = 0.8cm of c2] {$m$};
-   \node[use] (sseq1u2) [right = 1.5cm of sseq1] {$m$};
-   %c4
-   \node[component] (c4) [right = 1cm of sseq1u2] {$c_4$};
-   \node[use] (c4u) [above = 0.8cm of c4] {$m$};
-   %c6
-   \node[component] (c6) [below = 1.5cm of sseq1] {$c_6$};
-   \node[use] (c6u) [above = 0.8cm of c6] {$m$};
+   \node[use] (c5u) [above = 0.8cm of c5] {$m$};
+   \node[] (r5) [right = 0.5cm of c5u] {$R_5$};
   %c7
-   \node[component] (c7) [below = 5cm of par0] {$c_7$};
+   \node[component] (c7) [below = 1.5cm of par0] {$c_7$};
    \node[use] (c7u) [above = 0.8cm of c7] {$m$};
-   %sseq2
-   \node[component] (sseq2) [below = 1.5cm of c7] {$SSEQ$};
-   \node[use] (sseq2u1) [below = 0.8cm of c7] {$m$};
-   \node[use] (sseq2u2) [right = 1.5cm of sseq2] {$m$};
+   \node[] (r7) [right = 0.5cm of c7u] {$R_7$};
+  %sync2
+   \node[component] (sync2) [below = 1.5cm of c7] {$SYNC$};
+   \node[use] (sync2u) [above = 0.8cm of sync2] {$m$};
+   \node[] (star8) [right = 0.5cm of sync2u] {$*_8$};
   %c8
-   \node[component] (c8) [right = 1cm of sseq2u2] {$c_8$};
+   \node[component] (c8) [below = 1.5cm of sync2] {$c_8^*$};
    \node[use] (c8u) [above = 0.8cm of c8] {$m$};
+   \node[] (r8) [right = 0.5cm of c8u] {$R_8$};
 
@@
     (c0) edge node {} (c0u)
-   %seq1
-     (seq0u) edge node {} (sseq0.west)
-     (sseq0) edge node {} (sseq0u1)
-     (sseq0) edge node {} (sseq0u2)
+    %sync0
+    (seq0u) edge node {} (sync0.west)
+    (sync0) edge node {} (sync0u)
     %c1
-     (sseq0u2) edge node {} (c1)
+    (seq0u) edge node {} (c1)
     (c1) edge node {} (c1u)
@@
     (par0) edge node {} (par0u)
+  %seq1
+    (par0u) edge node {} (seq1)
+    (seq1) edge node {} (seq1u)
+  %par1
+    (seq1u) edge node {} (par1.west)
+    (par1) edge node {} (par1u)
+  %c2
+    (par1u) edge node {} (c2)
+    (c2) edge node {} (c2u)
+  %sync0
+    (par1u) edge node {} (sync1.west)
+    (sync1) edge node {} (sync1u)
+  %c4
+    (seq1u) edge node {} (c4.west)
+    (c4) edge node {} (c4u)
+  %c6
+    (seq1u) edge node {} (c6.west)
+    (c6) edge node {} (c6u)
   %seq2
@@
     (c3) edge node {} (c3u)
-   %c5
-    (seq2u) edge node {} (c5)
+  %c6
+    (seq2u) edge node {} (c5.west)
     (c5) edge node {} (c5u)
-   %seq3
-    (par0u) edge node {} (seq3.west)
-    (seq3) edge node {} (seq3u)
-   %c2
-    (seq3u) edge node {} (c2)
-    (c2) edge node {} (c2u)
-   %sseq1
-    (seq3u) edge node {} (sseq1.west)
-    (sseq1) edge node {} (sseq1u1)
-    (sseq1) edge node {} (sseq1u2)
-   %c4
-    (sseq1u2) edge node {} (c4)
-    (c4) edge node {} (c4u)
   %c6
-    (seq3u) edge node {} (c6)
-    (c6) edge node {} (c6u)
-   %c7
     (seq0u) edge node {} (c7.west)
     (c7) edge node {} (c7u)
-   %sseq2
-    (seq0u) edge node {} (sseq2.west)
-    (sseq2) edge node {} (sseq2u1)
-    (sseq2) edge node {} (sseq2u2)
+  %sync2
+    (seq0u) edge node {} (sync2.west)
+    (sync2) edge node {} (sync2u)
   %c8
-    (sseq2u2) edge node {} (c8)
-    (c8) edge node {} (c8u);
+    (seq0u) edge [bend right] node {} (c8.west)
+    (c8) edge node {} (c8u)
+   ;
 \end{tikzpicture}
@@
 However, there is not a unique way to design the rest of the assembly and to make accessible the different services (data, update, mesh etc.).
-The work presented until know can be considered as a generic solution to solve the parallel computation part, but the rest of the assembly is dependant of the implementation choice.
-However, the implementation presented in this work, and detailed in the next Section, follows the SIPSim model~\cite{} which seems to be an interesting solution to dump onto component models.
+The work presented until now can be considered as a generic solution to solve the parallel computation part, but the rest of the assembly is dependant of the implementation choice.
+However, the implementation presented in this work, and detailed in the next Section, is inspired by the SIPSim model~\cite{} which is an interesting \emph{implicit parallelism model} to dump onto component models.

### 2015-06-12 rapport/msp.tex
rapport_12_06

@@
    \node[] (c0) at (0,0) {$c_0$};
-   \node[] (c1) at (1,0) {$c_1^*$};
-   \node[] (c2) at (2,0.5) {$c_2$};
-   \node[] (c3) at (2,-0.5) {$c_3$};
-   \node[] (c4) at (3,0.5) {$c_4^*$};
-   \node[] (c5) at (3,-0.5) {$c_5$};
-   \node[] (c6) at (4,0.5) {$c_6$};
-   \node[] (c7) at (5,0) {$c_7$};
-   \node[] (c8) at (6,0) {$c_8^*$};
+   \node[] (star1) at (1,0) {$*_1$};
+   \node[] (c1) at (2,0) {$c_1^*$};
+   \node[] (c2) at (3,0.5) {$c_2$};
+   \node[] (star4) at (3,1.5) {$*_4$};
+   \node[] (c3) at (3,-0.5) {$c_3$};
+   \node[] (c4) at (4,0.5) {$c_4^*$};
+   \node[] (c5) at (4,-0.5) {$c_5$};
+   \node[] (c6) at (5,0.5) {$c_6$};
+   \node[] (c7) at (6,0) {$c_7$};
+   \node[] (star8) at (7,0) {$*_8$};
+   \node[] (c8) at (8,0) {$c_8^*$};
 
   \path[->]
-    (c0) edge node {} (c1)
+    (c0) edge node {} (star1)
+    (star1) edge node {} (c1)
     (c1) edge node {} (c2)
          edge node {} (c3)
+         edge node {} (star4)
+    (star4) edge node {} (c4)
     (c2) edge node {} (c4)
@@
     (c6) edge node {} (c7)
-    (c7) edge node {} (c8);
+    (c7) edge node {} (star8)
+    (star8) edge node {} (c8);
   \end{tikzpicture}
@@
    \node[circle,draw=black,fill=black,scale=0.3] (c0d) at (1,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (3,0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (3,-0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (4,0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c7d) at (6,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (star1d) at (2,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c7d) at (7,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (star8d) at (8,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
 
@@
     (c0s) edge node [above] {$c_0$} (c0d)
-    (c0d) edge node [above] {$c_1^*$} (c1d)
+    (c0d) edge node [above] {$*_1$} (star1d)
+    (star1d) edge node [above] {$c_1^*$} (c1d)
     (c1d) edge node [above] {$c_2$} (c2d)
           edge node [above] {$c_3$} (c3d)
+          edge [bend left=80] node [above] {$*_4$} (c2d)
     (c2d) edge node [above] {$c_4^*$} (c4d)
@@
     (c6d) edge node [above] {$c_7$} (c7d)
-    (c7d) edge node [above] {$c_8$} (c8d)
+    (c7d) edge node [above] {$c_8$} (star8d)
+    (star8d) edge node [above] {$c_8$} (c8d)
     (c3d) edge node [above] {$c_5$} (c6d);
@@
    \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (3,0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (3,-0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (4,0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
 
    %reduction c0 c1
-   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,1.5) {$c_0$};
-   %\node[] (c1) at (1.5,1.5) {$c_1^*$};
-   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
-   \node[] (star1) at (1,2.5) {$*_1$};
-   \node[] (c1) at (2,2.5) {$c_1$};
+   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
+   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,2.5) {$c_0$};
+   \node[] (star1) at (1.5,2.5) {$*_1$};
+   \node[] (c1) at (2,1.5) {$c_1^*$};
 
    %reduction c7 c8
-   \node[] (s3) at (6,0.5) {$\mathcal{S}$};
-   \node[] (c7) at (5.5,1.5) {$c_7$};
-   %\node[] (c8) at (6.5,1.5) {$c_8^*$};
-   \node[] (s4) at (6.5,1.5) {$\mathcal{S}$};
-   \node[] (star8) at (6,2.5) {$*_8$};
-   \node[] (c8) at (7,2.5) {$c_8$};
+   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
+   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
+   \node[] (c7) at (6.5,2.5) {$c_7$};
+   \node[] (star8) at (7.5,2.5) {$*_8$};
+   \node[] (c8) at (8,1.5) {$c_8$};
 
@@
           edge node [above] {$c_3$} (c3d)
+          edge [bend left=80] node [above] {$*_4$} (c2d)
     (c2d) edge node [above] {$c_4^*$} (c4d)
@@
     %reduction c0 c1
+    (s0) edge node {} (s1)
+         edge node {} (c1)
     (s1) edge node {} (c0)
-    %     edge node {} (c1)
-        edge node {} (s2)
-    (s2) edge node {} (star1)
-        edge node {} (c1)
+         edge node {} (star1)
     %reduction c7 c8
+    (s2) edge node {} (s3)
+         edge node {} (c8)
     (s3) edge node {} (c7)
-    %     edge node {} (c8);
-        edge node {} (s4)
-    (s4) edge node {} (star8)
-        edge node {} (c8);
+        edge node {} (star8);
   \end{tikzpicture}
@@
    \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (4,0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
 
    %reduction c0 c1
-   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,1.5) {$c_0$};
-   %\node[] (c1) at (1.5,1.5) {$c_1^*$};
-   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
-   \node[] (star1) at (1,2.5) {$*_1$};
-   \node[] (c1) at (2,2.5) {$c_1$};
+   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
+   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,2.5) {$c_0$};
+   \node[] (star1) at (1.5,2.5) {$*_1$};
+   \node[] (c1) at (2,1.5) {$c_1^*$};
 
    %reduction c7 c8
-   \node[] (s3) at (6,0.5) {$\mathcal{S}$};
-   \node[] (c7) at (5.5,1.5) {$c_7$};
-   %\node[] (c8) at (6.5,1.5) {$c_8^*$};
-   \node[] (s4) at (6.5,1.5) {$\mathcal{S}$};
-   \node[] (star8) at (6,2.5) {$*_8$};
-   \node[] (c8) at (7,2.5) {$c_8$};
+   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
+   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
+   \node[] (c7) at (6.5,2.5) {$c_7$};
+   \node[] (star8) at (7.5,2.5) {$*_8$};
+   \node[] (c8) at (8,1.5) {$c_8$};
 
    %reduction c3 c5
-   \node[] (s5) at (3.5,-2) {$\mathcal{S}$};
-   \node[] (c3) at (3,-1) {$c_3$};
-   \node[] (c5) at (4,-1) {$c_5$};
+   \node[] (s5) at (4.5,-2) {$\mathcal{S}$};
+   \node[] (c3) at (4,-1) {$c_3$};
+   \node[] (c5) at (5,-1) {$c_5$};
 
-   %reduction c2 c4
-   \node[] (s6) at (3.5,1) {$\mathcal{S}$};
+   %reduction c2 *4
+   \node[] (p0) at (3.5,1) {$\mathcal{P}$};
    \node[] (c2) at (3,2) {$c_2$};
-   %\node[] (c4) at (4,2) {$c_4^*$};
-   \node[] (s7) at (4,2) {$\mathcal{S}$};
-   \node[] (star4) at (3.5,3) {$*_4$};
-   \node[] (c4) at (4.5,3) {$c_4$};
+   \node[] (star4) at (4,2) {$*_4$};
 
@@
     (c0s) edge node {} (c1d)
-    (c1d) edge [bend left] node [above] {} (c4d)
+    (c1d) edge node [above] {} (c2d)
+    (c2d) edge node [above] {$c_4^*$} (c4d)
     (c4d) edge node [above] {$c_6$} (c6d)
@@
     %reduction c0 c1
+    (s0) edge node {} (s1)
+         edge node {} (c1)
     (s1) edge node {} (c0)
-    %     edge node {} (c1)
-         edge node {} (s2)
-    (s2) edge node {} (star1)
-         edge node {} (c1)
+         edge node {} (star1)
     %reduction c7 c8
+    (s2) edge node {} (s3)
+         edge node {} (c8)
     (s3) edge node {} (c7)
-    %     edge node {} (c8)
-         edge node {} (s4)
-    (s4) edge node {} (star8)
-         edge node {} (c8)
+        edge node {} (star8)
 
@@
 
-    %reduction c7 c8
-    (s6) edge node {} (c2)
-    %     edge node {} (c4);
-         edge node {} (s7)
-    (s7) edge node {} (star4)
-         edge node {} (c4);
+    %reduction c2 *4
+    (p0) edge node {} (c2)
+         edge node {} (star4);
   \end{tikzpicture}
@@
    \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
 
    %reduction c0 c1
-   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,1.5) {$c_0$};
-   %\node[] (c1) at (1.5,1.5) {$c_1^*$};
-   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
-   \node[] (star1) at (1,2.5) {$*_1$};
-   \node[] (c1) at (2,2.5) {$c_1$};
+   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
+   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,2.5) {$c_0$};
+   \node[] (star1) at (1.5,2.5) {$*_1$};
+   \node[] (c1) at (2,1.5) {$c_1^*$};
 
    %reduction c7 c8
-   \node[] (s3) at (6,0.5) {$\mathcal{S}$};
-   \node[] (c7) at (5.5,1.5) {$c_7$};
-   %\node[] (c8) at (6.5,1.5) {$c_8^*$};
-   \node[] (s4) at (6.5,1.5) {$\mathcal{S}$};
-   \node[] (star8) at (6,2.5) {$*_8$};
-   \node[] (c8) at (7,2.5) {$c_8$};
+   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
+   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
+   \node[] (c7) at (6.5,2.5) {$c_7$};
+   \node[] (star8) at (7.5,2.5) {$*_8$};
+   \node[] (c8) at (8,1.5) {$c_8$};
 
    %reduction c3 c5
-   \node[] (s5) at (3.5,-2) {$\mathcal{S}$};
-   \node[] (c3) at (3,-1) {$c_3$};
-   \node[] (c5) at (4,-1) {$c_5$};
+   \node[] (s5) at (4.5,-2) {$\mathcal{S}$};
+   \node[] (c3) at (4,-1) {$c_3$};
+   \node[] (c5) at (5,-1) {$c_5$};
 
    %reduction c2 c6
-   \node[] (s8) at (3.5,1) {$\mathcal{S}$};
-   \node[] (c6) at (4,2) {$c_6$};
-   \node[] (s6) at (3,2) {$\mathcal{S}$};
-   \node[] (c2) at (2.5,3) {$c_2$};
-   %\node[] (c4) at (3.5,3) {$c_4^*$};
-   \node[] (s7) at (3.5,3) {$\mathcal{S}$};
-   \node[] (star4) at (3,4) {$*_4$};
-   \node[] (c4) at (4,4) {$c_4$};
+   \node[] (s6) at (4.5,1) {$\mathcal{S}$};
+   \node[] (p0) at (3.5,2) {$\mathcal{P}$};
+   \node[] (c2) at (3,3) {$c_2$};
+   \node[] (star4) at (4,3) {$*_4$};
+   \node[] (s7) at (5.5,2) {$\mathcal{S}$};
+   \node[] (c4) at (5,3) {$c_4^*$};
+   \node[] (c6) at (6,3) {$c_6$};
 
@@
     %reduction c2 c6
-    (s8) edge node {} (s6)
-         edge node {} (c6)
-    (s6) edge node {} (c2)
-    %     edge node {} (c4);
+    (s6) edge node {} (p0)
          edge node {} (s7)
-    (s7) edge node {} (star4)
-         edge node {} (c4);
+    %reduction c2 *4
+    (p0) edge node {} (c2)
+         edge node {} (star4)
+    (s7) edge node {} (c4)
+         edge node {} (c6);
   \end{tikzpicture}
@@
    \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (8,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (7,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
 
    %reduction c0 c1
-   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,1.5) {$c_0$};
-   %\node[] (c1) at (1.5,1.5) {$c_1^*$};
-   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
-   \node[] (star1) at (1,2.5) {$*_1$};
-   \node[] (c1) at (2,2.5) {$c_1$};
+   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
+   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,2.5) {$c_0$};
+   \node[] (star1) at (1.5,2.5) {$*_1$};
+   \node[] (c1) at (2,1.5) {$c_1^*$};
 
    %reduction c7 c8
-   \node[] (s3) at (7,0.5) {$\mathcal{S}$};
-   \node[] (c7) at (6.5,1.5) {$c_7$};
-   %\node[] (c8) at (7.5,1.5) {$c_8^*$};
-   \node[] (s4) at (7.5,1.5) {$\mathcal{S}$};
-   \node[] (star8) at (7,2.5) {$*_8$};
-   \node[] (c8) at (8,2.5) {$c_8$};
+   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
+   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
+   \node[] (c7) at (6.5,2.5) {$c_7$};
+   \node[] (star8) at (7.5,2.5) {$*_8$};
+   \node[] (c8) at (8,1.5) {$c_8$};
 
-   %reduction c2 c4 c6 c3 c5
-   \node[] (p1) at (4,0.5) {$\mathcal{P}$};
-   \node[] (s5) at (3,1) {$\mathcal{S}$};
-   \node[] (c3) at (2.5,2) {$c_3$};
-   \node[] (c5) at (3.5,2) {$c_5$};
-   \node[] (s8) at (5,1) {$\mathcal{S}$};
-   \node[] (c6) at (5.5,2) {$c_6$};
-   \node[] (s6) at (4.5,2) {$\mathcal{S}$};
-   \node[] (c2) at (4,3) {$c_2$};
-   %\node[] (c4) at (5,3) {$c_4^*$};
-   \node[] (s7) at (5,3) {$\mathcal{S}$};
-   \node[] (star4) at (4.5,4) {$*_4$};
-   \node[] (c4) at (5.5,4) {$c_4$};
+   \node[] (p1) at (4.5,0.5) {$\mathcal{P}$};
+   %reduction c3 c5
+   \node[] (s5) at (5.5,1.5) {$\mathcal{S}$};
+   \node[] (c3) at (5,2.5) {$c_3$};
+   \node[] (c5) at (6,2.5) {$c_5$};
+   %reduction c2 c6
+   \node[] (s6) at (3.5,1.5) {$\mathcal{S}$};
+   \node[] (p0) at (2.8,2.5) {$\mathcal{P}$};
+   \node[] (c2) at (2.3,3.5) {$c_2$};
+   \node[] (star4) at (3.3,3.5) {$*_4$};
+   \node[] (s7) at (4.2,2.5) {$\mathcal{S}$};
+   \node[] (c4) at (3.8,3.5) {$c_4^*$};
+   \node[] (c6) at (4.8,3.5) {$c_6$};
 
@@
     %reduction c0 c1
+    (s0) edge node {} (s1)
+         edge node {} (c1)
     (s1) edge node {} (c0)
-    %     edge node {} (c1)
-         edge node {} (s2)
-    (s2) edge node {} (star1)
-         edge node {} (c1)
+         edge node {} (star1)
     %reduction c7 c8
+    (s2) edge node {} (s3)
+         edge node {} (c8)
     (s3) edge node {} (c7)
-    %     edge node {} (c8)
-         edge node {} (s4)
-    (s4) edge node {} (star8)
-         edge node {} (c8)
+        edge node {} (star8)
 
-    %reduction c2 c4 c6 c3 c5
-    (p1) edge node {} (s5)
-         edge node {} (s8)
+    (p1) edge node {} (s6)
+         edge node {} (s5)
+    %reduction c3 c5
     (s5) edge node {} (c3)
          edge node {} (c5)
-    (s8) edge node {} (s6)
-         edge node {} (c6)
-    (s6) edge node {} (c2)
-         %edge node {} (c4);
+    %reduction c2 c6
+    (s6) edge node {} (p0)
          edge node {} (s7)
-    (s7) edge node {} (star4)
-         edge node {} (c4);
+    %reduction c2 *4
+    (p0) edge node {} (c2)
+         edge node {} (star4)
+    (s7) edge node {} (c4)
+         edge node {} (c6);
   \end{tikzpicture}
@@
    \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (8,0) {};
+   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
 
-   %all
-   \node[] (s9) at (2.5,1.5) {$\mathcal{S}$};
-   \node[] (s10) at (4,0.5) {$\mathcal{S}$};
+   \node[] (s8) at (4.5,0.5) {$\mathcal{S}$};
+   \node[] (s9) at (3,1.5) {$\mathcal{S}$};
 
    %reduction c0 c1
-   \node[] (s1) at (1,2) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,3) {$c_0$};
-   \node[] (s2) at (1.5,3) {$\mathcal{S}$};
-   \node[] (star1) at (1,4) {$*_1$};
-   \node[] (c1) at (2,4) {$c_1$};
+   \node[] (s0) at (1.5,2.5) {$\mathcal{S}$};
+   \node[] (s1) at (1,3.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,4.5) {$c_0$};
+   \node[] (star1) at (1.5,4.5) {$*_1$};
+   \node[] (c1) at (2,3.5) {$c_1^*$};
 
    %reduction c7 c8
-   \node[] (s3) at (7,1) {$\mathcal{S}$};
-   \node[] (c7) at (6.5,2) {$c_7$};
-   \node[] (s4) at (7.5,2) {$\mathcal{S}$};
-   \node[] (star8) at (7,3) {$*_8$};
-   \node[] (c8) at (8,3) {$c_8$};
+   \node[] (s2) at (7.5,1.5) {$\mathcal{S}$};
+   \node[] (s3) at (7,2.5) {$\mathcal{S}$};
+   \node[] (c7) at (6.5,3.5) {$c_7$};
+   \node[] (star8) at (7.5,3.5) {$*_8$};
+   \node[] (c8) at (8,2.5) {$c_8$};
 
-   %reduction c2 c4 c6 c3 c5
-   \node[] (p1) at (4,2) {$\mathcal{P}$};
-   \node[] (s5) at (3,2.5) {$\mathcal{S}$};
-   \node[] (c3) at (2.5,3.5) {$c_3$};
-   \node[] (c5) at (3.5,3.5) {$c_5$};
-   \node[] (s8) at (5,2.5) {$\mathcal{S}$};
-   \node[] (c6) at (5.5,3.5) {$c_6$};
-   \node[] (s6) at (4.5,3.5) {$\mathcal{S}$};
-   \node[] (c2) at (4,4.5) {$c_2$};
-   \node[] (s7) at (5,4.5) {$\mathcal{S}$};
-   \node[] (star4) at (4.5,5.5) {$*_4$};
-   \node[] (c4) at (5.5,5.5) {$c_4$};
+   \node[] (p1) at (4.5,2.5) {$\mathcal{P}$};
+   %reduction c3 c5
+   \node[] (s5) at (5.5,3.5) {$\mathcal{S}$};
+   \node[] (c3) at (5,4.5) {$c_3$};
+   \node[] (c5) at (6,4.5) {$c_5$};
+   %reduction c2 c6
+   \node[] (s6) at (3.5,3.5) {$\mathcal{S}$};
+   \node[] (p0) at (2.8,4.5) {$\mathcal{P}$};
+   \node[] (c2) at (2.3,5.5) {$c_2$};
+   \node[] (star4) at (3.3,5.5) {$*_4$};
+   \node[] (s7) at (4.2,4.5) {$\mathcal{S}$};
+   \node[] (c4) at (3.8,5.5) {$c_4^*$};
+   \node[] (c6) at (4.8,5.5) {$c_6$};
 
@@
 
+    (s8) edge node {} (s9)
+         edge node {} (s2)
+    (s9) edge node {} (p1)
+         edge node {} (s0)
+
     %reduction c0 c1
+    (s0) edge node {} (s1)
+         edge node {} (c1)
     (s1) edge node {} (c0)
-         edge node {} (s2)
-    (s2) edge node {} (star1)
-         edge node {} (c1)
+         edge node {} (star1)
     %reduction c7 c8
+    (s2) edge node {} (s3)
+         edge node {} (c8)
     (s3) edge node {} (c7)
-         edge node {} (s4)
-    (s4) edge node {} (star8)
-         edge node {} (c8)
+        edge node {} (star8)
 
-    %reduction c2 c4 c6 c3 c5
-    (p1) edge node {} (s5)
-         edge node {} (s8)
+    (p1) edge node {} (s6)
+         edge node {} (s5)
+    %reduction c3 c5
     (s5) edge node {} (c3)
          edge node {} (c5)
-    (s8) edge node {} (s6)
-         edge node {} (c6)
-    (s6) edge node {} (c2)
+    %reduction c2 c6
+    (s6) edge node {} (p0)
          edge node {} (s7)
-    (s7) edge node {} (star4)
-         edge node {} (c4)
-
-    %all
-    (s9) edge node {} (s1)
-         edge node {} (p1)
-    (s10) edge node {} (s9)
-          edge node {} (s3);
+    %reduction c2 *4
+    (p0) edge node {} (c2)
+         edge node {} (star4)
+    (s7) edge node {} (c4)
+         edge node {} (c6);
   \end{tikzpicture}

### 2015-06-12 rapport/parallelism.tex
rapport_12_06

@@
 %------------------------------
-\subsection{Coarse-grain data parallelism}
+\subsection{Data parallelism}
 \label{sect:dataparal}
@@
 
-More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
+More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
 This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
-As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\{\Delta_0,\dots,\Delta_{p-1}\}$.
+As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\Delta=\{\Delta_0,\dots,\Delta_{p-1}\}$.
 The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma_{data}$.
@@
 \begin{mydef}
-For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j=s_j(R_j,w_j,D_j,\text{exp}_j)$, and $w_i \subset R_j$.
-Moreover, the data to update is $w_i \cap R_j = w_i$.
+For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j=s_j(R_j,w_j,\text{exp}_j)$, and $\{w_i\} \subset R_j$.
+Moreover, the data to update is $\{w_i\} \cap R_j = w_i$.
 \end{mydef}
@@
 \begin{mydef}
-A synchronization between two computations $c_i \pprec c_j$ is defined as a specific computation $\text{update}(w_i \cap R_j)=\text{update}(w_i)$.
+A synchronization between two computations $c_i \pprec c_j$ is defined as a specific computation $\text{update}(\{w_i\} \cap R_j)=\text{update}(w_i)$.
 \end{mydef}
@@
 \begin{mydef}
-An \textit{updated-computation} a computation which needs a data synchronization before it can be performed.
-If $c_i \pprec c_j$, with $i<j$, the computation $c_j$ is transformed to an updated-computation $c^*_j(R_j,w_j,D_j,\text{exp}_j,\text{update}(w_i))$ such that $\text{update}(w_i)$ is performed before the evaluation of $exp_j$.
-The updated-computation can also be denoted as $c^*_j(c_j,\text{update}(w_i))$.
+An \textit{updated-computation} is a computation which needs a set of data synchronizations before it can be performed.
 \end{mydef}
+
+\begin{mydef}
+If $c_j$ is an \textit{updated-computation}, it is replaced in $\Gamma_{data}$ by the ordered list of two elements
+\begin{equation*}
+[*_j(R^*_j,w^*_j),c^*_j(R_j \cup \{w^*_j\},w_j,\text{exp}_j)].
+\end{equation*}
+The computation $*_j$ is defined as the pair $*_j(R^*_j,w^*_j)$, where $R^*_j = \{\bigcup_{i} \{w_i\} \text{, }\forall i<j\text{, and }c_i \pprec c_j\}$, and $w^*_j$ is a dummy data which does not represent a quantity of the $\mathcal{MSP}$.
+This dummy $w^*_j$ is used, however, to keep a relation between the update phase $*_j$ and the computation $c_j$.
+For this reason, $c_j$ is replaced by $c^*_j$ where data read are modified to $R_j \cup \{w^*_j\}$.
+Finally, $*_j$ performs the computation $update(R^*_j)$ and $R^*_j \subseteq R_j \subseteq \Delta$.
+\end{mydef}
+
+ %$c^*_j(R_j,w_j,\text{exp}_j,*_j)$ if $\bigcup_{i} \{w_i\} \neq \emptyset \text{, }\forall i<j\text{, and }c_i \pprec c_j$.
+In such a case $*_j$ is itself a computation wich performs $update(\bigcup_{i} \{w_i\})$.
 
@@
 T_{data}(\Gamma,i) = 
-\begin{cases} 	[\Gamma[i]] \cdot T_{data}(\Gamma,i+1) & \mbox{if }\forall j\leq i\mbox{, } c_j \not\pprec c_i\\
-				[c^*_i(\Gamma[i],\text{update}(\{w_k\})] \cdot T_{data}(\Gamma,i+1) & \forall k < i\mbox{, }c_k \pprec c_i\\
+\begin{cases} 	[\Gamma[i]] \cdot T_{data}(\Gamma,i+1) & \mbox{if }\forall j< i\mbox{, } c_j \not\pprec c_i\\
+				[*_i(R^*_i,w^*_i),\Gamma[i]^*(R_i \cup \{w^*_i\},w_i,\text{exp}_i)] \cdot T_{data}(\Gamma,i+1) & \text{where } R^*_i = \{\bigcup_{k} \{w_k\} \text{, }\\
+        & \forall k<i\text{, and }c_k \pprec c_i\}\\
 				[] & \mbox{if }i=|\Gamma|.
@@
 
-The construction of $\Gamma_{data}$ implies that if a synchronization is needed it is always proceeded before the computation wich needs it.
-However, an optimization is possible to combine synchronizations together, and thus to synchronize more than one data at a time.
-If the number of synchronizations is reduced, final performance of the program will be better.
-For this reason, we define a reduction to apply to $\Gamma_{data}$, and which is performed as a post-transformation.
+% The construction of $\Gamma_{data}$ implies that if a synchronization is needed it is always proceeded before the computation wich needs it.
+One can notice that a combination of synchronizations is possible in some cases, and thus it is possible to synchronize more than one data at a time.
+%If the number of synchronizations is reduced, final performance of the program will be better.
+% For this reason, we define a reduction which can be applied to $\Gamma_{data}$, and which is performed as a post-transformation.
+This reduction, could propose better performance depending on the used network speed.
 
-\begin{myprop}
-Denoting four computations $c_l$, $c_k$, $c_i$ and $c_j$ of $\Gamma$, with $l<k$, $i<j$, and where $c_l \pprec c_k$ and $c_i \pprec c_j$.
-Then, computations $c_k$ and $c_j$ are tansformed to $c^*_k(c_k,update(w_l)$ and $c^*_j(c_j,update(w_i)$.
-If $i<k$, a single update call can be performed at $k$ such that $c^*_k(c_k,update(w_l\cup w_i)$, and in this case, the computation $c_j$ is not transformed.
-\end{myprop}
+% \begin{myprop}
+% Denoting four computations $c_l$, $c_k$, $c_i$ and $c_j$ of $\Gamma$, with $l<k$, $i<j$, and where $c_l \pprec c_k$ and $c_i \pprec c_j$.
+Then, computations $c_k$ and $c_j$ are tansformed to $c^*_k(c_k,update(w_l))$ and $c^*_j(c_j,update(w_i))$.
+If $i<k$, a single update call can be performed at $k$ such that $c^*_k(c_k,update(w_l\cup w_i))$, and in this case, the computation $c_j$ is not transformed.
+% \end{myprop}
 
-\begin{proof}
-By definition, the synchronization of $w_i$ can be performed between $c_i$ and $c_j$.
+% \begin{proof}
+% By definition, the synchronization of $w_i$ can be performed between $c_i$ and $c_j$.
 As $i<k$ the computation $c_i$ is not performed between $k$ and $j$.
 As a result, the synchronization of $w_i$ can be performed between $c_i$ and $c_k$.
-\end{proof}
+% \end{proof}
 
-In other words, for each synchronization in $\Gamma_{data}$, if the computation $c_i$ from which $c_j$ is synchronized occurs before the last synchronization, the two synchronizations can be reduced to one with the union of the data to update.
+% In other words, for each synchronization in $\Gamma_{data}$, if the computation $c_i$ from which $c_j$ is synchronized occurs before the last synchronization, the two synchronizations can be reduced to one with the union of the data to update.
 
- The final step of this parallelization is to run $\Gamma_{data}$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program defined by
+ The final step of this parallelization is to run $\Gamma_{data}$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program, defined by
 \begin{equation}
@@
 
-We denote this parallelization technique a coarse-grain data parallelization in contrast with the same technique applied to the finer level of a single computation.
+%We denote this parallelization technique a coarse-grain data parallelization in contrast with the same technique applied to the finer level of a single computation.
 In this case, for a computation $c(R,w,D,\text{exp})$, the local domain $D$ is partitionned for $p$ resources $\{D_0,\dots,D_{p-1}\}$, and each resource $k$ is responsible for a sub-computation $c_k(R,w,D_k,\text{exp}')$.
-%This finer data parallelization is particularly adapted to shared memory architectures (multi-cores, many-cores or GPUs), and in such case it does not require the addition of synchronizations because the computed element is local to the sub-domain $D_k$.
-Actually, if the computation is a stencil, the computation only read neighborhood elements but does not write them which does not require a synchronization, and if the computation is local nothing is read or written in the sub-domain of another resource.
-This finer data parallelization technique is not directly adressed by the work presented in this paper, but is exhibited, as it will be explained later.
+%This finer data parallelization technique is not directly adressed by the work presented in this paper, but is exhibited, as it will be explained later.
 
@@
 A dependency graph exhibits parallel tasks, or on the contrary sequential execution of tasks.
-In a multi-stencil program a task is defined as a computation $c(R,w,D,\text{exp})$.
+In a multi-stencil program a task is defined as a computation $c(R,w,\text{exp})$.
 As a result, the dependency graph of a multi-stencil program represents the dependencies between local and stencil computations of $\Gamma$.
@@
 \begin{mydef}
-For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $i < j$, it is said that $c_j$ is data dependant from $c_i$, denoted $c_i\prec c_j$, if $w_i \cap R_j \neq \emptyset$.
+For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, with $i < j$, it is said that $c_j$ is data dependant from $c_i$, denoted $c_i\prec c_j$, if $\{w_i\} \cap R_j \neq \emptyset$.
 In this case, $c_i$ has to be computed before $c_j$.
@@
 \begin{proof}
-Considering three computations $c_0(R_0,w_0,D_0,e_0)$, $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, where $w_0\cap R_1 \neq \emptyset$ and $w_1\cap R_2 \neq \emptyset$.
+Considering three computations $c_0(R_0,w_0,\text{exp}_0)$, $c_1(R_1,w_1,\text{exp}_1)$ and $c_2(R_2,w_2,\text{exp}_2)$, where $\{w_0\}\cap R_1 \neq \emptyset$ and $\{w_1\}\cap R_2 \neq \emptyset$.
 In this case two dependencies can be extracted $c_0 \prec c_1$ and $c_1 \prec c_2$.
-However, the dependency $c_0 \prec c_2$ is not true as $w_0\cap R_2 = \emptyset$.
+However, the dependency $c_0 \prec c_2$ is not true as $\{w_0\}\cap R_2 = \emptyset$.
 \end{proof}
@@
 This property is due to the fact that $\prec$ is defined as a data dependency between computations.
-A dependency chain, however, creates another type of dependency only due to time.
+However, another type of dependency, only due to time, is created from $\prec$.
 
 \begin{mydef}
-For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, a time dependency is denoted $c_i \blacktriangleleft c_j$ and means that $c_i$ has to be computed before $c_j$.
+For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, a time dependency is denoted $c_i \blacktriangleleft c_j$ and means that $c_i$ has to be computed before $c_j$.
 \end{mydef}
@@
 \begin{myprop}
-For two computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$ such that $c_i \prec c_j$, $c_i \blacktriangleleft c_j$ is verified.
+For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$ such that $c_i \prec c_j$, $c_i \blacktriangleleft c_j$ is verified.
 \end{myprop}
@@
 \begin{proof}
-By definition $c_i \prec c_j$ means that $w_i\cap R_j \neq \emptyset$ and that $c_i$ has to computed before $c_j$. %Thus, the relation $c_i \prec c_j$ is stronger than $c_i \blacktriangleleft c_j$.
+By definition $c_i \prec c_j$ means that $\{w_i\}\cap R_j \neq \emptyset$ and that $c_i$ has to computed before $c_j$.
+%Thus, the relation $c_i \prec c_j$ is stronger than $c_i \blacktriangleleft c_j$.
 \end{proof}
@@
 \begin{proof}
-Considering three computations $c_0(R_0,w_0,D_0,e_0)$, $c_1(R_1,w_1,D_1,e_1)$ and $c_2(R_2,w_2,D_2,e_2)$, where $c_0 \blacktriangleleft c_1$ and $c_1 \blacktriangleleft c_2$.
+Considering three computations $c_0(R_0,w_0,\text{exp}_0)$, $c_1(R_1,w_1,\text{exp}_1)$ and $c_2(R_2,w_2,\text{exp}_2)$, where $c_0 \blacktriangleleft c_1$ and $c_1 \blacktriangleleft c_2$.
 $c_0$ is computed before $c_1$ and $c_1$ is computed before $c_2$, as a result $c_0$ is computed before $c_2$ and the relation $c_0 \blacktriangleleft c_2$ is verified.
@@
 \begin{mydef}
-A directed acyclic graph (DAG) $G(V,A)$ is a graph where the edges are directed from a source to a destination vertex, and where following the direction of edges, no cycle can be found from a vertex $u$ to itself.
+A directed acyclic graph (DAG) $G(V,A)$ is a graph where the edges are directed from a source to a destination vertex and where, following the direction of edges, no cycle can be found from a vertex $u$ to itself.
 A directed edge is called an arc, and for two vertices $v,u \in V$ an arc from $u$ to $v$ is denoted $(\overset{\frown}{u,v}) \in A$.
@@
 
-From an ordered list of computations, a directed dependency graph $\Gamma_{task}(V,A)$ can be built finding all pairs of computations $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $i<j$, such that $c_i \blacktriangleleft c_j$.
+From an ordered list of computations, a directed dependency graph $\Gamma_{task}(V,A)$ can be built finding all pairs of computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, with $i<j$, such that $c_i \blacktriangleleft c_j$.
 Those time dependencies can be found from pairs of data dependencies $c_i \prec c_j$.
@@
 \begin{mydef}
-From an ordered list $\Gamma$ of computations $c(R,w,D,\text{exp})$, a directed dependency graph $\Gamma_{task}(V,A)$ is obtained from the call $T_{task}(\Gamma,0)$, where $T_{task}$ is the recursive function
+From an ordered list $\Gamma$ of computations $c(R,w,\text{exp})$, a directed dependency graph $\Gamma_{task}(V,A)$ is obtained from the call $T_{task}(\Gamma,0)$, where $T_{task}$ is the recursive function
 \begin{equation*}
@@
 Using the function $T_{task}$ to build $\Gamma_{task}$, however, duplication of dependencies may occur because of the transitivity of the relation $\blacktriangleleft$.
-Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,D_k,e_k)$, $c_i(R_i,w_i,D_i,e_i)$ and $c_j(R_j,w_j,D_j,e_j)$, with $k<i<j$ and where $c_k \prec c_i$, $c_i \prec c_j$ and $c_k \prec c_j$, then the transitivity $c_k \blacktriangleleft c_j$ is verified because of the relations $c_k \blacktriangleleft c_i$ and $c_i \blacktriangleleft c_j$.
+Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,\text{exp}_k)$, $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, with $k<i<j$ and where $c_k \prec c_i$, $c_i \prec c_j$ and $c_k \prec c_j$, then the transitivity $c_k \blacktriangleleft c_j$ is verified because of the relations $c_k \blacktriangleleft c_i$ and $c_i \blacktriangleleft c_j$.
 However, $c_k \blacktriangleleft c_j$ is also directly represented by the binary relation $c_k \prec c_j$.
@@
 \subsection{Hybrid parallelism}
-It is also possible to combine coarse-grain data and task parallelization techniques to get hybrid parallelism, sometimes more efficient on hybrid architectures.
-A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list of computations $c$ and updated-computations $c^*$.
+It is also possible to combine data and task parallelization techniques to get hybrid parallelism, sometimes more efficient on hybrid architectures.
+A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list of computations $c_i$ and updated-computations $*_i$.
 On the other hand, the task parallelization technique builds, from an ordered list of computation, a graph dependency $\Gamma_{task}$.
 As a result $T_{data}$ and $T_{task}$ can be composed to build $\Gamma_{hybrid}$.
-Thus, $k$ multi-stencil programs
+
+However, we have to clarify the dependency operator $\prec$ for the specific case of updated-computations.
+As explained in Section~\ref{sect:dataparal}, if $c_j$ is an updated-computations, which means that at least one $c_i$ exists such that $i<j$ and $c_i \pprec c_j$, $c_j$ is transformed to the ordered list of two computations
+
 \begin{equation*}
-\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{hybrid})
+[*_j(R^*_j,w^*_j),c^*_j(R_j \cup \{w^*_j\},w_j,\text{exp}_j)].
 \end{equation*}
-are responsible for an hybrid parallelization of a multi-stencil program.
-The set of computation $\Gamma_{hybrid}$ is a dependency graph between computations $c$ and updated-computations $c^*$ and can be built from the call to
+
+As a result, and because $*_j$ is itself a computation, it appears dependencies relations.
+First, by definition, the dependency $c^*_j \prec *_j$ is created because $\{w^*_j\} \subset R_j \cup \{w^*_j\}$.
+Second, as $R^*_j = \{\bigcup_{i} \{w_i\} \text{, }\forall i<j\text{, and }c_i \pprec c_j\}$, for each $i<j$ the dependency $c_i \prec *_j$ is also created.
+
+The choice of separating $*_j$ from $c_j$ has been done to entirely expose the eventual task parallelism in the hybrid parallelization.
+For example, Figure~\ref{fig:more} give an exemple where initially $c_0 \pprec c_2$ and $c_0 \prec c_1$, $c_1 \prec c_2$.
+Thus, $c_2$ is replaced first in $\Gamma_{data}$ by $[*_2(\{w_0\},w^*_2),c^*_2(R_2 \cup \{w^*_2\},\text{exp}_2)]$.
+As a result, when applying $T_{task}$ on $\Gamma_{data}$, it appears that $*_2$ can be performed in parallel with $c_1$.
+This is due to the fact that $*_2$ is splitted from from $c_2$ and that $R^*_2=\{w_0\}$ is a subset of $R_2=\{w_0,w_1\}$ and not equal to $R_2$.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (c0) at (0,0) {$c_0$};
+   \node[] (c1) at (2,0) {$c_1$};
+   \node[] (star2) at (2,1) {$*_2$};
+   \node[] (c2) at (4,0) {$c^*_2$};
+
+  \path[->]
+    (c0) edge node {} (c1)
+         edge node {} (star2)
+    (c1) edge node {} (c2)
+    (star2) edge node {} (c2);
+  \end{tikzpicture}
+  \caption{Example of hybrid parallelization.}
+  \label{fig:more}
+\end{center}
+\end{figure}
+
+As a result, the hybrid parallelization of a multi-stencil program $\mathcal{MSP}$, produces $k$ multi-stencil programs
+\begin{equation*}
+\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{hybrid}).
+\end{equation*}
+The set of computations $\Gamma_{hybrid}$ is a dependency graph between computations $c_i$ and update computations $*_i$, and can be built from the call to 
 \begin{equation*}
@@
 The rest of this paper deals with the hybrid parallelization of multi-stencil programs described above.
-The first reason for this choice is that this paper does not adress the portability problem, which however will be a perspective.
-Second, by adressing the hybrid parallelism case, we also are able to adress data parallelism and task parallelism, from which hybrid parallelism is built.
+The reason of this choice is the fact that by adressing the hybrid parallelism case, we also are able to adress data parallelism and task parallelism, from which hybrid parallelism is built.

### 2015-06-12 rapport/component.tex
rapport_12_06

@@
    \node[use] (u) at (0,1) {$m$};
+   \node[] (r) [right = 1cm of u] {$R_i \cup \{w_i\}$};
 
@@
    \node[use] (u) at (0,1) {$m$};
+   \node[] (r) [right = 0.5cm of u] {$R^*_i$};
 
@@
 For example, the canonical decomposition tree of Figure~\ref{fig:canon} is dumped into the component assembly of Figure~\ref{fig:assembly}.
+The text next to computation and synchronization components indicates to which data they have to be linked to.
+Actually, any computation $c_i$ or $c_i^*$ has to be linked to the data it reads and writes $R_i \cup w_i$.
+On the other hand, a synchronization $*_i$, has to be linked to the data it updates $R^*_i$.
 
@@
    \node[use] (c0u) [above = 0.8cm of c0] {$m$};
-   \node[] (r0) [right = 0.5cm of c0u] {$R_0$};
+   \node[] (r0) [right = 1cm of c0u] {\textcolor{gray}{$R_0 \cup \{w_0\}$}};
    %sync0
@@
    \node[use] (sync0u) [above = 0.8cm of sync0] {$m$};
-   \node[] (star1) [right = 0.5cm of sync0u] {$*_1$};
+   \node[] (star1) [right = 0.5cm of sync0u] {\textcolor{gray}{$R^*_1$}};
    %c1
@@
    \node[use] (c1u) [above = 0.8cm of c1] {$m$};
-   \node[] (r1) [right = 0.5cm of c1u] {$R_1$};
+   \node[] (r1) [right = 1cm of c1u] {\textcolor{gray}{$R_1 \cup \{w_1\}$}};
    %par0
@@
    \node[use] (c2u) [above = 0.8cm of c2] {$m$};
-   \node[] (r2) [right = 0.5cm of c2u] {$R_2$};
+   \node[] (r2) [right = 1cm of c2u] {\textcolor{gray}{$R_2 \cup \{w_2\}$}};
   %sync1
@@
    \node[use] (sync1u) [above = 0.8cm of sync1] {$m$};
-   \node[] (star4) [right = 0.5cm of sync1u] {$*_4$};
+   \node[] (star4) [right = 0.5cm of sync1u] {\textcolor{gray}{$R^*_4$}};
   %c4
@@
    \node[use] (c4u) [above = 0.8cm of c4] {$m$};
-   \node[] (r4) [right = 0.5cm of c4u] {$R_4$};
+   \node[] (r4) [right = 1cm of c4u] {\textcolor{gray}{$R_4 \cup \{w_4\}$}};
   %c6
@@
    \node[use] (c6u) [above = 0.8cm of c6] {$m$};
-   \node[] (r6) [right = 0.5cm of c6u] {$R_6$};
+   \node[] (r6) [right = 1cm of c6u] {\textcolor{gray}{$R_6 \cup \{w_6\}$}};
   %seq2
@@
    \node[use] (c3u) [above = 0.8cm of c3] {$m$};
-   \node[] (r3) [right = 0.5cm of c3u] {$R_3$};
+   \node[] (r3) [right = 1cm of c3u] {\textcolor{gray}{$R_3 \cup \{w_3\}$}};
   %c5
@@
    \node[use] (c5u) [above = 0.8cm of c5] {$m$};
-   \node[] (r5) [right = 0.5cm of c5u] {$R_5$};
+   \node[] (r5) [right = 1cm of c5u] {\textcolor{gray}{$R_5 \cup \{w_5\}$}};
   %c7
@@
    \node[use] (c7u) [above = 0.8cm of c7] {$m$};
-   \node[] (r7) [right = 0.5cm of c7u] {$R_7$};
+   \node[] (r7) [right = 1cm of c7u] {\textcolor{gray}{$R_7 \cup \{w_7\}$}};
   %sync2
@@
    \node[use] (sync2u) [above = 0.8cm of sync2] {$m$};
-   \node[] (star8) [right = 0.5cm of sync2u] {$*_8$};
+   \node[] (star8) [right = 0.5cm of sync2u] {\textcolor{gray}{$R^*_8$}};
   %c8
@@
    \node[use] (c8u) [above = 0.8cm of c8] {$m$};
-   \node[] (r8) [right = 0.5cm of c8u] {$R_8$};
+   \node[] (r8) [right = 1cm of c8u] {\textcolor{gray}{$R_8 \cup \{w_8\}$}};
 
@@
 \subsection{Complete component assembly}
-At this point of the work, the computation part of the parallel $\mathcal{MSP}$ program is solved and dumped to a component assembly.
-Of course, this component assembly does not represent the entire $\mathcal{MSP}$ program.
+At this point of the work, the computation part $\Gamma$ of the parallel $\mathcal{MSP}$ program is solved and dumped to a component assembly.
+Of course, this component assembly does not represent the entire $\mathcal{MSP}$ program which also contains $\mathcal{M}$, $\Delta$ and $T$.
 Actually, additionnal services are needed.
@@
 Those quantities are mapped onto a distributed mesh, which also means that the mesh has to be accessible by the data.
-In addition to this, and as explained in Section~\ref{sect:dataparal}, the data parallelization needs the management of specific updated-computations $c_i^*$, which means that the complete assembly have to manage an update service, as well as the partitioning of the mesh and the quantities mapped onto it.
+In addition to this, and as explained in Section~\ref{sect:dataparal}, the data parallelization needs the management of specific update computations $*_i$, which means that the complete assembly have to manage an update service, as well as the partitioning of the mesh and the quantities mapped onto it.
 
-However, there is not a unique way to design the rest of the assembly and to make accessible the different services (data, update, mesh etc.).
+to be continued...
+
+%However, there is not a unique way to design the rest of the assembly and to make accessible the different services (data, update, mesh etc.).
 The work presented until now can be considered as a generic solution to solve the parallel computation part, but the rest of the assembly is dependant of the implementation choice.

### 2015-07-20 WOLFHPC15/article/article.tex
article wolfhpc15

\documentclass{acm_proc_article-sp}
\usepackage{tikz}
\usepackage{listings}
%----------TIKZ
\usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}

\tikzset{
    use/.style={
    circle,draw=black,fill=black,scale=0.5,text=white
    },
    provide/.style={
    circle,draw=black,fill=white,scale=0.5
    },
    component/.style={
    rectangle,rounded corners=3pt,draw=black
    },
    dcomponent/.style={
    rectangle,rounded corners=3pt,dashed,draw=black
    },
    smallcp/.style={
    rectangle,rounded corners=3pt,draw=black,font=\tiny,
    },
    resource/.style={
    rectangle,rounded corners=3pt, dashed,draw=black,font=\tiny,
    },
    connection/.style={
    decoration={ markings,
      mark=at position 1 with {\node[fill=black,draw=black,circle,inner sep=2pt](conn){};}
      },
    postaction={decorate}
    },
    bconnection/.style={
    decoration={ markings,
      mark=at position 1 with {\node[fill=black,draw=black,circle,inner sep=2pt](conn){};}
      },
    postaction={decorate}
    },
    mpiconn/.style={
    decoration={ markings,
      mark=at position 0.99 with {\node[fill=black,draw=black,rectangle,,inner sep=2pt](mpi){};}
      },
    postaction={decorate}
    },
}
%-------------

\begin{document}

\title{A component-based DSL for multi-stencil programs}

\numberofauthors{2} 
\author{
% 1st. author
\alignauthor
Christian Perez\\
       \affaddr{INRIA}\\
       \affaddr{ENS Lyon, 46 Allee d'Italie}\\
       \affaddr{Lyon, France}\\
       \email{christian.perez@inria.fr}
% 2nd. author
\alignauthor
Helene Coullon\\
       \affaddr{INRIA}\\
       \affaddr{ENS Lyon, 46 Allee d'Italie}\\
       \affaddr{Lyon, France}\\
       \email{helene.coullon@inria.fr}
}

\maketitle

%----------------------------------------
\begin{abstract}

\end{abstract}
%----------------------------------------

\keywords{Component models, domain specific languages, multi-stencil, numerical simulations} % NOT required for Proceedings

%----------------------------------------
\section{Introduction}
\label{sect:intro}
\input{introduction.tex}
%----------------------------------------
\section{Underlying concepts}
\label{sect:concept}
\input{concept.tex}
%----------------------------------------
\section{The Multi-Stencil Component Assembly Compiler}
\label{sect:mscac}
\input{mscac.tex}
%----------------------------------------
\section{MS Language and MSC compiler}
\label{sect:msmsc}
\input{ms_msc.tex}
%----------------------------------------
\section{Control components and component assembly}
\label{sect:component}
\input{component.tex}
%----------------------------------------
\section{Evaluation}
\label{sect:eval}
\input{evaluation.tex}
%----------------------------------------
\section{Related Work}
\label{sect:related}
\input{related.tex}
%----------------------------------------
\section{Conclusion}
\label{sect:conclusion}
\input{conclusion.tex}
%----------------------------------------

\bibliographystyle{abbrv}
\bibliography{biblio}

\end{document}

### 2015-07-20 WOLFHPC15/article/component.tex
article wolfhpc15

%----------------------------------------
\subsection{Control components}
%----------------------------------------
The series-parallel tree decomposition $\Gamma_{tsp}$ represents the control of the dependencies of the simulation. As a result to be able to dump it to a component assembly it is needed to introduce what can be called \emph{control components}. Those components can be used for any case of simulation which increases code reuse between simulations. A control component is composed of a single \emph{provide} port linked to a single execution method, most of the time called \emph{go}. We introduce three types of control components described bellow:

\begin{itemize}
\item a sequence component, called SEQ, is the direct dump to a sequence node of $\Gamma_{tsp}$. The role of this component is to call an ordered list of other components. For this reason, it is composed of a \emph{use-multiple} port connected to the components to call as a sequence.

\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (seq) at (0,0) {$SEQ$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {$m$};

  \path[->]
    (p) edge node {} (seq)
    (seq) edge node {} (u);
\end{tikzpicture}

\item A parallel component, called PAR, is the direct dump to a parallel node of $\Gamma_{tsp}$. The role of this component is to call simultaneously a set of other components. For this reason, it is composed, as SEQ, of a \emph{use-multiple} port connected to the components to call in parallel.

\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (seq) at (0,0) {$PAR$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {$m$};

  \path[->]
    (p) edge node {} (seq)
    (seq) edge node {} (u);
\end{tikzpicture}

\item A synchronization component, called SYNC, is the direct dump to an update leaf of $\Gamma_{tsp}$. The role of this component is to call the synchronization of a given data. For this reason, it is composed of a \emph{use} port connected to the data to update.

\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (sync) at (0,0) {$SYNC$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};

  \path[->]
    (p) edge node {} (sync)
    (sync) edge node {} (u);
\end{tikzpicture}
\end{itemize}

In the computation part of the simulation a final component is needed, the computation component, denoted K. This type of component represent a single computation of the simulation, i.e. the computation of a single output data (written) using a set of input data (read). Using a \emph{use-multiple}, however, a list of component is used and the explicit identification of each used component is lost. If for SEQ or PAR, the identification of components is not usefull, it is usefull to manipulate data in a computation. For this reason a computation component has as much \emph{use} ports as input and output data to manipulate. We denote this in the bellow representation as a star $*$ on the use port.

\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (k) at (0,0) {$K$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[right=0.2 of u] (star) {*};

  \path[->]
    (p) edge node {} (k)
    (k) edge node {} (u);
\end{tikzpicture}

%----------------------------------------
\subsection{Dump to a component assembly}
%----------------------------------------
From $\Gamma_{tsp}$ and using the control components described above, a direct dump can be done to a component assembly for an hybrid (data and task) parallelization of a simulation. However, it is also possible to dump to a data parallelization only. In such a case the computation of $\Gamma_{data}$ is enough to generate the component assembly. Actually a single SEQ component is thus needed and this component is linked to all computations and synchronizations of $\Gamma_{data}$ directly.

As an example, from $\Gamma_{tsp}$ illustrated in Figure~\ref{fig:canon} a direct dump example is illustrated in Figure~\ref{fig:assembly}. In this Figure, the ports linked to data (ports of SYNC and K components) are represented but not their connections. Moreover, each computation component is an instance of the component K presented before but takes the identification name of the computation in the MSL file.

\begin{figure*}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   %seq0
   \node[component] (seq0) at (0,0) {$SEQ$};
   \node[provide] (seq0p) [left = 1cm of seq0] {};
   \node[use] (seq0u) [right = 1cm of seq0] {$m$};
   %c0
   \node[component] (c0) [right = 2.5cm of seq0u] {$c_0$};
   \node[use] (c0u) [right = 0.8cm of c0] {};
   \node[right=0.2 of c0u] (star) {*};
   %sync0
   \node[component] (sync0) [below = 1cm of c0] {$SYNC$};
   \node[use] (sync0u) [right = 1cm of sync0] {$m$};
   %c1
   \node[component] (c1) [below = 1cm of sync0] {$c_1^*$};
   \node[use] (c1u) [right = 0.8cm of c1] {};
   \node[right=0.2 of c1u] (star) {*};
   %par0
   \node[component] (par0) [below = 1cm of c1] {$PAR$};
   \node[use] (par0u) [right = 1cm of par0] {$m$};
  %seq1
   \node[component] (seq1) [right = 1cm of par0u] {$SEQ$};
   \node[use] (seq1u) [right = 1cm of seq1] {$m$};
   %par1
   \node[component] (par1) [right = 1cm of seq1u] {$PAR$};
   \node[use] (par1u) [right = 1cm of par1] {$m$};
  %c2
   \node[component] (c2) [right = 1cm of par1u] {$c_2$};
   \node[use] (c2u) [right = 0.8cm of c2] {};
   \node[right=0.2 of c2u] (star) {*};
  %sync1
   \node[component] (sync1) [below = 1cm of c2] {$SYNC$};
   \node[use] (sync1u) [right = 1cm of sync1] {$m$};
  %c4
   \node[component] (c4) [below = 1cm of par1] {$c_4^*$};
   \node[use] (c4u) [right = 0.8cm of c4] {};
   \node[right=0.2 of c4u] (star) {*};
  %c6
   \node[component] (c6) [below = 1cm of c4] {$c_6$};
   \node[use] (c6u) [right = 0.8cm of c6] {};
   \node[right=0.2 of c6u] (star) {*};
  %seq2
   \node[component] (seq2) [below = 3cm of seq1] {$SEQ$};
   \node[use] (seq2u) [right = 1cm of seq2] {$m$};
  %c3
   \node[component] (c3) [right = 1cm of seq2u] {$c_3$};
   \node[use] (c3u) [right = 0.8cm of c3] {};
   \node[right=0.2 of c3u] (star) {*};
  %c5
   \node[component] (c5) [below = 1cm of c3] {$c_5$};
   \node[use] (c5u) [right = 0.8cm of c5] {};
   \node[right=0.2 of c5u] (star) {*};
  %c7
   \node[component] (c7) [below = 1cm of par0] {$c_7$};
   \node[use] (c7u) [right = 0.8cm of c7] {};
   \node[right=0.2 of c7u] (star) {*};
  %sync2
   \node[component] (sync2) [below = 1cm of c7] {$SYNC$};
   \node[use] (sync2u) [right = 1cm of sync2] {$m$};
  %c8
   \node[component] (c8) [below = 1cm of sync2] {$c_8^*$};
   \node[use] (c8u) [right = 0.8cm of c8] {};
   \node[right=0.2 of c8u] (star) {*};

   \path[->]
   %seq0
    (seq0) edge node {} (seq0u)
    (seq0p) edge node {} (seq0)
   %c0
    (seq0u) edge node {} (c0)
    (c0) edge node {} (c0u)
    %sync0
    (seq0u) edge node {} (sync0.west)
    (sync0) edge node {} (sync0u)
    %c1
    (seq0u) edge node {} (c1)
    (c1) edge node {} (c1u)
   %par0
    (seq0u) edge node {} (par0.west)
    (par0) edge node {} (par0u)
  %seq1
    (par0u) edge node {} (seq1)
    (seq1) edge node {} (seq1u)
  %par1
    (seq1u) edge node {} (par1.west)
    (par1) edge node {} (par1u)
  %c2
    (par1u) edge node {} (c2)
    (c2) edge node {} (c2u)
  %sync0
    (par1u) edge node {} (sync1.west)
    (sync1) edge node {} (sync1u)
  %c4
    (seq1u) edge node {} (c4.west)
    (c4) edge node {} (c4u)
  %c6
    (seq1u) edge node {} (c6.west)
    (c6) edge node {} (c6u)
  %seq2
    (par0u) edge node {} (seq2.west)
    (seq2) edge node {} (seq2u)
  %c3
    (seq2u) edge node {} (c3)
    (c3) edge node {} (c3u)
  %c6
    (seq2u) edge node {} (c5.west)
    (c5) edge node {} (c5u)
  %c6
    (seq0u) edge node {} (c7.west)
    (c7) edge node {} (c7u)
  %sync2
    (seq0u) edge node {} (sync2.west)
    (sync2) edge node {} (sync2u)
  %c8
    (seq0u) edge [bend right] node {} (c8.west)
    (c8) edge node {} (c8u)
   ;
\end{tikzpicture}
\caption{Direct dump to a component assembly.}
\label{fig:assembly}
\end{center}
\end{figure*}

### 2015-07-20 WOLFHPC15/article/concept.tex
article wolfhpc15

%----------------------------------------
\subsection{Multi-stencil programs}
\label{sect:multistencil}
%----------------------------------------
\begin{itemize}
\item textual introduction on the problem, what is a stencil + example. What is a real simulation (more than one stencil and additionnal computations). 
\item Existing solutions with prgramming freedom (instead of PETSc and Trilinos) = stencil compilers which give optimized code for one stencil code. Thus, for distributed memory and HPC, still have to use MPI between optimized codes to get an overall parallelization of the simulation.
\item Short definition of a multi-stencil program and its computations (cite research report for more details)
\end{itemize}

%----------------------------------------
\subsection{Component models}
%----------------------------------------
\begin{itemize}
\item Interest of component models, references.
\item Textual definition of a component, a port (use, provide and use-multiple), (an interface) and a component assembly
\end{itemize}

### 2015-07-20 WOLFHPC15/article/evaluation.tex
article wolfhpc15

\subsection{Shallow water equations}
\subsection{MS description}
\subsection{Series-parallel tree decomposition}
\subsection{Performances}
\begin{itemize}
\item for data part - SIPSim and SkelGIS
\item speedup Figures
\item number of lines (counting computation components)
\end{itemize}

### 2015-07-20 WOLFHPC15/article/introduction.tex
article wolfhpc15

\begin{itemize}
\item DSL interest: simplified HPC programming + perf
\item DSL limits: Lots of DSLs difficult to write, to maintain, to make portable etc.
\item Components introduction, this paper takes a first example of DSL to components
\item The domain targeted is multi-stencil simulations
\item Contributions: DSL to components try + agnostic coarse grain DSL for multi-stencil programs
\item Outline
\end{itemize}

### 2015-07-20 WOLFHPC15/article/ms_msc.tex
article wolfhpc15

%----------------------------------------
\subsection{MS language}
%----------------------------------------
The MS language (MSL) is an agnostic descriptive language for multi-stencil simulations. Agnostic means that no knowledge on the mesh type and the neighborhood description of stencil computations are needed in the language. As the language consists in a description of the simulation, without any computation details and programming details, and as no computation is directly done from the language, this language is also called a descriptive language. It is composed of five description sections:
\begin{enumerate}
\item meshes description;
\item domains description;
\item data description;
\item time loop description;
\item computations description.
\end{enumerate}

This paper deals with the MSC compiler which only needs the three last enumerations above. For this reason we describe in the rest of this section MSL for data, time loop and computations descriptions.

\paragraph{Data description} Data description starts with the tag $data:$. Each line after this tag represents a single data description. This description is composed of an identification name, and the domain identification name it is applied on.

\begin{lstlisting}[basicstyle=\footnotesize,mathescape,frame=single,language=C++]
data:
  name_id,domain_id
\end{lstlisting}

\paragraph{Time description} The time of a numerical simulation consists in a set of iterations. The end of the simulation time can be controlled by a convergence function or by a manual configuration of the number of iteration. In this first version of MSL, the iteration number is handled.

\begin{lstlisting}[basicstyle=\footnotesize,mathescape,frame=single,language=C++]
time:nb_iteration
\end{lstlisting}

\paragraph{Computations description} The most important part of MSL is the description of the computations of the simulation. The language follows the definitions given in Section~\ref{sect:multistencil}. Thus, the description of a computation consists of a type of computation, an identification name of the computation, a set of identifiers for the data read by the computation, a single identifier for the data written by the computation. Moreover, if the type of computation is a stencil, a neighborhood identifier is needed.

\begin{lstlisting}[basicstyle=\footnotesize,mathescape,frame=single,language=C++]
computations:
  type:name_id({read1_id,read2_id,...},
 	    written_id[,neighborhood_id])
\end{lstlisting}

%----------------------------------------
\subsection{MSC compiler}
%----------------------------------------
The MSC compiler, which is a subpart of the overall compiler MSCAC, performs an algorithm composed of six main steps described bellow:

\begin{enumerate}
\item creation of $\Gamma$, the ordered list of computations, from the MSL input file;
\item creation of $\Gamma_{data}$ to insert synchronizations in the ordered list of computations;
\item creation of the dependency graph $\Gamma_{hybrid}$;
\item transform $\Gamma_{hybrid}$ to a minimal series-parallel graph $\Gamma_{msp}$;
\item creation of the series-parallel tree decomposition $\Gamma_{tsp}$;
\item dump $\Gamma_{tsp}$ to a component assembly.
\end{enumerate}

\paragraph{Creation of the ordered list of computations} This step of the MSC compiler parse the MSL input file which describes the simulation. From the ordered lines of computations, it builds the ordered list of computations $\Gamma$ to perform in the simulation.

\paragraph{Creation of the synchronized ordered list of computations} From the ordered list of computations $\Gamma$ it is possible to detect needed synchronizations if a \emph{data parallelization} \textbf{(PB: to explain !!!)} is performed on the simulation. A synchronization is needed each time a data read by a stencil computation has been written by a previous computation different from a boundary computation. Actually, as a stencil computation needs neighborhood values, in a domain decomposition values computed by different processors or cores are needed.

\paragraph{Creation of the dependency graph} From the synchronized ordered list of computations $\Gamma_{data}$, a dependency graph can be build. A dependency exists between two computations (including synchronizations) if and only if a data read has been written by a previous computation in $\Gamma_{data}$. Nodes of the dependency graph are computations and synchronizations, while edges are dependencies between them. $\Gamma_{hybrid}$ is a directed acyclic graph (\emph{dag}) as it is build from an ordered list.

\paragraph{Transformation to a minimal series-parallel graph} As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if the forbidden \emph{N-shape} is not found in the graph. To transform $\Gamma_{hybrid}$ to the minimal series-parallel graph $\Gamma_{msp}$, a BFS algorithm is applied to build a complete bipartite between two levels of $\Gamma_{hybrid}$~\cite{Mitchell:2004:CMV:1082101.1082117}.

\paragraph{Creation of a tree decomposition} As shown in different works~\cite{Valdes:1979:RSP:800135.804393,Schoenmakers95anew}, from a minimal series-parallel graph a tree decomposition can be built. A series-parallel tree decomposition consists in the decomposition of the graph as a set of \emph{sequences} and \emph{parallel} sections in a tree.

\paragraph{Dump to a component assembly} From the series-parallel tree decomposition $\Gamma_{tsp}$ can be built a program using any parallel language containing \emph{parallel sections} and \emph{sequences of instructions} such as OpenMP~\cite{660313}, HPF~\cite{219857}, UPC~\cite{El-Ghazawi:2006:UUP:1188455.1188483} etc. In this paper we study the dump of $\Gamma_{tsp}$ to a component assembly which needs the introduction of \emph{control components} described in the next section.

For more details on MSC, a full description of the compilation is described in~\cite{}.

%----------------------------------------
\subsection{Example}
%----------------------------------------
In this Section we give an example of an input file written with MSL, the $\Gamma_{tsp}$ series-parallel tree decomposition computed and the final component dump.

The input MSL file, shown in Figure~\ref{fig:msinput} is composed of nine computations and ten data.

\begin{figure}[h!]
\begin{lstlisting}[basicstyle=\footnotesize,mathescape,frame=single,language=C++]
data:
  a,d1
  b,d1
  c,d2
  d,d2
  e,d3
  f,d1
  g,d3
  h,d2
  i,d1
  j,d2
time:500
computations:
  local:c0({a},b)
  stencil:c1({b},c,n1)
  local:c2({c},d)
  local:c3({c},e)
  stencil:c4({d},f)
  local:c5({e},g)
  local:c6({f},h)
  local:c7({g,h},i)
  stencil:c8({i},j,n1)
\end{lstlisting}
\caption{Example of MSL input file}
\label{fig:msinput}
\end{figure}

The first step of the compilation is to build the synchronized ordered list of computations $\Gamma_{data}$. In this example each stencil computation actually needs a synchronization. For example, the stencil computation $c_1$ read the data $b$ which has been written by the computation $c_0$. For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed to the sublist $[c_0,sync_1,c_1]$ in $\Gamma_{data}$. The new computation $sync_1$ read and write $b$. As a result a dependency is kept between $sync_1$, and $c_1$. The same is performed for the stencils $c_4$ and $c_8$.

Once the synchronized list $\Gamma_{data}$ is obtained, the dependency graph $\Gamma_{hybrid}$ is built and illustrated in Figure~\ref{}. Each computation (and synchronization) is a node of the graph, while each edge represent a read/write data dependency between computations.

After a set of transformation briefly explained in the previous Section, a series-parallel tree decomposition $\Gamma_{tsp}$ is obtained and illustrated in Figure~\ref{fig:canon}.

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0) at (0,0) {$\mathcal{S}$};
   \node[] (c0) at (-3,1) {$c_0$};
   \node[] (star1) at (-2,1) {$sync_1$};
   \node[] (c1) at (-1,1) {$c_1^*$};

   \node[] (p0) at (0,1) {$\mathcal{P}$};
   \node[] (s1) at (-1,2) {$\mathcal{S}$};
   \node[] (p1) at (-2,3) {$\mathcal{P}$};
   \node[] (c4) at (-1,3) {$c_4^*$};
   \node[] (c6) at (-0,3) {$c_6$};
   \node[] (c2) at (-2.5,4) {$c_2$};
   \node[] (star4) at (-1.5,4) {$sync_4$};
   \node[] (s2) at (1,2) {$\mathcal{S}$};
   \node[] (c3) at (0.5,3) {$c_3$};
   \node[] (c5) at (1.5,3) {$c_5$};

   \node[] (c7) at (1,1) {$c_7$};
   \node[] (star8) at (2,1) {$sync_8$};
   \node[] (c8) at (3,1) {$c_8^*$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (p1)
         edge node {} (c4)
         edge node {} (c6)
    (p1) edge node {} (c2)
         edge node {} (star4)
    (s2) edge node {} (c3)
         edge node {} (c5);
  \end{tikzpicture}
\caption{$\Gamma_{tsp}$ of Figure~\ref{fig:msinput}.}
\label{fig:canon}
\end{center}
\end{figure}

### 2015-07-20 WOLFHPC15/article/mscac.tex
article wolfhpc15

The MS language (MSL) is an agnostic descriptive language for multi-stencil simulations. Agnostic in our context means that without any knowledge on the mesh type and the neighborhood description of stencil computations, the language consists in data and computations descriptions. From this agnostic and descriptive language can be compiled the structure of the parallel simulation in two ways:
\begin{itemize}
\item data parallelism;
\item hybrid (data and task) parallelism.
\end{itemize}

As illustrated in Figure~\ref{fig:mscac}, the MS language compiler, called MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of two different part:
 \begin{itemize}
\item first, the computation compiler, called MSC, which is responsible for the extraction of the computations parallelization;
\item second, the data compiler, called MSD, which handles a distributed data structure, data mapped onto it, and their relation with the computation part of the final program.
\end{itemize}

MSCAC produces from the MSL source file the structure of the parallel simulation as a component assembly. Thus, the coarse-grain parallelization of the simulation is produced. In this final component assembly, an additionnal work is asked to the user: write computation components, using interfaces of the distributed data structure choosen by MSD. Thus, this work is complementary to stencil compilers or to distributed data structure solutions. Actually, it produces a ready-to-fill parallel structure pattern of the simulation.

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[remember picture,
  inner/.style={rectangle,rounded corners=3pt,thick,inner sep=5pt},
  outer/.style={rectangle,rounded corners=3pt,thick,inner sep=5pt}
  ]
  \node (MSL) {MS Language};
  \node[outer,below=of MSL,draw=black] (MSCAC) {
    \begin{tikzpicture}
      \node [inner,draw=black,thick] (MSC)  {MSC};
      \node [inner,right=of MSC,draw=black,thick,,fill=gray!30!white,dotted] (MSD)  {MSD};
    \end{tikzpicture}
  };
  \node [below=of MSCAC] (CA) {Component Assembly};
  \node [left=0.1 of MSCAC] (CAid) {MSCAC};
  \draw[->] (MSL) -- (MSCAC);
  \draw[->] (MSCAC) -- (CA);
\end{tikzpicture}
\caption{MS Language and MSCAC compiler}
\label{fig:mscac}
\end{center}
\end{figure}

This paper presents the MS language and the MSC part of the compiler. The evaluation part will shortly explain how the MSD part of the compiler has been acheived for this paper.

### 2015-07-21 WOLFHPC15/article/concept.tex
wolfhpc15 21_07

@@
 %----------------------------------------
+\subsection{Parallelization techniques}
+\label{sect:parallel}
+%----------------------------------------
+\begin{itemize}
+\item data parallelism
+\item task parallelism
+\item hybrid parallelism
+\end{itemize}
+
+%----------------------------------------
 \subsection{Component models}
 %----------------------------------------
-\begin{itemize}
-\item Interest of component models, references.
-\item Textual definition of a component, a port (use, provide and use-multiple), (an interface) and a component assembly
-\end{itemize}
+Component model is an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228} which improves code re-use, scalability and maintainability of applications~\cite{Szyperski:2002:CSB:515228,bigot:inria-00388508}.
+Moreover, component models bring a separation of concerns in the application by a clear division of functionnalities in different components.
+In addition to this, recent work on component models have shown a simultaneous answer to performance, maintainability and portability of applications~\cite{l2c}, which makes this domain an interesting solution to bring maintainability and portability in HPC programming.
+
+Component-based software engineering~\cite{Szyperski:2002:CSB:515228} extends the concept of class by specifying in its interfaces not only the services it offers, or public methods, but also all its possible interactions with outer world, including the services it requires.
+As a result, each component is an independant entity composed of a set of services its provides, and a set of services it requires (and uses).
+
+A \emph{port} is an entity embbedded in the component which makes possible an \emph{assembly} of components.
+An assembly of components (of instanciated components) is a way to actually connect components together and to produce a complete application, as a set of components and their interactions.
+A provided service inside a component is associated to a \emph{provide} port, while a required service is associated to a \emph{use} port.
+It is also possible to group more than one required service into a single port, as a list, called a \emph{use-multiple} port.
+As illustrated in Figures~\ref{fig:ports}, a provide port will be represented by a white circle, a use port by a black circle and a use-multiple port by a black circle with a white $m$ in it.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (C) at (0,0) {$Comp_0$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (1.5,0) {};
+   \node[use] (um) at (0,1) {$m$};
+   \node[provide,right=of u] (p1) {};
+   \node[component,right=of p1] (C1) {$Comp_1$};
+
+  \path[-]
+    (p) edge node {} (C)
+    (C) edge node {} (u)
+    	edge node {} (um)
+    (p1) edge node {} (C1);
+\end{tikzpicture}
+\caption{Two components, one with a provide, use and use-multiple ports, the second with a single provide port}
+\label{fig:ports}
+\end{center}
+\end{figure}
+
+In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (C) at (0,0) {$Comp_0$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (1.5,0) {};
+   \node[use] (um) at (0,1) {$m$};
+   \node[component,right=of u] (C1) {$Comp_1$};
+
+  \path[-]
+    (p) edge node {} (C)
+    (C) edge node {} (u)
+    	edge node {} (um)
+    (u) edge node {} (C1);
+\end{tikzpicture}
+\caption{Component assembly of Figure~\ref{fig:ports}}
+\label{fig:assembly}
+\end{center}
+\end{figure}
+
+Many component models exist, each of them with its own specifications and functionnalities, like CCM~\cite{corba:omg06} (CORBA Component Model), GCM~\cite{Baude} (Grid Component Model) or CCA~\cite{Armstrong:1999:TCC:822084.823232} (Common Component Architecture) for example.
+However, conepts of component, port, interface and assembly are common to many component models.
+The rest of this paper will only use those concepts to present the MS Language and its compiler whatever the component model is.

### 2015-07-21 WOLFHPC15/article/ms_msc.tex
wolfhpc15 21_07

@@
 
-Once the synchronized list $\Gamma_{data}$ is obtained, the dependency graph $\Gamma_{hybrid}$ is built and illustrated in Figure~\ref{}.
+Once the synchronized list $\Gamma_{data}$ is obtained, the dependency graph $\Gamma_{hybrid}$ is built and illustrated in Figure~\ref{fig:hyb}.
 Each computation (and synchronization) is a node of the graph, while each edge represent a read/write data dependency between computations.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node (c0) at (0,0) {$c_0$};
+   \node[right=1 of c0] (sy1) {$sync_1$};
+   \node[right=1 of sy1] (c1) {$c_1$};
+   \node[above right=1 of c1] (c2) {$c_2$};
+   \node[below right=1 of c1] (c3) {$c_3$};
+   \node[right=1 of c2] (sy4) {$sync_4$};
+   \node[right=1 of c3] (c5) {$c_5$};
+   \node[right=1 of sy4] (c4) {$c_4$};
+   \node[right=1 of c4] (c6) {$c_6$};
+   \node[below right=1 of c6] (c7) {$c_7$};
+   \node[right=1 of c7] (sy8) {$sync_8$};
+   \node[right=1 of sy8] (c8) {$c_8$};
+
+  \path[->]
+    (c0) edge node {} (sy1)
+    (sy1) edge node {} (c1)
+    (c1)  edge node {} (c2)
+          edge node {} (c3)
+    (c2) edge node {} (sy4)
+    (sy4) edge node {} (c4)
+    (c4) edge node {} (c6)
+    (c3) edge node {} (c5)
+    (c5) edge node {} (c7)
+    (c6) edge node {} (c7)
+    (c7) edge node {} (sy8)
+    (sy8) edge node {} (c8);
+\end{tikzpicture}
+\caption{$\Gamma_{hybrid}$ of Figure~\ref{fig:msinput}}
+\label{fig:hyb}
+\end{center}
+\end{figure}
 
@@
   \end{tikzpicture}
-\caption{$\Gamma_{tsp}$ of Figure~\ref{fig:msinput}.}
+\caption{$\Gamma_{tsp}$ of Figure~\ref{fig:hyb}.}
 \label{fig:canon}

### 2015-07-21 WOLFHPC15/article/concept.tex
wolfhpc15 21_07

@@
 %----------------------------------------
-\begin{itemize}
-\item textual introduction on the problem, what is a stencil + example. What is a real simulation (more than one stencil and additionnal computations). 
-\item Existing solutions with prgramming freedom (instead of PETSc and Trilinos) = stencil compilers which give optimized code for one stencil code.
-Thus, for distributed memory and HPC, still have to use MPI between optimized codes to get an overall parallelization of the simulation.
-\item Short definition of a multi-stencil program and its computations (cite research report for more details)
-\end{itemize}
+To numerically solve a set of PDEs, iterative methods (finite difference, finite volume or finite element methods) are frequently used to approximate the solution through a discretized (step by step) phenomena.
+Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied onto a mesh (space discretization).
+In other words, in a mesh-based numerical simulation, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain (the mesh).
+Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}~\cite{spaaTangCKLL11}.
+
+If $x$ represents an element of the mesh, if $N(x)$ is a neighborhood shape around the element $x$, and if $u_t(x)$ is the data (or quantity) to write.
+Moreover if $V_t(x)$ is the set of data to simulate such that $V_t(x) \cap u_t(x) = \emptyset$, a stencil computation is of the following form
+
+\begin{equation} 
+\{u_{t-1}(x), u_{t-1}(N(x)), V_t(x), V_t(N(x)\} \longmapsto u_{t}(x).
+\label{space}
+\end{equation}
+
+In other words, in a stencil computation the quantity written can be computed using the same quantity at a previous time iteration at $x$ and at $N(x)$.
+And, any other quantity can also be used in the computation at $x$ and $N(x)$.
+
+A real case numerical simulation is most of the time composed of a set of numerical schemes (stencils computations), with one or more stencil shape (neighboorhood shape), and onto one or more data.
+Moreover, a numerical simulation also performs additionnal auxiliary computations which do not involve neighborhood, called local computations.
+In this paper we target this kind of complex simulations with more than one stencil computation and also additionnal auxiliary computations, that we call \emph{multi-stencil programs}.
+The work presented produces a general parallelization structure of the overall simulation, while most stencil solutions produce optimized codes for a single stencil code~\cite{citeulike12258902, spaaTangCKLL11, Giles2011}.
+As it will be detailed in the related work, this work is complementary to most stencil compilers and implicit parallelism languages.
+
+\textbf{Do I need the following definitions?}
+\begin{mydef}
+A \textit{stencil computation} is defined as a quadruplet $s(R,w,\text{exp},\mathcal{N})$,
+\end{mydef}
+where $R$ is the set of data read by the computation, $w$ is the data written by the computation, \textit{exp} is the numerical expression to compute $w$, and $\mathcal{N}$ is a neighborhood shape of the stencil computation.
+
+\begin{mydef}
+A local computation is a triplet $l(R,w,\text{exp})$, where $exp$ does not involve a neighborhood function $\mathcal{N}$.
+\end{mydef}
 
@@
 %----------------------------------------
-\begin{itemize}
-\item data parallelism
-\item task parallelism
-\item hybrid parallelism
-\end{itemize}
+Three parallelization techniques are used in this paper and are described in this section, the data parallelism, the task parallelism, and the hybrid data and task parallelism.
+
+\paragraph{data parallelism} explain data decomposition and distribution, with a single program to apply on each processor
+
+\paragraph{task parallelism} explain task dag
+
+\paragraph{hybrid parallelism} explain the combination
 

### 2015-07-21 WOLFHPC15/article/ms_msc.tex
wolfhpc15 21_07

@@
 
-\paragraph{Creation of the synchronized ordered list of computations} From the ordered list of computations $\Gamma$ it is possible to detect needed synchronizations if a \emph{data parallelization} \textbf{(PB: to explain !!!)} is performed on the simulation.
+\paragraph{Creation of the synchronized ordered list of computations} From the ordered list of computations $\Gamma$ it is possible to detect needed synchronizations if a \emph{data parallelization} is performed on the simulation.
 A synchronization is needed each time a data read by a stencil computation has been written by a previous computation different from a boundary computation.

### 2015-07-21 WOLFHPC15/article/related.tex
wolfhpc15 21_07

%----------------------------------------
\subsection{Stencil solutions}
%----------------------------------------

Many solutions exists to ease parallel programming of stencil codes, as for example PATUS~\cite{citeulike12258902}, Pochoir~\cite{spaaTangCKLL11}, OP2~\cite{Giles2011}. Those solutions are powerfull, let the user implement their own stencil codes, and produce high performance codes. Those solutions can be considered as stencil compilers. As a result it produces optimized (cache tiling etc.) or parallel (CUDA, OpenCL etc.) codes for a single stencil computations.

A real case numerical simulation is most of the time composed of a set of numerical schemes (stencils computations), with one or more stencil shape (neighboorhood shape), and onto one or more data. Moreover, a numerical simulation also performs additionnal auxiliary computations which do not involve neighborhood, called local computations. Solutions like Pochoir or PATUS consider the parallelization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations. However, by not taking into account the parallelization of the overall numerical simulation, those compilers also reduce the codes to shared memory systems. Actually using a distributed memory system, the MPI (Message Passing Interface)~\cite{Graham2009MSE} library still have to be used by the user, as well as the distribution of the mesh onto the different distant processors.

%----------------------------------------
\subsection{Component models and control}
%----------------------------------------

%----------------------------------------
\subsection{Languages and component models}
%----------------------------------------

### 2015-07-21 WOLFHPC15/article/related.tex
wolfhpc15 21_07

@@
 
-A real case numerical simulation is most of the time composed of a set of numerical schemes (stencils computations), with one or more stencil shape (neighboorhood shape), and onto one or more data.
-Moreover, a numerical simulation also performs additionnal auxiliary computations which do not involve neighborhood, called local computations.
-Solutions like Pochoir or PATUS consider the parallelization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations.
-However, by not taking into account the parallelization of the overall numerical simulation, those compilers also reduce the codes to shared memory systems.
+However, a real case numerical simulation is not composed of a single stencil code as explained in Section~\ref{sect:multistencil}.
+A multi-stencil simulation computes more than one stencil computation, involving one or more stencil shapes, and additionnal local coputations.
+Solutions like Pochoir or PATUS handle the parallelization or the optimization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations.
+However, not taking into account the parallelization of the overall numerical simulation implies that those compilers are reduced to shared memory systems.
 Actually using a distributed memory system, the MPI (Message Passing Interface)~\cite{Graham2009MSE} library still have to be used by the user, as well as the distribution of the mesh onto the different distant processors.

### 2015-07-21 WOLFHPC15/article/related.tex
wolfhpc15 21_07

@@
 
+On the other hand, the work presented in this paper does not propose to optimize and parallelize a single stencil code for shared memory machines, GPGPUs or many-core architectures.
+The MS language and the MSC compiler produce a coarse-grain parallel structure of the overall simulation which could be combined to existing stencil compilers.
+
 %----------------------------------------

### 2015-07-22 WOLFHPC15/article/concept.tex
wolfhpc15 22_07

@@
 %----------------------------------------
-\subsection{Multi-stencil programs}
-\label{sect:multistencil}
+\subsection{Stencil codes}
+\label{sect:stencil}
 %----------------------------------------
@@
 
-If $x$ represents an element of the mesh, if $N(x)$ is a neighborhood shape around the element $x$, and if $u_t(x)$ is the data (or quantity) to write.
-Moreover if $V_t(x)$ is the set of data to simulate such that $V_t(x) \cap u_t(x) = \emptyset$, a stencil computation is of the following form
-
-\begin{equation} 
-\{u_{t-1}(x), u_{t-1}(N(x)), V_t(x), V_t(N(x)\} \longmapsto u_{t}(x).
-\label{space}
-\end{equation}
+\begin{figure}[!h]\begin{center}
+  \resizebox{9cm}{!}{\includegraphics{./images/stencil.pdf}}
+  \caption{Example of a stencil computation.}
+  \label{fig:ex}
+\end{center}\end{figure}
 
@@
 
+%----------------------------------------
+\subsection{Multi-stencil programs}
+\label{sect:multistencil}
+%----------------------------------------
 A real case numerical simulation is most of the time composed of a set of numerical schemes (stencils computations), with one or more stencil shape (neighboorhood shape), and onto one or more data.
@@
 
-\textbf{Do I need the following definitions?}
 \begin{mydef}
@@
 \end{mydef}
+
+A multi-stencil program in this work is defined as a sextuplet
+\begin{equation*} 
+\mathcal{MSP}(T,\mathcal{M},\mathcal{D}_m,\mathcal{D}_c,\Delta,\Gamma)
+\end{equation*}
 

### 2015-07-23 WOLFHPC15/article/concept.tex
wolfhpc15 23_07

--- 2015-07-21 WOLFHPC15/article/concept.tex
@@
+%----------------------------------------
+\subsection{Stencil codes}
+\label{sect:stencil}
+%----------------------------------------
+To numerically solve a set of PDEs, iterative methods (finite difference, finite volume, finite element methods etc.) are frequently used to approximate the solution through a discretized (step by step) phenomena.
+Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied onto a mesh (space discretization).
+In other words, in a mesh-based numerical simulation, the PDEs are transformed to a set of numerical computations applied at each time step on elements of the discretized space domain (the mesh).
+Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}~\cite{spaaTangCKLL11}.
+To ease programming of such stencil computations, and because of the relative simplicity of such a computation, many solutions (languages or libraries) propose an easy way to describe the stencil code, while producing transparently an optimized and parallelized stencil code.
+For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, proposing an optimized cache tiling technique in one case, and a separated parallelization strategy in the other case.
+
+Figure~\ref{fig:ex} gives two examples of stencil computations and introduce the different concepts used in this paper.
+First, a stencil code is based on a \emph{mesh} which is the discretization of the physical real domain.
+A mesh is a connected undirected graph without bridges (an edge is a bridge if its removal results in two disconnected graphs), where nodes and edges are linked to form cells (closure).
+For example, a structured mesh is illustrated in Figure~\ref{fig:ex}, each cell contains four nodes and four edges.
+Second, \emph{mesh domains} are subset elements of the mesh, such as for example the center of the cell, edges or nodes.
+In Figure~\ref{fig:ex} two mesh domains are used, the center of the cells and the edge on the horizontal axe (on the left and right of each cell).
+A \emph{data} is a quantity to simulate or a temporary data to perform computations.
+A data is mapped onto a given mesh domain.
+For example, in the figure, data $A$ and $B$ are mapped onto the center of cells, while $C$ is mapped onto the edges.
+
+\begin{figure}[!h]\begin{center}
+  \resizebox{7cm}{!}{\includegraphics{./images/stencil.pdf}}
+  \caption{Examples of a stencil computations}
+  \label{fig:ex}
+\end{center}\end{figure}
+
+The first stencil computation of Figure~\ref{fig:ex} computes $B$ using $A$, while the second computes $B$ using $C$.
+Thus, a stencil computation is defined by a set of input data (in those examples only one input data), and a single output data, the result.
+The \emph{computation domain} is a subset of elements to compute on the output data.
+In other words a computation domain is a subset of a mesh domain.
+For example, in the first stencil, the computation of $B$ is performed on the plain elements, while dotted elements are not computed.
+On the other hand, on the second computation, the computation domain of $B$ is exactly equal to the cell mesh domain.
+The computation domain represents the set of elements on which the computation is performed, in other words, the space loop.
+
+Finally, a stencil computation is composed of a \emph{numerical expression} which indicates how to compute an element of the computation domain (on the output data), using elements of the set of input data.
+The numerical expression of a stencil computation has the particularity to use not only the element to compute but also a \emph{neighborhood} of this element, also called a \emph{stencil shape}.
+For example, the stencil shape of the first computation of Figure~\ref{fig:ex} contains direct neighbors on the right, left, top and bottom.
+Sometimes, the neighborhood shape can also access a different mesh domain, as for example in the second computation example.
+Actually, in this computation, the neighborhood contains edges on the left and on the right.
+As a result, we could imagine of example that the numerical expression of the first example is
+\begin{equation*} 
+B(x,y) = A(x+1,y)+A(x-1,y)+A(x,y+1)+A(x,y-1),
+\end{equation*}
+and that the numerical expression of the second example is
+\begin{equation*} 
+B(x,y) = B(x,y)+C(x1,y1)+C(x1+1,y1).
+\end{equation*}
+
+To conclude thise section, we define a stencil $s$ computation as the quintuplet
+\begin{equation*} 
+s(R,w,exp,\mathcal{N},d_c),
+\end{equation*}
+where $R$ is the set of input data, $w$ the output data, $exp$ the numerical expression of the stencil computation, which is applied on the computation domain $d_c$, using a stencil shape of neighboorhood $\mathcal{N}$.
+
 %----------------------------------------
@@
 %----------------------------------------
-\begin{itemize}
-\item textual introduction on the problem, what is a stencil + example. What is a real simulation (more than one stencil and additionnal computations). 
-\item Existing solutions with prgramming freedom (instead of PETSc and Trilinos) = stencil compilers which give optimized code for one stencil code.
-Thus, for distributed memory and HPC, still have to use MPI between optimized codes to get an overall parallelization of the simulation.
-\item Short definition of a multi-stencil program and its computations (cite research report for more details)
-\end{itemize}
+If a stencil computation is well defined and a lot studied in the literrature, it is not the case for real case numerical simulations.
+In this paper, we define what we call a \emph{multi-stencil program}, which actually is a general numerical simulation.
+
+A real case numerical simulation is most of the time not composed of a single stencil computation, but of a set of stencil computations, with one or more stencil shapes (neighboorhood), and onto one or more data.
+Moreover, a numerical simulation also performs additionnal auxiliary computations which do not involve neighborhood, called local computations.
+In this paper is targetted this general case of numerical simulations that we call \emph{multi-stencil programs}.
+The work presented produces a general parallelization structure of the overall simulation, while most stencil solutions produce optimized codes.
+As it will be detailed in the related work, this work is complementary to the optimization and parallelization of a single stencil computation.
+
+An auxiliary computation is almost the same than a stencil computation, however its numerical expression does not involve a neighborhood, which makes this kind of computation purely local.
+A local computation $l$ is defined as the quadruplet
+\begin{equation*} 
+l(R,w,exp,d_c),
+\end{equation*}
+where $R$ is the set of input data, $w$ the output data, $exp$ the numerical expression of the stencil computation, which is applied on the computation domain $d_c$.
+
+A multi-stencil program in this work is defined as a sextuplet
+\begin{equation*} 
+\mathcal{MSP}(T,\mathcal{M},\mathcal{D}_m,\mathcal{D}_c,\Delta,\Gamma),
+\end{equation*}
+where $T$ is the set of time iteration to run the simulation, $\mathcal{M}$ is the mesh of the simulation, $\mathcal{D}_m$ is the set of mesh domains, $\mathcal{D}_c$ is the set of computation domains used into computations, $\Delta$ is the set of data, each one mapped onto a mesh domain, and finally $\Gamma$ is the set of computations.
+The set of computations $\Gamma$ is composed of an ordered list of stencil and local computations.
+One can notice that this work is limited to a single type of mesh in a given simulation.
 
@@
 %----------------------------------------
-\begin{itemize}
-\item data parallelism
-\item task parallelism
-\item hybrid parallelism
-\end{itemize}
+Three classical parallelization techniques are used in this paper and are described in this section, the data parallelism, the task parallelism, and the hybrid data and task parallelism.
+
+\paragraph{data parallelism} explain data decomposition and distribution, with a single program to apply on each processor
+
+\paragraph{task parallelism} explain task dag
+
+\paragraph{hybrid parallelism} explain the combination
 

### 2015-07-23 WOLFHPC15/article/concept.tex
wolfhpc15 23_07

@@
 %----------------------------------------
-Component model is an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228} which improves code re-use, scalability and maintainability of applications~\cite{Szyperski:2002:CSB:515228,bigot:inria-00388508}.
-Moreover, component models bring a separation of concerns in the application by a clear division of functionnalities in different components.
+Component model is an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228}, where functionnalities of an application are dispatched in different entities: components.
+Component models improves code re-use by a separation of concerns, scalability and maintainability of applications~\cite{Szyperski:2002:CSB:515228,bigot:inria-00388508}.
 In addition to this, recent work on component models have shown a simultaneous answer to performance, maintainability and portability of applications~\cite{l2c}, which makes this domain an interesting solution to bring maintainability and portability in HPC programming.
 
-Component-based software engineering~\cite{Szyperski:2002:CSB:515228} extends the concept of class by specifying in its interfaces not only the services it offers, or public methods, but also all its possible interactions with outer world, including the services it requires.
+Many component models exist, each of them with its own specification and singularity, like CCM~\cite{corba:omg06} (CORBA Component Model), GCM~\cite{Baude} (Grid Component Model) or CCA~\cite{Armstrong:1999:TCC:822084.823232} (Common Component Architecture), for example.
+However, conepts of component, port, interface and assembly are most of the time found in component models.
+In the rest of this paper will be used concepts of the component model $L^2C$~\cite{l2c}.
+$L^2C$ is a minimalist HPC-oriented component model where a component extends the concept of class by specifying in its interfaces not only the services it offers, but also all its possible interactions with outer world, including the services it requires.
 As a result, each component is an independant entity composed of a set of services its provides, and a set of services it requires (and uses).
 
-A \emph{port} is an entity embbedded in the component which makes possible an \emph{assembly} of components.
+A \emph{port} in $L^2C$ is an entity embbedded in the component which makes possible an \emph{assembly} of components.
 An assembly of components (of instanciated components) is a way to actually connect components together and to produce a complete application, as a set of components and their interactions.
@@
 It is also possible to group more than one required service into a single port, as a list, called a \emph{use-multiple} port.
-As illustrated in Figures~\ref{fig:ports}, a provide port will be represented by a white circle, a use port by a black circle and a use-multiple port by a black circle with a white $m$ in it.
+As illustrated in Figures~\ref{fig:ports}, a provide port will be represented by a white circle on the left of the component, a use port by a black circle and a use-multiple port by a black circle with a white $m$ in it, both on the right of the component.
 
@@
    \node[use] (u) at (1.5,0) {};
-   \node[use] (um) at (0,1) {$m$};
-   \node[provide,right=of u] (p1) {};
-   \node[component,right=of p1] (C1) {$Comp_1$};
+   \node[provide,right=1.5cm of u] (p1) {};
+   \node[component,right=1.5cm of p1] (C1) {$Comp_1$};
+   \node[use,right=1.5cm of C1] (um) {$m$};
 
@@
     (p) edge node {} (C)
-    (C) edge node {} (u)
-    	edge node {} (um)
+    (C.east) edge node {} (u)
+    (C1)	edge node {} (um)
     (p1) edge node {} (C1);
@@
 In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}.
+As the use port if on the right of its component, and the provide port on the left, in a component assembly, the component on the left uses the provide port of the right component.
 
@@
    \node[use] (u) at (1.5,0) {};
-   \node[use] (um) at (0,1) {$m$};
-   \node[component,right=of u] (C1) {$Comp_1$};
+   \node[component,right=1.5 of u] (C1) {$Comp_1$};
+   \node[use,right=1.5 of C1] (um) {$m$};
 
@@
     (C) edge node {} (u)
-    	edge node {} (um)
+    (C1)	edge node {} (um)
     (u) edge node {} (C1);
@@
 
-Many component models exist, each of them with its own specifications and functionnalities, like CCM~\cite{corba:omg06} (CORBA Component Model), GCM~\cite{Baude} (Grid Component Model) or CCA~\cite{Armstrong:1999:TCC:822084.823232} (Common Component Architecture) for example.
-However, conepts of component, port, interface and assembly are common to many component models.
-The rest of this paper will only use those concepts to present the MS Language and its compiler whatever the component model is.
+With its minimal set of ports, $L^2C$ is a light component model which does not introduce overheads at runtime, and which results in a HPC-oriented component model.
+Another HPC-oriented component model is CCA~\cite{Armstrong:1999:TCC:822084.823232}, however the main difference between those component models is the way a dsitribution of component on distributed architectures is managed.
+Actually, in CCA, MPI~\cite{} (Message Passing Interfaces) have to be used and managed manually inside components, while $L^2C$ proposes a specific MPI port to link distant components and ease their communications.
+Moreover, by exposing MPI connections, $L^2C$ also exposes the parallel structure of a program inside an assembly, which increase separation of concerns and code reuse for parallel applications.

### 2015-07-23 WOLFHPC15/article/concept.tex
wolfhpc15 23_07

@@
 Three classical parallelization techniques are used in this paper and are described in this section, the data parallelism, the task parallelism, and the hybrid data and task parallelism.
+Those parallelization techniques are independent from the actual parallel hardware used.
 
-\paragraph{data parallelism} explain data decomposition and distribution, with a single program to apply on each processor
+\paragraph{Data parallelism} The idea of this parallelization technique is to split, or partition, the data on which computations are applied among available processors (or cores).
+Each processor then applies the same progam or instruction onto its subpart of data.
+If a neighborhood information is needed from another processor communications or synchronizations are performed.
 
-\paragraph{task parallelism} explain task dag
+In the domain of numerical simulations, this technique is most of the time called a domain decomposition.
+This parallelization technique produce efficient programs up to thousands processors or cores, but on certain conditions.
+First, each subpart of data has to be big enough to overlap communication time.
+Second, the partitioning of data has to be balanced among processors.
+Thus, if this parallelization technique is clearly adequate to structured meshes, easy to balance, it is not for unstructured meshes or irregular structure where the amount of work is not heterogenous.
 
-\paragraph{hybrid parallelism} explain the combination
+\paragraph{Task parallelism} Another well known parallelization technique is to identify in a program the different tasks and which one are independant and can be launched concurrently.
+Most of the time, such a prallelization technique create a dependency directed acyclic graph (dependency dag) from a set of ordered tasks.
+Dependencies are found from read/write information for each task.
+Actually, if a task $i$ write a data $a$ and if a task $j$ read $a$, then $i$ has to be finished before $j$ is performed.
+From a dependency graph too different solutions are available.
+First, a static schedule of tasks is built, which could be a good solution if the tasks are regular, or use a dynamic scheduler to dynamically decide at runtime which task is executed on which processor.
+
+\paragraph{Hybrid parallelism} Finally, it is also possible to combine both those parallelization techniques to get what is called an hybrid parallelization.
+The interest of an hybrid parallelization is to bring another source of parallelism in a program if limits of a given technique are reach.
 
@@
 Another HPC-oriented component model is CCA~\cite{Armstrong:1999:TCC:822084.823232}, however the main difference between those component models is the way a dsitribution of component on distributed architectures is managed.
-Actually, in CCA, MPI~\cite{} (Message Passing Interfaces) have to be used and managed manually inside components, while $L^2C$ proposes a specific MPI port to link distant components and ease their communications.
+Actually, in CCA, MPI~\cite{Graham2009MSE} (Message Passing Interfaces) have to be used and managed manually inside components, while $L^2C$ proposes a specific MPI port to link distant components and ease their communications.
 Moreover, by exposing MPI connections, $L^2C$ also exposes the parallel structure of a program inside an assembly, which increase separation of concerns and code reuse for parallel applications.

### 2015-07-23 WOLFHPC15/article/mscac.tex
wolfhpc15 23_07

@@
+The solution presented in this paper is a domain specific language, to describe a multi-stencil program (a numerical simulation), and a compiler which transforms this description to a parallel component-based structure of the overall simulation.
+
 The MS language (MSL) is an agnostic descriptive language for multi-stencil simulations.
-Agnostic in our context means that without any knowledge on the mesh type and the neighborhood description of stencil computations, the language consists in data and computations descriptions.
-From this agnostic and descriptive language can be compiled the structure of the parallel simulation in two ways:
-\begin{itemize}
-\item data parallelism;
-\item hybrid (data and task) parallelism.
-\end{itemize}
+Agnostic means that the description of a numerical simulation, using MSL, does not need any knowledge on the mesh type, the stencil shape description, and the final implementation constraints.
+Descriptive, on the other hand, means that the terminal states of the language grammar are string identifiers and that the language does not handle description of numerical computations, but description of a multi-stencil simulation, in a coarse-grain fashion.
 
-As illustrated in Figure~\ref{fig:mscac}, the MS language compiler, called MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of two different part:
- \begin{itemize}
-\item first, the computation compiler, called MSC, which is responsible for the extraction of the computations parallelization;
-\item second, the data compiler, called MSD, which handles a distributed data structure, data mapped onto it, and their relation with the computation part of the final program.
-\end{itemize}
-
-MSCAC produces from the MSL source file the structure of the parallel simulation as a component assembly.
-Thus, the coarse-grain parallelization of the simulation is produced.
-In this final component assembly, an additionnal work is asked to the user: write computation components, using interfaces of the distributed data structure choosen by MSD.
-Thus, this work is complementary to stencil compilers or to distributed data structure solutions.
-Actually, it produces a ready-to-fill parallel structure pattern of the simulation.
+As illustrated in Figure~\ref{fig:mscac}, the MS language compiler, called MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of two different parts.
+First, the computation compiler, called \emph{MSC}, is responsible for the computations parallelization and for its dump to a component assembly; second, the data compiler, called \emph{MSD}, handles a distributed data structure, data mapped onto it, and their relations with computations, as a final overall component assembly.
 
@@
 
-This paper presents the MS language and the MSC part of the compiler. The evaluation part will shortly explain how the MSD part of the compiler has been acheived for this paper.
+This paper presents the MS language and the MSC part of the compiler.
+The MSC compiler produces a static scheduling of the set of computations, ready to dump to a parallel language, such as OpenMP~\cite{660313} or HPF~\cite{219857}, but also ready to dump to a component assembly which is the main contribution of this work.
+
+It has to be noticed that in this paper, the MSD compiler is a first adhoc version which works for a precise distributed data structure, the SkelGIS labrary~\cite{HeleneLS13,HeleneLS14,HeleneEuroPar14,CPE:CPE3494}.
+However, using the system of python string templates~\footnote{\url{https://docs.python.org/2/library/string.html}} to define the data component assembly, it is already possible to use another distributed data structure, as for example Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.
+This part of the compiler takes place in a larger research project to propose domain decomposition skeletons using component models.
+
+The overall compiler MSCAC produces a ready-to-fill component-based parallel structure of the simulation.
+In this final component assembly, it is needed to write computation components, using interfaces of the distributed data structure choosen by MSD.
+Thus, this work is complementary to stencil compilers or to distributed data structure solutions.

### 2015-07-23 WOLFHPC15/article/concept.tex
wolfhpc15 23_07

@@
 To conclude thise section, we define a stencil $s$ computation as the quintuplet
-\begin{equation*} 
+\begin{equation} 
 s(R,w,exp,\mathcal{N},d_c),
-\end{equation*}
+\label{eq:st}
+\end{equation}
 where $R$ is the set of input data, $w$ the output data, $exp$ the numerical expression of the stencil computation, which is applied on the computation domain $d_c$, using a stencil shape of neighboorhood $\mathcal{N}$.
@@
 A local computation $l$ is defined as the quadruplet
-\begin{equation*} 
+\begin{equation} 
 l(R,w,exp,d_c),
-\end{equation*}
+\label{eq:loc}
+\end{equation}
 where $R$ is the set of input data, $w$ the output data, $exp$ the numerical expression of the stencil computation, which is applied on the computation domain $d_c$.
@@
 A multi-stencil program in this work is defined as a sextuplet
-\begin{equation*} 
+\begin{equation} 
 \mathcal{MSP}(T,\mathcal{M},\mathcal{D}_m,\mathcal{D}_c,\Delta,\Gamma),
-\end{equation*}
+\label{eq:msp}
+\end{equation}
 where $T$ is the set of time iteration to run the simulation, $\mathcal{M}$ is the mesh of the simulation, $\mathcal{D}_m$ is the set of mesh domains, $\mathcal{D}_c$ is the set of computation domains used into computations, $\Delta$ is the set of data, each one mapped onto a mesh domain, and finally $\Gamma$ is the set of computations.

### 2015-07-23 WOLFHPC15/article/ms_msc.tex
wolfhpc15 23_07

@@
 The MS language (MSL) is an agnostic descriptive language for multi-stencil simulations.
-Agnostic means that no knowledge on the mesh type and the neighborhood description of stencil computations are needed in the language.
-As the language consists in a description of the simulation, without any computation details and programming details, and as no computation is directly done from the language, this language is also called a descriptive language.
-It is composed of five description sections:
-\begin{enumerate}
-\item meshes description;
-\item domains description;
-\item data description;
-\item time loop description;
-\item computations description.
-\end{enumerate}
+MSL follows the definition of a multi-stencil program~(\ref{eq:msp}), and of stencil and local computations~(\ref{eq:st}) and~(\ref{eq:loc}).
+Thus, it is composed of six sections: the mesh description; the mesh domains description; the computation domains description; the data description; the time loop description; and finally, the computations description.
 
-This paper deals with the MSC compiler which only needs the three last enumerations above.
-For this reason we describe in the rest of this section MSL for data, time loop and computations descriptions.
+The overall grammar of a MSL program is given in Figure~\ref{fig:grammar}, where the terminals $meshid$, $meshdom$, $compdomid$, $dataid$, $compid$ are string identifiers, and $num$ is an integer value.
+An example is then given to explain each section.
 
-\paragraph{Data description} Data description starts with the tag $data:$.
-Each line after this tag represents a single data description.
-This description is composed of an identification name, and the domain identification name it is applied on.
+\begin{figure}[h!]
+\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
+program ::= "mesh:" meshid 
+            "mesh domains:" listmeshdom
+            "computation domains:" 
+                       listcompdom
+            "intersections:"
+                       listintersection
+            "data:" listdata
+            "time:" iteration
+            "computations:" listcomp
+listmeshdom ::= meshdom listmeshdom
+             |  meshdom
+listcompdom ::= compdom listcompdom
+             |  compdom
+compdom ::= compdomid "in" listmeshdom
+listintersection ::= inter listintersections
+                  |  inter
+inter ::= compdomid and compdomid
+listdata ::= data listdata
+          |  data
+data ::= dataid "," meshdom
+iteration ::= num
+listcomp ::= comp listcomp
+          |  comp
+comp ::= comptype ":" compid "=({" 
+            listdataid "}," dataid 
+            "," compdomid ")"
+listdataid ::= dataid listdataid
+            |  dataid
+comptype ::= "stencil"
+      |  "local"
+      |  "bound"
+\end{lstlisting}
+\caption{Grammar of the MS language}
+\label{fig:grammar}
+\end{figure}
 
-\begin{lstlisting}[basicstyle=\footnotesize,mathescape,frame=single,language=C++]
-data:
-  name_id,domain_id
+\paragraph{Mesh and mesh domains} A mesh is defined by a string identifier.
+As already explain, this work is for now limited to a single mesh.
+Each mesh domain is also represented by a string identifier.
+For example
+\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
+mesh: cart
+mesh domains: cell,edgex,edgey
 \end{lstlisting}
 
-\paragraph{Time description} The time of a numerical simulation consists in a set of iterations.
-The end of the simulation time can be controlled by a convergence function or by a manual configuration of the number of iteration.
-In this first version of MSL, the iteration number is handled.
+\paragraph{Computation domains} A computation domain can be seen as a subpart of a meh domain which will be used during computation phase.
+For each computation domain is indicated the mesh domain identifier from which it is built.
+After definitions of computation domains, the intersections between those are given.
+For example
+\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
+computation domains:
+  allcell in cell
+  alledgex in edgex
+  alledgey in edgey
+  partedgex in edgex
+interactions:
+  alledgex and partedgex
+\end{lstlisting}
+In this example, an entire computation domain is defined for each mesh domain, and additionnally a part of the mesh domain edgex is defined.
+As a result computation domains alledgex and partedgex intersect.
 
-\begin{lstlisting}[basicstyle=\footnotesize,mathescape,frame=single,language=C++]
-time:nb_iteration
+\paragraph{Data and time} A data is a quantity to simulate and is mapped onto a mesh domain. The time section simply indicates a number of iterations to perform in the simulation. For example:
+\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
+data:
+  a,cell
+  b,cell
+  c,edgex
+  d,edgex
+  e,edgey
+  f,cell
+  g,edgey
+  h,edgex
+  i,cell
+  j,edgex
+time:500
 \end{lstlisting}
@@
 \paragraph{Computations description} The most important part of MSL is the description of the computations of the simulation.
-The language follows the definitions given in Section~\ref{sect:multistencil}.
+The language follows the definitions given in Section~\ref{sect:multistencil}, however as MSL is not made to express numerical computations, the element $exp$ of the definition is not given in the language.
+Moreover, an additionnal information is asked to the user, the type of computation.
+A computation can be \emph{stencil}, \emph{local} or \emph{bound}.
+\textbf{OU INTRODUIRE bound et refs?}
 Thus, the description of a computation consists of a type of computation, an identification name of the computation, a set of identifiers for the data read by the computation, a single identifier for the data written by the computation.
 Moreover, if the type of computation is a stencil, a neighborhood identifier is needed.
-
-\begin{lstlisting}[basicstyle=\footnotesize,mathescape,frame=single,language=C++]
+\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
 computations:
-  type:name_id({read1_id,read2_id,...},
- 	    written_id[,neighborhood_id])
+  local:c0({a},b)
+  stencil:c1({b},c,n1)
+  local:c2({c},d)
+  local:c3({c},e)
+  stencil:c4({d},f)
+  local:c5({e},g)
+  local:c6({f},h)
+  local:c7({g,h},i)
+  stencil:c8({i},j,n1)
 \end{lstlisting}
@@
 \item creation of $\Gamma$, the ordered list of computations, from the MSL input file;
-\item creation of $\Gamma_{data}$ to insert synchronizations in the ordered list of computations;
+\item creation of $\Gamma_{data}$ the ordered list of computations including synchronizations;
 \item creation of the dependency graph $\Gamma_{hybrid}$;
-\item transform $\Gamma_{hybrid}$ to a minimal series-parallel graph $\Gamma_{msp}$;
+\item transformation of $\Gamma_{hybrid}$ to a minimal series-parallel graph $\Gamma_{msp}$;
 \item creation of the series-parallel tree decomposition $\Gamma_{tsp}$;

### 2015-07-23 WOLFHPC15/article/ms_msc.tex
wolfhpc15 23_07

@@
                        listcompdom
-            "intersections:"
-                       listintersection
+            "no intersections:"
+                       listnointer
             "data:" listdata
@@
 compdom ::= compdomid "in" listmeshdom
-listintersection ::= inter listintersections
-                  |  inter
-inter ::= compdomid and compdomid
+listnointer ::= nointer listnointer
+                  |  nointer
+nointer ::= compdomid and compdomid
 listdata ::= data listdata
@@
           |  comp
-comp ::= comptype ":" compid "=({" 
-            listdataid "}," dataid 
-            "," compdomid ")"
+comp ::= dataid "=" compid "=({" 
+            listdataid "}," compdomid 
+            ["," neighborid] ")"
 listdataid ::= dataid listdataid
@@
 For each computation domain is indicated the mesh domain identifier from which it is built.
-After definitions of computation domains, the intersections between those are given.
+By default two computation domains always intersect except when precise in the section \emph{no intersections}.
 For example
@@
   alledgey in edgey
-  partedgex in edgex
-interactions:
-  alledgex and partedgex
+  part1edgex in edgex
+  part2edgex in edgex
+no intersections:
+  part1edgex and part2edgex
 \end{lstlisting}
-In this example, an entire computation domain is defined for each mesh domain, and additionnally a part of the mesh domain edgex is defined.
-As a result computation domains alledgex and partedgex intersect.
+In this example, an entire computation domain is defined for each mesh domain, and additionnally two subpart of the mesh domain edgex are defined.
+It is then precised that part1edgex and part2edgex do not intersect.
 
@@
 Moreover, an additionnal information is asked to the user, the type of computation.
-A computation can be \emph{stencil}, \emph{local} or \emph{bound}.
-\textbf{OU INTRODUIRE bound et refs?}
+A computation can be \emph{stencil}, \emph{local} and if a stencil is computed a neighborhood identifier is given.
 Thus, the description of a computation consists of a type of computation, an identification name of the computation, a set of identifiers for the data read by the computation, a single identifier for the data written by the computation.
@@
 computations:
-  local:c0({a},b)
-  stencil:c1({b},c,n1)
-  local:c2({c},d)
-  local:c3({c},e)
-  stencil:c4({d},f)
-  local:c5({e},g)
-  local:c6({f},h)
-  local:c7({g,h},i)
-  stencil:c8({i},j,n1)
+  b=c0({a},allcell)
+  c=c1({b},alledgex,n1)
+  d=c2({c},alledgex)
+  e=c3({c},alledgey)
+  f=c4({d},allcell,n1)
+  g=c5({e},alledgey)
+  h=c6({f},alledgex)
+  i=c7({g,h},allcell)
+  j=c8({i},partedgex,n1)
 \end{lstlisting}
@@
 From the ordered lines of computations, it builds the ordered list of computations $\Gamma$ to perform in the simulation.
+For example the ordered list of computation of the example above is $[c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$.
 
 \paragraph{Creation of the synchronized ordered list of computations} From the ordered list of computations $\Gamma$ it is possible to detect needed synchronizations if a \emph{data parallelization} is performed on the simulation.
-A synchronization is needed each time a data read by a stencil computation has been written by a previous computation different from a boundary computation.
+A synchronization is needed each time a data read by a stencil computation has been written by a previous computation and if the computation domains intersect.
 Actually, as a stencil computation needs neighborhood values, in a domain decomposition values computed by different processors or cores are needed.
 
-\paragraph{Creation of the dependency graph} From the synchronized ordered list of computations $\Gamma_{data}$, a dependency graph can be build.
-A dependency exists between two computations (including synchronizations) if and only if a data read has been written by a previous computation in $\Gamma_{data}$.
-Nodes of the dependency graph are computations and synchronizations, while edges are dependencies between them.
-$\Gamma_{hybrid}$ is a directed acyclic graph (\emph{dag}) as it is build from an ordered list.
-
-\paragraph{Transformation to a minimal series-parallel graph} As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if the forbidden \emph{N-shape} is not found in the graph.
-To transform $\Gamma_{hybrid}$ to the minimal series-parallel graph $\Gamma_{msp}$, a BFS algorithm is applied to build a complete bipartite between two levels of $\Gamma_{hybrid}$~\cite{Mitchell:2004:CMV:1082101.1082117}.
-
-\paragraph{Creation of a tree decomposition} As shown in different works~\cite{Valdes:1979:RSP:800135.804393,Schoenmakers95anew}, from a minimal series-parallel graph a tree decomposition can be built.
-A series-parallel tree decomposition consists in the decomposition of the graph as a set of \emph{sequences} and \emph{parallel} sections in a tree.
-
-\paragraph{Dump to a component assembly} From the series-parallel tree decomposition $\Gamma_{tsp}$ can be built a program using any parallel language containing \emph{parallel sections} and \emph{sequences of instructions} such as OpenMP~\cite{660313}, HPF~\cite{219857}, UPC~\cite{El-Ghazawi:2006:UUP:1188455.1188483} etc.
-In this paper we study the dump of $\Gamma_{tsp}$ to a component assembly which needs the introduction of \emph{control components} described in the next section.
-
-For more details on MSC, a full description of the compilation is described in~\cite{}.
-
-%----------------------------------------
-\subsection{Example}
-%----------------------------------------
-In this Section we give an example of an input file written with MSL, the $\Gamma_{tsp}$ series-parallel tree decomposition computed and the final component dump.
-
-The input MSL file, shown in Figure~\ref{fig:msinput} is composed of nine computations and ten data.
-
-\begin{figure}[h!]
-\begin{lstlisting}[basicstyle=\footnotesize,mathescape,frame=single,language=C++]
-data:
-  a,d1
-  b,d1
-  c,d2
-  d,d2
-  e,d3
-  f,d1
-  g,d3
-  h,d2
-  i,d1
-  j,d2
-time:500
-computations:
-  local:c0({a},b)
-  stencil:c1({b},c,n1)
-  local:c2({c},d)
-  local:c3({c},e)
-  stencil:c4({d},f)
-  local:c5({e},g)
-  local:c6({f},h)
-  local:c7({g,h},i)
-  stencil:c8({i},j,n1)
-\end{lstlisting}
-\caption{Example of MSL input file}
-\label{fig:msinput}
-\end{figure}
-
-The first step of the compilation is to build the synchronized ordered list of computations $\Gamma_{data}$.
-In this example each stencil computation actually needs a synchronization.
 For example, the stencil computation $c_1$ read the data $b$ which has been written by the computation $c_0$.
+Moreover the mesh domains are different which means that the computation domains intersect.
 For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed to the sublist $[c_0,sync_1,c_1]$ in $\Gamma_{data}$.
@@
 
-Once the synchronized list $\Gamma_{data}$ is obtained, the dependency graph $\Gamma_{hybrid}$ is built and illustrated in Figure~\ref{fig:hyb}.
-Each computation (and synchronization) is a node of the graph, while each edge represent a read/write data dependency between computations.
+\paragraph{Creation of the dependency graph} From the synchronized ordered list of computations $\Gamma_{data}$, a dependency graph can be build.
+A dependency exists between two computations (including synchronizations) if and only if a data read has been written by a previous computation in $\Gamma_{data}$, and if the computation domains intersect.
+In other words, the only case where a data dependency exists but is not taking into account is when the computation domains do not intersect.
+Nodes of the dependency graph represent computations and synchronizations, while edges are dependencies between them.
+$\Gamma_{hybrid}$ is a directed acyclic graph (\emph{dag}) as it is build from an ordered list.
+For example, the dependency dag $\Gamma_{hybrid}$ of the example is illustrated in Figure~\ref{fig:hyb}.
 
@@
 
-After a set of transformation briefly explained in the previous Section, a series-parallel tree decomposition $\Gamma_{tsp}$ is obtained and illustrated in Figure~\ref{fig:canon}.
+\paragraph{Transformation to a minimal series-parallel graph} As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if the forbidden \emph{N-shape} is not found in the graph.
+To transform $\Gamma_{hybrid}$ to the minimal series-parallel graph $\Gamma_{msp}$, a breadth first search~\cite{} traversing algorithm is applied to build a complete bipartite graph between two levels of $\Gamma_{hybrid}$~\cite{Mitchell:2004:CMV:1082101.1082117}.
+
+\paragraph{Creation of a tree decomposition} As shown in different works~\cite{Valdes:1979:RSP:800135.804393,Schoenmakers95anew}, from a minimal series-parallel graph a tree decomposition can be built.
+A series-parallel tree decomposition consists in the decomposition of the graph as a set of \emph{sequences} and \emph{parallel} sections in a tree.
+For example, the series-parallel tree decomposition $\Gamma_{tsp}$ of the example is illustrated in Figure~\ref{fig:canon}.
 
@@
    \node[] (star1) at (-2,1) {$sync_1$};
-   \node[] (c1) at (-1,1) {$c_1^*$};
+   \node[] (c1) at (-1,1) {$c_1$};
 
@@
    \node[] (p1) at (-2,3) {$\mathcal{P}$};
-   \node[] (c4) at (-1,3) {$c_4^*$};
+   \node[] (c4) at (-1,3) {$c_4$};
    \node[] (c6) at (-0,3) {$c_6$};
@@
    \node[] (star8) at (2,1) {$sync_8$};
-   \node[] (c8) at (3,1) {$c_8^*$};
+   \node[] (c8) at (3,1) {$c_8$};
 
@@
 \end{figure}
+
+\paragraph{Dump to a component assembly} From the series-parallel tree decomposition $\Gamma_{tsp}$ can be built a program using any parallel language containing \emph{parallel sections} and \emph{sequences of instructions} such as OpenMP~\cite{660313}, HPF~\cite{219857}, UPC~\cite{El-Ghazawi:2006:UUP:1188455.1188483} etc.
+In this paper we study the dump of $\Gamma_{tsp}$ to a component assembly which needs the introduction of \emph{control components} described in the next section.
+
+For more details on MSC, a full description of the compilation is described in the research report~\cite{}.

### 2015-07-23 WOLFHPC15/article/ms_msc.tex
wolfhpc15 23_07

@@
                        listcompdom
-            "no intersections:"
-                       listnointer
+            "independent:"
+                       listinde
             "data:" listdata
@@
 compdom ::= compdomid "in" listmeshdom
-listnointer ::= nointer listnointer
-                  |  nointer
-nointer ::= compdomid and compdomid
+listinde ::= inde listinde
+          |  inde
+inde ::= compdomid and compdomid
 listdata ::= data listdata
@@
   part2edgex in edgex
-no intersections:
+independent:
   part1edgex and part2edgex

### 2015-07-24 WOLFHPC15/article/ms_msc.tex
wolfhpc15 24_07

@@
           |  comp
-comp ::= dataid "=" compid "=({" 
-            listdataid "}," compdomid 
-            ["," neighborid] ")"
-listdataid ::= dataid listdataid
-            |  dataid
-comptype ::= "stencil"
-          |  "local"
-          |  "bound"
+comp ::= dataid "[" compdomid "]=" compid "(" 
+            listdataread ")"
+listdataread ::= dataread listdataread
+              |  dataread
+dataread ::= dataid "[" neighborid "]"
 \end{lstlisting}
@@
 The language follows the definitions given in Section~\ref{sect:multistencil}, however as MSL is not made to express numerical computations, the element $exp$ of the definition is not given in the language.
-Moreover, an additionnal information is asked to the user, the type of computation.
-A computation can be \emph{stencil}, \emph{local} and if a stencil is computed a neighborhood identifier is given.
-Thus, the description of a computation consists of a type of computation, an identification name of the computation, a set of identifiers for the data read by the computation, a single identifier for the data written by the computation.
-Moreover, if the type of computation is a stencil, a neighborhood identifier is needed.
+Moreover, to propose a syntax close to an imperative code, information are given in a different order.
+Thus, the description of a computation consists of the identifier of the data to compute, the computation domain (which is a subpart of the mesh domain of the data) on which it is computed into the brackets, an identification name of the computation, and a set of identifiers for the data read by the computation.
+For each data read by the computation brackets indicate the stencil shape identifier used on the data.
+If the computation is local brackets stay empty.
+In the example below, the data $b$ is computed on the computation domain $allcell$ by the computation $c_0$ which read data $a$ without neighborhood shape.
 \begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
 computations:
-  b=c0({a},allcell)
-  c=c1({b},alledgex,n1)
-  d=c2({c},alledgex)
-  e=c3({c},alledgey)
-  f=c4({d},allcell,n1)
-  g=c5({e},alledgey)
-  h=c6({f},alledgex)
-  i=c7({g,h},allcell)
-  j=c8({i},partedgex,n1)
+  b[allcell]=c0(a[])
+  c[alledgex]=c1(b[n1])
+  d[alledgex]=c2(c[])
+  e[alledgey]=c3(c[])
+  f[allcell]=c4(d[n1])
+  g[alledgey]=c5(e)
+  h[alledgex]=c6(f)
+  i[allcell]=c7(g[],h[])
+  j[partedgex]=c8(i[n1])
 \end{lstlisting}

### 2015-07-24 WOLFHPC15/article/component.tex
wolfhpc15 21_07
wolfhpc15 24_07

@@
 Those components can be used for any case of simulation which increases code reuse between simulations.
-A control component is composed of a single \emph{provide} port linked to a single execution method, most of the time called \emph{go}.
+A control component is composed of a single \emph{provide} port linked to a single execution method, most of the time called the \emph{go} method.
 We introduce three types of control components described bellow:
@@
 The role of this component is to call an ordered list of other components.
-For this reason, it is composed of a \emph{use-multiple} port connected to the components to call as a sequence.
+For this reason, it is composed of an ordered \emph{use-multiple} port connected to the components to call as a sequence.
 
@@
 
-  \path[->]
+  \path[-]
     (p) edge node {} (seq)
@@
 
-  \path[->]
+  \path[-]
     (p) edge node {} (sync)
@@
 In the computation part of the simulation a final component is needed, the computation component, denoted K.
-This type of component represent a single computation of the simulation, i.e.
+This type of component represents a single computation of the simulation, i.e.
 the computation of a single output data (written) using a set of input data (read).
-Using a \emph{use-multiple}, however, a list of component is used and the explicit identification of each used component is lost.
+Using a \emph{use-multiple}, however, a list of component is used and the explicit identification of each component to use is lost.
 If for SEQ or PAR, the identification of components is not usefull, it is usefull to manipulate data in a computation.
 For this reason a computation component has as much \emph{use} ports as input and output data to manipulate.
-We denote this in the bellow representation as a star $*$ on the use port.
+We denote this in the bellow representation as a star on the use port.
 
@@
 
-  \path[->]
+  \path[-]
     (p) edge node {} (k)
@@
 As an example, from $\Gamma_{tsp}$ illustrated in Figure~\ref{fig:canon} a direct dump example is illustrated in Figure~\ref{fig:assembly}.
-In this Figure, the ports linked to data (ports of SYNC and K components) are represented but not their connections.
+In this Figure, the ports linked to data (use and use-multiple ports of SYNC and K) are represented but are not connected.
 Moreover, each computation component is an instance of the component K presented before but takes the identification name of the computation in the MSL file.
 
-\begin{figure*}[h!]
+\begin{figure*}
 \begin{center}
@@
 
-   \path[->]
+   \path[-]
    %seq0

### 2015-07-24 WOLFHPC15/article/concept.tex
wolfhpc15 24_07

@@
 Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}~\cite{spaaTangCKLL11}.
-To ease programming of such stencil computations, and because of the relative simplicity of such a computation, many solutions (languages or libraries) propose an easy way to describe the stencil code, while producing transparently an optimized and parallelized stencil code.
-For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, proposing an optimized cache tiling technique in one case, and a separated parallelization strategy in the other case.
+To ease programming of stencil computations, and because of the relative regularity of such a computation, many solutions (languages or libraries) propose an easy way to describe the stencil code, while producing transparently an optimized and parallelized stencil code.
+For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, respectively proposing an optimized cache tiling technique, and a separation of concerns by separating parallelization strategy from the computation description.
 
@@
 For example, a structured mesh is illustrated in Figure~\ref{fig:ex}, each cell contains four nodes and four edges.
-Second, \emph{mesh domains} are subset elements of the mesh, such as for example the center of the cell, edges or nodes.
-In Figure~\ref{fig:ex} two mesh domains are used, the center of the cells and the edge on the horizontal axe (on the left and right of each cell).
+Second, \emph{mesh domains} are subset elements of the mesh, such as the center of cells, edges or nodes, for example.
+In Figure~\ref{fig:ex} two mesh domains are used, the center of the cells (in red) and the edge on the horizontal axe (in blue on the left and right of each cell).
 A \emph{data} is a quantity to simulate or a temporary data to perform computations.
@@
 
-The first stencil computation of Figure~\ref{fig:ex} computes $B$ using $A$, while the second computes $B$ using $C$.
+The first stencil computation of Figure~\ref{fig:ex} computes $B$ using $A$, while the second one computes $B$ using $C$.
 Thus, a stencil computation is defined by a set of input data (in those examples only one input data), and a single output data, the result.
@@
 In other words a computation domain is a subset of a mesh domain.
-For example, in the first stencil, the computation of $B$ is performed on the plain elements, while dotted elements are not computed.
+For example, in the first stencil, the computation of $B$ is performed on plain elements, while dotted elements are not computed.
 On the other hand, on the second computation, the computation domain of $B$ is exactly equal to the cell mesh domain.
@@
 Sometimes, the neighborhood shape can also access a different mesh domain, as for example in the second computation example.
-Actually, in this computation, the neighborhood contains edges on the left and on the right.
-As a result, we could imagine of example that the numerical expression of the first example is
+Actually, in this computation, the neighborhood contains edges on the left and on the right of the cell.
+As a result, we could imagine for example that the numerical expression of the first example is
 \begin{equation*} 
@@
 
-To conclude thise section, we define a stencil $s$ computation as the quintuplet
+To conclude this section, we define a stencil computation $s$ as the quintuplet
 \begin{equation} 
@@
 \end{equation}
-where $R$ is the set of input data, $w$ the output data, $exp$ the numerical expression of the stencil computation, which is applied on the computation domain $d_c$, using a stencil shape of neighboorhood $\mathcal{N}$.
+where $R$ is the set of input data, $w$ the output data, $exp$ the numerical expression of the stencil computation, which is applied on the computation domain $d_c$, using a stencil shape $\mathcal{N}$.
 
@@
 %----------------------------------------
-If a stencil computation is well defined and a lot studied in the literrature, it is not the case for real case numerical simulations.
+If stencil computations are well defined and a lot studied in the literrature, it is not the case for real case numerical simulations.
 In this paper, we define what we call a \emph{multi-stencil program}, which actually is a general numerical simulation.
@@
 In this paper is targetted this general case of numerical simulations that we call \emph{multi-stencil programs}.
-The work presented produces a general parallelization structure of the overall simulation, while most stencil solutions produce optimized codes.
+The work presented produces a general parallelization structure of the overall simulation, while most stencil solutions produce optimized codes for a single stencil code.
 As it will be detailed in the related work, this work is complementary to the optimization and parallelization of a single stencil computation.
@@
 
-\paragraph{Data parallelism} The idea of this parallelization technique is to split, or partition, the data on which computations are applied among available processors (or cores).
+\paragraph{Data parallelism} The idea of this parallelization technique is to split, or partition, data on which computations are applied among available processors (or cores).
 Each processor then applies the same progam or instruction onto its subpart of data.
-If a neighborhood information is needed from another processor communications or synchronizations are performed.
+Moreover, if a neighborhood information is needed from another processor communications or synchronizations are performed.
 
@@
 Second, the partitioning of data has to be balanced among processors.
-Thus, if this parallelization technique is clearly adequate to structured meshes, easy to balance, it is not for unstructured meshes or irregular structure where the amount of work is not heterogenous.
+Thus, if this parallelization technique is clearly adequate to structured meshes, easy to balance, it is not for unstructured meshes or irregular structures where the amount of work is not heterogenous.
 
@@
 Actually, if a task $i$ write a data $a$ and if a task $j$ read $a$, then $i$ has to be finished before $j$ is performed.
-From a dependency graph too different solutions are available.
+From a dependency graph two different solutions are available.
 First, a static schedule of tasks is built, which could be a good solution if the tasks are regular, or use a dynamic scheduler to dynamically decide at runtime which task is executed on which processor.
@@
 \paragraph{Hybrid parallelism} Finally, it is also possible to combine both those parallelization techniques to get what is called an hybrid parallelization.
-The interest of an hybrid parallelization is to bring another source of parallelism in a program if limits of a given technique are reach.
+The interest of an hybrid parallelization is to bring another source of parallelism if limits of a given technique are reach.
 
@@
 %----------------------------------------
-Component model is an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228}, where functionnalities of an application are dispatched in different entities: components.
+Component model is an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228}, where functionnalities of an application are dispatched in different entities called components.
 Component models improves code re-use by a separation of concerns, scalability and maintainability of applications~\cite{Szyperski:2002:CSB:515228,bigot:inria-00388508}.
 In addition to this, recent work on component models have shown a simultaneous answer to performance, maintainability and portability of applications~\cite{l2c}, which makes this domain an interesting solution to bring maintainability and portability in HPC programming.
+Many component models exist, each of them with its own specification and singularity, like CCM~\cite{corba:omg06} (CORBA Component Model), GCM~\cite{Baude} (Grid Component Model) or CCA~\cite{Armstrong:1999:TCC:822084.823232} (Common Component Architecture), for example.
 
-Many component models exist, each of them with its own specification and singularity, like CCM~\cite{corba:omg06} (CORBA Component Model), GCM~\cite{Baude} (Grid Component Model) or CCA~\cite{Armstrong:1999:TCC:822084.823232} (Common Component Architecture), for example.
-However, conepts of component, port, interface and assembly are most of the time found in component models.
 In the rest of this paper will be used concepts of the component model $L^2C$~\cite{l2c}.
-$L^2C$ is a minimalist HPC-oriented component model where a component extends the concept of class by specifying in its interfaces not only the services it offers, but also all its possible interactions with outer world, including the services it requires.
+$L^2C$ is a minimalist HPC-oriented component model where a component extends the concept of class by specifying in its interfaces not only the services it offers, but also all its possible interactions with outer world.
 As a result, each component is an independant entity composed of a set of services its provides, and a set of services it requires (and uses).
-
 A \emph{port} in $L^2C$ is an entity embbedded in the component which makes possible an \emph{assembly} of components.
-An assembly of components (of instanciated components) is a way to actually connect components together and to produce a complete application, as a set of components and their interactions.
+An assembly of components is a way to actually connect components together and to produce a complete application, as a set of components and their interactions.
 A provided service inside a component is associated to a \emph{provide} port, while a required service is associated to a \emph{use} port.
@@
 In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}.
-As the use port if on the right of its component, and the provide port on the left, in a component assembly, the component on the left uses the provide port of the right component.
+As the use port is on the right of its component, and the provide port on the left, in a component assembly the component on the left uses the provide port of the component on the right.
 
@@
 With its minimal set of ports, $L^2C$ is a light component model which does not introduce overheads at runtime, and which results in a HPC-oriented component model.
-Another HPC-oriented component model is CCA~\cite{Armstrong:1999:TCC:822084.823232}, however the main difference between those component models is the way a dsitribution of component on distributed architectures is managed.
-Actually, in CCA, MPI~\cite{Graham2009MSE} (Message Passing Interfaces) have to be used and managed manually inside components, while $L^2C$ proposes a specific MPI port to link distant components and ease their communications.
-Moreover, by exposing MPI connections, $L^2C$ also exposes the parallel structure of a program inside an assembly, which increase separation of concerns and code reuse for parallel applications.
+Another HPC-oriented component model is CCA~\cite{Armstrong:1999:TCC:822084.823232}, however the main difference between those component models is the way a distribution of component on distributed architectures is managed.
+Actually, in CCA distant components (on different processors or machines) have to be managed manually inside components using MPI~\cite{Graham2009MSE} (Message Passing Interfaces), while $L^2C$ proposes a specific MPI port to link distant components and ease their communications.
+Moreover, by exposing MPI connections, $L^2C$ also exposes the parallel structure of a program inside the assembly, which increase separation of concerns and code reuse for parallel applications.

### 2015-07-24 WOLFHPC15/article/ms_msc.tex
wolfhpc15 24_07

@@
 
-The overall grammar of a MSL program is given in Figure~\ref{fig:grammar}, where the terminals $meshid$, $meshdom$, $compdomid$, $dataid$, $compid$ are string identifiers, and $num$ is an integer value.
+The overall grammar of a MSL program is given in Figure~\ref{fig:grammar}, where the terminals $meshid$, $meshdom$, $compdomid$, $dataid$, $compid$ ans $neighborid$ are string identifiers, and $num$ is an integer value.
 An example is then given to explain each section.
@@
 \paragraph{Mesh and mesh domains} A mesh is defined by a string identifier.
-As already explain, this work is for now limited to a single mesh.
+As already explained, this work is for now limited to a single mesh.
 Each mesh domain is also represented by a string identifier.
-For example
+For example for a Cartesian mesh and three mesh domains we can define
 \begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
@@
 
-\paragraph{Computation domains} A computation domain can be seen as a subpart of a meh domain which will be used during computation phase.
+\paragraph{Computation domains} A computation domain can be seen as a subpart of a meh domain which will be used during the computation phase.
 For each computation domain is indicated the mesh domain identifier from which it is built.
-By default two computation domains always intersect except when precise in the section \emph{no intersections}.
-For example
+By default two computation domains always intersect except when an independent relation is precised in the section \emph{independent} of the language.
+In the example below, an entire computation domain is defined for each mesh domain, and additionnally two subpart of the mesh domain edgex are defined.
+It is then precised that part1edgex and part2edgex do not intersect.
 \begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
@@
 \end{lstlisting}
-In this example, an entire computation domain is defined for each mesh domain, and additionnally two subpart of the mesh domain edgex are defined.
-It is then precised that part1edgex and part2edgex do not intersect.
 
-\paragraph{Data and time} A data is a quantity to simulate and is mapped onto a mesh domain. The time section simply indicates a number of iterations to perform in the simulation. For example:
+\paragraph{Data and time} A data is a quantity to simulate and is mapped onto a mesh domain.
+The time section simply indicates a number of iterations to perform in the simulation.
+For example we can define ten data and 500 iterations as below
 \begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
@@
 
-\paragraph{Computations description} The most important part of MSL is the description of the computations of the simulation.
-The language follows the definitions given in Section~\ref{sect:multistencil}, however as MSL is not made to express numerical computations, the element $exp$ of the definition is not given in the language.
+\paragraph{Computations description} The most important part of MSL is the description of the simulation computations.
+The language follows the three definitions given in Section~\ref{sect:multistencil}, however as MSL is not made to express numerical computations, the element $exp$ of definitions~\ref{eq:st} and~\ref{eq:loc} is not given in the language.
 Moreover, to propose a syntax close to an imperative code, information are given in a different order.
@@
 
-\paragraph{Creation of the synchronized ordered list of computations} From the ordered list of computations $\Gamma$ it is possible to detect needed synchronizations if a \emph{data parallelization} is performed on the simulation.
+\paragraph{Creation of the synchronized ordered list of computations} COnsidering a data parallelization of the simulation, from the ordered list of computations $\Gamma$ can be detect needed synchronizations.
 A synchronization is needed each time a data read by a stencil computation has been written by a previous computation and if the computation domains intersect.
-Actually, as a stencil computation needs neighborhood values, in a domain decomposition values computed by different processors or cores are needed.
+Actually, as a stencil computation needs neighborhood values, values computed by different processors or cores are needed.
 
 For example, the stencil computation $c_1$ read the data $b$ which has been written by the computation $c_0$.
-Moreover the mesh domains are different which means that the computation domains intersect.
 For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed to the sublist $[c_0,sync_1,c_1]$ in $\Gamma_{data}$.
@@
 
-\paragraph{Creation of the dependency graph} From the synchronized ordered list of computations $\Gamma_{data}$, a dependency graph can be build.
+\paragraph{Creation of the dependency graph} From the synchronized ordered list of computations $\Gamma_{data}$, a dependency graph is built.
 A dependency exists between two computations (including synchronizations) if and only if a data read has been written by a previous computation in $\Gamma_{data}$, and if the computation domains intersect.
@@
 \paragraph{Transformation to a minimal series-parallel graph} As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if the forbidden \emph{N-shape} is not found in the graph.
-To transform $\Gamma_{hybrid}$ to the minimal series-parallel graph $\Gamma_{msp}$, a breadth first search~\cite{} traversing algorithm is applied to build a complete bipartite graph between two levels of $\Gamma_{hybrid}$~\cite{Mitchell:2004:CMV:1082101.1082117}.
+To transform $\Gamma_{hybrid}$ to the minimal series-parallel graph $\Gamma_{msp}$, a breadth first search traversing algorithm is applied to build a complete bipartite graph between two levels of $\Gamma_{hybrid}$~\cite{Mitchell:2004:CMV:1082101.1082117}.
 
@@
 
-\paragraph{Dump to a component assembly} From the series-parallel tree decomposition $\Gamma_{tsp}$ can be built a program using any parallel language containing \emph{parallel sections} and \emph{sequences of instructions} such as OpenMP~\cite{660313}, HPF~\cite{219857}, UPC~\cite{El-Ghazawi:2006:UUP:1188455.1188483} etc.
-In this paper we study the dump of $\Gamma_{tsp}$ to a component assembly which needs the introduction of \emph{control components} described in the next section.
+\paragraph{Dump to a component assembly} Finally, from the series-parallel tree decomposition $\Gamma_{tsp}$ we could build a program using any parallel language containing \emph{parallel sections} and \emph{sequences of instructions} such as OpenMP~\cite{660313}, HPF~\cite{219857}, UPC~\cite{El-Ghazawi:2006:UUP:1188455.1188483} etc.
+In this paper we study the dump of $\Gamma_{tsp}$ to a component assembly which needs the introduction of \emph{control components}, described in the next section.
 

### 2015-07-24 WOLFHPC15/article/mscac.tex
wolfhpc15 24_07

@@
-The solution presented in this paper is a domain specific language, to describe a multi-stencil program (a numerical simulation), and a compiler which transforms this description to a parallel component-based structure of the overall simulation.
+The solution presented in this paper is a domain specific language, to describe a multi-stencil program (a numerical simulation), and a compiler which transforms this description to a parallel component-based application of the overall simulation.
 
 The MS language (MSL) is an agnostic descriptive language for multi-stencil simulations.
-Agnostic means that the description of a numerical simulation, using MSL, does not need any knowledge on the mesh type, the stencil shape description, and the final implementation constraints.
-Descriptive, on the other hand, means that the terminal states of the language grammar are string identifiers and that the language does not handle description of numerical computations, but description of a multi-stencil simulation, in a coarse-grain fashion.
+Agnostic means that the description of a numerical simulation, using MSL, does not need any knowledge on the mesh type, the stencil shapes, or the final implementation constraints.
+Descriptive, on the other hand, means that the terminal states of the language grammar are string identifiers and that the language does not handle description of numerical computations, but only description of a multi-stencil simulation, in a coarse-grain fashion.
 
 As illustrated in Figure~\ref{fig:mscac}, the MS language compiler, called MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of two different parts.
-First, the computation compiler, called \emph{MSC}, is responsible for the computations parallelization and for its dump to a component assembly; second, the data compiler, called \emph{MSD}, handles a distributed data structure, data mapped onto it, and their relations with computations, as a final overall component assembly.
+First, the computation compiler, called \emph{MSC}, is responsible for the parallelization of computations and for its dump to a component assembly; second, the data compiler, called \emph{MSD}, handles a distributed data structure, data mapped onto it, and their relations with computations, as a final overall component assembly.
 
@@
 
-It has to be noticed that in this paper, the MSD compiler is a first adhoc version which works for a precise distributed data structure, the SkelGIS labrary~\cite{HeleneLS13,HeleneLS14,HeleneEuroPar14,CPE:CPE3494}.
-However, using the system of python string templates~\footnote{\url{https://docs.python.org/2/library/string.html}} to define the data component assembly, it is already possible to use another distributed data structure, as for example Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.
+It has to be noticed that in this paper, the MSD compiler is a first adhoc version which works for a precise distributed data structure, the SkelGIS library~\cite{HeleneLS13,HeleneLS14,HeleneEuroPar14,CPE:CPE3494}.
+However, using the system of python string templates~\footnote{\url{https://docs.python.org/2/library/string.html}} to define the component assembly part which manages data, it is already possible to use another distributed data structure, as for example Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.
 This part of the compiler takes place in a larger research project to propose domain decomposition skeletons using component models.

### 2015-07-24 WOLFHPC15/article/related.tex
wolfhpc15 21_07
wolfhpc15 24_07

--- 2015-07-21 WOLFHPC15/article/related.tex
@@
 
-Many solutions exists to ease parallel programming of stencil codes, as for example PATUS~\cite{citeulike12258902}, Pochoir~\cite{spaaTangCKLL11}, OP2~\cite{Giles2011}.
-Those solutions are powerfull, let the user implement their own stencil codes, and produce high performance codes.
+Many specific solutions exist to ease parallel programming of numerical simulations.
+Solutions such as PETSc~\cite{petsc-efficient} and Trilinos~\cite{Trilinos-Overview} are specific libraries.
+They are made to solve many kinds of PDEs using well-known methods and solvers which are efficiently parallelized inside the library.
+However, using those solutions the user has to use and follow existing tools.
+In other words, it is not possible for a mathematician or a physician to write his own numerical method if it differs from existing ones.
+The purpose of those solutions are much more specific than our solution.
+
+Other solutions try to do a step behind in the abstraction and propose a way to hide parallel codes for stencils while keeping a programming freedom.
+For example PATUS~\cite{citeulike12258902} and Pochoir~\cite{spaaTangCKLL11} are proposed for general structured meshes, and OP2~\cite{Giles2011} and Liszt~\cite{DeVito2011LDS} for general unstructured meshes.
 Those solutions can be considered as stencil compilers.
-As a result it produces optimized (cache tiling etc.) or parallel (CUDA, OpenCL etc.) codes for a single stencil computations.
+Actually they produce optimized (cache tiling etc.) or parallel (CUDA, OpenCL etc.) codes for a single stencil computations.
+Recently, domain specific languages for more specific numerical methods such as ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} for multigrid solvers, and OPS~\cite{Reguly:2014:ODS:2691166.2691173} for meshes of structured meshes (hybrid mesh) have also been proposed.
+All those solutions are powerfull, let the user implement their own stencil codes in a sequential programming style, and produce high performance codes for shared memory architectures and sometimes distributed memory architectures.
 
-However, a real case numerical simulation is not composed of a single stencil code as explained in Section~\ref{sect:multistencil}.
-A multi-stencil simulation computes more than one stencil computation, involving one or more stencil shapes, and additionnal local coputations.
+Stencil compilers handle the parallelization or the optimization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations.
+However, real case numerical simulation is not composed of a single stencil code, as explained in Section~\ref{sect:multistencil}, but of more than one stencil computation, involving one or more stencil shapes, and additionnal local coputations.
 Solutions like Pochoir or PATUS handle the parallelization or the optimization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations.
-However, not taking into account the parallelization of the overall numerical simulation implies that those compilers are reduced to shared memory systems.
-Actually using a distributed memory system, the MPI (Message Passing Interface)~\cite{Graham2009MSE} library still have to be used by the user, as well as the distribution of the mesh onto the different distant processors.
+Liszt~\cite{DeVito2011LDS} is the closest solution to our work as it seems to take into account the entire program and not only stencil kernels.
+
+The work presented in this paper takes place at a different level of numerical simulations and can be seen as a complementary to stencil compilers.
+Actually, the MSL language is agnostic and works for any type of mesh and any neighborhood shape in stencil codes.
+The MSC part of the compiler only needs this agnostic description of the overall simulation to generate a parallel structure of the program, which needs to be filled with computation codes.
+Depending on the underlying distributed data structure used, those computation codes could be generated by stencil compilers.
+As a result, MSL and its compiler are not new solution to compile optimized stencil codes, but another level of parallelization and a different approach.
+Finally, the solution proposed in this paper brings an interesting separation of concerns between the overall conception of the simulation, and the actual implementation of computations using existing libraries or specific tools.
 
 %----------------------------------------
-\subsection{Component models and control}
+\subsection{Language, control and component models}
 %----------------------------------------
 
-%----------------------------------------
-\subsection{Languages and component models}
-%----------------------------------------
+\textbf{Put it somewhere else !}
+
+As far as we know, no existing DSL has been transformed to a HPC component assembly until know.
+This paper is a contribution to study the interest of using component models to increase code re-use, separation of concerns and maintainability in DSLs.
+Actually, as the code is not generated inside the DSL itself but is built from the assembly of existing components, many advantages can get out of this work.
+First, the code-reuse from one DSL to another is increased if the components are well studied and defined.
+For example, it seems that control components introduced in this paper can be used for many different DSLs in many different domains.
+Second, separation of concerns in the final code is also improved as the generated component assembly split the different functionnalities of the application in different components.
+As a result, some components can be written by computer-scientists, while other are specific to the application, as for example computation components.
+Finally, the maintainability of the application is also increased, as each functionnality is separated.
+In the final code if a computation component has to be modified by a numerician, intricates part of the code will be unchanged and not seen by the numerician.
+
+If the study of a DSL dumped to a component assembly has not been studied until now, some component models have already proposed the introduction of time control and data control in the assembly, as for example STCM~\cite{DBLP:confeuropar2008} and Xman~\cite{He12component-baseddesign}.
+However, both those solutions does not seem to target high performance computing and distributed components.

### 2015-07-24 WOLFHPC15/article/related.tex
wolfhpc15 24_07

@@
 
-If the study of a DSL dumped to a component assembly has not been studied until now, some component models have already proposed the introduction of time control and data control in the assembly, as for example STCM~\cite{DBLP:confeuropar2008} and Xman~\cite{He12component-baseddesign}.
+If the study of a DSL dumped to a component assembly has not been studied until now, some component models have already proposed the introduction of time control and data control in the assembly, as for example STCM~\cite{DBLP:confeuropar2008} and XMan~\cite{He12component-baseddesign}.
 However, both those solutions does not seem to target high performance computing and distributed components.

### 2015-07-28 WOLFHPC15/article/concept.tex
minor fixes

@@
 
-Figure~\ref{fig:ex} gives two examples of stencil computations and introduce the different concepts used in this paper.
+Let introduce some concepts used in this paper.
 First, a stencil code is based on a \emph{mesh} which is the discretization of the physical real domain.
@@
 A data is mapped onto a given mesh domain.
-For example, in the figure, data $A$ and $B$ are mapped onto the center of cells, while $C$ is mapped onto the edges.
+For example, in Figure~\ref{fig:ex}, data $A$ and $B$ are mapped onto the center of cells, while $C$ is mapped onto the edges.
 
@@
 
-The first stencil computation of Figure~\ref{fig:ex} computes $B$ using $A$, while the second one computes $B$ using $C$.
+The first stencil computation in Figure~\ref{fig:ex} computes $B$ using $A$, while the second one computes $B$ using $C$.
 Thus, a stencil computation is defined by a set of input data (in those examples only one input data), and a single output data, the result.
@@
 The numerical expression of a stencil computation has the particularity to use not only the element to compute but also a \emph{neighborhood} of this element, also called a \emph{stencil shape}.
-For example, the stencil shape of the first computation of Figure~\ref{fig:ex} contains direct neighbors on the right, left, top and bottom.
+For example, the stencil shape of the first computation in Figure~\ref{fig:ex} contains direct neighbors on the right, left, top and bottom.
 Sometimes, the neighborhood shape can also access a different mesh domain, as for example in the second computation example.
@@
 %----------------------------------------
-If stencil computations are well defined and a lot studied in the literrature, it is not the case for real case numerical simulations.
+While stencil computations are well defined and a lot studied in the literrature, it is not the case for real case numerical simulations.
 In this paper, we define what we call a \emph{multi-stencil program}, which actually is a general numerical simulation.
@@
 Moreover, a numerical simulation also performs additionnal auxiliary computations which do not involve neighborhood, called local computations.
-In this paper is targetted this general case of numerical simulations that we call \emph{multi-stencil programs}.
+This paper targets this general case of numerical simulations that we call \emph{multi-stencil programs}.
 The work presented produces a general parallelization structure of the overall simulation, while most stencil solutions produce optimized codes for a single stencil code.

### 2015-07-28 WOLFHPC15/article/ms_msc.tex
minor fixes

@@
 \paragraph{Computations description} The most important part of MSL is the description of the simulation computations.
-The language follows the three definitions given in Section~\ref{sect:multistencil}, however as MSL is not made to express numerical computations, the element $exp$ of definitions~\ref{eq:st} and~\ref{eq:loc} is not given in the language.
+The language follows the three definitions given in Section~\ref{sect:multistencil}.
+However as MSL is not made to express numerical computations, the term $exp$ of Definition~\ref{eq:st} and Definition~\ref{eq:loc} is not given in the language.
 Moreover, to propose a syntax close to an imperative code, information are given in a different order.
@@
 
-\paragraph{Creation of the synchronized ordered list of computations} COnsidering a data parallelization of the simulation, from the ordered list of computations $\Gamma$ can be detect needed synchronizations.
+\paragraph{Creation of the synchronized ordered list of computations} Considering a data parallelization of the simulation, from the ordered list of computations $\Gamma$ can be detect needed synchronizations.
 A synchronization is needed each time a data read by a stencil computation has been written by a previous computation and if the computation domains intersect.

### 2015-07-28 rapport/formalism.tex
minor fixes

@@
 Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}, and a set of auxiliary computations also needed to perform the simulation, and also called \emph{local computations}.
-In the following Section, formal definitions of a \textit{stencil program} and its computations are given.
-Then, the different parallelization techniques which can be applied on such program, are presented.
+This section gives formal definitions of a \textit{stencil program} and its computations. Then, the different parallelization techniques which can be applied on such program, are presented.
 
@@
 
+\rmq{Pas très clair la suite; motivation? dom == some sub elements of M ?}
 \begin{mydef}
@@
 For example, the set of cells $D_0$ in a Cartesian 2D mesh could be defined by exactly four vertices and four edges connected as a cycle.
-But we could also define another set of elements $D_1$ as the simple set of vertices $V$ etc.
+But we could also define another set of elements $D_1$ as the simple set of vertices $V$, etc.
 

### 2015-07-28 WOLFHPC15/article/article.tex
wolfhpc15 21_07
wolfhpc15 23_07
wolfhpc15 27_07
minor fixes
fix compiler 28_07

@@
 \usepackage{listings}
+\usepackage{url}
+\usepackage{cite}
 %----------TIKZ
@@
 %-------------
+\newenvironment{mydef}[1][Definition]{\begin{trivlist}
+\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
 
@@
 
-\title{A component-based DSL for multi-stencil programs}
+\title{From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study}
 
@@
 % 1st. author
+\alignauthor
+Julien Bigot\\
+       \affaddr{CEA, Maison de la Simulation}\\
+       \affaddr{USR 3441}\\
+       \affaddr{91191 Gif-sur-Yvette Cedex, France}\\
+       \email{julien.bigot@cea.fr}
+% 1st. author
+\alignauthor
+Helene Coullon\\
+       \affaddr{INRIA}\\
+       \affaddr{ENS Lyon, 46 Allee d'Italie}\\
+       \affaddr{Lyon, France}\\
+       \email{helene.coullon@inria.fr}
+% 2nd. author
+\and
 \alignauthor
@@
        \email{christian.perez@inria.fr}
-% 2nd. author
-\alignauthor
-Helene Coullon\\
-       \affaddr{INRIA}\\
-       \affaddr{ENS Lyon, 46 Allee d'Italie}\\
-       \affaddr{Lyon, France}\\
-       \email{helene.coullon@inria.fr}
 }
@@
 
-\keywords{Component models, domain specific languages, multi-stencil, numerical simulations} % NOT required for Proceedings
+%\keywords{Component models, domain specific languages, multi-stencil, numerical simulations} % NOT required for Proceedings
 
@@
 %----------------------------------------
-\section{Underlying concepts}
+\section{Stencil and multi-stencil}
 \label{sect:concept}
@@
 %----------------------------------------
-\section{The Multi-Stencil Component Assembly Compiler}
+\section{Overview}
 \label{sect:mscac}
@@
 %----------------------------------------
-\section{Control components and component assembly}
+\section{Dump to a component assembly}
 \label{sect:component}

### 2015-07-28 WOLFHPC15/article/component.tex
fix compiler 28_07

@@
+%----------------------------------------
+\subsection{Component models}
+%----------------------------------------
+Component model is an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228}, where functionnalities of an application are dispatched in different entities called components.
+Component models improves code re-use by a separation of concerns, scalability and maintainability of applications~\cite{Szyperski:2002:CSB:515228,bigot:inria-00388508}.
+In addition to this, recent work on component models have shown a simultaneous answer to performance, maintainability and portability of applications~\cite{l2c}, which makes this domain an interesting solution to bring maintainability and portability in HPC programming.
+Many component models exist, each of them with its own specification and singularity, like CCM~\cite{corba:omg06} (CORBA Component Model), GCM~\cite{Baude} (Grid Component Model) or CCA~\cite{Armstrong:1999:TCC:822084.823232} (Common Component Architecture), for example.
+
+In the rest of this paper will be used concepts of the component model $L^2C$~\cite{l2c}.
+$L^2C$ is a minimalist HPC-oriented component model where a component extends the concept of class by specifying in its interfaces not only the services it offers, but also all its possible interactions with outer world.
+As a result, each component is an independant entity composed of a set of services its provides, and a set of services it requires (and uses).
+A \emph{port} in $L^2C$ is an entity embbedded in the component which makes possible an \emph{assembly} of components.
+An assembly of components is a way to actually connect components together and to produce a complete application, as a set of components and their interactions.
+A provided service inside a component is associated to a \emph{provide} port, while a required service is associated to a \emph{use} port.
+It is also possible to group more than one required service into a single port, as a list, called a \emph{use-multiple} port.
+As illustrated in Figures~\ref{fig:ports}, a provide port will be represented by a white circle on the left of the component, a use port by a black circle and a use-multiple port by a black circle with a white $m$ in it, both on the right of the component.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (C) at (0,0) {$Comp_0$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (1.5,0) {};
+   \node[provide,right=1.5cm of u] (p1) {};
+   \node[component,right=1.5cm of p1] (C1) {$Comp_1$};
+   \node[use,right=1.5cm of C1] (um) {$m$};
+
+  \path[-]
+    (p) edge node {} (C)
+    (C.east) edge node {} (u)
+    (C1)  edge node {} (um)
+    (p1) edge node {} (C1);
+\end{tikzpicture}
+\caption{Two components, one with a provide, use and use-multiple ports, the second with a single provide port}
+\label{fig:ports}
+\end{center}
+\end{figure}
+
+In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}.
+As the use port is on the right of its component, and the provide port on the left, in a component assembly the component on the left uses the provide port of the component on the right.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (C) at (0,0) {$Comp_0$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (1.5,0) {};
+   \node[component,right=1.5 of u] (C1) {$Comp_1$};
+   \node[use,right=1.5 of C1] (um) {$m$};
+
+  \path[-]
+    (p) edge node {} (C)
+    (C) edge node {} (u)
+    (C1)  edge node {} (um)
+    (u) edge node {} (C1);
+\end{tikzpicture}
+\caption{Component assembly of Figure~\ref{fig:ports}}
+\label{fig:assembly}
+\end{center}
+\end{figure}
+
+With its minimal set of ports, $L^2C$ is a light component model which does not introduce overheads at runtime, and which results in a HPC-oriented component model.
+Another HPC-oriented component model is CCA~\cite{Armstrong:1999:TCC:822084.823232}, however the main difference between those component models is the way a distribution of component on distributed architectures is managed.
+Actually, in CCA distant components (on different processors or machines) have to be managed manually inside components using MPI~\cite{Graham2009MSE} (Message Passing Interfaces), while $L^2C$ proposes a specific MPI port to link distant components and ease their communications.
+Moreover, by exposing MPI connections, $L^2C$ also exposes the parallel structure of a program inside the assembly, which increase separation of concerns and code reuse for parallel applications.
+
 %----------------------------------------
@@
    \node[component] (seq0) at (0,0) {$SEQ$};
-   \node[provide] (seq0p) [left = 1cm of seq0] {};
+   \node[provide] (seq0p) [left = 0.8cm of seq0] {};
    \node[use] (seq0u) [right = 1cm of seq0] {$m$};
    %c0
-   \node[component] (c0) [right = 2.5cm of seq0u] {$c_0$};
+   \node[component] (c0) [right = 2cm of seq0u] {$c_0$};
    \node[use] (c0u) [right = 0.8cm of c0] {};
@@
    %c1
-   \node[component] (c1) [below = 1cm of sync0] {$c_1^*$};
+   \node[component] (c1) [below = 0.8cm of sync0] {$c_1^*$};
    \node[use] (c1u) [right = 0.8cm of c1] {};
@@
    %par0
-   \node[component] (par0) [below = 1cm of c1] {$PAR$};
+   \node[component] (par0) [below = 0.8cm of c1] {$PAR$};
    \node[use] (par0u) [right = 1cm of par0] {$m$};
@@
   %sync1
-   \node[component] (sync1) [below = 1cm of c2] {$SYNC$};
+   \node[component] (sync1) [below = 0.8cm of c2] {$SYNC$};
    \node[use] (sync1u) [right = 1cm of sync1] {$m$};
   %c4
-   \node[component] (c4) [below = 1cm of par1] {$c_4^*$};
+   \node[component] (c4) [below = 0.8cm of par1] {$c_4^*$};
    \node[use] (c4u) [right = 0.8cm of c4] {};
@@
   %c6
-   \node[component] (c6) [below = 1cm of c4] {$c_6$};
+   \node[component] (c6) [below = 0.8cm of c4] {$c_6$};
    \node[use] (c6u) [right = 0.8cm of c6] {};
@@
   %seq2
-   \node[component] (seq2) [below = 3cm of seq1] {$SEQ$};
+   \node[component] (seq2) [below = 2.5cm of seq1] {$SEQ$};
    \node[use] (seq2u) [right = 1cm of seq2] {$m$};
@@
   %c5
-   \node[component] (c5) [below = 1cm of c3] {$c_5$};
+   \node[component] (c5) [below = 0.8cm of c3] {$c_5$};
    \node[use] (c5u) [right = 0.8cm of c5] {};
@@
   %c7
-   \node[component] (c7) [below = 1cm of par0] {$c_7$};
+   \node[component] (c7) [below = 0.8cm of par0] {$c_7$};
    \node[use] (c7u) [right = 0.8cm of c7] {};
@@
   %sync2
-   \node[component] (sync2) [below = 1cm of c7] {$SYNC$};
+   \node[component] (sync2) [below = 0.8cm of c7] {$SYNC$};
    \node[use] (sync2u) [right = 1cm of sync2] {$m$};
   %c8
-   \node[component] (c8) [below = 1cm of sync2] {$c_8^*$};
+   \node[component] (c8) [below = 0.8cm of sync2] {$c_8^*$};
    \node[use] (c8u) [right = 0.8cm of c8] {};

### 2015-07-28 WOLFHPC15/article/concept.tex
fix compiler 28_07

@@
 
+% !!!!!!!!!!!! a mettre en plus court autre part !!!!!!!!!!!!!!!
 %----------------------------------------
-\subsection{Parallelization techniques}
-\label{sect:parallel}
+% \subsection{Parallelization techniques}
+% \label{sect:parallel}
 %----------------------------------------
-Three classical parallelization techniques are used in this paper and are described in this section, the data parallelism, the task parallelism, and the hybrid data and task parallelism.
+% Three classical parallelization techniques are used in this paper and are described in this section, the data parallelism, the task parallelism, and the hybrid data and task parallelism.
 Those parallelization techniques are independent from the actual parallel hardware used.
 
-\paragraph{Data parallelism} The idea of this parallelization technique is to split, or partition, data on which computations are applied among available processors (or cores).
+% \paragraph{Data parallelism} The idea of this parallelization technique is to split, or partition, data on which computations are applied among available processors (or cores).
 Each processor then applies the same progam or instruction onto its subpart of data.
@@
 
-In the domain of numerical simulations, this technique is most of the time called a domain decomposition.
+% In the domain of numerical simulations, this technique is most of the time called a domain decomposition.
 This parallelization technique produce efficient programs up to thousands processors or cores, but on certain conditions.
@@
 
-\paragraph{Task parallelism} Another well known parallelization technique is to identify in a program the different tasks and which one are independant and can be launched concurrently.
+% \paragraph{Task parallelism} Another well known parallelization technique is to identify in a program the different tasks and which one are independant and can be launched concurrently.
 Most of the time, such a prallelization technique create a dependency directed acyclic graph (dependency dag) from a set of ordered tasks.
@@
 
-\paragraph{Hybrid parallelism} Finally, it is also possible to combine both those parallelization techniques to get what is called an hybrid parallelization.
+% \paragraph{Hybrid parallelism} Finally, it is also possible to combine both those parallelization techniques to get what is called an hybrid parallelization.
 The interest of an hybrid parallelization is to bring another source of parallelism if limits of a given technique are reach.
-
-%----------------------------------------
-\subsection{Component models}
-%----------------------------------------
-Component model is an interesting domain of software engineering~\cite{Szyperski:2002:CSB:515228}, where functionnalities of an application are dispatched in different entities called components.
-Component models improves code re-use by a separation of concerns, scalability and maintainability of applications~\cite{Szyperski:2002:CSB:515228,bigot:inria-00388508}.
-In addition to this, recent work on component models have shown a simultaneous answer to performance, maintainability and portability of applications~\cite{l2c}, which makes this domain an interesting solution to bring maintainability and portability in HPC programming.
-Many component models exist, each of them with its own specification and singularity, like CCM~\cite{corba:omg06} (CORBA Component Model), GCM~\cite{Baude} (Grid Component Model) or CCA~\cite{Armstrong:1999:TCC:822084.823232} (Common Component Architecture), for example.
-
-In the rest of this paper will be used concepts of the component model $L^2C$~\cite{l2c}.
-$L^2C$ is a minimalist HPC-oriented component model where a component extends the concept of class by specifying in its interfaces not only the services it offers, but also all its possible interactions with outer world.
-As a result, each component is an independant entity composed of a set of services its provides, and a set of services it requires (and uses).
-A \emph{port} in $L^2C$ is an entity embbedded in the component which makes possible an \emph{assembly} of components.
-An assembly of components is a way to actually connect components together and to produce a complete application, as a set of components and their interactions.
-A provided service inside a component is associated to a \emph{provide} port, while a required service is associated to a \emph{use} port.
-It is also possible to group more than one required service into a single port, as a list, called a \emph{use-multiple} port.
-As illustrated in Figures~\ref{fig:ports}, a provide port will be represented by a white circle on the left of the component, a use port by a black circle and a use-multiple port by a black circle with a white $m$ in it, both on the right of the component.
-
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (C) at (0,0) {$Comp_0$};
-   \node[provide] (p) at (-1.5,0) {};
-   \node[use] (u) at (1.5,0) {};
-   \node[provide,right=1.5cm of u] (p1) {};
-   \node[component,right=1.5cm of p1] (C1) {$Comp_1$};
-   \node[use,right=1.5cm of C1] (um) {$m$};
-
-  \path[-]
-    (p) edge node {} (C)
-    (C.east) edge node {} (u)
-    (C1)	edge node {} (um)
-    (p1) edge node {} (C1);
-\end{tikzpicture}
-\caption{Two components, one with a provide, use and use-multiple ports, the second with a single provide port}
-\label{fig:ports}
-\end{center}
-\end{figure}
-
-In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}.
-As the use port is on the right of its component, and the provide port on the left, in a component assembly the component on the left uses the provide port of the component on the right.
-
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (C) at (0,0) {$Comp_0$};
-   \node[provide] (p) at (-1.5,0) {};
-   \node[use] (u) at (1.5,0) {};
-   \node[component,right=1.5 of u] (C1) {$Comp_1$};
-   \node[use,right=1.5 of C1] (um) {$m$};
-
-  \path[-]
-    (p) edge node {} (C)
-    (C) edge node {} (u)
-    (C1)	edge node {} (um)
-    (u) edge node {} (C1);
-\end{tikzpicture}
-\caption{Component assembly of Figure~\ref{fig:ports}}
-\label{fig:assembly}
-\end{center}
-\end{figure}
-
-With its minimal set of ports, $L^2C$ is a light component model which does not introduce overheads at runtime, and which results in a HPC-oriented component model.
-Another HPC-oriented component model is CCA~\cite{Armstrong:1999:TCC:822084.823232}, however the main difference between those component models is the way a distribution of component on distributed architectures is managed.
-Actually, in CCA distant components (on different processors or machines) have to be managed manually inside components using MPI~\cite{Graham2009MSE} (Message Passing Interfaces), while $L^2C$ proposes a specific MPI port to link distant components and ease their communications.
-Moreover, by exposing MPI connections, $L^2C$ also exposes the parallel structure of a program inside the assembly, which increase separation of concerns and code reuse for parallel applications.

### 2015-07-28 WOLFHPC15/article/introduction.tex
wolfhpc15 23_07
fix compiler 28_07

@@
 \item DSL interest: simplified HPC programming + perf
+\item A DSL for multi-stencil simulations + motivation fluid dynamics + bolzman + ...
 \item DSL limits: Lots of DSLs difficult to write, to maintain, to make portable etc.
 \item Components introduction, this paper takes a first example of DSL to components
-\item The domain targeted is multi-stencil simulations
-\item Contributions: DSL to components try + agnostic coarse grain DSL for multi-stencil programs
+\item Contributions: agnostic coarse grain DSL for multi-stencil programs + DSL to components try
 \item Outline

### 2015-07-28 WOLFHPC15/article/mscac.tex
fix compiler 28_07

@@
 
+\begin{figure*}
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (D) at (0,0) {Driver};
+   \node[provide] (Dp) at (-1,0) {};
+   \node (Ds) at (-1.5,0) {start};
+   \node[use] (Du1) at (1,0.5) {};
+   \node[use] (Du2) at (1,-0.5) {};
+
+   \node[component,right=1cm of Du2] (DA) {DriverApp};
+   \node[use] (DAu1) at (3.5,0) {$m$};
+   \node[use] (DAu2) at (3.5,-1) {};
+
+   \node[component] (T) at (4.5,-1) {Time};
+   \node[use,right=1cm of T] (Tu) {};
+   \node[component,right=2cm of Tu] (C) {Computations};
+   \node[use,right=2cm of C] (Cu) {};
+   \node[below=0.2cm of Cu] (star) {$*$};
+
+   \node[component,right=8cm of DAu1] (Data) {Data};
+   \node[use,right=1cm of Data] (Datau) {};
+
+   \node[component,right=13cm of Du1] (DDS) {DDS};
+
+  \path[-]
+    (Dp) edge node {} (D)
+    (D.east) edge node {} (Du1)
+      edge node {} (Du2)
+    (Du1) edge node {} (DDS.west)
+    (Du2) edge node {} (DA)
+    (DA.east) edge node {} (DAu1)
+      edge node {} (DAu2)
+    (DAu1) edge node {} (Data.west)
+    (DAu2) edge node {} (T)
+    (T) edge node {} (Tu)
+    (Tu) edge node {} (C)
+    (C) edge node {} (Cu)
+    (Cu) edge node {} (Data.west)
+    (Data) edge node {} (Datau)
+    (Datau) edge node {} (DDS.west);
+\end{tikzpicture}
+\caption{Shape of the generated component assembly.}
+\label{fig:assembly}
+\end{center}
+\end{figure*}
+
 It has to be noticed that in this paper, the MSD compiler is a first adhoc version which works for a precise distributed data structure, the SkelGIS library~\cite{HeleneLS13,HeleneLS14,HeleneEuroPar14,CPE:CPE3494}.

### 2015-07-29 WOLFHPC15/article/mscac.tex
fix compiler 29_07

@@
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (D) at (0,0) {Driver};
+   \node[component] (D) at (0,0) {$Driver$};
    \node[provide] (Dp) at (-1,0) {};
@@
 
-   \node[component,right=1cm of Du2] (DA) {DriverApp};
+   \node[component,right=1cm of Du2] (DA) {$DriverApp$};
    \node[use] (DAu1) at (3.5,0) {$m$};
@@
 
-   \node[component] (T) at (4.5,-1) {Time};
+   \node[component] (T) at (4.5,-1) {$Time$};
    \node[use,right=1cm of T] (Tu) {};
-   \node[component,right=2cm of Tu] (C) {Computations};
+   \node[component,right=2cm of Tu] (C) {$Computations$};
    \node[use,right=2cm of C] (Cu) {};
@@
 
-   \node[component,right=8cm of DAu1] (Data) {Data};
+   \node[component,right=8cm of DAu1] (Data) {$Data$};
    \node[use,right=1cm of Data] (Datau) {};
 
-   \node[component,right=13cm of Du1] (DDS) {DDS};
+   \node[component,right=13cm of Du1] (DDS) {$DDS$};
 

### 2015-07-29 WOLFHPC15/article/concept.tex
fix compiler 29_07

@@
 A mesh is a connected undirected graph without bridges (an edge is a bridge if its removal results in two disconnected graphs), where nodes and edges are linked to form cells (closure).
-For example, a structured mesh is illustrated in Figure~\ref{fig:ex}, each cell contains four nodes and four edges.
+For example, a structured mesh is illustrated in Figure~\ref{fig:mesh}, each cell contains four nodes and four edges.
 Second, \emph{mesh domains} are subset elements of the mesh, such as the center of cells, edges or nodes, for example.
-In Figure~\ref{fig:ex} two mesh domains are used, the center of the cells (in red) and the edge on the horizontal axe (in blue on the left and right of each cell).
+In Figure~\ref{fig:mesh} two mesh domains are illustrated, the center of the cells (in red) and edges on the horizontal axe (in blue on left and right of each cell).
 A \emph{data} is a quantity to simulate or a temporary data to perform computations.
 A data is mapped onto a given mesh domain.
-For example, in Figure~\ref{fig:ex}, data $A$ and $B$ are mapped onto the center of cells, while $C$ is mapped onto the edges.
+For example, in Figure~\ref{fig:ex1}, data $A$ and $B$ are mapped onto the center of cells, while in Figure~\ref{fig:ex2} $C$ is mapped onto the edges.
 
-\begin{figure}[!h]\begin{center}
-  \resizebox{7cm}{!}{\includegraphics{./images/stencil.pdf}}
-  \caption{Examples of a stencil computations}
-  \label{fig:ex}
-\end{center}\end{figure}
+\begin{figure}
+\begin{center}
+\subfloat[Mesh and mesh domains\label{fig:mesh}]{
+\resizebox{7cm}{!}{\includegraphics{./images/mesh.pdf}}
+}
+\hspace{10pt}
+\subfloat[First stencil example\label{fig:ex1}]{
+\resizebox{5cm}{!}{\includegraphics{./images/stencil1.pdf}}
+}
+\hspace{10pt}
+\subfloat[Second stencil example\label{fig:ex2}]{
+\resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
+}
+\end{center}
+\caption{Stencil examples}
+\label{fig:gspmsp}
+\end{figure}
 
-The first stencil computation in Figure~\ref{fig:ex} computes $B$ using $A$, while the second one computes $B$ using $C$.
+% \begin{figure}[!h]\begin{center}
+%   \resizebox{7cm}{!}{\includegraphics{./images/stencil.pdf}}
+%   \caption{Examples of a stencil computations}
+%   \label{fig:ex}
+% \end{center}\end{figure}
+
+The first stencil computation in Figure~\ref{fig:ex1} computes $A$ using $B$, while the second one in Figure~\ref{fig:ex2} computes $A$ using $C$.
 Thus, a stencil computation is defined by a set of input data (in those examples only one input data), and a single output data, the result.
@@
 In other words a computation domain is a subset of a mesh domain.
-For example, in the first stencil, the computation of $B$ is performed on plain elements, while dotted elements are not computed.
-On the other hand, on the second computation, the computation domain of $B$ is exactly equal to the cell mesh domain.
+For example, in the first stencil, the computation of $A$ is performed on plain elements, while dotted elements are not computed.
+On the other hand, on the second computation, the computation domain of $A$ is exactly equal to the cell mesh domain.
 The computation domain represents the set of elements on which the computation is performed, in other words, the space loop.
@@
 The numerical expression of a stencil computation has the particularity to use not only the element to compute but also a \emph{neighborhood} of this element, also called a \emph{stencil shape}.
-For example, the stencil shape of the first computation in Figure~\ref{fig:ex} contains direct neighbors on the right, left, top and bottom.
-Sometimes, the neighborhood shape can also access a different mesh domain, as for example in the second computation example.
+For example, the stencil shape of the first computation in Figure~\ref{fig:ex1} contains direct neighbors on the right, left, top and bottom.
+Sometimes, the neighborhood shape can also access a different mesh domain, as for example in Figure~\ref{fig:ex2}.
 Actually, in this computation, the neighborhood contains edges on the left and on the right of the cell.
-As a result, we could imagine for example that the numerical expression of the first example is
+As an example, the numerical expression of the first example could be
 \begin{equation*} 
@@
 \end{equation*}
-and that the numerical expression of the second example is
+and the numerical expression of the second example could be
 \begin{equation*} 

### 2015-07-29 WOLFHPC15/article/concept.tex
fix compiler 29_07

@@
 \begin{equation*} 
-B(x,y) = A(x+1,y)+A(x-1,y)+A(x,y+1)+A(x,y-1),
+A(x,y) = B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1),
 \end{equation*}
@@
 \begin{equation*} 
-B(x,y) = B(x,y)+C(x1,y1)+C(x1+1,y1).
+A(x,y) = A(x,y)+C(x1,y1)+C(x1+1,y1).
 \end{equation*}
-
-To conclude this section, we define a stencil computation $s$ as the quintuplet
-\begin{equation} 
-s(R,w,exp,\mathcal{N},d_c),
-\label{eq:st}
-\end{equation}
-where $R$ is the set of input data, $w$ the output data, $exp$ the numerical expression of the stencil computation, which is applied on the computation domain $d_c$, using a stencil shape $\mathcal{N}$.
 
@@
 %----------------------------------------
-While stencil computations are well defined and a lot studied in the literrature, it is not the case for real case numerical simulations.
-In this paper, we define what we call a \emph{multi-stencil program}, which actually is a general numerical simulation.
+A real case numerical simulation is most of the time composed of more than one stencil computation, with one or more stencil shapes, and onto one or more data, and of additionnal local computations.
+Let introduce formal definitions used in the rest of the paper.
 
-A real case numerical simulation is most of the time not composed of a single stencil computation, but of a set of stencil computations, with one or more stencil shapes (neighboorhood), and onto one or more data.
-Moreover, a numerical simulation also performs additionnal auxiliary computations which do not involve neighborhood, called local computations.
-This paper targets this general case of numerical simulations that we call \emph{multi-stencil programs}.
-The work presented produces a general parallelization structure of the overall simulation, while most stencil solutions produce optimized codes for a single stencil code.
-As it will be detailed in the related work, this work is complementary to the optimization and parallelization of a single stencil computation.
+For $\Delta$ the set of data of the simulation, a stencil computation $s$ is defined as the quintuplet
+\begin{equation} 
+s(R,w,exp,d_c),
+\label{eq:st}
+\end{equation}
+where $R$ is a list of pair $(r,n)$ with $r \in \Delta$ is a data read by the computation, and $n$ is a stencil shape (a neighborhood).
+The data written by the computation is denoted $w \in \Delta$, $exp$ is the numerical expression of the stencil computation.
+Finally the numerical expression is applied on the computation domain $d_c$.
 
-An auxiliary computation is almost the same than a stencil computation, however its numerical expression does not involve a neighborhood, which makes this kind of computation purely local.
+For example, in the example of Figure~\ref{fig:ex1}, if we call the computation domain (plain lines) $dc1$ and the stencil shape $n1$, the stencil computation can be defined with
+\begin{equation*}
+R: {(B,n1)} \quad w: A \quad d_c: dc1
+\end{equation*}
+\begin{equation*}
+exp: A(x,y)=B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1).
+\end{equation*}
+On the other hand, in the second example of Figure~\ref{fig:ex2}, if we call the computation domain $dc2$ and the stencil shape $n2$, the stencil computation can be defined as
+\begin{equation*}
+R: {(C,n2),(A,null)} \quad w: A \quad d_c: dc2
+\end{equation*}
+\begin{equation*}
+exp: A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1)
+\end{equation*}
+One can notice that the input data $A$ is associated to the stencil shape $null$. This means that no stencil shape is applied on $A$, but only a local access.
+
+A local, or auxiliary, computation is almost the same than a stencil computation without neighborhood information in the numerical expression, which makes this kind of computation purely local.
 A local computation $l$ is defined as the quadruplet
 \begin{equation} 
-l(R,w,exp,d_c),
+l(R_l,w,exp,d_c),
 \label{eq:loc}
 \end{equation}
-where $R$ is the set of input data, $w$ the output data, $exp$ the numerical expression of the stencil computation, which is applied on the computation domain $d_c$.
+where $R_l$ is the set of input data, and other parameters are the same than for a stencil computation.
 
@@
 \end{equation}
-where $T$ is the set of time iteration to run the simulation, $\mathcal{M}$ is the mesh of the simulation, $\mathcal{D}_m$ is the set of mesh domains, $\mathcal{D}_c$ is the set of computation domains used into computations, $\Delta$ is the set of data, each one mapped onto a mesh domain, and finally $\Gamma$ is the set of computations.
+where $T$ is the set of time iteration to run the simulation, $\mathcal{M}$ is the mesh of the simulation, $\mathcal{D}_m$ is the set of mesh domains, $\mathcal{D}_c$ is the set of computation domains used into computations, $\Delta$ is the set of data of the simulation, each one mapped onto a mesh domain, and finally $\Gamma$ is the set of computations.
 The set of computations $\Gamma$ is composed of an ordered list of stencil and local computations.

### 2015-07-29 WOLFHPC15/article/introduction.tex
fix compiler 29_07

@@
 \end{itemize}
+
+While stencil computations are well defined and a lot studied in the literrature, it is not the case for real case numerical simulations.
+In this section are introduced  we define what we call a \emph{multi-stencil program}, which actually is a general numerical simulation.
+
+A real case numerical simulation is most of the time not composed of a single stencil computation, but of a set of stencil computations, with one or more stencil shapes (neighboorhood), and onto one or more data.
+Moreover, a numerical simulation also performs additionnal auxiliary computations which do not involve neighborhood, called local computations.
+This paper targets this general case of numerical simulations that we call \emph{multi-stencil programs}.
+The work presented produces a general parallelization structure of the overall simulation, while most stencil solutions produce optimized codes for a single stencil code.
+As it will be detailed in the related work, this work is complementary to the optimization and parallelization of a single stencil computation.

### 2015-07-29 WOLFHPC15/article/ms_msc.tex
fix compiler 29_07

@@
 dataread ::= dataid "[" neighborid "]"
+          |  dataid
 \end{lstlisting}
@@
 computations:
-  b[allcell]=c0(a[])
+  b[allcell]=c0(a)
   c[alledgex]=c1(b[n1])
-  d[alledgex]=c2(c[])
-  e[alledgey]=c3(c[])
+  d[alledgex]=c2(c)
+  e[alledgey]=c3(c)
   f[allcell]=c4(d[n1])
@@
   h[alledgex]=c6(f)
-  i[allcell]=c7(g[],h[])
+  i[allcell]=c7(g,h)
   j[partedgex]=c8(i[n1])

### 2015-07-29 WOLFHPC15/article/component.tex
WIP Section Components

--- 2015-07-24 WOLFHPC15/article/component.tex
@@
+%----------------------------------------
+\subsection{Overview}
+%----------------------------------------
+
+This section describes both the projection of $\Gamma_{tsp}$ to
+components, and the remaining needed components to execute a
+multi-stencil application. But, first the section gives an overview of
+component models, and then it details \llc, the component model used in
+this work.
+
+%----------------------------------------
+\subsection{Component Model Overview}
+%----------------------------------------
+Component model is an domain of software
+engineering~\cite{Szyperski:2002:CSB:515228}, which promotes code
+re-use and separations of concerns. An application is made by a set of
+component; a component being a kind of black box that interacts only
+through ports: ports specify the services provided and required.
+
+Component models are known to improved code re-use, separation of
+concerns, and maintainability of
+applications~\cite{Szyperski:2002:CSB:515228}. More recent works have
+also shown that component models can handled performance, scalability,
+and portability of applications~\cite{Bernholdt01052006,
+  bigot:inria-00388508, l2c}. Many component models exist, each of
+them with its own specifications and singularities. In distributed
+computing, well known models are CCM~\cite{corba:omg06} (CORBA
+Component Model), and GCM~\cite{Baude} (Grid Component Model); in HPC,
+there are CCA~\cite{Bernholdt01052006} (Common Component
+Architecture), and \llc~\cite{l2c} (Low Level Component), for example.
+
+%----------------------------------------
+\subsection{\llc}
+%----------------------------------------
+
+The remaining of this paper focuses on \llc, in particular for the
+experiments. Let describe it in more details.
+
+\llc is a minimalist \texttt{C++} based HPC component model where a
+component extends the concept of class by specifying in its interfaces
+the services that it offers ($provide$ ports) and that it needs,
+either a single service instance ($use$ ports), or multiple service
+instances ($use-multiple$ ports). Services are \texttt{C++}
+interfaces. \llc also offer $MPI$ ports that enable components to
+share a MPI communicators.
+%
+As illustrated in Figures~\ref{fig:ports}, a $provide$ port is
+represented by a white circle on the left of the component, a $use$
+port by a black circle, a $use-multiple$ port by a black circle with a
+white $m$ in it, both on the right of the component. MPI PORT!!!
+
+A classical \llc-based application is a static \emph{assembly} of
+components made of instances of component and connections between
+ports. Such an assembly is describe in LAD (\llc Assembly Descriptor),
+an XML dialect~\cite{l2c}.
+
+\begin{figure}[t]
+a)\hspace{2mm} \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (C) at (0,0) {$C_0$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (1.5,0) {};
+   \node[provide,right=1.5cm of u] (p1) {};
+   \node[component,right=1.5cm of p1] (C1) {$C_1$};
+   \node[use,right=1.5cm of C1] (um) {$m$};
+
+  \path[-]
+    (p) edge node {$p$} (C)
+    (C.east) edge node {$u$} (u)
+    (C1)  edge node {$u'$} (um)
+    (p1) edge node {$p'$} (C1);
+\end{tikzpicture}
+\\\\
+b)\hspace{2mm}\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (C) at (0,0) {$C_0$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (1.5,0) {};
+   \node[component,right=1.5 of u] (C1) {$C_1$};
+   \node[use,right=1.5 of C1] (um) {$m$};
+
+  \path[-]
+    (p) edge node {} (C)
+    (C) edge node {} (u)
+    (C1)  edge node {} (um)
+    (u) edge node {} (C1);
+\end{tikzpicture}
+\\\\
+c) MPI!!!
+\\\\
+\caption{a) Component $C_0$ has a provide port ($p$) and a use port ($u$); Component $C_1$ has also a provide port ($p'$) but a use multiple port ($u'$).
+b) When a use port is connected to a provide port, only the use port is shown and by convention it has to be the left component.}
+\label{fig:ports}
+\end{figure}
+
+%% In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}.
+As the use port is on the right of its component, and the provide port on the left, in a component assembly the component on the left uses the provide port of the component on the right.
+
 %----------------------------------------

### 2015-07-29 WOLFHPC15/article/article.tex
fix compiler 29_07
fixing author
Add xspace
wolfhpc 29_07

@@
 \usepackage{cite}
+\usepackage{subfig}
+\usepackage{amsmath}
+\usepackage{xspace}
 %----------TIKZ
@@
     circle,draw=black,fill=black,scale=0.5,text=white
+    },
+    mpi/.style={
+    rectangle,draw=black,fill=black,scale=0.8,text=white
     },
@@
 
+\newcommand{\llc}{L\textsuperscript{2}C\xspace}
+
 \begin{document}
@@
 
-\numberofauthors{2} 
+\numberofauthors{3} 
 \author{
@@
        \affaddr{USR 3441}\\
-       \affaddr{91191 Gif-sur-Yvette Cedex, France}\\
+       \affaddr{91191 Gif-sur-Yvette, France}\\ % remove cedex
        \email{julien.bigot@cea.fr}

### 2015-07-29 WOLFHPC15/article/article.tex
overview helene 29_07

@@
     rectangle,rounded corners=3pt,dashed,draw=black
-    },
-    smallcp/.style={
-    rectangle,rounded corners=3pt,draw=black,font=\tiny,
-    },
-    resource/.style={
-    rectangle,rounded corners=3pt, dashed,draw=black,font=\tiny,
-    },
-    connection/.style={
-    decoration={ markings,
-      mark=at position 1 with {\node[fill=black,draw=black,circle,inner sep=2pt](conn){};}
-      },
-    postaction={decorate}
-    },
-    bconnection/.style={
-    decoration={ markings,
-      mark=at position 1 with {\node[fill=black,draw=black,circle,inner sep=2pt](conn){};}
-      },
-    postaction={decorate}
-    },
-    mpiconn/.style={
-    decoration={ markings,
-      mark=at position 0.99 with {\node[fill=black,draw=black,rectangle,,inner sep=2pt](mpi){};}
-      },
-    postaction={decorate}
     },

### 2015-07-29 WOLFHPC15/article/mscac.tex
overview helene 29_07

@@
-The solution presented in this paper is a domain specific language, to describe a multi-stencil program (a numerical simulation), and a compiler which transforms this description to a parallel component-based application of the overall simulation.
+The solution presented in this paper contains a domain specific language to describe a multi-stencil program, i.e.
+an overall numerical simulation, and a compiler which successively parses the front-end language, transforms the internal representation, and finally dumps the representation to a back-end code.
 
-The MS language (MSL) is an agnostic descriptive language for multi-stencil simulations.
-Agnostic means that the description of a numerical simulation, using MSL, does not need any knowledge on the mesh type, the stencil shapes, or the final implementation constraints.
-Descriptive, on the other hand, means that the terminal states of the language grammar are string identifiers and that the language does not handle description of numerical computations, but only description of a multi-stencil simulation, in a coarse-grain fashion.
+The proposed DSL, called MSL, is an agnostic descriptive language for multi-stencil simulations.
+Agnostic means that the description of a numerical simulation does not depend on implementation constraints as, for example, the type of mesh (structured, unstructured) and its associated interfaces (how to define a stencil shape etc.).
+Descriptive, on the other hand, means that the terminal states of the language grammar are string identifiers.
+As a result, MSL is not made to handle the expression of numerical computations, but only the description of a multi-stencil simulation, in a coarse-grain fashion.
+In a MSL description, six main sections are defined, as the six tuples of the definition of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\mathcal{D}_m,\mathcal{D}_c,\Delta,\Gamma)$.
 
-As illustrated in Figure~\ref{fig:mscac}, the MS language compiler, called MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of two different parts.
+As illustrated in Figure~\ref{fig:mscac}, the compiler, called MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of three different phases.
+The first phase of the compiler is to \emph{parse} the MSL description.
+Second, the compiler analyzes and transforms the computation part of the description, in other words $\Gamma$.
+This phase of compilation, called MSC, is responsible for the automatic parallelization of the overall numerical simulation.
+It transforms $\Gamma$ to a new parallel representation.
+Finally, the parallelized representation can be dumped to an actual implementation of the parallel structure of the simulation.
+
+% As illustrated in Figure~\ref{fig:mscac}, the MS language compiler, called MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of two different parts.
 First, the computation compiler, called \emph{MSC}, is responsible for the parallelization of computations and for its dump to a component assembly; second, the data compiler, called \emph{MSD}, handles a distributed data structure, data mapped onto it, and their relations with computations, as a final overall component assembly.
-
 \begin{figure}[h!]
@@
   \node (MSL) {MS Language};
-  \node[outer,below=of MSL,draw=black] (MSCAC) {
+  \node[outer,below=0.5 of MSL,draw=black] (MSCAC) {
     \begin{tikzpicture}
-      \node [inner,draw=black,thick] (MSC)  {MSC};
-      \node [inner,right=of MSC,draw=black,thick,,fill=gray!30!white,dotted] (MSD)  {MSD};
+      \node [inner,draw=black,thick] (pars)  {Parser};
+      \node [inner,draw=black,thick,below=0.5cm of pars] (MSC)  {MSC};
+      \node [inner,below=0.5cm of MSC,draw=black,thick] (Dump)  {Dump};
+
+      \draw[->] (pars) -- (MSC);
+      \draw[->] (MSC) -- (Dump);
     \end{tikzpicture}
   };
-  \node [below=of MSCAC] (CA) {Component Assembly};
-  \node [left=0.1 of MSCAC] (CAid) {MSCAC};
-  \draw[->] (MSL) -- (MSCAC);
-  \draw[->] (MSCAC) -- (CA);
+  \node [below=0.5 of MSCAC] (CA) {Component Assembly};
+  \node [rotate=90,above left=0.5 of MSC] (CAid) {MSCAC};
+  \draw[->] (MSL) -- (pars.north);
+  \draw[->] (Dump.south) -- (CA);
 \end{tikzpicture}
-\caption{MS Language and MSCAC compiler}
+\caption{MSCAC compiler}
 \label{fig:mscac}
@@
 
-This paper presents the MS language and the MSC part of the compiler.
+In this paper are presented the language MSL, the MSC compilation phase, and the dump compilation phase.
+Parallelization techniques handled by MSC are \emph{data parallelism}, where data are splitted among processors or cores while applying a single program everywhere, and \emph{task parallelism} where the program is transformed into a dependency graph of tasks.
+From a dependency graph different strategies can be used to schedule tasks, in other words to control their execution in a valid order.
+The MSC compiler presented in this paper handles the transformation of the dependency graph to a static schedule of the tasks.
+
+The final phase of compilation, the dump, uses the static schedule of the tasks, identified from $\Gamma$, and additionnal information on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{D}_m$ and $\mathcal{D}_c$, from the \emph{parser}, to produce a back-end solution which represents the parallel pattern or skeleton of the overall simulation.
+We could have propose a dump to OpenMP~\cite{660313}, however this paper presents a first contribution to dump a DSL to a component-based runtime.
+Component models, where a component represents a clear and independent functionnality of the application, have first been proposed in Software engineering and have proved many times good properties for code re-use, separation of concerns, maintainability and productivity of codes.
+Those properties, on the other hand, are recurrent stated problems of parallel applications.
+
+The back-end of the compiler is to build a parallel assembly of components, in other words appropriate connections between functionnalities of the numerical simulation.
+This assembly is illustrated in Figure~\ref{fig:assembly} which is explained in details in Section~\ref{sect:component}.
+As illustrated in the figure, each component is linked to one or more tuple of the multi-stencil program definition: \emph{DDS} component represents $\mathcal{M}$, $\mathcal{D}_m$; \emph{Data} component represents $\Delta$ and $\mathcal{D}_c$; \emph{Time} component represents $T$, and \emph{Computations} component (which generated by MSC) represents $\Gamma$.
+
+The final back-end component assembly is a ready-to-fill parallel structure where the user only has to actually write computation codes described in the MSL description.
+To write those computations, interfaces of the choosen implementation have to be used.
+For example, the back-end evaluated in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}, however a work in progress replace this DDS component by another one using Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} data structures.
+
+% This paper presents the MS language and the MSC part of the compiler.
 The MSC compiler produces a static scheduling of the set of computations, ready to dump to a parallel language, such as OpenMP~\cite{660313} or HPF~\cite{219857}, but also ready to dump to a component assembly which is the main contribution of this work.
@@
 
-It has to be noticed that in this paper, the MSD compiler is a first adhoc version which works for a precise distributed data structure, the SkelGIS library~\cite{HeleneLS13,HeleneLS14,HeleneEuroPar14,CPE:CPE3494}.
+% It has to be noticed that in this paper, the MSD compiler is a first adhoc version which works for a precise distributed data structure, the SkelGIS library~\cite{HeleneLS13,HeleneLS14,HeleneEuroPar14,CPE:CPE3494}.
 However, using the system of python string templates~\footnote{\url{https://docs.python.org/2/library/string.html}} to define the component assembly part which manages data, it is already possible to use another distributed data structure, as for example Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.
@@
 
-The overall compiler MSCAC produces a ready-to-fill component-based parallel structure of the simulation.
+% The overall compiler MSCAC produces a ready-to-fill component-based parallel structure of the simulation.
 In this final component assembly, it is needed to write computation components, using interfaces of the distributed data structure choosen by MSD.

### 2015-07-29 WOLFHPC15/article/mscac.tex
overview helene 29_07

@@
 The first phase of the compiler is to \emph{parse} the MSL description.
-Second, the compiler analyzes and transforms the computation part of the description, in other words $\Gamma$.
+Second, the compiler analyzes and transforms the computation part of the description, in other words $\Gamma$, to an intermediate representation.
 This phase of compilation, called MSC, is responsible for the automatic parallelization of the overall numerical simulation.
-It transforms $\Gamma$ to a new parallel representation.
-Finally, the parallelized representation can be dumped to an actual implementation of the parallel structure of the simulation.
+Finally, the parallel intermediate representation can be dumped to an actual implementation of the parallel structure of the simulation.
 
@@
 This assembly is illustrated in Figure~\ref{fig:assembly} which is explained in details in Section~\ref{sect:component}.
-As illustrated in the figure, each component is linked to one or more tuple of the multi-stencil program definition: \emph{DDS} component represents $\mathcal{M}$, $\mathcal{D}_m$; \emph{Data} component represents $\Delta$ and $\mathcal{D}_c$; \emph{Time} component represents $T$, and \emph{Computations} component (which generated by MSC) represents $\Gamma$.
+As illustrated in the figure, each component is linked to one or more tuple of the multi-stencil program definition: \emph{DDS} component represents $\mathcal{M}$ and $\mathcal{D}_m$; \emph{Data} component represents $\Delta$ and $\mathcal{D}_c$; \emph{Time} component represents $T$, and \emph{Computations} component (which is generated by MSC) represents $\Gamma$.
 
-The final back-end component assembly is a ready-to-fill parallel structure where the user only has to actually write computation codes described in the MSL description.
+The final back-end component assembly is a ready-to-fill parallel structure where the user has to write computation codes described in the MSL description.
 To write those computations, interfaces of the choosen implementation have to be used.
-For example, the back-end evaluated in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}, however a work in progress replace this DDS component by another one using Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} data structures.
+For example, the back-end evaluated in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}, however a work in progress replaces this DDS component by another one using Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} data structure.
 

### 2015-07-29 WOLFHPC15/article/mscac.tex
overview helene 29_07

@@
 
-In this paper are presented the language MSL, the MSC compilation phase, and the dump compilation phase.
+%In this paper are presented the language MSL, the MSC compilation phase, and the dump compilation phase. 
 Parallelization techniques handled by MSC are \emph{data parallelism}, where data are splitted among processors or cores while applying a single program everywhere, and \emph{task parallelism} where the program is transformed into a dependency graph of tasks.
@@
 
-% This paper presents the MS language and the MSC part of the compiler.
-The MSC compiler produces a static scheduling of the set of computations, ready to dump to a parallel language, such as OpenMP~\cite{660313} or HPF~\cite{219857}, but also ready to dump to a component assembly which is the main contribution of this work.
-
-\begin{figure*}
+\begin{figure}
 \begin{center}
@@
    \node (Ds) at (-1.5,0) {start};
-   \node[use] (Du1) at (1,0.5) {};
-   \node[use] (Du2) at (1,-0.5) {};
+   \node[use,right=1.5cm of D] (Du1) {};
+   \node[use,below=0.8cm of D] (Du2) {};
 
-   \node[component,right=1cm of Du2] (DA) {$DriverApp$};
-   \node[use] (DAu1) at (3.5,0) {$m$};
-   \node[use] (DAu2) at (3.5,-1) {};
+   \node[provide,below=0.15 of Du2] (DAp) {};
+   \node[component,below=0.8cm of DAp] (DA) {$DriverApp$};
+   \node[use,right=1.7cm of DA] (DAu1) {$m$};
+   \node[use,below=0.8cm of DA] (DAu2) {};
 
-   \node[component] (T) at (4.5,-1) {$Time$};
+   \node[provide,below=0.15 of DAu2] (Tp) {};
+   \node[component,below=0.8cm of Tp] (T) {$Time$};
    \node[use,right=1cm of T] (Tu) {};
-   \node[component,right=2cm of Tu] (C) {$Computations$};
-   \node[use,right=2cm of C] (Cu) {};
-   \node[below=0.2cm of Cu] (star) {$*$};
 
-   \node[component,right=8cm of DAu1] (Data) {$Data$};
-   \node[use,right=1cm of Data] (Datau) {};
+   \node[provide,right=0.15 of Tu] (Cp) {};
+   \node[component,right=2cm of Cp] (C) {$Computations$};
+   \node[use,above=0.8cm of C] (Cu) {};
+   \node[right=0.2cm of Cu] (star) {$*$};
 
-   \node[component,right=13cm of Du1] (DDS) {$DDS$};
+   \node[provide,right=0.2 of DAu1] (Datap1) {};
+   \node[provide,above=0.15 of Cu] (Datap2) {};
+   \node[component,above=0.8cm of Datap2] (Data) {$Data$};
+   \node[use,above=0.8cm of Data] (Datau) {};
+
+   \node[provide,right=0.15 of Du1] (DDSp1) {};
+   \node[provide,above=0.15 of Datau] (DDSp2) {};
+   \node[component,above=0.8cm of DDSp2] (DDS) {$DDS$};
 
@@
     (Dp) edge node {} (D)
-    (D.east) edge node {} (Du1)
+    (D) edge node {} (Du1)
         edge node {} (Du2)
-    (Du1) edge node {} (DDS.west)
-    (Du2) edge node {} (DA)
-    (DA.east) edge node {} (DAu1)
+    (DDSp1) edge node {} (DDS)
+    (DA) edge node {} (DAu1)
            edge node {} (DAu2)
-    (DAu1) edge node {} (Data.west)
-    (DAu2) edge node {} (T)
+    (DAp) edge node {} (DA)
+    (Tp) edge node {} (T)
     (T)  edge node {} (Tu)
-    (Tu) edge node {} (C)
+    (Cp) edge node {} (C)
     (C) edge node {} (Cu)
-    (Cu) edge node {} (Data.west)
+    (Datap2) edge node {} (Data)
+    (Datap1) edge node {} (Data)
     (Data) edge node {} (Datau)
-    (Datau) edge node {} (DDS.west);
+    (DDSp2) edge node {} (DDS);
 \end{tikzpicture}
@@
 \end{center}
-\end{figure*}
+\end{figure}
+
+% This paper presents the MS language and the MSC part of the compiler.
+The MSC compiler produces a static scheduling of the set of computations, ready to dump to a parallel language, such as OpenMP~\cite{660313} or HPF~\cite{219857}, but also ready to dump to a component assembly which is the main contribution of this work.
+
+% \begin{figure*}
+% \begin{center}
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[component] (D) at (0,0) {$Driver$};
+%    \node[provide] (Dp) at (-1,0) {};
+%    \node (Ds) at (-1.5,0) {start};
+%    \node[use] (Du1) at (1,0.5) {};
+%    \node[use] (Du2) at (1,-0.5) {};
+
+%    \node[component,right=1cm of Du2] (DA) {$DriverApp$};
+%    \node[use] (DAu1) at (3.5,0) {$m$};
+%    \node[use] (DAu2) at (3.5,-1) {};
+
+%    \node[component] (T) at (4.5,-1) {$Time$};
+%    \node[use,right=1cm of T] (Tu) {};
+%    \node[component,right=2cm of Tu] (C) {$Computations$};
+%    \node[use,right=2cm of C] (Cu) {};
+%    \node[below=0.2cm of Cu] (star) {$*$};
+
+%    \node[component,right=8cm of DAu1] (Data) {$Data$};
+%    \node[use,right=1cm of Data] (Datau) {};
+
+%    \node[component,right=13cm of Du1] (DDS) {$DDS$};
+
+%   \path[-]
+%     (Dp) edge node {} (D)
+%     (D.east) edge node {} (Du1)
+%       edge node {} (Du2)
+%     (Du1) edge node {} (DDS.west)
+%     (Du2) edge node {} (DA)
+%     (DA.east) edge node {} (DAu1)
+%       edge node {} (DAu2)
+%     (DAu1) edge node {} (Data.west)
+%     (DAu2) edge node {} (T)
+%     (T) edge node {} (Tu)
+%     (Tu) edge node {} (C)
+%     (C) edge node {} (Cu)
+%     (Cu) edge node {} (Data.west)
+%     (Data) edge node {} (Datau)
+%     (Datau) edge node {} (DDS.west);
+% \end{tikzpicture}
+% \caption{Shape of the generated component assembly.}
+% \label{fig:assembly}
+% \end{center}
+% \end{figure*}
 

### 2015-07-29 WOLFHPC15/article/introduction.tex
introduction helene 29_07

--- 2015-07-20 WOLFHPC15/article/introduction.tex
@@
-\begin{itemize}
-\item DSL interest: simplified HPC programming + perf
-\item DSL limits: Lots of DSLs difficult to write, to maintain, to make portable etc.
-\item Components introduction, this paper takes a first example of DSL to components
-\item The domain targeted is multi-stencil simulations
-\item Contributions: DSL to components try + agnostic coarse grain DSL for multi-stencil programs
-\item Outline
-\end{itemize}
+Programming complex scientific applications on high performance heterogenous architectures, such as clusters of nodes containing multicores and accelerators, requires for the scientists to be an expert in his own domain, but also to be an expert in low level parallel and high performance programming.
+It can takes months to years for a scientist to learn HPC programming, as it is not his main activity.
+And while learning HPC, hardware continue to change to be more powerfull, but also more difficult to understand and to use.
+
+Domain Specific Languages (DSLs) are a promising solution to hide intricate details of HPC programming from the scientist, while producing automatically an optimized and efficient final code.
+Thus, it seems that the clue for simplicity and efficiency on modern architectures is to specialize a language to a given domain, from which one or more efficient parallelizations and optimizations can be extracted.
+However, if the specific nature of a DSL simplifies HPC programming for scientists, its implementation cost (language, compiler, and runtime) is difficult to amortize.
+Actually, programming difficulties, as well as non-maintainability and non-portability seems to be deported from the scientist to the DSL designer.
+Moreover, one can consider that it is needed to write as much DSLs as scientific domains, which keeps the amount of work to write and maintain up-to-date DSLs out of hands.
+
+For this reason, solutions to ease the design and implementation of DSLs have recently appear~\cite{Fernandez:2014:DFL:2691166.2691168}.
+The main idea of those solutions is to propose a single DSL to write new DSLs.
+In this paper, we study a first contribution to a new approach to bring maintainability and code-reuse in DSLs.
+This approach proposes a transformation from a DSL to a component-based runtime which has recently prove its advantages for HPC programming~\cite{l2c}.
+
+To discuss and analyze advantages of such an approach, a proper new DSL for \emph{multi-stencil} programs, and its compiler are presented in this paper.
+Many solutions to automatically optimize and parallelize partial differential equations (PDEs) solvers have been proposed in the litterature, as specific libraries~\cite{petsc-efficient,Trilinos-Overview,CPE:CPE3494}, and DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
+Most of the time, proposed DSLs focuss on the parallelization and the optimization of a single numerical computation, also called a stencil.
+However, a real case numerical simulation is most of the time not composed of a single stencil computation, but of a set of stencil computations and a set of additionnal auxiliary local computations.
+The DSL proposed in this paper, called MSL, is a DSL to generate a parallel component-based structure of an overall numerical simulation, that we call a \emph{multi-stencil programs}.
+As it will be detailed in the related work, this work is complementary to the optimization and parallelization of a single stencil computation.
+
+The rest of this paper is organized as follows.
+First concepts of \emph{stencil} and \emph{multi-stencil porgams} will be formalized in Section~\ref{sect:concept}.
+Sections~\ref{sect:mscac} and~\ref{sect:msmsc} respectively give an overview of the solution presented in this paper, and details of the MSL language and its compiler.
+Section~\ref{sect:component} introduce component models and runtimes and explains the automatic generation of the component-based parallel structure.
+The end of this section also discuss on the interest of this new approach.
+Section~\ref{sect:eval} takes a real-case simulation, solving the Shallow-water equations, and analyze some performance results.
+Finally, Sections~\ref{sect:related} states on related work, and Section~\ref{sect:conclusion} gives conclusions and perspectives on this work.

### 2015-07-30 WOLFHPC15/article/introduction.tex
introduction fix helene 30_07

@@
 Programming complex scientific applications on high performance heterogenous architectures, such as clusters of nodes containing multicores and accelerators, requires for the scientists to be an expert in his own domain, but also to be an expert in low level parallel and high performance programming.
-It can takes months to years for a scientist to learn HPC programming, as it is not his main activity.
-And while learning HPC, hardware continue to change to be more powerfull, but also more difficult to understand and to use.
+It can takes months to years to a scientist to learn HPC programming, as it is not his main activity.
+And while learning HPC programming, hardware continue to grow, to be more powerfull, but growing at the same time its complexity and the difficulty to use it.
 
@@
 Actually, programming difficulties, as well as non-maintainability and non-portability seems to be deported from the scientist to the DSL designer.
-Moreover, one can consider that it is needed to write as much DSLs as scientific domains, which keeps the amount of work to write and maintain up-to-date DSLs out of hands.
+Moreover, one can consider that sientists need as much DSLs as domains, which keeps the amount of work to write and maintain up-to-date DSLs out of hands.
 
@@
 Many solutions to automatically optimize and parallelize partial differential equations (PDEs) solvers have been proposed in the litterature, as specific libraries~\cite{petsc-efficient,Trilinos-Overview,CPE:CPE3494}, and DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
-Most of the time, proposed DSLs focuss on the parallelization and the optimization of a single numerical computation, also called a stencil.
+Most of the time, DSLs focuss on the parallelization and the optimization of a single numerical computation, also called a stencil.
 However, a real case numerical simulation is most of the time not composed of a single stencil computation, but of a set of stencil computations and a set of additionnal auxiliary local computations.
@@
 Sections~\ref{sect:mscac} and~\ref{sect:msmsc} respectively give an overview of the solution presented in this paper, and details of the MSL language and its compiler.
-Section~\ref{sect:component} introduce component models and runtimes and explains the automatic generation of the component-based parallel structure.
-The end of this section also discuss on the interest of this new approach.
+Section~\ref{sect:component} introduces component models and runtimes and explains the automatic generation of the component-based parallel structure.
+The end of Section~\ref{sect:component} also discusses on interest of this new approach.
 Section~\ref{sect:eval} takes a real-case simulation, solving the Shallow-water equations, and analyze some performance results.
-Finally, Sections~\ref{sect:related} states on related work, and Section~\ref{sect:conclusion} gives conclusions and perspectives on this work.
+Sections~\ref{sect:related} states on related work, and finally Section~\ref{sect:conclusion} gives conclusions and perspectives on this work.

### 2015-07-30 WOLFHPC15/article/concept.tex
concepts fix helene 30_07

@@
 To ease programming of stencil computations, and because of the relative regularity of such a computation, many solutions (languages or libraries) propose an easy way to describe the stencil code, while producing transparently an optimized and parallelized stencil code.
-For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, respectively proposing an optimized cache tiling technique, and a separation of concerns by separating parallelization strategy from the computation description.
+For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, respectively proposing an optimized cache tiling technique, and a separation of concerns between parallelization strategies and computation descriptions.
 
@@
 A mesh is a connected undirected graph without bridges (an edge is a bridge if its removal results in two disconnected graphs), where nodes and edges are linked to form cells (closure).
-For example, a structured mesh is illustrated in Figure~\ref{fig:mesh}, each cell contains four nodes and four edges.
-Second, \emph{mesh domains} are subset elements of the mesh, such as the center of cells, edges or nodes, for example.
+For example, a structured mesh is illustrated in Figure~\ref{fig:mesh}.
+Each cell contains four nodes and four edges.
+Second, \emph{mesh domains} are subset elements of the mesh, such as the center of cells, edges or nodes.
 In Figure~\ref{fig:mesh} two mesh domains are illustrated, the center of the cells (in red) and edges on the horizontal axe (in blue on left and right of each cell).
@@
 Sometimes, the neighborhood shape can also access a different mesh domain, as for example in Figure~\ref{fig:ex2}.
-Actually, in this computation, the neighborhood contains edges on the left and on the right of the cell.
+Actually, in this computation, the neighborhood contains edges on the left and on the right of a cell.
 As an example, the numerical expression of the first example could be
@@
 \end{equation}
-where $R$ is a list of pair $(r,n)$ with $r \in \Delta$ is a data read by the computation, and $n$ is a stencil shape (a neighborhood).
+where $R$ is a set of pair $(r,n)$ with $r \in \Delta$ is a data read by the computation, and $n$ is a stencil shape (a neighborhood) applied on $r$.
 The data written by the computation is denoted $w \in \Delta$, $exp$ is the numerical expression of the stencil computation.

### 2015-07-30 WOLFHPC15/article/mscac.tex
overview helene 29_07
introduction helene 29_07
overview fix helene 30_07

@@
 The first phase of the compiler is to \emph{parse} the MSL description.
-Second, the compiler analyzes and transforms the computation part of the description, in other words $\Gamma$, to an intermediate representation.
+Second, the compiler analyzes and transforms the computation part of the description $\Gamma$, to an intermediate representation.
 This phase of compilation, called MSC, is responsible for the automatic parallelization of the overall numerical simulation.
-Finally, the parallel intermediate representation can be dumped to an actual implementation of the parallel structure of the simulation.
+Finally, the parallel intermediate representation is dumped to an actual implementation of the parallel pattern of the simulation.
 
@@
 From a dependency graph different strategies can be used to schedule tasks, in other words to control their execution in a valid order.
-The MSC compiler presented in this paper handles the transformation of the dependency graph to a static schedule of the tasks.
+The MSC compiler presented in this paper handles the transformation of the dependency graph to a static scheduling.
 
-The final phase of compilation, the dump, uses the static schedule of the tasks, identified from $\Gamma$, and additionnal information on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{D}_m$ and $\mathcal{D}_c$, from the \emph{parser}, to produce a back-end solution which represents the parallel pattern or skeleton of the overall simulation.
+The final phase of compilation, the dump, uses the static schedule of the tasks, created from $\Gamma$, and additionnal information on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{D}_m$ and $\mathcal{D}_c$, given by the \emph{parser}, to produce a back-end solution which represents the parallel pattern or skeleton of the overall simulation.
 We could have propose a dump to OpenMP~\cite{660313}, however this paper presents a first contribution to dump a DSL to a component-based runtime.
@@
 
-The back-end of the compiler is to build a parallel assembly of components, in other words appropriate connections between functionnalities of the numerical simulation.
+The back-end of the compiler is to build a parallel assembly of components, in other words the appropriate connections between functionnalities of the numerical simulation.
 This assembly is illustrated in Figure~\ref{fig:assembly} which is explained in details in Section~\ref{sect:component}.
@@
 To write those computations, interfaces of the choosen implementation have to be used.
-For example, the back-end evaluated in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}, however a work in progress replaces this DDS component by another one using Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} data structure.
+For example, the back-end evaluated in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}.
+A work in progress replaces this DDS component by another one using Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} data structure.
 
@@
    \node[use,right=1cm of T] (Tu) {};
+   \node[left=0.8cm of T] (tt) {$T$};
 
@@
    \node[right=0.2cm of Cu] (star) {$*$};
+   \node[right=1.5cm of C] (gamma) {$\Gamma$};
 
@@
    \node[use,above=0.8cm of Data] (Datau) {};
+   \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}_c$};
 
@@
    \node[component,above=0.8cm of DDSp2] (DDS) {$DDS$};
+   \node[right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{D}_m$};
 
@@
 \end{tikzpicture}
-\caption{Shape of the generated component assembly.}
+\caption{Overall back-end component assembly}
 \label{fig:assembly}

### 2015-07-30 WOLFHPC15/article/introduction.tex
WIP intro

@@
 It can takes months to years to a scientist to learn HPC programming, as it is not his main activity.
-And while learning HPC programming, hardware continue to grow, to be more powerfull, but growing at the same time its complexity and the difficulty to use it.
+And while learning HPC programming, hardware continue to evolve, becoming more powerful but usually also more  complex and difficult to use efficiently.
 
-Domain Specific Languages (DSLs) are a promising solution to hide intricate details of HPC programming from the scientist, while producing automatically an optimized and efficient final code.
-Thus, it seems that the clue for simplicity and efficiency on modern architectures is to specialize a language to a given domain, from which one or more efficient parallelizations and optimizations can be extracted.
+Domain Specific Languages (DSLs) are a promising solution to hide intricate details of HPC programming from non experts, while producing automatically an optimized and efficient final code.
+Thus, it seems that a clue for simplicity and efficiency on modern architectures is to specialize a language to a given domain, from which one or more efficient parallelizations and optimizations can be extracted.
 However, if the specific nature of a DSL simplifies HPC programming for scientists, its implementation cost (language, compiler, and runtime) is difficult to amortize.
-Actually, programming difficulties, as well as non-maintainability and non-portability seems to be deported from the scientist to the DSL designer.
-Moreover, one can consider that sientists need as much DSLs as domains, which keeps the amount of work to write and maintain up-to-date DSLs out of hands.
+Actually, programming difficulties, as well as non-maintainability and non-portability seems to be deported from the scientist to DSL designers.
+Moreover, one can consider that end users need as much DSLs as domains, which keeps the amount of work to write and maintain up-to-date DSLs out of hands.
 
@@
 The main idea of those solutions is to propose a single DSL to write new DSLs.
-In this paper, we study a first contribution to a new approach to bring maintainability and code-reuse in DSLs.
-This approach proposes a transformation from a DSL to a component-based runtime which has recently prove its advantages for HPC programming~\cite{l2c}.
+In this paper, we study a first contribution to another approach to bring maintainability and code-reuse in DSLs.
+This approach proposes a transformation from a DSL to a component-based runtime.
+TO DEV BENEFITS OF COMPONENTS.
 
-To discuss and analyze advantages of such an approach, a proper new DSL for \emph{multi-stencil} programs, and its compiler are presented in this paper.
+To evaluate such an approach, a proper new DSL for \emph{multi-stencil} programs, and its compiler are presented in this paper.
 Many solutions to automatically optimize and parallelize partial differential equations (PDEs) solvers have been proposed in the litterature, as specific libraries~\cite{petsc-efficient,Trilinos-Overview,CPE:CPE3494}, and DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
@@
 The rest of this paper is organized as follows.
-First concepts of \emph{stencil} and \emph{multi-stencil porgams} will be formalized in Section~\ref{sect:concept}.
-Sections~\ref{sect:mscac} and~\ref{sect:msmsc} respectively give an overview of the solution presented in this paper, and details of the MSL language and its compiler.
-Section~\ref{sect:component} introduces component models and runtimes and explains the automatic generation of the component-based parallel structure.
-The end of Section~\ref{sect:component} also discusses on interest of this new approach.
-Section~\ref{sect:eval} takes a real-case simulation, solving the Shallow-water equations, and analyze some performance results.
-Sections~\ref{sect:related} states on related work, and finally Section~\ref{sect:conclusion} gives conclusions and perspectives on this work.
+First concepts of \emph{stencil} and \emph{multi-stencil porgams} are formalized in Section~\ref{sect:concept}.
+Section~\ref{sect:mscac} gives an overview of the solution presented in this paper while Section~\ref{sect:msmsc} details the MSL language and its compiler.
+Section~\ref{sect:component} introduces component models and then it describes the proposed component-based runtime and the automatic generation of its parallel compuration part.
+%The end of Section~\ref{sect:component} also discusses on interest of this new approach.
+Section~\ref{sect:eval} focuses on a real-case simulation, solving the Shallow-water equations, and analyzes some performance results.
+Sections~\ref{sect:related} states on related work, and finally Section~\ref{sect:conclusion} concludes and gives some perspectives.

### 2015-07-30 WOLFHPC15/article/ms_msc.tex
section 4 helene 30_07

--- 2015-07-24 WOLFHPC15/article/ms_msc.tex
@@
 %----------------------------------------
-\subsection{MS language}
+\subsection{Multi Stencil Language}
 %----------------------------------------
-The MS language (MSL) is an agnostic descriptive language for multi-stencil simulations.
-MSL follows the definition of a multi-stencil program~(\ref{eq:msp}), and of stencil and local computations~(\ref{eq:st}) and~(\ref{eq:loc}).
-Thus, it is composed of six sections: the mesh description; the mesh domains description; the computation domains description; the data description; the time loop description; and finally, the computations description.
-
-The overall grammar of a MSL program is given in Figure~\ref{fig:grammar}, where the terminals $meshid$, $meshdom$, $compdomid$, $dataid$, $compid$ ans $neighborid$ are string identifiers, and $num$ is an integer value.
-An example is then given to explain each section.
+As already mentionned, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations.
+Six main sections are required in a MSL description, as the six tuples of the formal definition of a multi-stencil program: the mesh description; the mesh entities description; the computation domains description and their dependencies; the data description; the time loop description; and finally, the computations description.
+Figure~\ref{fig:grammar} shows the grammar of MSL.
+Lines 1 to 9 of this figure define what is a MSL program with the different sections described above.
+In this section are described the different sections of a MSL program and an example.
+As MSL is a descriptive language, almost all terminals $meshid$, $meshdom$, $compdomid$, $dataid$, $compid$ ans $neighborid$ are string identifiers. Only $num$ is an integer terminal.
 
 \begin{figure}[h!]
-\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
+\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++,numbers=left]
 program ::= "mesh:" meshid 
-            "mesh domains:" listmeshdom
+            "mesh entities:" listmeshent
             "computation domains:" 
@@
             "computations:" listcomp
-listmeshdom ::= meshdom listmeshdom
-             |  meshdom
+listmeshent ::= meshent listmeshent
+             |  meshent
 listcompdom ::= compdom listcompdom
              |  compdom
-compdom ::= compdomid "in" listmeshdom
+compdom ::= compdomid "in" listmeshent
 listinde ::= inde listinde
           |  inde
-inde ::= compdomid and compdomid
+inde ::= compdomid "and" compdomid
 listdata ::= data listdata
           |  data
-data ::= dataid "," meshdom
+data ::= dataid "," meshent
 iteration ::= num
@@
 
-\paragraph{Mesh and mesh domains} A mesh is defined by a string identifier.
-As already explained, this work is for now limited to a single mesh.
-Each mesh domain is also represented by a string identifier.
-For example for a Cartesian mesh and three mesh domains we can define
-\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
+\paragraph{Mesh and mesh entities} As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is directly defined by a string identifier.
+Mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
+In the example given in Figure~\ref{fig:mslex} a Cartesian mesh, called $cart$, and three mesh entities, called $cell$, $edgex$ and $edgey$, are defined.
+
+\paragraph{Computation domains and their dependencies} A computation domain can be seen as a subpart of a mesh entity which will be used during one or more computations.
+For each computation domain is indicated the mesh entity from which it is built (lines 3-4 and 12-14 in Figure~\ref{fig:grammar}).
+By default two computation domains always intersect except when an independency relation is precised in the section \emph{independent} of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
+In the example of Figure~\ref{fig:mslex}, on lines 4 to 6 an entire computation domain is defined for each mesh entity, called $allcell$, $alledgex$ and $alledgey$, and on lines 7 and 8 two subparts of the mesh entity edgex are also defined, called $part1edgex$ and $part2edgex$.
+It is then precised on line 10 that $part1edgex$ and $part2edgex$ do not intersect.
+
+\paragraph{Data and time} A data is a quantity to simulate and is mapped onto a mesh entity, as illustrated in lines 7 and 18-20 f Figure~\ref{fig:grammar}.
+The time section simply indicates a number of iterations to perform in the simulation.
+In the example of Figure~\ref{fig:mslex} are defined ten data and 500 iterations (lines 11-22).
+
+\paragraph{Computations description} Finally the language follows the two definitions of a stencil computation and a local computation given in Section~\ref{sect:multistencil}.
+However as MSL does not handle numerical expressions, the term $exp$ of Definition~\ref{eq:st} and Definition~\ref{eq:loc} is not used in the language.
+The exact code of the computation as to be writtenin the final ready-to-fill parallel pattern of the simulation, using external languages or interfaces.
+Moreover, to propose an intuitive syntax, information are given in a different order than in formal definitions.
+Thus, the description of a computation, as indicated in lines 24 and 25 of Figure~\ref{fig:grammar}), consists in first the identifier of the data to compute ($w$ in the formal description), brackets indicated the computation domain on which the computation is applied ($d_c$ in the formal definition), an identifier for the computation, and the set of identifiers for the data read by the computation.
+For each data read by the computation brackets indicate the stencil shape identifier used on the data ($R$ in the formal definition of a stencil).
+If the computation is local no brackets appear ($R_l$ in the formal definition).
+In the example of Figure~\ref{fig:mslex}, nine computations are described.
+For example, on line 24, the data $b$ is computed on the computation domain $allcell$ by the computation $c_0$ which read data $a$ without neighborhood shape.
+\begin{figure}
+\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++,numbers=left]
 mesh: cart
 mesh domains: cell,edgex,edgey
-\end{lstlisting}
-
-\paragraph{Computation domains} A computation domain can be seen as a subpart of a meh domain which will be used during the computation phase.
-For each computation domain is indicated the mesh domain identifier from which it is built.
-By default two computation domains always intersect except when an independent relation is precised in the section \emph{independent} of the language.
-In the example below, an entire computation domain is defined for each mesh domain, and additionnally two subpart of the mesh domain edgex are defined.
-It is then precised that part1edgex and part2edgex do not intersect.
-\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
 computation domains:
@@
   part1edgex and part2edgex
-\end{lstlisting}
-
-\paragraph{Data and time} A data is a quantity to simulate and is mapped onto a mesh domain.
-The time section simply indicates a number of iterations to perform in the simulation.
-For example we can define ten data and 500 iterations as below
-\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
 data:
@@
 time:500
-\end{lstlisting}
-
-\paragraph{Computations description} The most important part of MSL is the description of the simulation computations.
-The language follows the three definitions given in Section~\ref{sect:multistencil}, however as MSL is not made to express numerical computations, the element $exp$ of definitions~\ref{eq:st} and~\ref{eq:loc} is not given in the language.
-Moreover, to propose a syntax close to an imperative code, information are given in a different order.
-Thus, the description of a computation consists of the identifier of the data to compute, the computation domain (which is a subpart of the mesh domain of the data) on which it is computed into the brackets, an identification name of the computation, and a set of identifiers for the data read by the computation.
-For each data read by the computation brackets indicate the stencil shape identifier used on the data.
-If the computation is local brackets stay empty.
-In the example below, the data $b$ is computed on the computation domain $allcell$ by the computation $c_0$ which read data $a$ without neighborhood shape.
-\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++]
 computations:
-  b[allcell]=c0(a[])
+  b[allcell]=c0(a)
   c[alledgex]=c1(b[n1])
-  d[alledgex]=c2(c[])
-  e[alledgey]=c3(c[])
+  d[alledgex]=c2(c)
+  e[alledgey]=c3(c)
   f[allcell]=c4(d[n1])
@@
   h[alledgex]=c6(f)
-  i[allcell]=c7(g[],h[])
+  i[allcell]=c7(g,h)
   j[partedgex]=c8(i[n1])
 \end{lstlisting}
+\caption{Example of a MSL program}
+\label{fig:mslex}
+\end{figure}
 
 %----------------------------------------
-\subsection{MSC compiler}
+\subsection{MSC transformation}
 %----------------------------------------
-The MSC compiler, which is a subpart of the overall compiler MSCAC, performs an algorithm composed of six main steps described bellow:
+The MSC transformation, which is a subpart of the overall compiler MSCAC, use the ordered list of computations $\Gamma$, which can directly been extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
+This transformation phase of the compilation is divided in four different steps and transform $\Gamma$ to a series-parallel tree decomposition~\ref{}.
+This transformation phase handles data parallelism through the first step, and task parallelism through the three next steps.
+In this section are described those four steps of MSC.
+The detailed formalism and associated algorithms of those steps are described in the research report~\cite{}.
 
-\begin{enumerate}
-\item creation of $\Gamma$, the ordered list of computations, from the MSL input file;
-\item creation of $\Gamma_{data}$ the ordered list of computations including synchronizations;
-\item creation of the dependency graph $\Gamma_{hybrid}$;
-\item transformation of $\Gamma_{hybrid}$ to a minimal series-parallel graph $\Gamma_{msp}$;
-\item creation of the series-parallel tree decomposition $\Gamma_{tsp}$;
-\item dump $\Gamma_{tsp}$ to a component assembly.
-\end{enumerate}
+The creation of $\Gamma$ is directly given from the parser.
+Actually, the list of computations in the MSL program is already ordered, as a result $\Gamma$ is a direct map of this list.
+In the example of Figure~\ref{fig:mslex}, $\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$.
 
-\paragraph{Creation of the ordered list of computations} This step of the MSC compiler parse the MSL input file which describes the simulation.
-From the ordered lines of computations, it builds the ordered list of computations $\Gamma$ to perform in the simulation.
-For example the ordered list of computation of the example above is $[c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$.
-
-\paragraph{Creation of the synchronized ordered list of computations} COnsidering a data parallelization of the simulation, from the ordered list of computations $\Gamma$ can be detect needed synchronizations.
-A synchronization is needed each time a data read by a stencil computation has been written by a previous computation and if the computation domains intersect.
-Actually, as a stencil computation needs neighborhood values, values computed by different processors or cores are needed.
-
-For example, the stencil computation $c_1$ read the data $b$ which has been written by the computation $c_0$.
+\paragraph{Creation of the synchronized ordered list of computations} Considering a data parallelization of the simulation, the mesh and data of the simulation are splitted among available processors (or cores) and a single program is applied on each sub-part with additionnal synchronizations between processors or cores.
+For example, as a stencil code access neighborhood value, values computed by another processor are needed and synchronized.
+From the ordered list of computations $\Gamma$ can be automatically detected the needed synchronizations.
+A synchronization is needed each time a data read by a stencil computation has been written by a previous computation.
+In such a case, a computation is added before the stencil computation.
+This \emph{synchronization computation} read the data to synchronize, and write the same data.
+The computation domain of such a synchronization computation is the mesh entity on which the data is declared.
+As a result $\Gamma$ is transformed to a synchronized ordered list of computations $\Gamma_{data}$.
+For example, in Figure~\ref{fig:mslex}, the stencil computation $c_1$ read the data $b$ which has been written by the computation $c_0$.
 For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed to the sublist $[c_0,sync_1,c_1]$ in $\Gamma_{data}$.
-The new computation $sync_1$ read and write $b$.
-As a result a dependency is kept between $sync_1$, and $c_1$.
+The new computation $sync_1$ read and write $b$ and is applied on the computation domain of $c_1$.
+As a result, a dependency is kept between $c_0$ and $sync_1$ and between $sync_1$, and $c_1$.
 The same is performed for the stencils $c_4$ and $c_8$.
+Thus $\Gamma_{data} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
 
@@
 In other words, the only case where a data dependency exists but is not taking into account is when the computation domains do not intersect.
-Nodes of the dependency graph represent computations and synchronizations, while edges are dependencies between them.
-$\Gamma_{hybrid}$ is a directed acyclic graph (\emph{dag}) as it is build from an ordered list.
-For example, the dependency dag $\Gamma_{hybrid}$ of the example is illustrated in Figure~\ref{fig:hyb}.
+Nodes of the dependency graph represent computations of $\Gamma_{data}$, while edges are dependencies between them.
+The dependency graph $\Gamma_{hybrid}$ is a directed acyclic graph (\emph{dag}).
+For example, the dependency dag $\Gamma_{hybrid}$ of the example of Figure~\ref{fig:mslex} is illustrated in Figure~\ref{fig:hyb}.
 
@@
 \end{tikzpicture}
-\caption{$\Gamma_{hybrid}$ of Figure~\ref{fig:msinput}}
+\caption{$\Gamma_{hybrid}$ of Figure~\ref{fig:hyb}}
 \label{fig:hyb}
@@
 
-\paragraph{Transformation to a minimal series-parallel graph} As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if the forbidden \emph{N-shape} is not found in the graph.
-To transform $\Gamma_{hybrid}$ to the minimal series-parallel graph $\Gamma_{msp}$, a breadth first search traversing algorithm is applied to build a complete bipartite graph between two levels of $\Gamma_{hybrid}$~\cite{Mitchell:2004:CMV:1082101.1082117}.
+\paragraph{Transformation to a minimal series-parallel graph} Once a dependency graph is built many solutions can be used to build a parallel application, as for example dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}.
+In this work a static scheduling of the dependency graph is built.
+To do so the dependency graph is transformed to a minimal series-parallel graph~\cite{Valdes:1979:RSP:800135.804393}.
+As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape}, is not found in the graph.
+To transform $\Gamma_{hybrid}$ to the minimal series-parallel graph $\Gamma_{msp}$, a breadth first search algorithm is applied to build a complete bipartite graph between two levels of $\Gamma_{hybrid}$~\cite{Mitchell:2004:CMV:1082101.1082117}.
 
-\paragraph{Creation of a tree decomposition} As shown in different works~\cite{Valdes:1979:RSP:800135.804393,Schoenmakers95anew}, from a minimal series-parallel graph a tree decomposition can be built.
-A series-parallel tree decomposition consists in the decomposition of the graph as a set of \emph{sequences} and \emph{parallel} sections in a tree.
-For example, the series-parallel tree decomposition $\Gamma_{tsp}$ of the example is illustrated in Figure~\ref{fig:canon}.
+\paragraph{Creation of a tree decomposition} As shown in different works~\cite{Valdes:1979:RSP:800135.804393,Schoenmakers95anew}, from a minimal series-parallel graph a series-parallel tree decomposition can be built.
+A series-parallel tree decomposition consists in the decomposition of the minimal series-parallel graph as a tree representing a set of \emph{sequences}, indicated by a $S$ node, and \emph{parallel} sections, indicated as a $P$ node.
+For example, the series-parallel tree decomposition $\Gamma_{tsp}$ of Figure~\ref{fig:hyb} is illustrated in Figure~\ref{fig:tsp}.
 
@@
 \caption{$\Gamma_{tsp}$ of Figure~\ref{fig:hyb}.}
-\label{fig:canon}
+\label{fig:tsp}
 \end{center}
@@
 
-\paragraph{Dump to a component assembly} Finally, from the series-parallel tree decomposition $\Gamma_{tsp}$ we could build a program using any parallel language containing \emph{parallel sections} and \emph{sequences of instructions} such as OpenMP~\cite{660313}, HPF~\cite{219857}, UPC~\cite{El-Ghazawi:2006:UUP:1188455.1188483} etc.
+% \paragraph{Dump to a component assembly} Finally, from the series-parallel tree decomposition $\Gamma_{tsp}$ we could build a program using any parallel language containing \emph{parallel sections} and \emph{sequences of instructions} such as OpenMP~\cite{660313}, HPF~\cite{219857}, UPC~\cite{El-Ghazawi:2006:UUP:1188455.1188483} etc.
 In this paper we study the dump of $\Gamma_{tsp}$ to a component assembly which needs the introduction of \emph{control components}, described in the next section.
-
-For more details on MSC, a full description of the compilation is described in the research report~\cite{}.

### 2015-07-30 WOLFHPC15/article/ms_msc.tex
section 4 fix helene 30_07

@@
 A dependency exists between two computations (including synchronizations) if and only if a data read has been written by a previous computation in $\Gamma_{data}$, and if the computation domains intersect.
-In other words, the only case where a data dependency exists but is not taking into account is when the computation domains do not intersect.
+In other words, the only case where a data dependency exists but is not taken into account is when the computation domains do not intersect.
 Nodes of the dependency graph represent computations of $\Gamma_{data}$, while edges are dependencies between them.
@@
 \end{tikzpicture}
-\caption{$\Gamma_{hybrid}$ of Figure~\ref{fig:hyb}}
+\caption{$\Gamma_{hybrid}$ of the example of Figure~\ref{fig:mslex}}
 \label{fig:hyb}
@@
   \end{tikzpicture}
-\caption{$\Gamma_{tsp}$ of Figure~\ref{fig:hyb}.}
+\caption{$\Gamma_{tsp}$ of Figure~\ref{fig:hyb}. Nodes $S$ represent sequences, nodes $P$ represent parallel sections.}
 \label{fig:tsp}

### 2015-07-30 WOLFHPC15/article/concept.tex
Relecture...

@@
 
+\fix{cp2hc: transition a revoir.}
+
 Let introduce some concepts used in this paper.
@@
 Second, \emph{mesh domains} are subset elements of the mesh, such as the center of cells, edges or nodes.
-In Figure~\ref{fig:mesh} two mesh domains are illustrated, the center of the cells (in red) and edges on the horizontal axe (in blue on left and right of each cell).
+In Figure~\ref{fig:mesh}, two mesh domains are illustrated, the center of the cells (named \texttt{Cells}, in red) and edges on the horizontal axe (named \texttt{Edgex}, in blue on left and right of each cell).
 A \emph{data} is a quantity to simulate or a temporary data to perform computations.
-A data is mapped onto a given mesh domain.
-For example, in Figure~\ref{fig:ex1}, data $A$ and $B$ are mapped onto the center of cells, while in Figure~\ref{fig:ex2} $C$ is mapped onto the edges.
+A data is mapped onto a mesh domain.
+For example, in Figure~\ref{fig:ex1}, data $A$ and $B$ are mapped onto the center of cells, while, in Figure~\ref{fig:ex2}, $C$ is mapped onto the edges.
 
@@
 
-The first stencil computation in Figure~\ref{fig:ex1} computes $A$ using $B$, while the second one in Figure~\ref{fig:ex2} computes $A$ using $C$.
+The stencil computation in Figure~\ref{fig:ex1} computes $A$ using $B$, while the one in Figure~\ref{fig:ex2} computes $A$ using $C$.
 Thus, a stencil computation is defined by a set of input data (in those examples only one input data), and a single output data, the result.
@@
 Actually, in this computation, the neighborhood contains edges on the left and on the right of a cell.
-As an example, the numerical expression of the first example could be
+As an example, the numerical expression of the first example could be:
 \begin{equation*} 
@@
 \end{equation*}
-and the numerical expression of the second example could be
+and the numerical expression of the second example could be:
 \begin{equation*} 
@@
 %----------------------------------------
-\subsection{Multi-stencil programs}
+\subsection{A Multi-Stencil Formlization}
 \label{sect:multistencil}
@@
 A real case numerical simulation is most of the time composed of more than one stencil computation, with one or more stencil shapes, and onto one or more data, and of additionnal local computations.
-Let introduce formal definitions used in the rest of the paper.
+\fix{cp2hc: [optionnelle] Par example? domaine applicatif?}
+Let introduce some formal definitions used in the rest of the paper.
 
@@
 \end{equation}
-where $R$ is a set of pair $(r,n)$ with $r \in \Delta$ is a data read by the computation, and $n$ is a stencil shape (a neighborhood) applied on $r$.
+where $R$ is a set of pair $(r,n)$, with $r \in \Delta$ is a data read by the computation, and $n$ is a stencil shape (a neighborhood) applied on $r$.
 The data written by the computation is denoted $w \in \Delta$, $exp$ is the numerical expression of the stencil computation.
@@
 
-For example, in the example of Figure~\ref{fig:ex1}, if we call the computation domain (plain lines) $dc1$ and the stencil shape $n1$, the stencil computation can be defined with
+For example, in the example of Figure~\ref{fig:ex1}, assuming the computation domain (plain lines) is $dc1$ and the stencil shape is $n1$, the stencil computation can be defined with:
 \begin{equation*}
@@
 \end{equation*}
-On the other hand, in the second example of Figure~\ref{fig:ex2}, if we call the computation domain $dc2$ and the stencil shape $n2$, the stencil computation can be defined as
+On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil computation can be defined with:
 \begin{equation*}
@@
 \end{equation*}
-One can notice that the input data $A$ is associated to the stencil shape $null$. This means that no stencil shape is applied on $A$, but only a local access.
+One can notice that the input data $A$ is associated to the stencil shape $null$. This means that no stencil shape is applied on $A$, but only local accesses.
 
 A local, or auxiliary, computation is almost the same than a stencil computation without neighborhood information in the numerical expression, which makes this kind of computation purely local.
-A local computation $l$ is defined as the quadruplet
+A local computation $l$ is defined as the quadruplet:
 \begin{equation} 
@@
 
-A multi-stencil program in this work is defined as a sextuplet
+A multi-stencil program in this work is defined as a sextuplet:
 \begin{equation} 

### 2015-07-30 WOLFHPC15/article/related.tex

--- 2015-07-21 WOLFHPC15/article/related.tex
@@
-%----------------------------------------
-\subsection{Stencil solutions}
-%----------------------------------------
+% %----------------------------------------
+% \subsection{Stencil solutions}
+% %----------------------------------------
 
-Many solutions exists to ease parallel programming of stencil codes, as for example PATUS~\cite{citeulike12258902}, Pochoir~\cite{spaaTangCKLL11}, OP2~\cite{Giles2011}.
-Those solutions are powerfull, let the user implement their own stencil codes, and produce high performance codes.
+Many specific solutions exist to ease parallel programming of numerical simulations.
+Solutions such as PETSc~\cite{petsc-efficient} and Trilinos~\cite{Trilinos-Overview} are specific libraries.
+They are made to solve many kinds of PDEs using well-known methods and solvers which are efficiently parallelized inside the library.
+Those libraries also include solvers for implicit numerical schemes, not adressed in this work.
+However, using those solutions the user has to use and follow existing tools.
+In other words, it is not possible for a mathematician or a physician to write his own numerical method if it differs from existing ones.
+Moreover as PETSc and Trilinos are libraries they do not propose a domain specific language to define computations.
+The purpose of those solutions are much more specific than our solution.
+
+Other solutions try to take a step back in the abstraction to propose a way to hide parallel codes for stencils while keeping a sequential programming style.
+For example PATUS~\cite{citeulike12258902} and Pochoir~\cite{spaaTangCKLL11} are proposed to write stencil codes for general structured meshes, and OP2~\cite{Giles2011} and Liszt~\cite{DeVito2011LDS} for general unstructured meshes.
 Those solutions can be considered as stencil compilers.
-As a result it produces optimized (cache tiling etc.) or parallel (CUDA, OpenCL etc.) codes for a single stencil computations.
+Actually they produce optimized (cache tiling etc.) or parallel (CUDA, OpenCL etc.) codes for a single stencil computation.
+Recently, domain specific languages for very specific numerical methods such as ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} for multigrid solvers, and OPS~\cite{Reguly:2014:ODS:2691166.2691173} for meshes of structured meshes (hybrid mesh) have also been proposed.
+All those solutions are powerfull, let the user implement their own stencil codes in a sequential programming style, and produce high performance codes for shared memory architectures and sometimes distributed memory architectures.
 
-However, a real case numerical simulation is not composed of a single stencil code as explained in Section~\ref{sect:multistencil}.
-A multi-stencil simulation computes more than one stencil computation, involving one or more stencil shapes, and additionnal local coputations.
+Most stencil compilers handle the parallelization or the optimization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations.
+However, many real case numerical simulation is not composed of a single stencil code, as explained in Section~\ref{sect:multistencil}, but of more than one stencil computation, involving one or more stencil shapes, and additionnal local coputations.
 Solutions like Pochoir or PATUS handle the parallelization or the optimization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations.
-However, not taking into account the parallelization of the overall numerical simulation implies that those compilers are reduced to shared memory systems.
-Actually using a distributed memory system, the MPI (Message Passing Interface)~\cite{Graham2009MSE} library still have to be used by the user, as well as the distribution of the mesh onto the different distant processors.
+Liszt~\cite{DeVito2011LDS} is the closest solution to our work as it seems to take into account the entire program and not only stencil kernels.
 
-On the other hand, the work presented in this paper does not propose to optimize and parallelize a single stencil code for shared memory machines, GPGPUs or many-core architectures.
-The MS language and the MSC compiler produce a coarse-grain parallel structure of the overall simulation which could be combined to existing stencil compilers.
+The work presented in this paper takes place at a different level of numerical simulations and can be seen as complementary to stencil compilers.
+Actually, the MSL language is agnostic and should work for any type of mesh and any neighborhood shape in stencil codes.
+The MSC part of the compiler only needs this agnostic description of the overall simulation to generate a parallel pattern of the program, which needs to be filled with computation codes.
+Thus, depending on the underlying distributed data structure used, those computation codes could be generated by stencil compilers.
+As a result, MSL and its compiler are not a new contribution to compile optimized stencil codes, but propose another level of parallelization and a different approach of back-end solution.
 
-%----------------------------------------
-\subsection{Component models and control}
-%----------------------------------------
+% %----------------------------------------
+% \subsection{Language, control and component models}
+% %----------------------------------------
 
-%----------------------------------------
-\subsection{Languages and component models}
-%----------------------------------------
+% \textbf{Put it somewhere else !}
+
+% As far as we know, no existing DSL has been transformed to a HPC component assembly until know.
+This paper is a contribution to study the interest of using component models to increase code re-use, separation of concerns and maintainability in DSLs.
+Actually, as the code is not generated inside the DSL itself but is built from the assembly of existing components, many advantages can get out of this work.
+First, the code-reuse from one DSL to another is increased if the components are well studied and defined.
+For example, it seems that control components introduced in this paper can be used for many different DSLs in many different domains.
+Second, separation of concerns in the final code is also improved as the generated component assembly split the different functionnalities of the application in different components.
+As a result, some components can be written by computer-scientists, while other are specific to the application, as for example computation components.
+Finally, the maintainability of the application is also increased, as each functionnality is separated.
+In the final code if a computation component has to be modified by a numerician, intricates part of the code will be unchanged and not seen by the numerician.
+
+% If the study of a DSL dumped to a component assembly has not been studied until now, some component models have already proposed the introduction of time control and data control in the assembly, as for example STCM~\cite{DBLP:confeuropar2008} and XMan~\cite{He12component-baseddesign}.
+However, both those solutions does not seem to target high performance computing and distributed components.

### 2015-07-30 WOLFHPC15/article/mscac.tex
Relecture section 3

@@
-The solution presented in this paper contains a domain specific language to describe a multi-stencil program, i.e.
+The solution proposed in this paper contains a domain specific language to describe a multi-stencil program, i.e.
 an overall numerical simulation, and a compiler which successively parses the front-end language, transforms the internal representation, and finally dumps the representation to a back-end code.
+This back-end code is component based.
 
-The proposed DSL, called MSL, is an agnostic descriptive language for multi-stencil simulations.
-Agnostic means that the description of a numerical simulation does not depend on implementation constraints as, for example, the type of mesh (structured, unstructured) and its associated interfaces (how to define a stencil shape etc.).
-Descriptive, on the other hand, means that the terminal states of the language grammar are string identifiers.
+The proposed DSL, named MSL (\emph{Multi-Stencil Language}), is an agnostic descriptive language for multi-stencil simulations.
+Agnostic means that the description of a numerical simulation does not depend on implementation choices as, for example, the type of mesh (structured, unstructured), or its associated interfaces (how to define a stencil shape, etc.).
+Descriptive, on the other hand, means that the terminal states of the language grammar are (string) identifiers.
 As a result, MSL is not made to handle the expression of numerical computations, but only the description of a multi-stencil simulation, in a coarse-grain fashion.
-In a MSL description, six main sections are defined, as the six tuples of the definition of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\mathcal{D}_m,\mathcal{D}_c,\Delta,\Gamma)$.
+Numerical computations are encapsulated in specific components that users have to develop.
+In a MSL description, six main sections are defined, as the six tuples of the definition of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\mathcal{D}_m,\mathcal{D}_c,\Delta,\Gamma)$ (\cf Equation~\ref{eq:msp}).
 
-As illustrated in Figure~\ref{fig:mscac}, the compiler, called MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of three different phases.
-The first phase of the compiler is to \emph{parse} the MSL description.
-Second, the compiler analyzes and transforms the computation part of the description $\Gamma$, to an intermediate representation.
-This phase of compilation, called MSC, is responsible for the automatic parallelization of the overall numerical simulation.
-Finally, the parallel intermediate representation is dumped to an actual implementation of the parallel pattern of the simulation.
-
-% As illustrated in Figure~\ref{fig:mscac}, the MS language compiler, called MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of two different parts.
-First, the computation compiler, called \emph{MSC}, is responsible for the parallelization of computations and for its dump to a component assembly; second, the data compiler, called \emph{MSD}, handles a distributed data structure, data mapped onto it, and their relations with computations, as a final overall component assembly.
-\begin{figure}[h!]
+\begin{figure}[t]
 \begin{center}
@@
 \end{tikzpicture}
-\caption{MSCAC compiler}
+\caption{The three phases of the MSCAC compiler.}
 \label{fig:mscac}
@@
 
+As illustrated in Figure~\ref{fig:mscac}, the compiler, named MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of three different phases.
+The first phase of the compiler is to \emph{parse} the MSL description.
+Second, the compiler analyzes and transforms the computation part of the description $\Gamma$, to an intermediate representation.
+This phase of compilation, called MSC, is responsible for the automatic parallelization of the overall numerical simulation.
+Finally, the parallel intermediate representation is dumped to an actual implementation of the parallel pattern of the simulation.
+
+% As illustrated in Figure~\ref{fig:mscac}, the MS language compiler, called MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of two different parts.
+First, the computation compiler, called \emph{MSC}, is responsible for the parallelization of computations and for its dump to a component assembly; second, the data compiler, called \emph{MSD}, handles a distributed data structure, data mapped onto it, and their relations with computations, as a final overall component assembly.
+
 %In this paper are presented the language MSL, the MSC compilation phase, and the dump compilation phase. 
-Parallelization techniques handled by MSC are \emph{data parallelism}, where data are splitted among processors or cores while applying a single program everywhere, and \emph{task parallelism} where the program is transformed into a dependency graph of tasks.
-From a dependency graph different strategies can be used to schedule tasks, in other words to control their execution in a valid order.
+The parallelization techniques handled by MSC are \emph{data parallelism}, where data are splitted among processors or cores while applying a single program everywhere, and \emph{task parallelism} where the program is transformed into a dependency graph of tasks.
+From a dependency graph, different strategies can be used to schedule tasks, in other words to control their execution in a valid order.
 The MSC compiler presented in this paper handles the transformation of the dependency graph to a static scheduling.
@@
 We could have propose a dump to OpenMP~\cite{660313}, however this paper presents a first contribution to dump a DSL to a component-based runtime.
-Component models, where a component represents a clear and independent functionnality of the application, have first been proposed in Software engineering and have proved many times good properties for code re-use, separation of concerns, maintainability and productivity of codes.
+Component models, where a component represents a clear and independent functionnality of the application, have first been proposed in software engineering and have proved many times good properties for code re-use, separation of concerns, maintainability and productivity of codes.
 Those properties, on the other hand, are recurrent stated problems of parallel applications.
 
-The back-end of the compiler is to build a parallel assembly of components, in other words the appropriate connections between functionnalities of the numerical simulation.
-This assembly is illustrated in Figure~\ref{fig:assembly} which is explained in details in Section~\ref{sect:component}.
+The back-end of the compiler builds a component assembly, that can be divided in two parts: the components implementing the actual computation and the static schedule, and the components needed needed to manage it, \ie the runtime support.
+A general view of this assembly is proposed in Figure~\ref{fig:assembly}; it is explained in details in Section~\ref{sect:component}.
 As illustrated in the figure, each component is linked to one or more tuple of the multi-stencil program definition: \emph{DDS} component represents $\mathcal{M}$ and $\mathcal{D}_m$; \emph{Data} component represents $\Delta$ and $\mathcal{D}_c$; \emph{Time} component represents $T$, and \emph{Computations} component (which is generated by MSC) represents $\Gamma$.
@@
 For example, the back-end evaluated in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}.
-A work in progress replaces this DDS component by another one using Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} data structure.
+A work in progress replaces this DDS component by another one making using of Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.
 
-\begin{figure}
+\begin{figure}[t]
 \begin{center}

### 2015-07-30 WOLFHPC15/article/conclusion.tex
conclusion items helene 30_07

Conclusion :
\begin{itemize}
\item Multi-stencil formalism
\item MSL and MSC
\item Discussion dump to components
\item Evaluation
\end{itemize}

Perspectives :
\begin{itemize}
\item Dump to OpenMP for comparison
\item MSC to dynamic scheduler (starPU, XKaapi)
\item Formalization and abstraction of the data management in component-based applications
\item Parallel component language
\end{itemize}

### 2015-07-30 WOLFHPC15/article/ms_msc.tex
relecture section 4

@@
 %----------------------------------------
-As already mentionned, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations.
-Six main sections are required in a MSL description, as the six tuples of the formal definition of a multi-stencil program: the mesh description; the mesh entities description; the computation domains description and their dependencies; the data description; the time loop description; and finally, the computations description.
-Figure~\ref{fig:grammar} shows the grammar of MSL.
-Lines 1 to 9 of this figure define what is a MSL program with the different sections described above.
-In this section are described the different sections of a MSL program and an example.
-As MSL is a descriptive language, almost all terminals $meshid$, $meshdom$, $compdomid$, $dataid$, $compid$ ans $neighborid$ are string identifiers. Only $num$ is an integer terminal.
 
-\begin{figure}[h!]
+\begin{figure}[t]
 \begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++,numbers=left]
@@
 
-\paragraph{Mesh and mesh entities} As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is directly defined by a string identifier.
+As mentionned in the previous section, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations.
+Six main sections are required in a MSL description, as the six tuples of the formal definition of a multi-stencil program: the mesh description; the mesh entities description; the computation domains description and their dependencies; the data description; the time loop description; and finally, the computations description.
+Figure~\ref{fig:grammar} shows the grammar of MSL.
+Lines 1 to 9 define what is a MSL program with the different parts memntionned above.
+The remaining of this section describes the parts of a MSL program and presents an example.
+We can note that, as MSL is a descriptive language, almost all terminals $meshid$, $meshdom$, $compdomid$, $dataid$, $compid$ ans $neighborid$ are string identifiers.
+Only $num$ is an integer terminal.
+
+\subsubsection*{Mesh and mesh entities}
+As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is directly defined by a string identifier.
 Mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
@@
 
-\paragraph{Computation domains and their dependencies} A computation domain can be seen as a subpart of a mesh entity which will be used during one or more computations.
-For each computation domain is indicated the mesh entity from which it is built (lines 3-4 and 12-14 in Figure~\ref{fig:grammar}).
-By default two computation domains always intersect except when an independency relation is precised in the section \emph{independent} of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
-In the example of Figure~\ref{fig:mslex}, on lines 4 to 6 an entire computation domain is defined for each mesh entity, called $allcell$, $alledgex$ and $alledgey$, and on lines 7 and 8 two subparts of the mesh entity edgex are also defined, called $part1edgex$ and $part2edgex$.
-It is then precised on line 10 that $part1edgex$ and $part2edgex$ do not intersect.
+\subsubsection*{Computation domains and their dependencies}
+A computation domain can be seen as a subpart of a mesh entity which will be used during one or more computations.
+For each computation domain, the mesh entity from which it is built is indicated (lines 3-4 and 12-14 in Figure~\ref{fig:grammar}).
+By default two computation domains always intersect except when an independency relation is mentionned in the section \emph{independent} of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
+In the example of Figure~\ref{fig:mslex}, on lines 4 to 6, an whole computation domain is defined for each mesh entity, called $allcell$, $alledgex$ and $alledgey$, and on lines 7 and 8 two subparts of the mesh entity $edgex$ are also defined, called $part1edgex$ and $part2edgex$.
+Line 10 means that $part1edgex$ and $part2edgex$ do not intersect.
 
-\paragraph{Data and time} A data is a quantity to simulate and is mapped onto a mesh entity, as illustrated in lines 7 and 18-20 f Figure~\ref{fig:grammar}.
-The time section simply indicates a number of iterations to perform in the simulation.
+\subsubsection*{Data and time}
+A data is a quantity to simulate and is mapped onto a mesh entity, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}.
+The $time$ section simplies indicates a number of iterations to perform in the simulation.
 In the example of Figure~\ref{fig:mslex} are defined ten data and 500 iterations (lines 11-22).
+An extension could be to support a convergence criterion.
 
-\paragraph{Computations description} Finally the language follows the two definitions of a stencil computation and a local computation given in Section~\ref{sect:multistencil}.
+\subsubsection*{Computations description}
+Finally, the language follows the two definitions of a stencil computation and a local computation given in Section~\ref{sect:multistencil}.
 However as MSL does not handle numerical expressions, the term $exp$ of Definition~\ref{eq:st} and Definition~\ref{eq:loc} is not used in the language.
-The exact code of the computation as to be writtenin the final ready-to-fill parallel pattern of the simulation, using external languages or interfaces.
+The exact code of the computation as to be writtenin the final ready-to-fill parallel pattern of the simulation, using external languages or interfaces (\cf Section~\ref{sect:component}).
 Moreover, to propose an intuitive syntax, information are given in a different order than in formal definitions.
-Thus, the description of a computation, as indicated in lines 24 and 25 of Figure~\ref{fig:grammar}), consists in first the identifier of the data to compute ($w$ in the formal description), brackets indicated the computation domain on which the computation is applied ($d_c$ in the formal definition), an identifier for the computation, and the set of identifiers for the data read by the computation.
+Thus, the description of a computation, as indicated in lines 24 and 25 of Figure~\ref{fig:grammar}, consists in first the identifier of the data to compute ($w$ in the formal description), in brackets the identifier of the computation domain on which the computation is applied ($d_c$ in the formal definition), an identifier for the computation, and the set of identifiers for the data read by the computation with parenthesis.
 For each data read by the computation brackets indicate the stencil shape identifier used on the data ($R$ in the formal definition of a stencil).
@@
 For example, on line 24, the data $b$ is computed on the computation domain $allcell$ by the computation $c_0$ which read data $a$ without neighborhood shape.
-\begin{figure}
+
+\begin{figure}[t]
 \begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++,numbers=left]
@@
 %----------------------------------------
-The MSC transformation, which is a subpart of the overall compiler MSCAC, use the ordered list of computations $\Gamma$, which can directly been extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
-This transformation phase of the compilation is divided in four different steps and transform $\Gamma$ to a series-parallel tree decomposition~\ref{}.
+The MSC transformation, which is a subpart of the overall compiler MSCAC, makes use of the ordered list of computations $\Gamma$, which can directly been extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
+This transformation phase of the compilation is divided in four different steps; it transforms $\Gamma$ to a series-parallel tree decomposition~\ref{???}.
 This transformation phase handles data parallelism through the first step, and task parallelism through the three next steps.
-In this section are described those four steps of MSC.
-The detailed formalism and associated algorithms of those steps are described in the research report~\cite{}.
+This section gives an overview of these four steps of MSC; the detailed formalism and associated algorithms of those steps are described in the research report~\cite{????}.
 
-The creation of $\Gamma$ is directly given from the parser.
-Actually, the list of computations in the MSL program is already ordered, as a result $\Gamma$ is a direct map of this list.
+\subsubsection*{The Computation Ordered List: $\Gamma$}
+$\Gamma$ is directly obtained from the parser.
+Actually, the list of computations in the MSL program is already ordered: $\Gamma$ is a direct map of this list.
 In the example of Figure~\ref{fig:mslex}, $\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$.
 
-\paragraph{Creation of the synchronized ordered list of computations} Considering a data parallelization of the simulation, the mesh and data of the simulation are splitted among available processors (or cores) and a single program is applied on each sub-part with additionnal synchronizations between processors or cores.
+\subsubsection*{The Computation Synchronized Ordered List: $\Gamma_{data}$}
+Considering a data parallelization of the simulation, the mesh and data of the simulation are splitted among available processors (or cores) and a single program is applied on each sub-part with additionnal synchronizations between processors or cores.
 For example, as a stencil code access neighborhood value, values computed by another processor are needed and synchronized.
-From the ordered list of computations $\Gamma$ can be automatically detected the needed synchronizations.
+From the ordered list of computations $\Gamma$, the needed synchronizations can be automatically detected.
 A synchronization is needed each time a data read by a stencil computation has been written by a previous computation.
@@
 As a result, a dependency is kept between $c_0$ and $sync_1$ and between $sync_1$, and $c_1$.
-The same is performed for the stencils $c_4$ and $c_8$.
-Thus $\Gamma_{data} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
+The same transformation is performed for the stencils $c_4$ and $c_8$.
+Thus, $\Gamma_{data} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
 
-\paragraph{Creation of the dependency graph} From the synchronized ordered list of computations $\Gamma_{data}$, a dependency graph is built.
+\subsubsection*{The Dependency Graph: $\Gamma_{hybrid}$}
+From the synchronized ordered list of computations $\Gamma_{data}$, a dependency graph $\Gamma_{hybrid}$ is built.
 A dependency exists between two computations (including synchronizations) if and only if a data read has been written by a previous computation in $\Gamma_{data}$, and if the computation domains intersect.
@@
 The dependency graph $\Gamma_{hybrid}$ is a directed acyclic graph (\emph{dag}).
-For example, the dependency dag $\Gamma_{hybrid}$ of the example of Figure~\ref{fig:mslex} is illustrated in Figure~\ref{fig:hyb}.
+For example, the dependency dag $\Gamma_{hybrid}$ of the example of Figure~\ref{fig:mslex} is displayed in Figure~\ref{fig:hyb}.
 
-\begin{figure}[h!]
+\begin{figure}[t]
 \begin{center}
@@
 
-\paragraph{Transformation to a minimal series-parallel graph} Once a dependency graph is built many solutions can be used to build a parallel application, as for example dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}.
+\subsubsection*{The Minimal Series-Parallel Graph: $\Gamma_{msp}$}
+Once a dependency graph is built, many solutions can be used to build a parallel application, as for example dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}.
 In this work a static scheduling of the dependency graph is built.
@@
 
-\paragraph{Creation of a tree decomposition} As shown in different works~\cite{Valdes:1979:RSP:800135.804393,Schoenmakers95anew}, from a minimal series-parallel graph a series-parallel tree decomposition can be built.
+\subsubsection*{The Tree Decomposition: $\Gamma_{tsp}$}
+Many works~\cite{Valdes:1979:RSP:800135.804393,Schoenmakers95anew} explains how to build a series-parallel tree decomposition from a minimal series-parallel graph.
 A series-parallel tree decomposition consists in the decomposition of the minimal series-parallel graph as a tree representing a set of \emph{sequences}, indicated by a $S$ node, and \emph{parallel} sections, indicated as a $P$ node.
@@
 
-\begin{figure}[h!]
+\begin{figure}[t]
 \begin{center}

### 2015-07-30 WOLFHPC15/article/introduction.tex
add a fix?
intro

@@
-Programming complex scientific applications on high performance heterogenous architectures, such as clusters of nodes containing multicores and accelerators, requires for the scientists to be an expert in his own domain, but also to be an expert in low level parallel and high performance programming.
-It can takes months to years to a scientist to learn HPC programming, as it is not his main activity.
+Programming complex scientific applications on high performance heterogeneous architectures, such as clusters of multicores nodes containing accelerators, requires for the scientists to be an expert in his own domain, but also in low level parallel and high performance programming.
+It can take months to years for a scientist to learn HPC programming, as it is not his main activity.
 And while learning HPC programming, hardware continue to evolve, becoming more powerful but usually also more  complex and difficult to use efficiently.
 
-Domain Specific Languages (DSLs) are a promising solution to hide intricate details of HPC programming from non experts, while producing automatically an optimized and efficient final code.
-Thus, it seems that a clue for simplicity and efficiency on modern architectures is to specialize a language to a given domain, from which one or more efficient parallelizations and optimizations can be extracted.
-However, if the specific nature of a DSL simplifies HPC programming for scientists, its implementation cost (language, compiler, and runtime) is difficult to amortize.
-Actually, programming difficulties, as well as non-maintainability and non-portability seems to be deported from the scientist to DSL designers.
-Moreover, one can consider that end users need as much DSLs as domains, which keeps the amount of work to write and maintain up-to-date DSLs out of hands.
+Domain Specific Languages (DSLs) are a promising solution to hide intricate details of HPC programming from non experts.
+By being specific to a well-defined set of problems, these languages enable the generation of well parallelized and optimized code for various architectures while requiring only a limited amount of information from the developer; part of the knowledge is directly embedded in the compiler.
+A critical question however concerns the specificity of DSLs.
+The less specific the DSL is, the more informations its users must provide and the less automatic optimizations can be embedded in the compiler.
+On the other side of the scope, the more specific the DSL is, the less different code can use it and the more difficult it is to amortize the implementation cost of the DSL (language, compiler, and runtime).
+Programming difficulties, as well as non-maintainability and non-portability often seem to be simply deported from the scientist to DSL designers.
+% Moreover, one can consider that end users need as much DSLs as domains, which keeps the amount of work to write and maintain up-to-date DSLs out of hands.
 
-For this reason, solutions to ease the design and implementation of DSLs have recently appear~\cite{Fernandez:2014:DFL:2691166.2691168}.
+For this reason, solutions to ease the design and implementation of DSLs have recently appeared~\cite{Fernandez:2014:DFL:2691166.2691168}.
 The main idea of those solutions is to propose a single DSL to write new DSLs.
-In this paper, we study a first contribution to another approach to bring maintainability and code-reuse in DSLs.
+In this paper, we study a first contribution to another approach to improve maintainability and code-reuse in DSLs.
 This approach proposes a transformation from a DSL to a component-based runtime.
-TO DEV BENEFITS OF COMPONENTS.
+To evaluate this approach, a proper new DSL for \emph{multi-stencil} programs, and its compiler are presented in this paper.
+\fix{cp2cp+hc: to dev benefits of components.}
 
-To evaluate such an approach, a proper new DSL for \emph{multi-stencil} programs, and its compiler are presented in this paper.
-Many solutions to automatically optimize and parallelize partial differential equations (PDEs) solvers have been proposed in the litterature, as specific libraries~\cite{petsc-efficient,Trilinos-Overview,CPE:CPE3494}, and DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
-Most of the time, DSLs focuss on the parallelization and the optimization of a single numerical computation, also called a stencil.
-However, a real case numerical simulation is most of the time not composed of a single stencil computation, but of a set of stencil computations and a set of additionnal auxiliary local computations.
-The DSL proposed in this paper, called MSL, is a DSL to generate a parallel component-based structure of an overall numerical simulation, that we call a \emph{multi-stencil programs}.
-As it will be detailed in the related work, this work is complementary to the optimization and parallelization of a single stencil computation.
+Many solutions to automatically optimize and parallelize partial differential equations (PDEs) solvers have been proposed in the literature, either as specific libraries~\cite{petsc-efficient,Trilinos-Overview,CPE:CPE3494} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
+Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil.
+However, a real case numerical simulation is most of the time not composed of a single stencil computation, but of a set of stencil computations and a set of additional local auxiliary computations.
+The DSL proposed in this paper, called MSL, is a DSL to generate a parallel component-based structure of an overall numerical simulation, that we call a \emph{multi-stencil program}.
+As it will be detailed in the related work, this is complementary to the optimization and parallelization of a single stencil computation.
+\fix{JB: j'utiliserais bien le terme "kernel" ou "stencil kernel" pour ce que tu apelles "stencil" et "stencil" tout court pour ce que tu apelles "stencil shape"}
 
 The rest of this paper is organized as follows.
-First concepts of \emph{stencil} and \emph{multi-stencil porgams} are formalized in Section~\ref{sect:concept}.
+First, the concepts of \emph{stencil} and \emph{multi-stencil programs} are formalized in Section~\ref{sect:concept}.
 Section~\ref{sect:mscac} gives an overview of the solution presented in this paper while Section~\ref{sect:msmsc} details the MSL language and its compiler.
-Section~\ref{sect:component} introduces component models and then it describes the proposed component-based runtime and the automatic generation of its parallel compuration part.
+Section~\ref{sect:component} introduces component models and describes the proposed component-based runtime and the automatic generation of its parallel computation part.
 %The end of Section~\ref{sect:component} also discusses on interest of this new approach.

### 2015-07-30 WOLFHPC15/article/evaluation.tex
wolfhpc15 21_07
wolfhpc15 24_07
evaluation helene 30_07

@@
-\subsection{Shallow water equations}
-\subsection{MS description}
-\subsection{Series-parallel tree decomposition}
+Navier-Stokes equations is a well known set of PDEs in fluid dynamics to simulate a flow evolution in time.
+At the University of Orl\'eans, France, the MAPMO laboratory works on a software, called FullSWOF~\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}, which handles simulations of geophysical flows, for example rivers, channels, dambreak problems, ocean currents, estuarine systems, etc.
+This software solves the Shallow water equations obtained from the three dimensional incompressible Navier-Stokes equations, by averaging on the vertical direction (see e.g.~\cite{Ferrari2004}).
+Those equations are solved using a two-dimensional Cartesian discretization of the space domain, and a finite volume numerical methods more described in~\cite{HeleneLS13}.
+
+The MSL program of FullSWOF contains three mesh entities, seven computation domains, fourty-eight data and ninty-eight computations composed of thirty-two stencil computations and sixty-six local computations.
+
+The series-parallel tree decomposition extracted by the MSC transformation is composed of seventeen sequences and eighteen parallel sections.
+In Figure~\ref{fig:freq} represents the level of parallelism, \ie the number of tasks which can be executed concurrently, and gives the number of time the level is observed in the final tree.
+
+\begin{figure}[!h]
+ \begin{center}
+ \begin{tabular}{c|c|c|c|c|c|c|c|c|}
+   level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
+   \hline
+   frequence & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
+ \end{tabular}
+\caption{Parallelism level (number of parallel tasks) and the number of times this level appear.}
+\label{fig:freq}
+ \end{center}
+\end{figure}
+
+Figure~\ref{fig:freq} illustrates the execution time for each step of the MSC transformation for all overall execution time of ten seconds.
+Execution times have been computed on a laptop with a bi-core Intel Core i5 1,4 GHz, and 8Gb LPDDR3.
+
+\begin{figure}[!h]
+ \begin{center}
+ \begin{tabular}{c|c|c|c|c|}
+   step & $\Gamma_{sync}$ & $\Gamma_{dep}$ & $\Gamma_{msp}$ & $\Gamma_{tsp}$\\
+   \hline
+   time (ms) & 2 & 530 & 8297 & 1133\\
+   \hline
+   \% & 0.02 & 5.3 & 83.3 & 11.37\\
+ \end{tabular}
+\caption{Execution times of the MSC transformation steps}
+\label{fig:freq}
+ \end{center}
+\end{figure}
+
+One can notice that the transformation of $\Gamma_{dep}$ to a minimal series-parallel graph is the longest step of MSC, because of the removal of the forbidden shapes in the graph.
+Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted, because the algorithm use a general solution instead of finding each forbidden shape, but it seems that many of them appeared.
+The forbidden shape, called \emph{N-shape}, is represented in Figure~\ref{fig:n}.
+This shape is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.
+To remove a \emph{N-shape} an additionnal dependency constraint is added as illustrated in Figure~\ref{fig:over}.
+
+\begin{figure}[h!]
+\begin{center}
+\subfloat[][The forbidden $N$ shape.\label{fig:n}]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (a) at (0,0) {$c_0$};
+   \node[] (b) at (2,0) {$c_1$};
+   \node[] (c) at (0,-1) {$c_2$};
+   \node[] (d) at (2,-1) {$c_3$};
+
+  \path[->]
+    (a) edge node {} (b)
+    (c) edge node {} (b)
+        edge node {} (d);
+  \end{tikzpicture}
+  }
+  \hspace{25pt}
+  \subfloat[][Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
+	\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (c0) at (0,0) {$c_0$};
+   \node[] (bc0) at (-1,0) {};
+   \node[] (c1) at (2,0) {$c_1$};
+   \node[] (c2) at (0,-1) {$c_2$};
+   \node[] (bc2) at (-1,-1) {};
+   \node[] (c3) at (2,-1) {$c_3$};
+
+  \path[->]
+    (bc0) edge [dotted] node {} (c0)
+    (bc2) edge [dotted] node {} (c2)
+    (c0) edge node {} (c1)
+          edge [dashed] node [swap] {} (c3)
+    (c2) edge node {} (c1)
+        edge node {} (c3);
+  \end{tikzpicture}
+}
+  \caption{Forbidden $N$ subgraph shape for a DAG to be minimal series-parallel.}
+  \label{fig:forbidden}
+\end{center}
+\end{figure}
+
+The fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static shedule of tasks may not be the best solution for complex simulations.
+Actually, if a \emph{N-shape} cannot be represented by sequences and parallel sections, this shape can perfectly be handled by a dynamic scheduler or by the direct static representation of $\Gamma_{dep}$ where tasks wait for their input before starting.
+
+%----------------------------------------
+\subsection{A real case simulation}
+%----------------------------------------
+SkelGIS and MSD\\
+Description of shallow water equations\\
+MSL language description\\
+The obtained series-parallel graph, comments.
+
+%----------------------------------------
 \subsection{Performances}
+%----------------------------------------
 \begin{itemize}
-\item for data part - SIPSim and SkelGIS
 \item speedup Figures

### 2015-07-30 WOLFHPC15/article/concept.tex
fix 1,2,3,4 helene 30_07

@@
 In other words, in a mesh-based numerical simulation, the PDEs are transformed to a set of numerical computations applied at each time step on elements of the discretized space domain (the mesh).
-Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}~\cite{spaaTangCKLL11}.
-To ease programming of stencil computations, and because of the relative regularity of such a computation, many solutions (languages or libraries) propose an easy way to describe the stencil code, while producing transparently an optimized and parallelized stencil code.
-For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, respectively proposing an optimized cache tiling technique, and a separation of concerns between parallelization strategies and computation descriptions.
+Among those numerical computations is found a set of numerical schemes, also called \textit{stencil kernels}~\cite{spaaTangCKLL11}.
+%To ease programming of stencil kernels, and because of the relative regularity of such a computation, many solutions (languages or libraries) propose an easy way to describe the stencil code, while producing transparently an optimized and parallelized stencil kernel.
+For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil kernel on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, respectively proposing an optimized cache tiling technique, and a separation of concerns between parallelization strategies and computation descriptions.
 
-\fix{cp2hc: transition a revoir.}
+%\fix{cp2hc: transition a revoir.}
 
-Let introduce some concepts used in this paper.
-First, a stencil code is based on a \emph{mesh} which is the discretization of the physical real domain.
+In this paper are used rather unusual concepts to define a stencil, let introduce those concepts.
+First, a stencil kernel is based on a \emph{mesh} which is the discretization of the physical real domain.
 A mesh is a connected undirected graph without bridges (an edge is a bridge if its removal results in two disconnected graphs), where nodes and edges are linked to form cells (closure).
@@
 Each cell contains four nodes and four edges.
-Second, \emph{mesh domains} are subset elements of the mesh, such as the center of cells, edges or nodes.
-In Figure~\ref{fig:mesh}, two mesh domains are illustrated, the center of the cells (named \texttt{Cells}, in red) and edges on the horizontal axe (named \texttt{Edgex}, in blue on left and right of each cell).
+Second, \emph{mesh entities} are subset elements of the mesh, such as the center of cells, edges or nodes.
+In Figure~\ref{fig:mesh}, two mesh entities are illustrated, the center of the cells (named \texttt{Cells}, in red) and edges on the horizontal axe (named \texttt{Edgex}, in blue on left and right of each cell).
 A \emph{data} is a quantity to simulate or a temporary data to perform computations.
-A data is mapped onto a mesh domain.
+A data is mapped onto a given mesh entity.
 For example, in Figure~\ref{fig:ex1}, data $A$ and $B$ are mapped onto the center of cells, while, in Figure~\ref{fig:ex2}, $C$ is mapped onto the edges.
@@
 
-The stencil computation in Figure~\ref{fig:ex1} computes $A$ using $B$, while the one in Figure~\ref{fig:ex2} computes $A$ using $C$.
-Thus, a stencil computation is defined by a set of input data (in those examples only one input data), and a single output data, the result.
+The stencil kernel of Figure~\ref{fig:ex1} computes $A$ using $B$, while the one of Figure~\ref{fig:ex2} computes $A$ using $C$.
+Thus, a stencil kernel is defined by a set of input data (in those examples only one input data), and a single output data, the result.
 The \emph{computation domain} is a subset of elements to compute on the output data.
-In other words a computation domain is a subset of a mesh domain.
-For example, in the first stencil, the computation of $A$ is performed on plain elements, while dotted elements are not computed.
-On the other hand, on the second computation, the computation domain of $A$ is exactly equal to the cell mesh domain.
-The computation domain represents the set of elements on which the computation is performed, in other words, the space loop.
+In other words a computation domain is a subset of a mesh entity.
+For example, in the first stencil kernel, the computation of $A$ is performed on elements represented with full lines, while dotted elements are not computed.
+On the other hand, on the second kernel, the computation domain of $A$ is exactly equal to the cell mesh entity.
+The computation domain represents the set of elements on which the kernel is performed, in other words, the space loop.
 
-Finally, a stencil computation is composed of a \emph{numerical expression} which indicates how to compute an element of the computation domain (on the output data), using elements of the set of input data.
-The numerical expression of a stencil computation has the particularity to use not only the element to compute but also a \emph{neighborhood} of this element, also called a \emph{stencil shape}.
+Finally, a stencil kernel is composed of a \emph{numerical expression} which indicates how to compute an element of the computation domain (on the output data), using elements of the set of input data.
+The numerical expression of a stencil kernel has the particularity to use not only the element to compute but also a \emph{neighborhood} of this element, also called a \emph{stencil shape}.
 For example, the stencil shape of the first computation in Figure~\ref{fig:ex1} contains direct neighbors on the right, left, top and bottom.
-Sometimes, the neighborhood shape can also access a different mesh domain, as for example in Figure~\ref{fig:ex2}.
-Actually, in this computation, the neighborhood contains edges on the left and on the right of a cell.
-As an example, the numerical expression of the first example could be:
+Sometimes, the neighborhood shape can also access a different mesh entity, as for example in Figure~\ref{fig:ex2}.
+Actually, in this second kernel, the neighborhood contains edges on the left and on the right of a cell.
+As an example, the numerical expression of the first kernel could be:
 \begin{equation*} 
@@
 \end{equation*}
-and the numerical expression of the second example could be:
+and the numerical expression of the second kernel could be:
 \begin{equation*} 
@@
 
+To resume, concepts used in this paper to define a stencil kernel are \emph{mesh}, \emph{mesh entity}, \emph{data}, \emph{computation domain}, \emph{numerical expression} and \emph{stencil shape}. 
+If stencil kernels have been studied a lot, the formalization of real overall mesh-based numerical simulations is a rather new contribution.
+Actually, paying attention to complex numerical simulations, it appears that most of them are composed of more than one stencil kernel, with one or more stencil shapes and of additionnal local computations.
+For example, we would like to formalize and parallelize a numerical simulation which chains stencil kernels of Figures~\ref{fig:ex1} and~\ref{fig:ex2}.
+The next section formalize concepts of \emph{stencil kernel}, \emph{local computation} and what we call a \emph{multi-stencil program}.
+%\fix{cp2hc: [optionnelle] Par example? domaine applicatif?}
+
 %----------------------------------------
-\subsection{A Multi-Stencil Formlization}
+\subsection{A Multi-Stencil Formalization}
 \label{sect:multistencil}
 %----------------------------------------
-A real case numerical simulation is most of the time composed of more than one stencil computation, with one or more stencil shapes, and onto one or more data, and of additionnal local computations.
-\fix{cp2hc: [optionnelle] Par example? domaine applicatif?}
 Let introduce some formal definitions used in the rest of the paper.
-
-For $\Delta$ the set of data of the simulation, a stencil computation $s$ is defined as the quintuplet
+For $\Delta$ the set of data of the simulation, a stencil kernel $s$ is defined as the quintuplet
 \begin{equation} 
-s(R,w,exp,d_c),
+s(R,w,exp,d),
 \label{eq:st}
@@
 where $R$ is a set of pair $(r,n)$, with $r \in \Delta$ is a data read by the computation, and $n$ is a stencil shape (a neighborhood) applied on $r$.
-The data written by the computation is denoted $w \in \Delta$, $exp$ is the numerical expression of the stencil computation.
-Finally the numerical expression is applied on the computation domain $d_c$.
+The data written by the kernel is denoted $w \in \Delta$, $exp$ is the numerical expression of the stencil kernel.
+Finally the numerical expression is applied on the computation domain $d$.
 
-For example, in the example of Figure~\ref{fig:ex1}, assuming the computation domain (plain lines) is $dc1$ and the stencil shape is $n1$, the stencil computation can be defined with:
+For example, in Figure~\ref{fig:ex1}, assuming the computation domain (full lines) is $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined with:
 \begin{equation*}
-R: {(B,n1)} \quad w: A \quad d_c: dc1
+R: {(B,n1)} \quad w: A \quad d: dc1
 \end{equation*}
@@
 \end{equation*}
-On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil computation can be defined with:
+On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil kernel can be defined with:
 \begin{equation*}
-R: {(C,n2),(A,null)} \quad w: A \quad d_c: dc2
+R: {(C,n2),(A,null)} \quad w: A \quad d: dc2
 \end{equation*}
@@
 
-A local, or auxiliary, computation is almost the same than a stencil computation without neighborhood information in the numerical expression, which makes this kind of computation purely local.
+A local, or auxiliary, computation is a particular case of stencil kernel without neighborhood information in the numerical expression, which makes this kind of computation purely local.
 A local computation $l$ is defined as the quadruplet:
 \begin{equation} 
-l(R_l,w,exp,d_c),
+l(R_l,w,exp,d),
 \label{eq:loc}
@@
 \begin{equation} 
-\mathcal{MSP}(T,\mathcal{M},\mathcal{D}_m,\mathcal{D}_c,\Delta,\Gamma),
+\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma),
 \label{eq:msp}
 \end{equation}
-where $T$ is the set of time iteration to run the simulation, $\mathcal{M}$ is the mesh of the simulation, $\mathcal{D}_m$ is the set of mesh domains, $\mathcal{D}_c$ is the set of computation domains used into computations, $\Delta$ is the set of data of the simulation, each one mapped onto a mesh domain, and finally $\Gamma$ is the set of computations.
-The set of computations $\Gamma$ is composed of an ordered list of stencil and local computations.
-One can notice that this work is limited to a single type of mesh in a given simulation.
+where $T$ is the set of time iteration to run the simulation, $\mathcal{M}$ is the mesh of the simulation, $\mathcal{E}$ is the set of mesh entities, $\mathcal{D}$ is the set of computation domains used into computations, $\Delta$ is the set of data of the simulation, each one mapped onto a mesh entity, and finally $\Gamma$ is the set of computations.
+The set of computations $\Gamma$ is an ordered list of stencil kernels and local computations.
+
+\fix{HC 2 all : Est-ce que je laisse cette phrase : One can notice that this work is limited to a single type of mesh in a given simulation.}
 

### 2015-07-30 WOLFHPC15/article/introduction.tex
fix 1,2,3,4 helene 30_07

@@
 A critical question however concerns the specificity of DSLs.
-The less specific the DSL is, the more informations its users must provide and the less automatic optimizations can be embedded in the compiler.
-On the other side of the scope, the more specific the DSL is, the less different code can use it and the more difficult it is to amortize the implementation cost of the DSL (language, compiler, and runtime).
+Actually, the less specific the DSL is, the more information its users must provide and the less automatic optimizations can be embedded in the compiler.
+%On the other side of the scope
+Conversely, the more specific the DSL is, the less number of different applications can use it and the more difficult it is to amortize the implementation cost of the DSL (language, compiler, and runtime).
 Programming difficulties, as well as non-maintainability and non-portability often seem to be simply deported from the scientist to DSL designers.
@@
 The main idea of those solutions is to propose a single DSL to write new DSLs.
-In this paper, we study a first contribution to another approach to improve maintainability and code-reuse in DSLs.
+In this paper, we study a first contribution to another approach to improve maintainability and portability in DSLs.
 This approach proposes a transformation from a DSL to a component-based runtime.
+Component-based Software Engineering have proved many times good properties for code re-use, separation of concerns, maintainability and productivity of codes.
+For this reason, we think that mixing DSLs and component-based runtimes could enable those properties, by inheritance, in DSLs.
 To evaluate this approach, a proper new DSL for \emph{multi-stencil} programs, and its compiler are presented in this paper.
-\fix{cp2cp+hc: to dev benefits of components.}
+%\fix{cp2cp+hc: to dev benefits of components.}
 
 Many solutions to automatically optimize and parallelize partial differential equations (PDEs) solvers have been proposed in the literature, either as specific libraries~\cite{petsc-efficient,Trilinos-Overview,CPE:CPE3494} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
-Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil.
-However, a real case numerical simulation is most of the time not composed of a single stencil computation, but of a set of stencil computations and a set of additional local auxiliary computations.
+Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil kernel.
+However, a real case numerical simulation is most of the time not composed of a single stencil kernel, but of a set of stencil kernels and a set of additional local auxiliary computations.
 The DSL proposed in this paper, called MSL, is a DSL to generate a parallel component-based structure of an overall numerical simulation, that we call a \emph{multi-stencil program}.
-As it will be detailed in the related work, this is complementary to the optimization and parallelization of a single stencil computation.
-\fix{JB: j'utiliserais bien le terme "kernel" ou "stencil kernel" pour ce que tu apelles "stencil" et "stencil" tout court pour ce que tu apelles "stencil shape"}
+As it will be detailed in the related work, this is complementary to the optimization and parallelization of a single stencil kernel.
+%\fix{JB: j'utiliserais bien le terme "kernel" ou "stencil kernel" pour ce que tu apelles "stencil" et "stencil" tout court pour ce que tu apelles "stencil shape"}
 

### 2015-07-30 WOLFHPC15/article/ms_msc.tex
fix 1,2,3,4 helene 30_07

@@
 Lines 1 to 9 define what is a MSL program with the different parts memntionned above.
-The remaining of this section describes the parts of a MSL program and presents an example.
-We can note that, as MSL is a descriptive language, almost all terminals $meshid$, $meshdom$, $compdomid$, $dataid$, $compid$ ans $neighborid$ are string identifiers.
+The remaining of this section describes those different parts of a MSL program and presents an example.
+We can note that, as MSL is a descriptive language, almost all terminals $meshid$, $meshdom$, $compdomid$, $dataid$, $compid$ and $neighborid$, are string identifiers.
 Only $num$ is an integer terminal.
@@
 As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is directly defined by a string identifier.
-Mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
+Then, mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
 In the example given in Figure~\ref{fig:mslex} a Cartesian mesh, called $cart$, and three mesh entities, called $cell$, $edgex$ and $edgey$, are defined.
@@
 \subsubsection*{Computation domains and their dependencies}
-A computation domain can be seen as a subpart of a mesh entity which will be used during one or more computations.
+A computation domain can be seen as a subpart of a mesh entity which will be used during one or more computation.
 For each computation domain, the mesh entity from which it is built is indicated (lines 3-4 and 12-14 in Figure~\ref{fig:grammar}).
 By default two computation domains always intersect except when an independency relation is mentionned in the section \emph{independent} of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
-In the example of Figure~\ref{fig:mslex}, on lines 4 to 6, an whole computation domain is defined for each mesh entity, called $allcell$, $alledgex$ and $alledgey$, and on lines 7 and 8 two subparts of the mesh entity $edgex$ are also defined, called $part1edgex$ and $part2edgex$.
-Line 10 means that $part1edgex$ and $part2edgex$ do not intersect.
+In the example of Figure~\ref{fig:mslex}, on lines 4 to 6, a whole computation domain is defined for each mesh entity, called $allcell$, $alledgex$ and $alledgey$, and on lines 7 and 8 two subparts of the mesh entity $edgex$ are also defined, called $part1edgex$ and $part2edgex$.
+Finally, line 10 means that $part1edgex$ and $part2edgex$ do not intersect.
 
@@
 A data is a quantity to simulate and is mapped onto a mesh entity, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}.
-The $time$ section simplies indicates a number of iterations to perform in the simulation.
+The $time$ section simply indicates a number of iterations to perform in the simulation.
 In the example of Figure~\ref{fig:mslex} are defined ten data and 500 iterations (lines 11-22).
@@
 \subsubsection*{Computations description}
-Finally, the language follows the two definitions of a stencil computation and a local computation given in Section~\ref{sect:multistencil}.
-However as MSL does not handle numerical expressions, the term $exp$ of Definition~\ref{eq:st} and Definition~\ref{eq:loc} is not used in the language.
-The exact code of the computation as to be writtenin the final ready-to-fill parallel pattern of the simulation, using external languages or interfaces (\cf Section~\ref{sect:component}).
+Finally, the language follows the two definitions of a stencil kernel and a local computation given in Section~\ref{sect:multistencil}.
+However, as MSL does not handle numerical expressions, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} is not used in the language.
+The exact code of the computation as to be writtenin into the final ready-to-fill parallel pattern of the simulation, using external languages or interfaces (\cf Section~\ref{sect:component}).
 Moreover, to propose an intuitive syntax, information are given in a different order than in formal definitions.
-Thus, the description of a computation, as indicated in lines 24 and 25 of Figure~\ref{fig:grammar}, consists in first the identifier of the data to compute ($w$ in the formal description), in brackets the identifier of the computation domain on which the computation is applied ($d_c$ in the formal definition), an identifier for the computation, and the set of identifiers for the data read by the computation with parenthesis.
+Thus, the description of a computation, as indicated in lines 24 and 25 of Figure~\ref{fig:grammar}, consists in first the identifier of the data to compute ($w$ in the formal description), in brackets the identifier of the computation domain on which the computation is applied ($d$ in the formal definition), an identifier for the computation kernel, and the set of identifiers for the data read by the computation, with parenthesis.
 For each data read by the computation brackets indicate the stencil shape identifier used on the data ($R$ in the formal definition of a stencil).
@@
 %----------------------------------------
-The MSC transformation, which is a subpart of the overall compiler MSCAC, makes use of the ordered list of computations $\Gamma$, which can directly been extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
-This transformation phase of the compilation is divided in four different steps; it transforms $\Gamma$ to a series-parallel tree decomposition~\ref{???}.
+The MSC transformation, which is a subpart of the overall compiler MSCAC, makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
+This transformation phase of the compilation is divided in five different steps; it transforms $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
 This transformation phase handles data parallelism through the first step, and task parallelism through the three next steps.
-This section gives an overview of these four steps of MSC; the detailed formalism and associated algorithms of those steps are described in the research report~\cite{????}.
+This section gives an overview of the four steps of the MSC transformation; the detailed formalism and the associated algorithms of those steps are described in the research report~\cite{????}.
 
-\subsubsection*{The Computation Ordered List: $\Gamma$}
+\subsubsection*{The Ordered List of Computations: $\Gamma$}
 $\Gamma$ is directly obtained from the parser.
@@
 
-\subsubsection*{The Computation Synchronized Ordered List: $\Gamma_{data}$}
-Considering a data parallelization of the simulation, the mesh and data of the simulation are splitted among available processors (or cores) and a single program is applied on each sub-part with additionnal synchronizations between processors or cores.
-For example, as a stencil code access neighborhood value, values computed by another processor are needed and synchronized.
-From the ordered list of computations $\Gamma$, the needed synchronizations can be automatically detected.
+\subsubsection*{The Synchronized Ordered List of Computations: $\Gamma_{sync}$}
+Considering a data parallelization of the simulation, the mesh and data of the simulation are splitted among available processors (or cores) and a single program is applied on each sub-part with additionnal synchronizations between processors.
+For example, as a stencil kernel accesses neighborhood values, values computed by another processor are needed and synchronized.
+From the ordered list of computations $\Gamma$, the needed synchronizations can be automatically added.
 A synchronization is needed each time a data read by a stencil computation has been written by a previous computation.
@@
 The computation domain of such a synchronization computation is the mesh entity on which the data is declared.
-As a result $\Gamma$ is transformed to a synchronized ordered list of computations $\Gamma_{data}$.
+As a result $\Gamma$ is transformed to a synchronized ordered list of computations $\Gamma_{sync}$.
 For example, in Figure~\ref{fig:mslex}, the stencil computation $c_1$ read the data $b$ which has been written by the computation $c_0$.
-For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed to the sublist $[c_0,sync_1,c_1]$ in $\Gamma_{data}$.
+For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed to the sublist $[c_0,sync_1,c_1]$ in $\Gamma_{sync}$.
 The new computation $sync_1$ read and write $b$ and is applied on the computation domain of $c_1$.
 As a result, a dependency is kept between $c_0$ and $sync_1$ and between $sync_1$, and $c_1$.
-The same transformation is performed for the stencils $c_4$ and $c_8$.
-Thus, $\Gamma_{data} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
+The same transformation is performed for all stencils: $c_4$ and $c_8$.
+Thus, $\Gamma_{sync} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
 
-\subsubsection*{The Dependency Graph: $\Gamma_{hybrid}$}
-From the synchronized ordered list of computations $\Gamma_{data}$, a dependency graph $\Gamma_{hybrid}$ is built.
-A dependency exists between two computations (including synchronizations) if and only if a data read has been written by a previous computation in $\Gamma_{data}$, and if the computation domains intersect.
+\subsubsection*{The Dependency Graph: $\Gamma_{dep}$}
+From the synchronized ordered list of computations $\Gamma_{sync}$, a dependency graph $\Gamma_{dep}$ is built.
+A dependency exists between two computations (including synchronizations) if and only if a data read has been written by a previous computation in $\Gamma_{sync}$, and if the computation domains intersect.
 In other words, the only case where a data dependency exists but is not taken into account is when the computation domains do not intersect.
-Nodes of the dependency graph represent computations of $\Gamma_{data}$, while edges are dependencies between them.
-The dependency graph $\Gamma_{hybrid}$ is a directed acyclic graph (\emph{dag}).
-For example, the dependency dag $\Gamma_{hybrid}$ of the example of Figure~\ref{fig:mslex} is displayed in Figure~\ref{fig:hyb}.
+Nodes of the dependency graph represent computations of $\Gamma_{sync}$, while edges are dependencies between them.
+The dependency graph $\Gamma_{dep}$ is a directed acyclic graph (\emph{dag}).
+For example, the dependency dag $\Gamma_{dep}$ of the example of Figure~\ref{fig:mslex} is displayed in Figure~\ref{fig:hyb}.
 
@@
 To do so the dependency graph is transformed to a minimal series-parallel graph~\cite{Valdes:1979:RSP:800135.804393}.
-As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape}, is not found in the graph.
-To transform $\Gamma_{hybrid}$ to the minimal series-parallel graph $\Gamma_{msp}$, a breadth first search algorithm is applied to build a complete bipartite graph between two levels of $\Gamma_{hybrid}$~\cite{Mitchell:2004:CMV:1082101.1082117}.
+As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape}, is not found in the dag.
+Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes}.
 

### 2015-07-30 WOLFHPC15/article/mscac.tex
fix 1,2,3,4 helene 30_07

@@
-The solution proposed in this paper contains a domain specific language to describe a multi-stencil program, i.e.
-an overall numerical simulation, and a compiler which successively parses the front-end language, transforms the internal representation, and finally dumps the representation to a back-end code.
-This back-end code is component based.
+The solution proposed in this paper contains a domain specific language to describe a multi-stencil program, \ie an overall numerical simulation, and a compiler which successively parses the front-end language, transforms the internal representation, and finally dumps the representation to a component-based back-end code.
 
@@
 As a result, MSL is not made to handle the expression of numerical computations, but only the description of a multi-stencil simulation, in a coarse-grain fashion.
-Numerical computations are encapsulated in specific components that users have to develop.
-In a MSL description, six main sections are defined, as the six tuples of the definition of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\mathcal{D}_m,\mathcal{D}_c,\Delta,\Gamma)$ (\cf Equation~\ref{eq:msp}).
+%Numerical computations are encapsulated in specific components that users have to develop.
+In a MSL description, six main sections are defined, as the six tuples of the definition of a multi-stencil program of Equation~\ref{eq:msp}.
 
@@
   ]
-  \node (MSL) {MS Language};
+  \node (MSL) {MSL};
   \node[outer,below=0.5 of MSL,draw=black] (MSCAC) {
@@
   };
-  \node [below=0.5 of MSCAC] (CA) {Component Assembly};
+  \node [below=0.5 of MSCAC] (CA) {Component-based runtime};
   \node [rotate=90,above left=0.5 of MSC] (CAid) {MSCAC};
@@
 
-The final phase of compilation, the dump, uses the static schedule of the tasks, created from $\Gamma$, and additionnal information on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{D}_m$ and $\mathcal{D}_c$, given by the \emph{parser}, to produce a back-end solution which represents the parallel pattern or skeleton of the overall simulation.
+The final phase of compilation, the dump, uses the static schedule of the tasks, created from $\Gamma$, and additionnal information on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{E}$ and $\mathcal{D}$, given by the \emph{parser}, to produce a back-end solution which represents the parallel pattern or skeleton of the overall simulation.
 We could have propose a dump to OpenMP~\cite{660313}, however this paper presents a first contribution to dump a DSL to a component-based runtime.
@@
 
-The back-end of the compiler builds a component assembly, that can be divided in two parts: the components implementing the actual computation and the static schedule, and the components needed needed to manage it, \ie the runtime support.
+The back-end of the compiler builds a component assembly, that can be divided in two parts: the components implementing the actual computation and the static schedule, and the components needed to manage it, \ie the runtime support.
 A general view of this assembly is proposed in Figure~\ref{fig:assembly}; it is explained in details in Section~\ref{sect:component}.
-As illustrated in the figure, each component is linked to one or more tuple of the multi-stencil program definition: \emph{DDS} component represents $\mathcal{M}$ and $\mathcal{D}_m$; \emph{Data} component represents $\Delta$ and $\mathcal{D}_c$; \emph{Time} component represents $T$, and \emph{Computations} component (which is generated by MSC) represents $\Gamma$.
+As illustrated in the figure, each component is linked to one or more tuple of the multi-stencil program definition: \emph{DDS} component represents $\mathcal{M}$ and $\mathcal{E}$; \emph{Data} component represents $\Delta$ and $\mathcal{D}$; \emph{Time} component represents $T$, and \emph{Computations} component (which is generated by MSC) represents $\Gamma$.
 
@@
 For example, the back-end evaluated in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}.
-A work in progress replaces this DDS component by another one making using of Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.
+A work in progress replaces this DDS component by another one making use of Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.
 
@@
    \node[use,above=0.8cm of Data] (Datau) {};
-   \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}_c$};
+   \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
 
@@
    \node[component,above=0.8cm of DDSp2] (DDS) {$DDS$};
-   \node[right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{D}_m$};
+   \node[right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
 

### 2015-07-30 WOLFHPC15/article/evaluation.tex
fix 1,2,3,4 helene 30_07
fix evaluation helene 30_07

@@
 
-The MSL program of FullSWOF contains three mesh entities, seven computation domains, fourty-eight data and ninty-eight computations composed of thirty-two stencil computations and sixty-six local computations.
+The MSL program of FullSWOF contains three mesh entities, seven computation domains, fourty-eight data and ninty-eight computations composed of thirty-two stencil kernels and sixty-six local computations.
 
@@
 
-%----------------------------------------
-\subsection{A real case simulation}
-%----------------------------------------
-SkelGIS and MSD\\
-Description of shallow water equations\\
-MSL language description\\
-The obtained series-parallel graph, comments.
-
-%----------------------------------------
-\subsection{Performances}
-%----------------------------------------
 \begin{itemize}

### 2015-07-31 WOLFHPC15/article/concept.tex
2.1

--- 2015-07-30 WOLFHPC15/article/concept.tex
@@
 %----------------------------------------
-To numerically solve a set of PDEs, iterative methods (finite difference, finite volume, finite element methods etc.) are frequently used to approximate the solution through a discretized (step by step) phenomena.
+To numerically solve a set of PDEs, iterative methods (finite difference, finite volume, finite element methods etc.) are frequently used to approximate the solution through a discretized (step by step) process.
 Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied onto a mesh (space discretization).
 In other words, in a mesh-based numerical simulation, the PDEs are transformed to a set of numerical computations applied at each time step on elements of the discretized space domain (the mesh).
-Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}~\cite{spaaTangCKLL11}.
-To ease programming of stencil computations, and because of the relative regularity of such a computation, many solutions (languages or libraries) propose an easy way to describe the stencil code, while producing transparently an optimized and parallelized stencil code.
-For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, respectively proposing an optimized cache tiling technique, and a separation of concerns between parallelization strategies and computation descriptions.
+This paper focuses on one specific category of such numerical schemes based on \textit{stencil kernels}~\cite{spaaTangCKLL11}.
+This section introduces some definitions related to those.
 
-\fix{cp2hc: transition a revoir.}
-
-Let introduce some concepts used in this paper.
-First, a stencil code is based on a \emph{mesh} which is the discretization of the physical real domain.
-A mesh is a connected undirected graph without bridges (an edge is a bridge if its removal results in two disconnected graphs), where nodes and edges are linked to form cells (closure).
-For example, a structured mesh is illustrated in Figure~\ref{fig:mesh}.
+A \emph{mesh} is the discretization of a physical domain.
+It is a connected undirected graph without bridges (an edge is a bridge if its removal results in two disconnected graphs), where nodes and edges are linked to form cells (closure).
+An example of structured mesh is illustrated in Figure~\ref{fig:mesh}.
 Each cell contains four nodes and four edges.
-Second, \emph{mesh domains} are subset elements of the mesh, such as the center of cells, edges or nodes.
-In Figure~\ref{fig:mesh}, two mesh domains are illustrated, the center of the cells (named \texttt{Cells}, in red) and edges on the horizontal axe (named \texttt{Edgex}, in blue on left and right of each cell).
-A \emph{data} is a quantity to simulate or a temporary data to perform computations.
-A data is mapped onto a mesh domain.
+\emph{Mesh entities} are elements of the mesh, such as the center of cells, edges or nodes.
+In Figure~\ref{fig:mesh}, two mesh entities are illustrated, the center of the cells (named \texttt{Cells}, in red) and edges on the horizontal axis (named \texttt{Edgex}, in blue on left and right of each cell).
+A \emph{data} is a simulated quantity on which computations are performed.
+A data is mapped onto mesh entities.
 For example, in Figure~\ref{fig:ex1}, data $A$ and $B$ are mapped onto the center of cells, while, in Figure~\ref{fig:ex2}, $C$ is mapped onto the edges.
+\todo[inline]{JB: structured/unstructured pas defini, pas de notion de coordonees des elements du maillage dans l'espace reel?}
 
@@
 
-The stencil computation in Figure~\ref{fig:ex1} computes $A$ using $B$, while the one in Figure~\ref{fig:ex2} computes $A$ using $C$.
-Thus, a stencil computation is defined by a set of input data (in those examples only one input data), and a single output data, the result.
-The \emph{computation domain} is a subset of elements to compute on the output data.
-In other words a computation domain is a subset of a mesh domain.
+A \emph{stencil kernel} computes the value of one data or a subpart of it (its \emph{computation domain}) using a numerical expression taking as input on one or more other data.
+For example, the stencil kernel illustrated in Figure~\ref{fig:ex1} computes $A$ using $B$, while the one in Figure~\ref{fig:ex2} computes $A$ using $C$.
+Thus, a stencil kernel is defined by a set of input data (in those examples only one input data), and a single output data, the result.
+The computation domain is a subset of the mesh entities on which the output data is mapped.
 For example, in the first stencil, the computation of $A$ is performed on plain elements, while dotted elements are not computed.
-On the other hand, on the second computation, the computation domain of $A$ is exactly equal to the cell mesh domain.
-The computation domain represents the set of elements on which the computation is performed, in other words, the space loop.
+On the other hand, on the second computation, the computation domain of $A$ contains all the mesh entities on which it is mapped.
+This computation domain defines the elements aver which the space loop iterates.
 
-Finally, a stencil computation is composed of a \emph{numerical expression} which indicates how to compute an element of the computation domain (on the output data), using elements of the set of input data.
-The numerical expression of a stencil computation has the particularity to use not only the element to compute but also a \emph{neighborhood} of this element, also called a \emph{stencil shape}.
+The numerical expression of a stencil kernel has the particularity to use not only the element to compute but also a \emph{neighborhood} of this element, also called a \emph{stencil shape}.
 For example, the stencil shape of the first computation in Figure~\ref{fig:ex1} contains direct neighbors on the right, left, top and bottom.
-Sometimes, the neighborhood shape can also access a different mesh domain, as for example in Figure~\ref{fig:ex2}.
+Sometimes, the neighborhood can also access different mesh entities, as for example in Figure~\ref{fig:ex2}.
 Actually, in this computation, the neighborhood contains edges on the left and on the right of a cell.
@@
 \end{equation*}
+
+As can be seen from these definitions, stencil kernels implementations share many properties from an algorithmic point of view that can be used to apply well known optimization and parallelization strategies.
+Many solutions (languages or libraries) thus propose to ease their programming by producing optimized and parallelized code from a simple description of the local computation to apply on each element.
+For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, respectively proposing an optimized cache tiling technique, and a separation of concerns between parallelization strategies and computation descriptions.
 

### 2015-07-31 WOLFHPC15/article/introduction.tex
2.1

--- 2015-07-30 WOLFHPC15/article/introduction.tex
@@
 Many solutions to automatically optimize and parallelize partial differential equations (PDEs) solvers have been proposed in the literature, either as specific libraries~\cite{petsc-efficient,Trilinos-Overview,CPE:CPE3494} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
-Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil.
-However, a real case numerical simulation is most of the time not composed of a single stencil computation, but of a set of stencil computations and a set of additional local auxiliary computations.
+Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil kernel.
+However, a real case numerical simulation is most of the time not composed of a single stencil kernel, but of a set of stencil kernels and a set of additional local auxiliary computations.
 The DSL proposed in this paper, called MSL, is a DSL to generate a parallel component-based structure of an overall numerical simulation, that we call a \emph{multi-stencil program}.
-As it will be detailed in the related work, this is complementary to the optimization and parallelization of a single stencil computation.
-\fix{JB: j'utiliserais bien le terme "kernel" ou "stencil kernel" pour ce que tu apelles "stencil" et "stencil" tout court pour ce que tu apelles "stencil shape"}
+As it will be detailed in the related work, this is complementary to the optimization and parallelization of a single stencil kernel.
 

### 2015-07-31 WOLFHPC15/article/concept.tex
wip

@@
 
-A \emph{stencil kernel} computes the value of one data or a subpart of it (its \emph{computation domain}) using a numerical expression taking as input on one or more other data.
+A \emph{stencil kernel} computes the value of one data or a subpart of it (its \emph{computation domain}) using a \emph{numerical expression} taking as input on one or more other data.
 For example, the stencil kernel illustrated in Figure~\ref{fig:ex1} computes $A$ using $B$, while the one in Figure~\ref{fig:ex2} computes $A$ using $C$.
@@
 For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, respectively proposing an optimized cache tiling technique, and a separation of concerns between parallelization strategies and computation descriptions.
+
+To resume, concepts used in this paper to define a stencil kernel are \emph{mesh}, \emph{mesh entity}, \emph{data}, \emph{computation domain}, \emph{numerical expression} and \emph{stencil shape}. 
+If stencil kernels have been studied a lot, the formalization of real overall mesh-based numerical simulations is a rather new contribution.
+Actually, paying attention to complex numerical simulations, it appears that most of them are composed of more than one stencil kernel, with one or more stencil shapes and of additional local computations.
+For example, we would like to formalize and parallelize a numerical simulation which chains stencil kernels of Figures~\ref{fig:ex1} and~\ref{fig:ex2}.
+The next section formalizes concepts of \emph{stencil kernel}, \emph{local computation} and what we call a \emph{multi-stencil program}.
 

### 2015-07-31 WOLFHPC15/article/ms_msc.tex
Fix section & subsection title

@@
 
-\subsubsection*{Mesh and mesh entities}
+\subsubsection*{Mesh and Mesh Entities}
 As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is directly defined by a string identifier.
@@
 
-\subsubsection*{Computation domains and their dependencies}
+\subsubsection*{Computation Domains and Their Dependencies}
 A computation domain can be seen as a subpart of a mesh entity which will be used during one or more computation.
@@
 
-\subsubsection*{Data and time}
+\subsubsection*{Data and Time}
 A data is a quantity to simulate and is mapped onto a mesh entity, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}.
@@
 
-\subsubsection*{Computations description}
+\subsubsection*{Computation Description}
 Finally, the language follows the two definitions of a stencil kernel and a local computation given in Section~\ref{sect:multistencil}.
@@
 %----------------------------------------
-\subsection{MSC transformation}
+\subsection{MSC Parallelization}
 %----------------------------------------
-The MSC transformation, which is a subpart of the overall compiler MSCAC, makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
+The MSC parallelization, which is a subpart of the overall compiler MSCAC, makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
 This transformation phase of the compilation is divided in five different steps; it transforms $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.

### 2015-07-31 WOLFHPC15/article/concept.tex

@@
 A \emph{data} is a simulated quantity on which computations are performed.
-A data is mapped onto mesh entities.
+Each data is mapped onto mesh entities.
 For example, in Figure~\ref{fig:ex1}, data $A$ and $B$ are mapped onto the center of cells, while, in Figure~\ref{fig:ex2}, $C$ is mapped onto the edges.
@@
 
-A \emph{stencil kernel} computes the value of one data or a subpart of it (its \emph{computation domain}) using a \emph{numerical expression} taking as input on one or more other data.
+A \emph{stencil kernel} computes the value of one data or a subpart of it (the kernel \emph{computation domain}) using a \emph{numerical expression} taking as input on one or more data.
 For example, the stencil kernel illustrated in Figure~\ref{fig:ex1} computes $A$ using $B$, while the one in Figure~\ref{fig:ex2} computes $A$ using $C$.
-Thus, a stencil kernel is defined by a set of input data (in those examples only one input data), and a single output data, the result.
+Thus, a stencil kernel is defined by its numerical expression, a set of input data (only one in the examples), and its unique output data, the result.
 The computation domain is a subset of the mesh entities on which the output data is mapped.
@@
 On the other hand, on the second computation, the computation domain of $A$ contains all the mesh entities on which it is mapped.
-This computation domain defines the elements aver which the space loop iterates.
+This computation domain defines the elements over which the space loop iterates.
 
-The numerical expression of a stencil kernel has the particularity to use not only the element to compute but also a \emph{neighborhood} of this element, also called a \emph{stencil shape}.
+The numerical expression of a stencil kernel has the particularity to compute each element of the result independently using only elements of the inputs whose coordinates are close to that of the result element.
+The elements accessed form a \emph{neighborhood} of the output element coordinate known as the \emph{stencil shape}.
 For example, the stencil shape of the first computation in Figure~\ref{fig:ex1} contains direct neighbors on the right, left, top and bottom.
@@
 For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, respectively proposing an optimized cache tiling technique, and a separation of concerns between parallelization strategies and computation descriptions.
+\todo[inline]{JB: a virer?}
 
-To resume, concepts used in this paper to define a stencil kernel are \emph{mesh}, \emph{mesh entity}, \emph{data}, \emph{computation domain}, \emph{numerical expression} and \emph{stencil shape}. 
+To resume, concepts used in this paper to define a stencil kernel are \emph{mesh}, \emph{mesh entities}, \emph{data}, \emph{computation domain}, \emph{numerical expression} and \emph{stencil shape}. 
 If stencil kernels have been studied a lot, the formalization of real overall mesh-based numerical simulations is a rather new contribution.
@@
 %----------------------------------------
-\subsection{A Multi-Stencil Formlization}
+\subsection{A Multi-Stencil Formalization}
 \label{sect:multistencil}
 %----------------------------------------
-A real case numerical simulation is most of the time composed of more than one stencil computation, with one or more stencil shapes, and onto one or more data, and of additionnal local computations.
-\fix{cp2hc: [optionnelle] Par example? domaine applicatif?}
 Let introduce some formal definitions used in the rest of the paper.
-
-For $\Delta$ the set of data of the simulation, a stencil computation $s$ is defined as the quintuplet
+For $\Delta$ the set of data of the simulation, a stencil kernel $s$ is defined as the quintuplet
 \begin{equation} 
-s(R,w,exp,d_c),
+s(R,w,exp,d),
 \label{eq:st}

### 2015-07-31 WOLFHPC15/article/component.tex
wolfhpc 29_07
Re-work of S4

@@
 multi-stencil application. But, first the section gives an overview of
-component models, and then it details \llc, the component model used in
+component models, and it details \llc, the component model used in
 this work.
@@
 engineering~\cite{Szyperski:2002:CSB:515228}, which promotes code
-re-use and separations of concerns. An application is made by a set of
-component; a component being a kind of black box that interacts only
-through ports: ports specify the services provided and required.
+re-use, separations of concerns, and thus maintainability. An
+application is made by a set of component; a component being a kind of
+black box that interacts only through ports: ports specify the
+services provided and required.
+%
+With respect to high performance computing, some works have also shown
+that component models can achieve the needed level of performance, and
+scalability while also helping in application
+portability~\cite{Bernholdt01052006, bigot:inria-00388508, UCHPC2015}
 
-Component models are known to improved code re-use, separation of
-concerns, and maintainability of
-applications~\cite{Szyperski:2002:CSB:515228}. More recent works have
-also shown that component models can handled performance, scalability,
-and portability of applications~\cite{Bernholdt01052006,
-  bigot:inria-00388508, l2c}. Many component models exist, each of
-them with its own specifications and singularities. In distributed
-computing, well known models are CCM~\cite{corba:omg06} (CORBA
-Component Model), and GCM~\cite{Baude} (Grid Component Model); in HPC,
-there are CCA~\cite{Bernholdt01052006} (Common Component
-Architecture), and \llc~\cite{l2c} (Low Level Component), for example.
+Many component models exist, each of them with its own specifications
+and singularities. In distributed computing, well known component
+models are CCM~\cite{corba:omg06} (CORBA Component Model), and
+GCM~\cite{Baude} (Grid Component Model); in HPC, there are
+CCA~\cite{Bernholdt01052006} (Common Component Architecture), and
+\llc~\cite{l2c} (Low Level Component), for example.
+%
+As this work makes use of \llc, in particular for the experiments, let
+describe it in more details.
 
@@
 
-The remaining of this paper focuses on \llc, in particular for the
-experiments. Let describe it in more details.
-
-\llc is a minimalist \texttt{C++} based HPC component model where a
-component extends the concept of class by specifying in its interfaces
-the services that it offers ($provide$ ports) and that it needs,
-either a single service instance ($use$ ports), or multiple service
-instances ($use-multiple$ ports). Services are \texttt{C++}
+\llc is a minimalist \texttt{C++} based HPC-oriented component model
+where a component extends the concept of class by specifying in its
+interfaces the services that it offers ($provide$ ports) and that it
+needs, either a single service instance ($use$ ports), or multiple
+service instances ($use-multiple$ ports). Services are \texttt{C++}
 interfaces. \llc also offer $MPI$ ports that enable components to
-share a MPI communicators.
+share MPI communicators. Components can also have attribute ports to
+be configured.
 %
 As illustrated in Figures~\ref{fig:ports}, a $provide$ port is
-represented by a white circle on the left of the component, a $use$
-port by a black circle, a $use-multiple$ port by a black circle with a
-white $m$ in it, both on the right of the component. MPI PORT!!!
+represented with a white circle on the left of the component, a $use$
+port with a black circle, a $use-multiple$ port by a black circle with
+a white $m$ in it, both on the right of the component. MPI port are
+connected with a black rectangle.
 
 A classical \llc-based application is a static \emph{assembly} of
-components made of instances of component and connections between
-ports. Such an assembly is describe in LAD (\llc Assembly Descriptor),
-an XML dialect~\cite{l2c}.
+components made of instances of components and of connections between
+component ports. Such an assembly is described in LAD, an XML
+dialect~\cite{l2c}.
 
@@
     (C.east) edge node {$u$} (u)
-    (C1)  edge node {$u'$} (um)
-    (p1) edge node {$p'$} (C1);
+    (C1)  edge node {$v$} (um)
+    (p1) edge node {$q$} (C1);
 \end{tikzpicture}
@@
   \path[-]
-    (p) edge node {} (C)
-    (C) edge node {} (u)
-    (C1)  edge node {} (um)
-    (u) edge node {} (C1);
+    (p) edge node {$p$} (C)
+    (C) edge node {$u$} (u)
+    (C1)  edge node {$v$} (um)
+    (u) edge node {$q$} (C1);
 \end{tikzpicture}
 \\\\
-c) MPI!!!
+c) \hspace{2mm}\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (C) at (0,0) {$C_2$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[mpi] (u) at (1.5,0) {};
+   \node[component,right=1.5 of u] (C1) {$C_3$};
+
+  \path[-]
+    (p) edge node {$r$} (C)
+    (C) edge node {$m_1$} (u)
+    (u) edge node {$m_2$} (C1);
+\end{tikzpicture}
 \\\\
-\caption{a) Component $C_0$ has a provide port ($p$) and a use port ($u$); Component $C_1$ has also a provide port ($p'$) but a use multiple port ($u'$).
-b) When a use port is connected to a provide port, only the use port is shown and by convention it has to be the left component.}
+\caption{Example components and ports representation.
+a) Component $C_0$ has a provide port ($p$) and a use port ($u$); Component $C_1$ has also a provide port ($p'$) but a use multiple port ($u'$).
+b) A use port is connected to a (compatible) provide port.
+c) Component $C_2$ and $C_3$ shares an MPI communicator.}
 \label{fig:ports}
@@
 %----------------------------------------
-\subsection{Control components}
+\subsection{MSCAC Runtime Overview}
 %----------------------------------------
-The series-parallel tree decomposition $\Gamma_{tsp}$ represents the control of the dependencies of the simulation.
-As a result to be able to dump it to a component assembly it is needed to introduce what can be called \emph{control components}.
-Those components can be used for any case of simulation which increases code reuse between simulations.
-A control component is composed of a single \emph{provide} port linked to a single execution method, most of the time called the \emph{go} method.
-We introduce three types of control components described bellow:
 
-\begin{itemize}
-\item a sequence component, called SEQ, is the direct dump to a sequence node of $\Gamma_{tsp}$.
-The role of this component is to call an ordered list of other components.
-For this reason, it is composed of an ordered \emph{use-multiple} port connected to the components to call as a sequence.
+Figure~\ref{fig:assembly} has presented an overview of the component
+assembly that is executed. As discusses in Section~\ref{sect:mscac},
+there are different types of components. Components~\texttt{Driver},
+\texttt{DriverApp}, and \texttt{Time} are always present; they do not
+depend on the input but for the configuration of an attribute port of
+Component~\texttt{Time} with the number of iterations. There is also a
+single Component~\texttt{DDS} as in the current version only one mesh
+type is supported. All these components are provided with the
+compiler; dumping this part of the assembly is straightforward.
 
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (seq) at (0,0) {$SEQ$};
-   \node[provide] (p) at (-1.5,0) {};
-   \node[use] (u) at (1.5,0) {$m$};
+While Component~\texttt{DDS} is managing the structure of the mesh,
+each data of the simulation is handled by a component of type
+\texttt{Data}. Therefore, the compiler generates in the assembly as
+many instances of such component as needed, from the \texttt{data}
+section of an MSL program.
 
-  \path[-]
-    (p) edge node {} (seq)
-    (seq) edge node {} (u);
-\end{tikzpicture}
-
-\item A parallel component, called PAR, is the direct dump to a parallel node of $\Gamma_{tsp}$.
-The role of this component is to call simultaneously a set of other components.
-For this reason, it is composed, as SEQ, of a \emph{use-multiple} port connected to the components to call in parallel.
-
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (seq) at (0,0) {$PAR$};
-   \node[provide] (p) at (-1.5,0) {};
-   \node[use] (u) at (1.5,0) {$m$};
-
-  \path[-]
-    (p) edge node {} (seq)
-    (seq) edge node {} (u);
-\end{tikzpicture}
-
-\item A synchronization component, called SYNC, is the direct dump to an update leaf of $\Gamma_{tsp}$.
-The role of this component is to call the synchronization of a given data.
-For this reason, it is composed of a \emph{use} port connected to the data to update.
-
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (sync) at (0,0) {$SYNC$};
-   \node[provide] (p) at (-1.5,0) {};
-   \node[use] (u) at (1.5,0) {};
-
-  \path[-]
-    (p) edge node {} (sync)
-    (sync) edge node {} (u);
-\end{tikzpicture}
-\end{itemize}
-
-In the computation part of the simulation a final component is needed, the computation component, denoted K.
-This type of component represents a single computation of the simulation, i.e.
-the computation of a single output data (written) using a set of input data (read).
-Using a \emph{use-multiple}, however, a list of component is used and the explicit identification of each component to use is lost.
-If for SEQ or PAR, the identification of components is not usefull, it is usefull to manipulate data in a computation.
-For this reason a computation component has as much \emph{use} ports as input and output data to manipulate.
-We denote this in the bellow representation as a star on the use port.
-
+The last part of the assembly is the computations. It contains actual
+computations in kernel components, denoted K, that have to be provided
+to the compiler. A kernel component encapsulates a computations. To
+access data, they are connected by the compiler to the adequate
+\texttt{Data} components; their data ports name must respect the names
+given in the \texttt{data} section of MSL. We denote the fact that they may have
+several use port by adding a star on the use port:
+\begin{center}
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
 \end{tikzpicture}
+\end{center}
+
+%% In the computation part of the simulation a final component is needed,
+%% the computation component, denoted K. This type of component
+%% represents a single computation of the simulation, i.e. the
+%% computation of a single output data (written) using a set of input
+%% data (read). Using a \emph{use-multiple}, however, a list of component
+%% is used and the explicit identification of each component to use is
+%% lost. If for SEQ or PAR, the identification of components is not
+%% usefull, it is usefull to manipulate data in a computation. For this
+%% reason a computation component has as much \emph{use} ports as input
+%% and output data to manipulate. We denote this in the bellow
+%% representation as a star on the use port.
+
+The computation part of the assembly also contains the static schedule
+of the computations, \ie $\Gamma_{tsp}$. This is achieved by using
+three specific components for managing the $P$, $S$, and $sync$
+operations of $\Gamma_{tsp}$ as explained hereafter.
+
+%----------------------------------------
+\subsection{Control components}
+%----------------------------------------
+The series-parallel tree decomposition $\Gamma_{tsp}$ represents the
+control of the dependencies of the simulation. As a result to be able
+to dump it to a component assembly it is needed to introduce what can
+be called \emph{control components}. Those components can be used for
+any case of simulation which increases code reuse between
+simulations. A control component is composed of a single
+\emph{provide} port linked to a single execution method, most of the
+time called the \emph{go} method. We introduce three types of control
+components, whole graphical representation is given in Figure~\ref{fig:ctrlcomponents}.
+
+\begin{figure}[t]
+\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
+   \node[component] (seq) at (0,0) {$SEQ$}; \node[provide] (p) at
+   (-1,0) {}; \node[use] (u) at (1,0) {$m$};
+
+  \path[-]
+    (p) edge node {} (seq)
+    (seq) edge node {} (u);
+\end{tikzpicture}
+\hspace{\fill}
+\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
+   \node[component] (seq) at (0,0) {$PAR$};
+   \node[provide] (p) at (-1,0) {};
+   \node[use] (u) at (1,0) {$m$};
+
+  \path[-]
+    (p) edge node {} (seq)
+    (seq) edge node {} (u);
+\end{tikzpicture}
+\hspace{\fill}
+\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
+   \node[component] (sync) at (0,0) {$SYNC$};
+   \node[provide] (p) at (-1,0) {};
+   \node[use] (u) at (1,0) {};
+
+  \path[-]
+    (p) edge node {} (sync)
+    (sync) edge node {} (u);
+\end{tikzpicture}
+\\
+a) Comp. SEQ
+\hspace{\fill}
+b) Comp. PAR
+\hspace{\fill}
+c) Comp. SYNC
+\\
+\caption{The three control components used by the back-end.}
+\label{fig:ctrlcomponents}
+\end{figure}
+
+\begin{description}
+\item[Sequence component (SEQ)] It is the direct representation of a sequence node of $\Gamma_{tsp}$.
+The role of this component is to call an ordered list of other components.
+Its interface contains an ordered \emph{use-multiple} port to be connected to the components to call in sequence.
+
+\item[Parallel component (PAR)] It is the direct representation of a parallel node of $\Gamma_{tsp}$.
+The role of this component is to call simultaneously a set of other components.
+It offers a \emph{use-multiple} port to be connected to the components to call in parallel.
+
+\item[Synchronization component (SYNC)].
+It is the direct representation to an update leaf of $\Gamma_{tsp}$.
+The role of this component is to call the synchronization of a given data.
+It offers an \emph{use} port to be connected to the data to update.
+
+\end{description}
 
@@
 %----------------------------------------
-From $\Gamma_{tsp}$ and using the control components described above, a direct dump can be done to a component assembly for an hybrid (data and task) parallelization of a simulation.
-However, it is also possible to dump to a data parallelization only.
-In such a case the computation of $\Gamma_{data}$ is enough to generate the component assembly.
-Actually a single SEQ component is thus needed and this component is linked to all computations and synchronizations of $\Gamma_{data}$ directly.
+From $\Gamma_{tsp}$ and using the control components described above,
+a direct dump can be done to a component assembly for an hybrid (data
+and task) parallelization of a simulation. However, it is also
+possible to dump to a data parallelization only. In such a case the
+computation of $\Gamma_{data}$ is enough to generate the component
+assembly. Actually a single SEQ component is thus needed and this
+component is linked to all computations and synchronizations of
+$\Gamma_{data}$ directly.
 
-As an example, from $\Gamma_{tsp}$ illustrated in Figure~\ref{fig:canon} a direct dump example is illustrated in Figure~\ref{fig:assembly}.
-In this Figure, the ports linked to data (use and use-multiple ports of SYNC and K) are represented but are not connected.
-Moreover, each computation component is an instance of the component K presented before but takes the identification name of the computation in the MSL file.
+Figure~\ref{fig:assembly} displays the assembly part corresponding to
+$\Gamma_{tsp}$ of Figure~\ref{fig:tsp}. In this figure, the ports
+linked to data (use and use-multiple ports of SYNC and K) are
+represented but are not connected. Moreover, each computation
+component is an instance of the component K presented before but takes
+the identification name of the computation in the MSL file.
 
-\begin{figure*}
+\begin{figure*}[t]
 \begin{center}
@@
 \end{tikzpicture}
-\caption{Direct dump to a component assembly.}
+\caption{Component assembly representing the computation part generated from $\Gamma_{tsp}$ of Figure~\ref{fig:tsp}.}
 \label{fig:assembly}
@@
 \end{figure*}
+
+%----------------------------------------

### 2015-07-31 WOLFHPC15/article/component.tex
revolving 2 fig:assembly ref

@@
 
-Figure~\ref{fig:assembly} has presented an overview of the component
+Figure~\ref{fig:mscac:assembly} has presented an overview of the component
 assembly that is executed. As discusses in Section~\ref{sect:mscac},
@@
 
-Figure~\ref{fig:assembly} displays the assembly part corresponding to
+Figure~\ref{fig:tsp:assembly} displays the assembly part corresponding to
 $\Gamma_{tsp}$ of Figure~\ref{fig:tsp}. In this figure, the ports
@@
 \caption{Component assembly representing the computation part generated from $\Gamma_{tsp}$ of Figure~\ref{fig:tsp}.}
-\label{fig:assembly}
+\label{fig:tsp:assembly}
 \end{center}

### 2015-07-31 WOLFHPC15/article/mscac.tex
revolving 2 fig:assembly ref

@@
 The back-end of the compiler builds a component assembly, that can be divided in two parts: the components implementing the actual computation and the static schedule, and the components needed to manage it, \ie the runtime support.
-A general view of this assembly is proposed in Figure~\ref{fig:assembly}; it is explained in details in Section~\ref{sect:component}.
+A general view of this assembly is proposed in Figure~\ref{fig:mscac:assembly}; it is explained in details in Section~\ref{sect:component}.
 As illustrated in the figure, each component is linked to one or more tuple of the multi-stencil program definition: \emph{DDS} component represents $\mathcal{M}$ and $\mathcal{E}$; \emph{Data} component represents $\Delta$ and $\mathcal{D}$; \emph{Time} component represents $T$, and \emph{Computations} component (which is generated by MSC) represents $\Gamma$.
@@
 \end{tikzpicture}
-\caption{Overall back-end component assembly}
-\label{fig:assembly}
+\caption{MSCAC Component Assembly Back-End Overview.}
+\label{fig:mscac:assembly}
 \end{center}

### 2015-07-31 WOLFHPC15/article/concept.tex

@@
 %----------------------------------------
-Let introduce some formal definitions used in the rest of the paper.
-For $\Delta$ the set of data of the simulation, a stencil kernel $s$ is defined as the quintuplet
+This section introduces some formal definitions used in the rest of the paper.
+Let $\Delta$ be the set of data of the simulation
+A stencil kernel $s$ is defined as the quintuplet
 \begin{equation} 
@@
 \end{equation}
-where $R$ is a set of pair $(r,n)$, with $r \in \Delta$ is a data read by the computation, and $n$ is a stencil shape (a neighborhood) applied on $r$.
-The data written by the kernel is denoted $w \in \Delta$, $exp$ is the numerical expression of the stencil kernel.
+where $R$ is a set of pairs $(r,n)$, with $r \in \Delta$ a data read by the computation, and $n$ the stencil shape (neighborhood) used to access $r$.
+The data written by the kernel is denoted $w \in \Delta$.
+$exp$ is the numerical expression of the stencil kernel.
 Finally the numerical expression is applied on the computation domain $d$.
 
-For example, in Figure~\ref{fig:ex1}, assuming the computation domain (full lines) is $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined with:
+For example, in Figure~\ref{fig:ex1}, assuming the computation domain (full lines) is $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined as:
 \begin{equation*}
-R: {(B,n1)} \quad w: A \quad d: dc1
+R: \{(B,n1)\} \quad w: A \quad d: dc1
 \end{equation*}
@@
 \end{equation*}
-On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil kernel can be defined with:
+On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil kernel can be defined as:
 \begin{equation*}
-R: {(C,n2),(A,null)} \quad w: A \quad d: dc2
+R: \{(C,n2),(A,local)\} \quad w: A \quad d: dc2
 \end{equation*}
@@
 \end{equation*}
-One can notice that the input data $A$ is associated to the stencil shape $null$. This means that no stencil shape is applied on $A$, but only local accesses.
+One can notice that the input data $A$ is associated to the stencil shape $local$.
+This means that the stencil shape applied on $A$ is limited to the element with the exact same coordinate as the result element.
 
-A local, or auxiliary, computation is a particular case of stencil kernel without neighborhood information in the numerical expression, which makes this kind of computation purely local.
-A local computation $l$ is defined as the quadruplet:
+The specificities of kernels whose stencil shapes are all local makes it possible to apply specific optimizations.
+We therefore provide a specific definition in this case.
+A local (or auxiliary) kernel $l$ is defined as the quadruplet:
 \begin{equation} 
@@
 \end{equation}
-where $R_l$ is the set of input data, and other parameters are the same than for a stencil computation.
+where $R_l$ is the set of input data, and other parameters are the same as for normal stencil kernels.
 

### 2015-07-31 WOLFHPC15/article/component.tex
fix all helene 30_07

@@
 %----------------------------------------
-\subsection{Overview}
+%\subsection{Overview}
 %----------------------------------------
-
 This section describes both the projection of $\Gamma_{tsp}$ to
 components, and the remaining needed components to execute a
-multi-stencil application. But, first the section gives an overview of
-component models, and it details \llc, the component model used in
-this work.
+multi-stencil application. Before those details the section gives an overview of
+component models, and introduce the concepts of the Low Level Component Model \llc, which is used in this paper.
 
@@
 %----------------------------------------
-Component model is an domain of software
+Component model is a domain of software
 engineering~\cite{Szyperski:2002:CSB:515228}, which promotes code
 re-use, separations of concerns, and thus maintainability. An
-application is made by a set of component; a component being a kind of
-black box that interacts only through ports: ports specify the
-services provided and required.
+application is made of a set of components; a component being a kind of
+black box representing an independent functionnality of the application, and which interacts only through its ports. A port specifies the services provided and required by the component.
 %
@@
 As this work makes use of \llc, in particular for the experiments, let
-describe it in more details.
+introduce concepts of this component model with more details.
 
@@
 service instances ($use-multiple$ ports). Services are \texttt{C++}
-interfaces. \llc also offer $MPI$ ports that enable components to
+interfaces. \llc also offers $MPI$ ports that enable components to
 share MPI communicators. Components can also have attribute ports to
@@
 As illustrated in Figures~\ref{fig:ports}, a $provide$ port is
-represented with a white circle on the left of the component, a $use$
+represented with a white circle, a $use$
 port with a black circle, a $use-multiple$ port by a black circle with
-a white $m$ in it, both on the right of the component. MPI port are
+a white $m$ in it. MPI port are
 connected with a black rectangle.
@@
 \begin{figure}[t]
-a)\hspace{2mm} \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+\begin{center}
+\subfloat[][\label{fig:2comp}]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
    \node[component] (C) at (0,0) {$C_0$};
@@
 \end{tikzpicture}
-\\\\
-b)\hspace{2mm}\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (C) at (0,0) {$C_0$};
+}
+\subfloat[][\label{fig:ass}]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (C) at (0,0) {$c_0$};
    \node[provide] (p) at (-1.5,0) {};
    \node[use] (u) at (1.5,0) {};
-   \node[component,right=1.5 of u] (C1) {$C_1$};
+   \node[component,right=1.5 of u] (C1) {$c_1$};
    \node[use,right=1.5 of C1] (um) {$m$};
@@
 \end{tikzpicture}
-\\\\
-c) \hspace{2mm}\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (C) at (0,0) {$C_2$};
+}
+\subfloat[][\label{fig:mpi}]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (C) at (0,0) {$c_2$};
    \node[provide] (p) at (-1.5,0) {};
    \node[mpi] (u) at (1.5,0) {};
-   \node[component,right=1.5 of u] (C1) {$C_3$};
+   \node[component,right=1.5 of u] (C1) {$c_3$};
 
@@
 \end{tikzpicture}
-\\\\
-\caption{Example components and ports representation.
-a) Component $C_0$ has a provide port ($p$) and a use port ($u$); Component $C_1$ has also a provide port ($p'$) but a use multiple port ($u'$).
+}
+\caption{Example of components and their ports representation.
+a) Component $c_0$ has a provide port ($p$) and a use port ($u$); Component $c_1$ has also a provide port ($q$) but also a use multiple port ($v$).
 b) A use port is connected to a (compatible) provide port.
-c) Component $C_2$ and $C_3$ shares an MPI communicator.}
+c) Component $c_2$ and $c_3$ shares an MPI communicator.}
 \label{fig:ports}
@@
 
-Figure~\ref{fig:mscac:assembly} has presented an overview of the component
+Figure~\ref{fig:mscac:assembly} gives an overview of the back-end component
 assembly that is executed. As discusses in Section~\ref{sect:mscac},
 there are different types of components. Components~\texttt{Driver},
-\texttt{DriverApp}, and \texttt{Time} are always present; they do not
-depend on the input but for the configuration of an attribute port of
-Component~\texttt{Time} with the number of iterations. There is also a
-single Component~\texttt{DDS} as in the current version only one mesh
-type is supported. All these components are provided with the
+\texttt{DriverApp}, and \texttt{Time} are always instanciated in the final component assembly; the input MSL file is not used by them except the component \texttt{Time} which uses the terminal \texttt{iteration} of the grammar.
+There is also a
+single Component~\texttt{DDS} (the current version handles a single mesh
+type). All these components are provided with the
 compiler; dumping this part of the assembly is straightforward.
@@
 
-The last part of the assembly is the computations. It contains actual
-computations in kernel components, denoted K, that have to be provided
+The last part of the assembly is the computation part. Each computation kernel is embedded in a kernel component, denoted K, which have to be written by the user and provided
 to the compiler. A kernel component encapsulates a computations. To
-access data, they are connected by the compiler to the adequate
-\texttt{Data} components; their data ports name must respect the names
+access data, a kernel component is connected, by the compiler, to the adequate
+\texttt{Data} components such that their ports name respect the names
 given in the \texttt{data} section of MSL. We denote the fact that they may have
-several use port by adding a star on the use port:
+several use port by adding a star on the use port, as illustrated in Figure~\ref{fig:k}.
+\begin{figure}
 \begin{center}
@@
 \end{center}
+\caption{Kernel component.}
+\label{fig:k}
+\end{figure}
 
@@
 
-The computation part of the assembly also contains the static schedule
-of the computations, \ie $\Gamma_{tsp}$. This is achieved by using
-three specific components for managing the $P$, $S$, and $sync$
+The rest of the computation part of the assembly contains the static parallel schedule
+computed by MSC, \ie $\Gamma_{tsp}$. This is achieved through the use of
+three specific components which manage the $P$, $S$, and $sync$
 operations of $\Gamma_{tsp}$ as explained hereafter.
@@
 %----------------------------------------
-\subsection{Control components}
+\subsection{Control Components}
 %----------------------------------------
@@
 time called the \emph{go} method. We introduce three types of control
-components, whole graphical representation is given in Figure~\ref{fig:ctrlcomponents}.
+components represented in Figure~\ref{fig:ctrlcomponents}.
 
@@
 The role of this component is to call the synchronization of a given data.
-It offers an \emph{use} port to be connected to the data to update.
+It offers a \emph{use} port to be connected to the data to update.
 
@@
 %----------------------------------------
-\subsection{Dump to a component assembly}
+\subsection{Dump To Component Assembly}
 %----------------------------------------
@@
 a direct dump can be done to a component assembly for an hybrid (data
-and task) parallelization of a simulation. However, it is also
+and task) parallelization of a simulation. 
+\fix{HC: ce qui suit est a changer avec la fusion du code de plusieurs composants}
+However, it is also
 possible to dump to a data parallelization only. In such a case the
-computation of $\Gamma_{data}$ is enough to generate the component
+computation of $\Gamma_{sync}$ is enough to generate the component
 assembly. Actually a single SEQ component is thus needed and this
 component is linked to all computations and synchronizations of
-$\Gamma_{data}$ directly.
+$\Gamma_{sync}$ directly.
 
@@
 represented but are not connected. Moreover, each computation
-component is an instance of the component K presented before but takes
+component is an instance of the component K presented before but using
 the identification name of the computation in the MSL file.

### 2015-07-31 WOLFHPC15/article/concept.tex
fix all helene 30_07

@@
 %----------------------------------------
-\subsection{Stencil codes}
+\subsection{Stencil Kernels}
 \label{sect:stencil}
@@
 In other words, in a mesh-based numerical simulation, the PDEs are transformed to a set of numerical computations applied at each time step on elements of the discretized space domain (the mesh).
-This paper focuses on one specific category of such numerical schemes based on \textit{stencil kernels}~\cite{spaaTangCKLL11}.
-This section introduces some definitions related to those.
+This paper focuses on one specific category of such numerical schemes based on \textit{stencil kernels}~\cite{spaaTangCKLL11}, also called in applied mathematics explicit numerical schemes.
+This section introduces some definitions related to stencil kernels.
 
@@
 It is a connected undirected graph without bridges (an edge is a bridge if its removal results in two disconnected graphs), where nodes and edges are linked to form cells (closure).
+Most of the time, nodes of a mesh are linked to coordinates in the real space domain.
 An example of structured mesh is illustrated in Figure~\ref{fig:mesh}.
@@
 For example, in Figure~\ref{fig:ex1}, data $A$ and $B$ are mapped onto the center of cells, while, in Figure~\ref{fig:ex2}, $C$ is mapped onto the edges.
-\todo[inline]{JB: structured/unstructured pas defini, pas de notion de coordonees des elements du maillage dans l'espace reel?}
+%\todo[inline]{JB: structured/unstructured pas defini, pas de notion de coordonees des elements du maillage dans l'espace reel?}
 
@@
 
-A \emph{stencil kernel} computes the value of one data or a subpart of it (the kernel \emph{computation domain}) using a \emph{numerical expression} taking as input on one or more data.
+A \emph{stencil kernel} computes the value of one data or a subpart of it (the kernel \emph{computation domain}) using a \emph{numerical expression} which takes as input one or more data.
 For example, the stencil kernel illustrated in Figure~\ref{fig:ex1} computes $A$ using $B$, while the one in Figure~\ref{fig:ex2} computes $A$ using $C$.
 Thus, a stencil kernel is defined by its numerical expression, a set of input data (only one in the examples), and its unique output data, the result.
-The computation domain is a subset of the mesh entities on which the output data is mapped.
-For example, in the first stencil, the computation of $A$ is performed on plain elements, while dotted elements are not computed.
+The \emph{computation domain} is a subset of the mesh entities on which the output data is mapped.
+For example, in the first stencil, the computation of $A$ is performed on elements represented with full lines, while dotted elements are not computed.
 On the other hand, on the second computation, the computation domain of $A$ contains all the mesh entities on which it is mapped.
@@
 
-The numerical expression of a stencil kernel has the particularity to compute each element of the result independently using only elements of the inputs whose coordinates are close to that of the result element.
-The elements accessed form a \emph{neighborhood} of the output element coordinate known as the \emph{stencil shape}.
+The numerical expression of a stencil kernel has the particularity to compute each element of the result independently using some elements of the inputs in a given neighborhood.
+Accessed elements form a \emph{neighborhood} of the output known as the \emph{stencil shape}.
 For example, the stencil shape of the first computation in Figure~\ref{fig:ex1} contains direct neighbors on the right, left, top and bottom.
@@
 
-As can be seen from these definitions, stencil kernels implementations share many properties from an algorithmic point of view that can be used to apply well known optimization and parallelization strategies.
+As it can be seen from those definitions, stencil kernels implementations share many properties from an algorithmic point of view that can be used to apply well known optimization and parallelization strategies.
 Many solutions (languages or libraries) thus propose to ease their programming by producing optimized and parallelized code from a simple description of the local computation to apply on each element.
 For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, respectively proposing an optimized cache tiling technique, and a separation of concerns between parallelization strategies and computation descriptions.
-\todo[inline]{JB: a virer?}
+\todo[inline]{JB et HC: etat de l'art a mettre ailleurs ?}
 
-To resume, concepts used in this paper to define a stencil kernel are \emph{mesh}, \emph{mesh entities}, \emph{data}, \emph{computation domain}, \emph{numerical expression} and \emph{stencil shape}. 
+%To resume, concepts used in this paper to define a stencil kernel are \emph{mesh}, \emph{mesh entities}, \emph{data}, \emph{computation domain}, \emph{numerical expression} and \emph{stencil shape}. 
 If stencil kernels have been studied a lot, the formalization of real overall mesh-based numerical simulations is a rather new contribution.

### 2015-07-31 WOLFHPC15/article/evaluation.tex
fix evaluation helene 30_07
fix all helene 30_07

@@
 
+\fix{HC 2 all : Je ne sais pas si ce qui suit a sa place dans cet article.}
+
 One can notice that the transformation of $\Gamma_{dep}$ to a minimal series-parallel graph is the longest step of MSC, because of the removal of the forbidden shapes in the graph.
 Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted, because the algorithm use a general solution instead of finding each forbidden shape, but it seems that many of them appeared.
-The forbidden shape, called \emph{N-shape}, is represented in Figure~\ref{fig:n}.
-This shape is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.
-To remove a \emph{N-shape} an additionnal dependency constraint is added as illustrated in Figure~\ref{fig:over}.
-
-\begin{figure}[h!]
-\begin{center}
-\subfloat[][The forbidden $N$ shape.\label{fig:n}]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (a) at (0,0) {$c_0$};
-   \node[] (b) at (2,0) {$c_1$};
-   \node[] (c) at (0,-1) {$c_2$};
-   \node[] (d) at (2,-1) {$c_3$};
-
-  \path[->]
-    (a) edge node {} (b)
-    (c) edge node {} (b)
-        edge node {} (d);
-  \end{tikzpicture}
-  }
-  \hspace{25pt}
-  \subfloat[][Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
-	\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (c0) at (0,0) {$c_0$};
-   \node[] (bc0) at (-1,0) {};
-   \node[] (c1) at (2,0) {$c_1$};
-   \node[] (c2) at (0,-1) {$c_2$};
-   \node[] (bc2) at (-1,-1) {};
-   \node[] (c3) at (2,-1) {$c_3$};
-
-  \path[->]
-    (bc0) edge [dotted] node {} (c0)
-    (bc2) edge [dotted] node {} (c2)
-    (c0) edge node {} (c1)
-          edge [dashed] node [swap] {} (c3)
-    (c2) edge node {} (c1)
-        edge node {} (c3);
-  \end{tikzpicture}
-}
-  \caption{Forbidden $N$ subgraph shape for a DAG to be minimal series-parallel.}
-  \label{fig:forbidden}
-\end{center}
-\end{figure}
-
+The \emph{N-shape}, represented in Figure~\ref{fig:n} is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.
 The fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static shedule of tasks may not be the best solution for complex simulations.

### 2015-07-31 WOLFHPC15/article/introduction.tex
fix all helene 30_07

--- 2015-07-30 WOLFHPC15/article/introduction.tex
@@
 For this reason, solutions to ease the design and implementation of DSLs have recently appeared~\cite{Fernandez:2014:DFL:2691166.2691168}.
-The main idea of those solutions is to propose a single DSL to write new DSLs.
+The main idea of those solutions is to propose a single DSL or a framework to write new DSLs.
 In this paper, we study a first contribution to another approach to improve maintainability and portability in DSLs.
@@
 As it will be detailed in the related work, this is complementary to the optimization and parallelization of a single stencil kernel.
-%\fix{JB: j'utiliserais bien le terme "kernel" ou "stencil kernel" pour ce que tu apelles "stencil" et "stencil" tout court pour ce que tu apelles "stencil shape"}
 
 The rest of this paper is organized as follows.
-First, the concepts of \emph{stencil} and \emph{multi-stencil programs} are formalized in Section~\ref{sect:concept}.
+First, the concepts of \emph{stencil kernels} and \emph{multi-stencil programs} are formalized in Section~\ref{sect:concept}.
 Section~\ref{sect:mscac} gives an overview of the solution presented in this paper while Section~\ref{sect:msmsc} details the MSL language and its compiler.

### 2015-07-31 WOLFHPC15/article/ms_msc.tex
fix all helene 30_07

@@
 Figure~\ref{fig:grammar} shows the grammar of MSL.
-Lines 1 to 9 define what is a MSL program with the different parts memntionned above.
+Lines 1 to 9 define what is a MSL program with the different parts mentionned above.
 The remaining of this section describes those different parts of a MSL program and presents an example.
-We can note that, as MSL is a descriptive language, almost all terminals $meshid$, $meshdom$, $compdomid$, $dataid$, $compid$ and $neighborid$, are string identifiers.
-Only $num$ is an integer terminal.
+We can note that, as MSL is a descriptive language, almost all terminals \texttt{meshid}, \texttt{meshdom}, \texttt{compdomid}, \texttt{dataid}, \texttt{compid} and \texttt{neighborid}, are string identifiers.
+Only \texttt{num} is an integer terminal.
 
@@
 Then, mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
-In the example given in Figure~\ref{fig:mslex} a Cartesian mesh, called $cart$, and three mesh entities, called $cell$, $edgex$ and $edgey$, are defined.
+In the example given in Figure~\ref{fig:mslex} a Cartesian mesh, called \texttt{cart}, and three mesh entities, called \texttt{cell}, \texttt{edgex} and \texttt{edgey}, are defined.
 
@@
 By default two computation domains always intersect except when an independency relation is mentionned in the section \emph{independent} of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
-In the example of Figure~\ref{fig:mslex}, on lines 4 to 6, a whole computation domain is defined for each mesh entity, called $allcell$, $alledgex$ and $alledgey$, and on lines 7 and 8 two subparts of the mesh entity $edgex$ are also defined, called $part1edgex$ and $part2edgex$.
-Finally, line 10 means that $part1edgex$ and $part2edgex$ do not intersect.
+In the example of Figure~\ref{fig:mslex}, on lines 4 to 6, a whole computation domain is defined for each mesh entity, called \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}, and on lines 7 and 8 two subparts of the mesh entity \texttt{edgex} are also defined, called \texttt{part1edgex} and \texttt{part2edgex}.
+Finally, line 10 means that \texttt{part1edgex} and \texttt{part2edgex} do not intersect.
 
@@
 A data is a quantity to simulate and is mapped onto a mesh entity, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}.
-The $time$ section simply indicates a number of iterations to perform in the simulation.
+The \texttt{time} section simply indicates a number of iterations to perform in the simulation.
 In the example of Figure~\ref{fig:mslex} are defined ten data and 500 iterations (lines 11-22).
@@
 However, as MSL does not handle numerical expressions, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} is not used in the language.
-The exact code of the computation as to be writtenin into the final ready-to-fill parallel pattern of the simulation, using external languages or interfaces (\cf Section~\ref{sect:component}).
+The exact code of the computation as to be written into the final ready-to-fill parallel pattern of the simulation, using external languages or interfaces (\cf Section~\ref{sect:component}).
 Moreover, to propose an intuitive syntax, information are given in a different order than in formal definitions.
@@
 In the example of Figure~\ref{fig:mslex}, nine computations are described.
-For example, on line 24, the data $b$ is computed on the computation domain $allcell$ by the computation $c_0$ which read data $a$ without neighborhood shape.
+For example, on line 24, the data \texttt{b} is computed on the computation domain \texttt{allcell} by the computation \texttt{c_0} which read data \texttt{a} without neighborhood shape.
 
@@
 The MSC parallelization, which is a subpart of the overall compiler MSCAC, makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
-This transformation phase of the compilation is divided in five different steps; it transforms $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
+This parallelization phase of the compilation is divided in five different steps; it transforms $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
 This transformation phase handles data parallelism through the first step, and task parallelism through the three next steps.
-This section gives an overview of the four steps of the MSC transformation; the detailed formalism and the associated algorithms of those steps are described in the research report~\cite{????}.
+This section gives an overview of the five steps of the MSC transformation; the detailed formalism and the associated algorithms of those steps are described in the research report~\cite{????}.
 
@@
 To do so the dependency graph is transformed to a minimal series-parallel graph~\cite{Valdes:1979:RSP:800135.804393}.
-As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape}, is not found in the dag.
-Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes}.
+As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape} illustrated in Figure~\ref{fig:n}, is not found in the dag.
+Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes} by an over-constraint, as illustrated in Figure~\ref{fg:over}.
+
+\begin{figure}[h!]
+\begin{center}
+\subfloat[][The forbidden $N$ shape.\label{fig:n}]{
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (a) at (0,0) {$c_0$};
+   \node[] (b) at (2,0) {$c_1$};
+   \node[] (c) at (0,-1) {$c_2$};
+   \node[] (d) at (2,-1) {$c_3$};
+
+  \path[->]
+    (a) edge node {} (b)
+    (c) edge node {} (b)
+        edge node {} (d);
+  \end{tikzpicture}
+  }
+  \hspace{25pt}
+  \subfloat[][Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
+  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (c0) at (0,0) {$c_0$};
+   \node[] (bc0) at (-1,0) {};
+   \node[] (c1) at (2,0) {$c_1$};
+   \node[] (c2) at (0,-1) {$c_2$};
+   \node[] (bc2) at (-1,-1) {};
+   \node[] (c3) at (2,-1) {$c_3$};
+
+  \path[->]
+    (bc0) edge [dotted] node {} (c0)
+    (bc2) edge [dotted] node {} (c2)
+    (c0) edge node {} (c1)
+          edge [dashed] node [swap] {} (c3)
+    (c2) edge node {} (c1)
+        edge node {} (c3);
+  \end{tikzpicture}
+}
+  \caption{Forbidden $N$ subgraph shape for a DAG to be minimal series-parallel.}
+  \label{fig:forbidden}
+\end{center}
+\end{figure}
 

### 2015-08-01 WOLFHPC15/article/mscac.tex
Compilation + sec 3

@@
-The solution proposed in this paper contains a domain specific language to describe a multi-stencil program, \ie an overall numerical simulation, and a compiler which successively parses the front-end language, transforms the internal representation, and finally dumps the representation to a component-based back-end code.
+The solution proposed in this paper is based on a domain specific language to describe a multi-stencil program, \ie a complete numerical simulation, and a compiler which successively parses the front-end language, transforms the internal representation, and finally dumps the representation to a component-based back-end code.
 
@@
 %Numerical computations are encapsulated in specific components that users have to develop.
-In a MSL description, six main sections are defined, as the six tuples of the definition of a multi-stencil program of Equation~\ref{eq:msp}.
+A MSL description is made of six main sections that match the six tuples elements of the definition of a multi-stencil program of Equation~\ref{eq:msp}.
+\todo[inline]{JB: dans une grammaire, les identifiants sont toujours des chaines de caracteres, je ne comprends pas}
 
@@
 As illustrated in Figure~\ref{fig:mscac}, the compiler, named MSCAC for \emph{Multi-Stencil Component Assembly Compiler}, is composed of three different phases.
-The first phase of the compiler is to \emph{parse} the MSL description.
-Second, the compiler analyzes and transforms the computation part of the description $\Gamma$, to an intermediate representation.
-This phase of compilation, called MSC, is responsible for the automatic parallelization of the overall numerical simulation.
-Finally, the parallel intermediate representation is dumped to an actual implementation of the parallel pattern of the simulation.
+The first phase \emph{parses} the MSL description into an intermediate representation (IR).
+The second phase called MSC transforms the computation part ($\Gamma$) of the IR in order to automatically introduce parallelism of the overall numerical simulation.
+Finally, the third phase dumps the IR to an actual implementation of the parallel pattern of the simulation.
 
@@
 %In this paper are presented the language MSL, the MSC compilation phase, and the dump compilation phase. 
-The parallelization techniques handled by MSC are \emph{data parallelism}, where data are splitted among processors or cores while applying a single program everywhere, and \emph{task parallelism} where the program is transformed into a dependency graph of tasks.
+The parallelization techniques handled by MSC are \emph{data parallelism}, where data is split among processors or cores while applying the same computation on every piece of data, and \emph{task parallelism} where the program is transformed into a dependency graph of tasks.
 From a dependency graph, different strategies can be used to schedule tasks, in other words to control their execution in a valid order.
@@
 
-The final phase of compilation, the dump, uses the static schedule of the tasks, created from $\Gamma$, and additionnal information on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{E}$ and $\mathcal{D}$, given by the \emph{parser}, to produce a back-end solution which represents the parallel pattern or skeleton of the overall simulation.
-We could have propose a dump to OpenMP~\cite{660313}, however this paper presents a first contribution to dump a DSL to a component-based runtime.
-Component models, where a component represents a clear and independent functionnality of the application, have first been proposed in software engineering and have proved many times good properties for code re-use, separation of concerns, maintainability and productivity of codes.
+The final phase of compilation, the dump, uses the static schedule of the tasks, created from $\Gamma$, and additional informations on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{E}$ and $\mathcal{D}$, given by the \emph{parser}, to produce a back-end solution which represents the parallel pattern or skeleton of the overall simulation.
+We could have proposed a dump to OpenMP~\cite{660313}, however this paper presents a first contribution to dump a DSL to a component-based runtime.
+Component models, where a component represents a clear and independent functionality of the application, have first been proposed in software engineering and have proved many times good properties for code re-use, separation of concerns, maintainability and productivity of codes.
 Those properties, on the other hand, are recurrent stated problems of parallel applications.

### 2015-08-02 WOLFHPC15/article/ms_msc.tex
Forgotten files

@@
 In the example of Figure~\ref{fig:mslex}, nine computations are described.
-For example, on line 24, the data \texttt{b} is computed on the computation domain \texttt{allcell} by the computation \texttt{c_0} which read data \texttt{a} without neighborhood shape.
+For example, on line 24, the data \texttt{b} is computed on the computation domain \texttt{allcell} by the computation \texttt{c\_0} which read data \texttt{a} without neighborhood shape.
 
@@
 As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape} illustrated in Figure~\ref{fig:n}, is not found in the dag.
-Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes} by an over-constraint, as illustrated in Figure~\ref{fg:over}.
+Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes} by an over-constraint, as illustrated in Figure~\ref{fig:over}.
 

### 2015-08-02 WOLFHPC15/article/mscac.tex
Forgotten files

@@
-The solution proposed in this paper is based on a domain specific language to describe a multi-stencil program, \ie a complete numerical simulation, and a compiler which successively parses the front-end language, transforms the internal representation, and finally dumps the representation to a component-based back-end code.
+The solution proposed in this paper is based on \textbf{a)} a domain specific language to describe a multi-stencil program, \ie a complete numerical simulation, and \textbf{b)} a compiler which successively parses the language, applies transformations for automatic parallelization, and finally dumps the representation to a component-based code.
 
@@
 The first phase \emph{parses} the MSL description into an intermediate representation (IR).
-The second phase called MSC transforms the computation part ($\Gamma$) of the IR in order to automatically introduce parallelism of the overall numerical simulation.
+The second phase called MSC transforms the computation part ($\Gamma$) of the IR in order to automatically introduce parallelism in the overall numerical simulation.
 Finally, the third phase dumps the IR to an actual implementation of the parallel pattern of the simulation.
@@
 %In this paper are presented the language MSL, the MSC compilation phase, and the dump compilation phase. 
-The parallelization techniques handled by MSC are \emph{data parallelism}, where data is split among processors or cores while applying the same computation on every piece of data, and \emph{task parallelism} where the program is transformed into a dependency graph of tasks.
-From a dependency graph, different strategies can be used to schedule tasks, in other words to control their execution in a valid order.
+The parallelization techniques handled by MSC are \emph{data parallelism}, where data is split among processors or cores while applying the same computation on every piece of data, and \emph{task parallelism} where the program is transformed into a graph of tasks with dependencies.
+Different strategies can be used to schedule the tasks of this graph; in other words to plan their execution in a valid order.
 The MSC compiler presented in this paper handles the transformation of the dependency graph to a static scheduling.
 
-The final phase of compilation, the dump, uses the static schedule of the tasks, created from $\Gamma$, and additional informations on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{E}$ and $\mathcal{D}$, given by the \emph{parser}, to produce a back-end solution which represents the parallel pattern or skeleton of the overall simulation.
-We could have proposed a dump to OpenMP~\cite{660313}, however this paper presents a first contribution to dump a DSL to a component-based runtime.
-Component models, where a component represents a clear and independent functionality of the application, have first been proposed in software engineering and have proved many times good properties for code re-use, separation of concerns, maintainability and productivity of codes.
-Those properties, on the other hand, are recurrent stated problems of parallel applications.
+The final phase of compilation, the dump, uses the static scheduling of the tasks, created from $\Gamma$, and additional informations on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{E}$ and $\mathcal{D}$.
+It generates a program in the back-end language which embeds the parallel pattern or skeleton of the overall simulation.
+Various models could be used as back-end, such as OpenMP~\cite{660313} for example, however this paper adopts a novel approach where the dump targets a component-based runtime.
+This approach makes it possible to leverage good properties of software component models such as increased code re-use, eased separation of concerns, and increased maintainability and productivity.
+Those properties counter some known limitations of DSLs. % JB: was "parallel applications".
+%where a component represents a clear and independent functionality of the application, have first been proposed in software engineering and have proved many times good properties for
+
+\todo[inline]{JB: je déplacerais les 2 derniers paragraphes et la figure dans la sec.~\ref{sect:component}}
 

### 2015-08-02 WOLFHPC15/article/ms_msc.tex
4.1

@@
 
-As mentionned in the previous section, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations.
-Six main sections are required in a MSL description, as the six tuples of the formal definition of a multi-stencil program: the mesh description; the mesh entities description; the computation domains description and their dependencies; the data description; the time loop description; and finally, the computations description.
+As mentioned in the previous section, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations.
+Six main sections are required in a MSL description that match the six tuples of the formal definition of a multi-stencil program: \textbf{1.}\,the mesh description, \textbf{2.}\,the mesh entities description, \textbf{3.}\,the computation domains description and their dependencies, \textbf{5.}\,the data description, \textbf{5.}\,the time loop description, and finally \textbf{6.} the computations description.
 Figure~\ref{fig:grammar} shows the grammar of MSL.
-Lines 1 to 9 define what is a MSL program with the different parts mentionned above.
+Lines 1 to 9 define what is a MSL program with the different parts mentioned above.
 The remaining of this section describes those different parts of a MSL program and presents an example.
@@
 \subsubsection*{Mesh and Mesh Entities}
-As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is directly defined by a string identifier.
+As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is simply defined by a string identifier.
 Then, mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
-In the example given in Figure~\ref{fig:mslex} a Cartesian mesh, called \texttt{cart}, and three mesh entities, called \texttt{cell}, \texttt{edgex} and \texttt{edgey}, are defined.
+Figure~\ref{fig:mslex} presents an example defining a Cartesian mesh, called \texttt{cart}, and three mesh entities, called \texttt{cell}, \texttt{edgex} and \texttt{edgey}.
 
 \subsubsection*{Computation Domains and Their Dependencies}
-A computation domain can be seen as a subpart of a mesh entity which will be used during one or more computation.
+A computation domain can be seen as a subpart of entities of a mesh which will be used during one or more computation.
 For each computation domain, the mesh entity from which it is built is indicated (lines 3-4 and 12-14 in Figure~\ref{fig:grammar}).
-By default two computation domains always intersect except when an independency relation is mentionned in the section \emph{independent} of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
-In the example of Figure~\ref{fig:mslex}, on lines 4 to 6, a whole computation domain is defined for each mesh entity, called \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}, and on lines 7 and 8 two subparts of the mesh entity \texttt{edgex} are also defined, called \texttt{part1edgex} and \texttt{part2edgex}.
-Finally, line 10 means that \texttt{part1edgex} and \texttt{part2edgex} do not intersect.
+By default two computation domains always intersect except when an independence relation is mentioned in the section \emph{independent} of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
+In the lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each mesh entity, called \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they represent the whole domain.
+On lines 7 and 8 two other subparts of the mesh entity \texttt{edgex} are defined, called \texttt{part1edgex} and \texttt{part2edgex}; they represent the boundaries of the domain.
+On line 10 the independence relation between \texttt{part1edgex} and \texttt{part2edgex} is specified since they do not intersect.
 
 \subsubsection*{Data and Time}
-A data is a quantity to simulate and is mapped onto a mesh entity, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}.
+A data is a quantity to simulate.
+It is mapped on mesh entities, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}.
 The \texttt{time} section simply indicates a number of iterations to perform in the simulation.
-In the example of Figure~\ref{fig:mslex} are defined ten data and 500 iterations (lines 11-22).
-An extension could be to support a convergence criterion.
+In the example of Figure~\ref{fig:mslex} ten data identifiers are defined and 500 iterations (lines 11-22).
+%An extension could be to support a convergence criterion.
 
 \subsubsection*{Computation Description}
-Finally, the language follows the two definitions of a stencil kernel and a local computation given in Section~\ref{sect:multistencil}.
-However, as MSL does not handle numerical expressions, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} is not used in the language.
-The exact code of the computation as to be written into the final ready-to-fill parallel pattern of the simulation, using external languages or interfaces (\cf Section~\ref{sect:component}).
-Moreover, to propose an intuitive syntax, information are given in a different order than in formal definitions.
-Thus, the description of a computation, as indicated in lines 24 and 25 of Figure~\ref{fig:grammar}, consists in first the identifier of the data to compute ($w$ in the formal description), in brackets the identifier of the computation domain on which the computation is applied ($d$ in the formal definition), an identifier for the computation kernel, and the set of identifiers for the data read by the computation, with parenthesis.
-For each data read by the computation brackets indicate the stencil shape identifier used on the data ($R$ in the formal definition of a stencil).
-If the computation is local no brackets appear ($R_l$ in the formal definition).
+The last part of the language contains the specifications of stencil kernels and local computations as defined in Section~\ref{sect:multistencil}.
+This description does not contain a direct expression of the numerical expressions.
+Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} simply corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
+The specification starts by the identifier of the data to compute ($w$ in the formal description) with the computation domain between brackets ($d$ in the formal definition).
+Then, after the equal sign is specified the kernel identifier.
+Finally, between the parenthesis is the list of identifiers of data read by the computation with their stencil shape between brackets ($R$ in the formal definition).
+If the computation is local, no brackets appear ($R_l$ in the formal definition).
 In the example of Figure~\ref{fig:mslex}, nine computations are described.
-For example, on line 24, the data \texttt{b} is computed on the computation domain \texttt{allcell} by the computation \texttt{c\_0} which read data \texttt{a} without neighborhood shape.
+On line 24, the data \texttt{b} is computed on the computation domain \texttt{allcell} by the computation \texttt{c\_0} which read data \texttt{a} without neighborhood shape.
 

### 2015-08-02 WOLFHPC15/article/ms_msc.tex
4.2

@@
 %----------------------------------------
-The MSC parallelization, which is a subpart of the overall compiler MSCAC, makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
-This parallelization phase of the compilation is divided in five different steps; it transforms $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
-This transformation phase handles data parallelism through the first step, and task parallelism through the three next steps.
-This section gives an overview of the five steps of the MSC transformation; the detailed formalism and the associated algorithms of those steps are described in the research report~\cite{????}.
+The MSC parallelization is a subpart of the overall compiler MSCAC.
+It makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
+This parallelization phase is divided in five different steps to transform $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
+Data parallelism is handled in the first two step, and task parallelism in the following three steps. This section gives an overview of the five steps of the MSC transformation.
+The detailed formalism of these steps and the associated algorithms are described in a research report~\cite{????}.
 
@@
 \subsubsection*{The Synchronized Ordered List of Computations: $\Gamma_{sync}$}
-Considering a data parallelization of the simulation, the mesh and data of the simulation are splitted among available processors (or cores) and a single program is applied on each sub-part with additionnal synchronizations between processors.
-For example, as a stencil kernel accesses neighborhood values, values computed by another processor are needed and synchronized.
-From the ordered list of computations $\Gamma$, the needed synchronizations can be automatically added.
+In a data parallelization, data is split among processors and the same program is applied on each sub-part of the data.
+This does however require additional synchronizations between processors.
+For example, as a stencil kernel accesses neighbor values, values computed by another processor are needed and must be synchronized. 
+The required synchronizations can be automatically computed from the ordered list of computations $\Gamma$.
 A synchronization is needed each time a data read by a stencil computation has been written by a previous computation.
 In such a case, a computation is added before the stencil computation.
-This \emph{synchronization computation} read the data to synchronize, and write the same data.
+This \emph{synchronization computation} reads the data to synchronize, and write the same data.
 The computation domain of such a synchronization computation is the mesh entity on which the data is declared.
@@
 A dependency exists between two computations (including synchronizations) if and only if a data read has been written by a previous computation in $\Gamma_{sync}$, and if the computation domains intersect.
-In other words, the only case where a data dependency exists but is not taken into account is when the computation domains do not intersect.
+In other words, the only case where the same data is manipulated by distinct kernels without inducing a dependency is when the computation domains do not intersect.
 Nodes of the dependency graph represent computations of $\Gamma_{sync}$, while edges are dependencies between them.
-The dependency graph $\Gamma_{dep}$ is a directed acyclic graph (\emph{dag}).
-For example, the dependency dag $\Gamma_{dep}$ of the example of Figure~\ref{fig:mslex} is displayed in Figure~\ref{fig:hyb}.
+The dependency graph $\Gamma_{dep}$ is a directed acyclic graph (\emph{DAG}).
+For example, the dependency DAG $\Gamma_{dep}$ of the example of Figure~\ref{fig:mslex} is presented in Figure~\ref{fig:hyb}.
 
@@
 To do so the dependency graph is transformed to a minimal series-parallel graph~\cite{Valdes:1979:RSP:800135.804393}.
-As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a dag is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape} illustrated in Figure~\ref{fig:n}, is not found in the dag.
+As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a DAG is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape} illustrated in Figure~\ref{fig:n}, is not found in the DAG.
 Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes} by an over-constraint, as illustrated in Figure~\ref{fig:over}.

### 2015-08-03 WOLFHPC15/article/component.tex
Forgotten files
5.1,5.2

@@
 %----------------------------------------
-This section describes both the projection of $\Gamma_{tsp}$ to
-components, and the remaining needed components to execute a
-multi-stencil application. Before those details the section gives an overview of
-component models, and introduce the concepts of the Low Level Component Model \llc, which is used in this paper.
+This section describes the projection of a MSP program to components, including the components used to implement the static scheduling of $\Gamma_{tsp}$.
+Before those details the section gives an overview of component models, and introduce the Low Level Components (\llc), the model used in this paper.
 
 %----------------------------------------
-\subsection{Component Model Overview}
+\subsection{Component-Based Software Engineering}
 %----------------------------------------
-Component model is a domain of software
-engineering~\cite{Szyperski:2002:CSB:515228}, which promotes code
-re-use, separations of concerns, and thus maintainability. An
-application is made of a set of components; a component being a kind of
-black box representing an independent functionnality of the application, and which interacts only through its ports. A port specifies the services provided and required by the component.
-%
+Component-based software engineering (CBSE) is a domain of software engineering~\cite{Szyperski:2002:CSB:515228} which promotes code
+re-use, separations of concerns, and thus maintainability.
+An application is made of a set of component instances.
+A component is a black box that implements an independent functionality of the application, and which interacts with its environment only through well defined interfaces: its ports.
+A port can for example specify services provided or required by the component.
 With respect to high performance computing, some works have also shown
@@
 
-Many component models exist, each of them with its own specifications
-and singularities. In distributed computing, well known component
-models are CCM~\cite{corba:omg06} (CORBA Component Model), and
-GCM~\cite{Baude} (Grid Component Model); in HPC, there are
-CCA~\cite{Bernholdt01052006} (Common Component Architecture), and
-\llc~\cite{l2c} (Low Level Component), for example.
-%
-As this work makes use of \llc, in particular for the experiments, let
-introduce concepts of this component model with more details.
+Many component models exist, each of them with its own specificities.
+Well known component models include for example the CORBA Component Model (CCM)~\cite{corba:omg06}, and the Grid Component Model (GCM)~\cite{Baude} in distributed computing, the Common Component Architecture (CCA)~\cite{Bernholdt01052006}, and Low Level Components (\llc)~\cite{l2c} in HPC.
+This work makes use of \llc for the experiments.
+Let us therefore introduce the concepts of this model in more details.
 
@@
 \llc is a minimalist \texttt{C++} based HPC-oriented component model
-where a component extends the concept of class by specifying in its
-interfaces the services that it offers ($provide$ ports) and that it
-needs, either a single service instance ($use$ ports), or multiple
-service instances ($use-multiple$ ports). Services are \texttt{C++}
-interfaces. \llc also offers $MPI$ ports that enable components to
-share MPI communicators. Components can also have attribute ports to
-be configured.
+where a component extends the concept of class.
+The services offered by the components are specified trough $provide$ ports,
+those used either by $use$ ports for a single service instance,
+or $use-multiple$ ports for multiple service instances.
+Services are specified as \texttt{C++} interfaces.
+\llc also offers $MPI$ ports that enable components to share MPI communicators.
+Components can also have attribute ports to be configured.
 %
 As illustrated in Figures~\ref{fig:ports}, a $provide$ port is
-represented with a white circle, a $use$
+represented by a white circle, a $use$
 port with a black circle, a $use-multiple$ port by a black circle with
@@
 
-A classical \llc-based application is a static \emph{assembly} of
-components made of instances of components and of connections between
-component ports. Such an assembly is described in LAD, an XML
-dialect~\cite{l2c}.
+A \llc-based application is a static \emph{assembly} of components instances and connections between their ports.
+Such an assembly is described in LAD, an XML dialect~\cite{l2c}.
 
@@
 \label{fig:ports}
+\end{center}
 \end{figure}

### 2015-08-03 WOLFHPC15/article/component.tex
5.4,5.5

@@
 connected with a black rectangle.
+
+\todo[inline]{JB: chez moi la figure\,\ref{fig:ports} deborde...}
 
@@
 
-The last part of the assembly is the computation part. Each computation kernel is embedded in a kernel component, denoted K, which have to be written by the user and provided
-to the compiler. A kernel component encapsulates a computations. To
+The last part of the assembly is the computation part.
+Each computation kernel is embedded in a different kernel component which has to be written by the user and provided
+to the compiler.
+We denote this kind of components K.
+A kernel component encapsulates a computations. To
 access data, a kernel component is connected, by the compiler, to the adequate
@@
 %----------------------------------------
-The series-parallel tree decomposition $\Gamma_{tsp}$ represents the
-control of the dependencies of the simulation. As a result to be able
-to dump it to a component assembly it is needed to introduce what can
-be called \emph{control components}. Those components can be used for
-any case of simulation which increases code reuse between
-simulations. A control component is composed of a single
-\emph{provide} port linked to a single execution method, most of the
-time called the \emph{go} method. We introduce three types of control
-components represented in Figure~\ref{fig:ctrlcomponents}.
+The series-parallel tree decomposition $\Gamma_{tsp}$ represents the static scheduling of the simulation.
+It is dumped to components by generating an assembly that exactly matches the tree.
+We introduce what we call \emph{control components} to represent the nodes of the tree.
+These components can be used for any simulation, which increases code reuse between simulations.
+A control component exposes a single \emph{provide} port containing a single method, typically a control-only method without any parameter.
+We introduce three control components represented in Figure~\ref{fig:ctrlcomponents}.
 
@@
 \item[Sequence component (SEQ)] It is the direct representation of a sequence node of $\Gamma_{tsp}$.
-The role of this component is to call an ordered list of other components.
-Its interface contains an ordered \emph{use-multiple} port to be connected to the components to call in sequence.
+This component sequentially calls an ordered list of other components.
+It exposes an ordered \emph{use-multiple} port to be connected to the components to call in sequence.
 
 \item[Parallel component (PAR)] It is the direct representation of a parallel node of $\Gamma_{tsp}$.
-The role of this component is to call simultaneously a set of other components.
-It offers a \emph{use-multiple} port to be connected to the components to call in parallel.
+This component simultaneously calls a set of other components.
+It exposes a \emph{use-multiple} port to be connected to the components to call in parallel.
 
@@
 It is the direct representation to an update leaf of $\Gamma_{tsp}$.
-The role of this component is to call the synchronization of a given data.
-It offers a \emph{use} port to be connected to the data to update.
+This component calls the synchronization of a given data.
+It exposes a \emph{use} port to be connected to the data to update.
 
@@
 represented but are not connected. Moreover, each computation
-component is an instance of the component K presented before but using
+component is an instance of a specific K component using
 the identification name of the computation in the MSL file.
@@
    %c0
-   \node[component] (c0) [right = 2cm of seq0u] {$c_0$};
+   \node[component] (c0) [right = 2cm of seq0u] {$K(c_0)$};
    \node[use] (c0u) [right = 0.8cm of c0] {};
@@
    %c1
-   \node[component] (c1) [below = 0.8cm of sync0] {$c_1^*$};
+   \node[component] (c1) [below = 0.8cm of sync0] {$K(c_1^*)$};
    \node[use] (c1u) [right = 0.8cm of c1] {};
@@
   %c2
-   \node[component] (c2) [right = 1cm of par1u] {$c_2$};
+   \node[component] (c2) [right = 1cm of par1u] {$K(c_2)$};
    \node[use] (c2u) [right = 0.8cm of c2] {};
@@
   %c4
-   \node[component] (c4) [below = 0.8cm of par1] {$c_4^*$};
+   \node[component] (c4) [below = 0.8cm of par1] {$K(c_4^*)$};
    \node[use] (c4u) [right = 0.8cm of c4] {};
@@
   %c6
-   \node[component] (c6) [below = 0.8cm of c4] {$c_6$};
+   \node[component] (c6) [below = 0.8cm of c4] {$K(c_6)$};
    \node[use] (c6u) [right = 0.8cm of c6] {};
@@
   %c3
-   \node[component] (c3) [right = 1cm of seq2u] {$c_3$};
+   \node[component] (c3) [right = 1cm of seq2u] {$K(c_3)$};
    \node[use] (c3u) [right = 0.8cm of c3] {};
@@
   %c5
-   \node[component] (c5) [below = 0.8cm of c3] {$c_5$};
+   \node[component] (c5) [below = 0.8cm of c3] {$K(c_5)$};
    \node[use] (c5u) [right = 0.8cm of c5] {};
@@
   %c7
-   \node[component] (c7) [below = 0.8cm of par0] {$c_7$};
+   \node[component] (c7) [below = 0.8cm of par0] {$K(c_7)$};
    \node[use] (c7u) [right = 0.8cm of c7] {};
@@
   %c8
-   \node[component] (c8) [below = 0.8cm of sync2] {$c_8^*$};
+   \node[component] (c8) [below = 0.8cm of sync2] {$K(c_8^*)$};
    \node[use] (c8u) [right = 0.8cm of c8] {};

### 2015-08-03 WOLFHPC15/article/related.tex
sec. 7

@@
 They are made to solve many kinds of PDEs using well-known methods and solvers which are efficiently parallelized inside the library.
-Those libraries also include solvers for implicit numerical schemes, not adressed in this work.
+Those libraries also include solvers for implicit numerical schemes, not addressed in this work.
 However, using those solutions the user has to use and follow existing tools.
@@
 The purpose of those solutions are much more specific than our solution.
+\todo[inline]{JB: j'utiliserais plutot le terme "framework" que "library"}
 
@@
 Recently, domain specific languages for very specific numerical methods such as ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} for multigrid solvers, and OPS~\cite{Reguly:2014:ODS:2691166.2691173} for meshes of structured meshes (hybrid mesh) have also been proposed.
-All those solutions are powerfull, let the user implement their own stencil codes in a sequential programming style, and produce high performance codes for shared memory architectures and sometimes distributed memory architectures.
+All those solutions are powerful, let the user implement their own stencil codes in a sequential programming style, and produce high performance codes for shared memory architectures and sometimes distributed memory architectures.
 
-Most stencil compilers handle the parallelization or the optimization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations.
-However, many real case numerical simulation is not composed of a single stencil code, as explained in Section~\ref{sect:multistencil}, but of more than one stencil computation, involving one or more stencil shapes, and additionnal local coputations.
+Most stencil compilers handle the parallelization or the optimization of a single stencil kernel, considering that it represent the main computation time of numerical simulations.
+However, many real case numerical simulation are not composed of a single stencil kernel, as explained in Section~\ref{sect:multistencil}, but of more than one stencil computation, involving one or more stencil shapes, and additional local computations.
 Solutions like Pochoir or PATUS handle the parallelization or the optimization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations.
@@
 For example, it seems that control components introduced in this paper can be used for many different DSLs in many different domains.
-Second, separation of concerns in the final code is also improved as the generated component assembly split the different functionnalities of the application in different components.
+Second, separation of concerns in the final code is also improved as the generated component assembly split the different functionalities of the application in different components.
 As a result, some components can be written by computer-scientists, while other are specific to the application, as for example computation components.
-Finally, the maintainability of the application is also increased, as each functionnality is separated.
-In the final code if a computation component has to be modified by a numerician, intricates part of the code will be unchanged and not seen by the numerician.
+Finally, the maintainability of the application is also increased, as each functionality is separated.
+In the final code if a computation component has to be modified by a numerician, intricate part of the code will be unchanged and not seen by the numerician.
 

### 2015-08-03 WOLFHPC15/article/concept.tex
Removed related work in sec.2

@@
 Many solutions (languages or libraries) thus propose to ease their programming by producing optimized and parallelized code from a simple description of the local computation to apply on each element.
-For example, Liszt~\cite{DeVito2011LDS} and OP2~\cite{Giles2011} are domain specific languages to define a stencil code on unstructured meshes, while Pochoir~\cite{spaaTangCKLL11} and PATUS~\cite{citeulike12258902} address structured meshes, respectively proposing an optimized cache tiling technique, and a separation of concerns between parallelization strategies and computation descriptions.
-\todo[inline]{JB et HC: etat de l'art a mettre ailleurs ?}
+These are further discussed in Sections~\ref{sect:related}.
 

### 2015-08-04 WOLFHPC15/article/component.tex
wip Julien

@@
 This section describes the projection of a MSP program to components, including the components used to implement the static scheduling of $\Gamma_{tsp}$.
-Before those details the section gives an overview of component models, and introduce the Low Level Components (\llc), the model used in this paper.
+Before those details, the section gives an overview of component models, and especially the Low Level Components (\llc) used in this paper.
 
@@
 connected with a black rectangle.
-
-\todo[inline]{JB: chez moi la figure\,\ref{fig:ports} deborde...}
-
 A \llc-based application is a static \emph{assembly} of components instances and connections between their ports.
 Such an assembly is described in LAD, an XML dialect~\cite{l2c}.
+\todo[inline]{JB: chez moi la figure\,\ref{fig:ports} deborde...}
 
@@
 
+\todo[inline]{JB: debut WIP}
+
+The back-end of the compiler builds a component assembly, that can be divided in two parts: the components implementing the actual computation and the static schedule, and the components needed to manage it, \ie the runtime support.
+A general view of this assembly is proposed in Figure~\ref{fig:mscac:assembly}; it is explained in details in Section~\ref{sect:component}.
+As illustrated in the figure, each component is linked to one or more tuple of the multi-stencil program definition: \emph{DDS} component represents $\mathcal{M}$ and $\mathcal{E}$; \emph{Data} component represents $\Delta$ and $\mathcal{D}$; \emph{Time} component represents $T$, and \emph{Computations} component (which is generated by MSC) represents $\Gamma$.
+
+The final back-end component assembly is a ready-to-fill parallel structure where the user has to write computation codes described in the MSL description.
+To write those computations, interfaces of the choosen implementation have to be used.
+For example, the back-end evaluated in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}.
+A work in progress replaces this DDS component by another one making use of Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.
+
+\begin{figure}[t]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (D) at (0,0) {$Driver$};
+   \node[provide] (Dp) at (-1,0) {};
+   \node (Ds) at (-1.5,0) {start};
+   \node[use,right=1.5cm of D] (Du1) {};
+   \node[use,below=0.8cm of D] (Du2) {};
+
+   \node[provide,below=0.15 of Du2] (DAp) {};
+   \node[component,below=0.8cm of DAp] (DA) {$DriverApp$};
+   \node[use,right=1.7cm of DA] (DAu1) {$m$};
+   \node[use,below=0.8cm of DA] (DAu2) {};
+
+   \node[provide,below=0.15 of DAu2] (Tp) {};
+   \node[component,below=0.8cm of Tp] (T) {$Time$};
+   \node[use,right=1cm of T] (Tu) {};
+   \node[left=0.8cm of T] (tt) {$T$};
+
+   \node[provide,right=0.15 of Tu] (Cp) {};
+   \node[component,right=2cm of Cp] (C) {$Computations$};
+   \node[use,above=0.8cm of C] (Cu) {};
+   \node[right=0.2cm of Cu] (star) {$*$};
+   \node[right=1.5cm of C] (gamma) {$\Gamma$};
+
+   \node[provide,right=0.2 of DAu1] (Datap1) {};
+   \node[provide,above=0.15 of Cu] (Datap2) {};
+   \node[component,above=0.8cm of Datap2] (Data) {$Data$};
+   \node[use,above=0.8cm of Data] (Datau) {};
+   \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
+
+   \node[provide,right=0.15 of Du1] (DDSp1) {};
+   \node[provide,above=0.15 of Datau] (DDSp2) {};
+   \node[component,above=0.8cm of DDSp2] (DDS) {$DDS$};
+   \node[right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
+
+  \path[-]
+    (Dp) edge node {} (D)
+    (D) edge node {} (Du1)
+        edge node {} (Du2)
+    (DDSp1) edge node {} (DDS)
+    (DA) edge node {} (DAu1)
+           edge node {} (DAu2)
+    (DAp) edge node {} (DA)
+    (Tp) edge node {} (T)
+    (T)  edge node {} (Tu)
+    (Cp) edge node {} (C)
+    (C) edge node {} (Cu)
+    (Datap2) edge node {} (Data)
+    (Datap1) edge node {} (Data)
+    (Data) edge node {} (Datau)
+    (DDSp2) edge node {} (DDS);
+\end{tikzpicture}
+\caption{MSCAC Component Assembly Back-End Overview.}
+\label{fig:mscac:assembly}
+\end{center}
+\end{figure}
+
 Figure~\ref{fig:mscac:assembly} gives an overview of the back-end component
@@
 operations of $\Gamma_{tsp}$ as explained hereafter.
+
+\todo[inline]{JB: fin WIP}
 

### 2015-08-04 WOLFHPC15/article/concept.tex
wip Julien

@@
 \end{equation}
-where $T$ is the set of time iteration to run the simulation, $\mathcal{M}$ is the mesh of the simulation, $\mathcal{E}$ is the set of mesh entities, $\mathcal{D}$ is the set of computation domains used into computations, $\Delta$ is the set of data of the simulation, each one mapped onto a mesh entity, and finally $\Gamma$ is the set of computations.
+where $T$ is the set of time iteration to run the simulation, $\mathcal{M}$ is the mesh of the simulation, $\mathcal{E}$ is the set of mesh entities, $\mathcal{D}$ is the set of computation domains used for computations, $\Delta$ is the set of data of the simulation, each one mapped onto a mesh entity, and finally $\Gamma$ is the set of computations.
 The set of computations $\Gamma$ is an ordered list of stencil kernels and local computations.

### 2015-08-04 WOLFHPC15/article/ms_msc.tex
wip Julien

@@
 As mentioned in the previous section, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations.
-Six main sections are required in a MSL description that match the six tuples of the formal definition of a multi-stencil program: \textbf{1.}\,the mesh description, \textbf{2.}\,the mesh entities description, \textbf{3.}\,the computation domains description and their dependencies, \textbf{5.}\,the data description, \textbf{5.}\,the time loop description, and finally \textbf{6.} the computations description.
+Six main sections are required in a MSL description that match the six elements of the formal definition; they describe: \textbf{1.}\,the mesh, \textbf{2.}\,the mesh entities, \textbf{3.}\,the computation domains with their inter-dependencies, \textbf{4.}\,the data, \textbf{5.}\,the time iterations, and finally \textbf{6.} the computations.
 Figure~\ref{fig:grammar} shows the grammar of MSL.
@@
 As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is simply defined by a string identifier.
-Then, mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
-Figure~\ref{fig:mslex} presents an example defining a Cartesian mesh, called \texttt{cart}, and three mesh entities, called \texttt{cell}, \texttt{edgex} and \texttt{edgey}.
+Kind of mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
+Figure~\ref{fig:mslex} presents an example of a mesh named \texttt{cart} used to represent a Cartesian mesh, and three kind of mesh entities: \texttt{cell}, \texttt{edgex}, and \texttt{edgey}.
 
 \subsubsection*{Computation Domains and Their Dependencies}
-A computation domain can be seen as a subpart of entities of a mesh which will be used during one or more computation.
-For each computation domain, the mesh entity from which it is built is indicated (lines 3-4 and 12-14 in Figure~\ref{fig:grammar}).
-By default two computation domains always intersect except when an independence relation is mentioned in the section \emph{independent} of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
-In the lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each mesh entity, called \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they represent the whole domain.
-On lines 7 and 8 two other subparts of the mesh entity \texttt{edgex} are defined, called \texttt{part1edgex} and \texttt{part2edgex}; they represent the boundaries of the domain.
-On line 10 the independence relation between \texttt{part1edgex} and \texttt{part2edgex} is specified since they do not intersect.
+A computation domain is a subset of a given kind of entities from a mesh used during one or more computations.
+For each computation domain, the kind of mesh entities on which it is mapped is specified (lines 3-4 and 12-14 on Figure~\ref{fig:grammar}).
+By default two computation domains of the same entities are handled as intersecting except when an explicit independence relation is mentioned in the \emph{independent} section of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
+On lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each kind of mesh entities: \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they are used to represent the whole domain.
+On lines 7 and 8 two other domains of \texttt{edgex} are defined: \texttt{part1edgex} and \texttt{part2edgex}; they are used to represent the boundaries of the domain.
+On line 10 the independence relation between \texttt{part1edgex} and \texttt{part2edgex} is specified.
 
 \subsubsection*{Data and Time}
-A data is a quantity to simulate.
-It is mapped on mesh entities, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}.
+A data element is a quantity to simulate.
+It is mapped on a given kind of mesh entities, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}.
 The \texttt{time} section simply indicates a number of iterations to perform in the simulation.
-In the example of Figure~\ref{fig:mslex} ten data identifiers are defined and 500 iterations (lines 11-22).
+In the example of Figure~\ref{fig:mslex} ten data elements are defined and 500 iterations (lines 11-22).
 %An extension could be to support a convergence criterion.
@@
 Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} simply corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
-The specification starts by the identifier of the data to compute ($w$ in the formal description) with the computation domain between brackets ($d$ in the formal definition).
-Then, after the equal sign is specified the kernel identifier.
-Finally, between the parenthesis is the list of identifiers of data read by the computation with their stencil shape between brackets ($R$ in the formal definition).
-If the computation is local, no brackets appear ($R_l$ in the formal definition).
+The specification starts by the identifier of the computed data element ($w$ in the formal description) with the computation domain between brackets ($d$).
+Then, after the equal sign, the kernel identifier ($exp$) is specified.
+Finally, between the parenthesis is the list of identifiers of the data elements read by the computation with their stencil shape between brackets ($R$).
+If the computation is local, no brackets appear ($R_l$).
 In the example of Figure~\ref{fig:mslex}, nine computations are described.
-On line 24, the data \texttt{b} is computed on the computation domain \texttt{allcell} by the computation \texttt{c\_0} which read data \texttt{a} without neighborhood shape.
+On line 24, the data element \texttt{b} is computed on the domain \texttt{allcell} by the kernel \texttt{c\_0} which read data \texttt{a} without neighborhood shape.
 
@@
 This parallelization phase is divided in five different steps to transform $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
-Data parallelism is handled in the first two step, and task parallelism in the following three steps. This section gives an overview of the five steps of the MSC transformation.
-The detailed formalism of these steps and the associated algorithms are described in a research report~\cite{????}.
+Data parallelism is handled in the first two steps, and task parallelism in the following three steps.
+This section gives an overview of these five steps , but they are more formally defined with their associated algorithms in a research report~\cite{????}.
 
@@
 This \emph{synchronization computation} reads the data to synchronize, and write the same data.
-The computation domain of such a synchronization computation is the mesh entity on which the data is declared.
+The computation domain of such a synchronization computation encompasses the whole mesh entities on which the data is declared.
 As a result $\Gamma$ is transformed to a synchronized ordered list of computations $\Gamma_{sync}$.
-For example, in Figure~\ref{fig:mslex}, the stencil computation $c_1$ read the data $b$ which has been written by the computation $c_0$.
-For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed to the sublist $[c_0,sync_1,c_1]$ in $\Gamma_{sync}$.
-The new computation $sync_1$ read and write $b$ and is applied on the computation domain of $c_1$.
+For example, in Figure~\ref{fig:mslex}, the stencil computation $c_1$ reads the data element $b$ which has been written by $c_0$.
+For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed into $[c_0,sync_1,c_1]$ in $\Gamma_{sync}$.
+The new computation $sync_1$ reads and writes $b$ and is applied on the computation domain of $c_1$.
 As a result, a dependency is kept between $c_0$ and $sync_1$ and between $sync_1$, and $c_1$.
@@
 Thus, $\Gamma_{sync} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
+\todo[inline]{JB: "is applied on the computation domain of $c_1$" vraiment, c'est pas ce que tu dis juste avant ...}
 
@@
 From the synchronized ordered list of computations $\Gamma_{sync}$, a dependency graph $\Gamma_{dep}$ is built.
-A dependency exists between two computations (including synchronizations) if and only if a data read has been written by a previous computation in $\Gamma_{sync}$, and if the computation domains intersect.
-In other words, the only case where the same data is manipulated by distinct kernels without inducing a dependency is when the computation domains do not intersect.
+A dependency exists from computation $a$ to $b$ (including synchronizations) if and only if a data element read by $b$ has been written by $a$ in $\Gamma_{sync}$ with intersecting domains.
+%In other words, the only case where the same data is manipulated by distinct kernels without inducing a dependency is when the computation domains do not intersect.
 Nodes of the dependency graph represent computations of $\Gamma_{sync}$, while edges are dependencies between them.
@@
 Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes} by an over-constraint, as illustrated in Figure~\ref{fig:over}.
+
+\todo[inline]{JB: dans le DAG il y a une dependance de c2 a sync4, le graphe serie parallele les execute en parallele ...}
 

### 2015-08-04 WOLFHPC15/article/mscac.tex
5.4,5.5
wip Julien

@@
 This approach makes it possible to leverage good properties of software component models such as increased code re-use, eased separation of concerns, and increased maintainability and productivity.
+% JB: la phrase en dessous est pas tres heureuse ...
 Those properties counter some known limitations of DSLs. % JB: was "parallel applications".
@@
 
-\todo[inline]{JB: je déplacerais les 2 derniers paragraphes et la figure dans la sec.~\ref{sect:component}}
-
-The back-end of the compiler builds a component assembly, that can be divided in two parts: the components implementing the actual computation and the static schedule, and the components needed to manage it, \ie the runtime support.
-A general view of this assembly is proposed in Figure~\ref{fig:mscac:assembly}; it is explained in details in Section~\ref{sect:component}.
-As illustrated in the figure, each component is linked to one or more tuple of the multi-stencil program definition: \emph{DDS} component represents $\mathcal{M}$ and $\mathcal{E}$; \emph{Data} component represents $\Delta$ and $\mathcal{D}$; \emph{Time} component represents $T$, and \emph{Computations} component (which is generated by MSC) represents $\Gamma$.
-
-The final back-end component assembly is a ready-to-fill parallel structure where the user has to write computation codes described in the MSL description.
-To write those computations, interfaces of the choosen implementation have to be used.
-For example, the back-end evaluated in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}.
-A work in progress replaces this DDS component by another one making use of Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.
-
-\begin{figure}[t]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (D) at (0,0) {$Driver$};
-   \node[provide] (Dp) at (-1,0) {};
-   \node (Ds) at (-1.5,0) {start};
-   \node[use,right=1.5cm of D] (Du1) {};
-   \node[use,below=0.8cm of D] (Du2) {};
-
-   \node[provide,below=0.15 of Du2] (DAp) {};
-   \node[component,below=0.8cm of DAp] (DA) {$DriverApp$};
-   \node[use,right=1.7cm of DA] (DAu1) {$m$};
-   \node[use,below=0.8cm of DA] (DAu2) {};
-
-   \node[provide,below=0.15 of DAu2] (Tp) {};
-   \node[component,below=0.8cm of Tp] (T) {$Time$};
-   \node[use,right=1cm of T] (Tu) {};
-   \node[left=0.8cm of T] (tt) {$T$};
-
-   \node[provide,right=0.15 of Tu] (Cp) {};
-   \node[component,right=2cm of Cp] (C) {$Computations$};
-   \node[use,above=0.8cm of C] (Cu) {};
-   \node[right=0.2cm of Cu] (star) {$*$};
-   \node[right=1.5cm of C] (gamma) {$\Gamma$};
-
-   \node[provide,right=0.2 of DAu1] (Datap1) {};
-   \node[provide,above=0.15 of Cu] (Datap2) {};
-   \node[component,above=0.8cm of Datap2] (Data) {$Data$};
-   \node[use,above=0.8cm of Data] (Datau) {};
-   \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
-
-   \node[provide,right=0.15 of Du1] (DDSp1) {};
-   \node[provide,above=0.15 of Datau] (DDSp2) {};
-   \node[component,above=0.8cm of DDSp2] (DDS) {$DDS$};
-   \node[right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
-
-  \path[-]
-    (Dp) edge node {} (D)
-    (D) edge node {} (Du1)
-        edge node {} (Du2)
-    (DDSp1) edge node {} (DDS)
-    (DA) edge node {} (DAu1)
-           edge node {} (DAu2)
-    (DAp) edge node {} (DA)
-    (Tp) edge node {} (T)
-    (T)  edge node {} (Tu)
-    (Cp) edge node {} (C)
-    (C) edge node {} (Cu)
-    (Datap2) edge node {} (Data)
-    (Datap1) edge node {} (Data)
-    (Data) edge node {} (Datau)
-    (DDSp2) edge node {} (DDS);
-\end{tikzpicture}
-\caption{MSCAC Component Assembly Back-End Overview.}
-\label{fig:mscac:assembly}
-\end{center}
-\end{figure}
+Thanks to the software components based approach, the generated code is made of independent parts:
+A first set of components encodes the static scheduling provided by MSC.
+The other parts loosely match the elements of the multi-stencil program definition, these include the components implementing: \textbf{a)} the kernels ($\Gamma$), \textbf{b)} the data structures ($\mathcal{M}$ and $\mathcal{E}$), \textbf{c)} their parallelization ($\Delta$ and $\mathcal{D}$), and \textbf{d)} the time loop ($T$).
+This enables to change the technology used for each part while limiting the impact on the other parts as further discussed in Section~\ref{sect:component}.
 

### 2015-08-05 WOLFHPC15/article/component.tex
5.3, discussion modularite solution

--- 2015-08-03 WOLFHPC15/article/component.tex
@@
 
-Figure~\ref{fig:mscac:assembly} gives an overview of the back-end component
-assembly that is executed. As discusses in Section~\ref{sect:mscac},
-there are different types of components. Components~\texttt{Driver},
-\texttt{DriverApp}, and \texttt{Time} are always instanciated in the final component assembly; the input MSL file is not used by them except the component \texttt{Time} which uses the terminal \texttt{iteration} of the grammar.
-There is also a
-single Component~\texttt{DDS} (the current version handles a single mesh
-type). All these components are provided with the
-compiler; dumping this part of the assembly is straightforward.
+\begin{figure}[t]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[provide] (Dp) at (-1,0) {};
+   \node (Ds) at (-1.5,0) {start};
+   \node[component] (D) at (0,0) {$Driver$};
+   \node[use,right=1.5cm of D] (Du1) {};
+   \node[use,below=0.8cm of D] (Du2) {};
+   \node[use,right=1.7cm of Du2] (Du3) {$m$};
 
-While Component~\texttt{DDS} is managing the structure of the mesh,
-each data of the simulation is handled by a component of type
-\texttt{Data}. Therefore, the compiler generates in the assembly as
-many instances of such component as needed, from the \texttt{data}
-section of an MSL program.
+   \node[provide,below=0.15 of Du2] (Tp) {};
+   \node[component,below=0.8cm of Tp] (T) {$Time$};
+   \node[use,below=0.8cm of T] (Tu) {};
+   \node[left=0.8cm of T] (tt) {$T$};
 
-The last part of the assembly is the computation part.
-Each computation kernel is embedded in a different kernel component which has to be written by the user and provided
-to the compiler.
-We denote this kind of components K.
-A kernel component encapsulates a computations. To
-access data, a kernel component is connected, by the compiler, to the adequate
-\texttt{Data} components such that their ports name respect the names
-given in the \texttt{data} section of MSL. We denote the fact that they may have
-several use port by adding a star on the use port, as illustrated in Figure~\ref{fig:k}.
+   \node[provide,below=0.15 of Tu] (Sp) {};
+   \node[component,below=0.8cm of Sp] (S) {$Scheduling$};
+   \node[use,right=1.5cm of S] (Su) {};
+   \node[use,right=1.5cm of Tu] (Su2) {};
+   \node[left=1.5cm of S] (gamma) {$\Gamma_{tsp}$};
+
+   \node[provide,right=0.15 of Su] (Cp) {};
+   \node[component,right=1.5cm of Cp] (C) {$Kernels$};
+   \node[use,above=0.8cm of C] (Cu) {};
+   \node[right=0.2cm of Cu] (star) {$*$};
+
+   \node[provide,right=0.2 of Du3] (Datap1) {};
+   \node[provide,above=0.15 of Cu] (Datap2) {};
+   \node[provide,right=0.15 of Su2] (Datap3) {};
+   \node[component,above=0.8cm of Datap2] (Data) {$Data$};
+   \node[use,above=0.8cm of Data] (Datau) {};
+   \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
+
+   \node[provide,right=0.15 of Du1] (DDSp1) {};
+   \node[provide,above=0.15 of Datau] (DDSp2) {};
+   \node[component,above=0.8cm of DDSp2] (DDS) {$DDS$};
+   \node[right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
+
+  \path[-]
+    (Dp)     edge node {} (D)
+    (D)      edge node {} (Du1)
+             edge node {} (Du2)
+             edge node {} (Du3)
+    (DDSp1)  edge node {} (DDS)
+    (T)      edge node {} (Tu)
+    (Tp)     edge node {} (T)
+    (Sp)     edge node {} (S)
+    (S)      edge node {} (Su)
+             edge node {} (Su2)
+    (Cp)     edge node {} (C)
+    (C)      edge node {} (Cu)
+    (Datap1) edge node {} (Data)
+    (Datap2) edge node {} (Data)
+    (Datap3) edge node {} (Data)
+    (Data)   edge node {} (Datau)
+    (DDSp2)  edge node {} (DDS);
+\end{tikzpicture}
+\vspace*{.5em}
+\caption{MSCAC Component Assembly Back-End Overview.}
+\label{fig:mscac:assembly}
+\end{center}
+\end{figure}
+
+As described in Section~\ref{sect:mscac}, the compiler generates five independent parts in a component assembly in addition to a static part.
+This is represented on Figure~\ref{fig:mscac:assembly} where each rounded box represents one or more component instances.
+\texttt{Driver} is static and drives the whole application while \texttt{Kernels} contains code directly written by the user.
+The four other parts are generated based on the content of sections of the MSL program: \texttt{Scheduling} from $\Gamma$, \emph{DDS} from $\mathcal{M}$ and $\mathcal{E}$, \emph{Data} from $\Delta$ and $\mathcal{D}$, and \emph{Time} from $T$.
+
+\texttt{DDS} manages the structure of the mesh, a single instance is generated as MSL only supports a single mesh for now.
+The back-end presented in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}.
+\texttt{DDS} is used by \texttt{Data} containing one component instance for each data element in the MSL program.
+
+\texttt{Time} is parametrized by the the \texttt{iteration} value from MSL and calls the \texttt{Scheduling} part as many times as required.
+\texttt{Scheduling} contains the static scheduling computed by MSC, \ie $\Gamma_{tsp}$.
+The series-parallel tree is encoded using three specific components which manage the $P$, $S$, and $sync$ operations of $\Gamma_{tsp}$ as explained hereafter.
+At the leafs of the tree, the components instances make call either to \texttt{Data} for synchronization of to \texttt{Kernels} for computation.
+
+\texttt{Kernels} contains code written by the user to implement the actual computation.
+As presented on Figure~\ref{fig:k}, these components provide a single port used to launch the computation and expose one use port for each data element manipulated in the computation.
+We denote the fact that they may have several use port by adding a star on the use port.
+Each of these use ports is connected to the matching component instance in \texttt{Data}.
+We denote this kind of component K.
+
+As a result of this approach, each part of the generated code is rather independent.
+Changing the main loop to use a convergence criteria rather than a fixed number of iterations would only require changing the \texttt{Time} component.
+Similarly, changing the approach used for scheduling would only impact the \texttt{Scheduling} components.
+The technology used for data parallelism can also be changed by replacing the \texttt{DDS} and \texttt{Data} components; as a matter of fact work is in progress to propose an alternative based on Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} instead of SkelGIS.
+In this case however, the \texttt{Kernels} components written by the user have to be adapted to the new interface, but this is the only impact.
+
 \begin{figure}
@@
 \end{figure}
-
-%% In the computation part of the simulation a final component is needed,
-%% the computation component, denoted K. This type of component
-%% represents a single computation of the simulation, i.e. the
-%% computation of a single output data (written) using a set of input
-%% data (read). Using a \emph{use-multiple}, however, a list of component
-%% is used and the explicit identification of each component to use is
-%% lost. If for SEQ or PAR, the identification of components is not
-%% usefull, it is usefull to manipulate data in a computation. For this
-%% reason a computation component has as much \emph{use} ports as input
-%% and output data to manipulate. We denote this in the bellow
-%% representation as a star on the use port.
-
-The rest of the computation part of the assembly contains the static parallel schedule
-computed by MSC, \ie $\Gamma_{tsp}$. This is achieved through the use of
-three specific components which manage the $P$, $S$, and $sync$
-operations of $\Gamma_{tsp}$ as explained hereafter.
 

### 2015-08-05 WOLFHPC15/article/mscac.tex
5.3, discussion modularite solution

@@
 
-Thanks to the software components based approach, the generated code is made of independent parts:
-A first set of components encodes the static scheduling provided by MSC.
-The other parts loosely match the elements of the multi-stencil program definition, these include the components implementing: \textbf{a)} the kernels ($\Gamma$), \textbf{b)} the data structures ($\mathcal{M}$ and $\mathcal{E}$), \textbf{c)} their parallelization ($\Delta$ and $\mathcal{D}$), and \textbf{d)} the time loop ($T$).
+Thanks to the software components based approach, the generated code is made of five independent parts that loosely match the elements of the multi-stencil program definition.
+These are the components implementing: \textbf{a)} the scheduling ($\Gamma$), \textbf{b)} the data structures ($\mathcal{M}$ and $\mathcal{E}$), \textbf{c)} their parallelization ($\Delta$ and $\mathcal{D}$), \textbf{d)} the time loop ($T$), and \textbf{e)} the kernels directly provided by the user.
 This enables to change the technology used for each part while limiting the impact on the other parts as further discussed in Section~\ref{sect:component}.

### 2015-08-05 WOLFHPC15/article/component.tex
back to the future

@@
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (D) at (0,0) {$Driver$};
    \node[provide] (Dp) at (-1,0) {};
    \node (Ds) at (-1.5,0) {start};
-   \node[component] (D) at (0,0) {$Driver$};
    \node[use,right=1.5cm of D] (Du1) {};
    \node[use,below=0.8cm of D] (Du2) {};
-   \node[use,right=1.7cm of Du2] (Du3) {$m$};
 
-   \node[provide,below=0.15 of Du2] (Tp) {};
+   \node[provide,below=0.15 of Du2] (DAp) {};
+   \node[component,below=0.8cm of DAp] (DA) {$DriverApp$};
+   \node[use,right=1.7cm of DA] (DAu1) {$m$};
+   \node[use,below=0.8cm of DA] (DAu2) {};
+
+   \node[provide,below=0.15 of DAu2] (Tp) {};
    \node[component,below=0.8cm of Tp] (T) {$Time$};
-   \node[use,below=0.8cm of T] (Tu) {};
+   \node[use,right=1cm of T] (Tu) {};
    \node[left=0.8cm of T] (tt) {$T$};
 
-   \node[provide,below=0.15 of Tu] (Sp) {};
-   \node[component,below=0.8cm of Sp] (S) {$Scheduling$};
-   \node[use,right=1.5cm of S] (Su) {};
-   \node[use,right=1.5cm of Tu] (Su2) {};
-   \node[left=1.5cm of S] (gamma) {$\Gamma_{tsp}$};
-
-   \node[provide,right=0.15 of Su] (Cp) {};
-   \node[component,right=1.5cm of Cp] (C) {$Kernels$};
+   \node[provide,right=0.15 of Tu] (Cp) {};
+   \node[component,right=2cm of Cp] (C) {$Computations$};
    \node[use,above=0.8cm of C] (Cu) {};
    \node[right=0.2cm of Cu] (star) {$*$};
+   \node[right=1.5cm of C] (gamma) {$\Gamma$};
 
-   \node[provide,right=0.2 of Du3] (Datap1) {};
+   \node[provide,right=0.2 of DAu1] (Datap1) {};
    \node[provide,above=0.15 of Cu] (Datap2) {};
-   \node[provide,right=0.15 of Su2] (Datap3) {};
    \node[component,above=0.8cm of Datap2] (Data) {$Data$};
@@
         edge node {} (Du2)
-             edge node {} (Du3)
     (DDSp1) edge node {} (DDS)
+    (DA) edge node {} (DAu1)
+           edge node {} (DAu2)
+    (DAp) edge node {} (DA)
+    (Tp) edge node {} (T)
     (T)  edge node {} (Tu)
-    (Tp)     edge node {} (T)
-    (Sp)     edge node {} (S)
-    (S)      edge node {} (Su)
-             edge node {} (Su2)
     (Cp) edge node {} (C)
     (C) edge node {} (Cu)
+    (Datap2) edge node {} (Data)
     (Datap1) edge node {} (Data)
-    (Datap2) edge node {} (Data)
-    (Datap3) edge node {} (Data)
     (Data) edge node {} (Datau)
@@
 This is represented on Figure~\ref{fig:mscac:assembly} where each rounded box represents one or more component instances.
-\texttt{Driver} is static and drives the whole application while \texttt{Kernels} contains code directly written by the user.
-The four other parts are generated based on the content of sections of the MSL program: \texttt{Scheduling} from $\Gamma$, \emph{DDS} from $\mathcal{M}$ and $\mathcal{E}$, \emph{Data} from $\Delta$ and $\mathcal{D}$, and \emph{Time} from $T$.
+\texttt{Driver} \& \texttt{DriverApp} are static and drive the whole application.
+The four other parts are generated based on the content of sections of the MSL program: \texttt{Computations} from $\Gamma$ (with code provided by the user), \emph{DDS} from $\mathcal{M}$ and $\mathcal{E}$, \emph{Data} from $\Delta$ and $\mathcal{D}$, and \emph{Time} from $T$.
 
@@
 \texttt{DDS} is used by \texttt{Data} containing one component instance for each data element in the MSL program.
+\texttt{Time} is parametrized by the the \texttt{iteration} value from MSL and calls the \texttt{Scheduling} part as many times as required.
 
-\texttt{Time} is parametrized by the the \texttt{iteration} value from MSL and calls the \texttt{Scheduling} part as many times as required.
-\texttt{Scheduling} contains the static scheduling computed by MSC, \ie $\Gamma_{tsp}$.
+\texttt{Computations} contains the static scheduling computed by MSC, \ie $\Gamma_{tsp}$.
 The series-parallel tree is encoded using three specific components which manage the $P$, $S$, and $sync$ operations of $\Gamma_{tsp}$ as explained hereafter.
-At the leafs of the tree, the components instances make call either to \texttt{Data} for synchronization of to \texttt{Kernels} for computation.
-
-\texttt{Kernels} contains code written by the user to implement the actual computation.
+At the leafs of the tree, the components instances make call either to \texttt{Data} for synchronization of user-provided components for the computations.
 As presented on Figure~\ref{fig:k}, these components provide a single port used to launch the computation and expose one use port for each data element manipulated in the computation.
@@
 Changing the main loop to use a convergence criteria rather than a fixed number of iterations would only require changing the \texttt{Time} component.
-Similarly, changing the approach used for scheduling would only impact the \texttt{Scheduling} components.
+Similarly, changing the approach used for scheduling would only impact the code generated in \texttt{Computations}.
 The technology used for data parallelism can also be changed by replacing the \texttt{DDS} and \texttt{Data} components; as a matter of fact work is in progress to propose an alternative based on Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} instead of SkelGIS.
-In this case however, the \texttt{Kernels} components written by the user have to be adapted to the new interface, but this is the only impact.
+In this case however, the \texttt{Computations} part might have to be altered to match the new interface of \texttt{Data}, but this is the only impact.
 

### 2015-08-05 WOLFHPC15/article/component.tex
Fuse Driver & DriverApp

@@
    \node[use,right=1.5cm of D] (Du1) {};
-   \node[use,below=0.8cm of D] (Du2) {};
+   \node[use,below=1.75cm of D] (Du2) {};
+   \node[use,right=1.7cm of Du2] (Du3) {$m$};
 
-   \node[provide,below=0.15 of Du2] (DAp) {};
-   \node[component,below=0.8cm of DAp] (DA) {$DriverApp$};
-   \node[use,right=1.7cm of DA] (DAu1) {$m$};
-   \node[use,below=0.8cm of DA] (DAu2) {};
-
-   \node[provide,below=0.15 of DAu2] (Tp) {};
-   \node[component,below=0.8cm of Tp] (T) {$Time$};
+   \node[provide,below=0.15 of Du2] (Tp) {};
+   \node[component,below=1.6cm of Tp] (T) {$Time$};
    \node[use,right=1cm of T] (Tu) {};
@@
 
-   \node[provide,right=0.2 of DAu1] (Datap1) {};
+   \node[provide,right=0.2 of Du3] (Datap1) {};
    \node[provide,above=0.15 of Cu] (Datap2) {};
@@
         edge node {} (Du2)
+        edge node {} (Du3)
     (DDSp1) edge node {} (DDS)
-    (DA) edge node {} (DAu1)
-           edge node {} (DAu2)
-    (DAp) edge node {} (DA)
     (Tp) edge node {} (T)
@@
 This is represented on Figure~\ref{fig:mscac:assembly} where each rounded box represents one or more component instances.
-\texttt{Driver} \& \texttt{DriverApp} are static and drive the whole application.
+\texttt{Driver} is static and drives the whole application.
 The four other parts are generated based on the content of sections of the MSL program: \texttt{Computations} from $\Gamma$ (with code provided by the user), \emph{DDS} from $\mathcal{M}$ and $\mathcal{E}$, \emph{Data} from $\Delta$ and $\mathcal{D}$, and \emph{Time} from $T$.

### 2015-08-05 WOLFHPC15/article/component.tex
corrected figs

@@
   %seq1
-   \node[component] (seq1) [right = 1cm of par0u] {$SEQ$};
+   \node[component] (seq1) [right = 1cm of sync0u] {$SEQ$};
    \node[use] (seq1u) [right = 1cm of seq1] {$m$};
-   %par1
-   \node[component] (par1) [right = 1cm of seq1u] {$PAR$};
-   \node[use] (par1u) [right = 1cm of par1] {$m$};
   %c2
-   \node[component] (c2) [right = 1cm of par1u] {$K(c_2)$};
+   \node[component] (c2) [right = 4cm of c0] {$K(c_2)$};
    \node[use] (c2u) [right = 0.8cm of c2] {};
@@
   %c4
-   \node[component] (c4) [below = 0.8cm of par1] {$K(c_4^*)$};
+   \node[component] (c4) [below = 0.8cm of sync1] {$K(c_4^*)$};
    \node[use] (c4u) [right = 0.8cm of c4] {};
@@
   %seq2
-   \node[component] (seq2) [below = 2.5cm of seq1] {$SEQ$};
+   \node[component] (seq2) [below = 3.2cm of seq1] {$SEQ$};
    \node[use] (seq2u) [right = 1cm of seq2] {$m$};
@@
   %seq1
-    (par0u) edge node {} (seq1)
+    (par0u) edge node {} (seq1.west)
     (seq1) edge node {} (seq1u)
-  %par1
-    (seq1u) edge node {} (par1.west)
-    (par1) edge node {} (par1u)
   %c2
-    (par1u) edge node {} (c2)
+    (seq1u) edge node {} (c2.west)
     (c2) edge node {} (c2u)
   %sync0
-    (par1u) edge node {} (sync1.west)
+    (seq1u) edge node {} (sync1.west)
     (sync1) edge node {} (sync1u)
@@
 \end{tikzpicture}
+\vspace*{.5em}
 \caption{Component assembly representing the computation part generated from $\Gamma_{tsp}$ of Figure~\ref{fig:tsp}.}

### 2015-08-05 WOLFHPC15/article/ms_msc.tex
corrected figs

@@
    \node[] (s0)    at (0,0) {$\mathcal{S}$};
+
    \node[] (c0)    at (-3,1) {$c_0$};
@@
    \node[] (c1)    at (-1,1) {$c_1$};
+   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
+   \node[] (c7)    at ( 1,1) {$c_7$};
+   \node[] (star8) at ( 2,1) {$sync_8$};
+   \node[] (c8)    at ( 3,1) {$c_8$};
 
-   \node[] (p0) at (0,1) {$\mathcal{P}$};
    \node[] (s1)    at (-1,2) {$\mathcal{S}$};
-   \node[] (p1) at (-2,3) {$\mathcal{P}$};
+   \node[] (s2)    at (1,2) {$\mathcal{S}$};
+
+   \node[] (c2)    at (-3,3) {$c_2$};
+   \node[] (star4) at (-2,3) {$sync_4$};
    \node[] (c4)    at (-1,3) {$c_4$};
-   \node[] (c6) at (-0,3) {$c_6$};
-   \node[] (c2) at (-2.5,4) {$c_2$};
-   \node[] (star4) at (-1.5,4) {$sync_4$};
-   \node[] (s2) at (1,2) {$\mathcal{S}$};
-   \node[] (c3) at (0.5,3) {$c_3$};
-   \node[] (c5) at (1.5,3) {$c_5$};
-
-   \node[] (c7) at (1,1) {$c_7$};
-   \node[] (star8) at (2,1) {$sync_8$};
-   \node[] (c8) at (3,1) {$c_8$};
+   \node[] (c6)    at ( 0,3) {$c_6$};
+   \node[] (c3)    at ( 1,3) {$c_3$};
+   \node[] (c5)    at ( 2,3) {$c_5$};
 
@@
          edge node {} (s2)
-    (s1) edge node {} (p1)
+    (s1) edge node {} (c2)
+         edge node {} (star4)
+         edge node {} (c4)
          edge node {} (c4)
          edge node {} (c6)
-    (p1) edge node {} (c2)
-         edge node {} (star4)
     (s2) edge node {} (c3)

### 2015-08-06 WOLFHPC15/article/component.tex
wolfhpc fix all parts helene 06_08

@@
 A component is a black box that implements an independent functionality of the application, and which interacts with its environment only through well defined interfaces: its ports.
-A port can for example specify services provided or required by the component.
+A port can, for example, specify services provided or required by the component.
 With respect to high performance computing, some works have also shown
@@
 Many component models exist, each of them with its own specificities.
-Well known component models include for example the CORBA Component Model (CCM)~\cite{corba:omg06}, and the Grid Component Model (GCM)~\cite{Baude} in distributed computing, the Common Component Architecture (CCA)~\cite{Bernholdt01052006}, and Low Level Components (\llc)~\cite{l2c} in HPC.
+Well known component models include, for example, the CORBA Component Model (CCM)~\cite{corba:omg06}, and the Grid Component Model (GCM)~\cite{Baude} for distributed computing, while the Common Component Architecture (CCA)~\cite{Bernholdt01052006}, and Low Level Components (\llc)~\cite{l2c} are HPC-oriented.
 This work makes use of \llc for the experiments.
@@
 Such an assembly is described in LAD, an XML dialect~\cite{l2c}.
-\todo[inline]{JB: chez moi la figure\,\ref{fig:ports} deborde...}
 
@@
 }
+\\
 \subfloat[][\label{fig:ass}]{
@@
    \node[use] (u) at (1.5,0) {};
-   \node[component,right=1.5 of u] (C1) {$c_1$};
+   \node[provide,right=0.15 of u] (p2) {};
+   \node[component,right=1.5 of p2] (C1) {$c_1$};
    \node[use,right=1.5 of C1] (um) {$m$};
@@
     (C1)  edge node {$v$} (um)
-    (u) edge node {$q$} (C1);
+    (p2) edge node {$q$} (C1);
 \end{tikzpicture}
 }
+\\
 \subfloat[][\label{fig:mpi}]{
@@
    \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
+   \node[mpi,above right=1cm of Data] (mpidata) {};
 
@@
     (Data) edge node {} (Datau)
+          edge node {} (mpidata)
     (DDSp2) edge node {} (DDS);
@@
 
-As described in Section~\ref{sect:mscac}, the compiler generates five independent parts in a component assembly in addition to a static part.
+As described in Section~\ref{sect:mscac}, the compiler generates four independent parts in a component assembly in addition to a static part.
 This is represented on Figure~\ref{fig:mscac:assembly} where each rounded box represents one or more component instances.
@@
 \texttt{DDS} is used by \texttt{Data} containing one component instance for each data element in the MSL program.
-\texttt{Time} is parametrized by the the \texttt{iteration} value from MSL and calls the \texttt{Scheduling} part as many times as required.
+\texttt{Time} is parametrized by the the \texttt{iteration} value from MSL and calls the \texttt{Computations} part as many times as required.
 
-\texttt{Computations} contains the static scheduling computed by MSC, \ie $\Gamma_{tsp}$.
-The series-parallel tree is encoded using three specific components which manage the $P$, $S$, and $sync$ operations of $\Gamma_{tsp}$ as explained hereafter.
-At the leafs of the tree, the components instances make call either to \texttt{Data} for synchronization of user-provided components for the computations.
+\texttt{Computations} contains the static scheduling computed by MSC, \ie $\Gamma_{tsp}$, or the data parallel transformation $\Gamma_{data}$.
+Both transformations $\Gamma_{tsp}$ and $\Gamma_{data}$ are encoded using three specific components which manage the $P$, $S$, and $sync$ operations as explained hereafter.
+At the leaves of the tree, the computation components instances ($K$), provided by the user, are connected to \texttt{Data}.
 As presented on Figure~\ref{fig:k}, these components provide a single port used to launch the computation and expose one use port for each data element manipulated in the computation.
@@
 Similarly, changing the approach used for scheduling would only impact the code generated in \texttt{Computations}.
-The technology used for data parallelism can also be changed by replacing the \texttt{DDS} and \texttt{Data} components; as a matter of fact work is in progress to propose an alternative based on Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} instead of SkelGIS.
-In this case however, the \texttt{Computations} part might have to be altered to match the new interface of \texttt{Data}, but this is the only impact.
+The technology used for data parallelism can also be changed by replacing the \texttt{DDS} and \texttt{Data} components; as a matter of fact a work in progress proposes an alternative based on Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} instead of SkelGIS.
+In this case however, the kernel components of the \texttt{Computations} part, provided by the user, might have to be written with interfaces of the alternative distributed data structure.
 
@@
 %----------------------------------------
-\subsection{Control Components}
+\subsection{Control Components and Dump}
 %----------------------------------------
-The series-parallel tree decomposition $\Gamma_{tsp}$ represents the static scheduling of the simulation.
-It is dumped to components by generating an assembly that exactly matches the tree.
-We introduce what we call \emph{control components} to represent the nodes of the tree.
+The series-parallel tree decomposition $\Gamma_{tsp}$ represents a static scheduling of the simulation.
+It is dumped to components by generating an assembly that exactly matches the tree, as well as for $\Gamma_{data}$.
+We introduce what we call \emph{control components} to represent all nodes of $\Gamma_{tsp}$ or $\Gamma_{data}$.
 These components can be used for any simulation, which increases code reuse between simulations.
@@
 %----------------------------------------
-\subsection{Dump To Component Assembly}
+%\subsection{Dump To Component Assembly}
 %----------------------------------------
-From $\Gamma_{tsp}$ and using the control components described above,
-a direct dump can be done to a component assembly for an hybrid (data
+Then from $\Gamma_{tsp}$ or $\Gamma_{data}$, and using the control components described above,
+a direct dump can be done to a component assembly for a data or hybrid (data
 and task) parallelization of a simulation.
-\fix{HC: ce qui suit est a changer avec la fusion du code de plusieurs composants}
-However, it is also
-possible to dump to a data parallelization only. In such a case the
-computation of $\Gamma_{sync}$ is enough to generate the component
-assembly. Actually a single SEQ component is thus needed and this
-component is linked to all computations and synchronizations of
-$\Gamma_{sync}$ directly.
+%However, it is also
+%possible to dump to a data parallelization only. In such a case the
+%computation of $\Gamma_{sync}$ is enough to generate the component
+%assembly. Actually a single SEQ component is thus needed and this
+%component is linked to all computations and synchronizations of
+%$\Gamma_{sync}$ directly.
 

### 2015-08-06 WOLFHPC15/article/concept.tex
wolfhpc fix all parts helene 06_08

@@
 \end{center}
-\caption{Stencil examples}
+\caption{(a) a Cartesian mesh and two kind of mesh entities, (b) an example of stencil kernel on cells, (c) an example of stencil kernel on different entities of the mesh.}
 \label{fig:gspmsp}
@@
 As it can be seen from those definitions, stencil kernels implementations share many properties from an algorithmic point of view that can be used to apply well known optimization and parallelization strategies.
-Many solutions (languages or libraries) thus propose to ease their programming by producing optimized and parallelized code from a simple description of the local computation to apply on each element.
+Many solutions (languages or frameworks) thus propose to ease their programming by producing optimized and parallelized code from a simple description of the local computation to apply on each element.
 These are further discussed in Sections~\ref{sect:related}.
@@
 For example, we would like to formalize and parallelize a numerical simulation which chains stencil kernels of Figures~\ref{fig:ex1} and~\ref{fig:ex2}.
-The next section formalizes concepts of \emph{stencil kernel}, \emph{local computation} and what we call a \emph{multi-stencil program}.
+The next section formalizes concepts of \emph{stencil kernel}, \emph{local kernel} and \emph{multi-stencil program}.
 
@@
 where $T$ is the set of time iteration to run the simulation, $\mathcal{M}$ is the mesh of the simulation, $\mathcal{E}$ is the set of mesh entities, $\mathcal{D}$ is the set of computation domains used for computations, $\Delta$ is the set of data of the simulation, each one mapped onto a mesh entity, and finally $\Gamma$ is the set of computations.
-The set of computations $\Gamma$ is an ordered list of stencil kernels and local computations.
-
-\fix{HC 2 all : Est-ce que je laisse cette phrase : One can notice that this work is limited to a single type of mesh in a given simulation.}
+The set of computations $\Gamma$ is an ordered list of stencil and local kernels.
+One can notice that this work, for now, is limited to a single type of mesh in a given simulation.
 

### 2015-08-06 WOLFHPC15/article/evaluation.tex
Forgotten files
wolfhpc fix all parts helene 06_08

@@
 The series-parallel tree decomposition extracted by the MSC transformation is composed of seventeen sequences and eighteen parallel sections.
-In Figure~\ref{fig:freq} represents the level of parallelism, \ie the number of tasks which can be executed concurrently, and gives the number of time the level is observed in the final tree.
+Figure~\ref{fig:freq} represents for a given level of parallelism, \ie the number of tasks which can be executed concurrently, the number of time this level is observed in the final component assembly.
+One can notive that the level of task parallelism extracted from the Shallow water equations is limited by at the two sequential parts of the application (level 1).
+However, as a level of sixteen parallel tasks is reach too times, but also five times for the level six, sequential restrictions could be amortize.
+Moreover, as the level of parallelism in the application is heterogenous, the choice of the number of threads to launch for task parallelism is difficult.
 
@@
  \end{tabular}
-\caption{Parallelism level (number of parallel tasks) and the number of times this level appear.}
+\caption{Parallelism level (number of parallel tasks) and the number of times this level appears.}
 \label{fig:freq}
@@
 
-Figure~\ref{fig:freq} illustrates the execution time for each step of the MSC transformation for all overall execution time of ten seconds.
+Figure~\ref{fig:exectime} illustrates the execution time for each step of the MSC transformation for an overall execution time of ten seconds.
 Execution times have been computed on a laptop with a bi-core Intel Core i5 1,4 GHz, and 8Gb LPDDR3.
@@
 \caption{Execution times of the MSC transformation steps}
-\label{fig:freq}
+\label{fig:exectime}
  \end{center}
@@
 
-\fix{HC 2 all : Je ne sais pas si ce qui suit a sa place dans cet article.}
-
 One can notice that the transformation of $\Gamma_{dep}$ to a minimal series-parallel graph is the longest step of MSC, because of the removal of the forbidden shapes in the graph.
-Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted, because the algorithm use a general solution instead of finding each forbidden shape, but it seems that many of them appeared.
+Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted, because the algorithm uses a general solution instead of finding each forbidden shape, but it seems that many of them appeared.
 The \emph{N-shape}, represented in Figure~\ref{fig:n} is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.

### 2015-08-06 WOLFHPC15/article/introduction.tex
wolfhpc fix all parts helene 06_08

@@
 %On the other side of the scope
-Conversely, the more specific the DSL is, the less number of different applications can use it and the more difficult it is to amortize the implementation cost of the DSL (language, compiler, and runtime).
+Conversely, the more specific the DSL is, the less number of different applications can use it, and the more difficult it is to amortize the implementation cost of the DSL (language, compiler, and runtime).
 Programming difficulties, as well as non-maintainability and non-portability often seem to be simply deported from the scientist to DSL designers.
@@
 This approach proposes a transformation from a DSL to a component-based runtime.
-Component-based Software Engineering have proved many times good properties for code re-use, separation of concerns, maintainability and productivity of codes.
-For this reason, we think that mixing DSLs and component-based runtimes could enable those properties, by inheritance, in DSLs.
+Component-based Software Engineering has proved many times good properties for code re-use, separation of concerns, maintainability and productivity of codes.
+For this reason, we think that combining DSLs and component-based runtimes could enable those properties, by inheritance, in DSLs.
 To evaluate this approach, a proper new DSL for \emph{multi-stencil} programs, and its compiler are presented in this paper.
@@
 
-Many solutions to automatically optimize and parallelize partial differential equations (PDEs) solvers have been proposed in the literature, either as specific libraries~\cite{petsc-efficient,Trilinos-Overview,CPE:CPE3494} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
+Many solutions to automatically optimize and parallelize partial differential equations (PDEs) solvers have been proposed in the literature, either as specific frameworks~\cite{petsc-efficient,Trilinos-Overview,CPE:CPE3494} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
 Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil kernel.
-However, a real case numerical simulation is most of the time not composed of a single stencil kernel, but of a set of stencil kernels and a set of additional local auxiliary computations.
+However, a real case numerical simulation (such as fluid dynamics, magnetohydrodynamics, molecular synamics etc.) is most of the time not composed of a single stencil kernel, but of a set of stencil kernels and a set of additional local auxiliary computations.
 The DSL proposed in this paper, called MSL, is a DSL to generate a parallel component-based structure of an overall numerical simulation, that we call a \emph{multi-stencil program}.
@@
 %The end of Section~\ref{sect:component} also discusses on interest of this new approach.
-Section~\ref{sect:eval} focuses on a real-case simulation, solving the Shallow-water equations, and analyzes some performance results.
+Section~\ref{sect:eval} focuses on a real-case simulation, solving the Shallow-water equations, and analyzes preliminary performance results.
 Sections~\ref{sect:related} states on related work, and finally Section~\ref{sect:conclusion} concludes and gives some perspectives.

### 2015-08-06 WOLFHPC15/article/ms_msc.tex
wolfhpc fix all parts helene 06_08

@@
 \end{lstlisting}
-\caption{Grammar of the MS language}
+\caption{Grammar of the Multi-Stencil Language}
 \label{fig:grammar}
@@
 The remaining of this section describes those different parts of a MSL program and presents an example.
-We can note that, as MSL is a descriptive language, almost all terminals \texttt{meshid}, \texttt{meshdom}, \texttt{compdomid}, \texttt{dataid}, \texttt{compid} and \texttt{neighborid}, are string identifiers.
+One can notice that almost all terminals \texttt{meshid}, \texttt{meshdom}, \texttt{compdomid}, \texttt{dataid}, \texttt{compid} and \texttt{neighborid}, are string identifiers.
 Only \texttt{num} is an integer terminal.
@@
 As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is simply defined by a string identifier.
-Kind of mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
+The different kind of mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
 Figure~\ref{fig:mslex} presents an example of a mesh named \texttt{cart} used to represent a Cartesian mesh, and three kind of mesh entities: \texttt{cell}, \texttt{edgex}, and \texttt{edgey}.
@@
 \subsubsection*{Computation Domains and Their Dependencies}
-A computation domain is a subset of a given kind of entities from a mesh used during one or more computations.
+A computation domain is a subset of a given kind of mesh entities, used during one or more computations.
 For each computation domain, the kind of mesh entities on which it is mapped is specified (lines 3-4 and 12-14 on Figure~\ref{fig:grammar}).
-By default two computation domains of the same entities are handled as intersecting except when an explicit independence relation is mentioned in the \emph{independent} section of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
+By default, two computation domains of the same entities are handled as intersecting except when an explicit independence relation is mentioned in the \emph{independent} section of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
 On lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each kind of mesh entities: \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they are used to represent the whole domain.
@@
 \subsubsection*{Computation Description}
-The last part of the language contains the specifications of stencil kernels and local computations as defined in Section~\ref{sect:multistencil}.
-This description does not contain a direct expression of the numerical expressions.
+The last part of the language contains the specifications of stencil and local kernels as defined in Section~\ref{sect:multistencil}.
+This description does not contain a direct expression of the numerical expressions $exp$.
 Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} simply corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
@@
 In the example of Figure~\ref{fig:mslex}, nine computations are described.
-On line 24, the data element \texttt{b} is computed on the domain \texttt{allcell} by the kernel \texttt{c\_0} which read data \texttt{a} without neighborhood shape.
+On line 24, the data element \texttt{b} is computed on the domain \texttt{allcell} by the kernel expression \texttt{c\_0} which read data \texttt{a} without neighborhood shape.
 
@@
 This parallelization phase is divided in five different steps to transform $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
-Data parallelism is handled in the first two steps, and task parallelism in the following three steps.
+%Data parallelism is handled in the first two steps, and task parallelism in the following three steps.
 This section gives an overview of these five steps , but they are more formally defined with their associated algorithms in a research report~\cite{????}.
@@
 In a data parallelization, data is split among processors and the same program is applied on each sub-part of the data.
-This does however require additional synchronizations between processors.
+However, this parallelization technique requires additional synchronizations between processors.
 For example, as a stencil kernel accesses neighbor values, values computed by another processor are needed and must be synchronized. 
@@
 The computation domain of such a synchronization computation encompasses the whole mesh entities on which the data is declared.
-As a result $\Gamma$ is transformed to a synchronized ordered list of computations $\Gamma_{sync}$.
+As a result, $\Gamma$ is transformed to a synchronized ordered list of computations $\Gamma_{sync}$.
 For example, in Figure~\ref{fig:mslex}, the stencil computation $c_1$ reads the data element $b$ which has been written by $c_0$.
 For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed into $[c_0,sync_1,c_1]$ in $\Gamma_{sync}$.
-The new computation $sync_1$ reads and writes $b$ and is applied on the computation domain of $c_1$.
+The new computation $sync_1$ reads and writes $b$ and is applied on \texttt{edgex}.
 As a result, a dependency is kept between $c_0$ and $sync_1$ and between $sync_1$, and $c_1$.
@@
 Thus, $\Gamma_{sync} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
-\todo[inline]{JB: "is applied on the computation domain of $c_1$" vraiment, c'est pas ce que tu dis juste avant ...}
 
@@
 To do so the dependency graph is transformed to a minimal series-parallel graph~\cite{Valdes:1979:RSP:800135.804393}.
-As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a DAG is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape} illustrated in Figure~\ref{fig:n}, is not found in the DAG.
+As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a DAG is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape} and illustrated in Figure~\ref{fig:n}, is not found in the DAG.
 Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes} by an over-constraint, as illustrated in Figure~\ref{fig:over}.
-
-\todo[inline]{JB: dans le DAG il y a une dependance de c2 a sync4, le graphe serie parallele les execute en parallele ...}
 
@@
   \path[->]
-    (bc0) edge [dotted] node {} (c0)
-    (bc2) edge [dotted] node {} (c2)
+    %(bc0) edge [dotted] node {} (c0)
+    %(bc2) edge [dotted] node {} (c2)
     (c0) edge node {} (c1)
@@
 
+\subsubsection*{Data parallelism optimizations: $\Gamma_{data}$}
+Building the series-parallel tree decomposition $\Gamma_{tsp}$ is not only limited to the creation of a static scheduling of tasks.
+Actually, the tree can also be used to optimize the automatic generation of the data parallel structure of the simulation.
+In fact, creating a data parallelization of the simulation, it first seems natural to create the parallel structure directly from $\Gamma_{sync}$, as the program is parallelized from data decomposition and applying the same sequential algorithm on each processor.
+However, as a computation kernel in MSL writes a single data $w$ at a time (in order to entirely manage and order the dependencies between tasks), the same computation domain might be iterated more than once to compute different numerical expressions, while those expressions could be computed inside a single space loop.
+For this reason, an optimized and \emph{fusionned} ordered list of computation $\Gamma_{data}$ is built from $\Gamma_{tsp}$.
+Actually, in $\Gamma_{tsp}$, computations of a parallel subtree with a same computation domain can be computed in the same space loop, in any order.
+Moreover, in $\Gamma_{tsp}$, two consecutive computations of a sequence subtree with a same computation domain can be computed in the same space loop, if the order of computation of the sequence is kept.
+
 % \paragraph{Dump to a component assembly} Finally, from the series-parallel tree decomposition $\Gamma_{tsp}$ we could build a program using any parallel language containing \emph{parallel sections} and \emph{sequences of instructions} such as OpenMP~\cite{660313}, HPF~\cite{219857}, UPC~\cite{El-Ghazawi:2006:UUP:1188455.1188483} etc.

### 2015-08-06 WOLFHPC15/article/mscac.tex
wolfhpc fix all parts helene 06_08

@@
 Agnostic means that the description of a numerical simulation does not depend on implementation choices as, for example, the type of mesh (structured, unstructured), or its associated interfaces (how to define a stencil shape, etc.).
-Descriptive, on the other hand, means that the terminal states of the language grammar are (string) identifiers.
-As a result, MSL is not made to handle the expression of numerical computations, but only the description of a multi-stencil simulation, in a coarse-grain fashion.
+Descriptive, on the other hand, means that MSL is not made to handle the expression of numerical computations, but only the description of a multi-stencil simulation, in a coarse-grain fashion.
 %Numerical computations are encapsulated in specific components that users have to develop.
 A MSL description is made of six main sections that match the six tuples elements of the definition of a multi-stencil program of Equation~\ref{eq:msp}.
-\todo[inline]{JB: dans une grammaire, les identifiants sont toujours des chaines de caracteres, je ne comprends pas}
 
@@
 
-The final phase of compilation, the dump, uses the static scheduling of the tasks, created from $\Gamma$, and additional informations on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{E}$ and $\mathcal{D}$.
+The final phase of compilation, the dump, uses the static scheduling of the tasks, created from $\Gamma$, and additional information on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{E}$ and $\mathcal{D}$.
 It generates a program in the back-end language which embeds the parallel pattern or skeleton of the overall simulation.
@@
 
-Thanks to the software components based approach, the generated code is made of five independent parts that loosely match the elements of the multi-stencil program definition.
-These are the components implementing: \textbf{a)} the scheduling ($\Gamma$), \textbf{b)} the data structures ($\mathcal{M}$ and $\mathcal{E}$), \textbf{c)} their parallelization ($\Delta$ and $\mathcal{D}$), \textbf{d)} the time loop ($T$), and \textbf{e)} the kernels directly provided by the user.
-This enables to change the technology used for each part while limiting the impact on the other parts as further discussed in Section~\ref{sect:component}.
+Thanks to the software components based approach, the generated code is made of four independent parts that loosely match the elements of the multi-stencil program definition.
+These are the components implementing: \textbf{a)} the computations scheduling ($\Gamma$), \textbf{b)} the data structures ($\mathcal{M}$ and $\mathcal{E}$), \textbf{c)} data of the simulation ($\Delta$ and $\mathcal{D}$), and \textbf{d)} the time loop ($T$).
+The actual implementation of the numerical expressions are provided by the user and used in the computations scheduling.
+This component-based parallel structure enables to change the technology used for each part while limiting the impact on the other parts as further discussed in Section~\ref{sect:component}.
 

### 2015-08-06 WOLFHPC15/article/related.tex
wolfhpc fix all parts helene 06_08

@@
 Many specific solutions exist to ease parallel programming of numerical simulations.
-Solutions such as PETSc~\cite{petsc-efficient} and Trilinos~\cite{Trilinos-Overview} are specific libraries.
+Solutions such as PETSc~\cite{petsc-efficient} and Trilinos~\cite{Trilinos-Overview} are specific frameworks.
 They are made to solve many kinds of PDEs using well-known methods and solvers which are efficiently parallelized inside the library.
-Those libraries also include solvers for implicit numerical schemes, not addressed in this work.
+Those frameworks also include solvers for implicit numerical schemes, not addressed in this work.
 However, using those solutions the user has to use and follow existing tools.
 In other words, it is not possible for a mathematician or a physician to write his own numerical method if it differs from existing ones.
-Moreover as PETSc and Trilinos are libraries they do not propose a domain specific language to define computations.
+Moreover as PETSc and Trilinos are frameworks they do not propose a domain specific language to define computations.
 The purpose of those solutions are much more specific than our solution.
-\todo[inline]{JB: j'utiliserais plutot le terme "framework" que "library"}
 
@@
 However, many real case numerical simulation are not composed of a single stencil kernel, as explained in Section~\ref{sect:multistencil}, but of more than one stencil computation, involving one or more stencil shapes, and additional local computations.
-Solutions like Pochoir or PATUS handle the parallelization or the optimization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations.
+%Solutions like Pochoir or PATUS handle the parallelization or the optimization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations.
 Liszt~\cite{DeVito2011LDS} is the closest solution to our work as it seems to take into account the entire program and not only stencil kernels.

### 2015-08-06 WOLFHPC15/article/ms_msc.tex
wolfhpc fix language figures helene 06_08

@@
 
-\begin{figure}[t]
-\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++,numbers=left]
+\begin{filecontents*}{grammar.txt}
 program ::= "mesh:" meshid 
@@
           |  dataid
-\end{lstlisting}
-\caption{Grammar of the Multi-Stencil Language}
-\label{fig:grammar}
-\end{figure}
+\end{filecontents*}
+
+\begin{filecontents*}{example.txt}
+mesh: cart
+mesh domains: cell,edgex,edgey
+computation domains:
+  allcell in cell
+  alledgex in edgex
+  alledgey in edgey
+  part1edgex in edgex
+  part2edgex in edgex
+independent:
+  part1edgex and part2edgex
+data:
+  a,cell
+  b,cell
+  c,edgex
+  d,edgex
+  e,edgey
+  f,cell
+  g,edgey
+  h,edgex
+  i,cell
+  j,edgex
+time:500
+computations:
+  b[allcell]=c0(a)
+  c[alledgex]=c1(b[n1])
+  d[alledgex]=c2(c)
+  e[alledgey]=c3(c)
+  f[allcell]=c4(d[n1])
+  g[alledgey]=c5(e)
+  h[alledgex]=c6(f)
+  i[allcell]=c7(g,h)
+  j[partedgex]=c8(i[n1])
+\end{filecontents*}
+
+\begin{figure*}[t]
+\subfloat[\label{fig:grammar}]
+{\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left,linewidth=0.45\textwidth]{grammar.txt}}
+%\hfill
+\hspace{1.5cm}
+\subfloat[\label{fig:mslex}]
+{\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left,linewidth=0.45\textwidth]{example.txt}}
+\caption{(a) Grammar of the Multi-Stencil Language, (b) Example of a MSL program}
+\end{figure*}
 
@@
 On line 24, the data element \texttt{b} is computed on the domain \texttt{allcell} by the kernel expression \texttt{c\_0} which read data \texttt{a} without neighborhood shape.
-
-\begin{figure}[t]
-\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++,numbers=left]
-mesh: cart
-mesh domains: cell,edgex,edgey
-computation domains:
-  allcell in cell
-  alledgex in edgex
-  alledgey in edgey
-  part1edgex in edgex
-  part2edgex in edgex
-independent:
-  part1edgex and part2edgex
-data:
-  a,cell
-  b,cell
-  c,edgex
-  d,edgex
-  e,edgey
-  f,cell
-  g,edgey
-  h,edgex
-  i,cell
-  j,edgex
-time:500
-computations:
-  b[allcell]=c0(a)
-  c[alledgex]=c1(b[n1])
-  d[alledgex]=c2(c)
-  e[alledgey]=c3(c)
-  f[allcell]=c4(d[n1])
-  g[alledgey]=c5(e)
-  h[alledgex]=c6(f)
-  i[allcell]=c7(g,h)
-  j[partedgex]=c8(i[n1])
-\end{lstlisting}
-\caption{Example of a MSL program}
-\label{fig:mslex}
-\end{figure}
 

### 2015-08-06 WOLFHPC15/article/evaluation.tex
evaluations helene 06_08

@@
 
-The MSL program of FullSWOF contains three mesh entities, seven computation domains, fourty-eight data and ninty-eight computations composed of thirty-two stencil kernels and sixty-six local computations.
+\subsection*{Compiler Evaluation}
 
@@
 Execution times have been computed on a laptop with a bi-core Intel Core i5 1,4 GHz, and 8Gb LPDDR3.
-
 \begin{figure}[!h]
@@
 \end{figure}
-
 One can notice that the transformation of $\Gamma_{dep}$ to a minimal series-parallel graph is the longest step of MSC, because of the removal of the forbidden shapes in the graph.
@@
 
-\begin{itemize}
-\item speedup Figures
-\item number of lines (counting computation components)
-\end{itemize}
+\subsection*{Preliminary Performances}
+To evaluate if performances of the overall automatic parallelization are promising we have proceeded as follows.
+Our current implementation of the MSCAC compiler generates a back-end code using the SkelGIS library.
+For this reason, our evaluation compare the shallow water equations first parallelized with a pure SkelGIS code (data structures, applicators and interfaces of SkelGIS~\cite{CPE:CPE3494}), and second parallelized with MSL (which uses the SkelGIS data structure).
+As the SkelGIS library has proved its scalability on the Shallow water equations compared to an MPI parallelization, the evaluation is relevant to evaluate scalability of MSL.
+Moreover, as SkelGIS library handles data parallelization for distributed memory architectures, we have limited for now our evaluations to the fusionned data parallelization of MSCAC (dumped from $\Gamma_{data}$).
+
+\fix{HC : preciser la machine}
+
+Evaluations have been performed on ...
+.
+Figure~\ref{fig:times} shows the execution times in seconds as a function of the number of cores.
+Figure~\ref{fig:speedup} illustrates a speedup comparison, using the minimum sequential reference time (which is the MSL sequential time).
+Finally, Figure~\ref{fig:log2} illustrates the logarithmic scale of execution times as a function of the logarithmic scale of the number of cores.
+\begin{figure}[!h]
+ \begin{center}
+ \begin{tabular}{c|c|c|c|c|c|c}
+   cores & 8 & 16 & 32 & 64 & 128 & 256\\
+   \hline
+   MSL & 998.10 & 544.86 & 285.73 & 139.8 & 76.45 & 48.05\\
+   \hline
+   SkelGIS & 1324.9 & 638.5 & 371.62 & 197.32 & 117.05 & 64.96\\
+ \end{tabular}
+\caption{Execution times for the shallow water equations using MSL and SkelGIS (in seconds).}
+\label{fig:times}
+ \end{center}
+\end{figure}
+\begin{figure*}
+\begin{center}
+\subfloat[\label{fig:speedup}]{
+\resizebox{8cm}{!}{\includegraphics{./images/speedup.pdf}}
+}
+\hspace{10pt}
+\subfloat[\label{fig:log2}]{
+\resizebox{8cm}{!}{\includegraphics{./images/logtime.pdf}}
+}
+\end{center}
+\caption{(a) Speedup comparison, and (b) logarithmic execution times, both for the Shallow water equations, using pure SkelGIS and MSL with the SkelGIS data structure.}
+\label{fig:perfs}
+\end{figure*}
+
+One can notive that execution times using MSL, which itself uses the SkelGIS distributed data structures, is improved compared to a pure SkelGIS parallelization.
+Moreover, the scalability stays close to the scalability of SkelGIS even if the makespan is improved.

### 2015-08-06 rapport/formalism.tex
WIP relecture rapport recherche

@@
 
-A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ of a set of PDEs and is defined as followed.
+A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ of a set of PDEs and is defined as follows.
 
 \begin{mydef}
-\textit{A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V$ is the set of vertices and $E$ the set of edges.
+\textit{A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V\in \Omega$ is the set of vertices and $E\in V^2$ the set of edges.
 The set of edges $E$ of a mesh $\mathcal{M}=(V,E)$ does not contain bridges.}
 \end{mydef}
-
-\rmq{Pas très clair la suite; motivation? dom == some sub elements of M ?}
-\begin{mydef}
-$D_i$ is a set of elements of a mesh $\mathcal{M}=(V,E)$, constructed by a function $dom_i$ which defines a precise association between $V$ and $E$, $dom_i : V \times E \rightarrow D_i$.
-\end{mydef}
-For example, the set of cells $D_0$ in a Cartesian 2D mesh could be defined by exactly four vertices and four edges connected as a cycle.
-But we could also define another set of elements $D_1$ as the simple set of vertices $V$, etc.
-
-A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) and hybrid.
-
 \begin{figure}[!h]\begin{center}
@@
 \end{center}\end{figure}
+A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) and hybrid.
+
+%\rmq{Pas très clair la suite; motivation? dom == some sub elements of M ?}
+For simulations, data is discretized based on the mesh; but for a given mesh, there are various ways of mapping the data depending on the mesh entities on which the data is mapped.
+\begin{mydef}
+\todo{JB: vraiment, ta def dit que pour chaque couple (noeud, arete), tu associes un element de $D_i$ ...}
+$D_i$ is a set of elements of a mesh $\mathcal{M}=(V,E)$, constructed by a function $dom_i$ which defines a precise association between $V$ and $E$, $dom_i : V \times E \rightarrow D_i$.
+\end{mydef}
 
 \begin{mydef}
+\todo{JB: def alternative}
+An entity $d$ of a mesh $\mathcal{M}=(V,E)$ is a subset of the mesh elements: $d\subset (V\cup E)$.
+A kind of mesh entities $D$ is a set of entities: $D\subset \mathcal{P}(V\cup E)$
+\end{mydef}
+For example, in a 2D Cartesian mesh a first kind of entities are the cells ($D_0$), they can be defined as all sets containing exactly four vertices and four edges connected as a cycle.
+Another kind of entities are the vertices ($D_1$) that can be defined as all singletons formed of a single vertex each.
+
+This covers the space discretization, however another dimension manipulated in simulations that has to be discretized is the time dimension.
+\begin{mydef}
 The discretization of the continuous time domain $\mathcal{T}$ is denoted $T$ such that $\forall\mbox{ }t_i\mbox{, }t_{i+1} \in T\mbox{, }\exists\mbox{ }\Delta t \in \mathbb{R}$\mbox{, }$t_{i+1} = t_i + \Delta t$.
-Thus, $T$ is responsible for the iteration time steps of the numerical simulation.
 \end{mydef}
+$T$ is responsible for the iteration time steps of the numerical simulation.
 
-In a numerical simulation a set of data, or quantities, are applied onto the mesh and represent the set of values to compute, or to use, for computation.
-
+In a numerical simulation a set of data elements, or quantities, are applied onto the mesh and represent the set of values to compute, or to use, for computation.
 \begin{mydef}
-The set of data applied on the mesh is denoted by $\Delta$, such that $\delta \in \Delta$ is a function which associates each element $d \in D_i$ (the domain it is applied on) to a value $v \in V$, $\delta : D_i \rightarrow V$.
-In the rest of this paper, the domain of a data $\delta$ can be given by the function $domain(\delta)=D_i$.
+A quantity is a function $\delta_t: D_{\delta t} \mapsto V_{\delta t}$ which associates each entity of its definition domain $d\in D_{\delta t}$ to a value $v\in V_{\delta t}$ where $V_{\delta t}$ is the data type, typically $\mathbb{R}$ or $\mathbb{C}$.
+The set of quantities applied on the mesh is denoted $\Delta$.
+In the rest of this paper, the entities on which a quantity $\delta_t$ is mapped is denoted $dom(\delta_t)=D_{\delta t}$.
 \end{mydef}
-One can notice that in applied mathematics, the signature of $\delta$ would be $\delta : D_i \times T \rightarrow V$, however when programming a numerical simulation it is not wise to store all values of each time iteration.
+Another approach closer to that taken in applied mathematics would have been to define $\delta$ functions over $D \times T$; however the approach chosen in this paper where a distinct function is defined for each time-step better matches the implementation of numerical simulations where the outer loop iterates over time and where only a few time-steps of quantity values are stored.
 
@@
 \begin{mydef}
-A numerical expression $\text{exp}$ is a function which represents how to compute a value for an element $d \in domain(w)=D_i$, using the set $R \subset \Delta$ of input data (read data), $\text{exp} : R \times D_i \rightarrow w \times D_i$.
+A numerical expression $\text{exp}_{wt}: dom(w_t)\times R\mapsto V_{wt}$ is a function that computes the value of a quantity $w_t(d)$ (write quantity) for a given entity $d \in dom(w_t)=D$, using the set of its input data $R=\{\delta 1_{t-1}, \delta 2_{t-1}, ...\} \subset \Delta$ (read quantities).
+It is defined as $\text{exp}_{wt}(d,r) \rightarrow w_t(d)$.
 \end{mydef}
+
+%JB: fin de relecture
 
@@
 Figure~\ref{fig:ex} gives an example of a stencil computation $s(R,w,\text{exp},\mathcal{N})$, where $\mathcal{M}(V,E)$ is a two dimensional Cartesian mesh.
-A single domain $D=domain(w)$ is defined in this example and is composed of cells formed by a cycle of four vertices $v \in V$ and four edges $e \in E$.
+A single domain $D=dom(w)$ is defined in this example and is composed of cells formed by a cycle of four vertices $v \in V$ and four edges $e \in E$.
 Furthermore, in this example $R=\{A\}$, $w=B$, and for $(x,y) \in D$ the neighborhood function is

### 2015-08-06 rapport/formalism.tex
un tout petit peu de relecture du rapport

@@
 It is defined as $\text{exp}_{wt}(d,r) \rightarrow w_t(d)$.
+\todo{JB: tu n'introduis pas la notion de stencil ici?}
 \end{mydef}
-
-%JB: fin de relecture
 
@@
 \end{mydef}
-It has to be noticed that at each time iteration, all the elements of a mesh are computed.
-However, it happens that computations of the mesh elements are splitted in different domains, as for example the computation of the physical border.
+It has to be noticed that at each time iteration, all the values of a quantity are computed.
+However, it happens that computations of the quantity is split in different domains, as for example the computation of the physical border.
 In this case additional $D_i$ can be specified for the mesh $\mathcal{M}$.
+\todo{JB: je comprends pas vraiment...}
 
@@
 The set of $n$ ordered computations of a numerical simulation is denoted $\Gamma = [c_i]_{0 \leq i \leq n-1}$, such that $\forall c_i,c_j \in \Gamma$, if $i \leq j$, then $c_i$ is computed before $c_j$, and $c_j$ can be computed only when $c_i$ is finished.
+\todo{JB: je dirais que $c_j$ peut etre calculé quand $c_j$ l'est, j'eleverais le "only", c'est justement parce qu'il n'y a pas le only que tu peux faire du task parallelism}
 \end{mydef}
+
+%JB: fin de relecture
 

### 2015-08-07 WOLFHPC15/article/evaluation.tex
fix helene 07_08

@@
 Navier-Stokes equations is a well known set of PDEs in fluid dynamics to simulate a flow evolution in time.
-At the University of Orl\'eans, France, the MAPMO laboratory works on a software, called FullSWOF~\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}, which handles simulations of geophysical flows, for example rivers, channels, dambreak problems, ocean currents, estuarine systems, etc.
-This software solves the Shallow water equations obtained from the three dimensional incompressible Navier-Stokes equations, by averaging on the vertical direction (see e.g.~\cite{Ferrari2004}).
+At the University of Orl\'eans, France, the MAPMO laboratory works on a software, called FullSWOF~\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}, which solves the Shallow water equations obtained from the three dimensional Navier-Stokes equations, by averaging on the vertical direction (see e.g.~\cite{Ferrari2004}).
 Those equations are solved using a two-dimensional Cartesian discretization of the space domain, and a finite volume numerical methods more described in~\cite{HeleneLS13}.
 
-\subsection*{Compiler Evaluation}
+\subsubsection*{Compiler Evaluation}
 
@@
 
-\subsection*{Preliminary Performances}
+\subsubsection*{Preliminary Performances}
 To evaluate if performances of the overall automatic parallelization are promising we have proceeded as follows.

### 2015-08-07 WOLFHPC15/article/ms_msc.tex
fix helene 07_08

@@
 \subsubsection*{Data parallelism optimizations: $\Gamma_{data}$}
-Building the series-parallel tree decomposition $\Gamma_{tsp}$ is not only limited to the creation of a static scheduling of tasks.
-Actually, the tree can also be used to optimize the automatic generation of the data parallel structure of the simulation.
-In fact, creating a data parallelization of the simulation, it first seems natural to create the parallel structure directly from $\Gamma_{sync}$, as the program is parallelized from data decomposition and applying the same sequential algorithm on each processor.
-However, as a computation kernel in MSL writes a single data $w$ at a time (in order to entirely manage and order the dependencies between tasks), the same computation domain might be iterated more than once to compute different numerical expressions, while those expressions could be computed inside a single space loop.
+The series-parallel tree decomposition $\Gamma_{tsp}$ is used to build the hybrid parallelization of the simulation, but it can also be used to optimize the data parallelization only.
+In fact, creating a data parallelization of the simulation, it first seems natural to create the parallel structure directly from $\Gamma_{sync}$, as the program is parallelized from data decomposition and applying the same sequential tasks on each processor.
+However, as a computation kernel writes a single data $w$ (to finely manage dependencies between tasks), using $\Gamma_{sync}$ the same computation domain might be iterated more than once, while some numerical expressions could be computed inside a joint space loop.
 For this reason, an optimized and \emph{fusionned} ordered list of computation $\Gamma_{data}$ is built from $\Gamma_{tsp}$.
 Actually, in $\Gamma_{tsp}$, computations of a parallel subtree with a same computation domain can be computed in the same space loop, in any order.
-Moreover, in $\Gamma_{tsp}$, two consecutive computations of a sequence subtree with a same computation domain can be computed in the same space loop, if the order of computation of the sequence is kept.
+Moreover, in $\Gamma_{tsp}$, two consecutive computations of a sequence subtree with a same computation domain can be computed in the same space loop, keeping the computation order of the sequence.
 

### 2015-08-07 WOLFHPC15/article/mscac.tex
fix helene 07_08

@@
   \node (MSL) {MSL};
-  \node[outer,below=0.5 of MSL,draw=black] (MSCAC) {
+  \node[outer,right=0.5cm of MSL,draw=black] (MSCAC) {
     \begin{tikzpicture}
       \node [inner,draw=black,thick] (pars)  {Parser};
-      \node [inner,draw=black,thick,below=0.5cm of pars] (MSC)  {MSC};
-      \node [inner,below=0.5cm of MSC,draw=black,thick] (Dump)  {Dump};
+      \node [inner,draw=black,thick,right=0.5cm of pars] (MSC)  {MSC};
+      \node [inner,right=0.5cm of MSC,draw=black,thick] (Dump)  {Dump};
 
@@
   };
-  \node [below=0.5 of MSCAC] (CA) {Component-based runtime};
-  \node [rotate=90,above left=0.5 of MSC] (CAid) {MSCAC};
-  \draw[->] (MSL) -- (pars.north);
-  \draw[->] (Dump.south) -- (CA);
+  \node [right=0.5cm of MSCAC] (CA) {CBR};
+  \node [above =0.2cm of MSCAC] (CAid) {MSCAC};
+  \draw[->] (MSL) -- (pars.west);
+  \draw[->] (Dump.east) -- (CA);
 \end{tikzpicture}
-\caption{The three phases of the MSCAC compiler.}
+\caption{The three phases of the MSCAC compiler. First the parser of the MSL description, second the MSC parallelization, third the dump to the Component-Based Runtime (CBR).}
 \label{fig:mscac}

### 2015-08-08 WOLFHPC15/article/component.tex
wolfhpc fix scheduling figure helene 06_08
fix helene 07_08
fix component.tex helene 08_08

@@
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (C) at (0,0) {$C_0$};
+   \node[component] (C) at (0,0) {$c_0$};
    \node[provide] (p) at (-1.5,0) {};
@@
    \node[provide,right=1.5cm of u] (p1) {};
-   \node[component,right=1.5cm of p1] (C1) {$C_1$};
+   \node[component,right=1.5cm of p1] (C1) {$c_1$};
    \node[use,right=1.5cm of C1] (um) {$m$};
@@
   \path[-]
-    (p) edge node {$r$} (C)
-    (C) edge node {$m_1$} (u)
-    (u) edge node {$m_2$} (C1);
+    (p) edge node {} (C)
+    (C) edge node {} (u)
+    (u) edge node {} (C1);
 \end{tikzpicture}
@@
    \node[use,below=1.75cm of D] (Du2) {};
-   \node[use,right=1.7cm of Du2] (Du3) {$m$};
+   \node[use,below=0.8cm of Du1] (Du3) {$m$};
 
@@
 
-   \node[provide,right=0.2 of Du3] (Datap1) {};
+   \node[provide,below right=0.2 of Du3] (Datap1) {};
    \node[provide,above=0.15 of Cu] (Datap2) {};
-   \node[component,above=0.8cm of Datap2] (Data) {$Data$};
+   \node[component,double,above=0.8cm of Datap2] (Data) {$Data$};
    \node[use,above=0.8cm of Data] (Datau) {};
-   \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
-   \node[mpi,above right=1cm of Data] (mpidata) {};
+   \node[left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
+   \node[mpi,right=1cm of Data] (mpidata) {};
+   \node[above=0.2cm of mpidata] (star2) {$*$};
 
@@
 As described in Section~\ref{sect:mscac}, the compiler generates four independent parts in a component assembly in addition to a static part.
-This is represented on Figure~\ref{fig:mscac:assembly} where each rounded box represents one or more component instances.
+This is represented on Figure~\ref{fig:mscac:assembly} where each rounded box represents one or more component instances, respectively drawn with simple lines or double lines.
+For example in the figure, the component \texttt{Data} is instanciated more than once.
 \texttt{Driver} is static and drives the whole application.
-The four other parts are generated based on the content of sections of the MSL program: \texttt{Computations} from $\Gamma$ (with code provided by the user), \emph{DDS} from $\mathcal{M}$ and $\mathcal{E}$, \emph{Data} from $\Delta$ and $\mathcal{D}$, and \emph{Time} from $T$.
+The four other parts are generated based on the content of the different sections of the MSL program: \texttt{Computations}, from the parallelization of $\Gamma$ (done by MSC as explained in Section~\ref{sect:msc}), \emph{DDS} from $\mathcal{M}$ and $\mathcal{E}$, \emph{Data} from $\Delta$ and $\mathcal{D}$, and \emph{Time} from $T$.
 
-\texttt{DDS} manages the structure of the mesh, a single instance is generated as MSL only supports a single mesh for now.
+%, and using the kernel codes provided by the user
+
+\texttt{DDS} manages the structure and the partitioning of the mesh, a single instance is generated as MSL only supports a single mesh for now.
 The back-end presented in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}.
-\texttt{DDS} is used by \texttt{Data} containing one component instance for each data element in the MSL program.
-\texttt{Time} is parametrized by the the \texttt{iteration} value from MSL and calls the \texttt{Computations} part as many times as required.
+\texttt{DDS} is used by \texttt{Data}, which represents one component instance for each data element in the MSL program.
+\texttt{Time} is parametrized by the the \texttt{iteration} value from MSL and uses the \texttt{Computations} component as many times as required.
 
-\texttt{Computations} contains the static scheduling computed by MSC, \ie $\Gamma_{tsp}$, or the data parallel transformation $\Gamma_{data}$.
-Both transformations $\Gamma_{tsp}$ and $\Gamma_{data}$ are encoded using three specific components which manage the $P$, $S$, and $sync$ operations as explained hereafter.
-At the leaves of the tree, the computation components instances ($K$), provided by the user, are connected to \texttt{Data}.
-As presented on Figure~\ref{fig:k}, these components provide a single port used to launch the computation and expose one use port for each data element manipulated in the computation.
-We denote the fact that they may have several use port by adding a star on the use port.
-Each of these use ports is connected to the matching component instance in \texttt{Data}.
-We denote this kind of component K.
-
-As a result of this approach, each part of the generated code is rather independent.
-Changing the main loop to use a convergence criteria rather than a fixed number of iterations would only require changing the \texttt{Time} component.
-Similarly, changing the approach used for scheduling would only impact the code generated in \texttt{Computations}.
-The technology used for data parallelism can also be changed by replacing the \texttt{DDS} and \texttt{Data} components; as a matter of fact a work in progress proposes an alternative based on Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} instead of SkelGIS.
-In this case however, the kernel components of the \texttt{Computations} part, provided by the user, might have to be written with interfaces of the alternative distributed data structure.
+\texttt{Computations} contains the static scheduling $\Gamma_{tsp}$, or the data parallel transformation $\Gamma_{data}$, both computed by MSC.
+Both transformations, $\Gamma_{tsp}$ and $\Gamma_{data}$, are encoded using three specific components,detailed in Section~\ref{sect:control}, which manage the $P$, $S$, and $sync$ operations.
+At the leaves of $\Gamma_{tsp}$ and $\Gamma_{data}$, the computation components instances (denoted $K$), provided by the user, are connected to \texttt{Data}.
+As presented on Figure~\ref{fig:k}, these $K$ components provide a single port used to launch the computation and expose one use port for each data element manipulated in the computation.
+We denote the fact that they may have several use port by adding a star on the use port to \texttt{Data}.
 
@@
 
+Moreover, to finalize the data parallelization of the back-end component-based runtime, which is partially managed by \texttt{DDS} for the partitioning, and $sync$ operations in \texttt{Computations}, we have chosen to duplicate the component assembly of Figure~\ref{fig:mscac:assembly} on available processors (or cores), connecting them by MPI ports between their \texttt{Data}.
+
+As a result of this approach, each part of the generated code is rather independent.
+Changing the main loop to use a convergence criteria rather than a fixed number of iterations would only require changing the \texttt{Time} component.
+Similarly, changing the approach used for scheduling would only impact the code generated in \texttt{Computations}.
+The technology used for data parallelism can also be changed by replacing the \texttt{DDS} and \texttt{Data} components; and by using the adequate interfaces in $K$.
+As a matter of fact, a work in progress proposes an alternative based on Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} instead of SkelGIS.
+
 %----------------------------------------
 \subsection{Control Components and Dump}
+\label{sect:control}
 %----------------------------------------
 The series-parallel tree decomposition $\Gamma_{tsp}$ represents a static scheduling of the simulation.
-It is dumped to components by generating an assembly that exactly matches the tree, as well as for $\Gamma_{data}$.
+$\Gamma_{data}$, on the other hand, is an optimized fusionned ordered list of computations.
+Both are dumped to components by generating an assembly that exactly matches their structure.
 We introduce what we call \emph{control components} to represent all nodes of $\Gamma_{tsp}$ or $\Gamma_{data}$.
 These components can be used for any simulation, which increases code reuse between simulations.
-A control component exposes a single \emph{provide} port containing a single method, typically a control-only method without any parameter.
+A control component exposes a single \emph{provide} port containing a \emph{control-only} method, without any parameter.
 We introduce three control components represented in Figure~\ref{fig:ctrlcomponents}.
@@
 \begin{description}
-\item[Sequence component (SEQ)] It is the direct representation of a sequence node of $\Gamma_{tsp}$.
+\item[Sequence component (SEQ)] It is the direct representation of a sequence node of $\Gamma_{tsp}$ or $\Gamma_{data}$.
 This component sequentially calls an ordered list of other components.
@@
 
-\item[Parallel component (PAR)] It is the direct representation of a parallel node of $\Gamma_{tsp}$.
+\item[Parallel component (PAR)] It is the direct representation of a parallel node of $\Gamma_{tsp}$ (not $\Gamma_{data}$).
 This component simultaneously calls a set of other components.
@@
 \item[Synchronization component (SYNC)].
-It is the direct representation to an update leaf of $\Gamma_{tsp}$.
+It is the direct representation to an update leaf of $\Gamma_{tsp}$ or $\Gamma_{data}$.
 This component calls the synchronization of a given data.
 It exposes a \emph{use} port to be connected to the data to update.
-
 \end{description}
 
-%----------------------------------------
-%\subsection{Dump To Component Assembly}
-%----------------------------------------
 Then from $\Gamma_{tsp}$ or $\Gamma_{data}$, and using the control components described above,
@@
 and task) parallelization of a simulation.
-%However, it is also
-%possible to dump to a data parallelization only. In such a case the
-%computation of $\Gamma_{sync}$ is enough to generate the component
-%assembly. Actually a single SEQ component is thus needed and this
-%component is linked to all computations and synchronizations of
-%$\Gamma_{sync}$ directly.
-
 Figure~\ref{fig:tsp:assembly} displays the assembly part corresponding to
@@
 the identification name of the computation in the MSL file.
+Thus, this figure represents what the component \texttt{Computations} of Figure~\ref{fig:mscac:assembly} contains.
 
@@
    %c1
-   \node[component] (c1) [below = 0.8cm of sync0] {$K(c_1^*)$};
+   \node[component] (c1) [below = 0.8cm of sync0] {$K(c_1)$};
    \node[use] (c1u) [right = 0.8cm of c1] {};
@@
   %c4
-   \node[component] (c4) [below = 0.8cm of sync1] {$K(c_4^*)$};
+   \node[component] (c4) [below = 0.8cm of sync1] {$K(c_4)$};
    \node[use] (c4u) [right = 0.8cm of c4] {};
@@
   %c8
-   \node[component] (c8) [below = 0.8cm of sync2] {$K(c_8^*)$};
+   \node[component] (c8) [below = 0.8cm of sync2] {$K(c_8)$};
    \node[use] (c8u) [right = 0.8cm of c8] {};

### 2015-08-08 WOLFHPC15/article/evaluation.tex
fix all helene 08_08

@@
 
-The series-parallel tree decomposition extracted by the MSC transformation is composed of seventeen sequences and eighteen parallel sections.
-Figure~\ref{fig:freq} represents for a given level of parallelism, \ie the number of tasks which can be executed concurrently, the number of time this level is observed in the final component assembly.
-One can notive that the level of task parallelism extracted from the Shallow water equations is limited by at the two sequential parts of the application (level 1).
-However, as a level of sixteen parallel tasks is reach too times, but also five times for the level six, sequential restrictions could be amortize.
-Moreover, as the level of parallelism in the application is heterogenous, the choice of the number of threads to launch for task parallelism is difficult.
+The series-parallel tree decomposition $\Gamma_{tsp}$ of this simulation, extracted by the MSC transformation, is composed of seventeen sequence nodes and eighteen parallel nodes.
+Figure~\ref{fig:freq} represents for a given level of parallelism, \ie the number of tasks to perform concurrently, the number of time this level is observed in the final component assembly.
+One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1).
+As a level of sixteen parallel tasks is reached too times, and also five times for the level six, sequential restrictions could be amortized.
+However, it is rather clear that the task parallelization technique should be used to improve the data parallelism when reaching at its limits, but not to directly get performance.
+Moreover, as the level of parallelism in the application is heterogenous, the number of threads to launch for task parallelism, and the number of cores to keep for data parallelism is difficult to choose.
 
@@
 
-Figure~\ref{fig:exectime} illustrates the execution time for each step of the MSC transformation for an overall execution time of ten seconds.
-Execution times have been computed on a laptop with a bi-core Intel Core i5 1,4 GHz, and 8Gb LPDDR3.
 \begin{figure}[!h]
@@
 \end{figure}
+
+Figure~\ref{fig:exectime} illustrates the execution time for each step of the MSC transformation for an overall execution time of ten seconds.
+Execution times have been computed on a laptop with a bi-core Intel Core i5 1,4 GHz, and 8Gb DDR3.
 One can notice that the transformation of $\Gamma_{dep}$ to a minimal series-parallel graph is the longest step of MSC, because of the removal of the forbidden shapes in the graph.
-Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted, because the algorithm uses a general solution instead of finding each forbidden shape, but it seems that many of them appeared.
+Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted by the algorithm, which applies a general solution (instead of finding each forbidden shape), but it seems that many of them appear.
 The \emph{N-shape}, represented in Figure~\ref{fig:n} is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.
 The fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static shedule of tasks may not be the best solution for complex simulations.
-Actually, if a \emph{N-shape} cannot be represented by sequences and parallel sections, this shape can perfectly be handled by a dynamic scheduler or by the direct static representation of $\Gamma_{dep}$ where tasks wait for their input before starting.
+Actually, a \emph{N-shape} could perfectly be handled by a dynamic scheduler for example.
+It has to be added, however, that the computation of $\Gamma_{tsp}$ is still interesting to detect how space loops can be fusionned in $\Gamma_{data}$.
 
 \subsubsection*{Preliminary Performances}
-To evaluate if performances of the overall automatic parallelization are promising we have proceeded as follows.
+To evaluate performance of the generated component-based parallel structure, we have proceeded as follows.
 Our current implementation of the MSCAC compiler generates a back-end code using the SkelGIS library.
 For this reason, our evaluation compare the shallow water equations first parallelized with a pure SkelGIS code (data structures, applicators and interfaces of SkelGIS~\cite{CPE:CPE3494}), and second parallelized with MSL (which uses the SkelGIS data structure).
-As the SkelGIS library has proved its scalability on the Shallow water equations compared to an MPI parallelization, the evaluation is relevant to evaluate scalability of MSL.
+As the SkelGIS library has proved its scalability on the Shallow water equations compared to an MPI parallelization, this performance evaluation should be relevant.
 Moreover, as SkelGIS library handles data parallelization for distributed memory architectures, we have limited for now our evaluations to the fusionned data parallelization of MSCAC (dumped from $\Gamma_{data}$).
 
-\fix{HC : preciser la machine}
+Evaluations have been performed on TGCC Curie \emph{thin nodes}.
+Each node is equiped with an Intel E5-2680 (Sandy-Bridge EP) processor containing 16 cores, 64GB of memory, and is linked to an Infiniband network.
+\fix{HC: what are the experiments presented ? Size of the mesh, number of time iterations}
 
-Evaluations have been performed on ...
-.
 Figure~\ref{fig:times} shows the execution times in seconds as a function of the number of cores.
@@
  \begin{tabular}{c|c|c|c|c|c|c}
-   cores & 8 & 16 & 32 & 64 & 128 & 256\\
+   Cores & 8 & 16 & 32 & 64 & 128 & 256\\
    \hline
@@
 
-One can notive that execution times using MSL, which itself uses the SkelGIS distributed data structures, is improved compared to a pure SkelGIS parallelization.
-Moreover, the scalability stays close to the scalability of SkelGIS even if the makespan is improved.
+One can notice that execution times using MSL, which itself uses the SkelGIS distributed data structures, are improved compared to a pure SkelGIS parallelization.
+Moreover, the scalability stays close to the scalability of SkelGIS.
+Thus, using a component-based runtime, performance of the back-end code is still HPC-oriented and does not introduce overheads.
+
+\fix{more figures, more conclusions}

### 2015-08-08 WOLFHPC15/article/introduction.tex
fix all helene 08_08

@@
 
-Many solutions to automatically optimize and parallelize partial differential equations (PDEs) solvers have been proposed in the literature, either as specific frameworks~\cite{petsc-efficient,Trilinos-Overview,CPE:CPE3494} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
+Many solutions to automatically optimize and parallelize partial differential equations (PDEs) codes have been proposed in the literature, either as specific frameworks~\cite{CPE:CPE3494} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
 Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil kernel.

### 2015-08-08 WOLFHPC15/article/related.tex
fix all helene 08_08

@@
 Many specific solutions exist to ease parallel programming of numerical simulations.
-Solutions such as PETSc~\cite{petsc-efficient} and Trilinos~\cite{Trilinos-Overview} are specific frameworks.
+%Solutions such as PETSc~\cite{petsc-efficient} and Trilinos~\cite{Trilinos-Overview} are specific frameworks.
 They are made to solve many kinds of PDEs using well-known methods and solvers which are efficiently parallelized inside the library.
@@
 The purpose of those solutions are much more specific than our solution.
-
-Other solutions try to take a step back in the abstraction to propose a way to hide parallel codes for stencils while keeping a sequential programming style.
-For example PATUS~\cite{citeulike12258902} and Pochoir~\cite{spaaTangCKLL11} are proposed to write stencil codes for general structured meshes, and OP2~\cite{Giles2011} and Liszt~\cite{DeVito2011LDS} for general unstructured meshes.
+The closest solutions to MSL are DSLs proposing to generate parallel codes from a sequential imperative stencil description.
+For example PATUS~\cite{citeulike12258902} and Pochoir~\cite{spaaTangCKLL11} propose a simple description to write stencil codes for general structured meshes, and OP2~\cite{Giles2011} and Liszt~\cite{DeVito2011LDS} for general unstructured meshes.
 Those solutions can be considered as stencil compilers.
 Actually they produce optimized (cache tiling etc.) or parallel (CUDA, OpenCL etc.) codes for a single stencil computation.
-Recently, domain specific languages for very specific numerical methods such as ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} for multigrid solvers, and OPS~\cite{Reguly:2014:ODS:2691166.2691173} for meshes of structured meshes (hybrid mesh) have also been proposed.
-All those solutions are powerful, let the user implement their own stencil codes in a sequential programming style, and produce high performance codes for shared memory architectures and sometimes distributed memory architectures.
+Recently, domain specific languages for very specific numerical methods such as ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} for multigrid solvers, and OPS~\cite{Reguly:2014:ODS:2691166.2691173} for meshes of meshes (hybrid mesh) have also been proposed.
+All those solutions let the user implement their own stencil codes in a sequential programming style, and produce high performance codes for shared memory architectures and sometimes distributed memory architectures.
+Most stencil compilers, though, handle the parallelization or the optimization of a single stencil kernel, considering that it represents the main computation time of numerical simulations.
+However, most real case numerical simulation are composed of more than one stencil kernel, involving one or more stencil shapes, and additional local kernels, as explained in Section~\ref{sect:multistencil}.
 
-Most stencil compilers handle the parallelization or the optimization of a single stencil kernel, considering that it represent the main computation time of numerical simulations.
-However, many real case numerical simulation are not composed of a single stencil kernel, as explained in Section~\ref{sect:multistencil}, but of more than one stencil computation, involving one or more stencil shapes, and additional local computations.
-%Solutions like Pochoir or PATUS handle the parallelization or the optimization of a single stencil code, considering that stencil codes represent the main computation time of numerical simulations.
-Liszt~\cite{DeVito2011LDS} is the closest solution to our work as it seems to take into account the entire program and not only stencil kernels.
+Other solutions, as SkelGIS library~\cite{CPE:CPE3494} or Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} propose more or less specialized distributed data structures (specific meshes for SkelGIS, general arrays for Global Arrays), which simplifies or totally hide communications and synchronizations between processes.
 
-The work presented in this paper takes place at a different level of numerical simulations and can be seen as complementary to stencil compilers.
+The work presented in this paper takes place at a different abstraction level of numerical simulations and can be seen as complementary to stencil compilers or distributed data structures.
 Actually, the MSL language is agnostic and should work for any type of mesh and any neighborhood shape in stencil codes.
-The MSC part of the compiler only needs this agnostic description of the overall simulation to generate a parallel pattern of the program, which needs to be filled with computation codes.
-Thus, depending on the underlying distributed data structure used, those computation codes could be generated by stencil compilers.
-As a result, MSL and its compiler are not a new contribution to compile optimized stencil codes, but propose another level of parallelization and a different approach of back-end solution.
+The MSCAC compiler only needs this agnostic description of the overall simulation to generate a parallel pattern of the program.
+This parallel pattern then needs to be filled with computation codes.
+%. Thus, depending on the underlying distributed data structure used, those computation codes 
+Computation codes could be generated by stencil compilers or could be written using interfaces of distributed data structures (as it is done using SkelGIS in this paper).
+As a result, MSL is not a new contribution to stencil compilers, or to distributed data structures, but proposes a coarse-grain automatic parallelization of an overall multi-stencil program.
+Finally, MSL is a first case study to show that component-based runtimes could improve code-reuse, separation of concerns and thus maintainability of DSLs.
 

### 2015-08-08 WOLFHPC15/article/conclusion.tex
evaluations helene 06_08
conclusion helene 08_08

@@
-Conclusion :
-\begin{itemize}
-\item Multi-stencil formalism
-\item MSL and MSC
-\item Discussion dump to components
-\item Evaluation
-\end{itemize}
+In this paper, we have presented the domain specific language MSL and its compiler MSCAC designed to produce a parallel component-based runtime of an overall multi-stencil program, \ie a mesh-based numerical simulation reduced to explicit schemes.
+MSL has been evaluated by the description of a real case numerical simulation: the shallow water equations.
+The component-based data parallelization of the simulation has been compared to a pure SkelGIS parallelization.
+Execution times of the simulation have been improved and its scalability has been shown.
+Those results demonstrate that component-based runtimes may be relevant back-end codes for DSLs.
+Moreover interests of component-based runtimes has been discussed in the paper.
 
-Perspectives :
-\begin{itemize}
-\item Dump to OpenMP for comparison
-\item MSC to dynamic scheduler (starPU, XKaapi)
-\item Formalization and abstraction of the data management in component-based applications
-\item Parallel component language
-\end{itemize}
+Although MSL is a promising case study from DSLs to component-based runtimes, many works in progress aims at improving this first contribution.
+First, to more clearly show the improvement of DSLs maintainability using component-based back-end, an alternative \texttt{DDS} component is under study, using Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.
+In addition to this, alternatives for the \texttt{Computations} component, computed by MSC, are under study such as a dump to an OpenMP~\cite{660313} code or the use of dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}.
+This last work on dynamic schedulers could also improve the expressivity of task parallelism in MSL, as it has been discussed in Section~\ref{sect:eval}, and thus could bring interesting performance for the hybrid parallelization.

### 2015-08-08 WOLFHPC15/article/introduction.tex
fix accents et taille references helene 08_08

@@
 
-Many solutions to automatically optimize and parallelize partial differential equations (PDEs) codes have been proposed in the literature, either as specific frameworks~\cite{CPE:CPE3494} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
+Many solutions to automatically optimize and parallelize partial differential equations (PDEs) codes have been proposed in the literature, either as specific frameworks~\cite{CPE:CPE3494,Trilinos-Overview,petsc-efficient} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
 Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil kernel.

### 2015-08-12 WOLFHPC15/article/evaluation.tex
fix evaluation helene 08_08
new evaluation helene 12_08

@@
 At the University of Orl\'eans, France, the MAPMO laboratory works on a software, called FullSWOF~\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}, which solves the Shallow water equations obtained from the three dimensional Navier-Stokes equations, by averaging on the vertical direction (see e.g.~\cite{Ferrari2004}).
-Those equations are solved using a two-dimensional Cartesian discretization of the space domain, and a finite volume numerical methods more described in~\cite{HeleneLS13}.
+Those equations are solved using a two-dimensional Cartesian discretization of the space domain, and a finite volume numerical methods more described in~\cite{CPE:CPE3494}.
+The MSL program of FullSWOF contains three mesh entities, seven computation domains, fourty-eight data and ninty-eight computations composed of thirty-two stencil kernels and sixty-six local computations.
 
@@
 One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1).
-As a level of sixteen parallel tasks is reached too times, and also five times for the level six, sequential restrictions could be amortized.
-However, it is rather clear that the task parallelization technique should be used to improve the data parallelism when reaching at its limits, but not to directly get performance.
+As a level of sixteen parallel tasks is reached too times, and also five times for the level four, sequential restrictions could be amortized.
+However, it is rather clear that the task parallelization technique should be used to improve the data parallelism when reaching its limits, but not to use alone.
 Moreover, as the level of parallelism in the application is heterogenous, the number of threads to launch for task parallelism, and the number of cores to keep for data parallelism is difficult to choose.
@@
 The \emph{N-shape}, represented in Figure~\ref{fig:n} is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.
-The fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static shedule of tasks may not be the best solution for complex simulations.
-Actually, a \emph{N-shape} could perfectly be handled by a dynamic scheduler for example.
-It has to be added, however, that the computation of $\Gamma_{tsp}$ is still interesting to detect how space loops can be fusionned in $\Gamma_{data}$.
+However, a \emph{N-shape} could perfectly be handled by a dynamic scheduler, for example.
+Thus, the fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static shedule of tasks may not be the best solution for complex simulations.
+It has to be notice, however, that the all computation of $\Gamma_{tsp}$ is still usefull to detect how space loops can be fusionned in $\Gamma_{data}$.
 
@@
 
-Evaluations have been performed on TGCC Curie \emph{thin nodes}.
-Each node is equiped with an Intel E5-2680 (Sandy-Bridge EP) processor containing 16 cores, 64GB of memory, and is linked to an Infiniband network.
-\fix{HC: what are the experiments presented ? Size of the mesh, number of time iterations}
+Evaluations have been performed on two clusters of \emph{Grid'5000}~\footnote{\url{https://www.grid5000.fr/mediawiki/index.php/Grid5000:Home}}.
+Figure~\ref{tab:hard} shows hardware configuration of tohse clusters.
+On the cluster \emph{Edel} (Grenoble site) have been performed three different experiments, each one described with the size of the Cartesian mesh, and the number of time iterations, as follows: (1) $5k \times 5k$, $500$ iterations; (2) $10k \times 10k$, $500$ iterations; (3) $10k \times 10k$, $2k$ iterations.
+On the cluster \emph{Paravence} (Rennes sites), the two last experiments have been performed.
 
-Figure~\ref{fig:times} shows the execution times in seconds as a function of the number of cores.
+\begin{figure}[!h]
+\begin{center}
+ \begin{tabular}{|c|c|c|}
+\hline
+    Cluster & \textbf{Edel} & \textbf{Paravance}\\
+     \hline         
+    Proc. & 2 Intel Xeon E5520 & 2 Intel Xeon E52630v3\\
+    & (2.27 GHz) & (2.4 GHz)\\
+    Cores/n & 8 & 8\\
+    Mem./n & 24 GB & 128 GB\\
+    Compiler & OpenMPI & OpenMPI\\
+    Network & InfiniBand 40G & 10 Gigabit Ethernet\\
+\hline
+ \end{tabular}
+   \caption{\label{tab:grappe}Hardware of clusters used in the experiments.}
+ \end{center}
+\end{figure}
+
+%Figure~\ref{fig:times} shows the execution times in seconds as a function of the number of cores. 
 Figure~\ref{fig:speedup} illustrates a speedup comparison, using the minimum sequential reference time (which is the MSL sequential time).
 Finally, Figure~\ref{fig:log2} illustrates the logarithmic scale of execution times as a function of the logarithmic scale of the number of cores.
-\begin{figure}[!h]
- \begin{center}
- \begin{tabular}{c|c|c|c|c|c|c}
-   Cores & 8 & 16 & 32 & 64 & 128 & 256\\
-   \hline
-   MSL & 998.10 & 544.86 & 285.73 & 139.8 & 76.45 & 48.05\\
-   \hline
-   SkelGIS & 1324.9 & 638.5 & 371.62 & 197.32 & 117.05 & 64.96\\
- \end{tabular}
-\caption{Execution times for the shallow water equations using MSL and SkelGIS (in seconds).}
-\label{fig:times}
- \end{center}
-\end{figure}
+% \begin{figure}[!h]
+%  \begin{center}
+%  \begin{tabular}{c|c|c|c|c|c|c}
+%    Cores & 8 & 16 & 32 & 64 & 128 & 256\\
+%    \hline
+%    MSL & 998.10 & 544.86 & 285.73 & 139.8 & 76.45 & 48.05\\
+%    \hline
+%    SkelGIS & 1324.9 & 638.5 & 371.62 & 197.32 & 117.05 & 64.96\\
+%  \end{tabular}
+% \caption{Execution times for the shallow water equations using MSL and SkelGIS (in seconds).}
+% \label{fig:times}
+%  \end{center}
+% \end{figure}
 \begin{figure*}
 \begin{center}
-\subfloat[\label{fig:speedup}]{
-\resizebox{8cm}{!}{\includegraphics{./images/speedup.pdf}}
+\subfloat[Edel\label{fig:speedup1}]{
+\resizebox{4cm}{!}{\includegraphics{./images/itpersec_5K_500.pdf}}
 }
-\hspace{10pt}
-\subfloat[\label{fig:log2}]{
-\resizebox{8cm}{!}{\includegraphics{./images/logtime.pdf}}
+\hspace{5pt}
+\subfloat[Edel\label{fig:log21}]{
+\resizebox{4cm}{!}{\includegraphics{./images/logtimes_5K_500.pdf}}
+}
+\hspace{5pt}
+\subfloat[Paravance\label{fig:speedup2}]{
+\resizebox{4cm}{!}{\includegraphics{./images/itpersec_5K_500.pdf}}
+}
+\hspace{5pt}
+\subfloat[Paravance\label{fig:log22}]{
+\resizebox{4cm}{!}{\includegraphics{./images/logtimes_5K_500.pdf}}
+}
+
+\subfloat[Edel\label{fig:speedup3}]{
+\resizebox{4cm}{!}{\includegraphics{./images/itpersec_10K_500.pdf}}
+}
+\hspace{5pt}
+\subfloat[Edel\label{fig:log23}]{
+\resizebox{4cm}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
+}
+\hspace{5pt}
+\subfloat[Paravance\label{fig:speedup4}]{
+\resizebox{4cm}{!}{\includegraphics{./images/itpersec_10K_500.pdf}}
+}
+\hspace{5pt}
+\subfloat[Paravance\label{fig:log24}]{
+\resizebox{4cm}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
 }

### 2015-08-13 WOLFHPC15/article/evaluation.tex
evaluation introduction helene 13_08

@@
  \begin{center}
- \begin{tabular}{c|c|c|c|c|c|c|c|c|}
+ \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+    \hline 
    level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
@@
    frequence & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
+   \hline
  \end{tabular}
@@
  \begin{center}
- \begin{tabular}{c|c|c|c|c|}
+ \begin{tabular}{|c|c|c|c|c|}
+  \hline
    step & $\Gamma_{sync}$ & $\Gamma_{dep}$ & $\Gamma_{msp}$ & $\Gamma_{tsp}$\\
@@
    \% & 0.02 & 5.3 & 83.3 & 11.37\\
+   \hline
  \end{tabular}
@@
 To evaluate performance of the generated component-based parallel structure, we have proceeded as follows.
-Our current implementation of the MSCAC compiler generates a back-end code using the SkelGIS library.
+Our current implementation of the MSCAC compiler generates a back-end code using the SkelGIS distributed data structure.
 For this reason, our evaluation compare the shallow water equations first parallelized with a pure SkelGIS code (data structures, applicators and interfaces of SkelGIS~\cite{CPE:CPE3494}), and second parallelized with MSL (which uses the SkelGIS data structure).
-As the SkelGIS library has proved its scalability on the Shallow water equations compared to an MPI parallelization, this performance evaluation should be relevant.
 Moreover, as SkelGIS library handles data parallelization for distributed memory architectures, we have limited for now our evaluations to the fusionned data parallelization of MSCAC (dumped from $\Gamma_{data}$).
 
-Evaluations have been performed on two clusters of \emph{Grid'5000}~\footnote{\url{https://www.grid5000.fr/mediawiki/index.php/Grid5000:Home}}.
-Figure~\ref{tab:hard} shows hardware configuration of tohse clusters.
-On the cluster \emph{Edel} (Grenoble site) have been performed three different experiments, each one described with the size of the Cartesian mesh, and the number of time iterations, as follows: (1) $5k \times 5k$, $500$ iterations; (2) $10k \times 10k$, $500$ iterations; (3) $10k \times 10k$, $2k$ iterations.
-On the cluster \emph{Paravence} (Rennes sites), the two last experiments have been performed.
+Evaluations have been performed on the cluster \emph{Edel} of \emph{Grid'5000}~\footnote{\url{https://www.grid5000.fr/mediawiki/index.php/Grid5000:Home}}.
+Figure~\ref{tab:hard} shows hardware configuration of this cluster.
+On \emph{Edel} have been performed three different experiments, each one described with the size of the Cartesian mesh and the number of time iterations, as follows: (1) $5k \times 5k$, $500$ iterations; (2) $10k \times 10k$, $500$ iterations; (3) $10k \times 10k$, $2k$ iterations.
 
@@
     \hline
-    Cluster & \textbf{Edel} & \textbf{Paravance}\\
+    Cluster & \textbf{Edel}\\
     \hline         
-    Proc. & 2 Intel Xeon E5520 & 2 Intel Xeon E52630v3\\
-    & (2.27 GHz) & (2.4 GHz)\\
-    Cores/n & 8 & 8\\
-    Mem./n & 24 GB & 128 GB\\
-    Compiler & OpenMPI & OpenMPI\\
-    Network & InfiniBand 40G & 10 Gigabit Ethernet\\
+    Processor & 2 Intel Xeon E5520\\
+    & (2.27 GHz)\\
+    Cores/node & 8 \\
+    Memory/node & 24 GB \\
+    Compiler & OpenMPI [-O3] \\
+    Network & InfiniBand 40G\\
     \hline
  \end{tabular}
-   \caption{\label{tab:grappe}Hardware of clusters used in the experiments.}
+   \caption{\label{tab:hard}Hardware of clusters used in the experiments.}
  \end{center}
@@
 
-%Figure~\ref{fig:times} shows the execution times in seconds as a function of the number of cores. 
-Figure~\ref{fig:speedup} illustrates a speedup comparison, using the minimum sequential reference time (which is the MSL sequential time).
-Finally, Figure~\ref{fig:log2} illustrates the logarithmic scale of execution times as a function of the logarithmic scale of the number of cores.
-% \begin{figure}[!h]
-%  \begin{center}
-%  \begin{tabular}{c|c|c|c|c|c|c}
-%    Cores & 8 & 16 & 32 & 64 & 128 & 256\\
-%    \hline
-%    MSL & 998.10 & 544.86 & 285.73 & 139.8 & 76.45 & 48.05\\
-%    \hline
-%    SkelGIS & 1324.9 & 638.5 & 371.62 & 197.32 & 117.05 & 64.96\\
-%  \end{tabular}
-% \caption{Execution times for the shallow water equations using MSL and SkelGIS (in seconds).}
-% \label{fig:times}
-%  \end{center}
-% \end{figure}
+Figure~\ref{fig:perfs} shows results on the three different experiments.
+On the first line are represented, for each experiment, the number of iteration computed per second for a given number of cores.
+On the second line are represented, for each experiment, the execution time  for a given number of cores, using a logarithmic scale.
+
 \begin{figure*}
 \begin{center}
-\subfloat[Edel\label{fig:speedup1}]{
-\resizebox{4cm}{!}{\includegraphics{./images/itpersec_5K_500.pdf}}
+\subfloat[$5k \times 5k$ mesh, $500$ iterations\label{fig:speedup1}]{
+\resizebox{5.5cm}{!}{\includegraphics{./images/itpersec_5K_500.pdf}}
 }
 \hspace{5pt}
-\subfloat[Edel\label{fig:log21}]{
-\resizebox{4cm}{!}{\includegraphics{./images/logtimes_5K_500.pdf}}
+\subfloat[$10k \times 10k$ mesh, $500$ iterations\label{fig:speedup2}]{
+\resizebox{5.5cm}{!}{\includegraphics{./images/itpersec_10K_500.pdf}}
 }
 \hspace{5pt}
-\subfloat[Paravance\label{fig:speedup2}]{
-\resizebox{4cm}{!}{\includegraphics{./images/itpersec_5K_500.pdf}}
+\subfloat[$10k \times 10k$ mesh, $2k$ iterations\label{fig:speedup3}]{
+\resizebox{5.5cm}{!}{\includegraphics{./images/itpersec_10K_500.pdf}}
+}
+
+\subfloat[$5k \times 5k$ mesh, $500$ iterations\label{fig:log21}]{
+\resizebox{5.5cm}{!}{\includegraphics{./images/logtimes_5K_500.pdf}}
 }
 \hspace{5pt}
-\subfloat[Paravance\label{fig:log22}]{
-\resizebox{4cm}{!}{\includegraphics{./images/logtimes_5K_500.pdf}}
-}
-
-\subfloat[Edel\label{fig:speedup3}]{
-\resizebox{4cm}{!}{\includegraphics{./images/itpersec_10K_500.pdf}}
+\subfloat[$10k \times 10k$ mesh, $500$ iterations\label{fig:log22}]{
+\resizebox{5.5cm}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
 }
 \hspace{5pt}
-\subfloat[Edel\label{fig:log23}]{
-\resizebox{4cm}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
-}
-\hspace{5pt}
-\subfloat[Paravance\label{fig:speedup4}]{
-\resizebox{4cm}{!}{\includegraphics{./images/itpersec_10K_500.pdf}}
-}
-\hspace{5pt}
-\subfloat[Paravance\label{fig:log24}]{
-\resizebox{4cm}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
+\subfloat[$10k \times 10k$ mesh, $2k$ iterations\label{fig:log23}]{
+\resizebox{5.5cm}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
 }
 \end{center}
-\caption{(a) Speedup comparison, and (b) logarithmic execution times, both for the Shallow water equations, using pure SkelGIS and MSL with the SkelGIS data structure.}
+\caption{The first line represents, for each experiment, the number of iteration computed per second for a given number of cores.
+The second line represents, for each experiment, the execution time for a given number of cores, using a logarithmic scale.}
 \label{fig:perfs}
@@
 
-One can notice that execution times using MSL, which itself uses the SkelGIS distributed data structures, are improved compared to a pure SkelGIS parallelization.
-Moreover, the scalability stays close to the scalability of SkelGIS.
-Thus, using a component-based runtime, performance of the back-end code is still HPC-oriented and does not introduce overheads.
+One can notice that, for the three different experiments, the execution times of MSL, which itself uses the SkelGIS distributed data structure, are improved compared to a pure SkelGIS parallelization.
+This result can be explained by the automatic generation of the parallel component-based runtime.
+Actually, writing a pure SkelGIS code, communications between processes are hidden from the user through specific interfaces called \emph{applicators}~\cite{CPE:CPE3494}.
+Using MSL, on the other hand, communications are directly managed by $SYNC$ components, which are automatically placed in the component-based runtime by MSCAC.
+Thus, no overheads are introduced to hide communications.
 
-\fix{more figures, more conclusions}
+In addition to this, the scalability slope shown by the number of iterations per second is close to SkelGIS.
+This is an important result as the execution time of MSL is at the same time improved compared to SkelGIS.
+Actually, the less the execution time is, the less the scalability capacity is.
+Thus, these results show that using a component-based runtime does not damage performance of the back-end code.
+Finally, we notice that SkelGIS has proved its scalability on the Shallow water equations compared to an MPI parallelization~\cite{CPE:CPE3494}, which makes these performance results relevant.

### 2015-08-13 WOLFHPC15/article/introduction.tex
evaluation introduction helene 13_08

--- 2015-08-08 WOLFHPC15/article/introduction.tex
@@
-Programming complex scientific applications on high performance heterogeneous architectures, such as clusters of multicores nodes containing accelerators, requires for the scientists to be an expert in his own domain, but also in low level parallel and high performance programming.
+Programming complex scientific applications on high performance heterogeneous architectures, such as clusters of multicores nodes containing accelerators, requires for a scientist to be an expert in his own domain, but also in low level parallel and high performance programming.
 It can take months to years for a scientist to learn HPC programming, as it is not his main activity.
-And while learning HPC programming, hardware continue to evolve, becoming more powerful but usually also more complex and difficult to use efficiently.
+And while learning HPC programming, hardware continue to evolve, becoming more powerful but also usually more complex and difficult to use efficiently.
 
@@
 A critical question however concerns the specificity of DSLs.
-Actually, the less specific the DSL is, the more information its users must provide and the less automatic optimizations can be embedded in the compiler.
+Actually, the less specific the DSL is, the more information its users must provide, and the less automatic optimizations can be embedded in the compiler.
 %On the other side of the scope
@@
 For this reason, solutions to ease the design and implementation of DSLs have recently appeared~\cite{Fernandez:2014:DFL:2691166.2691168}.
-The main idea of those solutions is to propose a single DSL or a framework to write new DSLs.
+The main idea of these solutions is to propose a single DSL or a framework to write new DSLs.
 In this paper, we study a first contribution to another approach to improve maintainability and portability in DSLs.
@@
 For this reason, we think that combining DSLs and component-based runtimes could enable those properties, by inheritance, in DSLs.
-To evaluate this approach, a proper new DSL for \emph{multi-stencil} programs, and its compiler are presented in this paper.
+To evaluate this approach, a proper new DSL for \emph{multi-stencil} programs, \ie a subclass of numerically solved partial differential equations (PDEs), and its compiler are presented in this paper.
 %\fix{cp2cp+hc: to dev benefits of components.}
 
-Many solutions to automatically optimize and parallelize partial differential equations (PDEs) codes have been proposed in the literature, either as specific frameworks~\cite{CPE:CPE3494} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
+Many solutions to automatically optimize and parallelize numerical simulations have been proposed in the literature, either as specific frameworks~\cite{CPE:CPE3494,Trilinos-Overview,petsc-efficient} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
 Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil kernel.
-However, a real case numerical simulation (such as fluid dynamics, magnetohydrodynamics, molecular synamics etc.) is most of the time not composed of a single stencil kernel, but of a set of stencil kernels and a set of additional local auxiliary computations.
+However, a real case numerical simulation (such as fluid dynamics, magnetohydrodynamics, molecular dynamics etc.) is most of the time not composed of a single stencil kernel, but of a set of stencil kernels and a set of additional local auxiliary computations.
 The DSL proposed in this paper, called MSL, is a DSL to generate a parallel component-based structure of an overall numerical simulation, that we call a \emph{multi-stencil program}.
-As it will be detailed in the related work, this is complementary to the optimization and parallelization of a single stencil kernel.
+As it will be detailed in the related work, this work is complementary to the optimization and parallelization of a single stencil kernel.
 
@@
 First, the concepts of \emph{stencil kernels} and \emph{multi-stencil programs} are formalized in Section~\ref{sect:concept}.
-Section~\ref{sect:mscac} gives an overview of the solution presented in this paper while Section~\ref{sect:msmsc} details the MSL language and its compiler.
-Section~\ref{sect:component} introduces component models and describes the proposed component-based runtime and the automatic generation of its parallel computation part.
+Section~\ref{sect:mscac} gives an overview of the solution presented in this paper while Section~\ref{sect:msmsc} details the MSL language and its compiler which automatically generates the parallel computation part.
+Section~\ref{sect:component} introduces component models and describes the proposed component-based runtime of the language.
 %The end of Section~\ref{sect:component} also discusses on interest of this new approach.

### 2015-08-13 WOLFHPC15/article/article.tex
section 4 helene 30_07
Adding \fix
\cf \ie
evaluation helene 30_07
Fix section & subsection title
Removed related work in sec.2
todo
wolfhpc fix language figures helene 06_08
evaluations helene 06_08
Ack++
conclusion helene 08_08
fix accents et taille references helene 08_08
evaluation introduction helene 13_08
v1 abstract helene 13_08

@@
 \usepackage{subfig}
+\usepackage{filecontents}
 \usepackage{amsmath}
 \usepackage{xspace}
+\usepackage{todonotes}
+
 %----------TIKZ
@@
 
+\newcommand{\fix}[1]{\todo[inline,color=red!80]{#1}}
+% \renewcommand{\todo}[2][]{}
+
+\newcommand{\cf}{\emph{cf.}\xspace}
+\newcommand{\ie}{\emph{i.e.,}\xspace}
 \newcommand{\llc}{L\textsuperscript{2}C\xspace}
@@
        \affaddr{CEA, Maison de la Simulation}\\
-       \affaddr{USR 3441}\\
+       \affaddr{USR 3441, CEA Saclay}\\
        \affaddr{91191 Gif-sur-Yvette, France}\\ % remove cedex
@@
 \alignauthor
-Helene Coullon\\
+H\'el\`ene Coullon\\
        \affaddr{INRIA}\\
-       \affaddr{ENS Lyon, 46 Allee d'Italie}\\
+       \affaddr{ENS Lyon, 46 All\'ee d'Italie}\\
        \affaddr{Lyon, France}\\
@@
 \alignauthor
-Christian Perez\\
+Christian P\'erez\\
        \affaddr{INRIA}\\
-       \affaddr{ENS Lyon, 46 Allee d'Italie}\\
+       \affaddr{ENS Lyon, 46 All\'ee d'Italie}\\
        \affaddr{Lyon, France}\\
@@
 \begin{abstract}
-
+High performance heterogeneous architectures continuously evolve to be more powerfull.
+Such architectures also usually become more difficult to use efficiently.
+As a scientist is not a low level and high performance programming expert, Domain Specific Languages (DSLs) are a promising solution to automatically and efficiently write high performance codes.
+However, if DSLs ease programming for scientists, it seems that non-maintainability and non-portability problems are deported from the scientist to the DSL designer.
+In this paper is presented a new approach to improve maintainability and programming productivity of DSLs through the generation of a component-based parallel runtime.
+To study this approach the paper presents a proper new DSL for \emph{multi-stencil} programs, which is evaluated on the real-case shallow water equations.
 \end{abstract}
@@
 %----------------------------------------
-\section{Stencil and multi-stencil}
+\section{Stencil and Multi-Stencil}
 \label{sect:concept}
@@
 %----------------------------------------
-\section{Overview}
+\section{MSL and MSCAC Overview }
 \label{sect:mscac}
@@
 %----------------------------------------
-\section{MS Language and MSC compiler}
+\section{Language and Parallelization Description}
 \label{sect:msmsc}
@@
 %----------------------------------------
-\section{Dump to a component assembly}
+\section{A Component based Back-End}
 \label{sect:component}
@@
 %----------------------------------------
-\section{Evaluation}
+\section{Real-case Evaluation}
 \label{sect:eval}
@@
 
+\section{Acknowledgments}
+
+We acknowledge Grid'5000 for evaluations.
+
+\small
 \bibliographystyle{abbrv}

### 2015-08-13 WOLFHPC15/article/component.tex
relectures/corrections helene 13_08

@@
 %----------------------------------------
-This section describes the projection of a MSP program to components, including the components used to implement the static scheduling of $\Gamma_{tsp}$.
+This section describes the projection of a MSP program to a component-based runtime, including the components used to implement the static scheduling of $\Gamma_{tsp}$.
 Before those details, the section gives an overview of component models, and especially the Low Level Components (\llc) used in this paper.
@@
 %----------------------------------------
-\subsection{MSCAC Runtime Overview}
+\subsection{Runtime Overview}
 %----------------------------------------
@@
 As described in Section~\ref{sect:mscac}, the compiler generates four independent parts in a component assembly in addition to a static part.
-This is represented on Figure~\ref{fig:mscac:assembly} where each rounded box represents one or more component instances, respectively drawn with simple lines or double lines.
-For example in the figure, the component \texttt{Data} is instanciated more than once.
+This is represented on Figure~\ref{fig:mscac:assembly} where each rounded box represents one component instance, drawn with simple lines, or more, drawn with double lines.
+For example, the component \texttt{Data} is instanciated more than once.
 \texttt{Driver} is static and drives the whole application.
@@
 The back-end presented in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}.
-\texttt{DDS} is used by \texttt{Data}, which represents one component instance for each data element in the MSL program.
+\texttt{DDS} is used by \texttt{Data}, for which one instance represents one data element in the MSL program.
 \texttt{Time} is parametrized by the the \texttt{iteration} value from MSL and uses the \texttt{Computations} component as many times as required.
@@
 \texttt{Computations} contains the static scheduling $\Gamma_{tsp}$, or the data parallel transformation $\Gamma_{data}$, both computed by MSC.
-Both transformations, $\Gamma_{tsp}$ and $\Gamma_{data}$, are encoded using three specific components,detailed in Section~\ref{sect:control}, which manage the $P$, $S$, and $sync$ operations.
+Both transformations, $\Gamma_{tsp}$ and $\Gamma_{data}$, are encoded using three specific components, detailed in Section~\ref{sect:control}, which manage the $P$, $S$, and $sync$ operations.
 At the leaves of $\Gamma_{tsp}$ and $\Gamma_{data}$, the computation components instances (denoted $K$), provided by the user, are connected to \texttt{Data}.
-As presented on Figure~\ref{fig:k}, these $K$ components provide a single port used to launch the computation and expose one use port for each data element manipulated in the computation.
-We denote the fact that they may have several use port by adding a star on the use port to \texttt{Data}.
+As presented on Figure~\ref{fig:k}, a $K$ component provides a single port used to launch the computation and exposes one use port for each data element manipulated in the computation.
+We denote the fact that it may have several use port by adding a star on the use port to \texttt{Data}.
 
@@
 As a result of this approach, each part of the generated code is rather independent.
-Changing the main loop to use a convergence criteria rather than a fixed number of iterations would only require changing the \texttt{Time} component.
+Changing the main time loop to use a convergence criteria rather than a fixed number of iterations would only require changing the \texttt{Time} component.
 Similarly, changing the approach used for scheduling would only impact the code generated in \texttt{Computations}.
@@
 
-\item[Synchronization component (SYNC)].
-It is the direct representation to an update leaf of $\Gamma_{tsp}$ or $\Gamma_{data}$.
+\item[Synchronization component (SYNC)] It is the direct representation to an update leaf of $\Gamma_{tsp}$ or $\Gamma_{data}$.
 This component calls the synchronization of a given data.
@@
 the identification name of the computation in the MSL file.
-Thus, this figure represents what the component \texttt{Computations} of Figure~\ref{fig:mscac:assembly} contains.
+Thus, this figure represents what the component \texttt{Computations} of Figure~\ref{fig:mscac:assembly} contains if an hybrid parallelization is performed.
+On the other hand, if a data parallelization is performed, a single $SEQ$ component is used by \texttt{Time}, which itself uses in the appropriate order the computation kernels and synchronizations.
+Among those kernels, a single one \emph{fusionned} represents $c_3$ and $c_5$.
 

### 2015-08-13 WOLFHPC15/article/concept.tex
relectures/corrections helene 13_08

@@
 
-As it can be seen from those definitions, stencil kernels implementations share many properties from an algorithmic point of view that can be used to apply well known optimization and parallelization strategies.
+As it can be seen from these definitions, stencil kernels implementations share many properties from an algorithmic point of view that can be used to apply well known optimization and parallelization strategies.
 Many solutions (languages or frameworks) thus propose to ease their programming by producing optimized and parallelized code from a simple description of the local computation to apply on each element.
@@
 \end{equation}
-where $R_l$ is the set of input data, and other parameters are the same as for normal stencil kernels.
+where $R_l$ is the set of input data locally accessed, and other parameters are the same as for normal stencil kernels.
 

### 2015-08-13 WOLFHPC15/article/ms_msc.tex
fix component.tex helene 08_08
fix all helene 08_08
relectures/corrections helene 13_08

@@
 Lines 1 to 9 define what is a MSL program with the different parts mentioned above.
-The remaining of this section describes those different parts of a MSL program and presents an example.
-One can notice that almost all terminals \texttt{meshid}, \texttt{meshdom}, \texttt{compdomid}, \texttt{dataid}, \texttt{compid} and \texttt{neighborid}, are string identifiers.
-Only \texttt{num} is an integer terminal.
+The remaining of this section describes these different parts of a MSL program and presents an example.
+One can notice that terminals are the string identifiers \texttt{meshid}, \texttt{meshdom}, \texttt{compdomid}, \texttt{dataid}, \texttt{compid} and \texttt{neighborid}.
+The terminal \texttt{num} is an integer terminal.
 
@@
 \subsubsection*{Computation Domains and Their Dependencies}
-A computation domain is a subset of a given kind of mesh entities, used during one or more computations.
+A computation domain is a subset of a given kind of mesh entities, used for at least one computation.
 For each computation domain, the kind of mesh entities on which it is mapped is specified (lines 3-4 and 12-14 on Figure~\ref{fig:grammar}).
@@
 Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} simply corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
-The specification starts by the identifier of the computed data element ($w$ in the formal description) with the computation domain between brackets ($d$).
+The description starts by the identifier of the computed data element ($w$ in the formal description) with the computation domain between brackets ($d$).
 Then, after the equal sign, the kernel identifier ($exp$) is specified.
@@
 \subsection{MSC Parallelization}
+\label{sect:msc}
 %----------------------------------------
@@
 This parallelization phase is divided in five different steps to transform $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
-%Data parallelism is handled in the first two steps, and task parallelism in the following three steps.
-This section gives an overview of these five steps , but they are more formally defined with their associated algorithms in a research report~\cite{????}.
+These steps are described in this section.
+%This section gives an overview of these five steps , but they are more formally defined with their associated algorithms in a research report~\cite{????}.
 
@@
 However, this parallelization technique requires additional synchronizations between processors.
-For example, as a stencil kernel accesses neighbor values, values computed by another processor are needed and must be synchronized. 
+%For example, as a stencil kernel accesses neighbor values, values computed by another processor are needed and must be synchronized. 
 The required synchronizations can be automatically computed from the ordered list of computations $\Gamma$.
@@
 As a result, a dependency is kept between $c_0$ and $sync_1$ and between $sync_1$, and $c_1$.
-The same transformation is performed for all stencils: $c_4$ and $c_8$.
+The same transformation is performed for the two other stencils: $c_4$ and $c_8$.
 Thus, $\Gamma_{sync} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
@@
 From the synchronized ordered list of computations $\Gamma_{sync}$, a dependency graph $\Gamma_{dep}$ is built.
-A dependency exists from computation $a$ to $b$ (including synchronizations) if and only if a data element read by $b$ has been written by $a$ in $\Gamma_{sync}$ with intersecting domains.
+A dependency exists between computations $a$ and $b$ (including synchronizations) if and only if a data element read by $b$ has been written by $a$ in $\Gamma_{sync}$ with intersecting domains.
 %In other words, the only case where the same data is manipulated by distinct kernels without inducing a dependency is when the computation domains do not intersect.
@@
 To do so the dependency graph is transformed to a minimal series-parallel graph~\cite{Valdes:1979:RSP:800135.804393}.
-As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a DAG is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape} and illustrated in Figure~\ref{fig:n}, is not found in the DAG.
-Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes} by an over-constraint, as illustrated in Figure~\ref{fig:over}.
+As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a DAG is a minimal series-parallel graph if and only if the \emph{forbidden} shape called \emph{N-shape}, illustrated in Figure~\ref{fig:n}, is not found in the DAG.
+Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm is applied to remove all \emph{N-shapes} by an over-constraint~\cite{Mitchell:2004:CMV:1082101.1082117}, as illustrated in Figure~\ref{fig:over}.
 
@@
 Many works~\cite{Valdes:1979:RSP:800135.804393,Schoenmakers95anew} explains how to build a series-parallel tree decomposition from a minimal series-parallel graph.
-A series-parallel tree decomposition consists in the decomposition of the minimal series-parallel graph as a tree representing a set of \emph{sequences}, indicated by a $S$ node, and \emph{parallel} sections, indicated as a $P$ node.
+This transformation decomposes the minimal series-parallel graph as a tree where internal nodes are \emph{sequences}, indicated by a $S$ node, and \emph{parallel} sections, indicated as a $P$ node, while leaves of the tree are the nodes of $\Gamma_{msp}$.
 For example, the series-parallel tree decomposition $\Gamma_{tsp}$ of Figure~\ref{fig:hyb} is illustrated in Figure~\ref{fig:tsp}.
@@
 \subsubsection*{Data parallelism optimizations: $\Gamma_{data}$}
-The series-parallel tree decomposition $\Gamma_{tsp}$ is used to build the hybrid parallelization of the simulation, but it can also be used to optimize the data parallelization only.
-In fact, creating a data parallelization of the simulation, it first seems natural to create the parallel structure directly from $\Gamma_{sync}$, as the program is parallelized from data decomposition and applying the same sequential tasks on each processor.
-However, as a computation kernel writes a single data $w$ (to finely manage dependencies between tasks), using $\Gamma_{sync}$ the same computation domain might be iterated more than once, while some numerical expressions could be computed inside a joint space loop.
+The series-parallel tree decomposition $\Gamma_{tsp}$ is used to build the hybrid parallelization of the simulation, but it can also be used to optimize the data parallelization.
+In fact, it first seems natural to create the data parallel structure directly from $\Gamma_{sync}$, as the program is parallelized from data decomposition and applying the same sequential tasks on each processor.
+However, as a computation kernel writes a single data $w$ (to finely manage dependencies between tasks), the same computation domain might be iterated more than once, while some numerical expressions could be computed inside a joint space loop.
 For this reason, an optimized and \emph{fusionned} ordered list of computation $\Gamma_{data}$ is built from $\Gamma_{tsp}$.
 Actually, in $\Gamma_{tsp}$, computations of a parallel subtree with a same computation domain can be computed in the same space loop, in any order.
-Moreover, in $\Gamma_{tsp}$, two consecutive computations of a sequence subtree with a same computation domain can be computed in the same space loop, keeping the computation order of the sequence.
+Moreover, in $\Gamma_{tsp}$, two consecutive computations of a sequence subtree with a same computation domain can be computed in the same space loop, if the computation order of the sequence is kept.
+For example, from $\Gamma_{tsp}$ of Figure~\ref{fig:tsp}, as kernels $c_3$ and $c_5$ are executed inside a parallel section, and as their computation domains are the same (\texttt{alledgey}), a single component $K$ can be written peforming both computations in the same domain loop.
 

### 2015-08-13 WOLFHPC15/article/mscac.tex
relectures/corrections helene 13_08

@@
-The solution proposed in this paper is based on \textbf{a)} a domain specific language to describe a multi-stencil program, \ie a complete numerical simulation, and \textbf{b)} a compiler which successively parses the language, applies transformations for automatic parallelization, and finally dumps the representation to a component-based code.
+The solution proposed in this paper is based on \textbf{a)} a domain specific language to describe a multi-stencil program, \ie a complete numerical simulation, and \textbf{b)} a compiler which successively parses the language, applies transformations for automatic parallelization, and finally dumps the representation to a component-based runtime.
 
@@
 The actual implementation of the numerical expressions are provided by the user and used in the computations scheduling.
-This component-based parallel structure enables to change the technology used for each part while limiting the impact on the other parts as further discussed in Section~\ref{sect:component}.
+This component-based parallel structure enables to change the technology used for each part while limiting the impact on the other parts, as further discussed in Section~\ref{sect:component}.
 

### 2015-08-13 WOLFHPC15/article/introduction.tex
clean git helene 13_08

@@
 Domain Specific Languages (DSLs) are a promising solution to hide intricate details of HPC programming from non experts.
-By being specific to a well-defined set of problems, these languages enable the generation of well parallelized and optimized code for various architectures while requiring only a limited amount of information from the developer; part of the knowledge is directly embedded in the compiler.
+By being specific to a well-defined set of problems, those languages enable the generation of well parallelized and optimized code for various architectures while requiring only a limited amount of information from the developer; part of the knowledge is directly embedded in the compiler.
 A critical question however concerns the specificity of DSLs.
@@
 For this reason, solutions to ease the design and implementation of DSLs have recently appeared~\cite{Fernandez:2014:DFL:2691166.2691168}.
-The main idea of these solutions is to propose a single DSL or a framework to write new DSLs.
+The main idea of those solutions is to propose a single DSL or a framework to write new DSLs.
 In this paper, we study a first contribution to another approach to improve maintainability and portability in DSLs.

### 2015-08-13 WOLFHPC15/article/mscac.tex
clean git helene 13_08

@@
 % JB: la phrase en dessous est pas tres heureuse ...
-Those properties counter some known limitations of DSLs. % JB: was "parallel applications".
+These properties counter some known limitations of DSLs. % JB: was "parallel applications".
 %where a component represents a clear and independent functionality of the application, have first been proposed in software engineering and have proved many times good properties for
@@
 Thanks to the software components based approach, the generated code is made of four independent parts that loosely match the elements of the multi-stencil program definition.
-These are the components implementing: \textbf{a)} the computations scheduling ($\Gamma$), \textbf{b)} the data structures ($\mathcal{M}$ and $\mathcal{E}$), \textbf{c)} data of the simulation ($\Delta$ and $\mathcal{D}$), and \textbf{d)} the time loop ($T$).
+Those are the components implementing: \textbf{a)} the computations scheduling ($\Gamma$), \textbf{b)} the data structures ($\mathcal{M}$ and $\mathcal{E}$), \textbf{c)} data of the simulation ($\Delta$ and $\mathcal{D}$), and \textbf{d)} the time loop ($T$).
 The actual implementation of the numerical expressions are provided by the user and used in the computations scheduling.

### 2015-08-13 WOLFHPC15/article/conclusion.tex
relectures/corrections helene 13_08

@@
-In this paper, we have presented the domain specific language MSL and its compiler MSCAC designed to produce a parallel component-based runtime of an overall multi-stencil program, \ie a mesh-based numerical simulation reduced to explicit schemes.
-MSL has been evaluated by the description of a real case numerical simulation: the shallow water equations.
-The component-based data parallelization of the simulation has been compared to a pure SkelGIS parallelization.
-Execution times of the simulation have been improved and its scalability has been shown.
-Those results demonstrate that component-based runtimes may be relevant back-end codes for DSLs.
-Moreover interests of component-based runtimes has been discussed in the paper.
+In this paper, we have presented the domain specific language MSL and its compiler MSCAC designed to produce a parallel (data and hybrid) component-based runtime of an overall multi-stencil program, \ie a mesh-based numerical simulation reduced to explicit schemes.
+After details on the language and its compiler, MSL has been evaluated by the description of a real case numerical simulation: the shallow water equations.
+The component-based data parallelization of the simulation has been compared to a pure SkelGIS parallelization, and has shown improved execution times as well as a promising scalability.
+Those results demonstrate that component-based runtimes may be relevant back-end codes for DSLs as they do not introduce performance damage.
+Moreover, interests of component-based runtimes has been discussed in the paper.
 

### 2015-08-13 WOLFHPC15/article/evaluation.tex
relectures/corrections helene 13_08

@@
 Those equations are solved using a two-dimensional Cartesian discretization of the space domain, and a finite volume numerical methods more described in~\cite{CPE:CPE3494}.
-The MSL program of FullSWOF contains three mesh entities, seven computation domains, fourty-eight data and ninty-eight computations composed of thirty-two stencil kernels and sixty-six local computations.
+The MSL program of FullSWOF contains three mesh entities, seven computation domains, fourty-eight data and ninty-eight computations.
+Among these computations thirty-two stencil kernels and sixty-six local kernels are performed in the simulation.
 
@@
 Thus, the fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static shedule of tasks may not be the best solution for complex simulations.
-It has to be notice, however, that the all computation of $\Gamma_{tsp}$ is still usefull to detect how space loops can be fusionned in $\Gamma_{data}$.
+It has to be noticed, however, that the all computation of $\Gamma_{tsp}$ is still usefull to detect how space loops can be fusionned in $\Gamma_{data}$.
 
@@
 Figure~\ref{tab:hard} shows hardware configuration of this cluster.
-On \emph{Edel} have been performed three different experiments, each one described with the size of the Cartesian mesh and the number of time iterations, as follows: (1) $5k \times 5k$, $500$ iterations; (2) $10k \times 10k$, $500$ iterations; (3) $10k \times 10k$, $2k$ iterations.
+On \emph{Edel} have been performed three different experiments, each one described with the size of the Cartesian mesh and the number of time iterations: (1) $5k \times 5k$, $500$ iterations; (2) $10k \times 10k$, $500$ iterations; (3) $10k \times 10k$, $2k$ iterations.
 
@@
 
-One can notice that, for the three different experiments, the execution times of MSL, which itself uses the SkelGIS distributed data structure, are improved compared to a pure SkelGIS parallelization.
+One can notice that, for the three different experiments, execution times of MSL, which itself uses the SkelGIS distributed data structure, are improved compared to a pure SkelGIS parallelization.
 This result can be explained by the automatic generation of the parallel component-based runtime.
@@
 
-In addition to this, the scalability slope shown by the number of iterations per second is close to SkelGIS.
+In addition to this, the scalability slope, shown by the number of iterations per second, is close to SkelGIS.
 This is an important result as the execution time of MSL is at the same time improved compared to SkelGIS.

### 2015-08-13 WOLFHPC15/article/related.tex
fix accents et taille references helene 08_08
relectures/corrections helene 13_08

@@
 
-Many specific solutions exist to ease parallel programming of numerical simulations.
+Many specific solutions exist to ease parallel programming of numerical simulations~\cite{Trilinos-Overview,petsc-efficient}.
 %Solutions such as PETSc~\cite{petsc-efficient} and Trilinos~\cite{Trilinos-Overview} are specific frameworks.
@@
 The purpose of those solutions are much more specific than our solution.
-The closest solutions to MSL are DSLs proposing to generate parallel codes from a sequential imperative stencil description.
-For example PATUS~\cite{citeulike12258902} and Pochoir~\cite{spaaTangCKLL11} propose a simple description to write stencil codes for general structured meshes, and OP2~\cite{Giles2011} and Liszt~\cite{DeVito2011LDS} for general unstructured meshes.
+The closest solutions to MSL are DSLs which propose to generate parallel codes from a sequential imperative stencil description.
+For example, PATUS~\cite{citeulike12258902} and Pochoir~\cite{spaaTangCKLL11} propose a simple description to write stencil codes for general structured meshes, and OP2~\cite{Giles2011} and Liszt~\cite{DeVito2011LDS} for general unstructured meshes.
 Those solutions can be considered as stencil compilers.
@@
 The MSCAC compiler only needs this agnostic description of the overall simulation to generate a parallel pattern of the program.
-This parallel pattern then needs to be filled with computation codes.
+This parallel pattern then needs to be filled with implementation parameters and computation codes.
 %. Thus, depending on the underlying distributed data structure used, those computation codes 

### 2015-08-21 WOLFHPC15/article/article.tex
fix affiliation LIP
fix abstract

@@
 H\'el\`ene Coullon\\
-       \affaddr{INRIA}\\
+       \affaddr{INRIA, LIP}\\
        \affaddr{ENS Lyon, 46 All\'ee d'Italie}\\
@@
 Christian P\'erez\\
-       \affaddr{INRIA}\\
+       \affaddr{INRIA, LIP}\\
        \affaddr{ENS Lyon, 46 All\'ee d'Italie}\\
@@
 \begin{abstract}
-High performance heterogeneous architectures continuously evolve to be more powerfull.
+  High performance architectures evolve continuously to be more powerful.
 Such architectures also usually become more difficult to use efficiently.
 As a scientist is not a low level and high performance programming expert, Domain Specific Languages (DSLs) are a promising solution to automatically and efficiently write high performance codes.
-However, if DSLs ease programming for scientists, it seems that non-maintainability and non-portability problems are deported from the scientist to the DSL designer.
-In this paper is presented a new approach to improve maintainability and programming productivity of DSLs through the generation of a component-based parallel runtime.
-To study this approach the paper presents a proper new DSL for \emph{multi-stencil} programs, which is evaluated on the real-case shallow water equations.
+However, if DSLs ease programming for scientists, maintainability and portability issues are deported from scientists to DSL designers.
+This paper deals with an approach to improve maintainability and programming productivity of DSLs through the generation of a component-based parallel runtime.
+To study it, the paper presents a  DSL for \emph{multi-stencil} programs, that is evaluated on a real-case of shallow water equations.
 \end{abstract}

### 2015-08-21 WOLFHPC15/article/introduction.tex
rel. S1. Intro

@@
-Programming complex scientific applications on high performance heterogeneous architectures, such as clusters of multicores nodes containing accelerators, requires for a scientist to be an expert in his own domain, but also in low level parallel and high performance programming.
+Programming complex scientific applications on high performance architectures, such as clusters of multi-core nodes containing accelerators, requires for a scientist to be an expert in his own domain, but also in low level parallel and high performance programming.
 It can take months to years for a scientist to learn HPC programming, as it is not his main activity.
-And while learning HPC programming, hardware continue to evolve, becoming more powerful but also usually more complex and difficult to use efficiently.
+As hardware continue to evolve to become more powerful but also usually more complex and difficult to use efficiently, the learning is never over.
 
@@
 %On the other side of the scope
-Conversely, the more specific the DSL is, the less number of different applications can use it, and the more difficult it is to amortize the implementation cost of the DSL (language, compiler, and runtime).
-Programming difficulties, as well as non-maintainability and non-portability often seem to be simply deported from the scientist to DSL designers.
+Conversely, the more specific the DSL is, the less applications can use it, and the more difficult it is to amortize the implementation cost of the DSL (language, compiler, and runtime).
+Programming complexity, as well as non-maintainability and non-portability often seem to be simply deported from scientists to DSL designers.
 % Moreover, one can consider that end users need as much DSLs as domains, which keeps the amount of work to write and maintain up-to-date DSLs out of hands.
@@
 For this reason, we think that combining DSLs and component-based runtimes could enable those properties, by inheritance, in DSLs.
-To evaluate this approach, a proper new DSL for \emph{multi-stencil} programs, \ie a subclass of numerically solved partial differential equations (PDEs), and its compiler are presented in this paper.
+This paper proposes a first evaluation of this approach though a use-case study that is a DSL for \emph{multi-stencil} programs, \ie a subclass of numerically solved partial differential equations (PDEs), and its compiler.
 %\fix{cp2cp+hc: to dev benefits of components.}
@@
 Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil kernel.
-However, a real case numerical simulation (such as fluid dynamics, magnetohydrodynamics, molecular dynamics etc.) is most of the time not composed of a single stencil kernel, but of a set of stencil kernels and a set of additional local auxiliary computations.
-The DSL proposed in this paper, called MSL, is a DSL to generate a parallel component-based structure of an overall numerical simulation, that we call a \emph{multi-stencil program}.
-As it will be detailed in the related work, this work is complementary to the optimization and parallelization of a single stencil kernel.
+However, a real case numerical simulation (such as fluid dynamics, magneto-hydrodynamics, molecular dynamics etc.) is most of the time not composed of a single stencil kernel, but of a set of stencil kernels and a set of additional local auxiliary computations.
+The DSL proposed in this paper, called \emph{Multi-Stencil Language} or MSL, is a DSL to generate a parallel component-based structure of a numerical simulation, that we call a \emph{multi-stencil program}.
+As discussed in the related work, this work is complementary to the optimization and parallelization of a single stencil kernel.
 
 The rest of this paper is organized as follows.
-First, the concepts of \emph{stencil kernels} and \emph{multi-stencil programs} are formalized in Section~\ref{sect:concept}.
-Section~\ref{sect:mscac} gives an overview of the solution presented in this paper while Section~\ref{sect:msmsc} details the MSL language and its compiler which automatically generates the parallel computation part.
-Section~\ref{sect:component} introduces component models and describes the proposed component-based runtime of the language.
+The concepts of \emph{stencil kernels} and \emph{multi-stencil programs} are formalized in Section~\ref{sect:concept}.
+Section~\ref{sect:mscac} gives an overview of the proposed solution while Section~\ref{sect:msmsc} details the MSL language and its compiler which automatically generates the parallel computation part.
+Section~\ref{sect:component} introduces component models and describes the proposed component-based runtime of MSL.
 %The end of Section~\ref{sect:component} also discusses on interest of this new approach.
-Section~\ref{sect:eval} focuses on a real-case simulation, solving the Shallow-water equations, and analyzes preliminary performance results.
-Sections~\ref{sect:related} states on related work, and finally Section~\ref{sect:conclusion} concludes and gives some perspectives.
+Section~\ref{sect:eval} focuses on a real-case simulation, solving the Shallow-water equations, and analyzes preliminary performance and usability results.
+Sections~\ref{sect:related} states on related work, while Section~\ref{sect:conclusion} concludes the paper.

### 2015-08-21 WOLFHPC15/article/concept.tex
rel. S2.

@@
 \begin{center}
-\subfloat[Mesh and mesh domains\label{fig:mesh}]{
+\subfloat[Mesh and mesh domains.\label{fig:mesh}]{
 \resizebox{7cm}{!}{\includegraphics{./images/mesh.pdf}}
@@
 \hspace{10pt}
-\subfloat[First stencil example\label{fig:ex1}]{
+\subfloat[4-neighborhood stencil.\label{fig:ex1}]{
 \resizebox{5cm}{!}{\includegraphics{./images/stencil1.pdf}}
@@
 \hspace{10pt}
-\subfloat[Second stencil example\label{fig:ex2}]{
+\subfloat[2-neighborhood 2-entity stencil.\label{fig:ex2}]{
 \resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
@@
 \end{center}
-\caption{(a) a Cartesian mesh and two kind of mesh entities, (b) an example of stencil kernel on cells, (c) an example of stencil kernel on different entities of the mesh.}
+\caption{(a) a Cartesian mesh and two kind of mesh entities, (b) an example of stencil kernel on cells, (c) an example of stencil kernel on two different entities of the mesh.}
 \label{fig:gspmsp}
@@
 Actually, paying attention to complex numerical simulations, it appears that most of them are composed of more than one stencil kernel, with one or more stencil shapes and of additional local computations.
-For example, we would like to formalize and parallelize a numerical simulation which chains stencil kernels of Figures~\ref{fig:ex1} and~\ref{fig:ex2}.
+For example, how to formalize and parallelize a numerical simulation which chains the stencil kernels of Figures~\ref{fig:ex1} and~\ref{fig:ex2}.
 The next section formalizes concepts of \emph{stencil kernel}, \emph{local kernel} and \emph{multi-stencil program}.
@@
 Let $\Delta$ be the set of data of the simulation
-A stencil kernel $s$ is defined as the quintuplet
+A stencil kernel $s$ is defined as the quintuplet:
 \begin{equation} 
@@
 \begin{equation*}
-R: \{(B,n1)\} \quad w: A \quad d: dc1
+R: \{(B,n1)\}, \quad w: A, \quad d: dc1,
 \end{equation*}
@@
 \end{equation*}
-On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil kernel can be defined as:
+On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil kernel is defined as:
 \begin{equation*}
-R: \{(C,n2),(A,local)\} \quad w: A \quad d: dc2
+R: \{(C,n2),(A,local)\}, \quad w: A, \quad d: dc2,
 \end{equation*}
 \begin{equation*}
-exp: A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1)
+exp: A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1).
 \end{equation*}
@@
 
-A multi-stencil program in this work is defined as a sextuplet:
+We define a multi-stencil program as a sextuplet:
 \begin{equation} 

### 2015-08-21 WOLFHPC15/article/mscac.tex
Rel. S3

@@
-The solution proposed in this paper is based on \textbf{a)} a domain specific language to describe a multi-stencil program, \ie a complete numerical simulation, and \textbf{b)} a compiler which successively parses the language, applies transformations for automatic parallelization, and finally dumps the representation to a component-based runtime.
+The solution proposed in this paper is based on both a domain specific language to describe a multi-stencil program, \ie a complete numerical simulation, and on a compiler which parses the language, applies transformations for automatic parallelization, and dumps the representation to a component-based runtime.
 
 The proposed DSL, named MSL (\emph{Multi-Stencil Language}), is an agnostic descriptive language for multi-stencil simulations.
-Agnostic means that the description of a numerical simulation does not depend on implementation choices as, for example, the type of mesh (structured, unstructured), or its associated interfaces (how to define a stencil shape, etc.).
+Agnostic means that the description of a numerical simulation does not depend on implementation choices as, for example, the type of mesh (structured or unstructured), or its associated interfaces (how to define a stencil shape, etc.).
 Descriptive, on the other hand, means that MSL is not made to handle the expression of numerical computations, but only the description of a multi-stencil simulation, in a coarse-grain fashion.
@@
 \end{tikzpicture}
-\caption{The three phases of the MSCAC compiler. First the parser of the MSL description, second the MSC parallelization, third the dump to the Component-Based Runtime (CBR).}
-\label{fig:mscac}
+\caption{The three phases of the MSCAC compiler: \emph{i)} the parser of the MSL description; \emph{ii)} the MSC parallelization; \emph{iii)} the dump to the Component-Based Runtime (CBR).\label{fig:mscac}}
 \end{center}
@@
 The parallelization techniques handled by MSC are \emph{data parallelism}, where data is split among processors or cores while applying the same computation on every piece of data, and \emph{task parallelism} where the program is transformed into a graph of tasks with dependencies.
-Different strategies can be used to schedule the tasks of this graph; in other words to plan their execution in a valid order.
+Different strategies can be used to schedule the tasks of this graph, \ie to plan their execution in a valid order.
 The MSC compiler presented in this paper handles the transformation of the dependency graph to a static scheduling.
@@
 It generates a program in the back-end language which embeds the parallel pattern or skeleton of the overall simulation.
-Various models could be used as back-end, such as OpenMP~\cite{660313} for example, however this paper adopts a novel approach where the dump targets a component-based runtime.
+Various models could be used as back-end, such as OpenMP~\cite{660313} for example. However this paper adopts a novel approach where the dump targets a component-based runtime.
 This approach makes it possible to leverage good properties of software component models such as increased code re-use, eased separation of concerns, and increased maintainability and productivity.
@@
 Thanks to the software components based approach, the generated code is made of four independent parts that loosely match the elements of the multi-stencil program definition.
-Those are the components implementing: \textbf{a)} the computations scheduling ($\Gamma$), \textbf{b)} the data structures ($\mathcal{M}$ and $\mathcal{E}$), \textbf{c)} data of the simulation ($\Delta$ and $\mathcal{D}$), and \textbf{d)} the time loop ($T$).
+Those are the components implementing: \emph{a)} the computations scheduling ($\Gamma$), \textbf{b)} the data structures ($\mathcal{M}$ and $\mathcal{E}$), \textbf{c)} data of the simulation ($\Delta$ and $\mathcal{D}$), and \textbf{d)} the time loop ($T$).
 The actual implementation of the numerical expressions are provided by the user and used in the computations scheduling.
-This component-based parallel structure enables to change the technology used for each part while limiting the impact on the other parts, as further discussed in Section~\ref{sect:component}.
+This component-based parallel structure enables to change the technology used for each part while limiting the impact on the other parts, as discussed in Section~\ref{sect:component}.
 

### 2015-08-21 WOLFHPC15/article/ms_msc.tex
Big Fig 3 splits into 2.

@@
-%----------------------------------------
-\subsection{Multi Stencil Language}
-%----------------------------------------
-
 \begin{filecontents*}{grammar.txt}
@@
           |  comp
-comp ::= dataid "[" compdomid "]=" compid "(" 
-            listdataread ")"
+comp ::= dataid "[" compdomid "]=" compid
+            "(" listdataread ")"
 listdataread ::= dataread listdataread
@@
 
-\begin{figure*}[t]
-\subfloat[\label{fig:grammar}]
-{\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left,linewidth=0.45\textwidth]{grammar.txt}}
-%\hfill
-\hspace{1.5cm}
-\subfloat[\label{fig:mslex}]
-{\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left,linewidth=0.45\textwidth]{example.txt}}
-\caption{(a) Grammar of the Multi-Stencil Language, (b) Example of a MSL program}
-\end{figure*}
+\begin{figure}[t]
+  \hspace{5mm}
+  \begin{minipage}[t]{.49\textwidth}
+    %  \subfloat[\label{fig:grammar}]
+    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left,linewidth=.87\textwidth]{grammar.txt}}   
+    \caption{Grammar of the Multi-Stencil Language. \label{fig:grammar}}
+  \end{minipage}
+\end{figure}
+
+\begin{figure}[t]
+\hspace{5mm}
+  \begin{minipage}[t]{.49\textwidth}
+    %\subfloat[\label{fig:mslex}]
+    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left,linewidth=0.87\textwidth]{example.txt}}
+    \caption{Example of a MSL program. \label{fig:mslex}}
+      \end{minipage}
+\end{figure}
+
+%----------------------------------------
+\section{Language and Parallelization}
+\label{sect:msmsc}
+%----------------------------------------
+\subsection{Multi Stencil Language}
+%----------------------------------------
 
 As mentioned in the previous section, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations.
-Six main sections are required in a MSL description that match the six elements of the formal definition; they describe: \textbf{1.}\,the mesh, \textbf{2.}\,the mesh entities, \textbf{3.}\,the computation domains with their inter-dependencies, \textbf{4.}\,the data, \textbf{5.}\,the time iterations, and finally \textbf{6.} the computations.
+Six main sections are required in a MSL description that match the six elements of the formal definition; they describe: \textbf{1)}\,the mesh, \textbf{2)}\,the mesh entities, \textbf{3)}\,the computation domains with their inter-dependencies, \textbf{4)}\,the data, \textbf{5)}\,the time iterations, and finally \textbf{6)} the computations.
 Figure~\ref{fig:grammar} shows the grammar of MSL.
@@
 As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is simply defined by a string identifier.
-The different kind of mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
-Figure~\ref{fig:mslex} presents an example of a mesh named \texttt{cart} used to represent a Cartesian mesh, and three kind of mesh entities: \texttt{cell}, \texttt{edgex}, and \texttt{edgey}.
+The different kind of mesh entities are described as a list of identifiers (Lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
+Figure~\ref{fig:mslex} gives an example of a mesh named \texttt{cart} used to represent a Cartesian mesh, and three kinds of mesh entities: \texttt{cell}, \texttt{edgex}, and \texttt{edgey}.
 

### 2015-08-21 WOLFHPC15/article/ms_msc.tex
Rel S4

@@
 On lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each kind of mesh entities: \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they are used to represent the whole domain.
-On lines 7 and 8 two other domains of \texttt{edgex} are defined: \texttt{part1edgex} and \texttt{part2edgex}; they are used to represent the boundaries of the domain.
+On lines 7 and 8, two other domains of \texttt{edgex} are defined: \texttt{part1edgex} and \texttt{part2edgex}; they are used to represent the boundaries of the domain.
 On line 10 the independence relation between \texttt{part1edgex} and \texttt{part2edgex} is specified.
@@
 This description does not contain a direct expression of the numerical expressions $exp$.
-Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} simply corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
+Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
 The description starts by the identifier of the computed data element ($w$ in the formal description) with the computation domain between brackets ($d$).
@@
 In the example of Figure~\ref{fig:mslex}, nine computations are described.
-On line 24, the data element \texttt{b} is computed on the domain \texttt{allcell} by the kernel expression \texttt{c\_0} which read data \texttt{a} without neighborhood shape.
+On line 24, the data element \texttt{b} is computed on the domain \texttt{allcell} by the kernel expression \texttt{c\_0} which reads data \texttt{a} without any neighborhood shape.
 
@@
 It makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
-This parallelization phase is divided in five different steps to transform $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
-These steps are described in this section.
+This parallelization phase is divided in five different steps to transform $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}. This sections describes these steps.
 %This section gives an overview of these five steps , but they are more formally defined with their associated algorithms in a research report~\cite{????}.
@@
 \subsubsection*{The Synchronized Ordered List of Computations: $\Gamma_{sync}$}
-In a data parallelization, data is split among processors and the same program is applied on each sub-part of the data.
+Data parallelization splits data among processors and the same program is applied on each sub-part of the data.
 However, this parallelization technique requires additional synchronizations between processors.
@@
 The same transformation is performed for the two other stencils: $c_4$ and $c_8$.
-Thus, $\Gamma_{sync} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
+Thus, for the example of Figure~\ref{fig:mslex}, $\Gamma_{sync} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
 
@@
 \end{tikzpicture}
-\caption{$\Gamma_{hybrid}$ of the example of Figure~\ref{fig:mslex}}
+\caption{$\Gamma_{hybrid}$ of the example of Figure~\ref{fig:mslex}.}
 \label{fig:hyb}
@@
 \begin{center}
-\subfloat[][The forbidden $N$ shape.\label{fig:n}]{
+\subfloat[][The forbidden $N$ shape.\\\label{fig:n}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]

### 2015-08-21 WOLFHPC15/article/component.tex
Rel S5

@@
 %----------------------------------------
-This section describes the projection of a MSP program to a component-based runtime, including the components used to implement the static scheduling of $\Gamma_{tsp}$.
-Before those details, the section gives an overview of component models, and especially the Low Level Components (\llc) used in this paper.
+This section describes the projection of a MSL program to a component-based runtime, including the components used to implement the static scheduling of $\Gamma_{tsp}$.
+The section first gives an overview of component models, and especially the Low Level Components (\llc) used in this work.
 
@@
 This work makes use of \llc for the experiments.
-Let us therefore introduce the concepts of this model in more details.
+%Let therefore introduce this model in more details.
 
@@
 
-\llc is a minimalist \texttt{C++} based HPC-oriented component model
+\llc~\cite{l2c} is a minimalist \texttt{C++} based HPC-oriented component model
 where a component extends the concept of class.
@@
 A \llc-based application is a static \emph{assembly} of components instances and connections between their ports.
-Such an assembly is described in LAD, an XML dialect~\cite{l2c}.
+Such an assembly is described in LAD, an XML dialect.
 
@@
 %----------------------------------------
-\subsection{Runtime Overview}
+\subsection{MSP Component Runtime Overview}
 %----------------------------------------
@@
 \vspace*{.5em}
-\caption{MSCAC Component Assembly Back-End Overview.}
+\caption{MSCAC Component-based Runtime Overview.}
 \label{fig:mscac:assembly}
@@
 
-As described in Section~\ref{sect:mscac}, the compiler generates four independent parts in a component assembly in addition to a static part.
-This is represented on Figure~\ref{fig:mscac:assembly} where each rounded box represents one component instance, drawn with simple lines, or more, drawn with double lines.
-For example, the component \texttt{Data} is instanciated more than once.
+As described in Section~\ref{sect:mscac}, the compiler generates four independent parts in a component assembly in addition to a static part as represented in Figure~\ref{fig:mscac:assembly}: each rounded box represents one component instance, drawn with simple lines, or more, drawn with double lines.
+For example, the component \texttt{Data} is instantiated more than once.
 \texttt{Driver} is static and drives the whole application.
@@
 \texttt{DDS} is used by \texttt{Data}, for which one instance represents one data element in the MSL program.
-\texttt{Time} is parametrized by the the \texttt{iteration} value from MSL and uses the \texttt{Computations} component as many times as required.
+\texttt{Time} is parametrized by the \texttt{iteration} value from MSL and uses the \texttt{Computations} component as many times as required.
 
@@
 At the leaves of $\Gamma_{tsp}$ and $\Gamma_{data}$, the computation components instances (denoted $K$), provided by the user, are connected to \texttt{Data}.
-As presented on Figure~\ref{fig:k}, a $K$ component provides a single port used to launch the computation and exposes one use port for each data element manipulated in the computation.
-We denote the fact that it may have several use port by adding a star on the use port to \texttt{Data}.
-
+As shown in Figure~\ref{fig:k}, a $K$ component provides a port to launch the computation and exposes a use port for each data element manipulated in the computation; a star is added on the use port to \texttt{Data} to denote it.
 \begin{figure}
@@
 \end{center}
-\caption{Kernel component.}
+\caption{A kernel component $K$.}
 \label{fig:k}
@@
 
-Moreover, to finalize the data parallelization of the back-end component-based runtime, which is partially managed by \texttt{DDS} for the partitioning, and $sync$ operations in \texttt{Computations}, we have chosen to duplicate the component assembly of Figure~\ref{fig:mscac:assembly} on available processors (or cores), connecting them by MPI ports between their \texttt{Data}.
+A complete assembly is obtained by replicating the component assembly of Figure~\ref{fig:mscac:assembly} on available processors (or cores).
+Connections is done with MPI ports of \texttt{Data} components.
 
@@
 Similarly, changing the approach used for scheduling would only impact the code generated in \texttt{Computations}.
-The technology used for data parallelism can also be changed by replacing the \texttt{DDS} and \texttt{Data} components; and by using the adequate interfaces in $K$.
-As a matter of fact, a work in progress proposes an alternative based on Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} instead of SkelGIS.
+The technology used for data parallelism can also be changed by replacing the \texttt{DDS} and \texttt{Data} components, and by using the adequate interfaces in $K$.
+As a matter of fact, we have started evaluating an alternative based on Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} instead of SkelGIS.
 
@@
 The series-parallel tree decomposition $\Gamma_{tsp}$ represents a static scheduling of the simulation.
-$\Gamma_{data}$, on the other hand, is an optimized fusionned ordered list of computations.
+$\Gamma_{data}$, on the other hand, is an optimized merged ordered list of computations.
 Both are dumped to components by generating an assembly that exactly matches their structure.
@@
 A control component exposes a single \emph{provide} port containing a \emph{control-only} method, without any parameter.
-We introduce three control components represented in Figure~\ref{fig:ctrlcomponents}.
+Three control components have been needed: SEQ, PAR, and SYNC. They are represented in Figure~\ref{fig:ctrlcomponents}.
+
+\begin{description}
+\item[Sequence component (SEQ)] It is the direct representation of a sequence node of $\Gamma_{tsp}$ or $\Gamma_{data}$.
+This component sequentially calls an ordered list of other components.
+It exposes an ordered \emph{use-multiple} port to be connected to the components to call in sequence.
+
+\item[Parallel component (PAR)] It is the direct representation of a parallel node of $\Gamma_{tsp}$ (not $\Gamma_{data}$).
+This component simultaneously calls a set of other components.
+It exposes a \emph{use-multiple} port to be connected to the components to call in parallel.
+
+\item[Synchronization component (SYNC)] It is the direct representation to an update leaf of $\Gamma_{tsp}$ or $\Gamma_{data}$.
+This component calls the synchronization of a given data.
+It exposes a \emph{use} port to be connected to the data to be updated.
+\end{description}
+
+Then, from $\Gamma_{tsp}$ or $\Gamma_{data}$, and using the control components described above,
+a direct dump can be done to a component assembly for a data or hybrid (data
+and task) parallelization of a simulation.
+Figure~\ref{fig:tsp:assembly} displays the assembly part corresponding to
+$\Gamma_{tsp}$ of Figure~\ref{fig:tsp}. In this figure, the ports
+linked to data (use and use-multiple ports of SYNC and K) are
+represented but are not connected for readability. Moreover, each computation
+component is an instance of a specific K component using
+the identification name of the computation of an MSL instance.
+Thus, this figure represents what the component \texttt{Computations} of Figure~\ref{fig:mscac:assembly} contains if an hybrid parallelization is performed.
+On the other hand, if a data parallelization is performed, a single $SEQ$ component is used by \texttt{Time}, which invokes in the appropriate order the computation kernels and synchronizations.
+Among those kernels, a single \emph{merged} kernel would represent $c_3$ and $c_5$.
 
@@
 \\
-\caption{The three control components used by the back-end.}
+\caption{The three control components.}
 \label{fig:ctrlcomponents}
 \end{figure}
-
-\begin{description}
-\item[Sequence component (SEQ)] It is the direct representation of a sequence node of $\Gamma_{tsp}$ or $\Gamma_{data}$.
-This component sequentially calls an ordered list of other components.
-It exposes an ordered \emph{use-multiple} port to be connected to the components to call in sequence.
-
-\item[Parallel component (PAR)] It is the direct representation of a parallel node of $\Gamma_{tsp}$ (not $\Gamma_{data}$).
-This component simultaneously calls a set of other components.
-It exposes a \emph{use-multiple} port to be connected to the components to call in parallel.
-
-\item[Synchronization component (SYNC)] It is the direct representation to an update leaf of $\Gamma_{tsp}$ or $\Gamma_{data}$.
-This component calls the synchronization of a given data.
-It exposes a \emph{use} port to be connected to the data to update.
-\end{description}
-
-Then from $\Gamma_{tsp}$ or $\Gamma_{data}$, and using the control components described above,
-a direct dump can be done to a component assembly for a data or hybrid (data
-and task) parallelization of a simulation.
-Figure~\ref{fig:tsp:assembly} displays the assembly part corresponding to
-$\Gamma_{tsp}$ of Figure~\ref{fig:tsp}. In this figure, the ports
-linked to data (use and use-multiple ports of SYNC and K) are
-represented but are not connected. Moreover, each computation
-component is an instance of a specific K component using
-the identification name of the computation in the MSL file.
-Thus, this figure represents what the component \texttt{Computations} of Figure~\ref{fig:mscac:assembly} contains if an hybrid parallelization is performed.
-On the other hand, if a data parallelization is performed, a single $SEQ$ component is used by \texttt{Time}, which itself uses in the appropriate order the computation kernels and synchronizations.
-Among those kernels, a single one \emph{fusionned} represents $c_3$ and $c_5$.
 

### 2015-08-21 WOLFHPC15/article/evaluation.tex
last results g5k helene 13_08
Rel S6

@@
 Navier-Stokes equations is a well known set of PDEs in fluid dynamics to simulate a flow evolution in time.
-At the University of Orl\'eans, France, the MAPMO laboratory works on a software, called FullSWOF~\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}, which solves the Shallow water equations obtained from the three dimensional Navier-Stokes equations, by averaging on the vertical direction (see e.g.~\cite{Ferrari2004}).
+At the University of Orl\'eans, France, the MAPMO laboratory works on a software, called FullSWOF\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}, which solves the Shallow water equations obtained from the three dimensional Navier-Stokes equations, by averaging on the vertical direction~\cite{Ferrari2004}.
 Those equations are solved using a two-dimensional Cartesian discretization of the space domain, and a finite volume numerical methods more described in~\cite{CPE:CPE3494}.
-The MSL program of FullSWOF contains three mesh entities, seven computation domains, fourty-eight data and ninty-eight computations.
-Among these computations thirty-two stencil kernels and sixty-six local kernels are performed in the simulation.
+We have developed a MSL version of FullSWOF that contains 3~mesh entities, 7~computation domains, 48~data and 98~computations.
+These computations are made of 32~stencil kernels and 66~local kernels.
 
-\subsubsection*{Compiler Evaluation}
+\subsection{Compiler Evaluation}
 
-The series-parallel tree decomposition $\Gamma_{tsp}$ of this simulation, extracted by the MSC transformation, is composed of seventeen sequence nodes and eighteen parallel nodes.
+The series-parallel tree decomposition $\Gamma_{tsp}$ of this simulation, extracted by the MSC transformation, is composed of 17~sequence nodes and 18~parallel nodes.
 Figure~\ref{fig:freq} represents for a given level of parallelism, \ie the number of tasks to perform concurrently, the number of time this level is observed in the final component assembly.
 One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1).
-As a level of sixteen parallel tasks is reached too times, and also five times for the level four, sequential restrictions could be amortized.
+As a level of 16~parallel tasks is reached two times, and also five times for the fourth level, sequential restrictions could be amortized.
 However, it is rather clear that the task parallelization technique should be used to improve the data parallelism when reaching its limits, but not to use alone.
@@
     \hline 
-   level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
+   Level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
    \hline
-   frequence & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
+   Frequence & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
    \hline
@@
   \hline
-   step & $\Gamma_{sync}$ & $\Gamma_{dep}$ & $\Gamma_{msp}$ & $\Gamma_{tsp}$\\
+   Step & $\Gamma_{sync}$ & $\Gamma_{dep}$ & $\Gamma_{msp}$ & $\Gamma_{tsp}$\\
    \hline
-   time (ms) & 2 & 530 & 8297 & 1133\\
+   Time (ms) & 2 & 530 & 8297 & 1133\\
    \hline
@@
 Figure~\ref{fig:exectime} illustrates the execution time for each step of the MSC transformation for an overall execution time of ten seconds.
-Execution times have been computed on a laptop with a bi-core Intel Core i5 1,4 GHz, and 8Gb DDR3.
-One can notice that the transformation of $\Gamma_{dep}$ to a minimal series-parallel graph is the longest step of MSC, because of the removal of the forbidden shapes in the graph.
+Execution times have been computed on a laptop with a bi-core Intel Core i5 1.4~GHz, and 8~GB of DDR3.
+One can notice that the transformation of $\Gamma_{dep}$ to a minimal series-parallel graph is the longest step of MSC, because of the removal of the forbidden N-shapes in the graph.
 Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted by the algorithm, which applies a general solution (instead of finding each forbidden shape), but it seems that many of them appear.
-The \emph{N-shape}, represented in Figure~\ref{fig:n} is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.
+%The \emph{N-shape}, represented in Figure~\ref{fig:n} is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.
 However, a \emph{N-shape} could perfectly be handled by a dynamic scheduler, for example.
 Thus, the fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static shedule of tasks may not be the best solution for complex simulations.
-It has to be noticed, however, that the all computation of $\Gamma_{tsp}$ is still usefull to detect how space loops can be fusionned in $\Gamma_{data}$.
+It has to be noticed, however, that the all computation of $\Gamma_{tsp}$ is still usefull to detect how space loops can be merged in $\Gamma_{data}$.
 
-\subsubsection*{Preliminary Performances}
+\subsection{Preliminary Performance Evaluation}
 To evaluate performance of the generated component-based parallel structure, we have proceeded as follows.
@@
 For this reason, our evaluation compare the shallow water equations first parallelized with a pure SkelGIS code (data structures, applicators and interfaces of SkelGIS~\cite{CPE:CPE3494}), and second parallelized with MSL (which uses the SkelGIS data structure).
-Moreover, as SkelGIS library handles data parallelization for distributed memory architectures, we have limited for now our evaluations to the fusionned data parallelization of MSCAC (dumped from $\Gamma_{data}$).
+Moreover, as SkelGIS library handles data parallelization for distributed memory architectures, we have limited for now our evaluations to the merged data parallelization of MSCAC (dumped from $\Gamma_{data}$).
 
-Evaluations have been performed on the cluster \emph{Edel} of \emph{Grid'5000}~\footnote{\url{https://www.grid5000.fr/mediawiki/index.php/Grid5000:Home}}.
-Figure~\ref{tab:hard} shows hardware configuration of this cluster.
-On \emph{Edel} have been performed three different experiments, each one described with the size of the Cartesian mesh and the number of time iterations: (1) $5k \times 5k$, $500$ iterations; (2) $10k \times 10k$, $500$ iterations; (3) $10k \times 10k$, $2k$ iterations.
+Evaluations have been performed on the cluster \emph{Edel} of \emph{Grid'5000}\footnote{\url{https://www.grid5000.fr/}}.
+Figure~\ref{tab:hard} displays the features of this cluster.
+On \emph{Edel}, three different experiments have been performed that differ in the size of the Cartesian mesh and the number of time iterations: (1) $5,000 \times 5,000$, $500$ iterations; (2) $10,000 \times 10,000$, $500$ iterations; (3) $10,000 \times 10,000$, $2,000$ iterations.
 
@@
 
-Figure~\ref{fig:perfs} shows results on the three different experiments.
-On the first line are represented, for each experiment, the number of iteration computed per second for a given number of cores.
-On the second line are represented, for each experiment, the execution time  for a given number of cores, using a logarithmic scale.
+Figure~\ref{fig:perfs} reports the results of the three experiments.
+The first row of graphs is the number of iterations computed per second for various number of cores.
+The second row of graphs represents the execution time for various number of cores, using a logarithmic scale.
 
@@
 \subfloat[$10k \times 10k$ mesh, $2k$ iterations\label{fig:speedup3}]{
-\resizebox{5.5cm}{!}{\includegraphics{./images/itpersec_10K_500.pdf}}
+\resizebox{5.5cm}{!}{\includegraphics{./images/itpersec_10K_2K.pdf}}
 }
@@
 \subfloat[$10k \times 10k$ mesh, $2k$ iterations\label{fig:log23}]{
-\resizebox{5.5cm}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
+\resizebox{5.5cm}{!}{\includegraphics{./images/logtimes_10K_2K.pdf}}
 }
@@
 This result can be explained by the automatic generation of the parallel component-based runtime.
-Actually, writing a pure SkelGIS code, communications between processes are hidden from the user through specific interfaces called \emph{applicators}~\cite{CPE:CPE3494}.
+Actually, in a pure SkelGIS code, communications between processes are hidden from the user through specific interfaces called \emph{applicators}~\cite{CPE:CPE3494}.
 Using MSL, on the other hand, communications are directly managed by $SYNC$ components, which are automatically placed in the component-based runtime by MSCAC.
-Thus, no overheads are introduced to hide communications.
+Thus, no overhead is introduced to hide communications.
 
@@
 Actually, the less the execution time is, the less the scalability capacity is.
-Thus, these results show that using a component-based runtime does not damage performance of the back-end code.
-Finally, we notice that SkelGIS has proved its scalability on the Shallow water equations compared to an MPI parallelization~\cite{CPE:CPE3494}, which makes these performance results relevant.
+Thus, these results confirm the result of~\cite{l2c} that using a component-based runtime does not damage performance of the back-end code.
+Finally, we notice that SkelGIS has proved its scalability on the Shallow water equations compared to an MPI parallelization~\cite{CPE:CPE3494}, that makes out performance results relevant.

### 2015-08-21 WOLFHPC15/article/related.tex
Rel S7

@@
 Those solutions can be considered as stencil compilers.
-Actually they produce optimized (cache tiling etc.) or parallel (CUDA, OpenCL etc.) codes for a single stencil computation.
+Actually they produce optimized (cache tiling, etc.) or parallel (CUDA, OpenCL, etc.) codes for a \emph{single} stencil computation.
 Recently, domain specific languages for very specific numerical methods such as ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} for multigrid solvers, and OPS~\cite{Reguly:2014:ODS:2691166.2691173} for meshes of meshes (hybrid mesh) have also been proposed.
@@
 As a result, MSL is not a new contribution to stencil compilers, or to distributed data structures, but proposes a coarse-grain automatic parallelization of an overall multi-stencil program.
-Finally, MSL is a first case study to show that component-based runtimes could improve code-reuse, separation of concerns and thus maintainability of DSLs.
+Finally, MSL is a quite complex case study that indicates that component-based runtimes could improve code-reuse, separation of concerns and thus maintainability of DSLs without impacting performance.
 

### 2015-08-21 WOLFHPC15/article/conclusion.tex
Rel S8

@@
-In this paper, we have presented the domain specific language MSL and its compiler MSCAC designed to produce a parallel (data and hybrid) component-based runtime of an overall multi-stencil program, \ie a mesh-based numerical simulation reduced to explicit schemes.
+This paper has presented the domain specific language MSL and its compiler MSCAC designed to produce a parallel (data and hybrid) component-based runtime of an overall multi-stencil program, \ie a mesh-based numerical simulation reduced to explicit schemes.
 After details on the language and its compiler, MSL has been evaluated by the description of a real case numerical simulation: the shallow water equations.
@@
 Those results demonstrate that component-based runtimes may be relevant back-end codes for DSLs as they do not introduce performance damage.
-Moreover, interests of component-based runtimes has been discussed in the paper.
+Moreover, components bring software engineering benefits such as separation of concern, code re-use, improving maintainability.
 
@@
 In addition to this, alternatives for the \texttt{Computations} component, computed by MSC, are under study such as a dump to an OpenMP~\cite{660313} code or the use of dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}.
-This last work on dynamic schedulers could also improve the expressivity of task parallelism in MSL, as it has been discussed in Section~\ref{sect:eval}, and thus could bring interesting performance for the hybrid parallelization.
+This last work on dynamic schedulers could also improve the expressivity of task parallelism in MSL, as it has been discussed in Section~\ref{sect:eval}, and thus it could bring interesting performance for the hybrid parallelization.

### 2015-08-21 WOLFHPC15/article/conclusion.tex
minor diff

@@
 Those results demonstrate that component-based runtimes may be relevant back-end codes for DSLs as they do not introduce performance damage.
-Moreover, components bring software engineering benefits such as separation of concern, code re-use, improving maintainability.
+Moreover, components bring software engineering benefits such as separation of concerns, code re-use, improving maintainability.
 
@@
 In addition to this, alternatives for the \texttt{Computations} component, computed by MSC, are under study such as a dump to an OpenMP~\cite{660313} code or the use of dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}.
-This last work on dynamic schedulers could also improve the expressivity of task parallelism in MSL, as it has been discussed in Section~\ref{sect:eval}, and thus it could bring interesting performance for the hybrid parallelization.
+This last work on dynamic schedulers could also improve the expressivity of task parallelism in MSL, as discussed in Section~\ref{sect:eval}.
+Thus, it could bring interesting performance for the hybrid parallelization.

### 2015-08-21 WOLFHPC15/article/mscac.tex
minor diff

@@
 The first phase \emph{parses} the MSL description into an intermediate representation (IR).
-The second phase called MSC transforms the computation part ($\Gamma$) of the IR in order to automatically introduce parallelism in the overall numerical simulation.
+The second phase called MSC transforms the computation part ($\Gamma$) of the IR into a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393} to introduce parallelism in the overall numerical simulation.
 Finally, the third phase dumps the IR to an actual implementation of the parallel pattern of the simulation.

### 2015-09-21 WOLFHPC15/article/article.tex
G5k ack
corrections article reviewer 3 + bib en plus Halide DSL -  Helene

@@
 As a scientist is not a low level and high performance programming expert, Domain Specific Languages (DSLs) are a promising solution to automatically and efficiently write high performance codes.
-However, if DSLs ease programming for scientists, maintainability and portability issues are deported from scientists to DSL designers.
+However, if DSLs ease programming for scientists, maintainability and portability issues are transferred from scientists to DSL designers.
 This paper deals with an approach to improve maintainability and programming productivity of DSLs through the generation of a component-based parallel runtime.
@@
 %----------------------------------------
-\section{Language and Parallelization Description}
-\label{sect:msmsc}
 \input{ms_msc.tex}
@@
 
-We acknowledge Grid'5000 for evaluations.
+Experiments presented in this paper were carried out using the
+Grid'5000 testbed, supported by a scientific interest group hosted by
+Inria and including CNRS, RENATER and several Universities as well as
+other organizations (see \url{https://www.grid5000.fr}).
 

### 2015-09-21 WOLFHPC15/article/concept.tex
corrections article reviewer 3 + bib en plus Halide DSL -  Helene

@@
 %To resume, concepts used in this paper to define a stencil kernel are \emph{mesh}, \emph{mesh entities}, \emph{data}, \emph{computation domain}, \emph{numerical expression} and \emph{stencil shape}. 
-If stencil kernels have been studied a lot, the formalization of real overall mesh-based numerical simulations is a rather new contribution.
+While stencil kernels have been studied a lot, the formalization of real overall mesh-based numerical simulations is a rather new contribution.
 Actually, paying attention to complex numerical simulations, it appears that most of them are composed of more than one stencil kernel, with one or more stencil shapes and of additional local computations.
@@
 Let $\Delta$ be the set of data of the simulation
-A stencil kernel $s$ is defined as the quintuplet:
+A stencil kernel $s$ is defined as the quadruplet:
 \begin{equation} 

### 2015-09-21 WOLFHPC15/article/evaluation.tex
corrections article reviewer 3 + bib en plus Halide DSL -  Helene

@@
    \hline
-   Frequence & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
+   Frequency & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
    \hline
@@
 Thus, these results confirm the result of~\cite{l2c} that using a component-based runtime does not damage performance of the back-end code.
-Finally, we notice that SkelGIS has proved its scalability on the Shallow water equations compared to an MPI parallelization~\cite{CPE:CPE3494}, that makes out performance results relevant.
+Finally, we notice that SkelGIS has proved its scalability on the Shallow water equations compared to an MPI parallelization~\cite{CPE:CPE3494}, that makes our performance results relevant.

### 2015-09-21 WOLFHPC15/article/introduction.tex
corrections article reviewer 3 + bib en plus Halide DSL -  Helene

@@
 
-Many solutions to automatically optimize and parallelize numerical simulations have been proposed in the literature, either as specific frameworks~\cite{CPE:CPE3494,Trilinos-Overview,petsc-efficient} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
+Many solutions to automatically optimize and parallelize numerical simulations have been proposed in the literature, either as specific frameworks~\cite{CPE:CPE3494,Trilinos-Overview,petsc-efficient} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS, Ragan-Kelley:2013:HLC:2491956.2462176}.
 Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil kernel.

### 2015-09-21 WOLFHPC15/article/component.tex
corrections article reviewer2 + related + reduction taille -  Helene

@@
 
-\begin{description}
-\item[Sequence component (SEQ)] It is the direct representation of a sequence node of $\Gamma_{tsp}$ or $\Gamma_{data}$.
+% \begin{description}
+% \item[Sequence component (SEQ)] It is the direct representation of a sequence node of $\Gamma_{tsp}$ or $\Gamma_{data}$.
 This component sequentially calls an ordered list of other components.
@@
 
-\item[Parallel component (PAR)] It is the direct representation of a parallel node of $\Gamma_{tsp}$ (not $\Gamma_{data}$).
+% \item[Parallel component (PAR)] It is the direct representation of a parallel node of $\Gamma_{tsp}$ (not $\Gamma_{data}$).
 This component simultaneously calls a set of other components.
@@
 
-\item[Synchronization component (SYNC)] It is the direct representation to an update leaf of $\Gamma_{tsp}$ or $\Gamma_{data}$.
+% \item[Synchronization component (SYNC)] It is the direct representation to an update leaf of $\Gamma_{tsp}$ or $\Gamma_{data}$.
 This component calls the synchronization of a given data.
 It exposes a \emph{use} port to be connected to the data to be updated.
-\end{description}
+% \end{description}
+
+\textbf{Sequence component (SEQ).} It is the direct representation of a sequence node of $\Gamma_{tsp}$ or $\Gamma_{data}$.
+This component sequentially calls an ordered list of other components.
+It exposes an ordered \emph{use-multiple} port to be connected to the components to call in sequence.
+
+\textbf{Parallel component (PAR).} It is the direct representation of a parallel node of $\Gamma_{tsp}$ (not $\Gamma_{data}$).
+This component simultaneously calls a set of other components.
+It exposes a \emph{use-multiple} port to be connected to the components to call in parallel.
+
+\textbf{Synchronization component (SYNC).} It is the direct representation to an update leaf of $\Gamma_{tsp}$ or $\Gamma_{data}$.
+This component calls the synchronization of a given data.
+It exposes a \emph{use} port to be connected to the data to be updated.
 

### 2015-09-21 WOLFHPC15/article/concept.tex
corrections article reviewer2 + related + reduction taille -  Helene

@@
 %To resume, concepts used in this paper to define a stencil kernel are \emph{mesh}, \emph{mesh entities}, \emph{data}, \emph{computation domain}, \emph{numerical expression} and \emph{stencil shape}. 
-While stencil kernels have been studied a lot, the formalization of real overall mesh-based numerical simulations is a rather new contribution.
+While stencil kernels have been studied a lot, the formalization of real overall mesh-based numerical simulations is poorly studied and not exactly for the same kind of appication~\cite{Ragan-Kelley:2013:HLC:2491956.2462176}.
 Actually, paying attention to complex numerical simulations, it appears that most of them are composed of more than one stencil kernel, with one or more stencil shapes and of additional local computations.
-For example, how to formalize and parallelize a numerical simulation which chains the stencil kernels of Figures~\ref{fig:ex1} and~\ref{fig:ex2}.
+For example, we would like to formalize and parallelize a numerical simulation which chains the stencil kernels of Figures~\ref{fig:ex1} and~\ref{fig:ex2}.
 The next section formalizes concepts of \emph{stencil kernel}, \emph{local kernel} and \emph{multi-stencil program}.
@@
 The set of computations $\Gamma$ is an ordered list of stencil and local kernels.
-One can notice that this work, for now, is limited to a single type of mesh in a given simulation.
+One can notice that this work, for now, is limited to a single mesh type for a given simulation.
 

### 2015-09-21 WOLFHPC15/article/conclusion.tex
corrections article reviewer2 + related + reduction taille -  Helene

@@
 First, to more clearly show the improvement of DSLs maintainability using component-based back-end, an alternative \texttt{DDS} component is under study, using Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.
-In addition to this, alternatives for the \texttt{Computations} component, computed by MSC, are under study such as a dump to an OpenMP~\cite{660313} code or the use of dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}.
-This last work on dynamic schedulers could also improve the expressivity of task parallelism in MSL, as discussed in Section~\ref{sect:eval}.
+In addition to this, alternatives for the \texttt{Computations} component, computed by MSC, are under study such as a dump to a pure OpenMP~\cite{660313} code or the use of dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}.
+%This last work on dynamic schedulers could also improve the expressivity of task parallelism in MSL, as discussed in Section~\ref{sect:eval}.
 Thus, it could bring interesting performance for the hybrid parallelization.

### 2015-09-21 WOLFHPC15/article/evaluation.tex
corrections article reviewer2 + related + reduction taille -  Helene

@@
 As a level of 16~parallel tasks is reached two times, and also five times for the fourth level, sequential restrictions could be amortized.
-However, it is rather clear that the task parallelization technique should be used to improve the data parallelism when reaching its limits, but not to use alone.
+Moreover, each identified task can itself be parallelized, introducing data parallelism on loops.
+Thus, it is rather clear that the task parallelization technique should be used to improve the data parallelism when reaching its limits, but not to use alone.
 Moreover, as the level of parallelism in the application is heterogenous, the number of threads to launch for task parallelism, and the number of cores to keep for data parallelism is difficult to choose.
@@
 One can notice that the transformation of $\Gamma_{dep}$ to a minimal series-parallel graph is the longest step of MSC, because of the removal of the forbidden N-shapes in the graph.
-Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted by the algorithm, which applies a general solution (instead of finding each forbidden shape), but it seems that many of them appear.
+%Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted by the algorithm, which applies a general solution (instead of finding each forbidden shape), but it seems that many of them appear.
 %The \emph{N-shape}, represented in Figure~\ref{fig:n} is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.
@@
 To evaluate performance of the generated component-based parallel structure, we have proceeded as follows.
-Our current implementation of the MSCAC compiler generates a back-end code using the SkelGIS distributed data structure.
-For this reason, our evaluation compare the shallow water equations first parallelized with a pure SkelGIS code (data structures, applicators and interfaces of SkelGIS~\cite{CPE:CPE3494}), and second parallelized with MSL (which uses the SkelGIS data structure).
+Our current implementation of the MSCAC compiler generates a back-end code using the SkelGIS distributed data structure as an underlying library.
+For this reason, our evaluation compare the shallow water equations first parallelized with a pure SkelGIS code (data structures, applicators and interfaces of SkelGIS~\cite{CPE:CPE3494}), and second parallelized with MSL (which uses the SkelGIS data structure only).
 Moreover, as SkelGIS library handles data parallelization for distributed memory architectures, we have limited for now our evaluations to the merged data parallelization of MSCAC (dumped from $\Gamma_{data}$).
 
-Evaluations have been performed on the cluster \emph{Edel} of \emph{Grid'5000}\footnote{\url{https://www.grid5000.fr/}}.
-Figure~\ref{tab:hard} displays the features of this cluster.
+Evaluations have been performed on the cluster \emph{Edel} of \emph{Grid'5000}\footnote{\url{https://www.grid5000.fr/}} for which each node is equiped with two 8-cores Intel Xeon E5520, 24GB of memory, and is interconnected with an Infiniband 40G network.
+%Figure~\ref{tab:hard} displays the features of this cluster.
 On \emph{Edel}, three different experiments have been performed that differ in the size of the Cartesian mesh and the number of time iterations: (1) $5,000 \times 5,000$, $500$ iterations; (2) $10,000 \times 10,000$, $500$ iterations; (3) $10,000 \times 10,000$, $2,000$ iterations.
-
-\begin{figure}[!h]
-\begin{center}
- \begin{tabular}{|c|c|c|}
-    \hline
-    Cluster & \textbf{Edel}\\
-    \hline         
-    Processor & 2 Intel Xeon E5520\\
-    & (2.27 GHz)\\
-    Cores/node & 8 \\
-    Memory/node & 24 GB \\
-    Compiler & OpenMPI [-O3] \\
-    Network & InfiniBand 40G\\
-    \hline
- \end{tabular}
-   \caption{\label{tab:hard}Hardware of clusters used in the experiments.}
- \end{center}
-\end{figure}
-
+% \begin{figure}[!h]
+% \begin{center}
+%  \begin{tabular}{|c|c|c|}
+%     \hline
+%     Cluster & \textbf{Edel}\\
+%     \hline         
+%     Processor & 2 Intel Xeon E5520\\
+%     & (2.27 GHz)\\
+%     Cores/node & 8 \\
+%     Memory/node & 24 GB \\
+%     Compiler & OpenMPI [-O3] \\
+%     Network & InfiniBand 40G\\
+%     \hline
+%  \end{tabular}
+%    \caption{\label{tab:hard}Hardware of clusters used in the experiments.}
+%  \end{center}
+% \end{figure}
 Figure~\ref{fig:perfs} reports the results of the three experiments.
@@
 Thus, no overhead is introduced to hide communications.
-
 In addition to this, the scalability slope, shown by the number of iterations per second, is close to SkelGIS.

### 2015-09-21 WOLFHPC15/article/introduction.tex
corrections article reviewer2 + related + reduction taille -  Helene

--- 2015-08-21 WOLFHPC15/article/introduction.tex
@@
 Conversely, the more specific the DSL is, the less applications can use it, and the more difficult it is to amortize the implementation cost of the DSL (language, compiler, and runtime).
-Programming complexity, as well as non-maintainability and non-portability often seem to be simply deported from scientists to DSL designers.
+Programming complexity, as well as non-maintainability and non-portability often seem to be simply transferred from scientists to DSL designers.
 % Moreover, one can consider that end users need as much DSLs as domains, which keeps the amount of work to write and maintain up-to-date DSLs out of hands.
@@
 Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil kernel.
-However, a real case numerical simulation (such as fluid dynamics, magneto-hydrodynamics, molecular dynamics etc.) is most of the time not composed of a single stencil kernel, but of a set of stencil kernels and a set of additional local auxiliary computations.
+However, a real case numerical simulation (such as fluid dynamics, magneto-hydrodynamics, molecular dynamics etc.) is most of the time not composed of a single stencil kernel, but of a set of stencil kernels and a set of additional local auxiliary computations~\cite{Ragan-Kelley:2013:HLC:2491956.2462176}.
 The DSL proposed in this paper, called \emph{Multi-Stencil Language} or MSL, is a DSL to generate a parallel component-based structure of a numerical simulation, that we call a \emph{multi-stencil program}.
-As discussed in the related work, this work is complementary to the optimization and parallelization of a single stencil kernel.
+As discussed in the related work, this work is complementary to the optimization and parallelization of a single stencil kernel and is independent from implementation choices (as for example the form of the studied mesh).
 

### 2015-09-21 WOLFHPC15/article/ms_msc.tex
minor diff
corrections article reviewer2 + related + reduction taille -  Helene

@@
 On lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each kind of mesh entities: \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they are used to represent the whole domain.
-On lines 7 and 8, two other domains of \texttt{edgex} are defined: \texttt{part1edgex} and \texttt{part2edgex}; they are used to represent the boundaries of the domain.
+On lines 7 and 8, two other domains of \texttt{edgex} are defined: \texttt{part1edgex} and \texttt{part2edgex}; they could be used to represent the boundaries of the domain, for example, or any other sub-part of the whole domain.
+It is the user's responsability to define the needed domains for his simulation.
 On line 10 the independence relation between \texttt{part1edgex} and \texttt{part2edgex} is specified.
@@
 However, as a computation kernel writes a single data $w$ (to finely manage dependencies between tasks), the same computation domain might be iterated more than once, while some numerical expressions could be computed inside a joint space loop.
-For this reason, an optimized and \emph{fusionned} ordered list of computation $\Gamma_{data}$ is built from $\Gamma_{tsp}$.
+For this reason, an optimized and \emph{merged} ordered list of computation $\Gamma_{data}$ is built from $\Gamma_{tsp}$.
 Actually, in $\Gamma_{tsp}$, computations of a parallel subtree with a same computation domain can be computed in the same space loop, in any order.

### 2015-09-21 WOLFHPC15/article/related.tex
corrections article reviewer2 + related + reduction taille -  Helene

@@
 
-Many specific solutions exist to ease parallel programming of numerical simulations~\cite{Trilinos-Overview,petsc-efficient}.
+Many specific solutions exist to ease parallel programming of numerical simulations, %~\cite{Trilinos-Overview,petsc-efficient}.
 %Solutions such as PETSc~\cite{petsc-efficient} and Trilinos~\cite{Trilinos-Overview} are specific frameworks.
@@
 The purpose of those solutions are much more specific than our solution.
-The closest solutions to MSL are DSLs which propose to generate parallel codes from a sequential imperative stencil description.
+as for example DSLs which propose to generate parallel codes from a sequential imperative stencil description.
 For example, PATUS~\cite{citeulike12258902} and Pochoir~\cite{spaaTangCKLL11} propose a simple description to write stencil codes for general structured meshes, and OP2~\cite{Giles2011} and Liszt~\cite{DeVito2011LDS} for general unstructured meshes.
@@
 
+Halide~\cite{Ragan-Kelley:2013:HLC:2491956.2462176} is a DSL which focusses on the automatic parallelization and optimization of stencil pipelines in image processing.
+Thus, this work is the closest to MSL.
+However, at least one important difference can be noticed: the specific domain it is applied on.
+Actually, image processing is applied onto 2D or 3D grids, while explicit numerical schemes to solve PDEs can be applied on many different kind of meshes (2D, 3D grids, but also unstructured or hybrid meshes).
+This main difference is the reason why MSL is a descriptive and agnostic language, and why its optimizations capabilities are limited compared to DSLs applied on known data structures (making available tiling optimizations etc.).
+
 The work presented in this paper takes place at a different abstraction level of numerical simulations and can be seen as complementary to stencil compilers or distributed data structures.
-Actually, the MSL language is agnostic and should work for any type of mesh and any neighborhood shape in stencil codes.
-The MSCAC compiler only needs this agnostic description of the overall simulation to generate a parallel pattern of the program.
+The MSCAC compiler only needs the agnostic description of the overall simulation to generate a parallel pattern of the program.
 This parallel pattern then needs to be filled with implementation parameters and computation codes.
@@
 Computation codes could be generated by stencil compilers or could be written using interfaces of distributed data structures (as it is done using SkelGIS in this paper).
-As a result, MSL is not a new contribution to stencil compilers, or to distributed data structures, but proposes a coarse-grain automatic parallelization of an overall multi-stencil program.
+As a result, MSL is not a new contribution to stencil compilers, to tiling optimizations, or to distributed data structures, but proposes a coarse-grain automatic parallelization of an overall multi-stencil program.
 Finally, MSL is a quite complex case study that indicates that component-based runtimes could improve code-reuse, separation of concerns and thus maintainability of DSLs without impacting performance.

### 2015-09-25 WOLFHPC15/article/evaluation.tex
copyright ACM template + corrections Helene

@@
 %Actually, the number of forbidden shapes removed in $\Gamma_{dep}$ is not counted by the algorithm, which applies a general solution (instead of finding each forbidden shape), but it seems that many of them appear.
-%The \emph{N-shape}, represented in Figure~\ref{fig:n} is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.
+The \emph{N-shape}, represented in Figure~\ref{fig:n} is forbidden in a minimal series-parallel graph as it is not possible to exactly express it using sequences and parallel sections.
 However, a \emph{N-shape} could perfectly be handled by a dynamic scheduler, for example.
-Thus, the fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static shedule of tasks may not be the best solution for complex simulations.
+Thus, the fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static schedule of tasks may not be the best solution for complex simulations.
 It has to be noticed, however, that the all computation of $\Gamma_{tsp}$ is still usefull to detect how space loops can be merged in $\Gamma_{data}$.
 
+%\pagebreak
 \subsection{Preliminary Performance Evaluation}
@@
 
-Evaluations have been performed on the cluster \emph{Edel} of \emph{Grid'5000}\footnote{\url{https://www.grid5000.fr/}} for which each node is equiped with two 8-cores Intel Xeon E5520, 24GB of memory, and is interconnected with an Infiniband 40G network.
+Evaluations have been performed on the cluster \emph{Edel} of \emph{Grid'5000}\footnote{\url{https://www.grid5000.fr/}} 
+for which each node is equiped with two 8-cores Intel Xeon E5520, 24GB of memory, and is interconnected with an Infiniband 40G network. 
 %Figure~\ref{tab:hard} displays the features of this cluster. 

### 2015-09-25 WOLFHPC15/article/related.tex
copyright ACM template + corrections Helene

@@
 The purpose of those solutions are much more specific than our solution.
-as for example DSLs which propose to generate parallel codes from a sequential imperative stencil description.
-For example, PATUS~\cite{citeulike12258902} and Pochoir~\cite{spaaTangCKLL11} propose a simple description to write stencil codes for general structured meshes, and OP2~\cite{Giles2011} and Liszt~\cite{DeVito2011LDS} for general unstructured meshes.
+as for example %DSLs which propose to generate parallel codes from a sequential imperative stencil description. For example, 
+PATUS~\cite{citeulike12258902} and Pochoir~\cite{spaaTangCKLL11} which propose a simple description to write stencil codes for general structured meshes, and OP2~\cite{Giles2011} and Liszt~\cite{DeVito2011LDS} for general unstructured meshes.
 Those solutions can be considered as stencil compilers.
 Actually they produce optimized (cache tiling, etc.) or parallel (CUDA, OpenCL, etc.) codes for a \emph{single} stencil computation.
-Recently, domain specific languages for very specific numerical methods such as ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} for multigrid solvers, and OPS~\cite{Reguly:2014:ODS:2691166.2691173} for meshes of meshes (hybrid mesh) have also been proposed.
+Recently, domain specific languages for very specific numerical methods such as ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} for multigrid solvers, have also been proposed.%, and OPS~\cite{Reguly:2014:ODS:2691166.2691173} for meshes of meshes (hybrid mesh) have also been proposed.
 All those solutions let the user implement their own stencil codes in a sequential programming style, and produce high performance codes for shared memory architectures and sometimes distributed memory architectures.
@@
 However, at least one important difference can be noticed: the specific domain it is applied on.
-Actually, image processing is applied onto 2D or 3D grids, while explicit numerical schemes to solve PDEs can be applied on many different kind of meshes (2D, 3D grids, but also unstructured or hybrid meshes).
-This main difference is the reason why MSL is a descriptive and agnostic language, and why its optimizations capabilities are limited compared to DSLs applied on known data structures (making available tiling optimizations etc.).
+Actually, image processing is applied onto 2D or 3D grids, while explicit numerical schemes to solve PDEs can be applied on many different meshes (grids, but also unstructured or hybrid meshes).
+This main difference is the reason why MSL is a descriptive and agnostic language, and why it does not target specific optimizations, as tiling for example.%optimizations capabilities are limited compared to DSLs applied on known data structures (making available tiling optimizations etc.).
 
-The work presented in this paper takes place at a different abstraction level of numerical simulations and can be seen as complementary to stencil compilers or distributed data structures.
+The work presented in this paper takes place at a different abstraction level and can be seen as complementary to stencil compilers or distributed data structures.
 The MSCAC compiler only needs the agnostic description of the overall simulation to generate a parallel pattern of the program.
@@
 %. Thus, depending on the underlying distributed data structure used, those computation codes 
-Computation codes could be generated by stencil compilers or could be written using interfaces of distributed data structures (as it is done using SkelGIS in this paper).
+Those computation codes could be generated by stencil compilers or could be written using existing distributed data structures (as SkelGIS in this paper).
 As a result, MSL is not a new contribution to stencil compilers, to tiling optimizations, or to distributed data structures, but proposes a coarse-grain automatic parallelization of an overall multi-stencil program.
-Finally, MSL is a quite complex case study that indicates that component-based runtimes could improve code-reuse, separation of concerns and thus maintainability of DSLs without impacting performance.
+Finally, MSL is a case study to show that component-based runtimes could improve maintainability adn flexibility of DSLs with a low impact on performance.
 

### 2015-09-25 WOLFHPC15/article/evaluation.tex
copyright ACM template + corrections Helene

--- 2015-09-21 WOLFHPC15/article/evaluation.tex
@@
 However, a \emph{N-shape} could perfectly be handled by a dynamic scheduler, for example.
-Thus, the fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static shedule of tasks may not be the best solution for complex simulations.
+Thus, the fact that many \emph{N-shape} are removed in $\Gamma_{dep}$ shows that the creation of a static schedule of tasks may not be the best solution for complex simulations.
 It has to be noticed, however, that the all computation of $\Gamma_{tsp}$ is still usefull to detect how space loops can be merged in $\Gamma_{data}$.
 
+%\pagebreak
 \subsection{Preliminary Performance Evaluation}

### 2015-09-30 WOLFHPC15/article/ms_msc.tex
figures fonttype=truetype corrections Helene

@@
 mesh: cart
-mesh domains: cell,edgex,edgey
+mesh entities: cell,edgex,edgey
 computation domains:

### 2015-09-30 WOLFHPC15/results_g5k/SW_10K_2K/results.py

@@ -3,6 +3,9 @@ import matplotlib
 import matplotlib.pyplot as plt
 from matplotlib.backends.backend_pdf import PdfPages
 
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
 cores = [16,32,64,128,256]
 timesMSL = [9035.94,4519.9,2390.61,1300.08,693.294]
 timesSK = [13692.4,6946.07,3235.88,1861.24,892.882]

### 2015-09-30 WOLFHPC15/results_g5k/SW_5K_500/results.py

@@ -3,6 +3,9 @@ import matplotlib
 import matplotlib.pyplot as plt
 from matplotlib.backends.backend_pdf import PdfPages
 
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
 cores = [8,16,32,64,128,256]
 timesMSL = [998.1,544.87,285.73,139.8,76.45,48.05]
 timesSK = [1324.9,638.5,371.62,197.32,117.047,64.96]

### 2015-10-06 WOLFHPC15/article/introduction.tex

@@
 
-Many solutions to automatically optimize and parallelize numerical simulations have been proposed in the literature, either as specific frameworks~\cite{CPE:CPE3494,Trilinos-Overview,petsc-efficient} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
+Many solutions to automatically optimize and parallelize numerical simulations have been proposed in the literature, either as specific frameworks~\cite{CPE:CPE3494,petsc-efficient} or as DSLs~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}.
 Most of the time, DSLs focus on the parallelization and the optimization of a single numerical computation, also called a stencil kernel.

### 2015-10-06 WOLFHPC15/article/mscac.tex

@@
-The solution proposed in this paper is based on both a domain specific language to describe a multi-stencil program, \ie a complete numerical simulation, and on a compiler which parses the language, applies transformations for automatic parallelization, and dumps the representation to a component-based runtime.
+The solution proposed in this paper is based on both a domain specific language to describe a multi-stencil program, \ie a complete numerical simulation, and on a compiler which parses the language, applies transformations for automatic parallelization, and dumps the representation to a component-based runtime, ready-to-fill with the user code.
 

### 2015-10-06 rapport/formalism.tex

--- 2015-08-06 rapport/formalism.tex
@@
 %\rmq{Pas très clair la suite; motivation? dom == some sub elements of M ?}
-For simulations, data is discretized based on the mesh; but for a given mesh, there are various ways of mapping the data depending on the mesh entities on which the data is mapped.
-\begin{mydef}
-\todo{JB: vraiment, ta def dit que pour chaque couple (noeud, arete), tu associes un element de $D_i$ ...}
-$D_i$ is a set of elements of a mesh $\mathcal{M}=(V,E)$, constructed by a function $dom_i$ which defines a precise association between $V$ and $E$, $dom_i : V \times E \rightarrow D_i$.
-\end{mydef}
+For simulations, data is discretized based on the mesh; but for a given mesh, there are various ways of mapping the data depending on the mesh entities on which the data is applied.
+% \begin{mydef}
+% \todo{JB: vraiment, ta def dit que pour chaque couple (noeud, arete), tu associes un element de $D_i$ ...}
+% $D_i$ is a set of elements of a mesh $\mathcal{M}=(V,E)$, constructed by a function $dom_i$ which defines a precise association between $V$ and $E$, $dom_i : V \times E \rightarrow D_i$.
+% \end{mydef}
 
 \begin{mydef}
-\todo{JB: def alternative}
-An entity $d$ of a mesh $\mathcal{M}=(V,E)$ is a subset of the mesh elements: $d\subset (V\cup E)$.
-A kind of mesh entities $D$ is a set of entities: $D\subset \mathcal{P}(V\cup E)$
+%\todo{JB: def alternative}
+An entity $e$ of a mesh $\mathcal{M}=(V,E)$ is a subset of the mesh elements: $e\subset (V\cup E)$.
+A kind of mesh entities $E$ is a set of entities: $E\subset \mathcal{P}(V\cup E)$
 \end{mydef}
 For example, in a 2D Cartesian mesh a first kind of entities are the cells ($D_0$), they can be defined as all sets containing exactly four vertices and four edges connected as a cycle.
-Another kind of entities are the vertices ($D_1$) that can be defined as all singletons formed of a single vertex each.
+Another kind of entities are the vertices ($D_1$) that can be defined as all singletons formed of a single vertex.
 
@@
 \begin{mydef}
-A quantity is a function $\delta_t: D_{\delta t} \mapsto V_{\delta t}$ which associates each entity of its definition domain $d\in D_{\delta t}$ to a value $v\in V_{\delta t}$ where $V_{\delta t}$ is the data type, typically $\mathbb{R}$ or $\mathbb{C}$.
+A quantity is a function $\delta: E_{\delta} \mapsto V_{\delta}$ which associates each entity of its definition domain $e\in E_{\delta}$ to a value $v\in V_{\delta}$ where $V_{\delta}$ is the data type, typically $\mathbb{R}$ or $\mathbb{C}$.
 The set of quantities applied on the mesh is denoted $\Delta$.
-In the rest of this paper, the entities on which a quantity $\delta_t$ is mapped is denoted $dom(\delta_t)=D_{\delta t}$.
+In the rest of this paper, the entities on which a quantity $\delta$ is mapped is denoted $entity(\delta)=E_{\delta}$.
 \end{mydef}
-Another approach closer to that taken in applied mathematics would have been to define $\delta$ functions over $D \times T$; however the approach chosen in this paper where a distinct function is defined for each time-step better matches the implementation of numerical simulations where the outer loop iterates over time and where only a few time-steps of quantity values are stored.
+Another approach closer to that taken in applied mathematics would have been to define $\delta$ functions over $E_{\delta} \times T$; however the approach chosen in this paper where a distinct function is defined for each time-step better matches the implementation of numerical simulations where the outer loop iterates over time and where only a few time-steps of quantity values are stored.
 
@@
 \begin{mydef}
-A numerical expression $\text{exp}_{wt}: dom(w_t)\times R\mapsto V_{wt}$ is a function that computes the value of a quantity $w_t(d)$ (write quantity) for a given entity $d \in dom(w_t)=D$, using the set of its input data $R=\{\delta 1_{t-1}, \delta 2_{t-1}, ...\} \subset \Delta$ (read quantities).
-It is defined as $\text{exp}_{wt}(d,r) \rightarrow w_t(d)$.
+A numerical expression $\text{exp}_{w}: entity(w)\times R\mapsto V_{w}$ is a function that computes the value of a quantity $w(e)$ (written quantity) for a given entity $e \in entity(w)=E$, using the set of its input data $R=\{e_{1}, e_{2}, ...\} \subset \Delta$ (read quantities).
+It is defined as $\text{exp}_{w}(e,r) \rightarrow w(e)$.
+\todo{JB: tu n'introduis pas la notion de stencil ici?}
+\end{mydef}
+
+\begin{mydef}
+A computation kernel $k$ of a numerical simulation is defined as $k(R,w,\text{exp})$, where $R \subset \Delta, w \in \Delta$ and $\text{exp}$ a numerical expression.
+%$D$ is one of the subsets $D_i \subset \mathcal{M}$, such that $w : D \rightarrow V$.
+\end{mydef}
+% It has to be noticed that at each time iteration, all the values of a quantity are computed.
+However, it happens that computations of the quantity is split in different domains, as for example the computation of the physical border.
+In this case additional $D_i$ can be specified for the mesh $\mathcal{M}$.
+% \todo{JB: je comprends pas vraiment...}
+
+\begin{mydef}
+The set of $n$ ordered computations of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i \leq j$, then $k_i$ is computed before $k_j$, and $k_j$ can be computed when $k_i$ is finished.
+%\todo{JB: je dirais que $c_j$ peut etre calculé quand $c_j$ l'est, j'eleverais le "only", c'est justement parce qu'il n'y a pas le only que tu peux faire du task parallelism}
 \end{mydef}
@@
 \begin{mydef}
-A computation $c$ of a numerical simulation is defined as $c(R,w,\text{exp})$, where $R \subset \Delta, w \in \Delta$ and $\text{exp}$ a numerical expression.
-%$D$ is one of the subsets $D_i \subset \mathcal{M}$, such that $w : D \rightarrow V$.
-\end{mydef}
-It has to be noticed that at each time iteration, all the elements of a mesh are computed.
-However, it happens that computations of the mesh elements are splitted in different domains, as for example the computation of the physical border.
-In this case additional $D_i$ can be specified for the mesh $\mathcal{M}$.
-
-\begin{mydef}
-The set of $n$ ordered computations of a numerical simulation is denoted $\Gamma = [c_i]_{0 \leq i \leq n-1}$, such that $\forall c_i,c_j \in \Gamma$, if $i \leq j$, then $c_i$ is computed before $c_j$, and $c_j$ can be computed only when $c_i$ is finished.
-\end{mydef}
-
-\begin{mydef}
-A \textit{multi-stencil program} is defined by the quadruplet $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$.
+A \textit{multi-stencil program} is defined by the sextuplet $\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$.
 \end{mydef}

### 2015-11-11 rapport/formalism.tex
rapport, presentation WOLF / Helene

--- 2015-08-06 rapport/formalism.tex
@@
 
+$\Omega=\mathbb{R}^n$ is the continuous space domain of a numerical simulation. Any geo-referenced point in this space can be represented.
+
 A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ of a set of PDEs and is defined as follows.
@@
 \begin{mydef}
-\textit{A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V\in \Omega$ is the set of vertices and $E\in V^2$ the set of edges.
+\textit{A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V\subset \Omega$ is the set of vertices and $E\subseteq V^2$ the set of edges.
 The set of edges $E$ of a mesh $\mathcal{M}=(V,E)$ does not contain bridges.}
@@
 \end{center}\end{figure}
-A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) and hybrid.
+A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) and hybrid as illustrated in Figure~\ref{fig:mesh}.
 
-%\rmq{Pas très clair la suite; motivation? dom == some sub elements of M ?}
-For simulations, data is discretized based on the mesh; but for a given mesh, there are various ways of mapping the data depending on the mesh entities on which the data is mapped.
-\begin{mydef}
-\todo{JB: vraiment, ta def dit que pour chaque couple (noeud, arete), tu associes un element de $D_i$ ...}
-$D_i$ is a set of elements of a mesh $\mathcal{M}=(V,E)$, constructed by a function $dom_i$ which defines a precise association between $V$ and $E$, $dom_i : V \times E \rightarrow D_i$.
-\end{mydef}
+\noindent \textbf{Definitions}
+%\begin{mydef}
+\begin{itemize}
+\item An entity $e$ of a mesh $\mathcal{M}=(V,E)$ is a subset of its vertices and edges, $e\subset (V\cup E)$. 
+\item The set of mesh entities, of a same given type, is denoted $E$ such that $E\subset \mathcal{P}(V\cup E)$. 
+\item Finally the set of mesh entities of a simulation is denoted $\mathcal{E}$.
+\end{itemize}
+%\end{mydef}
 
-\begin{mydef}
-\todo{JB: def alternative}
-An entity $d$ of a mesh $\mathcal{M}=(V,E)$ is a subset of the mesh elements: $d\subset (V\cup E)$.
-A kind of mesh entities $D$ is a set of entities: $D\subset \mathcal{P}(V\cup E)$
-\end{mydef}
-For example, in a 2D Cartesian mesh a first kind of entities are the cells ($D_0$), they can be defined as all sets containing exactly four vertices and four edges connected as a cycle.
-Another kind of entities are the vertices ($D_1$) that can be defined as all singletons formed of a single vertex each.
+For example, in a 2D Cartesian mesh a first kind of entities are the cells, they can be defined as all sets containing exactly four vertices and four edges connected as a cycle.
+Another kind of entities simply are the vertices ($V$) and can be defined as all singletons formed of a single vertex of $V$.
 
+\medskip
 This covers the space discretization, however another dimension manipulated in simulations that has to be discretized is the time dimension.
-\begin{mydef}
-The discretization of the continuous time domain $\mathcal{T}$ is denoted $T$ such that $\forall\mbox{ }t_i\mbox{, }t_{i+1} \in T\mbox{, }\exists\mbox{ }\Delta t \in \mathbb{R}$\mbox{, }$t_{i+1} = t_i + \Delta t$.
-\end{mydef}
+
+\noindent \textbf{Definitions}
+%\begin{mydef}
+\begin{itemize}
+\item $\mathcal{T}=\mathbb{R}$ is the continuous time domain of a numerical simulation.
+\item The discretization of the continuous time domain $\mathcal{T}$ is denoted $T$ such that $\forall\mbox{ }t_i\mbox{, }t_{i+1} \in T\mbox{, }\exists\mbox{ }\Delta t \in \mathbb{R}$\mbox{, }$t_{i+1} = t_i + \Delta t$.
+\end{itemize}
+%\end{mydef}
+
 $T$ is responsible for the iteration time steps of the numerical simulation.
+A numerical simulation though does not run forever and must be stopped at some point.
+Some simulations choose a fix number of time iteration, others compute a \emph{convergence} criteria at the end of each time iteration to determine if the simulation has to continue.
+A convergence criteria is computed by a reduction computation that will be detailed in the next section.
 
 In a numerical simulation a set of data elements, or quantities, are applied onto the mesh and represent the set of values to compute, or to use, for computation.
-\begin{mydef}
-A quantity is a function $\delta_t: D_{\delta t} \mapsto V_{\delta t}$ which associates each entity of its definition domain $d\in D_{\delta t}$ to a value $v\in V_{\delta t}$ where $V_{\delta t}$ is the data type, typically $\mathbb{R}$ or $\mathbb{C}$.
-The set of quantities applied on the mesh is denoted $\Delta$.
-In the rest of this paper, the entities on which a quantity $\delta_t$ is mapped is denoted $dom(\delta_t)=D_{\delta t}$.
-\end{mydef}
-Another approach closer to that taken in applied mathematics would have been to define $\delta$ functions over $D \times T$; however the approach chosen in this paper where a distinct function is defined for each time-step better matches the implementation of numerical simulations where the outer loop iterates over time and where only a few time-steps of quantity values are stored.
+
+\noindent \textbf{Definitions}
+%\begin{mydef}
+\begin{itemize}
+\item A quantity is a function $\delta: E_{\delta} \mapsto V_{\delta}$ which associates each entity of a type $e\in E_{\delta}$ to a value $v\in V_{\delta}$ where $V_{\delta}$ is the data type, typically $\mathbb{R}$ or $\mathbb{C}$.
+\item The set of quantities applied onto the mesh is denoted $\Delta$.
+\item In the rest of this paper, the type of mesh entities on which a quantity $\delta$ is mapped is denoted $entity(\delta)=E_{\delta}$.
+\end{itemize}
+%\end{mydef}
+
+Another approach closer to that taken in applied mathematics would have been to define $\delta$ functions over $E_{\delta} \times T$; however the approach chosen in this paper where a distinct function is defined for each time-step better matches the implementation of numerical simulations where the outer loop iterates over time and where only a few time-steps of quantity values are stored.
 
@@
 
-\begin{mydef}
-A numerical expression $\text{exp}_{wt}: dom(w_t)\times R\mapsto V_{wt}$ is a function that computes the value of a quantity $w_t(d)$ (write quantity) for a given entity $d \in dom(w_t)=D$, using the set of its input data $R=\{\delta 1_{t-1}, \delta 2_{t-1}, ...\} \subset \Delta$ (read quantities).
-It is defined as $\text{exp}_{wt}(d,r) \rightarrow w_t(d)$.
-\end{mydef}
+In this section are considered two different types of computations, a reduction or a numerical kernel.
 
-%JB: fin de relecture
-
-\begin{mydef}
-A computation $c$ of a numerical simulation is defined as $c(R,w,\text{exp})$, where $R \subset \Delta, w \in \Delta$ and $\text{exp}$ a numerical expression.
-%$D$ is one of the subsets $D_i \subset \mathcal{M}$, such that $w : D \rightarrow V$.
-\end{mydef}
-It has to be noticed that at each time iteration, all the elements of a mesh are computed.
-However, it happens that computations of the mesh elements are splitted in different domains, as for example the computation of the physical border.
-In this case additional $D_i$ can be specified for the mesh $\mathcal{M}$.
-
-\begin{mydef}
-The set of $n$ ordered computations of a numerical simulation is denoted $\Gamma = [c_i]_{0 \leq i \leq n-1}$, such that $\forall c_i,c_j \in \Gamma$, if $i \leq j$, then $c_i$ is computed before $c_j$, and $c_j$ can be computed only when $c_i$ is finished.
-\end{mydef}
-
-\begin{mydef}
-A \textit{multi-stencil program} is defined by the quadruplet $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$.
-\end{mydef}
+\noindent \textbf{Definitions}
+%\begin{mydef}
+\begin{itemize}
+\item A computation domain $D$ is a subpart of mesh entities of a given type, $D \subset E \in \mathcal{E}$.
+\item The set of computation domains of a numerical simulation is denoted $\mathcal{D}$.
+\item A neighborhood $n$ is a function which for a given entity $e \in E_i$, returns a set of $m$ entities in $E_j$, $n : E_i \rightarrow E_j^m$. It is possible that $E_i = E_j$.
+\item The set of neihborhood functions in a numerical simulation is denoted $\mathcal{N}$.
+\item A set of data to read and used in a computation is denoted $R \subset \Delta \times \mathcal{N}$.
+It is the set of pairs $(r,n)$, where $r \in \Delta$ and $n$ is a neighborhood function such that $n : E_i \rightarrow entity(r)^m$.
+\item A numerical expression $\text{exp}_{w}: entity(w) \times R \rightarrow V_{w}$ is a function that computes the value of the written quantity $w \in \Delta$ for a given entity $e \in entity(w)=E$, using a set of input data $R$.
+\item A computation kernel $k$ of a numerical simulation is defined as $k(R,w,\text{exp},D)$, where $R \subset \Delta \times \mathcal{N}$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$.
+\end{itemize}
+%\end{mydef}
 %If the number of computations in $\Gamma$ is $card(\Gamma)=n$, such that $\bigcup_{i=0}^{n-1}c_i = \Gamma$, then $\bigcup_{i=0}^{m-1}R_i \cup w_i \subseteq \Delta$.
 
-As already mentioned, the ordered list $\Gamma$ can be composed of two different types of computations, stencil and local computations, which will be defined in the rest of this Section.
+Most of the time we can dissociate two types of kernel computations.
 
-\begin{mydef}
-The neighborhood $\mathcal{N}$ of an element $d \in D_i$ is a function to obtain a set of elements in any $D_k \subset \mathcal{M}$, $\mathcal{N} : D_i \rightarrow D_k \times D_k \times \dots$.
-\end{mydef}
-The function $\mathcal{N}$ is also sometimes called the \textit{stencil shape}, or the \textit{stencil} in applied mathematics.
-In this paper we distinguish a stencil shape from a \textit{stencil computation} defined as followed:
+\noindent \textbf{Definitions}
+%\begin{mydef}
+\begin{itemize}
+\item We denote by $identity$ the identity function $x=x$.
+\item A kernel computation $k(R,w,\text{exp},D) \in \Gamma$ is a stencil kernel $\iff \exists (r,n) \in R$ such that $n \neq identity$.
+\item A kernel computation $k(R,w,\text{exp},D) \in \Gamma$ is a local kernel $\iff \forall (r,n) \in R$, $n = identity$.
+\item The set of $n$ ordered computations of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i \leq j$, then $k_i$ is computed before $k_j$.
+\end{itemize}
 
-\begin{mydef}
-A \textit{stencil computation} is defined as a quadruplet $s(R,w,\text{exp},\mathcal{N})$, where $R \subset \Delta$, and $w : D \rightarrow V \in \Delta$.
-\end{mydef}
-In a stencil computation $s$, $\forall d \in D$, the stencil numerical expression $\text{exp}$ is applied such that $w(d) = \text{exp}(R(d),R(\mathcal{N}(d))$.
-In this work, a stencil computation $s(R,w,\text{exp},\mathcal{N})$ always verifies $R \cap \{w\} = \emptyset$, otherwize an implicit numerical scheme has to be solve which is over the scope of this paper.
-As a result, the ordered list $\Gamma$ of a multi-stencil program can be composed of a set of stencil computations applied on one or more stencil shapes.
+A reduction computation is typically used to compute the convergence criteria of the time loop of the simulation.
+Occasionally reductions can also be performed during a time iteration.
+This last usage is typically done to place a conditionnal branch to chose one computation or another.
+This paper only study the use of reduction computations as convergence citerion.
 
-Figure~\ref{fig:ex} gives an example of a stencil computation $s(R,w,\text{exp},\mathcal{N})$, where $\mathcal{M}(V,E)$ is a two dimensional Cartesian mesh.
-A single domain $D=dom(w)$ is defined in this example and is composed of cells formed by a cycle of four vertices $v \in V$ and four edges $e \in E$.
-Furthermore, in this example $R=\{A\}$, $w=B$, and for $(x,y) \in D$ the neighborhood function is
+\noindent \textbf{Definitions}
+%\begin{mydef}
+\begin{itemize}
+\item A reduction computation is defined as the triplet $red(R,)$
+\item ...
+\item A \textit{multi-stencil program} is defined by the sextuplet $\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$.
+\end{itemize}
+%\end{mydef}
+
+For example, in Figure~\ref{fig:ex1}, assuming the computation domain (full lines) is denoted $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined as:
 \begin{equation*}
-\mathcal{N} : (x,y) \rightarrow \{(x,y+1),(x,y-1),(x+1,y),(x-1,y)\}.
+R: \{(B,n1)\}, \quad w: A, \quad d: dc1,
 \end{equation*}
-Finally, the numerical expression of this example is 
 \begin{equation*}
-\text{exp}(A(x,y),A(\mathcal{N}(x,y)) = B(x,y) = A(x,y)+(A(x,y+1)+A(x,y-1)+A(x+1,y)+A(x-1,y))/4.
+exp: A(x,y)=B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1).
+\end{equation*}
+On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil kernel is defined as:
+\begin{equation*}
+R: \{(C,n2),(A,identity)\}, \quad w: A, \quad d: dc2,
+\end{equation*}
+\begin{equation*}
+exp: A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1).
 \end{equation*}
 
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{./images/example.pdf}}
-  \caption{Example of a stencil computation.}
-  \label{fig:ex}
-\end{center}\end{figure}
-
-The second type of numerical computation is a local computation.
-\begin{mydef}
-A local computation is a triplet $l(R,w,\text{exp})$, where $exp$ does not involve a neighborhood function $\mathcal{N}$.
-\end{mydef}
+\begin{figure}
+\begin{center}
+\subfloat[Mesh and mesh domains.\label{fig:mesh}]{
+\resizebox{8cm}{!}{\includegraphics{./images/mesh.pdf}}
+}\\
+\hspace{10pt}
+\subfloat[4-neighborhood stencil.\label{fig:ex1}]{
+\resizebox{5cm}{!}{\includegraphics{./images/stencil1.pdf}}
+}
+\vspace{20pt}
+\subfloat[4-neighborhood stencil.\label{fig:ex2}]{
+\resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
+}
+\end{center}
+\caption{(a) a Cartesian mesh and two kind of mesh entities, (b) an example of stencil kernel on cells, (c) an example of stencil kernel on two different entities of the mesh.}
+\label{fig:gspmsp}
+\end{figure}
 

### 2015-11-12 WOLFHPC15/presentation/slides.tex

\documentclass{beamer}

\usepackage{beamerthemeDresden}
\usepackage[frenchb]{babel}
\usepackage[utf8]{inputenc}
\usecolortheme{beaver}

\usepackage{listings}
\usepackage{xcolor}
\lstdefinestyle{base}{
  basicstyle=\ttfamily\color{black} \footnotesize,
  moredelim=**[is][\color{blue}]{@}{@},
  moredelim=**[is][\color{red}]{&}{&}
}

\setbeamertemplate{navigation symbols}{
\insertframenumber/
\inserttotalframenumber
}

\usepackage{color}
\definecolor{vert}{rgb}{0,0.5,0}
\definecolor{violet}{rgb}{0.5,0,0.5}

\usepackage{tikz}
\usetikzlibrary{calc,positioning,matrix,arrows,shapes.geometric,shapes.symbols,shapes.misc,shapes,automata,petri,decorations.markings,shadows}
\usepackage{array}
\usepackage{subfig}

%----------TIKZ
\usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}

\tikzset{
    use/.style={
    circle,draw=black,fill=black,scale=0.5,text=white
    },
    mpi/.style={
    rectangle,draw=black,fill=black,scale=0.8,text=white
    },
    provide/.style={
    circle,draw=black,fill=white,scale=0.5
    },
    component/.style={
    rectangle,rounded corners=3pt,draw=black
    },
    dcomponent/.style={
    rectangle,rounded corners=3pt,dashed,draw=black
    },
    progm/.style={rectangle,dashed, draw=black, thin, text width=10em, text centered,rounded corners, minimum height=2em},
  line/.style={draw, thin, ->, shorten >=2pt},
  purp/.style={rectangle, draw=violet,fill=violet!20, text width=10em, text centered,thin,rounded corners,inner sep=3pt},
  purp2/.style={rectangle, draw=violet,fill=violet!20, text width=15em, text centered,thin,rounded corners,inner sep=3pt},
}
%-------------

\def\pprec{\mathrel{\scalebox{.9}[1]{$\prec$}\mkern-3mu%
  \scalebox{.4}[1]{$\prec$}\mkern-5.5mu\scalebox{.4}[1]{$\prec$}}}

\makeatletter
    \newenvironment{withoutheadline}{
        \setbeamertemplate{headline}[default]
        \def\beamer@entrycode{\vspace*{-\headheight}}
    }{}
\makeatother

%-------------------------------------------------------------------
\title[From DSL to HPC Component-Based Runtime: A Multi-Stencil DSL Case Study]{From DSL to HPC Component-Based Runtime: A Multi-Stencil DSL Case Study}
\author[Julien Bigot (CEA), Hélène Coullon (INRIA), Christian Perez (INRIA)]{Julien Bigot, \underline{Hélène Coullon}, Christian Perez}
\institute[INRIA]{INRIA team Avalon\\Maison de la simulation (CEA)}
\date{$16^{th}$ November 2015}
%-------------------------------------------------------------------

\begin{document}

%---------------
%---------------
\begin{withoutheadline}

\begin{frame}
    \titlepage
\end{frame}

% INTRODUCTION / MOTIVATION
%-------------------------------------------------------------
\begin{frame}
\frametitle{Motivation} % HPC DSL
\begin{block}{+ Domain Specific Languages}
\begin{itemize}
\item Easy language for the user
\item Implicit optimizations
\item Implicit parallelization
\end{itemize}
\end{block}
\begin{alertblock}{- Domain Specific Languages}
\begin{itemize}
\item Difficulties deported to the DSL designer
\begin{itemize}
\item Low level high performance programming
\item Maintainability and portability
\end{itemize}
\item As many DSLs as domains
\begin{itemize}
\item Productivity and code-reuse
\item Separation of concerns
\end{itemize}
\end{itemize}
\end{alertblock}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Motivation} % components
\begin{block}{Component models}
\begin{itemize}
\item Divide an application into several functionnal black boxes
\item Each component defines its connections with outer world
\item Each component are built independently from the others
\item Application = Assembly of components
\end{itemize}
\end{block}
\begin{block}{+ Component models}
\begin{itemize}
\item Maintainability through separation of concerns
\item Code-reuse and productivity
\item Dynamic assembly of components
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Motivation} % components
What if a DSL produces a component-based runtime ?

\begin{itemize}
\item Is it efficient ?
\item Is it feasible ?
\item Does it improve issues of DSLs ?
\begin{itemize}
\item maintainability
\item portability
\item productivity
\end{itemize}
\end{itemize}

Let's take a useful example: \textit{the Multi-Stencil Language }!
\end{frame}
%-------------------------------------------------------------

% presentation of the table of contents
%-------------------------------------------------------------
\begin{frame}
\frametitle{Table of contents}
    \tableofcontents[hideallsubsections]
\end{frame}
%-------------------------------------------------------------

\end{withoutheadline}
%---------------
%---------------

%-------------------------------------------------------------
\section{Multi-Stencil Language}
% related work, multi-stencil problem, language
%-------------------------------------------------------------
%----------------
\subsection{Multi-Stencil Language}
%----------------
\begin{frame}
\frametitle{Numerical simulation = Multi-Stencil application}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}
\def \n {5}
\def \radius {3cm}
\def \margin {8} % margin in angles, depends on the radius

  \node[purp2,align=center] (edp) at ({360/\n * (2 - 1)}:\radius) {Partial Differential Equations\\\scriptsize $\frac{\partial u(x,y,t)}{\partial t} = \frac{\partial^2 u(x,y,t)}{\partial x^2} + \frac{\partial^2 u(x,y,t)}{\partial y^2}$};
  \node[progm,align=center,right of=edp,node distance=6cm] (cond) {\scriptsize + specific behavior for\\\scriptsize boundary conditions};
  \path[line,dashed] (edp) -- (cond);
  \node[purp,align=center] (maillage) at ({360/\n * (1 - 1)}:\radius) {Mesh\\Time iterations};
  \path[->] (maillage) edge[<-,thin,shorten <=2pt,shorten >=2pt, bend right=40]  node [fill=white,align=center] {\scriptsize Time and space\\\scriptsize discretization} (edp.east);
  \path[] (edp.west) edge[thin,shorten <=2pt,shorten >=2pt, bend right=90]  node [draw,circle,fill=white] (plus) {\scriptsize +} (maillage.west);
  \node[purp,align=center, below of=plus, node distance=3.5cm] (sche) {Explicit\\numerical schemes\\\textcolor{red}{stencils}};
  \node[progm,align=center,right of=sche,node distance=5.5cm] (expl) {\scriptsize Step by step approximation of the phenomena};
  \path[line,dashed] (sche) -- (expl);
  \path[line] (plus) -- node [fill=white,align=center] {\scriptsize Numerical mathods\\\scriptsize Finite difference/volume/elment} (sche);
\end{tikzpicture}
}
\end{center}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Time and Mesh}
\begin{block}{Time}
At each time iteration of the simulation are applied the \emph{computations kernels} of the application.
\end{block}
\begin{block}{Mesh}
\begin{itemize}
\item A Mesh is a connected undirected graph $\mathcal{M}=(V,E)$ without bridges
\item Mesh entities are a subset of $V \cup E$
\end{itemize}
\end{block}
\begin{center}
\resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
\end{center}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Data and Computation Kernels}
\begin{block}{Data}
A Data is a set of numerical values, each one attached to a given mesh entity
\end{block}
\begin{block}{Computation kernel}
\begin{itemize}
\item Set of data read for the computation
\begin{itemize}
\item Each one associated to a stencil shape
\end{itemize}
\item Data written by the computaion
\item A numerical expression
\item A computation domain
\begin{itemize}
\item Subset of mesh entities
\end{itemize}
\end{itemize}
\end{block}
\end{frame}
%----------------

%----------------
% \begin{frame}
% \frametitle{Computation kernel: Examples}
% \begin{center}
% \resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
% \hspace{10pt}
% \resizebox{5cm}{!}{\includegraphics{./images/stencil1.pdf}}
% \end{center}
% Compute $A$ using $\{(B,n1)\}$ on the computation domain called $dc1$ with the numerical expression
% \tiny
% \begin{equation*}
% A(x,y)=B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1).
% \end{equation*}
% \end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Computation kernel: Examples}
\begin{center}
\resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
\hspace{10pt}
\resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
\end{center}
Compute $A$ using $\{(A,local),(C,n2)\}$ on the computation domain called $dc2$ with the numerical expression
\tiny
\begin{equation*}
A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1).
\end{equation*}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Multi-Stencil program}
\begin{center}
$\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$
\end{center}
\begin{itemize}
\item $T$ the set of time iterations to tun the simulation
\item $\mathcal{M}$ the mesh of the simulation
\item $\mathcal{E}$ the set of mesh entities
\item $\mathcal{D}$ the set of computation domains
\item $\Delta$ the set of data
\item $\Gamma$ the set of computations
\end{itemize}
\center \textit{= the six sections of a \emph{Multi-Stencil Language} program !}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Multi-Stencil Language}
\begin{alertblock}{MSL is not}
\begin{itemize}
\item a new stencil optimizer/compiler
\item a new distributed data structure
\end{itemize}
\end{alertblock}
\begin{block}{MSL is}
\begin{itemize}
\item a high-level language for multi-stencil simulations
\item agnostic from the type of mesh used (data structure)
\item based on identifiers only
\end{itemize}
\end{block}
\begin{center}
\textit{MSL produces a ready-to-fill\\component-based parallel pattern of the simulation !}
\end{center}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Related Work}
\begin{block}{Complementary work}
\begin{itemize}
\item Distributed data structures: SkelGIS, Global Arrays
\item Stencil DSLs (on grids): Pochoir, PATUS
\item Stencil DSLs (on unstructured meshes): OP2, Liszt
\end{itemize}
\end{block}
\begin{alertblock}{Similar work}
\begin{itemize}
\item Pipeline of stencil computations for image processing: Halide
\item DSL to component-based runtime: ?
\end{itemize}
\end{alertblock}
\end{frame}
%----------------
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Overview}
% take the same notations inside components + user
%-------------------------------------------------------------
\subsection{Overview}

%----------------
\begin{frame}
\frametitle{MSL to Component-based runtime}
\begin{block}{Ready-to-fill parallel pattern}
\begin{itemize}
\item Data parallelism
\begin{itemize}
\item External distributed data structure
\item Automatic detection of synchronizations
\end{itemize}
\item Task parallelism (mid-grain)
\begin{itemize}
\item Compile a static scheduling of computation kernels
\end{itemize}
\end{itemize}
\end{block}
The fine grain task parallelism is left to other languages: 
\begin{itemize}
\item OpenMP in the kernels
\item Kernels generated by stencil compilers (Pochoir, PATUS, Liszt etc.)
\end{itemize}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{MSL to Component-based runtime}
\begin{center}
$\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$\\

\bigskip
\begin{tikzpicture}[remember picture]
  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
    \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
     \node[component,solid,draw=red] (D) at (0,0) {$Driver$};
     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
     \node[solid] (Ds) at (-1.5,0) {start};
     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};

     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
     \node[solid,left=0.8cm of T] (tt) {$T$};

     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
     \node[component,dashed,right=2cm of Cp] (C) {$Computations$};
     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
     \node[solid,right=0.2cm of Cu] (star) {$*$};
     \node[solid,right=1.5cm of C] (gamma) {$\Gamma$};

     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
     \node[solid,left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};

     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
     \node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};

    \path[-]
      (Dp) edge [solid] node {} (D)
      (D) edge [solid] node {} (Du1)
          edge [solid] node {} (Du2)
          edge [solid] node {} (Du3)
      (DDSp1) edge [solid] node {} (DDS)
      (Tp) edge [solid] node {} (T)
      (T)  edge [solid] node {} (Tu)
      (Cp) edge [solid] node {} (C)
      (C) edge [solid] node {} (Cu)
      (Datap2) edge [solid] node {} (Data)
      (Datap1) edge [solid] node {} (Data)
      (Data) edge [solid] node {} (Datau)
      (DDSp2) edge [solid] node {} (DDS);
    \end{tikzpicture}
    };
    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=0.5cm of proc] (mpidata) {};
    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
    \node[solid,below=2.5cm of proc.west,anchor=west] (core) {\textit{\tiny{Duplicated on each processor/core}}};
    \path[-]
      (Data.east) edge node {} (mpidata);
  \end{tikzpicture}
\end{center}
\end{frame}
%-------------------------------------------------------------

%----------------
\begin{frame}
\frametitle{MSL to Component-based runtime}
\begin{center}
$\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$\\

\bigskip
\begin{tikzpicture}[remember picture]
  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
    \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
     \node[component,solid] (D) at (0,0) {$Driver$};
     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
     \node[solid] (Ds) at (-1.5,0) {start};
     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};

     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
     \node[solid,left=0.8cm of T] (tt) {$T$};

     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
     \node[component,dashed,right=2cm of Cp] (C) {$Computations$};
     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
     \node[solid,right=0.2cm of Cu] (star) {$*$};
     \node[solid,right=1.5cm of C] (gamma) {$\Gamma$};

     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
     \node[component,solid,draw=red,double,above=0.8cm of Datap2] (Data) {$Data$};
     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
     \node[solid,left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};

     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
     \node[component,draw=red,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
     \node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};

    \path[-]
      (Dp) edge [solid] node {} (D)
      (D) edge [solid] node {} (Du1)
          edge [solid] node {} (Du2)
          edge [solid] node {} (Du3)
      (DDSp1) edge [solid] node {} (DDS)
      (Tp) edge [solid] node {} (T)
      (T)  edge [solid] node {} (Tu)
      (Cp) edge [solid] node {} (C)
      (C) edge [solid] node {} (Cu)
      (Datap2) edge [solid] node {} (Data)
      (Datap1) edge [solid] node {} (Data)
      (Data) edge [solid] node {} (Datau)
      (DDSp2) edge [solid] node {} (DDS);
    \end{tikzpicture}
    };
    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=0.5cm of proc] (mpidata) {};
    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
    \node[solid,below=2.5cm of proc.west,anchor=west] (core) {\textit{\tiny{Duplicated on each processor/core}}};
    \path[-]
      (Data.east) edge node {} (mpidata);
  \end{tikzpicture}
\end{center}
\end{frame}
%-------------------------------------------------------------

%----------------
\begin{frame}
\frametitle{MSL to Component-based runtime}
\begin{center}
$\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$\\

\bigskip
\begin{tikzpicture}[remember picture]
  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
    \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
     \node[component,solid] (D) at (0,0) {$Driver$};
     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
     \node[solid] (Ds) at (-1.5,0) {start};
     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};

     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
     \node[component,solid,draw=red,below=1.6cm of Tp] (T) {$Time$};
     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
     \node[solid,left=0.8cm of T] (tt) {$T$};

     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
     \node[component,dashed,draw=red,right=2cm of Cp] (C) {$Computations$};
     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
     \node[solid,right=0.2cm of Cu] (star) {$*$};
     \node[solid,right=1.5cm of C] (gamma) {$\Gamma$};

     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
     \node[solid,left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};

     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
     \node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};

    \path[-]
      (Dp) edge [solid] node {} (D)
      (D) edge [solid] node {} (Du1)
          edge [solid] node {} (Du2)
          edge [solid] node {} (Du3)
      (DDSp1) edge [solid] node {} (DDS)
      (Tp) edge [solid] node {} (T)
      (T)  edge [solid] node {} (Tu)
      (Cp) edge [solid] node {} (C)
      (C) edge [solid] node {} (Cu)
      (Datap2) edge [solid] node {} (Data)
      (Datap1) edge [solid] node {} (Data)
      (Data) edge [solid] node {} (Datau)
      (DDSp2) edge [solid] node {} (DDS);
    \end{tikzpicture}
    };
    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=0.5cm of proc] (mpidata) {};
    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
    \node[solid,below=2.5cm of proc.west,anchor=west] (core) {\textit{\tiny{Duplicated on each processor/core}}};
    \path[-]
      (Data.east) edge node {} (mpidata);
  \end{tikzpicture}
\end{center}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Compiler}
% A slide example MS, 6 slides on graph transformations
%-------------------------------------------------------------
\subsection{Compiler}
%----------------
\begin{frame}[fragile]
\frametitle{Example}
\begin{columns}
\begin{column}{.52\textwidth}
\begin{lstlisting}[style=base]
@mesh:@ cart
@mesh entities:@ cell,edgex,edgey
@computation domains:@
  allcell in cell
  alledgex in edgex
  alledgey in edgey
  part1edgex in edgex
  part2edgex in edgex
@data:@
  a,cell
  b,cell
  c,edgex
  d,edgex
  e,edgey
\end{lstlisting}
\end{column}
\begin{column}{.48\textwidth}
\begin{lstlisting}[style=base]
  f,cell
  g,edgey
  h,edgex
  i,cell
  j,edgex
@time:@500
@computations:@
  b[allcell]=&c0&(a)
  c[alledgex]=&c1&(b[n1])
  d[alledgex]=&c2&(c)
  e[alledgey]=&c3&(c)
  f[allcell]=&c4&(d[n1])
  g[alledgey]=&c5&(e)
  h[alledgex]=&c6&(f)
  i[allcell]=&c7&(g,h)
  j[partedgex]=&c8&(i[n1])
\end{lstlisting}
\end{column}
\end{columns}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Data parallelism}
$\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$\\
$\hookrightarrow [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$

\medskip
\begin{block}{Synchronization}
For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronization} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j$ read a data written by $c_i$ with a stencil shape and not locally.
\end{block}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Mid-grain task parallelism}

Build the read/write dependency graph.

\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node (c0) at (0,0) {$c_0$};
   \node[right=1 of c0] (sy1) {$sync_1$};
   \node[right=1 of sy1] (c1) {$c_1$};
   \node[above right=1 of c1] (c2) {$c_2$};
   \node[below right=1 of c1] (c3) {$c_3$};
   \node[right=1 of c2] (sy4) {$sync_4$};
   \node[right=1 of c3] (c5) {$c_5$};
   \node[right=1 of sy4] (c4) {$c_4$};
   \node[right=1 of c4] (c6) {$c_6$};
   \node[below right=1 of c6] (c7) {$c_7$};
   \node[right=1 of c7] (sy8) {$sync_8$};
   \node[right=1 of sy8] (c8) {$c_8$};

  \path[->]
    (c0) edge node {} (sy1)
    (sy1) edge node {} (c1)
    (c1)  edge node {} (c2)
          edge node {} (c3)
    (c2) edge node {} (sy4)
    (sy4) edge node {} (c4)
    (c4) edge node {} (c6)
    (c3) edge node {} (c5)
    (c5) edge node {} (c7)
    (c6) edge node {} (c7)
    (c7) edge node {} (sy8)
    (sy8) edge node {} (c8);
\end{tikzpicture}
\end{center}

\begin{center}
\textbf{Data parallelism + Task parallelism}
\end{center}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Which scheduling ?}
From the dependency graph we can:
\begin{itemize}
\item Use a dynamic scheduler
\item Compute a static scheduling
\end{itemize}

In this work,
\begin{itemize}
\item We build a static scheduling of the multi-stencil application
\item This static scheduling can be dumped to a component-based runtime
\end{itemize}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Series-Parallel Tree}
\begin{columns}
\begin{column}{.48\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

   \node[] (c0)    at (-3,1) {$c_0$};
   \node[] (star1) at (-2,1) {$sync_1$};
   \node[] (c1)    at (-1,1) {$c_1$};
   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
   \node[] (c7)    at ( 1,1) {$c_7$};
   \node[] (star8) at ( 2,1) {$sync_8$};
   \node[] (c8)    at ( 3,1) {$c_8$};

   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
   \node[] (s2)    at (1,2) {$\mathcal{S}$};

   \node[] (c2)    at (-3,3) {$c_2$};
   \node[] (star4) at (-2,3) {$sync_4$};
   \node[] (c4)    at (-1,3) {$c_4$};
   \node[] (c6)    at ( 0,3) {$c_6$};
   \node[] (c3)    at ( 1,3) {$c_3$};
   \node[] (c5)    at ( 2,3) {$c_5$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3)
         edge node {} (c5);
  \end{tikzpicture}
  }
\end{center}
\end{column}
\begin{column}{.52\textwidth}
\begin{block}{Specific components}
\begin{itemize}
\item $SEQ$ to directly replace $\mathcal{S}$ nodes
\item $PAR$ to directly replace $\mathcal{P}$ nodes
\item $SYNC$ for synchronizations
\item $K$ for computation kernels
\end{itemize}
\end{block}
\end{column}
\end{columns}
\begin{center}
\textbf{Direct dump to the component assembly}
\end{center}
\end{frame}
%----------------

% %----------------
% \begin{frame}
% \frametitle{Component-based runtime}
% \begin{block}{Specific components}
% \begin{itemize}
% \item $SEQ$ to directly replace $\mathcal{S}$ nodes
% \item $PAR$ to directly replace $\mathcal{P}$ nodes
% \item $SYNC$ for synchronizations
% \item $K$ for computation kernels
% \end{itemize}
% \end{block}

% \begin{center}
% \textbf{Direct dump to the component assembly}\\
% \textbf{Parallel component-based skeleton of the multi-stencil application}
% \end{center}

% \end{frame}
% %----------------

%----------------
\begin{frame}
\frametitle{Component-based runtime}

\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
     \node[component,solid] (D) at (0,0) {$Driver$};
     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
     %\node[solid] (Ds) at (-1.5,0) {start};
     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};

     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
     %\node[solid,left=0.8cm of T] (tt) {$T$};

     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
     \node[component,dashed,right=2cm of Cp] (C) {$Computations$};
     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
     \node[solid,right=0.2cm of Cu] (star) {$*$};
     %\node[solid,right=1.5cm of C] (gamma) {$\Gamma$};

     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
     %\node[solid,left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};

     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
     %\node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};

    \path[-]
      (Dp) edge [solid] node {} (D)
      (D) edge [solid] node {} (Du1)
          edge [solid] node {} (Du2)
          edge [solid] node {} (Du3)
      (DDSp1) edge [solid] node {} (DDS)
      (Tp) edge [solid] node {} (T)
      (T)  edge [solid] node {} (Tu)
      (Cp) edge [solid] node {} (C)
      (C) edge [solid] node {} (Cu)
      (Datap2) edge [solid] node {} (Data)
      (Datap1) edge [solid] node {} (Data)
      (Data) edge [solid] node {} (Datau)
      (DDSp2) edge [solid] node {} (DDS);
    %\end{tikzpicture}
    %}
% \end{column}
% \begin{column}{.70\textwidth}
% \begin{center}
% \resizebox{.80\textwidth}{!}{%
% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   %seq0
   \node[component,right=3.5cm of C] (seq0) {$SEQ$};
   \node[provide] (seq0p) [left = 0.8cm of seq0] {};
   \node[use] (seq0u) [right = 1cm of seq0] {$m$};
   %c0
   \node[component] (c0) [right = 2cm of seq0u] {$K(c_0)$};
   \node[use] (c0u) [right = 0.8cm of c0] {};
   \node[right=0.2 of c0u] (star) {*};
   %sync0
   \node[component] (sync0) [below = 1cm of c0] {$SYNC$};
   \node[use] (sync0u) [right = 1cm of sync0] {$m$};
   %c1
   \node[component] (c1) [below = 0.8cm of sync0] {$K(c_1)$};
   \node[use] (c1u) [right = 0.8cm of c1] {};
   \node[right=0.2 of c1u] (star) {*};
   %par0
   \node[component] (par0) [below = 0.8cm of c1] {$PAR$};
   \node[use] (par0u) [right = 1cm of par0] {$m$};
  %seq1
   \node[component] (seq1) [right = 1cm of sync0u] {$SEQ$};
   \node[use] (seq1u) [right = 1cm of seq1] {$m$};
  %c2
   \node[component] (c2) [right = 4cm of c0] {$K(c_2)$};
   \node[use] (c2u) [right = 0.8cm of c2] {};
   \node[right=0.2 of c2u] (star) {*};
  %sync1
   \node[component] (sync1) [below = 0.8cm of c2] {$SYNC$};
   \node[use] (sync1u) [right = 1cm of sync1] {$m$};
  %c4
   \node[component] (c4) [below = 0.8cm of sync1] {$K(c_4)$};
   \node[use] (c4u) [right = 0.8cm of c4] {};
   \node[right=0.2 of c4u] (star) {*};
  %c6
   \node[component] (c6) [below = 0.8cm of c4] {$K(c_6)$};
   \node[use] (c6u) [right = 0.8cm of c6] {};
   \node[right=0.2 of c6u] (star) {*};
  %seq2
   \node[component] (seq2) [below = 3.2cm of seq1] {$SEQ$};
   \node[use] (seq2u) [right = 1cm of seq2] {$m$};
  %c3
   \node[component] (c3) [right = 1cm of seq2u] {$K(c_3)$};
   \node[use] (c3u) [right = 0.8cm of c3] {};
   \node[right=0.2 of c3u] (star) {*};
  %c5
   \node[component] (c5) [below = 0.8cm of c3] {$K(c_5)$};
   \node[use] (c5u) [right = 0.8cm of c5] {};
   \node[right=0.2 of c5u] (star) {*};
  %c7
   \node[component] (c7) [below = 0.8cm of par0] {$K(c_7)$};
   \node[use] (c7u) [right = 0.8cm of c7] {};
   \node[right=0.2 of c7u] (star) {*};
  %sync2
   \node[component] (sync2) [below = 0.8cm of c7] {$SYNC$};
   \node[use] (sync2u) [right = 1cm of sync2] {$m$};
  %c8
   \node[component] (c8) [below = 0.8cm of sync2] {$K(c_8)$};
   \node[use] (c8u) [right = 0.8cm of c8] {};
   \node[right=0.2 of c8u] (star) {*};

   \path[-]
   %seq0
    (seq0) edge node {} (seq0u)
    (seq0p) edge node {} (seq0)
   %c0
    (seq0u) edge node {} (c0)
    (c0) edge node {} (c0u)
    %sync0
    (seq0u) edge node {} (sync0.west)
    (sync0) edge node {} (sync0u)
    %c1
    (seq0u) edge node {} (c1)
    (c1) edge node {} (c1u)
   %par0
    (seq0u) edge node {} (par0.west)
    (par0) edge node {} (par0u)
  %seq1
    (par0u) edge node {} (seq1.west)
    (seq1) edge node {} (seq1u)
  %c2
    (seq1u) edge node {} (c2.west)
    (c2) edge node {} (c2u)
  %sync0
    (seq1u) edge node {} (sync1.west)
    (sync1) edge node {} (sync1u)
  %c4
    (seq1u) edge node {} (c4.west)
    (c4) edge node {} (c4u)
  %c6
    (seq1u) edge node {} (c6.west)
    (c6) edge node {} (c6u)
  %seq2
    (par0u) edge node {} (seq2.west)
    (seq2) edge node {} (seq2u)
  %c3
    (seq2u) edge node {} (c3)
    (c3) edge node {} (c3u)
  %c6
    (seq2u) edge node {} (c5.west)
    (c5) edge node {} (c5u)
  %c6
    (seq0u) edge node {} (c7.west)
    (c7) edge node {} (c7u)
  %sync2
    (seq0u) edge node {} (sync2.west)
    (sync2) edge node {} (sync2u)
  %c8
    (seq0u) edge [bend right] node {} (c8.west)
    (c8) edge node {} (c8u)
   ;

   \path[-]
    (C) edge[dashed] node {} (seq0);
\end{tikzpicture}
}
\end{center}
% \end{column}
% \end{columns}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Back to Data Parallelism}
\begin{center}
\textbf{Back to Data Parallelism only ?}\\
\end{center}
\textbf{Loop Fusion}: Same computation domain under a sequence or a parallel node

\medskip
\begin{center}
\resizebox{0.5\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

   \node[] (c0)    at (-3,1) {$c_0$};
   \node[] (star1) at (-2,1) {$sync_1$};
   \node[] (c1)    at (-1,1) {$c_1$};
   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
   \node[] (c7)    at ( 1,1) {$c_7$};
   \node[] (star8) at ( 2,1) {$sync_8$};
   \node[] (c8)    at ( 3,1) {$c_8$};

   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
   \node[] (s2)    at (1,2) {$\mathcal{S}$};

   \node[] (c2)    at (-3,3) {$c_2$};
   \node[] (star4) at (-2,3) {$sync_4$};
   \node[] (c4)    at (-1,3) {$c_4$};
   \node[] (c6)    at ( 0,3) {$c_6$};
   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
   %\node[] (c5)    at ( 2,3) {$c_5$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3);
  \end{tikzpicture}
  }
\end{center}

\small
\textbf{Canonical form}: $[c_0,sync_1,c_1,c_2,sync_4,c_4,c_6,\{c_3,c_5\},c_7,sync_8,c_8]$
\end{frame}
%----------------

%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Evaluation}
% Present Shallow-water case, present results
%-------------------------------------------------------------
\subsection{Evaluation}
\begin{frame}
\frametitle{Implementation and evaluation}
\textbf{Implementation of MSL}: Python, SkelGIS and $L^2C$

\medskip
\textbf{Shallow-water equations}: 1 mesh, 3 mesh entities, 7 computation domains, 48 data, 98 computations (32 stencils, 66 local kernels)

\medskip
\textbf{Evaluation of the data parallelism}
\begin{itemize}
\item Full SkelGIS implementation (DDS + specific interfaces to hide communications)
\item MSL implementation which uses the SkelGIS DDS
\item Cluster Edel of Grid'5000: 2 8-cores Intel Xeon E5520, 24GB RAM / nodes
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Evaluations}
\begin{columns}
\begin{column}{.5\textwidth}
\begin{center}
\small Mesh $10k \times 10k$, $500$ iterations\\
\resizebox{\textwidth}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
\end{center}
\end{column}
\begin{column}{.5\textwidth}
\begin{center}
\small Mesh $10k \times 10k$, $2K$ iterations\\
\resizebox{\textwidth}{!}{\includegraphics{./images/logtimes_10K_2K.pdf}}
\end{center}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Conclusion and perspectives}
%-------------------------------------------------------------
\subsection{Conclusion and perspectives}
\begin{frame}
\frametitle{Conclusion}
\begin{itemize}
\item A DSL for Multi-Stencil applications (MSL)
\item The compilation of MSL to get a static scheduling
\item The dump to a component-based runtime
\item No overhead introduced by the solution
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Perspectives}
\begin{itemize}
\item A comparison with an improved version of SkelGIS
\item More evaluations on TGCC Curie (CEA) up to $32k$ cores
\item An evaluation of the Data+Task parallelism (mixing the static scheduling and OpenMP)
\item A new dump with a component responsible for the dynamic scheduling of the application (using OpenMP 4.x)
\item A new dump to CPU+GPGPUs (using existing stencil compilers Pochoir, PATUS etc.)
\item An alternative to the SkelGIS DDS using Global Arrays (PGAS)
\end{itemize}
\end{frame}
%-------------------------------------------------------------

\begin{withoutheadline}
\begin{frame}{}
\begin{center}
\huge Thank You !
\end{center}
\end{frame}
\end{withoutheadline}

\end{document}

### 2016-01-18 rapport/formalism.tex
maj formalism Helene 18/01/2016

@@
 Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied onto a mesh (space discretization).
-In other words, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain.
+In other words, the PDEs are transformed to a set of numerical computations applied at each time step on elements of the discretized space domain.
 Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}, and a set of auxiliary computations also needed to perform the simulation, and also called \emph{local computations}.
-This section gives formal definitions of a \textit{stencil program} and its computations. Then, the different parallelization techniques which can be applied on such program, are presented.
+This section gives formal definitions of what we call a \textit{stencil program} and its computations.
+Then, the different mid-grain parallelization techniques which can be applied on such program, are presented.
 
@@
 
-$\Omega=\mathbb{R}^n$ is the continuous space domain of a numerical simulation. Any geo-referenced point in this space can be represented.
-
+$\Omega=\mathbb{R}^n$ is the continuous space domain of a numerical simulation.
+%For example, if $n=3$, any geographic point on earth could be represented in this space in a given geographic coordinate system.
 A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ of a set of PDEs and is defined as follows.
@@
 \textit{A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V\subset \Omega$ is the set of vertices and $E\subseteq V^2$ the set of edges.
-The set of edges $E$ of a mesh $\mathcal{M}=(V,E)$ does not contain bridges.}
+The set of edges $E$ of a mesh $\mathcal{M}=(V,E)$ does not contain bridges.
+It is said that the mesh is applied onto $\Omega$}.
 \end{mydef}
@@
 \end{center}\end{figure}
+\begin{mydef}
+\textit{The dimension of a mesh $\mathcal{M}=(V,E)$ applied onto $\Omega=\mathbb{R}^n$ is denoted $dim(\mathcal{M})=n$.}
+\end{mydef}
 A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) and hybrid as illustrated in Figure~\ref{fig:mesh}.
+
+\medskip
+\noindent \textbf{Definitions}
+%\begin{mydef}
+\begin{itemize}
+\item \textit{An entity $e$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $e\subset (V\cup E)$.
+\item A mesh entities group is denoted $E$ such that $E \in \mathcal{P}(V\cup E)$. It represents a set of entities of the same type.
+\item Finally the set of mesh entities groups of a simulation is denoted $\mathcal{E}$.}
+\end{itemize}
+%\end{mydef}
+
+For example, in a 2D Cartesian mesh an entity could be a cell, made of four vertices and four edges, or simply a vertex.
+As a result a group of cells and a group of vertices could be defined in $\mathcal{E}$.
+%This type can be defined as the sets containing exactly four vertices and four edges connected as a cycle.
+Another type of entities simply are the vertices ($V$) and can be defined as all singletons formed of a single vertex of $V$.
+
+\medskip
+This covers the space discretization, however there is also the time dimension which has to be discretized in the simulation.
+
+\medskip
+\noindent \textbf{Definitions}
+%\begin{mydef}
+\begin{itemize}
+\item \textit{We denote a scalar as an identifier associated to a numerical value and applied onto a mesh $\mathcal{M}$ where $dim(\mathcal{M})=0$.
+In other words a scalar can be seen as a variable containing a numerical value.
+The set of scalars is denoted $\mathcal{S}$.}
+\item \textit{$\mathcal{T}=\mathbb{R}$ is the continuous time domain of a numerical simulation.}
+\item \textit{The discretization of the continuous time domain $\mathcal{T}$ is denoted as the pair $T(\Delta t,conv)$.} 
+\begin{itemize}
+\item \textit{$\Delta t \in \mathcal{S}$ represents the time interval of the numerical simulation, such that for a current time iteration $t_i$, the next time iteration is $t_{i+1} = t_i + \Delta t$.
+The default value of $\Delta t$ is $1$.
+\item $conv:\mathcal{S}^n \rightarrow bool$ is a function which returns a boolean from a set of scalar variables. This function represents the convergence criteria of the simulation. 
+\item At a given time step, the convergence criteria is evaluated such that if $conv$ returns $true$ the next time step can start such that $t_{i+1} = t_i + \Delta t$, otherwise the simulation ends.
+For example, the simplest $conv$ is typically to have two scalars: $t$ the time, and a fix number of iterations denoted $it$.
+At each time step $\Delta t=1$, and $conv$ returns the boolean expression $t<it$.}
+\end{itemize}
+\end{itemize}
+%\end{mydef}
+
+%$T$ is responsible for the iteration time steps of the numerical simulation.
+A numerical simulation though does not run forever and must be stopped at some point.
+Some simulations choose a fix number of time iteration, others compute a \emph{convergence} criteria at the end of each time iteration to determine if the simulation has to continue.
+A convergence criteria is computed by a reduction computation that will be detailed in the next section.
+
+In a numerical simulation, as a set of scalars can be applied onto a mesh with a nil dimension, a set of data elements, or quantities, can be applied onto meshes of dimension superior or equal to zero.
+Those quantities represent, as well as scalars, the set of values to compute, or to use, for computations.
 
@@
 \begin{itemize}
-\item An entity $e$ of a mesh $\mathcal{M}=(V,E)$ is a subset of its vertices and edges, $e\subset (V\cup E)$. 
-\item The set of mesh entities, of a same given type, is denoted $E$ such that $E\subset \mathcal{P}(V\cup E)$. 
-\item Finally the set of mesh entities of a simulation is denoted $\mathcal{E}$.
+\item \textit{A quantity is a function $\delta: E_{\delta} \mapsto V_{\delta}$ which associates each entity of a group $e\in E_{\delta}$ to a value $v\in V_{\delta}$ where $V_{\delta}$ is a data type, typically $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.
+\item The set of quantities applied onto the mesh is denoted $\Delta$.
+\item In the rest of this paper, the group of mesh entities on which a quantity $\delta$ is mapped is denoted $entity(\delta)=E_{\delta}$.}
 \end{itemize}
@@
 
-For example, in a 2D Cartesian mesh a first kind of entities are the cells, they can be defined as all sets containing exactly four vertices and four edges connected as a cycle.
-Another kind of entities simply are the vertices ($V$) and can be defined as all singletons formed of a single vertex of $V$.
-
-\medskip
-This covers the space discretization, however another dimension manipulated in simulations that has to be discretized is the time dimension.
-
-\noindent \textbf{Definitions}
-%\begin{mydef}
-\begin{itemize}
-\item $\mathcal{T}=\mathbb{R}$ is the continuous time domain of a numerical simulation.
-\item The discretization of the continuous time domain $\mathcal{T}$ is denoted $T$ such that $\forall\mbox{ }t_i\mbox{, }t_{i+1} \in T\mbox{, }\exists\mbox{ }\Delta t \in \mathbb{R}$\mbox{, }$t_{i+1} = t_i + \Delta t$.
-\end{itemize}
-%\end{mydef}
-
-$T$ is responsible for the iteration time steps of the numerical simulation.
-A numerical simulation though does not run forever and must be stopped at some point.
-Some simulations choose a fix number of time iteration, others compute a \emph{convergence} criteria at the end of each time iteration to determine if the simulation has to continue.
-A convergence criteria is computed by a reduction computation that will be detailed in the next section.
-
-In a numerical simulation a set of data elements, or quantities, are applied onto the mesh and represent the set of values to compute, or to use, for computation.
-
-\noindent \textbf{Definitions}
-%\begin{mydef}
-\begin{itemize}
-\item A quantity is a function $\delta: E_{\delta} \mapsto V_{\delta}$ which associates each entity of a type $e\in E_{\delta}$ to a value $v\in V_{\delta}$ where $V_{\delta}$ is the data type, typically $\mathbb{R}$ or $\mathbb{C}$.
-\item The set of quantities applied onto the mesh is denoted $\Delta$.
-\item In the rest of this paper, the type of mesh entities on which a quantity $\delta$ is mapped is denoted $entity(\delta)=E_{\delta}$.
-\end{itemize}
-%\end{mydef}
-
-Another approach closer to that taken in applied mathematics would have been to define $\delta$ functions over $E_{\delta} \times T$; however the approach chosen in this paper where a distinct function is defined for each time-step better matches the implementation of numerical simulations where the outer loop iterates over time and where only a few time-steps of quantity values are stored.
+Another option, closer to the applied mathematics domain, would have been to define $\delta$ as the function over $E_{\delta} \times T$; however the approach chosen for now in this work is to let the user be aware of the number of data he is using and what exactly for.
 
@@
 \begin{itemize}
-\item A computation domain $D$ is a subpart of mesh entities of a given type, $D \subset E \in \mathcal{E}$.
+\item \textit{A computation domain $D$ is a subpart  a mesh entities group, $D \subset E \in \mathcal{E}$.
 \item The set of computation domains of a numerical simulation is denoted $\mathcal{D}$.
-\item A neighborhood $n$ is a function which for a given entity $e \in E_i$, returns a set of $m$ entities in $E_j$, $n : E_i \rightarrow E_j^m$. It is possible that $E_i = E_j$.
+\item A neighborhood $n$ is a function which for a given entity $e \in E_i$, returns a set of $m$ entities in $E_j$, $n : E_i \rightarrow E_j^m$. One can notice that $i = j$ is possible.
 \item The set of neihborhood functions in a numerical simulation is denoted $\mathcal{N}$.
-\item A set of data to read and used in a computation is denoted $R \subset \Delta \times \mathcal{N}$.
-It is the set of pairs $(r,n)$, where $r \in \Delta$ and $n$ is a neighborhood function such that $n : E_i \rightarrow entity(r)^m$.
-\item A numerical expression $\text{exp}_{w}: entity(w) \times R \rightarrow V_{w}$ is a function that computes the value of the written quantity $w \in \Delta$ for a given entity $e \in entity(w)=E$, using a set of input data $R$.
-\item A computation kernel $k$ of a numerical simulation is defined as $k(R,w,\text{exp},D)$, where $R \subset \Delta \times \mathcal{N}$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$.
+\item A single data is written during a computation and is denoted $w \in \Delta$. Thus, the computation domain of the computation is $D \subset E_w=entity(w)$.
+\item A set of data to read and use in a computation is denoted $R$.
+Each element of $R$ is a tuple $(r,n)$, where $r \in \Delta$ and $n$ is a neighborhood function such that $n : E_w \rightarrow entity(r)^m$.
+The neighborhood indicates which entities, on which $r$ is applied, will be read during the computation.
+\item A numerical expression $\text{exp}_{w}: entity(w) \times R^n \rightarrow V_{w}$ is a function that computes the value of the written quantity $w \in \Delta$ for a given entity $e \in entity(w)=E_w$, using a set of input data $R$.
+\item A computation kernel $k$ of a numerical simulation is defined as $k(R,w,\text{exp},D)$, where $R$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$ is the computation domain.}
 \end{itemize}
@@
 \begin{itemize}
-\item We denote by $identity$ the identity function $x=x$.
+\item \textit{We denote by $identity$ the identity function $x=x$.
 \item A kernel computation $k(R,w,\text{exp},D) \in \Gamma$ is a stencil kernel $\iff \exists (r,n) \in R$ such that $n \neq identity$.
-\item A kernel computation $k(R,w,\text{exp},D) \in \Gamma$ is a local kernel $\iff \forall (r,n) \in R$, $n = identity$.
-\item The set of $n$ ordered computations of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i \leq j$, then $k_i$ is computed before $k_j$.
+\item A kernel computation $k(R,w,\text{exp},D) \in \Gamma$ is a local kernel $\iff \forall (r,n) \in R$, $n = identity$.}
 \end{itemize}
 
-A reduction computation is typically used to compute the convergence criteria of the time loop of the simulation.
+\noindent \textbf{Property}
+\textit{A kernel for which all data read and written are applied onto a mesh of dimension $0$ is a local kernel.}
+
+A reduction computation is a computation which takes data applied onto meshes and returns a data with a dimension reduced to zero, in other words a scalar.
+A reduction is typically used to compute the convergence criteria of the time loop of the simulation.
 Occasionally reductions can also be performed during a time iteration.
-This last usage is typically done to place a conditionnal branch to chose one computation or another.
-This paper only study the use of reduction computations as convergence citerion.
+%This last usage is typically done to have a conditionnal branch to choose one computation or another, or to compute a dynamic time step for example.
 
+\medskip
 \noindent \textbf{Definitions}
@@
 \begin{itemize}
-\item A reduction computation is defined as the triplet $red(R,)$
-\item ...
-\item A \textit{multi-stencil program} is defined by the sextuplet $\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$.
+\item \textit{A reduction computation $r$ of a numerical simulation is defined as $r(R,w,red)$, where $R$ is the set of data read, $w \in \mathcal{S}$ is the scalar written, and $\text{red}$ is an associative reduction function.
+% \item For any $i,j$ such that $r_i \in R$ and $r_j \in R$, if $dim(mesh(r_i))=dim(mesh(r_j))=n$.
+% \item The data written by the reduction $w$ verifies $dim(mesh(w))<n$.
+\item A reduction function $\text{red}: \Delta^n \rightarrow \mathcal{S}$ is an operation which reduces a set of data in $\Delta$ (on it sentities) to a single scalar variable in $\mathcal{S}$.
+The reduction function must be associative.
+\item The set of $n$ ordered computations (kernel or reduction) of a numerical simulation is denoted $\Gamma = [c_i]_{0 \leq i \leq n-1}$, such that $\forall c_i,c_j \in \Gamma$, if $i \leq j$, then $c_i$ is computed before $c_j$.
+\item Finally, a \textit{multi-stencil program} is defined by the septuplet $\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta, \mathcal{S},\Gamma)$.}
 \end{itemize}
+%!!!!!!!  est-ce qu'on fusionne le maillage, les entites et les domaines ?????
 %\end{mydef}
 
-For example, in Figure~\ref{fig:ex1}, assuming the computation domain (full lines) is denoted $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined as:
+For example, in Figure~\ref{fig:ex1}, assuming that the computation domain (full lines) is denoted $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined as:
 \begin{equation*}
@@
 \begin{center}
-\subfloat[Mesh and mesh domains.\label{fig:mesh}]{
+\subfloat[Mesh and mesh domains.\label{fig:meshbase}]{
 \resizebox{8cm}{!}{\includegraphics{./images/mesh.pdf}}
@@
 
-A stencil program and stencil and local computations have been formally defined in this section.
-This formalism is used in the next Section to define two parallelization techniques of a multi-stencil program.
+A stencil program has been formally defined in this section. This formalism is used in the next Section to define two mid-grain parallelization techniques of a multi-stencil program.

### 2016-01-20 rapport/formalism.tex
maj formalism+parallelism Helene 20/01/2016

@@
 \item A numerical expression $\text{exp}_{w}: entity(w) \times R^n \rightarrow V_{w}$ is a function that computes the value of the written quantity $w \in \Delta$ for a given entity $e \in entity(w)=E_w$, using a set of input data $R$.
-\item A computation kernel $k$ of a numerical simulation is defined as $k(R,w,\text{exp},D)$, where $R$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$ is the computation domain.}
+\item A computation kernel $k$ of a numerical simulation is defined as $c(R,w,\text{exp},D)$, where $R$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$ is the computation domain.}
 \end{itemize}

### 2016-01-20 rapport/parallelism.tex
maj formalism+parallelism Helene 20/01/2016

@@
 \label{sect:dataparal}
-In a data parallelization technique, the idea is to split the data on which the program is computed in balanced sub-parts, one for each available resource.
+In a data parallelization technique, the idea is to split the data on which the program is computed into balanced sub-parts, one for each available resource.
 The same sequential program can afterwards be applied on each sub-part simultaneously, with some additioinal synchronizations between resources to update the data not computed locally, and thus to guarantee a correct result.
 
-More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
+More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta, \mathcal{S},\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
 This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
 As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\Delta=\{\Delta_0,\dots,\Delta_{p-1}\}$.
-The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma_{data}$.
+The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma_{sync}$.
 
 \begin{mydef}
-For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j=s_j(R_j,w_j,\text{exp}_j)$, and $\{w_i\} \subset R_j$.
-Moreover, the data to update is $\{w_i\} \cap R_j = w_i$.
+\textit{
+For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $\exists (r,n) \in R_j$ with $n\neq identity$ ($c_j$ is a stencil computation), and such that $\{w_i\} \subset R_j$.
+Moreover, the data to update is $\{w_i\} \cap R_j = \{w_i$\}.}
 \end{mydef}
@@
 Actually, a synchronization can only be needed by the data read by a stencil computation (not local), and only if this data has been modified before, which means that it has been written before.
-This synchronization is needed because the neighborhood function $\mathcal{N}$ of a stencil computation involves values computed on different resources.
+This synchronization is needed because a neighborhood function $n \in \mathcal{N}$ of a stencil computation involves values computed on different resources.
 
 \begin{mydef}
-A synchronization between two computations $c_i \pprec c_j$ is defined as a specific computation $\text{update}(\{w_i\} \cap R_j)=\text{update}(w_i)$.
+\textit{A synchronization between two computations $c_i \pprec c_j$ is defined as a specific computation $sync(R,w,exp,D)$ where $R=\{w_i\}$, $w=w_i$, $exp=identity$ and $D=\emptyset$.}
 \end{mydef}
@@
 \begin{mydef}
-An \textit{updated-computation} is a computation which needs a set of data synchronizations before it can be performed.
+If a synchronization is needed before $c_j$ (on the data $w_i$), $c_j$ is replaced in $\Gamma_{sync}$ by the list
+\begin{equation*}
+\Gamma_{sync} = [sync(\{w_i\},w_i,identity,\emptyset), c_j]
+\end{equation*}
 \end{mydef}
-
-\begin{mydef}
-If $c_j$ is an \textit{updated-computation}, it is replaced in $\Gamma_{data}$ by the ordered list of two elements
-\begin{equation*}
-[*_j(R^*_j,w^*_j),c^*_j(R_j \cup \{w^*_j\},w_j,\text{exp}_j)].
-\end{equation*}
-The computation $*_j$ is defined as the pair $*_j(R^*_j,w^*_j)$, where $R^*_j = \{\bigcup_{i} \{w_i\} \text{, }\forall i<j\text{, and }c_i \pprec c_j\}$, and $w^*_j$ is a dummy data which does not represent a quantity of the $\mathcal{MSP}$.
-This dummy $w^*_j$ is used, however, to keep a relation between the update phase $*_j$ and the computation $c_j$.
-For this reason, $c_j$ is replaced by $c^*_j$ where data read are modified to $R_j \cup \{w^*_j\}$.
-Finally, $*_j$ performs the computation $update(R^*_j)$ and $R^*_j \subseteq R_j \subseteq \Delta$.
-\end{mydef}
-
- %$c^*_j(R_j,w_j,\text{exp}_j,*_j)$ if $\bigcup_{i} \{w_i\} \neq \emptyset \text{, }\forall i<j\text{, and }c_i \pprec c_j$.
-In such a case $*_j$ is itself a computation wich performs $update(\bigcup_{i} \{w_i\})$.
 
@@
 \begin{mydef}
-From the ordered list of computation $\Gamma$, a new ordered list $\Gamma_{data}$ is obtained from the call $\Gamma_{data} = T_{data}(\Gamma,0)$, where $T_{data}$ is the recursive function defined as
+From the ordered list of computation $\Gamma$, a new ordered list $\Gamma_{sync}$ is obtained from the call $\Gamma_{sync} = F_{sync}(\Gamma,0)$, where $F_{sync}$ is the recursive function defined as
 \begin{equation*}
-T_{data}(\Gamma,i) = 
-\begin{cases} 	[\Gamma[i]] \cdot T_{data}(\Gamma,i+1) & \mbox{if }\forall j< i\mbox{, } c_j \not\pprec c_i\\
-				[*_i(R^*_i,w^*_i),\Gamma[i]^*(R_i \cup \{w^*_i\},w_i,\text{exp}_i)] \cdot T_{data}(\Gamma,i+1) & \text{where } R^*_i = \{\bigcup_{k} \{w_k\} \text{, }\\
-        & \forall k<i\text{, and }c_k \pprec c_i\}\\
-				[] & \mbox{if }i=|\Gamma|.
+F_{sync}(\Gamma,j) = 
+\begin{cases} 	[\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) & \mbox{if }\forall i<j \mbox{, } c_i \not\pprec c_j\\
+				[sync(\{w_i\},w_i,identity,\emptyset), c_j] \cdot F_{sync}(\Gamma,j+1) & \forall i<j \text{, and }c_i \pprec c_j\}\\
+				[] & \mbox{if }j=|\Gamma|.
 \end{cases}
@@
 
- The final step of this parallelization is to run $\Gamma_{data}$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program, defined by
+ The final step of this parallelization is to run $\Gamma_{sync}$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program, defined by
 \begin{equation}
-\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data}),
+\mathcal{MSP}_k(T,\mathcal{M}_k,\mathcal{E},\mathcal{D},\Delta_k, \mathcal{S},\Gamma_{sync}),
 \end{equation}
@@
 %------------------------------
-\subsection{Task parallelism}
+\subsection{Hybrid parallelism}
 A task parallelization technique is a technique to transform a program as a dependency graph of different tasks.
 A dependency graph exhibits parallel tasks, or on the contrary sequential execution of tasks.
-In a multi-stencil program a task is defined as a computation $c(R,w,\text{exp})$.
-As a result, the dependency graph of a multi-stencil program represents the dependencies between local and stencil computations of $\Gamma$.
+Such a dependency graph can directly be given to a dynamic scheduler, or can be statically scheduled.
+In this paper, we introduce task parallelism by building the dependency graph between computations of the sequential list $\Gamma_{sync}$ (which itself take into account data parallelism).
 
 \begin{mydef}
-For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, with $i < j$, it is said that $c_j$ is data dependant from $c_i$, denoted $c_i\prec c_j$, if $\{w_i\} \cap R_j \neq \emptyset$.
+For two computations $c_i(R_i,w_i,\text{exp}_i,D_i)$ and $c_j(R_j,w_j,\text{exp}_j,D_j)$, with $i < j$, it is said that $c_j$ is data dependant from $c_i$, denoted $c_i\prec c_j$, if $\{w_i\} \cap R_j \neq \emptyset$.
 In this case, $c_i$ has to be computed before $c_j$.
@@
 
-\begin{myprop}
-The binary relation $\prec$ is not transitive. 
-\end{myprop}
+% \begin{myprop}
+% The binary relation $\prec$ is not transitive. 
+% \end{myprop}
 
-\begin{proof}
-Considering three computations $c_0(R_0,w_0,\text{exp}_0)$, $c_1(R_1,w_1,\text{exp}_1)$ and $c_2(R_2,w_2,\text{exp}_2)$, where $\{w_0\}\cap R_1 \neq \emptyset$ and $\{w_1\}\cap R_2 \neq \emptyset$.
+% \begin{proof}
+% Considering three computations $c_0(R_0,w_0,\text{exp}_0)$, $c_1(R_1,w_1,\text{exp}_1)$ and $c_2(R_2,w_2,\text{exp}_2)$, where $\{w_0\}\cap R_1 \neq \emptyset$ and $\{w_1\}\cap R_2 \neq \emptyset$.
 In this case two dependencies can be extracted $c_0 \prec c_1$ and $c_1 \prec c_2$.
 However, the dependency $c_0 \prec c_2$ is not true as $\{w_0\}\cap R_2 = \emptyset$.
-\end{proof}
+% \end{proof}
 
-It has been proved that the binary relation $\prec$ is not transitive.
+% It has been proved that the binary relation $\prec$ is not transitive.
 This property is due to the fact that $\prec$ is defined as a data dependency between computations.
@@
 
-\begin{mydef}
-For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, a time dependency is denoted $c_i \blacktriangleleft c_j$ and means that $c_i$ has to be computed before $c_j$.
-\end{mydef}
+% \begin{mydef}
+% For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, a time dependency is denoted $c_i \blacktriangleleft c_j$ and means that $c_i$ has to be computed before $c_j$.
+% \end{mydef}
 
-As a result, the relation $\blacktriangleleft$ is more general than $\prec$.
+% As a result, the relation $\blacktriangleleft$ is more general than $\prec$.
 
-\begin{myprop}
-For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$ such that $c_i \prec c_j$, $c_i \blacktriangleleft c_j$ is verified.
-\end{myprop}
+% \begin{myprop}
+% For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$ such that $c_i \prec c_j$, $c_i \blacktriangleleft c_j$ is verified.
+% \end{myprop}
 
-\begin{proof}
-By definition $c_i \prec c_j$ means that $\{w_i\}\cap R_j \neq \emptyset$ and that $c_i$ has to computed before $c_j$.
+% \begin{proof}
+% By definition $c_i \prec c_j$ means that $\{w_i\}\cap R_j \neq \emptyset$ and that $c_i$ has to computed before $c_j$.
 %Thus, the relation $c_i \prec c_j$ is stronger than $c_i \blacktriangleleft c_j$.
-\end{proof}
+% \end{proof}
 
-\begin{myprop}
-The binary relation $\blacktriangleleft$ is transitive. 
-\end{myprop}
+% \begin{myprop}
+% The binary relation $\blacktriangleleft$ is transitive. 
+% \end{myprop}
 
-\begin{proof}
-Considering three computations $c_0(R_0,w_0,\text{exp}_0)$, $c_1(R_1,w_1,\text{exp}_1)$ and $c_2(R_2,w_2,\text{exp}_2)$, where $c_0 \blacktriangleleft c_1$ and $c_1 \blacktriangleleft c_2$.
+% \begin{proof}
+% Considering three computations $c_0(R_0,w_0,\text{exp}_0)$, $c_1(R_1,w_1,\text{exp}_1)$ and $c_2(R_2,w_2,\text{exp}_2)$, where $c_0 \blacktriangleleft c_1$ and $c_1 \blacktriangleleft c_2$.
 $c_0$ is computed before $c_1$ and $c_1$ is computed before $c_2$, as a result $c_0$ is computed before $c_2$ and the relation $c_0 \blacktriangleleft c_2$ is verified.
-\end{proof}
+% \end{proof}
 
@@
 
-From an ordered list of computations, a directed dependency graph $\Gamma_{task}(V,A)$ can be built finding all pairs of computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, with $i<j$, such that $c_i \blacktriangleleft c_j$.
-Those time dependencies can be found from pairs of data dependencies $c_i \prec c_j$.
+From an ordered list of computations $\Gamma_{sync}$, a directed dependency graph $\Gamma_{dep}(V,A)$ can be built finding all pairs of computations $c_i(R_i,w_i,\text{exp}_i,D_i)$ and $c_j(R_j,w_j,\text{exp}_j,D_j)$, with $i<j$, such that $c_i \prec c_j$.
+%Those time dependencies can be found from pairs of data dependencies $c_i \prec c_j$.
 
@@
 \begin{mydef}
-From an ordered list $\Gamma$ of computations $c(R,w,\text{exp})$, a directed dependency graph $\Gamma_{task}(V,A)$ is obtained from the call $T_{task}(\Gamma,0)$, where $T_{task}$ is the recursive function
+From the ordered list $\Gamma_{sync}$ of computations $c(R,w,\text{exp},D)$, a directed dependency graph $\Gamma_{dep}(V,A)$ is obtained from the call $T_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function
 \begin{equation*}
-T_{task}(\Gamma,i) = 
-\begin{cases} 	(\{\},\{\}) & \mbox{if }i=|\Gamma|\\
-				(c_i, \{(\overset{\frown}{c_k,c_i})\mbox{, }\forall k < i \mbox{, } c_k\prec c_i \})\cup T_{task}(\Gamma,i+1) & \mbox{if }i<|\Gamma|
+F_{dep}(\Gamma_{sync},j) = 
+\begin{cases} 	(\{\},\{\}) & \mbox{if }j=|\Gamma_{sync}|\\
+				(c_j, \{(\overset{\frown}{c_i,c_j})\mbox{, }\forall i < j \mbox{, } c_i\prec c_j \})\cup F_{dep}(\Gamma_{sync},j+1) & \mbox{if }j<|\Gamma_{sync}|
 \end{cases}
@@
 This constructive function is possible because the input is an ordered list.
-Actually, if $c_k\prec c_i$ then $k<i$.
-As a result, $c_k$ is already in $V$ when the arc $(\overset{\frown}{c_k,c_i})$ is built.
+Actually, if $c_i\prec c_j$ then $i<j$.
+As a result, $c_i$ is already in $V$ when the arc $(\overset{\frown}{c_i,c_j})$ is built.
 
 \begin{myprop}
-The directed graph $\Gamma_{task}$ is an acyclic graph.
+The directed graph $\Gamma_{dep}$ is an acyclic graph.
 \end{myprop}
@@
 \begin{proof}
-$\Gamma_{task}$ is built from $\Gamma$ which is an ordered and sequential list of computations.
-Moreover, each computation of the list $\Gamma$ is associated to a vertex of $V$, even if the same computation is represented more than once in $\Gamma$.
+$\Gamma_{dep}$ is built from $\Gamma_{sync}$ which is an ordered and sequential list of computations.
+Moreover, each computation of the list $\Gamma_{sync}$ is associated to a vertex of $V$, even if the same computation is represented more than once in $\Gamma_{sync}$.
 As a result it is not possible to go back to a previous computation and to create a cycle.
@@
 
-Using the function $T_{task}$ to build $\Gamma_{task}$, however, duplication of dependencies may occur because of the transitivity of the relation $\blacktriangleleft$.
+As a result, the hybrid parallelization of a multi-stencil program $\mathcal{MSP}$, produces $k$ multi-stencil programs
+\begin{equation*}
+\mathcal{MSP}_k(T,\mathcal{M}_k,\mathcal{E},\mathcal{D},\Delta_k,\Gamma_{dep}).
+\end{equation*}
+The set of computations $\Gamma_{dep}$ is a dependency graph between computations $c_i$ of $\Gamma$ and the synchronization computations $sync_i$ added into $\Gamma_{sync}$.
+$\Gamma_{dep}$ can be built from the call to
+\begin{equation*}
+F_{dep}(F_{sync}(\Gamma,0),0).
+\end{equation*}
+
+% Using the function $T_{task}$ to build $\Gamma_{task}$, however, duplication of dependencies may occur because of the transitivity of the relation $\blacktriangleleft$.
 Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,\text{exp}_k)$, $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, with $k<i<j$ and where $c_k \prec c_i$, $c_i \prec c_j$ and $c_k \prec c_j$, then the transitivity $c_k \blacktriangleleft c_j$ is verified because of the relations $c_k \blacktriangleleft c_i$ and $c_i \blacktriangleleft c_j$.
@@
 
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (c0) at (0,0) {$c_0$};
-   \node[] (c1) at (1,0) {$c_1$};
-   \node[] (c2) at (2,0) {$c_2$};
-   \node[] (c3) at (3,0) {$c_3$};
-   \node[] (c4) at (4,0) {$c_4$};
+% \begin{figure}[h!]
+% \begin{center}
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[] (c0) at (0,0) {$c_0$};
+%    \node[] (c1) at (1,0) {$c_1$};
+%    \node[] (c2) at (2,0) {$c_2$};
+%    \node[] (c3) at (3,0) {$c_3$};
+%    \node[] (c4) at (4,0) {$c_4$};
 
-  \path[->]
-    (c0) edge node {} (c1)
-    (c1.east) edge node {} (c2)
-    	 edge [bend right=50] node [swap] {} (c4)
-    (c2) edge node {} (c3)
-    (c3) edge node {} (c4);
-  \end{tikzpicture}
-  \caption{Useless duplication of the dependency $c_1 \blacktriangleleft c_4$}
-  \label{fig:duplication}
-\end{center}
-\end{figure}
+%   \path[->]
+%     (c0) edge node {} (c1)
+%     (c1.east) edge node {} (c2)
+%     	 edge [bend right=50] node [swap] {} (c4)
+%     (c2) edge node {} (c3)
+%     (c3) edge node {} (c4);
+%   \end{tikzpicture}
+%   \caption{Useless duplication of the dependency $c_1 \blacktriangleleft c_4$}
+%   \label{fig:duplication}
+% \end{center}
+% \end{figure}
 
-Another view of the relations $\prec$ and $\blacktriangleleft$ is that $c_i\blacktriangleleft c_j$ in $\Gamma_{task}$ is a path from $c_i$ to $c_j$ of any size, while $c_i \prec c_j$ is a path of size $1$ from $c_i$ to $c_j$.
+% Another view of the relations $\prec$ and $\blacktriangleleft$ is that $c_i\blacktriangleleft c_j$ in $\Gamma_{task}$ is a path from $c_i$ to $c_j$ of any size, while $c_i \prec c_j$ is a path of size $1$ from $c_i$ to $c_j$.
 And in any case a path in the graph represents a time dependency between computations.
 Scheduling a graph dependency which do not contains duplication of information is easier.
-%For this reason, we define a reduction to apply to $\Gamma_{task}$, and which is performed as a post-transformation.
+% %For this reason, we define a reduction to apply to $\Gamma_{task}$, and which is performed as a post-transformation.
 In the example of Figure~\ref{fig:duplication} the dependency represented by the path $c_1 \blacktriangleleft c_2 \blacktriangleleft c_3 \blacktriangleleft c_4$ gives more information than the one directly represented by $c_1 \prec c_4$.
@@
 
+% % \begin{mydef}
+% % The dependency graph $\Gamma_{task}(V,A)$ is reduced to the dependency graph $\Gamma_{task}(V,A')$, where $A' \subseteq A$.
+A direct arc $(\overset{\frown}{c_i,c_j})$, $i<j$, of $A$, which means that $c_i \prec c_j$, is included in $A'$ if no longer path $c_i \blacktriangleleft c_j$ exists in $A$ because of transitivity.
+% % \end{mydef}
 % \begin{mydef}
-% The dependency graph $\Gamma_{task}(V,A)$ is reduced to the dependency graph $\Gamma_{task}(V,A')$, where $A' \subseteq A$.
-A direct arc $(\overset{\frown}{c_i,c_j})$, $i<j$, of $A$, which means that $c_i \prec c_j$, is included in $A'$ if no longer path $c_i \blacktriangleleft c_j$ exists in $A$ because of transitivity.
+% A DAG is transitive if it contains an arc $(\overset{\frown}{u,v})$ between any two vertices such that there is a path from $u$ to $v$.
+The transitive closure of a DAG $G=(V,E)$, is the DAG $G_T=(V,E_T)$ for which $E_T$ is the minimal subset of $V \times V$ that includes $E$ and makes $G_T$ transitive.
 % \end{mydef}
-\begin{mydef}
-A DAG is transitive if it contains an arc $(\overset{\frown}{u,v})$ between any two vertices such that there is a path from $u$ to $v$.
-The transitive closure of a DAG $G=(V,E)$, is the DAG $G_T=(V,E_T)$ for which $E_T$ is the minimal subset of $V \times V$ that includes $E$ and makes $G_T$ transitive.
-\end{mydef}
 
-\begin{mydef}
-An arc $(\overset{\frown}{u,v})$ of a DAG is redundant if there is a path from $u$ to $v$ in the DAG that does not include the edge.
+% \begin{mydef}
+% An arc $(\overset{\frown}{u,v})$ of a DAG is redundant if there is a path from $u$ to $v$ in the DAG that does not include the edge.
 A DAG that does not contain any redundant arc is called minimal.
 The transitive reduction of a DAG $G$ is its unique minimal sub-graph having the same transitive closure.
-\end{mydef}
+% \end{mydef}
 
-Thus, the removal of the redundant dependencies of $\Gamma_{task}$ consists in applying a transitive reduction.
-
-%et alors qu'est-ce qu'on fait de Gamma_task ? Il faut le scheduler, donner les travail aux ressources ! On va voir plus tard dans cet article...
+% Thus, the removal of the redundant dependencies of $\Gamma_{task}$ consists in applying a transitive reduction.
 
 %------------------------------
-\subsection{Hybrid parallelism}
-It is also possible to combine data and task parallelization techniques to get hybrid parallelism, sometimes more efficient on hybrid architectures.
+% \subsection{Hybrid parallelism}
+% It is also possible to combine data and task parallelization techniques to get hybrid parallelism, sometimes more efficient on hybrid architectures.
 A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list of computations $c_i$ and updated-computations $*_i$.
@@
 
-However, we have to clarify the dependency operator $\prec$ for the specific case of updated-computations.
+% However, we have to clarify the dependency operator $\prec$ for the specific case of updated-computations.
 As explained in Section~\ref{sect:dataparal}, if $c_j$ is an updated-computations, which means that at least one $c_i$ exists such that $i<j$ and $c_i \pprec c_j$, $c_j$ is transformed to the ordered list of two computations
 
-\begin{equation*}
-[*_j(R^*_j,w^*_j),c^*_j(R_j \cup \{w^*_j\},w_j,\text{exp}_j)].
-\end{equation*}
+% \begin{equation*}
+% [*_j(R^*_j,w^*_j),c^*_j(R_j \cup \{w^*_j\},w_j,\text{exp}_j)].
+% \end{equation*}
 
-As a result, and because $*_j$ is itself a computation, it appears dependencies relations.
+% As a result, and because $*_j$ is itself a computation, it appears dependencies relations.
 First, by definition, the dependency $c^*_j \prec *_j$ is created because $\{w^*_j\} \subset R_j \cup \{w^*_j\}$.
@@
 
-The choice of separating $*_j$ from $c_j$ has been done to entirely expose the eventual task parallelism in the hybrid parallelization.
+% The choice of separating $*_j$ from $c_j$ has been done to entirely expose the eventual task parallelism in the hybrid parallelization.
 For example, Figure~\ref{fig:more} give an exemple where initially $c_0 \pprec c_2$ and $c_0 \prec c_1$, $c_1 \prec c_2$.
@@
 
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (c0) at (0,0) {$c_0$};
-   \node[] (c1) at (2,0) {$c_1$};
-   \node[] (star2) at (2,1) {$*_2$};
-   \node[] (c2) at (4,0) {$c^*_2$};
+% \begin{figure}[h!]
+% \begin{center}
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[] (c0) at (0,0) {$c_0$};
+%    \node[] (c1) at (2,0) {$c_1$};
+%    \node[] (star2) at (2,1) {$*_2$};
+%    \node[] (c2) at (4,0) {$c^*_2$};
 
-  \path[->]
-    (c0) edge node {} (c1)
-         edge node {} (star2)
-    (c1) edge node {} (c2)
-    (star2) edge node {} (c2);
-  \end{tikzpicture}
-  \caption{Example of hybrid parallelization.}
-  \label{fig:more}
-\end{center}
-\end{figure}
+%   \path[->]
+%     (c0) edge node {} (c1)
+%          edge node {} (star2)
+%     (c1) edge node {} (c2)
+%     (star2) edge node {} (c2);
+%   \end{tikzpicture}
+%   \caption{Example of hybrid parallelization.}
+%   \label{fig:more}
+% \end{center}
+% \end{figure}
 
-As a result, the hybrid parallelization of a multi-stencil program $\mathcal{MSP}$, produces $k$ multi-stencil programs
-\begin{equation*}
-\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{hybrid}).
-\end{equation*}
-The set of computations $\Gamma_{hybrid}$ is a dependency graph between computations $c_i$ and update computations $*_i$, and can be built from the call to 
-\begin{equation*}
-T_{task}(T_{data}(\Gamma,0),0).
-\end{equation*}
+% As a result, the hybrid parallelization of a multi-stencil program $\mathcal{MSP}$, produces $k$ multi-stencil programs
+% \begin{equation*}
+% \mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{hybrid}).
+% \end{equation*}
+% The set of computations $\Gamma_{hybrid}$ is a dependency graph between computations $c_i$ and update computations $*_i$, and can be built from the call to 
+% \begin{equation*}
+% T_{task}(T_{data}(\Gamma,0),0).
+% \end{equation*}
 
-The rest of this paper deals with the hybrid parallelization of multi-stencil programs described above.
+% The rest of this paper deals with the hybrid parallelization of multi-stencil programs described above.
 The reason of this choice is the fact that by adressing the hybrid parallelism case, we also are able to adress data parallelism and task parallelism, from which hybrid parallelism is built.

### 2016-01-20 IJPP/results/strong_scaling/10K_1K/strong_scaling.py
all results Helene 20/01/2016

@@ -1,0 +1,161 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+import glob, os, string, math
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [32,64,128,256,512,1024,2048,4096,8192,16384] #1,2,4,8,
+
+meanMSL = []
+medianMSL = []
+meanSK = []
+medianSK = []
+
+#---------------------------EXTRACT INFO MSL
+print("**************************** EXTRACT MSL")
+os.chdir("./MSL")
+for c in cores:
+    filesearch = "NF_MSL_10k_1k_"+str(c)+"_"
+
+    vallist = []
+    for file in glob.glob("*.o"):
+        if file.find(filesearch)>=0 :
+            print(file)
+            with open(file) as f:
+                lines = f.readlines()
+                listVal = []
+                for line in lines :
+                    if line.find("Time TOTAL CALCULATION")>=0 :
+                        spl = string.split(line," : ")
+                        listVal.append(string.strip(spl[1],"\n"))
+                if listVal:
+                    print(max(listVal))
+                    vallist.append(max(listVal))
+    medianarray = np.array(vallist).astype(np.float)
+    median = np.median(medianarray)
+    mean = np.mean(medianarray)
+    print("****************************")
+    print("Execution time average = "+str(mean))
+    print("****************************")
+    meanMSL.append(median)
+
+    print("****************************")
+    print("Execution time median = "+str(median))
+    print("****************************")
+    medianMSL.append(median)
+#---------------------------
+
+#---------------------------EXTRACT INFO SKMPI
+print("**************************** EXTRACT SKMPI")
+os.chdir("../SKMPI")
+for c in cores:
+    filesearch = "strongSK_10k_1k_"+str(c)+"_"
+
+    vallist = []
+    for file in glob.glob("*.o"):
+        if file.find(filesearch)>=0 :
+            print(file)
+            with open(file) as f:
+                lines = f.readlines()
+                listVal = []
+                for line in lines :
+                    if line.find("Time TOTAL CALCULATION")>=0 :
[…]

### 2016-01-20 IJPP/results/task_scaling/500_200/analytic/results.py

@@ -1,0 +1,137 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+import glob, os, string, math
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [1,2,4,8,16,32,64,128,256,512,1024]#,2048]
+
+# COMPmean = []
+COMMmean = []
+
+# COMPmedian = []
+COMMmedian = []
+
+# #---------------------------EXTRACT INFO COMP
+# print("**************************** EXTRACT COMP")
+# os.chdir("./comp_1")
+
+# for c in cores:
+#     filesearch = "base_MSL_500_1_"+str(c)+"_"
+
+#     Tcomp = []
+
+#     for file in glob.glob("*.o"):
+#         if file.find(filesearch)>=0 :
+#             print(file)
+#             with open(file) as f:
+#                 lines = f.readlines()
+#                 tcomp = [0] * c
+#                 for line in lines :
+#                     proc = []
+#                     if line.find("compute")>=0 :
+#                         proc = string.split(line," compute ")
+#                     if len(proc) > 0 :
+#                         for p in range(0,c):
+#                             if int(proc[0])==p:
+#                                 if line.find("compute")>=0 :
+#                                     spl = string.split(line," : ")
+#                                     tcomp[p] = tcomp[p] + float(string.strip(spl[1],"\n"))
+#                                     break
+#                 #print(tcomp)
+#                 Tcomp.append(max(tcomp))
+
+#     TcompArray = np.array(Tcomp).astype(np.float)
+
+#     print("****************************")
+#     print("Computation time average = "+str(np.mean(TcompArray)))
+#     print("****************************")
+#     COMPmean.append(np.mean(TcompArray))
+
+#     print("****************************")
+#     print("Computation time median = "+str(np.median(TcompArray)))
+#     print("****************************")
+#     COMPmedian.append(np.median(TcompArray))
+# #---------------------------
+
+#---------------------------EXTRACT INFO COMM
+print("**************************** EXTRACT COMM")
+os.chdir("./comm_1")
+
+for c in cores:
+    filesearch = "base_MSL_500_1_"+str(c)+"_"
+
+    Tcomm = []
+
+    for file in glob.glob("*.o"):
+        if file.find(filesearch)>=0 :
+            print(file)
+            with open(file) as f:
[…]

### 2016-01-20 IJPP/results/task_scaling/500_200/fusVSbase.py

@@ -1,0 +1,56 @@
+import math
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [1,2,4,8,16,32,64,128]
+
+#median
+timesMSL = [30.5,15.86,8.53,4.83,2.35,1.04,0.54,0.3]
+errminMSL = [30.1,15.85,8.23,4.81,2.34,1.04,0.53,0.297]
+errmaxMSL = [30.9,15.87,8.83,4.84,2.37,1.06,0.56,0.307]
+
+timesMSLF = [17.08,8.95,4.68,3.2,1.38,0.54,0.28,0.17]
+errminMSLF = [17,8.9,4.58,3.201,1.38,0.53,0.28,0.17]
+errmaxMSLF = [17.2,9.01,4.77,3.207,1.59,0.56,0.3,0.18]
+
+itsMSL = []
+itsMSLF = []
+
+for i in timesMSL:
+    itsMSL.append(500/i)
+for i in timesMSLF:
+    itsMSLF.append(500/i)
+
+matplotlib.rcParams.update({'font.size': 16})
+
+#plt.xkcd()
+
+# widths = []
+# for c in cores :
+# 	widths.append(math.log(c,2)*(0.5))
+
+# print(widths)
+
+plt1 = np.log2(cores)
+
+p2 = plt.bar(plt1, itsMSLF, color='b', alpha=0.75,label='With fusion') #yerr=[errminMSLF,errmaxMSLF],
+p1 = plt.bar(plt1, itsMSL, color='r', alpha=0.75,label='Without fusion') #yerr=[errminMSL,errmaxMSL], 
+
+plt.xticks(plt1, ('1', '2', '4', '8', '16','32','64','128'))
+
+legend = plt.legend(loc='upper left', shadow=True,prop={'size':12})
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('cores')
+plt.ylabel('iterations per second')
+pp = PdfPages('fusVSbase.pdf')
+plt.savefig(pp, format='pdf', bbox_inches='tight')
+pp.close()
+
+plt.cla()
+plt.clf()
+plt.close()

### 2016-01-20 IJPP/results/task_scaling/500_500/resultsF.py

@@ -1,0 +1,60 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [16,32,64,128,256,512,1024,2048]
+
+# fusion close
+timesMSL = [3.57,1.41,0.72,0.43,0.31,0.31,0.74,2.43]
+timesT2 = [3.74,1.45,0.74,0.43,0.27,0.4,1.06,6.05]
+
+# fusion spread
+timesT2spread = [5.83,2.46,1.27,0.7,0.42,0.57,1.23,6.27]
+
+itsMSL = []
+itsT2 = []
+itsT4 = []
+itsT2spread = []
+itsT4spread = []
+
+for i in timesMSL:
+    itsMSL.append(500/i)
+for i in timesT2:
+	if i!=0:
+		itsT2.append(500/i)
+	else :
+		itsT2.append(0)
+for i in timesT2spread:
+	if i!=0:
+		itsT2spread.append(500/i)
+	else :
+		itsT2spread.append(0)
+
+relatives_its2 = []
+for i in cores:
+    relatives_its2.append(500/(timesMSL[0]/(i/cores[0])))
+
+matplotlib.rcParams.update({'font.size': 16})
+
+#plt.xkcd()
+
+#plt.plot(cores, relatives_its2, 'b--', label='Ideal')
+plt.loglog(cores, itsMSL, 'b-+', label='MSL MPI', basex=2,basey=2)
+plt.loglog(cores, itsT2, 'r-*', label='MSL MPI + 2 threads close', basex=2,basey=2)
+plt.loglog(cores, itsT2spread, 'g-*', label='MSL MPI + 2 threads spread', basex=2,basey=2)
+
+legend = plt.legend(loc='lower right', shadow=True,prop={'size':12})
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('cores')
+plt.ylabel('iterations per second')
+pp = PdfPages('itpersecF.pdf')
+plt.savefig(pp, format='pdf', bbox_inches='tight')
+pp.close()
+
+plt.cla()
+plt.clf()
+plt.close()

### 2016-01-20 IJPP/results/task_scaling/500_500/resultsWF.py

@@ -1,0 +1,75 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [16,32,64,128,256,512,1024,2048]
+
+# without fusion close
+timesMSL = [6.08,2.69,1.38,0.74,0.45,0.34,0.35,0.56]
+timesT2 = [6.02,3.1,1.38,0.87,0.49,0.31,0.26,0.31]
+6.07,2.82,1.53,0.84,0.75,1
+timesT4 = [8.28,3.92,2,1.07,0.61,0.36,0.28,0.43]
+
+#without fusion spread
+timesT2spread = [9.69,4.68,2.38,1.26,0.69,0.42,0.31,0.31]
+timesT4spread = [11.8,5.85,2.97,1.56,0.86,0.5,0.3,0.36]
+
+itsMSL = []
+itsT2 = []
+itsT4 = []
+itsT2spread = []
+itsT4spread = []
+
+for i in timesMSL:
+    itsMSL.append(500/i)
+for i in timesT2:
+	if i!=0:
+		itsT2.append(500/i)
+	else :
+		itsT2.append(0)
+for i in timesT4:
+	if i!=0:
+		itsT4.append(500/i)
+	else :
+		itsT4.append(0)
+for i in timesT2spread:
+	if i!=0:
+		itsT2spread.append(500/i)
+	else :
+		itsT2spread.append(0)
+for i in timesT4spread:
+	if i!=0:
+		itsT4spread.append(500/i)
+	else :
+		itsT4spread.append(0)
+
+relatives_its2 = []
+for i in cores:
+    relatives_its2.append(500/(timesMSL[0]/(i/cores[0])))
+
+matplotlib.rcParams.update({'font.size': 16})
+
+#plt.xkcd()
+
+#plt.plot(cores, relatives_its2, 'b--', label='Ideal')
+plt.loglog(cores, itsMSL, 'b-+', label='MSL MPI', basex=2,basey=2)
+plt.loglog(cores, itsT2, 'r-*', label='MSL MPI + 2 threads close', basex=2,basey=2)
+plt.loglog(cores, itsT4, 'r-x', label='MSL MPI+ 4 threads close', basex=2,basey=2)
[…]

### 2016-01-20 IJPP/results/weak_scaling/400/debug.py

@@ -1,0 +1,27 @@
+import numpy as np
+import glob, os, string
+os.chdir("./test")
+filesearch = "weakMSL_25600_500_4096_"
+
+vallist = []
+for file in glob.glob("*.o"):
+    if file.find(filesearch)>=0 :
+        print(file)
+        with open(file) as f:
+            lines = f.readlines()
+            listVal = []
+            for line in lines :
+                if line.find("Time SYNC")>=0 :
+                    spl = string.split(line," : ")
+                    if len(spl)==2:
+                        point = string.split(spl[1],".")
+                        if len(point)==2:
+                            listVal.append(string.strip(spl[1],"\n"))
+            if listVal:
+                print("****************************")
+                print("Execution sync min = "+str(min(listVal)))
+                print("****************************")
+
+                print("****************************")
+                print("Execution sync max = "+str(max(listVal)))
+                print("****************************")

### 2016-01-20 IJPP/results/weak_scaling/400/weak_scaling.py

@@ -1,0 +1,177 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+import glob, os, string, math
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [16,32,64,128,256,512,1024,2048,4096,8192,16384] #1,2,4,8,
+sizes = [1600,1600,3200,3200,6400,6400,12800,12800,25600,25600,51200] #400,400,800,800,
+
+meanMSL = []
+medianMSL = []
+timesMSLerrmin=[]
+timesMSLerrmax=[]
+meanSK = []
+medianSK = []
+timesSKerrmin=[]
+timesSKerrmax=[]
+
+#---------------------------EXTRACT INFO MSL
+print("**************************** EXTRACT MSL")
+os.chdir("./MSL")
+for c,s in zip(cores,sizes):
+    filesearch = "weakMSL_"+str(s)+"_500_"+str(c)+"_"
+
+    vallist = []
+    for file in glob.glob("*.o"):
+        if file.find(filesearch)>=0 :
+            print(file)
+            with open(file) as f:
+                lines = f.readlines()
+                listVal = []
+                for line in lines :
+                    if line.find("Time TOTAL CALCULATION")>=0 :
+                        spl = string.split(line," : ")
+                        listVal.append(string.strip(spl[1],"\n"))
+                if listVal:
+                    print(max(listVal))
+                    vallist.append(max(listVal))
+    medianarray = np.array(vallist).astype(np.float)
+    median = np.median(medianarray)
+    mean = np.mean(medianarray)
+    print("****************************")
+    print("Execution time average = "+str(mean))
+    print("****************************")
+    meanMSL.append(median)
+
+    print("****************************")
+    print("Execution time median = "+str(median))
+    print("****************************")
+    medianMSL.append(median)
+
+    print("****************************")
+    print("Execution time min = "+str(min(vallist)))
+    print("****************************")
+    timesMSLerrmin.append(math.fabs(float(min(vallist))-median))
+
+    print("****************************")
+    print("Execution time max = "+str(max(vallist)))
+    print("****************************")
[…]

### 2016-01-20 IJPP/results/weak_scaling/800/weak_scaling.py

@@ -1,0 +1,177 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+import glob, os, string, math
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [16,32,64,128,256,512,1024,2048,4096,8192,16384] #1,2,4,8,
+sizes = [3200,3200,6400,6400,12800,12800,25600,25600,51200,51200,102400] #400,400,800,800,
+
+meanMSL = []
+medianMSL = []
+timesMSLerrmin=[]
+timesMSLerrmax=[]
+meanSK = []
+medianSK = []
+timesSKerrmin=[]
+timesSKerrmax=[]
+
+#---------------------------EXTRACT INFO MSL
+print("**************************** EXTRACT MSL")
+os.chdir("./MSL")
+for c,s in zip(cores,sizes):
+    filesearch = "weakMSL_"+str(s)+"_200_"+str(c)+"_"
+
+    vallist = []
+    for file in glob.glob("*.o"):
+        if file.find(filesearch)>=0 :
+            print(file)
+            with open(file) as f:
+                lines = f.readlines()
+                listVal = []
+                for line in lines :
+                    if line.find("Time TOTAL CALCULATION")>=0 :
+                        spl = string.split(line," : ")
+                        listVal.append(string.strip(spl[1],"\n"))
+                if listVal:
+                    print(max(listVal))
+                    vallist.append(max(listVal))
+    medianarray = np.array(vallist).astype(np.float)
+    median = np.median(medianarray)
+    mean = np.mean(medianarray)
+    print("****************************")
+    print("Execution time average = "+str(mean))
+    print("****************************")
+    meanMSL.append(median)
+
+    print("****************************")
+    print("Execution time median = "+str(median))
+    print("****************************")
+    medianMSL.append(median)
+
+    print("****************************")
+    print("Execution time min = "+str(min(vallist)))
+    print("****************************")
+    timesMSLerrmin.append(math.fabs(float(min(vallist))-median))
+
+    print("****************************")
+    print("Execution time max = "+str(max(vallist)))
+    print("****************************")
[…]

### 2016-01-20 ARTICLES/OTHER/LaMHA_261115/slides.tex
clean Helene 20/01/2016

--- 2015-11-12 WOLFHPC15/presentation/slides.tex
@@
 \institute[INRIA]{INRIA team Avalon\\Maison de la simulation (CEA)}
-\date{$16^{th}$ November 2015}
+\date{WOLFHPC 2015 - $16^{th}$ November 2015}
 %-------------------------------------------------------------------
@@
 \begin{itemize}
+\item Separation of concerns (domain/implementation)
 \item Easy language for the user
@@
 \begin{itemize}
-\item Productivity and code-reuse
-\item Separation of concerns
+\item DSL composition ?
 \end{itemize}
@@
 \begin{itemize}
-\item Divide an application into several functionnal black boxes
-\item Each component defines its connections with outer world
-\item Each component are built independently from the others
+\item Divide an application into several independent black boxes
+\item Each component defines its interactions with outer world
 \item Application = Assembly of components
@@
 \begin{itemize}
+\item Is it feasible ?
 \item Is it efficient ?
-\item Is it feasible ?
-\item Does it improve issues of DSLs ?
-\begin{itemize}
-\item maintainability
-\item portability
-\item productivity
+\item \textcolor{gray}{Does it improve issues of DSLs ?}
+\begin{itemize}
+\item \textcolor{gray}{maintainability}
+\item \textcolor{gray}{portability}
+\item \textcolor{gray}{productivity}
 \end{itemize}
@@
   \path[line,dashed] (sche) -- (expl);
-  \path[line] (plus) -- node [fill=white,align=center] {\scriptsize Numerical mathods\\\scriptsize Finite difference/volume/elment} (sche);
+  \path[line] (plus) -- node [fill=white,align=center] {\scriptsize Numerical methods\\\scriptsize Finite difference/volume/elment} (sche);
 \end{tikzpicture}
@@
 \begin{block}{Time}
-At each time iteration of the simulation are applied the \emph{computations kernels} of the application.
+At each time iteration of the simulation are applied the \emph{computation kernels} of the application.
 \end{block}
@@
 \begin{block}{Data}
-A Data is a set of numerical values, each one attached to a given mesh entity
+Data is a set of numerical values, each one attached to a given mesh entity
 \end{block}
@@
 \end{itemize}
-\item Data written by the computaion
+\item Data written by the computation
 \item A numerical expression
@@
 %----------------
-\begin{frame}
-\frametitle{Computation kernel: Examples}
-\begin{center}
-\resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
-\hspace{10pt}
-\resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
-\end{center}
-Compute $A$ using $\{(A,local),(C,n2)\}$ on the computation domain called $dc2$ with the numerical expression
-\tiny
-\begin{equation*}
-A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1).
-\end{equation*}
-\end{frame}
+% \begin{frame}
+% \frametitle{Computation kernel: Example}
+% \begin{center}
+% \resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
+% \hspace{10pt}
+% \resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
+% \end{center}
+% \end{frame}
 %----------------
@@
 %----------------
+\begin{frame}[fragile]
+\frametitle{Example}
+\begin{columns}
+\begin{column}{.52\textwidth}
+\resizebox{\textwidth}{!}{\includegraphics{./images/mesh.pdf}}
+\hspace{10pt}
+\resizebox{\textwidth}{!}{\includegraphics{./images/stencil2.pdf}}
+\end{column}
+\begin{column}{.48\textwidth}
+$\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$\\
+
+\begin{lstlisting}[style=base]
+@mesh:@ cart
+@mesh entities:@ cell@,@edgex
+@computation domains:@
+  allcell @in@ cell
+  alledgex @in@ edgex
+@data:@
+  A@,@cell
+  C@,@edgex
+@time:@500
+@computations:@
+  A@[@allcell@]@=&comp&@(@C@[@n1@]@@)@
+\end{lstlisting}
+\end{column}
+\end{columns}
+\end{frame}
+
+% %----------------
+% \begin{frame}[fragile]
+% \frametitle{MSL Example}
+% \begin{columns}
+% \begin{column}{.52\textwidth}
+% \begin{lstlisting}[style=base]
+% @mesh:@ cart
+% @mesh entities:@ cell,edgex,edgey
+% @computation domains:@
+%   allcell in cell
+%   alledgex in edgex
+%   alledgey in edgey
+%   part1edgex in edgex
+%   part2edgex in edgex
+% @data:@
+%   a,cell
+%   b,cell
+%   c,edgex
+%   d,edgex
+%   e,edgey
+% \end{lstlisting}
+% \end{column}
+% \begin{column}{.48\textwidth}
+% \begin{lstlisting}[style=base]
+%   f,cell
+%   g,edgey
+%   h,edgex
+%   i,cell
+%   j,edgex
+% @time:@500
+% @computations:@
+%   b[allcell]=&c0&(a)
+%   c[alledgex]=&c1&(b[n1])
+%   d[alledgex]=&c2&(c)
+%   e[alledgey]=&c3&(c)
+%   f[allcell]=&c4&(d[n1])
+%   g[alledgey]=&c5&(e)
+%   h[alledgex]=&c6&(f)
+%   i[allcell]=&c7&(g,h)
+%   j[partedgex]=&c8&(i[n1])
+% \end{lstlisting}
+% \end{column}
+% \end{columns}
+% \end{frame}
+% %----------------
+
+%----------------
 \begin{frame}
@@
 \begin{center}
-\textit{MSL produces a ready-to-fill\\component-based parallel pattern of the simulation !}
+\textit{MSL produces a "ready-to-fill"\\component-based parallel scheduling of the simulation}
 \end{center}
@@
 \begin{itemize}
+\item The SIPSim model (DDS,Data,applicators and iterators)
+\begin{itemize}
+\item Abstraction of a distributed data structure
+\end{itemize}
 \item Pipeline of stencil computations for image processing: Halide
+\begin{itemize}
+\item On grids (image), different abstraction level
+\end{itemize}
 \item DSL to component-based runtime: ?
@@
 \frametitle{MSL to Component-based runtime}
-\begin{block}{Ready-to-fill parallel pattern}
+\begin{block}{Ready-to-fill parallel scheduling: mid-grain parallelism}
 \begin{itemize}
@@
 \end{itemize}
-\item Task parallelism (mid-grain)
+\item Task parallelism
 \begin{itemize}
@@
 \end{block}
-The fine grain task parallelism is left to other languages: 
+The fine grain parallelism is left to other languages: 
 \begin{itemize}
 \item OpenMP in the kernels
-\item Kernels generated by stencil compilers (Pochoir, PATUS, Liszt etc.)
+\item Kernels generated by stencil compilers for CPU or GPU (Pochoir, Liszt etc.)
 \end{itemize}
@@
 @computation domains:@
-  allcell in cell
-  alledgex in edgex
-  alledgey in edgey
-  part1edgex in edgex
-  part2edgex in edgex
+  allcell @in@ cell
+  alledgex @in@ edgex
+  alledgey @in@ edgey
+  part1edgex @in@ edgex
+  part2edgex @in@ edgex
 @data:@
@@
 \frametitle{Data parallelism}
+\begin{enumerate}
+\item Assembly of components duplicated on each resource
+\item External Distributed Data Structure to split data among resources
+\item Detect when synchronizations are needed
+\end{enumerate}
+
+\medskip
+\begin{block}{Synchronization}
+When a computation read a data, usign a stencil shape, that has been written by a previous computation.
+\end{block}
+
 $\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$\\
 $\hookrightarrow [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$
-
-\medskip
-\begin{block}{Synchronization}
-For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronization} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j$ read a data written by $c_i$ with a stencil shape and not locally.
-\end{block}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Mid-grain task parallelism}
-
-Build the read/write dependency graph.
+\end{frame}
+%----------------
+
+%----------------
+\begin{frame}
+\frametitle{Data and task parallelism}
+
+\begin{block}{Dependency graph}
+\begin{enumerate}
+\item Each node is a computation or a synchronization
+\item Each edge is a dependency: a computation read a data that has been written before.
+\end{enumerate}
+\end{block}
 
@@
 \begin{center}
-\textbf{Data parallelism + Task parallelism}
-\end{center}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Which scheduling ?}
-From the dependency graph we can:
-\begin{itemize}
-\item Use a dynamic scheduler
-\item Compute a static scheduling
-\end{itemize}
-
-In this work,
-\begin{itemize}
-\item We build a static scheduling of the multi-stencil application
-\item This static scheduling can be dumped to a component-based runtime
-\end{itemize}
-\end{frame}
+\textcolor{gray}{Dynamic} or static scheduling ?
+\end{center}
+\end{frame}
+%----------------
+
+%----------------
+% \begin{frame}
+% \frametitle{Which scheduling ?}
+% From the dependency graph we can:
+% \begin{itemize}
+% \item Use a dynamic scheduler
+% \item Compute a static scheduling
+% \end{itemize}
+
+% In this work,
+% \begin{itemize}
+% \item We build a static scheduling of the multi-stencil application
+% \item This static scheduling can be dumped to a component-based runtime
+% \end{itemize}
+% \end{frame}
 %----------------
@@
 \frametitle{Series-Parallel Tree}
+\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
+\normalsize
 \begin{columns}
@@
 \begin{column}{.52\textwidth}
+
+\end{column}
+\end{columns}
+\end{frame}
+%----------------
+
+%----------------
+\begin{frame}
+\frametitle{Series-Parallel Tree}
+\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
+\normalsize
+\begin{columns}
+\begin{column}{.48\textwidth}
+\begin{center}
+\resizebox{\textwidth}{!}{%
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (s0)    at (0,0) {$\mathcal{S}$};
+
+   \node[] (c0)    at (-3,1) {$c_0$};
+   \node[] (star1) at (-2,1) {$sync_1$};
+   \node[] (c1)    at (-1,1) {$c_1$};
+   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
+   \node[] (c7)    at ( 1,1) {$c_7$};
+   \node[] (star8) at ( 2,1) {$sync_8$};
+   \node[] (c8)    at ( 3,1) {$c_8$};
+
+   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
+   \node[] (s2)    at (1,2) {$\mathcal{S}$};
+
+   \node[] (c2)    at (-3,3) {$c_2$};
+   \node[] (star4) at (-2,3) {$sync_4$};
+   \node[] (c4)    at (-1,3) {$c_4$};
+   \node[] (c6)    at ( 0,3) {$c_6$};
+   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
+   %\node[] (c5)    at ( 2,3) {$c_5$};
+
+  \path[->]
+    (s0) edge node {} (c0)
+         edge node {} (star1)
+         edge node {} (c1)
+         edge node {} (p0)
+         edge node {} (c7)
+         edge node {} (star8)
+         edge node {} (c8)
+    (p0) edge node {} (s1)
+         edge node {} (s2)
+    (s1) edge node {} (c2)
+         edge node {} (star4)
+         edge node {} (c4)
+         edge node {} (c4)
+         edge node {} (c6)
+    (s2) edge node {} (c3);
+  \end{tikzpicture}
+  }
+\end{center}
+\end{column}
+\begin{column}{.52\textwidth}
+\end{column}
+\end{columns}
+Loop fusion optimization possible
+\end{frame}
+%----------------
+
+%----------------
+\begin{frame}
+\frametitle{Series-Parallel Tree}
+\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
+\normalsize
+\begin{columns}
+\begin{column}{.48\textwidth}
+\begin{center}
+\resizebox{\textwidth}{!}{%
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (s0)    at (0,0) {$\mathcal{S}$};
+
+   \node[] (c0)    at (-3,1) {$c_0$};
+   \node[] (star1) at (-2,1) {$sync_1$};
+   \node[] (c1)    at (-1,1) {$c_1$};
+   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
+   \node[] (c7)    at ( 1,1) {$c_7$};
+   \node[] (star8) at ( 2,1) {$sync_8$};
+   \node[] (c8)    at ( 3,1) {$c_8$};
+
+   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
+   \node[] (s2)    at (1,2) {$\mathcal{S}$};
+
+   \node[] (c2)    at (-3,3) {$c_2$};
+   \node[] (star4) at (-2,3) {$sync_4$};
+   \node[] (c4)    at (-1,3) {$c_4$};
+   \node[] (c6)    at ( 0,3) {$c_6$};
+   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
+   %\node[] (c5)    at ( 2,3) {$c_5$};
+
+  \path[->]
+    (s0) edge node {} (c0)
+         edge node {} (star1)
+         edge node {} (c1)
+         edge node {} (p0)
+         edge node {} (c7)
+         edge node {} (star8)
+         edge node {} (c8)
+    (p0) edge node {} (s1)
+         edge node {} (s2)
+    (s1) edge node {} (c2)
+         edge node {} (star4)
+         edge node {} (c4)
+         edge node {} (c4)
+         edge node {} (c6)
+    (s2) edge node {} (c3);
+  \end{tikzpicture}
+  }
+\end{center}
+\end{column}
+\begin{column}{.52\textwidth}
 \begin{block}{Specific components}
@@
 \end{columns}
-\begin{center}
-\textbf{Direct dump to the component assembly}
-\end{center}
+Loop fusion optimization possible
 \end{frame}
@@
 %----------------
-\begin{frame}
-\frametitle{Back to Data Parallelism}
-\begin{center}
-\textbf{Back to Data Parallelism only ?}\\
-\end{center}
-\textbf{Loop Fusion}: Same computation domain under a sequence or a parallel node
-
-\medskip
-\begin{center}
-\resizebox{0.5\textwidth}{!}{%
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (s0)    at (0,0) {$\mathcal{S}$};
-
-   \node[] (c0)    at (-3,1) {$c_0$};
-   \node[] (star1) at (-2,1) {$sync_1$};
-   \node[] (c1)    at (-1,1) {$c_1$};
-   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
-   \node[] (c7)    at ( 1,1) {$c_7$};
-   \node[] (star8) at ( 2,1) {$sync_8$};
-   \node[] (c8)    at ( 3,1) {$c_8$};
-
-   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
-   \node[] (s2)    at (1,2) {$\mathcal{S}$};
-
-   \node[] (c2)    at (-3,3) {$c_2$};
-   \node[] (star4) at (-2,3) {$sync_4$};
-   \node[] (c4)    at (-1,3) {$c_4$};
-   \node[] (c6)    at ( 0,3) {$c_6$};
-   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
-   %\node[] (c5)    at ( 2,3) {$c_5$};
-
-  \path[->]
-    (s0) edge node {} (c0)
-         edge node {} (star1)
-         edge node {} (c1)
-         edge node {} (p0)
-         edge node {} (c7)
-         edge node {} (star8)
-         edge node {} (c8)
-    (p0) edge node {} (s1)
-         edge node {} (s2)
-    (s1) edge node {} (c2)
-         edge node {} (star4)
-         edge node {} (c4)
-         edge node {} (c4)
-         edge node {} (c6)
-    (s2) edge node {} (c3);
-  \end{tikzpicture}
-  }
-\end{center}
-
-\small
-\textbf{Canonical form}: $[c_0,sync_1,c_1,c_2,sync_4,c_4,c_6,\{c_3,c_5\},c_7,sync_8,c_8]$
-\end{frame}
+% \begin{frame}
+% \frametitle{Data Parallelism optimization}
+% \begin{block}{Loop Fusion}
+% Same computation domain under a sequence or a parallel node
+% \end{block}
+
+% \medskip
+% \begin{center}
+% \resizebox{0.5\textwidth}{!}{%
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%    \node[] (s0)    at (0,0) {$\mathcal{S}$};
+
+%    \node[] (c0)    at (-3,1) {$c_0$};
+%    \node[] (star1) at (-2,1) {$sync_1$};
+%    \node[] (c1)    at (-1,1) {$c_1$};
+%    \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
+%    \node[] (c7)    at ( 1,1) {$c_7$};
+%    \node[] (star8) at ( 2,1) {$sync_8$};
+%    \node[] (c8)    at ( 3,1) {$c_8$};
+
+%    \node[] (s1)    at (-1,2) {$\mathcal{S}$};
+%    \node[] (s2)    at (1,2) {$\mathcal{S}$};
+
+%    \node[] (c2)    at (-3,3) {$c_2$};
+%    \node[] (star4) at (-2,3) {$sync_4$};
+%    \node[] (c4)    at (-1,3) {$c_4$};
+%    \node[] (c6)    at ( 0,3) {$c_6$};
+%    \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
+%    %\node[] (c5)    at ( 2,3) {$c_5$};
+
+%   \path[->]
+%     (s0) edge node {} (c0)
+%          edge node {} (star1)
+%          edge node {} (c1)
+%          edge node {} (p0)
+%          edge node {} (c7)
+%          edge node {} (star8)
+%          edge node {} (c8)
+%     (p0) edge node {} (s1)
+%          edge node {} (s2)
+%     (s1) edge node {} (c2)
+%          edge node {} (star4)
+%          edge node {} (c4)
+%          edge node {} (c4)
+%          edge node {} (c6)
+%     (s2) edge node {} (c3);
+%   \end{tikzpicture}
+%   }
+% \end{center}
+
+% \small
+% %\textbf{Canonical form}: $[c_0,sync_1,c_1,c_2,sync_4,c_4,c_6,\{c_3,c_5\},c_7,sync_8,c_8]$
+% \end{frame}
 %----------------
@@
 \subsection{Evaluation}
+
+%----------------
+\begin{frame}
+\frametitle{Resume}
+The MSL compiler can produce:
+\begin{itemize}
+\item A data parallel pattern of the multi-stencil application
+\item \textcolor{gray}{An hybrid (data + task) pattern of the multi-stencil application}
+\end{itemize}
+\end{frame}
+%----------------
+
 \begin{frame}
@@
 \item MSL implementation which uses the SkelGIS DDS
-\item Cluster Edel of Grid'5000: 2 8-cores Intel Xeon E5520, 24GB RAM / nodes
+%\item Cluster Edel of Grid'5000: 2 8-cores Intel Xeon E5520, 24GB RAM / nodes, Infiniband
+\item Thin Nodes TGCC Curie: two 8-cores Intel Sandy Bridge 2.7GHz, 64GB RAM, Infiniband
 \end{itemize}
@@
 \frametitle{Evaluations}
+Mesh size: $10k \times 10k$
+Number of iterations: $500$
 \begin{columns}
@@
 \begin{center}
-\small Mesh $10k \times 10k$, $500$ iterations\\
-\resizebox{\textwidth}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
+\resizebox{\textwidth}{!}{\includegraphics{./images/logtimes_NF.pdf}}
 \end{center}
@@
 \begin{center}
-\small Mesh $10k \times 10k$, $2K$ iterations\\
-\resizebox{\textwidth}{!}{\includegraphics{./images/logtimes_10K_2K.pdf}}
+\resizebox{\textwidth}{!}{\includegraphics{./images/itpersec_NF.pdf}}
 \end{center}
@@
 \frametitle{Conclusion}
+\begin{block}{Conclusion}
 \begin{itemize}
 \item A DSL for Multi-Stencil applications (MSL)
-\item The compilation of MSL to get a static scheduling
+\item The compilation of MSL to get a parallel scheduling pattern of the simulation
+\begin{itemize}
+\item Data parallelism
+\item Task parallelism
+\end{itemize}
 \item The dump to a component-based runtime
-\item No overhead introduced by the solution
-\end{itemize}
-\end{frame}
-
-\begin{frame}
+\item Data parallelism evaluation: no overhead introduced 
+\end{itemize}
+\end{block}
+\end{frame}
+%-------------------------------------------------------------
+
+%-------------------------------------------------------------
+\begin{frame}
+\begin{block}{Perspectives}
 \frametitle{Perspectives}
 \begin{itemize}
-\item A comparison with an improved version of SkelGIS
-\item More evaluations on TGCC Curie (CEA) up to $32k$ cores
-\item An evaluation of the Data+Task parallelism (mixing the static scheduling and OpenMP)
-\item A new dump with a component responsible for the dynamic scheduling of the application (using OpenMP 4.x)
-\item A new dump to CPU+GPGPUs (using existing stencil compilers Pochoir, PATUS etc.)
-\item An alternative to the SkelGIS DDS using Global Arrays (PGAS)
-\end{itemize}
-\end{frame}
-%-------------------------------------------------------------
-
-\begin{withoutheadline}
-\begin{frame}{}
-\begin{center}
-\huge Thank You !
-\end{center}
-\end{frame}
-\end{withoutheadline}
+\item Improvment of the language (convergence criteria, reduction etc.)
+\item Scalability up to $32k$ cores on TGCC Curie (CEA)
+\begin{itemize}
+\item Compared with SkelGIS and MPI only
+\end{itemize}
+\item Evaluations on Data+Task parallelism
+\begin{itemize}
+\item OpenMP 3 inside kernels
+\end{itemize}
+\item Dynamic scheduling
+\begin{itemize}
+\item OpenMP 4 with a scheduling component (libgomp)
+\item Kstar for StarPU and XKaapi runtimes
+\end{itemize}
+\item CPU+GPGPUs using stencil compilers (Pochoir, PATUS etc.)
+\end{itemize}
+$\hookrightarrow$ Show portability, maintainability introduced by components
+\end{block}
+\end{frame}
+%-------------------------------------------------------------
+
+%\begin{withoutheadline}
+% \begin{frame}{}
+% \begin{center}
+% \huge Thank You !
+% \end{center}
+% \end{frame}
+%\end{withoutheadline}
 

### 2016-01-20 ARTICLES/OTHER/WG_Avalon/slides.tex
clean Helene 20/01/2016

--- 2016-01-20 ARTICLES/OTHER/LaMHA_261115/slides.tex
@@
 \usepackage{subfig}
+
+\usepackage[absolute,overlay]{textpos}
+\usepackage{tcolorbox}
 
@@
 \frametitle{Motivation} % components
+\begin{tcolorbox}[width=\textwidth,title={Motivation}]
 What if a DSL produces a component-based runtime ?
-
 \begin{itemize}
@@
 \item Is it efficient ?
-\item \textcolor{gray}{Does it improve issues of DSLs ?}
-\begin{itemize}
-\item \textcolor{gray}{maintainability}
-\item \textcolor{gray}{portability}
-\item \textcolor{gray}{productivity}
-\end{itemize}
-\end{itemize}
-
-Let's take a useful example: \textit{the Multi-Stencil Language }!
+\item Does it improve issues of DSLs ? (maintainability etc.)
+\end{itemize}
+\end{tcolorbox}
+
+\begin{block}{Contributions}
+\begin{itemize}
+\item The Multi-Stencil Language (MSL)
+\item A first component-based back end
+\item No overheads introduced
+\end{itemize}
+\end{block}
+
 \end{frame}
@@
 %-------------------------------------------------------------
-\section{Multi-Stencil Language}
+\section{Multi-Stencil Applications}
 % related work, multi-stencil problem, language
@@
 %----------------
-\subsection{Multi-Stencil Language}
+\subsection{Multi-Stencil Applications}
 %----------------
@@
 \begin{frame}
-\frametitle{Multi-Stencil program}
+\frametitle{Multi-Stencil application}
 \begin{center}
@@
 %----------------
+
+%-------------------------------------------------------------
+\section{MSL Overview}
+% take the same notations inside components + user
+%-------------------------------------------------------------
+\subsection{MSL Overview}
 
@@
 
+\medskip
+\textit{\textcolor{blue}{blue} = MSL and black = user identifiers}
+
+\medskip
 \begin{lstlisting}[style=base]
@@
 \end{columns}
+\end{frame}
+
+%----------------
+\begin{frame}[fragile]
+\frametitle{Where is the code ?}
+Keywords + identifiers ? That's it ? Where is the code ???
+\end{frame}
+
+%----------------
+\begin{frame}[fragile]
+\frametitle{Where is the code ?}
+\begin{block}{Data parallelism}
+Whatever the mesh is and the exact computations are:
+\begin{itemize}
+\item mesh and data are splitted among resources
+\item synchronizations detected from \textbf{computation dependencies}
+\end{itemize}
+\end{block}
+
+\begin{block}{Mid-grain task parallelism}
+Task = complete computation\\
+Whatever the mesh is and the exact computations are:
+\begin{itemize}
+\item task dependencies detected from \textbf{computation dependencies}
+\end{itemize}
+\end{block}
+
+\begin{alertblock}{Fine grain optimizations and parallelizations}
+Stencil compilers already exist !
+\end{alertblock}
+
+\end{frame}
+
+%----------------
+\begin{frame}[fragile]
+\frametitle{Where is the code ?}
+\begin{block}{Data parallelism}
+Whatever the mesh is and the exact computations are:
+\begin{itemize}
+\item mesh and data are splitted among resources
+\item synchronizations detected from computation dependencies
+\end{itemize}
+\end{block}
+
+\begin{block}{Mid-grain task parallelism}
+Task = complete computation\\
+Whatever the mesh is and the exact computations are:
+\begin{itemize}
+\item task dependencies detected from computation dependencies
+\end{itemize}
+\end{block}
+
+\begin{alertblock}{Fine grain optimizations and parallelizations}
+Stencil compilers already exist !
+\end{alertblock}
+
+\begin{textblock*}{0.8\textwidth}(20mm,32mm)
+\begin{tcolorbox}[width=\textwidth,title={MSL = computation dependencies}]
+  \begin{itemize}  
+  \item produces a component based parallel scheduling of the simulation
+  \item it is agnostic from the mesh
+  \item it is agnostic from the actual numerical code
+  \end{itemize}
+\end{tcolorbox} 
+\end{textblock*}
+% based parallel scheduling of the simulation\\ It is agnostic from the mesh\\it is agnostic from the actual numerical code
+
 \end{frame}
@@
 
-%----------------
-\begin{frame}
-\frametitle{Multi-Stencil Language}
-\begin{alertblock}{MSL is not}
-\begin{itemize}
-\item a new stencil optimizer/compiler
-\item a new distributed data structure
-\end{itemize}
-\end{alertblock}
-\begin{block}{MSL is}
-\begin{itemize}
-\item a high-level language for multi-stencil simulations
-\item agnostic from the type of mesh used (data structure)
-\item based on identifiers only
-\end{itemize}
-\end{block}
-\begin{center}
-\textit{MSL produces a "ready-to-fill"\\component-based parallel scheduling of the simulation}
-\end{center}
-\end{frame}
-%----------------
+% %----------------
+% \begin{frame}
+% \frametitle{Multi-Stencil Language}
+% \begin{alertblock}{MSL is not}
+% \begin{itemize}
+% \item a new stencil optimizer/compiler
+% \item a new distributed data structure
+% \end{itemize}
+% \end{alertblock}
+% \begin{block}{MSL is}
+% \begin{itemize}
+% \item a high-level language for multi-stencil simulations
+% \item agnostic from the type of mesh used (data structure)
+% \end{itemize}
+% \end{block}
+% \begin{center}
+% \textit{MSL produces an empty\\component-based parallel scheduling of the simulation}
+% \end{center}
+% \end{frame}
+% %----------------
 
@@
 \begin{itemize}
-\item The SIPSim model (DDS,Data,applicators and iterators)
-\begin{itemize}
-\item Abstraction of a distributed data structure
-\end{itemize}
 \item Pipeline of stencil computations for image processing: Halide
@@
 \end{itemize}
-\item DSL to component-based runtime: ?
+\item DSL to component-based runtime ?
 \end{itemize}
@@
 
-%-------------------------------------------------------------
-\section{Overview}
-% take the same notations inside components + user
-%-------------------------------------------------------------
-\subsection{Overview}
-
-%----------------
-\begin{frame}
-\frametitle{MSL to Component-based runtime}
-\begin{block}{Ready-to-fill parallel scheduling: mid-grain parallelism}
-\begin{itemize}
-\item Data parallelism
-\begin{itemize}
-\item External distributed data structure
-\item Automatic detection of synchronizations
-\end{itemize}
-\item Task parallelism
-\begin{itemize}
-\item Compile a static scheduling of computation kernels
-\end{itemize}
-\end{itemize}
-\end{block}
-The fine grain parallelism is left to other languages: 
-\begin{itemize}
-\item OpenMP in the kernels
-\item Kernels generated by stencil compilers for CPU or GPU (Pochoir, Liszt etc.)
-\end{itemize}
-\end{frame}
+% %----------------
+% \begin{frame}
+% \frametitle{MSL to Component-based runtime}
+% \begin{block}{Ready-to-fill parallel scheduling: mid-grain parallelism}
+% \begin{itemize}
+% \item Data parallelism
+% \begin{itemize}
+% \item External distributed data structure
+% \item Automatic detection of synchronizations
+% \end{itemize}
+% \item Task parallelism
+% \begin{itemize}
+% \item Compile a static scheduling of computation kernels
+% \end{itemize}
+% \end{itemize}
+% \end{block}
+% The fine grain parallelism is left to other languages: 
+% \begin{itemize}
+% \item OpenMP in the kernels
+% \item Kernels generated by stencil compilers for CPU or GPU (Pochoir, Liszt etc.)
+% \end{itemize}
+% \end{frame}
+% %----------------
+
+%----------------
+% \begin{frame}
+% \frametitle{MSL to Component-based runtime}
+% \begin{block}{Empty component-based parallel scheduling of the simulation}
+% \begin{itemize}
+% \item Data parallelism
+% \begin{itemize}
+% \item Use of an external distributed data structure
+% \item Automatic detection of synchronizations
+% \end{itemize}
+% \item Mid-grain task parallelism
+% \begin{itemize}
+% \item A task is a complete computation (space loop)
+% \end{itemize}
+% \end{itemize}
+% \end{block}
+% The fine grain optimizations/parallelizations are left to other languages: 
+% \begin{itemize}
+% \item OpenMP in the kernels
+% \item Kernels generated by stencil compilers (Pochoir, Liszt etc.)
+% \end{itemize}
+% \end{frame}
 %----------------
@@
 \begin{tikzpicture}[remember picture]
-  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
+  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=red,dotted] (proc) {
     \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
 \begin{enumerate}
-\item Assembly of components duplicated on each resource
+\item Assembly of components duplicated on each resource (SPMD)
 \item External Distributed Data Structure to split data among resources
@@
 \begin{block}{Synchronization}
-When a computation read a data, usign a stencil shape, that has been written by a previous computation.
+When a computation read a data, using a stencil shape, that has been written by a previous computation.
 \end{block}
@@
 %----------------
+\begin{frame}[fragile]
+\frametitle{Separation of concerns}
+\begin{itemize}
+\item Non-computer scientist who uses the DSL
+\item The numerician who writes numerical codes
+\item The computer-scientists who writes
+\begin{itemize}
+\item parallel components
+\item DDS + Data
+\end{itemize}
+\end{itemize}
+\end{frame}
 
@@
 %----------------
-\begin{frame}
-\frametitle{Resume}
-The MSL compiler can produce:
-\begin{itemize}
-\item A data parallel pattern of the multi-stencil application
-\item \textcolor{gray}{An hybrid (data + task) pattern of the multi-stencil application}
-\end{itemize}
-\end{frame}
+% \begin{frame}
+% \frametitle{Resume}
+% The MSL compiler can produce:
+% \begin{itemize}
+% \item A data parallel pattern of the multi-stencil application
+% \item \textcolor{gray}{An hybrid (data + task) pattern of the multi-stencil application}
+% \end{itemize}
+% \end{frame}
 %----------------
@@
 \begin{itemize}
-\item Full SkelGIS implementation (DDS + specific interfaces to hide communications)
-\item MSL implementation which uses the SkelGIS DDS
-%\item Cluster Edel of Grid'5000: 2 8-cores Intel Xeon E5520, 24GB RAM / nodes, Infiniband
+\item SkelGIS DDS + manual synchronizations and fusions
+\item SkelGIS DDS + MSL
 \item Thin Nodes TGCC Curie: two 8-cores Intel Sandy Bridge 2.7GHz, 64GB RAM, Infiniband
@@
 \item Scalability up to $32k$ cores on TGCC Curie (CEA)
-\begin{itemize}
-\item Compared with SkelGIS and MPI only
-\end{itemize}
 \item Evaluations on Data+Task parallelism
@@
 \begin{itemize}
-\item OpenMP 4 with a scheduling component (libgomp)
+\item OpenMP 4 with a scheduling component
 \item Kstar for StarPU and XKaapi runtimes
@@
 \end{block}
+\end{frame}
+%-------------------------------------------------------------
+
+%-------------------------------------------------------------
+\begin{frame}
+\frametitle{Photos !}
+\resizebox{0.9\textwidth}{!}{\includegraphics{./images/sc15.jpg}}
 \end{frame}

### 2016-01-20 ARTICLES/RR/rapport/formalism_old.tex
clean Helene 20/01/2016

To numerically solve a set of PDEs, iterative methods (finite difference, finite volume or finite element methods) are frequently used to approximate the solution through a discretized (step by step) phenomena. Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied onto a mesh (space discretization). In other words, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain. Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}, and a set of auxiliary computations also needed to perform the simulation, and also called \emph{local computations}.

A stencil numerical expression~\cite{spaaTangCKLL11} represents the computation of an element of the mesh, at the time step $t$ and for a given data (or quantity) of the simulation, as a function of the quantities computed at $t-1$ mapped onto the same mesh element and to its neighboorhood. For example, if $A$ and $B$ are data of the simulation, if $(x,y)$ represents the Cartesian coordinates of an element of a two dimensional grid (Cartesian mesh), thus Figure~\ref{fig:ex} is an example of a stencil computation, where the neighborhood is the set of elements $(x-1,y)$, $(x,y-1)$, $(x+1,y)$ and $(x,y+1)$.

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{./images/example.pdf}}
  \caption{Example of a stencil computation.}
  \label{fig:ex}
\end{center}\end{figure}

Many languages and libraries exist to transparently get an optimized and parallelized code for a stencil computation~\cite{spaaTangCKLL11,citeulike12258902,Giles2011,DeVito2011LDS}. Most of those works can be considered as stencil compilers. However, a real case complex simulation is composed of more than one stencil computation and of additionnal local computations, and only a few related work handle the parallelization of an overall simulation instead of a single stencil computation. This paper presents a new approach to transparently write parallel simulations from a coarse-grain description of the numerical simulation.

In the following Section, formal definitions are given to clearly state on what we call a \textit{multi-stencil program}. Those definitions will be used in the rest of the paper.

%-------------------------------------
\subsection{Time, mesh and data}

A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ of a set of PDEs and is defined as followed.

\begin{mydef}
\textit{A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V$ is the set of vertices and $E$ the set of edges. The set of edges $E$ of a mesh $\mathcal{M}=(V,E)$ does not contain bridges.}
\end{mydef}

\begin{mydef}
$D_i$ is a set of elements of a mesh $\mathcal{M}=(V,E)$, constructed by a function $dom_i$ which defines a precise association between $V$ and $E$, $dom_i : V \times E \rightarrow D_i$. $D_i$ is called a \emph{domain} of the simulation.
\end{mydef}
For example, the set of cells $D_0$ in a Cartesian 2D mesh could be defined by exactly four vertices and four edges as a closure~\cite{}. But we could also define another set of elements $D_1$ as the simple set of vertices $V$ etc.

A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) and hybrid.

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{./images/maillages.pdf}}
  \caption{From left to right, Cartesian, curvilinear and unstructured meshes.}
  \label{fig:mesh}
\end{center}\end{figure}

\begin{mydef}
The discretization of the continuous time domain $\mathcal{T}$ is denoted $T$ such that $\forall\mbox{ }t_i\mbox{, }t_{i+1} \in T\mbox{, }\exists\mbox{ }\Delta t \in \mathbb{R}$\mbox{, }$t_{i+1} = t_i + \Delta t$. Thus, $T$ is responsible for the iteration time steps of the numerical simulation. 
\end{mydef}

In a numerical simulation a set of data, or quantities, are applied onto the mesh and represent the set of values to compute, or to use, for computation. A data is more precisely mapped onto a \emph{domain} of the mesh $D_i$, as follows:

\begin{mydef}
The set of data applied onto the mesh is denoted by $\Delta$, such that $\delta \in \Delta$ is a function which associates each element $d \in D_i$ (the domain it is applied on) to a value $v \in V$, $\delta : D_i \rightarrow V$. In the rest of this paper, the domain of a data $\delta \in \Delta$ can be given by the function $domain(\delta)=D_i$.
\end{mydef}
One can notice that in applied mathematics, the signature of $\delta$ would be $\delta : D_i \times T \rightarrow V$, however when programming a numerical simulation it is not wise to store all values of each time iteration.

%-----------------------
\subsection{Computations}

\begin{mydef}
A computation $c$ of a numerical simulation is defined as $c(R,w,\text{exp})$,
\end{mydef}
where $R \subset \Delta$ is the set of data read and used by the computation, $w \in \Delta$ is the data written by the computation and \textit{exp} is a numerical expression defined as follows:

\begin{mydef}
For $w \in \Delta$, and $R \subset \Delta$, a numerical expression \textit{exp} is a function which represents how to compute a value for an element $d \in domain(w)=D_i$, using the set $R \subset \Delta$ of input data (read data), $\text{exp} : R \times D_i \rightarrow w \times D_i$.
\end{mydef}

For example, the expression $B(x,y) = A(x,y)+(A(x,y+1)+A(x,y-1)+A(x+1,y)+A(x-1,y))/4$ of Figure~\ref{fig:ex} is a numerical expression, as $B(x,y) = A(x,y)$ is also one.

%It has to be noticed that at each time iteration, all the elements of a mesh are computed. However, it happens that computations of the mesh elements are splitted in different domains, as for example the computation of the physical border. In this case additional $D_i$ can be specified for the mesh $\mathcal{M}$.

\begin{mydef}
The set of $n$ ordered computations of a numerical simulation is denoted $\Gamma = [c_i]_{0 \leq i \leq n-1}$, such that $\forall c_i,c_j \in \Gamma$, if $i \leq j$, then $c_i$ is computed before $c_j$.
\end{mydef}

\begin{mydef}
A \textit{multi-stencil program} is defined by the quadruplet $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$.
\end{mydef}
%If the number of computations in $\Gamma$ is $card(\Gamma)=n$, such that $\bigcup_{i=0}^{n-1}c_i = \Gamma$, then $\bigcup_{i=0}^{m-1}R_i \cup w_i \subseteq \Delta$.

As introduced before, a numerical simulation is composed of stencil and local computations. Thus, the ordered list $\Gamma$ is also composed of stencil and local computations. Those types of computations are defined in the rest of this Section.

\begin{mydef}
The neighborhood $\mathcal{N}$ of an element of a domain $d \in D_i$ is a function to obtain a set of $m$ elements in any domain $D_k \subset \mathcal{M}$, where $k$ could be equal to $i$. Thus the neighborhood is defined by the function $\mathcal{N} : D_i \rightarrow D_k^m$.
\end{mydef}
The function $\mathcal{N}$ is also sometimes called the \textit{stencil shape}, or the \textit{stencil} in applied mathematics. In this paper we distinguish a stencil shape from a \textit{stencil computation} defined as follows:

\begin{mydef}
A \textit{stencil computation} is defined as a quadruplet $s(R,w,\text{exp},\mathcal{N})$,
\end{mydef}
where $R \subset \Delta$ is the set of data read by the computation, $w : D_i \rightarrow V \in \Delta$ is the data written by the computation, \textit{exp} is the numerical expression $\text{exp} : R \times D_i \rightarrow w \times D_i$, $D_i=domain(w)$, and $\mathcal{N}$ is a neighborhood function.

In a stencil computation $s$, $\forall d \in D_i$, the stencil numerical expression $\text{exp}$ is applied such that $w(d) = \text{exp}(R(d),R(\mathcal{N}(d))$. In this work, a stencil computation $s(R,w,\text{exp},\mathcal{N})$ always verifies $R \cap \{w\} = \emptyset$, otherwize an implicit numerical scheme has to be solve which is not handled by this paper. As a result, the ordered list $\Gamma$ of a multi-stencil program can be composed of a set of stencil computations applied on one or more stencil shapes.

For example, in the example of Figure~\ref{fig:ex}, the numerical expression $B(x,y) = A(x,y)+(A(x,y+1)+A(x,y-1)+A(x+1,y)+A(x-1,y))/4$

%!!!!!!!!!!!!!!!!!!!!!!!!!

Figure~\ref{fig:ex} gives an example of a stencil computation $s(R,w,\text{exp},\mathcal{N})$, where $\mathcal{M}(V,E)$ is a two dimensional Cartesian mesh. A single domain $D=domain(w)$ is defined in this example and is composed of cells formed by a cycle of four vertices $v \in V$ and four edges $e \in E$. Furthermore, in this example $R=\{A\}$, $w=B$, and for $(x,y) \in D$ the neighborhood function is 
\begin{equation*}
\mathcal{N} : (x,y) \rightarrow \{(x,y+1),(x,y-1),(x+1,y),(x-1,y)\}.
\end{equation*}
Finally, the numerical expression of this example is 
\begin{equation*}
\text{exp}(A(x,y),A(\mathcal{N}(x,y)) = B(x,y) = A(x,y)+(A(x,y+1)+A(x,y-1)+A(x+1,y)+A(x-1,y))/4.
\end{equation*}

% \begin{figure}[!h]\begin{center}
%   \resizebox{8cm}{!}{\includegraphics{./images/example.pdf}}
%   \caption{Example of a stencil computation.}
%   \label{fig:ex}
% \end{center}\end{figure}

The second type of numerical computation is a local computation.
\begin{mydef}
A local computation is a triplet $l(R,w,\text{exp})$, where $exp$ does not involve a neighborhood function $\mathcal{N}$.
\end{mydef}

A stencil program and stencil and local computations have been formally defined in this section. This formalism is used in the next Section to define two parallelization techniques of a multi-stencil program.

### 2016-01-20 ARTICLES/RR/rapport/save_11_06/component.tex
clean Helene 20/01/2016

--- 2015-06-05 rapport/component.tex
@@
 Those components are called control-components because they are used in a different way than usual components.
-Actually, as explained before, a component represents a precise functionnality of the application, while control-components represent a way to control the execution time of other components.
-Each control-component contains a single provide port, this port is associated to a single method which correspond to the execution of the control-component.
+Actually, as explained before, a component represents a precise functionnality of the application, while control-components represent a way to control the execution of components.
+Each control-component contains a single provide port, this port is associated to a single method which corresponds to the execution of the control-component.
 
@@
 \paragraph{Synchronized-sequence component.} A synchronized-sequence component, denoted $SSEQ$, is a control-component for which the provided service is to garantee that a set of data synchronizations is performed before a sequence of other required services.
-A $SSEQ$ component is thus composed of two use-multiple ports, one not necessarily ordered to link the $SSEQ$ to data to synchronize (could be more than one), and the second, ordered, to be connect $SSEQ$ to other services, as $SEQ$ does.
+A $SSEQ$ component is thus composed of two use-multiple ports, one not necessarily ordered to link the $SSEQ$ to data to synchronize (could be more than one), and the second, ordered, to connect $SSEQ$ to other services, as $SEQ$ does.
 
@@
 
+\medskip
 To be able to dump the series-parallel tree decomposition of $\Gamma_{hybrid}$ to a component assembly, each computation $c_i$ has to be associated to a component.
 Such a component contains a single provide port for which the service is to perform the computation.
-As a computation $c_i(R_i,w_i,D_i,\text{exp}_i)$ manipulates data $R$ and $w$, the component needs a use-multiple port to be linked to data.
+As a computation $c_i(R_i,w_i,\text{exp}_i)$ manipulates data $R$ and $w$, the component needs a use-multiple port to be linked to data.
 
@@
 In the series-parallel binary tree decomposition of $\Gamma_{hybrid}$, $\mathcal{S}$ and $\mathcal{P}$ are binary operators.
-On the ther hand, definitions of $SEQ$, $SSEQ$ and $PAR$ are not binary, mainly because of HPC concerns.
+On the other hand, definitions of $SEQ$, $SSEQ$ and $PAR$ are not binary (use-multiple), mainly because of HPC concerns.
 Actually, as less as the number of calls is, the better the performance is.

### 2016-01-20 ARTICLES/RR/rapport/save_11_06/parallelism.tex
clean Helene 20/01/2016

Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.

%------------------------------
\subsection{Data parallelism}
\label{sect:dataparal}
In a data parallelization technique, the idea is to split the data on which the program is computed in balanced sub-parts, one for each available resource. The same sequential program can afterwards be applied on each sub-part simultaneously, with some additioinal synchronizations between resources to update the data not computed locally, and thus to guarantee a correct result.

More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$. This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper. As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\Delta=\{\Delta_0,\dots,\Delta_{p-1}\}$. The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma_{data}$.

\begin{mydef}
For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j=s_j(R_j,w_j,\text{exp}_j)$, and $\{w_i\} \subset R_j$. Moreover, the data to update is $\{w_i\} \cap R_j = w_i$.
\end{mydef}

Actually, a synchronization can only be needed by the data read by a stencil computation (not local), and only if this data has been modified before, which means that it has been written before. This synchronization is needed because the neighborhood function $\mathcal{N}$ of a stencil computation involves values computed on different resources.

\begin{mydef}
A synchronization between two computations $c_i \pprec c_j$ is defined as a specific computation $\text{update}(\{w_i\} \cap R_j)=\text{update}(w_i)$.
\end{mydef}

\begin{mydef}
An \textit{updated-computation} is a computation which needs a set of data synchronizations before it can be performed. The computation $c_j$ is transformed to an updated-computation $c^*_j(R_j,w_j,\text{exp}_j,*_j)$ if $\bigcup_{i} \{w_i\} \neq \emptyset \text{, }\forall i<j\text{, and }c_i \pprec c_j$. In such a case $*_j=update(\bigcup_{i} \{w_i\})$ is performed before the evaluation of $exp_j$.
\end{mydef}

\begin{mydef}
The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[c_i]_{0 \leq i \leq n-1}$ and $l_2=[c'_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[c_0,\dots,c_{n-1},c'_0,\dots,c'_{m-1}]$.
\end{mydef}

\begin{mydef}
From the ordered list of computation $\Gamma$, a new ordered list $\Gamma_{data}$ is obtained from the call $\Gamma_{data} = T_{data}(\Gamma,0)$, where $T_{data}$ is the recursive function defined as
\begin{equation*}
T_{data}(\Gamma,i) = 
\begin{cases} 	[\Gamma[i]] \cdot T_{data}(\Gamma,i+1) & \mbox{if }\forall j< i\mbox{, } c_j \not\pprec c_i\\
				[c^*_i(\Gamma[i],\text{update}(\bigcup_{k}\{w_k\})] \cdot T_{data}(\Gamma,i+1) & \forall k < i\mbox{, }c_k \pprec c_i\\
				[] & \mbox{if }i=|\Gamma|.
\end{cases}
\end{equation*}
\end{mydef}

The construction of $\Gamma_{data}$ implies that if a synchronization is needed it is always proceeded before the computation wich needs it. One can notice that a combination of synchronizations is possible in some cases, and thus it is possible to synchronize more than one data at a time. %If the number of synchronizations is reduced, final performance of the program will be better. 
For this reason, we define a reduction which can be applied to $\Gamma_{data}$, and which is performed as a post-transformation. This reduction, could propose better performance depending on the used network speed.

\begin{myprop}
Denoting four computations $c_l$, $c_k$, $c_i$ and $c_j$ of $\Gamma$, with $l<k$, $i<j$, and where $c_l \pprec c_k$ and $c_i \pprec c_j$. Then, computations $c_k$ and $c_j$ are tansformed to $c^*_k(c_k,update(w_l))$ and $c^*_j(c_j,update(w_i))$. If $i<k$, a single update call can be performed at $k$ such that $c^*_k(c_k,update(w_l\cup w_i))$, and in this case, the computation $c_j$ is not transformed.
\end{myprop}

\begin{proof}
By definition, the synchronization of $w_i$ can be performed between $c_i$ and $c_j$. As $i<k$ the computation $c_i$ is not performed between $k$ and $j$. As a result, the synchronization of $w_i$ can be performed between $c_i$ and $c_k$.
\end{proof}

In other words, for each synchronization in $\Gamma_{data}$, if the computation $c_i$ from which $c_j$ is synchronized occurs before the last synchronization, the two synchronizations can be reduced to one with the union of the data to update.

 The final step of this parallelization is to run $\Gamma_{data}$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program, defined by
\begin{equation}
\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data}),
\end{equation}
is performed.

%We denote this parallelization technique a coarse-grain data parallelization in contrast with the same technique applied to the finer level of a single computation. In this case, for a computation $c(R,w,D,\text{exp})$, the local domain $D$ is partitionned for $p$ resources $\{D_0,\dots,D_{p-1}\}$, and each resource $k$ is responsible for a sub-computation $c_k(R,w,D_k,\text{exp}')$.
%This finer data parallelization technique is not directly adressed by the work presented in this paper, but is exhibited, as it will be explained later.

%------------------------------
\subsection{Task parallelism}
A task parallelization technique is a technique to transform a program as a dependency graph of different tasks. A dependency graph exhibits parallel tasks, or on the contrary sequential execution of tasks. In a multi-stencil program a task is defined as a computation $c(R,w,\text{exp})$. As a result, the dependency graph of a multi-stencil program represents the dependencies between local and stencil computations of $\Gamma$.

\begin{mydef}
For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, with $i < j$, it is said that $c_j$ is data dependant from $c_i$, denoted $c_i\prec c_j$, if $\{w_i\} \cap R_j \neq \emptyset$. In this case, $c_i$ has to be computed before $c_j$. The binary relation $c_i\prec c_j$ represents a \textit{dependency}.
\end{mydef}

\begin{myprop}
The binary relation $\prec$ is not transitive. 
\end{myprop}

\begin{proof}
Considering three computations $c_0(R_0,w_0,\text{exp}_0)$, $c_1(R_1,w_1,\text{exp}_1)$ and $c_2(R_2,w_2,\text{exp}_2)$, where $\{w_0\}\cap R_1 \neq \emptyset$ and $\{w_1\}\cap R_2 \neq \emptyset$. In this case two dependencies can be extracted $c_0 \prec c_1$ and $c_1 \prec c_2$. However, the dependency $c_0 \prec c_2$ is not true as $\{w_0\}\cap R_2 = \emptyset$.
\end{proof}

It has been proved that the binary relation $\prec$ is not transitive. This property is due to the fact that $\prec$ is defined as a data dependency between computations. However, another type of dependency, only due to time, is created from $\prec$.

\begin{mydef}
For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, a time dependency is denoted $c_i \blacktriangleleft c_j$ and means that $c_i$ has to be computed before $c_j$.
\end{mydef}

As a result, the relation $\blacktriangleleft$ is more general than $\prec$.

\begin{myprop}
For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$ such that $c_i \prec c_j$, $c_i \blacktriangleleft c_j$ is verified.
\end{myprop}

\begin{proof}
By definition $c_i \prec c_j$ means that $\{w_i\}\cap R_j \neq \emptyset$ and that $c_i$ has to computed before $c_j$. %Thus, the relation $c_i \prec c_j$ is stronger than $c_i \blacktriangleleft c_j$.
\end{proof}

\begin{myprop}
The binary relation $\blacktriangleleft$ is transitive. 
\end{myprop}

\begin{proof}
Considering three computations $c_0(R_0,w_0,\text{exp}_0)$, $c_1(R_1,w_1,\text{exp}_1)$ and $c_2(R_2,w_2,\text{exp}_2)$, where $c_0 \blacktriangleleft c_1$ and $c_1 \blacktriangleleft c_2$. $c_0$ is computed before $c_1$ and $c_1$ is computed before $c_2$, as a result $c_0$ is computed before $c_2$ and the relation $c_0 \blacktriangleleft c_2$ is verified. 
\end{proof}

\begin{mydef}
A directed acyclic graph (DAG) $G(V,A)$ is a graph where the edges are directed from a source to a destination vertex and where, following the direction of edges, no cycle can be found from a vertex $u$ to itself. A directed edge is called an arc, and for two vertices $v,u \in V$ an arc from $u$ to $v$ is denoted $(\overset{\frown}{u,v}) \in A$.
\end{mydef}

From an ordered list of computations, a directed dependency graph $\Gamma_{task}(V,A)$ can be built finding all pairs of computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, with $i<j$, such that $c_i \blacktriangleleft c_j$. Those time dependencies can be found from pairs of data dependencies $c_i \prec c_j$.

\begin{mydef}
For two directed graphs $G(V,A)$ and $G'(V',A')$, the union $(V,A)\cup (V',A')$ is defined as the union of each set $(V\cup V', A \cup A')$.
\end{mydef}

\begin{mydef}
From an ordered list $\Gamma$ of computations $c(R,w,\text{exp})$, a directed dependency graph $\Gamma_{task}(V,A)$ is obtained from the call $T_{task}(\Gamma,0)$, where $T_{task}$ is the recursive function
\begin{equation*}
T_{task}(\Gamma,i) = 
\begin{cases} 	(\{\},\{\}) & \mbox{if }i=|\Gamma|\\
				(c_i, \{(\overset{\frown}{c_k,c_i})\mbox{, }\forall k < i \mbox{, } c_k\prec c_i \})\cup T_{task}(\Gamma,i+1) & \mbox{if }i<|\Gamma|
\end{cases}
\end{equation*}
\end{mydef}

This constructive function is possible because the input is an ordered list. Actually, if $c_k\prec c_i$ then $k<i$. As a result, $c_k$ is already in $V$ when the arc $(\overset{\frown}{c_k,c_i})$ is built.

\begin{myprop}
The directed graph $\Gamma_{task}$ is an acyclic graph.
\end{myprop}

\begin{proof}
$\Gamma_{task}$ is built from $\Gamma$ which is an ordered and sequential list of computations. Moreover, each computation of the list $\Gamma$ is associated to a vertex of $V$, even if the same computation is represented more than once in $\Gamma$. As a result it is not possible to go back to a previous computation and to create a cycle.
\end{proof}

Using the function $T_{task}$ to build $\Gamma_{task}$, however, duplication of dependencies may occur because of the transitivity of the relation $\blacktriangleleft$. Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,\text{exp}_k)$, $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, with $k<i<j$ and where $c_k \prec c_i$, $c_i \prec c_j$ and $c_k \prec c_j$, then the transitivity $c_k \blacktriangleleft c_j$ is verified because of the relations $c_k \blacktriangleleft c_i$ and $c_i \blacktriangleleft c_j$. However, $c_k \blacktriangleleft c_j$ is also directly represented by the binary relation $c_k \prec c_j$. As a result, a duplication of the dependency is created in $\Gamma_{task}$. In Figure~\ref{fig:duplication} an example of duplication is given for the relation $c_1 \blacktriangleleft c_4$.

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$c_0$};
   \node[] (c1) at (1,0) {$c_1$};
   \node[] (c2) at (2,0) {$c_2$};
   \node[] (c3) at (3,0) {$c_3$};
   \node[] (c4) at (4,0) {$c_4$};

  \path[->]
    (c0) edge node {} (c1)
    (c1.east) edge node {} (c2)
    	 edge [bend right=50] node [swap] {} (c4)
    (c2) edge node {} (c3)
    (c3) edge node {} (c4);
  \end{tikzpicture}
  \caption{Useless duplication of the dependency $c_1 \blacktriangleleft c_4$}
  \label{fig:duplication}
\end{center}
\end{figure}

Another view of the relations $\prec$ and $\blacktriangleleft$ is that $c_i\blacktriangleleft c_j$ in $\Gamma_{task}$ is a path from $c_i$ to $c_j$ of any size, while $c_i \prec c_j$ is a path of size $1$ from $c_i$ to $c_j$. And in any case a path in the graph represents a time dependency between computations. Scheduling a graph dependency which do not contains duplication of information is easier. 
%For this reason, we define a reduction to apply to $\Gamma_{task}$, and which is performed as a post-transformation. In the example of Figure~\ref{fig:duplication} the dependency represented by the path $c_1 \blacktriangleleft c_2 \blacktriangleleft c_3 \blacktriangleleft c_4$ gives more information than the one directly represented by $c_1 \prec c_4$. The proposed reduction is based on this property.

% \begin{mydef}
% The dependency graph $\Gamma_{task}(V,A)$ is reduced to the dependency graph $\Gamma_{task}(V,A')$, where $A' \subseteq A$. A direct arc $(\overset{\frown}{c_i,c_j})$, $i<j$, of $A$, which means that $c_i \prec c_j$, is included in $A'$ if no longer path $c_i \blacktriangleleft c_j$ exists in $A$ because of transitivity.
% \end{mydef}
\begin{mydef}
A DAG is transitive if it contains an arc $(\overset{\frown}{u,v})$ between any two vertices such that there is a path from $u$ to $v$. The transitive closure of a DAG $G=(V,E)$, is the DAG $G_T=(V,E_T)$ for which $E_T$ is the minimal subset of $V \times V$ that includes $E$ and makes $G_T$ transitive.
\end{mydef}

\begin{mydef}
An arc $(\overset{\frown}{u,v})$ of a DAG is redundant if there is a path from $u$ to $v$ in the DAG that does not include the edge. A DAG that does not contain any redundant arc is called minimal. The transitive reduction of a DAG $G$ is its unique minimal sub-graph having the same transitive closure.
\end{mydef}

Thus, the removal of the redundant dependencies of $\Gamma_{task}$ consists in applying a transitive reduction.

%et alors qu'est-ce qu'on fait de Gamma_task ? Il faut le scheduler, donner les travail aux ressources ! On va voir plus tard dans cet article...

%------------------------------
\subsection{Hybrid parallelism}
It is also possible to combine data and task parallelization techniques to get hybrid parallelism, sometimes more efficient on hybrid architectures. A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list of computations $c$ and updated-computations $c^*$. On the other hand, the task parallelization technique builds, from an ordered list of computation, a graph dependency $\Gamma_{task}$. As a result $T_{data}$ and $T_{task}$ can be composed to build $\Gamma_{hybrid}$. Thus, $k$ multi-stencil programs
\begin{equation*}
\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{hybrid})
\end{equation*}
are responsible for an hybrid parallelization of a multi-stencil program. The set of computation $\Gamma_{hybrid}$ is a dependency graph between computations $c$ and updated-computations $c^*$ and can be built from the call to 
\begin{equation*}
T_{task}(T_{data}(\Gamma,0),0).
\end{equation*}

The rest of this paper deals with the hybrid parallelization of multi-stencil programs described above. The first reason for this choice is that this paper does not adress the portability problem, which however will be a perspective. Second, by adressing the hybrid parallelism case, we also are able to adress data parallelism and task parallelism, from which hybrid parallelism is built.

### 2016-01-20 ARTICLES/RR/rapport_v0/csm.tex
clean Helene 20/01/2016

--- 2015-04-10 rapport/csm.tex
@@
 \subsection{Phase, group and kernel}
-Using the formalism described in Sections~\ref{sect:formalism} and~\ref{sect:component}, the following definitions are given for the solution.
+Using the formalism described in the Section~\ref{sect:formalism}, the following definitions are given for the solution.
 
@@
 Thus, a dependency exists between two different \emph{groups} of a stencil program $\mathcal{P}$, but the computations inside a \emph{group} are unordered, without dependencies.
-As a result, the Group component is defined as $PAR$
-\begin{equation}
-Group(\{provide,use-multiple\},\{group\}),
-\end{equation}
- where the function $group$ creates a thread for each component to use, and join all threads at the end.
-
-\begin{algorithm}[H]
- \ForAll{component $cp$ to use}{
- create thread $t$
- $t$ uses the provide interface of $cp$
- }
- join all threads
- \label{alg:par}
- \caption{group function}
- \end{algorithm}
+As a result, the Group component is equivalent to $PAR$.
 
@@
  }
- \caption{kernel function}
+ \caption{phase function}
  \end{algorithm}
@@
 %-------------------------------------
-\subsection{Overall model}
-
-%---------
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (phase) at (0,0) {Phase};
-   \node[component] (group) [right=of phase] {Group};
-   \node[component] (kernel) [right=of group] {Kernel};
-   \node[dcomponent] (init) [above=1.5cm of group] {Initializer};
-   \node[component] (data) [above=1.5cm of phase] {Data};
-   \node[component] (time) [left=of phase] {Time};
-   \node[component] (drapp) [above=1.5cm of time] {DrApp};
-   \node[dcomponent] (conv) [left=of time] {Conv};
-   \node[component] (dr) [above=1.5cm of drapp] {Driver};
-   \node[component] (dds) [above=1.5cm of data] {DDS};
-   \node[component] (connec) [above=1.5cm of init] {Connector};
-
-  \path[->]
-  	(dr)		edge  [connection]	node	{}	(dds)
-  	(dr)		edge  [connection]	node	{}	(drapp)
-  	(drapp)		edge  [connection]	node	{1,n}	(data)
-  	(drapp)		edge  [connection]	node	{}	(time)
-  	(data)  	edge  [connection]	node	{}	(dds)
-  	(data)  	edge  [connection]	node	{}	(connec)
-  	(connec)	edge  [connection]	node	{}	(dds)
-  				edge  [mpiconn]		node	{}	(4,3)
-  	(data)		edge  [connection]	node	{1,n}	(init)
-  	(time)		edge  [connection]	node	{}	(conv)
-  	(time)		edge  [connection]	node	{1,n}	(phase)
-  	(phase)		edge  [connection]	node	{1,n}	(data)
-  	(phase)		edge  [connection]	node	{1,n}	(group)
-  	(phase)		edge  [bend right=50,connection]   node  [swap]   {1,n}   (kernel)
-  	(group)		edge  [connection]	node	{1,n}	(kernel)
-  	(kernel)	edge  [connection]	node	{1,n}	(data);
-\end{tikzpicture}
-\caption{Overall assembly of the Component Stencil Model.}
-\label{phgpk}
-\end{center}
-\end{figure}
+\subsection{Overall assembly and evaluation}

### 2016-01-20 ARTICLES/WOLFHPC15/presentation/slides.tex
clean Helene 20/01/2016

--- 2016-01-20 ARTICLES/OTHER/LaMHA_261115/slides.tex
@@
 \item As many DSLs as domains
-\begin{itemize}
-\item DSL composition ?
-\end{itemize}
 \end{itemize}
@@
 @mesh:@ cart
-@mesh entities:@ cell@,@edgex
+@mesh entities:@ cell,edgex
 @computation domains:@
-  allcell @in@ cell
-  alledgex @in@ edgex
+  allcell in cell
+  alledgex in edgex
 @data:@
-  A@,@cell
-  C@,@edgex
+  A,cell
+  C,edgex
 @time:@500
 @computations:@
-  A@[@allcell@]@=&comp&@(@C@[@n1@]@@)@
+  A[allcell]=&comp&(C[n1])
 \end{lstlisting}
@@
 \begin{center}
-\textit{MSL produces a "ready-to-fill"\\component-based parallel scheduling of the simulation}
+\textit{MSL produces a ready-to-fill\\component-based parallel pattern of the simulation !}
 \end{center}
@@
 \frametitle{MSL to Component-based runtime}
-\begin{block}{Ready-to-fill parallel scheduling: mid-grain parallelism}
+\begin{block}{Ready-to-fill parallel pattern}
 \begin{itemize}
@@
 \end{itemize}
-\item Task parallelism
+\item Task parallelism (mid-grain)
 \begin{itemize}
@@
 \end{block}
-The fine grain parallelism is left to other languages: 
+The fine grain task parallelism is left to other languages: 
 \begin{itemize}
 \item OpenMP in the kernels
-\item Kernels generated by stencil compilers for CPU or GPU (Pochoir, Liszt etc.)
+\item Kernels generated by stencil compilers (Pochoir, PATUS, Liszt etc.)
 \end{itemize}
@@
 @computation domains:@
-  allcell @in@ cell
-  alledgex @in@ edgex
-  alledgey @in@ edgey
-  part1edgex @in@ edgex
-  part2edgex @in@ edgex
+  allcell in cell
+  alledgex in edgex
+  alledgey in edgey
+  part1edgex in edgex
+  part2edgex in edgex
 @data:@
@@
 \begin{column}{.52\textwidth}
-
+\begin{block}{Specific components}
+\begin{itemize}
+\item $SEQ$ to directly replace $\mathcal{S}$ nodes
+\item $PAR$ to directly replace $\mathcal{P}$ nodes
+\item $SYNC$ for synchronizations
+\item $K$ for computation kernels
+\end{itemize}
+\end{block}
 \end{column}
 \end{columns}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Series-Parallel Tree}
-\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
-\normalsize
-\begin{columns}
-\begin{column}{.48\textwidth}
-\begin{center}
-\resizebox{\textwidth}{!}{%
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (s0)    at (0,0) {$\mathcal{S}$};
-
-   \node[] (c0)    at (-3,1) {$c_0$};
-   \node[] (star1) at (-2,1) {$sync_1$};
-   \node[] (c1)    at (-1,1) {$c_1$};
-   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
-   \node[] (c7)    at ( 1,1) {$c_7$};
-   \node[] (star8) at ( 2,1) {$sync_8$};
-   \node[] (c8)    at ( 3,1) {$c_8$};
-
-   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
-   \node[] (s2)    at (1,2) {$\mathcal{S}$};
-
-   \node[] (c2)    at (-3,3) {$c_2$};
-   \node[] (star4) at (-2,3) {$sync_4$};
-   \node[] (c4)    at (-1,3) {$c_4$};
-   \node[] (c6)    at ( 0,3) {$c_6$};
-   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
-   %\node[] (c5)    at ( 2,3) {$c_5$};
-
-  \path[->]
-    (s0) edge node {} (c0)
-         edge node {} (star1)
-         edge node {} (c1)
-         edge node {} (p0)
-         edge node {} (c7)
-         edge node {} (star8)
-         edge node {} (c8)
-    (p0) edge node {} (s1)
-         edge node {} (s2)
-    (s1) edge node {} (c2)
-         edge node {} (star4)
-         edge node {} (c4)
-         edge node {} (c4)
-         edge node {} (c6)
-    (s2) edge node {} (c3);
-  \end{tikzpicture}
-  }
-\end{center}
-\end{column}
-\begin{column}{.52\textwidth}
-\end{column}
-\end{columns}
-Loop fusion optimization possible
 \end{frame}
@@
 \frametitle{Resume}
-The MSL compiler can produce:
+The MSL compiler can produces :
+\begin{itemize}
+\item A data parallel pattern of the multi-stencil application
+\item An hybrid (data + task) pattern of the multi-stencil application
+\end{itemize}
+\end{frame}
+%----------------
+
+%----------------
+\begin{frame}
+\frametitle{Resume}
+The MSL compiler can produces :
 \begin{itemize}
@@
 \item A DSL for Multi-Stencil applications (MSL)
-\item The compilation of MSL to get a parallel scheduling pattern of the simulation
-\begin{itemize}
-\item Data parallelism
-\item Task parallelism
-\end{itemize}
+\item The compilation of MSL to get data + task parallelism
 \item The dump to a component-based runtime
@@
 \end{block}
-\end{frame}
-%-------------------------------------------------------------
-
-%-------------------------------------------------------------
-\begin{frame}
+
 \begin{block}{Perspectives}
-\frametitle{Perspectives}
-\begin{itemize}
-\item Improvment of the language (convergence criteria, reduction etc.)
+\begin{itemize}
 \item Scalability up to $32k$ cores on TGCC Curie (CEA)
-\begin{itemize}
-\item Compared with SkelGIS and MPI only
-\end{itemize}
-\item Evaluations on Data+Task parallelism
-\begin{itemize}
-\item OpenMP 3 inside kernels
-\end{itemize}
-\item Dynamic scheduling
-\begin{itemize}
-\item OpenMP 4 with a scheduling component (libgomp)
-\item Kstar for StarPU and XKaapi runtimes
-\end{itemize}
-\item CPU+GPGPUs using stencil compilers (Pochoir, PATUS etc.)
+\item Evaluations of the Data+Task parallelism (OpenMP 3)
+\item Dynamic scheduling (OpenMP 4), CPU+GPU (Pochoir etc.)
 \end{itemize}

### 2016-01-20 ARTICLES/WOLFHPC15/results_g5k/results.Rmd
clean Helene 20/01/2016

---
title: "Results on Edel cluster Grid'5000"
output: pdf_document
---

On a 5000*5000 Cartesian grid and 500 time iterations, from 8 to 256 cores (32 nodes) :
For MSL :
```{r, echo=FALSE}
data2=read.table('/Users/hcoullon/Documents/Postdoc/WORK/csm/WOLFHPC15/results_g5k/5k_500_base.data',header=FALSE)
data2
```
For SkelGIS :
```{r, echo=FALSE}
data3=read.table('/Users/hcoullon/Documents/Postdoc/WORK/csm/WOLFHPC15/results_g5k/5k_500_FS.data',header=FALSE)
data3
```

Speedup where the reference time is the minimum one (MSL) :

```{r, echo=FALSE}
#calcul des speedups
xy = rep(0,length(data2$V1))
spV2 = rep(0,length(data2$V2))
for(i in 1:length(data2$V2)){
  spV2[i] = data2$V2[1]/data2$V2[i]
  xy[i] = data2$V1[i]/data2$V1[1]
}
spV3 = rep(0,length(data3$V2))
for(i in 1:length(data3$V2)){
  spV3[i] = data2$V2[1]/data3$V2[i]
  xy[i] = data3$V1[i]/data3$V1[1]
}

par(mar=c(4,4,0.5,0.5))
plot(data2$V1, spV2, type = "n", ylim = range(c(spV2, xy)), xlab = "Cores", ylab = "Speedup",axes=FALSE) 
box()
grid(col = "black")
lines(data2$V1, spV2, col = "blue")
lines(data3$V1, spV3, col = "red")
lines(data2$V1, xy, lty=2, col = "green") 
points(data2$V1,spV2, pch=21, col = "blue")
points(data3$V1,spV3, pch=21, col = "red")
points(data2$V1,xy, pch=22, col = "green")

legend(200, 8, c("MSL","SkelGIS","ideal"), cex=1, col=c("blue","red","green"), pch=21:22, lty=1:2, bty="n");
axis(1, at=1, lab="1",cex.axis=1)
axis(1, at=2, lab="2",cex.axis=1)
axis(1, at=4, lab="4",cex.axis=1)
axis(1, at=8, lab="8",cex.axis=1)
axis(1, at=16, lab="16",cex.axis=1)
axis(1, at=32, lab="32",cex.axis=1)
axis(1, at=64, lab="64",cex.axis=1)
axis(1, at=128, lab="128",cex.axis=1)
axis(1, at=256, lab="256",cex.axis=1)
axis(2, at=1, lab="1",cex.axis=1)
axis(2, at=2, lab="2",cex.axis=1)
axis(2, at=4, lab="4",cex.axis=1)
axis(2, at=8, lab="8",cex.axis=1)
axis(2, at=16, lab="16",cex.axis=1)
axis(2, at=32, lab="32",cex.axis=1)
axis(2, at=64, lab="64",cex.axis=1)
axis(2, at=128, lab="128",cex.axis=1)
axis(2, at=256, lab="256",cex.axis=1)
```

Independent speedups where each application has its own reference time :

```{r, echo=FALSE}
#calcul des speedups
xy = rep(0,length(data2$V1))
spV2 = rep(0,length(data2$V2))
for(i in 1:length(data2$V2)){
  spV2[i] = data2$V2[1]/data2$V2[i]
  xy[i] = data2$V1[i]/data2$V1[1]
}
spV3 = rep(0,length(data3$V2))
for(i in 1:length(data3$V2)){
  spV3[i] = data3$V2[1]/data3$V2[i]
  xy[i] = data3$V1[i]/data3$V1[1]
}

par(mar=c(4,4,0.5,0.5))
plot(data2$V1, spV2, type = "n", ylim = range(c(spV2, xy)), xlab = "Cores", ylab = "Speedup",axes=FALSE) 
box()
grid(col = "black")
lines(data2$V1, spV2, col = "blue")
lines(data3$V1, spV3, col = "red")
lines(data2$V1, xy, lty=2, col = "green") 
points(data2$V1,spV2, pch=21, col = "blue")
points(data3$V1,spV3, pch=21, col = "red")
points(data2$V1,xy, pch=22, col = "green")

legend(200, 8, c("MSL","SkelGIS","ideal"), cex=1, col=c("blue","red","green"), pch=21:22, lty=1:2, bty="n");
axis(1, at=1, lab="1",cex.axis=1)
axis(1, at=2, lab="2",cex.axis=1)
axis(1, at=4, lab="4",cex.axis=1)
axis(1, at=8, lab="8",cex.axis=1)
axis(1, at=16, lab="16",cex.axis=1)
axis(1, at=32, lab="32",cex.axis=1)
axis(1, at=64, lab="64",cex.axis=1)
axis(1, at=128, lab="128",cex.axis=1)
axis(1, at=256, lab="256",cex.axis=1)
axis(2, at=1, lab="1",cex.axis=1)
axis(2, at=2, lab="2",cex.axis=1)
axis(2, at=4, lab="4",cex.axis=1)
axis(2, at=8, lab="8",cex.axis=1)
axis(2, at=16, lab="16",cex.axis=1)
axis(2, at=32, lab="32",cex.axis=1)
axis(2, at=64, lab="64",cex.axis=1)
axis(2, at=128, lab="128",cex.axis=1)
axis(2, at=256, lab="256",cex.axis=1)
```

log2 of execution times :

```{r, echo=FALSE}
par(mar=c(4,4,0.5,0.5))
plot(log2(data2$V1), log2(data2$V2), type = "n", xlim = range(c(log2(data3$V1), log2(data2$V1))), xlab = "Cores", ylab = "Speedup",axes=FALSE) 
box()
grid(col = "black")
lines(log2(data2$V1), log2(data2$V2), col = "blue")
lines(log2(data3$V1), log2(data3$V2), col = "red")
points(log2(data2$V1),log2(data2$V2), pch=21, col = "blue")
points(log2(data3$V1), log2(data3$V2), pch=21, col = "red")

legend(0, 0, c("MSL","SkelGIS"), cex=1, col=c("blue","red"), pch=21:22, lty=1:2, bty="n");
axis(1, at=1, lab="1",cex.axis=1)
axis(1, at=2, lab="2",cex.axis=1)
axis(1, at=4, lab="4",cex.axis=1)
axis(1, at=log2(8), lab="8",cex.axis=1)
axis(1, at=log2(16), lab="16",cex.axis=1)
axis(1, at=log2(32), lab="32",cex.axis=1)
axis(1, at=log2(64), lab="64",cex.axis=1)
axis(1, at=log2(128), lab="128",cex.axis=1)
axis(1, at=log2(256), lab="256",cex.axis=1)
```

### 2016-01-20 ARTICLES/IJPP/results/task_scaling/500_200/max_mean.py

@@ -1,0 +1,44 @@
+import numpy as np
+import glob, os, string
+# os.chdir("./base")
+# filesearch = "base_MSL_500_200_128_"
+
+os.chdir("./fus")
+filesearch = "fus_MSL_500_200_128_"
+
+# os.chdir("./task16")
+# filesearch = "task16MSL_500_200_128_"
+
+
+vallist = []
+for file in glob.glob("*.o"):
+    if file.find(filesearch)>=0 :
+        print(file)
+        with open(file) as f:
+            lines = f.readlines()
+            listVal = []
+            for line in lines :
+                if line.find("Time TOTAL CALCULATION")>=0 :
+                    spl = string.split(line," : ")
+                    listVal.append(string.strip(spl[1],"\n"))
+            if listVal:
+                print(max(listVal))
+                vallist.append(max(listVal))
+medianarray = np.array(vallist).astype(np.float)
+median = np.median(medianarray)
+mean = np.mean(medianarray)
+print("****************************")
+print("Execution time average = "+str(mean))
+print("****************************")
+
+print("****************************")
+print("Execution time median = "+str(median))
+print("****************************")
+
+print("****************************")
+print("Execution time min = "+str(min(vallist)))
+print("****************************")
+
+print("****************************")
+print("Execution time max = "+str(max(vallist)))
+print("****************************")

### 2016-01-20 ARTICLES/IJPP/results/task_scaling/500_200/resultsF.py

@@ -1,0 +1,60 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [16,32,64,128,256,512,1024,2048]
+
+# fusion close
+timesMSL = [1.42,0.54,0.28,0.17,0.15,0.19,0.54,2.54]
+timesT2 = [1.45,0.57,0.29,0.18,0.12,0.23,1.02,6.15]
+
+# fusion spread
+timesT2spread = [2.27,0.96,0.5,0.28,0.17,0.26,1.02,5.8]
+
+itsMSL = []
+itsT2 = []
+itsT4 = []
+itsT2spread = []
+itsT4spread = []
+
+for i in timesMSL:
+    itsMSL.append(500/i)
+for i in timesT2:
+	if i!=0:
+		itsT2.append(500/i)
+	else :
+		itsT2.append(0)
+for i in timesT2spread:
+	if i!=0:
+		itsT2spread.append(500/i)
+	else :
+		itsT2spread.append(0)
+
+relatives_its2 = []
+for i in cores:
+    relatives_its2.append(500/(timesMSL[0]/(i/cores[0])))
+
+matplotlib.rcParams.update({'font.size': 16})
+
+#plt.xkcd()
+
+#plt.plot(cores, relatives_its2, 'b--', label='Ideal')
+plt.loglog(cores, itsMSL, 'b-+', label='MSL MPI', basex=2,basey=2)
+plt.loglog(cores, itsT2, 'r-*', label='MSL MPI + 2 threads close', basex=2,basey=2)
+plt.loglog(cores, itsT2spread, 'g-*', label='MSL MPI + 2 threads spread', basex=2,basey=2)
+
+legend = plt.legend(loc='lower right', shadow=True,prop={'size':12})
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('cores')
+plt.ylabel('iterations per second')
+pp = PdfPages('itpersecF.pdf')
+plt.savefig(pp, format='pdf', bbox_inches='tight')
+pp.close()
+
+plt.cla()
+plt.clf()
+plt.close()

### 2016-01-20 ARTICLES/IJPP/results/task_scaling/500_200/resultsWF.py

@@ -1,0 +1,124 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [16,32,64,128,256,512,1024,2048]
+
+# median 10 runs
+# without fusion close
+timesMSL = [2.35,1.05,0.54,0.3,0.2,0.19,0.24,0.49]
+timesT2 = [2.32,1.05,0.54,0.3,0.18,0.14,0.16,0.22]
+timesT4 = [2.8,1.32,0.68,0.37,0.23,0.14,0.13,0.22]
+timesT8 = [3.06,1.48,0.77,0.41,0.23,0.14,0.12,0.078]
+timesT16 = [5.7,2.9,1.49,0.8,0.47,0.3,0.24,0.3]
+
+#without fusion spread
+timesT2spread = [3.78,1.82,0.93,0.5,0.28,0.2,0.17,0.22]
+timesT4spread = [4.5,2.26,1.16,0.61,0.35,0.21,0.14,0.19]
+timesT8spread = [5.16,2.53,1.3,0.68,0.37,0.21,0.13,0.09]
+timesT16spread = [9.3,4.8,2.37,1.24,0.69,0.39,0.26,0.25]
+
+itsMSL = []
+itsT2 = []
+itsT4 = []
+itsT8 = []
+itsT16 = []
+itsT2spread = []
+itsT4spread = []
+itsT8spread = []
+itsT16spread = []
+
+for i in timesMSL:
+    itsMSL.append(500/i)
+for i in timesT2:
+	if i!=0:
+		itsT2.append(500/i)
+	else :
+		itsT2.append(0)
+for i in timesT4:
+	if i!=0:
+		itsT4.append(500/i)
+	else :
+		itsT4.append(0)
+for i in timesT8:
+	if i!=0:
+		itsT8.append(500/i)
+	else :
+		itsT8.append(0)
+for i in timesT16:
+	if i!=0:
+		itsT16.append(500/i)
+	else :
+		itsT16.append(0)
+
+
+for i in timesT2spread:
+	if i!=0:
+		itsT2spread.append(500/i)
+	else :
+		itsT2spread.append(0)
+for i in timesT4spread:
+	if i!=0:
+		itsT4spread.append(500/i)
+	else :
+		itsT4spread.append(0)
[…]

### 2016-01-20 ARTICLES/IJPP/results/weak_scaling/400/max_mean.py

@@ -1,0 +1,40 @@
+import numpy as np
+import glob, os, string
+os.chdir("./test")
+filesearch = "weakMSL_25600_500_8192_"
+
+# os.chdir("./SKMPI")
+# filesearch = "weakSK_25600_500_16384_"
+
+vallist = []
+for file in glob.glob("*.o"):
+    if file.find(filesearch)>=0 :
+        print(file)
+        with open(file) as f:
+            lines = f.readlines()
+            listVal = []
+            for line in lines :
+                if line.find("Time TOTAL CALCULATION")>=0 :
+                    spl = string.split(line," : ")
+                    listVal.append(string.strip(spl[1],"\n"))
+            if listVal:
+                print(max(listVal))
+                vallist.append(max(listVal))
+medianarray = np.array(vallist).astype(np.float)
+median = np.median(medianarray)
+mean = np.mean(medianarray)
+print("****************************")
+print("Execution time average = "+str(mean))
+print("****************************")
+
+print("****************************")
+print("Execution time median = "+str(median))
+print("****************************")
+
+print("****************************")
+print("Execution time min = "+str(min(vallist)))
+print("****************************")
+
+print("****************************")
+print("Execution time max = "+str(max(vallist)))
+print("****************************")

### 2016-01-20 ARTICLES/WOLFHPC15/results_curie/results/10K_5K/results.py

@@ -1,0 +1,49 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+
+cores = [256,512,1024,2048,4096,8192,16384]
+timesMSL = [979.18,565,]
+timesSK = [1472.16,783.46,396.73,203.97,107.81,58.68,31.95]
+timesSK3 = []
+
+spdMSL = []
+spdSK = []
+
+for i in timesMSL:
+    spdMSL.append(timesMSL[0]/i)
+for i in timesSK:
+    spdSK.append(timesMSL[0]/i)
+    
+relatives = []
+for i in cores:
+    relatives.append(i/cores[0])
+
+plt.plot(cores, relatives, 'g--', label='Ideal')
+plt.plot(cores, spdMSL, 'b-+', label='MSL')
+plt.plot(cores, spdSK, 'r-x', label='SkelGIS')
+legend = plt.legend(loc='upper left', shadow=True)
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('cores')
+plt.ylabel('speedup')
+pp = PdfPages('speedups.pdf')
+plt.savefig(pp, format='pdf', bbox_inches='tight')
+pp.close()
+
+plt.cla()
+plt.clf()
+plt.close()
+
+plt.loglog(cores, timesMSL,'b-+',label='MSL', basex=2,basey=2)
+plt.loglog(cores, timesSK,'r-x',label='SkelGIS', basex=2,basey=2)
+legend = plt.legend(shadow=True)
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('log cores')
+plt.ylabel('log execution time')
+pp2 = PdfPages('logtimes.pdf')
+plt.savefig(pp2, format='pdf', bbox_inches='tight')
+pp2.close()
+
+plt.cla()
+plt.clf()
+plt.close()

### 2016-01-20 ARTICLES/WOLFHPC15/results_curie/results/5K_500/results.py

@@ -1,0 +1,49 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+
+cores = [8,16,32,64,128,256,512]
+timesMSL = [818.69,404.02,197.106,98.95,49.07,29.44,19.88,]
+timesSK = [1073.31,542.26,273.65,138.87,80.11,44.32,19.69]#,11.15,5.75,3.77]
+#timesSK = [674.75,342.73,173.31,87.64,47.76,36.45,21.15]#,14.18,18.67,17,03]
+
+spdMSL = []
+spdSK = []
+
+for i in timesMSL:
+    spdMSL.append(timesMSL[0]/i)
+for i in timesSK:
+    spdSK.append(timesSK[0]/i)
+    
+relatives = []
+for i in cores:
+    relatives.append(i/cores[0])
+
+plt.plot(cores, relatives, 'g--', label='Ideal')
+plt.plot(cores, spdMSL, 'b-+', label='MSL')
+plt.plot(cores, spdSK, 'r-x', label='SkelGIS')
+legend = plt.legend(loc='upper left', shadow=True)
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('cores')
+plt.ylabel('speedup')
+pp = PdfPages('speedups.pdf')
+plt.savefig(pp, format='pdf', bbox_inches='tight')
+pp.close()
+
+plt.cla()
+plt.clf()
+plt.close()
+
+plt.loglog(cores, timesMSL,'b-+',label='MSL', basex=2,basey=2)
+plt.loglog(cores, timesSK,'r-x',label='SkelGIS', basex=2,basey=2)
+legend = plt.legend(shadow=True)
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('log cores')
+plt.ylabel('log execution time')
+pp2 = PdfPages('logtimes.pdf')
+plt.savefig(pp2, format='pdf', bbox_inches='tight')
+pp2.close()
+
+plt.cla()
+plt.clf()
+plt.close()

### 2016-01-20 ARTICLES/WOLFHPC15/results_g5k/SW_10K_2K/results.py

@@ -1,0 +1,82 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [16,32,64,128,256]
+timesMSL = [9035.94,4519.9,2390.61,1300.08,693.294]
+timesSK = [13692.4,6946.07,3235.88,1861.24,892.882]
+
+spdMSL = []
+spdSK = []
+
+itsMSL = []
+itsSK = []
+
+for i in timesMSL:
+    spdMSL.append(timesMSL[0]/i)
+    itsMSL.append(500/i)
+for i in timesSK:
+    spdSK.append(timesMSL[0]/i)
+    itsSK.append(500/i)
+    
+relatives = []
+relatives_its = []
+for i in cores:
+    relatives.append(i/cores[0])
+    relatives_its.append(500/(timesMSL[0]/(i/cores[0])))
+
+matplotlib.rcParams.update({'font.size': 16})
+
+#plt.xkcd()
+
+plt.plot(cores, relatives, 'g--', label='Ideal')
+plt.plot(cores, spdMSL, 'b-+', label='MSL')
+plt.plot(cores, spdSK, 'r-x', label='SkelGIS')
+legend = plt.legend(loc='upper left', shadow=True)
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('cores')
+plt.ylabel('speedup')
+pp = PdfPages('speedups.pdf')
+plt.savefig(pp, format='pdf', bbox_inches='tight')
+pp.close()
+
+plt.cla()
+plt.clf()
+plt.close()
+
+#plt.xkcd()
+
+plt.plot(cores, relatives_its, 'g--', label='Ideal')
+plt.plot(cores, itsMSL, 'b-+', label='MSL')
+plt.plot(cores, itsSK, 'r-x', label='SkelGIS')
+legend = plt.legend(loc='upper left', shadow=True)
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('cores')
+plt.ylabel('iterations per second')
+pp = PdfPages('itpersec.pdf')
+plt.savefig(pp, format='pdf', bbox_inches='tight')
+pp.close()
+
+plt.cla()
+plt.clf()
+plt.close()
+
+#plt.xkcd()
+
[…]

### 2016-01-20 ARTICLES/WOLFHPC15/results_g5k/SW_10K_500/results.py

@@ -1,0 +1,82 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [16,32,64,128,256]
+timesMSL = [2089.33,1155.82,601.274,286.427,154.084]
+timesSK = [3284.03,1775.92,789.858,410.984,215.731]
+
+spdMSL = []
+spdSK = []
+
+itsMSL = []
+itsSK = []
+
+for i in timesMSL:
+    spdMSL.append(timesMSL[0]/i)
+    itsMSL.append(500/i)
+for i in timesSK:
+    spdSK.append(timesMSL[0]/i)
+    itsSK.append(500/i)
+    
+relatives = []
+relatives_its = []
+for i in cores:
+    relatives.append(i/cores[0])
+    relatives_its.append(500/(timesMSL[0]/(i/cores[0])))
+
+matplotlib.rcParams.update({'font.size': 16})
+
+#plt.xkcd()
+
+plt.plot(cores, relatives, 'g--', label='Ideal')
+plt.plot(cores, spdMSL, 'b-+', label='MSL')
+plt.plot(cores, spdSK, 'r-x', label='SkelGIS')
+legend = plt.legend(loc='upper left', shadow=True)
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('cores')
+plt.ylabel('speedup')
+pp = PdfPages('speedups.pdf')
+plt.savefig(pp, format='pdf', bbox_inches='tight')
+pp.close()
+
+plt.cla()
+plt.clf()
+plt.close()
+
+#plt.xkcd()
+
+plt.plot(cores, relatives_its, 'g--', label='Ideal')
+plt.plot(cores, itsMSL, 'b-+', label='MSL')
+plt.plot(cores, itsSK, 'r-x', label='SkelGIS')
+legend = plt.legend(loc='upper left', shadow=True)
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('cores')
+plt.ylabel('iterations per second')
+pp = PdfPages('itpersec.pdf')
+plt.savefig(pp, format='pdf', bbox_inches='tight')
+pp.close()
+
+plt.cla()
+plt.clf()
+plt.close()
+
+#plt.xkcd()
+
[…]

### 2016-01-20 ARTICLES/WOLFHPC15/results_g5k/SW_5K_500/results.py

@@ -1,0 +1,82 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [8,16,32,64,128,256]
+timesMSL = [998.1,544.87,285.73,139.8,76.45,48.05]
+timesSK = [1324.9,638.5,371.62,197.32,117.047,64.96]
+
+spdMSL = []
+spdSK = []
+
+itsMSL = []
+itsSK = []
+
+for i in timesMSL:
+    spdMSL.append(timesMSL[0]/i)
+    itsMSL.append(500/i)
+for i in timesSK:
+    spdSK.append(timesMSL[0]/i)
+    itsSK.append(500/i)
+    
+relatives = []
+relatives_its = []
+for i in cores:
+    relatives.append(i/cores[0])
+    relatives_its.append(500/(timesMSL[0]/(i/cores[0])))
+
+matplotlib.rcParams.update({'font.size': 16})
+
+#plt.xkcd()
+
+plt.plot(cores, relatives, 'g--', label='Ideal')
+plt.plot(cores, spdMSL, 'b-+', label='MSL')
+plt.plot(cores, spdSK, 'r-x', label='SkelGIS')
+legend = plt.legend(loc='upper left', shadow=True)
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('cores')
+plt.ylabel('speedup')
+pp = PdfPages('speedups.pdf')
+plt.savefig(pp, format='pdf', bbox_inches='tight')
+pp.close()
+
+plt.cla()
+plt.clf()
+plt.close()
+
+#plt.xkcd()
+
+plt.plot(cores, relatives_its, 'g--', label='Ideal')
+plt.plot(cores, itsMSL, 'b-+', label='MSL')
+plt.plot(cores, itsSK, 'r-x', label='SkelGIS')
+legend = plt.legend(loc='upper left', shadow=True)
+plt.grid(True, which="both", linestyle="dotted")
+plt.xlabel('cores')
+plt.ylabel('iterations per second')
+pp = PdfPages('itpersec.pdf')
+plt.savefig(pp, format='pdf', bbox_inches='tight')
+pp.close()
+
+plt.cla()
+plt.clf()
+plt.close()
+
+#plt.xkcd()
+
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/assemblybuilding.h

@@ -1,0 +1,44 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_ASSEMBLYBUILDING_H
+#define LLCMCPP_ASSEMBLYBUILDING_H
+
+namespace llcmcpp {
+
+class AssemblyBuilding
+{
+public:
+	virtual void instantiation () = 0;
+	
+	virtual void configuration () = 0;
+	
+	virtual void activation () = 0;
+	
+	virtual void execution () = 0;
+	
+	virtual void deactivation () = 0;
+	
+	virtual void destruction () = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_ASSEMBLYBUILDING_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/clusterinstantiationmanagement.h

@@ -1,0 +1,45 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_CLUSTERINSTANTIATIONMANAGEMENT_H
+#define LLCMCPP_CLUSTERINSTANTIATIONMANAGEMENT_H
+
+#include <string>
+
+#include "hostid.h"
+
+namespace llcmcpp {
+
+class ClusterInstantiationManagement
+{
+public:
+	/*! \brief Creates a new component instance in the cluser.
+	 * \param host the host in the cluster where the instance should be created
+	 * \param name the name (identifier) of the instance
+	 * \param type the type of the instance
+	 * \param from the library to load the instance from, defaults to $LD_LIBRARY_PATH+"lib"+type+".so"
+	 * \post id.state == CREATED
+	 */
+	virtual void instantiate ( HostId host, std::string id, std::string type, std::string from = "" ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_CLUSTERINSTANTIATIONMANAGEMENT_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/clusterplacementmanagement.h

@@ -1,0 +1,46 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_CLUSTERPLACEMENTMANAGEMENT_H
+#define LLCMCPP_CLUSTERPLACEMENTMANAGEMENT_H
+
+#include <set>
+#include <string>
+
+#include "hostid.h"
+
+namespace llcmcpp {
+
+/*! \brief Manages a (cluster local) namespace of component instances.
+ */
+class ClusterPlacementManagement
+{
+public:
+	/*! \brief Retreives the host on which a given instance is hosted.
+	 * \param instance the name (identifier) of the instance to look for
+	 * \returns the identifier of the host for the instance
+	 * \throws std::out_of_range if the instance doesn't exist
+	 */
+	virtual HostId get_rank ( std::string instance ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_CLUSTERPLACEMENTMANAGEMENT_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/corbaaccess.h

@@ -1,0 +1,40 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_CORBAACCESS_H
+#define LLCMCPP_CORBAACCESS_H
+
+#include <string>
+
+#include <omniORB4/CORBA.h>
+
+#include "endpoint.h"
+
+namespace llcmcpp {
+
+class CorbaAccess
+{
+public:
+	virtual CORBA::Object_ptr get ( std::string hostw, EndPoint property ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_CORBAACCESS_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/cppmanagement.h

@@ -1,0 +1,137 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_CPPMANAGEMENT_H
+#define LLCMCPP_CPPMANAGEMENT_H
+
+#include <complex>
+
+#include <stdint.h>
+
+#include "endpoint.h"
+
+namespace llcmcpp {
+
+/*! \brief manages the local (C++) aspects of an LLCMc++ assembly
+ */
+class CppManagement
+{
+public:
+	/*! \brief connects two components by a C++ use/provide interaction
+	 * \param user the user endpoint
+	 * \param provider the provider endpoint
+	 * \pre user.state == created && provider.state == created
+	 */
+	virtual void connect ( EndPoint user, EndPoint provider ) = 0;
+	
+	/*! \brief Configures an input property of a component using a system value.
+	 * \param property the property to configure
+	 * \param value_name the name of the system value to use
+	 * \pre id.state == created
+	 */
+	virtual void system_configure ( EndPoint property, std::string value_name ) = 0;
+	
+	/*! \brief Configures an input property of a component.
+	 * \param property the property to configure
+	 * \param provider the value to set for the property
+	 * \pre id.state == created
+	 */
+	virtual void configure ( EndPoint property, int8_t value ) = 0;
+	
+	/*! \brief Configures an input property of a component.
+	 * \param property the property to configure
+	 * \param provider the value to set for the property
+	 * \pre id.state == created
+	 */
+	virtual void configure ( EndPoint property, uint8_t value ) = 0;
+	
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/cppmanagement2.h

@@ -1,0 +1,133 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_CPPMANAGEMENT2_H
+#define LLCMCPP_CPPMANAGEMENT2_H
+
+#include <complex>
+
+#include <stdint.h>
+
+#include "endpoint.h"
+
+namespace llcmcpp {
+
+/*! \brief manages the local (C++) aspects of an LLCMc++ assembly
+ * Unlike CppManagement, this interface does no handle the starting of
+ * component instances that should be done through the start interface.
+ */
+class CppManagement2
+{
+public:
+	/*! \brief connects two components by a C++ use/provide interaction
+	 * \param user the user endpoint
+	 * \param provider the provider endpoint
+	 * \pre user.state == created && provider.state == created
+	 */
+	virtual void connect ( EndPoint user, EndPoint provider ) = 0;
+	
+	/*! \brief Configures an input property of a component using a system value.
+	 * \param property the property to configure
+	 * \param value_name the name of the system value to use
+	 * \pre id.state == created
+	 */
+	virtual void system_configure ( EndPoint property, std::string value_name ) = 0;
+	
+	/*! \brief Configures an input property of a component.
+	 * \param property the property to configure
+	 * \param provider the value to set for the property
+	 * \pre id.state == created
+	 */
+	virtual void configure ( EndPoint property, int8_t value ) = 0;
+	
+	/*! \brief Configures an input property of a component.
+	 * \param property the property to configure
+	 * \param provider the value to set for the property
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/distributedinstantiationmanagement.h

@@ -1,0 +1,36 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_DISTRIBUTEDINSTANTIATIONMANAGEMENT_H
+#define LLCMCPP_DISTRIBUTEDINSTANTIATIONMANAGEMENT_H
+
+#include <string>
+
+namespace llcmcpp {
+
+class DistributedInstantiationManagement
+{
+public:
+	virtual void instantiate ( std::string host, std::string id, std::string type, std::string from = "" ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_DISTRIBUTEDINSTANTIATIONMANAGEMENT_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/distributedplacementmanagement.h

@@ -1,0 +1,43 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_DISTRIBUTEDPLACEMENTMANAGEMENT_H
+#define LLCMCPP_DISTRIBUTEDPLACEMENTMANAGEMENT_H
+
+#include <set>
+#include <string>
+
+namespace llcmcpp {
+
+/*! \brief Manages a (distributed) namespace of component instances.
+ */
+class DistributedPlacementManagement
+{
+public:
+	/*! \brief Retreives the host on which a given instance is hosted.
+	 * \param instance the name (identifier) of the instance to look for
+	 * \returns the identifier of the host for the instance
+	 */
+	virtual std::string get_host ( std::string instance ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_DISTRIBUTEDPLACEMENTMANAGEMENT_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/endpoint.h

@@ -1,0 +1,37 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_ENDPOINT_H
+#define LLCMCPP_ENDPOINT_H
+
+#include <string>
+
+namespace llcmcpp {
+
+struct EndPoint
+{
+	EndPoint ( std::string instance="", std::string port="" ): instance(instance), port(port) {}
+	operator bool () const { return !instance.empty() && !port.empty(); }
+	std::string instance;
+	std::string port;
+};
+
+}
+
+#endif // LLCMCPP_ENDPOINT_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/go.h

@@ -1,0 +1,35 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_GO_H
+#define LLCMCPP_GO_H
+
+namespace llcmcpp
+{
+
+class Go
+{
+public:
+	virtual void go () = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_GO_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/gridinstantiationmanagement.h

@@ -1,0 +1,38 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_GRIDINSTANTIATIONMANAGEMENT_H
+#define LLCMCPP_GRIDINSTANTIATIONMANAGEMENT_H
+
+#include <string>
+
+#include "hostid.h"
+
+namespace llcmcpp {
+
+class GridInstantiationManagement
+{
+public:
+	virtual void instantiate ( std::string cluster, HostId host, std::string id, std::string type, std::string from = "" ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_GRIDINSTANTIATIONMANAGEMENT_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/host.h

@@ -1,0 +1,69 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_HOST_H
+#define LLCMCPP_HOST_H
+
+#include <stdint.h>
+#include <string>
+
+namespace llcmcpp
+{
+
+/*! \brief a Host manages a set of component instances in one process.
+ */
+class Host
+{
+public:
+	/*! \brief a opaque handle identifying component instances handled by this host
+	 */
+	typedef uint16_t ComponentInstance;
+	
+	/*! \brief Instanciates a component
+	 * \param type the type of the component to instanciate
+	 * \param library path to the library from which to load the component, defaults to $LD_LIBRARY_PATH+"lib"+type+".so"
+	 * \return an opaque handle on the component instance
+	 */
+	virtual ComponentInstance instantiate ( std::string type, std::string library = "" ) = 0;
+	
+	/*! \brief Destroys a given component instance
+	 * Destroys an instance: i.e. calls its destructor and deallocates the memory.
+	 * \param instance the component instance identifier
+	 */
+	virtual void destroy ( ComponentInstance instance ) = 0;
+
+	/*! \brief Gets a pointer to the value of a component's property
+	 * \param instance the component instance identifier
+	 * \param propname the name on the property
+	 * \return the value of the property
+	 */
+	virtual void* get ( ComponentInstance instance, std::string propname ) = 0;
+	
+	/*! \brief Sets the value of a component's property
+	 * \param instance the component instance identifier
+	 * \param propname the name on the property
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/host2.h

@@ -1,0 +1,41 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_HOST2_H
+#define LLCMCPP_HOST2_H
+
+#include "host.h"
+
+namespace llcmcpp {
+
+class Host2:
+		virtual public Host
+{
+public:
+	/*! \brief Preloads a component type.
+	 * \param type the type of the component to preload
+	 * \param library path to the library from which to load the component, defaults to $LD_LIBRARY_PATH+"lib"+type+".so"
+	 */
+	virtual void preload ( std::string type, std::string library = "" ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_HOST2_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/host3.h

@@ -1,0 +1,76 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_HOST3_H
+#define LLCMCPP_HOST3_H
+
+#include <stdint.h>
+#include <string>
+
+namespace llcmcpp
+{
+
+/*! \brief a Host3 manages a set of up to 2^64 component instances in one process.
+ */
+class Host3
+{
+public:
+	/*! \brief a opaque handle identifying component instances managed by this host
+	 */
+	typedef uint64_t ComponentInstance;
+
+	/*! \brief Instanciates a component
+	 * \param type the type of the component to instanciate
+	 * \param library path to the library from which to load the component, defaults to $LD_LIBRARY_PATH+"lib"+type+".so"
+	 * \return an opaque handle on the component instance
+	 */
+	virtual ComponentInstance instantiate ( std::string type, std::string library = "" ) = 0;
+
+	/*! \brief Destroys a given component instance
+	 * Destroys an instance: i.e. calls its destructor and deallocates the memory.
+	 * \param instance the component instance identifier
+	 */
+	virtual void destroy ( ComponentInstance instance ) = 0;
+
+	/*! \brief Gets a pointer to the value of a component's property
+	 * \param instance the component instance identifier
+	 * \param propname the name on the property
+	 * \return the value of the property
+	 */
+	virtual void* get ( ComponentInstance instance, std::string propname ) = 0;
+
+	/*! \brief Sets the value of a component's property
+	 * \param instance the component instance identifier
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/hostid.h

@@ -1,0 +1,32 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_HOSTID_H
+#define LLCMCPP_HOSTID_H
+
+#include <stdint.h>
+
+namespace llcmcpp {
+
+//! \brief An identifier for hosts inside an (homogeneous) cluser.
+typedef uint16_t HostId;
+
+}
+
+#endif // LLCMCPP_HOSTID_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/initializer.h

@@ -1,0 +1,34 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_INITIALIZATIONMANAGEMENT_H
+#define LLCMCPP_INITIALIZATIONMANAGEMENT_H
+
+namespace llcmcpp {
+
+class Initializer
+{
+public:
+	virtual void initialize ( int* argc, char*** argv ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_INITIALIZATIONMANAGEMENT_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/llcmcommon.h

@@ -1,0 +1,104 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_LLCMCOMMON_H
+#define LLCMCPP_LLCMCOMMON_H
+
+#include <iostream>
+#include <cstdlib>
+#include <map>
+#include <string>
+
+#include "component.h"
+
+namespace llcmcpp {
+
+namespace impl {
+
+template < class Cmp >
+struct Setter
+{
+	virtual void operator() ( Cmp*, void* ) = 0;
+	
+};
+
+template < class Cmp >
+struct Getter
+{
+	virtual void* operator() ( Cmp* ) = 0;
+	
+};
+
+template < class Cmp >
+class BasicComponent:
+		public Component
+{
+public:
+	Cmp instance;
+	
+	const char* name;
+	
+	BasicComponent (): name(0) {}
+	
+	BasicComponent ( const char* name ): name(name) {}
+	
+	BasicComponent ( const char* name, const Cmp& inst  ): name(name), instance(inst) {}
+	
+	std::map< std::string, Setter< Cmp >* > input_properties;
+	
+	std::map< std::string, Getter< Cmp >* > output_properties;
+	
+	void* get(const char* propname)
+	{
+		typename std::map< std::string, Getter< Cmp >* >::iterator it = output_properties.find(std::string(propname));
+		if ( it == output_properties.end() ) {
+			if ( std::string("lifecycle") != propname ) {
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/llcmcorba.h

@@ -1,0 +1,155 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_LLCMCORBA_H
+#define LLCMCPP_LLCMCORBA_H
+
+#include <string>
+
+#include <omniORB4/CORBA.h>
+
+#include "llcmcommon.h"
+
+namespace llcmcpp {
+
+namespace impl {
+
+template < class Cmp, class C >
+struct SetterCorbaUse:
+		public Setter< Cmp >
+{
+	SetterCorbaUse ( void (Cmp::*mth) ( typename C::_ptr_type ) ):
+			m_mth(mth)
+	{}
+	
+	virtual void operator() ( Cmp* c, void* v )
+	{
+		(c->*m_mth)(C::_narrow(static_cast< ::CORBA::Object_ptr >(v)));
+	}
+	
+	void (Cmp::*m_mth) ( typename C::_ptr_type );
+	
+};
+
+template < class Cmp, class C >
+struct AttributeCorbaUse:
+		public Setter< Cmp >
+{
+	AttributeCorbaUse ( typename C::_ptr_type Cmp::*dat ):
+			m_dat(dat)
+	{}
+	
+	virtual void operator() ( Cmp* c, void* v )
+	{
+		c->*m_dat = (C::_narrow(static_cast< ::CORBA::Object_ptr >(v)));
+	}
+	
+	typename C::_ptr_type Cmp::*m_dat;
+	
+};
+
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/llcmcpp.h

@@ -1,0 +1,35 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_LLCMCPP_H
+#define LLCMCPP_LLCMCPP_H
+
+#include "component.h"
+
+#warning The <llcmcpp.h> header is deprecated, consider using <llcmcpp3.h> instead
+
+#define LLCMCPP_CREATE(ctype) \
+extern "C" { \
+llcmcpp::Component* _llcmcpp_create_##ctype() { \
+	return new ctype; \
+} \
+}
+
+
+#endif // LLCMCPP_LLCMCPP_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/llcmcpp2.h

@@ -1,0 +1,50 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_LLCMCPP2_H
+#define LLCMCPP_LLCMCPP2_H
+
+#include <string>
+
+#include "llcmcpp3.h"
+
+namespace llcmcpp {
+
+namespace impl {
+
+template < class Cmp, class C >
+void cpp_use ( BasicComponent< Cmp >* cmp, std::string name, C Cmp::* dat )
+{
+	property< Cmp, C >(cmp, name, dat);
+}
+  
+}
+
+}
+
+#define LUSE(type, name) \
+	::llcmcpp::impl::cpp_use< Cmp, type >(cmp, #name, &Cmp::name);
+
+#define LPROVIDE(type, name) \
+	::llcmcpp::impl::cpp_provide< Cmp, type >(cmp, #name);
+
+#define LPROVIDE2(type, name) \
+	::llcmcpp::impl::cpp_provide< Cmp, type >(cmp, #name, &Cmp::name);
+
+#endif // LLCMCPP_LLCMCPP2_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/llcmcpp3.h

@@ -1,0 +1,202 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_LLCMCPP3_H
+#define LLCMCPP_LLCMCPP3_H
+
+#include <string>
+
+#include "llcmcommon.h"
+
+namespace llcmcpp {
+
+namespace impl {
+
+template < class Cmp, class C >
+struct SetterCppUse:
+		public Setter< Cmp >
+{
+	SetterCppUse ( void (Cmp::*mth) ( C* ) ):
+			m_mth(mth)
+	{}
+	
+	virtual void operator() ( Cmp* c, void* v )
+	{
+		(c->*m_mth)(static_cast< C* >(v));
+	}
+	
+	void (Cmp::*m_mth) ( C* );
+	
+};
+
+template < class Cmp, class C >
+struct AttributeCppUse:
+		public Setter< Cmp >
+{
+	AttributeCppUse ( C* Cmp::*dat ):
+			m_dat(dat)
+	{}
+	
+	virtual void operator() ( Cmp* c, void* v )
+	{
+		c->*m_dat = static_cast< C* >(v);
+	}
+	
+	C* Cmp::*m_dat;
+	
+};
+
+template < class Cmp, class C >
+struct SetterProperty:
+		public Setter< Cmp >
+{
+	SetterProperty ( void (Cmp::*mth) ( C ) ):
+			m_mth(mth)
+	{}
+	
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/llcmf.h

@@ -1,0 +1,28 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_LLCMF_H
+#define LLCMCPP_LLCMF_H
+
+class fortranitf {
+ public:
+};
+
+
+#endif // LLCMCPP_LLCMF_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/llcmmpi.h

@@ -1,0 +1,80 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_LLCMMPI_H
+#define LLCMCPP_LLCMMPI_H
+
+#include <mpi.h>
+
+#include "llcmcommon.h"
+
+#include "component.h"
+
+namespace llcmcpp {
+
+namespace impl {
+
+template < class Cmp >
+struct SetterMpiComm:
+		public Setter< Cmp >
+{
+	SetterMpiComm ( void (Cmp::*mth) ( MPI_Comm ) ):
+			m_mth(mth)
+	{}
+	
+	virtual void operator() ( Cmp* c, void* v )
+	{
+		(c->*m_mth)(*static_cast< MPI_Comm* >(v));
+	}
+	
+	void (Cmp::*m_mth) ( MPI_Comm );
+	
+};
+
+template < class Cmp >
+struct AttributeMpiComm:
+		public Setter< Cmp >
+{
+	AttributeMpiComm ( MPI_Comm Cmp::*dat ):
+			m_dat(dat)
+	{}
+	
+	virtual void operator() ( Cmp* c, void* v )
+	{
+		c->*m_dat = *static_cast< MPI_Comm* >(v);
+	}
+	
+	MPI_Comm Cmp::*m_dat;
+	
+};
+
+template < class Cmp >
+void mpi_comm ( BasicComponent< Cmp >* cmp, std::string name, MPI_Comm Cmp::*dat )
+{
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/mpiaccess.h

@@ -1,0 +1,40 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_MPIACCESS_H
+#define LLCMCPP_MPIACCESS_H
+
+#include <list>
+
+#include <mpi.h>
+
+#include "hostid.h"
+
+namespace llcmcpp {
+
+class MpiAccess
+{
+public:
+	virtual MPI_Comm communicator ( std::list< HostId > hosts ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_MPIACCESS_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/mpimanagement.h

@@ -1,0 +1,45 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_MPIMANAGEMENT_H
+#define LLCMCPP_MPIMANAGEMENT_H
+
+#include <list>
+
+#include "endpoint.h"
+
+namespace llcmcpp {
+
+/*! \brief manages the cluster-wide (MPI) aspects of an LLCMc++ assembly
+ */
+class MpiManagement
+{
+public:
+	/*! \brief connects some components with an MPI communicator.
+	 * \warning There can be at most one instance so connected by process.
+	 * \param group the endpoints to be configured with the communicator
+	 * \pre each(endpoint:group) { endpoint.state == created }
+	 */
+	virtual void mpi_connect ( std::list< EndPoint > group ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_MPIMANAGEMENT_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/namespacemanagement.h

@@ -1,0 +1,42 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_NAMESPACEMANAGEMENT_H
+#define LLCMCPP_NAMESPACEMANAGEMENT_H
+
+#include <string>
+#include <set>
+
+namespace llcmcpp {
+
+/*! \brief Manages a (process local) namespace of component instances.
+ */
+class NamespaceManagement
+{
+public:
+	/*! \brief Lists the component instances.
+	 * \returns the set of component instances existing in this namespace
+	 */
+	virtual const std::set< std::string >& instances () = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_NAMESPACEMANAGEMENT_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/print.h

@@ -1,0 +1,45 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_PRINT_H
+#define LLCMCPP_PRINT_H
+
+#include <string>
+#include <iostream>
+
+namespace llcmcpp
+{
+
+class Print
+{
+public:
+	virtual void printout (const ::std::string&) = 0;
+	virtual void printlnout (const ::std::string&) = 0;
+	virtual void printerr (const ::std::string&) = 0;
+	virtual void printlnerr (const ::std::string&) = 0;
+};
+
+}
+
+#define PRINTOUT(port, msg)   { std::stringstream str; str << msg; if (port) { port->printout(str.str());   } else { std::cout<<str.str(); }}
+#define PRINTLNOUT(port, msg) { std::stringstream str; str << msg; if (port) { port->printlnout(str.str()); } else { std::cout<<str.str() << "\n"; }}
+#define PRINTERR(port, msg)   { std::stringstream str; str << msg; if (port) { port->printerr(str.str());   } else { std::cerr<<str.str(); }}
+#define PRINTLNERR(port, msg) { std::stringstream str; str << msg; if (port) { port->printlnerr(str.str()); } else { std::cerr<<str.str() << "\n"; }}
+
+#endif // LLCMCPP_PRINT_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/rawcomponentmanagement.h

@@ -1,0 +1,58 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_RAWCOMPONENTMANAGEMENT_H
+#define LLCMCPP_RAWCOMPONENTMANAGEMENT_H
+
+#include <string>
+
+#include "endpoint.h"
+
+namespace llcmcpp {
+
+/*! \brief Provides a raw access to some component instances.
+ */
+class RawComponentManagement
+{
+public:
+	/*! \brief Destroys a component instance.
+	 * \param name the name (identifier) of the instance
+	 * \pre id.state == DESTROYABLE
+	 */
+	virtual void destroy ( std::string name ) = 0;
+	
+	/*! \brief Provides raw acess to components input properties.
+	 * \param property the property to access
+	 * \param value a pointer to the value to set for the property, this should have the exact type expected by the component.
+	 * \pre id.state == CREATED
+	 */
+	virtual void set ( EndPoint property, void* value ) = 0;
+	
+	/*! \brief Provides raw acess to components output properties.
+	 * \param property the property to access
+	 * \returns a pointer to the value of the property
+	 * \pre id.state == CREATED | CONFIGURED | ACTIVATED
+	 */
+	virtual void* get ( EndPoint property ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_RAWCOMPONENTMANAGEMENT_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/systemproperty.h

@@ -1,0 +1,36 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_SYSTEMPROPERTY_H
+#define LLCMCPP_SYSTEMPROPERTY_H
+
+#include <string>
+
+namespace llcmcpp {
+
+class SystemProperty
+{
+public:
+	virtual void set_property ( std::string name, void* value ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_SYSTEMPROPERTY_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/include/typesetter.h

@@ -1,0 +1,38 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef LLCMCPP_TYPESETTER_H
+#define LLCMCPP_TYPESETTER_H
+
+#include <string>
+
+#include "endpoint.h"
+
+namespace llcmcpp {
+
+class TypeSetter
+{
+public:
+	virtual void configure ( EndPoint property, std::string type, std::string value ) = 0;
+	
+};
+
+}
+
+#endif // LLCMCPP_TYPESETTER_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/AssemblyBuilder/assemblybuilder.cpp

@@ -1,0 +1,478 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <complex>
+#include <iostream>
+#include <list>
+#include <map>
+#include <string>
+#include <set>
+
+#include <tinyxml.h>
+
+#include <assemblybuilding.h>
+#include <clusterinstantiationmanagement.h>
+#include <collective.h>
+#include <corbamanagement.h>
+#include <cppmanagement.h>
+#include <distributedinstantiationmanagement.h>
+#include <endpoint.h>
+#include <go.h>
+#include <gridinstantiationmanagement.h>
+#include <lifecycle.h>
+#include <llcmcpp3.h>
+#include <localinstantiationmanagement.h>
+#include <mpimanagement.h>
+#include <rmimanagement.h>
+#include <typesetter.h>
+
+using ::std::cerr;
+using ::std::complex;
+using ::std::endl;
+using ::std::list;
+using ::std::make_pair;
+using ::std::map;
+using ::std::multimap;
+using ::std::pair;
+using ::std::set;
+using ::std::string;
+
+namespace llcmcpp {
+
+namespace impl {
+
+/*! \brief A component that stores the instruction for assembly creation and plays it when required.
+ */
+class AssemblyBuilder:
+		virtual public AssemblyBuilding,
+		virtual public ClusterInstantiationManagement,
+		virtual public CorbaManagement,
+		virtual public CppManagement,
+		virtual public DistributedInstantiationManagement,
+		virtual public Go,
+		virtual public GridInstantiationManagement,
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/AssemblyBuilder2/assemblybuilder2.cpp

@@ -1,0 +1,176 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <string>
+#include <vector>
+
+#include <assemblybuilding.h>
+#include <collective.h>
+#include <go.h>
+#include <lifecycle.h>
+#include <llcmcpp3.h>
+
+namespace llcmcpp {
+
+namespace impl {
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::string;
+using ::std::vector;
+
+class AssemblyBuilder2:
+		virtual public AssemblyBuilding,
+		virtual public Go
+{
+public:
+	Collective* collective_access;
+	
+	LifeCycle* lifecycle_management;
+	
+	void instantiations ( Go* instantiation )
+	{
+		m_instantiations.push_back(instantiation);
+	}
+	
+	vector< Go* > m_instantiations;
+	
+	void configurations ( Go* configuration )
+	{
+		m_configurations.push_back(configuration);
+	}
+	
+	vector< Go* > m_configurations;
+	
+	void executions ( Go* execution )
+	{
+		m_executions.push_back(execution);
+	}
+	
+	vector< Go* > m_executions;
+	
+	AssemblyBuilder2 ():
+			collective_access(0),
+			lifecycle_management(0)
+	{}
+	
+	virtual void instantiation ()
+	{
+#ifndef NDEBUG
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/Cad3Handler/cad3handler.cpp

@@ -1,0 +1,372 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <list>
+#include <map>
+#include <string>
+
+#include <tinyxml.h>
+
+#include <clusterinstantiationmanagement.h>
+#include <corbamanagement.h>
+#include <cppmanagement.h>
+#include <distributedinstantiationmanagement.h>
+#include <endpoint.h>
+#include <gridinstantiationmanagement.h>
+#include <lifecycle.h>
+#include <llcmcpp3.h>
+#include <localinstantiationmanagement.h>
+#include <mpimanagement.h>
+#include <rmimanagement.h>
+#include <typesetter.h>
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::make_pair;
+using ::std::map;
+using ::std::multimap;
+using ::std::pair;
+using ::std::string;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class Cad3Handler :
+		public TiXmlVisitor
+{
+public:
+	RmiManagement* charm_management;
+	
+	ClusterInstantiationManagement* cluster_instantiation_management;
+	
+	CorbaManagement* corba_management;
+	
+	RmiManagement* corba_management2;
+	
+	CppManagement* cpp_management;
+	
+	DistributedInstantiationManagement* distributed_instantiation_management;
+	
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/ClusterInstantiationManagementBuffer/clusterinstantiationmanagementbuffer.cpp

@@ -1,0 +1,87 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <list>
+#include <string>
+
+#include <clusterinstantiationmanagement.h>
+#include <go.h>
+#include <llcmcpp3.h>
+
+namespace llcmcpp {
+
+namespace impl {
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::string;
+
+/*! \brief A component that stores the instruction for assembly creation and plays it when required.
+ */
+class ClusterInstantiationManagementBuffer:
+		virtual public ClusterInstantiationManagement,
+		virtual public Go
+{
+public:
+	ClusterInstantiationManagement* instantiation_management;
+	
+	ClusterInstantiationManagementBuffer ():
+			instantiation_management(0)
+	{}
+	
+	virtual void instantiate ( HostId host, string id, string type, string from = "" )
+	{
+		m_buffer.push_back(ClusterInstantiation(host, id, type, from));
+	}
+	
+	virtual void go ()
+	{
+#ifndef NDEBUG
+		cerr << "[ClusterInstantiationManagementBuffer] instantiating!" << endl;
+#endif
+		for ( list< ClusterInstantiation >::iterator it = m_buffer.begin(); it != m_buffer.end(); ++it ) {
+			instantiation_management->instantiate(it->host, it->name, it->type, it->from);
+		}
+		m_buffer.clear();
+#ifndef NDEBUG
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/CppManagementBuffer/cppmanagementbuffer.cpp

@@ -1,0 +1,170 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <complex>
+#include <iostream>
+#include <list>
+#include <string>
+
+#include <cppmanagement2.h>
+#include <go.h>
+#include <llcmcpp3.h>
+
+namespace llcmcpp {
+
+namespace impl {
+
+using ::std::cerr;
+using ::std::complex;
+using ::std::endl;
+using ::std::list;
+using ::std::string;
+
+class CppManagementBuffer:
+		virtual public CppManagement2,
+		virtual public Go
+{
+public:
+	CppManagement2* cpp_management;
+	
+	CppManagementBuffer ():
+			cpp_management(0)
+	{}
+	
+public:
+	virtual void configure ( EndPoint property, int8_t value )
+	{
+		m_buffer.push_back(new Configuration< int8_t >(property, value));
+	}
+	
+	virtual void configure ( EndPoint property, uint8_t value )
+	{
+		m_buffer.push_back(new Configuration< uint8_t >(property, value));
+	}
+	
+	virtual void configure ( EndPoint property, int16_t value )
+	{
+		m_buffer.push_back(new Configuration< int16_t >(property, value));
+	}
+	
+	virtual void configure ( EndPoint property, uint16_t value )
+	{
+		m_buffer.push_back(new Configuration< uint16_t >(property, value));
+	}
+	
+	virtual void configure ( EndPoint property, int32_t value )
+	{
+		m_buffer.push_back(new Configuration< int32_t >(property, value));
+	}
+	
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/DistributedNamespaceManager/distributednamespacemanager.cpp

@@ -1,0 +1,120 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <map>
+#include <set>
+#include <string>
+
+#include <distributedinstantiationmanagement.h>
+#include <distributedplacementmanagement.h>
+#include <limits.h>
+#include <llcmcpp3.h>
+#include <localinstantiationmanagement.h>
+#include <namespacemanagement.h>
+#include <unistd.h>
+
+#ifdef HOST_NAME_MAX
+#define L2C_DNM_HOST_NAME_MAX (HOST_NAME_MAX+1)
+#else
+#ifdef _POSIX_HOST_NAME_MAX
+#define L2C_DNM_HOST_NAME_MAX (_POSIX_HOST_NAME_MAX+1)
+#else
+#define L2C_DNM_HOST_NAME_MAX (256+1)
+#endif
+#endif
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::map;
+using ::std::set;
+using ::std::string;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class DistributedNamespaceManager:
+		virtual public DistributedInstantiationManagement,
+		virtual public DistributedPlacementManagement,
+		virtual public NamespaceManagement
+{
+public:
+	LocalInstantiationManagement* local_instantiation_management;
+	
+	string local_host;
+	
+	DistributedNamespaceManager ():
+			local_instantiation_management(0)
+	{
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/Driver/driver.cpp

@@ -1,0 +1,161 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <cstdlib>
+#include <iostream>
+#include <list>
+#include <string>
+#include <vector>
+
+#include <go.h>
+#include <host.h>
+#include <initializer.h>
+#include <llcmcpp3.h>
+#include <systemproperty.h>
+#include <typesetter.h>
+#include <xmlparser.h>
+
+using ::std::cerr;
+using ::std::cout;
+using ::std::endl;
+using ::std::exit;
+using ::std::list;
+using ::std::string;
+using ::std::vector;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class Driver:
+		virtual public Go
+{
+public:
+	int argc;
+	
+	char** argv;
+	
+	Go* assembly_building;
+	
+	Host* host;
+	
+	void initializations ( Initializer* initialization )
+	{
+		m_initializers.push_back(initialization);
+	}
+	
+	SystemProperty* system_properties;
+	
+	TypeSetter* typed_configuration;
+	
+	XmlParser* xml_parsing;
+	
+	Driver ():
+			argc(0),
+			argv(0),
+			assembly_building(0),
+			system_properties(0),
+			typed_configuration(0),
+			xml_parsing(0)
+	{}
+	
+	virtual void go ()
+	{
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/Driver2/driver2.cpp

@@ -1,0 +1,76 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <list>
+
+#include <go.h>
+#include <initializer.h>
+#include <llcmcpp3.h>
+
+using ::std::list;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class Driver2:
+		virtual public Go
+{
+public:
+	int argc;
+	
+	char** argv;
+	
+	Go* assembly_building;
+	
+	void initializations ( Initializer* initialization )
+	{
+		m_initializers.push_back(initialization);
+	}
+	
+	Driver2 ():
+			assembly_building(0)
+	{}
+	
+	virtual void go ()
+	{
+		for ( list< Initializer* >::iterator it = m_initializers.begin(); it != m_initializers.end(); ++it ) {
+			(*it)->initialize(&argc, &argv);
+		}
+		
+		assembly_building->go();
+	}
+	
+private:
+	list< Initializer* > m_initializers;
+	
+};
+
+LCMP(Driver2)
+	L_PROPERTY(int, argc)
+	L_PROPERTY(char**, argv)
+	L_CPP_PROVIDE(Go, go);
+	L_CPP_USE(Go, assembly_building);
+	L_CPP_USE(Initializer, initializations);
+LEND
+
+}
+
+}
+

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/GridInstantiationManagementBuffer/gridinstantiationmanagementbuffer.cpp

@@ -1,0 +1,86 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <list>
+#include <string>
+
+#include <go.h>
+#include <gridinstantiationmanagement.h>
+#include <llcmcpp3.h>
+
+namespace llcmcpp {
+
+namespace impl {
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::string;
+
+class GridInstantiationManagementBuffer:
+		virtual public Go,
+		virtual public GridInstantiationManagement
+{
+public:
+	GridInstantiationManagement* instantiation_management;
+	
+	GridInstantiationManagementBuffer ():
+			instantiation_management(0)
+	{}
+	
+private:
+	struct Instantiation {
+		string cluster; HostId host; string name; string type; string from;
+		Instantiation ( string cluster, HostId host, string name, string type, string from = "" ): cluster(cluster), host(host), name(name), type(type), from(from) {}
+	};
+	
+	list< Instantiation > m_buffer;
+	
+public:
+	virtual void instantiate ( string cluster, HostId host, string id, string type, string from = "" )
+	{
+		m_buffer.push_back(Instantiation(cluster, host, id, type, from));
+	}
+	
+	virtual void go ()
+	{
+#ifndef NDEBUG
+		cerr << "[GridInstantiationManagementBuffer] instantiating!" << endl;
+#endif
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/InstanceHost/instancehost.cpp

@@ -1,0 +1,192 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <cstdlib>
+#include <sstream>
+
+#include <dlfcn.h>
+
+#include <llcmcpp3.h>
+
+#include "instancehost.h"
+
+namespace llcmcpp {
+
+using ::std::map;
+using ::std::string;
+
+InstanceHost::InstanceHost():
+		m_available_compid(0),
+		print(0),
+		m_host_legacy(*this)
+{
+#ifndef NDEBUG
+PRINTLNERR(print, "[InstanceHost-"<<this<<"] Created");
+#endif
+}
+
+InstanceHost::~InstanceHost()
+{
+	for ( std::map< ComponentInstance, Component* >::const_iterator it = m_components.begin(); it != m_components.end(); ++ it ) {
+		delete it->second;
+	}
+#ifndef NDEBUG
+PRINTLNERR(print, "[InstanceHost-"<<this<<"] Destroyed");
+#endif
+}
+
+typedef void (*Registration)();
+
+ComponentFactory 
+InstanceHost::factory ( string name, string from )
+{
+#ifndef NDEBUG
+PRINTLNERR(print, "[InstanceHost-"<<this<<"] getting factory for component " << name);
+#endif
+	
+	map<string, ComponentFactory>::iterator it = m_factories.find(name);
+	if ( it != m_factories.end() ) {
+#ifndef NDEBUG
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/LifecycleManager/lifecyclemanager.cpp

@@ -1,0 +1,105 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <set>
+#include <string>
+
+#include <llcmcpp3.h>
+#include <lifecycle.h>
+#include <namespacemanagement.h>
+#include <rawcomponentmanagement.h>
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::set;
+using ::std::string;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class LifecycleManager:
+		virtual public LifeCycle
+{
+public:
+	NamespaceManagement* managed_instances;
+	
+	RawComponentManagement* raw_component_management;
+	
+	LifecycleManager ():
+			managed_instances(0),
+			raw_component_management(0)
+	{}
+	
+	void configure_complete ()
+	{
+#ifndef NDEBUG
+		cerr << "[LifecycleManager] configure_complete" << endl;
+#endif
+		for ( set< string >::const_iterator it = managed_instances->instances().begin(); it!=managed_instances->instances().end(); ++it ) {
+			void* life = raw_component_management->get(EndPoint(*it, "lifecycle"));
+			if ( life ) static_cast<LifeCycle*>(life)->configure_complete();
+		}
+	}
+	
+	virtual void activate ()
+	{
+#ifndef NDEBUG
+		cerr << "[LifecycleManager] activate" << endl;
+#endif
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/LocalCppManagementFilter/localcppmanagementfilter.cpp

@@ -1,0 +1,264 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <cstdlib>
+#include <iostream>
+#include <string>
+
+#include <cppmanagement.h>
+#include <cppmanagement2.h>
+#include <llcmcpp3.h>
+#include <namespacemanagement.h>
+#include <start.h>
+
+using ::std::abort;
+using ::std::cerr;
+using ::std::endl;
+using ::std::string;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class LocalCppManagementFilter:
+		virtual public CppManagement,
+		virtual public CppManagement2,
+		virtual public Start
+{
+public:
+	NamespaceManagement* global_namespace_management;
+	
+	CppManagement* local_cpp_management;
+	
+	NamespaceManagement* local_namespace_management;
+	
+	LocalCppManagementFilter ():
+			global_namespace_management(0),
+			local_cpp_management(0),
+			local_namespace_management(0)
+	{
+	}
+	
+	virtual void configure ( EndPoint property, string value)
+	{
+		if ( local_namespace_management->instances().find(property.instance) != local_namespace_management->instances().end() ) {
+			local_cpp_management->configure(property, value);
+		} else {
+			if ( global_namespace_management->instances().find(property.instance) == global_namespace_management->instances().end() ) {
+				cerr << "[LocalCppManagementFilter] *** Error: invalid instance: " << property.instance << endl;
+				abort();
+			}
+		}
+	}
+	
+	virtual void configure ( EndPoint property, int8_t value)
+	{
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/LocalInstantiationManagementBuffer/localinstantiationmanagementbuffer.cpp

@@ -1,0 +1,88 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <list>
+#include <string>
+
+#include <go.h>
+#include <llcmcpp3.h>
+#include <localinstantiationmanagement.h>
+
+namespace llcmcpp {
+
+namespace impl {
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::string;
+
+class LocalInstantiationManagementBuffer:
+		virtual public Go,
+		virtual public LocalInstantiationManagement
+{
+public:
+	
+	LocalInstantiationManagement* instantiation_management;
+	
+	LocalInstantiationManagementBuffer ():
+			instantiation_management(0)
+	{}
+	
+private:
+	struct LocalInstantiation {
+		string name; string type; string from;
+		LocalInstantiation ( string name, string type, string from = "" ): name(name), type(type), from(from) {}
+		virtual void execute ( LocalInstantiationManagementBuffer* builder ) { builder->instantiation_management->instantiate(name, type, from); }
+	};
+	
+	list< LocalInstantiation > m_buffer;
+	
+public:
+	virtual void instantiate ( string name, string type, string from = "" )
+	{
+		m_buffer.push_back(LocalInstantiation(name, type, from));
+	}
+	
+	virtual void go ()
+	{
+#ifndef NDEBUG
+		cerr << "[LocalInstantiationManagementBuffer] instantiating!" << endl;
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/LocalNamespaceManager/localnamespacemanager.cpp

@@ -1,0 +1,159 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <complex>
+#include <cstdlib>
+#include <stdexcept>
+#include <iostream>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+
+#include <tinyxml.h>
+
+#include <go.h>
+#include <host.h>
+#include <host3.h>
+#include <llcmcpp3.h>
+#include <localinstantiationmanagement.h>
+#include <namespacemanagement.h>
+#include <rawcomponentmanagement.h>
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::map;
+using ::std::out_of_range;
+using ::std::set;
+using ::std::string;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class LocalNamespaceManager :
+		virtual public LocalInstantiationManagement,
+		virtual public NamespaceManagement,
+		virtual public RawComponentManagement
+{
+public:
+	Host* host;
+
+	Host3* host3;
+
+	LocalNamespaceManager ():
+		host(0),
+		host3(0)
+	{}
+	
+	virtual void instantiate ( string name, string type, string from )
+	{
+#ifndef NDEBUG
+		cerr << "[LocalNamespaceManager] Creating instance id=" << name << " type=" << type << " from \"" << from << '"' << endl;
+#endif
+		Host3::ComponentInstance inst;
+		if ( host3 ) {
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/MpiManagementBuffer/mpimanagementbuffer.cpp

@@ -1,0 +1,80 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <list>
+#include <string>
+
+#include <go.h>
+#include <llcmcpp3.h>
+#include <mpimanagement.h>
+
+namespace llcmcpp {
+
+namespace impl {
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::string;
+
+class MpiManagementBuffer:
+		virtual public Go,
+		virtual public MpiManagement
+{
+public:
+	MpiManagement* mpi_management;
+	
+	MpiManagementBuffer ():
+			mpi_management(0)
+	{}
+	
+    virtual void mpi_connect( list< EndPoint > group )
+    {
+		m_configurations.push_back(group);
+	}
+	
+	virtual void go ()
+	{
+#ifndef NDEBUG
+		cerr << "[MpiManagementBuffer] configurating!" << endl;
+#endif
+		for ( list< list< EndPoint > >::iterator it = m_configurations.begin(); it != m_configurations.end(); ++it ) {
+			mpi_management->mpi_connect(*it);
+		}
+		m_configurations.clear();
+#ifndef NDEBUG
+		cerr << "[MpiManagementBuffer] configuration done!" << endl;
+#endif
+	}
+	
+private:
+	list< list< EndPoint > > m_configurations;
+	
+};
+
+LCMP(MpiManagementBuffer)
+	L_CPP_PROVIDE(Go, go);
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/StartBuffer/startbuffer.cpp

@@ -1,0 +1,88 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <complex>
+#include <iostream>
+#include <list>
+#include <string>
+
+#include <start.h>
+#include <go.h>
+#include <llcmcpp3.h>
+
+namespace llcmcpp {
+
+namespace impl {
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::string;
+
+class StartBuffer:
+		virtual public Start,
+		virtual public Go
+{
+public:
+	Start* start_management;
+	
+	StartBuffer ():
+			start_management(0)
+	{}
+	
+	virtual void start ( EndPoint property )
+	{
+		m_buffer.push_back(Start(property));
+	}
+	
+	virtual void go ()
+	{
+#ifndef NDEBUG
+		cerr << "[StartBuffer] executing!" << endl;
+#endif
+		for ( list< Start >::iterator it = m_buffer.begin(); it != m_buffer.end(); ++it ) {
+			start_management->start(it->property);
+		}
+		m_buffer.clear();
+#ifndef NDEBUG
+		cerr << "[StartBuffer] executions done!" << endl;
+#endif
+	}
+	
+private:
+	struct Start
+	{
+		EndPoint property;
+		Start ( EndPoint property ): property(property) {}
+	};
+	
+	list< Start > m_buffer;
+	
+};
+
+LCMP(StartBuffer)
+	L_CPP_PROVIDE(Start, start_manager);
+	L_CPP_PROVIDE(Go, go);
+	L_CPP_USE(Start, start_management);
+LEND
+
+}
+
+}
+

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/TypeInterpreter/typeinterpreter.cpp

@@ -1,0 +1,244 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <complex>
+#include <cstdlib>
+#include <sstream>
+
+#include <cppmanagement.h>
+#include <llcmcpp3.h>
+#include <typesetter.h>
+#include <cppmanagement2.h>
+
+using ::std::cerr;
+using ::std::complex;
+using ::std::endl;
+using ::std::string;
+using ::std::stringstream;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class TypeInterpreter :
+		virtual public TypeSetter
+{
+public:
+	TypeInterpreter():
+			cpp_management(0),
+			cpp_management2(0)
+	{
+	}
+	
+	CppManagement* cpp_management;
+	
+	CppManagement2* cpp_management2;
+	
+	
+	
+	virtual void configure ( EndPoint property, string type, string value )
+	{
+		stringstream valstr(value);
+		
+		if ( type == "int8" ) {
+			int8_t tval;
+			valstr >> tval;
+			if ( valstr ) {
+#ifndef NDEBUG
+				cerr << "[TypeInterpreter] " << property.instance << '.' << property.port << " := (" << type << ") " << value << endl;
+#endif // NDEBUG
+				if ( cpp_management) {
+					cpp_management->configure(property, tval);
+				} else {
+					cpp_management2->configure(property, tval);
+				}
+			} else {
+				cerr << "[TypeInterpreter] ** ERROR: Invalid value of type " << type << ": " << value << endl;
+				abort();
+			}
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/TypeSetterBuffer/typesetterbuffer.cpp

@@ -1,0 +1,85 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <list>
+#include <string>
+
+#include <go.h>
+#include <llcmcpp3.h>
+#include <typesetter.h>
+
+namespace llcmcpp {
+
+namespace impl {
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::string;
+
+class TypeSetterBuffer:
+		virtual public Go,
+		virtual public TypeSetter
+{
+public:
+	TypeSetter* typed_configuration;
+	
+	TypeSetterBuffer ():
+			typed_configuration(0)
+	{}
+	
+	virtual void configure ( EndPoint property, string type, string value )
+	{
+		m_buffer.push_back(TypedConfiguration(property, type, value));
+	}
+	
+	virtual void go ()
+	{
+#ifndef NDEBUG
+		cerr << "[TypeSetterBuffer] configurating!" << endl;
+#endif
+		for ( list< TypedConfiguration >::iterator it = m_buffer.begin(); it != m_buffer.end(); ++it ) {
+			typed_configuration->configure(it->property, it->type, it->value);
+		}
+		m_buffer.clear();
+#ifndef NDEBUG
+		cerr << "[TypeSetterBuffer] configuration done!" << endl;
+#endif
+	}
+	
+private:
+	struct TypedConfiguration {
+		EndPoint property; string type; string value;
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/base/src/llcmcpp_loader.cpp

@@ -1,0 +1,111 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifdef WITH_MPI
+#include <mpi.h>
+#endif
+
+#include <cstdlib>
+#include <iostream>
+#include <string>
+
+#include <llcmcpp/go.h>
+#include <llcmcpp/host2.h>
+#include <llcmcpp/host3.h>
+#include <llcmcpp/lifecycle.h>
+#include <llcmcpp/systemproperty.h>
+#include <llcmcpp/xmlparser.h>
+#include <llcmcpp/instancehost.h>
+#include <llcmcpp/print.h>
+
+using ::llcmcpp::Go;
+using ::llcmcpp::Host;
+using ::llcmcpp::Host3;
+using ::llcmcpp::InstanceHost;
+using ::llcmcpp::LifeCycle;
+using ::llcmcpp::SystemProperty;
+using ::llcmcpp::XmlParser;
+using ::std::cout;
+using ::std::endl;
+
+int main(int argc, char **argv)
+{
+  //int required = MPI_THREAD_SINGLE;
+  //int provided;
+  //MPI_Init_thread(&argc, &argv, required, &provided);
+  MPI_Init(&argc,&argv);
+  //std::cout<<"required serialized "<<required<<" - provided "<<provided<<std::endl;
+	
+	InstanceHost host;
+
+	Host3::ComponentInstance l2cprint = host.instantiate("L2CPrint");
+	host.print = static_cast<llcmcpp::Print*>(host.get(l2cprint, "print"));
+	
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/cmake/ReadCharmConfig.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+#usage $0 includedir VARNAME
+
+if [ "$#" != 2 ]
+then
+	echo "Error: invalid usage $0 $@" >&2
+	exit 1
+fi
+CHARMINC="$1"
+VAR="$2"
+
+. $CHARMINC/conv-config.sh $CHARMINC
+
+eval "printf \"%s\\n\" \"\${${VAR}}\""

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/corba/src/CorbaBridgeMaster/corbabridgemaster.cpp

@@ -1,0 +1,293 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <cstdlib>
+#include <fstream>
+#include <iostream>
+#include <list>
+#include <map>
+#include <string>
+#include <sstream>
+
+#include <boost/thread/condition_variable.hpp>
+#include <boost/thread/mutex.hpp>
+#include <omniORB4/CORBA.h>
+#include <stdint.h>
+
+#include <collective.h>
+#include <corbaaccess.h>
+#include <initializer.h>
+#include <interbridge.h>
+#include <llcmcpp3.h>
+#include <rawcomponentmanagement.h>
+#include <master.h>
+#include <slave.h>
+
+using ::boost::condition_variable;
+using ::boost::lock_guard;
+using ::boost::mutex;
+using ::boost::unique_lock;
+using ::CORBA::Object;
+using ::CORBA::Object_ptr;
+using ::CORBA::ORB_init;
+using ::CORBA::ORB_ptr;
+using ::PortableServer::POA;
+using ::PortableServer::POA_var;
+using ::std::abort;
+using ::std::cerr;
+using ::std::cout;
+using ::std::endl;
+using ::std::list;
+using ::std::make_pair;
+using ::std::map;
+using ::std::ofstream;
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/corba/src/CorbaManager/corbamanager.cpp

@@ -1,0 +1,96 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <cstdlib>
+#include <iostream>
+#include <map>
+#include <set>
+
+#include <corbaaccess.h>
+#include <corbamanagement.h>
+#include <distributedplacementmanagement.h>
+#include <llcmcpp3.h>
+#include <namespacemanagement.h>
+#include <rawcomponentmanagement.h>
+#include <rmimanagement.h>
+
+using ::std::abort;
+using ::std::cerr;
+using ::std::endl;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class CorbaManager :
+		virtual public CorbaManagement,
+		virtual public RmiManagement
+{
+public:
+	CorbaAccess* corba_access;
+	
+	NamespaceManagement* global_namespace_management;
+	
+	NamespaceManagement* local_namespace_management;
+	
+	DistributedPlacementManagement* namespace_management;
+	
+	RawComponentManagement* raw_component_management;
+	
+	CorbaManager ():
+			corba_access(0),
+			global_namespace_management(0),
+			local_namespace_management(0),
+			namespace_management(0),
+			raw_component_management(0)
+	{}
+	
+	virtual void corba_connect ( EndPoint user, EndPoint provider )
+	{
+		connect(user, provider);
+	}
+	
+	virtual void connect ( EndPoint user, EndPoint provider )
+	{
+		// the connector local to the user is responsible for connection
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/corba_mpi/src/GridNamespaceManager/gridnamespacemanager.cpp

@@ -1,0 +1,147 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <map>
+#include <set>
+#include <string>
+#include <sstream>
+#include <utility>
+
+#include <corbaaccess.h>
+#include <clusterinstantiationmanagement.h>
+#include <clusterplacementmanagement.h>
+#include <distributedplacementmanagement.h>
+#include <gridinstantiationmanagement.h>
+#include <gridplacementmanagement.h>
+#include <initializer.h>
+#include <llcmcpp3.h>
+#include <mpiaccess.h>
+#include <namespacemanagement.h>
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::map;
+using ::std::pair;
+using ::std::set;
+using ::std::string;
+using ::std::ostringstream;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class GridNamespaceManager:
+		virtual public ClusterPlacementManagement,
+		virtual public DistributedPlacementManagement,
+		virtual public GridInstantiationManagement,
+		virtual public GridPlacementManagement,
+		virtual public Initializer,
+		virtual public NamespaceManagement
+{
+public:
+	string local_host;
+	
+	string local_cluster;
+	
+	HostId* local_rank;
+	
+	ClusterInstantiationManagement* cluster_instantiation_management;
+	
+	GridNamespaceManager ():
+			local_rank(0),
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/gluon/include/asynccollective.h

@@ -1,0 +1,36 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+
+#ifndef GLUON_ASYNCCOLLECTIVE_H
+#define GLUON_ASYNCCOLLECTIVE_H
+
+#include "callback.h"
+
+namespace gluon {
+
+class AsyncCollective
+{
+public:
+	virtual void barrier ( Callback& cb ) = 0;
+};
+
+}
+
+#endif // GLUON_ASYNCCOLLECTIVE_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/gluon/include/callback.h

@@ -1,0 +1,34 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+
+#ifndef _GLUON_CALLBACK_H
+#define _GLUON_CALLBACK_H
+
+namespace gluon {
+
+class Callback
+{
+public:
+	virtual void operator() () = 0;
+};
+
+}
+
+#endif // _GLUON_CALLBACK_H

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/gluon/include/charminstantiationmanagement.h

@@ -1,0 +1,66 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+
+#ifndef GLUON_CHARMINSTANTIATIONMANAGEMENT_H
+#define GLUON_CHARMINSTANTIATIONMANAGEMENT_H
+
+#include <string>
+
+#include <stdint.h>
+
+namespace gluon {
+
+/*! \brief Instantition of Charm++ specific entities (groups, arrays & co.)
+ */
+class CharmInstantiationManagement
+{
+public:
+	/*! \brief Creates a new group component instance.
+	 * \param name the name (identifier) of the instance
+	 * \param type the type of the instance
+	 * \param from the library to load the instance from, defaults to $LD_LIBRARY_PATH+"/lib"+type+".so"
+	 * \post id.state == CREATED
+	 */
+	virtual void instantiate_group ( std::string name, std::string type, std::string from = "" ) = 0;
+	
+	/*! \brief Creates a new array component instance.
+	 * \param name the name (identifier) of the instance
+	 * \param size the size (number of elements) of the instance
+	 * \param type the type of the instance
+	 * \param from the library to load the instance from, defaults to $LD_LIBRARY_PATH+"/lib"+type+".so"
+	 * \post id.state == CREATED
+	 */
+	virtual void instantiate_array ( std::string name, uint32_t size, std::string type, std::string from = "" ) = 0;
+	
+	/*! \brief Creates a new array component instance.
+	 * \param name the name (identifier) of the instance
+	 * \param size the size (number of elements) of the instance
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/gluon/src/CharmCollective/charmcollective.cpp

@@ -1,0 +1,203 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <list>
+
+#include <llcmcpp/hostid.h>
+#include <llcmcpp/initializer.h>
+#include <llcmcpp/llcmcpp3.h>
+
+#include <asynccollective.h>
+
+#include <CharmCollective.decl.h>
+
+using ::llcmcpp::HostId;
+using ::llcmcpp::Initializer;
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+
+namespace gluon {
+
+namespace impl {
+
+class CharmCollective:
+		virtual public AsyncCollective,
+		virtual public Initializer
+{
+public:
+	HostId local_rank;
+	
+	CharmCollective ():
+			local_rank(CkMyPe()),
+			m_backbone_access(0),
+			m_called_barrier(0)
+	{
+		_registerCharmCollective();
+		s_waiting_collectives.push_back(this);
+	}
+	
+	static void set_backbone ( CollectiveBackbone* backbone );
+	
+	virtual void initialize ( int* argc, char*** argv )
+	{
+#ifndef NDEBUG
+		cerr << "[CharmCollective|"<<CkMyPe()<<"] initialize()"<<endl;
+#endif
+		if ( CkMyPe() == 0 ) CProxy_CollectiveBackbone::ckNew();
+	}
+	
+	virtual void barrier ( Callback& cb )
+	{
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/gluon/src/CharmInstantiationManagementBuffer/charminstantiationmanagementbuffer.cpp

@@ -1,0 +1,120 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <list>
+#include <string>
+
+#include <llcmcpp/go.h>
+#include <llcmcpp/llcmcpp3.h>
+
+#include <charminstantiationmanagement.h>
+
+namespace gluon {
+
+namespace impl {
+
+using ::llcmcpp::Go;
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::string;
+
+class CharmInstantiationManagementBuffer:
+		virtual public Go,
+		virtual public CharmInstantiationManagement
+{
+public:
+	CharmInstantiationManagement* instantiation_management;
+	
+	CharmInstantiationManagementBuffer ():
+			instantiation_management(0)
+	{}
+	
+private:
+	struct Instantiation {
+		string name; string type; string from;
+		Instantiation ( string name, string type, string from = "" ): name(name), type(type), from(from) {}
+	};
+	
+	struct ArrayInstantiation: public Instantiation {
+		uint32_t size;
+		ArrayInstantiation ( string name, uint32_t size, string type, string from = "" ): Instantiation(name, type, from), size(size) {}
+	};
+	
+	list< ArrayInstantiation > m_array_buffer;
+	
+	struct BoundArrayInstantiation: public ArrayInstantiation {
+		string bound_to;
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/gluon/src/GadHandler/gadhandler.cpp

@@ -1,0 +1,322 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <list>
+#include <map>
+#include <string>
+#include <sstream>
+
+#include <tinyxml.h>
+
+#include <llcmcpp/clusterinstantiationmanagement.h>
+#include <llcmcpp/corbamanagement.h>
+#include <llcmcpp/cppmanagement2.h>
+#include <llcmcpp/endpoint.h>
+#include <llcmcpp/lifecycle.h>
+#include <llcmcpp/llcmcpp3.h>
+#include <llcmcpp/localinstantiationmanagement.h>
+#include <llcmcpp/mpimanagement.h>
+#include <llcmcpp/rmimanagement.h>
+#include <llcmcpp/typesetter.h>
+#include <llcmcpp/start.h>
+
+#include <charminstantiationmanagement.h>
+
+namespace gluon {
+
+namespace impl {
+
+using ::llcmcpp::ClusterInstantiationManagement;
+using ::llcmcpp::EndPoint;
+using ::llcmcpp::HostId;
+using ::llcmcpp::CppManagement2;
+using ::llcmcpp::RmiManagement;
+using ::llcmcpp::Start;
+using ::llcmcpp::TypeSetter;
+using ::std::cerr;
+using ::std::endl;
+using ::std::istringstream;
+using ::std::list;
+using ::std::make_pair;
+using ::std::map;
+using ::std::multimap;
+using ::std::pair;
+using ::std::string;
+
+class GadHandler :
+		public TiXmlVisitor
+{
+public:
+	RmiManagement* charm_management;
+	
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/gluon/src/GluonDriver/gluondriver.cpp

@@ -1,0 +1,157 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <cstring>
+#include <list>
+
+#include <llcmcpp/assemblybuilding.h>
+#include <llcmcpp/go.h>
+#include <llcmcpp/initializer.h>
+#include <llcmcpp/llcmcpp3.h>
+
+#include <asynccollective.h>
+
+namespace gluon {
+
+namespace impl {
+
+using ::llcmcpp::AssemblyBuilding;
+using ::llcmcpp::Go;
+using ::llcmcpp::Initializer;
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::strcpy;
+using ::std::strlen;
+
+class GluonDriver:
+		virtual public Initializer,
+		virtual public Go,
+		virtual private Callback
+{
+public:
+	AssemblyBuilding* assembly_builder;
+	
+	AsyncCollective* collective;
+	
+	GluonDriver ():
+			assembly_builder(0),
+			collective(0),
+			m_argc(0),
+			m_argv(0),
+			m_step(INST)
+	{}
+	
+	void initializations ( Initializer* initialization )
+	{
+		m_initializers.push_back(initialization);
+	}
+	
+	void runtime_initializations ( Initializer* runtime_initialization )
+	{
+		m_runtime_initializers.push_back(runtime_initialization);
+	}
+	
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/gluon/src/ModuleRegistrar/moduleregistrar.cpp

@@ -1,0 +1,94 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <charm.h>
+#include <dlfcn.h>
+#include <llcmcpp/clusterinstantiationmanagement.h>
+#include <llcmcpp/host2.h>
+#include <llcmcpp/llcmcpp3.h>
+
+#include <charminstantiationmanagement.h>
+
+namespace gluon {
+
+namespace impl {
+
+using ::llcmcpp::ClusterInstantiationManagement;
+using ::llcmcpp::Host2;
+using ::llcmcpp::HostId;
+using ::std::string;
+
+class ModuleRegistrar:
+		virtual public CharmInstantiationManagement,
+		virtual public ClusterInstantiationManagement
+{
+public:
+	CharmInstantiationManagement* charm_instantiation_management;
+	
+	Host2* host;
+	
+	ClusterInstantiationManagement* instantiation_management;
+	
+	ModuleRegistrar ():
+			charm_instantiation_management(0),
+			host(0),
+			instantiation_management(0)
+	{
+	}
+	
+	virtual void instantiate ( HostId hostid, string id, string type, string from )
+	{
+#ifndef NDEBUG
+	  CkPrintf("[ModuleRegistrat] instantiate %s\n", type.c_str());
+#endif
+		host->preload(type, from);
+		instantiation_management->instantiate(hostid, id, type, from);
+	}
+	
+	virtual void instantiate_array ( string id, uint32_t size, string type, string from )
+	{
+		host->preload(type, from);
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/gluon/src/gluon_loader.cpp

@@ -1,0 +1,203 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <sstream>
+
+#include <llcmcpp/go.h>
+#include <llcmcpp/host.h>
+#include <llcmcpp/initializer.h>
+#include <llcmcpp/instancehost.h>
+
+#include <gluon_loader.decl.h>
+
+using ::llcmcpp::Go;
+using ::llcmcpp::Host;
+using ::llcmcpp::Host2;
+using ::llcmcpp::Initializer;
+using ::llcmcpp::InstanceHost;
+using ::std::cerr;
+using ::std::endl;
+using ::std::endl;
+
+InstanceHost host;
+
+Host::ComponentInstance gluonprint;
+
+Host::ComponentInstance local_namespace_manager;
+Host::ComponentInstance lifecycle_manager;
+Host::ComponentInstance namespace_manager;
+Host::ComponentInstance collective;
+Host::ComponentInstance charm_manager;
+Host::ComponentInstance cpp_manager;
+Host::ComponentInstance cpp_global_manager;
+Host::ComponentInstance type_interpreter;
+Host::ComponentInstance charm_buffer;
+Host::ComponentInstance charm_array_buffer;
+Host::ComponentInstance charm_group_buffer;
+Host::ComponentInstance start_buffer;
+Host::ComponentInstance cpp_mgt_buffer;
+Host::ComponentInstance charm_instantiation_buffer;
+Host::ComponentInstance instantiation_buffer;
+Host::ComponentInstance assembly_builder;
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/l2c/lad/cppmpicorbaloader_master.lad.sh

@@ -1,0 +1,362 @@
+#!/bin/sh
+#     This file is part of LLCMc++ (a.k.a. llcmcpp)
+#     Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+#     Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+# 
+#     LLCMc++ is free software: you can redistribute it and/or modify
+#     it under the terms of the GNU Lesser General Public License as published by
+#     the Free Software Foundation, either version 3 of the License, or
+#     (at your option) any later version.
+# 
+#     LLCMc++ is distributed in the hope that it will be useful,
+#     but WITHOUT ANY WARRANTY; without even the implied warranty of
+#     MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+#     GNU Lesser General Public License for more details.
+# 
+#     You should have received a copy of the GNU Lesser General Public License
+#     along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+
+part1()
+{
+cat << EOF
+<?xml version='1.0'?>
+<!--
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+-->
+<lad xmlns="http://www.inria.fr/graal/lad/1.0.0">
+	<mpi>
+		<process>
+			<start_property instance="driver_0" property="go"/>
+			
+			<instance id="local_namespace_manager_0" type="LocalNamespaceManager" >
+				<property id="host3"><systemref name="host3"/></property>
+			</instance>
+			
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/l2c/lad/cppmpicorbaloader_slave.lad.sh

@@ -1,0 +1,360 @@
+#!/bin/sh
+#     This file is part of LLCMc++ (a.k.a. llcmcpp)
+#     Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+#     Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+# 
+#     LLCMc++ is free software: you can redistribute it and/or modify
+#     it under the terms of the GNU Lesser General Public License as published by
+#     the Free Software Foundation, either version 3 of the License, or
+#     (at your option) any later version.
+# 
+#     LLCMc++ is distributed in the hope that it will be useful,
+#     but WITHOUT ANY WARRANTY; without even the implied warranty of
+#     MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+#     GNU Lesser General Public License for more details.
+# 
+#     You should have received a copy of the GNU Lesser General Public License
+#     along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+
+part1()
+{
+cat << EOF
+<?xml version='1.0'?>
+<!--
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+-->
+<lad xmlns="http://www.inria.fr/graal/lad/1.0.0">
+	<mpi>
+		<process>
+			<start_property instance="driver_0" property="go"/>
+			
+			<instance id="local_namespace_manager_0" type="LocalNamespaceManager" >
+				<property id="host3"><systemref name="host3"/></property>
+			</instance>
+			
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/l2c/src/L2CPrint/l2cprint.cpp

@@ -1,0 +1,66 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <llcmcpp/llcmcpp3.h>
+#include "llcmcpp/print.h"
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::string;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class L2CPrint :
+		virtual public Print
+{
+public:
+	
+	L2CPrint ()	{}
+	
+	virtual void printout ( const string& msg)
+	{
+	  ::std::cout << msg;
+	}
+
+	virtual void printlnout ( const string& msg)
+	{
+	  ::std::cout << msg << ::std::endl;
+	}
+
+	virtual void printerr ( const string& msg)
+	{
+	  ::std::cerr << msg;
+	}
+
+	virtual void printlnerr ( const string& msg)
+	{
+	  ::std::cerr << msg << ::std::endl;
+	}
+};
+
+LCMP(L2CPrint)
+	L_CPP_PROVIDE(Print, print);
+LEND
+
+}
+
+}

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/l2c/src/LadHandler/ladhandler.cpp

@@ -1,0 +1,312 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <iostream>
+#include <list>
+#include <map>
+#include <string>
+
+#include <tinyxml.h>
+
+#include <llcmcpp/corbamanagement.h>
+#include <llcmcpp/cppmanagement.h>
+#include <llcmcpp/distributedinstantiationmanagement.h>
+#include <llcmcpp/endpoint.h>
+#include <llcmcpp/gridinstantiationmanagement.h>
+#include <llcmcpp/lifecycle.h>
+#include <llcmcpp/llcmcpp3.h>
+#include <llcmcpp/mpimanagement.h>
+#include <llcmcpp/rmimanagement.h>
+#include <llcmcpp/typesetter.h>
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::make_pair;
+using ::std::map;
+using ::std::multimap;
+using ::std::pair;
+using ::std::string;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class LadHandler :
+		public TiXmlVisitor
+{
+public:
+	RmiManagement* charm_management;
+	
+	RmiManagement* corba_management;
+	
+	CppManagement* cpp_management;
+	
+	DistributedInstantiationManagement* distributed_instantiation_management;
+	
+	GridInstantiationManagement* grid_instantiation_management;
+	
+	MpiManagement* mpi_management;
+	
+	TypeSetter* typed_configuration;
+	
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/l2c/src/fllcmcc.sh

@@ -1,0 +1,131 @@
+#!/bin/sh
+
+# $1 = .f08 file with
+# ! PORT_DECL $$ procedure, pass :: get_provide_##name => compo##_provide_##portname $$
+# 
+# & PORT_METHOD_DECL
+
+SF=$1
+DF=$2
+
+DFT1="$DF"_tmp1
+DFT2="$DF"_tmp2
+DFT3="$DF"_tmp3
+DFT4="$DF"_tmp4
+
+function run_cpp()
+{
+    # cpp interprets Fortran // as comments :<
+    sed 's/\/\//__DOUBLE_SLASH__/g' $1 | cpp -x c -E - | sed "s/@@/'/g" | sed 's/_LLCM_CR/\n/g' | sed 's/__DOUBLE_SLASH__/\/\//g' > $2
+}
+
+function component_port_method_generate() 
+{
+    if=$1
+    of=$2
+
+    # generating port decl
+    grep PORT_DECL $if |  sed 's/.*PORT_DECL \$\$ \([^$]*\)\$\$.*/\1/g' | while read line 
+    do
+	cat $DFT1 | perl -npe "s/! PORT_METHOD_DECL/$line\n! PORT_METHOD_DECL/g" > $DFT2
+	mv $DFT2 $DFT1
+    done
+
+    # cleaning
+    mv $DFT1 $of
+    if [ -e $DFT1 ]; then rm $DFT1; fi
+    if [ -e $DFT2 ]; then rm $DFT2; fi
+}
+
+function prop_generate()
+{
+    # in/ou: DFT1
+    cp $DFT1 $DFT3
+
+    # generating component implementation
+    grep PORT_PROP_IMPL $DFT1 | sed 's/.*PORT_PROP_IMPL\(.*\)/\1/g' | while read line 
+    do
+	cat $DFT3 | perl -npe "s/!_LLCMF_COMPONENT_IMPLEMENTATION/LLCMF_PROPERTY$line\n!_LLCMF_COMPONENT_IMPLEMENTATION/g" > $DFT2
+	mv $DFT2 $DFT3
+    done
+    mv $DFT3 $DFT1
+}
+
+function component_generate()
+{
+    # in/ou: DFT1
+    cp $DFT1 $DFT3
+
+    # generating port prop implementation
+    grep TYPE_COMPONENT_IMPL $DFT1 | sed 's/.*TYPE_COMPONENT_IMPL\(.*\)/\1/g' | while read line 
+    do
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/mpi/src/MpiBarrier/mpibarrier.cpp

@@ -1,0 +1,64 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <mpi.h>
+
+#include <iostream>
+#include <list>
+#include <vector>
+
+#include <collective.h>
+#include <llcmcpp3.h>
+#include <llcmmpi.h>
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::vector;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class MpiBarrier:
+		virtual public Collective
+{
+public:
+	MPI_Comm comm;
+	
+	MpiBarrier ():
+			comm(MPI_COMM_NULL)
+	{}
+	
+	virtual void barrier ()
+	{
+		MPI_Barrier(comm);
+	}
+	
+};
+
+LCMP(MpiBarrier)
+	L_CPP_PROVIDE(Collective, collective);
+	L_MPI_COMM(comm);
+LEND
+
+}
+
+}
+

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/mpi/src/MpiInitializer/mpiinitializer.cpp

@@ -1,0 +1,113 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <mpi.h>
+
+#include <iostream>
+#include <string>
+
+
+#include <stdint.h>
+
+#include <initializer.h>
+#include <llcmcpp3.h>
+
+namespace llcmcpp {
+
+namespace impl {
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::string;
+
+class MpiInitializer:
+		virtual public Initializer
+{
+public:
+	string thread_level;
+	
+	MPI_Comm world;
+	
+	MpiInitializer ():
+		thread_level("SINGLE"),
+		world(MPI_COMM_WORLD)
+	{}
+	
+	~MpiInitializer ()
+	{
+		MPI_Finalize();
+	}
+	
+	static int mpi_level( string level )
+	{
+		if ( level == "SINGLE" ) return MPI_THREAD_SINGLE;
+		if ( level == "SERIALIZED" ) return MPI_THREAD_SERIALIZED;
+		if ( level == "FUNNELED" ) return MPI_THREAD_FUNNELED;
+		if ( level == "MULTIPLE" ) return MPI_THREAD_MULTIPLE;
+		cerr << "[MpiInitializer] *** Error: unknown thread level: "<<level<<endl;
+		cerr << "                 valid levels are SINGLE, SERIALIZED, FUNNELED, MULTIPLE"<<endl;
+		abort();
+	}
+	
[…]

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/mpi/src/MpiIorBroadcastSlave/mpiiorbroadcastslave.cpp

@@ -1,0 +1,71 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <mpi.h>
+
+#include <iostream>
+#include <string>
+
+#include <initializer.h>
+#include <llcmcpp3.h>
+#include <llcmmpi.h>
+
+namespace llcmcpp {
+
+namespace impl {
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::string;
+
+class MpiIorBroadcastSlave:
+		virtual public Initializer
+{
+public:
+	MPI_Comm comm;
+	
+	string* ior;
+	
+	MpiIorBroadcastSlave ():
+			comm(MPI_COMM_NULL),
+			ior(0)
+	{}
+	
+    virtual void initialize(int* argc, char*** argv)
+	{
+		unsigned long size = 0;
+		MPI_Bcast(&size, 1, MPI_UNSIGNED_LONG, 0, comm);
+		char* received_ior = new char[size+1]; // copy needed because string::data() returns a readonly pointer
+		MPI_Bcast(received_ior, size+1, MPI_CHAR, 0, comm);
+		received_ior[size] = 0;
+		*ior = received_ior;
+		delete[] received_ior;
+	}
+	
+};
+
+LCMP(MpiIorBroadcastSlave)
+	L_CPP_PROVIDE(Initializer, initializer);
+	L_MPI_COMM(comm);
+	L_CPP_USE(string, ior);
+LEND
+
+}
+
+}

### 2016-01-20 LIBRARIES/L2CLIGHT/l2c.git/mpi/src/MpiManager/mpimanager.cpp

@@ -1,0 +1,150 @@
+/*
+    This file is part of LLCMc++ (a.k.a. llcmcpp)
+    Copyright (C) 2011 INRIA Julien Bigot <julien.bigot@inria.fr>
+    Copyright (C) 2011 INRIA Christian Pérez <christian.perez@inria.fr>
+
+    LLCMc++ is free software: you can redistribute it and/or modify
+    it under the terms of the GNU Lesser General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    LLCMc++ is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public License
+    along with LLCMc++.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#include <mpi.h>
+
+#include <iostream>
+#include <list>
+#include <set>
+#include <stdexcept>
+#include <string>
+#include <vector>
+
+#include <clusterplacementmanagement.h>
+#include <initializer.h>
+#include <llcmcpp3.h>
+#include <llcmmpi.h>
+#include <mpiaccess.h>
+#include <mpimanagement.h>
+#include <namespacemanagement.h>
+#include <rawcomponentmanagement.h>
+
+using ::std::cerr;
+using ::std::endl;
+using ::std::list;
+using ::std::out_of_range;
+using ::std::string;
+using ::std::set;
+using ::std::vector;
+
+namespace llcmcpp {
+
+namespace impl {
+
+class MpiManager:
+		virtual public Initializer,
+		virtual public MpiManagement
+{
+public:
+	ClusterPlacementManagement* cluster_namespace_management;
+	
+	MPI_Comm comm;
+	
+	HostId local_rank;
+	
+	RawComponentManagement* raw_component_management;
+	
+	MpiManager ():
+			cluster_namespace_management(0),
+			comm(MPI_COMM_NULL),
+			raw_component_management(0)
+	{}
+	
+	virtual void initialize(int* argc, char*** argv)
+	{
+		// this should be done with LifeCycle in "activate"
[…]

### 2016-01-20 MSL/compiler/cac_parser.py

@@ -1,0 +1,148 @@
+#===============================================================
+# Helene Coullon
+# LIP INRIA
+# Parser of Python MSP_CAC
+#===============================================================
+
+#===============================================================
+# Imports
+#===============================================================
+import sys
+import os
+sys.path.append(os.path.dirname("./utils/"))
+from cac_types import *
+
+#===============================================================
+# format of the input :
+# mesh:string
+# data:
+# id,domain
+# ...
+# time: int
+# computations:
+# type:id(R,w)
+# ...
+#===============================================================
+
+#===============================================================
+# Funtions
+#===============================================================
+#-----------------------
+def parse(pfile,mesh,data,time,computations,duplicate):
+#-----------------------
+	#open the file to read
+	toparse = open(pfile,'r')
+	
+	##### MESH
+	#do not take empty lines at the beginning
+	line = toparse.readline()
+	while line.find('mesh:') < 0:
+		line = toparse.readline()
+	#mesh type
+	mesh = line.split(':')[1].strip(' ')
+	#####
+	
+	##### DATA
+	#do not take empty lines between mesh and data
+	line = toparse.readline()
+	while line.find('data:')<0:
+		line = toparse.readline()
+	line = toparse.readline() #first line of data does not contain anything, next one
+	#data extraction
+	while line.find('time:')<0 and len(line)>0:
+		tmp = line.split(',')
+		ident = tmp[0].strip('\t').strip(' ')
+		domain = tmp[1].strip('\n').strip(' ')
+		data.append(Data(ident,domain))
+		line = toparse.readline()
+	#####
+		
+	##### TIME
+	while line.find('time:')<0:
+		line = toparse.readline()
+	#time
+	time = int(line.split(':')[1].strip(' '))
+	######
+	
+	##### COMPUTATIONS
+	#do not take empty lines between time and computations
+	line = toparse.readline()
+	while line.find('computations:')<0:
+		line = toparse.readline()
+	line = toparse.readline() #first line of computations does not contain anything, next one
+	#computation extraction
+	while len(line)>0:
+		tmp = line.split(':')
[…]

### 2016-01-20 MSL/compiler/utils/cac_drawer.py

@@ -1,0 +1,24 @@
+#===============================================================
+# Helene Coullon
+# LIP INRIA
+# Drawer of Python MSP_CAC
+#===============================================================
+
+#===============================================================
+# Imports
+#===============================================================
+import sys
+import os
+sys.path.append(os.path.dirname("./utils/"))
+from cac_types import *
+
+import matplotlib.pyplot as plt
+
+#===============================================================
+# Funtions
+#===============================================================
+#-----------------------
+def saveGraph(graph,pfile):
+#-----------------------
+    nx.draw_graphviz(graph)
+    nx.write_dot(graph,pfile)

### 2016-01-20 MSL/compiler/utils/cac_import.py

@@ -1,0 +1,15 @@
+#===============================================================
+# Helene Coullon
+# LIP INRIA
+# Imports of Python MSP_CAC
+#===============================================================
+
+#===============================================================
+# Imports
+#===============================================================
+import os
+import string
+from collections import namedtuple
+import networkx as nx
+from string import Template
+import math

### 2016-01-20 MSL/compiler/utils/cac_types.py

@@ -1,0 +1,16 @@
+#===============================================================
+# Helene Coullon
+# LIP INRIA
+# Types of Python MSP_CAC
+#===============================================================
+
+#===============================================================
+# Imports
+#===============================================================
+import sys
+import os
+sys.path.append(os.path.dirname("./utils/"))
+from cac_import import *
+
+Computation = namedtuple("Computation", "id type R w N d")
+Data = namedtuple("Data", "id domain")

### 2016-01-20 MSL/old_compiler/utils.py

@@ -1,0 +1,137 @@
+import sys
+import xml.dom.minidom
+import xml.etree.ElementTree
+
+
+useFastXmlImplem = True # False => Slow
+indentXML = False # True => Very Slow
+
+
+class InternalXmlImplem:
+    class Element:
+        def __init__(self, name, attr={}):
+            self.name = name
+            self.attr = attr
+            self.tail = ''
+            self.text = ''
+            self.children = []
+        def append(self, elem):
+            self.children.append(elem)
+        def str(self):
+            if len(self.attr) > 0:
+                attrStr = ' ' + ' '.join([k+'="'+v+'"' for k,v in self.attr.iteritems()])
+            else:
+                attrStr = ''
+            if len(self.children) > 0:
+                return '<' + self.name + attrStr + '>' + self.text + ''.join([e.str() for e in self.children]) + '</' + self.name + '>' + self.tail
+            elif len(self.text) > 0:
+                return '<' + self.name + attrStr + '>' + self.text + '</' + self.name + '>' + self.tail
+            return '<' + self.name + attrStr + '/>' + self.tail
+
+    class SubElement(Element):
+        def __init__(self, parentNode, name, attr={}):
+            self.name = name
+            self.attr = attr
+            self.tail = ''
+            self.text = ''
+            self.children = []
+            parentNode.children.append(self)
+    class Comment:
+        def __init__(self, text):
+            self.text = text
+        def str(self):
+            return '<!--' + self.text + '-->\n'
+    @staticmethod
+    def tostring(elem):
+        return elem.str()
+
+
+if useFastXmlImplem:
+    xmlImplem = InternalXmlImplem
+else:
+    xmlImplem = xml.etree.ElementTree
+
+
+def makeRootNode():
+    root = xmlImplem.Element('lad', {'xmlns': 'http://www.inria.fr/avalon/lad'})
+    child = xmlImplem.SubElement(root, 'mpi')
+    #root.tail = '\n'
+    #root.text = '\n'
+    #child.tail = '\n'
+    #child.text = '\n'
+    return (root, child)
+
+
+def makeProcess(parentNode):
+    node = xmlImplem.SubElement(parentNode, 'process')
+    #node.tail = '\n'
+    #node.text = '\n'
+    return node
+
+
+def makeStartProperty(parentNode, instance, instanceProperty):
[…]

### 2016-01-21 ARTICLES/RR/rapport/formalism.tex
rapport Helene 21/01/2016

--- 2016-01-20 rapport/formalism.tex
@@
 The neighborhood indicates which entities, on which $r$ is applied, will be read during the computation.
-\item A numerical expression $\text{exp}_{w}: entity(w) \times R^n \rightarrow V_{w}$ is a function that computes the value of the written quantity $w \in \Delta$ for a given entity $e \in entity(w)=E_w$, using a set of input data $R$.
+\item A numerical expression $\text{exp}_{w}: entity(w) \times R^n \rightarrow V_{w}$ is a function which computes the values of the written quantity $w \in \Delta$ for a given entity $e \in entity(w)=E_w$, using a set of input data $R$.
 \item A computation kernel $k$ of a numerical simulation is defined as $c(R,w,\text{exp},D)$, where $R$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$ is the computation domain.}

### 2016-01-21 ARTICLES/IJPP/article/article.tex
article IJPP Helene 21/01/2016

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{subfig}
\usepackage{caption}
\usepackage{cite}
\usepackage{filecontents}
\usepackage{listings}
\usepackage{tikz}
\usepackage[ruled,noend]{algorithm2e}
\usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}

\usepackage{todonotes}

%\newtheorem{theorem}{Theorem}[section]
\newtheorem{myprop}[theorem]{Proposition}
% \newenvironment{proof}[1][Proof]{\begin{trivlist}
% \item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{mydef}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{myth}[1][Theorem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\def\pprec{\mathrel{\scalebox{.9}[1]{$\prec$}\mkern-3mu%
  \scalebox{.4}[1]{$\prec$}\mkern-5.5mu\scalebox{.4}[1]{$\prec$}}}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{The Multi-Stencil Language}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Helene Coullon         \and
        Chirtsian Perez \and
        Julien Bigot
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{Helene Coullon \and Christian Perez \at
              INRIA \\
              46 Allée d'Italie\\
              69007 Lyon\\
              \email{helene.coullon@inria.fr}
           \and
           Julien Bigot \at
              CEA
}

%\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Insert your abstract here.
\keywords{First keyword \and Second keyword \and More}
\end{abstract}

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
Contributions :
\begin{itemize}
\item Formalism of a multi-stencil program and its parallelization
\item The Multi-stencil Language: agnostic from the type of mesh and of the choosen implementation
\item A first implementation of the MSL compiler (static scheduling + fusion + dump to SkelGIS/components)
\item Evaluation of the different types of parallelism introduced + compiler + fusion
\end{itemize}
%------------------------------------------------------------------------------
\section{Computational model of Multi-Stencil Programs}
\label{sect:formalism}
\input{formalism.tex}
% Formalism :
% \begin{itemize}
% \item mesh / mesh entities and groups / computations domains
% \item data and scalars
% \item computations : stencil, local and reduction
% \item Multi-Stencil Program
% \end{itemize}
%------------------------------------------------------------------------------
\section{Parallelization of Multi-Stencil Programs}
\label{sect:parallelism}
\input{parallelism.tex}
% Parallelization formalism :
% \begin{itemize}
% \item Data parallelism (introduction of synchronizations)
% \item Task parallelism (computation of the dependency graph)
% \end{itemize}
%------------------------------------------------------------------------------
\section{The Multi-Stencil Language}
\label{sect:msl}
\input{msl.tex}
% The parallel empty skeleton of the application can be computed automatically from the language :
% \begin{itemize}
% \item grammar
% \item example from the language to the dependency graph
% \end{itemize}
%------------------------------------------------------------------------------
\section{Static Scheduling}
\label{sect:msp}
\input{msp.tex}
% From the dependency graph we study a first dump: static scheduling
% \begin{itemize}
% \item description of the algorithm to get the binary tree decomposition of a minimal serie-parallel graph
% \item use of the serie-parallel tree decomposition to detect fusions of kernels
% \item example (the same one) from the dependency graph to the binary tree
% \end{itemize}
%------------------------------------------------------------------------------
\section{Evaluation}
\label{sect:eval}
\begin{itemize}
\item short description of the implementation (compiler in Python, DDS = SkelGIS, Components)
\item evaluation of the compiler (execution times of the different steps)
\item evaluation of the data parallelism (weak/strong scaling)

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/400/median_weak.pdf}}
  \caption{weak-scaling 400x400 blocks.}
  \label{fig:mesh}
\end{center}\end{figure}

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/800/median_weak.pdf}}
  \caption{weak-scaling 800x800 blocks}
  \label{fig:mesh}
\end{center}\end{figure}

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/strong_scaling/10K_1K/median_strong.pdf}}
  \caption{strong-scaling 10kx10k, 1k iterations}
  \label{fig:mesh}
\end{center}\end{figure}

\item evaluation of the fusion

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/fusVSbase.pdf}}
  \caption{strong scaling 500x500, 200 iterations, with and without fusion}
  \label{fig:mesh}
\end{center}\end{figure}

\item evaluation of the hybrid parallelism

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_close_median.pdf}}
  \caption{MPI only vs MPI+threads with \emph{close} scheduling policy (OpenMP)}
  \label{fig:mesh}
\end{center}\end{figure}

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_spread_median.pdf}}
  \caption{MPI only vs MPI+threads with \emph{spread} scheduling policy (OpenMP)}
  \label{fig:mesh}
\end{center}\end{figure}

\item analytic model for the results

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/times.pdf}}
  \caption{Computation vs communication times in the MPI only application}
  \label{fig:mesh}
\end{center}\end{figure}

\end{itemize}
%------------------------------------------------------------------------------
\section{Related work}
\label{sect:rel}
\begin{itemize}
\item Lizst, Pochoir, PATUS for optimizations/parallelization of a single stencil kernel + mesh dependent
\item Halide for pipeline of stencil computations = mesh dependent
\item Reuse of external solutions as SkelGIS, GA, Lizst, Pochoir etc.
\end{itemize}
%------------------------------------------------------------------------------
\section{Conclusion}
\label{sect:concl}
%------------------------------------------------------------------------------

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

\bibliographystyle{spbasic}
\bibliography{biblio}

\end{document}
% end of file template.tex

### 2016-01-21 ARTICLES/IJPP/article/msl.tex
article IJPP Helene 21/01/2016

\begin{filecontents*}{grammar.txt}
program ::= "mesh:" meshid 
            "mesh entities:" listmeshent
            "computation domains:" 
                      listcompdom
            "independent:"
                      listinde
            "stencil shapes:"
                      liststencil
            "data:" listdata
            "scalar" listscalar
            listloop

listmeshent ::= meshent listmeshent 
                        |  meshent
listcompdom ::= compdom listcompdom 
                        |  compdom
compdom ::= compdomid "in" listmeshent
listinde ::= inde listinde |  inde
inde ::= compdomid "and" compdomid
liststencil ::= stencil liststencil 
                        | stencil
listdata ::= data listdata |  data
data ::= dataid "," meshent
listscalar ::= scalar listscalar 
                      | scalar

listloop ::= loop listloop | loop
loop ::=  "time:" iteration
          "computations:" listcomp

iteration ::= num | criteria
criteria ::= scalar "=" criteriaid "(" 
                    listscalarread "," 
                    listdataread ")"
listscalarread ::= scalar listscalarread 
                          |  scalar
listcomp ::= comp listcomp |  comp
comp ::= dataid "[" compdomid "]=" 
                    compid "(" 
                    listscalarread "," 
                    listdataread ")"
listdataread ::= dataread listdataread 
                          |  dataread
dataread ::= dataid "[" neighborid "]" 
              |  dataid
\end{filecontents*}

\begin{figure}[t]
  \hspace{5mm}
  \begin{minipage}[t]{\textwidth}
    %  \subfloat[\label{fig:grammar}]
    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left,linewidth=.87\textwidth]{grammar.txt}}   
    \caption{Grammar of the Multi-Stencil Language. \label{fig:grammar}}
  \end{minipage}
\end{figure}

### 2016-01-21 ARTICLES/IJPP/article/formalism.tex
article IJPP Helene 21/01/2016

--- 2016-01-21 ARTICLES/RR/rapport/formalism.tex
@@
 In other words, the PDEs are transformed to a set of numerical computations applied at each time step on elements of the discretized space domain.
-Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}, and a set of auxiliary computations also needed to perform the simulation, and also called \emph{local computations}.
+Among those numerical computations are found a set of numerical schemes, also called \textit{stencil computations}, a set of auxiliary computations also called \emph{local computations}, and finally a set of \emph{reductions} to reduce a bunch of values to a single scalar value.
 This section gives formal definitions of what we call a \textit{stencil program} and its computations.
@@
 \begin{mydef}
-\textit{A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V\subset \Omega$ is the set of vertices and $E\subseteq V^2$ the set of edges.
+A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V\subset \Omega$ is the set of vertices and $E\subseteq V^2$ the set of edges.
 The set of edges $E$ of a mesh $\mathcal{M}=(V,E)$ does not contain bridges.
-It is said that the mesh is applied onto $\Omega$}.
+It is said that the mesh is applied onto $\Omega$.
 \end{mydef}
@@
 \begin{mydef}
-\textit{The dimension of a mesh $\mathcal{M}=(V,E)$ applied onto $\Omega=\mathbb{R}^n$ is denoted $dim(\mathcal{M})=n$.}
+The dimension of a mesh $\mathcal{M}=(V,E)$ applied onto $\Omega=\mathbb{R}^n$ is denoted $dim(\mathcal{M})=n$.
 \end{mydef}
@@
 \begin{itemize}
-\item \textit{An entity $e$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $e\subset (V\cup E)$.
+\item An entity $e$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $e\subset (V\cup E)$.
 \item A mesh entities group is denoted $E$ such that $E \in \mathcal{P}(V\cup E)$. It represents a set of entities of the same type.
-\item Finally the set of mesh entities groups of a simulation is denoted $\mathcal{E}$.}
+\item Finally the set of mesh entities groups of a simulation is denoted $\mathcal{E}$.
 \end{itemize}
@@
 \begin{itemize}
-\item \textit{We denote a scalar as an identifier associated to a numerical value and applied onto a mesh $\mathcal{M}$ where $dim(\mathcal{M})=0$.
+\item We denote a scalar as an identifier associated to a numerical value and applied onto a mesh $\mathcal{M}$ where $dim(\mathcal{M})=0$.
 In other words a scalar can be seen as a variable containing a numerical value.
-The set of scalars is denoted $\mathcal{S}$.}
-\item \textit{$\mathcal{T}=\mathbb{R}$ is the continuous time domain of a numerical simulation.}
-\item \textit{The discretization of the continuous time domain $\mathcal{T}$ is denoted as the pair $T(\Delta t,conv)$.} 
+The set of scalars is denoted $\mathcal{S}$.
+\item $\mathcal{T}=\mathbb{R}$ is the continuous time domain of a numerical simulation.
+\item The discretization of the continuous time domain $\mathcal{T}$ is denoted as the pair $T(\Delta t,conv)$.
 \begin{itemize}
-\item \textit{$\Delta t \in \mathcal{S}$ represents the time interval of the numerical simulation, such that for a current time iteration $t_i$, the next time iteration is $t_{i+1} = t_i + \Delta t$.
+\item $\Delta t \in \mathcal{S}$ represents the time interval of the numerical simulation, such that for a current time iteration $t_i$, the next time iteration is $t_{i+1} = t_i + \Delta t$.
 The default value of $\Delta t$ is $1$.
@@
 For example, the simplest $conv$ is typically to have two scalars: $t$ the time, and a fix number of iterations denoted $it$.
-At each time step $\Delta t=1$, and $conv$ returns the boolean expression $t<it$.}
+At each time step $\Delta t=1$, and $conv$ returns the boolean expression $t<it$.
 \end{itemize}
@@
 \begin{itemize}
-\item \textit{A quantity is a function $\delta: E_{\delta} \mapsto V_{\delta}$ which associates each entity of a group $e\in E_{\delta}$ to a value $v\in V_{\delta}$ where $V_{\delta}$ is a data type, typically $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.
+\item \textit{A quantity (or data) is a function $\delta: E_{\delta} \mapsto V_{\delta}$ which associates each entity of a group $e\in E_{\delta}$ to a value $v\in V_{\delta}$ where $V_{\delta}$ is a data type, typically $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.
 \item The set of quantities applied onto the mesh is denoted $\Delta$.
@@
 
-In this section are considered two different types of computations, a reduction or a numerical kernel.
+In this section are considered three different types of kernel computations, stencil kernels, local kernels, and reduction kernels.
 
@@
 \begin{itemize}
-\item \textit{A computation domain $D$ is a subpart  a mesh entities group, $D \subset E \in \mathcal{E}$.
+\item A computation domain $D$ is a subpart  a mesh entities group, $D \subset E \in \mathcal{E}$.
 \item The set of computation domains of a numerical simulation is denoted $\mathcal{D}$.
@@
 \item A numerical expression $\text{exp}_{w}: entity(w) \times R^n \rightarrow V_{w}$ is a function which computes the values of the written quantity $w \in \Delta$ for a given entity $e \in entity(w)=E_w$, using a set of input data $R$.
-\item A computation kernel $k$ of a numerical simulation is defined as $c(R,w,\text{exp},D)$, where $R$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$ is the computation domain.}
+\item A computation kernel $k$ of a numerical simulation is defined as $c(R,w,\text{exp},D)$, where $R$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$ is the computation domain.
 \end{itemize}
@@
 \begin{itemize}
-\item \textit{We denote by $identity$ the identity function $x=x$.
+\item We denote by $identity$ the identity function $x=x$.
 \item A kernel computation $k(R,w,\text{exp},D) \in \Gamma$ is a stencil kernel $\iff \exists (r,n) \in R$ such that $n \neq identity$.
-\item A kernel computation $k(R,w,\text{exp},D) \in \Gamma$ is a local kernel $\iff \forall (r,n) \in R$, $n = identity$.}
+\item A kernel computation $k(R,w,\text{exp},D) \in \Gamma$ is a local kernel $\iff \forall (r,n) \in R$, $n = identity$.
 \end{itemize}
@@
 \noindent \textbf{Property}
-\textit{A kernel for which all data read and written are applied onto a mesh of dimension $0$ is a local kernel.}
+A kernel for which all data read and written are applied onto a mesh of dimension $0$ is a local kernel.
 
@@
 \begin{itemize}
-\item \textit{A reduction computation $r$ of a numerical simulation is defined as $r(R,w,red)$, where $R$ is the set of data read, $w \in \mathcal{S}$ is the scalar written, and $\text{red}$ is an associative reduction function.
+\item A reduction computation $r$ of a numerical simulation is defined as $r(R,w,red)$, where $R$ is the set of data read, $w \in \mathcal{S}$ is the scalar written, and $\text{red}$ is an associative reduction function.
 % \item For any $i,j$ such that $r_i \in R$ and $r_j \in R$, if $dim(mesh(r_i))=dim(mesh(r_j))=n$.
@@
 \item The set of $n$ ordered computations (kernel or reduction) of a numerical simulation is denoted $\Gamma = [c_i]_{0 \leq i \leq n-1}$, such that $\forall c_i,c_j \in \Gamma$, if $i \leq j$, then $c_i$ is computed before $c_j$.
-\item Finally, a \textit{multi-stencil program} is defined by the septuplet $\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta, \mathcal{S},\Gamma)$.}
+\item Finally, a \textit{multi-stencil program} is defined by the septuplet $\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta, \mathcal{S},\Gamma)$.
 \end{itemize}

### 2016-01-21 ARTICLES/IJPP/article/formalism.tex
article IJPP Helene 21/01/2016

@@
 \begin{itemize}
-\item A computation domain $D$ is a subpart  a mesh entities group, $D \subset E \in \mathcal{E}$.
+\item A computation domain $D$ is a subpart of a mesh entities group, $D \subset E \in \mathcal{E}$.
 \item The set of computation domains of a numerical simulation is denoted $\mathcal{D}$.
-\item A neighborhood $n$ is a function which for a given entity $e \in E_i$, returns a set of $m$ entities in $E_j$, $n : E_i \rightarrow E_j^m$. One can notice that $i = j$ is possible.
+\item A neighborhood $n$ is a function which for a given entity $e \in E_i$, returns a set of $m$ entities in $E_j$, $n : E_i \rightarrow E_j^m$.
+One can notice that $i = j$ is possible.
+Most of the time, such a neighborhood is called a stencil shape.
 \item The set of neihborhood functions in a numerical simulation is denoted $\mathcal{N}$.
-\item A single data is written during a computation and is denoted $w \in \Delta$. Thus, the computation domain of the computation is $D \subset E_w=entity(w)$.
-\item A set of data to read and use in a computation is denoted $R$.
-Each element of $R$ is a tuple $(r,n)$, where $r \in \Delta$ and $n$ is a neighborhood function such that $n : E_w \rightarrow entity(r)^m$.
-The neighborhood indicates which entities, on which $r$ is applied, will be read during the computation.
-\item A numerical expression $\text{exp}_{w}: entity(w) \times R^n \rightarrow V_{w}$ is a function which computes the values of the written quantity $w \in \Delta$ for a given entity $e \in entity(w)=E_w$, using a set of input data $R$.
-\item A computation kernel $k$ of a numerical simulation is defined as $c(R,w,\text{exp},D)$, where $R$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$ is the computation domain.
 \end{itemize}
-%\end{mydef}
-%If the number of computations in $\Gamma$ is $card(\Gamma)=n$, such that $\bigcup_{i=0}^{n-1}c_i = \Gamma$, then $\bigcup_{i=0}^{m-1}R_i \cup w_i \subseteq \Delta$.
 
-Most of the time we can dissociate two types of kernel computations.
+\begin{mydef}
+A computation kernel $k$ of a numerical simulation is defined as $k(R,(w,D),comp)$, where 
+\begin{itemize}
+\item $(w,D) \in \Delta \times \mathcal{D}$ is the single data modified by the computation kernel onto a given computation domain.
+Thus, $w \in \Delta$ and $D \in \mathcal{D}$ is the computation domain on which $w$ is computed, $D \subset E_w=entity(w)$.
+\item $R$ is the set of tuples $(r,n)$, where $r \in \Delta$ and $n \in \mathcal{N}$ is a neighborhood function such that $n : E_w \rightarrow entity(r)^m$.
+The neighborhood indicates which entities of $r$ will be read in order to compute a single entity of $w$.
+\item $comp$ is the numerical computation which returns a value from a set of input values, $comp: V^n \rightarrow V$.
+\end{itemize}
+% \item A numerical expression $\text{exp}_{w}: entity(w) \times R^n \rightarrow V_{w}$ is a function which computes the values of the written quantity $w \in \Delta$ for a given entity $e \in entity(w)=E_w$, using a set of input data $R$.
+% \item A computation kernel $k$ of a numerical simulation is defined as $c(R,w,\text{exp},D)$, where $R$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$ is the computation domain.
+% \end{itemize}
+%$R$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$ is the computation domain.
+\end{mydef}
+
+$comp$ represents the actual numerical expression which is computed by a kernel. We can dissociate three types of kernel computations.
 
@@
 \item We denote by $identity$ the identity function $x=x$.
-\item A kernel computation $k(R,w,\text{exp},D) \in \Gamma$ is a stencil kernel $\iff \exists (r,n) \in R$ such that $n \neq identity$.
-\item A kernel computation $k(R,w,\text{exp},D) \in \Gamma$ is a local kernel $\iff \forall (r,n) \in R$, $n = identity$.
+\item A kernel computation $k(R,(w,D),comp)$ is a \emph{stencil kernel} $\iff \exists (r,n) \in R$ such that $n \neq identity$.
+\item A kernel computation $k(R,(w,D),comp)$ is a \emph{local kernel} $\iff \forall (r,n) \in R$, $n = identity$.
+\item A kernel computation $k(R,(w,D),comp)$ is a \emph{reduction kernel} if $w$ is a scalar, in other words $\iff dim(mesh(entity(w)))=0$.
 \end{itemize}
@@
 
+\medskip
 A reduction computation is a computation which takes data applied onto meshes and returns a data with a dimension reduced to zero, in other words a scalar.
@@
 
-\medskip
+\noindent \textbf{Property}
+Considering a reduction kernel $k(R,(w,D),comp)$, $comp$ must be a binary and associative operation on the type $V$, $comp: V \times V \rightarrow V$.
+
 \noindent \textbf{Definitions}
-%\begin{mydef}
 \begin{itemize}
-\item A reduction computation $r$ of a numerical simulation is defined as $r(R,w,red)$, where $R$ is the set of data read, $w \in \mathcal{S}$ is the scalar written, and $\text{red}$ is an associative reduction function.
-% \item For any $i,j$ such that $r_i \in R$ and $r_j \in R$, if $dim(mesh(r_i))=dim(mesh(r_j))=n$.
-% \item The data written by the reduction $w$ verifies $dim(mesh(w))<n$.
-\item A reduction function $\text{red}: \Delta^n \rightarrow \mathcal{S}$ is an operation which reduces a set of data in $\Delta$ (on it sentities) to a single scalar variable in $\mathcal{S}$.
-The reduction function must be associative.
-\item The set of $n$ ordered computations (kernel or reduction) of a numerical simulation is denoted $\Gamma = [c_i]_{0 \leq i \leq n-1}$, such that $\forall c_i,c_j \in \Gamma$, if $i \leq j$, then $c_i$ is computed before $c_j$.
-\item Finally, a \textit{multi-stencil program} is defined by the septuplet $\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta, \mathcal{S},\Gamma)$.
+\item The set of $n$ ordered computation kernels of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i \leq j$, then $k_i$ is computed before $k_j$.
+\item Finally, a \textit{multi-stencil program} is defined by the septuplet 
+\begin{equation*}
+\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta, \mathcal{S},\Gamma)
+\end{equation*}
 \end{itemize}
-%!!!!!!!  est-ce qu'on fusionne le maillage, les entites et les domaines ?????
-%\end{mydef}
 

### 2016-01-21 ARTICLES/IJPP/article/parallelism.tex
article IJPP Helene 21/01/2016

--- 2016-01-20 rapport/parallelism.tex
@@
-Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
+% Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
 
@@
 \textit{
-For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $\exists (r,n) \in R_j$ with $n\neq identity$ ($c_j$ is a stencil computation), and such that $\{w_i\} \subset R_j$.
+For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $k_i$ and $k_j$, denoted $k_i \pprec k_j$, if $\exists (r,n) \in R_j$ with $n\neq identity$ ($k_j$ is a stencil computation), and such that $\{w_i\} \subset R_j$.
 Moreover, the data to update is $\{w_i\} \cap R_j = \{w_i$\}.}
@@
 \begin{mydef}
-\textit{A synchronization between two computations $c_i \pprec c_j$ is defined as a specific computation $sync(R,w,exp,D)$ where $R=\{w_i\}$, $w=w_i$, $exp=identity$ and $D=\emptyset$.}
+\textit{A synchronization between two computations $k_i \pprec k_j$ is defined as a specific computation $sync(R,w,exp,D)$ where $R=\{w_i\}$, $w=w_i$, $exp=identity$ and $D=\emptyset$.}
 \end{mydef}
@@
 \begin{mydef}
-If a synchronization is needed before $c_j$ (on the data $w_i$), $c_j$ is replaced in $\Gamma_{sync}$ by the list
+If a synchronization is needed before $k_j$ (on the data $w_i$), $k_j$ is replaced in $\Gamma_{sync}$ by the list
 \begin{equation*}
-\Gamma_{sync} = [sync(\{w_i\},w_i,identity,\emptyset), c_j]
+\Gamma_{sync} = [sync(\{w_i\},w_i,identity,\emptyset), k_j]
 \end{equation*}
@@
 \begin{mydef}
-The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[c_i]_{0 \leq i \leq n-1}$ and $l_2=[c'_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[c_0,\dots,c_{n-1},c'_0,\dots,c'_{m-1}]$.
+The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[k_i]_{0 \leq i \leq n-1}$ and $l_2=[c'_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[c_0,\dots,c_{n-1},c'_0,\dots,c'_{m-1}]$.
 \end{mydef}
@@
 F_{sync}(\Gamma,j) = 
-\begin{cases} 	[\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) & \mbox{if }\forall i<j \mbox{, } c_i \not\pprec c_j\\
-				[sync(\{w_i\},w_i,identity,\emptyset), c_j] \cdot F_{sync}(\Gamma,j+1) & \forall i<j \text{, and }c_i \pprec c_j\}\\
+\begin{cases} 	[\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) & \mbox{if }\forall i<j \mbox{, } k_i \not\pprec k_j\\
+				[sync(\{w_i\},w_i,identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1) & \forall i<j \text{, and }k_i \pprec k_j\}\\
 				[] & \mbox{if }j=|\Gamma|.
@@
 % \begin{myprop}
-% Denoting four computations $c_l$, $c_k$, $c_i$ and $c_j$ of $\Gamma$, with $l<k$, $i<j$, and where $c_l \pprec c_k$ and $c_i \pprec c_j$.
-Then, computations $c_k$ and $c_j$ are tansformed to $c^*_k(c_k,update(w_l))$ and $c^*_j(c_j,update(w_i))$.
-If $i<k$, a single update call can be performed at $k$ such that $c^*_k(c_k,update(w_l\cup w_i))$, and in this case, the computation $c_j$ is not transformed.
+% Denoting four computations $c_l$, $c_k$, $k_i$ and $k_j$ of $\Gamma$, with $l<k$, $i<j$, and where $c_l \pprec c_k$ and $k_i \pprec k_j$.
+Then, computations $c_k$ and $k_j$ are tansformed to $c^*_k(c_k,update(w_l))$ and $c^*_j(k_j,update(w_i))$.
+If $i<k$, a single update call can be performed at $k$ such that $c^*_k(c_k,update(w_l\cup w_i))$, and in this case, the computation $k_j$ is not transformed.
 % \end{myprop}
@@
 % \begin{proof}
-% By definition, the synchronization of $w_i$ can be performed between $c_i$ and $c_j$.
-As $i<k$ the computation $c_i$ is not performed between $k$ and $j$.
-As a result, the synchronization of $w_i$ can be performed between $c_i$ and $c_k$.
+% By definition, the synchronization of $w_i$ can be performed between $k_i$ and $k_j$.
+As $i<k$ the computation $k_i$ is not performed between $k$ and $j$.
+As a result, the synchronization of $w_i$ can be performed between $k_i$ and $c_k$.
 % \end{proof}
 
-% In other words, for each synchronization in $\Gamma_{data}$, if the computation $c_i$ from which $c_j$ is synchronized occurs before the last synchronization, the two synchronizations can be reduced to one with the union of the data to update.
+% In other words, for each synchronization in $\Gamma_{data}$, if the computation $k_i$ from which $k_j$ is synchronized occurs before the last synchronization, the two synchronizations can be reduced to one with the union of the data to update.
 
@@
 \begin{mydef}
-For two computations $c_i(R_i,w_i,\text{exp}_i,D_i)$ and $c_j(R_j,w_j,\text{exp}_j,D_j)$, with $i < j$, it is said that $c_j$ is data dependant from $c_i$, denoted $c_i\prec c_j$, if $\{w_i\} \cap R_j \neq \emptyset$.
-In this case, $c_i$ has to be computed before $c_j$.
-The binary relation $c_i\prec c_j$ represents a \textit{dependency}.
+For two computations $k_i(R_i,w_i,\text{exp}_i,D_i)$ and $k_j(R_j,w_j,\text{exp}_j,D_j)$, with $i < j$, it is said that $k_j$ is data dependant from $k_i$, denoted $k_i\prec k_j$, if $\{w_i\} \cap R_j \neq \emptyset$.
+In this case, $k_i$ has to be computed before $k_j$.
+The binary relation $k_i\prec k_j$ represents a \textit{dependency}.
 \end{mydef}
@@
 % \begin{mydef}
-% For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, a time dependency is denoted $c_i \blacktriangleleft c_j$ and means that $c_i$ has to be computed before $c_j$.
+% For two computations $k_i(R_i,w_i,\text{exp}_i)$ and $k_j(R_j,w_j,\text{exp}_j)$, a time dependency is denoted $k_i \blacktriangleleft k_j$ and means that $k_i$ has to be computed before $k_j$.
 % \end{mydef}
@@
 % \begin{myprop}
-% For two computations $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$ such that $c_i \prec c_j$, $c_i \blacktriangleleft c_j$ is verified.
+% For two computations $k_i(R_i,w_i,\text{exp}_i)$ and $k_j(R_j,w_j,\text{exp}_j)$ such that $k_i \prec k_j$, $k_i \blacktriangleleft k_j$ is verified.
 % \end{myprop}
@@
 % \begin{proof}
-% By definition $c_i \prec c_j$ means that $\{w_i\}\cap R_j \neq \emptyset$ and that $c_i$ has to computed before $c_j$.
-%Thus, the relation $c_i \prec c_j$ is stronger than $c_i \blacktriangleleft c_j$.
+% By definition $k_i \prec k_j$ means that $\{w_i\}\cap R_j \neq \emptyset$ and that $k_i$ has to computed before $k_j$.
+%Thus, the relation $k_i \prec k_j$ is stronger than $k_i \blacktriangleleft k_j$.
 % \end{proof}
@@
 
-From an ordered list of computations $\Gamma_{sync}$, a directed dependency graph $\Gamma_{dep}(V,A)$ can be built finding all pairs of computations $c_i(R_i,w_i,\text{exp}_i,D_i)$ and $c_j(R_j,w_j,\text{exp}_j,D_j)$, with $i<j$, such that $c_i \prec c_j$.
-%Those time dependencies can be found from pairs of data dependencies $c_i \prec c_j$.
+From an ordered list of computations $\Gamma_{sync}$, a directed dependency graph $\Gamma_{dep}(V,A)$ can be built finding all pairs of computations $k_i(R_i,w_i,\text{exp}_i,D_i)$ and $k_j(R_j,w_j,\text{exp}_j,D_j)$, with $i<j$, such that $k_i \prec k_j$.
+%Those time dependencies can be found from pairs of data dependencies $k_i \prec k_j$.
 
@@
 \begin{cases} 	(\{\},\{\}) & \mbox{if }j=|\Gamma_{sync}|\\
-				(c_j, \{(\overset{\frown}{c_i,c_j})\mbox{, }\forall i < j \mbox{, } c_i\prec c_j \})\cup F_{dep}(\Gamma_{sync},j+1) & \mbox{if }j<|\Gamma_{sync}|
+				(k_j, \{(\overset{\frown}{k_i,k_j})\mbox{, }\forall i < j \mbox{, } k_i\prec k_j \})\cup F_{dep}(\Gamma_{sync},j+1) & \mbox{if }j<|\Gamma_{sync}|
 \end{cases}
@@
 This constructive function is possible because the input is an ordered list.
-Actually, if $c_i\prec c_j$ then $i<j$.
-As a result, $c_i$ is already in $V$ when the arc $(\overset{\frown}{c_i,c_j})$ is built.
+Actually, if $k_i\prec k_j$ then $i<j$.
+As a result, $k_i$ is already in $V$ when the arc $(\overset{\frown}{k_i,k_j})$ is built.
 
@@
 \end{equation*}
-The set of computations $\Gamma_{dep}$ is a dependency graph between computations $c_i$ of $\Gamma$ and the synchronization computations $sync_i$ added into $\Gamma_{sync}$.
+The set of computations $\Gamma_{dep}$ is a dependency graph between computations $k_i$ of $\Gamma$ and the synchronization computations $synk_i$ added into $\Gamma_{sync}$.
 $\Gamma_{dep}$ can be built from the call to
@@
 % Using the function $T_{task}$ to build $\Gamma_{task}$, however, duplication of dependencies may occur because of the transitivity of the relation $\blacktriangleleft$.
-Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,\text{exp}_k)$, $c_i(R_i,w_i,\text{exp}_i)$ and $c_j(R_j,w_j,\text{exp}_j)$, with $k<i<j$ and where $c_k \prec c_i$, $c_i \prec c_j$ and $c_k \prec c_j$, then the transitivity $c_k \blacktriangleleft c_j$ is verified because of the relations $c_k \blacktriangleleft c_i$ and $c_i \blacktriangleleft c_j$.
-However, $c_k \blacktriangleleft c_j$ is also directly represented by the binary relation $c_k \prec c_j$.
+Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,\text{exp}_k)$, $k_i(R_i,w_i,\text{exp}_i)$ and $k_j(R_j,w_j,\text{exp}_j)$, with $k<i<j$ and where $c_k \prec k_i$, $k_i \prec k_j$ and $c_k \prec k_j$, then the transitivity $c_k \blacktriangleleft k_j$ is verified because of the relations $c_k \blacktriangleleft k_i$ and $k_i \blacktriangleleft k_j$.
+However, $c_k \blacktriangleleft k_j$ is also directly represented by the binary relation $c_k \prec k_j$.
 As a result, a duplication of the dependency is created in $\Gamma_{task}$.
@@
 
-% Another view of the relations $\prec$ and $\blacktriangleleft$ is that $c_i\blacktriangleleft c_j$ in $\Gamma_{task}$ is a path from $c_i$ to $c_j$ of any size, while $c_i \prec c_j$ is a path of size $1$ from $c_i$ to $c_j$.
+% Another view of the relations $\prec$ and $\blacktriangleleft$ is that $k_i\blacktriangleleft k_j$ in $\Gamma_{task}$ is a path from $k_i$ to $k_j$ of any size, while $k_i \prec k_j$ is a path of size $1$ from $k_i$ to $k_j$.
 And in any case a path in the graph represents a time dependency between computations.
@@
 % % The dependency graph $\Gamma_{task}(V,A)$ is reduced to the dependency graph $\Gamma_{task}(V,A')$, where $A' \subseteq A$.
-A direct arc $(\overset{\frown}{c_i,c_j})$, $i<j$, of $A$, which means that $c_i \prec c_j$, is included in $A'$ if no longer path $c_i \blacktriangleleft c_j$ exists in $A$ because of transitivity.
+A direct arc $(\overset{\frown}{k_i,k_j})$, $i<j$, of $A$, which means that $k_i \prec k_j$, is included in $A'$ if no longer path $k_i \blacktriangleleft k_j$ exists in $A$ because of transitivity.
 % % \end{mydef}
@@
 % It is also possible to combine data and task parallelization techniques to get hybrid parallelism, sometimes more efficient on hybrid architectures.
-A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list of computations $c_i$ and updated-computations $*_i$.
+A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list of computations $k_i$ and updated-computations $*_i$.
 On the other hand, the task parallelization technique builds, from an ordered list of computation, a graph dependency $\Gamma_{task}$.
@@
 % However, we have to clarify the dependency operator $\prec$ for the specific case of updated-computations.
-As explained in Section~\ref{sect:dataparal}, if $c_j$ is an updated-computations, which means that at least one $c_i$ exists such that $i<j$ and $c_i \pprec c_j$, $c_j$ is transformed to the ordered list of two computations
+As explained in Section~\ref{sect:dataparal}, if $k_j$ is an updated-computations, which means that at least one $k_i$ exists such that $i<j$ and $k_i \pprec k_j$, $k_j$ is transformed to the ordered list of two computations
 
@@
 First, by definition, the dependency $c^*_j \prec *_j$ is created because $\{w^*_j\} \subset R_j \cup \{w^*_j\}$.
-Second, as $R^*_j = \{\bigcup_{i} \{w_i\} \text{, }\forall i<j\text{, and }c_i \pprec c_j\}$, for each $i<j$ the dependency $c_i \prec *_j$ is also created.
+Second, as $R^*_j = \{\bigcup_{i} \{w_i\} \text{, }\forall i<j\text{, and }k_i \pprec k_j\}$, for each $i<j$ the dependency $k_i \prec *_j$ is also created.
 
-% The choice of separating $*_j$ from $c_j$ has been done to entirely expose the eventual task parallelism in the hybrid parallelization.
+% The choice of separating $*_j$ from $k_j$ has been done to entirely expose the eventual task parallelism in the hybrid parallelization.
 For example, Figure~\ref{fig:more} give an exemple where initially $c_0 \pprec c_2$ and $c_0 \prec c_1$, $c_1 \prec c_2$.
@@
 % \end{equation*}
-% The set of computations $\Gamma_{hybrid}$ is a dependency graph between computations $c_i$ and update computations $*_i$, and can be built from the call to 
+% The set of computations $\Gamma_{hybrid}$ is a dependency graph between computations $k_i$ and update computations $*_i$, and can be built from the call to 
 % \begin{equation*}

### 2016-01-22 ARTICLES/IJPP/article/formalism.tex
article IJPP add reduction Helene 22/01/2016

@@
 \begin{mydef}
-A computation kernel $k$ of a numerical simulation is defined as $k(R,(w,D),comp)$, where 
+A computation kernel $k$ of a numerical simulation is defined as $k(S,R,(w,D),comp)$, where 
 \begin{itemize}
+\item $S \in \mathcal{S}$ is the set of scalar to read, 
 \item $(w,D) \in \Delta \times \mathcal{D}$ is the single data modified by the computation kernel onto a given computation domain.
@@
 \item We denote by $identity$ the identity function $x=x$.
-\item A kernel computation $k(R,(w,D),comp)$ is a \emph{stencil kernel} $\iff \exists (r,n) \in R$ such that $n \neq identity$.
-\item A kernel computation $k(R,(w,D),comp)$ is a \emph{local kernel} $\iff \forall (r,n) \in R$, $n = identity$.
-\item A kernel computation $k(R,(w,D),comp)$ is a \emph{reduction kernel} if $w$ is a scalar, in other words $\iff dim(mesh(entity(w)))=0$.
+\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{stencil kernel} $\iff \exists (r,n) \in R$ such that $n \neq identity$.
+\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{local kernel} $\iff \forall (r,n) \in R$, $n = identity$.
+\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} $\iff w$ is a scalar, $R=\{(r,n)\}$, and if $n=entity(r)$.
 \end{itemize}
@@
 
+\noindent \textbf{Property}
+In a reduction kernel $k(S,R,(w,D),comp)$, $D=entity(w)$ as a single entity exists for a scalar.
+
 \medskip
-A reduction computation is a computation which takes data applied onto meshes and returns a data with a dimension reduced to zero, in other words a scalar.
+A reduction computation is a computation which takes a single data applied onto a mesh and returns from all its entities a single data with a dimension reduced to zero, in other words a scalar.
 A reduction is typically used to compute the convergence criteria of the time loop of the simulation.
@@
 \noindent \textbf{Property}
-Considering a reduction kernel $k(R,(w,D),comp)$, $comp$ must be a binary and associative operation on the type $V$, $comp: V \times V \rightarrow V$.
+Considering a reduction kernel $k(S,R,(w,D),comp)$, $comp$ must be a binary and associative operation on the type $V$, $comp: V \times V \rightarrow V$.
 
@@
 
-A stencil program has been formally defined in this section. This formalism is used in the next Section to define two mid-grain parallelization techniques of a multi-stencil program.
+A stencil program has been formally defined in this section. This formalism is used in the next Section to define two parallelization techniques of a multi-stencil program.

### 2016-01-22 ARTICLES/IJPP/article/parallelism.tex
article IJPP add reduction Helene 22/01/2016

@@
 This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
-As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\Delta=\{\Delta_0,\dots,\Delta_{p-1}\}$.
+As entities and as data are mapped onto the mesh, the set of entities and the set of data $\Delta$ are partitionned the same way than the mesh: $\mathcal{E}=\{\mathcal{E},\dots,\mathcal{E}{p-1}\}$, $\Delta=\{\Delta_0,\dots,\Delta_{p-1}\}$.
 The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma_{sync}$.
@@
 \begin{mydef}
-\textit{
-For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $k_i$ and $k_j$, denoted $k_i \pprec k_j$, if $\exists (r,n) \in R_j$ with $n\neq identity$ ($k_j$ is a stencil computation), and such that $\{w_i\} \subset R_j$.
-Moreover, the data to update is $\{w_i\} \cap R_j = \{w_i$\}.}
+For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $k_i$ and $k_j$, denoted $k_i \pprec k_j$, if $\exists (r_j,n_j) \in R_j$ such that $w_i=r_j$ and $n_j\neq identity$ ($k_j$ is a stencil computation).
+Moreover, the data to update is $\{w_i\}$.
 \end{mydef}
 
-Actually, a synchronization can only be needed by the data read by a stencil computation (not local), and only if this data has been modified before, which means that it has been written before.
+Actually, a synchronization is needed by the data read by a stencil computation (not local), and only if this data has been modified before, which means that it has been written before.
 This synchronization is needed because a neighborhood function $n \in \mathcal{N}$ of a stencil computation involves values computed on different resources.
@@
 \begin{mydef}
-\textit{A synchronization between two computations $k_i \pprec k_j$ is defined as a specific computation $sync(R,w,exp,D)$ where $R=\{w_i\}$, $w=w_i$, $exp=identity$ and $D=\emptyset$.}
+A synchronization between two computations $k_i \pprec k_j$ is defined as a specific computation 
+\begin{equation*}
+sync(S,R,w,comp)
+\end{equation*}
+\end{mydef}
+where $S=\emptyset$, $R=\{(w_i,n_j)\}$, $w=(w_i,n_j)$ and $comp=identity$. In other words, $w_i$ has to be synchronized for the neighborhood $n_j$.
+
+\begin{mydef}
+If $k_i \pprec k_j$, $k_j$ is replaced by the list
+\begin{equation*}
+[sync(\emptyset,\{(w_i,n_j)\},(w_i,n_j),identity), k_j]
+\end{equation*}
 \end{mydef}
 
+When a data parallelism is applied, the other type of computation which is responsible for additional synchronizations is the reduction.
+Actually, the reduction is applied locally on each subset of entities, on each resource, which results in $p$ (number of resources) scalar value, one on each resource.
+Thus, to perform the final reduction a synchronization is needed to get the $p$ scalars onto a single resource, to finally get the reduction scalar.
+
 \begin{mydef}
-If a synchronization is needed before $k_j$ (on the data $w_i$), $k_j$ is replaced in $\Gamma_{sync}$ by the list
+For a reduction kernel $k(S,R,(w,D),comp)$, if we denote by $w^i$, $0 \leq i<p$, the local $w$ computed on the resource $i$, a reduction synchronization is defined as the specific computation 
 \begin{equation*}
-\Gamma_{sync} = [sync(\{w_i\},w_i,identity,\emptyset), k_j]
+sync(S,R_s,w,comp)
 \end{equation*}
 \end{mydef}
+where, $S=\emptyset$, $R=\{(w^0,entity(w^0)) \dots (w^{p-1},entity(w^{p-1}))\}$ and $w$ and $comp$ are the same than the actual reduction.
+
+One can notice that both types of synchronizations are performed by all resources.
 
@@
 F_{sync}(\Gamma,j) = 
-\begin{cases} 	[\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) & \mbox{if }\forall i<j \mbox{, } k_i \not\pprec k_j\\
-				[sync(\{w_i\},w_i,identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1) & \forall i<j \text{, and }k_i \pprec k_j\}\\
-				[] & \mbox{if }j=|\Gamma|.
+\begin{cases} 	[] \mbox{ if }j=|\Gamma|\\
+				[sync(\emptyset,\{w_i\},w_i,identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1)\\
+				\text{ } \qquad \forall i<j \text{, and }k_i \pprec k_j\\
+				[k_j, sync(\emptyset,,w_j,comp_j)] \cdot F_{sync}(\Gamma,j+1)\\ 
+				\text{ } \qquad \mbox{if } w_j \in \mathcal{S} \text{, and } R_j=\{(r_j,entity(r_j))\}\\
+				[\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) \text{ else}.
 \end{cases}
@@
 \begin{equation}
-\mathcal{MSP}_k(T,\mathcal{M}_k,\mathcal{E},\mathcal{D},\Delta_k, \mathcal{S},\Gamma_{sync}),
+\mathcal{MSP}_k(T,\mathcal{M}_k,\mathcal{E},\mathcal{D},\Delta_k,\mathcal{S},\Gamma_{sync}),
 \end{equation}

### 2016-01-22 ARTICLES/IJPP/article/parallelism.tex
article IJPP add reduction Helene 22/01/2016

@@
 \begin{equation*}
-sync(S,R,w,comp)
+k_{sync}(S,R,w,comp)
 \end{equation*}
@@
 \begin{equation*}
-[sync(\emptyset,\{(w_i,n_j)\},(w_i,n_j),identity), k_j]
+[k_{sync}(\emptyset,\{(w_i,n_j)\},(w_i,n_j),identity), k_j]
 \end{equation*}
@@
 \begin{equation*}
-sync(S,R_s,w,comp)
+k_{sync}(S,R_s,w,comp)
 \end{equation*}
@@
 \begin{cases} 	[] \mbox{ if }j=|\Gamma|\\
-				[sync(\emptyset,\{w_i\},w_i,identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1)\\
+				[k_{sync}(\emptyset,\{w_i\},w_i,identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1)\\
 				\text{ } \qquad \forall i<j \text{, and }k_i \pprec k_j\\
-				[k_j, sync(\emptyset,,w_j,comp_j)] \cdot F_{sync}(\Gamma,j+1)\\ 
+				[k_j, k_{sync}(\emptyset,,w_j,comp_j)] \cdot F_{sync}(\Gamma,j+1)\\ 
 				\text{ } \qquad \mbox{if } w_j \in \mathcal{S} \text{, and } R_j=\{(r_j,entity(r_j))\}\\
@@
 
-% The construction of $\Gamma_{data}$ implies that if a synchronization is needed it is always proceeded before the computation wich needs it.
-One can notice that a combination of synchronizations is possible in some cases, and thus it is possible to synchronize more than one data at a time.
-%If the number of synchronizations is reduced, final performance of the program will be better.
-% For this reason, we define a reduction which can be applied to $\Gamma_{data}$, and which is performed as a post-transformation.
-This reduction, could propose better performance depending on the used network speed.
-
-% \begin{myprop}
-% Denoting four computations $c_l$, $c_k$, $k_i$ and $k_j$ of $\Gamma$, with $l<k$, $i<j$, and where $c_l \pprec c_k$ and $k_i \pprec k_j$.
-Then, computations $c_k$ and $k_j$ are tansformed to $c^*_k(c_k,update(w_l))$ and $c^*_j(k_j,update(w_i))$.
-If $i<k$, a single update call can be performed at $k$ such that $c^*_k(c_k,update(w_l\cup w_i))$, and in this case, the computation $k_j$ is not transformed.
-% \end{myprop}
-
-% \begin{proof}
-% By definition, the synchronization of $w_i$ can be performed between $k_i$ and $k_j$.
-As $i<k$ the computation $k_i$ is not performed between $k$ and $j$.
-As a result, the synchronization of $w_i$ can be performed between $k_i$ and $c_k$.
-% \end{proof}
-
-% In other words, for each synchronization in $\Gamma_{data}$, if the computation $k_i$ from which $k_j$ is synchronized occurs before the last synchronization, the two synchronizations can be reduced to one with the union of the data to update.
-
- The final step of this parallelization is to run $\Gamma_{sync}$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program, defined by
+ The final step of this parallelization is to run $\Gamma_{sync}$ on each resource. Thus, for each resource $0 \leq m \leq p-1$ a multi-stencil program, defined by
 \begin{equation}
-\mathcal{MSP}_k(T,\mathcal{M}_k,\mathcal{E},\mathcal{D},\Delta_k,\mathcal{S},\Gamma_{sync}),
+\mathcal{MSP}_m(T,\mathcal{M}_m,\mathcal{E}_m,\mathcal{D}_m,\Delta_m,\mathcal{S},\Gamma_{sync}),
 \end{equation}
 is performed.
-
-%We denote this parallelization technique a coarse-grain data parallelization in contrast with the same technique applied to the finer level of a single computation.
-In this case, for a computation $c(R,w,D,\text{exp})$, the local domain $D$ is partitionned for $p$ resources $\{D_0,\dots,D_{p-1}\}$, and each resource $k$ is responsible for a sub-computation $c_k(R,w,D_k,\text{exp}')$.
-%This finer data parallelization technique is not directly adressed by the work presented in this paper, but is exhibited, as it will be explained later.
 
@@
 \begin{mydef}
-For two computations $k_i(R_i,w_i,\text{exp}_i,D_i)$ and $k_j(R_j,w_j,\text{exp}_j,D_j)$, with $i < j$, it is said that $k_j$ is data dependant from $k_i$, denoted $k_i\prec k_j$, if $\{w_i\} \cap R_j \neq \emptyset$.
+For two computations $k_i(S_i,R_i,(w_i,D_i),comp_i)$ and $k_j(S_j,R_j,(w_j,D_j),comp_j)$, with $i < j$, it is said that $k_j$ is data dependant from $k_i$, denoted $k_i\prec k_j$, if $w_i \in R_j$.
 In this case, $k_i$ has to be computed before $k_j$.
@@
 \end{mydef}
-
-% \begin{myprop}
-% The binary relation $\prec$ is not transitive. 
-% \end{myprop}
-
-% \begin{proof}
-% Considering three computations $c_0(R_0,w_0,\text{exp}_0)$, $c_1(R_1,w_1,\text{exp}_1)$ and $c_2(R_2,w_2,\text{exp}_2)$, where $\{w_0\}\cap R_1 \neq \emptyset$ and $\{w_1\}\cap R_2 \neq \emptyset$.
-In this case two dependencies can be extracted $c_0 \prec c_1$ and $c_1 \prec c_2$.
-However, the dependency $c_0 \prec c_2$ is not true as $\{w_0\}\cap R_2 = \emptyset$.
-% \end{proof}
-
-% It has been proved that the binary relation $\prec$ is not transitive.
-This property is due to the fact that $\prec$ is defined as a data dependency between computations.
-However, another type of dependency, only due to time, is created from $\prec$.
-
-% \begin{mydef}
-% For two computations $k_i(R_i,w_i,\text{exp}_i)$ and $k_j(R_j,w_j,\text{exp}_j)$, a time dependency is denoted $k_i \blacktriangleleft k_j$ and means that $k_i$ has to be computed before $k_j$.
-% \end{mydef}
-
-% As a result, the relation $\blacktriangleleft$ is more general than $\prec$.
-
-% \begin{myprop}
-% For two computations $k_i(R_i,w_i,\text{exp}_i)$ and $k_j(R_j,w_j,\text{exp}_j)$ such that $k_i \prec k_j$, $k_i \blacktriangleleft k_j$ is verified.
-% \end{myprop}
-
-% \begin{proof}
-% By definition $k_i \prec k_j$ means that $\{w_i\}\cap R_j \neq \emptyset$ and that $k_i$ has to computed before $k_j$.
-%Thus, the relation $k_i \prec k_j$ is stronger than $k_i \blacktriangleleft k_j$.
-% \end{proof}
-
-% \begin{myprop}
-% The binary relation $\blacktriangleleft$ is transitive. 
-% \end{myprop}
-
-% \begin{proof}
-% Considering three computations $c_0(R_0,w_0,\text{exp}_0)$, $c_1(R_1,w_1,\text{exp}_1)$ and $c_2(R_2,w_2,\text{exp}_2)$, where $c_0 \blacktriangleleft c_1$ and $c_1 \blacktriangleleft c_2$.
-$c_0$ is computed before $c_1$ and $c_1$ is computed before $c_2$, as a result $c_0$ is computed before $c_2$ and the relation $c_0 \blacktriangleleft c_2$ is verified.
-% \end{proof}
 
@@
 
-From an ordered list of computations $\Gamma_{sync}$, a directed dependency graph $\Gamma_{dep}(V,A)$ can be built finding all pairs of computations $k_i(R_i,w_i,\text{exp}_i,D_i)$ and $k_j(R_j,w_j,\text{exp}_j,D_j)$, with $i<j$, such that $k_i \prec k_j$.
-%Those time dependencies can be found from pairs of data dependencies $k_i \prec k_j$.
+From an ordered list of computations $\Gamma_{sync}$, a directed dependency graph $\Gamma_{dep}(V,A)$ can be built finding all pairs of computations $k_i(S_i,R_i,(w_i,D_i),comp_i)$ and $k_j(S_j,R_j,(w_j,D_j),comp_j)$, with $i<j$, such that $k_i \prec k_j$.
 
@@
 \begin{mydef}
-From the ordered list $\Gamma_{sync}$ of computations $c(R,w,\text{exp},D)$, a directed dependency graph $\Gamma_{dep}(V,A)$ is obtained from the call $T_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function
+From the ordered list $\Gamma_{sync}$ of computation kernels, a directed dependency graph $\Gamma_{dep}(V,A)$ is obtained from the call $T_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function
 \begin{equation*}
 F_{dep}(\Gamma_{sync},j) = 
-\begin{cases} 	(\{\},\{\}) & \mbox{if }j=|\Gamma_{sync}|\\
-				(k_j, \{(\overset{\frown}{k_i,k_j})\mbox{, }\forall i < j \mbox{, } k_i\prec k_j \})\cup F_{dep}(\Gamma_{sync},j+1) & \mbox{if }j<|\Gamma_{sync}|
+\begin{cases} 	(\{\},\{\}) \mbox{ if }j=|\Gamma_{sync}|\\
+				(k_j, \{(\overset{\frown}{k_i,k_j})\mbox{, }\forall i < j \mbox{, } k_i\prec k_j \})\\
+				\text{ } \qquad \cup F_{dep}(\Gamma_{sync},j+1) \mbox{ if }j<|\Gamma_{sync}|
 \end{cases}
@@
 
-As a result, the hybrid parallelization of a multi-stencil program $\mathcal{MSP}$, produces $k$ multi-stencil programs
+As a result of the hybrid parallelization, each resource $0 \leq m \leq p-1$ perform a multi-stencil program, defined by
 \begin{equation*}
-\mathcal{MSP}_k(T,\mathcal{M}_k,\mathcal{E},\mathcal{D},\Delta_k,\Gamma_{dep}).
+\mathcal{MSP}_m(T,\mathcal{M}_m,\mathcal{E}_m,\mathcal{D}_m,\Delta_m,\Gamma_{dep}).
 \end{equation*}
-The set of computations $\Gamma_{dep}$ is a dependency graph between computations $k_i$ of $\Gamma$ and the synchronization computations $synk_i$ added into $\Gamma_{sync}$.
+The set of computations $\Gamma_{dep}$ is a dependency graph between computations $k_i$ of $\Gamma$ and the synchronization kernels added into $\Gamma_{sync}$.
 $\Gamma_{dep}$ can be built from the call to
@@
 \end{equation*}
-
-% Using the function $T_{task}$ to build $\Gamma_{task}$, however, duplication of dependencies may occur because of the transitivity of the relation $\blacktriangleleft$.
-Actually, as the relation $\prec$ verifies the relation $\blacktriangleleft$, and as $\blacktriangleleft$ is transitive, if for three computations $c_k(R_k,w_k,\text{exp}_k)$, $k_i(R_i,w_i,\text{exp}_i)$ and $k_j(R_j,w_j,\text{exp}_j)$, with $k<i<j$ and where $c_k \prec k_i$, $k_i \prec k_j$ and $c_k \prec k_j$, then the transitivity $c_k \blacktriangleleft k_j$ is verified because of the relations $c_k \blacktriangleleft k_i$ and $k_i \blacktriangleleft k_j$.
-However, $c_k \blacktriangleleft k_j$ is also directly represented by the binary relation $c_k \prec k_j$.
-As a result, a duplication of the dependency is created in $\Gamma_{task}$.
-In Figure~\ref{fig:duplication} an example of duplication is given for the relation $c_1 \blacktriangleleft c_4$.
-
-% \begin{figure}[h!]
-% \begin{center}
-% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%    \node[] (c0) at (0,0) {$c_0$};
-%    \node[] (c1) at (1,0) {$c_1$};
-%    \node[] (c2) at (2,0) {$c_2$};
-%    \node[] (c3) at (3,0) {$c_3$};
-%    \node[] (c4) at (4,0) {$c_4$};
-
-%   \path[->]
-%     (c0) edge node {} (c1)
-%     (c1.east) edge node {} (c2)
-%     	 edge [bend right=50] node [swap] {} (c4)
-%     (c2) edge node {} (c3)
-%     (c3) edge node {} (c4);
-%   \end{tikzpicture}
-%   \caption{Useless duplication of the dependency $c_1 \blacktriangleleft c_4$}
-%   \label{fig:duplication}
-% \end{center}
-% \end{figure}
-
-% Another view of the relations $\prec$ and $\blacktriangleleft$ is that $k_i\blacktriangleleft k_j$ in $\Gamma_{task}$ is a path from $k_i$ to $k_j$ of any size, while $k_i \prec k_j$ is a path of size $1$ from $k_i$ to $k_j$.
-And in any case a path in the graph represents a time dependency between computations.
-Scheduling a graph dependency which do not contains duplication of information is easier.
-% %For this reason, we define a reduction to apply to $\Gamma_{task}$, and which is performed as a post-transformation.
-In the example of Figure~\ref{fig:duplication} the dependency represented by the path $c_1 \blacktriangleleft c_2 \blacktriangleleft c_3 \blacktriangleleft c_4$ gives more information than the one directly represented by $c_1 \prec c_4$.
-The proposed reduction is based on this property.
-
-% % \begin{mydef}
-% % The dependency graph $\Gamma_{task}(V,A)$ is reduced to the dependency graph $\Gamma_{task}(V,A')$, where $A' \subseteq A$.
-A direct arc $(\overset{\frown}{k_i,k_j})$, $i<j$, of $A$, which means that $k_i \prec k_j$, is included in $A'$ if no longer path $k_i \blacktriangleleft k_j$ exists in $A$ because of transitivity.
-% % \end{mydef}
-% \begin{mydef}
-% A DAG is transitive if it contains an arc $(\overset{\frown}{u,v})$ between any two vertices such that there is a path from $u$ to $v$.
-The transitive closure of a DAG $G=(V,E)$, is the DAG $G_T=(V,E_T)$ for which $E_T$ is the minimal subset of $V \times V$ that includes $E$ and makes $G_T$ transitive.
-% \end{mydef}
-
-% \begin{mydef}
-% An arc $(\overset{\frown}{u,v})$ of a DAG is redundant if there is a path from $u$ to $v$ in the DAG that does not include the edge.
-A DAG that does not contain any redundant arc is called minimal.
-The transitive reduction of a DAG $G$ is its unique minimal sub-graph having the same transitive closure.
-% \end{mydef}
-
-% Thus, the removal of the redundant dependencies of $\Gamma_{task}$ consists in applying a transitive reduction.
-
-%------------------------------
-% \subsection{Hybrid parallelism}
-% It is also possible to combine data and task parallelization techniques to get hybrid parallelism, sometimes more efficient on hybrid architectures.
-A coarse-grain data parallelism creates for $k$ resources $k$ multi-stencil computations $\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{data})$, where $\Gamma_{data}$ is built from an ordered list of computations and is itself an ordered list of computations $k_i$ and updated-computations $*_i$.
-On the other hand, the task parallelization technique builds, from an ordered list of computation, a graph dependency $\Gamma_{task}$.
-As a result $T_{data}$ and $T_{task}$ can be composed to build $\Gamma_{hybrid}$.
-
-% However, we have to clarify the dependency operator $\prec$ for the specific case of updated-computations.
-As explained in Section~\ref{sect:dataparal}, if $k_j$ is an updated-computations, which means that at least one $k_i$ exists such that $i<j$ and $k_i \pprec k_j$, $k_j$ is transformed to the ordered list of two computations
-
-% \begin{equation*}
-% [*_j(R^*_j,w^*_j),c^*_j(R_j \cup \{w^*_j\},w_j,\text{exp}_j)].
-% \end{equation*}
-
-% As a result, and because $*_j$ is itself a computation, it appears dependencies relations.
-First, by definition, the dependency $c^*_j \prec *_j$ is created because $\{w^*_j\} \subset R_j \cup \{w^*_j\}$.
-Second, as $R^*_j = \{\bigcup_{i} \{w_i\} \text{, }\forall i<j\text{, and }k_i \pprec k_j\}$, for each $i<j$ the dependency $k_i \prec *_j$ is also created.
-
-% The choice of separating $*_j$ from $k_j$ has been done to entirely expose the eventual task parallelism in the hybrid parallelization.
-For example, Figure~\ref{fig:more} give an exemple where initially $c_0 \pprec c_2$ and $c_0 \prec c_1$, $c_1 \prec c_2$.
-Thus, $c_2$ is replaced first in $\Gamma_{data}$ by $[*_2(\{w_0\},w^*_2),c^*_2(R_2 \cup \{w^*_2\},\text{exp}_2)]$.
-As a result, when applying $T_{task}$ on $\Gamma_{data}$, it appears that $*_2$ can be performed in parallel with $c_1$.
-This is due to the fact that $*_2$ is splitted from from $c_2$ and that $R^*_2=\{w_0\}$ is a subset of $R_2=\{w_0,w_1\}$ and not equal to $R_2$.
-
-% \begin{figure}[h!]
-% \begin{center}
-% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%    \node[] (c0) at (0,0) {$c_0$};
-%    \node[] (c1) at (2,0) {$c_1$};
-%    \node[] (star2) at (2,1) {$*_2$};
-%    \node[] (c2) at (4,0) {$c^*_2$};
-
-%   \path[->]
-%     (c0) edge node {} (c1)
-%          edge node {} (star2)
-%     (c1) edge node {} (c2)
-%     (star2) edge node {} (c2);
-%   \end{tikzpicture}
-%   \caption{Example of hybrid parallelization.}
-%   \label{fig:more}
-% \end{center}
-% \end{figure}
-
-% As a result, the hybrid parallelization of a multi-stencil program $\mathcal{MSP}$, produces $k$ multi-stencil programs
-% \begin{equation*}
-% \mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma_{hybrid}).
-% \end{equation*}
-% The set of computations $\Gamma_{hybrid}$ is a dependency graph between computations $k_i$ and update computations $*_i$, and can be built from the call to 
-% \begin{equation*}
-% T_{task}(T_{data}(\Gamma,0),0).
-% \end{equation*}
-
-% The rest of this paper deals with the hybrid parallelization of multi-stencil programs described above.
-The reason of this choice is the fact that by adressing the hybrid parallelism case, we also are able to adress data parallelism and task parallelism, from which hybrid parallelism is built.

### 2016-01-22 ARTICLES/IJPP/article/article.tex
article IJPP add reduction Helene 22/01/2016

@@
 \usepackage{listings}
+\usepackage{mathtools, cuted}
 \usepackage{tikz}
@@
 \input{msl.tex}
-% The parallel empty skeleton of the application can be computed automatically from the language :
-% \begin{itemize}
-% \item grammar
-% \item example from the language to the dependency graph
-% \end{itemize}
+The parallel empty skeleton of the application can be computed automatically from the language :
+\begin{itemize}
+\item grammar
+\item example from the language to the dependency graph
+\end{itemize}
 %------------------------------------------------------------------------------
@@
 \label{sect:msp}
-\input{msp.tex}
-% From the dependency graph we study a first dump: static scheduling
-% \begin{itemize}
-% \item description of the algorithm to get the binary tree decomposition of a minimal serie-parallel graph
-% \item use of the serie-parallel tree decomposition to detect fusions of kernels
-% \item example (the same one) from the dependency graph to the binary tree
-% \end{itemize}
+%\input{msp.tex}
+From the dependency graph we study a first dump: static scheduling
+\begin{itemize}
+\item description of the algorithm to get the binary tree decomposition of a minimal serie-parallel graph
+\item use of the serie-parallel tree decomposition to detect fusions of kernels
+\item example (the same one) from the dependency graph to the binary tree
+\end{itemize}
 %------------------------------------------------------------------------------

### 2016-01-25 ARTICLES/IJPP/article/formalism.tex
article IJPP corrections formalism et parallelism Helene 25/01/2016

@@
 \begin{itemize}
-\item An entity $e$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $e\subset (V\cup E)$.
+\item An entity $e$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $e\subset V\cup E$.
 \item A mesh entities group is denoted $E$ such that $E \in \mathcal{P}(V\cup E)$. It represents a set of entities of the same type.
-\item Finally the set of mesh entities groups of a simulation is denoted $\mathcal{E}$.
+\item The set of mesh entities groups of a simulation is denoted $\mathcal{E}$.
+\item Finally, the mesh on which is applied a group of mesh entities $E$ is denoted $mesh(E)$
 \end{itemize}
@@
 
-In a numerical simulation, as a set of scalars can be applied onto a mesh with a nil dimension, a set of data elements, or quantities, can be applied onto meshes of dimension superior or equal to zero.
+In a numerical simulation, as a set of scalars can be applied onto a mesh with a nil dimension, a set of data elements, or quantities, can be applied onto meshes of dimension superior to zero.
 Those quantities represent, as well as scalars, the set of values to compute, or to use, for computations.
 
+\medskip
 \noindent \textbf{Definitions}
@@
 \begin{itemize}
-\item \textit{A quantity (or data) is a function $\delta: E_{\delta} \mapsto V_{\delta}$ which associates each entity of a group $e\in E_{\delta}$ to a value $v\in V_{\delta}$ where $V_{\delta}$ is a data type, typically $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.
+\item A quantity (or data) is a function $\delta: E_{\delta} \mapsto V_{\delta}$ which associates each entity of a group $e\in E_{\delta}$ to a value $v\in V_{\delta}$ where $V_{\delta}$ is a data type, typically $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.
 \item The set of quantities applied onto the mesh is denoted $\Delta$.
-\item In the rest of this paper, the group of mesh entities on which a quantity $\delta$ is mapped is denoted $entity(\delta)=E_{\delta}$.}
+\item In the rest of this paper, the group of mesh entities on which a quantity $\delta$ is mapped is denoted $entity(\delta)=E_{\delta}$.
 \end{itemize}
@@
 
+\medskip
 \noindent \textbf{Definitions}
@@
 
+\medskip
 \noindent \textbf{Property}

### 2016-01-25 ARTICLES/IJPP/article/parallelism.tex
article IJPP corrections formalism et parallelism Helene 25/01/2016

@@
 \begin{equation*}
-k_{sync}(S,R,w,comp)
+k_{sync}(S,R,(w,D),comp)
 \end{equation*}
 \end{mydef}
-where $S=\emptyset$, $R=\{(w_i,n_j)\}$, $w=(w_i,n_j)$ and $comp=identity$. In other words, $w_i$ has to be synchronized for the neighborhood $n_j$.
+where $S=\emptyset$, $R=\{(w_i,\bigcup_{e \in entity(w_i)} n_j(e))\}$, $w=(w_i,\bigcup_{e \in entity(w_i)} n_j(e)))$ and $comp=identity$.
+In other words, $w_i$ has to be synchronized for the neighborhood $n_j$ for all entities of $w_i$.
 
@@
 
-When a data parallelism is applied, the other type of computation which is responsible for additional synchronizations is the reduction.
-Actually, the reduction is applied locally on each subset of entities, on each resource, which results in $p$ (number of resources) scalar value, one on each resource.
+When data parallelism is applied, the other type of computation which is responsible for additional synchronizations is the reduction.
+Actually, the reduction is applied locally on each subset of entities, on each resource, which results in $p$ (number of resources) scalar values, one on each resource.
 Thus, to perform the final reduction a synchronization is needed to get the $p$ scalars onto a single resource, to finally get the reduction scalar.
@@
 \begin{equation*}
-k_{sync}(S,R_s,w,comp)
+k_{sync}(S,R_s,(w,D),comp)
 \end{equation*}
 \end{mydef}
-where, $S=\emptyset$, $R=\{(w^0,entity(w^0)) \dots (w^{p-1},entity(w^{p-1}))\}$ and $w$ and $comp$ are the same than the actual reduction.
+where, $S=\emptyset$, $R_s=\{(w^0,entity(w^0)) \dots (w^{p-1},entity(w^{p-1}))\}$, and $w$, $D=entity(w)$ and $comp$ are the same than the actual reduction kernel $k$.
 
@@
 \begin{mydef}
-The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[k_i]_{0 \leq i \leq n-1}$ and $l_2=[c'_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[c_0,\dots,c_{n-1},c'_0,\dots,c'_{m-1}]$.
+The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[k_i]_{0 \leq i \leq n-1}$ and $l_2=[k'_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[k_0,\dots,k_{n-1},k'_0,\dots,k'_{m-1}]$.
 \end{mydef}
@@
 F_{sync}(\Gamma,j) = 
-\begin{cases} 	[] \mbox{ if }j=|\Gamma|\\
-				[k_{sync}(\emptyset,\{w_i\},w_i,identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1)\\
+\begin{cases} 	\bullet [] \mbox{ if }j=|\Gamma|\\
+				\bullet [k_{sync}(\emptyset,\\
+				\{(w_i,\bigcup_{e \in entity(w_i)} n_j(e))\},\\
+				(w_i,\bigcup_{e \in entity(w_i)} n_j(e))),\\
+				identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1)\\
 				\text{ } \qquad \forall i<j \text{, and }k_i \pprec k_j\\
-				[k_j, k_{sync}(\emptyset,,w_j,comp_j)] \cdot F_{sync}(\Gamma,j+1)\\ 
+				\bullet [k_j, k_{sync}(\emptyset,\\
+				\{(w^0_j,entity(w^0_j)) \dots (w^{p-1}_j,entity(w^{p-1}_j))\},\\
+				(w_j,entity(w_j)),comp_j)] \cdot F_{sync}(\Gamma,j+1)\\ 
 				\text{ } \qquad \mbox{if } w_j \in \mathcal{S} \text{, and } R_j=\{(r_j,entity(r_j))\}\\
-				[\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) \text{ else}.
+				\bullet [\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) \text{ else}.
 \end{cases}
@@
 \begin{mydef}
-From the ordered list $\Gamma_{sync}$ of computation kernels, a directed dependency graph $\Gamma_{dep}(V,A)$ is obtained from the call $T_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function
+From the ordered list $\Gamma_{sync}$ of computation kernels, a directed dependency graph $\Gamma_{dep}(V,A)$ is obtained from the call $F_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function
 \begin{equation*}
 F_{dep}(\Gamma_{sync},j) = 
-\begin{cases} 	(\{\},\{\}) \mbox{ if }j=|\Gamma_{sync}|\\
-				(k_j, \{(\overset{\frown}{k_i,k_j})\mbox{, }\forall i < j \mbox{, } k_i\prec k_j \})\\
+\begin{cases} 	\bullet (\{\},\{\}) \mbox{ if }j=|\Gamma_{sync}|\\
+				\bullet (k_j, \{(\overset{\frown}{k_i,k_j})\mbox{, }\forall i < j \mbox{, } k_i\prec k_j \})\\
 				\text{ } \qquad \cup F_{dep}(\Gamma_{sync},j+1) \mbox{ if }j<|\Gamma_{sync}|
@@
 \end{equation*}
-The set of computations $\Gamma_{dep}$ is a dependency graph between computations $k_i$ of $\Gamma$ and the synchronization kernels added into $\Gamma_{sync}$.
+The set of computations $\Gamma_{dep}$ is a dependency graph between computations $k_i$ of $\Gamma$ and synchronizations of kernels added into $\Gamma_{sync}$.
 $\Gamma_{dep}$ can be built from the call to

### 2016-01-26 ARTICLES/IJPP/article/msl.tex
article IJPP Helene 26/01/2016

@@
 iteration ::= num | criteria
-criteria ::= scalar "=" criteriaid "(" 
-                    listscalarread "," 
-                    listdataread ")"
+criteria ::= criteriaid "(" listscalarread ")"
 listscalarread ::= scalar listscalarread 
@@
                           |  dataread
-dataread ::= dataid "[" neighborid "]" 
+dataread ::= dataid "[" stencil "]" 
               |  dataid
@@
 
-\begin{figure}[t]
+\begin{figure}[!h]
   \hspace{5mm}
-  \begin{minipage}[t]{\textwidth}
-    %  \subfloat[\label{fig:grammar}]
-    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left,linewidth=.87\textwidth]{grammar.txt}}   
+  \begin{minipage}[!h]{0.5\textwidth}
+    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=right]{grammar.txt}}   
     \caption{Grammar of the Multi-Stencil Language. \label{fig:grammar}}

### 2016-01-26 ARTICLES/IJPP/article/msp.tex
article IJPP Helene 26/01/2016

In this section we argue that the graphs $\Gamma_{dep}$, previously defined, can be transformed to a minimal series-parallel directed acyclic graph. As a result, this graphs can be represented as a binary tree of parallel and series compositions of sub-graphs, also called \emph{binary decomposition tree}~\cite{Valdes:1979:RSP:800135.804393}. The obtained binary tree is a static scheduling of $\Gamma_{dep}$ and can also be used to optimize the code including data parallelism only.

%--------------------
\subsection{Minimal series-parallel DAG}
\begin{mydef}
A vertex $v$ of a DAG $G$ is a \emph{source} if no edge of $G$ enters $v$. Similarly, a vertex $v$ is a \emph{sink} if no edge of $G$ leaves $v$. 
\end{mydef}

In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of minimal series-parallel DAGs (SPD).

\begin{mydef}Minimal Series Parallel DAG
\begin{itemize}
\item The DAG having a single vertex and no edges is SPD.
\item If $G_1=(V_1,E_1)$ and $G_2=(V_2,E_2)$ are two SPD DAGs, so is either of the DAGs constructed by the following operations:
\begin{itemize}
\item Parallel composition: $G_p=(V_1\cup V_2,E_1\cup E_2)$.
\item Series composition: $G_s=(V_1\cup V_2,E_1\cup E_2\cup (N_1 \times R_2))$, where $N_1$ is the set of sinks of $G_1$ and $R_2$ is the set of sources of $G_2$.
\end{itemize}
\end{itemize}
\end{mydef}

Finally, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or subgraph, called $N$ and represented in Figure~\ref{fig:over}, such that

A possible definition of a N-Shape is
\begin{mydef}
For two vertices $v_1$ and $v_2$ of a DAG which have a Lowest Common Ancestor $v_{lca}$, for all $v_p$ which is either a parent vertex of $v_1$ or $v_2$ and which is different from $v_{lca}$, a N-Shape is formed by the induced subgraph containing vertices $v_1$, $v_2$, $v_{lca}$ and $v_p$.
\end{mydef}

\begin{myth}
The transitive reduction of a DAG $G$ is SPD if and only if it does not contain $N$ as an induced subgraph.
\end{myth}

\begin{myprop}
The transitive reduction of $\Gamma_{dep}$ is not SPD. Actually, it is possible in a multi-stencil program to have a set of computations such that their dependencies form the forbidden $N$ subgraph.
\end{myprop}

\begin{proof}
Considering four computation kernels $k_0$, $k_1$, $k_2$ and $k_3$ of a $\mathcal{MSP}$ such that $k_0 \prec k_1$, $k_2 \prec k_1$ and $k_2 \prec k_3$ are verified. Thus, the \emph{zigzag} relation $k_0 \prec k_1 \succ k_2 \prec k_3$ which form a forbidden subgraph of a SPD is found in $\Gamma_{dep}$.
\end{proof}

To remove the forbidden N-shapes from $\Gamma_{dep}=(V,E)$, we have chosen to apply an over-constraint with the relation $k_0 \prec k_3$, such that a complete bipartite graph is created for the sub-dag, and can be translated to a series-parallel decomposition as illustrated in Figure~\ref{fig:allover}. The algorithm is inspired from the one proposed by~\cite{Mitchell:2004:CMV:1082101.1082117}, and is described in Algorithm~\ref{alg:nshape}. We precisely describe how we find edges to add to get rid of N-Shapes. One can notice that another solution would have been to remove the extra-edge from the N-Shape instead of adding one~\cite{}. However, we have chosen to exactly keep dependencies computed in $\Gamma_{dep}$, to not approximate the static scheduling. Finally, algorithm proposed by~\cite{Bender200575} to find a Lowest Common Ancestor in a DAG (in $\mathcal{O}(|V|^3)$, could have been also used.

The first step of Algorithm~\ref{alg:nshape} is to create a new DAG $\Gamma_{dep}'$ from $\Gamma_{dep}$, such that vertices having exactly one source and one destination in $\Gamma_{dep}$ is not added to $\Gamma_{dep}'$. This step is called $RTV(\Gamma_{dep})$ (line 2) and is done in $\mathcal{O}(|V|)$. The second step is to get source vertices in $\Gamma_{dep}'$. The third step calls $BFSnshape$ (line 4). This procedure is based on a Breadth First Search recursive algorithm (in $\mathcal{O}(|V|+|E|)$) and takes as input a list of source vertices $sources$, $\Gamma_{dep}$ and $\Gamma_{dep}'$. At each recursive call the induced sub-graph of $\Gamma_{dep}'$ containing $sources$ vertices and the list of their successors $successors(sources)$ is created and called $SG$. If this sub-graph $SG$ is connected and is not a complete bipartite graph, the algorithm make the subgraph $SG$ complete by adding needed edges into $\Gamma_{dep}'$ (line 12 in $\mathcal{O}(|V|^2$). The algorithm also add same edges to $\Gamma_{dep}$ (line 13). Finally, line 5, the transitive closure of $\Gamma_{dep}$ is computed.

\begin{algorithm}
\caption{Remove N-Shapes}
\label{alg:nshape}
\begin{algorithmic}[1]
\Procedure{removeNSH} {$\Gamma_{dep}$}
\\$\Gamma_{dep}'$ = RTV($\Gamma_{dep}$)
\\$sources$ = sources($\Gamma_{dep}'$)
\\BFSnshape($sources$,$\Gamma_{dep}'$,$\Gamma_{dep}$)
\\TC($\Gamma_{dep}$)
\EndProcedure

\Procedure{BFSnshape} {$sources$, $\Gamma_{dep}'$, $\Gamma_{dep}$}
\\$successors$ = successors($sources$)
\\$SG$ = subgraph($sources \cup successors$)
\If{isconnected($SG$)}
\If{! complete($SG$)}
\\addedges($\Gamma_{dep}'$)
\\addedges($\Gamma_{dep}$)
\\BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
\EndIf
\If{! isconnected($SG$)}
\\BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
\EndIf
\EndIf
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{figure}[h!]
\begin{center}
\subfloat[][Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$k_0$};
   \node[] (bc0) at (-1,0) {};
   \node[] (c1) at (2,0) {$k_1$};
   \node[] (c2) at (0,-1) {$k_2$};
   \node[] (bc2) at (-1,-1) {};
   \node[] (c3) at (2,-1) {$k_3$};

  \path[->]
    (bc0) edge [dotted] node {} (c0)
    (bc2) edge [dotted] node {} (c2)
    (c0) edge node {} (c1)
          edge [dashed] node [swap] {} (c3)
    (c2) edge node {} (c1)
        edge node {} (c3);
  \end{tikzpicture}
}
\hspace{50pt}
\subfloat[][Series-parallel tree associated to the over-constraint\label{fig:treeover}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (S0) at (0,0) {$\mathcal{S}$};
   \node[] (bS0) at (-1,-1) {};
   \node[] (P1) at (-1,1) {$\mathcal{P}$};
   \node[] (aP1) at (-1.5,2) {};
   \node[] (aaP1) at (-0.5,2) {};
   \node[] (P2) at (1,1) {$\mathcal{P}$};
   \node[] (c0) at (-2,3) {$k_0$};
   \node[] (c1) at (0.5,3) {$k_1$};
   \node[] (c2) at (0,3) {$k_2$};
   \node[] (c3) at (2,3) {$k_3$};

  \path[->]
    (bS0) edge [dotted] node {} (S0)
    (S0) edge node {} (P1)
         edge node {} (P2)
    (P1) edge [dotted] node [swap] {} (aP1)
          edge [dotted] node [swap] {} (aaP1)
    (aP1) edge node {} (c0)
    (aaP1) edge node {} (c2)
    (P2) edge node {} (c1)
        edge node {} (c3);
  \end{tikzpicture}
}
\caption{Deletion of forbidden subgraphs.}
\label{fig:allover}
\end{center}
\end{figure}

After these over-constraints are applied, $\Gamma_{dep}$ is a SPD. As a result, the binary tree decomposition algorithm of Valdes \& Al can be applied on $\Gamma_{dep}$ to get a static scheduling of the multi-stencil program.

%--------------------
\subsection{Series-parallel binary tree decomposition}
A \emph{binary decomposition tree} is a tree having a leaf for each vertex of the SPD DAG it represents, and whose internal nodes are labelled $S$ or $P$ to indicate respectively the series or parallel composition of the SPD sub-DAGs represented by the subtrees rooted at $S$ or $P$.

Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have also proposed a linear algorithm to know if a DAG is SPD and, if it is, to decompose it to its associated binary decomposition tree. This algorithm is based on the duality of the class of SPD DAGs, with the class of \emph{Two Terminal Series Parallel} DAGs (TTSP) from which a binary tree decomposition can be performed in linear time.

\begin{mydef}
The \emph{line digraph} of a digraph (directed graph) $G$ is a digraph $L(G)$ that has:
\begin{itemize}
\item a vertex $f(e)$ for each edge $e$ of $G$; and
\item an edge $(f(e_1),f(e_2))$ for each pair of edges of $G$ of the form $e_1=(u,v)$, $e_2=(v,w)$.
\end{itemize}
\end{mydef}

\begin{mydef}Two Terminal Series Parallel
\begin{itemize}
\item A digraph consisting of two vertices joined by a single edge is TTSP.
\item If $G_1$ and $G_2$ are TTSP digraphs, so is the digraph obtained by either of the following operations:
\begin{itemize}
\item \emph{Two terminal parallel composition}: identify the sourc of $G_1$ with the source of $G_2$ and the sink of $G_1$ with the sink of $G_2$.
\item \emph{Two terminal series composition}: identify the sink of $G_1$ with the source of $G_2$.
\end{itemize}
\end{itemize}
\end{mydef}

\begin{myth}
If the DAG $G$ is a SPD graph, its \emph{inverse line DAG} $L^{-1}(G)$ is TTSP.
\end{myth}

It has to be indicated that for a DAG $G$, $L^{-1}(G)$ is not unique. Thus, in the work of Valdes \& Al, and in this work, $L^{-1}(G)$ will refer to the unique digraph having a single source and a single sink whose line digraph is $L(L^{-1}(G))=G$.

Once the removal of N-Shapes has been applied onto $\Gamma_{dep}$, the linear series-parallel tree decomposition can be applied following the algorithm of~\cite{Valdes:1979:RSP:800135.804393}, called in this paper $BDT(L^{-1}(\Gamma_{dep}))$.

The binary decomposition tree algorithm is based on the fact that the decomposition can be obtained as a byproduct of a reduction process on $L^{-1}(\Gamma_{dep})$, which is TTSP. In order to obtain the decomposition, Valdes \& Al associate a label with each edge of the digraph being reduced. Initially the label of each edge is a trivial binary tree consisting of a single node. As the reduction process introduces new edges the rules of Figure~\ref{fig:rules} are used to compute the binary trees used to label them. The algorithm ends when the TTSP graph is reduced to its minimum, i.e.\ two vertices and a single edge between them.

\begin{figure}[h!]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
%FIRST RULE
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};

  \path[->]
    (x) edge node [above] {$T_1$} (y)
    (y) edge node [above] {$T_2$} (z);
  \end{tikzpicture}
  }
\hspace{5pt}
%ARROW
\subfloat[]{
$\Rightarrow$
}
\hspace{5pt}
%SECOND RULE
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
   \node[] (s) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (t1) at (1,1.5) {$T_1$};
   \node[] (t2) at (2,1.5) {$T_2$};

  \path[->]
    (x) edge node {} (y)
    (s) edge node {} (t1)
        edge node {} (t2);
  \end{tikzpicture}
}
\\
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};

  \path[->]
    (x) edge [bend left] node [above] {$T_1$} (y)
        edge [bend right] node [below] {$T_2$} (y);
  \end{tikzpicture}
  }
\hspace{5pt}
\subfloat[]{
$\Rightarrow$
}
\hspace{5pt}
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
   \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
   \node[] (t1) at (1,1.5) {$T_1$};
   \node[] (t2) at (2,1.5) {$T_2$};

  \path[->]
    (x) edge node {} (y)
    (p) edge node {} (t1)
        edge node {} (t2);
  \end{tikzpicture}
}
\caption{Reduction rules of the decomposition tree algorithm.}
\label{fig:rules}
\end{center}
\end{figure}

%--------------------
\subsection{Canonical Form and Optimizations}

### 2016-01-26 ARTICLES/IJPP/article/parallelism.tex
article IJPP Helene 26/01/2016

@@
 
-More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta, \mathcal{S},\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
+More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\mathcal{N},\Delta, \mathcal{S},\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
 This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
@@
 \begin{equation}
-\mathcal{MSP}_m(T,\mathcal{M}_m,\mathcal{E}_m,\mathcal{D}_m,\Delta_m,\mathcal{S},\Gamma_{sync}),
+\mathcal{MSP}_m(T,\mathcal{M}_m,\mathcal{E}_m,\mathcal{D}_m,\mathcal{N},\Delta_m,\mathcal{S},\Gamma_{sync}),
 \end{equation}
@@
 \begin{equation*}
-\mathcal{MSP}_m(T,\mathcal{M}_m,\mathcal{E}_m,\mathcal{D}_m,\Delta_m,\Gamma_{dep}).
+\mathcal{MSP}_m(T,\mathcal{M}_m,\mathcal{E}_m,\mathcal{D}_m,\mathcal{N},\Delta_m,\Gamma_{dep}).
 \end{equation*}

### 2016-01-26 ARTICLES/IJPP/article/compiler.tex
article IJPP compiler Helene 26/01/2016

--- 2015-06-12 rapport/msp.tex
@@
-In this section we argue that the graphs $\Gamma_{hybrid}$ or $\Gamma_{task}$, previously defined, on which an approximation will be defined, are minimal series-parallel graphs.
-For this reason, the structure of those graphs can be represented as binary trees of parallel and series compositions of sub-graphs, also called \emph{binary decomposition tree}~\cite{Valdes:1979:RSP:800135.804393}.
-First, the needed definitions on series-parallel graphs are given, and second the case of the dag $\Gamma_{hybrid}$ is studied.
+%--------------------
+\subsection{Overall compiler algorithm}
+
+The overall compiler algorithm is represented in Algorithm~\ref{alg:compiler}.
+First, the parsing of the input file (written with the MSL language) gave enough information to build $\Gamma$, the ordered list of computations.
+Second, $\Gamma_{sync}$ and $\Gamma_{dep}$ can be built from $\Gamma$ using the recursive functions $F_{sync}$ and $F_{dep}$.
+Third, N-Shapes are removed from $\Gamma_{dep}$ to get a SPD graph, and to be able to create its serie-parallel binary tree decomposition $TSP$.
+
+\begin{algorithm}
+\caption{MSL Compiler}
+\label{alg:compiler}
+\begin{algorithmic}[1]
+\Procedure{compileMSL} {file}
+\\$\Gamma$ = parser(file)
+\\$\Gamma_{sync} = F_{sync}(\Gamma,0)$
+\\$\Gamma_{dep} = F_{dep}(\Gamma_{sync},0)$
+\\removeNSH($\Gamma_{dep}$)
+\\$TSP = BDT(L^{-1}(\Gamma_{dep}))$
+\\$TSP = F_{tsp}(TSP,root(TSP))$
+\If {data parallel}
+\\$TSP = F_{tsp}(TSP)$
+\EndIf
+\\dump(TSP)
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+Lines 7 to 9 are explained bellow.
 
 %--------------------
-\subsection{GSP and MSP classes}
-A vertex $v$ of a DAG $G$ is a \emph{source} if no edge of $G$ enters $v$.
-Similarly, a vertex $v$ is a \emph{sink} if no edge of $G$ leaves $v$.
-In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of minimal series-parallel DAGs (MSP).
+\subsection{Canonical form and fusions}
 
-\begin{mydef}Minimal Series Parallel
+\todo{à améliorer}
+
+The serie-parallel binary tree decomposition $TSP$ obtained from $BDT(L^{-1}(\Gamma_{dep}))$ can be reduced to its canonical form which consists in recursively merging successive $S$ vertices or successive $P$ vertices following these rules:
+
+\medskip
+\noindent \textbf{Definitions}
+%\begin{mydef}
 \begin{itemize}
-\item The DAG having a single vertex and no edges is MSP.
-\item If $G_1=(V_1,E_1)$ and $G_2=(V_2,E_2)$ are two MSP DAGs, so is either of the DAGs constructed by the following operations:
-\begin{itemize}
-\item Parallel composition: $G_p=(V_1\cup V_2,E_1\cup E_2)$.
-\item Series composition: $G_s=(V_1\cup V_2,E_1\cup E_2\cup (N_1 \times R_2))$, where $N_1$ is the set of sinks of $G_1$ and $R_2$ is the set of sources of $G_2$.
-\end{itemize}
-\end{itemize}
-\end{mydef}
-
-\begin{mydef}
-A DAG is \emph{General Series Parallel} (GSP) if and only if its transitive reduction is a MSP DAG.
-\end{mydef}
-
-A \emph{binary decomposition tree} is a tree having a leaf for each vertex of the MSP DAG it represents, and whose internal nodes are labelled $S$ or $P$ to indicate respectively the series or parallel composition of the MSP sub-DAGs represented by the subtrees rooted at $S$ or $P$.
-Figures~\ref{fig:gsp}, ~\ref{fig:msp} and~\ref{fig:t} respectively give an example of a GSP DAG, its transitive reduction which is MSP, and its tree decomposition.
-
-\begin{figure}[h!]
-\begin{center}
-\subfloat[][\label{fig:gsp}]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (a) at (0,0) {$a$};
-   \node[] (b) at (1,0) {$b$};
-   \node[] (c) at (2,0.5) {$c$};
-   \node[] (d) at (3,0.5) {$d$};
-   \node[] (e) at (2,-0.5) {$e$};
-   \node[] (f) at (3,-0.5) {$f$};
-   \node[] (g) at (4,0) {$g$};
-   \node[] (h) at (1,-1) {$h$};
-   \node[] (i) at (2,-1) {$i$};
-   \node[] (j) at (3,-1) {$j$};
-
-  \path[->]
-    (a) edge node {} (b)
-        edge [bend left=50] node [swap] {} (d)
-        edge node {} (h)
-    (b) edge node {} (c)
-        edge node {} (e)
-    (c) edge node {} (d)
-        edge node {} (f)
-    (e) edge node {} (d)
-        edge node {} (f)
-    (d) edge node {} (g)
-    (f) edge node {} (g)
-    (h) edge node {} (i)
-        edge [bend right=50] node [swap] {} (j)
-    (i) edge node {} (j);
-  \end{tikzpicture}
-}
-\hspace{10pt}
-\subfloat[][\label{fig:msp}]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-  \node[] (a) at (0,0) {$a$};
-   \node[] (b) at (1,0) {$b$};
-   \node[] (c) at (2,0.5) {$c$};
-   \node[] (d) at (3,0.5) {$d$};
-   \node[] (e) at (2,-0.5) {$e$};
-   \node[] (f) at (3,-0.5) {$f$};
-   \node[] (g) at (4,0) {$g$};
-   \node[] (h) at (1,-1) {$h$};
-   \node[] (i) at (2,-1) {$i$};
-   \node[] (j) at (3,-1) {$j$};
-
-  \path[->]
-    (a) edge node {} (b)
-        edge node {} (h)
-    (b) edge node {} (c)
-        edge node {} (e)
-    (c) edge node {} (d)
-        edge node {} (f)
-    (e) edge node {} (d)
-        edge node {} (f)
-    (d) edge node {} (g)
-    (f) edge node {} (g)
-    (h) edge node {} (i)
-    (i) edge node {} (j);
-  \end{tikzpicture}
-}
-\end{center}
-\caption{(a) GSP and (b) MSP DAGs}
-\label{fig:gspmsp}
-\end{figure}
-
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (S1) at (0,0) {$\mathcal{S}$};
-   \node[] (P1) at (0.5,1) {$\mathcal{P}$};
-   \node[] (a) at (-0.5,1) {$a$};
-   \node[] (S2) at (-1,2) {$\mathcal{S}$}; %+1
-   \node[] (S3) at (2,2) {$\mathcal{S}$}; %-1
-   \node[] (b) at (-2,3) {$b$};
-   \node[] (S4) at (0,3) {$\mathcal{S}$};
-   \node[] (P2) at (-1,4) {$\mathcal{P}$};
-   \node[] (g) at (1,4) {$g$};
-   \node[] (S5) at (-2,5) {$\mathcal{S}$};
-   \node[] (S6) at (0,5) {$\mathcal{S}$};
-   \node[] (c) at (-2.5,6) {$c$};
-   \node[] (d) at (-1.5,6) {$d$};
-   \node[] (e) at (-0.5,6) {$e$};
-   \node[] (f) at (0.5,6) {$f$};
-
-   \node[] (h) at (1,3) {$h$};
-   \node[] (S7) at (3,3) {$\mathcal{S}$};
-   \node[] (i) at (2.5,4) {$i$};
-   \node[] (j) at (3.5,4) {$j$};
-
-  \path[->]
-    (S1) edge node {} (a)
-          edge node {} (P1)
-    (P1) edge node {} (S2)
-          edge node {} (S3)
-    (S2) edge node {} (b)
-        edge node {} (S4)
-    (S4) edge node {} (P2)
-          edge node {} (g)
-    (P2) edge node {} (S5)
-          edge node {} (S6)
-    (S5) edge node {} (c)
-          edge node {} (d)
-    (S6) edge node {} (e)
-          edge node {} (f)
-    (S3) edge node {} (h)
-         edge node {} (S7)
-    (S7) edge node {} (i)
-          edge node {} (j);
-  \end{tikzpicture}
-  \caption{Binary decomposition tree of the MSP of Figure~\ref{fig:msp}}
-  \label{fig:t}
-\end{center}
-\end{figure}
-
-Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have also proposed a linear algorithm to know if a DAG is MSP and, if it is, to decompose it to its associated binary decomposition tree.
-This algorithm is based on the duality of the class of MSP DAGs, with the class of \emph{Two Terminal Series Parallel} DAGs (TTSP) from which a binary tree decomposition can be performed in linear time.
-As it will be explained later, in this paper we are interested in DAGs that we already know to be MSP.
-Thus, we only give the needed definitions to understand the binary decomposition algorithm, and not the recognition algorithm.
-To let the reader understand this algorithm, which is modified in this work, the needed definitions are given.
-%  Other algorithms have also been proposed, all of them in linear time~\cite{Schoenmakers95anew}.
-
-\begin{mydef}
-The \emph{line digraph} of a digraph $G$ is a digraph $L(G)$ that has:
-\begin{itemize}
-\item a vertex $f(e)$ for each edge $e$ of $G$; and
-\item an edge $(f(e_1),f(e_2))$ for each pair of edges of $G$ of the form $e_1=(u,v)$, $e_2=(v,w)$.
-\end{itemize}
-\end{mydef}
-
-\begin{mydef}Two Terminal Series Parallel
-\begin{itemize}
-\item A digraph consisting of two vertices joined by a single edge is TTSP.
-\item If $G_1$ and $G_2$ are TTSP digraphs, so is the digraph obtained by either of the following operations:
-\begin{itemize}
-\item \emph{Two terminal parallel composition}: identify the sourc of $G_1$ with the source of $G_2$ and the sink of $G_1$ with the sink of $G_2$.
-\item \emph{Two terminal series composition}: identify the sink of $G_1$ with the source of $G_2$.
-\end{itemize}
-\end{itemize}
-\end{mydef}
-
-\begin{myth}
-If the DAG $G$ is a MSP graph, its \emph{inverse line DAG} $L^{-1}(G)$ is TTSP.
-\end{myth}
-
-It has to be indicated that for a DAG $G$, $L^{-1}(G)$ is not unique.
-Thus, in the work of Valdes \& Al, and in this work, $L^{-1}(G)$ will refer to the unique digraph having a single source and a single sink whose line digraph is $L(L^{-1}(G))=G$.
-
-The binary decomposition tree algorithm is based on the fact that the decomposition can be obtained as a byproduct of a reduction process on $L^{-1}(G)$, which is TTSP.
-In order to obtain the decomposition, Valdes \& Al associate a label with each edge of the digraph being reduced.
-Initially the label of each edge is a trivial binary tree consisting of a single node.
-As the reduction process introduces new edges the rules of Figure~\ref{fig:rules} are used to compute the binary trees used to label them.
-The algorithm ends when the TTSP graph is reduced to its minimum, i.e.\ two vertices and a single edge between them.
-
-\begin{figure}[h!]
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-%FIRST RULE
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};
-
-  \path[->]
-    (x) edge node [above] {$T_1$} (y)
-    (y) edge node [above] {$T_2$} (z);
-  \end{tikzpicture}
-  }
-\hspace{50pt}
-%ARROW
-\subfloat[]{
-$\Rightarrow$
-}
-\hspace{50pt}
-%SECOND RULE
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-   \node[] (s) at (1.5,0.5) {$\mathcal{S}$};
-   \node[] (t1) at (1,1.5) {$T_1$};
-   \node[] (t2) at (2,1.5) {$T_2$};
-
-  \path[->]
-    (x) edge node {} (y)
-    (s) edge node {} (t1)
-        edge node {} (t2);
-  \end{tikzpicture}
-}
-\\
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
-
-  \path[->]
-    (x) edge [bend left] node [above] {$T_1$} (y)
-        edge [bend right] node [below] {$T_2$} (y);
-  \end{tikzpicture}
-  }
-\hspace{50pt}
-\subfloat[]{
-$\Rightarrow$
-}
-\hspace{50pt}
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-   \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
-   \node[] (t1) at (1,1.5) {$T_1$};
-   \node[] (t2) at (2,1.5) {$T_2$};
-
-  \path[->]
-    (x) edge node {} (y)
-    (p) edge node {} (t1)
-        edge node {} (t2);
-  \end{tikzpicture}
-}
-\caption{Reduction rules of the decomposition tree algorithm.}
-\label{fig:rules}
-\end{center}
-\end{figure}
-
-Finally, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identify a forbidden shape, or subgraph, called $N$ and represented in Figure~\ref{fig:n}, such that
-
-\begin{myth}
-A DAG $G$ is GSP if and only if its transitive closure does not contain $N$ as a subgraph.
-\end{myth}
-
-\begin{figure}[h!]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (a) at (0,0) {$a$};
-   \node[] (b) at (1,0) {$b$};
-   \node[] (c) at (0,-1) {$c$};
-   \node[] (d) at (1,-1) {$d$};
-
-  \path[->]
-    (a) edge node {} (b)
-    (c) edge node {} (b)
-        edge node {} (d);
-  \end{tikzpicture}
-  \caption{Forbidden $N$ subgraph shape for a DAG to be GSP.}
-  \label{fig:n}
-\end{center}
-\end{figure}
-
-%------------------
-\subsection{Multi-stencil programs.}
-We are interested in GSP and MSP classes of graphs to be able to represent computations of a multi-stencil program as a set of sequences and parallel executions.
-Actually, such a representation can directly be dumped to a parallel language, in which \emph{sequence} of instructions and \emph{parallel} execution of instructions are defined.
-Series-parallel trees can also be used as input of scheduling optimizations~\cite{Finta1996323,Wang20082684} to improve task parallelism efficiency, which opens more perspectives to this work.
-Finally, in this paper is presented that such a series-parallel representation can also be dumped to component models by defining specific \emph{control components}.
-Thus, the proposed DSL inheritates software engineering advantages of component models, such as code re-use, productivity and maintainability.
-
-Returning back to the parallelization formalism of Section~\ref{sect:parallelism}, two important steps have to be performed to make $\Gamma_{hybrid}$ a MSP graph:
-\begin{itemize}
-\item transitive reduction,
-\item deletion of the forbidden $N$ shape.
+\item We denote $root(T)$, the function which returns the root vertex of the tree $T$.
+\item We denote as a sequence sub-tree $T(V,E)$ and a parallel sub-tree $T(V,E)$ sub-trees for which the root vertex respectively verifies $type(root(T))=S$ or $type(root(T))=P$.
+\item We denote $replace(E,v_1,v_2)$, the function which replaces for all edges of the set $E$, $v_1$ by $v_2$.
+\item A sequence of sequence is a sequence if the order is kept. Thus for two sequence sub-trees $T_1$ and $T_2$, if $T_2 \in V_1$, $T_1$ can be replaced by 
+\begin{equation*}
+T_1'(V_1 \setminus root(T_2),E_1 \cup replace(E_2,root(T_2),root(T_1)))
+\end{equation*}
+\item A parallel sub-tree of parallel sub-tree is a parallel sub-tree in any order. Thus for two parallel sub-trees $T_1$ and $T_2$, if $T_2 \in V_1$, $T_1$ can be replaced by 
+\begin{equation*}
+T_1'(V_1 \setminus root(T_2),E_1 \cup replace(E_2,root(T_2),root(T_1)))
+\end{equation*}
 \end{itemize}
 
-The transitive reduction has already been applied on $\Gamma_{hybrid}$ by the transitivity of the relation $\blacktriangleleft$.
-However, even by applying this transitive reduction it is possible to obtain a DAG which is not MSP.
-Actually, it is possible in a multi-stencil program to have a set of computations such that their dependencies form the forbidden $N$ subgraph.
-For example, if $c_0$, $c_1$, $c_2$ and $c_3$ are computations of a $\mathcal{MSP}$ such that $c_0 \prec c_1$, $c_2 \prec c_1$ and $c_2 \prec c_3$, the \emph{zigzag} relation $c_0 \prec c_1 \succ c_2 \prec c_3$ which form a forbidden subgraph of MSP DAGs is found in $\Gamma_{hybrid}$.
-For this reason, we have made the choice to over-constrain such a case by adding the relation $c_0 \prec c_3$ such that a complete graph is created and can be translated to a series-parallel decomposition as illustrated in Figure~\ref{fig:allover}.
+These reductions applied recursively on the serie-parallel binary tree decomposition of $\Gamma_{dep}$ produce the canonical serie-parallel tree decomposition $TSP = F_{tsp}(TSP,root(TSP),successors(root(TSP))$.
+The recursive function $F_{tsp}$ is based on a Depth First Search algorithm and defined by Algorithm~\ref{alg:tsp}.
 
-\begin{figure}[h!]
-\begin{center}
-\subfloat[][Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (c0) at (0,0) {$c_0$};
-   \node[] (bc0) at (-1,0) {};
-   \node[] (c1) at (2,0) {$c_1$};
-   \node[] (c2) at (0,-1) {$c_2$};
-   \node[] (bc2) at (-1,-1) {};
-   \node[] (c3) at (2,-1) {$c_3$};
+\begin{algorithm}
+\caption{$F_{tsp}$}
+\label{alg:tsp}
+\begin{algorithmic}[1]
+\Procedure{$F_{tsp}$} {$TSP$,root,succ}
+\For {$s \in succ$}
+\If {type($s$)=$S$ or type($s$)=$P$}
+\\$F_{tsp}(TSP,s,successors(s))$
+\If {type($s$)=type(root)}
+\\$TSP = (V_{tsp} \setminus s,E_{tsp} \cup replace(E_{rooted(s)},s,root))$
+\EndIf
+\EndIf
+\EndFor
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 
-  \path[->]
-    (bc0) edge [dotted] node {} (c0)
-    (bc2) edge [dotted] node {} (c2)
-    (c0) edge node {} (c1)
-          edge [dashed] node [swap] {} (c3)
-    (c2) edge node {} (c1)
-        edge node {} (c3);
-  \end{tikzpicture}
-}
-\hspace{50pt}
-\subfloat[][Series-parallel tree associated to the over-constraint\label{fig:treeover}]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (S0) at (0,0) {$\mathcal{S}$};
-   \node[] (bS0) at (-1,-1) {};
-   \node[] (P1) at (-1,1) {$\mathcal{P}$};
-   \node[] (aP1) at (-1.5,2) {};
-   \node[] (aaP1) at (-0.5,2) {};
-   \node[] (P2) at (1,1) {$\mathcal{P}$};
-   \node[] (c0) at (-2,3) {$c_0$};
-   \node[] (c1) at (0.5,3) {$c_1$};
-   \node[] (c2) at (0,3) {$c_2$};
-   \node[] (c3) at (2,3) {$c_3$};
+Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization.
+Even though the MSL language is not dedicated to produce very optimized stencil codes, but to produces the parallel pattern of the application, building the $TSP$ tree make available an easy optimization when the data parallelization technique is the only one used.
+This optimization consists in proposing a valid merge of some computation kernels in a single space loop.
+As a result, the user can use this valid fusion of kernels or not when implementing those.
+The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and $iscomp(k)$ indicates if the vertex is a computation kernel.
+If the kernel $k$ is a synchronization kernel or $S$ or $P$, $iscomp(k)$ returns False.
 
-  \path[->]
-    (bS0) edge [dotted] node {} (S0)
-    (S0) edge node {} (P1)
-         edge node {} (P2)
-    (P1) edge [dotted] node [swap] {} (aP1)
-          edge [dotted] node [swap] {} (aaP1)
-    (aP1) edge node {} (c0)
-    (aaP1) edge node {} (c2)
-    (P2) edge node {} (c1)
-        edge node {} (c3);
-  \end{tikzpicture}
-}
-\caption{Deletion of forbidden subgraphs.}
-\label{fig:allover}
-\end{center}
-\end{figure}
+\begin{algorithm}
+\caption{$F_{fus}$}
+\label{alg:fus}
+\begin{algorithmic}[1]
+\Procedure{$F_{fus}$} {$TSP(V,E)$}
+\For {$(k_i,k_j) \in V^2$}
+\If {$iscomp(k_i)$ and $iscomp(k_j$)}
+\If {parent($k_i$)==parent($k_j$)}
+\If {parent($k_i$)==$S$}
+\If {$D_i==D_j$}
+\\propose the fusion $k_{i;j}$
+\EndIf
+\EndIf
+\If {parent($k_i$)==$P$}
+\If {$D_i==D_j$ and $R_i \cap R_j \neq \emptyset$}
+\\propose the fusion $k_{i;j}$
+\EndIf
+\EndIf
+\EndIf
+\EndIf
+\EndFor
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 
-We consider this approximation acceptable for multi-stencil programs, because of the relative homogeneity of computations.
-Actually, all computations, except the ones on the physical border of the domain, are performed on an entire domain of the mesh, and as a computation performs a single quantity at a time ($w_i$ of $c_i$ is a singleton), the amount of arithmetic operations in a computation are quite homogeneous.
-However, this approximation can also be responsible for a useless wait for computation $c_3$ and less performance.
-For this reason, perspectives of this work open to a new kind of component models.
+We are not arguing that such a simple fusion algorithm could be as good as complex cache optimizations which can be found in stencil DSLs for example.
+However, this fusion takes place at a different level and can bring performance improvments as it will be illustrated in Section~\ref{sect:eval}.
+This algorithm relies on very simple statements:
+\begin{itemize}
+\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependant.
+As a result, what is written by the first one is read by the second one.
+Thus, at least one data is common to those computations (the one written by $k_i$).
+The fusion of $k_i$ and $k_j$ will decrease cache misses.
+\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependant.
+However, if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}$.
+\end{itemize}
 
-After these over-constraints are applied, $\Gamma_{hybrid}$ is a MSP DAG.
-As a result, the binary tree decomposition algorithm of Valdes \& Al can be applied on $\Gamma_{hybrid}$.
-However, because of updated-computations inserted in $\Gamma_{data}$, we have to define additional reduction rules for the algorithm.
-%For an updated-computation $c^*_j(c_j,\text{update}(\{w_i\text{, }\forall i\text{ }c_i \prec c_j\}))$, we define $*_j$ as the update needed by $c_j$, $*_j = \text{update}(\{w_i\text{, }\forall i\text{ }c_i \prec c_j\})$.
-As a result, the updated-computation $c^*_j$ is defined by $c^*_j(c_j,*_j)$.
-The initial label of the edges of the graph $L^{-1}(\Gamma_{hybrid})$ to reduce are the computations of the multi-stencil program.
-%Four new rules are needed to perform the binary tree decomposition of $\Gamma_{hybrid}$ and are defined in Figure~\ref{fig:newrules}.
-The symetric transforations are not represented.
+\todo{solve order problem in $\Gamma_{dep}$ and $TSP$}
 
-% \begin{figure}[h!]
-% \captionsetup[subfigure]{labelformat=empty}
-% \begin{center}
-% %FIRST RULE
-% \subfloat[]{
-% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
-%    \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};
-
-%   \path[->]
-%     (x) edge node [above] {$c_1^*$} (y)
-%     (y) edge node [above] {$c_2$} (z);
-%   \end{tikzpicture}
-%   }
-% \hspace{50pt}
-% %ARROW
-% \subfloat[]{
-% $\Rightarrow$
-% }
-% \hspace{50pt}
-% %TRANSFORMATION
-% \subfloat[]{
-% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-%    \node[] (s1) at (1.5,0.5) {$\mathcal{S}$};
-%    \node[] (s2) at (1,1.5) {$\mathcal{S}$};
-%    \node[] (star) at (0.5,2.5) {$*_1$};
-%    \node[] (t1) at (1.5,2.5) {$c_1$};
-%    \node[] (t2) at (2,1.5) {$c_2$};
-
-%   \path[->]
-%     (x) edge node {} (y)
-%     (s1) edge node {} (s2)
-%         edge node {} (t2)
-%     (s2) edge node {} (star)
-%          edge node {} (t1);
-%   \end{tikzpicture}
-% }
-% \\
-% %SECOND RULE
-% %FIRST RULE
-% \subfloat[]{
-% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
-%    \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};
-
-%   \path[->]
-%     (x) edge node [above] {$c_1^*$} (y)
-%     (y) edge node [above] {$c_2^*$} (z);
-%   \end{tikzpicture}
-%   }
-% \hspace{50pt}
-% %ARROW
-% \subfloat[]{
-% $\Rightarrow$
-% }
-% \hspace{50pt}
-% %TRANSFORMATION
-% \subfloat[]{
-% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-%    \node[] (s1) at (1.5,0.5) {$\mathcal{S}$};
-%    \node[] (s2) at (0.5,1.5) {$\mathcal{S}$};
-%    \node[] (s3) at (2.5,1.5) {$\mathcal{S}$};
-%    \node[] (star1) at (0,2.5) {$*_1$};
-%    \node[] (t1) at (1,2.5) {$c_1$};
-%    \node[] (star2) at (2,2.5) {$*_2$};
-%    \node[] (t2) at (3,2.5) {$c_2$};
-
-%   \path[->]
-%     (x) edge node {} (y)
-%     (s1) edge node {} (s2)
-%         edge node {} (s3)
-%     (s2) edge node {} (star1)
-%          edge node {} (t1)
-%     (s3) edge node {} (star2)
-%          edge node {} (t2);
-%   \end{tikzpicture}
-% }
-% \\
-% %THIRD RULE
-% \subfloat[]{
-% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
-
-%   \path[->]
-%     (x) edge [bend left] node [above] {$c_1^*$} (y)
-%         edge [bend right] node [below] {$c_2$} (y);
-%   \end{tikzpicture}
-%   }
-% \hspace{50pt}
-% %ARROW
-% \subfloat[]{
-% $\Rightarrow$
-% }
-% \hspace{50pt}
-% %TRANSFORMATION
-% \subfloat[]{
-% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-%    \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
-%    \node[] (s) at (1,1.5) {$\mathcal{S}$};
-%    \node[] (star) at (0.5,2.5) {$*_1$};
-%    \node[] (t1) at (1.5,2.5) {$c_1$};
-%    \node[] (t2) at (2,1.5) {$c_2$};
-
-%   \path[->]
-%     (x) edge node {} (y)
-%     (p) edge node {} (s)
-%         edge node {} (t2)
-%     (s) edge node {} (star)
-%         edge node {} (t1);
-%   \end{tikzpicture}
-% }
-% \\
-% %FOURTH RULE
-% \subfloat[]{
-% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
-
-%   \path[->]
-%     (x) edge [bend left] node [above] {$c_1^*$} (y)
-%         edge [bend right] node [below] {$c_2^*$} (y);
-%   \end{tikzpicture}
-%   }
-% \hspace{50pt}
-% %ARROW
-% \subfloat[]{
-% $\Rightarrow$
-% }
-% \hspace{50pt}
-% %TRANSFORMATION
-% \subfloat[]{
-% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%    \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-%    \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-%    \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
-%    \node[] (s2) at (0.5,1.5) {$\mathcal{S}$};
-%    \node[] (s3) at (2.5,1.5) {$\mathcal{S}$};
-%    \node[] (star1) at (0,2.5) {$*_1$};
-%    \node[] (t1) at (1,2.5) {$c_1$};
-%    \node[] (star2) at (2,2.5) {$*_2$};
-%    \node[] (t2) at (3,2.5) {$c_2$};
-
-%   \path[->]
-%     (x) edge node {} (y)
-%     (p) edge node {} (s2)
-%         edge node {} (s3)
-%     (s2) edge node {} (star1)
-%          edge node {} (t1)
-%     (s3) edge node {} (star2)
-%          edge node {} (t2);
-%   \end{tikzpicture}
-% }
-% \caption{Reduction rules of the decomposition tree algorithm.}
-% \label{fig:newrules}
-% \end{center}
-% \end{figure}
+%--------------------
+\subsection{Overall example}
 
@@
 
-At this point, the $\mathcal{MSP}$ program can be dumped to any parallel language in which a sequence and a parallel execution are available.
-For example, we could imagine a parallel functionnal language with a binary \emph{sequence} function and a binary \emph{parallel} function.% such that the tree of Figure~\ref{fig:t} corresponds the following nested call:
-
-% \begin{equation}
-% \begin{split}
-% \text{sequence}(a,\text{parallel}(&\\
-% &\text{sequence}(b,\text{sequence}(\text{parallel}(\text{sequence}(c,d),\text{sequence}(e,f)),g)),\\
-% &\text{sequence}(h,\text{sequence}(i,j))\\
-% &))
-% \end{split}
-% \end{equation}
+% At this point, the $\mathcal{MSP}$ program can be dumped to any parallel language in which a sequence and a parallel execution are available.
+For example, we could imagine a parallel functionnal language with a binary \emph{sequence} function and a binary \emph{parallel} function.

### 2016-01-26 ARTICLES/IJPP/results/task_scaling/500_200/analytic/results.py

@@ -127,6 +127,7 @@ legend = plt.legend(loc='upper right', shadow=True,prop={'size':12})
 plt.grid(True, which="both", linestyle="dotted")
 plt.xlabel('cores')
 plt.ylabel('time (s)')
+plt.yscale("log", basey=2)
 pp = PdfPages('times.pdf')
 plt.savefig(pp, format='pdf', bbox_inches='tight')
 pp.close()

### 2016-01-27 ARTICLES/IJPP/article/compiler.tex
article IJPP compiler corrections Helene 27/01/2016

@@
 Second, $\Gamma_{sync}$ and $\Gamma_{dep}$ can be built from $\Gamma$ using the recursive functions $F_{sync}$ and $F_{dep}$.
-Third, N-Shapes are removed from $\Gamma_{dep}$ to get a SPD graph, and to be able to create its serie-parallel binary tree decomposition $TSP$.
+Third, N-Shapes are removed from $\Gamma_{dep}$ to get a MSPD graph, and to be able to create its serie-parallel binary tree decomposition $TSP$ (line 6).
+Lines 7 to 9 are explained in the next section.
+Finally a dump of $TSP$ is performed by the compiler and will be described in Section~\ref{sect:eval}.
 
@@
 
-Lines 7 to 9 are explained bellow.
-
 %--------------------
@@
 
-\todo{à améliorer}
-
-The serie-parallel binary tree decomposition $TSP$ obtained from $BDT(L^{-1}(\Gamma_{dep}))$ can be reduced to its canonical form which consists in recursively merging successive $S$ vertices or successive $P$ vertices following these rules:
+The serie-parallel binary tree decomposition $TSP$ obtained from $BDT(L^{-1}(\Gamma_{dep}))$ can be reduced to its canonical form which consists in recursively merging successive $S$ vertices or successive $P$ vertices.
 
 \medskip
-\noindent \textbf{Definitions}
+\noindent \textbf{Definitions} For a tree $T(V,E)$
 %\begin{mydef}
 \begin{itemize}
-\item We denote $root(T)$, the function which returns the root vertex of the tree $T$.
-\item We denote as a sequence sub-tree $T(V,E)$ and a parallel sub-tree $T(V,E)$ sub-trees for which the root vertex respectively verifies $type(root(T))=S$ or $type(root(T))=P$.
+\item we denote $root(T)$, the function which returns the root vertex of the tree $T$;
+\item we denote $type(v)$ with $v \in V$ the function which returns the type of a vertex $v$. This type is either $S$, $P$, $comp$ or $sync$;
 \item We denote $replace(E,v_1,v_2)$, the function which replaces for all edges of the set $E$, $v_1$ by $v_2$.
-\item A sequence of sequence is a sequence if the order is kept. Thus for two sequence sub-trees $T_1$ and $T_2$, if $T_2 \in V_1$, $T_1$ can be replaced by 
+\end{itemize}
+
+\begin{mydef}
+We denote as a sequence or parallel sub-tree $T(V,E)$ sub-trees for which the root vertex respectively verifies $type(root(T))=S$ or $type(root(T))=P$.
+\end{mydef}
+
+\begin{myth}
+\begin{itemize}
+\item A sequence of sequence is a sequence if the order is kept. Thus for two sequence sub-trees $T_i$ and $T_j$, if $T_j \in V_i$, $T_i$ can be replaced by 
 \begin{equation*}
-T_1'(V_1 \setminus root(T_2),E_1 \cup replace(E_2,root(T_2),root(T_1)))
+T_i'(V_i \setminus root(T_j),E_i \cup replace(E_j,root(T_j),root(T_i)))
 \end{equation*}
-\item A parallel sub-tree of parallel sub-tree is a parallel sub-tree in any order. Thus for two parallel sub-trees $T_1$ and $T_2$, if $T_2 \in V_1$, $T_1$ can be replaced by 
+\item A parallel sub-tree of parallel sub-tree is a parallel sub-tree in any order. Thus for two parallel sub-trees $T_i$ and $T_j$, if $T_j \in V_i$, $T_i$ can be replaced by 
 \begin{equation*}
-T_1'(V_1 \setminus root(T_2),E_1 \cup replace(E_2,root(T_2),root(T_1)))
+T_i'(V_i \setminus root(T_j),E_i \cup replace(E_j,root(T_j),root(T_i)))
 \end{equation*}
 \end{itemize}
+\end{myth}
 
-These reductions applied recursively on the serie-parallel binary tree decomposition of $\Gamma_{dep}$ produce the canonical serie-parallel tree decomposition $TSP = F_{tsp}(TSP,root(TSP),successors(root(TSP))$.
-The recursive function $F_{tsp}$ is based on a Depth First Search algorithm and defined by Algorithm~\ref{alg:tsp}.
+These reductions applied recursively on the serie-parallel binary tree decomposition of $\Gamma_{dep}$ produce the canonical serie-parallel tree decomposition 
+\begin{equation*}
+TSP = F_{tsp}(TSP,root(TSP),successors(root(TSP)).
+\end{equation*}
+The recursive function $F_{tsp}$ is based on a Depth First Search algorithm and defined by Algorithm~\ref{alg:tsp}, where $E_{rooted(s)}$ represents the set of edges of the sub-tree rooted by the vertex $s$.
 
@@
 Even though the MSL language is not dedicated to produce very optimized stencil codes, but to produces the parallel pattern of the application, building the $TSP$ tree make available an easy optimization when the data parallelization technique is the only one used.
-This optimization consists in proposing a valid merge of some computation kernels in a single space loop.
+This optimization consists in proposing a valid merge of some computation kernels inside a single space loop.
 As a result, the user can use this valid fusion of kernels or not when implementing those.
-The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and $iscomp(k)$ indicates if the vertex is a computation kernel.
-If the kernel $k$ is a synchronization kernel or $S$ or $P$, $iscomp(k)$ returns False.
+The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and where $k_{i;j}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.
 
@@
 \For {$(k_i,k_j) \in V^2$}
-\If {$iscomp(k_i)$ and $iscomp(k_j$)}
+\If {$type(k_i)=type(k_j)=comp$}
 \If {parent($k_i$)==parent($k_j$)}
@@
 
-We are not arguing that such a simple fusion algorithm could be as good as complex cache optimizations which can be found in stencil DSLs for example.
+We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs for example~\cite{spaaTangCKLL11}.
 However, this fusion takes place at a different level and can bring performance improvments as it will be illustrated in Section~\ref{sect:eval}.
-This algorithm relies on very simple statements:
+This fusion algorithm relies on very simple statements:
 \begin{itemize}
@@
 Thus, at least one data is common to those computations (the one written by $k_i$).
-The fusion of $k_i$ and $k_j$ will decrease cache misses.
+Thus, if the computation domains verify $D_i=D_j$, the fusion of $k_i$ and $k_j$ will decrease cache misses.
 \item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependant.
-However, if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}$.
+However, if the computation domains verify $D_i=D_j$, and if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}$.
 \end{itemize}

### 2016-01-27 ARTICLES/IJPP/article/msp.tex
article IJPP compiler Helene 26/01/2016
article IJPP compiler corrections Helene 27/01/2016

@@
 In this section we argue that the graphs $\Gamma_{dep}$, previously defined, can be transformed to a minimal series-parallel directed acyclic graph.
-As a result, this graphs can be represented as a binary tree of parallel and series compositions of sub-graphs, also called \emph{binary decomposition tree}~\cite{Valdes:1979:RSP:800135.804393}.
-The obtained binary tree is a static scheduling of $\Gamma_{dep}$ and can also be used to optimize the code including data parallelism only.
+As a result, this graph can be represented as a binary tree of parallel and series compositions of sub-graphs, also called \emph{binary decomposition tree}~\cite{Valdes:1979:RSP:800135.804393}.
+The obtained binary tree is a valid static scheduling of $\Gamma_{dep}$ and can also be used to optimize the code including data parallelism only.
 
@@
 
-In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of minimal series-parallel DAGs (SPD).
+In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of minimal series-parallel DAGs (MSPD).
 
@@
 \begin{itemize}
-\item The DAG having a single vertex and no edges is SPD.
-\item If $G_1=(V_1,E_1)$ and $G_2=(V_2,E_2)$ are two SPD DAGs, so is either of the DAGs constructed by the following operations:
+\item The DAG having a single vertex and no edges is MSPD.
+\item If $G_1=(V_1,E_1)$ and $G_2=(V_2,E_2)$ are two MSPD DAGs, so is either of the DAGs constructed by the following operations:
 \begin{itemize}
@@
 
-Finally, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or subgraph, called $N$ and represented in Figure~\ref{fig:over}, such that
-
-A possible definition of a N-Shape is
+Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or subgraph, called $N$ and represented in Figure~\ref{fig:over}. A possible definition of a N-Shape is
 \begin{mydef}
-For two vertices $v_1$ and $v_2$ of a DAG which have a Lowest Common Ancestor $v_{lca}$, for all $v_p$ which is either a parent vertex of $v_1$ or $v_2$ and which is different from $v_{lca}$, a N-Shape is formed by the induced subgraph containing vertices $v_1$, $v_2$, $v_{lca}$ and $v_p$.
+For two vertices $v_1$ and $v_2$ of a DAG which have a Lowest Common Ancestor $v_{lca}$.
+For all $v_p$ which is either a parent vertex of $v_1$ or $v_2$ and which is different from $v_{lca}$, a N-Shape is formed by the induced subgraph containing vertices $v_1$, $v_2$, $v_{lca}$ and $v_p$.
 \end{mydef}
@@
 \begin{myth}
-The transitive reduction of a DAG $G$ is SPD if and only if it does not contain $N$ as an induced subgraph.
+The transitive reduction of a DAG $G$ is MSPD if and only if it does not contain $N$ as an induced subgraph.
 \end{myth}
 
+\begin{proof} \cite{Valdes:1979:RSP:800135.804393} \end{proof}
+
 \begin{myprop}
-The transitive reduction of $\Gamma_{dep}$ is not SPD. Actually, it is possible in a multi-stencil program to have a set of computations such that their dependencies form the forbidden $N$ subgraph.
+The transitive reduction of $\Gamma_{dep}$ is not MSPD.
 \end{myprop}
@@
 \begin{proof}
+Actually, it is possible in a multi-stencil program to have a set of computations such that their dependencies form the forbidden $N$ subgraph.
 Considering four computation kernels $k_0$, $k_1$, $k_2$ and $k_3$ of a $\mathcal{MSP}$ such that $k_0 \prec k_1$, $k_2 \prec k_1$ and $k_2 \prec k_3$ are verified.
-Thus, the \emph{zigzag} relation $k_0 \prec k_1 \succ k_2 \prec k_3$ which form a forbidden subgraph of a SPD is found in $\Gamma_{dep}$.
+Thus, the \emph{zigzag} relation $k_0 \prec k_1 \succ k_2 \prec k_3$ which form a forbidden subgraph of a MSPD is found in $\Gamma_{dep}$.
 \end{proof}
 
-To remove the forbidden N-shapes from $\Gamma_{dep}=(V,E)$, we have chosen to apply an over-constraint with the relation $k_0 \prec k_3$, such that a complete bipartite graph is created for the sub-dag, and can be translated to a series-parallel decomposition as illustrated in Figure~\ref{fig:allover}.
-The algorithm is inspired from the one proposed by~\cite{Mitchell:2004:CMV:1082101.1082117}, and is described in Algorithm~\ref{alg:nshape}.
-We precisely describe how we find edges to add to get rid of N-Shapes.
-One can notice that another solution would have been to remove the extra-edge from the N-Shape instead of adding one~\cite{}.
-However, we have chosen to exactly keep dependencies computed in $\Gamma_{dep}$, to not approximate the static scheduling.
-Finally, algorithm proposed by~\cite{Bender200575} to find a Lowest Common Ancestor in a DAG (in $\mathcal{O}(|V|^3)$, could have been also used.
-
-The first step of Algorithm~\ref{alg:nshape} is to create a new DAG $\Gamma_{dep}'$ from $\Gamma_{dep}$, such that vertices having exactly one source and one destination in $\Gamma_{dep}$ is not added to $\Gamma_{dep}'$.
-This step is called $RTV(\Gamma_{dep})$ (line 2) and is done in $\mathcal{O}(|V|)$.
-The second step is to get source vertices in $\Gamma_{dep}'$.
-The third step calls $BFSnshape$ (line 4).
-This procedure is based on a Breadth First Search recursive algorithm (in $\mathcal{O}(|V|+|E|)$) and takes as input a list of source vertices $sources$, $\Gamma_{dep}$ and $\Gamma_{dep}'$.
-At each recursive call the induced sub-graph of $\Gamma_{dep}'$ containing $sources$ vertices and the list of their successors $successors(sources)$ is created and called $SG$.
-If this sub-graph $SG$ is connected and is not a complete bipartite graph, the algorithm make the subgraph $SG$ complete by adding needed edges into $\Gamma_{dep}'$ (line 12 in $\mathcal{O}(|V|^2$).
-The algorithm also add same edges to $\Gamma_{dep}$ (line 13).
-Finally, line 5, the transitive closure of $\Gamma_{dep}$ is computed.
-
-\begin{algorithm}
-\caption{Remove N-Shapes}
-\label{alg:nshape}
-\begin{algorithmic}[1]
-\Procedure{removeNSH} {$\Gamma_{dep}$}
-\\$\Gamma_{dep}'$ = RTV($\Gamma_{dep}$)
-\\$sources$ = sources($\Gamma_{dep}'$)
-\\BFSnshape($sources$,$\Gamma_{dep}'$,$\Gamma_{dep}$)
-\\TC($\Gamma_{dep}$)
-\EndProcedure
-
-\Procedure{BFSnshape} {$sources$, $\Gamma_{dep}'$, $\Gamma_{dep}$}
-\\$successors$ = successors($sources$)
-\\$SG$ = subgraph($sources \cup successors$)
-\If{isconnected($SG$)}
-\If{! complete($SG$)}
-\\addedges($\Gamma_{dep}'$)
-\\addedges($\Gamma_{dep}$)
-\\BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
-\EndIf
-\If{! isconnected($SG$)}
-\\BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
-\EndIf
-\EndIf
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
+To remove the forbidden N-shapes from $\Gamma_{dep}=(V,E)$, we have chosen to apply an over-constraint with the relation $k_0 \prec k_3$, such that a complete bipartite graph is created for the sub-dag, and can be translated to a series-parallel decomposition, as illustrated in Figure~\ref{fig:allover}.
 
@@
 
-After these over-constraints are applied, $\Gamma_{dep}$ is a SPD.
+The algorithm is inspired from the one proposed by~\cite{Mitchell:2004:CMV:1082101.1082117}, and is described in Algorithm~\ref{alg:nshape}.
+We precisely describe how we find edges to add to get rid of N-Shapes.
+One can notice that another solution would have been to remove the extra-edge from the N-Shape instead of adding one~\cite{}.
+However, we have chosen to exactly keep dependencies computed in $\Gamma_{dep}$, to not approximate the static scheduling.
+Finally, algorithm proposed by~\cite{Bender200575} to find a Lowest Common Ancestor in a DAG (in $\mathcal{O}(|V|^3)$, could have been also used.
+
+The first step of Algorithm~\ref{alg:nshape} is to create a new DAG $\Gamma_{dep}'$ from $\Gamma_{dep}$, such that vertices having exactly one source and one destination in $\Gamma_{dep}$ is not added to $\Gamma_{dep}'$, and its source and destination vertices are directly linked by an edge.
+This step is called $RTV(\Gamma_{dep})$ (line 2) and is done in $\mathcal{O}(|V|)$.
+The second step is to get root vertices in $\Gamma_{dep}'$.
+The third step calls $BFSnshape$ (line 4).
+This procedure is based on a Breadth First Search recursive algorithm (in $\mathcal{O}(|V|+|E|)$) and takes as input a list of source vertices $sources$, $\Gamma_{dep}$ and $\Gamma_{dep}'$.
+At each recursive call the induced sub-graph of $\Gamma_{dep}'$ containing $sources$ vertices and the list of their successors $successors(sources)$ is created and called $SG$.
+If this sub-graph $SG$ is connected and is not a complete bipartite graph, the algorithm make the subgraph $SG$ complete by adding needed edges into $\Gamma_{dep}'$ (line 12 in $\mathcal{O}(|V|^2$).
+The algorithm also add same edges to $\Gamma_{dep}$ (line 13).
+Finally, line 5, the transitive closure of $\Gamma_{dep}$ is computed.
+
+\begin{algorithm}
+\caption{Remove N-Shapes}
+\label{alg:nshape}
+\begin{algorithmic}[1]
+\Procedure{removeNSH} {$\Gamma_{dep}$}
+\\$\Gamma_{dep}'$ = RTV($\Gamma_{dep}$)
+\\$sources$ = roots($\Gamma_{dep}'$)
+\\BFSnshape($sources$,$\Gamma_{dep}'$,$\Gamma_{dep}$)
+\\TC($\Gamma_{dep}$)
+\EndProcedure
+
+\Procedure{BFSnshape} {$sources$, $\Gamma_{dep}'$, $\Gamma_{dep}$}
+\\$successors$ = successors($sources$)
+\\$SG$ = subgraph($sources \cup successors$)
+\If{isconnected($SG$)}
+\If{! complete($SG$)}
+\\addedges($\Gamma_{dep}'$)
+\\addedges($\Gamma_{dep}$)
+\\BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
+\EndIf
+\If{! isconnected($SG$)}
+\\BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
+\EndIf
+\EndIf
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+After these over-constraints are applied, $\Gamma_{dep}$ is MSPD.
 As a result, the binary tree decomposition algorithm of Valdes \& Al can be applied on $\Gamma_{dep}$ to get a static scheduling of the multi-stencil program.
@@
 \subsection{Series-parallel binary tree decomposition}
-A \emph{binary decomposition tree} is a tree having a leaf for each vertex of the SPD DAG it represents, and whose internal nodes are labelled $S$ or $P$ to indicate respectively the series or parallel composition of the SPD sub-DAGs represented by the subtrees rooted at $S$ or $P$.
+A \emph{binary decomposition tree} is a tree where each leaf is a vertex of the MSPD DAG it represents, and whose internal nodes are labelled $S$ or $P$ to indicate respectively the series or parallel composition of sub-trees.
 
-Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have also proposed a linear algorithm to know if a DAG is SPD and, if it is, to decompose it to its associated binary decomposition tree.
-This algorithm is based on the duality of the class of SPD DAGs, with the class of \emph{Two Terminal Series Parallel} DAGs (TTSP) from which a binary tree decomposition can be performed in linear time.
+Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have also proposed a linear algorithm to know if a DAG is MSPD and, if it is, to decompose it to its associated binary decomposition tree.
+This algorithm is based on the duality of the class of MSPD DAGs, with the class of \emph{Two Terminal Series Parallel} DAGs (TTSP) from which a binary tree decomposition can be performed in linear time.
 
@@
 \begin{itemize}
-\item \emph{Two terminal parallel composition}: identify the sourc of $G_1$ with the source of $G_2$ and the sink of $G_1$ with the sink of $G_2$.
+\item \emph{Two terminal parallel composition}: identify the source of $G_1$ with the source of $G_2$ and the sink of $G_1$ with the sink of $G_2$.
 \item \emph{Two terminal series composition}: identify the sink of $G_1$ with the source of $G_2$.
@@
 \begin{myth}
-If the DAG $G$ is a SPD graph, its \emph{inverse line DAG} $L^{-1}(G)$ is TTSP.
+If the DAG $G$ is a MSPD graph, its \emph{inverse line DAG} $L^{-1}(G)$ is TTSP.
 \end{myth}
@@
 As the reduction process introduces new edges the rules of Figure~\ref{fig:rules} are used to compute the binary trees used to label them.
+The sequencial composition rule keep the order $T_1$ then $T_2$.
 The algorithm ends when the TTSP graph is reduced to its minimum, i.e.\ two vertices and a single edge between them.
@@
 
-%--------------------
-\subsection{Canonical Form and Optimizations}
+We have explained in this section how a valid static scheduling of $\Gamma_{dep}$ can be built using existing graph transformations of the MSPD class.
+In the next section, the overall compiler algorithm and a complete example is given.

### 2016-01-27 ARTICLES/IJPP/article/msl.tex
article IJPP language corrections Helene 27/01/2016

@@
                         | stencil
+stencil ::= stencilid "from" meshent "to" meshent
 listdata ::= data listdata |  data
@@
 comp ::= dataid "[" compdomid "]=" 
-                    compid "(" 
-                    listscalarread "," 
-                    listdataread ")"
+                    compid "({" 
+                    listscalarread "},{" 
+                    listdataread "})"
 listdataread ::= dataread listdataread 
@@
 \end{figure}
+
+\begin{filecontents*}{exmsl.txt}
+mesh : cart
+mesh entities : cell, edgex
+computation domains :
+  d1 in cell
+  d2 in edgex
+independent :
+  d1 and d2
+stencil shapes : 
+  n1 from cell to cell
+  n2 from cell to edgex
+  n3 from edgex to cell
+data :
+  A,cell
+  B,cell
+  C,edgex
+  D,cell
+  E,edgex
+  F,cell
+  G,cell
+  H,edgex
+  I,cell
+  J,cell
+scalar : mu, tau
+time : 500
+computations :
+  B[d1] = k0({tau},{A})
+  C[d2] = k1({},{B[n2]})
+  D[d1] = k2({},{C})
+  E[d2] = k3({},{C})
+  F[d1] = k4({},{D,C[n3]})
+  G[d1] = k0({mu,tau},{E})
+  H[d2] = k6({},{F})
+  I[d1] = k7({},{G,H})
+  J[d1] = k8({mu},{I[n1]})
+\end{filecontents*}
+
+\begin{figure}[!h]
+  \hspace{5mm}
+  \begin{minipage}[!h]{0.4\textwidth}
+    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left]{exmsl.txt}}   
+    \caption{Example of program using the Multi-Stencil Language. \label{fig:exmsl}}
+  \end{minipage}
+\end{figure}

### 2016-01-29 ARTICLES/IJPP/article/compiler.tex
article IJPP language corrections Helene 27/01/2016
article IJPP corrections Helene 29/01/2016

@@
 \Procedure{compileMSL} {file}
-\\$\Gamma$ = parser(file)
-\\$\Gamma_{sync} = F_{sync}(\Gamma,0)$
-\\$\Gamma_{dep} = F_{dep}(\Gamma_{sync},0)$
-\\removeNSH($\Gamma_{dep}$)
-\\$TSP = BDT(L^{-1}(\Gamma_{dep}))$
-\\$TSP = F_{tsp}(TSP,root(TSP))$
+\State $\Gamma$ = parser(file)
+\State $\Gamma_{sync} = F_{sync}(\Gamma,0)$
+\State $\Gamma_{dep} = F_{dep}(\Gamma_{sync},0)$
+\State removeNSH($\Gamma_{dep}$)
+\State $TSP = BDT(L^{-1}(\Gamma_{dep}))$
+\State $TSP = F_{tsp}(TSP,root(TSP))$
 \If {data parallel}
-\\$TSP = F_{tsp}(TSP)$
+\State $TSP = F_{tsp}(TSP)$
 \EndIf
-\\dump(TSP)
+\State dump(TSP)
 \EndProcedure
@@
 \If {type($s$)=$S$ or type($s$)=$P$}
-\\$F_{tsp}(TSP,s,successors(s))$
+\State $F_{tsp}(TSP,s,successors(s))$
 \If {type($s$)=type(root)}
-\\$TSP = (V_{tsp} \setminus s,E_{tsp} \cup replace(E_{rooted(s)},s,root))$
+\State $V_{tsp} = V_{tsp} \setminus s$
+\State $E_{tsp} = E_{tsp} \cup replace(E_{rooted(s)},s,root)$
+\State $TSP = (V_{tsp},E_{tsp})$
 \EndIf
@@
 As a result, the user can use this valid fusion of kernels or not when implementing those.
-The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and where $k_{i;j}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.
+The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and where $k_{i;j}^{fus}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.
 
@@
 \For {$(k_i,k_j) \in V^2$}
+\If {parent($k_i$)==parent($k_j$)}
 \If {$type(k_i)=type(k_j)=comp$}
-\If {parent($k_i$)==parent($k_j$)}
 \If {parent($k_i$)==$S$}
 \If {$D_i==D_j$}
-\\propose the fusion $k_{i;j}$
+\State propose the fusion $k_{i;j}^{fus}$
 \EndIf
-\EndIf
-\If {parent($k_i$)==$P$}
+\ElsIf {parent($k_i$)==$P$}
 \If {$D_i==D_j$ and $R_i \cap R_j \neq \emptyset$}
-\\propose the fusion $k_{i;j}$
+\State propose the fusion $k_{i;j}^{fus}$
 \EndIf
@@
 \item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependant.
-However, if the computation domains verify $D_i=D_j$, and if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}$.
+However, if the computation domains verify $D_i=D_j$, and if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}^{fus}$.
 \end{itemize}
-
-\todo{solve order problem in $\Gamma_{dep}$ and $TSP$}
 
@@
 \subsection{Overall example}
+
+MSL code~\ref{fig:exmsl}. 
+\begin{equation*}
+\Gamma = [k_0,k_1,k_2,k_3,k_4,k_0,k_6,k_7,k_8]
+\end{equation*}
+
+\begin{equation*}
+\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_0,k_6,k_7,k_{7;8}^{sync},k_8]
+\end{equation*}
+
+where $k_{0;1}^{sync}$...
+
+$\Gamma_{dep}$~\ref{} and $L^{-1}(\Gamma_{dep})$~\ref{}
 
@@
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (c0) at (0,0) {$c_0$};
-   \node[] (star1) at (1,0) {$*_1$};
-   \node[] (c1) at (2,0) {$c_1^*$};
-   \node[] (c2) at (3,0.5) {$c_2$};
-   \node[] (star4) at (3,1.5) {$*_4$};
-   \node[] (c3) at (3,-0.5) {$c_3$};
-   \node[] (c4) at (4,0.5) {$c_4^*$};
-   \node[] (c5) at (4,-0.5) {$c_5$};
-   \node[] (c6) at (5,0.5) {$c_6$};
-   \node[] (c7) at (6,0) {$c_7$};
-   \node[] (star8) at (7,0) {$*_8$};
-   \node[] (c8) at (8,0) {$c_8^*$};
+   \node[] (c0) at (0,0) {$k_0$};
+   \node[] (star1) at (1,0) {$k_{0;1}^{sync}$};
+   \node[] (c1) at (2,0) {$k_1$};
+   \node[] (c2) at (3,0.5) {$k_2$};
+   \node[] (star4) at (3,1.5) {$k_{1;4}^{sync}$};
+   \node[] (c3) at (3,-0.5) {$k_3$};
+   \node[] (c4) at (4,0.5) {$k_4$};
+   \node[] (c5) at (4,-0.5) {$k_5$};
+   \node[] (c6) at (5,0.5) {$k_6$};
+   \node[] (c7) at (6,0) {$k_7$};
+   \node[] (star8) at (7,0) {$k_{7;8}^{sync}$};
+   \node[] (c8) at (8,0) {$k_8$};
 
@@
   \path[->]
-    (c0s) edge node [above] {$c_0$} (c0d)
-    (c0d) edge node [above] {$*_1$} (star1d)
-    (star1d) edge node [above] {$c_1^*$} (c1d)
-    (c1d) edge node [above] {$c_2$} (c2d)
-          edge node [above] {$c_3$} (c3d)
-          edge [bend left=80] node [above] {$*_4$} (c2d)
-    (c2d) edge node [above] {$c_4^*$} (c4d)
-    (c4d) edge node [above] {$c_6$} (c6d)
-    (c6d) edge node [above] {$c_7$} (c7d)
-    (c7d) edge node [above] {$c_8$} (star8d)
-    (star8d) edge node [above] {$c_8$} (c8d)
-    (c3d) edge node [above] {$c_5$} (c6d);
+    (c0s) edge node [above] {$k_0$} (c0d)
+    (c0d) edge node [above] {$k_{0;1}^{sync}$} (star1d)
+    (star1d) edge node [above] {$k_1$} (c1d)
+    (c1d) edge node [above] {$k_2$} (c2d)
+          edge node [above] {$k_3$} (c3d)
+          edge [bend left=80] node [above] {$k_{1;4}^{sync}$} (c2d)
+    (c2d) edge node [above] {$k_4$} (c4d)
+    (c4d) edge node [above] {$k_6$} (c6d)
+    (c6d) edge node [above] {$k_7$} (c7d)
+    (c7d) edge node [above] {$k_{7;8}^{sync}$} (star8d)
+    (star8d) edge node [above] {$k_8$} (c8d)
+    (c3d) edge node [above] {$k_0$} (c6d);
   \end{tikzpicture}
 }
-\\
+\end{center}
+\caption{$\Gamma_{dep}$ and $L^{-1}(\Gamma_{dep})$}
+\label{fig:dep}
+\end{figure}
+
+\begin{figure}[ht!]
+\begin{center}
 % TRANSFORMATIONS
@@
    \node[] (s1) at (1,1.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,2.5) {$c_0$};
-   \node[] (star1) at (1.5,2.5) {$*_1$};
-   \node[] (c1) at (2,1.5) {$c_1^*$};
+   \node[] (c0) at (0.5,2.5) {$k_0$};
+   \node[] (star1) at (1.5,2.5) {$k_{0;1}^{sync}$};
+   \node[] (c1) at (2,1.5) {$k_1$};
 
@@
    \node[] (s3) at (7,1.5) {$\mathcal{S}$};
-   \node[] (c7) at (6.5,2.5) {$c_7$};
-   \node[] (star8) at (7.5,2.5) {$*_8$};
-   \node[] (c8) at (8,1.5) {$c_8$};
+   \node[] (c7) at (6.5,2.5) {$k_7$};
+   \node[] (star8) at (7.5,2.5) {$k_{7;8}^{sync}$};
+   \node[] (c8) at (8,1.5) {$k_8$};
 
@@
     (c0s) edge node {} (c1d)
-    (c1d) edge node [above] {$c_2$} (c2d)
-          edge node [above] {$c_3$} (c3d)
-          edge [bend left=80] node [above] {$*_4$} (c2d)
-    (c2d) edge node [above] {$c_4^*$} (c4d)
-    (c4d) edge node [above] {$c_6$} (c6d)
+    (c1d) edge node [above] {$k_2$} (c2d)
+          edge node [above] {$k_3$} (c3d)
+          edge [bend left=80] node [above] {$k_{1;4}^{sync}$} (c2d)
+    (c2d) edge node [above] {$k_4$} (c4d)
+    (c4d) edge node [above] {$k_6$} (c6d)
     (c6d) edge node [above] {} (c8d)
-    (c3d) edge node [above] {$c_5$} (c6d)
+    (c3d) edge node [above] {$k_0$} (c6d)
 
@@
    \node[] (s5) at (4.5,-2) {$\mathcal{S}$};
-   \node[] (c3) at (4,-1) {$c_3$};
-   \node[] (c5) at (5,-1) {$c_5$};
+   \node[] (c3) at (4,-1) {$k_3$};
+   \node[] (c5) at (5,-1) {$k_0$};
 
@@
    \node[] (p0) at (3.5,1) {$\mathcal{P}$};
-   \node[] (c2) at (3,2) {$c_2$};
-   \node[] (star4) at (4,2) {$*_4$};
+   \node[] (c2) at (3,2) {$k_2$};
+   \node[] (star4) at (4,2) {$k_{1;4}^{sync}$};
 
@@
     (c1d) edge node [above] {} (c2d)
-    (c2d) edge node [above] {$c_4^*$} (c4d)
-    (c4d) edge node [above] {$c_6$} (c6d)
+    (c2d) edge node [above] {$k_4$} (c4d)
+    (c4d) edge node [above] {$k_6$} (c6d)
     (c6d) edge node [above] {} (c8d)
@@
 \end{center}
-\caption{Big example}
-\label{fig:example}
+\caption{Sequence and parallel rules applications}
+\label{fig:SP}
 \end{figure}
@@
 \begin{figure}[ht!]
-\ContinuedFloat
 \begin{center}
@@
    \node[] (p0) at (3.5,2) {$\mathcal{P}$};
-   \node[] (c2) at (3,3) {$c_2$};
-   \node[] (star4) at (4,3) {$*_4$};
+   \node[] (c2) at (3,3) {$k_2$};
+   \node[] (star4) at (4,3) {$k_{1;4}^{sync}$};
    \node[] (s7) at (5.5,2) {$\mathcal{S}$};
-   \node[] (c4) at (5,3) {$c_4^*$};
-   \node[] (c6) at (6,3) {$c_6$};
+   \node[] (c4) at (5,3) {$k_4$};
+   \node[] (c6) at (6,3) {$k_6$};
 
@@
    \node[] (s5) at (5.5,1.5) {$\mathcal{S}$};
-   \node[] (c3) at (5,2.5) {$c_3$};
-   \node[] (c5) at (6,2.5) {$c_5$};
+   \node[] (c3) at (5,2.5) {$k_3$};
+   \node[] (c5) at (6,2.5) {$k_0$};
    %reduction c2 c6
@@
    \node[] (p0) at (2.8,2.5) {$\mathcal{P}$};
-   \node[] (c2) at (2.3,3.5) {$c_2$};
-   \node[] (star4) at (3.3,3.5) {$*_4$};
+   \node[] (c2) at (2.3,3.5) {$k_2$};
+   \node[] (star4) at (3.3,3.5) {$k_{1;4}^{sync}$};
    \node[] (s7) at (4.2,2.5) {$\mathcal{S}$};
-   \node[] (c4) at (3.8,3.5) {$c_4^*$};
-   \node[] (c6) at (4.8,3.5) {$c_6$};
+   \node[] (c4) at (3.8,3.5) {$k_4$};
+   \node[] (c6) at (4.8,3.5) {$k_6$};
 
@@
    \node[] (s1) at (1,3.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,4.5) {$c_0$};
-   \node[] (star1) at (1.5,4.5) {$*_1$};
-   \node[] (c1) at (2,3.5) {$c_1^*$};
+   \node[] (c0) at (0.5,4.5) {$k_0$};
+   \node[] (star1) at (1.5,4.5) {$k_{0;1}^{sync}$};
+   \node[] (c1) at (2,3.5) {$k_1$};
 
@@
    \node[] (s3) at (7,2.5) {$\mathcal{S}$};
-   \node[] (c7) at (6.5,3.5) {$c_7$};
-   \node[] (star8) at (7.5,3.5) {$*_8$};
-   \node[] (c8) at (8,2.5) {$c_8$};
+   \node[] (c7) at (6.5,3.5) {$k_7$};
+   \node[] (star8) at (7.5,3.5) {$k_{7;8}^{sync}$};
+   \node[] (c8) at (8,2.5) {$k_8$};
 
@@
    \node[] (s5) at (5.5,3.5) {$\mathcal{S}$};
-   \node[] (c3) at (5,4.5) {$c_3$};
-   \node[] (c5) at (6,4.5) {$c_5$};
+   \node[] (c3) at (5,4.5) {$k_3$};
+   \node[] (c5) at (6,4.5) {$k_0$};
    %reduction c2 c6
@@
    \node[] (p0) at (2.8,4.5) {$\mathcal{P}$};
-   \node[] (c2) at (2.3,5.5) {$c_2$};
-   \node[] (star4) at (3.3,5.5) {$*_4$};
+   \node[] (c2) at (2.3,5.5) {$k_2$};
+   \node[] (star4) at (3.3,5.5) {$k_{1;4}^{sync}$};
    \node[] (s7) at (4.2,4.5) {$\mathcal{S}$};
-   \node[] (c4) at (3.8,5.5) {$c_4^*$};
-   \node[] (c6) at (4.8,5.5) {$c_6$};
+   \node[] (c4) at (3.8,5.5) {$k_4$};
+   \node[] (c6) at (4.8,5.5) {$k_6$};
 
@@
 \end{center}
-\caption{Big example}
+\caption{Rest of sequence and parallel rules.}
 \label{fig:example}

### 2016-01-29 ARTICLES/IJPP/article/formalism.tex
article IJPP Helene 26/01/2016
article IJPP corrections Helene 29/01/2016

@@
 
-$comp$ represents the actual numerical expression which is computed by a kernel. We can dissociate three types of kernel computations.
+$comp$ represents the actual numerical expression which is computed by a kernel. We can dissociate four types of kernel computations.
 
@@
 \item We denote by $identity$ the identity function $x=x$.
-\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{stencil kernel} $\iff \exists (r,n) \in R$ such that $n \neq identity$.
+\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{stencil kernel} $\iff \exists (r,n) \in R$ such that $n \neq identity$ and if $\exists (r,n)$ with $r=w$ then $n=identity$.
+\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{boundary kernel} $\iff \exists (r,n) \in R$ such that $r=w$ and such that $n \neq identity$.
 \item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{local kernel} $\iff \forall (r,n) \in R$, $n = identity$.
@@
 \end{itemize}
+
+In our formalism, it is forbidden for a stencil kernel to find $(r,n) \in R$ for which $r=w$ and $n \neq identity$.
+Actually, if such a property is possible implicit numerical schemes are needed into the simulation.
+Such a scheme is not a stencil and is over the scope of this paper.
+The problem does not happened for local kernels as $\forall (r,n) \in R$, $n = identity$.
+Computations for which this property is authorized are boundary kernels only.
+This type of kernel is a particular case as it commputes boundary conditions, which represents what should happened outside the space domain but still impact stencil kernels at the next time step.
 
@@
 \item The set of $n$ ordered computation kernels of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i \leq j$, then $k_i$ is computed before $k_j$.
-\item Finally, a \textit{multi-stencil program} is defined by the septuplet 
+\item Finally, a \textit{multi-stencil program} is defined by the octuplet 
 \begin{equation*}
-\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta, \mathcal{S},\Gamma)
+\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\mathcal{N},\Delta, \mathcal{S},\Gamma)
 \end{equation*}

### 2016-01-29 ARTICLES/IJPP/article/msl.tex
article IJPP corrections Helene 29/01/2016

@@
   D,cell
-  E,edgex
+  E,cell
   F,cell
@@
   D[d1] = k2({},{C})
-  E[d2] = k3({},{C})
+  E[d1] = k3({},{C})
   F[d1] = k4({},{D,C[n3]})

### 2016-01-29 ARTICLES/IJPP/article/parallelism.tex
article IJPP corrections Helene 29/01/2016

@@
 \begin{equation*}
-k_{sync}(S,R,(w,D),comp)
+k_{i,j}^{sync}(S,R,(w,D),comp)
 \end{equation*}
 \end{mydef}
-where $S=\emptyset$, $R=\{(w_i,\bigcup_{e \in entity(w_i)} n_j(e))\}$, $w=(w_i,\bigcup_{e \in entity(w_i)} n_j(e)))$ and $comp=identity$.
+where $S=\emptyset$, $R=\{(w_i,\bigcup_{e \in D_i} n_j(e))\}$, $w=(w_i,\bigcup_{e \in D_i} n_j(e)))$ and $comp=identity$.
 In other words, $w_i$ has to be synchronized for the neighborhood $n_j$ for all entities of $w_i$.
@@
 \begin{equation*}
-[k_{sync}(\emptyset,\{(w_i,n_j)\},(w_i,n_j),identity), k_j]
+[k_{i,j}^{sync}, k_j]
 \end{equation*}
@@
 \begin{mydef}
-For a reduction kernel $k(S,R,(w,D),comp)$, if we denote by $w^i$, $0 \leq i<p$, the local $w$ computed on the resource $i$, a reduction synchronization is defined as the specific computation 
+For a reduction kernel $k_j(S_j,R_j,(w_j,D_j),comp)$, if we denote by $w^r$, $0 \leq r<p$, the local $w$ computed on the resource $r$, a reduction synchronization is defined as the specific computation
 \begin{equation*}
-k_{sync}(S,R_s,(w,D),comp)
+k_{j}^{sync}(S,R,(w,D),comp)
 \end{equation*}
 \end{mydef}
-where, $S=\emptyset$, $R_s=\{(w^0,entity(w^0)) \dots (w^{p-1},entity(w^{p-1}))\}$, and $w$, $D=entity(w)$ and $comp$ are the same than the actual reduction kernel $k$.
+where, $S=\emptyset$, $R=\{(w^0,entity(w^0)) \dots (w^{p-1},entity(w^{p-1}))\}$, and $w=w_i$, $D=entity(w)=D_i$ and $comp=comp_i$.
 
@@
 \begin{mydef}
-From the ordered list of computation $\Gamma$, a new ordered list $\Gamma_{sync}$ is obtained from the call $\Gamma_{sync} = F_{sync}(\Gamma,0)$, where $F_{sync}$ is the recursive function defined as
-\begin{equation*}
-F_{sync}(\Gamma,j) = 
-\begin{cases} 	\bullet [] \mbox{ if }j=|\Gamma|\\
-				\bullet [k_{sync}(\emptyset,\\
-				\{(w_i,\bigcup_{e \in entity(w_i)} n_j(e))\},\\
-				(w_i,\bigcup_{e \in entity(w_i)} n_j(e))),\\
-				identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1)\\
-				\text{ } \qquad \forall i<j \text{, and }k_i \pprec k_j\\
-				\bullet [k_j, k_{sync}(\emptyset,\\
-				\{(w^0_j,entity(w^0_j)) \dots (w^{p-1}_j,entity(w^{p-1}_j))\},\\
-				(w_j,entity(w_j)),comp_j)] \cdot F_{sync}(\Gamma,j+1)\\ 
-				\text{ } \qquad \mbox{if } w_j \in \mathcal{S} \text{, and } R_j=\{(r_j,entity(r_j))\}\\
-				\bullet [\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) \text{ else}.
-\end{cases}
-\end{equation*}
+From the ordered list of computation $\Gamma$, a new ordered list $\Gamma_{sync}$ is obtained from the call $\Gamma_{sync} = F_{sync}(\Gamma,0)$, where $F_{sync}$ is the recursive function defined in Algorithm~\ref{alg:sync}.
+
+\begin{algorithm}
+\caption{$F_{sync}$ recursive function}
+\label{alg:sync}
+\begin{algorithmic}[1]
+\Procedure{$F_{sync}$} {$\Gamma$,$j$}
+\State $k_j = \Gamma[j]$
+\State $list = []$
+\If {$j=|\Gamma|$}
+\State return $list$
+\ElsIf {$\exists (r_j,n_j) \in R_j$ such that $n_j\neq identity$}
+\For {$0 \leq i<j$}
+\State $k_i = \Gamma[i]$
+\If {$k_i \pprec k_j$}
+\State $S = \emptyset$
+\State $R = \{(w_i,\bigcup_{e \in D_i} n_j(e))\}$
+\State $(w,D) = (w_i,\bigcup_{e \in D_i} n_j(e)))$
+\State $comp = identity$
+\State $list.[k_{i;j}^{sync}(S,R,(w,D),comp)$
+\State break
+\EndIf
+\EndFor
+\State $list \cdot [k_j]$
+\ElsIf {$w_j \in \mathcal{S} \text{, and } R_j=\{(r_j,entity(r_j))\}$}
+\State $S = \emptyset$
+\State $R = \{(w^0_j,entity(w^0_j)) \dots (w^{p-1}_j,entity(w^{p-1}_j))\}$
+\State $(w,D) = (w_j,entity(w_j))$
+\State $comp = comp_j$
+\State $list.[k_j, k_{j}^{sync}(S,R,(w,D),comp)]$
+\Else
+\State $list.[k_j]$
+\EndIf
+\State return $list \cdot F_{sync}(\Gamma,j+1)$
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+% \begin{equation*}
+% F_{sync}(\Gamma,j) = 
+% \begin{cases} 	\bullet [] \mbox{ if }j=|\Gamma|\\
+% 				\bullet [k_{sync}(\emptyset,\\
+% 				\{(w_i,\bigcup_{e \in entity(w_i)} n_j(e))\},\\
+% 				(w_i,\bigcup_{e \in entity(w_i)} n_j(e))),\\
+% 				identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1)\\
+% 				\text{ } \qquad \forall i<j \text{, and }k_i \pprec k_j\\
+% 				\bullet [k_j, k_{sync}(\emptyset,\\
+% 				\{(w^0_j,entity(w^0_j)) \dots (w^{p-1}_j,entity(w^{p-1}_j))\},\\
+% 				(w_j,entity(w_j)),comp_j)] \cdot F_{sync}(\Gamma,j+1)\\ 
+% 				\text{ } \qquad \mbox{if } w_j \in \mathcal{S} \text{, and } R_j=\{(r_j,entity(r_j))\}\\
+% 				\bullet [\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) \text{ else}.
+% \end{cases}
+% \end{equation*}
 \end{mydef}
@@
 
+% \begin{mydef}
+% For two computations domains $D_i$ and $D_j$ both subsets of the same entities group $E$.
+We say that $D_i$ and $D_j$ are independent, denoted $D_i \oplus D_j$, if and only if $D_i \cap D_j = \emptyset$.
+% \end{mydef}
+
 \begin{mydef}
-For two computations $k_i(S_i,R_i,(w_i,D_i),comp_i)$ and $k_j(S_j,R_j,(w_j,D_j),comp_j)$, with $i < j$, it is said that $k_j$ is data dependant from $k_i$, denoted $k_i\prec k_j$, if $w_i \in R_j$.
+For two computations $k_i(S_i,R_i,(w_i,D_i),comp_i)$ and $k_j(S_j,R_j,(w_j,D_j),comp_j)$, with $i < j$ and such that neither $k_i$ or $k_j$ are boundary kernels, it is said that $k_j$ is data read/write dependant from $k_i$, denoted $k_i \prec_{rw} k_j$, if $w_i \in R_j$.
 In this case, $k_i$ has to be computed before $k_j$.
-The binary relation $k_i\prec k_j$ represents a \textit{dependency}.
+The binary relation $k_i \prec_{rw} k_j$ represents a \textit{dependency}.
+\end{mydef}
+
+\begin{mydef}
+For two computations $k_i(S_i,R_i,(w_i,D_i),comp_i)$ and $k_j(S_j,R_j,(w_j,D_j),comp_j)$, with $i < j$, it is said that $k_j$ is data write dependant from $k_i$, denoted $k_i \prec_{w} k_j$, if $w_i = w_j$ and $D_i \cap D_j \neq \emptyset$.
+In this case, $k_i$ also has to be computed before $k_j$.
+The binary relation $k_i \prec_{w} k_j$ represents a \textit{dependency}.
 \end{mydef}
@@
 
-From an ordered list of computations $\Gamma_{sync}$, a directed dependency graph $\Gamma_{dep}(V,A)$ can be built finding all pairs of computations $k_i(S_i,R_i,(w_i,D_i),comp_i)$ and $k_j(S_j,R_j,(w_j,D_j),comp_j)$, with $i<j$, such that $k_i \prec k_j$.
+From an ordered list of computations $\Gamma_{sync}$, a directed dependency graph $\Gamma_{dep}(V,A)$ can be built finding all pairs of computations $k_i(S_i,R_i,(w_i,D_i),comp_i)$ and $k_j(S_j,R_j,(w_j,D_j),comp_j)$, with $i<j$, such that $k_i \prec_{rw} k_j$ or $k_i \prec_w k_j$.
 
@@
 From the ordered list $\Gamma_{sync}$ of computation kernels, a directed dependency graph $\Gamma_{dep}(V,A)$ is obtained from the call $F_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function
-\begin{equation*}
-F_{dep}(\Gamma_{sync},j) = 
-\begin{cases} 	\bullet (\{\},\{\}) \mbox{ if }j=|\Gamma_{sync}|\\
-				\bullet (k_j, \{(\overset{\frown}{k_i,k_j})\mbox{, }\forall i < j \mbox{, } k_i\prec k_j \})\\
-				\text{ } \qquad \cup F_{dep}(\Gamma_{sync},j+1) \mbox{ if }j<|\Gamma_{sync}|
-\end{cases}
-\end{equation*}
+
+\begin{algorithm}
+\caption{$F_{dep}$ recursive function}
+\label{alg:sync}
+\begin{algorithmic}[1]
+\Procedure{$F_{dep}$} {$\Gamma_{sync}$,$j$}
+\State $k_j = \Gamma_{sync}[j]$
+\If {$j=|\Gamma_{sync}|$}
+\State return $(\{\},\{\})$
+\ElsIf {$j<|\Gamma_{sync}|$}
+\State $G=(\{\},\{\})$
+\For {$0 \leq i<j$}
+\State $k_i = \Gamma_{sync}[i]$
+\If {$k_i \prec_{rw} k_j$ or $k_i \prec_{w} k_j$}
+\State $G = G \cup (k_j, \{(\overset{\frown}{k_i,k_j} \})$
+\EndIf
+\EndFor
+\State return $G \cup F_{dep}(\Gamma_{sync},j+1)$
+\EndIf
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+% \begin{equation*}
+% F_{dep}(\Gamma_{sync},j) = 
+% \begin{cases} 	\bullet (\{\},\{\}) \mbox{ if }j=|\Gamma_{sync}|\\
+% 				\bullet (k_j, \{(\overset{\frown}{k_i,k_j})\mbox{, }\forall i < j \mbox{, } k_i\prec k_j \})\\
+% 				\text{ } \qquad \cup F_{dep}(\Gamma_{sync},j+1) \mbox{ if }j<|\Gamma_{sync}|
+% \end{cases}
+% \end{equation*}
 \end{mydef}

### 2016-01-29 ARTICLES/IJPP/article/article.tex
article IJPP Helene 26/01/2016

--- 2016-01-21 ARTICLES/IJPP/article/article.tex
@@
 \usepackage{listings}
+\usepackage{mathtools, cuted}
 \usepackage{tikz}
-\usepackage[ruled,noend]{algorithm2e}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 \usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}
@@
            Julien Bigot \at
-              CEA
+           Maison de la Simulation, CEA, CNRS, Univ. Paris-Sud, UVSQ, Universit\'e Paris-Saclay, 91191 Gif-sur-Yvette, France\\
+           \email{julien.bigot@cea.fr}
 }
@@
 \input{msl.tex}
-% The parallel empty skeleton of the application can be computed automatically from the language :
-% \begin{itemize}
-% \item grammar
-% \item example from the language to the dependency graph
-% \end{itemize}
+The parallel empty skeleton of the application can be computed automatically from the language :
+\begin{itemize}
+\item grammar
+\item example from the language to $\Gamma$
+\end{itemize}
 %------------------------------------------------------------------------------
@@
 % \end{itemize}
+%------------------------------------------------------------------------------
+\section{MSL Compiler}
+\label{sect:comp}
+\input{compiler.tex}
+\begin{itemize}
+\item Overall algrotihm~\ref{alg:compiler}
+\item Big example from the language to the static sheduling/optimization~\ref{fig:example}
+\end{itemize}
 %------------------------------------------------------------------------------
@@
 
-\bibliographystyle{spbasic}
+\bibliographystyle{plain}
 \bibliography{biblio}

### 2016-01-29 ARTICLES/IJPP/article/msl.tex
article IJPP corrections Helene 29/01/2016

@@
   \begin{minipage}[!h]{0.5\textwidth}
-    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=right]{grammar.txt}}   
+    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left]{grammar.txt}}   
     \caption{Grammar of the Multi-Stencil Language. \label{fig:grammar}}
@@
   \begin{minipage}[!h]{0.4\textwidth}
-    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left]{exmsl.txt}}   
+    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=right]{exmsl.txt}}   
     \caption{Example of program using the Multi-Stencil Language. \label{fig:exmsl}}

### 2016-02-01 ARTICLES/IJPP/article/article.tex
article IJPP corrections Helene 29/01/2016
article IJPP corrections+example Helene 01/02/2016

@@
 \author{Helene Coullon         \and
-        Chirtsian Perez \and
+        Christian Perez \and
         Julien Bigot
@@
 \input{compiler.tex}
-\begin{itemize}
-\item Overall algrotihm~\ref{alg:compiler}
-\item Big example from the language to the static sheduling/optimization~\ref{fig:example}
-\end{itemize}
+% \begin{itemize}
+% \item Overall algrotihm~\ref{alg:compiler}
+% \item Big example from the language to the static sheduling/optimization~\ref{fig:example}
+% \end{itemize}
 %------------------------------------------------------------------------------
@@
 \begin{itemize}
-\item Lizst, Pochoir, PATUS for optimizations/parallelization of a single stencil kernel + mesh dependent
-\item Halide for pipeline of stencil computations = mesh dependent
+\item Cartesian meshes : Pochoir, PATUS for optimizations/parallelization of a single stencil kernel
+\item Unstructured meshes : Lizst for optimizations/parallelization of a single stencil kernel
+\item Halide for pipeline of stencil computations in image (Cartesian) = mesh dependent
+\item SYCL same Halide mais mesh dependent
 \item Reuse of external solutions as SkelGIS, GA, Lizst, Pochoir etc.

### 2016-02-01 ARTICLES/IJPP/article/compiler.tex
article IJPP corrections+example Helene 01/02/2016

@@
 The overall compiler algorithm is represented in Algorithm~\ref{alg:compiler}.
-First, the parsing of the input file (written with the MSL language) gave enough information to build $\Gamma$, the ordered list of computations.
+First, the parsing of the input file (written with the MSL language) gave enough information to build $\Gamma$, the ordered list of computation kernels.
 Second, $\Gamma_{sync}$ and $\Gamma_{dep}$ can be built from $\Gamma$ using the recursive functions $F_{sync}$ and $F_{dep}$.
@@
 \begin{myth}
-\begin{itemize}
-\item A sequence of sequence is a sequence if the order is kept. Thus for two sequence sub-trees $T_i$ and $T_j$, if $T_j \in V_i$, $T_i$ can be replaced by 
+A sequence of sequence is a sequence if the order is kept. Thus for two sequence sub-trees $T_i$ and $T_j$, if $T_j \in successors(root(T_i))$, $T_i$ can be replaced by 
 \begin{equation*}
@@
 \end{equation*}
-\item A parallel sub-tree of parallel sub-tree is a parallel sub-tree in any order. Thus for two parallel sub-trees $T_i$ and $T_j$, if $T_j \in V_i$, $T_i$ can be replaced by 
+\end{myth}
+
+\begin{myth}
+A parallel sub-tree of parallel sub-tree is a parallel sub-tree in any order. Thus for two parallel sub-trees $T_i$ and $T_j$, if $T_j \in successors(root(T_i))$, $T_i$ can be replaced by 
 \begin{equation*}
@@
 \end{equation*}
-\end{itemize}
 \end{myth}
@@
 
-MSL code~\ref{fig:exmsl}. 
+Figure~\ref{fig:exmsl} gives an example of $\mathcal{MSP}$ program.
+From this example, and using a very simple parser, the following ordered list of computation kernels can be extracted, including all information on mesh, entities, computation domains and data:
 \begin{equation*}
@@
 
+From this ordered list of computation kernels $\Gamma$ and included information, synchronizations can be automatically detected from the call to $F_{sync}(\Gamma,0)$ to get the synchronized ordered list of kernels:
+
 \begin{equation*}
-\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_0,k_6,k_7,k_{7;8}^{sync},k_8]
+\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_0,k_6,k_7,k_{7;8}^{sync},k_8],
 \end{equation*}
 
-where $k_{0;1}^{sync}$...
+\begin{subequations}
+where
+\begin{align}
+        k_{0;1}^{sync}=(\emptyset,\{(B,\bigcup_{e \in D_0} nce(e))\},(B,\bigcup_{e \in D_0} nce(e)),identity),\\
+        k_{1;4}^{sync}=(\emptyset,\{(C,\bigcup_{e \in D_1} nec(e))\},(C,\bigcup_{e \in D_1} nec(e)),identity),\\
+        k_{7;8}^{sync}=(\emptyset,\{(I,\bigcup_{e \in D_7} ncc(e))\},(I,\bigcup_{e \in D_7} ncc(e)),identity).
+\end{align}
+\end{subequations}
 
-$\Gamma_{dep}$~\ref{} and $L^{-1}(\Gamma_{dep})$~\ref{}
-
-% Using those additional reduction rules, the binary series-parallel tree decomposition can be computed in linear time. 
-A complete example of reduction is given in Figure~\ref{fig:example}.
-The initial $\Gamma_{hybrid}$ DAG and its inverse line graph are given first, followed by the set of reductions.
-The label of the final single edge corresponds to the binary tree decomposition of the initial $\Gamma_{hybrid}$ graph.
+From $\Gamma_{sync}$, the dependency DAG can be built.
+For example, as $k_0$ computes $B$ and $k_1$ reads $B$, $k_0$ and $k_1$ becomes vertices of $\Gamma_{dep}$, and an arc $(\overset{\frown}{k_0,k_1})$ is added to $\Gamma_{dep}$.
+The overall $\Gamma_{dep}$ built from the call to $F_{dep}(\Gamma_{sync},0)$ is drawn in Figure~\ref{fig:depdep}.
+$L^{-1}(\Gamma_{dep})$ also represented in Figure~\ref{fig:depline} is the line digraph of $\Gamma_{dep}$.
 
@@
 % INITIAL GRAPH
-\subfloat[]{
+\subfloat[$\Gamma_{dep}$\label{fig:depdep}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
 % INVERSE LINE GRAPH
-\subfloat[]{
+\subfloat[$L^{-1}(\Gamma_{dep})$\label{fig:depline}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
 
+Figure~\ref{fig:SPS} illustrates two binary tree decomposition following the sequence rule of Figure~\ref{fig:rules}.
+This rule is first applied to $k_0$ and $k_{0;1}^{sync}$, then to the resulting tree to $k_1$.
+The same is applied onto $k_7$ and $k_{7;8}^{sync}$ and then to the resulting tree to $k_8$.
+In Figure~\ref{fig:SPP} the parallel rule of Figure~\ref{fig:rules} is applied onto $k_{1;4}^{sync}$ and $k_2$.
+
 \begin{figure}[ht!]
@@
 % TRANSFORMATIONS
-\subfloat[]{
+\subfloat[Sequence rule applied onto $k_0$ $k_{0;1}^{sync}$ and $k_1$, and onto $k_7$ $k_{7;8}^{sync}$ and $k_8$\label{fig:SPS}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
 \\
-\subfloat[]{
+\subfloat[Parallel rule applied onto $k_{1;4}^{sync}$ and $k_2$\label{fig:SPP}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
 
+Finally, the rest of the successive transformations, following rules of Figure~\ref{fig:rules}, are illustrated in Figure~\ref{fig:example} until a single edge is reach which ends the algorithm.
+
 \begin{figure}[ht!]
 \begin{center}
-\subfloat[]{
+\subfloat{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
 \\
-\subfloat[]{
+\subfloat{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]

### 2016-02-01 ARTICLES/IJPP/article/formalism.tex
article IJPP corrections+example Helene 01/02/2016

@@
 Then, the different mid-grain parallelization techniques which can be applied on such program, are presented.
+
+We denote stencil kernels discretized explicit numerical schemes which involve neighborhood computations, whatever the type of space discrtization is (mesh).
 
@@
 \item An entity $e$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $e\subset V\cup E$.
-\item A mesh entities group is denoted $E$ such that $E \in \mathcal{P}(V\cup E)$. It represents a set of entities of the same type.
+\item A group of mesh entities is denoted $E$ such that $E \in \mathcal{P}(V\cup E)$. It represents a set of entities of the same type.
 \item The set of mesh entities groups of a simulation is denoted $\mathcal{E}$.
@@
 For example, in a 2D Cartesian mesh an entity could be a cell, made of four vertices and four edges, or simply a vertex.
-As a result a group of cells and a group of vertices could be defined in $\mathcal{E}$.
+As a result, a group of cells and a group of vertices could be defined in $\mathcal{E}$.
 %This type can be defined as the sets containing exactly four vertices and four edges connected as a cycle.
@@
 \begin{itemize}
-\item A quantity (or data) is a function $\delta: E_{\delta} \mapsto V_{\delta}$ which associates each entity of a group $e\in E_{\delta}$ to a value $v\in V_{\delta}$ where $V_{\delta}$ is a data type, typically $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.
+\item A quantity (or data) is a function $\delta: E_{\delta} \mapsto V_{\delta}$ which associates each entity of a group $e\in E_{\delta}$ to a value $v\in V_{\delta}$ where $V_{\delta}$ is a primitive data type, typically $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.
 \item The set of quantities applied onto the mesh is denoted $\Delta$.
@@
 
-Another option, closer to the applied mathematics domain, would have been to define $\delta$ as the function over $E_{\delta} \times T$; however the approach chosen for now in this work is to let the user be aware of the number of data he is using and what exactly for.
+Another option, closer to the applied mathematics domain, would have been to define $\delta$ as the function over $E_{\delta} \times T$.
+In this case a single quantity would represents all its occurences over time steps.
+This solution could be investigated in future work, however the approach chosen for now is to let the user be aware of the number of data he is using and what exactly for.
 
@@
 
-In this section are considered three different types of kernel computations, stencil kernels, local kernels, and reduction kernels.
+In this section are considered four different types of kernel computations, stencil kernels, boundary kernels, local kernels, and reduction kernels.
 
@@
 \begin{itemize}
-\item A computation domain $D$ is a subpart of a mesh entities group, $D \subset E \in \mathcal{E}$.
+\item A computation domain $D$ is a subpart of a mesh entities group, $D \subseteq E \in \mathcal{E}$.
 \item The set of computation domains of a numerical simulation is denoted $\mathcal{D}$.
@@
 \item $(w,D) \in \Delta \times \mathcal{D}$ is the single data modified by the computation kernel onto a given computation domain.
-Thus, $w \in \Delta$ and $D \in \mathcal{D}$ is the computation domain on which $w$ is computed, $D \subset E_w=entity(w)$.
+Thus, $w \in \Delta$ and $D \in \mathcal{D}$ is the computation domain on which $w$ is computed, $D \subseteq E_w=entity(w)$.
 \item $R$ is the set of tuples $(r,n)$, where $r \in \Delta$ and $n \in \mathcal{N}$ is a neighborhood function such that $n : E_w \rightarrow entity(r)^m$.
 The neighborhood indicates which entities of $r$ will be read in order to compute a single entity of $w$.
-\item $comp$ is the numerical computation which returns a value from a set of input values, $comp: V^n \rightarrow V$.
+\item $comp$ is the numerical computation which returns a value from a set of $n$ input values, $comp: V^n \rightarrow V$.
 \end{itemize}
@@
 In our formalism, it is forbidden for a stencil kernel to find $(r,n) \in R$ for which $r=w$ and $n \neq identity$.
-Actually, if such a property is possible implicit numerical schemes are needed into the simulation.
+Actually, if such a property was permitted, implicit numerical schemes would be needed into the simulation, which involves linear solvers.
 Such a scheme is not a stencil and is over the scope of this paper.
@@
 \medskip
-A reduction computation is a computation which takes a single data applied onto a mesh and returns from all its entities a single data with a dimension reduced to zero, in other words a scalar.
+A reduction computation is a computation which reads a single data applied onto a mesh and returns from all its entities a single scalar (mesh with a dimension reduced to zero).
 A reduction is typically used to compute the convergence criteria of the time loop of the simulation.
-Occasionally reductions can also be performed during a time iteration.
+Occasionally reductions can also be used during a time iteration.
 %This last usage is typically done to have a conditionnal branch to choose one computation or another, or to compute a dynamic time step for example.
@@
 \begin{itemize}
-\item The set of $n$ ordered computation kernels of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i \leq j$, then $k_i$ is computed before $k_j$.
+\item The set of $n$ ordered computation kernels of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i < j$, then $k_i$ is computed before $k_j$.
 \item Finally, a \textit{multi-stencil program} is defined by the octuplet 
@@
 \begin{equation*}
-R: \{(B,n1)\}, \quad w: A, \quad d: dc1,
+R: \{(B,n1)\}, \quad w: A, \quad D: dc1,
 \end{equation*}
 \begin{equation*}
-exp: A(x,y)=B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1).
+comp: A(x,y)=B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1).
 \end{equation*}
@@
 \begin{equation*}
-R: \{(C,n2),(A,identity)\}, \quad w: A, \quad d: dc2,
+R: \{(C,n2),(A,identity)\}, \quad w: A, \quad D: dc2,
 \end{equation*}
 \begin{equation*}
-exp: A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1).
+comp: A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1).
 \end{equation*}
@@
 A stencil program has been formally defined in this section. This formalism is used in the next Section to define two parallelization techniques of a multi-stencil program.
+
+One can note that all definitions given in this section are not dependent from the topology of the mesh. This property will be kept in the rest of this paper to propose the mesh-agnostic MSL language.

### 2016-02-01 ARTICLES/IJPP/article/msl.tex
article IJPP corrections+example Helene 01/02/2016

@@
 stencil shapes : 
-  n1 from cell to cell
-  n2 from cell to edgex
-  n3 from edgex to cell
+  ncc from cell to cell
+  nce from cell to edgex
+  nec from edgex to cell
 data :

### 2016-02-01 ARTICLES/IJPP/article/msp.tex
article IJPP corrections Helene 29/01/2016
article IJPP corrections+example Helene 01/02/2016

@@
 %--------------------
-\subsection{Minimal series-parallel DAG}
+\subsection{Minimal Series-Parallel DAG}
 \begin{mydef}
@@
 
-In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of minimal series-parallel DAGs (MSPD).
+In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of Minimal Series-Parallel DAGs (MSPD).
 
@@
 Actually, it is possible in a multi-stencil program to have a set of computations such that their dependencies form the forbidden $N$ subgraph.
-Considering four computation kernels $k_0$, $k_1$, $k_2$ and $k_3$ of a $\mathcal{MSP}$ such that $k_0 \prec k_1$, $k_2 \prec k_1$ and $k_2 \prec k_3$ are verified.
-Thus, the \emph{zigzag} relation $k_0 \prec k_1 \succ k_2 \prec k_3$ which form a forbidden subgraph of a MSPD is found in $\Gamma_{dep}$.
+For example, considering four computation kernels $k_0$, $k_1$, $k_2$ and $k_3$ of a $\mathcal{MSP}$ such that $k_0 \prec k_1$, $k_2 \prec k_1$ and $k_2 \prec k_3$ are verified.
+The \emph{zigzag} relation $k_0 \prec k_1 \succ k_2 \prec k_3$ which form a forbidden subgraph of a MSPD is found in $\Gamma_{dep}$.
 \end{proof}
@@
 However, we have chosen to exactly keep dependencies computed in $\Gamma_{dep}$, to not approximate the static scheduling.
-Finally, algorithm proposed by~\cite{Bender200575} to find a Lowest Common Ancestor in a DAG (in $\mathcal{O}(|V|^3)$, could have been also used.
+Finally, algorithm proposed by~\cite{Bender200575} to find a Lowest Common Ancestor in a DAG (in $\mathcal{O}(|V|^3)$, could have also been used.
 
-The first step of Algorithm~\ref{alg:nshape} is to create a new DAG $\Gamma_{dep}'$ from $\Gamma_{dep}$, such that vertices having exactly one source and one destination in $\Gamma_{dep}$ is not added to $\Gamma_{dep}'$, and its source and destination vertices are directly linked by an edge.
+The first step of Algorithm~\ref{alg:nshape} is to create a new DAG $\Gamma_{dep}'$ from $\Gamma_{dep}$, such that vertices having exactly one source and one destination in $\Gamma_{dep}$ are not added to $\Gamma_{dep}'$, and its source and destination vertices are directly linked by an edge.
 This step is called $RTV(\Gamma_{dep})$ (line 2) and is done in $\mathcal{O}(|V|)$.
@@
 \Procedure{removeNSH} {$\Gamma_{dep}$}
-\\$\Gamma_{dep}'$ = RTV($\Gamma_{dep}$)
-\\$sources$ = roots($\Gamma_{dep}'$)
-\\BFSnshape($sources$,$\Gamma_{dep}'$,$\Gamma_{dep}$)
-\\TC($\Gamma_{dep}$)
+\State $\Gamma_{dep}'$ = RTV($\Gamma_{dep}$)
+\State $sources$ = roots($\Gamma_{dep}'$)
+\State BFSnshape($sources$,$\Gamma_{dep}'$,$\Gamma_{dep}$)
+\State TC($\Gamma_{dep}$)
 \EndProcedure
@@
 \Procedure{BFSnshape} {$sources$, $\Gamma_{dep}'$, $\Gamma_{dep}$}
-\\$successors$ = successors($sources$)
-\\$SG$ = subgraph($sources \cup successors$)
+\State $successors$ = successors($sources$)
+\State $SG$ = $\Gamma_{dep}'$.subgraph($sources \cup successors$)
 \If{isconnected($SG$)}
 \If{! complete($SG$)}
-\\addedges($\Gamma_{dep}'$)
-\\addedges($\Gamma_{dep}$)
-\\BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
-\EndIf
-\If{! isconnected($SG$)}
-\\BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
+\State addedges($\Gamma_{dep}'$)
+\State addedges($\Gamma_{dep}$)
+\State BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
+\ElsIf{! isconnected($SG$)}
+\State BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
 \EndIf
@@
 As a result, the binary tree decomposition algorithm of Valdes \& Al can be applied on $\Gamma_{dep}$ to get a static scheduling of the multi-stencil program.
+
+Proof of the algorithm and complexity improvments could be studied in future works.
+However, as it will be shown in Section~\ref{sect:eval}, the complexity of this algorithm is not a problem because of the relatively small size of $\Gamma_{dep}$.
 

### 2016-02-01 ARTICLES/IJPP/article/parallelism.tex
article IJPP corrections+example Helene 01/02/2016

@@
 \label{sect:dataparal}
-In a data parallelization technique, the idea is to split the data on which the program is computed into balanced sub-parts, one for each available resource.
+In a data parallelization technique, the idea is to split quantities on which the program is computed into balanced sub-parts, one for each available resource.
 The same sequential program can afterwards be applied on each sub-part simultaneously, with some additioinal synchronizations between resources to update the data not computed locally, and thus to guarantee a correct result.
@@
 
-Actually, a synchronization is needed by the data read by a stencil computation (not local), and only if this data has been modified before, which means that it has been written before.
+Actually, a synchronization is needed by the data read by a stencil computation (not local), if this data has been modified before, which means that it has been written before.
 This synchronization is needed because a neighborhood function $n \in \mathcal{N}$ of a stencil computation involves values computed on different resources.
@@
 When data parallelism is applied, the other type of computation which is responsible for additional synchronizations is the reduction.
-Actually, the reduction is applied locally on each subset of entities, on each resource, which results in $p$ (number of resources) scalar values, one on each resource.
-Thus, to perform the final reduction a synchronization is needed to get the $p$ scalars onto a single resource, to finally get the reduction scalar.
+Actually, the reduction is first applied locally on each subset of entities, on each resource.
+This results in $p$ (number of resources) scalar values.
+Thus, to perform the final reduction a synchronization is needed to get the $p$ scalars onto each resource.
+The final reduction step is applied such that each resource get the result scalar.
 
 \begin{mydef}
-For a reduction kernel $k_j(S_j,R_j,(w_j,D_j),comp)$, if we denote by $w^r$, $0 \leq r<p$, the local $w$ computed on the resource $r$, a reduction synchronization is defined as the specific computation
+For a reduction kernel $k_j(S_j,R_j,(w_j,D_j),comp)$, if we denote by $w^r$, $0 \leq r<p$, the local scalar $w$ computed on the resource $r$, a reduction synchronization is defined as the specific computation
 \begin{equation*}
@@
 From the ordered list of computation $\Gamma$, a new ordered list $\Gamma_{sync}$ is obtained from the call $\Gamma_{sync} = F_{sync}(\Gamma,0)$, where $F_{sync}$ is the recursive function defined in Algorithm~\ref{alg:sync}.
+
+% \begin{equation*}
+% F_{sync}(\Gamma,j) = 
+% \begin{cases} 	\bullet [] \mbox{ if }j=|\Gamma|\\
+% 				\bullet [k_{sync}(\emptyset,\\
+% 				\{(w_i,\bigcup_{e \in entity(w_i)} n_j(e))\},\\
+% 				(w_i,\bigcup_{e \in entity(w_i)} n_j(e))),\\
+% 				identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1)\\
+% 				\text{ } \qquad \forall i<j \text{, and }k_i \pprec k_j\\
+% 				\bullet [k_j, k_{sync}(\emptyset,\\
+% 				\{(w^0_j,entity(w^0_j)) \dots (w^{p-1}_j,entity(w^{p-1}_j))\},\\
+% 				(w_j,entity(w_j)),comp_j)] \cdot F_{sync}(\Gamma,j+1)\\ 
+% 				\text{ } \qquad \mbox{if } w_j \in \mathcal{S} \text{, and } R_j=\{(r_j,entity(r_j))\}\\
+% 				\bullet [\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) \text{ else}.
+% \end{cases}
+% \end{equation*}
+\end{mydef}
 
@@
 \State $list.[k_{i;j}^{sync}(S,R,(w,D),comp)$
-\State break
 \EndIf
@@
 
-% \begin{equation*}
-% F_{sync}(\Gamma,j) = 
-% \begin{cases} 	\bullet [] \mbox{ if }j=|\Gamma|\\
-% 				\bullet [k_{sync}(\emptyset,\\
-% 				\{(w_i,\bigcup_{e \in entity(w_i)} n_j(e))\},\\
-% 				(w_i,\bigcup_{e \in entity(w_i)} n_j(e))),\\
-% 				identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1)\\
-% 				\text{ } \qquad \forall i<j \text{, and }k_i \pprec k_j\\
-% 				\bullet [k_j, k_{sync}(\emptyset,\\
-% 				\{(w^0_j,entity(w^0_j)) \dots (w^{p-1}_j,entity(w^{p-1}_j))\},\\
-% 				(w_j,entity(w_j)),comp_j)] \cdot F_{sync}(\Gamma,j+1)\\ 
-% 				\text{ } \qquad \mbox{if } w_j \in \mathcal{S} \text{, and } R_j=\{(r_j,entity(r_j))\}\\
-% 				\bullet [\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) \text{ else}.
-% \end{cases}
-% \end{equation*}
-\end{mydef}
-
- The final step of this parallelization is to run $\Gamma_{sync}$ on each resource. Thus, for each resource $0 \leq m \leq p-1$ a multi-stencil program, defined by
+ The final step of this parallelization is to run $\Gamma_{sync}$ on each resource. Thus, for each resource $0 \leq r \leq p-1$ a multi-stencil program, defined by
 \begin{equation}
-\mathcal{MSP}_m(T,\mathcal{M}_m,\mathcal{E}_m,\mathcal{D}_m,\mathcal{N},\Delta_m,\mathcal{S},\Gamma_{sync}),
+\mathcal{MSP}_r(T,\mathcal{M}_r,\mathcal{E}_r,\mathcal{D}_r,\mathcal{N},\Delta_r,\mathcal{S},\Gamma_{sync}),
 \end{equation}
@@
 A dependency graph exhibits parallel tasks, or on the contrary sequential execution of tasks.
-Such a dependency graph can directly be given to a dynamic scheduler, or can be statically scheduled.
-In this paper, we introduce task parallelism by building the dependency graph between computations of the sequential list $\Gamma_{sync}$ (which itself take into account data parallelism).
+Such a dependency graph can directly be given to a dynamic scheduler, or can statically be scheduled.
+In this paper, we introduce task parallelism by building the dependency graph between kernels of the sequential list $\Gamma_{sync}$.
+Thus, as $\Gamma_{sync}$ takes into account data parallelism, we introduce hybrid parallelism.
 
@@
 
+Those definitions are close to traditionnal read/write dependencies.
+However, specific conditions due to the specific domain can be introduced, such as the condition on boundary kernels, and conditions on the computation domain.
+
 \begin{mydef}
@@
 \begin{mydef}
-From the ordered list $\Gamma_{sync}$ of computation kernels, a directed dependency graph $\Gamma_{dep}(V,A)$ is obtained from the call $F_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function
+From the ordered list $\Gamma_{sync}$ of computation kernels, a directed dependency graph $\Gamma_{dep}(V,A)$ is obtained from the call $F_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function defined in Algorithm~\ref{alg:dep}.
+
+% \begin{equation*}
+% F_{dep}(\Gamma_{sync},j) = 
+% \begin{cases} 	\bullet (\{\},\{\}) \mbox{ if }j=|\Gamma_{sync}|\\
+% 				\bullet (k_j, \{(\overset{\frown}{k_i,k_j})\mbox{, }\forall i < j \mbox{, } k_i\prec k_j \})\\
+% 				\text{ } \qquad \cup F_{dep}(\Gamma_{sync},j+1) \mbox{ if }j<|\Gamma_{sync}|
+% \end{cases}
+% \end{equation*}
+\end{mydef}
 
@@
 \caption{$F_{dep}$ recursive function}
-\label{alg:sync}
+\label{alg:dep}
 \begin{algorithmic}[1]
@@
 
-% \begin{equation*}
-% F_{dep}(\Gamma_{sync},j) = 
-% \begin{cases} 	\bullet (\{\},\{\}) \mbox{ if }j=|\Gamma_{sync}|\\
-% 				\bullet (k_j, \{(\overset{\frown}{k_i,k_j})\mbox{, }\forall i < j \mbox{, } k_i\prec k_j \})\\
-% 				\text{ } \qquad \cup F_{dep}(\Gamma_{sync},j+1) \mbox{ if }j<|\Gamma_{sync}|
-% \end{cases}
-% \end{equation*}
-\end{mydef}
-
 This constructive function is possible because the input is an ordered list.
@@
 
-\begin{proof}
-$\Gamma_{dep}$ is built from $\Gamma_{sync}$ which is an ordered and sequential list of computations.
+% \begin{proof}
+% $\Gamma_{dep}$ is built from $\Gamma_{sync}$ which is an ordered and sequential list of computations.
 Moreover, each computation of the list $\Gamma_{sync}$ is associated to a vertex of $V$, even if the same computation is represented more than once in $\Gamma_{sync}$.
 As a result it is not possible to go back to a previous computation and to create a cycle.
-\end{proof}
+% \end{proof}
 
-As a result of the hybrid parallelization, each resource $0 \leq m \leq p-1$ perform a multi-stencil program, defined by
+As a result of the hybrid parallelization, each resource $0 \leq r \leq p-1$ perform a multi-stencil program, defined by
 \begin{equation*}
-\mathcal{MSP}_m(T,\mathcal{M}_m,\mathcal{E}_m,\mathcal{D}_m,\mathcal{N},\Delta_m,\Gamma_{dep}).
+\mathcal{MSP}_r(T,\mathcal{M}_r,\mathcal{E}_r,\mathcal{D}_r,\mathcal{N},\Delta_r,\Gamma_{dep}).
 \end{equation*}
-The set of computations $\Gamma_{dep}$ is a dependency graph between computations $k_i$ of $\Gamma$ and synchronizations of kernels added into $\Gamma_{sync}$.
+The set of computations $\Gamma_{dep}$ is a dependency graph between computation kernels $k_i$ of $\Gamma$ and synchronizations of kernels added into $\Gamma_{sync}$.
 $\Gamma_{dep}$ can be built from the call to

### 2016-05-20 ARTICLES/IJPP/article/article.tex
IJPP article modifications plan + partie 2 + partie 3

@@
 %\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
-%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
-\documentclass[twocolumn]{svjour3}          % twocolumn
+\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+%\documentclass[twocolumn]{svjour3}          % twocolumn
 %
@@
 \institute{Helene Coullon \and Christian Perez \at
-              INRIA \\
-              46 Allée d'Italie\\
-              69007 Lyon\\
+              INRIA, ENS Lyon, 46 Allée d'Italie, 69007 Lyon\\
               \email{helene.coullon@inria.fr}
@@
 \input{formalism.tex}
-% Formalism :
-% \begin{itemize}
-% \item mesh / mesh entities and groups / computations domains
-% \item data and scalars
-% \item computations : stencil, local and reduction
-% \item Multi-Stencil Program
-% \end{itemize}
+%------------------------------------------------------------------------------
+\section{The Multi-Stencil Language}
+\label{sect:msl}
+\input{msl.tex}
 %------------------------------------------------------------------------------
@@
 \input{parallelism.tex}
-% Parallelization formalism :
-% \begin{itemize}
-% \item Data parallelism (introduction of synchronizations)
-% \item Task parallelism (computation of the dependency graph)
-% \end{itemize}
-%------------------------------------------------------------------------------
-\section{The Multi-Stencil Language}
-\label{sect:msl}
-\input{msl.tex}
-The parallel empty skeleton of the application can be computed automatically from the language :
-\begin{itemize}
-\item grammar
-\item example from the language to $\Gamma$
-\end{itemize}
 %------------------------------------------------------------------------------

### 2016-05-20 ARTICLES/IJPP/article/formalism.tex
IJPP article modifications plan + partie 2 + partie 3

@@
 Among those numerical computations are found a set of numerical schemes, also called \textit{stencil computations}, a set of auxiliary computations also called \emph{local computations}, and finally a set of \emph{reductions} to reduce a bunch of values to a single scalar value.
-This section gives formal definitions of what we call a \textit{stencil program} and its computations.
-Then, the different mid-grain parallelization techniques which can be applied on such program, are presented.
+In the rest of this paper, we denote a \emph{stencil kernel} as a discretized explicit numerical schemes which involve neighborhood computations, whatever the type of space discretization is (mesh).
 
-We denote stencil kernels discretized explicit numerical schemes which involve neighborhood computations, whatever the type of space discrtization is (mesh).
+This section gives a complete formal description of what we call a \textit{stencil program} and its computations.
+In this paper, the proposed language MSL (Multi-Stencil Language) does not depend on the type of space discretization used.
+Thus, some details given in the following sections, are usefull to understand the domain but do not have to be kept in mind to understand the Multi-Stencil Language.
+As a result, at the end of each subsection a summary of what is usefull to understand the language is given.
 
@@
   \caption{From left to right, Cartesian, curvilinear and unstructured meshes.}
-  \label{fig:mesh}
+  \label{fig:mesh1}
 \end{center}\end{figure}
@@
 \end{mydef}
-A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) and hybrid as illustrated in Figure~\ref{fig:mesh}.
+A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) as illustrated in Figure~\ref{fig:mesh1}.
+One can notice that more than one type of mesh is also possible inside a single simulation.
+For example, an hybrid mesh can be defined as an unstructured mesh composed itself of a Cartesian mesh inside each of its vertices.
+Multi-grid numerical methods is another example.
+However, in this paper a single mesh simulation is addressed.
 
 \medskip
-\noindent \textbf{Definitions}
+\noindent \textbf{Definitions (mesh)}
 %\begin{mydef}
 \begin{itemize}
-\item An entity $e$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $e\subset V\cup E$.
-\item A group of mesh entities is denoted $E$ such that $E \in \mathcal{P}(V\cup E)$. It represents a set of entities of the same type.
-\item The set of mesh entities groups of a simulation is denoted $\mathcal{E}$.
-\item Finally, the mesh on which is applied a group of mesh entities $E$ is denoted $mesh(E)$
+\item An entity $\phi$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $\phi\subset V\cup E$.
+\item A group of mesh entities, denoted $\mathcal{G}$, is a set of mesh entities which are composed of the same topology and types of elements in $V\cup E$.
+\item The set of mesh entities groups of a simulation is denoted $\Phi$.
+\item Finally, the mesh on which is applied a group of mesh entities $\mathcal{G}$ is denoted $mesh(\mathcal{G})$
 \end{itemize}
@@
 For example, in a 2D Cartesian mesh an entity could be a cell, made of four vertices and four edges, or simply a vertex.
-As a result, a group of cells and a group of vertices could be defined in $\mathcal{E}$.
+As a result, a group of cells and a group of vertices could be defined in $\Phi$.
 %This type can be defined as the sets containing exactly four vertices and four edges connected as a cycle.
@@
 \medskip
-This covers the space discretization, however there is also the time dimension which has to be discretized in the simulation.
+This covers the space discretization, however there is also the time dimension which has to be discretized to perform a numerical simulation.
+
+\begin{mydef}
+We denote a scalar as a numerical variable. The set of scalars is denoted $\mathcal{S}$.
+\end{mydef}
 
 \medskip
-\noindent \textbf{Definitions}
+\noindent \textbf{Definitions (time)}
 %\begin{mydef}
 \begin{itemize}
-\item We denote a scalar as an identifier associated to a numerical value and applied onto a mesh $\mathcal{M}$ where $dim(\mathcal{M})=0$.
-In other words a scalar can be seen as a variable containing a numerical value.
-The set of scalars is denoted $\mathcal{S}$.
 \item $\mathcal{T}=\mathbb{R}$ is the continuous time domain of a numerical simulation.
@@
 \item $\Delta t \in \mathcal{S}$ represents the time interval of the numerical simulation, such that for a current time iteration $t_i$, the next time iteration is $t_{i+1} = t_i + \Delta t$.
-The default value of $\Delta t$ is $1$.
 \item $conv:\mathcal{S}^n \rightarrow bool$ is a function which returns a boolean from a set of scalar variables. This function represents the convergence criteria of the simulation. 
-\item At a given time step, the convergence criteria is evaluated such that if $conv$ returns $true$ the next time step can start such that $t_{i+1} = t_i + \Delta t$, otherwise the simulation ends.
-For example, the simplest $conv$ is typically to have two scalars: $t$ the time, and a fix number of iterations denoted $it$.
+\end{itemize}
+\end{itemize}
+
+At a given time step, the convergence criteria is evaluated such that if $conv$ returns $true$ the next time step $t_{i+1} = t_i + \Delta t$ can start.
+Otherwise the simulation ends.
+For example, the simplest $conv$ is typically to have two scalars: $t$ the time, and a constant number of iterations denoted $it$.
 At each time step $\Delta t=1$, and $conv$ returns the boolean expression $t<it$.
-\end{itemize}
+%\end{mydef}
+
+In a numerical simulation, a scalar can be associated to each entity of the mesh. In such a case, a single variable is used and is denoted a quantity.
+%A quantityas a set of scalars can be applied onto a mesh with a nil dimension, a set of data elements, or quantities, can be applied onto meshes of dimension superior to zero.
+Those quantities represent, as well as scalars, the set of values to compute, or to use, for computations.
+
+\medskip
+\noindent \textbf{Definitions (quantity)}
+%\begin{mydef}
+\begin{itemize}
+\item A quantity represents a variable applied onto a group of mesh entities $\mathcal{G}$.
+\item The set of quantities applied onto the mesh is denoted $\Delta$.
+\item In the rest of this paper, the group of mesh entities on which a quantity $\delta$ is mapped is denoted $entity(\delta)=\mathcal{G}_{\delta}$.
 \end{itemize}
@@
 
-%$T$ is responsible for the iteration time steps of the numerical simulation.
-A numerical simulation though does not run forever and must be stopped at some point.
-Some simulations choose a fix number of time iteration, others compute a \emph{convergence} criteria at the end of each time iteration to determine if the simulation has to continue.
-A convergence criteria is computed by a reduction computation that will be detailed in the next section.
+Another option, closer to the applied mathematics domain, would have been to apply a quantity to each entity of a group $\mathcal{G}$, but also to each time iteration.
+In this case a single quantity would represents all its occurences over time steps.
+This solution could be investigated in future work, however the approach chosen in this work is to let the user be aware of the number of data he is using and what exactly for.
 
-In a numerical simulation, as a set of scalars can be applied onto a mesh with a nil dimension, a set of data elements, or quantities, can be applied onto meshes of dimension superior to zero.
-Those quantities represent, as well as scalars, the set of values to compute, or to use, for computations.
-
-\medskip
-\noindent \textbf{Definitions}
-%\begin{mydef}
-\begin{itemize}
-\item A quantity (or data) is a function $\delta: E_{\delta} \mapsto V_{\delta}$ which associates each entity of a group $e\in E_{\delta}$ to a value $v\in V_{\delta}$ where $V_{\delta}$ is a primitive data type, typically $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.
-\item The set of quantities applied onto the mesh is denoted $\Delta$.
-\item In the rest of this paper, the group of mesh entities on which a quantity $\delta$ is mapped is denoted $entity(\delta)=E_{\delta}$.
-\end{itemize}
-%\end{mydef}
-
-Another option, closer to the applied mathematics domain, would have been to define $\delta$ as the function over $E_{\delta} \times T$.
-In this case a single quantity would represents all its occurences over time steps.
-This solution could be investigated in future work, however the approach chosen for now is to let the user be aware of the number of data he is using and what exactly for.
+\paragraph{\textbf{Summary}} In this subsection have been presented the general formalism of what is a mesh, its entities and groups of entities, of how the time is discretized, and finally of what is a scalar and a quantity.
+The language presented in this paper, the Multi-Stencil Language, offers a simple way to describe a numerical simulation while being able to automatically extract parallelism.
+Its aim is to be mesh-agnostic, thus to not have details about the topology of the mesh which is an implementation problem.
+As a result the details about the mesh topology is not needed in the rest of this paper.
 
@@
 \begin{itemize}
-\item A computation domain $D$ is a subpart of a mesh entities group, $D \subseteq E \in \mathcal{E}$.
+\item A computation domain $D$ is a subpart of a mesh entities group, $D \subseteq \mathcal{G} \in \Phi$.
 \item The set of computation domains of a numerical simulation is denoted $\mathcal{D}$.
-\item A neighborhood $n$ is a function which for a given entity $e \in E_i$, returns a set of $m$ entities in $E_j$, $n : E_i \rightarrow E_j^m$.
+\item A neighborhood $n$ is a function which for a given entity $\phi \in \mathcal{G}_i$, returns a set of $m$ entities in $\mathcal{G}_j$, $n : \mathcal{G}_i \rightarrow \mathcal{G}_j^m$.
 One can notice that $i = j$ is possible.
-Most of the time, such a neighborhood is called a stencil shape.
+Most of the time, such a neighborhood is called a \emph{stencil shape}.
 \item The set of neihborhood functions in a numerical simulation is denoted $\mathcal{N}$.
@@
 \item $(w,D) \in \Delta \times \mathcal{D}$ is the single data modified by the computation kernel onto a given computation domain.
-Thus, $w \in \Delta$ and $D \in \mathcal{D}$ is the computation domain on which $w$ is computed, $D \subseteq E_w=entity(w)$.
-\item $R$ is the set of tuples $(r,n)$, where $r \in \Delta$ and $n \in \mathcal{N}$ is a neighborhood function such that $n : E_w \rightarrow entity(r)^m$.
+Thus, $w \in \Delta$ and $D \in \mathcal{D}$ is the computation domain on which $w$ is computed, $D \subseteq \mathcal{G}_w=entity(w)$.
+\item $R$ is the set of tuples $(r,n)$, where $r \in \Delta$ and $n \in \mathcal{N}$ is a neighborhood function such that $n : \mathcal{G}_w \rightarrow entity(r)^m$.
 The neighborhood indicates which entities of $r$ will be read in order to compute a single entity of $w$.
-\item $comp$ is the numerical computation which returns a value from a set of $n$ input values, $comp: V^n \rightarrow V$.
+\item $comp$ is the numerical computation which returns a value from a set of $n$ input values, $comp: V^n \rightarrow V$, where $V$ is typically $\mathbb{N}$, $\mathbb{Z}$, or $\mathbb{R}$.
+$comp$ represents the actual numerical expression which is computed by a kernel.
 \end{itemize}
-% \item A numerical expression $\text{exp}_{w}: entity(w) \times R^n \rightarrow V_{w}$ is a function which computes the values of the written quantity $w \in \Delta$ for a given entity $e \in entity(w)=E_w$, using a set of input data $R$.
-% \item A computation kernel $k$ of a numerical simulation is defined as $c(R,w,\text{exp},D)$, where $R$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$ is the computation domain.
-% \end{itemize}
-%$R$ is the set of data read, $w \in \Delta$ is the unique data written, $\text{exp}$ is a numerical expression, and $D \in \mathcal{D}$ is the computation domain.
 \end{mydef}
 
-$comp$ represents the actual numerical expression which is computed by a kernel. We can dissociate four types of kernel computations.
+At each time iteration, a set of computations is performed.
+During a computation kernel, it can be considered that a set of old states ($t-1$) of quantities are read ($R$), and that a new state ($t$) of a single quantity is written ($w$).
+In this paper only explicit numerical schemes are considered, thus if we denote by $identity$ the identity function, we have $w \not\in R$ except if $w=(r,identity)$.
 
-\noindent \textbf{Definitions}
-%\begin{mydef}
+Such a definition of a computation kernel offers a large panel of different computations.
+For example, the four usual types of computations (stencil, local, boundary and reduction) performed into a simulation can be defined as follow :
 \begin{itemize}
-\item We denote by $identity$ the identity function $x=x$.
-\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{stencil kernel} $\iff \exists (r,n) \in R$ such that $n \neq identity$ and if $\exists (r,n)$ with $r=w$ then $n=identity$.
-\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{boundary kernel} $\iff \exists (r,n) \in R$ such that $r=w$ and such that $n \neq identity$.
-\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{local kernel} $\iff \forall (r,n) \in R$, $n = identity$.
-\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} $\iff w$ is a scalar, $R=\{(r,n)\}$, and if $n=entity(r)$.
+\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{stencil kernel} if $\exists (r,n) \in R$ such that $n \neq identity$ .
+\item A \emph{boundary kernel} is a kernel $k(S,R,(w,D),comp)$ where $D$ is a specific computation domain at the border of entities, and which does not intersect with any other computation domain.
+\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{local kernel} if $\forall (r,n) \in R$, $n = identity$.
+\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} if $w$ is a scalar, $D=\emptyset$, $R=\{(r,n)\}$ with $n=entity(r)$, and $comp$ is a reduction operation (usually $+$, $-$, $*$, $max$, $min$ etc.).
 \end{itemize}
 
-In our formalism, it is forbidden for a stencil kernel to find $(r,n) \in R$ for which $r=w$ and $n \neq identity$.
+%In our formalism, it is forbidden for a stencil kernel to find $(r,n) \in R$ for which $r=w$ and $n \neq identity$.
 Actually, if such a property was permitted, implicit numerical schemes would be needed into the simulation, which involves linear solvers.
@@
 
-\noindent \textbf{Property}
-A kernel for which all data read and written are applied onto a mesh of dimension $0$ is a local kernel.
+% \noindent \textbf{Property}
+% A kernel for which all data read and written are applied onto a mesh of dimension $0$ is a local kernel.
 
-\noindent \textbf{Property}
-In a reduction kernel $k(S,R,(w,D),comp)$, $D=entity(w)$ as a single entity exists for a scalar.
+% \noindent \textbf{Property}
+% In a reduction kernel $k(S,R,(w,D),comp)$, $D=entity(w)$ as a single entity exists for a scalar.
 
-\medskip
-A reduction computation is a computation which reads a single data applied onto a mesh and returns from all its entities a single scalar (mesh with a dimension reduced to zero).
-A reduction is typically used to compute the convergence criteria of the time loop of the simulation.
-Occasionally reductions can also be used during a time iteration.
-%This last usage is typically done to have a conditionnal branch to choose one computation or another, or to compute a dynamic time step for example.
+%\medskip
+%A reduction computation is a computation which reads a single data applied onto a mesh and returns from all its entities a single scalar (mesh with a dimension reduced to zero).
 
-\medskip
-\noindent \textbf{Property}
-Considering a reduction kernel $k(S,R,(w,D),comp)$, $comp$ must be a binary and associative operation on the type $V$, $comp: V \times V \rightarrow V$.
+A reduction is typically used to compute the convergence criteria of the time loop of the simulation. Occasionally reductions can also be used during a time iteration.
 
-\medskip
-\noindent \textbf{Definitions}
-\begin{itemize}
-\item The set of $n$ ordered computation kernels of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i < j$, then $k_i$ is computed before $k_j$.
-\item Finally, a \textit{multi-stencil program} is defined by the octuplet 
+% \medskip
+% \noindent \textbf{Property}
+% Considering a reduction kernel $k(S,R,(w,D),comp)$, $comp$ must be a binary and associative operation on the type $V$, $comp: V \times V \rightarrow V$.
+
+\begin{mydef}
+The set of $n$ ordered computation kernels of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i < j$, then $k_i$ is computed before $k_j$.
+\end{mydef}
+
+\begin{mydef}
+A \textit{multi-stencil program} is defined by the octuplet 
 \begin{equation*}
-\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\mathcal{N},\Delta, \mathcal{S},\Gamma)
+\mathcal{MSP}(\mathcal{M},\Phi,\mathcal{D},\mathcal{N},\Delta, \mathcal{S},T,\Gamma)
 \end{equation*}
-\end{itemize}
+\end{mydef}
 
+\paragraph{\textbf{Summary}} In this subsection have been presented the general formalism of what is a computation kernel and a multi-stencil program.
+The language presented in this paper, the Multi-Stencil Language, offers a simple way to describe a numerical simulation while being able to automatically extract parallelism.
+Its aim is to be mesh-agnostic, as previously described, but also to make a clear separation of concerns between the simulation description and its implementation.
+As a result, and as already described, details about the topology of the mesh, which is an implementation problem, is not used into the language.
+Moreover, in a computation kernel $k(S,R,(w,D),comp)$, $comp$ is not used by the language.
+
+\paragraph{\textbf{Example}}
 For example, in Figure~\ref{fig:ex1}, assuming that the computation domain (full lines) is denoted $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined as:

### 2016-05-20 ARTICLES/IJPP/article/msl.tex
IJPP article modifications plan + partie 2 + partie 3

--- 2016-01-29 ARTICLES/IJPP/article/msl.tex
@@
+From the fomalism detailed in the previous section, the Multi-Stencil Language and its grammar can already be given.
+The reason why this grammar is sufficient to automatically extract parallelism will be explained in the next section.
+
+\medskip
+The grammar of the Multi-Stencil Language is given in Figure~\ref{fig:grammar}. A Multi-Stencil program is composed of eight parts.
+
+\begin{enumerate}
+\item The first part represents the mesh of the simulation $\mathcal{M}$.
+As a single mesh is for now supported by the language, and as the topology of the mesh is not needed by the language compiler, \textit{meshid} (line 1) is a terminal symbol of the grammar.
+\item The second part (line 2) represents the set of groups of mesh entities needed by the simulation $\Phi$.
+It is defined as a list of groups $\mathcal{G}$.
+Again, as mesh entities are defined using the topology of the mesh, a group of mesh entities \textit{group} (defined line 13) is a terminal symbol of the grammar.
+\item The third part (lines 3 and 4) represents the set of computation domains of the simulation $\mathcal{D}$.
+It is defined as a list of domains, each one defined (lines 14 and 15) as a subpart of a group of mesh entities, using the symbol "in".
+\item The fourth part (lines 5 and 6) of the program is linked to the third one.
+It offers a way to declare that two computation domains are independants (lines 16 and 17).
+Two given computation domains $D_1$ and $D_2$ are independants if and only if $D_1 \cap D_2 = \emptyset$.
+It uses the symbol "and" to describe that $D_1$ and $D_2$ are independants.
+\item The fith part (lines 7 and 8) represents the set of stencil shapes used by the simulation $\mathcal{N}$.
+A stencil shape (lines 18 and 19) is defined as a function "from" a group of mesh entities "to" another group of mesh entities.
+\item The sixth part (line 9) represents the set of quantities of the simulation $\Delta$.
+It is composed of a list of quantities (lines 20 and 21) for which each line (line 22) begins with a group of mesh entities and a list of \textit{quanityid} which are applied onto the given group.
+\item The seventh part (line 10) represents the set of scalars of the simulation $\mathcal{S}$. Each scalar (line 23) is a terminal symbol.
+\item Finally, the eith part (line 11) represents the simulation computations.
+It is composed of a list of loop (line 24).
+Each \textit{loop} (line 25) is composed of a \textit{time} loop composed of \textit{iteration} (which represents $T$), and of a set of \textit{computations} (which represents $\Gamma$).
+\begin{itemize}
+\item A time loop, denoted \textit{iteration} in the grammar, can be a numerical constant value, \textit{num}, which directly indicates the number of time iterations, or a convergence \textit{criteria}.
+The \textit{criteria} symbol represents the function $conv:\mathcal{S}^n \rightarrow bool$ of the formalism.
+To define a criteria the terminal \textit{criteriaid} has to be set, and between bracket are indicated a list of scalars to read.
+\item Each computation of the list of \textit{computations} follows the information of the formalism $k(S,R,(w,D),comp)$ except the $comp$ function which is an implementation concern.
+To be more easy to write the syntax is different though.
+\end{itemize}
+\end{enumerate}
+
 \begin{filecontents*}{grammar.txt}
 program ::= "mesh:" meshid 
-            "mesh entities:" listmeshent
+            "mesh entities:" listgroup
             "computation domains:" 
@@
                       liststencil
-            "data:" listdata
+            "quantities:" listquantities
             "scalar" listscalar
@@
 
-listmeshent ::= meshent listmeshent 
-                        |  meshent
-listcompdom ::= compdom listcompdom 
-                        |  compdom
-compdom ::= compdomid "in" listmeshent
+listgroup ::= group listgroup |  group
+listcompdom ::= compdom listcompdom |  compdom
+compdom ::= compdomid "in" group
 listinde ::= inde listinde |  inde
 inde ::= compdomid "and" compdomid
-liststencil ::= stencil liststencil 
-                        | stencil
-stencil ::= stencilid "from" meshent "to" meshent
-listdata ::= data listdata |  data
-data ::= dataid "," meshent
-listscalar ::= scalar listscalar 
-                      | scalar
-
+liststencil ::= stencil liststencil | stencil
+stencil ::= stencilid "from" group "to" group
+listquantities ::= data listquantities |  quantity
+quantity ::= group listquantityid
+listquantityid ::= quantityid listquantityid |  quantityid
+listscalar ::= scalar listscalar | scalar
 listloop ::= loop listloop | loop
@@
           "computations:" listcomp
-
 iteration ::= num | criteria
 criteria ::= criteriaid "(" listscalarread ")"
-listscalarread ::= scalar listscalarread 
-                          |  scalar
+listscalarread ::= scalar listscalarread |  scalar
 listcomp ::= comp listcomp |  comp
-comp ::= dataid "[" compdomid "]=" 
-                    compid "({" 
-                    listscalarread "},{" 
-                    listdataread "})"
-listdataread ::= dataread listdataread 
-                          |  dataread
-dataread ::= dataid "[" stencil "]" 
-              |  dataid
+comp ::= dataid "[" compdomid "]=" compid "({"listscalarread
+                                          "},{"listdataread "})"
+listdataread ::= dataread listdataread |  dataread
+dataread ::= dataid "[" stencil "]" |  dataid
 \end{filecontents*}
@@
   \hspace{5mm}
-  \begin{minipage}[!h]{0.5\textwidth}
+  \begin{minipage}[!h]{0.98\textwidth}
     {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left]{grammar.txt}}   
@@
 \end{figure}
+
+\paragraph{\textbf{Example}} To give a better idea of a multi-stencil program, a short example is given in Figure~\ref{fig:exmsl}.
+In this example, the mesh is called \textit{cart}.
+Two groups of mesh entities are defined \textit{cell} and \textit{edgex}.
+As previously explained, it is not needed in the grammar to give details on the topology of the mesh and the group of mesh entities.
+Two computation domains are given, \textit{d1} is a subpart of the entities of the group \textit{cell}, and \textit{d2} is a subpart of the entities of the group \textit{edgex}.
+It is declared that \textit{d1} and \textit{d2} are independant.
+Three stencil shapes are given.
+The second one for example \textit{nce} returns mesh entities of the group \textit{edgex} from an entity of the group \textit{cell}.
+Eight quantities are applied onto the group \textit{cell}, and two onto the group \textit{edgex}.
+Two scalars are defined, \textit{mu} and \textit{tau}.
+The time loop will iterate 500 times.
+Finally, nine computations are declared.
+For example, the computation \textit{k8} write the quantity \textit{J} onto the computation domain \textit{d1} by reading the scalar \textit{mu} and the quantity \textit{I} with a stencil shape \textit{n1}.
 
@@
 stencil shapes : 
-  n1 from cell to cell
-  n2 from cell to edgex
-  n3 from edgex to cell
-data :
-  A,cell
-  B,cell
-  C,edgex
-  D,cell
-  E,cell
-  F,cell
-  G,cell
-  H,edgex
-  I,cell
-  J,cell
+  ncc from cell to cell
+  nce from cell to edgex
+  nec from edgex to cell
+quantities :
+  cell A,B,D,E,F,G,I,J
+  edgex C,H
 scalar : mu, tau
@@
   \hspace{5mm}
-  \begin{minipage}[!h]{0.4\textwidth}
-    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=right]{exmsl.txt}}   
+  \begin{minipage}[!h]{0.98\textwidth}
+    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left]{exmsl.txt}}   
     \caption{Example of program using the Multi-Stencil Language. \label{fig:exmsl}}
@@
 \end{figure}
+
+The aim of the Multi-Stencil Language is to offer a way to describe a numerical simulation based on explicit numerical schemes.
+The language is mesh-agnostic as no information is asked to the user about the actual topology of the mesh.
+Moreover, information that is usefull for the implementation is also split from the MSL description.
+For this reason, the language grammar is simple.
+
+In the next section will be shown how parallelization can be extracted from this simple language and how an empty parallel skeleton of the application can be generated, introducing implementation concerns in a second phase.

### 2016-05-20 ARTICLES/IJPP/article/formalism.tex
IJPP article modifications partie 4

@@
 \item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{local kernel} if $\forall (r,n) \in R$, $n = identity$.
-\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} if $w$ is a scalar, $D=\emptyset$, $R=\{(r,n)\}$ with $n=entity(r)$, and $comp$ is a reduction operation (usually $+$, $-$, $*$, $max$, $min$ etc.).
+\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} if $w$ is a scalar.%, $D=\emptyset$, $R=\{(r,n)\}$ with $n=entity(r)$, and $comp$ is a reduction operation (usually $+$, $-$, $*$, $max$, $min$ etc.).
 \end{itemize}

### 2016-05-20 ARTICLES/IJPP/article/parallelism.tex
IJPP article modifications partie 4

@@
 
-More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\mathcal{N},\Delta, \mathcal{S},\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
+\medskip
+More formally, the data parallelization of a multi-stencil program 
+\begin{equation*}
+\mathcal{MSP}(\mathcal{M},\Phi,\mathcal{D},\mathcal{N},\Delta, \mathcal{S},T,\Gamma)
+\end{equation*}
+
+consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
 This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
-As entities and as data are mapped onto the mesh, the set of entities and the set of data $\Delta$ are partitionned the same way than the mesh: $\mathcal{E}=\{\mathcal{E},\dots,\mathcal{E}{p-1}\}$, $\Delta=\{\Delta_0,\dots,\Delta_{p-1}\}$.
+
+As entities and quantities are mapped onto the mesh, the set of groups of mesh entities and the set of quantities $\Delta$ are partitionned the same way than the mesh: $\Phi=\{\Phi_0,\dots,\Phi_{p-1}\}$, $\Delta=\{\Delta_0,\dots,\Delta_{p-1}\}$.
+
 The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma_{sync}$.
@@
 For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $k_i$ and $k_j$, denoted $k_i \pprec k_j$, if $\exists (r_j,n_j) \in R_j$ such that $w_i=r_j$ and $n_j\neq identity$ ($k_j$ is a stencil computation).
-Moreover, the data to update is $\{w_i\}$.
+The quantity to synchronise is $\{w_i\}$.
 \end{mydef}
 
-Actually, a synchronization is needed by the data read by a stencil computation (not local), if this data has been modified before, which means that it has been written before.
+Actually, a synchronization is needed by the quantity read by a stencil computation (not local), if this quantity has been modified before, which means that it has been written before.
 This synchronization is needed because a neighborhood function $n \in \mathcal{N}$ of a stencil computation involves values computed on different resources.
+
+However, as a multi-stencil program is an iterative program, computations which happen after $k_j$ at the time iteration $t$ have also been computed before $k_j$ at the previous time iteration $t-1$.
+For this reason another case of synchronization has to be defined.
+
+\begin{mydef}
+For $n$ the number of computations in $\Gamma$ and $j<n$, if $\exists (r_j,n_j) \in R_j$ such that $n_j\neq identity$ and for all $i<j$, $k_i \not \pprec k_j$, a \textit{synchronisation} is needed between $k_l$ and $k_j$, where $j<l<n$, denoted $k_l \pprec k_j$, if $w_l=r_j$.
+The quantity to synchronise is $\{w_l\}$.
+\end{mydef}
 
@@
 \begin{equation*}
-k_{i,j}^{sync}(S,R,(w,D),comp)
+k_{i,j}^{sync}(S,R,(w,D),comp), 
 \end{equation*}
+where $S=\emptyset$, $R=\{(r,n)\}=\{(w_i,n_j \in \mathcal{N}\}$, $(w,D)=(w_i,\bigcup_{\phi \in D_j} n_j(\phi)))$ and $comp=identity$.
+In other words, $w_i$ has to be synchronized for the neighborhood $n_j$ for all entities of $D_j$.
 \end{mydef}
-where $S=\emptyset$, $R=\{(w_i,\bigcup_{e \in D_i} n_j(e))\}$, $w=(w_i,\bigcup_{e \in D_i} n_j(e)))$ and $comp=identity$.
-In other words, $w_i$ has to be synchronized for the neighborhood $n_j$ for all entities of $w_i$.
 
@@
 This results in $p$ (number of resources) scalar values.
-Thus, to perform the final reduction a synchronization is needed to get the $p$ scalars onto each resource.
-The final reduction step is applied such that each resource get the result scalar.
+Thus, to perform the final reduction, a set of synchronizations are needed to get the final reduction scalar.
+As most parallelism libraries (MPI, OpenMP) already propose a reduction synchronization with its own optimizations, we simply choose to replace the reduction computation by itself anotated by $red$.
 
 \begin{mydef}
-For a reduction kernel $k_j(S_j,R_j,(w_j,D_j),comp)$, if we denote by $w^r$, $0 \leq r<p$, the local scalar $w$ computed on the resource $r$, a reduction synchronization is defined as the specific computation
-\begin{equation*}
-k_{j}^{sync}(S,R,(w,D),comp)
-\end{equation*}
+A reduction kernel $k_j(S_j,R_j,(w_j,D_j),comp)$, where $w$ is a scalar, is replaced by $k^{red}_j(S_j,R_j,(w_j,D_j),comp)$.
+%if we denote by $w^r$, $0 \leq r<p$, the local scalar $w$ computed on the resource $r$, a reduction synchronization is defined as the specific computation
+% \begin{equation*}
+% k_{j}^{sync}(S,R,(w,D),comp)
+% \end{equation*}
 \end{mydef}
-where, $S=\emptyset$, $R=\{(w^0,entity(w^0)) \dots (w^{p-1},entity(w^{p-1}))\}$, and $w=w_i$, $D=entity(w)=D_i$ and $comp=comp_i$.
+% where, $S=\emptyset$, $R=\{(w^0,entity(w^0)) \dots (w^{p-1},entity(w^{p-1}))\}$, and $w=w_i$, $D=entity(w)=D_i$ and $comp=comp_i$.
 
@@
 \begin{mydef}
-From the ordered list of computation $\Gamma$, a new ordered list $\Gamma_{sync}$ is obtained from the call $\Gamma_{sync} = F_{sync}(\Gamma,0)$, where $F_{sync}$ is the recursive function defined in Algorithm~\ref{alg:sync}.
+From the ordered list of computation $\Gamma$, a new synchronized ordered list $\Gamma_{sync}$ is obtained from the call $\Gamma_{sync} = F_{sync}(\Gamma,0)$, where $F_{sync}$ is the recursive function defined in Algorithm~\ref{alg:sync}.
+\end{mydef}
 
-% \begin{equation*}
-% F_{sync}(\Gamma,j) = 
-% \begin{cases} 	\bullet [] \mbox{ if }j=|\Gamma|\\
-% 				\bullet [k_{sync}(\emptyset,\\
-% 				\{(w_i,\bigcup_{e \in entity(w_i)} n_j(e))\},\\
-% 				(w_i,\bigcup_{e \in entity(w_i)} n_j(e))),\\
-% 				identity,\emptyset), k_j] \cdot F_{sync}(\Gamma,j+1)\\
-% 				\text{ } \qquad \forall i<j \text{, and }k_i \pprec k_j\\
-% 				\bullet [k_j, k_{sync}(\emptyset,\\
-% 				\{(w^0_j,entity(w^0_j)) \dots (w^{p-1}_j,entity(w^{p-1}_j))\},\\
-% 				(w_j,entity(w_j)),comp_j)] \cdot F_{sync}(\Gamma,j+1)\\ 
-% 				\text{ } \qquad \mbox{if } w_j \in \mathcal{S} \text{, and } R_j=\{(r_j,entity(r_j))\}\\
-% 				\bullet [\Gamma[j]] \cdot F_{sync}(\Gamma,j+1) \text{ else}.
-% \end{cases}
-% \end{equation*}
-\end{mydef}
+Algorithm~\ref{alg:sync} follows previous definitions to include synchronizations into $\Gamma$.
 
@@
 \ElsIf {$\exists (r_j,n_j) \in R_j$ such that $n_j\neq identity$}
+\For {all $(r_j,n_j) \in R_j$ such that $n_j\neq identity$}
+\State found = false
 \For {$0 \leq i<j$}
@@
 \If {$k_i \pprec k_j$}
+\State found = true
 \State $S = \emptyset$
-\State $R = \{(w_i,\bigcup_{e \in D_i} n_j(e))\}$
-\State $(w,D) = (w_i,\bigcup_{e \in D_i} n_j(e)))$
+\State $R = \{(w_i,n_j)\}$
+\State $(w,D) = (w_i,\bigcup_{\phi \in D_j} n_j(\phi)))$
 \State $comp = identity$
-\State $list.[k_{i;j}^{sync}(S,R,(w,D),comp)$
+\State $list.[k_{i;j}^{sync}(S,R,(w,D),comp)]$
 \EndIf
 \EndFor
+\If {!found}
+\For {$j<i\leq n$}
+\State $k_i = \Gamma[i]$
+\If {$k_i \pprec k_j$}
+\State $S = \emptyset$
+\State $R = \{(w_i,n_j)\}$
+\State $(w,D) = (w_i,\bigcup_{\phi \in D_j} n_j(\phi)))$
+\State $comp = identity$
+\State $list.[k_{i;j}^{sync}(S,R,(w,D),comp)]$
+\EndIf
+\EndFor
+\EndIf
 \State $list \cdot [k_j]$
-\ElsIf {$w_j \in \mathcal{S} \text{, and } R_j=\{(r_j,entity(r_j))\}$}
-\State $S = \emptyset$
-\State $R = \{(w^0_j,entity(w^0_j)) \dots (w^{p-1}_j,entity(w^{p-1}_j))\}$
-\State $(w,D) = (w_j,entity(w_j))$
-\State $comp = comp_j$
-\State $list.[k_j, k_{j}^{sync}(S,R,(w,D),comp)]$
+\EndFor
+\ElsIf {$w_j \in \mathcal{S}$}
+\State $list.[k^{red}_j]$
 \Else
@@
 \begin{equation}
-\mathcal{MSP}_r(T,\mathcal{M}_r,\mathcal{E}_r,\mathcal{D}_r,\mathcal{N},\Delta_r,\mathcal{S},\Gamma_{sync}),
+\mathcal{MSP}_r(\mathcal{M}_r,\Phi_r,\mathcal{D}_r,\mathcal{N},\Delta_r,\mathcal{S},T,\Gamma_{sync}),
 \end{equation}
@@
 
-% \begin{mydef}
-% For two computations domains $D_i$ and $D_j$ both subsets of the same entities group $E$.
-We say that $D_i$ and $D_j$ are independent, denoted $D_i \oplus D_j$, if and only if $D_i \cap D_j = \emptyset$.
-% \end{mydef}
-
 \begin{mydef}
-For two computations $k_i(S_i,R_i,(w_i,D_i),comp_i)$ and $k_j(S_j,R_j,(w_j,D_j),comp_j)$, with $i < j$ and such that neither $k_i$ or $k_j$ are boundary kernels, it is said that $k_j$ is data read/write dependant from $k_i$, denoted $k_i \prec_{rw} k_j$, if $w_i \in R_j$.
+For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{read after write}, denoted $k_i \prec_{raw} k_j$, if $w_i \in R_j$.
 In this case, $k_i$ has to be computed before $k_j$.
-The binary relation $k_i \prec_{rw} k_j$ represents a \textit{dependency}.
+The binary relation $k_i \prec_{raw} k_j$ represents a \textit{dependency}.
 \end{mydef}
@@
 \begin{mydef}
-For two computations $k_i(S_i,R_i,(w_i,D_i),comp_i)$ and $k_j(S_j,R_j,(w_j,D_j),comp_j)$, with $i < j$, it is said that $k_j$ is data write dependant from $k_i$, denoted $k_i \prec_{w} k_j$, if $w_i = w_j$ and $D_i \cap D_j \neq \emptyset$.
+For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{write after write}, denoted $k_i \prec_{waw} k_j$, if $w_i = w_j$ and $D_i \cap D_j \neq \emptyset$.
 In this case, $k_i$ also has to be computed before $k_j$.
-The binary relation $k_i \prec_{w} k_j$ represents a \textit{dependency}.
+The binary relation $k_i \prec_{waw} k_j$ represents a \textit{dependency}.
 \end{mydef}
 
-Those definitions are close to traditionnal read/write dependencies.
-However, specific conditions due to the specific domain can be introduced, such as the condition on boundary kernels, and conditions on the computation domain.
+\begin{mydef}
+For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{write after read}, denoted $k_i \prec_{war} k_j$, if $w_j \in R_i$.
+In this case, $k_i$ also has to be computed before $k_j$ is started so that values read by $k_i$ are relevant.
+The binary relation $k_i \prec_{war} k_j$ represents a \textit{dependency}.
+\end{mydef}
+
+Those definitions are known as \emph{data hazards classification}.
+However, a specific condition on the computation domain, due to the specific domain of multi-stencils, is introduced for the write after write case.
 
 \begin{mydef}
-A directed acyclic graph (DAG) $G(V,A)$ is a graph where the edges are directed from a source to a destination vertex and where, following the direction of edges, no cycle can be found from a vertex $u$ to itself.
+A directed acyclic graph (DAG) $G(V,A)$ is a graph where the edges are directed from a source to a destination vertex, and where, by following edges direction, no cycle can be found from a vertex $u$ to itself.
 A directed edge is called an arc, and for two vertices $v,u \in V$ an arc from $u$ to $v$ is denoted $(\overset{\frown}{u,v}) \in A$.
@@
 
-From an ordered list of computations $\Gamma_{sync}$, a directed dependency graph $\Gamma_{dep}(V,A)$ can be built finding all pairs of computations $k_i(S_i,R_i,(w_i,D_i),comp_i)$ and $k_j(S_j,R_j,(w_j,D_j),comp_j)$, with $i<j$, such that $k_i \prec_{rw} k_j$ or $k_i \prec_w k_j$.
+From an ordered list of computations $\Gamma_{sync}$, a directed dependency graph $\Gamma_{dep}(V,A)$ can be built finding all pairs of computations $k_i$ and $k_j$, with $i<j$, such that $k_i \prec_{raw} k_j$ or $k_i \prec_{waw} k_j$ or $k_i \prec_{war} k_j$.
 
@@
 \begin{mydef}
-From the ordered list $\Gamma_{sync}$ of computation kernels, a directed dependency graph $\Gamma_{dep}(V,A)$ is obtained from the call $F_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function defined in Algorithm~\ref{alg:dep}.
+From the synchronized ordered list of computation kernels $\Gamma_{sync}$, the dependency graph of the computations $\Gamma_{dep}(V,A)$ is obtained from the call $F_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function defined in Algorithm~\ref{alg:dep}.
 
@@
 \State $k_i = \Gamma_{sync}[i]$
-\If {$k_i \prec_{rw} k_j$ or $k_i \prec_{w} k_j$}
+\If {$k_i \prec_{raw} k_j$ or $k_i \prec_{waw} k_j$ or $k_i \prec_{war} k_j$}
 \State $G = G \cup (k_j, \{(\overset{\frown}{k_i,k_j} \})$
@@
 
+One can notice that $\Gamma_{dep}$ is the dependency graph of the computations of a multi-stencil program, but it only takes into account a single time iteration.
+By representing more formally time iterations, a complete dependency graph of the simulation could be built.
+This is a possible extension of this work.
+
 \begin{myprop}
@@
 \begin{equation*}
-\mathcal{MSP}_r(T,\mathcal{M}_r,\mathcal{E}_r,\mathcal{D}_r,\mathcal{N},\Delta_r,\Gamma_{dep}).
+\mathcal{MSP}_r(\mathcal{M}_r,\Phi_r,\mathcal{D}_r,\mathcal{N},\Delta_r,T,\Gamma_{dep}).
 \end{equation*}

### 2016-05-20 ARTICLES/IJPP/article/article.tex
IJPP article modifications parties 4 5 et 6

@@
 %------------------------------------------------------------------------------
-\section{Static Scheduling}
+\section{Static Scheduling and performance model}
 \label{sect:msp}
 \input{msp.tex}
-% From the dependency graph we study a first dump: static scheduling
-% \begin{itemize}
-% \item description of the algorithm to get the binary tree decomposition of a minimal serie-parallel graph
-% \item use of the serie-parallel tree decomposition to detect fusions of kernels
-% \item example (the same one) from the dependency graph to the binary tree
-% \end{itemize}
 %------------------------------------------------------------------------------
-\section{MSL Compiler}
+\section{Implementation}
 \label{sect:comp}
 \input{compiler.tex}
-% \begin{itemize}
-% \item Overall algrotihm~\ref{alg:compiler}
-% \item Big example from the language to the static sheduling/optimization~\ref{fig:example}
-% \end{itemize}
 %------------------------------------------------------------------------------
@@
 \begin{itemize}
-\item short description of the implementation (compiler in Python, DDS = SkelGIS, Components)
 \item evaluation of the compiler (execution times of the different steps)
 \item evaluation of the data parallelism (weak/strong scaling)
+\item evaluation of fusions
+\item evaluation of the hybrid performance
+\item validation of the performance model
+\end{itemize}
 
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/400/median_weak.pdf}}
-  \caption{weak-scaling 400x400 blocks.}
-  \label{fig:mesh}
-\end{center}\end{figure}
+% \begin{figure}[!h]\begin{center}
+%   \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/400/median_weak.pdf}}
+%   \caption{weak-scaling 400x400 blocks.}
+%   \label{fig:mesh}
+% \end{center}\end{figure}
 
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/800/median_weak.pdf}}
-  \caption{weak-scaling 800x800 blocks}
-  \label{fig:mesh}
-\end{center}\end{figure}
+% \begin{figure}[!h]\begin{center}
+%   \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/800/median_weak.pdf}}
+%   \caption{weak-scaling 800x800 blocks}
+%   \label{fig:mesh}
+% \end{center}\end{figure}
 
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/strong_scaling/10K_1K/median_strong.pdf}}
-  \caption{strong-scaling 10kx10k, 1k iterations}
-  \label{fig:mesh}
-\end{center}\end{figure}
+% \begin{figure}[!h]\begin{center}
+%   \resizebox{8cm}{!}{\includegraphics{../results/strong_scaling/10K_1K/median_strong.pdf}}
+%   \caption{strong-scaling 10kx10k, 1k iterations}
+%   \label{fig:mesh}
+% \end{center}\end{figure}
 
-\item evaluation of the fusion
+% \item evaluation of the fusion
 
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/fusVSbase.pdf}}
-  \caption{strong scaling 500x500, 200 iterations, with and without fusion}
-  \label{fig:mesh}
-\end{center}\end{figure}
+% \begin{figure}[!h]\begin{center}
+%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/fusVSbase.pdf}}
+%   \caption{strong scaling 500x500, 200 iterations, with and without fusion}
+%   \label{fig:mesh}
+% \end{center}\end{figure}
 
-\item evaluation of the hybrid parallelism
+% \item evaluation of the hybrid parallelism
 
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_close_median.pdf}}
-  \caption{MPI only vs MPI+threads with \emph{close} scheduling policy (OpenMP)}
-  \label{fig:mesh}
-\end{center}\end{figure}
+% \begin{figure}[!h]\begin{center}
+%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_close_median.pdf}}
+%   \caption{MPI only vs MPI+threads with \emph{close} scheduling policy (OpenMP)}
+%   \label{fig:mesh}
+% \end{center}\end{figure}
 
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_spread_median.pdf}}
-  \caption{MPI only vs MPI+threads with \emph{spread} scheduling policy (OpenMP)}
-  \label{fig:mesh}
-\end{center}\end{figure}
+% \begin{figure}[!h]\begin{center}
+%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_spread_median.pdf}}
+%   \caption{MPI only vs MPI+threads with \emph{spread} scheduling policy (OpenMP)}
+%   \label{fig:mesh}
+% \end{center}\end{figure}
 
-\item analytic model for the results
+% \item analytic model for the results
 
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/times.pdf}}
-  \caption{Computation vs communication times in the MPI only application}
-  \label{fig:mesh}
-\end{center}\end{figure}
+% \begin{figure}[!h]\begin{center}
+%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/times.pdf}}
+%   \caption{Computation vs communication times in the MPI only application}
+%   \label{fig:mesh}
+% \end{center}\end{figure}
 
-\end{itemize}
+% \end{itemize}
 %------------------------------------------------------------------------------
@@
 \begin{itemize}
-\item Cartesian meshes : Pochoir, PATUS for optimizations/parallelization of a single stencil kernel
-\item Unstructured meshes : Lizst for optimizations/parallelization of a single stencil kernel
-\item Halide for pipeline of stencil computations in image (Cartesian) = mesh dependent
-\item SYCL same Halide mais mesh dependent
+\item stencil compilers cartesian : Pochoir, PATUS for optimizations/parallelization of a single stencil kernel
+\item stencil compilers unstructured meshes : Lizst for optimizations/parallelization of a single stencil kernel
+\item Halide and Nabla for pipeline of stencil computations in image (Cartesian) = mesh dependent
 \item Reuse of external solutions as SkelGIS, GA, Lizst, Pochoir etc.

### 2016-05-20 ARTICLES/IJPP/article/compiler.tex
IJPP article modifications parties 4 5 et 6

--- 2016-01-27 ARTICLES/IJPP/article/compiler.tex
@@
 The overall compiler algorithm is represented in Algorithm~\ref{alg:compiler}.
-First, the parsing of the input file (written with the MSL language) gave enough information to build $\Gamma$, the ordered list of computations.
+First, the parsing of the input file (written with the MSL language) gave enough information to build $\Gamma$, the ordered list of computation kernels.
 Second, $\Gamma_{sync}$ and $\Gamma_{dep}$ can be built from $\Gamma$ using the recursive functions $F_{sync}$ and $F_{dep}$.
 Third, N-Shapes are removed from $\Gamma_{dep}$ to get a MSPD graph, and to be able to create its serie-parallel binary tree decomposition $TSP$ (line 6).
-Lines 7 to 9 are explained in the next section.
-Finally a dump of $TSP$ is performed by the compiler and will be described in Section~\ref{sect:eval}.
+If only a data parallelization has been chosen, Line 8 performs the fusion of kernels.
+Finally a dump of $TSP$ is performed by the compiler.
+Implementation choices of the dump are described in the next subsection.
 
@@
 \Procedure{compileMSL} {file}
-\\$\Gamma$ = parser(file)
-\\$\Gamma_{sync} = F_{sync}(\Gamma,0)$
-\\$\Gamma_{dep} = F_{dep}(\Gamma_{sync},0)$
-\\removeNSH($\Gamma_{dep}$)
-\\$TSP = BDT(L^{-1}(\Gamma_{dep}))$
-\\$TSP = F_{tsp}(TSP,root(TSP))$
+\State $\Gamma$ = parser(file)
+\State $\Gamma_{sync} = F_{sync}(\Gamma,0)$
+\State $\Gamma_{dep} = F_{dep}(\Gamma_{sync},0)$
+\State removeNSH($\Gamma_{dep}$)
+\State $TSP = F_{tsp}(TSP,root(TSP))$
 \If {data parallel}
-\\$TSP = F_{tsp}(TSP)$
+\State $TSP = F_{fus}(TSP)$
 \EndIf
-\\dump(TSP)
+\State dump(TSP)
 \EndProcedure
@@
 %--------------------
-\subsection{Canonical form and fusions}
+\subsection{Implementation choices}
 
-The serie-parallel binary tree decomposition $TSP$ obtained from $BDT(L^{-1}(\Gamma_{dep}))$ can be reduced to its canonical form which consists in recursively merging successive $S$ vertices or successive $P$ vertices.
-
-\medskip
-\noindent \textbf{Definitions} For a tree $T(V,E)$
-%\begin{mydef}
-\begin{itemize}
-\item we denote $root(T)$, the function which returns the root vertex of the tree $T$;
-\item we denote $type(v)$ with $v \in V$ the function which returns the type of a vertex $v$. This type is either $S$, $P$, $comp$ or $sync$;
-\item We denote $replace(E,v_1,v_2)$, the function which replaces for all edges of the set $E$, $v_1$ by $v_2$.
-\end{itemize}
-
-\begin{mydef}
-We denote as a sequence or parallel sub-tree $T(V,E)$ sub-trees for which the root vertex respectively verifies $type(root(T))=S$ or $type(root(T))=P$.
-\end{mydef}
-
-\begin{myth}
-\begin{itemize}
-\item A sequence of sequence is a sequence if the order is kept. Thus for two sequence sub-trees $T_i$ and $T_j$, if $T_j \in V_i$, $T_i$ can be replaced by 
-\begin{equation*}
-T_i'(V_i \setminus root(T_j),E_i \cup replace(E_j,root(T_j),root(T_i)))
-\end{equation*}
-\item A parallel sub-tree of parallel sub-tree is a parallel sub-tree in any order. Thus for two parallel sub-trees $T_i$ and $T_j$, if $T_j \in V_i$, $T_i$ can be replaced by 
-\begin{equation*}
-T_i'(V_i \setminus root(T_j),E_i \cup replace(E_j,root(T_j),root(T_i)))
-\end{equation*}
-\end{itemize}
-\end{myth}
-
-These reductions applied recursively on the serie-parallel binary tree decomposition of $\Gamma_{dep}$ produce the canonical serie-parallel tree decomposition 
-\begin{equation*}
-TSP = F_{tsp}(TSP,root(TSP),successors(root(TSP)).
-\end{equation*}
-The recursive function $F_{tsp}$ is based on a Depth First Search algorithm and defined by Algorithm~\ref{alg:tsp}, where $E_{rooted(s)}$ represents the set of edges of the sub-tree rooted by the vertex $s$.
-
-\begin{algorithm}
-\caption{$F_{tsp}$}
-\label{alg:tsp}
-\begin{algorithmic}[1]
-\Procedure{$F_{tsp}$} {$TSP$,root,succ}
-\For {$s \in succ$}
-\If {type($s$)=$S$ or type($s$)=$P$}
-\\$F_{tsp}(TSP,s,successors(s))$
-\If {type($s$)=type(root)}
-\\$TSP = (V_{tsp} \setminus s,E_{tsp} \cup replace(E_{rooted(s)},s,root))$
-\EndIf
-\EndIf
-\EndFor
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization.
-Even though the MSL language is not dedicated to produce very optimized stencil codes, but to produces the parallel pattern of the application, building the $TSP$ tree make available an easy optimization when the data parallelization technique is the only one used.
-This optimization consists in proposing a valid merge of some computation kernels inside a single space loop.
-As a result, the user can use this valid fusion of kernels or not when implementing those.
-The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and where $k_{i;j}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.
-
-\begin{algorithm}
-\caption{$F_{fus}$}
-\label{alg:fus}
-\begin{algorithmic}[1]
-\Procedure{$F_{fus}$} {$TSP(V,E)$}
-\For {$(k_i,k_j) \in V^2$}
-\If {$type(k_i)=type(k_j)=comp$}
-\If {parent($k_i$)==parent($k_j$)}
-\If {parent($k_i$)==$S$}
-\If {$D_i==D_j$}
-\\propose the fusion $k_{i;j}$
-\EndIf
-\EndIf
-\If {parent($k_i$)==$P$}
-\If {$D_i==D_j$ and $R_i \cap R_j \neq \emptyset$}
-\\propose the fusion $k_{i;j}$
-\EndIf
-\EndIf
-\EndIf
-\EndIf
-\EndFor
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs for example~\cite{spaaTangCKLL11}.
-However, this fusion takes place at a different level and can bring performance improvments as it will be illustrated in Section~\ref{sect:eval}.
-This fusion algorithm relies on very simple statements:
-\begin{itemize}
-\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependant.
-As a result, what is written by the first one is read by the second one.
-Thus, at least one data is common to those computations (the one written by $k_i$).
-Thus, if the computation domains verify $D_i=D_j$, the fusion of $k_i$ and $k_j$ will decrease cache misses.
-\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependant.
-However, if the computation domains verify $D_i=D_j$, and if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}$.
-\end{itemize}
-
-\todo{solve order problem in $\Gamma_{dep}$ and $TSP$}
-
-%--------------------
-\subsection{Overall example}
-
-% Using those additional reduction rules, the binary series-parallel tree decomposition can be computed in linear time. 
-A complete example of reduction is given in Figure~\ref{fig:example}.
-The initial $\Gamma_{hybrid}$ DAG and its inverse line graph are given first, followed by the set of reductions.
-The label of the final single edge corresponds to the binary tree decomposition of the initial $\Gamma_{hybrid}$ graph.
-
-%====BIG EXAMPLE
-\begin{figure}[ht!]
-\begin{center}
-% INITIAL GRAPH
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (c0) at (0,0) {$c_0$};
-   \node[] (star1) at (1,0) {$*_1$};
-   \node[] (c1) at (2,0) {$c_1^*$};
-   \node[] (c2) at (3,0.5) {$c_2$};
-   \node[] (star4) at (3,1.5) {$*_4$};
-   \node[] (c3) at (3,-0.5) {$c_3$};
-   \node[] (c4) at (4,0.5) {$c_4^*$};
-   \node[] (c5) at (4,-0.5) {$c_5$};
-   \node[] (c6) at (5,0.5) {$c_6$};
-   \node[] (c7) at (6,0) {$c_7$};
-   \node[] (star8) at (7,0) {$*_8$};
-   \node[] (c8) at (8,0) {$c_8^*$};
-
-  \path[->]
-    (c0) edge node {} (star1)
-    (star1) edge node {} (c1)
-    (c1) edge node {} (c2)
-         edge node {} (c3)
-         edge node {} (star4)
-    (star4) edge node {} (c4)
-    (c2) edge node {} (c4)
-    (c4) edge node {} (c6)
-    (c3) edge node {} (c5)
-    (c5) edge node {} (c7)
-    (c6) edge node {} (c7)
-    (c7) edge node {} (star8)
-    (star8) edge node {} (c8);
-  \end{tikzpicture}
-}
-\\
-% INVERSE LINE GRAPH
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c0d) at (1,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (star1d) at (2,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c7d) at (7,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (star8d) at (8,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
-
-  \path[->]
-    (c0s) edge node [above] {$c_0$} (c0d)
-    (c0d) edge node [above] {$*_1$} (star1d)
-    (star1d) edge node [above] {$c_1^*$} (c1d)
-    (c1d) edge node [above] {$c_2$} (c2d)
-          edge node [above] {$c_3$} (c3d)
-          edge [bend left=80] node [above] {$*_4$} (c2d)
-    (c2d) edge node [above] {$c_4^*$} (c4d)
-    (c4d) edge node [above] {$c_6$} (c6d)
-    (c6d) edge node [above] {$c_7$} (c7d)
-    (c7d) edge node [above] {$c_8$} (star8d)
-    (star8d) edge node [above] {$c_8$} (c8d)
-    (c3d) edge node [above] {$c_5$} (c6d);
-  \end{tikzpicture}
-}
-\\
-% TRANSFORMATIONS
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
-
-   %reduction c0 c1
-   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
-   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,2.5) {$c_0$};
-   \node[] (star1) at (1.5,2.5) {$*_1$};
-   \node[] (c1) at (2,1.5) {$c_1^*$};
-
-   %reduction c7 c8
-   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
-   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
-   \node[] (c7) at (6.5,2.5) {$c_7$};
-   \node[] (star8) at (7.5,2.5) {$*_8$};
-   \node[] (c8) at (8,1.5) {$c_8$};
-
-  \path[->]
-    (c0s) edge node {} (c1d)
-    (c1d) edge node [above] {$c_2$} (c2d)
-          edge node [above] {$c_3$} (c3d)
-          edge [bend left=80] node [above] {$*_4$} (c2d)
-    (c2d) edge node [above] {$c_4^*$} (c4d)
-    (c4d) edge node [above] {$c_6$} (c6d)
-    (c6d) edge node [above] {} (c8d)
-    (c3d) edge node [above] {$c_5$} (c6d)
-
-    %reduction c0 c1
-    (s0) edge node {} (s1)
-         edge node {} (c1)
-    (s1) edge node {} (c0)
-         edge node {} (star1)
-    %reduction c7 c8
-    (s2) edge node {} (s3)
-         edge node {} (c8)
-    (s3) edge node {} (c7)
-        edge node {} (star8);
-  \end{tikzpicture}
-}
-\\
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
-
-   %reduction c0 c1
-   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
-   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,2.5) {$c_0$};
-   \node[] (star1) at (1.5,2.5) {$*_1$};
-   \node[] (c1) at (2,1.5) {$c_1^*$};
-
-   %reduction c7 c8
-   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
-   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
-   \node[] (c7) at (6.5,2.5) {$c_7$};
-   \node[] (star8) at (7.5,2.5) {$*_8$};
-   \node[] (c8) at (8,1.5) {$c_8$};
-
-   %reduction c3 c5
-   \node[] (s5) at (4.5,-2) {$\mathcal{S}$};
-   \node[] (c3) at (4,-1) {$c_3$};
-   \node[] (c5) at (5,-1) {$c_5$};
-
-   %reduction c2 *4
-   \node[] (p0) at (3.5,1) {$\mathcal{P}$};
-   \node[] (c2) at (3,2) {$c_2$};
-   \node[] (star4) at (4,2) {$*_4$};
-
-  \path[->]
-    (c0s) edge node {} (c1d)
-    (c1d) edge node [above] {} (c2d)
-    (c2d) edge node [above] {$c_4^*$} (c4d)
-    (c4d) edge node [above] {$c_6$} (c6d)
-    (c6d) edge node [above] {} (c8d)
-    (c1d) edge [bend right] node [above] {} (c6d)
-
-    %reduction c0 c1
-    (s0) edge node {} (s1)
-         edge node {} (c1)
-    (s1) edge node {} (c0)
-         edge node {} (star1)
-    %reduction c7 c8
-    (s2) edge node {} (s3)
-         edge node {} (c8)
-    (s3) edge node {} (c7)
-        edge node {} (star8)
-
-    %reduction c3 c5
-    (s5) edge node {} (c3)
-         edge node {} (c5)
-
-    %reduction c2 *4
-    (p0) edge node {} (c2)
-         edge node {} (star4);
-  \end{tikzpicture}
-}
-\end{center}
-\caption{Big example}
-\label{fig:example}
-\end{figure}
-
-\begin{figure}[ht!]
-\ContinuedFloat
-\begin{center}
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
-
-   %reduction c0 c1
-   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
-   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,2.5) {$c_0$};
-   \node[] (star1) at (1.5,2.5) {$*_1$};
-   \node[] (c1) at (2,1.5) {$c_1^*$};
-
-   %reduction c7 c8
-   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
-   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
-   \node[] (c7) at (6.5,2.5) {$c_7$};
-   \node[] (star8) at (7.5,2.5) {$*_8$};
-   \node[] (c8) at (8,1.5) {$c_8$};
-
-   %reduction c3 c5
-   \node[] (s5) at (4.5,-2) {$\mathcal{S}$};
-   \node[] (c3) at (4,-1) {$c_3$};
-   \node[] (c5) at (5,-1) {$c_5$};
-
-   %reduction c2 c6
-   \node[] (s6) at (4.5,1) {$\mathcal{S}$};
-   \node[] (p0) at (3.5,2) {$\mathcal{P}$};
-   \node[] (c2) at (3,3) {$c_2$};
-   \node[] (star4) at (4,3) {$*_4$};
-   \node[] (s7) at (5.5,2) {$\mathcal{S}$};
-   \node[] (c4) at (5,3) {$c_4^*$};
-   \node[] (c6) at (6,3) {$c_6$};
-
-  \path[->]
-    (c0s) edge node {} (c1d)
-    (c1d) edge [bend left] node [above] {} (c6d)
-    (c6d) edge node [above] {} (c8d)
-    (c1d) edge [bend right] node [above] {} (c6d)
-
-    %reduction c0 c1
-    (s0) edge node {} (s1)
-         edge node {} (c1)
-    (s1) edge node {} (c0)
-         edge node {} (star1)
-    %reduction c7 c8
-    (s2) edge node {} (s3)
-         edge node {} (c8)
-    (s3) edge node {} (c7)
-        edge node {} (star8)
-
-    %reduction c3 c5
-    (s5) edge node {} (c3)
-         edge node {} (c5)
-
-    %reduction c2 c6
-    (s6) edge node {} (p0)
-         edge node {} (s7)
-    %reduction c2 *4
-    (p0) edge node {} (c2)
-         edge node {} (star4)
-    (s7) edge node {} (c4)
-         edge node {} (c6);
-  \end{tikzpicture}
-}
-\\
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (7,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
-
-   %reduction c0 c1
-   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
-   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,2.5) {$c_0$};
-   \node[] (star1) at (1.5,2.5) {$*_1$};
-   \node[] (c1) at (2,1.5) {$c_1^*$};
-
-   %reduction c7 c8
-   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
-   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
-   \node[] (c7) at (6.5,2.5) {$c_7$};
-   \node[] (star8) at (7.5,2.5) {$*_8$};
-   \node[] (c8) at (8,1.5) {$c_8$};
-
-   \node[] (p1) at (4.5,0.5) {$\mathcal{P}$};
-   %reduction c3 c5
-   \node[] (s5) at (5.5,1.5) {$\mathcal{S}$};
-   \node[] (c3) at (5,2.5) {$c_3$};
-   \node[] (c5) at (6,2.5) {$c_5$};
-   %reduction c2 c6
-   \node[] (s6) at (3.5,1.5) {$\mathcal{S}$};
-   \node[] (p0) at (2.8,2.5) {$\mathcal{P}$};
-   \node[] (c2) at (2.3,3.5) {$c_2$};
-   \node[] (star4) at (3.3,3.5) {$*_4$};
-   \node[] (s7) at (4.2,2.5) {$\mathcal{S}$};
-   \node[] (c4) at (3.8,3.5) {$c_4^*$};
-   \node[] (c6) at (4.8,3.5) {$c_6$};
-
-  \path[->]
-    (c0s) edge node {} (c1d)
-    (c1d) edge node [above] {} (c6d)
-    (c6d) edge node [above] {} (c8d)
-
-    %reduction c0 c1
-    (s0) edge node {} (s1)
-         edge node {} (c1)
-    (s1) edge node {} (c0)
-         edge node {} (star1)
-    %reduction c7 c8
-    (s2) edge node {} (s3)
-         edge node {} (c8)
-    (s3) edge node {} (c7)
-        edge node {} (star8)
-
-    (p1) edge node {} (s6)
-         edge node {} (s5)
-    %reduction c3 c5
-    (s5) edge node {} (c3)
-         edge node {} (c5)
-    %reduction c2 c6
-    (s6) edge node {} (p0)
-         edge node {} (s7)
-    %reduction c2 *4
-    (p0) edge node {} (c2)
-         edge node {} (star4)
-    (s7) edge node {} (c4)
-         edge node {} (c6);
-  \end{tikzpicture}
-}
-\\
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
-
-   \node[] (s8) at (4.5,0.5) {$\mathcal{S}$};
-   \node[] (s9) at (3,1.5) {$\mathcal{S}$};
-
-   %reduction c0 c1
-   \node[] (s0) at (1.5,2.5) {$\mathcal{S}$};
-   \node[] (s1) at (1,3.5) {$\mathcal{S}$};
-   \node[] (c0) at (0.5,4.5) {$c_0$};
-   \node[] (star1) at (1.5,4.5) {$*_1$};
-   \node[] (c1) at (2,3.5) {$c_1^*$};
-
-   %reduction c7 c8
-   \node[] (s2) at (7.5,1.5) {$\mathcal{S}$};
-   \node[] (s3) at (7,2.5) {$\mathcal{S}$};
-   \node[] (c7) at (6.5,3.5) {$c_7$};
-   \node[] (star8) at (7.5,3.5) {$*_8$};
-   \node[] (c8) at (8,2.5) {$c_8$};
-
-   \node[] (p1) at (4.5,2.5) {$\mathcal{P}$};
-   %reduction c3 c5
-   \node[] (s5) at (5.5,3.5) {$\mathcal{S}$};
-   \node[] (c3) at (5,4.5) {$c_3$};
-   \node[] (c5) at (6,4.5) {$c_5$};
-   %reduction c2 c6
-   \node[] (s6) at (3.5,3.5) {$\mathcal{S}$};
-   \node[] (p0) at (2.8,4.5) {$\mathcal{P}$};
-   \node[] (c2) at (2.3,5.5) {$c_2$};
-   \node[] (star4) at (3.3,5.5) {$*_4$};
-   \node[] (s7) at (4.2,4.5) {$\mathcal{S}$};
-   \node[] (c4) at (3.8,5.5) {$c_4^*$};
-   \node[] (c6) at (4.8,5.5) {$c_6$};
-
-  \path[->]
-    (c0s) edge node {} (c8d)
-
-    (s8) edge node {} (s9)
-         edge node {} (s2)
-    (s9) edge node {} (p1)
-         edge node {} (s0)
-
-    %reduction c0 c1
-    (s0) edge node {} (s1)
-         edge node {} (c1)
-    (s1) edge node {} (c0)
-         edge node {} (star1)
-    %reduction c7 c8
-    (s2) edge node {} (s3)
-         edge node {} (c8)
-    (s3) edge node {} (c7)
-        edge node {} (star8)
-
-    (p1) edge node {} (s6)
-         edge node {} (s5)
-    %reduction c3 c5
-    (s5) edge node {} (c3)
-         edge node {} (c5)
-    %reduction c2 c6
-    (s6) edge node {} (p0)
-         edge node {} (s7)
-    %reduction c2 *4
-    (p0) edge node {} (c2)
-         edge node {} (star4)
-    (s7) edge node {} (c4)
-         edge node {} (c6);
-  \end{tikzpicture}
-}
-\end{center}
-\caption{Big example}
-\label{fig:example}
-\end{figure}
-
-% At this point, the $\mathcal{MSP}$ program can be dumped to any parallel language in which a sequence and a parallel execution are available.
-For example, we could imagine a parallel functionnal language with a binary \emph{sequence} function and a binary \emph{parallel} function.
+SkelGIS + OpenMP + L2C

### 2016-05-20 ARTICLES/IJPP/article/msp.tex
IJPP article modifications parties 4 5 et 6

--- 2016-01-27 ARTICLES/IJPP/article/msp.tex
@@
-In this section we argue that the graphs $\Gamma_{dep}$, previously defined, can be transformed to a minimal series-parallel directed acyclic graph.
-As a result, this graph can be represented as a binary tree of parallel and series compositions of sub-graphs, also called \emph{binary decomposition tree}~\cite{Valdes:1979:RSP:800135.804393}.
-The obtained binary tree is a valid static scheduling of $\Gamma_{dep}$ and can also be used to optimize the code including data parallelism only.
+In this section we detail a static scheduling of $\Gamma_{dep}$ by using minimal series-parallel directed acyclic graphs.
+Such a static scheduling may not be the most efficient one, but it offers a simple fork/join task model which make possible the design of a performance model.
+Moreover, such a scheduling offers a simple way to propose a fusion optimization.
 
 %--------------------
-\subsection{Minimal series-parallel DAG}
-\begin{mydef}
-A vertex $v$ of a DAG $G$ is a \emph{source} if no edge of $G$ enters $v$. Similarly, a vertex $v$ is a \emph{sink} if no edge of $G$ leaves $v$. 
-\end{mydef}
+\subsection{Series-Parallel graph}
 
-In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of minimal series-parallel DAGs (MSPD).
+In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of Minimal Series-Parallel DAGs (MSPD).
+Such a graph can be decomposed as a serie-parallel tree, denoted $TSP$, where each leaf is a vertex of the MSPD DAG it represents, and whose internal nodes are labelled $S$ or $P$ to indicate respectively the series or parallel composition of sub-trees.
+Such a tree can be considered as a fork-join model and as a static scheduling.
 
-\begin{mydef}Minimal Series Parallel DAG
-\begin{itemize}
-\item The DAG having a single vertex and no edges is MSPD.
-\item If $G_1=(V_1,E_1)$ and $G_2=(V_2,E_2)$ are two MSPD DAGs, so is either of the DAGs constructed by the following operations:
-\begin{itemize}
-\item Parallel composition: $G_p=(V_1\cup V_2,E_1\cup E_2)$.
-\item Series composition: $G_s=(V_1\cup V_2,E_1\cup E_2\cup (N_1 \times R_2))$, where $N_1$ is the set of sinks of $G_1$ and $R_2$ is the set of sources of $G_2$.
-\end{itemize}
-\end{itemize}
-\end{mydef}
-
-Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or subgraph, called $N$ and represented in Figure~\ref{fig:over}. A possible definition of a N-Shape is
-\begin{mydef}
-For two vertices $v_1$ and $v_2$ of a DAG which have a Lowest Common Ancestor $v_{lca}$.
-For all $v_p$ which is either a parent vertex of $v_1$ or $v_2$ and which is different from $v_{lca}$, a N-Shape is formed by the induced subgraph containing vertices $v_1$, $v_2$, $v_{lca}$ and $v_p$.
-\end{mydef}
+Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or subgraph, called $N$, such that the following property is verified :
 
@@
 
-\begin{proof} \cite{Valdes:1979:RSP:800135.804393} \end{proof}
-
-\begin{myprop}
-The transitive reduction of $\Gamma_{dep}$ is not MSPD.
-\end{myprop}
-
-\begin{proof}
-Actually, it is possible in a multi-stencil program to have a set of computations such that their dependencies form the forbidden $N$ subgraph.
-Considering four computation kernels $k_0$, $k_1$, $k_2$ and $k_3$ of a $\mathcal{MSP}$ such that $k_0 \prec k_1$, $k_2 \prec k_1$ and $k_2 \prec k_3$ are verified.
-Thus, the \emph{zigzag} relation $k_0 \prec k_1 \succ k_2 \prec k_3$ which form a forbidden subgraph of a MSPD is found in $\Gamma_{dep}$.
-\end{proof}
-
-To remove the forbidden N-shapes from $\Gamma_{dep}=(V,E)$, we have chosen to apply an over-constraint with the relation $k_0 \prec k_3$, such that a complete bipartite graph is created for the sub-dag, and can be translated to a series-parallel decomposition, as illustrated in Figure~\ref{fig:allover}.
+To remove the forbidden N-shapes from the transitive reduction of $\Gamma_{dep}=(V,E)$, we have chosen to apply an over-constraint with the relation $k_0 \prec k_3$, such that a complete bipartite graph is created for the sub-dag, and can be translated to a series-parallel decomposition, as illustrated in Figure~\ref{fig:allover}.
 
@@
 \begin{center}
-\subfloat[][Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
   \end{tikzpicture}
-}
-\hspace{50pt}
-\subfloat[][Series-parallel tree associated to the over-constraint\label{fig:treeover}]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (S0) at (0,0) {$\mathcal{S}$};
-   \node[] (bS0) at (-1,-1) {};
-   \node[] (P1) at (-1,1) {$\mathcal{P}$};
-   \node[] (aP1) at (-1.5,2) {};
-   \node[] (aaP1) at (-0.5,2) {};
-   \node[] (P2) at (1,1) {$\mathcal{P}$};
-   \node[] (c0) at (-2,3) {$k_0$};
-   \node[] (c1) at (0.5,3) {$k_1$};
-   \node[] (c2) at (0,3) {$k_2$};
-   \node[] (c3) at (2,3) {$k_3$};
-
-  \path[->]
-    (bS0) edge [dotted] node {} (S0)
-    (S0) edge node {} (P1)
-         edge node {} (P2)
-    (P1) edge [dotted] node [swap] {} (aP1)
-          edge [dotted] node [swap] {} (aaP1)
-    (aP1) edge node {} (c0)
-    (aaP1) edge node {} (c2)
-    (P2) edge node {} (c1)
-        edge node {} (c3);
-  \end{tikzpicture}
-}
-\caption{Deletion of forbidden subgraphs.}
-\label{fig:allover}
+\caption{Over-constraint on the forbidden $N$ shape.}
+\label{fig:over}
 \end{center}
@@
 
-The algorithm is inspired from the one proposed by~\cite{Mitchell:2004:CMV:1082101.1082117}, and is described in Algorithm~\ref{alg:nshape}.
-We precisely describe how we find edges to add to get rid of N-Shapes.
-One can notice that another solution would have been to remove the extra-edge from the N-Shape instead of adding one~\cite{}.
-However, we have chosen to exactly keep dependencies computed in $\Gamma_{dep}$, to not approximate the static scheduling.
-Finally, algorithm proposed by~\cite{Bender200575} to find a Lowest Common Ancestor in a DAG (in $\mathcal{O}(|V|^3)$, could have been also used.
+After these over-constraints are applied, $\Gamma_{dep}$ is MSPD.
+Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have proposed a linear algorithm to know if a DAG is MSPD and, if it is, to decompose it to its associated binary decomposition tree.
+As a result, the binary tree decomposition algorithm of Valdes \& Al can be applied on $\Gamma_{dep}$ to get the $TSP$ static scheduling of the multi-stencil program.
 
-The first step of Algorithm~\ref{alg:nshape} is to create a new DAG $\Gamma_{dep}'$ from $\Gamma_{dep}$, such that vertices having exactly one source and one destination in $\Gamma_{dep}$ is not added to $\Gamma_{dep}'$, and its source and destination vertices are directly linked by an edge.
-This step is called $RTV(\Gamma_{dep})$ (line 2) and is done in $\mathcal{O}(|V|)$.
-The second step is to get root vertices in $\Gamma_{dep}'$.
-The third step calls $BFSnshape$ (line 4).
-This procedure is based on a Breadth First Search recursive algorithm (in $\mathcal{O}(|V|+|E|)$) and takes as input a list of source vertices $sources$, $\Gamma_{dep}$ and $\Gamma_{dep}'$.
-At each recursive call the induced sub-graph of $\Gamma_{dep}'$ containing $sources$ vertices and the list of their successors $successors(sources)$ is created and called $SG$.
-If this sub-graph $SG$ is connected and is not a complete bipartite graph, the algorithm make the subgraph $SG$ complete by adding needed edges into $\Gamma_{dep}'$ (line 12 in $\mathcal{O}(|V|^2$).
-The algorithm also add same edges to $\Gamma_{dep}$ (line 13).
-Finally, line 5, the transitive closure of $\Gamma_{dep}$ is computed.
+\paragraph{\textbf{Example}} The Serie-Parallel tree decomposition of the example given in Figure~\ref{fig:exmsl}, which is built from the dependency graph of Figure~\ref{fig:depdep} is given in Figure~\ref{fig:tree}.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   %\node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
+   %\node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
+
+   \node[] (s8) at (4.5,0.5) {$\mathcal{S}$};
+   \node[] (s9) at (3,1.5) {$\mathcal{S}$};
+
+   %reduction c0 c1
+   \node[] (s0) at (1.5,2.5) {$\mathcal{S}$};
+   \node[] (s1) at (1,3.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,4.5) {$k_0$};
+   \node[] (star1) at (1.5,4.5) {$k_{0;1}^{sync}$};
+   \node[] (c1) at (2,3.5) {$k_1$};
+
+   %reduction c7 c8
+   \node[] (s2) at (7.5,1.5) {$\mathcal{S}$};
+   \node[] (s3) at (7,2.5) {$\mathcal{S}$};
+   \node[] (c7) at (6.5,3.5) {$k_7$};
+   \node[] (star8) at (7.5,3.5) {$k_{7;8}^{sync}$};
+   \node[] (c8) at (8,2.5) {$k_8$};
+
+   \node[] (p1) at (4.5,2.5) {$\mathcal{P}$};
+   %reduction c3 c5
+   \node[] (s5) at (5.5,3.5) {$\mathcal{S}$};
+   \node[] (c3) at (5,4.5) {$k_3$};
+   \node[] (c5) at (6,4.5) {$k_0$};
+   %reduction c2 c6
+   \node[] (s6) at (3.5,3.5) {$\mathcal{S}$};
+   \node[] (p0) at (2.8,4.5) {$\mathcal{P}$};
+   \node[] (c2) at (2.3,5.5) {$k_2$};
+   \node[] (star4) at (3.3,5.5) {$k_{1;4}^{sync}$};
+   \node[] (s7) at (4.2,4.5) {$\mathcal{S}$};
+   \node[] (c4) at (3.8,5.5) {$k_4$};
+   \node[] (c6) at (4.8,5.5) {$k_6$};
+
+  \path[->]
+    %(c0s) edge node {} (c8d)
+
+    (s8) edge node {} (s9)
+         edge node {} (s2)
+    (s9) edge node {} (p1)
+         edge node {} (s0)
+
+    %reduction c0 c1
+    (s0) edge node {} (s1)
+         edge node {} (c1)
+    (s1) edge node {} (c0)
+         edge node {} (star1)
+    %reduction c7 c8
+    (s2) edge node {} (s3)
+         edge node {} (c8)
+    (s3) edge node {} (c7)
+        edge node {} (star8)
+
+    (p1) edge node {} (s6)
+         edge node {} (s5)
+    %reduction c3 c5
+    (s5) edge node {} (c3)
+         edge node {} (c5)
+    %reduction c2 c6
+    (s6) edge node {} (p0)
+         edge node {} (s7)
+    %reduction c2 *4
+    (p0) edge node {} (c2)
+         edge node {} (star4)
+    (s7) edge node {} (c4)
+         edge node {} (c6);
+  \end{tikzpicture}
+\caption{Serie-Parallel tree decomposition of the example of program of Figure~\ref{fig:exmsl}}
+\label{fig:tree}
+\end{center}
+\end{figure}
+
+%--------------------
+\subsection{Performance model}
+
+%--------------------
+\subsection{Fusion optimization}
+
+Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization.
+Even though the MSL language is not dedicated to produce very optimized stencil codes, but to produces the parallel pattern of the application, building the $TSP$ tree make available an easy optimization when the data parallelization technique is the only one used.
+This optimization consists in proposing a valid merge of some computation kernels inside a single space loop.
+As a result, the user can use this valid fusion of kernels or not when implementing those.
+
+Those fusions can be computed from the canonical form of the $TSP$ tree decomposition. The canonical form consists in recursively merging successive $S$ vertices or successive $P$ vertices of $TSP$.
+
+The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and where $k_{i;j}^{fus}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.
+Finally, $type(k)$ returns $comp$ if the kernel is a computation kernel, and $sync$ otherwise.
 
 \begin{algorithm}
-\caption{Remove N-Shapes}
-\label{alg:nshape}
+\caption{$F_{fus}$}
+\label{alg:fus}
 \begin{algorithmic}[1]
-\Procedure{removeNSH} {$\Gamma_{dep}$}
-\\$\Gamma_{dep}'$ = RTV($\Gamma_{dep}$)
-\\$sources$ = roots($\Gamma_{dep}'$)
-\\BFSnshape($sources$,$\Gamma_{dep}'$,$\Gamma_{dep}$)
-\\TC($\Gamma_{dep}$)
-\EndProcedure
-
-\Procedure{BFSnshape} {$sources$, $\Gamma_{dep}'$, $\Gamma_{dep}$}
-\\$successors$ = successors($sources$)
-\\$SG$ = subgraph($sources \cup successors$)
-\If{isconnected($SG$)}
-\If{! complete($SG$)}
-\\addedges($\Gamma_{dep}'$)
-\\addedges($\Gamma_{dep}$)
-\\BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
+\Procedure{$F_{fus}$} {$TSP(V,E)$}
+\For {$(k_i,k_j) \in V^2$}
+\If {parent($k_i$)==parent($k_j$)}
+\If {$type(k_i)=type(k_j)=comp$}
+\If {parent($k_i$)==$S$}
+\If {$D_i==D_j$}
+\State propose the fusion $k_{i;j}^{fus}$
 \EndIf
-\If{! isconnected($SG$)}
-\\BFSnshape($successors$, $\Gamma_{dep}'$, $\Gamma_{dep}$)
+\ElsIf {parent($k_i$)==$P$}
+\If {$D_i==D_j$ and $R_i \cap R_j \neq \emptyset$}
+\State propose the fusion $k_{i;j}^{fus}$
 \EndIf
 \EndIf
+\EndIf
+\EndIf
+\EndFor
 \EndProcedure
@@
 
-After these over-constraints are applied, $\Gamma_{dep}$ is MSPD.
-As a result, the binary tree decomposition algorithm of Valdes \& Al can be applied on $\Gamma_{dep}$ to get a static scheduling of the multi-stencil program.
-
-%--------------------
-\subsection{Series-parallel binary tree decomposition}
-A \emph{binary decomposition tree} is a tree where each leaf is a vertex of the MSPD DAG it represents, and whose internal nodes are labelled $S$ or $P$ to indicate respectively the series or parallel composition of sub-trees.
-
-Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have also proposed a linear algorithm to know if a DAG is MSPD and, if it is, to decompose it to its associated binary decomposition tree.
-This algorithm is based on the duality of the class of MSPD DAGs, with the class of \emph{Two Terminal Series Parallel} DAGs (TTSP) from which a binary tree decomposition can be performed in linear time.
-
-\begin{mydef}
-The \emph{line digraph} of a digraph (directed graph) $G$ is a digraph $L(G)$ that has:
+We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs for example~\cite{spaaTangCKLL11}.
+However, this fusion takes place at a different level and can bring performance improvments as it will be illustrated in Section~\ref{sect:eval}.
+This fusion algorithm relies on very simple statements:
 \begin{itemize}
-\item a vertex $f(e)$ for each edge $e$ of $G$; and
-\item an edge $(f(e_1),f(e_2))$ for each pair of edges of $G$ of the form $e_1=(u,v)$, $e_2=(v,w)$.
+\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependant.
+As a result, what is written by the first one is read by the second one.
+Thus, at least one data is common to those computations (the one written by $k_i$).
+Thus, if the computation domains verify $D_i=D_j$, the fusion of $k_i$ and $k_j$ will decrease cache misses.
+\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependant.
+However, if the computation domains verify $D_i=D_j$, and if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}^{fus}$.
 \end{itemize}
-\end{mydef}
-
-\begin{mydef}Two Terminal Series Parallel
-\begin{itemize}
-\item A digraph consisting of two vertices joined by a single edge is TTSP.
-\item If $G_1$ and $G_2$ are TTSP digraphs, so is the digraph obtained by either of the following operations:
-\begin{itemize}
-\item \emph{Two terminal parallel composition}: identify the source of $G_1$ with the source of $G_2$ and the sink of $G_1$ with the sink of $G_2$.
-\item \emph{Two terminal series composition}: identify the sink of $G_1$ with the source of $G_2$.
-\end{itemize}
-\end{itemize}
-\end{mydef}
-
-\begin{myth}
-If the DAG $G$ is a MSPD graph, its \emph{inverse line DAG} $L^{-1}(G)$ is TTSP.
-\end{myth}
-
-It has to be indicated that for a DAG $G$, $L^{-1}(G)$ is not unique.
-Thus, in the work of Valdes \& Al, and in this work, $L^{-1}(G)$ will refer to the unique digraph having a single source and a single sink whose line digraph is $L(L^{-1}(G))=G$.
-
-Once the removal of N-Shapes has been applied onto $\Gamma_{dep}$, the linear series-parallel tree decomposition can be applied following the algorithm of~\cite{Valdes:1979:RSP:800135.804393}, called in this paper $BDT(L^{-1}(\Gamma_{dep}))$.
-
-The binary decomposition tree algorithm is based on the fact that the decomposition can be obtained as a byproduct of a reduction process on $L^{-1}(\Gamma_{dep})$, which is TTSP.
-In order to obtain the decomposition, Valdes \& Al associate a label with each edge of the digraph being reduced.
-Initially the label of each edge is a trivial binary tree consisting of a single node.
-As the reduction process introduces new edges the rules of Figure~\ref{fig:rules} are used to compute the binary trees used to label them.
-The sequencial composition rule keep the order $T_1$ then $T_2$.
-The algorithm ends when the TTSP graph is reduced to its minimum, i.e.\ two vertices and a single edge between them.
-
-\begin{figure}[h!]
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-%FIRST RULE
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};
-
-  \path[->]
-    (x) edge node [above] {$T_1$} (y)
-    (y) edge node [above] {$T_2$} (z);
-  \end{tikzpicture}
-  }
-\hspace{5pt}
-%ARROW
-\subfloat[]{
-$\Rightarrow$
-}
-\hspace{5pt}
-%SECOND RULE
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-   \node[] (s) at (1.5,0.5) {$\mathcal{S}$};
-   \node[] (t1) at (1,1.5) {$T_1$};
-   \node[] (t2) at (2,1.5) {$T_2$};
-
-  \path[->]
-    (x) edge node {} (y)
-    (s) edge node {} (t1)
-        edge node {} (t2);
-  \end{tikzpicture}
-}
-\\
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
-
-  \path[->]
-    (x) edge [bend left] node [above] {$T_1$} (y)
-        edge [bend right] node [below] {$T_2$} (y);
-  \end{tikzpicture}
-  }
-\hspace{5pt}
-\subfloat[]{
-$\Rightarrow$
-}
-\hspace{5pt}
-\subfloat[]{
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
-   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
-   \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
-   \node[] (t1) at (1,1.5) {$T_1$};
-   \node[] (t2) at (2,1.5) {$T_2$};
-
-  \path[->]
-    (x) edge node {} (y)
-    (p) edge node {} (t1)
-        edge node {} (t2);
-  \end{tikzpicture}
-}
-\caption{Reduction rules of the decomposition tree algorithm.}
-\label{fig:rules}
-\end{center}
-\end{figure}
-
-We have explained in this section how a valid static scheduling of $\Gamma_{dep}$ can be built using existing graph transformations of the MSPD class.
-In the next section, the overall compiler algorithm and a complete example is given.

### 2016-05-20 ARTICLES/IJPP/article/parallelism.tex
IJPP article modifications parties 4 5 et 6

@@
 
+\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl} gives an example of $\mathcal{MSP}$ program. From this example, the following ordered list of computation kernels can be extracted:
+\begin{equation*}
+\Gamma = [k_0,k_1,k_2,k_3,k_4,k_0,k_6,k_7,k_8]
+\end{equation*}
+From this ordered list of computation kernels $\Gamma$, and from the rest of the multi-stencil program, synchronizations can be automatically detected from the call to $F_{sync}(\Gamma,0)$ to get the synchronized ordered list of kernels:
+\begin{equation}
+\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_0,k_6,k_7,k_{7;8}^{sync},k_8],
+\label{eq:exsync}
+\end{equation}
+\begin{subequations}
+where
+\begin{align}
+        k_{0;1}^{sync}=(\emptyset,\{(B,nce)\},(B,\cup_{\phi \in D_1} nce(\phi)),identity),\\
+        k_{1;4}^{sync}=(\emptyset,\{(C,nec)\},(C,\cup_{\phi \in D_4} nec(\phi)),identity),\\
+        k_{7;8}^{sync}=(\emptyset,\{(I,ncc)\},(I,\cup_{\phi \in D_8} ncc(\phi)),identity).
+\end{align}
+\end{subequations}
+
 %------------------------------
@@
 \end{equation*}
+
+\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl} gives an example of $\mathcal{MSP}$ program.
+From $\Gamma_{sync}$ that has been built in Equation~\ref{eq:exsync}, the dependency DAG can be built.
+For example, as $k_0$ computes $B$ and $k_1$ reads $B$, $k_0$ and $k_1$ becomes vertices of $\Gamma_{dep}$, and an arc $(\overset{\frown}{k_0,k_1})$ is added to $\Gamma_{dep}$.
+The overall $\Gamma_{dep}$ built from the call to $F_{dep}(\Gamma_{sync},0)$ is drawn in Figure~\ref{fig:depdep}.
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (c0) at (0,0) {$k_0$};
+   \node[] (star1) at (1,0) {$k_{0;1}^{sync}$};
+   \node[] (c1) at (2,0) {$k_1$};
+   \node[] (c2) at (3,0.5) {$k_2$};
+   \node[] (star4) at (3,1.5) {$k_{1;4}^{sync}$};
+   \node[] (c3) at (3,-0.5) {$k_3$};
+   \node[] (c4) at (4,0.5) {$k_4$};
+   \node[] (c5) at (4,-0.5) {$k_5$};
+   \node[] (c6) at (5,0.5) {$k_6$};
+   \node[] (c7) at (6,0) {$k_7$};
+   \node[] (star8) at (7,0) {$k_{7;8}^{sync}$};
+   \node[] (c8) at (8,0) {$k_8$};
+
+  \path[->]
+    (c0) edge node {} (star1)
+    (star1) edge node {} (c1)
+    (c1) edge node {} (c2)
+         edge node {} (c3)
+         edge node {} (star4)
+    (star4) edge node {} (c4)
+    (c2) edge node {} (c4)
+    (c4) edge node {} (c6)
+    (c3) edge node {} (c5)
+    (c5) edge node {} (c7)
+    (c6) edge node {} (c7)
+    (c7) edge node {} (star8)
+    (star8) edge node {} (c8);
+  \end{tikzpicture}
+\caption{$\Gamma_{dep}$ of the example of program of Figure~\ref{fig:exmsl}}
+\label{fig:depdep}
+\end{center}
+\end{figure}

### 2016-05-24 ARTICLES/IJPP/article/msp.tex
IJPP article modifications performance model

@@
 
+In this subsection are introduced two performance model, one for the data parallelization technique, and one for the hybrid data and task parallelization technique, both previously explained.
+
+The performance model of a data parallelization technique is inspired from the Bulk Synchronous Parallel model.
+Actually data parallelization technique consider that each process has its own part of the domain.
+Thus the performance model reveals that the computation time is the sum of the reference sequential time divided by the number of processes, and of the time spent in communications between processes.
+Thus, for
+\begin{itemize}
+\item $T_{SEQ}$ the sequential reference time, 
+\item $P$ the total number of processes, 
+\item $T_{COM}$ the communications time, 
+\end{itemize}
+the total computation time is
+\begin{equation}
+T = \frac{T_{SEQ}}{P} + T_{COM}.
+\end{equation}
+
+% The time spent into communications between MPI processes can be represented as the time spent to send data onto the network and the time spent to synchronize processes. Thus, for
+% \begin{itemize}
+% \item $S$ the total number of bytes sent during the simulation
+% \item $T_s$ the time to send one byte
+% \item $N_{COM}$ the number of communication steps
+% \item $Sync$ the time to synchronize two MPI processes for communications
+% \end{itemize}
+% the communication time is
+% \begin{equation*}
+% T_{COM} = S \times T_s + N_{COM} \times Sync.
+% \end{equation*}
+
+Thus, when the number of processes $P$ increase in data parallelization, the performance model limit is $T_{COM}$
+\begin{equation}
+\lim\limits_{P \rightarrow +\infty} T = T_{COM}.
+\end{equation}
+
+As a result, the critical point for performance is when $T_{COM} \geq \frac{T_{SEQ}}{P}$, which happens naturally in data parallelization as $T_{COM}$ will increase with the number of processes, and $\frac{T_{SEQ}}{P}$ decrease with the number of processes.
+
+This limitation is always true, but can be delayed by different strategies.
+First, it is possible to perform communications through the network while computations are performed simultaneously.
+Second, it is possible to introduce another kind of parallelization, task parallelization.
+Thus, for the same total number of processes, only a part of them are used for data parallelization, and the rest are used for task parallelism.
+As a result, $\frac{T_{SEQ}}{P}$ will continue to decrease but $T_{COM}$ will increase later.
+This second strategy is the one studied in the following hybrid performance model.
+
+For an hybrid (data and task) parallelization technique, and for
+\begin{itemize}
+\item $P_{data}$ the number of processes used for data parallelization,
+\item $P_{task}$ the number of processes used for task parallelization,
+\item such that $P = P_{data} \times P_{task}$ is the total number of processes used,
+\item $T_{task}$ the overhead time due to task parallelization technique,
+\item and $F_{task}$ the task parallelization degree of the application,
+\end{itemize}
+the total computation time is
+\begin{equation}
+T = \frac{T_{SEQ}}{P_{data} \times F_{task}} + T_{COM} + T_{task}
+\end{equation}
+
+The time overhead due to task parallelization can be represented as the time spent to create a pool of threads and the time spent to synchronize those threads. Thus, for
+\begin{itemize}
+\item $T_{cr}$ the total time to create the pool of threads (may happened more than once), 
+\item $T_{sync}$ the total time spent to synchronize threads, 
+\end{itemize}
+the overhead is
+\begin{equation*}
+T_{task} = T_{cr} + T_{sync}.
+\end{equation*}
+
+The task parallelization degree of the application $F_{task}$ is the limitation of a task parallelization technique.
+As explained before, a task parallelization technique is based on the dependency graph of the application.
+Thus, this dependency graph must expose enough parallelism for the number of available threads.
+For this performance model we consider that
+\begin{equation*}
+F_{task} = P_{task}.
+\end{equation*}
+However, the upper bound of $F_{task}$ is constrained by the dependency graph of the application and by the time spent in each task.
+
+\medskip
+As a result when $P_{data}$ is small a data parallelization technique may be more efficient, while an hybrid parallelization could be interesting at some point to improve performance.
+The question asked here is when is it interesting to use hybrid parallelization.
+
+To answer this question lets consider the two parallelization techniques, data only and hybrid. We denote
+\begin{itemize}
+\item $P_{data1}$ the total number of processes entirely used by the data only parallelization,
+\item $P_{data2}$ the number of processes used for data parallelization in the hybrid parallelization,
+\item and $P_{task}$ the number of processes used for task parallelization in the hybrid parallelization,
+\item such that $P_{data1} = P_{data2} \times P_{task}$.
+\end{itemize}
+
+We search the point where the data parallelization is less efficient than the hybrid parallelization. Thus, 
+\begin{equation*}
+\frac{T_{SEQ}}{P_{data1}} + T_{COM1} \geq \frac{T_{SEQ}}{P_{data2} \times P_{task}} + T_{COM2} + T_{task}.
+\end{equation*}
+
+This happens when
+\begin{equation}
+T_{COM1} \geq T_{COM2} + T_{TH}
+\end{equation}
+
+This performance model will be validated and will help to explain results the Section~\ref{sect:eval}.
+
 %--------------------

### 2016-05-24 ARTICLES/IJPP/article/compiler.tex
IJPP article modifications implementation

@@
 
-SkelGIS + OpenMP + L2C
+The multi-stencil language is a descriptive language.
+MSL follows a light grammar which extracts information needed to build an empty parallel pattern of a numerical simulation (a multi-stencil program), also called an empty parallel skeleton of the simulation.
+From the language grammar are automatically detected where synchronizations are needed in order to apply a data parallelization technique, and how computations are dependent from each other to apply a task parallelization technique.
+
+The multi-stencil language splits the simulation description from its implementation concerns.
+For this reason, the multi-stencil language is mesh-agnostic, which means that no information on the actual topology of the mesh is given by the user.
+For this reason too, no numerical code is asked to the user.
+Those implementation concerns are deported to a second phase where other languages or libraries can be used.
+This phase is called the \emph{dump} in the overall compiler algorithm.
+
+In this paper, the MSL dump has been implemented using two existing languages, SkelGIS and OpenMP.
+
+SkelGIS is a C++ embedded language~\cite{} which offers distributed data structures for numerical simulations, and programming interfaces to easily write codes using them.
+The code produced by SkelGIS uses MPI (message Passing Interface~\cite{}).
+Thus, this code can be used on distributed memory architectures such as clusters of machines.
+The distributed data structures offered by SkelGIS are, first, a distributed two dimensional Cartesian mesh~\cite{}, and second, a distributed graph of Cartesian meshes (hybrid mesh)~\cite{}.
+SkelGIS is used in this paper for the dump of MSL and more precisely to produce the code associated to the data parallelization technique.
+
+OpenMP~\cite{} is a language based on directives added to C/C++ or Fortran.
+Those directives offers a way to take advantage of shared memory architectures such as multi-processors, multi-cores, many-cores or even GPUs (in OpenMP 4.0).
+Among other available directives, some of them offer the possibility to manage a pool of threads and to produce a task parallel program.
+In this paper OpenMP is used in its version 3.0, which means that we do not take advantage of the dynamic schedulers available for OpenMP 4.0 such as libgomp~\cite{}, StarPU~\cite{}, or XKaapi~\cite{}.
+Those dynamic schedulers will be the subject of future work.
+In this paper, however, is evaluated the performance of the static scheduling introduced in Section~\ref{sect:msp}.
+
+The dump step of the MSL compiler automatically generates an hybrid code which uses the SkelGIS Cartesian distributed data structure and an OpenMP static scheduling generated from the algorithm previously described in Section~\ref{sect:tsp}.
+To facilitate the maintainability of this hybrid code, a component-based programming model is used, the Low Level Component Model~\cite{}.
+
+A component is a piece of code equivalent to a class in object-oriented programming, but which exposes not only provided methods, but also methods used from other components.
+Such programming models increase maintainability, productivity and code-reuse.
+
+The advantage of using such models in MSL is to clearly separate the code responsible for data parallelization from the code responsible for the static scheduling of tasks.
+Moreover, one component is generated for each computation kernel which was described by the user in its MSL description.
+As a result, the user simply has to fill those components without any knowledge on the rest of the generated code.
+To fill its computation kernels, the user must use programming interfaces proposed by SkelGIS.
+
+One can notice that MSL could use different languages as it is mesh-agnostic.
+For example, a future work is to use the distributed unstructured mesh proposed by PamPA~\cite{}.
+Moreover, MSL could use pthreads instead of OpenMP to manage the static scheduling.

### 2016-05-26 ARTICLES/IJPP/article/evaluation.tex
IJPP article modifications evaluation

This section presents evaluations of the work presented in this paper. First, the compiler is evaluated. Second, the data prallelization technique is evaluated, as well as the fusion optimization, and compared to a full SkelGIS program. Third, the hybrid parallelization (data and task) is evaluated and validates the performance model detailed in Section~\ref{sect:tsp}.

All evaluations presented in this section are based on a real case study, the shallow-Water Equations. Navier-Stokes Equations is a well known set of partial differential equations in fluid dynamics to simulate a flow evolution in time. At the University of Orl\'eans, France, the MAPMO laboratory works on a software, called FullSWOF2D\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}, which solves the Shallow-Water Equations obtained from the three dimensional Navier-Stokes equations, by averaging on the vertical direction~\cite{Ferrari2004}. Those equations are solved using a two-dimensional Cartesian discretization of the space domain, and a finite volume numerical methods more described in~\cite{CPE:CPE3494}. We have developed a MSL version of FullSWOF2D that contains 3~mesh entities, 7~computation domains, 48~data and 98~computations (32~stencil kernels and 66~local kernels).

%-------------------------------------
\subsection{Compiler}

The series-parallel tree decomposition $TSP$ of this simulation, extracted by the MSL compiler, is composed of 17~sequence nodes and 18~parallel nodes. Table~\ref{fig:freq} represents the number of time a given level of parallelism, \ie the number of tasks to perform concurrently, is observed in the final tree. One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1). Moreover, a level of 16~parallel tasks is reached two times, and five times for the fourth level.

\begin{table}[!h]
 \begin{center}
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline 
   Level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
   \hline
   Frequency & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
   \hline
 \end{tabular}
\caption{Parallelism level (number of parallel tasks) and the number of times this level appears.}
\label{fig:freq}
 \end{center}
\end{table}

Table~\ref{fig:exectime} illustrates the execution time for each step of the MSL compiler for an overall execution time of 4.6 seconds. Execution times have been computed on a laptop with a bi-core Intel Core i5 1.4~GHz, and 8~GB of DDR3. One can notice that the computation of the $TSP$ tree is the longest step of the compiler.

\begin{table}[!h]
 \begin{center}
 \begin{tabular}{|c|c|c|c|c|}
  \hline
   Step & Parser & $\Gamma_{sync}$ & $\Gamma_{dep}$ & $TSP$\\
   \hline
   Time (ms) & 1 & 2 & 4.2 & 3998.5\\
   \hline
   \% & 0.022 & 0.043 & 0.09 & 86.6\\
   \hline
 \end{tabular}
\caption{Execution times of the MSL compiler}
\label{fig:exectime}
 \end{center}
\end{table}

%-------------------------------------
\subsection{Data parallelism}

The first parallelization technique proposed by MSL is a data parallelization. To implement this parallelization, the dump step of the MSL compiler uses the SkelGIS language. This language, as already described, offers a distributed Cartesian mesh and programming interfaces to use it as a sequential data structure. It has been evaluated itself compared to an MPI implementation on the same FullSWOF2D simulation~\cite{}. What MSL brings to SkelGIS is an automatic detection of where MPI synchronizations are needed during the numerical simulation, and how to make a fusion of some computations in order to reduce cache misses. This section gives a performance comparison of the code produced by MSL for FullSWOF2D, and of the equivalent manual SkelGIS program (with manual synchronizations and fusion choices). Moreover, as the code produced by MSL is a component-based code, this evaluation also shows that no overheads are introduced \llc~\cite{}. Following evaluations have been performed onto the machine of the TGCC Curie described in Table~\ref{tab:TGCC}. Each evaluation has been performed ten times and the median is presented in results.

\begin{table}[!h]
\begin{center}
 \begin{tabular}{|c|c|}
   \hline
    Cluster & \textbf{TGCC Curie Thin Nodes}\\
     \hline         
    Processor & 2$\times$SandyBridge\\
    & (2.7 GHz)\\
    Cores/node & 16 \\
    RAM/node & 64 GB\\
    RAM/core & 4GB\\
    Compiler [-O3] & gcc 4.9.1\\
    MPI & Bullxmpi\\
    Network & Infiniband\\
    \hline
 \end{tabular}
 \caption{\label{tab:TGCC}Hardware configuration of TGCC Curie Thin nodes.}
 \end{center}
\end{table}

\paragraph{\textbf{Weak scaling}} The first evaluation is a weak scaling evaluation. In a weak scaling, a constant amount of work is given to each core while increasing the overall number of cores. As a result, the overall domain size also increase with the number of cores. A weak scaling is a good evaluation to detect overheads which are independent from the amount of computation performed by each core, such as collective communications, or overheads introduced by \llc components, for example. Figures~\ref{fig:weak1} and~\ref{fig:weak2} respectively show the weak scaling evaluations for a $400 \times 400$ and a $800 \times 800$ domain for each core, from 16 cores to 16.384 cores. Minimum and maximum values are also shown as error bars in figures.

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/400/median_weak.pdf}}
  \caption{weak-scaling with $400 \times 400$ domain per core.}
  \label{fig:weak1}
\end{center}\end{figure}
\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/800/median_weak.pdf}}
  \caption{weak-scaling with $800 \times 800$ domain per core.}
  \label{fig:weak2}
\end{center}\end{figure}

One can notice that the MSL code produces a better execution time on the $400 \times 400$ domain, while it produces a small overhead on the $800 \times 800$ domain. Those results could be due to different optimizations of the gcc compiler due to components. Actually, a component in \llc is compiled independently, as an dynamic library, with the \emph{-fpic} compilation option, while the SkelGIS version is compiled entirely as a single code. However, one can also notice that except those not significant differences of execution times, weak scalings have the same behavior on both codes.

\paragraph{\textbf{Strong scaling}} The second evaluation is a strong scaling evaluation. Unlike a weak scaling, a strong scaling keeps a constant overall size of problem, while the number of cores increase. As a result, the sub-domain computed by each core reduces while the number of cores increases. The strong scaling is a good additional evaluation to the weak scaling, showing different kinds of overheads, such as cache misses (when the size of sub-domain becomes small enough to fit into cache memory), or small constant overheads. Figure~\ref{fig:strong} shows the strong scaling evaluation for a $10k \times 10k$ overall domain size, from 16 cores to 16.384 cores. This strong scaling is illustrated with the number of iterations per seconds as a function of the number of cores. The ideal strong scaling is also illustrated.

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/strong_scaling/10K_1K/median_strong.pdf}}
  \caption{Strong scaling on a $10k \times 10k$ domain.}
  \label{fig:strong}
\end{center}\end{figure}

First, one can notice that the strong scaling evaluated for the code generated by MSL is close to the ideal speedup up to 16.384 cores, which is a very good result. Moreover, no overheads are introduced by MSL which shows that automatic synchronization detections and automatic fusion detections are the same one that the one written manually into the SkelGIS code of FullSWOF2D. Finally, no overheads are introduced by components of \llc. A small behavior difference can be noticed with $2^9=512$ cores, however this variation is no longer observed with 1024 cores.

\paragraph{\textbf{Fusion}} Finally, to evaluate the data parallelization technique automatically introduced by MSL, the fusion optimization is evaluated. Figure~\ref{fig:fusion} shows the number of iterations per second as a function of the number of cores, for FullSWOF2D with and without fusion optimization and onto a $500 \time 500$ domain. As explained in Section~\ref{sect:fusion}, the MSL fusion happens at a high level and is most of the time done naturally by a computer scientist. However, for a non computer scientist which write its numerical codes, an automatic proposition of such fusions makes the implementation easier. Moreover, one can notice that the performance is clearly improved (around 40\%) by this fusion.

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/fusVSbase.pdf}}
  \caption{Strong scaling on a 500x500 domain, with and without the fusion optimization.}
  \label{fig:fusion}
\end{center}\end{figure}

%-------------------------------------
\subsection{Hybrid parallelism}

The second parallelization technique proposed by MSL is an hybrid parallelization which mixes data parallelization and task parallelization. To implement this hybrid parallelization, the dump step of the MSL compiler uses the SkelGIS language, and the OpenMP language.

First, Figure~\ref{fig:limit} illustrates limitations of data parallelization technique alone. As explained in Section~\ref{sect:perfs}, performance of data parallelism is limited by the communication time between resources. As a result, when the size of sub-domain for each core becomes too small, and as the communication time increases with the number of cores, the strong speedup of the data parallelization starts to bend down. To catch this phenomena without too much cores, a relatively small domain of $500 \times 500$ is used.

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/times.pdf}}
  \caption{Computation vs communication times in the data parallelization technique.}
  \label{fig:limit}
\end{center}\end{figure}

As expected, from 1 to 2048 cores the computation times decrease linearly. On the other hand, from 1 to 256 cores the communication time is almost constant, but from 256 to 2048 cores it increases. As a result from 512 cores to 2048 cores, the communication time becomes greater than the computation time. In Figure~\ref{fig:close}, the blue curve shows the strong scaling of the same example. Thus, from 256 to 2048 cores, the speedup bends down.

\medskip
In addition to the blue curve, Figure~\ref{fig:close} shows strong scalings for the same example but using an hybrid parallelization. For example, the purple curve shows the parallelization which uses 8 cores, to perform the task parallelization, for each process used for data parallelization (\ie MPI process). As a result, for example, when using 2 machines of the TGCC cluster, with a total of 32 cores, 4 cores are used for SkelGIS MPI processes, for data parallelization, and for each one 8 cores are used for task parallelization ($4 \times 8 = 32$).

As a result, and as explained in Section~\ref{sect:perfs}, data is less divided into sub-domains and the effect which is observe onto the blue curve is delayed. This figure shows a comparison with 2, 4, 8 and 16 cores per MPI process for task parallelization.

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_close_median.pdf}}
  \caption{Strong scaling comparisons between data parallelization and hybrid parallelization. A \emph{close} OpenMP clause is used to bind threads onto cores.}
  \label{fig:close}
\end{center}\end{figure}

From 2 to 8 cores, the improvement of the strong scaling is clear. However, reaching 16 cores, an important initial overhead appears and in addition to this, the curve bends down rapidly instead of improving the one with 8 cores for task parallelization. Two different phenomena happen.

First, thin nodes of the TGCC Curie are built with two NUMA nodes each of 8 cores. As a result, when increasing the number of OpenMP cores for task parallelization from 8 to 16 cores, an overhead is introduced by exchanges of data between memories of the two NUMA nodes. This phenomena is illutrated in Figure~\ref{fig:spread}. In this figure, a different strategy is used to bind threads onto available cores (using OpenMP). This strategy, called \emph{spread} instead of \emph{close} in Figure~\ref{fig:close}, binds threads on cores in order to spread as much as possible onto resources, which means that the two NUMA nodes are directly used whatever the number of cores kept for task parallelization. As a result, and as shown in the figure, using 2, 4 and 8 cores an initial overhead is introduced.

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_spread_median.pdf}}
  \caption{Strong scaling comparisons between data parallelization and hybrid parallelization. A \emph{spread} OpenMP clause is used to bind threads onto cores.}
  \label{fig:spread}
\end{center}\end{figure}

The second phenomena, which happens in Figure~\ref{fig:close} with 16 cores for tasks parallelization, is due to the level of parallelization introduced by the task parallelization technique. Actually, as illustrated in Table~\ref{fig:freq}, only two steps of the $TSP$ static scheduling generated by the MSL compiler can take advantage of 16 cores among a total of 18 steps. This phenomena has been explained in Section~\ref{sect:perfs} by the variable $F_{task}$ and the fact that it is not always true that $F_{task}=P_{task}$. This explains why using 16 cores for task parallelization in Figure~\ref{fig:spread} is still less efficient than using 8 cores even if the two NUMA nodes are always used in this evaluation.

Finally, to completely validate the performance model introduced in Section~\ref{sect:perfs}, and to understand when the hybrid parallelization becomes more interesting than the data parallelization, Figure~\ref{tth2} represents $T_{COM1}$ and $T_{COM2}+T_{task}$ of Equation~(\ref{eq:hyb}), for the best case, \ie when 8 cores are used for task parallelization, and with a \emph{close} OpenMP bind of threads onto cores. Table~\ref{fig:tth} gives time details.

\begin{table}[!h]
 \begin{center}
 \begin{tabular}{|c|c|c|c|c|}
    \hline 
    & $T_{COM1}$ & $T_{COM2}$ & $T_{task}$ & Equation~(\ref{eq:hyb})\\
   \hline
   16 cores ($2 \times 8$) & 0.0005 & 0.00032 & 0.013 & False\\
   32 cores ($4 \times 8$) & 0.0018 & 0.00045 & 0.0062 & False\\
   64 cores ($8 \times 8$) & 0.0013 & 0.00038 & 0.0034 & False\\
   128 cores ($16 \times 8$) & 0.00075 & 0.0005 & 0.0023 & False\\
   256 cores ($32 \times 8$) & 0.00077 & 0.0018 & 0.001 & False\\
   512 cores ($64 \times 8$) & 0.0029 & 0.0013 & 0.00052 & True\\
   1024 cores ($128 \times 8$) & 0.018 & 0.00075 & 0.00029 & True\\
   2048 cores ($256 \times 8$) & 0.0623 & 0.00077 & 0.00016 & True\\
   \hline
 \end{tabular}
\caption{Execution times (seconds) of $T_{COM1}$, $T_{COM2}$ and $T_{task}$ for 8 cores for task parallelization. Verification of the Equation~(\ref{eq:hyb}).}
\label{fig:tth}
 \end{center}
\end{table}

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/tth.pdf}}
  \caption{Execution times (seconds) of $T_{COM1}$ and $T_{COM2} + T_{task}$ for 8 cores for task parallelization. Verification of the Equation~(\ref{eq:hyb}).}
  \label{fig:tth2}
\end{center}\end{figure}

Figure~\ref{fig:tth2} and Table~\ref{fig:tth} perfectly matches results observed in Figure~\ref{fig:close} for 8 cores used for task parallelization per core used for data parallelization. As a result, the hybrid parallelization becomes better with a total of 512 cores.

### 2016-05-26 ARTICLES/IJPP/article/evaluation.tex
IJPP article modifications evaluation

@@
 
-Finally, to completely validate the performance model introduced in Section~\ref{sect:perfs}, and to understand when the hybrid parallelization becomes more interesting than the data parallelization, Figure~\ref{tth2} represents $T_{COM1}$ and $T_{COM2}+T_{task}$ of Equation~(\ref{eq:hyb}), for the best case, \ie when 8 cores are used for task parallelization, and with a \emph{close} OpenMP bind of threads onto cores.
+Finally, to completely validate the performance model introduced in Section~\ref{sect:perfs}, and to understand when the hybrid parallelization becomes more interesting than the data parallelization, Figure~\ref{fig:tth2} represents $T_{COM1}$ and $T_{COM2}+T_{task}$ of Equation~(\ref{eq:hyb}), for the best case, \ie when 8 cores are used for task parallelization, and with a \emph{close} OpenMP bind of threads onto cores.
 Table~\ref{fig:tth} gives time details.

### 2016-05-27 ARTICLES/IJPP/article/article.tex
IJPP article modifications performance model
IJPP article modifications evaluation
IJPP article related work

@@
 \usepackage{todonotes}
+\usepackage{url}
+\usepackage{xspace}
+
+\newcommand{\fix}[1]{\todo[inline,color=red!80]{#1}}
+\newcommand{\cf}{\emph{cf.}\xspace}
+\newcommand{\ie}{\emph{i.e.,}\xspace}
+\newcommand{\llc}{L\textsuperscript{2}C\xspace}
 
@@
 \author{Helene Coullon \and
-        Christian Perez \and
-        Julien Bigot
+        Julien Bigot \and
+        Christian Perez
 }
@@
 %------------------------------------------------------------------------------
+\section{Related work}
+\label{sect:rel}
+\input{related.tex}
+%------------------------------------------------------------------------------
 \section{Computational model of Multi-Stencil Programs}
@@
 \label{sect:eval}
-\begin{itemize}
-\item evaluation of the compiler (execution times of the different steps)
-\item evaluation of the data parallelism (weak/strong scaling)
-\item evaluation of fusions
-\item evaluation of the hybrid performance
-\item validation of the performance model
-\end{itemize}
-
-% \begin{figure}[!h]\begin{center}
-%   \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/400/median_weak.pdf}}
-%   \caption{weak-scaling 400x400 blocks.}
-%   \label{fig:mesh}
-% \end{center}\end{figure}
-
-% \begin{figure}[!h]\begin{center}
-%   \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/800/median_weak.pdf}}
-%   \caption{weak-scaling 800x800 blocks}
-%   \label{fig:mesh}
-% \end{center}\end{figure}
-
-% \begin{figure}[!h]\begin{center}
-%   \resizebox{8cm}{!}{\includegraphics{../results/strong_scaling/10K_1K/median_strong.pdf}}
-%   \caption{strong-scaling 10kx10k, 1k iterations}
-%   \label{fig:mesh}
-% \end{center}\end{figure}
-
-% \item evaluation of the fusion
-
-% \begin{figure}[!h]\begin{center}
-%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/fusVSbase.pdf}}
-%   \caption{strong scaling 500x500, 200 iterations, with and without fusion}
-%   \label{fig:mesh}
-% \end{center}\end{figure}
-
-% \item evaluation of the hybrid parallelism
-
-% \begin{figure}[!h]\begin{center}
-%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_close_median.pdf}}
-%   \caption{MPI only vs MPI+threads with \emph{close} scheduling policy (OpenMP)}
-%   \label{fig:mesh}
-% \end{center}\end{figure}
-
-% \begin{figure}[!h]\begin{center}
-%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_spread_median.pdf}}
-%   \caption{MPI only vs MPI+threads with \emph{spread} scheduling policy (OpenMP)}
-%   \label{fig:mesh}
-% \end{center}\end{figure}
-
-% \item analytic model for the results
-
-% \begin{figure}[!h]\begin{center}
-%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/times.pdf}}
-%   \caption{Computation vs communication times in the MPI only application}
-%   \label{fig:mesh}
-% \end{center}\end{figure}
-
-% \end{itemize}
-%------------------------------------------------------------------------------
-\section{Related work}
-\label{sect:rel}
-\begin{itemize}
-\item stencil compilers cartesian : Pochoir, PATUS for optimizations/parallelization of a single stencil kernel
-\item stencil compilers unstructured meshes : Lizst for optimizations/parallelization of a single stencil kernel
-\item Halide and Nabla for pipeline of stencil computations in image (Cartesian) = mesh dependent
-\item Reuse of external solutions as SkelGIS, GA, Lizst, Pochoir etc.
-\end{itemize}
+\input{evaluation.tex}
 %------------------------------------------------------------------------------

### 2016-05-27 ARTICLES/IJPP/article/related.tex
IJPP article related work

Many domain specific languages have been proposed for stencil computations. Each one has its own specificities and answers to a specific stencil case or to a specific additionnal optimization. For example, Pochoir~\cite{spaaTangCKLL11} works on cache optimization techniques for stencils applied onto Cartesian meshes. On the other hand, PATUS~\cite{citeulike12258902} proposes to add a parallelization strategy grammar to its stencil language to perform an auto-tuning parallelization strategy. PATUS also works for stencils onto Cartesian meshes. Moreover, Halide~\cite{Ragan-Kelley:2013:HLC:2491956.2462176} works for Cartesian meshes, and proposes an optimization and parallelization of a pipeline of stencil codes. Finally, ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} is specific to multigrid numerical methods.

The reason why each of those languages has its own particularities and its own optimizations is because each language is built and thought as a single block, which makes impossible or difficult code reuse from one language to another. Thus, each language starts from scratch to answer its own constraints. In other words domain specific languages (for stencils or other domains) suffer from a lack of software engineering properties, which could increase the productivity to build a new language, by code-reuse, and also the maintainability of languages, with more separation of concerns. The aim of the Mulsi-Stencil Language (MSL), presented in this paper, is to offer more separation of concerns, by splitting the description of a stencil-based numerical simulation, from its implementation concerns. As a result, MSL is mesh-agnostic and is able to reuse existing other languages (SkelGIS and OpenMP in this paper).

Liszt~\cite{DeVito:2011:LDS:2063384.2063396} and Nabla~\cite{Camier:2015:IPP:2820083.2820107}, both offer languages for stencils applied onto any kind of mesh, from Cartesian to unstructured meshes. The mesh which is needed into the simulation can be built from a set of available symbols in the grammar of each language. Thus, those languages generalize the definition of a mesh, as it is proposed into the MSL formalism. 
However, two main differences can be noticed. First, the formalism is more flexible in MSL. For example, a computation can be applied onto a subset of the space domain which is not possible with Liszt and Nabla. This functionnality is an important flexibility in numerical simulations yet. For example, many computations performed in the simulation studied in this paper have to be applied onto a subpart of the overall space domain.
Second, in Liszt and Nabla the description of a simulation is not splitted from its implementation concerns. In other words, the topology of a mesh as well as numerical codes are given at the same time than the description of the different computations to apply into the simulation.

MSL splits those two different phases of a simulation conception, which improves separation of concerns and maintainability of the simulation. As a result, a numerician could describes the different computations to perform into the simulation, while another one, later, could focuss on the implementation and the numerical code of the simulation. This also facilitates reuse of other languages. The formalism of a Multi-Stencil program, given in this paper, is new, as far as we know. Finally, the parallelization techniques proposed in this paper, take place at paradigm level, with data parallelism and hybrid (data plus task) parallelism. Thus those parallelization do not need details onto architectures (distributed or shared memories, with or without accelerators). This makes possible a large panel of backend architectures and languages.

### 2016-05-27 ARTICLES/IJPP/article/article.tex
IJPP article biblio

@@
 \label{sect:introduction}
-Contributions :
-\begin{itemize}
-\item Formalism of a multi-stencil program and its parallelization
-\item The Multi-stencil Language: agnostic from the type of mesh and of the choosen implementation
-\item A first implementation of the MSL compiler (static scheduling + fusion + dump to SkelGIS/components)
-\item Evaluation of the different types of parallelism introduced + compiler + fusion
-\end{itemize}
+\input{introduction.tex}
 %------------------------------------------------------------------------------

### 2016-05-27 ARTICLES/IJPP/article/compiler.tex
IJPP article biblio

@@
 
-SkelGIS is a C++ embedded language~\cite{} which offers distributed data structures for numerical simulations, and programming interfaces to easily write codes using them.
-The code produced by SkelGIS uses MPI (message Passing Interface~\cite{}).
+SkelGIS is a C++ embedded language~\cite{CPE:CPE3494} which offers distributed data structures for numerical simulations, and programming interfaces to easily write codes using them.
+The code produced by SkelGIS uses MPI (message Passing Interface~\cite{Graham2009MSE}).
 Thus, this code can be used on distributed memory architectures such as clusters of machines.
-The distributed data structures offered by SkelGIS are, first, a distributed two dimensional Cartesian mesh~\cite{}, and second, a distributed graph of Cartesian meshes (hybrid mesh)~\cite{}.
+The distributed data structures offered by SkelGIS are, first, a distributed two dimensional Cartesian mesh~\cite{DBLP:conf/ieeehpcs/HeleneS13}, and second, a distributed graph of Cartesian meshes (hybrid mesh)~\cite{DBLP:conf/europar/CoullonL14}.
 SkelGIS is used in this paper for the dump of MSL and more precisely to produce the code associated to the data parallelization technique.
 
-OpenMP~\cite{} is a language based on directives added to C/C++ or Fortran.
+OpenMP~\cite{660313} is a language based on directives added to C/C++ or Fortran.
 Those directives offers a way to take advantage of shared memory architectures such as multi-processors, multi-cores, many-cores or even GPUs (in OpenMP 4.0).
 Among other available directives, some of them offer the possibility to manage a pool of threads and to produce a task parallel program.
-In this paper OpenMP is used in its version 3.0, which means that we do not take advantage of the dynamic schedulers available for OpenMP 4.0 such as libgomp~\cite{}, StarPU~\cite{}, or XKaapi~\cite{}.
+In this paper OpenMP is used in its version 3.0, which means that we do not take advantage of the dynamic schedulers available for OpenMP 4.0 such as libgomp~\footnote{\url{https://gcc.gnu.org/projects/gomp/}}, StarPU~\cite{Augonnet2011}, or XKaapi~\cite{Gautier:2013:XRS:2510661.2511383}.
 Those dynamic schedulers will be the subject of future work.
@@
 The dump step of the MSL compiler automatically generates an hybrid code which uses the SkelGIS Cartesian distributed data structure and an OpenMP static scheduling generated from the algorithm previously described in Section~\ref{sect:tsp}.
-To facilitate the maintainability of this hybrid code, a component-based programming model is used, the Low Level Component Model~\cite{}.
+To facilitate the maintainability of this hybrid code, a component-based programming model is used, the Low Level Component Model~\cite{l2c} (\llc).
 
@@
 One can notice that MSL could use different languages as it is mesh-agnostic.
-For example, a future work is to use the distributed unstructured mesh proposed by PamPA~\cite{}.
+For example, a future work is to use the distributed unstructured mesh proposed by PaMPA~\cite{lachat:hal-00768916}.
 Moreover, MSL could use pthreads instead of OpenMP to manage the static scheduling.

### 2016-05-27 ARTICLES/IJPP/article/evaluation.tex
IJPP article biblio

@@
 This language, as already described, offers a distributed Cartesian mesh and programming interfaces to use it as a sequential data structure.
-It has been evaluated itself compared to an MPI implementation on the same FullSWOF2D simulation~\cite{}.
+It has been evaluated itself compared to an MPI implementation on the same FullSWOF2D simulation~\cite{CPE:CPE3494}.
 What MSL brings to SkelGIS is an automatic detection of where MPI synchronizations are needed during the numerical simulation, and how to make a fusion of some computations in order to reduce cache misses.
 This section gives a performance comparison of the code produced by MSL for FullSWOF2D, and of the equivalent manual SkelGIS program (with manual synchronizations and fusion choices).
-Moreover, as the code produced by MSL is a component-based code, this evaluation also shows that no overheads are introduced \llc~\cite{}.
+Moreover, as the code produced by MSL is a component-based code, this evaluation also shows that no overheads are introduced \llc~\cite{l2c}.
 Following evaluations have been performed onto the machine of the TGCC Curie described in Table~\ref{tab:TGCC}.

### 2016-05-27 ARTICLES/IJPP/article/introduction.tex
IJPP article introduction

Modern high performance architectures are becoming more and more heterogeneous and difficult to use. The current world's top supercomputer Tianhe-2~\footnote{url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a peak performance of thirty floating point operations par seconds. Heterogeneity and complexity of architectures will probably continue to increase in future.

One the big challenges to solve to be able to use exascale computers is to propose programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}. Actually, applications which runs on supercomputers and which needs such a computation power are physics, weather or genomic applications, which are not implemented by HPC specialists most of the time.

Many low level execution models for exascale computers focuss onto the dynamic scheduling of task graphs combined to message passing models to be able to use more than one machine~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011}. Those models increase HPC code portability and reach an interesting performance onto heterogeneous architectures. At a higher abstraction level, general purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} follow the direction of task graph scheduling proposed by those execution models. However, for non-experts end-users, general purpose languages still are difficult to use and to tune for a given application onto a given architecture. The current easiest (with the higher abstraction level), but still efficient, programming model is Domain Specific Language (DSL). Such a language is specific to the end-user domain and proposes a grammar which is easy to understand and completely sequential. The DSL compiler is able, by knowing the specific domain targetted, to automatically apply parallelization and specific optimizations to produce a high performance back-end code. Thus, a DSL is able to split end-user concerns from HPC concerns which is relevant to reach exascale programming models.

However, many DSLs have been proposed for many domains yet, but none of them has been able to reuse work already done by another language. In other words, software engineering properties have to be integrated into DSL conception, such that a new DSL can be seen as a composition of parallelization, optimizations or even languages semantics already proposed by others.

For example, to numerically solve a set of partial differential equations (PDEs), iterative methods are frequently used to approximate the exact solution through a discretized phenomena. A very well known and usual way to discretize PDEs is to transform them to explicit numerical schemes, also often called \emph{stencils}. Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107}, as it will be described in Section~\ref{sect:rel}. Many of them use same kind of parallelization, data structures or optimizations, however each one has been built from scratch to deal with another additionnal specific case.

We present the Multi-Stencil Language (MSL) DSL for stencil-based numerical simulations. MSL is a language with a light grammar to only describe such a numerical simulation without implementation details. From the description, the compiler has enough information to extract and build an empty parallel pattern of the simulation, which can be filled, in a second step, by implementation concerns. Te parallel pattern generated by the language is able to use different existing languages and libraries as it is independent from implementation choices. Moreover, the parallelization performed by the language is large enough to be compatible with many architectures and back-end languages. Contributions presented in this paper are : the computational model of a multi-stencil program and its parallelization formalism; the MSL grammar and its compiler; and a performance evaluation of a real case numerical simulation up to 16.384 cores.

Section~\ref{sect:rel} introduces related work on stencils DSLs. Section~\ref{sect:formalism} formally explains the targetted domain and its computational model. From this model can be extracted the grammar of MSL in Section~\ref{sect:msl}. Section~\ref{sect:parallelism} shows how parallelism can be extracted from this light grammar of MSL. Sections~\ref{sect:msp} and~\ref{sect:comp} detail choices that have been done in this paper to evaluate MSL, and Section~\ref{sect:eval} shows performance results of the language. Finally, Section~\ref{sect:concl} concludes and proposes perspectives on this work.

### 2016-05-28 ARTICLES/IJPP/article/article.tex
Updated "Time, mesh, data"

--- 2016-05-27 ARTICLES/IJPP/article/article.tex
@@
 \usepackage{amsmath}
-\usepackage{amssymb}
+\usepackage{amssymb, stmaryrd}
 \usepackage{subfig}
@@
 \usepackage{cite}
+\usepackage{enumitem}
 \usepackage{filecontents}
@@
 \item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+\newenvironment{mydefs}[1][Definitions]{\par\medskip\noindent\textbf{#1}\nopagebreak\begin{itemize}[noitemsep,topsep=.25\topsep]}{\end{itemize}}
 
@@
 %------------------------------------------------------------------------------
-\section{Related work}
-\label{sect:rel}
-\input{related.tex}
-%------------------------------------------------------------------------------
 \section{Computational model of Multi-Stencil Programs}
@@
 %------------------------------------------------------------------------------
+\section{Related work}
+\label{sect:rel}
+\begin{itemize}
+\item stencil compilers cartesian : Pochoir, PATUS for optimizations/parallelization of a single stencil kernel
+\item stencil compilers unstructured meshes : Lizst for optimizations/parallelization of a single stencil kernel
+\item Halide and Nabla for pipeline of stencil computations in image (Cartesian) = mesh dependent
+\item Reuse of external solutions as SkelGIS, GA, Lizst, Pochoir etc.
+\end{itemize}
+%------------------------------------------------------------------------------
 \section{Conclusion}

### 2016-05-28 ARTICLES/IJPP/article/formalism.tex

@@
-To numerically solve a set of PDEs, iterative methods (finite difference, finite volume or finite element methods) are frequently used to approximate the solution through a discretized (step by step) phenomena.
-Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied onto a mesh (space discretization).
-In other words, the PDEs are transformed to a set of numerical computations applied at each time step on elements of the discretized space domain.
-Among those numerical computations are found a set of numerical schemes, also called \textit{stencil computations}, a set of auxiliary computations also called \emph{local computations}, and finally a set of \emph{reductions} to reduce a bunch of values to a single scalar value.
-In the rest of this paper, we denote a \emph{stencil kernel} as a discretized explicit numerical schemes which involve neighborhood computations, whatever the type of space discretization is (mesh).
+Some approaches to numerically solve partial differential equations (PDEs) are based on direct iterative methods (\textit{e.g.} finite difference, finite volume or finite element methods).
+They approximate the solution through a discretized process where the continuous time and space domains are discretized and numerical computations are iteratively (time discretization) applied onto a mesh (space discretization).
+%In other words, the PDEs are transformed to a set of numerical computations applied at each time step on elements of the discretized space domain.
+While the computations can have various forms, we focus on three categories.
+\emph{Stencil} computations involve access to neighbor values only (the concept of neighborhood depending on the space discretization used).
+\emph{Local} computations depend on the computed location only (this can be seen as a stencil of size one).
+Finally, \emph{reductions} enable to transform values mapped on the mesh to a single scalar value.
 
 This section gives a complete formal description of what we call a \textit{stencil program} and its computations.
-In this paper, the proposed language MSL (Multi-Stencil Language) does not depend on the type of space discretization used.
-Thus, some details given in the following sections, are usefull to understand the domain but do not have to be kept in mind to understand the Multi-Stencil Language.
-As a result, at the end of each subsection a summary of what is usefull to understand the language is given.
+The proposed language MSL (Multi-Stencil Language) does not depend on the type of space discretization used.
+Thus, some details given in the following sections, are useful to understand the domain but do not have to be kept in mind to understand the Multi-Stencil Language.
+% As a result, at the end of each subsection a summary of what is useful to understand the language is given.
 
@@
 
-$\Omega=\mathbb{R}^n$ is the continuous space domain of a numerical simulation.
+$\Omega$ is the continuous space domain of a numerical simulation (typically $\mathbb{R}^n$).
 %For example, if $n=3$, any geographic point on earth could be represented in this space in a given geographic coordinate system.
@@
 For example, an hybrid mesh can be defined as an unstructured mesh composed itself of a Cartesian mesh inside each of its vertices.
-Multi-grid numerical methods is another example.
-However, in this paper a single mesh simulation is addressed.
+% Multi-grid numerical methods is another example. %JB: pas vraiment...
+However in this paper, only single mesh simulations are addressed.
 
 \medskip
-\noindent \textbf{Definitions (mesh)}
-%\begin{mydef}
-\begin{itemize}
-\item An entity $\phi$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $\phi\subset V\cup E$.
-\item A group of mesh entities, denoted $\mathcal{G}$, is a set of mesh entities which are composed of the same topology and types of elements in $V\cup E$.
-\item The set of mesh entities groups of a simulation is denoted $\Phi$.
-\item Finally, the mesh on which is applied a group of mesh entities $\mathcal{G}$ is denoted $mesh(\mathcal{G})$
-\end{itemize}
-%\end{mydef}
 
-For example, in a 2D Cartesian mesh an entity could be a cell, made of four vertices and four edges, or simply a vertex.
-As a result, a group of cells and a group of vertices could be defined in $\Phi$.
-%This type can be defined as the sets containing exactly four vertices and four edges connected as a cycle.
-Another type of entities simply are the vertices ($V$) and can be defined as all singletons formed of a single vertex of $V$.
+\begin{mydefs}[Definitions (mesh)]
+\item An \textit{entity} $\phi$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $\phi\subset V\cup E$.
+\item A \textit{group of mesh entities} denoted $\mathcal{G} \in \mathcal{P}(V\cup E)$ represents a set of entities of the same topology and the mesh of a group is denoted $mesh(\mathcal{G})$, \textit{i.e.} $\mathcal{G} \in \mathcal{P}(V\cup E) \Leftrightarrow mesh(\mathcal{G})=(V,E)$
+\item The \textit{set of entities groups} used in a simulation is denoted $\Phi$.
+\end{mydefs}
+
+For example, in a 2D Cartesian mesh, a group of entities made of four vertices and four edges each could form the cells.
+Another group of entities are the vertices defined as all singletons formed of a single vertex of $V$.
+Both groups would then be part of $Phi$.
 
 \medskip
-This covers the space discretization, however there is also the time dimension which has to be discretized to perform a numerical simulation.
 
-\begin{mydef}
-We denote a scalar as a numerical variable. The set of scalars is denoted $\mathcal{S}$.
+\begin{mydef}%[Definitions (time)]
+The finite sequence $T: (t_n)_{n\in\llbracket 0, Tmax \rrbracket}$ represents the discretization of the continuous time domain $\mathcal{T}=\mathbb{R}$.
+To each discrete time step $n$, it associates a time value $t_n\in\mathcal{T}$.
 \end{mydef}
 
-\medskip
-\noindent \textbf{Definitions (time)}
-%\begin{mydef}
-\begin{itemize}
-\item $\mathcal{T}=\mathbb{R}$ is the continuous time domain of a numerical simulation.
-\item The discretization of the continuous time domain $\mathcal{T}$ is denoted as the pair $T(\Delta t,conv)$.
-\begin{itemize}
-\item $\Delta t \in \mathcal{S}$ represents the time interval of the numerical simulation, such that for a current time iteration $t_i$, the next time iteration is $t_{i+1} = t_i + \Delta t$.
-\item $conv:\mathcal{S}^n \rightarrow bool$ is a function which returns a boolean from a set of scalar variables. This function represents the convergence criteria of the simulation. 
-\end{itemize}
-\end{itemize}
-
-At a given time step, the convergence criteria is evaluated such that if $conv$ returns $true$ the next time step $t_{i+1} = t_i + \Delta t$ can start.
-Otherwise the simulation ends.
-For example, the simplest $conv$ is typically to have two scalars: $t$ the time, and a constant number of iterations denoted $it$.
-At each time step $\Delta t=1$, and $conv$ returns the boolean expression $t<it$.
-%\end{mydef}
-
-In a numerical simulation, a scalar can be associated to each entity of the mesh. In such a case, a single variable is used and is denoted a quantity.
-%A quantityas a set of scalars can be applied onto a mesh with a nil dimension, a set of data elements, or quantities, can be applied onto meshes of dimension superior to zero.
-Those quantities represent, as well as scalars, the set of values to compute, or to use, for computations.
+The time discretization can be as simple as a constant time-step with a fixed number of steps.
+It can however also be defined by recursion with a both the time-step and the number of steps depending on the data variables (see definitions below).
 
 \medskip
-\noindent \textbf{Definitions (quantity)}
-%\begin{mydef}
-\begin{itemize}
-\item A quantity represents a variable applied onto a group of mesh entities $\mathcal{G}$.
-\item The set of quantities applied onto the mesh is denoted $\Delta$.
-\item In the rest of this paper, the group of mesh entities on which a quantity $\delta$ is mapped is denoted $entity(\delta)=\mathcal{G}_{\delta}$.
-\end{itemize}
-%\end{mydef}
 
-Another option, closer to the applied mathematics domain, would have been to apply a quantity to each entity of a group $\mathcal{G}$, but also to each time iteration.
-In this case a single quantity would represents all its occurences over time steps.
-This solution could be investigated in future work, however the approach chosen in this work is to let the user be aware of the number of data he is using and what exactly for.
+\begin{mydefs}[Definitions (quantity)]
+\item $\mathbb{V}=\Delta\cup\mathcal{S}$ is the set of \emph{variables} or \emph{quantities}.
+\item $\Delta$ are the \textit{mesh variables} that to each entity of a group and time-step associates a value $\delta: \mathcal{G}\times T\mapsto \mathcal{V}_\delta$ where $V_{\delta}$ is a value type.
+\item The group of entities a quantity is mapped on is denoted $entity(\delta)=\mathcal{G}$.
+\item $\mathcal{S}$ are the \textit{scalar variables} that to each time-step associates a value $s: T\mapsto \mathcal{V}_\delta$ where $V_{\delta}$ is a value type.
+\item Amongst the scalar variables is one specific variable $conv\in\mathcal{S}$, the convergence criteria whose value is $0$ except at the last iteration where it is 1 $\forall t\in \llbracket 0, Tmax-1 \rrbracket conv(t)=1, conv(Tmax)=1$.
+\end{mydefs}
 
-\paragraph{\textbf{Summary}} In this subsection have been presented the general formalism of what is a mesh, its entities and groups of entities, of how the time is discretized, and finally of what is a scalar and a quantity.
+%For example, the simplest $conv$ is typically to have two scalars: $t$ the time, and a constant number of iterations denoted $it$.
+%At each time step $\Delta t=1$, and $conv$ returns the boolean expression $t<it$.
+
+\medskip
+
+This section has presented the general formalism of what is a mesh, its entities and groups of entities, of how the time is discretized, and finally of what is a scalar and a quantity.
 The language presented in this paper, the Multi-Stencil Language, offers a simple way to describe a numerical simulation while being able to automatically extract parallelism.
@@
 \medskip
-\noindent \textbf{Definitions}
-%\begin{mydef}
-\begin{itemize}
+\begin{mydefs}
 \item A computation domain $D$ is a subpart of a mesh entities group, $D \subseteq \mathcal{G} \in \Phi$.
@@
 \item The set of neihborhood functions in a numerical simulation is denoted $\mathcal{N}$.
-\end{itemize}
+\end{mydefs}
 

### 2016-05-28 ARTICLES/IJPP/article/formalism.tex
Minor improvement

@@
 \item An \textit{entity} $\phi$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $\phi\subset V\cup E$.
-\item A \textit{group of mesh entities} denoted $\mathcal{G} \in \mathcal{P}(V\cup E)$ represents a set of entities of the same topology and the mesh of a group is denoted $mesh(\mathcal{G})$, \textit{i.e.} $\mathcal{G} \in \mathcal{P}(V\cup E) \Leftrightarrow mesh(\mathcal{G})=(V,E)$
+\item A \textit{group of mesh entities} $\mathcal{G} \in \mathcal{P}(V\cup E)$ represents a set of entities of the same topology and the mesh of a group is denoted $mesh(\mathcal{G})$, \textit{i.e.} $\mathcal{G} \in \mathcal{P}(V\cup E) \Leftrightarrow mesh(\mathcal{G})=(V,E)$
 \item The \textit{set of entities groups} used in a simulation is denoted $\Phi$.
@@
 
-\begin{mydef}%[Definitions (time)]
+\begin{mydef}[Definitions (time)]
 The finite sequence $T: (t_n)_{n\in\llbracket 0, Tmax \rrbracket}$ represents the discretization of the continuous time domain $\mathcal{T}=\mathbb{R}$.
-To each discrete time step $n$, it associates a time value $t_n\in\mathcal{T}$.
+To each discrete time step $n\in\llbracket 0, Tmax \rrbracket$, it associates a time value $t_n\in\mathcal{T}$.
 \end{mydef}
@@
 \item $\Delta$ are the \textit{mesh variables} that to each entity of a group and time-step associates a value $\delta: \mathcal{G}\times T\mapsto \mathcal{V}_\delta$ where $V_{\delta}$ is a value type.
-\item The group of entities a quantity is mapped on is denoted $entity(\delta)=\mathcal{G}$.
+\item The group of entities a variable is mapped on is denoted $entity(\delta)=\mathcal{G}$.
 \item $\mathcal{S}$ are the \textit{scalar variables} that to each time-step associates a value $s: T\mapsto \mathcal{V}_\delta$ where $V_{\delta}$ is a value type.
@@
 Its aim is to be mesh-agnostic, thus to not have details about the topology of the mesh which is an implementation problem.
-As a result the details about the mesh topology is not needed in the rest of this paper.
+As a result the details about the mesh topology are not needed in the rest of this paper.
 

### 2016-05-30 ARTICLES/IJPP/article/formalism.tex
Formalism

@@
 They approximate the solution through a discretized process where the continuous time and space domains are discretized and numerical computations are iteratively (time discretization) applied onto a mesh (space discretization).
-%In other words, the PDEs are transformed to a set of numerical computations applied at each time step on elements of the discretized space domain.
+%In other words, the PDEs are transformed to a set of numerical computations applied at each time-step on elements of the discretized space domain.
 While the computations can have various forms, we focus on three categories.
+%TODO[JB] 4 plus tard...
 \emph{Stencil} computations involve access to neighbor values only (the concept of neighborhood depending on the space discretization used).
@@
 The finite sequence $T: (t_n)_{n\in\llbracket 0, Tmax \rrbracket}$ represents the discretization of the continuous time domain $\mathcal{T}=\mathbb{R}$.
-To each discrete time step $n\in\llbracket 0, Tmax \rrbracket$, it associates a time value $t_n\in\mathcal{T}$.
+To each discrete time-step $n\in\llbracket 0, Tmax \rrbracket$, it associates a time value $t_n\in\mathcal{T}$.
 \end{mydef}
@@
 \item $\mathcal{S}$ are the \textit{scalar variables} that to each time-step associates a value $s: T\mapsto \mathcal{V}_\delta$ where $V_{\delta}$ is a value type.
-\item Amongst the scalar variables is one specific variable $conv\in\mathcal{S}$, the convergence criteria whose value is $0$ except at the last iteration where it is 1 $\forall t\in \llbracket 0, Tmax-1 \rrbracket conv(t)=1, conv(Tmax)=1$.
+\item Amongst the scalar variables is one specific variable $conv\in\mathcal{S}$, the convergence criteria whose value is $0$ except at the last step where it is 1 $\forall t\in \llbracket 0, Tmax-1 \rrbracket conv(t)=1, conv(Tmax)=1$.
 \end{mydefs}
@@
 %For example, the simplest $conv$ is typically to have two scalars: $t$ the time, and a constant number of iterations denoted $it$.
-%At each time step $\Delta t=1$, and $conv$ returns the boolean expression $t<it$.
+%At each time-step $\Delta t=1$, and $conv$ returns the boolean expression $t<it$.
 
@@
 
-This section has presented the general formalism of what is a mesh, its entities and groups of entities, of how the time is discretized, and finally of what is a scalar and a quantity.
-The language presented in this paper, the Multi-Stencil Language, offers a simple way to describe a numerical simulation while being able to automatically extract parallelism.
-Its aim is to be mesh-agnostic, thus to not have details about the topology of the mesh which is an implementation problem.
-As a result the details about the mesh topology are not needed in the rest of this paper.
+This section has presented the general formalism of meshes, their entities, groups of entities, time discretization, and finally quantities (mapped on mesh and scalars).
+While the details about mesh and their topologies were useful to specify the following concepts, the Multi-Stencil Language presented in this paper is mesh-agnostic.
+These details will therefore not be needed in the remaining of the paper.
 
@@
 
-In this section are considered four different types of kernel computations, stencil kernels, boundary kernels, local kernels, and reduction kernels.
+In this section four different types of kernel computations are considered: stencil kernels, boundary kernels, local kernels, and reduction kernels.
 
@@
 \item The set of computation domains of a numerical simulation is denoted $\mathcal{D}$.
-\item A neighborhood $n$ is a function which for a given entity $\phi \in \mathcal{G}_i$, returns a set of $m$ entities in $\mathcal{G}_j$, $n : \mathcal{G}_i \rightarrow \mathcal{G}_j^m$.
+\item $\mathcal{N}$ are the neighborhood functions $n: \mathcal{G}_i \mapsto \mathcal{G}_j^m$ which for a given entity $\phi \in \mathcal{G}_i$ returns a set of $m$ entities in $\mathcal{G}_j$.
 One can notice that $i = j$ is possible.
 Most of the time, such a neighborhood is called a \emph{stencil shape}.
-\item The set of neihborhood functions in a numerical simulation is denoted $\mathcal{N}$.
 \end{mydefs}
@@
 \begin{mydef}
-A computation kernel $k$ of a numerical simulation is defined as $k(S,R,(w,D),comp)$, where 
+A computation kernel $k$ of a numerical simulation is defined as $k=(S,R,(w,D),comp)$, where
 \begin{itemize}
 \item $S \in \mathcal{S}$ is the set of scalar to read,
-\item $(w,D) \in \Delta \times \mathcal{D}$ is the single data modified by the computation kernel onto a given computation domain.
-Thus, $w \in \Delta$ and $D \in \mathcal{D}$ is the computation domain on which $w$ is computed, $D \subseteq \mathcal{G}_w=entity(w)$.
-\item $R$ is the set of tuples $(r,n)$, where $r \in \Delta$ and $n \in \mathcal{N}$ is a neighborhood function such that $n : \mathcal{G}_w \rightarrow entity(r)^m$.
-The neighborhood indicates which entities of $r$ will be read in order to compute a single entity of $w$.
+\item $(w,D) \in \Delta \times \mathcal{D}$ is the single data written by the kernel:
+\begin{itemize}
+  \item $w$ is the single quantity modified by the computation kernel,
+  \item $D$ is the computation domain on which $w$ is computed, $D \subseteq entity(w)$,
+\end{itemize}
+\item $R \in \Delta \times \mathcal{N}$ is the set of tuples $(r,n)$ representing the data read where
+\begin{itemize}
+  \item $r$ is a quantity read by the kernel to compute $w$,
+  \item $n : entity(w) \rightarrow entity(r)^m$ is a neighborhood function that indicates whose values of $r$ are read to compute $w$.
+\end{itemize}
 \item $comp$ is the numerical computation which returns a value from a set of $n$ input values, $comp: V^n \rightarrow V$, where $V$ is typically $\mathbb{N}$, $\mathbb{Z}$, or $\mathbb{R}$.
@@
 
-At each time iteration, a set of computations is performed.
+At each time-step, a set of computations is performed.
 During a computation kernel, it can be considered that a set of old states ($t-1$) of quantities are read ($R$), and that a new state ($t$) of a single quantity is written ($w$).
-In this paper only explicit numerical schemes are considered, thus if we denote by $identity$ the identity function, we have $w \not\in R$ except if $w=(r,identity)$.
+%TODO[jb] introduire la contrainte plus tard
 
-Such a definition of a computation kernel offers a large panel of different computations.
+Such a definition of a computation kernel covers a large panel of different computations.
 For example, the four usual types of computations (stencil, local, boundary and reduction) performed into a simulation can be defined as follow :
@@
 \end{itemize}
+Since we only consider explicit numerical schemes in this paper, a kernel can not write the same quantity it reads, except locally, \textit{i.e.} $(w,n) \in R\Rightarrow n=identity$
+
+%In this paper only explicit numerical schemes are considered, thus if we denote by $identity$ the identity function, we have $w \not\in R$ except if $w=(r,identity)$.
 
@@
 Computations for which this property is authorized are boundary kernels only.
-This type of kernel is a particular case as it commputes boundary conditions, which represents what should happened outside the space domain but still impact stencil kernels at the next time step.
+This type of kernel is a particular case as it commputes boundary conditions, which represents what should happened outside the space domain but still impact stencil kernels at the next time-step.
 
@@
 
-\paragraph{\textbf{Summary}} In this subsection have been presented the general formalism of what is a computation kernel and a multi-stencil program.
-The language presented in this paper, the Multi-Stencil Language, offers a simple way to describe a numerical simulation while being able to automatically extract parallelism.
-Its aim is to be mesh-agnostic, as previously described, but also to make a clear separation of concerns between the simulation description and its implementation.
-As a result, and as already described, details about the topology of the mesh, which is an implementation problem, is not used into the language.
-Moreover, in a computation kernel $k(S,R,(w,D),comp)$, $comp$ is not used by the language.
+% This section has presented the general formalism of what is a computation kernel and a multi-stencil program.
+% The language presented in this paper, the Multi-Stencil Language, offers a simple way to describe a numerical simulation while being able to automatically extract parallelism.
+% Its aim is to be mesh-agnostic, as previously described, but also to make a clear separation of concerns between the simulation description and its implementation.
+% As a result, and as already described, details about the topology of the mesh, which is an implementation problem, is not used into the language.
+% Moreover, in a computation kernel $k(S,R,(w,D),comp)$, $comp$ is not used by the language.
 

### 2016-05-30 ARTICLES/IJPP/article/introduction.tex
IJPP article introduction

@@
-Modern high performance architectures are becoming more and more heterogeneous and difficult to use.
-The current world's top supercomputer Tianhe-2~\footnote{url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a peak performance of thirty floating point operations par seconds.
-Heterogeneity and complexity of architectures will probably continue to increase in future.
+As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
+For example, the current world's top supercomputer Tianhe-2~\footnote{url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance around thirty floating point operations par seconds.
+However, to be able to use such a machine, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA etc., and multiple optimization techniques, such as cache optimizations or vectorization optimizations, have to be combined.
+Moreover, current architectures evolution lets think that heterogeneity and complexity in HPC will continue to grow in future.
 
 One the big challenges to solve to be able to use exascale computers is to propose programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}.
-Actually, applications which runs on supercomputers and which needs such a computation power are physics, weather or genomic applications, which are not implemented by HPC specialists most of the time.
+Actually, applications which run on supercomputers and which need such a computation power are physics, weather or genomic applications, which are not implemented by HPC specialists most of the time.
 
-Many low level execution models for exascale computers focuss onto the dynamic scheduling of task graphs combined to message passing models to be able to use more than one machine~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011}.
-Those models increase HPC code portability and reach an interesting performance onto heterogeneous architectures.
+Many low level execution models focuss onto the dynamic scheduling of task graphs combined to message passing models to be able to use more than one machine~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359}.
+Those models increase HPC code portability and reach an interesting performance onto heterogeneous architectures, which is interesting to reach exascale programming.
 At a higher abstraction level, general purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} follow the direction of task graph scheduling proposed by those execution models.
 However, for non-experts end-users, general purpose languages still are difficult to use and to tune for a given application onto a given architecture.
-The current easiest (with the higher abstraction level), but still efficient, programming model is Domain Specific Language (DSL).
-Such a language is specific to the end-user domain and proposes a grammar which is easy to understand and completely sequential.
-The DSL compiler is able, by knowing the specific domain targetted, to automatically apply parallelization and specific optimizations to produce a high performance back-end code.
+The current easiest (with the higher abstraction level), but still efficient, programming model for end-users is Domain Specific Language (DSL).
+Such a language is specific to the end-user domain and proposes a grammar which is easy to understand.
+The DSL compiler is able, because of the specific knowledge onto the domain targetted, to automatically apply parallelization and specific optimizations to produce a high performance back-end code.
 Thus, a DSL is able to split end-user concerns from HPC concerns which is relevant to reach exascale programming models.
 
-However, many DSLs have been proposed for many domains yet, but none of them has been able to reuse work already done by another language.
+However, many DSLs have been proposed for many domains yet, and, as far as we know, only a few of them have been able to reuse work already done by another language~\cite{Sujeeth:2013:CRC:2524984.2524988}.
 In other words, software engineering properties have to be integrated into DSL conception, such that a new DSL can be seen as a composition of parallelization, optimizations or even languages semantics already proposed by others.
@@
 A very well known and usual way to discretize PDEs is to transform them to explicit numerical schemes, also often called \emph{stencils}.
-Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107}, as it will be described in Section~\ref{sect:rel}.
+Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107}, as it will be detailed in Section~\ref{sect:rel}.
 Many of them use same kind of parallelization, data structures or optimizations, however each one has been built from scratch to deal with another additionnal specific case.
-
-We present the Multi-Stencil Language (MSL) DSL for stencil-based numerical simulations.
-MSL is a language with a light grammar to only describe such a numerical simulation without implementation details.
+We present the Multi-Stencil Language (MSL) DSL, also for stencil-based numerical simulations.
+MSL is a language with a light grammar to describe a numerical simulation without implementation details.
 From the description, the compiler has enough information to extract and build an empty parallel pattern of the simulation, which can be filled, in a second step, by implementation concerns.

### 2016-05-30 ARTICLES/IJPP/article/introduction.tex
IJPP article related

@@
 From the description, the compiler has enough information to extract and build an empty parallel pattern of the simulation, which can be filled, in a second step, by implementation concerns.
-Te parallel pattern generated by the language is able to use different existing languages and libraries as it is independent from implementation choices.
+The parallel pattern generated by the language is able to use different existing languages and libraries as it is independent from implementation choices.
 Moreover, the parallelization performed by the language is large enough to be compatible with many architectures and back-end languages.

### 2016-05-30 ARTICLES/IJPP/article/msl.tex
MSL++

--- 2016-01-29 ARTICLES/IJPP/article/msl.tex
@@
+From the formalism detailed in the previous section, the Multi-Stencil Language and its grammar can already be given.
+This grammar is sufficient to automatically extract parallelism from the program as will be explained in the next section.
+
 \begin{filecontents*}{grammar.txt}
 program ::= "mesh:" meshid 
-            "mesh entities:" listmeshent
-            "computation domains:" 
-                      listcompdom
-            "independent:"
-                      listinde
-            "stencil shapes:"
-                      liststencil
-            "data:" listdata
+            "mesh entities:" listgroup
+            "computation domains:" listcompdom
+            "independent:" listinde
+            "stencil shapes:" liststencil
+            "quantities:" listquantities
             "scalar" listscalar
@@
 
-listmeshent ::= meshent listmeshent 
-                        |  meshent
-listcompdom ::= compdom listcompdom 
-                        |  compdom
-compdom ::= compdomid "in" listmeshent
+listgroup ::= groupid "," listgroup | groupid
+listcompdom ::= compdom listcompdom | compdom
+compdom ::= compdomid "in" groupid
 listinde ::= inde listinde | inde
 inde ::= compdomid "and" compdomid
-liststencil ::= stencil liststencil 
-                        | stencil
-stencil ::= stencilid "from" meshent "to" meshent
-listdata ::= data listdata |  data
-data ::= dataid "," meshent
-listscalar ::= scalar listscalar 
-                      | scalar
-
+liststencil ::= stencil liststencil | stencil
+stencil ::= stencilid "from" groupid "to" groupid
+listquantities ::= quantity listquantities |  quantity
+quantity ::= groupid listquantityid
+listquantityid ::= quantityid "," listquantityid |  quantityid
+listscalar ::= scalarid "," listscalar | scalarid
 listloop ::= loop listloop | loop
@@
           "computations:" listcomp
-
-iteration ::= num | criteria
-criteria ::= criteriaid "(" listscalarread ")"
-listscalarread ::= scalar listscalarread 
-                          |  scalar
+iteration ::= num | scalar
 listcomp ::= comp listcomp |  comp
-comp ::= dataid "[" compdomid "]=" 
-                    compid "({" 
-                    listscalarread "},{" 
-                    listdataread "})"
-listdataread ::= dataread listdataread 
-                          |  dataread
-dataread ::= dataid "[" stencil "]" 
-              |  dataid
+comp ::= quantityid "[" compdomid "]" "=" compid "(" listread ")"
+listread ::= dataread listread | dataread
+dataread ::= quantityid "[" stencilid "]" |  quantityid | scalar
 \end{filecontents*}
 
-\begin{figure}[!h]
+\begin{figure}[hbt]
   \hspace{5mm}
-  \begin{minipage}[!h]{0.5\textwidth}
+  \begin{minipage}[!h]{0.98\textwidth}
     {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left]{grammar.txt}}   
@@
 \end{figure}
+
+The grammar of the Multi-Stencil Language is given in Figure~\ref{fig:grammar} and an example is provided in Figure~\ref{fig:exmsl}.
+A Multi-Stencil program is composed of eight parts.
 
@@
 computations :
-  B[d1] = k0({tau},{A})
-  C[d2] = k1({},{B[n2]})
-  D[d1] = k2({},{C})
-  E[d1] = k3({},{C})
-  F[d1] = k4({},{D,C[n3]})
-  G[d1] = k0({mu,tau},{E})
-  H[d2] = k6({},{F})
-  I[d1] = k7({},{G,H})
-  J[d1] = k8({mu},{I[n1]})
+  B[d1] = k0(tau,A)
+  C[d2] = k1(B[nec])
+  D[d1] = k2(C)
+  E[d1] = k3(C)
+  F[d1] = k4(D,C[nce])
+  G[d1] = k0(mu,tau,E)
+  H[d2] = k6(F)
+  I[d1] = k7(G,H)
+  J[d1] = k8(mu,I[ncc])
 \end{filecontents*}
@@
 \end{figure}
+
+\begin{enumerate}
+
+\item The \texttt{mesh} keyword (Fig.\ref{fig:grammar},~l.1) introduces a identifier for $\mathcal{M}$, the single mesh of the simulation.
+For example \texttt{cart} in Fig.\ref{fig:exmsl},~l.1.
+Since the language is independent of the mesh topology, this is not used by the compiler and present for reference only.
+
+\item The \texttt{mesh entities} keyword (Fig.\ref{fig:grammar},~l.2) introduces identifiers for the groups of entities $\mathcal{G}\in\Phi$.
+For example \texttt{cell} or \texttt{edgex} in Fig.\ref{fig:exmsl},~l.2.
+Again, this is for reference only as the compiler does not rely on the mesh topology.
+
+\item The \texttt{computation domains} keyword (Fig.\ref{fig:grammar},~l.3) introduces identifiers for the computation domains $D\in\mathcal{D}$.
+For example \texttt{d1} and \texttt{d2} in Fig.\ref{fig:exmsl},~l.4-5.
+For reference, each domain is associated to a group of entities (Fig.\ref{fig:grammar},~l.15) such as \texttt{cell} for \texttt{d1} in Fig.\ref{fig:exmsl},~l.4.
+
+\item The \texttt{independent} keyword (Fig.\ref{fig:grammar},~l.5) offers a way to declare that computation domains do not overlap Such as \texttt{d1} and \texttt{d2} in  Fig.\ref{fig:exmsl},~l.7.
+This is used by the compiler to compute dependencies between computations.
+
+\item The \texttt{stencil} keyword (Fig.\ref{fig:grammar},~l.7) introduces identifiers for $n\in\mathcal{N}$, the stencil shapes.
+For each shape, the source and destination group of entities (Fig.\ref{fig:grammar},~l.19) are specified.
+For example \texttt{nec} in Fig.\ref{fig:exmsl},~l.11 associates \texttt{cell} entities to each \texttt{edgex} entity.
+
+\item The \texttt{quantities} keyword (Fig.\ref{fig:grammar},~l.9) introduces identifiers for $\delta\in\Delta$, the quantities with the group of entities they are mapped on (Fig.\ref{fig:grammar},~l.21).
+For example the quantities \texttt{C} and \texttt{H} are mapped on \texttt{edgex} entities.
+
+\item The \texttt{scalar} keyword (Fig.\ref{fig:grammar},~l.10) introduces identifiers for $s\in\mathcal{S}$, the scalars. For example \texttt{mu} and \texttt{tau} in Fig.\ref{fig:exmsl},~l.15.
+
+\item Finally, the last part (Fig.\ref{fig:grammar},~l.11) introduce the computations steps. Each step is made of two parts:
+\begin{itemize}
+\item the \texttt{time} keyword introduces $conv$, the convergence criteria either as a number of iterations (Fig.\ref{fig:grammar},~l.27) or as a function of scalars (Fig.\ref{fig:grammar},~l.28).
+500 iterations are for example specified in Fig.\ref{fig:exmsl},~l.16,
+\item the \texttt{computations} keyword (Fig.\ref{fig:grammar},~l.26) introduces identifiers for the computation $k\in\Gamma=(S,R,(w,D),comp)$.
+\end{itemize}
+
+\item each computation (Fig.\ref{fig:grammar},~l.31) specifies:
+\begin{itemize}
+ \item the quantity $w$ written and its domain $D$, for example in Fig.\ref{fig:exmsl},~l.22, kernel \texttt{k4} computes the quantity \texttt{F} on domain \texttt{d1},
+ \item the read scalars $S$ and mesh quantities with their associated stencil shape $R$, for example in Fig.\ref{fig:exmsl},~l.16, \texttt{k4} reads \texttt{C} with the shape \texttt{nce} and \texttt{D} with the default \textit{identity} shape, it uses no scalar.
+\end{itemize}
+
+\end{enumerate}
+
+One can notice that in the example of Figure~\ref{fig:exmsl}, there are no (reduction) kernel associated to the scalars.
+In this case, the scalars are in fact constants.
+One can also notice that the computation to execute for each kernel is not specified.
+This is indeed not handled by MSL that only generates a skeleton of the application.
+It can then be specified elsewhere by referencing the identifier chosen here.
+
+% To give a better idea of a multi-stencil program, a short example is given in Figure~\ref{fig:exmsl}.
+In this example, the mesh is called \textit{cart}.
+Two groups of mesh entities are defined \textit{cell} and \textit{edgex}.
+As previously explained, it is not needed in the grammar to give details on the topology of the mesh and the group of mesh entities.
+Two computation domains are given, \textit{d1} is a subpart of the entities of the group \textit{cell}, and \textit{d2} is a subpart of the entities of the group \textit{edgex}.
+It is declared that \textit{d1} and \textit{d2} are independant.
+Three stencil shapes are given.
+The second one for example \textit{nce} returns mesh entities of the group \textit{edgex} from an entity of the group \textit{cell}.
+Eight quantities are applied onto the group \textit{cell}, and two onto the group \textit{edgex}.
+Two scalars are defined, \textit{mu} and \textit{tau}.
+The time loop will iterate 500 times.
+Finally, nine computations are declared.
+For example, the computation \textit{k8} write the quantity \textit{J} onto the computation domain \textit{d1} by reading the scalar \textit{mu} and the quantity \textit{I} with a stencil shape \textit{n1}.
+
+% The aim of the Multi-Stencil Language is to offer a way to describe a numerical simulation based on explicit numerical schemes.
+The language is mesh-agnostic as no information is asked to the user about the actual topology of the mesh.
+Moreover, information that is usefull for the implementation is also split from the MSL description.
+For this reason, the language grammar is simple.
+
+This section has introduced the grammar of MSL.
+The next section will show how parallelization can be extracted from this simple language and how an empty parallel skeleton of the application can be generated.
+%, introducing implementation concerns in a second phase.

### 2016-05-30 ARTICLES/IJPP/article/related.tex

@@
 Each one has its own specificities and answers to a specific stencil case or to a specific additionnal optimization.
-For example, Pochoir~\cite{spaaTangCKLL11} works on cache optimization techniques for stencils applied onto Cartesian meshes.
+For example, applied on Cartesian meshes, Pochoir~\cite{spaaTangCKLL11} works on cache optimization techniques for stencils applied onto Cartesian meshes.
 On the other hand, PATUS~\cite{citeulike12258902} proposes to add a parallelization strategy grammar to its stencil language to perform an auto-tuning parallelization strategy.
-PATUS also works for stencils onto Cartesian meshes.
-Moreover, Halide~\cite{Ragan-Kelley:2013:HLC:2491956.2462176} works for Cartesian meshes, and proposes an optimization and parallelization of a pipeline of stencil codes.
-Finally, ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} is specific to multigrid numerical methods.
+Moreover, Halide~\cite{Ragan-Kelley:2013:HLC:2491956.2462176}  proposes an optimization and parallelization of a pipeline of stencil codes, while ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} is specific to multigrid numerical methods etc.
 
-The reason why each of those languages has its own particularities and its own optimizations is because each language is built and thought as a single block, which makes impossible or difficult code reuse from one language to another.
-Thus, each language starts from scratch to answer its own constraints.
+The reason why each of those languages is implemented from scratch to answer its own particularities and its own optimizations is because each language is built and thought as a single block, which makes impossible or difficult code reuse from one language to another.
 In other words domain specific languages (for stencils or other domains) suffer from a lack of software engineering properties, which could increase the productivity to build a new language, by code-reuse, and also the maintainability of languages, with more separation of concerns.
-The aim of the Mulsi-Stencil Language (MSL), presented in this paper, is to offer more separation of concerns, by splitting the description of a stencil-based numerical simulation, from its implementation concerns.
-As a result, MSL is mesh-agnostic and is able to reuse existing other languages (SkelGIS and OpenMP in this paper).
+A few works propose frameworks to create DSL, which afterwards will be easier to compose and combine together~\cite{Sujeeth:2013:CRC:2524984.2524988}.
+The Mulsi-Stencil Language (MSL), presented in this paper, is itself a domain specific language for stencil-based simulations.
+It offers a way to give a mesh-agnostic description of a simulation, which could be common to different cases of simulations, and thus facilitates reuse of existing underlying data structures, optimizations or parallelization techniques.
+In other words, MSL extracts where parallelization and optimizations are needed into the simulation, by producing an empty parallel pattern, but not how to do it, which is left to existing languages (SkelGIS and OpenMP in this paper), and which is an implementation concern.
 
@@
 However, two main differences can be noticed.
+
 First, the formalism is more flexible in MSL.
@@
 For example, many computations performed in the simulation studied in this paper have to be applied onto a subpart of the overall space domain.
+
 Second, in Liszt and Nabla the description of a simulation is not splitted from its implementation concerns.
-In other words, the topology of a mesh as well as numerical codes are given at the same time than the description of the different computations to apply into the simulation.
+In other words, the topology of a mesh as well as numerical codes are given at the same time than the description of the different computations to apply into the simulation, while MSL splits those two different phases.
+Thus, MSL improves separation of concerns and maintainability of a given simulation.
+Actually, a numerician could describes the different computations to perform into the simulation, while another one, later, could focuss on the implementation and the numerical code of the simulation.
+This also facilitates reuse of other languages.
 
-MSL splits those two different phases of a simulation conception, which improves separation of concerns and maintainability of the simulation.
-As a result, a numerician could describes the different computations to perform into the simulation, while another one, later, could focuss on the implementation and the numerical code of the simulation.
-This also facilitates reuse of other languages.
 The formalism of a Multi-Stencil program, given in this paper, is new, as far as we know.
 Finally, the parallelization techniques proposed in this paper, take place at paradigm level, with data parallelism and hybrid (data plus task) parallelism.
-Thus those parallelization do not need details onto architectures (distributed or shared memories, with or without accelerators).
+Thus the parallel pattern of the simulation does not need details onto parallel architectures (distributed or shared memories, with or without accelerators).
 This makes possible a large panel of backend architectures and languages.

### 2016-05-30 ARTICLES/IJPP/article/formalism.tex
Moved fig

@@
 \paragraph{\textbf{Example}}
-For example, in Figure~\ref{fig:ex1}, assuming that the computation domain (full lines) is denoted $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined as:
-\begin{equation*}
-R: \{(B,n1)\}, \quad w: A, \quad D: dc1,
-\end{equation*}
-\begin{equation*}
-comp: A(x,y)=B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1).
-\end{equation*}
-On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil kernel is defined as:
-\begin{equation*}
-R: \{(C,n2),(A,identity)\}, \quad w: A, \quad D: dc2,
-\end{equation*}
-\begin{equation*}
-comp: A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1).
-\end{equation*}
-
-\begin{figure}
+\begin{figure}[ht]
 \begin{center}
@@
 \end{figure}
+For example, in Figure~\ref{fig:ex1}, assuming that the computation domain (full lines) is denoted $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined as:
+\begin{equation*}
+R: \{(B,n1)\}, \quad w: A, \quad D: dc1,
+\end{equation*}
+\begin{equation*}
+comp: A(x,y)=B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1).
+\end{equation*}
+On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil kernel is defined as:
+\begin{equation*}
+R: \{(C,n2),(A,identity)\}, \quad w: A, \quad D: dc2,
+\end{equation*}
+\begin{equation*}
+comp: A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1).
+\end{equation*}
 

### 2016-05-30 ARTICLES/IJPP/article/msl.tex
Updated lines in refs

@@
 listcomp ::= comp listcomp |  comp
-comp ::= quantityid "[" compdomid "]" "=" compid "(" listread ")"
+comp ::= written "=" compid "(" listread ")"
+written ::= quantityid "[" compdomid "]" | scalar
 listread ::= dataread listread | dataread
@@
 For example \texttt{d1} and \texttt{d2} in Fig.\ref{fig:exmsl},~l.4-5.
-For reference, each domain is associated to a group of entities (Fig.\ref{fig:grammar},~l.15) such as \texttt{cell} for \texttt{d1} in Fig.\ref{fig:exmsl},~l.4.
+For reference, each domain is associated to a group of entities (Fig.\ref{fig:grammar},~l.12) such as \texttt{cell} for \texttt{d1} in Fig.\ref{fig:exmsl},~l.4.
 
-\item The \texttt{independent} keyword (Fig.\ref{fig:grammar},~l.5) offers a way to declare that computation domains do not overlap Such as \texttt{d1} and \texttt{d2} in  Fig.\ref{fig:exmsl},~l.7.
+\item The \texttt{independent} keyword (Fig.\ref{fig:grammar},~l.4) offers a way to declare that computation domains do not overlap Such as \texttt{d1} and \texttt{d2} in  Fig.\ref{fig:exmsl},~l.7.
 This is used by the compiler to compute dependencies between computations.
 
-\item The \texttt{stencil} keyword (Fig.\ref{fig:grammar},~l.7) introduces identifiers for $n\in\mathcal{N}$, the stencil shapes.
-For each shape, the source and destination group of entities (Fig.\ref{fig:grammar},~l.19) are specified.
+\item The \texttt{stencil shapes} keyword (Fig.\ref{fig:grammar},~l.5) introduces identifiers for $n\in\mathcal{N}$, the stencil shapes.
+For each shape, the source and destination group of entities (Fig.\ref{fig:grammar},~l.16) are specified.
 For example \texttt{nec} in Fig.\ref{fig:exmsl},~l.11 associates \texttt{cell} entities to each \texttt{edgex} entity.
 
-\item The \texttt{quantities} keyword (Fig.\ref{fig:grammar},~l.9) introduces identifiers for $\delta\in\Delta$, the quantities with the group of entities they are mapped on (Fig.\ref{fig:grammar},~l.21).
+\item The \texttt{quantities} keyword (Fig.\ref{fig:grammar},~l.6) introduces identifiers for $\delta\in\Delta$, the quantities with the group of entities they are mapped on (Fig.\ref{fig:grammar},~l.16).
 For example the quantities \texttt{C} and \texttt{H} are mapped on \texttt{edgex} entities.
 
-\item The \texttt{scalar} keyword (Fig.\ref{fig:grammar},~l.10) introduces identifiers for $s\in\mathcal{S}$, the scalars. For example \texttt{mu} and \texttt{tau} in Fig.\ref{fig:exmsl},~l.15.
+\item The \texttt{scalar} keyword (Fig.\ref{fig:grammar},~l.7) introduces identifiers for $s\in\mathcal{S}$, the scalars. For example \texttt{mu} and \texttt{tau} in Fig.\ref{fig:exmsl},~l.15.
 
-\item Finally, the last part (Fig.\ref{fig:grammar},~l.11) introduce the computations steps. Each step is made of two parts:
+\item Finally, the last part (Fig.\ref{fig:grammar},~l.8) introduce the computations loops. Each loop is made of two parts:
 \begin{itemize}
-\item the \texttt{time} keyword introduces $conv$, the convergence criteria either as a number of iterations (Fig.\ref{fig:grammar},~l.27) or as a function of scalars (Fig.\ref{fig:grammar},~l.28).
-500 iterations are for example specified in Fig.\ref{fig:exmsl},~l.16,
-\item the \texttt{computations} keyword (Fig.\ref{fig:grammar},~l.26) introduces identifiers for the computation $k\in\Gamma=(S,R,(w,D),comp)$.
+\item the \texttt{time} keyword (Fig.\ref{fig:grammar},~l.22) introduces $conv$, the convergence criteria either as a number of iterations or as ${0,1}$ valued scalar (Fig.\ref{fig:grammar},~l.24), for example 500 iterations are specified in Fig.\ref{fig:exmsl},~l.16,
+\item the \texttt{computations} keyword (Fig.\ref{fig:grammar},~l.23) introduces identifiers for the computation $k\in\Gamma=(S,R,(w,D),comp)$.
 \end{itemize}
 
-\item each computation (Fig.\ref{fig:grammar},~l.31) specifies:
+\item each computation (Fig.\ref{fig:grammar},~l.26) specifies:
 \begin{itemize}
@@
 \end{itemize}
-
 \end{enumerate}

### 2016-05-30 ARTICLES/IJPP/article/formalism.tex
IJPP article formalism + msl modifications

@@
 % Multi-grid numerical methods is another example. %JB: pas vraiment...
-However in this paper, only single mesh simulations are addressed.
+However in this paper, single mesh simulations are addressed.
 
@@
 
-For example, in a 2D Cartesian mesh, a group of entities made of four vertices and four edges each could form the cells.
-Another group of entities are the vertices defined as all singletons formed of a single vertex of $V$.
-Both groups would then be part of $Phi$.
+For example, in a 2D Cartesian mesh, a group of entities where each entity is made of four vertices and four edges could form the cells.
+Another group of entities are the vertices defined as all singletons formed of a single vertex of $V$. Both groups, then, would be part of $\Phi$.
 
@@
 \begin{mydef}[Definitions (time)]
-The finite sequence $T: (t_n)_{n\in\llbracket 0, Tmax \rrbracket}$ represents the discretization of the continuous time domain $\mathcal{T}=\mathbb{R}$.
-To each discrete time-step $n\in\llbracket 0, Tmax \rrbracket$, it associates a time value $t_n\in\mathcal{T}$.
+The finite sequence $T: (t_n)_{n\in\llbracket 0, T_{max} \rrbracket}$ represents the discretization of the continuous time domain $\mathcal{T}=\mathbb{R}$.
+To each discrete time-step $n\in\llbracket 0, T_{max} \rrbracket$, it associates a time value $t_n\in\mathcal{T}$.
 \end{mydef}
@@
 The time discretization can be as simple as a constant time-step with a fixed number of steps.
-It can however also be defined by recursion with a both the time-step and the number of steps depending on the data variables (see definitions below).
+It can, however, also be defined by recursion with both the time-step and the number of steps depending on the data variables (see definitions below).
 
@@
 \item The group of entities a variable is mapped on is denoted $entity(\delta)=\mathcal{G}$.
-\item $\mathcal{S}$ are the \textit{scalar variables} that to each time-step associates a value $s: T\mapsto \mathcal{V}_\delta$ where $V_{\delta}$ is a value type.
-\item Amongst the scalar variables is one specific variable $conv\in\mathcal{S}$, the convergence criteria whose value is $0$ except at the last step where it is 1 $\forall t\in \llbracket 0, Tmax-1 \rrbracket conv(t)=1, conv(Tmax)=1$.
+\item $\mathcal{S}$ are the \textit{scalar variables} that to each time-step associates a value $s: T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
+\item Amongst the scalar variables is one specific variable $conv\in\mathcal{S}$, the convergence criteria whose value is $0$ except at the last step where it is 1.
+Thus, $\forall t\in \llbracket 0, T_{max-1} \rrbracket$, $conv(t)=1$, and $conv(T_{max})=1$.
 \end{mydefs}
@@
 
-In this section four different types of kernel computations are considered: stencil kernels, boundary kernels, local kernels, and reduction kernels.
+%In this section four different types of kernel computations are considered: stencil kernels, boundary kernels, local kernels, and reduction kernels.
 
@@
 \begin{itemize}
-  \item $w$ is the single quantity modified by the computation kernel,
+  \item $w$ is the single quantity (variable) modified by the computation kernel,
   \item $D$ is the computation domain on which $w$ is computed, $D \subseteq entity(w)$,
@@
 \end{itemize}
-\item $comp$ is the numerical computation which returns a value from a set of $n$ input values, $comp: V^n \rightarrow V$, where $V$ is typically $\mathbb{N}$, $\mathbb{Z}$, or $\mathbb{R}$.
-$comp$ represents the actual numerical expression which is computed by a kernel.
+\item $comp$ is the numerical computation which returns a value from a set of $n$ input values, $comp: \mathcal{V}_i^n \rightarrow \mathcal{V}_j$, where $\mathcal{V}_i$ and $\mathcal{V}_j$ are value types.
+Thus, $comp$ represents the actual numerical expression which is computed by a kernel.
 \end{itemize}
@@
 \end{itemize}
-Since we only consider explicit numerical schemes in this paper, a kernel can not write the same quantity it reads, except locally, \textit{i.e.} $(w,n) \in R\Rightarrow n=identity$
+Since we only consider explicit numerical schemes in this paper, a kernel can not write the same quantity it reads, except locally, \textit{i.e.} if $\exists (w,n) \in R\Rightarrow n=identity$
 
@@
 
-A reduction is typically used to compute the convergence criteria of the time loop of the simulation. Occasionally reductions can also be used during a time iteration.
+A reduction is typically used to compute the convergence criteria of the time loop of the simulation. %Occasionally reductions can also be used to compute other scalars.
 
@@
 
+\medskip
 A stencil program has been formally defined in this section. This formalism is used in the next Section to define two parallelization techniques of a multi-stencil program.
 
-One can note that all definitions given in this section are not dependent from the topology of the mesh. This property will be kept in the rest of this paper to propose the mesh-agnostic MSL language.
+One can note that all definitions given in this section are independent from the topology of the mesh. This property will be kept in the rest of this paper to propose the mesh-agnostic MSL language.

### 2016-05-30 ARTICLES/IJPP/article/msl.tex
IJPP article formalism + msl modifications

@@
 
-\item The \texttt{mesh} keyword (Fig.\ref{fig:grammar},~l.1) introduces a identifier for $\mathcal{M}$, the single mesh of the simulation.
+\item The \texttt{mesh} keyword (Fig.\ref{fig:grammar},~l.1) introduces an identifier for $\mathcal{M}$, the single mesh of the simulation.
 For example \texttt{cart} in Fig.\ref{fig:exmsl},~l.1.
-Since the language is independent of the mesh topology, this is not used by the compiler and present for reference only.
+Since the language is independent of the mesh topology, this is not used by the compiler but is syntactic sugar for the user.%and presents for reference only.
 
@@
 For example \texttt{cell} or \texttt{edgex} in Fig.\ref{fig:exmsl},~l.2.
-Again, this is for reference only as the compiler does not rely on the mesh topology.
+Again, this is syntactic sugar only as the compiler does not rely on the mesh topology.
 
@@
 
-\item The \texttt{independent} keyword (Fig.\ref{fig:grammar},~l.4) offers a way to declare that computation domains do not overlap Such as \texttt{d1} and \texttt{d2} in  Fig.\ref{fig:exmsl},~l.7.
+\item The \texttt{independent} keyword (Fig.\ref{fig:grammar},~l.4) offers a way to declare that computation domains do not overlap, such as \texttt{d1} and \texttt{d2} in  Fig.\ref{fig:exmsl},~l.7.
 This is used by the compiler to compute dependencies between computations.
@@
 
-\item Finally, the last part (Fig.\ref{fig:grammar},~l.8) introduce the computations loops. Each loop is made of two parts:
+\item Finally, the last part (Fig.\ref{fig:grammar},~l.8) introduces the computations loops. Each loop is made of two parts:
 \begin{itemize}
-\item the \texttt{time} keyword (Fig.\ref{fig:grammar},~l.22) introduces $conv$, the convergence criteria either as a number of iterations or as ${0,1}$ valued scalar (Fig.\ref{fig:grammar},~l.24), for example 500 iterations are specified in Fig.\ref{fig:exmsl},~l.16,
-\item the \texttt{computations} keyword (Fig.\ref{fig:grammar},~l.23) introduces identifiers for the computation $k\in\Gamma=(S,R,(w,D),comp)$.
-\end{itemize}
-
-\item each computation (Fig.\ref{fig:grammar},~l.26) specifies:
+\item the \texttt{time} keyword (Fig.\ref{fig:grammar},~l.22) introduces $conv$, the convergence criteria either as a number of iterations or as ${0,1}$ valued scalar (Fig.\ref{fig:grammar},~l.24).
+For example, 500 iterations are specified in Fig.\ref{fig:exmsl},~l.16,
+\item the \texttt{computations} keyword (Fig.\ref{fig:grammar},~l.23) introduces identifiers for each computation $k\in\Gamma=(S,R,(w,D),comp)$.
+Each computation (Fig.\ref{fig:grammar},~l.26) specifies:
 \begin{itemize}
@@
  \end{itemize}
+\end{itemize}
 \end{enumerate}
 
-One can notice that in the example of Figure~\ref{fig:exmsl}, there are no (reduction) kernel associated to the scalars.
-In this case, the scalars are in fact constants.
+One can notice that in the example of Figure~\ref{fig:exmsl}, there are no (reduction) kernel associated to the scalars \texttt{mu} and \texttt{tau}.
+In this case, those scalars are in fact constants.
 One can also notice that the computation to execute for each kernel is not specified.
-This is indeed not handled by MSL that only generates a skeleton of the application.
-It can then be specified elsewhere by referencing the identifier chosen here.
+This is indeed not handled by MSL, which is mesh-agnostic, and which only generates a skeleton of the application (a parallel pattern of the application).
+Actually, a numerical computation is naturally mesh-dependent and is left to other languages.
+The numerical computation is specified elsewhere by referencing the identifier chosen here.
 
@@
 This section has introduced the grammar of MSL.
-The next section will show how parallelization can be extracted from this simple language and how an empty parallel skeleton of the application can be generated.
+The next section will show how parallelization can be extracted from this simple language and how an empty parallel skeleton (or pattern) of the application can be generated.
 %, introducing implementation concerns in a second phase.

### 2016-05-30 ARTICLES/IJPP/article/parallelism.tex
IJPP article parallelism.tex modifications

@@
 % Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
+As previously explained, in a computation $k(S,R,(w,D),comp)$, $comp$ is not handled by MSL.
+As a result, in the rest of this paper, and to simplify notations, we denote the same computation $k(S,R,(w,D))$.
 
@@
 \begin{mydef}
-For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronisation} is needed between $k_i$ and $k_j$, denoted $k_i \pprec k_j$, if $\exists (r_j,n_j) \in R_j$ such that $w_i=r_j$ and $n_j\neq identity$ ($k_j$ is a stencil computation).
-The quantity to synchronise is $\{w_i\}$.
+For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronization} is needed between $k_i$ and $k_j$, denoted $k_i \pprec k_j$, if $\exists (r_j,n_j) \in R_j$ such that $w_i=r_j$ and $n_j\neq identity$ ($k_j$ is a stencil computation).
+The quantity to synchronize is $\{w_i\}$.
+\label{def:sync}
 \end{mydef}
@@
 \begin{mydef}
-For $n$ the number of computations in $\Gamma$ and $j<n$, if $\exists (r_j,n_j) \in R_j$ such that $n_j\neq identity$ and for all $i<j$, $k_i \not \pprec k_j$, a \textit{synchronisation} is needed between $k_l$ and $k_j$, where $j<l<n$, denoted $k_l \pprec k_j$, if $w_l=r_j$.
-The quantity to synchronise is $\{w_l\}$.
+For $n$ the number of computations in $\Gamma$ and $j<n$, if $\exists (r_j,n_j) \in R_j$ such that $n_j\neq identity$ and for all $i<j$, $k_i \not \pprec k_j$, a \textit{synchronization} is needed between $k_l$ and $k_j$, where $j<l<n$, denoted $k_l \pprec k_j$, if $w_l=r_j$.
+The quantity to synchronize is $\{w_l\}$.
+\label{def:sync2}
 \end{mydef}
@@
 \begin{equation*}
-k_{i,j}^{sync}(S,R,(w,D),comp), 
+k_{i,j}^{sync}(S,R,(w,D)), 
 \end{equation*}
-where $S=\emptyset$, $R=\{(r,n)\}=\{(w_i,n_j \in \mathcal{N}\}$, $(w,D)=(w_i,\bigcup_{\phi \in D_j} n_j(\phi)))$ and $comp=identity$.
+where $S=\emptyset$, $R=\{(r,n)\}=\{(w_i,n_j \in \mathcal{N}\}$, $(w,D)=(w_i,\bigcup_{\phi \in D_j} n_j(\phi)))$.
 In other words, $w_i$ has to be synchronized for the neighborhood $n_j$ for all entities of $D_j$.
@@
 Actually, the reduction is first applied locally on each subset of entities, on each resource.
-This results in $p$ (number of resources) scalar values.
-Thus, to perform the final reduction, a set of synchronizations are needed to get the final reduction scalar.
+Thus, $p$ (number of resources) scalar values are obtained.
+For this reason, to perform the final reduction, a set of synchronizations are needed to get the final reduced scalar.
 As most parallelism libraries (MPI, OpenMP) already propose a reduction synchronization with its own optimizations, we simply choose to replace the reduction computation by itself anotated by $red$.
@@
 \begin{mydef}
-A reduction kernel $k_j(S_j,R_j,(w_j,D_j),comp)$, where $w$ is a scalar, is replaced by $k^{red}_j(S_j,R_j,(w_j,D_j),comp)$.
+A reduction kernel $k_j(S_j,R_j,(w_j,D_j))$, where $w$ is a scalar, is replaced by $k^{red}_j(S_j,R_j,(w_j,D_j))$.
 %if we denote by $w^r$, $0 \leq r<p$, the local scalar $w$ computed on the resource $r$, a reduction synchronization is defined as the specific computation
@@
 % \end{equation*}
+\label{def:red}
 \end{mydef}
@@
 
-Algorithm~\ref{alg:sync} follows previous definitions to include synchronizations into $\Gamma$.
+Algorithm~\ref{alg:sync} follows previous definitions to build a new ordered list which includes synchronizations.
+In this algorithm, lines 7 to 19 apply Definition~(\ref{def:sync}), lines 20 to 29 apply Definition~(\ref{def:sync2}), and finally lines 34 and 35 apply Definition~(\ref{def:red}).
+Finally, line 39 of the algorithm is the recursive call.
 
@@
 \State $(w,D) = (w_i,\bigcup_{\phi \in D_j} n_j(\phi)))$
-\State $comp = identity$
-\State $list.[k_{i;j}^{sync}(S,R,(w,D),comp)]$
+%\State $comp = identity$
+\State $list.[k_{i;j}^{sync}(S,R,(w,D))]$%,comp)]$
 \EndIf
@@
 
- The final step of this parallelization is to run $\Gamma_{sync}$ on each resource. Thus, for each resource $0 \leq r \leq p-1$ a multi-stencil program, defined by
+ The final step of this parallelization is to run $\Gamma_{sync}$ on each resource. Thus, for each resource $0 \leq r \leq p-1$ the multi-stencil program 
 \begin{equation}
@@
 
-\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl} gives an example of $\mathcal{MSP}$ program. From this example, the following ordered list of computation kernels can be extracted:
+\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl} gives an example of a $\mathcal{MSP}$ program. From this example, the following ordered list of computation kernels can be extracted:
 \begin{equation*}
@@
 \begin{mydef}
-For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{read after write}, denoted $k_i \prec_{raw} k_j$, if $w_i \in R_j$.
+For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{read after write} dependency, denoted $k_i \prec_{raw} k_j$, if $\exists (r_j,n_j) \in R_j$ such that $w_i=r_j$.
 In this case, $k_i$ has to be computed before $k_j$.
-The binary relation $k_i \prec_{raw} k_j$ represents a \textit{dependency}.
 \end{mydef}
@@
 \begin{mydef}
-For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{write after write}, denoted $k_i \prec_{waw} k_j$, if $w_i = w_j$ and $D_i \cap D_j \neq \emptyset$.
+For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{write after write} dependency, denoted $k_i \prec_{waw} k_j$, if $w_i = w_j$ and $D_i \cap D_j \neq \emptyset$.
 In this case, $k_i$ also has to be computed before $k_j$.
-The binary relation $k_i \prec_{waw} k_j$ represents a \textit{dependency}.
 \end{mydef}
@@
 \begin{mydef}
-For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{write after read}, denoted $k_i \prec_{war} k_j$, if $w_j \in R_i$.
+For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{write after read} dependency, denoted $k_i \prec_{war} k_j$, if $\exists (r_i,n_i) \in R_i$ such that $w_j=r_i$.
 In this case, $k_i$ also has to be computed before $k_j$ is started so that values read by $k_i$ are relevant.
-The binary relation $k_i \prec_{war} k_j$ represents a \textit{dependency}.
 \end{mydef}
@@
 One can notice that $\Gamma_{dep}$ is the dependency graph of the computations of a multi-stencil program, but it only takes into account a single time iteration.
-By representing more formally time iterations, a complete dependency graph of the simulation could be built.
+A complete dependency graph of the simulation could be built.
 This is a possible extension of this work.
@@
 \paragraph{\textbf{Example}} Figure~\ref{fig:exmsl} gives an example of $\mathcal{MSP}$ program.
-From $\Gamma_{sync}$ that has been built in Equation~\ref{eq:exsync}, the dependency DAG can be built.
+From $\Gamma_{sync}$ that has been built in Equation~(\ref{eq:exsync}), the dependency DAG can be built.
 For example, as $k_0$ computes $B$ and $k_1$ reads $B$, $k_0$ and $k_1$ becomes vertices of $\Gamma_{dep}$, and an arc $(\overset{\frown}{k_0,k_1})$ is added to $\Gamma_{dep}$.

### 2016-05-30 ARTICLES/IJPP/article/introduction.tex
IJPP article intro

@@
 As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
-For example, the current world's top supercomputer Tianhe-2~\footnote{url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance around thirty floating point operations par seconds.
+For example, the current world's top supercomputer Tianhe-2~\footnote{url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance around thirty peta floating point operations par seconds.
 However, to be able to use such a machine, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA etc., and multiple optimization techniques, such as cache optimizations or vectorization optimizations, have to be combined.

### 2016-05-30 ARTICLES/IJPP/article/introduction.tex
IJPP article intro

@@
 For example, the current world's top supercomputer Tianhe-2~\footnote{url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance around thirty peta floating point operations par seconds.
-However, to be able to use such a machine, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA etc., and multiple optimization techniques, such as cache optimizations or vectorization optimizations, have to be combined.
+However, to be able to use such a machine, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA etc., and multiple optimization techniques, such as cache or vectorization optimizations, have to be combined.
 Moreover, current architectures evolution lets think that heterogeneity and complexity in HPC will continue to grow in future.
 
-One the big challenges to solve to be able to use exascale computers is to propose programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}.
+One of the big challenges to solve to be able to use exascale computers is to propose programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}.
 Actually, applications which run on supercomputers and which need such a computation power are physics, weather or genomic applications, which are not implemented by HPC specialists most of the time.
 
-Many low level execution models focuss onto the dynamic scheduling of task graphs combined to message passing models to be able to use more than one machine~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359}.
+One possible runtime execution model for HPC is to propose dynamic scheduling of task graphs combined to message passing models~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359} (to be able to use more than one machine).
 Those models increase HPC code portability and reach an interesting performance onto heterogeneous architectures, which is interesting to reach exascale programming.
-At a higher abstraction level, general purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} follow the direction of task graph scheduling proposed by those execution models.
+At a highest abstraction level, general purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} follow the direction of task graph scheduling proposed by those execution models.
 However, for non-experts end-users, general purpose languages still are difficult to use and to tune for a given application onto a given architecture.
@@
 
-Section~\ref{sect:rel} introduces related work on stencils DSLs.
+Section~\ref{sect:rel} introduces the related work on DSLs for stencils.
 Section~\ref{sect:formalism} formally explains the targetted domain and its computational model.

### 2016-05-30 ARTICLES/IJPP/article/evaluation.tex
Minor re-wording

@@
 First, the compiler is evaluated.
-Second, the data prallelization technique is evaluated, as well as the fusion optimization, and compared to a full SkelGIS program.
-Third, the hybrid parallelization (data and task) is evaluated and validates the performance model detailed in Section~\ref{sect:tsp}.
+Second, the data parallelization and fusion optimization are evaluated and compared to a full SkelGIS program.
+Third, the hybrid parallelization (data and task) is evaluated to validate the performance model detailed in Section~\ref{sect:tsp}.
 
@@
 Moreover, a level of 16~parallel tasks is reached two times, and five times for the fourth level.
+This means that if two threads are dedicated to task parallelism, two parts of the code at least will not take advantage of this and that no part of the code would benefit from more than 16 threads.
 

### 2016-05-30 ARTICLES/IJPP/article/article.tex
IJPP article conclusion added
IJPP article parallelism.tex modifications

--- 2016-05-27 ARTICLES/IJPP/article/article.tex
@@
 \newtheorem{myprop}[theorem]{Proposition}
+\newtheorem{mydef}[theorem]{Definition}
+\newtheorem{myth}[theorem]{Theorem}
 % \newenvironment{proof}[1][Proof]{\begin{trivlist}
 % \item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-\newenvironment{mydef}[1][Definition]{\begin{trivlist}
-\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-\newenvironment{myth}[1][Theorem]{\begin{trivlist}
-\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+%\newenvironment{mydef}[1][Definition]{\begin{trivlist}
+%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+%\newenvironment{myth}[1][Theorem]{\begin{trivlist}
+%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+\newenvironment{mydefs}[1][Definitions]{\par\medskip\noindent\textbf{#1}\nopagebreak\begin{itemize}[noitemsep,topsep=.25\topsep]}{\end{itemize}}
 
@@
 
-\author{Helene Coullon \and
-        Julien Bigot \and
-        Christian Perez
-}
+\author{H\'el\`ene Coullon \and Julien Bigot \and Christian Perez}
 
@@
 
-\institute{Helene Coullon \and Christian Perez \at
-              INRIA, ENS Lyon, 46 Allée d'Italie, 69007 Lyon\\
+\institute{H\'el\`ene Coullon \and Christian Perez \at
+              INRIA, LIP, ENS Lyon, 46 Allée d'Italie, 69364 Lyon Cedex 07, France\\
               \email{helene.coullon@inria.fr}
@@
 \begin{abstract}
-Insert your abstract here.
-\keywords{First keyword \and Second keyword \and More}
+As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
+One of the big challenges of exascale is to get programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists.
+One relevant solution to ease parallel programming for scientists is Domain Specific Language (DSL).
+However, one problem to avoid with DSLs is to not design a new DSL each time a new domain or a new problem has to be solved.
+This phenomenon happens for stencil-based numerical simulations, for which a large number of languages has been proposed without code reuse between them.
+The Multi-Stencil Language (MSL) presented in this paper is a language common to any kind of mesh used into a stencil-based numerical simulation.
+It is said that MSL is mesh-agnostic.
+Actually, from the description of a numerical simulation, MSL produces an empty parallel pattern, or skeleton, of the simulation which will be filled using other existing parallel languages and libraries.
+Thus, MSL, by finding a common language for different kinds of stencil-based simulation, facilitates code reuse.
+MSL is evaluated on a real case simulation which solves shallow-water equations.
+It is shown that MSL does not introduce overheads on data parallelism up to 16.384 cores, and that the hybrid parallelism (data and task) introduced improves performance of the simulation.
+
+\keywords{Domain Specific Language (DSL) \and stencil \and numerical simulation \and mesh-agnostic \and data parallelism \and task parallelism \and scheduling \and MPI \and OpenMP}
 \end{abstract}
@@
 \label{sect:concl}
+\input{conclusion.tex}
 %------------------------------------------------------------------------------

### 2016-05-30 ARTICLES/IJPP/article/conclusion.tex

We have presented the Multi-Stencil Language (MSL), a domain specific language for stencil-based numerical simulations. Compared to other existing stencil DSLs, MSL proposes a higher abstraction level for end-users. MSL offers a way to describe a numerical simulation independently from the type of mesh used. As a result, the language can be used for a larger number of simulations. It helps to clearly separate the description of the simulation from implementation choices, which facilitates reuse of existing languages and libraries, for example languages for data structures optimizations and parallelization strategies.

In this paper, the formal definition of a multi-stencil program has been given. This formalism helps to understand what is defined into the light grammar of MSL, and also how parallelization can be extracted from it. MSL produces an empty parallel pattern of the numerical simulation described. This pattern indicates where synchronizations are needed, but not how they are performed, which is dependent from the type of mesh and from implementation choices. This second step is left to other existing languages, which are chosen as a back-end of MSL. For example, in this paper the chosen back-end is a code using SkelGIS~\cite{CPE:CPE3494,DBLP:conf/ieeehpcs/HeleneS13,DBLP:conf/europar/CoullonL14}, a templated C++ language for distributed meshes (using MPI), and OpenMP~\cite{660313}.

Experiments presented in this paper show that the back-end code produced by MSL, \ie by a mesh-agnostic language, does not introduce overheads up to 16.384 cores, and that the hybrid (data and task) parallelism automatically introduced by the language, which combines SkelGIS and OpenMP as a back-end, improves performance compared to a data parallelization only. Thus, with the same code asked to the user, two different parallel versions of the simulation can be obtained.

However, if performances are improved by the hybrid parallelization, this performance is also limited by the scheduling strategy used in this paper. Actually, as described in Section~\ref{sect:msp}, a static series-parallel (fork-join) scheduling is performed in this work. This helps to define a predictive performance model, which has been validated in experiments. However, getting rid of global synchronizations (join) could be an interesting improvement to get better performance, especially for irregular simulations, where tasks are not naturally balanced. Our immediate perspective, thus, is to use OpenMP dynamic schedulers~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359} as back-ends to study performance improvments. Using such dynamic schedulers also opens to more heterogeneous architectures, such as GPUs or many-cores accelerators.

Finally, to validate the separation of concerns introduced by MSL, a second back-end for unstructured meshes could be interesting. We think about PaMPA~\cite{lachat:hal-00768916} as a distributed data structure backend, instead of SkelGIS. Thus using a single language, the MSL language, two different kinds of numerical simulations could be defined: one onto Cartesian meshes; and one on unstructured meshes.

### 2016-05-30 ARTICLES/IJPP/article/msl.tex

@@
             "stencil shapes:" liststencil
-            "quantities:" listquantities
-            "scalar" listscalar
+            "mesh quantities:" listquantities
+            "scalars" listscalar
             listloop
@@
   nec from edgex to cell
-quantities :
+mesh quantities :
   cell A,B,D,E,F,G,I,J
   edgex C,H
-scalar : mu, tau
+scalars : mu, tau
 time : 500
@@
 
-\item The \texttt{quantities} keyword (Fig.\ref{fig:grammar},~l.6) introduces identifiers for $\delta\in\Delta$, the quantities with the group of entities they are mapped on (Fig.\ref{fig:grammar},~l.16).
+\item The \texttt{mesh quantities} keyword (Fig.\ref{fig:grammar},~l.6) introduces identifiers for $\delta\in\Delta$, the quantities with the group of entities they are mapped on (Fig.\ref{fig:grammar},~l.16).
 For example the quantities \texttt{C} and \texttt{H} are mapped on \texttt{edgex} entities.
 
-\item The \texttt{scalar} keyword (Fig.\ref{fig:grammar},~l.7) introduces identifiers for $s\in\mathcal{S}$, the scalars. For example \texttt{mu} and \texttt{tau} in Fig.\ref{fig:exmsl},~l.15.
+\item The \texttt{scalars} keyword (Fig.\ref{fig:grammar},~l.7) introduces identifiers for $s\in\mathcal{S}$, the scalars. For example \texttt{mu} and \texttt{tau} in Fig.\ref{fig:exmsl},~l.15.
 

### 2016-05-30 ARTICLES/IJPP/article/related.tex

@@
 Each one has its own specificities and answers to a specific stencil case or to a specific additionnal optimization.
-For example, applied on Cartesian meshes, Pochoir~\cite{spaaTangCKLL11} works on cache optimization techniques for stencils applied onto Cartesian meshes.
+For example, Pochoir~\cite{spaaTangCKLL11} works on cache optimization techniques for stencils applied onto Cartesian meshes.
 On the other hand, PATUS~\cite{citeulike12258902} proposes to add a parallelization strategy grammar to its stencil language to perform an auto-tuning parallelization strategy.
@@
 
-The reason why each of those languages is implemented from scratch to answer its own particularities and its own optimizations is because each language is built and thought as a single block, which makes impossible or difficult code reuse from one language to another.
+The reason why each of those languages are implemented from scratch to answer its own particularities and its own optimizations is because each language is built and thought as a single block, which makes impossible or difficult code reuse from one language to another.
 In other words domain specific languages (for stencils or other domains) suffer from a lack of software engineering properties, which could increase the productivity to build a new language, by code-reuse, and also the maintainability of languages, with more separation of concerns.
@@
 For example, a computation can be applied onto a subset of the space domain which is not possible with Liszt and Nabla.
-This functionnality is an important flexibility in numerical simulations yet.
+This functionnality is important in numerical simulations yet.
 For example, many computations performed in the simulation studied in this paper have to be applied onto a subpart of the overall space domain.
@@
 Thus, MSL improves separation of concerns and maintainability of a given simulation.
-Actually, a numerician could describes the different computations to perform into the simulation, while another one, later, could focuss on the implementation and the numerical code of the simulation.
+Actually, a numerician could describes the different computations to perform into the simulation, while another one, later, could focuss on the implementation and the numerical code.
 This also facilitates reuse of other languages.

### 2016-05-30 ARTICLES/IJPP/article/introduction.tex
IJPP article intro + related

@@
 Moreover, the parallelization performed by the language is large enough to be compatible with many architectures and back-end languages.
-Contributions presented in this paper are : the computational model of a multi-stencil program and its parallelization formalism; the MSL grammar and its compiler; and a performance evaluation of a real case numerical simulation up to 16.384 cores.
+Contributions presented in this paper are : the computational model of a multi-stencil program and its parallelization formalism; the MSL grammar and its compiler; a back-end implementation and its performance evaluation onto a real case numerical simulation up to 16.384 cores.
 

### 2016-05-30 ARTICLES/IJPP/article/compiler.tex
implem

--- 2016-05-20 ARTICLES/IJPP/article/compiler.tex
@@
 %--------------------
-\subsection{Overall compiler algorithm}
+%The multi-stencil language is a descriptive language.
+MSL follows a light grammar which extracts information needed to build an empty parallel pattern of a numerical simulation (a multi-stencil program), also called an empty parallel skeleton of the simulation.
+From the language grammar are automatically detected where synchronizations are needed in order to apply a data parallelization technique, and how computations are dependent from each other to apply a task parallelization technique.
 
-The overall compiler algorithm is represented in Algorithm~\ref{alg:compiler}.
-First, the parsing of the input file (written with the MSL language) gave enough information to build $\Gamma$, the ordered list of computation kernels.
-Second, $\Gamma_{sync}$ and $\Gamma_{dep}$ can be built from $\Gamma$ using the recursive functions $F_{sync}$ and $F_{dep}$.
-Third, N-Shapes are removed from $\Gamma_{dep}$ to get a MSPD graph, and to be able to create its serie-parallel binary tree decomposition $TSP$ (line 6).
-If only a data parallelization has been chosen, Line 8 performs the fusion of kernels.
-Finally a dump of $TSP$ is performed by the compiler.
-Implementation choices of the dump are described in the next subsection.
+%The multi-stencil language splits the simulation description from its implementation concerns.
+For this reason, the multi-stencil language is mesh-agnostic, which means that no information on the actual topology of the mesh is given by the user.
+For this reason too, no numerical code is asked to the user.
+Those implementation concerns are deported to a second phase where other languages or libraries can be used.
+This phase is called the \emph{dump} in the overall compiler algorithm.
 
-\begin{algorithm}
-\caption{MSL Compiler}
-\label{alg:compiler}
-\begin{algorithmic}[1]
-\Procedure{compileMSL} {file}
-\State $\Gamma$ = parser(file)
-\State $\Gamma_{sync} = F_{sync}(\Gamma,0)$
-\State $\Gamma_{dep} = F_{dep}(\Gamma_{sync},0)$
-\State removeNSH($\Gamma_{dep}$)
-\State $TSP = F_{tsp}(TSP,root(TSP))$
-\If {data parallel}
-\State $TSP = F_{fus}(TSP)$
-\EndIf
-\State dump(TSP)
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
+The MSL compiler takes a file written in the grammar described in Section~\ref{sect:msl} as input and generates an application skeleton.
+In this skeleton, the user still has to fill the functions corresponding to the various computation kernels.
+The overall behavior of the compiler is as follow: %presented in Algorithm~\ref{alg:compiler}.
+\begin{enumerate}
+ \item it parses the input file and generates $\Gamma$, the list of computation kernels,
+ \item from $\Gamma$, it builds $\Gamma_{sync}$, the list including synchronizations for data parallelism using Algorithm $F_{sync}$ introduced in Section~\ref{sect:parallelism},
+ \item from $\Gamma_{sync}$, it builds $\Gamma_{dep}$, the DAG supporting hybrid parallelism using Algorithm $F_{dep}$ introduced in Section~\ref{sect:parallelism},
+ \item it then removes the N-Shapes from $\Gamma_{dep}$ to get a MSPD graph, and generates its serie-parallel binary tree decomposition $TSP$,
+ \item it performs the fusion of kernels in $TSP$ if required,
+ \item finally it dumps an application structure matching $TSP$.
+% Implementation choices of the dump are described in the next subsection.
+\end{enumerate}
+
+%JB: un enumerate me semble suffisant
+% \begin{algorithm}
+% \caption{MSL Compiler}
+% \label{alg:compiler}
+% \begin{algorithmic}[1]
+% \Procedure{compileMSL} {file}
+% \State $\Gamma$ = parser(file)
+% \State $\Gamma_{sync} = F_{sync}(\Gamma,0)$
+% \State $\Gamma_{dep} = F_{dep}(\Gamma_{sync},0)$
+% \State removeNSH($\Gamma_{dep}$)
+% \State $TSP = F_{tsp}(TSP,root(TSP))$
+% \If {data parallel}
+% \State $TSP = F_{fus}(TSP)$
+% \EndIf
+% \State dump(TSP)
+% \EndProcedure
+% \end{algorithmic}
+% \end{algorithm}
 
 %--------------------
-\subsection{Implementation choices}
+% \subsection{Implementation choices}
 
-SkelGIS + OpenMP + L2C
+The abstract definition of MSL enables to use a very wide range of languages for the dump.
+In this paper, we use a compiler that targets the Low Level Components~\cite{l2c} (\llc), a component model for C++ supporting high-performance applications.
+Similar to classes in object-oriented models, components specify the services they provide, but in addition they also specify the services they require.
+This enables to build applications by assembling components in a second phase through the matching of requirements of some components with the services provided by others.
+In the case of MSL, this makes it possible to clearly separate in distinct components the features that are implemented:
+\begin{itemize}
+ \item by reusing the same code every-time such as for the time iteration logic,
+ \item by generating some code that depends on $TSP$ and thus from the input file such as for the task parallelism logic,
+ \item or by letting the user fill-in an empty skeleton component such as for the computation kernels.
+\end{itemize}
+By enforcing the use of well defined interfaces in the generated code, \llc makes it easier to replace an implementation by another to reconsider an implementation choice.
+It also simplifies the potential addition of new features to MSL or the coupling of MSL with another domain-specific language for example.
+
+The two main other choices in the code generation concern the technologies used for the data and task parallelizations.
+For the data-parallelization, we rely on SkelGIS, a C++ embedded DSL~\cite{CPE:CPE3494} which offers distributed data structures (meshes) for numerical simulations, and programming interfaces to easily write codes using them.
+SkelGIS is implemented over MPI~\cite{Graham2009MSE} and can therefore be used on distributed memory architectures such as clusters.
+Two kind of meshes are offered by SkelGIS: \textbf{a)} a distributed two dimensional Cartesian mesh~\cite{DBLP:conf/ieeehpcs/HeleneS13}, and \textbf{b)} a distributed graph of Cartesian meshes (hybrid mesh)~\cite{DBLP:conf/europar/CoullonL14}.
+The use of different back-ends for data-parallelism such as for example the distributed unstructured mesh proposed by PaMPA~\cite{lachat:hal-00768916} will be the subject of future work.
+
+For the task parallelism, we use OpenMP~\cite{660313}.
+OpenMP targets shared-memory platforms only, but the degree of task-parallelism is typically much lower than that of data parallelism.
+With the increasing number of cores of current machines this has not proved to be a problem in our experience.
+While the version 4 of OpenMP has introduced explicit support for tasks, our implementation only requires version 3 whose fork-join model is well suited for the static scheduling introduced in Section~\ref{sect:msp}.
+The use of dynamic schedulers such as provided by libgomp\footnote{\url{https://gcc.gnu.org/projects/gomp/}}, StarPU~\cite{Augonnet2011}, or XKaapi~\cite{Gautier:2013:XRS:2510661.2511383} to directly execute the DAG will be the subject of some future work.
+
+As a result, the MSL compiler generates an hybrid code which uses both SkelGIS and OpenMP.
+It also generates empty components where the user must provide local sequential implementations of the kernels using well defined interfaces of the component to iterate over the local data subdomain.
+The following section will evaluate this compiler as well as the code it generates.

### 2016-05-30 ARTICLES/IJPP/article/formalism.tex

@@
 For example, in a 2D Cartesian mesh, a group of entities where each entity is made of four vertices and four edges could form the cells.
-Another group of entities are the vertices defined as all singletons formed of a single vertex of $V$. Both groups, then, would be part of $\Phi$.
+Another group of entities are the vertices defined as all singletons formed of a single vertex of $V$. Both groups, then, would be part of $\Phi$. This example is illustrated in Figure~\ref{fig:ex1}.
 
@@
 \begin{center}
-\subfloat[Mesh and mesh domains.\label{fig:meshbase}]{
+\subfloat[Mesh and two groups of mesh entities.\label{fig:meshbase}]{
 \resizebox{8cm}{!}{\includegraphics{./images/mesh.pdf}}
@@
 \hspace{10pt}
-\subfloat[4-neighborhood stencil.\label{fig:ex1}]{
+\subfloat[A is computed with a 4-neighborhood stencil applied on B. A is computed onto a computation domain which does not include all entities of the group.\label{fig:ex1}]{
 \resizebox{5cm}{!}{\includegraphics{./images/stencil1.pdf}}
@@
 \vspace{20pt}
-\subfloat[4-neighborhood stencil.\label{fig:ex2}]{
+\hspace{10 pt}
+\subfloat[A is computed with a 2-neighborhood stencil applied on C. A is computed onto a computation domain includes all entities of the group.\label{fig:ex2}]{
 \resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
@@
 \end{center}
-\caption{(a) a Cartesian mesh and two kind of mesh entities, (b) an example of stencil kernel on cells, (c) an example of stencil kernel on two different entities of the mesh.}
+\caption{(a) a Cartesian mesh and two kind of groups of mesh entities, (b) an example of stencil kernel on cells, (c) an example of stencil kernel on two different entities of the mesh.}
 \label{fig:gspmsp}

### 2016-05-30 ARTICLES/IJPP/article/parallelism.tex

@@
 \begin{align}
-        k_{0;1}^{sync}=(\emptyset,\{(B,nce)\},(B,\cup_{\phi \in D_1} nce(\phi)),identity),\\
-        k_{1;4}^{sync}=(\emptyset,\{(C,nec)\},(C,\cup_{\phi \in D_4} nec(\phi)),identity),\\
-        k_{7;8}^{sync}=(\emptyset,\{(I,ncc)\},(I,\cup_{\phi \in D_8} ncc(\phi)),identity).
+        k_{0;1}^{sync}=(\emptyset,\{(B,nce)\},(B,\cup_{\phi \in D_1} nce(\phi))),\\
+        k_{1;4}^{sync}=(\emptyset,\{(C,nec)\},(C,\cup_{\phi \in D_4} nec(\phi))),\\
+        k_{7;8}^{sync}=(\emptyset,\{(I,ncc)\},(I,\cup_{\phi \in D_8} ncc(\phi))).
 \end{align}

### 2016-05-30 ARTICLES/IJPP/article/related.tex

@@
 In other words domain specific languages (for stencils or other domains) suffer from a lack of software engineering properties, which could increase the productivity to build a new language, by code-reuse, and also the maintainability of languages, with more separation of concerns.
-A few works propose frameworks to create DSL, which afterwards will be easier to compose and combine together~\cite{Sujeeth:2013:CRC:2524984.2524988}.
-The Mulsi-Stencil Language (MSL), presented in this paper, is itself a domain specific language for stencil-based simulations.
+As far as we know, a single work proposes a framework to create DSL, which afterwards will be easier to compose and combine together~\cite{Sujeeth:2013:CRC:2524984.2524988}, which is interesting for inter-domains applications.
+However, the point argued by this paper, and by the Multi-Stencil Language (MSL), is that DSL conception must be studied to maximize its usefulness for different types of applications, while keeping a good performance.
+This maximization is directly linked to the abstraction level proposed to the end-user, and also to separation of concerns and code reuse improvements.
+
+MSL is itself a domain specific language for stencil-based simulations.
 It offers a way to give a mesh-agnostic description of a simulation, which could be common to different cases of simulations, and thus facilitates reuse of existing underlying data structures, optimizations or parallelization techniques.
@@
 In other words, the topology of a mesh as well as numerical codes are given at the same time than the description of the different computations to apply into the simulation, while MSL splits those two different phases.
-Thus, MSL improves separation of concerns and maintainability of a given simulation.
+Thus, MSL improves separation of concerns.
 Actually, a numerician could describes the different computations to perform into the simulation, while another one, later, could focuss on the implementation and the numerical code.
@@
 
-The formalism of a Multi-Stencil program, given in this paper, is new, as far as we know.
 Finally, the parallelization techniques proposed in this paper, take place at paradigm level, with data parallelism and hybrid (data plus task) parallelism.
@@
 This makes possible a large panel of backend architectures and languages.
+In other words MSL improves portability.

### 2016-05-30 ARTICLES/IJPP/article/compiler.tex
IJPP article implementation corrections

@@
 This enables to build applications by assembling components in a second phase through the matching of requirements of some components with the services provided by others.
-In the case of MSL, this makes it possible to clearly separate in distinct components the features that are implemented:
+In the case of MSL, this assembly of components offers a way to clearly separate in distinct components the features that are implemented:
 \begin{itemize}
- \item by reusing the same code every-time such as for the time iteration logic,
- \item by generating some code that depends on $TSP$ and thus from the input file such as for the task parallelism logic,
- \item or by letting the user fill-in an empty skeleton component such as for the computation kernels.
+ \item by reusing the same code more than once, such as for the time iteration logic,
+ \item by generating some code that depends on $TSP$ and thus from the input file, such as for the task parallelism logic,
+ \item or by letting the user fill-in an empty skeleton component, such as for the computation kernels.
 \end{itemize}
-By enforcing the use of well defined interfaces in the generated code, \llc makes it easier to replace an implementation by another to reconsider an implementation choice.
-It also simplifies the potential addition of new features to MSL or the coupling of MSL with another domain-specific language for example.
+By enforcing the use of well defined interfaces in the generated code, \llc makes easier the substitution of an implementation by another, for example to reconsider an implementation choice.
+It also simplifies the addition of new features to MSL or the coupling of MSL with another domain-specific language, for example.
 
-The two main other choices in the code generation concern the technologies used for the data and task parallelizations.
+The two main other choices in the code generation concern the technologies used for data and task parallelizations.
 For the data-parallelization, we rely on SkelGIS, a C++ embedded DSL~\cite{CPE:CPE3494} which offers distributed data structures (meshes) for numerical simulations, and programming interfaces to easily write codes using them.
@@
 Two kind of meshes are offered by SkelGIS: \textbf{a)} a distributed two dimensional Cartesian mesh~\cite{DBLP:conf/ieeehpcs/HeleneS13}, and \textbf{b)} a distributed graph of Cartesian meshes (hybrid mesh)~\cite{DBLP:conf/europar/CoullonL14}.
-The use of different back-ends for data-parallelism such as for example the distributed unstructured mesh proposed by PaMPA~\cite{lachat:hal-00768916} will be the subject of future work.
+The use of different back-ends for data-parallelism, such as for example the distributed unstructured mesh proposed by PaMPA~\cite{lachat:hal-00768916}, will be the subject of future work.
 
@@
 OpenMP targets shared-memory platforms only, but the degree of task-parallelism is typically much lower than that of data parallelism.
-With the increasing number of cores of current machines this has not proved to be a problem in our experience.
+With the increasing number of cores of current machines this has not proved to be a problem in our experience. % LN : ?
 While the version 4 of OpenMP has introduced explicit support for tasks, our implementation only requires version 3 whose fork-join model is well suited for the static scheduling introduced in Section~\ref{sect:msp}.
-The use of dynamic schedulers such as provided by libgomp\footnote{\url{https://gcc.gnu.org/projects/gomp/}}, StarPU~\cite{Augonnet2011}, or XKaapi~\cite{Gautier:2013:XRS:2510661.2511383} to directly execute the DAG will be the subject of some future work.
+The use of dynamic schedulers such as provided by libgomp\footnote{\url{https://gcc.gnu.org/projects/gomp/}}, StarPU~\cite{Augonnet2011}, or XKaapi~\cite{Gautier:2013:XRS:2510661.2511383} to directly execute the DAG $\Gamma_{dep}$ will be the subject of some future work.
 

### 2016-05-30 ARTICLES/IJPP/article/evaluation.tex
allegement debut eval

@@
-This section presents evaluations of the work presented in this paper.
-First, the compiler is evaluated.
-Second, the data parallelization and fusion optimization are evaluated and compared to a full SkelGIS program.
-Third, the hybrid parallelization (data and task) is evaluated to validate the performance model detailed in Section~\ref{sect:tsp}.
+This section evaluates the MSL compiler and the code it generates.
+We first evaluate the generated code in the case where only data parallelization and fusion are applied and compare this to a full SkelGIS program.
+Then, we analyze a case taking advantage of the full the hybrid parallelization combining data and task parallelism and validate the performance model proposed in Section~\ref{sect:tsp}.
 
-All evaluations presented in this section are based on a real case study, the shallow-Water Equations.
-Navier-Stokes Equations is a well known set of partial differential equations in fluid dynamics to simulate a flow evolution in time.
-At the University of Orl\'eans, France, the MAPMO laboratory works on a software, called FullSWOF2D\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}, which solves the Shallow-Water Equations obtained from the three dimensional Navier-Stokes equations, by averaging on the vertical direction~\cite{Ferrari2004}.
-Those equations are solved using a two-dimensional Cartesian discretization of the space domain, and a finite volume numerical methods more described in~\cite{CPE:CPE3494}.
+All evaluations presented in this section are based on a real case study of the shallow-Water Equations as solved in  the FullSWOF2D\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}~\cite{Ferrari2004,CPE:CPE3494} code from the MAPMO laboratory, University of Orl\'eans.
 We have developed a MSL version of FullSWOF2D that contains 3~mesh entities, 7~computation domains, 48~data and 98~computations (32~stencil kernels and 66~local kernels).
@@
 
-The series-parallel tree decomposition $TSP$ of this simulation, extracted by the MSL compiler, is composed of 17~sequence nodes and 18~parallel nodes.
-Table~\ref{fig:freq} represents the number of time a given level of parallelism, \ie the number of tasks to perform concurrently, is observed in the final tree.
-One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1).
-Moreover, a level of 16~parallel tasks is reached two times, and five times for the fourth level.
-This means that if two threads are dedicated to task parallelism, two parts of the code at least will not take advantage of this and that no part of the code would benefit from more than 16 threads.
-
-\begin{table}[!h]
- \begin{center}
- \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-    \hline 
-   Level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
-   \hline
-   Frequency & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
-   \hline
- \end{tabular}
-\caption{Parallelism level (number of parallel tasks) and the number of times this level appears.}
-\label{fig:freq}
- \end{center}
-\end{table}
-
-Table~\ref{fig:exectime} illustrates the execution time for each step of the MSL compiler for an overall execution time of 4.6 seconds.
-Execution times have been computed on a laptop with a bi-core Intel Core i5 1.4~GHz, and 8~GB of DDR3.
-One can notice that the computation of the $TSP$ tree is the longest step of the compiler.
+Table~\ref{fig:exectime} illustrates the execution time for each step of the MSL compiler for the FullSWOF2D compiler.
+This has been computed on a laptop with a bi-core Intel Core i5 1.4~GHz, and 8~GB of DDR3.
+While the overall time of 4.6 seconds remains reasonable, one can notice that the computation of the $TSP$ tree is by far the longest step of the compiler.
+As a matter of fact, the complexity of the algorithm for N-shapes removal is $O(n^3)$.
+The replacement of the static scheduling by a direct scheduling of the DAG by dedicated tools should solve this in the future.
 
@@
 
+The series-parallel tree decomposition $TSP$ of this simulation, extracted by the MSL compiler, is composed of 17~sequence nodes and 18~parallel nodes.
+Table~\ref{fig:freq} represents the number of time a given level of parallelism, \ie the number of tasks to perform concurrently, is observed in the final tree.
+One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1).
+Moreover, a level of 16~parallel tasks is reached two times, and five times for the fourth level.
+This means that if two threads are dedicated to task parallelism, two parts of the code at least will not take advantage of this and that no part of the code would benefit from more than 16 threads.
+
+\begin{table}[!h]
+ \begin{center}
+ \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+    \hline 
+   Level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
+   \hline
+   Frequency & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
+   \hline
+ \end{tabular}
+\caption{Parallelism level (number of parallel tasks) and the number of times this level appears.}
+\label{fig:freq}
+ \end{center}
+\end{table}
+
 The second parallelization technique proposed by MSL is an hybrid parallelization which mixes data parallelization and task parallelization.

### 2016-05-30 ARTICLES/IJPP/article/evaluation.tex
typo
eval

@@
 
-Table~\ref{fig:exectime} illustrates the execution time for each step of the MSL compiler for the FullSWOF2D compiler.
-This has been computed on a laptop with a bi-core Intel Core i5 1.4~GHz, and 8~GB of DDR3.
-While the overall time of 4.6 seconds remains reasonable, one can notice that the computation of the $TSP$ tree is by far the longest step of the compiler.
+Table~\ref{fig:exectime} illustrates the execution time of each step of the MSL compiler for the FullSWOF2D example.
+This has been computed on a laptop with a dual-core Intel Core i5 1.4~GHz, and 8~GB of DDR3.
+While the overall time of 4.6 seconds remains reasonable, one can notice that the computation of the $TSP$ tree is by far the longest step.
 As a matter of fact, the complexity of the algorithm for N-shapes removal is $O(n^3)$.
@@
 
-The first parallelization technique proposed by MSL is a data parallelization.
-To implement this parallelization, the dump step of the MSL compiler uses the SkelGIS language.
-This language, as already described, offers a distributed Cartesian mesh and programming interfaces to use it as a sequential data structure.
-It has been evaluated itself compared to an MPI implementation on the same FullSWOF2D simulation~\cite{CPE:CPE3494}.
-What MSL brings to SkelGIS is an automatic detection of where MPI synchronizations are needed during the numerical simulation, and how to make a fusion of some computations in order to reduce cache misses.
-This section gives a performance comparison of the code produced by MSL for FullSWOF2D, and of the equivalent manual SkelGIS program (with manual synchronizations and fusion choices).
-Moreover, as the code produced by MSL is a component-based code, this evaluation also shows that no overheads are introduced \llc~\cite{l2c}.
-Following evaluations have been performed onto the machine of the TGCC Curie described in Table~\ref{tab:TGCC}.
-Each evaluation has been performed ten times and the median is presented in results.
+In this evaluation, we disable task-parallelism (fix its degree to 1) so as to focus on data-parallelism.
+In the implementation evaluated in this paper, the data parallelism is handled using SkelGIS that has already been evaluated in comparison with a native MPI implementation for the FullSWOF2D example~\cite{CPE:CPE3494}.
+In addition to SkelGIS, MSL automatically inserts the required MPI synchronizations and identifies potential fusion of kernels to reduce cache misses.
 
-\begin{table}[!h]
+\begin{table}[!ht]
 \begin{center}
@@
     MPI & Bullxmpi\\
-    Network & Infiniband\\
+    Network & fat-tree Infiniband\\
     \hline
@@
 
-\paragraph{\textbf{Weak scaling}} The first evaluation is a weak scaling evaluation.
-In a weak scaling, a constant amount of work is given to each core while increasing the overall number of cores.
-As a result, the overall domain size also increase with the number of cores.
-A weak scaling is a good evaluation to detect overheads which are independent from the amount of computation performed by each core, such as collective communications, or overheads introduced by \llc components, for example.
-Figures~\ref{fig:weak1} and~\ref{fig:weak2} respectively show the weak scaling evaluations for a $400 \times 400$ and a $800 \times 800$ domain for each core, from 16 cores to 16.384 cores.
+This section compares the performance of the code produced by MSL for FullSWOF2D with a plain SkelGIS program where synchronizations and fusion choices have been done by the developer without any automatic support.
+This enables to evaluate both the choices made by the compiler as well as the potential overheads of using \llc~\cite{l2c} that is not used in the plain SkelGIS version.
+The evaluations have been performed on the Curie supercomputer (TGCC, France) described in Table~\ref{tab:TGCC}.
+Each evaluation has been performed ten times and the median is presented in results.
+
+\paragraph{\textbf{Weak scaling}}
+\begin{figure}[t]
+\begin{minipage}{.475\textwidth}
+  \includegraphics[width=\textwidth]{../results/weak_scaling/400/median_weak.pdf}
+  \caption{weak-scaling with $400 \times 400$ domain per core.}
+  \label{fig:weak1}
+\end{minipage}
+\hfill
+\begin{minipage}{.475\textwidth}
+  \includegraphics[width=\textwidth]{../results/weak_scaling/800/median_weak.pdf}
+  \caption{weak-scaling with $800 \times 800$ domain per core.}
+  \label{fig:weak2}
+\end{minipage}\end{figure}
+%In a weak scaling, a constant amount of work is given to each core while increasing the overall number of cores.
+%As a result, the overall domain size also increase with the number of cores.
+%A weak scaling is a good evaluation to detect overheads which are independent from the amount of computation performed by each core, such as collective communications, or overheads introduced by \llc components, for example.
+Figures~\ref{fig:weak1} and~\ref{fig:weak2} show a weak scaling for a $400 \times 400$ points by core and a $800 \times 800$ points by core from 16 cores to 16.384 cores.
 Minimum and maximum values are also shown as error bars in figures.
 
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/400/median_weak.pdf}}
-  \caption{weak-scaling with $400 \times 400$ domain per core.}
-  \label{fig:weak1}
-\end{center}\end{figure}
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/800/median_weak.pdf}}
-  \caption{weak-scaling with $800 \times 800$ domain per core.}
-  \label{fig:weak2}
-\end{center}\end{figure}
+One can notice that the MSL code produces a slightly better execution time on the $400 \times 400$ domain and slightly worse a small overhead on the $800 \times 800$ domain.
+This might be explained by slightly different compiler optimization flags since \llc requires compilation as dynamic libraries with the \emph{-fpic} compilation flag while the SkelGIS version does not.
+This flag can have slight positive or negative effect on code performance depending on the situation and might be responsible for the observed difference.
+In any case, these difference are barely noticeable and often included in the error bar of the measure and both version globally perform similarly.
+The MPI synchronizations performed by MSL perform as good as manually specified ones at scale and \llc does not introduce any noticeable overhead.
 
-One can notice that the MSL code produces a better execution time on the $400 \times 400$ domain, while it produces a small overhead on the $800 \times 800$ domain.
-Those results could be due to different optimizations of the gcc compiler due to components.
-Actually, a component in \llc is compiled independently, as an dynamic library, with the \emph{-fpic} compilation option, while the SkelGIS version is compiled entirely as a single code.
-However, one can also notice that except those not significant differences of execution times, weak scalings have the same behavior on both codes.
-
-\paragraph{\textbf{Strong scaling}} The second evaluation is a strong scaling evaluation.
-Unlike a weak scaling, a strong scaling keeps a constant overall size of problem, while the number of cores increase.
-As a result, the sub-domain computed by each core reduces while the number of cores increases.
-The strong scaling is a good additional evaluation to the weak scaling, showing different kinds of overheads, such as cache misses (when the size of sub-domain becomes small enough to fit into cache memory), or small constant overheads.
-Figure~\ref{fig:strong} shows the strong scaling evaluation for a $10k \times 10k$ overall domain size, from 16 cores to 16.384 cores.
-This strong scaling is illustrated with the number of iterations per seconds as a function of the number of cores.
-The ideal strong scaling is also illustrated.
-
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/strong_scaling/10K_1K/median_strong.pdf}}
+\paragraph{\textbf{Strong scaling}}
+\begin{figure}[t]\begin{center}
+  \includegraphics[width=.6\textwidth]{../results/strong_scaling/10K_1K/median_strong.pdf}
   \caption{Strong scaling on a $10k \times 10k$ domain.}
@@
 \end{center}\end{figure}
+% The second evaluation is a strong scaling evaluation.
+% Unlike a weak scaling, a strong scaling keeps a constant overall size of problem, while the number of cores increase.
+% As a result, the sub-domain computed by each core reduces while the number of cores increases.
+% The strong scaling is a good additional evaluation to the weak scaling, showing different kinds of overheads, such as cache misses (when the size of sub-domain becomes small enough to fit into cache memory), or small constant overheads.
+Figure~\ref{fig:strong} shows the number of iteration per second for a 10k$\times$10k global domain size from 16 to 16.384 cores.
+The ideal strong scaling is also illustrated.
 
@@
 \begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/fusVSbase.pdf}}
+  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/fusVSbase.pdf}
   \caption{Strong scaling on a 500x500 domain, with and without the fusion optimization.}
@@
 
-The series-parallel tree decomposition $TSP$ of this simulation, extracted by the MSL compiler, is composed of 17~sequence nodes and 18~parallel nodes.
-Table~\ref{fig:freq} represents the number of time a given level of parallelism, \ie the number of tasks to perform concurrently, is observed in the final tree.
-One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1).
-Moreover, a level of 16~parallel tasks is reached two times, and five times for the fourth level.
-This means that if two threads are dedicated to task parallelism, two parts of the code at least will not take advantage of this and that no part of the code would benefit from more than 16 threads.
-
-\begin{table}[!h]
+\begin{table}[th]
  \begin{center}
@@
 
-The second parallelization technique proposed by MSL is an hybrid parallelization which mixes data parallelization and task parallelization.
-To implement this hybrid parallelization, the dump step of the MSL compiler uses the SkelGIS language, and the OpenMP language.
+In this evaluation, we introduce task parallelism to evaluate the hybrid parallelization offered by MSL that mixes data parallelization and task parallelization.
+In the implementation evaluated in this paper and in addition to SkelGIS, the hybrid parallelization relies on OpenMP.
 
-First, Figure~\ref{fig:limit} illustrates limitations of data parallelization technique alone.
-As explained in Section~\ref{sect:perfs}, performance of data parallelism is limited by the communication time between resources.
-As a result, when the size of sub-domain for each core becomes too small, and as the communication time increases with the number of cores, the strong speedup of the data parallelization starts to bend down.
-To catch this phenomena without too much cores, a relatively small domain of $500 \times 500$ is used.
+The series-parallel tree decomposition $TSP$ of this simulation, extracted by the MSL compiler, is composed of 17~sequence nodes and 18~parallel nodes.
+Table~\ref{fig:freq} represents the number of time a given level of parallelism, \ie the number of tasks to perform concurrently, is observed in the final tree.
+One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1).
+Moreover, a level of 16~parallel tasks is reached two times, and five times for the fourth level.
+This means that if two threads are dedicated to task parallelism, two parts of the code at least will not take advantage of this and that no part of the code would benefit from more than 16 threads.
+The task parallelism alone is therefore clearly not enough to take advantage of even a single node of a cluster that typically supports today between 16 and 32 threads (16 for Curie).
 
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/times.pdf}}
+\begin{figure}[bh]\begin{center}
+  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/analytic/times.pdf}
   \caption{Computation vs communication times in the data parallelization technique.}
@@
 
-As expected, from 1 to 2048 cores the computation times decrease linearly.
-On the other hand, from 1 to 256 cores the communication time is almost constant, but from 256 to 2048 cores it increases.
-As a result from 512 cores to 2048 cores, the communication time becomes greater than the computation time.
-In Figure~\ref{fig:close}, the blue curve shows the strong scaling of the same example.
-Thus, from 256 to 2048 cores, the speedup bends down.
+On the other hand, Figure~\ref{fig:limit} illustrates limitations of data parallelization technique alone.
+While the computation time decreases linearly with the number of core used, the communication behavior is much more erratic.
+Between 2 and 16 cores, the communication happen inside a single node and the time is small and nearly constant.
+There is a small oscillation that might be explained by difference between number of processors that are power of 4 where the communications are all the same and those that are not where there is an imbalance.
+Starting with 32 cores, there are multiple nodes and the communication time is typically modeled as $L+S/B$ where $L$ is the latency, $S$ the data size and $B$ the bandwidth.
+This explains the decrease of time from 32 to 128 cores where the data sizes communicated by each process decreases.
+The increased observed after that might be due to the fact that with the increased number of processes the fat-tree becomes deeper and the latencies increase.
 
-\medskip
+\begin{figure}[th]\begin{center}
+  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/base_close_median.pdf}
+  \caption{Strong scaling comparisons between data parallelization and hybrid parallelization. A \emph{close} OpenMP clause is used to bind threads onto cores.}
+  \label{fig:close}
+\end{center}\end{figure}
+
+All in all, when the number of core used increases, the computation/communication ratio becomes poorer and poorer.
+As a result, the data parallelism alone fails to provide enough parallelism to leverage the whole machine and other sources of parallelism have to be found.
+The blue curve of Figure~\ref{fig:close} shows the total execution time for the same example as in Figure~\ref{fig:limit} and as expected, the speedup bends down from 256 to 2048 cores.
+
 In addition to the blue curve, Figure~\ref{fig:close} shows strong scalings for the same example but using an hybrid parallelization.
@@
 This figure shows a comparison with 2, 4, 8 and 16 cores per MPI process for task parallelization.
-
-\begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_close_median.pdf}}
-  \caption{Strong scaling comparisons between data parallelization and hybrid parallelization. A \emph{close} OpenMP clause is used to bind threads onto cores.}
-  \label{fig:close}
-\end{center}\end{figure}
 

### 2016-05-30 ARTICLES/IJPP/article/introduction.tex

@@
 Those models increase HPC code portability and reach an interesting performance onto heterogeneous architectures, which is interesting to reach exascale programming.
-At a highest abstraction level, general purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} follow the direction of task graph scheduling proposed by those execution models.
+At a higher abstraction level, general purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} follow the direction of task graph scheduling proposed by those execution models.
 However, for non-experts end-users, general purpose languages still are difficult to use and to tune for a given application onto a given architecture.

### 2016-05-31 ARTICLES/IJPP/article/article.tex
IJPP article title and corrections
IJPP article evaluation correction

@@
 
-\title{The Multi-Stencil Language}
+\title{The Multi-Stencil Language: orchestrating stencils with a mesh-agnostic DSL}
 %\subtitle{Do you have a subtitle?\\ If so, write it here}
@@
               INRIA, LIP, ENS Lyon, 46 Allée d'Italie, 69364 Lyon Cedex 07, France\\
-              \email{helene.coullon@inria.fr}
+              \email{helene.coullon@inria.fr} \email{christian.perez@inria.fr}
            \and
@@
 
-%\begin{acknowledgements}
-%If you'd like to thank anyone, place your comments here
-%and remove the percent signs.
-%\end{acknowledgements}
+\begin{acknowledgements}
+This work has partially been supported by the PIA ELCI project of the French FSN.
+This  work  was  granted  access  to  the  HPC  resources  of TGCC under  the 
+allocations t2015067470 and x2016067617 made by GENCI.
+\end{acknowledgements}
 

### 2016-05-31 ARTICLES/IJPP/article/evaluation.tex
IJPP article evaluation correction

@@
 
-In this evaluation, we disable task-parallelism (fix its degree to 1) so as to focus on data-parallelism.
+In this evaluation, we disable task-parallelism (fix its degree to 1) to focus on data-parallelism.
 In the implementation evaluated in this paper, the data parallelism is handled using SkelGIS that has already been evaluated in comparison with a native MPI implementation for the FullSWOF2D example~\cite{CPE:CPE3494}.
-In addition to SkelGIS, MSL automatically inserts the required MPI synchronizations and identifies potential fusion of kernels to reduce cache misses.
+In addition to SkelGIS, MSL automatically inserts the required MPI synchronizations and identifies possible fusion of kernels to reduce cache misses.
 
@@
 
-One can notice that the MSL code produces a slightly better execution time on the $400 \times 400$ domain and slightly worse a small overhead on the $800 \times 800$ domain.
+One can notice that the MSL code produces a slightly better execution time on the $400 \times 400$ domain and a small overhead on the $800 \times 800$ domain.
 This might be explained by slightly different compiler optimization flags since \llc requires compilation as dynamic libraries with the \emph{-fpic} compilation flag while the SkelGIS version does not.
@@
 In any case, these difference are barely noticeable and often included in the error bar of the measure and both version globally perform similarly.
-The MPI synchronizations performed by MSL perform as good as manually specified ones at scale and \llc does not introduce any noticeable overhead.
+The MPI synchronizations performed by MSL perform as good as manually specified ones, and \llc does not introduce any noticeable overhead.
 
@@
 \paragraph{\textbf{Fusion}} Finally, to evaluate the data parallelization technique automatically introduced by MSL, the fusion optimization is evaluated.
-Figure~\ref{fig:fusion} shows the number of iterations per second as a function of the number of cores, for FullSWOF2D with and without fusion optimization and onto a $500 \time 500$ domain.
+Figure~\ref{fig:fusion} shows the number of iterations per second as a function of the number of cores, for FullSWOF2D with and without fusion optimization and onto a $500 \times 500$ domain.
 As explained in Section~\ref{sect:fusion}, the MSL fusion happens at a high level and is most of the time done naturally by a computer scientist.
@@
 Between 2 and 16 cores, the communication happen inside a single node and the time is small and nearly constant.
-There is a small oscillation that might be explained by difference between number of processors that are power of 4 where the communications are all the same and those that are not where there is an imbalance.
+There is a small oscillation that might be explained by difference between number of processors that are power of 4, where the communications are all the same, and those that are nota power of 4, where there is an imbalance.
 Starting with 32 cores, there are multiple nodes and the communication time is typically modeled as $L+S/B$ where $L$ is the latency, $S$ the data size and $B$ the bandwidth.
@@
 In this figure, a different strategy is used to bind threads onto available cores (using OpenMP).
-This strategy, called \emph{spread} instead of \emph{close} in Figure~\ref{fig:close}, binds threads on cores in order to spread as much as possible onto resources, which means that the two NUMA nodes are directly used whatever the number of cores kept for task parallelization.
+This strategy, called \emph{spread} (instead of \emph{close} in Figure~\ref{fig:close}), binds threads on cores in order to spread as much as possible onto resources, which means that the two NUMA nodes are directly used whatever the number of cores kept for task parallelization.
 As a result, and as shown in the figure, using 2, 4 and 8 cores an initial overhead is introduced.
@@
 
-Finally, to completely validate the performance model introduced in Section~\ref{sect:perfs}, and to understand when the hybrid parallelization becomes more interesting than the data parallelization, Figure~\ref{fig:tth2} represents $T_{COM1}$ and $T_{COM2}+T_{task}$ of Equation~(\ref{eq:hyb}), for the best case, \ie when 8 cores are used for task parallelization, and with a \emph{close} OpenMP bind of threads onto cores.
+Finally, to completely validate the performance model introduced in Section~\ref{sect:perfs}, and to understand when the hybrid parallelization becomes more interesting than the data parallelization, Figure~\ref{fig:tth2} represents $T_{COM1}$ and $T_{COM2}+T_{task}$ of Equation~(\ref{eq:hyb}), for the best case (\ie when 8 cores are used for task parallelization), and with a \emph{close} OpenMP bind of threads onto cores.
 Table~\ref{fig:tth} gives time details.

### 2016-05-31 ARTICLES/IJPP/results/task_scaling/500_200/analytic/results.py

@@ -133,6 +133,7 @@ p2 = plt.bar(plt1, COMPmedian,width, color='b', alpha=0.75,label='Computations')
 p1 = plt.bar(plt1+width, COMMmedian, width, color='r', alpha=0.75,label='Communications') 
 
 plt.xticks(plt1+width, ('1', '2', '4', '8', '16','32','64','128','256','512','1024','2048'))
+plt.ylim(ymin=0.0001)
 
 legend = plt.legend(loc='upper center', shadow=True,prop={'size':12})
 plt.grid(True, which="both", linestyle="dotted")

### 2016-06-27 ARTICLES/OTHER/JLESC/slides.tex
JLESC presentation

--- 2015-11-12 WOLFHPC15/presentation/slides.tex
@@
 %-------------------------------------------------------------------
-\title[From DSL to HPC Component-Based Runtime: A Multi-Stencil DSL Case Study]{From DSL to HPC Component-Based Runtime: A Multi-Stencil DSL Case Study}
-\author[Julien Bigot (CEA), Hélène Coullon (INRIA), Christian Perez (INRIA)]{Julien Bigot, \underline{Hélène Coullon}, Christian Perez}
+\title[The Multi-Stencil Language]{The Multi-Stencil Language: orchestrating stencils with a mesh-agnostic DSL}
+\author[Hélène Coullon (INRIA), Julien Bigot (CEA), Christian Perez (INRIA)]{\underline{Hélène Coullon}, Julien Bigot, Christian Perez}
 \institute[INRIA]{INRIA team Avalon\\Maison de la simulation (CEA)}
-\date{$16^{th}$ November 2015}
+\date{$5{th}$ JLESC Workshop - 28th June 2016, Lyon}
 %-------------------------------------------------------------------
@@
 %---------------
-%---------------
-\begin{withoutheadline}
 
@@
 
+%-------------------------------------------------------------
 % INTRODUCTION / MOTIVATION
 %-------------------------------------------------------------
+\section{Introduction}
+%-------------------------------------------------------------
 \begin{frame}
-\frametitle{Motivation} % HPC DSL
+\frametitle{Domain Specific Languages} % HPC DSL
 \begin{block}{+ Domain Specific Languages}
+Separation of concerns between the domain and its efficient implementation
 \begin{itemize}
-\item Easy language for the user
+\item Easy language for the end-user
 \item Implicit optimizations
@@
 \end{block}
+But none or only a few DSLs used in production !
+\end{frame}
+%-------------------------------------------------------------
+\begin{frame}
+\frametitle{Domain Specific Languages} % HPC DSL
 \begin{alertblock}{- Domain Specific Languages}
 \begin{itemize}
-\item Difficulties deported to the DSL designer
+\item Choose a domain and a language more or less specific
+\item Difficult to combine DSLs (interoperability)
 \begin{itemize}
-\item Low level high performance programming
-\item Maintainability and portability
+\item semantic / compilation / back-end codes
 \end{itemize}
-\item As many DSLs as domains
-\begin{itemize}
-\item Productivity and code-reuse
-\item Separation of concerns
-\end{itemize}
+\item Difficult to maintain
+\item Exascale applications = many specific domains and interactions
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Motivation} % components
-\begin{block}{Component models}
+\frametitle{How to improve it ?} % HPC DSL
 \begin{itemize}
-\item Divide an application into several functionnal black boxes
-\item Each component defines its connections with outer world
-\item Each component are built independently from the others
-\item Application = Assembly of components
+\item Choose the good abstraction level to
+\begin{itemize}
+\item stay efficient
+\item be used in many domains (find common meta-domains)
 \end{itemize}
-\end{block}
-\begin{block}{+ Component models}
+\item Choose good programming models for
 \begin{itemize}
-\item Maintainability through separation of concerns
-\item Code-reuse and productivity
-\item Dynamic assembly of components
+\item the compiler programming
+\item the back-end
 \end{itemize}
-\end{block}
-\end{frame}
-%-------------------------------------------------------------
-\begin{frame}
-\frametitle{Motivation} % components
-What if a DSL produces a component-based runtime ?
-
-\begin{itemize}
-\item Is it efficient ?
-\item Is it feasible ?
-\item Does it improve issues of DSLs ?
-\begin{itemize}
-\item maintainability
-\item portability
-\item productivity
+\item Try to reuse existing DSLs (compilation process or back-end codes)
 \end{itemize}
-\end{itemize}
-
-Let's take a useful example: \textit{the Multi-Stencil Language }!
+\textit{Delite (Scala)\\
+Composition and Reuse with Compiled Domain-Specific Languages\\
+ECOOP'13}
 \end{frame}
@@
 
-% presentation of the table of contents
+%-------------------------------------------------------------
+\section{Multi-Stencil Language : Yet another DSL for stencils !}
 %-------------------------------------------------------------
 \begin{frame}
-\frametitle{Table of contents}
-    \tableofcontents[hideallsubsections]
-\end{frame}
-%-------------------------------------------------------------
-
-\end{withoutheadline}
-%---------------
-%---------------
-
-%-------------------------------------------------------------
-\section{Multi-Stencil Language}
-% related work, multi-stencil problem, language
-%-------------------------------------------------------------
-%----------------
-\subsection{Multi-Stencil Language}
-%----------------
-\begin{frame}
-\frametitle{Numerical simulation = Multi-Stencil application}
+\frametitle{Multi-Stencil application}
 \begin{center}
@@
   \node[purp,align=center, below of=plus, node distance=3.5cm] (sche) {Explicit\\numerical schemes\\\textcolor{red}{stencils}};
-  \node[progm,align=center,right of=sche,node distance=5.5cm] (expl) {\scriptsize Step by step approximation of the phenomena};
+  \node[progm, text width=11em,align=center,right of=sche,node distance=5.5cm] (expl) {\scriptsize Iterative neighboorhood approximation of the phenomena};
   \path[line,dashed] (sche) -- (expl);
-  \path[line] (plus) -- node [fill=white,align=center] {\scriptsize Numerical mathods\\\scriptsize Finite difference/volume/elment} (sche);
+  \path[line] (plus) -- node [fill=white,align=center] {\scriptsize Numerical methods\\\scriptsize Finite difference/volume/elment} (sche);
 \end{tikzpicture}
@@
 %----------------
-
-%----------------
 \begin{frame}
-\frametitle{Time and Mesh}
-\begin{block}{Time}
-At each time iteration of the simulation are applied the \emph{computations kernels} of the application.
-\end{block}
-\begin{block}{Mesh}
+\frametitle{Multi-Stencil Language : Yet another DSL for stencils !}
+\begin{block}{Inputs}
+follow a descriptive grammar to simply declare
 \begin{itemize}
-\item A Mesh is a connected undirected graph $\mathcal{M}=(V,E)$ without bridges
-\item Mesh entities are a subset of $V \cup E$
+\item the sequential order of computations
+\item what is read and written by each computation
+\item whith or without a neighborhood access (stencil)
 \end{itemize}
 \end{block}
-\begin{center}
-\resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
-\end{center}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Data and Computation Kernels}
-\begin{block}{Data}
-A Data is a set of numerical values, each one attached to a given mesh entity
-\end{block}
-\begin{block}{Computation kernel}
+\begin{block}{Outputs}
 \begin{itemize}
-\item Set of data read for the computation
-\begin{itemize}
-\item Each one associated to a stencil shape
-\end{itemize}
-\item Data written by the computaion
-\item A numerical expression
-\item A computation domain
-\begin{itemize}
-\item Subset of mesh entities
-\end{itemize}
+\item parallel pattern of the multi-stencil application
+\item hybrid parallelism included: data and task parallelism
+\item choice of parallel languages (MPI, OpenMP, StarPU, SkelGIS, Global Arrays etc.)
+\item end-users have to fill computation functions into this pattern
 \end{itemize}
@@
 %----------------
-
-%----------------
-% \begin{frame}
-% \frametitle{Computation kernel: Examples}
-% \begin{center}
-% \resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
-% \hspace{10pt}
-% \resizebox{5cm}{!}{\includegraphics{./images/stencil1.pdf}}
-% \end{center}
-% Compute $A$ using $\{(B,n1)\}$ on the computation domain called $dc1$ with the numerical expression
-% \tiny
-% \begin{equation*}
-% A(x,y)=B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1).
-% \end{equation*}
-% \end{frame}
-%----------------
-
-%----------------
 \begin{frame}
-\frametitle{Computation kernel: Examples}
-\begin{center}
-\resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
-\hspace{10pt}
-\resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
-\end{center}
-Compute $A$ using $\{(A,local),(C,n2)\}$ on the computation domain called $dc2$ with the numerical expression
-\tiny
-\begin{equation*}
-A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1).
-\end{equation*}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Multi-Stencil program}
-\begin{center}
-$\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$
-\end{center}
+\frametitle{Multi-Stencil Language : Yet another DSL for stencils !}
 \begin{itemize}
-\item $T$ the set of time iterations to tun the simulation
-\item $\mathcal{M}$ the mesh of the simulation
-\item $\mathcal{E}$ the set of mesh entities
-\item $\mathcal{D}$ the set of computation domains
-\item $\Delta$ the set of data
-\item $\Gamma$ the set of computations
+\item Choosen abstraction level
+\begin{itemize}
+\item mesh-agnostic description language
+\begin{itemize}
+\item should be usable for unstructured, Cartesian, curvilinear meshes
 \end{itemize}
-\center \textit{= the six sections of a \emph{Multi-Stencil Language} program !}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Multi-Stencil Language}
-\begin{alertblock}{MSL is not}
+\item description concerns splited from implementation concerns
 \begin{itemize}
-\item a new stencil optimizer/compiler
-\item a new distributed data structure
-\end{itemize}
-\end{alertblock}
-\begin{block}{MSL is}
-\begin{itemize}
-\item a high-level language for multi-stencil simulations
-\item agnostic from the type of mesh used (data structure)
-\item based on identifiers only
-\end{itemize}
-\end{block}
-\begin{center}
-\textit{MSL produces a ready-to-fill\\component-based parallel pattern of the simulation !}
-\end{center}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Related Work}
-\begin{block}{Complementary work}
-\begin{itemize}
-\item Distributed data structures: SkelGIS, Global Arrays
-\item Stencil DSLs (on grids): Pochoir, PATUS
-\item Stencil DSLs (on unstructured meshes): OP2, Liszt
-\end{itemize}
-\end{block}
-\begin{alertblock}{Similar work}
-\begin{itemize}
-\item Pipeline of stencil computations for image processing: Halide
-\item DSL to component-based runtime: ?
-\end{itemize}
-\end{alertblock}
-\end{frame}
-%----------------
-%-------------------------------------------------------------
-
-%-------------------------------------------------------------
-\section{Overview}
-% take the same notations inside components + user
-%-------------------------------------------------------------
-\subsection{Overview}
-
-%----------------
-\begin{frame}
-\frametitle{MSL to Component-based runtime}
-\begin{block}{Ready-to-fill parallel pattern}
-\begin{itemize}
-\item Data parallelism
-\begin{itemize}
-\item External distributed data structure
-\item Automatic detection of synchronizations
-\end{itemize}
-\item Task parallelism (mid-grain)
-\begin{itemize}
-\item Compile a static scheduling of computation kernels
+\item usable with many different parallel libraries or languages back-ends
 \end{itemize}
 \end{itemize}
-\end{block}
-The fine grain task parallelism is left to other languages: 
+\item Choosen programming models for back-end
 \begin{itemize}
-\item OpenMP in the kernels
-\item Kernels generated by stencil compilers (Pochoir, PATUS, Liszt etc.)
+\item Component programming model (maintainability, portability and composability)
+\item Task scheduling model (portability and efficiency)
+\end{itemize}
+\item Reuse of existing languages and DSLs
+\begin{itemize}
+\item Use of a Distributed Data Structure DSL (MPI) : SkelGIS
+\item MPI and OpenMP (4.x)
+\end{itemize}
 \end{itemize}
 \end{frame}
-%----------------
-
 %----------------
@@
 \begin{center}
-$\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$\\
-
-\bigskip
-\begin{tikzpicture}[remember picture]
-  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
-    \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-     \node[component,solid,draw=red] (D) at (0,0) {$Driver$};
-     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
-     \node[solid] (Ds) at (-1.5,0) {start};
-     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
-     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
-     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};
-
-     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
-     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
-     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
-     \node[solid,left=0.8cm of T] (tt) {$T$};
-
-     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
-     \node[component,dashed,right=2cm of Cp] (C) {$Computations$};
-     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
-     \node[solid,right=0.2cm of Cu] (star) {$*$};
-     \node[solid,right=1.5cm of C] (gamma) {$\Gamma$};
-
-     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
-     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
-     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
-     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
-     \node[solid,left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
-     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
-     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};
-
-     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
-     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
-     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
-     \node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
-
-    \path[-]
-      (Dp) edge [solid] node {} (D)
-      (D) edge [solid] node {} (Du1)
-          edge [solid] node {} (Du2)
-          edge [solid] node {} (Du3)
-      (DDSp1) edge [solid] node {} (DDS)
-      (Tp) edge [solid] node {} (T)
-      (T)  edge [solid] node {} (Tu)
-      (Cp) edge [solid] node {} (C)
-      (C) edge [solid] node {} (Cu)
-      (Datap2) edge [solid] node {} (Data)
-      (Datap1) edge [solid] node {} (Data)
-      (Data) edge [solid] node {} (Datau)
-      (DDSp2) edge [solid] node {} (DDS);
-    \end{tikzpicture}
-    };
-    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=0.5cm of proc] (mpidata) {};
-    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
-    \node[solid,below=2.5cm of proc.west,anchor=west] (core) {\textit{\tiny{Duplicated on each processor/core}}};
-    \path[-]
-      (Data.east) edge node {} (mpidata);
-  \end{tikzpicture}
-\end{center}
-\end{frame}
-%-------------------------------------------------------------
-
-%----------------
-\begin{frame}
-\frametitle{MSL to Component-based runtime}
-\begin{center}
-$\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$\\
-
-\bigskip
 \begin{tikzpicture}[remember picture]
@@
      \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
-     \node[solid,left=0.8cm of T] (tt) {$T$};
 
-     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
-     \node[component,dashed,right=2cm of Cp] (C) {$Computations$};
+     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Sp) {};
+     \node[component,solid,right=1.5cm of Sp] (S) {$Scheduling$};
+     \node[use,scale=0.5,solid,right=1.5cm of S] (Su) {};
+
+     \node[provide,scale=0.5,solid,right=0.15 of Su] (Cp) {};
+     \node[component,solid,double,right=2cm of Cp] (C) {$Kernel$};
      \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
-     \node[solid,right=0.2cm of Cu] (star) {$*$};
-     \node[solid,right=1.5cm of C] (gamma) {$\Gamma$};
 
@@
      \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
-     \node[component,solid,draw=red,double,above=0.8cm of Datap2] (Data) {$Data$};
+     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
      \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
-     \node[solid,left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
-     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
-     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};
 
@@
      \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
-     \node[component,draw=red,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
-     \node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
+     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
 
@@
       (T)  edge [solid] node {} (Tu)
-      (Cp) edge [solid] node {} (C)
-      (C) edge [solid] node {} (Cu)
+      (Sp) edge [solid] node {} (S)
+      (S) edge [solid] node {} (Su)
       (Datap2) edge [solid] node {} (Data)
@@
       (Data) edge [solid] node {} (Datau)
+      (C) edge [solid] node {} (Cu)
+      	  edge [solid] node {} (Cp)
       (DDSp2) edge [solid] node {} (DDS);
@@
 %-------------------------------------------------------------
-
+\section{Few results}
 %----------------
 \begin{frame}
-\frametitle{MSL to Component-based runtime}
+\frametitle{Shallow-water equations}
 \begin{center}
-$\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$\\
+Cartesian mesh / finite volume method\\Strong scaling 10k x 10k mesh with 1k iterations\\TGCC Curie, Thin nodes
+\resizebox{7cm}{!}{\includegraphics{./images/median_strong.pdf}}
+\end{center}
+\end{frame}
+%----------------
+\begin{frame}
+\frametitle{Shallow-water equations}
+\begin{center}
+Cartesian mesh / finite volume method\\Weak scaling 400 x 400 per process\\TGCC Curie, Thin nodes
+\resizebox{7cm}{!}{\includegraphics{./images/median_weak.pdf}}
+\end{center}
+\end{frame}
+%----------------
+% \begin{frame}
+% \frametitle{Multi-Stencil Language : Yet another DSL for stencils !}
+% \textbf{Improvment with data+task scheduling}
+% \begin{center}
+% \resizebox{8cm}{!}{\includegraphics{./images/base_close_median.pdf}}
+% \end{center}
+% \end{frame}
 
-\bigskip
-\begin{tikzpicture}[remember picture]
-  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
-    \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-     \node[component,solid] (D) at (0,0) {$Driver$};
-     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
-     \node[solid] (Ds) at (-1.5,0) {start};
-     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
-     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
-     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};
-
-     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
-     \node[component,solid,draw=red,below=1.6cm of Tp] (T) {$Time$};
-     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
-     \node[solid,left=0.8cm of T] (tt) {$T$};
-
-     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
-     \node[component,dashed,draw=red,right=2cm of Cp] (C) {$Computations$};
-     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
-     \node[solid,right=0.2cm of Cu] (star) {$*$};
-     \node[solid,right=1.5cm of C] (gamma) {$\Gamma$};
-
-     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
-     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
-     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
-     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
-     \node[solid,left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
-     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
-     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};
-
-     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
-     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
-     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
-     \node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
-
-    \path[-]
-      (Dp) edge [solid] node {} (D)
-      (D) edge [solid] node {} (Du1)
-          edge [solid] node {} (Du2)
-          edge [solid] node {} (Du3)
-      (DDSp1) edge [solid] node {} (DDS)
-      (Tp) edge [solid] node {} (T)
-      (T)  edge [solid] node {} (Tu)
-      (Cp) edge [solid] node {} (C)
-      (C) edge [solid] node {} (Cu)
-      (Datap2) edge [solid] node {} (Data)
-      (Datap1) edge [solid] node {} (Data)
-      (Data) edge [solid] node {} (Datau)
-      (DDSp2) edge [solid] node {} (DDS);
-    \end{tikzpicture}
-    };
-    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=0.5cm of proc] (mpidata) {};
-    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
-    \node[solid,below=2.5cm of proc.west,anchor=west] (core) {\textit{\tiny{Duplicated on each processor/core}}};
-    \path[-]
-      (Data.east) edge node {} (mpidata);
-  \end{tikzpicture}
-\end{center}
+%-------------------------------------------------------------
+\begin{frame}
+\frametitle{Work in progress and collaborations}
+\begin{block}{Work in progress}
+\begin{itemize}
+\item different types of scheduling and back-end in MSL
+\item combine MSL implementation choices to an energy-aware framework
+\item new programming model back-end for MSL: combining component models and task models (Ph.D. Jérôme Richard)
+\end{itemize}
+\end{block}
+\begin{block}{Collaborations}
+\begin{itemize}
+\item More application cases
+\item Back-end programming model: component models and task scheduling (Components in OmpSs ?)
+\item DSLs interoperability
+\end{itemize}
+\end{block}
 \end{frame}
@@
 
-%-------------------------------------------------------------
-\section{Compiler}
-% A slide example MS, 6 slides on graph transformations
-%-------------------------------------------------------------
-\subsection{Compiler}
-%----------------
-\begin{frame}[fragile]
-\frametitle{Example}
-\begin{columns}
-\begin{column}{.52\textwidth}
-\begin{lstlisting}[style=base]
-@mesh:@ cart
-@mesh entities:@ cell,edgex,edgey
-@computation domains:@
-  allcell in cell
-  alledgex in edgex
-  alledgey in edgey
-  part1edgex in edgex
-  part2edgex in edgex
-@data:@
-  a,cell
-  b,cell
-  c,edgex
-  d,edgex
-  e,edgey
-\end{lstlisting}
-\end{column}
-\begin{column}{.48\textwidth}
-\begin{lstlisting}[style=base]
-  f,cell
-  g,edgey
-  h,edgex
-  i,cell
-  j,edgex
-@time:@500
-@computations:@
-  b[allcell]=&c0&(a)
-  c[alledgex]=&c1&(b[n1])
-  d[alledgex]=&c2&(c)
-  e[alledgey]=&c3&(c)
-  f[allcell]=&c4&(d[n1])
-  g[alledgey]=&c5&(e)
-  h[alledgex]=&c6&(f)
-  i[allcell]=&c7&(g,h)
-  j[partedgex]=&c8&(i[n1])
-\end{lstlisting}
-\end{column}
-\end{columns}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Data parallelism}
-$\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$\\
-$\hookrightarrow [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$
-
-\medskip
-\begin{block}{Synchronization}
-For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronization} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j$ read a data written by $c_i$ with a stencil shape and not locally.
-\end{block}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Mid-grain task parallelism}
-
-Build the read/write dependency graph.
-
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node (c0) at (0,0) {$c_0$};
-   \node[right=1 of c0] (sy1) {$sync_1$};
-   \node[right=1 of sy1] (c1) {$c_1$};
-   \node[above right=1 of c1] (c2) {$c_2$};
-   \node[below right=1 of c1] (c3) {$c_3$};
-   \node[right=1 of c2] (sy4) {$sync_4$};
-   \node[right=1 of c3] (c5) {$c_5$};
-   \node[right=1 of sy4] (c4) {$c_4$};
-   \node[right=1 of c4] (c6) {$c_6$};
-   \node[below right=1 of c6] (c7) {$c_7$};
-   \node[right=1 of c7] (sy8) {$sync_8$};
-   \node[right=1 of sy8] (c8) {$c_8$};
-
-  \path[->]
-    (c0) edge node {} (sy1)
-    (sy1) edge node {} (c1)
-    (c1)  edge node {} (c2)
-          edge node {} (c3)
-    (c2) edge node {} (sy4)
-    (sy4) edge node {} (c4)
-    (c4) edge node {} (c6)
-    (c3) edge node {} (c5)
-    (c5) edge node {} (c7)
-    (c6) edge node {} (c7)
-    (c7) edge node {} (sy8)
-    (sy8) edge node {} (c8);
-\end{tikzpicture}
-\end{center}
-
-\begin{center}
-\textbf{Data parallelism + Task parallelism}
-\end{center}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Which scheduling ?}
-From the dependency graph we can:
-\begin{itemize}
-\item Use a dynamic scheduler
-\item Compute a static scheduling
-\end{itemize}
-
-In this work,
-\begin{itemize}
-\item We build a static scheduling of the multi-stencil application
-\item This static scheduling can be dumped to a component-based runtime
-\end{itemize}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Series-Parallel Tree}
-\begin{columns}
-\begin{column}{.48\textwidth}
-\begin{center}
-\resizebox{\textwidth}{!}{%
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (s0)    at (0,0) {$\mathcal{S}$};
-
-   \node[] (c0)    at (-3,1) {$c_0$};
-   \node[] (star1) at (-2,1) {$sync_1$};
-   \node[] (c1)    at (-1,1) {$c_1$};
-   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
-   \node[] (c7)    at ( 1,1) {$c_7$};
-   \node[] (star8) at ( 2,1) {$sync_8$};
-   \node[] (c8)    at ( 3,1) {$c_8$};
-
-   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
-   \node[] (s2)    at (1,2) {$\mathcal{S}$};
-
-   \node[] (c2)    at (-3,3) {$c_2$};
-   \node[] (star4) at (-2,3) {$sync_4$};
-   \node[] (c4)    at (-1,3) {$c_4$};
-   \node[] (c6)    at ( 0,3) {$c_6$};
-   \node[] (c3)    at ( 1,3) {$c_3$};
-   \node[] (c5)    at ( 2,3) {$c_5$};
-
-  \path[->]
-    (s0) edge node {} (c0)
-         edge node {} (star1)
-         edge node {} (c1)
-         edge node {} (p0)
-         edge node {} (c7)
-         edge node {} (star8)
-         edge node {} (c8)
-    (p0) edge node {} (s1)
-         edge node {} (s2)
-    (s1) edge node {} (c2)
-         edge node {} (star4)
-         edge node {} (c4)
-         edge node {} (c4)
-         edge node {} (c6)
-    (s2) edge node {} (c3)
-         edge node {} (c5);
-  \end{tikzpicture}
-  }
-\end{center}
-\end{column}
-\begin{column}{.52\textwidth}
-\begin{block}{Specific components}
-\begin{itemize}
-\item $SEQ$ to directly replace $\mathcal{S}$ nodes
-\item $PAR$ to directly replace $\mathcal{P}$ nodes
-\item $SYNC$ for synchronizations
-\item $K$ for computation kernels
-\end{itemize}
-\end{block}
-\end{column}
-\end{columns}
-\begin{center}
-\textbf{Direct dump to the component assembly}
-\end{center}
-\end{frame}
-%----------------
-
-% %----------------
-% \begin{frame}
-% \frametitle{Component-based runtime}
-% \begin{block}{Specific components}
-% \begin{itemize}
-% \item $SEQ$ to directly replace $\mathcal{S}$ nodes
-% \item $PAR$ to directly replace $\mathcal{P}$ nodes
-% \item $SYNC$ for synchronizations
-% \item $K$ for computation kernels
-% \end{itemize}
-% \end{block}
-
+% \begin{withoutheadline}
+% \begin{frame}{}
 % \begin{center}
-% \textbf{Direct dump to the component assembly}\\
-% \textbf{Parallel component-based skeleton of the multi-stencil application}
+% \huge Thank You !
 % \end{center}
-
 % \end{frame}
-% %----------------
-
-%----------------
-\begin{frame}
-\frametitle{Component-based runtime}
-
-\begin{center}
-\resizebox{\textwidth}{!}{%
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-     \node[component,solid] (D) at (0,0) {$Driver$};
-     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
-     %\node[solid] (Ds) at (-1.5,0) {start};
-     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
-     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
-     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};
-
-     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
-     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
-     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
-     %\node[solid,left=0.8cm of T] (tt) {$T$};
-
-     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
-     \node[component,dashed,right=2cm of Cp] (C) {$Computations$};
-     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
-     \node[solid,right=0.2cm of Cu] (star) {$*$};
-     %\node[solid,right=1.5cm of C] (gamma) {$\Gamma$};
-
-     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
-     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
-     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
-     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
-     %\node[solid,left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
-     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
-     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};
-
-     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
-     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
-     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
-     %\node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
-
-    \path[-]
-      (Dp) edge [solid] node {} (D)
-      (D) edge [solid] node {} (Du1)
-          edge [solid] node {} (Du2)
-          edge [solid] node {} (Du3)
-      (DDSp1) edge [solid] node {} (DDS)
-      (Tp) edge [solid] node {} (T)
-      (T)  edge [solid] node {} (Tu)
-      (Cp) edge [solid] node {} (C)
-      (C) edge [solid] node {} (Cu)
-      (Datap2) edge [solid] node {} (Data)
-      (Datap1) edge [solid] node {} (Data)
-      (Data) edge [solid] node {} (Datau)
-      (DDSp2) edge [solid] node {} (DDS);
-    %\end{tikzpicture}
-    %}
-% \end{column}
-% \begin{column}{.70\textwidth}
-% \begin{center}
-% \resizebox{.80\textwidth}{!}{%
-% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   %seq0
-   \node[component,right=3.5cm of C] (seq0) {$SEQ$};
-   \node[provide] (seq0p) [left = 0.8cm of seq0] {};
-   \node[use] (seq0u) [right = 1cm of seq0] {$m$};
-   %c0
-   \node[component] (c0) [right = 2cm of seq0u] {$K(c_0)$};
-   \node[use] (c0u) [right = 0.8cm of c0] {};
-   \node[right=0.2 of c0u] (star) {*};
-   %sync0
-   \node[component] (sync0) [below = 1cm of c0] {$SYNC$};
-   \node[use] (sync0u) [right = 1cm of sync0] {$m$};
-   %c1
-   \node[component] (c1) [below = 0.8cm of sync0] {$K(c_1)$};
-   \node[use] (c1u) [right = 0.8cm of c1] {};
-   \node[right=0.2 of c1u] (star) {*};
-   %par0
-   \node[component] (par0) [below = 0.8cm of c1] {$PAR$};
-   \node[use] (par0u) [right = 1cm of par0] {$m$};
-  %seq1
-   \node[component] (seq1) [right = 1cm of sync0u] {$SEQ$};
-   \node[use] (seq1u) [right = 1cm of seq1] {$m$};
-  %c2
-   \node[component] (c2) [right = 4cm of c0] {$K(c_2)$};
-   \node[use] (c2u) [right = 0.8cm of c2] {};
-   \node[right=0.2 of c2u] (star) {*};
-  %sync1
-   \node[component] (sync1) [below = 0.8cm of c2] {$SYNC$};
-   \node[use] (sync1u) [right = 1cm of sync1] {$m$};
-  %c4
-   \node[component] (c4) [below = 0.8cm of sync1] {$K(c_4)$};
-   \node[use] (c4u) [right = 0.8cm of c4] {};
-   \node[right=0.2 of c4u] (star) {*};
-  %c6
-   \node[component] (c6) [below = 0.8cm of c4] {$K(c_6)$};
-   \node[use] (c6u) [right = 0.8cm of c6] {};
-   \node[right=0.2 of c6u] (star) {*};
-  %seq2
-   \node[component] (seq2) [below = 3.2cm of seq1] {$SEQ$};
-   \node[use] (seq2u) [right = 1cm of seq2] {$m$};
-  %c3
-   \node[component] (c3) [right = 1cm of seq2u] {$K(c_3)$};
-   \node[use] (c3u) [right = 0.8cm of c3] {};
-   \node[right=0.2 of c3u] (star) {*};
-  %c5
-   \node[component] (c5) [below = 0.8cm of c3] {$K(c_5)$};
-   \node[use] (c5u) [right = 0.8cm of c5] {};
-   \node[right=0.2 of c5u] (star) {*};
-  %c7
-   \node[component] (c7) [below = 0.8cm of par0] {$K(c_7)$};
-   \node[use] (c7u) [right = 0.8cm of c7] {};
-   \node[right=0.2 of c7u] (star) {*};
-  %sync2
-   \node[component] (sync2) [below = 0.8cm of c7] {$SYNC$};
-   \node[use] (sync2u) [right = 1cm of sync2] {$m$};
-  %c8
-   \node[component] (c8) [below = 0.8cm of sync2] {$K(c_8)$};
-   \node[use] (c8u) [right = 0.8cm of c8] {};
-   \node[right=0.2 of c8u] (star) {*};
-
-   \path[-]
-   %seq0
-    (seq0) edge node {} (seq0u)
-    (seq0p) edge node {} (seq0)
-   %c0
-    (seq0u) edge node {} (c0)
-    (c0) edge node {} (c0u)
-    %sync0
-    (seq0u) edge node {} (sync0.west)
-    (sync0) edge node {} (sync0u)
-    %c1
-    (seq0u) edge node {} (c1)
-    (c1) edge node {} (c1u)
-   %par0
-    (seq0u) edge node {} (par0.west)
-    (par0) edge node {} (par0u)
-  %seq1
-    (par0u) edge node {} (seq1.west)
-    (seq1) edge node {} (seq1u)
-  %c2
-    (seq1u) edge node {} (c2.west)
-    (c2) edge node {} (c2u)
-  %sync0
-    (seq1u) edge node {} (sync1.west)
-    (sync1) edge node {} (sync1u)
-  %c4
-    (seq1u) edge node {} (c4.west)
-    (c4) edge node {} (c4u)
-  %c6
-    (seq1u) edge node {} (c6.west)
-    (c6) edge node {} (c6u)
-  %seq2
-    (par0u) edge node {} (seq2.west)
-    (seq2) edge node {} (seq2u)
-  %c3
-    (seq2u) edge node {} (c3)
-    (c3) edge node {} (c3u)
-  %c6
-    (seq2u) edge node {} (c5.west)
-    (c5) edge node {} (c5u)
-  %c6
-    (seq0u) edge node {} (c7.west)
-    (c7) edge node {} (c7u)
-  %sync2
-    (seq0u) edge node {} (sync2.west)
-    (sync2) edge node {} (sync2u)
-  %c8
-    (seq0u) edge [bend right] node {} (c8.west)
-    (c8) edge node {} (c8u)
-   ;
-
-   \path[-]
-    (C) edge[dashed] node {} (seq0);
-\end{tikzpicture}
-}
-\end{center}
-% \end{column}
-% \end{columns}
-\end{frame}
-%----------------
-
-%----------------
-\begin{frame}
-\frametitle{Back to Data Parallelism}
-\begin{center}
-\textbf{Back to Data Parallelism only ?}\\
-\end{center}
-\textbf{Loop Fusion}: Same computation domain under a sequence or a parallel node
-
-\medskip
-\begin{center}
-\resizebox{0.5\textwidth}{!}{%
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[] (s0)    at (0,0) {$\mathcal{S}$};
-
-   \node[] (c0)    at (-3,1) {$c_0$};
-   \node[] (star1) at (-2,1) {$sync_1$};
-   \node[] (c1)    at (-1,1) {$c_1$};
-   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
-   \node[] (c7)    at ( 1,1) {$c_7$};
-   \node[] (star8) at ( 2,1) {$sync_8$};
-   \node[] (c8)    at ( 3,1) {$c_8$};
-
-   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
-   \node[] (s2)    at (1,2) {$\mathcal{S}$};
-
-   \node[] (c2)    at (-3,3) {$c_2$};
-   \node[] (star4) at (-2,3) {$sync_4$};
-   \node[] (c4)    at (-1,3) {$c_4$};
-   \node[] (c6)    at ( 0,3) {$c_6$};
-   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
-   %\node[] (c5)    at ( 2,3) {$c_5$};
-
-  \path[->]
-    (s0) edge node {} (c0)
-         edge node {} (star1)
-         edge node {} (c1)
-         edge node {} (p0)
-         edge node {} (c7)
-         edge node {} (star8)
-         edge node {} (c8)
-    (p0) edge node {} (s1)
-         edge node {} (s2)
-    (s1) edge node {} (c2)
-         edge node {} (star4)
-         edge node {} (c4)
-         edge node {} (c4)
-         edge node {} (c6)
-    (s2) edge node {} (c3);
-  \end{tikzpicture}
-  }
-\end{center}
-
-\small
-\textbf{Canonical form}: $[c_0,sync_1,c_1,c_2,sync_4,c_4,c_6,\{c_3,c_5\},c_7,sync_8,c_8]$
-\end{frame}
-%----------------
-
-%-------------------------------------------------------------
-
-%-------------------------------------------------------------
-\section{Evaluation}
-% Present Shallow-water case, present results
-%-------------------------------------------------------------
-\subsection{Evaluation}
-\begin{frame}
-\frametitle{Implementation and evaluation}
-\textbf{Implementation of MSL}: Python, SkelGIS and $L^2C$
-
-\medskip
-\textbf{Shallow-water equations}: 1 mesh, 3 mesh entities, 7 computation domains, 48 data, 98 computations (32 stencils, 66 local kernels)
-
-\medskip
-\textbf{Evaluation of the data parallelism}
-\begin{itemize}
-\item Full SkelGIS implementation (DDS + specific interfaces to hide communications)
-\item MSL implementation which uses the SkelGIS DDS
-\item Cluster Edel of Grid'5000: 2 8-cores Intel Xeon E5520, 24GB RAM / nodes
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Evaluations}
-\begin{columns}
-\begin{column}{.5\textwidth}
-\begin{center}
-\small Mesh $10k \times 10k$, $500$ iterations\\
-\resizebox{\textwidth}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
-\end{center}
-\end{column}
-\begin{column}{.5\textwidth}
-\begin{center}
-\small Mesh $10k \times 10k$, $2K$ iterations\\
-\resizebox{\textwidth}{!}{\includegraphics{./images/logtimes_10K_2K.pdf}}
-\end{center}
-\end{column}
-\end{columns}
-\end{frame}
-%-------------------------------------------------------------
-
-%-------------------------------------------------------------
-\section{Conclusion and perspectives}
-%-------------------------------------------------------------
-\subsection{Conclusion and perspectives}
-\begin{frame}
-\frametitle{Conclusion}
-\begin{itemize}
-\item A DSL for Multi-Stencil applications (MSL)
-\item The compilation of MSL to get a static scheduling
-\item The dump to a component-based runtime
-\item No overhead introduced by the solution
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Perspectives}
-\begin{itemize}
-\item A comparison with an improved version of SkelGIS
-\item More evaluations on TGCC Curie (CEA) up to $32k$ cores
-\item An evaluation of the Data+Task parallelism (mixing the static scheduling and OpenMP)
-\item A new dump with a component responsible for the dynamic scheduling of the application (using OpenMP 4.x)
-\item A new dump to CPU+GPGPUs (using existing stencil compilers Pochoir, PATUS etc.)
-\item An alternative to the SkelGIS DDS using Global Arrays (PGAS)
-\end{itemize}
-\end{frame}
-%-------------------------------------------------------------
-
-\begin{withoutheadline}
-\begin{frame}{}
-\begin{center}
-\huge Thank You !
-\end{center}
-\end{frame}
-\end{withoutheadline}
+% \end{withoutheadline}
 

### 2016-06-27 ARTICLES/OTHER/JLESC/slides.tex
JLESC presentation

@@
 \begin{itemize}
+\item Eexascale applications = many specific domains and interactions
 \item semantic / compilation / back-end codes
@@
 \item Difficult to maintain
-\item Exascale applications = many specific domains and interactions
 \end{itemize}
@@
 \end{itemize}
-\textit{Delite (Scala)\\
-Composition and Reuse with Compiled Domain-Specific Languages\\
-ECOOP'13}
+% \textit{Delite (Scala)\\
+% Composition and Reuse with Compiled Domain-Specific Languages\\
+% ECOOP'13}
 \end{frame}
@@
 \item what is read and written by each computation
-\item whith or without a neighborhood access (stencil)
+\item with or without a neighborhood access (stencil)
 \end{itemize}
@@
 \end{frame}
+%-------------------------------------------------------------
+\section{Few results}
+%----------------
+\begin{frame}
+\frametitle{Shallow-water equations}
+\begin{center}
+Cartesian mesh / finite volume method\\Strong scaling 10k x 10k mesh with 1k iterations\\TGCC Curie, Thin nodes
+\resizebox{7cm}{!}{\includegraphics{./images/median_strong.pdf}}
+\end{center}
+\end{frame}
+%----------------
+\begin{frame}
+\frametitle{Shallow-water equations}
+\begin{center}
+Cartesian mesh / finite volume method\\Weak scaling 400 x 400 per process\\TGCC Curie, Thin nodes
+\resizebox{7cm}{!}{\includegraphics{./images/median_weak.pdf}}
+\end{center}
+\end{frame}
+%----------------
+% \begin{frame}
+% \frametitle{Multi-Stencil Language : Yet another DSL for stencils !}
+% \textbf{Improvment with data+task scheduling}
+% \begin{center}
+% \resizebox{8cm}{!}{\includegraphics{./images/base_close_median.pdf}}
+% \end{center}
+% \end{frame}
+
+%-------------------------------------------------------------
+\begin{frame}
+\frametitle{Work in progress and collaborations}
+\begin{block}{Work in progress}
+\begin{itemize}
+\item different types of scheduling and back-end in MSL
+\item combine MSL implementation choices to an energy-aware framework
+\item new programming model back-end for MSL: combining component models and task models (Ph.D. Jérôme Richard)
+\end{itemize}
+\end{block}
+\begin{block}{Collaborations}
+\begin{itemize}
+\item More application cases
+\item Back-end programming model: component models and task scheduling (Components in OmpSs ?)
+\item DSLs interoperability
+\end{itemize}
+\end{block}
+\end{frame}
+%-------------------------------------------------------------
+
+% \begin{withoutheadline}
+% \begin{frame}{}
+% \begin{center}
+% \huge Thank You !
+% \end{center}
+% \end{frame}
+% \end{withoutheadline}
+
 %----------------
@@
 \end{frame}
-%-------------------------------------------------------------
-\section{Few results}
-%----------------
-\begin{frame}
-\frametitle{Shallow-water equations}
-\begin{center}
-Cartesian mesh / finite volume method\\Strong scaling 10k x 10k mesh with 1k iterations\\TGCC Curie, Thin nodes
-\resizebox{7cm}{!}{\includegraphics{./images/median_strong.pdf}}
-\end{center}
-\end{frame}
-%----------------
-\begin{frame}
-\frametitle{Shallow-water equations}
-\begin{center}
-Cartesian mesh / finite volume method\\Weak scaling 400 x 400 per process\\TGCC Curie, Thin nodes
-\resizebox{7cm}{!}{\includegraphics{./images/median_weak.pdf}}
-\end{center}
-\end{frame}
-%----------------
-% \begin{frame}
-% \frametitle{Multi-Stencil Language : Yet another DSL for stencils !}
-% \textbf{Improvment with data+task scheduling}
-% \begin{center}
-% \resizebox{8cm}{!}{\includegraphics{./images/base_close_median.pdf}}
-% \end{center}
-% \end{frame}
-
-%-------------------------------------------------------------
-\begin{frame}
-\frametitle{Work in progress and collaborations}
-\begin{block}{Work in progress}
-\begin{itemize}
-\item different types of scheduling and back-end in MSL
-\item combine MSL implementation choices to an energy-aware framework
-\item new programming model back-end for MSL: combining component models and task models (Ph.D. Jérôme Richard)
-\end{itemize}
-\end{block}
-\begin{block}{Collaborations}
-\begin{itemize}
-\item More application cases
-\item Back-end programming model: component models and task scheduling (Components in OmpSs ?)
-\item DSLs interoperability
-\end{itemize}
-\end{block}
-\end{frame}
-%-------------------------------------------------------------
-
-% \begin{withoutheadline}
-% \begin{frame}{}
-% \begin{center}
-% \huge Thank You !
-% \end{center}
-% \end{frame}
-% \end{withoutheadline}
 

### 2016-06-27 ARTICLES/OTHER/JLESC/slides.tex
JLESC presentation

@@
 \end{block}
-But none or only a few DSLs used in production !
+But none or only a few DSLs used in \textbf{production} !\\
+\textbf{ETP4HPC}: programming models for non experts to reach exascale !
 \end{frame}
@@
 \begin{itemize}
-\item Eexascale applications = many specific domains and interactions
+\item exascale applications = many specific domains and interactions
 \item semantic / compilation / back-end codes
@@
 \item hybrid parallelism included: data and task parallelism
-\item choice of parallel languages (MPI, OpenMP, StarPU, SkelGIS, Global Arrays etc.)
+\item choice of parallel languages (MPI, OpenMP 3.x/4.x, SkelGIS etc.)
 \item end-users have to fill computation functions into this pattern
@@
 \begin{itemize}
-\item Use of a Distributed Data Structure DSL (MPI) : SkelGIS
-\item MPI and OpenMP (4.x)
+\item Use of a Distributed Data Structure DSeL (C++,MPI) : SkelGIS
+\item MPI and OpenMP (3.x and 4.x)
 \end{itemize}
@@
 \begin{itemize}
-\item different types of scheduling and back-end in MSL
-\item combine MSL implementation choices to an energy-aware framework
-\item new programming model back-end for MSL: combining component models and task models (Ph.D. Jérôme Richard)
+\item Different types of scheduling and back-end in MSL
+\item Combine MSL implementation choices to an energy-aware framework
+\item New programming model back-end for MSL: combining component models and task models (Ph.D. Jérôme Richard)
 \end{itemize}

### 2016-10-12 ARTICLES/RR_IJPP/article.tex
Rapport de recherche avec contenu de la soumission IJPP

%%
%% This is file `squelette-rr.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% RR.dtx  (with options: `sample')
%% ********************************************************************
%% Copyright (C) 1997-1999 2004 2006-2011 INRIA/APICS/MARELLE by Jose' Grimm
%% This file may be distributed and/or modified under the
%% conditions of the LaTeX Project Public License, either version 1.3
%% of this license or (at your option) any later version.
%% The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.3 or later is part of all distributions of LaTeX
%% version 2003/12/01 or later.
%% An archive of the software can be found at
%%    ftp://ftp-sop.inria.fr/marelle/RR-INRIA

\documentclass[twoside]{article}
\usepackage[a4paper]{geometry}
\usepackage[latin1]{inputenc} % ou \usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc} % ou \usepackage[OT1]{fontenc}
\usepackage{RR}
\usepackage{hyperref}
%\RequirePackage{fix-cm}
%
%\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb, stmaryrd}
\usepackage{subfig}
\usepackage{caption}
\usepackage{cite}
\usepackage{enumitem}
\usepackage{filecontents}
\usepackage{listings}
\usepackage{mathtools, cuted}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}

\usepackage{todonotes}
\usepackage{url}
\usepackage{xspace}

\newcommand{\fix}[1]{\todo[inline,color=red!80]{#1}}
\newcommand{\cf}{\emph{cf.}\xspace}
\newcommand{\ie}{\emph{i.e.,}\xspace}
\newcommand{\llc}{L\textsuperscript{2}C\xspace}

\newtheorem{theorem}{Theorem}
\newtheorem{myprop}[theorem]{Proposition}
\newtheorem{mydef}[theorem]{Definition}
\newtheorem{myth}[theorem]{Theorem}
\newenvironment{mydefs}[1][Definitions]{\par\medskip\noindent\textbf{#1}\nopagebreak\begin{itemize}[noitemsep,topsep=.25\topsep]}{\end{itemize}}
\def\pprec{\mathrel{\scalebox{.9}[1]{$\prec$}\mkern-3mu%
  \scalebox{.4}[1]{$\prec$}\mkern-5.5mu\scalebox{.4}[1]{$\prec$}}}

\RRNo{7003}
%% \RTNo{0703}
%%
%% date de publication du rapport
\RRdate{October 2016}
%%
%% Cas d'une version deux
%% \RRversion{2}
%% date de publication de la version 2
%% \RRdater{November 2008}
%%
\RRauthor{H\'el\`ene Coullon\thanks[lip]{INRIA, LIP, ENS Lyon, 46 All\'e d'Italie, 69364 Lyon Cedex 07, France}\thanks{helene.coullon@inria.fr} \and Julien Bigot\thanks{Maison de la Simulation, CEA, CNRS, Univ. Paris-Sud, UVSQ, Universit\'e Paris-Saclay, 91191 Gif-sur-Yvette, France}\thanks{julien.bigot@cea.fr} \and Christian Perez\thanksref{lip}\thanks{christian.perez@inria.fr}}

% \institute{H\'el\`ene Coullon \and Christian Perez \at
%               INRIA, LIP, ENS Lyon, 46 Allée d'Italie, 69364 Lyon Cedex 07, France\\
%               \email{helene.coullon@inria.fr} \email{christian.perez@inria.fr}
%            \and
%            Julien Bigot \at
%            Maison de la Simulation, CEA, CNRS, Univ. Paris-Sud, UVSQ, Universit\'e Paris-Saclay, 91191 Gif-sur-Yvette, France\\
%            \email{julien.bigot@cea.fr}
% }

% Jos\'e Grimm\thanks{Footnote for first author}%
%   % note partag\'ee (optionnelle)
%   \thanks[sfn]{Shared foot note}%
%  % \and entre chaque auteur s'il y en a plusieurs
%   \and
% Philippe Louarn\thanks{Footnote for second author}%
%  % r\'ef\'erence \`a la note partag\'ee
% \thanksref{sfn}
%  % liste longue pour tests de mise en page
% \and Nicolas Bourbaki\thanks{etc} \and Andr\'e Lichnerowicz
% \and Marcel-Paul Sch\"utzenberger \and Jacques-Louis Lions}
%% Ceci apparait sur chaque page paire.
\authorhead{Coullon \& Bigot \& Perez}
%% titre francais long
\RRtitle{Le langage MSL: orchestration de stencils au travers d'un DSL maillage-agnostique}
%% English title
\RRetitle{The Multi-Stencil Language: orchestrating stencils with a mesh-agnostic DSL}
%%
\titlehead{The Multi-Stencil Language}
%%
%\RRnote{This is a note}
%\RRnote{This is a second note}
%%
\RRresume{

}
\RRabstract{
As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important. One of the big challenges of exascale is to get programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists. One relevant solution to ease parallel programming for scientists is Domain Specific Language (DSL). However, one problem to avoid with DSLs is to not design a new DSL each time a new domain or a new problem has to be solved. This phenomenon happens for stencil-based numerical simulations, for which a large number of languages has been proposed without code reuse between them. The Multi-Stencil Language (MSL) presented in this paper is a language common to any kind of mesh used into a stencil-based numerical simulation. It is said that MSL is mesh-agnostic. Actually, from the description of a numerical simulation, MSL produces an empty parallel pattern, or skeleton, of the simulation which will be filled using other existing parallel languages and libraries. Thus, MSL, by finding a common language for different kinds of stencil-based simulation, facilitates code reuse. MSL is evaluated on a real case simulation which solves shallow-water equations. It is shown that MSL does not introduce overheads on data parallelism up to 16.384 cores, and that the hybrid parallelism (data and task) introduced improves performance of the simulation.}
%%
\RRmotcle{Langage de domaine sp\'cifique, stencil, simulation num\'rique, parall\'lisme de donn\'es, parall\'lisme de t\^aches, ordonnancement, MPI, OpenMP}
\RRkeyword{Domain Specific Language (DSL), stencil, numerical simulation, data parallelism, task parallelism, scheduling, MPI, OpenMP}
%%
%% \RRprojet{Apics}  % cas d'un seul projet
\RRprojets{Avalon}
%%
%% \URLorraine % pour ceux qui sont \`a l'est
%% \URRennes  % pour ceux qui sont \`a l'ouest
%% \URRhoneAlpes % pour ceux qui sont dans les montagnes
%% \URRocq % pour ceux qui sont au centre de la France
%% \URFuturs % pour ceux qui sont dans le virtuel
%% \URSophia % pour ceux qui sont au Sud.
%%
%% \RCBordeaux % centre de recherche Bordeaux - Sud Ouest
%% \RCLille % centre de recherche Lille Nord Europe
%% \RCParis % Paris Rocquencourt
%% \RCSaclay % Saclay \^Ile de France
\RCGrenoble % Grenoble - Rh\^one-Alpes
%% \RCNancy % Nancy - Grand Est
%% \RCRennes % Rennes - Bretagne Atlantique
%% \RCSophia % Sophia Antipolis M\'editerran\'ee

%
\begin{document}

%\title{The Multi-Stencil Language: orchestrating stencils with a mesh-agnostic DSL}

%\author{H\'el\`ene Coullon \and Julien Bigot \and Christian Perez}

% \institute{H\'el\`ene Coullon \and Christian Perez \at
%               INRIA, LIP, ENS Lyon, 46 Allée d'Italie, 69364 Lyon Cedex 07, France\\
%               \email{helene.coullon@inria.fr} \email{christian.perez@inria.fr}
%            \and
%            Julien Bigot \at
%            Maison de la Simulation, CEA, CNRS, Univ. Paris-Sud, UVSQ, Universit\'e Paris-Saclay, 91191 Gif-sur-Yvette, France\\
%            \email{julien.bigot@cea.fr}
% }

\makeRR

% \begin{abstract}
% As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important. One of the big challenges of exascale is to get programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists. One relevant solution to ease parallel programming for scientists is Domain Specific Language (DSL). However, one problem to avoid with DSLs is to not design a new DSL each time a new domain or a new problem has to be solved. This phenomenon happens for stencil-based numerical simulations, for which a large number of languages has been proposed without code reuse between them. The Multi-Stencil Language (MSL) presented in this paper is a language common to any kind of mesh used into a stencil-based numerical simulation. It is said that MSL is mesh-agnostic. Actually, from the description of a numerical simulation, MSL produces an empty parallel pattern, or skeleton, of the simulation which will be filled using other existing parallel languages and libraries. Thus, MSL, by finding a common language for different kinds of stencil-based simulation, facilitates code reuse. MSL is evaluated on a real case simulation which solves shallow-water equations. It is shown that MSL does not introduce overheads on data parallelism up to 16.384 cores, and that the hybrid parallelism (data and task) introduced improves performance of the simulation.

% \keywords{Domain Specific Language (DSL) \and stencil \and numerical simulation \and mesh-agnostic \and data parallelism \and task parallelism \and scheduling \and MPI \and OpenMP}
% \end{abstract}

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
\input{introduction.tex}
%------------------------------------------------------------------------------
\section{Related work}
\label{sect:rel}
\input{related.tex}
%------------------------------------------------------------------------------
\section{Computational model of Multi-Stencil Programs}
\label{sect:formalism}
\input{formalism.tex}
%------------------------------------------------------------------------------
\section{The Multi-Stencil Language}
\label{sect:msl}
\input{msl.tex}
%------------------------------------------------------------------------------
\section{Parallelization of Multi-Stencil Programs}
\label{sect:parallelism}
\input{parallelism.tex}
%------------------------------------------------------------------------------
\section{Static Scheduling and performance model}
\label{sect:msp}
\input{msp.tex}
%------------------------------------------------------------------------------
\section{Implementation}
\label{sect:comp}
\input{compiler.tex}
%------------------------------------------------------------------------------
\section{Evaluation}
\label{sect:eval}
\input{evaluation.tex}
%------------------------------------------------------------------------------
\section{Conclusion}
\label{sect:concl}
\input{conclusion.tex}
%------------------------------------------------------------------------------

%\begin{acknowledgements}
This work has partially been supported by the PIA ELCI project of the French FSN.
This  work  was  granted  access  to  the  HPC  resources  of TGCC under  the 
allocations t2015067470 and x2016067617 made by GENCI.
%\end{acknowledgements}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}
% end of file template.tex

### 2016-10-12 ARTICLES/RR_IJPP/evaluation.tex
Rapport de recherche avec contenu de la soumission IJPP

--- 2016-05-31 ARTICLES/IJPP/article/evaluation.tex
@@
 \begin{minipage}{.475\textwidth}
-  \includegraphics[width=\textwidth]{../results/weak_scaling/400/median_weak.pdf}
+  \includegraphics[width=\textwidth]{../IJPP/results/weak_scaling/400/median_weak.pdf}
   \caption{weak-scaling with $400 \times 400$ domain per core.}
@@
 \begin{minipage}{.475\textwidth}
-  \includegraphics[width=\textwidth]{../results/weak_scaling/800/median_weak.pdf}
+  \includegraphics[width=\textwidth]{../IJPP/results/weak_scaling/800/median_weak.pdf}
   \caption{weak-scaling with $800 \times 800$ domain per core.}
@@
 \begin{figure}[t]\begin{center}
-  \includegraphics[width=.6\textwidth]{../results/strong_scaling/10K_1K/median_strong.pdf}
+  \includegraphics[width=.6\textwidth]{../IJPP/results/strong_scaling/10K_1K/median_strong.pdf}
   \caption{Strong scaling on a $10k \times 10k$ domain.}
@@
 \begin{figure}[!h]\begin{center}
-  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/fusVSbase.pdf}
+  \includegraphics[width=.6\textwidth]{../IJPP/results/task_scaling/500_200/fusVSbase.pdf}
   \caption{Strong scaling on a 500x500 domain, with and without the fusion optimization.}
@@
 \begin{figure}[bh]\begin{center}
-  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/analytic/times.pdf}
+  \includegraphics[width=.6\textwidth]{../IJPP/results/task_scaling/500_200/analytic/times.pdf}
   \caption{Computation vs communication times in the data parallelization technique.}
@@
 \begin{figure}[th]\begin{center}
-  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/base_close_median.pdf}
+  \includegraphics[width=.6\textwidth]{../IJPP/results/task_scaling/500_200/base_close_median.pdf}
   \caption{Strong scaling comparisons between data parallelization and hybrid parallelization. A \emph{close} OpenMP clause is used to bind threads onto cores.}
@@
 \begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_spread_median.pdf}}
+  \resizebox{8cm}{!}{\includegraphics{../IJPP/results/task_scaling/500_200/base_spread_median.pdf}}
   \caption{Strong scaling comparisons between data parallelization and hybrid parallelization. A \emph{spread} OpenMP clause is used to bind threads onto cores.}
@@
 \begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/tth.pdf}}
+  \resizebox{8cm}{!}{\includegraphics{../IJPP/results/task_scaling/500_200/analytic/tth.pdf}}
   \caption{Execution times (seconds) of $T_{COM1}$ and $T_{COM2} + T_{task}$ for 8 cores for task parallelization. Verification of the Equation~(\ref{eq:hyb}).}

### 2016-10-12 ARTICLES/RR_IJPP/article.tex
Rapport de recherche avec contenu de la soumission IJPP

@@
 \RRresume{
-
+Alors que la puissance de calcul des architectures hautes performances augmente, leur h\'et\'erog\'en\'eit\'e et leur complexit\'e augmente \'egalement ce qui rend de plus en plus difficile leur utilisation.
+Pour cette raison, l'un des sujet de recherche majeur pour pr\'etendre atteindre des machines exascale est de proposer des mod\`eles de programmation permettant l'acc\`es au calcul hautes performances au plus grand nombre et non pas seulement \`a une poign\'ee de sp\'ecialistes.
+L'une des solutions permettant de faciliter la programmation parall\`ele aux scientifiques tout en conservant les performances est l'utilisation de langages de domaine sp\'ecifiques (DSL).
+Cependant, l'un des probl\`emes li\'e aux DSL est leur conception qui n'est pas envisageable pour chaque nouveau domaine vis\'e car trop chronophage et complexe.
+Ce probl\`eme appara\^it notamment dans le domaine de la simulation num\'erique bas\'e sur les calculs de type stencil.
+Dans ce domaine un tr\`es large ensemble de langages ont \'et\'e propos\'es chacun avec ses sp\'ecificit\'es et sans aucune r\'e-utilisation de fonctionnalit\'es et de codes de conception de l'un vers l'autre.
+Le langage MSL (Multi-Stencil Language) pr\'esent\'e dans ce rapport de recherche est un langage qui extrait le fondement commun des diff\'erents types de simulation multi-stencil en proposant une abstraction du type de maillage utilis\'e.
+MSL produit un squelette (ou patron) de l'application d\'ecrite en entr\'ee et r\'e-utilise des langages hautes performances existants pour obtenir l'application finale.
+MSL est \'evalu\'e sur un cas r\'eel de simulation num\'erique par lequel il est montr\'e que MSL n'introduit pas de surco\^uts par rapport \`a une utilisation classique des langages sous-jacents.
+Le passage \`a l'\'echelle est \'evalu\'e jusqu'\`a 16.384 c\oe{urs} de calcul.
+Il est \'egalement montr\'e que la version hybride de code introduite par MSL am\'eliore les performances en suivant un mod\`ele de performances \'egalement d\'ecrit.
 }

### 2016-10-13 ARTICLES/RR_IJPP/article.tex
article.tex

@@
 
-\RRNo{7003}
-%% \RTNo{0703}
-%%
 %% date de publication du rapport
@@
 %%
-\RRauthor{H\'el\`ene Coullon\thanks[lip]{INRIA, LIP, ENS Lyon, 46 All\'e d'Italie, 69364 Lyon Cedex 07, France}\thanks{helene.coullon@inria.fr} \and Julien Bigot\thanks{Maison de la Simulation, CEA, CNRS, Univ.
-Paris-Sud, UVSQ, Universit\'e Paris-Saclay, 91191 Gif-sur-Yvette, France}\thanks{julien.bigot@cea.fr} \and Christian Perez\thanksref{lip}\thanks{christian.perez@inria.fr}}
+\RRauthor{H\'el\`ene Coullon\thanks[lip]{Univ.
+Lyon, Inria, CNRS, ENS de Lyon, Univ.
+Claude-Bernard Lyon 1, LIP}\thanks{\{helene.coullon\}\{christian.perez\}@inria.fr} \and Julien Bigot\thanks{Maison de la Simulation, CEA, CNRS, Univ.
+Paris-Sud, UVSQ, Univ.
+Paris-Saclay}\thanks{julien.bigot@cea.fr} \and Christian Perez}%\thanksref{lip}\thanks{christian.perez@inria.fr}}
 
@@
 \begin{document}
+\RRNo{8962}
 

### 2016-11-02 ARTICLES/OTHER/LIP_JL/slides.tex
renaming

\documentclass{beamer}

\usepackage{beamerthemeCambridgeUS}
\usepackage[frenchb]{babel}
\usepackage[utf8]{inputenc}
\usecolortheme{beaver}

\usepackage{listings}
\usepackage{xcolor}
\lstdefinestyle{base}{
  basicstyle=\ttfamily\color{black} \footnotesize,
  moredelim=**[is][\color{blue}]{@}{@},
  moredelim=**[is][\color{red}]{&}{&}
}

% \setbeamertemplate{navigation symbols}{
% \insertframenumber/
% \inserttotalframenumber
% }

\usepackage{color}
\definecolor{vert}{rgb}{0,0.5,0}
\definecolor{violet}{rgb}{0.5,0,0.5}

\usepackage{tikz}
\usetikzlibrary{calc,positioning,matrix,arrows,shapes.geometric,shapes.symbols,shapes.misc,shapes,automata,petri,decorations.markings,shadows}
\usepackage{array}
\usepackage{subfig}

\usepackage[absolute,overlay]{textpos}
\usepackage{tcolorbox}

\usepackage[lined]{algorithm2e}
\usepackage{ulem}

%----------TIKZ
\usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}

\tikzset{
    use/.style={
    circle,draw=black,fill=black,scale=0.5,text=white
    },
    mpi/.style={
    rectangle,draw=black,fill=black,scale=0.8,text=white
    },
    provide/.style={
    circle,draw=black,fill=white,scale=0.5
    },
    component/.style={
    rectangle,rounded corners=3pt,draw=black
    },
    dcomponent/.style={
    rectangle,rounded corners=3pt,dashed,draw=black
    },
    progm/.style={rectangle,dotted, draw=black, thin, text width=8em, text centered,rounded corners, minimum height=2em},
  line/.style={draw, thin, ->, shorten >=2pt},
  purp/.style={rectangle, draw=violet,fill=violet!20, text width=10em, text centered,thin,rounded corners,inner sep=3pt},
  purp2/.style={rectangle, draw=violet,fill=violet!20, text width=15em, text centered,thin,rounded corners,inner sep=3pt},
  prog/.style={rectangle, draw=black, thin, text width=10em, text centered,rounded corners, minimum height=2em},
  progbl/.style={rectangle, draw=blue, thin, text width=10em, text centered,rounded corners, minimum height=2em},
  progrg/.style={rectangle, draw=red, thin, text width=10em, text centered,rounded corners, minimum height=2em},
  proggr/.style={rectangle, draw=vert, thin, text width=10em, text centered,rounded corners, minimum height=2em},
  progbrg/.style={rectangle, draw=red, very thick, text width=10em, text centered,rounded corners, minimum height=2em},
  progm/.style={rectangle,dotted, draw=black, thin, text width=10em, text centered,rounded corners, minimum height=2em},
  line/.style={draw, thin, ->, shorten >=2pt},
  inner/.style={rectangle, draw=blue!50,fill=blue!20, text width=10em, text centered,thin,rounded corners,inner sep=3pt},
  purp/.style={rectangle, draw=violet,fill=violet!20, text width=10em, text centered,thin,rounded corners,inner sep=3pt},
  purp2/.style={rectangle, draw=violet,fill=violet!20, text width=15em, text centered,thin,rounded corners,inner sep=3pt},
  purpb/.style={rectangle, draw=violet,fill=violet!20, text width=10em, text centered,very thick,rounded corners,inner sep=3pt},
  outer/.style={draw=vert,fill=green!20,thin,rounded corners,inner sep=10pt},
  outer2/.style={draw=vert,fill=green!20,text width=10em, text centered,thin,rounded corners,inner sep=3pt},
  outer3/.style={draw=red,fill=red!20, text width=10em, text centered,thin,rounded corners,inner sep=3pt},
  outer2b/.style={draw=vert,fill=green!20,text width=10em, text centered,very thick,rounded corners,inner sep=3pt},
  outer3b/.style={draw=red,fill=red!20, text width=10em, text centered,very thick,rounded corners,inner sep=3pt},
  comp/.style={draw=red,fill=red!20, text width=10em, text centered,thin,rounded corners,inner sep=10pt},
  the/.style={rectangle, draw=black, thin, text width=16em, text centered,rounded corners, minimum height=2em},
  the2/.style={rectangle,dotted, draw=black, thin, text centered,rounded corners, minimum height=2em},
  the3/.style={rectangle, draw=black, thin, text width=12em, text centered,rounded corners, minimum height=2em},
  the4/.style={rectangle,draw=black, thin, text centered,rounded corners, minimum height=2em},
}
%-------------
% * * * * * * * * * * * * * * *  ANNEXES * * * * * * * * * * * * * * * * * *
\newcommand{\backupbegin}{
   \newcounter{framenumberappendix}
   \setcounter{framenumberappendix}{\value{framenumber}}
}
\newcommand{\backupend}{
   \addtocounter{framenumberappendix}{-\value{framenumber}}
   \addtocounter{framenumber}{\value{framenumberappendix}} 
}

\AtBeginSection[]
{
  \begin{withoutheadline}
   \begin{frame}
   \frametitle{Table of contents}
        \tableofcontents[currentsection,hideallsubsections]
   \end{frame}
   \end{withoutheadline}
}
%-------------------------------------------------------------

\def\pprec{\mathrel{\scalebox{.9}[1]{$\prec$}\mkern-3mu%
  \scalebox{.4}[1]{$\prec$}\mkern-5.5mu\scalebox{.4}[1]{$\prec$}}}

\makeatletter
    \newenvironment{withoutheadline}{
        \setbeamertemplate{headline}[default]
        \def\beamer@entrycode{\vspace*{-\headheight}}
    }{}
\makeatother

%-------------------------------------------------------------------
\title[Journée Langages LIP]{The Multi-Stencil Language}
\author[Hélène Coullon]{\underline{Hélène Coullon}, Christian Perez and Julien Bigot}
\institute[Inria]{Inria Avalon, LIP Lyon}
\date[$17^{th}$ March 2016]{$17^{th}$ March 2016\\Journée Langages du LIP}
%-------------------------------------------------------------------

\begin{document}

%---------------
%---------------
\begin{withoutheadline}

\begin{frame}
    \titlepage
\end{frame}

% INTRODUCTION / MOTIVATION
%--------
%----------
\begin{frame}
\frametitle{High performance computing and parallelism in 2015}
\begin{center}
\only<1>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Parallel\\Programming models}; \\
      \node[progrg] (mod) {Parallel libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
    \end{scope}
  \end{tikzpicture}
}
%--------
\only<2>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Parallel\\Programming models}; \\
      \node[progrg] (mod) {Parallel libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
    \end{scope}
  \end{tikzpicture}
}
%--------
\only<3>{

  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Parallel\\Programming models}; \\
      \node[progrg] (mod) {Parallel libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \node[fill=white,right of=mod,node distance=3cm] (bsp) {\scriptsize \textcolor{red}{BSP}};
    \node[fill=white,right of=bsp,node distance=1.5cm] (mpi) {\scriptsize \textcolor{red}{MPI}};
    \node[fill=white,right of=mpi,node distance=1.5cm] (openmp) {\scriptsize \textcolor{red}{OpenMP}};
    \node[fill=white,right of=openmp,node distance=1.5cm] (cuda) {\scriptsize \textcolor{red}{Cuda}};
    \node[fill=white,right of=cuda,node distance=1.5cm] (opencl) {\scriptsize \textcolor{red}{OpenCL ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
      \path (mpi) edge[->,thin, bend right=50] (openmp);
      \path (mpi) edge[->,thin, bend right=50] (cuda);
      \path (bsp) edge[->,thin, bend right=50] (mpi);
    \end{scope}

  \end{tikzpicture}
}
%--------
\only<4>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Parallel\\Programming models}; \\
      \node[progrg] (mod) {Parallel libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \node[fill=white,right of=mod,node distance=3cm] (bsp) {\scriptsize \textcolor{red}{BSP}};
    \node[fill=white,right of=bsp,node distance=1.5cm] (mpi) {\scriptsize \textcolor{red}{MPI}};
    \node[fill=white,right of=mpi,node distance=1.5cm] (openmp) {\scriptsize \textcolor{red}{OpenMP}};
    \node[fill=white,right of=openmp,node distance=1.5cm] (cuda) {\scriptsize \textcolor{red}{Cuda}};
    \node[fill=white,right of=cuda,node distance=1.5cm] (opencl) {\scriptsize \textcolor{red}{OpenCL ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
      \path (mpi) edge[->,thin, bend right=50] (openmp);
      \path (mpi) edge[->,thin, bend right=50] (cuda);
      \path (bsp) edge[->,thin, bend right=50] (mpi);

      \path (mpi) -- (donn);
      \path (mpi) -- (tach);
      \path (mpi) -- (mess);
      \path (bsp) -- (donn);
      \path (bsp) -- (mess);
      \path (openmp) -- (tach);
      \path (openmp) -- (donn);
      \path (cuda) -- (tach);
      \path (opencl) -- (tach);
      \path (opencl) -- (mess);
    \end{scope}
  \end{tikzpicture}
}
%---------
\only<5>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Parallel\\Programming models}; \\
      \node[progrg] (mod) {Parallel libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \node[fill=white,right of=mod,node distance=3cm] (bsp) {\scriptsize \textcolor{red}{BSP}};
    \node[fill=white,right of=bsp,node distance=1.5cm] (mpi) {\scriptsize \textcolor{red}{MPI}};
    \node[fill=white,right of=mpi,node distance=1.5cm] (openmp) {\scriptsize \textcolor{red}{OpenMP}};
    \node[fill=white,right of=openmp,node distance=1.5cm] (cuda) {\scriptsize \textcolor{red}{Cuda}};
    \node[fill=white,right of=cuda,node distance=1.5cm] (opencl) {\scriptsize \textcolor{red}{OpenCL ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
      \path (mpi) edge[->,thin, bend right=50] (openmp);
      \path (mpi) edge[->,thin, bend right=50] (cuda);
      \path (bsp) edge[->,thin, bend right=50] (mpi);

      \path (mpi) -- (donn);
      \path (mpi) -- (tach);
      \path (mpi) -- (mess);
      \path (bsp) -- (donn);
      \path (bsp) -- (mess);
      \path (openmp) -- (tach);
      \path (openmp) -- (donn);
      \path (cuda) -- (tach);
      \path (opencl) -- (tach);
      \path (opencl) -- (mess);

      \path (mpi) -- (grap);
      \path (mpi) -- (mc);
      \path (mpi) -- (many);
      \path (bsp) -- (grap);
      \path (bsp) -- (mc);
      \path (cuda) -- (gpu);
      \path (openmp) -- (many);
      \path (openmp) -- (mc);
      \path (opencl) -- (gpu);
      \path (opencl) -- (mc);
      \path (opencl) -- (many);
      \path (opencl) -- (grap);
    \end{scope}

  \end{tikzpicture}

  \medskip
  \textit{Very complex domain !}
}
  \end{center}
\end{frame}
%----------
%-------------------------------------------------------------
\begin{frame}
\frametitle{"Please Help"}
%implicitly parallel (one sep. of concerns) / specific / performance
\begin{block}{HPC}
\begin{itemize}
\item Heterogenous and complex parallel hardware
\item Many programming models to combine
\item Expert programming
\end{itemize}
\end{block}

\begin{block}{Users}
Weather, climat, physics, biology etc.
\end{block}

\begin{center}
Separation of concerns between domain/parallelization and optimization\\
Domain specific languages (DSL)\\
\medskip
\resizebox{2cm}{!}{\includegraphics{images/scientist.jpg}}
\end{center}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Contribution}
%implicitly parallel (one sep. of concerns) / specific / performance
\begin{block}{From MSL}
\begin{itemize}
\item Descriptive language,
\item for Multi-Stencil simulations,
\item without numerical code.
\end{itemize}
\end{block}

\begin{block}{To}
\begin{itemize}
\item Parallel pattern of the simulation,
\item with automatic synchronizations for data and task parallelism,
\item with empty functions to fill,
\item with good performance.
\end{itemize}
\end{block}

Separation of concerns between domain/implementation/parallelization

\end{frame}
%-------------------------------------------------------------

% presentation of the table of contents
%-------------------------------------------------------------
\begin{frame}
\frametitle{Table of contents}
    \tableofcontents[hideallsubsections]
\end{frame}
%-------------------------------------------------------------

\end{withoutheadline}

%-------------------------------------------------------------
\section{The domain: Multi-Stencil}
%-------------------------------------------------------------
\begin{frame}
\frametitle{What is a multi-stencil simulation ?}
\begin{center}
\begin{tikzpicture}
\def \n {5}
\def \radius {3cm}
\def \margin {8} % margin in angles, depends on the radius

  \node[purp2,align=center] (edp) at ({360/\n * (2 - 1)}:\radius) {set of PDEs\\\scriptsize $\frac{\partial u(x,y,t)}{\partial t} = \frac{\partial^2 u(x,y,t)}{\partial x^2} + \frac{\partial^2 u(x,y,t)}{\partial y^2}$};
  \node[rectangle,dotted, draw=black, thin, text width=7em, text centered,rounded corners, minimum height=2em,align=center,right of=edp,node distance=6cm] (cond) {\tiny Behavior of the phenomena\\and boundary behaviors};
  \path[line,dotted] (edp) -- (cond);
  \node[purp,align=center] (maillage) at ({360/\n * (1 - 1)}:\radius) {Mesh and\\time iterations};
  \path[->] (maillage) edge[<-,thin,shorten <=2pt,shorten >=2pt, bend right=40]  node [fill=white,align=center] {\scriptsize Discretization} (edp.east);
  \path[] (edp.west) edge[thin,shorten <=2pt,shorten >=2pt, bend right=90]  node [draw,circle,fill=white] (plus) {\scriptsize +} (maillage.west);
  \node[purp,align=center, below of=plus, node distance=3cm] (sche) {Numerical schemes\\Explicit = \textcolor{red}{stencil}};
  \node[fill=white,right of=sche,node distance=6cm,align=left] {\tiny Quantity $\sigma(x,t)$ computed from quantities at $(x,t)$ if different,\\\tiny $(x,t-1)$ and $(y,t-1)$, where $y$ is a neighborhood of $x$};
  \path[line] (plus) -- node [fill=white,align=center] {\scriptsize Numerical methods\\\scriptsize Finite difference/volume/elements} (sche);
\end{tikzpicture}
\end{center}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Multi-Stencil algorithm}
\begin{algorithm}[H]
 Create the mesh $\mu$\\
 Create the quantities to simulate mapped onto $\mu$\\
 Initialization of the quantities and parameters\\
 Create the time step $\delta t$\\
 Create the maximum time $tmax$\\
 \While{$criteria(t)==true$}{
 \For{each element of the boundary}{
   Numerical computations of boundary conditions
 }
  \For{each $x \in \mu$}{
  Compute stencils and auxiliary computations
  }
  $t = t+\delta t$
 }
 ...
\end{algorithm}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Definitions}
\begin{block}{Mesh}
\begin{itemize}
\item \textit{Mesh}: graph without bridges
\item \textit{Mesh entity}: sub-graph of the mesh
\item \textit{Group of mesh entities}: set of mesh entities of the same type
\item \textit{Computation domain}: sub-part of a group
\item \textit{Neighboorhood}: function from a mesh entity to a set of mesh entities (possibly from a different group than the input)
\end{itemize}
\end{block}

\begin{block}{Quantities}
\begin{itemize}
\item A \textit{quantity} is mapped onto a group of mesh entities
\item A \textit{scalar} is global to the simulation and is typically a numerical value constant or not
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Definitions}
\begin{block}{Time and computations}
\begin{itemize}
\item \textit{Time}: time step and convergence criteria
\item \textit{Computation}:
\begin{itemize}
\item Set of scalar read
\item Set of quantities read, on which neighborhood
\item Quantity written, on which computation domain
\item A numerical expression
\end{itemize}
\end{itemize}
\end{block}

\begin{block}{Four kinds of computations}
\begin{itemize}
\item \textit{Stencil computation}: it exists a quantity read with a neighborhood
\item \textit{Boundary computation}: it exists a quantity read equal to the quantity written with a neighborhood
\item \textit{Local computation}: for all quantity read, there is not neighborhood
\item \textit{Reduction}: the quantity written is a scalar
\end{itemize}
\end{block}

\begin{equation*}
\mathcal{MSP}(\mathcal{M},\mathcal{E},\mathcal{D},\mathcal{N},\Delta,\mathcal{S},T,\Gamma)
\end{equation*}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Wait a while}
One can notice that
\begin{itemize}
\item Definitions independant of the mesh topology
\item Definitions independant of the numerical expression of computations
\end{itemize}
\begin{block}{The Multi-Stencil Language}
\begin{itemize}
\item Descriptive language mesh-agnostic and without numerical code asked to the user
\item Separation of concerns: description / implementation / parallelization
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{The Multi-Stencil Language}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: the mesh}
\begin{center}
\resizebox{6cm}{!}{\includegraphics{images/mesh.pdf}}
\end{center}
\begin{columns}
\begin{column}{6.5cm}
\begin{lstlisting}[frame=single]
mesh : cart
mesh entities : cell, edgex
computation domains :
  d1 in cell
  d2 in edgex
stencil shapes : 
  ncc from cell to cell
  nce from cell to edgex
  nec from edgex to cell
\end{lstlisting}
\end{column}
\begin{column}{5.5cm}
\begin{itemize}
\item A cartesian mesh is used
\item Two kind of mesh entities are declared onto it
\item Two computations domains, one for each entity type
\item Three stencil shapes (neighborhood from mesh entity to mesh entity)
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: quantity}
\begin{columns}
\begin{column}{5cm}
\begin{lstlisting}[frame=single]
quantity :
  A,cell
  B,cell
  C,edgex
  D,cell
  E,cell
  F,cell
  G,cell
  H,edgex
  I,cell
  J,cell
scalar : mu, tau
\end{lstlisting}
\end{column}
\begin{column}{6cm}
\begin{itemize}
\item A is a quantity applied onto cell
\item C is a quantity applied onto edgex
\item mu and tau are scalar values
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: time and computations}
\begin{columns}
\begin{column}{6cm}
\begin{lstlisting}[frame=single]
time : 500
computations :
  B[d1] = k0({tau},{A})
  C[d2] = k1({},{B[nce]})
  D[d1] = k2({},{C})
  E[d1] = k3({},{C})
  F[d1] = k4({},{D,C[nec]})
  G[d1] = k0({mu,tau},{E})
  H[d2] = k6({},{F})
  I[d1] = k7({},{G,H})
  J[d1] = k8({mu},{I[ncc]})
\end{lstlisting}
\end{column}
\begin{column}{5cm}
\begin{itemize}
\item The time loop is composed of 500 iterations.
\item J is written
\item onto the computation domain d1,
\item by the computation k8,
\item which read the scalar mu
\item and the quantity I,
\item which is accessed with the neighborhood ncc.
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Introduction of parallelism}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Synchronizations}
\textit{\footnotesize Application = Quantity + Program\\
Data parallel Application = Split quantity + Program + Synchronizations\\}
\normalsize $\rightarrow$ Find where synchronizations are needed
\begin{block}{Synchronizations}
\begin{itemize}
\item For stencil computations (point to point synchronization)\\
      if for $k_i$ and $k_j$, with $k_j$ a stencil, $R_j \cap \{w_i\} \neq \emptyset$
\item For reduction computations (global synchronization)\\
\end{itemize}
\end{block}
$\Gamma = [k_0,k_1,k_2,k_3,k_4,k_0,k_6,k_7,k_8]$\\
\textbf{$\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_0,k_6,k_7,k_{7;8}^{sync},k_8]$}\\
\textbf{$\Gamma_{sync}$ is applied on each resource !}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Dependencies}
\textit{\footnotesize Program = sequence of tasks\\
Task parallel Program = schedule of parallel and sequential tasks}\\
\normalsize $\rightarrow$ Define dependency relations
\begin{block}{Read after write dependency}
\begin{itemize}
\item for $k_i$ and $k_j$, with $i < j$
\item if $R_j \cap \{w_i\} \neq \emptyset$, and domain computations intersect
\item $k_i$ has to be computed before $k_j$
\end{itemize}
\end{block}
\begin{block}{Write after write dependency}
\begin{itemize}
\item for $k_i$ and $k_j$, with $i < j$
\item if $w_j = w_i$, and domain computations intersect
\item $k_i$ has to be computed before $k_j$
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Dependency graph}
From read/write and write dependencies is built $\Gamma_{dep}$\\
$\Gamma_{dep}$ represents a \textbf{single time iteration}\\

\medskip
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$k_0$};
   \node[] (star1) at (1,0) {$k_{0;1}^{sync}$};
   \node[] (c1) at (2,0) {$k_1$};
   \node[] (c2) at (3,0.5) {$k_2$};
   \node[] (star4) at (3,1.5) {$k_{1;4}^{sync}$};
   \node[] (c3) at (3,-0.5) {$k_3$};
   \node[] (c4) at (4,0.5) {$k_4$};
   \node[] (c5) at (4,-0.5) {$k_5$};
   \node[] (c6) at (5,0.5) {$k_6$};
   \node[] (c7) at (6,0) {$k_7$};
   \node[] (star8) at (7,0) {$k_{7;8}^{sync}$};
   \node[] (c8) at (8,0) {$k_8$};

  \path[->]
    (c0) edge node {} (star1)
    (star1) edge node {} (c1)
    (c1) edge node {} (c2)
         edge node {} (c3)
         edge node {} (star4)
    (star4) edge node {} (c4)
    (c2) edge node {} (c4)
    (c4) edge node {} (c6)
    (c3) edge node {} (c5)
    (c5) edge node {} (c7)
    (c6) edge node {} (c7)
    (c7) edge node {} (star8)
    (star8) edge node {} (c8);
  \end{tikzpicture}
  \end{center}

\medskip
\textbf{$\Gamma_{dep}$ is applied on each resource !}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Implementation}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Component models}
\begin{itemize}
\item Black box with code
\item Use/provide interfaces
\item Improve code reuse (productivity)
\item Improve separation of concerns (maintainability and portability)
\item Improve composability of applications
\end{itemize}
\begin{center}
\begin{tikzpicture}[remember picture]
  \node[component,solid] (A) at (0,0) {A};
  \node[component,solid,right=5cm of A] (B) {B};
  \node[use,scale=0.5,solid,right=1.5cm of A] (Au) {};
  \node[provide,scale=0.5,solid,left=1.5cm of B] (Bp) {};
  \path[-]
    (A) edge node {} (Au)
    (Bp) edge node {} (B);
\end{tikzpicture}

\bigskip
\begin{tikzpicture}[remember picture]
  \node[component,solid] (A) at (0,0) {A};
  \node[use,scale=0.5,solid,right=1cm of A] (Au) {};
  \node[provide,scale=0.5,solid,right=0.01cm of Au] (Bp) {};
  \node[component,solid,right=1cm of Bp] (B) {B};
  \path[-]
    (A) edge node {} (Au)
    (Bp) edge node {} (B);
\end{tikzpicture}

\medskip
Components A and B into a component assembly
\end{center}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{What is produced ?}
A data parallel/task parallel component assembly\\

\bigskip
\begin{center}
\begin{tikzpicture}[remember picture]
  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
    \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
     \node[component,solid,draw=black] (D) at (0,0) {$Driver$};
     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
     \node[solid] (Ds) at (-1.5,0) {start};
     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};

     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
     \node[solid,left=0.8cm of T] (tt) {$T$};

     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
     \node[component,solid,right=2cm of Cp] (C) {$Scheduler$};
     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
     \node[use,scale=0.5,solid,right=1.5cm of C] (Cu2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Cu2] (Cu3) {};
     \node[use,scale=0.5,solid,below=0.8cm of Cu3] (Cu4) {};
     \node[solid,right=0.2cm of Cu] (star) {$*$};
     %\node[solid,right=1.5cm of C] (gamma) {$\Gamma$};

     \node[solid,below=1cm of C] (calc) {$\Gamma_{sync}$, $\Gamma_{dep}$};

     \node[provide,scale=0.5,solid,right=0.15 of Cu2] (k0p) {};
     \node[component,solid,right=1.5cm of k0p] (k0) {$k_0$};

     \node[provide,scale=0.5,solid,right=0.15 of Cu3] (k0sp) {};
     \node[component,solid,right=1.5cm of k0sp] (k0s) {$k_{0;1}^{sync}$};

     \node[provide,scale=0.5,solid,right=0.15 of Cu4] (kotherp) {};
     \node[component,solid,right=1.5cm of kotherp] (kother) {$\dots$};

     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Quantity$};
     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
     \node[solid,left=1.5cm of Data] (delta) {$\Delta,\mathcal{D}$};
     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};

     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
     \node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};

    \path[-]
      (Dp) edge [solid] node {} (D)
      (D) edge [solid] node {} (Du1)
          edge [solid] node {} (Du2)
          edge [solid] node {} (Du3)
      (DDSp1) edge [solid] node {} (DDS)
      (Tp) edge [solid] node {} (T)
      (T)  edge [solid] node {} (Tu)
      (Cp) edge [solid] node {} (C)
      (C) edge [solid] node {} (Cu)
      (Datap2) edge [solid] node {} (Data)
      (Datap1) edge [solid] node {} (Data)
      (Data) edge [solid] node {} (Datau)
      (DDSp2) edge [solid] node {} (DDS)
      (C)  edge [solid] node {} (Cu2)
          edge [solid] node {} (Cu3)
          edge [solid] node {} (Cu4)
      (Cu2)  edge [solid] node {} (k0p)
      (k0p)  edge [solid] node {} (k0)
      (Cu3)  edge [solid] node {} (k0sp)
      (k0sp)  edge [solid] node {} (k0s)
      (Cu4)  edge [solid] node {} (kotherp)
      (kotherp)  edge [solid] node {} (kother);
    \end{tikzpicture}
    };
    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=4cm of Data] (mpidata) {};
    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
    \node[solid,above=3.2cm of proc.west,anchor=west] (core) {\textit{\tiny{Applied on each resource}}};
    \path[-]
      (Data.east) edge node {} (mpidata);
  \end{tikzpicture}
\end{center}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{What is produced ?}
\begin{block}{Implementation}
\begin{itemize}
\item DDS + Quantity + $k_{*}^{sync}$ = written components using SkelGIS (distributed memory, MPI)
\item Scheduler = Serie-Parallel graph of $\Gamma_{dep}$ (OpenMP 3.0)
\item Time + Scheduler = generated components
\item $k_i$ = empty components to write
\end{itemize}
\end{block}

\begin{block}{Separation of concerns}
\begin{itemize}
\item numerician (MSL)
\item computer engineer (K)
\item HPC engineer (DDS, quantity, $k_{*}^{sync}$ etc.)
\item DSL designer (component assembly + generated components)
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Results}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Evaluations}
\begin{block}{Application}
\begin{itemize}
\item FullSWOF2D: developed at the MAPMO, Université d'Orléans
\item Solve the shallow-water equations using a finite volume method
\item Cartesian mesh
\item 3 mesh entities, 7 computation domains, 48 quantities
\item 98 computations (32 stencils, 66 local kernels)
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Evaluations}
\begin{block}{Evaluations}
\begin{itemize}
\item Data parallelization: no overhead
  \begin{itemize}
  \item weak scaling, fix size 400x400 (16.384 cores)
  \item strong scaling, 10kx10k:1k (16.384 cores)
  \end{itemize}
\item Hybrid parallelization: increase performance\\
  \begin{itemize}
  \item when data parallelism reaches its limits
  \item introduce new parallelism
  \end{itemize}
\end{itemize}
\end{block}

\begin{block}{Machines}
\begin{itemize}
\item Thin nodes TGCC Curie
\item 2 CPU, 8-cores, Intel Sandy Bridge EP (E5-2680) 2.7 GHz, 64 Go
\item OpenMPI and gcc 4.9
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Evaluations}
\begin{columns}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/median_weak.pdf}}
\end{column}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/median_strong.pdf}}
\end{column}
\end{columns}
\begin{center}
\textbf{No overhead introduced by components}
\end{center}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Evaluations}
\begin{columns}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/times.pdf}}
\end{column}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/base_close_median.pdf}}
\end{column}
\end{columns}
\begin{center}
\textbf{Good performance with 8 threads}
\end{center}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Perspectives}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Perspectives}
\begin{itemize}
\item Other schedulers (OpenMP 4, KStar)
\item Entire dependency graph
\item Other DDS + quantities + Sync (Global Arrays)
\item Other hardware (GPGPUs, MIC)
\item More than one mesh, multi-physics simulations
\item Unstructured meshes implementation (PamPA LaBRI + application)
\item Show interest of components
\end{itemize}
\end{frame}
%-------------------------------------------------------------

\begin{frame}
\begin{center}
\huge Thank You !
\end{center}
\end{frame}

%----------------------------------------------------------------------
\appendix
\backupbegin
%----------------------------------------------------------------------
\begin{frame}
\frametitle{From the dependency graph to a static scheduling}
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
  \node[] (s8) at (4.5,0.5) {$\mathcal{S}$};
   \node[] (s9) at (3,1.5) {$\mathcal{S}$};

   %reduction c0 c1
   \node[] (s0) at (1.5,2.5) {$\mathcal{S}$};
   \node[] (s1) at (1,3.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,4.5) {$k_0$};
   \node[] (star1) at (1.5,4.5) {$k_{0;1}^{sync}$};
   \node[] (c1) at (2,3.5) {$k_1$};

   %reduction c7 c8
   \node[] (s2) at (7.5,1.5) {$\mathcal{S}$};
   \node[] (s3) at (7,2.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,3.5) {$k_7$};
   \node[] (star8) at (7.5,3.5) {$k_{7;8}^{sync}$};
   \node[] (c8) at (8,2.5) {$k_8$};

   \node[] (p1) at (4.5,2.5) {$\mathcal{P}$};
   %reduction c3 c5
   \node[] (s5) at (5.5,3.5) {$\mathcal{S}$};
   \node[] (c3) at (5,4.5) {$k_3$};
   \node[] (c5) at (6,4.5) {$k_0$};
   %reduction c2 c6
   \node[] (s6) at (3.5,3.5) {$\mathcal{S}$};
   \node[] (p0) at (2.8,4.5) {$\mathcal{P}$};
   \node[] (c2) at (2.3,5.5) {$k_2$};
   \node[] (star4) at (3.3,5.5) {$k_{1;4}^{sync}$};
   \node[] (s7) at (4.2,4.5) {$\mathcal{S}$};
   \node[] (c4) at (3.8,5.5) {$k_4$};
   \node[] (c6) at (4.8,5.5) {$k_6$};

  \path[->]
    (s8) edge node {} (s9)
         edge node {} (s2)
    (s9) edge node {} (p1)
         edge node {} (s0)

    %reduction c0 c1
    (s0) edge node {} (s1)
         edge node {} (c1)
    (s1) edge node {} (c0)
         edge node {} (star1)
    %reduction c7 c8
    (s2) edge node {} (s3)
         edge node {} (c8)
    (s3) edge node {} (c7)
        edge node {} (star8)

    (p1) edge node {} (s6)
         edge node {} (s5)
    %reduction c3 c5
    (s5) edge node {} (c3)
         edge node {} (c5)
    %reduction c2 c6
    (s6) edge node {} (p0)
         edge node {} (s7)
    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4)
    (s7) edge node {} (c4)
         edge node {} (c6);
\end{tikzpicture}
\end{center}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{From the dependency graph to a static scheduling}
To build a static scheduling we build a \textbf{Serie-Parallel tree decomposition}
\begin{enumerate}
\item Transitive reduction
\item Remove the forbidden N-Shapes
\item Tree decomposition
\end{enumerate}

\medskip
\begin{center}
\begin{columns}
\begin{column}{5cm}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$k_0$};
   \node[] (bc0) at (-1,0) {};
   \node[] (c1) at (2,0) {$k_1$};
   \node[] (c2) at (0,-1) {$k_2$};
   \node[] (bc2) at (-1,-1) {};
   \node[] (c3) at (2,-1) {$k_3$};

  \path[->]
    (bc0) edge [dotted] node {} (c0)
    (bc2) edge [dotted] node {} (c2)
    (c0) edge node {} (c1)
          edge [dashed] node [swap] {} (c3)
    (c2) edge node {} (c1)
        edge node {} (c3);
\end{tikzpicture}
\end{column}
\begin{column}{5cm}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (S0) at (0,0) {$\mathcal{S}$};
   \node[] (bS0) at (-1,-1) {};
   \node[] (P1) at (-1,1) {$\mathcal{P}$};
   \node[] (aP1) at (-1.5,2) {};
   \node[] (aaP1) at (-0.5,2) {};
   \node[] (P2) at (1,1) {$\mathcal{P}$};
   \node[] (c0) at (-2,3) {$k_0$};
   \node[] (c1) at (0.5,3) {$k_1$};
   \node[] (c2) at (0,3) {$k_2$};
   \node[] (c3) at (2,3) {$k_3$};

  \path[->]
    (bS0) edge [dotted] node {} (S0)
    (S0) edge node {} (P1)
         edge node {} (P2)
    (P1) edge [dotted] node [swap] {} (aP1)
          edge [dotted] node [swap] {} (aaP1)
    (aP1) edge node {} (c0)
    (aaP1) edge node {} (c2)
    (P2) edge node {} (c1)
        edge node {} (c3);
\end{tikzpicture}
\end{column}
\end{columns}
\end{center}
\end{frame}
%----------------------------------------------------------------------
\backupend
%----------------------------------------------------------------------

\end{document}

### 2016-11-03 ARTICLES/OTHER/C2S_EXA/slides.tex
WIP...

\documentclass{beamer}

\usepackage{beamerthemeDresden}
\usepackage[utf8]{inputenc}
\usecolortheme{beaver}

\usepackage{listings}
\usepackage{xcolor}
\lstdefinestyle{base}{
  basicstyle=\ttfamily\color{black} \footnotesize,
  moredelim=**[is][\color{blue}]{@}{@},
  moredelim=**[is][\color{red}]{&}{&}
}

\setbeamertemplate{navigation symbols}{
\insertframenumber/
\inserttotalframenumber
}

\usepackage{color}
\definecolor{vert}{rgb}{0,0.5,0}
\definecolor{violet}{rgb}{0.5,0,0.5}

\usepackage{tikz}
\usetikzlibrary{calc,positioning,matrix,arrows,shapes.geometric,shapes.symbols,shapes.misc,shapes,automata,petri,decorations.markings,shadows}
\usepackage{array}
\usepackage{subfig}

%----------TIKZ
\usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}

\tikzset{
    use/.style={
    circle,draw=black,fill=black,scale=0.5,text=white
    },
    mpi/.style={
    rectangle,draw=black,fill=black,scale=0.8,text=white
    },
    provide/.style={
    circle,draw=black,fill=white,scale=0.5
    },
    component/.style={
    rectangle,rounded corners=3pt,draw=black
    },
    dcomponent/.style={
    rectangle,rounded corners=3pt,dashed,draw=black
    },
    progm/.style={rectangle,dashed, draw=black, thin, text width=10em, text centered,rounded corners, minimum height=2em},
  line/.style={draw, thin, ->, shorten >=2pt},
  purp/.style={rectangle, draw=violet,fill=violet!20, text width=10em, text centered,thin,rounded corners,inner sep=3pt},
  purp2/.style={rectangle, draw=violet,fill=violet!20, text width=15em, text centered,thin,rounded corners,inner sep=3pt},
  progbl/.style={rectangle, draw=blue, thin, text width=7em, text centered,rounded corners, minimum height=2em},
  progrg/.style={rectangle, draw=red, thin, text width=7em, text centered,rounded corners, minimum height=2em},
  proggr/.style={rectangle, draw=vert, thin, text width=7em, text centered,rounded corners, minimum height=2em},
  progbrg/.style={rectangle, draw=red, very thick, text width=10em, text centered,rounded corners, minimum height=2em}
}
%-------------

\def\pprec{\mathrel{\scalebox{.9}[1]{$\prec$}\mkern-3mu%
  \scalebox{.4}[1]{$\prec$}\mkern-5.5mu\scalebox{.4}[1]{$\prec$}}}

\makeatletter
    \newenvironment{withoutheadline}{
        \setbeamertemplate{headline}[default]
        \def\beamer@entrycode{\vspace*{-\headheight}}
    }{}
\makeatother

%-------------------------------------------------------------------
\title[The Multi-Stencil Language]{Using Component Model to implement a DSL: The Multi-Stencil Language Case Study}
\author[]{Hélène Coullon, Julien Bigot, \underline{Christian Perez}} % [Hélène Coullon (INRIA), Julien Bigot (CEA), Christian Perez (INRIA)]
\institute[]{Avalon project team\\Maison de la simulation}
\date{C2S@Exa -- Scientific Days \\November 8th, 2016\\ Paris}
%-------------------------------------------------------------------

\begin{document}
%gets rid of bottom navigation bars
\setbeamertemplate{footline}[page number]{}

%gets rid of navigation symbols
\setbeamertemplate{navigation symbols}{}
%---------------

\begin{frame}
    \titlepage
\end{frame}

%% \begin{frame}
%%   \begin{enumerate}
%%   \item On // langage => low level vs high level => a (temporary) solution: DSL ?
%%   \item On DSL (2 slides)
%%   \item how to productivively build efficient DSL (MDS + component) and their runtime
%%   \item this talk: a case study
%%   \item MSL overview
%%   \item Compiler overview
%%   \item runtime overview
%%   \item eval de perf
%%   \item conclusion
%%   \end{enumerate}
%% \end{frame}

%-------------------------------------------------------------
% INTRODUCTION / MOTIVATION
%-------------------------------------------------------------
\section{Introduction}
\begin{frame}
\frametitle{High performance computing and parallelism in 2016}
\begin{center}
\only<1>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Programming models}; \\
      \node[progrg] (mod) {Libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
    \end{scope}
  \end{tikzpicture}
}
%--------
\only<2>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Programming models}; \\
      \node[progrg] (mod) {Libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
    \end{scope}
  \end{tikzpicture}
}
%--------
\only<3>{

  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Programming models}; \\
      \node[progrg] (mod) {Libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \node[fill=white,right of=mod,node distance=3cm] (bsp) {\scriptsize \textcolor{red}{BSP}};
    \node[fill=white,right of=bsp,node distance=1.5cm] (mpi) {\scriptsize \textcolor{red}{MPI}};
    \node[fill=white,right of=mpi,node distance=1.5cm] (openmp) {\scriptsize \textcolor{red}{OpenMP}};
    \node[fill=white,right of=openmp,node distance=1.5cm] (cuda) {\scriptsize \textcolor{red}{Cuda}};
    \node[fill=white,right of=cuda,node distance=1.5cm] (opencl) {\scriptsize \textcolor{red}{OpenCL ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
      \path (mpi) edge[->,thin, bend right=50] (openmp);
      \path (mpi) edge[->,thin, bend right=50] (cuda);
      \path (bsp) edge[->,thin, bend right=50] (mpi);
    \end{scope}

  \end{tikzpicture}
}
%--------
\only<4>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Programming models}; \\
      \node[progrg] (mod) {Libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \node[fill=white,right of=mod,node distance=3cm] (bsp) {\scriptsize \textcolor{red}{BSP}};
    \node[fill=white,right of=bsp,node distance=1.5cm] (mpi) {\scriptsize \textcolor{red}{MPI}};
    \node[fill=white,right of=mpi,node distance=1.5cm] (openmp) {\scriptsize \textcolor{red}{OpenMP}};
    \node[fill=white,right of=openmp,node distance=1.5cm] (cuda) {\scriptsize \textcolor{red}{Cuda}};
    \node[fill=white,right of=cuda,node distance=1.5cm] (opencl) {\scriptsize \textcolor{red}{OpenCL ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
      \path (mpi) edge[->,thin, bend right=50] (openmp);
      \path (mpi) edge[->,thin, bend right=50] (cuda);
      \path (bsp) edge[->,thin, bend right=50] (mpi);

      \path (mpi) -- (donn);
      \path (mpi) -- (tach);
      \path (mpi) -- (mess);
      \path (bsp) -- (donn);
      \path (bsp) -- (mess);
      \path (openmp) -- (tach);
      \path (openmp) -- (donn);
      \path (cuda) -- (tach);
      \path (opencl) -- (tach);
      \path (opencl) -- (mess);
    \end{scope}
  \end{tikzpicture}
}
%---------
\only<5>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Programming models}; \\
      \node[progrg] (mod) {Libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \node[fill=white,right of=mod,node distance=3cm] (bsp) {\scriptsize \textcolor{red}{BSP}};
    \node[fill=white,right of=bsp,node distance=1.5cm] (mpi) {\scriptsize \textcolor{red}{MPI}};
    \node[fill=white,right of=mpi,node distance=1.5cm] (openmp) {\scriptsize \textcolor{red}{OpenMP}};
    \node[fill=white,right of=openmp,node distance=1.5cm] (cuda) {\scriptsize \textcolor{red}{Cuda}};
    \node[fill=white,right of=cuda,node distance=1.5cm] (opencl) {\scriptsize \textcolor{red}{OpenCL ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
      \path (mpi) edge[->,thin, bend right=50] (openmp);
      \path (mpi) edge[->,thin, bend right=50] (cuda);
      \path (bsp) edge[->,thin, bend right=50] (mpi);

      \path (mpi) -- (donn);
      \path (mpi) -- (tach);
      \path (mpi) -- (mess);
      \path (bsp) -- (donn);
      \path (bsp) -- (mess);
      \path (openmp) -- (tach);
      \path (openmp) -- (donn);
      \path (cuda) -- (tach);
      \path (opencl) -- (tach);
      \path (opencl) -- (mess);

      \path (mpi) -- (grap);
      \path (mpi) -- (mc);
      \path (mpi) -- (many);
      \path (bsp) -- (grap);
      \path (bsp) -- (mc);
      \path (cuda) -- (gpu);
      \path (openmp) -- (many);
      \path (openmp) -- (mc);
      \path (opencl) -- (gpu);
      \path (opencl) -- (mc);
      \path (opencl) -- (many);
      \path (opencl) -- (grap);
    \end{scope}

  \end{tikzpicture}

{\small\textbf{ETP4HPC}: programming models for non experts to reach exascale!}

}
  \end{center}
\end{frame}
%----------
%-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{"Please Help"}
%% %implicitly parallel (one sep. of concerns) / specific / performance
%% \begin{block}{HPC}
%% \begin{itemize}
%% \item Heterogenous and complex parallel hardware
%% \item Many programming models to combine
%% \item Expert programming
%% \end{itemize}
%% \end{block}

%% \begin{block}{Users}
%% Weather, climat, physics, biology etc.
%% \end{block}

%% \begin{center}
%% Separation of concerns between domain/parallelization and optimization\\
%% Domain specific languages (DSL)\\
%% \medskip
%% \resizebox{2cm}{!}{\includegraphics{images/scientist.jpg}}
%% \end{center}
%% \end{frame}
%-------------------------------------------------------------
\begin{frame}
  \frametitle{Domain Specific Languages} % HPC DSL
  \begin{block}{Advantages}
    \begin{itemize}
    \item Easy language for end users (can be application specific)
    \item Separation of concerns (domain/implementation)
    \item Implicit parallelization and optimizations
    \end{itemize}
  \end{block}
  \begin{alertblock}{Limitations}
    \begin{itemize}
    \item Difficulties deported to the DSL designer and implementer
      \begin{itemize}
      \item Low level high performance programming
      \item Maintainability and portability
      \end{itemize}
    \item Difficult to combine DSLs (interoperability)
      \begin{itemize}
      \item exascale applications = many specific domains and interactions
      \end{itemize}
    \end{itemize}
  \end{alertblock}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
  \frametitle{Component Models} % components
  \begin{block}{Overview}
    \begin{itemize}
    \item Component = black box with well defined interactions (provided/required)
    \item Application = Assembly of components
    \end{itemize}
  \end{block}
  \begin{block}{Benefits}
    \begin{itemize}
    \item Common model to define software architecture
    \item Maintainability through separation of concerns
    \item Code-reuse and productivity
    \end{itemize}
  \end{block}
  \begin{alertblock}{Open issues}
    \begin{itemize}
    \item Component for HPC, eg task graph?\\ {\color{red} Cf next talk of Jérôme Richard!}
    \end{itemize}
  \end{alertblock}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{This talk}
%implicitly parallel (one sep. of concerns) / specific / performance
\begin{block}{From a DSL: Multi Stencil Language}
\begin{itemize}
\item Descriptive language,
\item for Multi-Stencil simulations,
\item without numerical code.
\end{itemize}
\end{block}

\begin{block}{To a generated component assembly}
\begin{itemize}
\item with automatic synchronizations for data and task parallelism,
\item with empty functions to fill (components),
\item with good performance.
\end{itemize}
\end{block}

Separation of concerns between domain/implementation/parallelization

\end{frame}
%-------------------------------------------------------------
%% %-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{Motivation} % components
%% What if a DSL produces a component-based runtime ?

%% \begin{itemize}
%% \item Is it feasible ?
%% \item Is it efficient ?
%% \item \textcolor{gray}{Does it improve issues of DSLs ?}
%% \begin{itemize}
%% \item \textcolor{gray}{maintainability}
%% \item \textcolor{gray}{portability}
%% \item \textcolor{gray}{productivity}
%% \end{itemize}
%% \end{itemize}

%% Let's take a useful example: \textit{the Multi-Stencil Language }!
%% \end{frame}
%% %-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{How to improve it?} % HPC DSL
%% \begin{itemize}
%% \item Choose the good abstraction level to
%% \begin{itemize}
%% \item stay efficient
%% \item be used in many domains (find common meta-domains)
%% \end{itemize}
%% \item Choose good programming models for
%% \begin{itemize}
%% \item the compiler programming
%% \item the back-end
%% \end{itemize}
%% \item Try to reuse existing DSLs (compilation process or back-end codes)
%% \end{itemize}
%% % \textit{Delite (Scala)\\
%% % Composition and Reuse with Compiled Domain-Specific Languages\\
%% % ECOOP'13}
%% \end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{The Multi-Stencil Language}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: the mesh}
\begin{center}
\resizebox{6cm}{!}{\includegraphics{images/mesh.pdf}}
\end{center}
\begin{columns}
\begin{column}{6.5cm}
%    \fontsize{11}{11}\selectfont
  \begin{lstlisting}[frame=single]
mesh : cart
mesh entities : cell, edgex
computation domains :
  d1 in cell
  d2 in edgex
stencil shapes : 
  ncc from cell to cell
  nce from cell to edgex
  nec from edgex to cell
\end{lstlisting}
\end{column}
\begin{column}{5.5cm}
\begin{itemize}
\item A cartesian mesh is used
\item Two kinds of mesh entities are declared onto it
\item Two computations domains, one for each entity type
\item Three stencil shapes (neighborhood from mesh entity to mesh entity)
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: quantity}
\begin{columns}
\begin{column}{5cm}
\begin{lstlisting}[frame=single]
quantity :
  A,cell
  B,cell
  C,edgex
  D,cell
  E,cell
  F,cell
  G,cell
  H,edgex
  I,cell
  J,cell
scalar : mu, tau
\end{lstlisting}
\end{column}
\begin{column}{6cm}
\begin{itemize}
\item A is a quantity applied onto cell
\item C is a quantity applied onto edgex
\item mu and tau are scalar values
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: time and computations}
\begin{columns}
\begin{column}{6cm}
\begin{lstlisting}[frame=single]
time : 500
computations :
  B[d1] = k0({tau},{A})
  C[d2] = k1({},{B[nce]})
  D[d1] = k2({},{C})
  E[d1] = k3({},{C})
  F[d1] = k4({},{D,C[nec]})
  G[d1] = k0({mu,tau},{E})
  H[d2] = k6({},{F})
  I[d1] = k7({},{G,H})
  J[d1] = k8({mu},{I[ncc]})
\end{lstlisting}
\end{column}
\begin{column}{5cm}
\begin{itemize}
\item The time loop is composed of 500 iterations.
\item J is written
\item onto the computation domain d1,
\item by the computation k8,
\item which read the scalar mu
\item and the quantity I,
\item which is accessed with the neighborhood ncc.
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------

%% %-------------------------------------------------------------
%% \section{Multi-Stencil Language : Yet another DSL for stencils !}
%% %-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{Multi-Stencil application}
%% \begin{center}
%% \resizebox{\textwidth}{!}{%
%% \begin{tikzpicture}
%% \def \n {5}
%% \def \radius {3cm}
%% \def \margin {8} % margin in angles, depends on the radius

%%   \node[purp2,align=center] (edp) at ({360/\n * (2 - 1)}:\radius) {Partial Differential Equations\\\scriptsize $\frac{\partial u(x,y,t)}{\partial t} = \frac{\partial^2 u(x,y,t)}{\partial x^2} + \frac{\partial^2 u(x,y,t)}{\partial y^2}$};
%%   \node[progm,align=center,right of=edp,node distance=6cm] (cond) {\scriptsize + specific behavior for\\\scriptsize boundary conditions};
%%   \path[line,dashed] (edp) -- (cond);
%%   \node[purp,align=center] (maillage) at ({360/\n * (1 - 1)}:\radius) {Mesh\\Time iterations};
%%   \path[->] (maillage) edge[<-,thin,shorten <=2pt,shorten >=2pt, bend right=40]  node [fill=white,align=center] {\scriptsize Time and space\\\scriptsize discretization} (edp.east);
%%   \path[] (edp.west) edge[thin,shorten <=2pt,shorten >=2pt, bend right=90]  node [draw,circle,fill=white] (plus) {\scriptsize +} (maillage.west);
%%   \node[purp,align=center, below of=plus, node distance=3.5cm] (sche) {Explicit\\numerical schemes\\\textcolor{red}{stencils}};
%%   \node[progm, text width=11em,align=center,right of=sche,node distance=5.5cm] (expl) {\scriptsize Iterative neighboorhood approximation of the phenomena};
%%   \path[line,dashed] (sche) -- (expl);
%%   \path[line] (plus) -- node [fill=white,align=center] {\scriptsize Numerical methods\\\scriptsize Finite difference/volume/elment} (sche);
%% \end{tikzpicture}
%% }
%% \end{center}
%% \end{frame}
%% %----------------
%% \begin{frame}
%% \frametitle{Multi-Stencil Language : Yet another DSL for stencils !}
%% \begin{block}{Inputs}
%% follow a descriptive grammar to simply declare
%% \begin{itemize}
%% \item the sequential order of computations
%% \item what is read and written by each computation
%% \item with or without a neighborhood access (stencil)
%% \end{itemize}
%% \end{block}
%% \begin{block}{Outputs}
%% \begin{itemize}
%% \item parallel pattern of the multi-stencil application
%% \item hybrid parallelism included: data and task parallelism
%% \item choice of parallel languages (MPI, OpenMP 3.x/4.x, SkelGIS etc.)
%% \item end-users have to fill computation functions into this pattern
%% \end{itemize}
%% \end{block}
%% \end{frame}
%% %----------------
%% \begin{frame}
%% \frametitle{Multi-Stencil Language : Yet another DSL for stencils !}
%% \begin{itemize}
%% \item Choosen abstraction level
%% \begin{itemize}
%% \item mesh-agnostic description language
%% \begin{itemize}
%% \item should be usable for unstructured, Cartesian, curvilinear meshes
%% \end{itemize}
%% \item description concerns splited from implementation concerns
%% \begin{itemize}
%% \item usable with many different parallel libraries or languages back-ends
%% \end{itemize}
%% \end{itemize}
%% \item Choosen programming models for back-end
%% \begin{itemize}
%% \item Component programming model (maintainability, portability and composability)
%% \item Task scheduling model (portability and efficiency)
%% \end{itemize}
%% \item Reuse of existing languages and DSLs
%% \begin{itemize}
%% \item Use of a Distributed Data Structure DSeL (C++,MPI) : SkelGIS
%% \item MPI and OpenMP (3.x and 4.x)
%% \end{itemize}
%% \end{itemize}
%% \end{frame}
%% %-------------------------------------------------------------

%-------------------------------------------------------------
\section{From DSL to Component Assembly}
%-------------------------------------------------------------
%-------------------------------------------------------------
\begin{frame}
\frametitle{What is produced ?}
A data parallel/task parallel component assembly.\vspace{-0.2em}

\begin{center}
\begin{tikzpicture}[remember picture]
  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
    \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
     \node[component,solid,draw=black] (D) at (0,0) {$Driver$};
     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
     \node[solid] (Ds) at (-1.5,0) {start};
     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};

     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
     \node[solid,left=0.8cm of T] (tt) {$T$};

     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
     \node[component,solid,right=2cm of Cp] (C) {$Scheduler$};
     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
     \node[use,scale=0.5,solid,right=1.5cm of C] (Cu2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Cu2] (Cu3) {};
     \node[use,scale=0.5,solid,below=0.8cm of Cu3] (Cu4) {};
     \node[solid,right=0.2cm of Cu] (star) {$*$};
     %\node[solid,right=1.5cm of C] (gamma) {$\Gamma$};

     \node[solid,below=1cm of C] (calc) {$\Gamma_{sync}$, $\Gamma_{dep}$};

     \node[provide,scale=0.5,solid,right=0.15 of Cu2] (k0p) {};
     \node[component,solid,right=1.5cm of k0p] (k0) {$k_0$};

     \node[provide,scale=0.5,solid,right=0.15 of Cu3] (k0sp) {};
     \node[component,solid,right=1.5cm of k0sp] (k0s) {$k_{0;1}^{sync}$};

     \node[provide,scale=0.5,solid,right=0.15 of Cu4] (kotherp) {};
     \node[component,solid,right=1.5cm of kotherp] (kother) {$\dots$};

     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Quantity$};
     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
     \node[solid,left=1.5cm of Data] (delta) {$\Delta,\mathcal{D}$};
     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};

     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
     \node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};

    \path[-]
      (Dp) edge [solid] node {} (D)
      (D) edge [solid] node {} (Du1)
          edge [solid] node {} (Du2)
          edge [solid] node {} (Du3)
      (DDSp1) edge [solid] node {} (DDS)
      (Tp) edge [solid] node {} (T)
      (T)  edge [solid] node {} (Tu)
      (Cp) edge [solid] node {} (C)
      (C) edge [solid] node {} (Cu)
      (Datap2) edge [solid] node {} (Data)
      (Datap1) edge [solid] node {} (Data)
      (Data) edge [solid] node {} (Datau)
      (DDSp2) edge [solid] node {} (DDS)
      (C)  edge [solid] node {} (Cu2)
          edge [solid] node {} (Cu3)
          edge [solid] node {} (Cu4)
      (Cu2)  edge [solid] node {} (k0p)
      (k0p)  edge [solid] node {} (k0)
      (Cu3)  edge [solid] node {} (k0sp)
      (k0sp)  edge [solid] node {} (k0s)
      (Cu4)  edge [solid] node {} (kotherp)
      (kotherp)  edge [solid] node {} (kother);
    \end{tikzpicture}
    };
    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=4cm of Data] (mpidata) {};
    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
    \node[solid,above=3.2cm of proc.west,anchor=west] (core) {\textit{\tiny{Applied on each resource}}};
    \path[-]
      (Data.east) edge node {} (mpidata);
  \end{tikzpicture}
\end{center}
{\fontsize{8}{8}\selectfont DDS: Distributed Data Structure (using SkelGIS)\par}
\end{frame}
%% %-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{What is produced ?}
%% \begin{block}{Implementation}
%% \begin{itemize}
%% \item DDS + Quantity + $k_{*}^{sync}$ = written components using SkelGIS (distributed memory, MPI)
%% \item Scheduler = Serie-Parallel graph of $\Gamma_{dep}$ (OpenMP 3.0)
%% \item Time + Scheduler = generated components
%% \item $k_i$ = empty components to write
%% \end{itemize}
%% \end{block}

%% \begin{block}{Separation of concerns}
%% \begin{itemize}
%% \item numerician (MSL)
%% \item computer engineer (K)
%% \item HPC engineer (DDS, quantity, $k_{*}^{sync}$ etc.)
%% \item DSL designer (component assembly + generated components)
%% \end{itemize}
%% \end{block}
%% \end{frame}
%% %-------------------------------------------------------------

%----------------
\begin{frame}
\frametitle{Data and Task parallelism}
\begin{block}{Synchronizations}
When a computation read a data, using a stencil shape, that has been written by a previous computation.
\end{block}

$\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$\\
$\hookrightarrow [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$

\medskip

\begin{block}{Dependency graph}
%\begin{enumerate}
Node: a computation or a synchronization\\
Edge: a direct data dependency (Read / Write)
\end{block}

\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node (c0) at (0,0) {$c_0$};
   \node[right=1 of c0] (sy1) {$sync_1$};
   \node[right=1 of sy1] (c1) {$c_1$};
   \node[above right=1 of c1] (c2) {$c_2$};
   \node[below right=1 of c1] (c3) {$c_3$};
   \node[right=1 of c2] (sy4) {$sync_4$};
   \node[right=1 of c3] (c5) {$c_5$};
   \node[right=1 of sy4] (c4) {$c_4$};
   \node[right=1 of c4] (c6) {$c_6$};
   \node[below right=1 of c6] (c7) {$c_7$};
   \node[right=1 of c7] (sy8) {$sync_8$};
   \node[right=1 of sy8] (c8) {$c_8$};

  \path[->]
    (c0) edge node {} (sy1)
    (sy1) edge node {} (c1)
    (c1)  edge node {} (c2)
          edge node {} (c3)
    (c2) edge node {} (sy4)
    (sy4) edge node {} (c4)
    (c4) edge node {} (c6)
    (c3) edge node {} (c5)
    (c5) edge node {} (c7)
    (c6) edge node {} (c7)
    (c7) edge node {} (sy8)
    (sy8) edge node {} (c8);
\end{tikzpicture}
\end{center}

\begin{center}
\textcolor{red}{Static} or \textcolor{red}{dynamic} scheduling ?
\end{center}
\end{frame}
%----------------
%% \begin{frame}
%% \frametitle{Series-Parallel Tree}
%% \small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
%% \normalsize
%% \begin{columns}
%% \begin{column}{.48\textwidth}
%% \begin{center}
%% \resizebox{\textwidth}{!}{%
%% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
%%    \node[] (s0)    at (0,0) {$\mathcal{S}$};

%%    \node[] (c0)    at (-3,1) {$c_0$};
%%    \node[] (star1) at (-2,1) {$sync_1$};
%%    \node[] (c1)    at (-1,1) {$c_1$};
%%    \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
%%    \node[] (c7)    at ( 1,1) {$c_7$};
%%    \node[] (star8) at ( 2,1) {$sync_8$};
%%    \node[] (c8)    at ( 3,1) {$c_8$};

%%    \node[] (s1)    at (-1,2) {$\mathcal{S}$};
%%    \node[] (s2)    at (1,2) {$\mathcal{S}$};

%%    \node[] (c2)    at (-3,3) {$c_2$};
%%    \node[] (star4) at (-2,3) {$sync_4$};
%%    \node[] (c4)    at (-1,3) {$c_4$};
%%    \node[] (c6)    at ( 0,3) {$c_6$};
%%    \node[] (c3)    at ( 1,3) {$c_3$};
%%    \node[] (c5)    at ( 2,3) {$c_5$};

%%   \path[->]
%%     (s0) edge node {} (c0)
%%          edge node {} (star1)
%%          edge node {} (c1)
%%          edge node {} (p0)
%%          edge node {} (c7)
%%          edge node {} (star8)
%%          edge node {} (c8)
%%     (p0) edge node {} (s1)
%%          edge node {} (s2)
%%     (s1) edge node {} (c2)
%%          edge node {} (star4)
%%          edge node {} (c4)
%%          edge node {} (c4)
%%          edge node {} (c6)
%%     (s2) edge node {} (c3)
%%          edge node {} (c5);
%%   \end{tikzpicture}
%%   }
%% \end{center}
%% \end{column}
%% \begin{column}{.52\textwidth}

%% \end{column}
%% \end{columns}
%% \end{frame}
%% %----------------

%% %----------------
%% \begin{frame}
%% \frametitle{Series-Parallel Tree}
%% \small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
%% \normalsize
%% \begin{columns}
%% \begin{column}{.48\textwidth}
%% \begin{center}
%% \resizebox{\textwidth}{!}{%
%% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
%%    \node[] (s0)    at (0,0) {$\mathcal{S}$};

%%    \node[] (c0)    at (-3,1) {$c_0$};
%%    \node[] (star1) at (-2,1) {$sync_1$};
%%    \node[] (c1)    at (-1,1) {$c_1$};
%%    \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
%%    \node[] (c7)    at ( 1,1) {$c_7$};
%%    \node[] (star8) at ( 2,1) {$sync_8$};
%%    \node[] (c8)    at ( 3,1) {$c_8$};

%%    \node[] (s1)    at (-1,2) {$\mathcal{S}$};
%%    \node[] (s2)    at (1,2) {$\mathcal{S}$};

%%    \node[] (c2)    at (-3,3) {$c_2$};
%%    \node[] (star4) at (-2,3) {$sync_4$};
%%    \node[] (c4)    at (-1,3) {$c_4$};
%%    \node[] (c6)    at ( 0,3) {$c_6$};
%%    \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
%%    %\node[] (c5)    at ( 2,3) {$c_5$};

%%   \path[->]
%%     (s0) edge node {} (c0)
%%          edge node {} (star1)
%%          edge node {} (c1)
%%          edge node {} (p0)
%%          edge node {} (c7)
%%          edge node {} (star8)
%%          edge node {} (c8)
%%     (p0) edge node {} (s1)
%%          edge node {} (s2)
%%     (s1) edge node {} (c2)
%%          edge node {} (star4)
%%          edge node {} (c4)
%%          edge node {} (c4)
%%          edge node {} (c6)
%%     (s2) edge node {} (c3);
%%   \end{tikzpicture}
%%   }
%% \end{center}
%% \end{column}
%% \begin{column}{.52\textwidth}
%% \end{column}
%% \end{columns}
%% Loop fusion optimization possible
%% \end{frame}
%% %----------------

%% %----------------
%% \begin{frame}
%% \frametitle{Series-Parallel Tree}
%% \small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
%% \normalsize
%% \begin{columns}
%% \begin{column}{.48\textwidth}
%% \begin{center}
%% \resizebox{\textwidth}{!}{%
%% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
%%    \node[] (s0)    at (0,0) {$\mathcal{S}$};

%%    \node[] (c0)    at (-3,1) {$c_0$};
%%    \node[] (star1) at (-2,1) {$sync_1$};
%%    \node[] (c1)    at (-1,1) {$c_1$};
%%    \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
%%    \node[] (c7)    at ( 1,1) {$c_7$};
%%    \node[] (star8) at ( 2,1) {$sync_8$};
%%    \node[] (c8)    at ( 3,1) {$c_8$};

%%    \node[] (s1)    at (-1,2) {$\mathcal{S}$};
%%    \node[] (s2)    at (1,2) {$\mathcal{S}$};

%%    \node[] (c2)    at (-3,3) {$c_2$};
%%    \node[] (star4) at (-2,3) {$sync_4$};
%%    \node[] (c4)    at (-1,3) {$c_4$};
%%    \node[] (c6)    at ( 0,3) {$c_6$};
%%    \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
%%    %\node[] (c5)    at ( 2,3) {$c_5$};

%%   \path[->]
%%     (s0) edge node {} (c0)
%%          edge node {} (star1)
%%          edge node {} (c1)
%%          edge node {} (p0)
%%          edge node {} (c7)
%%          edge node {} (star8)
%%          edge node {} (c8)
%%     (p0) edge node {} (s1)
%%          edge node {} (s2)
%%     (s1) edge node {} (c2)
%%          edge node {} (star4)
%%          edge node {} (c4)
%%          edge node {} (c4)
%%          edge node {} (c6)
%%     (s2) edge node {} (c3);
%%   \end{tikzpicture}
%%   }
%% \end{center}
%% \end{column}
%% \begin{column}{.52\textwidth}
%% \begin{block}{Specific components}
%% \begin{itemize}
%% \item $SEQ$ to directly replace $\mathcal{S}$ nodes
%% \item $PAR$ to directly replace $\mathcal{P}$ nodes
%% \item $SYNC$ for synchronizations
%% \item $K$ for computation kernels
%% \end{itemize}
%% \end{block}
%% \end{column}
%% \end{columns}
%% Loop fusion optimization possible
%% \end{frame}
%% %----------------

%% %-------------------------------------------------------------
%% \section{Implementation}
%% %-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{Component models}
%% \begin{itemize}
%% \item Black box with code
%% \item Use/provide interfaces
%% \item Improve code reuse (productivity)
%% \item Improve separation of concerns (maintainability and portability)
%% \item Improve composability of applications
%% \end{itemize}
%% \begin{center}
%% \begin{tikzpicture}[remember picture]
%%   \node[component,solid] (A) at (0,0) {A};
%%   \node[component,solid,right=5cm of A] (B) {B};
%%   \node[use,scale=0.5,solid,right=1.5cm of A] (Au) {};
%%   \node[provide,scale=0.5,solid,left=1.5cm of B] (Bp) {};
%%   \path[-]
%%     (A) edge node {} (Au)
%%     (Bp) edge node {} (B);
%% \end{tikzpicture}

%% \bigskip
%% \begin{tikzpicture}[remember picture]
%%   \node[component,solid] (A) at (0,0) {A};
%%   \node[use,scale=0.5,solid,right=1cm of A] (Au) {};
%%   \node[provide,scale=0.5,solid,right=0.01cm of Au] (Bp) {};
%%   \node[component,solid,right=1cm of Bp] (B) {B};
%%   \path[-]
%%     (A) edge node {} (Au)
%%     (Bp) edge node {} (B);
%% \end{tikzpicture}

%% \medskip
%% Components A and B into a component assembly
%% \end{center}
%% \end{frame}

%-------------------------------------------------------------
\section{Evaluations}
%----------------
\begin{frame}
\frametitle{Evaluations}
\begin{block}{Application}
\begin{itemize}
\item FullSWOF2D: developed at the MAPMO, Université d'Orléans
\item Solve the shallow-water equations using a finite volume method
\item Cartesian mesh
\item 3 mesh entities, 7 computation domains, 48 quantities
\item 98 computations (32 stencils, 66 local kernels)
\end{itemize}
\end{block}
\end{frame}
%----------------

\begin{frame}
\frametitle{Evaluations}
\begin{block}{Evaluations}
\begin{itemize}
\item Data parallelization: no overhead
  \begin{itemize}
  \item weak scaling, fix size 400x400 (16.384 cores)
  \item strong scaling, 10kx10k:1k (16.384 cores)
  \end{itemize}
\item Hybrid parallelization: increase performance\\
  \begin{itemize}
  \item when data parallelism reaches its limits
  \item introduce new parallelism
  \end{itemize}
\end{itemize}
\end{block}

\begin{block}{Machines}
\begin{itemize}
\item Thin nodes TGCC Curie
\item 2 CPU, 8-cores, Intel Sandy Bridge EP (E5-2680) 2.7 GHz, 64 Go
\item OpenMPI and gcc 4.9
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Evaluations}
\begin{columns}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/median_weak.pdf}}
\end{column}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/median_strong.pdf}}
\end{column}
\end{columns}
\begin{center}
\textbf{No overhead introduced by components}
\end{center}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Evaluations}
\begin{columns}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/times.pdf}}
\end{column}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/base_close_median.pdf}}
\end{column}
\end{columns}
\begin{center}
\textbf{Good performance with 8 threads}
\end{center}
\end{frame}
%-------------------------------------------------------------

%----------------
\begin{frame}
\frametitle{Evaluations}
\textbf{Improvment with data+task scheduling}
\begin{center}
\resizebox{8cm}{!}{\includegraphics{./images/base_close_median.pdf}}
\end{center}
\end{frame}

%-------------------------------------------------------------
\section{Conclusion}
%-------------------------------------------------------------
%\subsection{Conclusion and perspectives}
\begin{frame}
\frametitle{Conclusion}
\begin{block}{Conclusion}
\begin{itemize}
\item A DSL for Multi-Stencil applications (MSL)
\item The compilation of MSL to get a parallel scheduling pattern of the simulation
\begin{itemize}
\item Data parallelism
\item Task parallelism
\end{itemize}
\item The dump to a component-based runtime
\item Data parallelism evaluation: no overhead introduced 
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\begin{frame}
\begin{block}{Perspectives}
\frametitle{Perspectives}
\begin{itemize}
\item Improvment of the language (convergence criteria, reduction etc.)
\item Scalability up to $32k$ cores on TGCC Curie (CEA)
\item Evaluations on Data+Task parallelism
\begin{itemize}
\item OpenMP 3 inside kernels
\end{itemize}
\item Dynamic scheduling
\begin{itemize}
\item OpenMP 4 with a scheduling component
\item Kstar for StarPU and XKaapi runtimes
\end{itemize}
\item CPU+GPGPUs using stencil compilers (Pochoir, PATUS etc.)
\end{itemize}
$\hookrightarrow$ Show portability, maintainability introduced by components
\end{block}
\end{frame}
%-------------------------------------------------------------

%% %-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{Conclusion}
%% \begin{block}{Work in progress}
%% \begin{itemize}
%% \item Different types of scheduling and back-end in MSL
%% \item Combine MSL implementation choices to an energy-aware framework
%% \item New programming model back-end for MSL: combining component models and task models (Ph.D. Jérôme Richard)
%% \end{itemize}
%% \end{block}
%% \begin{block}{Collaborations}
%% \begin{itemize}
%% \item More application cases
%% \item Back-end programming model: component models and task scheduling (Components in OmpSs ?)
%% \item DSLs interoperability
%% \end{itemize}
%% \end{block}
%% \end{frame}
%-------------------------------------------------------------

% \begin{withoutheadline}
% \begin{frame}{}
% \begin{center}
% \huge Thank You !
% \end{center}
% \end{frame}
% \end{withoutheadline}

%----------------
%% \begin{frame}
%% \frametitle{MSL to Component-based runtime}
%% \begin{center}
%% \begin{tikzpicture}[remember picture]
%%   \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
%%     \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
%%      \node[component,solid] (D) at (0,0) {$Driver$};
%%      \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
%%      \node[solid] (Ds) at (-1.5,0) {start};
%%      \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
%%      \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
%%      \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};

%%      \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
%%      \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
%%      \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};

%%      \node[provide,scale=0.5,solid,right=0.15 of Tu] (Sp) {};
%%      \node[component,solid,right=1.5cm of Sp] (S) {$Scheduling$};
%%      \node[use,scale=0.5,solid,right=1.5cm of S] (Su) {};

%%      \node[provide,scale=0.5,solid,right=0.15 of Su] (Cp) {};
%%      \node[component,solid,double,right=2cm of Cp] (C) {$Kernel$};
%%      \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};

%%      \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
%%      \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
%%      \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
%%      \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};

%%      \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
%%      \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
%%      \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};

%%     \path[-]
%%       (Dp) edge [solid] node {} (D)
%%       (D) edge [solid] node {} (Du1)
%%           edge [solid] node {} (Du2)
%%           edge [solid] node {} (Du3)
%%       (DDSp1) edge [solid] node {} (DDS)
%%       (Tp) edge [solid] node {} (T)
%%       (T)  edge [solid] node {} (Tu)
%%       (Sp) edge [solid] node {} (S)
%%       (S) edge [solid] node {} (Su)
%%       (Datap2) edge [solid] node {} (Data)
%%       (Datap1) edge [solid] node {} (Data)
%%       (Data) edge [solid] node {} (Datau)
%%       (C) edge [solid] node {} (Cu)
%%       	  edge [solid] node {} (Cp)
%%       (DDSp2) edge [solid] node {} (DDS);
%%     \end{tikzpicture}
%%     };
%%     \node[mpi,scale=0.8,rounded corners=0pt,solid,right=0.5cm of proc] (mpidata) {};
%%     \node[solid,above=0.1cm of mpidata] (star2) {$*$};
%%     \node[solid,below=2.5cm of proc.west,anchor=west] (core) {\textit{\tiny{Duplicated on each processor/core}}};
%%     \path[-]
%%       (Data.east) edge node {} (mpidata);
%%   \end{tikzpicture}
%% \end{center}
%% \end{frame}

\end{document}

### 2016-11-04 ARTICLES/OTHER/C2S_EXA/slides.tex
WIP ..

@@
 \institute[]{Avalon project team\\Maison de la simulation}
-\date{C2S@Exa -- Scientific Days \\November 8th, 2016\\ Paris}
+\date{C2S@Exa -- Scientific Days \\Paris, November 8th, 2016}
 %-------------------------------------------------------------------
@@
     \titlepage
+    \vspace{1.25cm}
+    \resizebox{2.5cm}{!}{\includegraphics{images/logo-inria.png}}\hspace{1cm}
+    \resizebox{1cm}{!}{\includegraphics{images/logo-mdls.png}}\hfill
+    \resizebox{1cm}{!}{\includegraphics{images/elci_logo.png}}
 \end{frame}
@@
 %-------------------------------------------------------------
-%-------------------------------------------------------------
+%----------------
 \begin{frame}
-\frametitle{What is produced ?}
+\frametitle{Data and Task parallelism}
+\begin{block}{Synchronizations}
+When a computation read a data, using a stencil shape, that has been written by a previous computation.
+\end{block}
+
+$\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$\\
+$\hookrightarrow [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$
+
+\medskip
+
+\begin{block}{Dependency graph}
+%\begin{enumerate}
+Node: a computation or a synchronization\\
+Edge: a direct data dependency (Read / Write)
+\end{block}
+
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node (c0) at (0,0) {$c_0$};
+   \node[right=1 of c0] (sy1) {$sync_1$};
+   \node[right=1 of sy1] (c1) {$c_1$};
+   \node[above right=1 of c1] (c2) {$c_2$};
+   \node[below right=1 of c1] (c3) {$c_3$};
+   \node[right=1 of c2] (sy4) {$sync_4$};
+   \node[right=1 of c3] (c5) {$c_5$};
+   \node[right=1 of sy4] (c4) {$c_4$};
+   \node[right=1 of c4] (c6) {$c_6$};
+   \node[below right=1 of c6] (c7) {$c_7$};
+   \node[right=1 of c7] (sy8) {$sync_8$};
+   \node[right=1 of sy8] (c8) {$c_8$};
+
+  \path[->]
+    (c0) edge node {} (sy1)
+    (sy1) edge node {} (c1)
+    (c1)  edge node {} (c2)
+          edge node {} (c3)
+    (c2) edge node {} (sy4)
+    (sy4) edge node {} (c4)
+    (c4) edge node {} (c6)
+    (c3) edge node {} (c5)
+    (c5) edge node {} (c7)
+    (c6) edge node {} (c7)
+    (c7) edge node {} (sy8)
+    (sy8) edge node {} (c8);
+\end{tikzpicture}
+\end{center}
+
+\begin{center}
+\textcolor{red}{Static} or \textcolor{red}{dynamic} scheduling ?
+\end{center}
+\end{frame}
+%----------------
+%----------------
+\begin{frame}
+\frametitle{MSL to Component-based runtime}
+\begin{block}{Ready-to-fill parallel pattern}
+\begin{itemize}
+\item Data parallelism
+\begin{itemize}
+\item External distributed data structure
+\item Automatic detection of synchronizations
+\end{itemize}
+\item Task parallelism (mid-grain)
+\begin{itemize}
+\item Static scheduling at the assembly level 
+\item Dynamic scheduling within a driver component
+  \begin{itemize}
+  \item OpenMP task (For loop also available)
+  \end{itemize}
+\end{itemize}
+\end{itemize}
+\end{block}
+The fine grain task parallelism is left to other languages: 
+\begin{itemize}
+\item OpenMP in the kernels
+\item Kernels generated by stencil compilers
+  \begin{itemize}
+  \item Pochoir, PATUS, Liszt etc.
+  \end{itemize}
+\end{itemize}
+\end{frame}
+%----------------
+%-------------------------------------------------------------
+\begin{frame}
+\frametitle{Component Assembly Runtime Template}
 A data parallel/task parallel component assembly.\vspace{-0.2em}
@@
 
-%----------------
-\begin{frame}
-\frametitle{Data and Task parallelism}
-\begin{block}{Synchronizations}
-When a computation read a data, using a stencil shape, that has been written by a previous computation.
-\end{block}
-
-$\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$\\
-$\hookrightarrow [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$
-
-\medskip
-
-\begin{block}{Dependency graph}
-%\begin{enumerate}
-Node: a computation or a synchronization\\
-Edge: a direct data dependency (Read / Write)
-\end{block}
-
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node (c0) at (0,0) {$c_0$};
-   \node[right=1 of c0] (sy1) {$sync_1$};
-   \node[right=1 of sy1] (c1) {$c_1$};
-   \node[above right=1 of c1] (c2) {$c_2$};
-   \node[below right=1 of c1] (c3) {$c_3$};
-   \node[right=1 of c2] (sy4) {$sync_4$};
-   \node[right=1 of c3] (c5) {$c_5$};
-   \node[right=1 of sy4] (c4) {$c_4$};
-   \node[right=1 of c4] (c6) {$c_6$};
-   \node[below right=1 of c6] (c7) {$c_7$};
-   \node[right=1 of c7] (sy8) {$sync_8$};
-   \node[right=1 of sy8] (c8) {$c_8$};
-
-  \path[->]
-    (c0) edge node {} (sy1)
-    (sy1) edge node {} (c1)
-    (c1)  edge node {} (c2)
-          edge node {} (c3)
-    (c2) edge node {} (sy4)
-    (sy4) edge node {} (c4)
-    (c4) edge node {} (c6)
-    (c3) edge node {} (c5)
-    (c5) edge node {} (c7)
-    (c6) edge node {} (c7)
-    (c7) edge node {} (sy8)
-    (sy8) edge node {} (c8);
-\end{tikzpicture}
-\end{center}
-
-\begin{center}
-\textcolor{red}{Static} or \textcolor{red}{dynamic} scheduling ?
-\end{center}
-\end{frame}
-%----------------
 %% \begin{frame}
@@
 \end{block}
+\begin{block}{Machines}
+\begin{itemize}
+\item Thin nodes TGCC Curie
+\item 2 CPU, 8 cores, Sandy Bridge EP (E5-2680) 2.7 GHz, 64 GB
+\item OpenMPI and gcc 4.9
+\end{itemize}
+\end{block}
 \end{frame}
 %----------------
-
+%-------------------------------------------------------------
 \begin{frame}
-\frametitle{Evaluations}
-\begin{block}{Evaluations}
-\begin{itemize}
-\item Data parallelization: no overhead
-  \begin{itemize}
-  \item weak scaling, fix size 400x400 (16.384 cores)
-  \item strong scaling, 10kx10k:1k (16.384 cores)
-  \end{itemize}
-\item Hybrid parallelization: increase performance\\
-  \begin{itemize}
-  \item when data parallelism reaches its limits
-  \item introduce new parallelism
-  \end{itemize}
-\end{itemize}
-\end{block}
-
-\begin{block}{Machines}
-\begin{itemize}
-\item Thin nodes TGCC Curie
-\item 2 CPU, 8-cores, Intel Sandy Bridge EP (E5-2680) 2.7 GHz, 64 Go
-\item OpenMPI and gcc 4.9
-\end{itemize}
-\end{block}
-\end{frame}
-%-------------------------------------------------------------
-\begin{frame}
-\frametitle{Evaluations}
+\frametitle{Data Parallelization: Weak and Strong Scaling}
 \begin{columns}
@@
 \resizebox{6cm}{!}{\includegraphics{images/median_weak.pdf}}
+\begin{center}
+\vspace{-1em}
+  Weak Scaling\\
+  400x400 domain/core
+\end{center}
 \end{column}
@@
 \resizebox{6cm}{!}{\includegraphics{images/median_strong.pdf}}
+\begin{center}
+\vspace{-1em}
+  Strong Scaling\\
+  10k x 10k domaine
+\end{center}
 \end{column}
 \end{columns}
+\bigskip
 \begin{center}
@@
 \begin{frame}
-\frametitle{Evaluations}
+\frametitle{Data Paralization to Hybrid Parallization: Strong Scaling}
 \begin{columns}
@@
   \resizebox{6cm}{!}{\includegraphics{images/times.pdf}}
+  \begin{center}
+\vspace{-1em}\small    Data Parallelization    
+  \end{center}
 \end{column}
@@
 \resizebox{6cm}{!}{\includegraphics{images/base_close_median.pdf}}
+\begin{center}
+\vspace{-1em}\small  Data and Hybrid Parallelizations
+\end{center}
 \end{column}
 \end{columns}
+\bigskip
 \begin{center}
-\textbf{Good performance with 8 threads}
+{\small
+  \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+    \hline 
+   Parallelism Level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
+   \hline
+   Frequency & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
+   \hline
+ \end{tabular}
+ \\
+ Task graph analysis
+}
 \end{center}
@@
 %-------------------------------------------------------------
-
-%----------------
-\begin{frame}
-\frametitle{Evaluations}
-\textbf{Improvment with data+task scheduling}
-\begin{center}
-\resizebox{8cm}{!}{\includegraphics{./images/base_close_median.pdf}}
-\end{center}
-\end{frame}
 
@@
 \begin{frame}
-\frametitle{Conclusion}
+\frametitle{Conclusion and Future Work}
 \begin{block}{Conclusion}
@@
 \item A DSL for Multi-Stencil applications (MSL)
-\item The compilation of MSL to get a parallel scheduling pattern of the simulation
+\item The generation of a component based runtime, including scheduling
 \begin{itemize}
@@
 \end{itemize}
-\item The dump to a component-based runtime
-\item Data parallelism evaluation: no overhead introduced 
+\item Evaluation on Curie: no overhead introduced 
 \end{itemize}
 \end{block}
-\end{frame}
-%-------------------------------------------------------------
-
-%-------------------------------------------------------------
-\begin{frame}
-\begin{block}{Perspectives}
-\frametitle{Perspectives}
-\begin{itemize}
-\item Improvment of the language (convergence criteria, reduction etc.)
-\item Scalability up to $32k$ cores on TGCC Curie (CEA)
-\item Evaluations on Data+Task parallelism
-\begin{itemize}
+
+\begin{block}{Future Work}
+  \begin{itemize}
+  \item Language improvement (convergence criteria, reduction etc.)
   \item OpenMP 3 inside kernels
-\end{itemize}
-\item Dynamic scheduling
-\begin{itemize}
-\item OpenMP 4 with a scheduling component
-\item Kstar for StarPU and XKaapi runtimes
-\end{itemize}
   \item CPU+GPGPUs using stencil compilers (Pochoir, PATUS etc.)
-\end{itemize}
 $\hookrightarrow$ Show portability, maintainability introduced by components
+
+  \end{itemize}
 \end{block}

### 2016-11-08 ARTICLES/OTHER/C2S_EXA/slides.tex
final

@@
 
+\AtBeginSection[]
+{
+\begin{frame}{Table of Contents}
+\tableofcontents[currentsection]
+\end{frame}
+}
 %-------------------------------------------------------------------
@@
     \resizebox{1cm}{!}{\includegraphics{images/elci_logo.png}}
+\end{frame}
+
+\begin{frame}{Table of Contents}
+  \tableofcontents
 \end{frame}
@@
     \begin{itemize}
-    \item Component = black box with well defined interactions (provided/required)
+    \item Technology advocating composition rather than development
+      \begin{itemize}
+      \item Old idea (late 60') but major development in the 90'
+      \item After Object and before Model Driven Engineering (MDE)
+      \end{itemize}
+    \item Component = black box with well defined interactions
+      \begin{itemize}
+      \item Provides / Requires
+      \end{itemize}
     \item Application = Assembly of components
@@
     \begin{itemize}
-    \item Common model to define software architecture
+      \item Application structure well defined
+    \item Code-reuse and productivity
     \item Maintainability through separation of concerns
-    \item Code-reuse and productivity
-    \end{itemize}
-  \end{block}
-  \begin{alertblock}{Open issues}
-    \begin{itemize}
-    \item Component for HPC, eg task graph?\\ {\color{red} Cf next talk of Jérôme Richard!}
+    \end{itemize}
+  \end{block}
+\vspace{-4cm}\hspace{8.15cm}\resizebox{2.5cm}{!}{\includegraphics{images/assembly.png}}
+\end{frame}
+%-------------------------------------------------------------
+\begin{frame}
+\frametitle{L2C: A HPC Component Model}
+\begin{block}{A minimalist component model for HPC}
+  \begin{itemize}
+  \item Component creation/deletion, configuration, and connection
+  \item No L2C code between components \@ runtime
+    \begin{itemize}
+      \item No language interoperability provided
+    \end{itemize}
+  \item  Support of native interactions
+    \begin{itemize}
+    \item Use/Provide: C++, CORBA, [ FORTRAN (2008)]
+    \item MPI communicator
+    \end{itemize}
+  \item \url{hlcm.gforge.inria.fr}
+  \end{itemize}
+\end{block}
+\begin{alertblock}{Ongoing extensions}
+  \begin{itemize}
+  \item Concurrent reconfiguration
+    \begin{itemize}
+    \item DirectL2C
+    \end{itemize}
+  \item Task support
+    \begin{itemize}
+    \item {\color{red} Cf next talk on Comet!}
+    \end{itemize}
   \end{itemize}
 \end{alertblock}
+\vspace{-4.2cm}\hspace{6.8cm}\resizebox{4cm}{!}{\includegraphics{images/jacobi-2X2.png}}
 \end{frame}
@@
 \begin{itemize}
-\item A cartesian mesh is used
+\item A Cartesian mesh is used
 \item Two kinds of mesh entities are declared onto it
@@
 \begin{frame}
+\frametitle{Series-Parallel Tree}
+\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
+\normalsize
+\begin{columns}
+\begin{column}{.48\textwidth}
+\begin{center}
+\resizebox{\textwidth}{!}{%
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (s0)    at (0,0) {$\mathcal{S}$};
+
+   \node[] (c0)    at (-3,1) {$c_0$};
+   \node[] (star1) at (-2,1) {$sync_1$};
+   \node[] (c1)    at (-1,1) {$c_1$};
+   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
+   \node[] (c7)    at ( 1,1) {$c_7$};
+   \node[] (star8) at ( 2,1) {$sync_8$};
+   \node[] (c8)    at ( 3,1) {$c_8$};
+
+   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
+   \node[] (s2)    at (1,2) {$\mathcal{S}$};
+
+   \node[] (c2)    at (-3,3) {$c_2$};
+   \node[] (star4) at (-2,3) {$sync_4$};
+   \node[] (c4)    at (-1,3) {$c_4$};
+   \node[] (c6)    at ( 0,3) {$c_6$};
+   \node[] (c3)    at ( 1,3) {$c_3$};
+   \node[] (c5)    at ( 2,3) {$c_5$};
+
+  \path[->]
+    (s0) edge node {} (c0)
+         edge node {} (star1)
+         edge node {} (c1)
+         edge node {} (p0)
+         edge node {} (c7)
+         edge node {} (star8)
+         edge node {} (c8)
+    (p0) edge node {} (s1)
+         edge node {} (s2)
+    (s1) edge node {} (c2)
+         edge node {} (star4)
+         edge node {} (c4)
+         edge node {} (c4)
+         edge node {} (c6)
+    (s2) edge node {} (c3)
+         edge node {} (c5);
+  \end{tikzpicture}
+  }
+\end{center}
+\end{column}
+\begin{column}{.52\textwidth}
+
+\end{column}
+\end{columns}
+\end{frame}
+%----------------
+
+%----------------
+\begin{frame}
+\frametitle{Series-Parallel Tree}
+\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
+\normalsize
+\begin{columns}
+\begin{column}{.48\textwidth}
+\begin{center}
+\resizebox{\textwidth}{!}{%
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (s0)    at (0,0) {$\mathcal{S}$};
+
+   \node[] (c0)    at (-3,1) {$c_0$};
+   \node[] (star1) at (-2,1) {$sync_1$};
+   \node[] (c1)    at (-1,1) {$c_1$};
+   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
+   \node[] (c7)    at ( 1,1) {$c_7$};
+   \node[] (star8) at ( 2,1) {$sync_8$};
+   \node[] (c8)    at ( 3,1) {$c_8$};
+
+   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
+   \node[] (s2)    at (1,2) {$\mathcal{S}$};
+
+   \node[] (c2)    at (-3,3) {$c_2$};
+   \node[] (star4) at (-2,3) {$sync_4$};
+   \node[] (c4)    at (-1,3) {$c_4$};
+   \node[] (c6)    at ( 0,3) {$c_6$};
+   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
+   %\node[] (c5)    at ( 2,3) {$c_5$};
+
+  \path[->]
+    (s0) edge node {} (c0)
+         edge node {} (star1)
+         edge node {} (c1)
+         edge node {} (p0)
+         edge node {} (c7)
+         edge node {} (star8)
+         edge node {} (c8)
+    (p0) edge node {} (s1)
+         edge node {} (s2)
+    (s1) edge node {} (c2)
+         edge node {} (star4)
+         edge node {} (c4)
+         edge node {} (c4)
+         edge node {} (c6)
+    (s2) edge node {} (c3);
+  \end{tikzpicture}
+  }
+\end{center}
+\end{column}
+\begin{column}{.52\textwidth}
+\end{column}
+\end{columns}
+Loop fusion optimization possible
+\end{frame}
+%----------------
+
+%----------------
+\begin{frame}
+\frametitle{Series-Parallel Tree}
+\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
+\normalsize
+\begin{columns}
+\begin{column}{.48\textwidth}
+\begin{center}
+\resizebox{\textwidth}{!}{%
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (s0)    at (0,0) {$\mathcal{S}$};
+
+   \node[] (c0)    at (-3,1) {$c_0$};
+   \node[] (star1) at (-2,1) {$sync_1$};
+   \node[] (c1)    at (-1,1) {$c_1$};
+   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
+   \node[] (c7)    at ( 1,1) {$c_7$};
+   \node[] (star8) at ( 2,1) {$sync_8$};
+   \node[] (c8)    at ( 3,1) {$c_8$};
+
+   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
+   \node[] (s2)    at (1,2) {$\mathcal{S}$};
+
+   \node[] (c2)    at (-3,3) {$c_2$};
+   \node[] (star4) at (-2,3) {$sync_4$};
+   \node[] (c4)    at (-1,3) {$c_4$};
+   \node[] (c6)    at ( 0,3) {$c_6$};
+   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
+   %\node[] (c5)    at ( 2,3) {$c_5$};
+
+  \path[->]
+    (s0) edge node {} (c0)
+         edge node {} (star1)
+         edge node {} (c1)
+         edge node {} (p0)
+         edge node {} (c7)
+         edge node {} (star8)
+         edge node {} (c8)
+    (p0) edge node {} (s1)
+         edge node {} (s2)
+    (s1) edge node {} (c2)
+         edge node {} (star4)
+         edge node {} (c4)
+         edge node {} (c4)
+         edge node {} (c6)
+    (s2) edge node {} (c3);
+  \end{tikzpicture}
+  }
+\end{center}
+\end{column}
+\begin{column}{.52\textwidth}
+\begin{block}{Specific components}
+\begin{itemize}
+\item $SEQ$ to directly replace $\mathcal{S}$ nodes
+\item $PAR$ to directly replace $\mathcal{P}$ nodes
+\item $SYNC$ for synchronizations
+\item $K$ for computation kernels
+\end{itemize}
+\end{block}
+\end{column}
+\end{columns}
+Loop fusion optimization possible
+\end{frame}
+%----------------
+
+%----------------
+\begin{frame}
 \frametitle{MSL to Component-based runtime}
@@
   \begin{itemize}
-  \item OpenMP task (For loop also available)
+  \item OpenMP task
   \end{itemize}
@@
 \end{frame}
-%% %-------------------------------------------------------------
-%% \begin{frame}
-%% \frametitle{What is produced ?}
-%% \begin{block}{Implementation}
-%% \begin{itemize}
-%% \item DDS + Quantity + $k_{*}^{sync}$ = written components using SkelGIS (distributed memory, MPI)
-%% \item Scheduler = Serie-Parallel graph of $\Gamma_{dep}$ (OpenMP 3.0)
-%% \item Time + Scheduler = generated components
-%% \item $k_i$ = empty components to write
-%% \end{itemize}
-%% \end{block}
-
-%% \begin{block}{Separation of concerns}
-%% \begin{itemize}
-%% \item numerician (MSL)
-%% \item computer engineer (K)
-%% \item HPC engineer (DDS, quantity, $k_{*}^{sync}$ etc.)
-%% \item DSL designer (component assembly + generated components)
-%% \end{itemize}
-%% \end{block}
-%% \end{frame}
-%% %-------------------------------------------------------------
-
-%% \begin{frame}
-%% \frametitle{Series-Parallel Tree}
-%% \small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
-%% \normalsize
-%% \begin{columns}
-%% \begin{column}{.48\textwidth}
-%% \begin{center}
-%% \resizebox{\textwidth}{!}{%
-%% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%%    \node[] (s0)    at (0,0) {$\mathcal{S}$};
-
-%%    \node[] (c0)    at (-3,1) {$c_0$};
-%%    \node[] (star1) at (-2,1) {$sync_1$};
-%%    \node[] (c1)    at (-1,1) {$c_1$};
-%%    \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
-%%    \node[] (c7)    at ( 1,1) {$c_7$};
-%%    \node[] (star8) at ( 2,1) {$sync_8$};
-%%    \node[] (c8)    at ( 3,1) {$c_8$};
-
-%%    \node[] (s1)    at (-1,2) {$\mathcal{S}$};
-%%    \node[] (s2)    at (1,2) {$\mathcal{S}$};
-
-%%    \node[] (c2)    at (-3,3) {$c_2$};
-%%    \node[] (star4) at (-2,3) {$sync_4$};
-%%    \node[] (c4)    at (-1,3) {$c_4$};
-%%    \node[] (c6)    at ( 0,3) {$c_6$};
-%%    \node[] (c3)    at ( 1,3) {$c_3$};
-%%    \node[] (c5)    at ( 2,3) {$c_5$};
-
-%%   \path[->]
-%%     (s0) edge node {} (c0)
-%%          edge node {} (star1)
-%%          edge node {} (c1)
-%%          edge node {} (p0)
-%%          edge node {} (c7)
-%%          edge node {} (star8)
-%%          edge node {} (c8)
-%%     (p0) edge node {} (s1)
-%%          edge node {} (s2)
-%%     (s1) edge node {} (c2)
-%%          edge node {} (star4)
-%%          edge node {} (c4)
-%%          edge node {} (c4)
-%%          edge node {} (c6)
-%%     (s2) edge node {} (c3)
-%%          edge node {} (c5);
-%%   \end{tikzpicture}
-%%   }
-%% \end{center}
-%% \end{column}
-%% \begin{column}{.52\textwidth}
-
-%% \end{column}
-%% \end{columns}
-%% \end{frame}
-%% %----------------
-
-%% %----------------
-%% \begin{frame}
-%% \frametitle{Series-Parallel Tree}
-%% \small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
-%% \normalsize
-%% \begin{columns}
-%% \begin{column}{.48\textwidth}
-%% \begin{center}
-%% \resizebox{\textwidth}{!}{%
-%% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%%    \node[] (s0)    at (0,0) {$\mathcal{S}$};
-
-%%    \node[] (c0)    at (-3,1) {$c_0$};
-%%    \node[] (star1) at (-2,1) {$sync_1$};
-%%    \node[] (c1)    at (-1,1) {$c_1$};
-%%    \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
-%%    \node[] (c7)    at ( 1,1) {$c_7$};
-%%    \node[] (star8) at ( 2,1) {$sync_8$};
-%%    \node[] (c8)    at ( 3,1) {$c_8$};
-
-%%    \node[] (s1)    at (-1,2) {$\mathcal{S}$};
-%%    \node[] (s2)    at (1,2) {$\mathcal{S}$};
-
-%%    \node[] (c2)    at (-3,3) {$c_2$};
-%%    \node[] (star4) at (-2,3) {$sync_4$};
-%%    \node[] (c4)    at (-1,3) {$c_4$};
-%%    \node[] (c6)    at ( 0,3) {$c_6$};
-%%    \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
-%%    %\node[] (c5)    at ( 2,3) {$c_5$};
-
-%%   \path[->]
-%%     (s0) edge node {} (c0)
-%%          edge node {} (star1)
-%%          edge node {} (c1)
-%%          edge node {} (p0)
-%%          edge node {} (c7)
-%%          edge node {} (star8)
-%%          edge node {} (c8)
-%%     (p0) edge node {} (s1)
-%%          edge node {} (s2)
-%%     (s1) edge node {} (c2)
-%%          edge node {} (star4)
-%%          edge node {} (c4)
-%%          edge node {} (c4)
-%%          edge node {} (c6)
-%%     (s2) edge node {} (c3);
-%%   \end{tikzpicture}
-%%   }
-%% \end{center}
-%% \end{column}
-%% \begin{column}{.52\textwidth}
-%% \end{column}
-%% \end{columns}
-%% Loop fusion optimization possible
-%% \end{frame}
-%% %----------------
-
-%% %----------------
-%% \begin{frame}
-%% \frametitle{Series-Parallel Tree}
-%% \small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
-%% \normalsize
-%% \begin{columns}
-%% \begin{column}{.48\textwidth}
-%% \begin{center}
-%% \resizebox{\textwidth}{!}{%
-%% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%%    \node[] (s0)    at (0,0) {$\mathcal{S}$};
-
-%%    \node[] (c0)    at (-3,1) {$c_0$};
-%%    \node[] (star1) at (-2,1) {$sync_1$};
-%%    \node[] (c1)    at (-1,1) {$c_1$};
-%%    \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
-%%    \node[] (c7)    at ( 1,1) {$c_7$};
-%%    \node[] (star8) at ( 2,1) {$sync_8$};
-%%    \node[] (c8)    at ( 3,1) {$c_8$};
-
-%%    \node[] (s1)    at (-1,2) {$\mathcal{S}$};
-%%    \node[] (s2)    at (1,2) {$\mathcal{S}$};
-
-%%    \node[] (c2)    at (-3,3) {$c_2$};
-%%    \node[] (star4) at (-2,3) {$sync_4$};
-%%    \node[] (c4)    at (-1,3) {$c_4$};
-%%    \node[] (c6)    at ( 0,3) {$c_6$};
-%%    \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
-%%    %\node[] (c5)    at ( 2,3) {$c_5$};
-
-%%   \path[->]
-%%     (s0) edge node {} (c0)
-%%          edge node {} (star1)
-%%          edge node {} (c1)
-%%          edge node {} (p0)
-%%          edge node {} (c7)
-%%          edge node {} (star8)
-%%          edge node {} (c8)
-%%     (p0) edge node {} (s1)
-%%          edge node {} (s2)
-%%     (s1) edge node {} (c2)
-%%          edge node {} (star4)
-%%          edge node {} (c4)
-%%          edge node {} (c4)
-%%          edge node {} (c6)
-%%     (s2) edge node {} (c3);
-%%   \end{tikzpicture}
-%%   }
-%% \end{center}
-%% \end{column}
-%% \begin{column}{.52\textwidth}
-%% \begin{block}{Specific components}
-%% \begin{itemize}
-%% \item $SEQ$ to directly replace $\mathcal{S}$ nodes
-%% \item $PAR$ to directly replace $\mathcal{P}$ nodes
-%% \item $SYNC$ for synchronizations
-%% \item $K$ for computation kernels
-%% \end{itemize}
-%% \end{block}
-%% \end{column}
-%% \end{columns}
-%% Loop fusion optimization possible
-%% \end{frame}
-%% %----------------
-
-%% %-------------------------------------------------------------
-%% \section{Implementation}
-%% %-------------------------------------------------------------
-%% \begin{frame}
-%% \frametitle{Component models}
-%% \begin{itemize}
-%% \item Black box with code
-%% \item Use/provide interfaces
-%% \item Improve code reuse (productivity)
-%% \item Improve separation of concerns (maintainability and portability)
-%% \item Improve composability of applications
-%% \end{itemize}
-%% \begin{center}
-%% \begin{tikzpicture}[remember picture]
-%%   \node[component,solid] (A) at (0,0) {A};
-%%   \node[component,solid,right=5cm of A] (B) {B};
-%%   \node[use,scale=0.5,solid,right=1.5cm of A] (Au) {};
-%%   \node[provide,scale=0.5,solid,left=1.5cm of B] (Bp) {};
-%%   \path[-]
-%%     (A) edge node {} (Au)
-%%     (Bp) edge node {} (B);
-%% \end{tikzpicture}
-
-%% \bigskip
-%% \begin{tikzpicture}[remember picture]
-%%   \node[component,solid] (A) at (0,0) {A};
-%%   \node[use,scale=0.5,solid,right=1cm of A] (Au) {};
-%%   \node[provide,scale=0.5,solid,right=0.01cm of Au] (Bp) {};
-%%   \node[component,solid,right=1cm of Bp] (B) {B};
-%%   \path[-]
-%%     (A) edge node {} (Au)
-%%     (Bp) edge node {} (B);
-%% \end{tikzpicture}
-
-%% \medskip
-%% Components A and B into a component assembly
-%% \end{center}
-%% \end{frame}
+%----------------
+\begin{frame}
+\frametitle{Component-based runtime: Static Scheduling}
+
+\begin{center}
+\resizebox{\textwidth}{!}{%
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+     \node[component,solid] (D) at (0,0) {$Driver$};
+     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
+     %\node[solid] (Ds) at (-1.5,0) {start};
+     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
+     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
+     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};
+
+     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
+     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
+     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
+     %\node[solid,left=0.8cm of T] (tt) {$T$};
+
+     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
+     \node[component,dashed,right=2cm of Cp] (C) {$Computations$};
+     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
+     \node[solid,right=0.2cm of Cu] (star) {$*$};
+     %\node[solid,right=1.5cm of C] (gamma) {$\Gamma$};
+
+     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
+     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
+     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
+     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
+     %\node[solid,left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
+     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
+     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};
+
+     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
+     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
+     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
+     %\node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
+
+    \path[-]
+      (Dp) edge [solid] node {} (D)
+      (D) edge [solid] node {} (Du1)
+          edge [solid] node {} (Du2)
+          edge [solid] node {} (Du3)
+      (DDSp1) edge [solid] node {} (DDS)
+      (Tp) edge [solid] node {} (T)
+      (T)  edge [solid] node {} (Tu)
+      (Cp) edge [solid] node {} (C)
+      (C) edge [solid] node {} (Cu)
+      (Datap2) edge [solid] node {} (Data)
+      (Datap1) edge [solid] node {} (Data)
+      (Data) edge [solid] node {} (Datau)
+      (DDSp2) edge [solid] node {} (DDS);
+    %\end{tikzpicture}
+    %}
+% \end{column}
+% \begin{column}{.70\textwidth}
+% \begin{center}
+% \resizebox{.80\textwidth}{!}{%
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   %seq0
+   \node[component,right=3.5cm of C] (seq0) {$SEQ$};
+   \node[provide] (seq0p) [left = 0.8cm of seq0] {};
+   \node[use] (seq0u) [right = 1cm of seq0] {$m$};
+   %c0
+   \node[component] (c0) [right = 2cm of seq0u] {$K(c_0)$};
+   \node[use] (c0u) [right = 0.8cm of c0] {};
+   \node[right=0.2 of c0u] (star) {*};
+   %sync0
+   \node[component] (sync0) [below = 1cm of c0] {$SYNC$};
+   \node[use] (sync0u) [right = 1cm of sync0] {$m$};
+   %c1
+   \node[component] (c1) [below = 0.8cm of sync0] {$K(c_1)$};
+   \node[use] (c1u) [right = 0.8cm of c1] {};
+   \node[right=0.2 of c1u] (star) {*};
+   %par0
+   \node[component] (par0) [below = 0.8cm of c1] {$PAR$};
+   \node[use] (par0u) [right = 1cm of par0] {$m$};
+  %seq1
+   \node[component] (seq1) [right = 1cm of sync0u] {$SEQ$};
+   \node[use] (seq1u) [right = 1cm of seq1] {$m$};
+  %c2
+   \node[component] (c2) [right = 4cm of c0] {$K(c_2)$};
+   \node[use] (c2u) [right = 0.8cm of c2] {};
+   \node[right=0.2 of c2u] (star) {*};
+  %sync1
+   \node[component] (sync1) [below = 0.8cm of c2] {$SYNC$};
+   \node[use] (sync1u) [right = 1cm of sync1] {$m$};
+  %c4
+   \node[component] (c4) [below = 0.8cm of sync1] {$K(c_4)$};
+   \node[use] (c4u) [right = 0.8cm of c4] {};
+   \node[right=0.2 of c4u] (star) {*};
+  %c6
+   \node[component] (c6) [below = 0.8cm of c4] {$K(c_6)$};
+   \node[use] (c6u) [right = 0.8cm of c6] {};
+   \node[right=0.2 of c6u] (star) {*};
+  %seq2
+   \node[component] (seq2) [below = 3.2cm of seq1] {$SEQ$};
+   \node[use] (seq2u) [right = 1cm of seq2] {$m$};
+  %c3
+   \node[component] (c3) [right = 1cm of seq2u] {$K(c_3)$};
+   \node[use] (c3u) [right = 0.8cm of c3] {};
+   \node[right=0.2 of c3u] (star) {*};
+  %c5
+   \node[component] (c5) [below = 0.8cm of c3] {$K(c_5)$};
+   \node[use] (c5u) [right = 0.8cm of c5] {};
+   \node[right=0.2 of c5u] (star) {*};
+  %c7
+   \node[component] (c7) [below = 0.8cm of par0] {$K(c_7)$};
+   \node[use] (c7u) [right = 0.8cm of c7] {};
+   \node[right=0.2 of c7u] (star) {*};
+  %sync2
+   \node[component] (sync2) [below = 0.8cm of c7] {$SYNC$};
+   \node[use] (sync2u) [right = 1cm of sync2] {$m$};
+  %c8
+   \node[component] (c8) [below = 0.8cm of sync2] {$K(c_8)$};
+   \node[use] (c8u) [right = 0.8cm of c8] {};
+   \node[right=0.2 of c8u] (star) {*};
+
+   \path[-]
+   %seq0
+    (seq0) edge node {} (seq0u)
+    (seq0p) edge node {} (seq0)
+   %c0
+    (seq0u) edge node {} (c0)
+    (c0) edge node {} (c0u)
+    %sync0
+    (seq0u) edge node {} (sync0.west)
+    (sync0) edge node {} (sync0u)
+    %c1
+    (seq0u) edge node {} (c1)
+    (c1) edge node {} (c1u)
+   %par0
+    (seq0u) edge node {} (par0.west)
+    (par0) edge node {} (par0u)
+  %seq1
+    (par0u) edge node {} (seq1.west)
+    (seq1) edge node {} (seq1u)
+  %c2
+    (seq1u) edge node {} (c2.west)
+    (c2) edge node {} (c2u)
+  %sync0
+    (seq1u) edge node {} (sync1.west)
+    (sync1) edge node {} (sync1u)
+  %c4
+    (seq1u) edge node {} (c4.west)
+    (c4) edge node {} (c4u)
+  %c6
+    (seq1u) edge node {} (c6.west)
+    (c6) edge node {} (c6u)
+  %seq2
+    (par0u) edge node {} (seq2.west)
+    (seq2) edge node {} (seq2u)
+  %c3
+    (seq2u) edge node {} (c3)
+    (c3) edge node {} (c3u)
+  %c6
+    (seq2u) edge node {} (c5.west)
+    (c5) edge node {} (c5u)
+  %c6
+    (seq0u) edge node {} (c7.west)
+    (c7) edge node {} (c7u)
+  %sync2
+    (seq0u) edge node {} (sync2.west)
+    (sync2) edge node {} (sync2u)
+  %c8
+    (seq0u) edge [bend right] node {} (c8.west)
+    (c8) edge node {} (c8u)
+   ;
+
+   \path[-]
+    (C) edge[dashed] node {} (seq0);
+\end{tikzpicture}
+}
+\end{center}
+% \end{column}
+% \end{columns}
+\end{frame}
+%----------------
+\begin{frame}
+  \frametitle{Component-based runtime: Dynamic Scheduling}
+\begin{center}
+  \begin{itemize}
+  \item \emph{mpiOmpDyn}
+    \begin{itemize}
+    \item \emph{DDS} and \emph{Quantity}: data parallelization.
+    \item \emph{Scheduler}: DAG encoded into OpenMP tasks
+    \end{itemize}
+  \end{itemize}
+  \medskip
+  \begin{tikzpicture}[remember picture]
+  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
+    \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+     \node[component,solid] (D) at (0,0) {$Driver$};
+     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
+     \node[solid] (Ds) at (-1.5,0) {start};
+     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
+     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
+     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};
+
+     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
+     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
+     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
+
+     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Sp) {};
+     \node[component,solid,right=1.5cm of Sp] (S) {$Scheduling$};
+     \node[use,scale=0.5,solid,right=1.5cm of S] (Su) {};
+
+     \node[provide,scale=0.5,solid,right=0.15 of Su] (Cp) {};
+     \node[component,solid,double,right=2cm of Cp] (C) {$Kernel$};
+     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
+
+     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
+     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
+     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
+     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
+
+     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
+     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
+     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
+
+    \path[-]
+      (Dp) edge [solid] node {} (D)
+      (D) edge [solid] node {} (Du1)
+          edge [solid] node {} (Du2)
+          edge [solid] node {} (Du3)
+      (DDSp1) edge [solid] node {} (DDS)
+      (Tp) edge [solid] node {} (T)
+      (T)  edge [solid] node {} (Tu)
+      (Sp) edge [solid] node {} (S)
+      (S) edge [solid] node {} (Su)
+      (Datap2) edge [solid] node {} (Data)
+      (Datap1) edge [solid] node {} (Data)
+      (Data) edge [solid] node {} (Datau)
+      (C) edge [solid] node {} (Cu)
+      	  edge [solid] node {} (Cp)
+      (DDSp2) edge [solid] node {} (DDS);
+    \end{tikzpicture}
+    };
+    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=0.5cm of proc] (mpidata) {};
+    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
+    \node[solid,below=2.5cm of proc.west,anchor=west] (core) {\textit{\tiny{Duplicated on each processor/core}}};
+    \path[-]
+      (Data.east) edge node {} (mpidata);
+  \end{tikzpicture}
+\end{center}
+
+\end{frame}
 
@@
 \begin{frame}
-\frametitle{Evaluations}
+\frametitle{Evaluation Setup}
 \begin{block}{Application}
 \begin{itemize}
-\item FullSWOF2D: developed at the MAPMO, Université d'Orléans
+\item FullSWOF2D: developed at the MAPMO, University of Orléans
 \item Solve the shallow-water equations using a finite volume method
@@
   Strong Scaling\\
-  10k x 10k domaine
+  10k x 10k domain
 \end{center}
@@
 \begin{frame}
-\frametitle{Data Paralization to Hybrid Parallization: Strong Scaling}
+\frametitle{Data Parallelization to Hybrid Parallelization: Strong Scaling}
 \begin{columns}
@@
   \begin{center}
-\vspace{-1em}\small    Data Parallelization    
+{\vspace{-1em}\small    Data Parallelization    }
   \end{center}
@@
 \begin{center}
-\vspace{-1em}\small  Data and Hybrid Parallelizations
+{\vspace{-1em}\small  Data and Hybrid Parallelizations}
 \end{center}
@@
 \end{columns}
-\bigskip
+\medskip
 \begin{center}
@@
   \\
+  \medskip
  Task graph analysis
@@
 \end{center}
+Mesh: 500x500
+\end{frame}
+%-------------------------------------------------------------
+
+\begin{frame}
+  \frametitle{Dynamic Scheduling and Energy Evaluations}
+
+  Application versions:
+  \small
+  \centering
+\begin{tabular}{|c||c|c|c|}
+  \hline
+  Variant Name  & DDS / Quantity & Kernels & Scheduler \\
+  \hline
+  \hline
+  mpiBase        & MPI & Sequential & Sequential \\
+  \hline
+  mpiOmpFor      & MPI & Parallel loops & Sequential \\
+  \hline
+  mpiOmpForkJoin & MPI & Sequential & SPT \\
+  \hline
+  mpiOmpDyn      & MPI & Sequential & OpenMP Tasks \\
+  \hline
+\end{tabular}
+
+\end{frame}
+
+%-------------------------------------------------------------
+
+  \begin{frame}
+\frametitle{Energy Usage on Grid'5000}
+\begin{columns}
+\begin{column}{6cm}
+\resizebox{6cm}{!}{\includegraphics{images/4000_100_4node_med_time_maxWatt_cloud.pdf}}
+    \begin{center}
+{\vspace{-1em}\small Time vs Max Watt on 4 nodes}
+  \end{center}
+\end{column}
+\begin{column}{6cm}
+  \resizebox{6cm}{!}{\includegraphics{images/4000_100_4node_med_time_avrgWatt.pdf}}
+\begin{center}
+{\vspace{-1em}\small Time vs Average Watt on 4 nodes}
+\end{center}
+\end{column}
+\end{columns}
+\bigskip
+{\small
+\begin{block}{Parameters:}
+\begin{itemize}
+\vspace{-0.7em}\item Domain: 4000 x 4000
+\item \#Iteration: 100
+\end{itemize}
+\end{block}
+}
+  \end{frame}
+%-------------------------------------------------------------
+  \begin{frame}
+
+\frametitle{Energy Usage on Curie}
+\begin{columns}
+\begin{column}{6cm}
+\resizebox{6cm}{!}{\includegraphics{images/Curie_20k_10k_time_avrgWatt_2048.pdf}}
+    \begin{center}
+{\vspace{-1em}\small Time vs Max Watt on 2048 cores}
+  \end{center}
+\end{column}
+\begin{column}{6cm}
+  \resizebox{6cm}{!}{\includegraphics{images/Curie_84k_500_time_avrgWatt.pdf}}
+\begin{center}
+{\vspace{-1em}\small Time vs Average Watt on 2048 cores}
+\end{center}
+\end{column}
+\end{columns}
+\bigskip
+{\small
+\begin{block}{Parameters:}
+\begin{itemize}
+\vspace{-0.7em}\item Domain: 20,000 x 20,000
+\item \#Iteration: 10,000
+\item One measure every 5 min!
+\end{itemize}
+\end{block}
+}
+
 \end{frame}
@@
 \begin{itemize}
-\item A DSL for Multi-Stencil applications (MSL)
+\vspace{-0.4em}\item A DSL for Multi-Stencil applications (MSL)
 \item The generation of a component based runtime, including scheduling
 \begin{itemize}
-\item Data parallelism
-\item Task parallelism
-\end{itemize}
-\item Evaluation on Curie: no overhead introduced 
-\end{itemize}
-\end{block}
-
-\begin{block}{Future Work}
-  \begin{itemize}
-  \item Language improvement (convergence criteria, reduction etc.)
-  \item OpenMP 3 inside kernels
+\item Data and task parallelism
+\end{itemize}
+\item Evaluation on Grid'5000 and Curie: no overhead introduced
+\item Basis for ongoing energy studies
+\end{itemize}
+\end{block}
+
+\begin{block}{Perspectives}
+  \begin{itemize}
+\vspace{-0.4em}  \item Language improvement (convergence criteria, reduction etc.)
+  \item OpenMP inside kernels/components
   \item CPU+GPGPUs using stencil compilers (Pochoir, PATUS etc.)
-$\hookrightarrow$ Show portability, maintainability introduced by components
-
-  \end{itemize}
-\end{block}
-\end{frame}
-%-------------------------------------------------------------
-
-%% %-------------------------------------------------------------
-%% \begin{frame}
-%% \frametitle{Conclusion}
-%% \begin{block}{Work in progress}
-%% \begin{itemize}
-%% \item Different types of scheduling and back-end in MSL
-%% \item Combine MSL implementation choices to an energy-aware framework
-%% \item New programming model back-end for MSL: combining component models and task models (Ph.D. Jérôme Richard)
-%% \end{itemize}
-%% \end{block}
-%% \begin{block}{Collaborations}
-%% \begin{itemize}
-%% \item More application cases
-%% \item Back-end programming model: component models and task scheduling (Components in OmpSs ?)
-%% \item DSLs interoperability
-%% \end{itemize}
-%% \end{block}
-%% \end{frame}
-%-------------------------------------------------------------
-
-% \begin{withoutheadline}
-% \begin{frame}{}
-% \begin{center}
-% \huge Thank You !
-% \end{center}
-% \end{frame}
-% \end{withoutheadline}
-
-%----------------
-%% \begin{frame}
-%% \frametitle{MSL to Component-based runtime}
-%% \begin{center}
-%% \begin{tikzpicture}[remember picture]
-%%   \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
-%%     \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-%%      \node[component,solid] (D) at (0,0) {$Driver$};
-%%      \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
-%%      \node[solid] (Ds) at (-1.5,0) {start};
-%%      \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
-%%      \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
-%%      \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};
-
-%%      \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
-%%      \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
-%%      \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
-
-%%      \node[provide,scale=0.5,solid,right=0.15 of Tu] (Sp) {};
-%%      \node[component,solid,right=1.5cm of Sp] (S) {$Scheduling$};
-%%      \node[use,scale=0.5,solid,right=1.5cm of S] (Su) {};
-
-%%      \node[provide,scale=0.5,solid,right=0.15 of Su] (Cp) {};
-%%      \node[component,solid,double,right=2cm of Cp] (C) {$Kernel$};
-%%      \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
-
-%%      \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
-%%      \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
-%%      \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
-%%      \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
-
-%%      \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
-%%      \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
-%%      \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
-
-%%     \path[-]
-%%       (Dp) edge [solid] node {} (D)
-%%       (D) edge [solid] node {} (Du1)
-%%           edge [solid] node {} (Du2)
-%%           edge [solid] node {} (Du3)
-%%       (DDSp1) edge [solid] node {} (DDS)
-%%       (Tp) edge [solid] node {} (T)
-%%       (T)  edge [solid] node {} (Tu)
-%%       (Sp) edge [solid] node {} (S)
-%%       (S) edge [solid] node {} (Su)
-%%       (Datap2) edge [solid] node {} (Data)
-%%       (Datap1) edge [solid] node {} (Data)
-%%       (Data) edge [solid] node {} (Datau)
-%%       (C) edge [solid] node {} (Cu)
-%%       	  edge [solid] node {} (Cp)
-%%       (DDSp2) edge [solid] node {} (DDS);
-%%     \end{tikzpicture}
-%%     };
-%%     \node[mpi,scale=0.8,rounded corners=0pt,solid,right=0.5cm of proc] (mpidata) {};
-%%     \node[solid,above=0.1cm of mpidata] (star2) {$*$};
-%%     \node[solid,below=2.5cm of proc.west,anchor=west] (core) {\textit{\tiny{Duplicated on each processor/core}}};
-%%     \path[-]
-%%       (Data.east) edge node {} (mpidata);
-%%   \end{tikzpicture}
-%% \end{center}
-%% \end{frame}
+  \item Composition of various DSLs by agreeing on component runtimes?
+  \end{itemize}
+\end{block}
+\end{frame}
+%-------------------------------------------------------------
 

### 2017-02-15 ARTICLES/IJPP/articleR1/article.tex
version révisée IJPP

--- 2016-05-31 ARTICLES/IJPP/article/article.tex
@@
 \newtheorem{myth}[theorem]{Theorem}
-% \newenvironment{proof}[1][Proof]{\begin{trivlist}
-% \item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-%\newenvironment{mydef}[1][Definition]{\begin{trivlist}
-%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-%\newenvironment{myth}[1][Theorem]{\begin{trivlist}
-%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
 \newenvironment{mydefs}[1][Definitions]{\par\medskip\noindent\textbf{#1}\nopagebreak\begin{itemize}[noitemsep,topsep=.25\topsep]}{\end{itemize}}
@@
   \scalebox{.4}[1]{$\prec$}\mkern-5.5mu\scalebox{.4}[1]{$\prec$}}}
-%
-% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
-%
-% insert here the call for the packages your document requires
-%\usepackage{latexsym}
-% etc.
-%
-% please place your own definitions here and don't use \def but
-% \newcommand{}{}
-%
-% Insert the name of "your journal" with
-% \journalname{myjournal}
-%
 \begin{document}
 
-\title{The Multi-Stencil Language: orchestrating stencils with a mesh-agnostic DSL}
-%\subtitle{Do you have a subtitle?\\ If so, write it here}
-
-%\titlerunning{Short form of title}        % if too long for running head
+\title{Orchestrating stencils with the Multi-Stencil Framework}
 
 \author{H\'el\`ene Coullon \and Julien Bigot \and Christian Perez}
-
-%\authorrunning{Short form of author list} % if too long for running head
 
@@
 }
-
-%\date{Received: date / Accepted: date}
-% The correct dates will be entered by the editor
 
@@
 %------------------------------------------------------------------------------
-\section{Related work}
-\label{sect:rel}
-\input{related.tex}
-%------------------------------------------------------------------------------
-\section{Computational model of Multi-Stencil Programs}
+\section{Meta-modelization of a Multi-Stencil Program}
 \label{sect:formalism}
 \input{formalism.tex}
+%------------------------------------------------------------------------------
+\section{The Component-Based Multi-Stencil Framework}
+\label{sect:msf}
+\input{msf.tex}
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-\section{Parallelization of Multi-Stencil Programs}
+\section{The Multi-Stencil Compiler}
 \label{sect:parallelism}
@@
 \input{evaluation.tex}
+%------------------------------------------------------------------------------
+\section{Related work}
+\label{sect:rel}
+\input{related.tex}
 %------------------------------------------------------------------------------

### 2017-02-15 ARTICLES/IJPP/articleR1/compiler.tex
version révisée IJPP

--- 2016-05-30 ARTICLES/IJPP/article/compiler.tex
@@
 For the task parallelism, we use OpenMP~\cite{660313}.
-OpenMP targets shared-memory platforms only, but the degree of task-parallelism is typically much lower than that of data parallelism.
-With the increasing number of cores of current machines this has not proved to be a problem in our experience. % LN : ?
+OpenMP targets shared-memory platforms only.
+However, as the level of parallelism introduced by task parallelism technique is low compared to data parallelism, shared-memory, or intra-node parallelism is a good architecture choice for task parallelism.% but the degree of task-parallelism is typically much lower than that of data parallelism.
+%With the increasing number of cores of current machines this has not proved to be a problem in our experience. % LN : ?
 While the version 4 of OpenMP has introduced explicit support for tasks, our implementation only requires version 3 whose fork-join model is well suited for the static scheduling introduced in Section~\ref{sect:msp}.
@@
 It also generates empty components where the user must provide local sequential implementations of the kernels using well defined interfaces of the component to iterate over the local data subdomain.
+These implementations can of course be written by hand, but ideally one should also be able to generate them using a stencil compiler such as Pochoir or PATUS thus enabling the combinations of approaches used by those compilers.
 The following section will evaluate this compiler as well as the code it generates.

### 2017-02-15 ARTICLES/IJPP/articleR1/article.tex
restructuration du plan + figure MSF

@@
 
+\tikzset{
+    use/.style={
+    circle,draw=black,fill=black,scale=0.5,text=white
+    },
+    mpi/.style={
+    rectangle,draw=black,fill=black,scale=0.8,text=white
+    },
+    provide/.style={
+    circle,draw=black,fill=white,scale=0.5
+    },
+    component/.style={
+    rectangle,rounded corners=3pt,draw=black
+    },
+    dcomponent/.style={
+    rectangle,rounded corners=3pt,dashed,draw=black
+    },
+}
+
 \usepackage{todonotes}
@@
 \usepackage{xspace}
+\usepackage{color}
 
@@
 \begin{abstract}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% à revoir, il ne faut plus passer le même message et être plus clair avec une vision framework plus que langage seul
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
@@
 %------------------------------------------------------------------------------
+\section{The Component-Based Multi-Stencil Framework}
+\label{sect:msf}
+\input{msf.tex}
+%------------------------------------------------------------------------------
 \section{Meta-modelization of a Multi-Stencil Program}
@@
 %------------------------------------------------------------------------------
-\section{The Component-Based Multi-Stencil Framework}
-\label{sect:msf}
-\input{msf.tex}
-%------------------------------------------------------------------------------
-\section{The Multi-Stencil Language}
+\section{Generic Asssembly and The Multi-Stencil Language}
 \label{sect:msl}

### 2017-02-15 ARTICLES/IJPP/articleR1/msf.tex
restructuration du plan + figure MSF

--- 2015-08-21 WOLFHPC15/article/component.tex
@@
-%----------------------------------------
-%\subsection{Overview}
-%----------------------------------------
-This section describes the projection of a MSL program to a component-based runtime, including the components used to implement the static scheduling of $\Gamma_{tsp}$.
-The section first gives an overview of component models, and especially the Low Level Components (\llc) used in this work.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% présenter le framework dans sa vision d'ensemble comme dans WOLFHPC
+% faire un joli schéma sous inkscape
+% éléments à représenter et expliquer :
+%	- MSL language
+%	- generic component assembly
+%	- MSC compiler
+%	- final component assembly
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 %----------------------------------------
-\subsection{Component-Based Software Engineering}
+\subsection{Background on component models}
 %----------------------------------------
@@
 that component models can achieve the needed level of performance, and
-scalability while also helping in application
-portability~\cite{Bernholdt01052006, bigot:inria-00388508, UCHPC2015}
+scalability while also helping in application portability~\cite{Bernholdt01052006, bigot:inria-00388508, UCHPC2015}
 
@@
 This work makes use of \llc for the experiments.
-%Let therefore introduce this model in more details.
-
-%----------------------------------------
-\subsection{\llc}
-%----------------------------------------
 
@@
 
-%% In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}.
-As the use port is on the right of its component, and the provide port on the left, in a component assembly the component on the left uses the provide port of the component on the right.
-
 %----------------------------------------
-\subsection{MSP Component Runtime Overview}
+\subsection{Multi-Stencil Framework overview}
 %----------------------------------------
@@
 \begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (D) at (0,0) {$Driver$};
-   \node[provide] (Dp) at (-1,0) {};
-   \node (Ds) at (-1.5,0) {start};
-   \node[use,right=1.5cm of D] (Du1) {};
-   \node[use,below=1.75cm of D] (Du2) {};
-   \node[use,below=0.8cm of Du1] (Du3) {$m$};
-
-   \node[provide,below=0.15 of Du2] (Tp) {};
-   \node[component,below=1.6cm of Tp] (T) {$Time$};
-   \node[use,right=1cm of T] (Tu) {};
-   \node[left=0.8cm of T] (tt) {$T$};
-
-   \node[provide,right=0.15 of Tu] (Cp) {};
-   \node[component,right=2cm of Cp] (C) {$Computations$};
-   \node[use,above=0.8cm of C] (Cu) {};
-   \node[right=0.2cm of Cu] (star) {$*$};
-   \node[right=1.5cm of C] (gamma) {$\Gamma$};
-
-   \node[provide,below right=0.2 of Du3] (Datap1) {};
-   \node[provide,above=0.15 of Cu] (Datap2) {};
-   \node[component,double,above=0.8cm of Datap2] (Data) {$Data$};
-   \node[use,above=0.8cm of Data] (Datau) {};
-   \node[left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
-   \node[mpi,right=1cm of Data] (mpidata) {};
-   \node[above=0.2cm of mpidata] (star2) {$*$};
-
-   \node[provide,right=0.15 of Du1] (DDSp1) {};
-   \node[provide,above=0.15 of Datau] (DDSp2) {};
-   \node[component,above=0.8cm of DDSp2] (DDS) {$DDS$};
-   \node[right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
-
-  \path[-]
-    (Dp) edge node {} (D)
-    (D) edge node {} (Du1)
-        edge node {} (Du2)
-        edge node {} (Du3)
-    (DDSp1) edge node {} (DDS)
-    (Tp) edge node {} (T)
-    (T)  edge node {} (Tu)
-    (Cp) edge node {} (C)
-    (C) edge node {} (Cu)
-    (Datap2) edge node {} (Data)
-    (Datap1) edge node {} (Data)
-    (Data) edge node {} (Datau)
-          edge node {} (mpidata)
-    (DDSp2) edge node {} (DDS);
-\end{tikzpicture}
-\vspace*{.5em}
-\caption{MSCAC Component-based Runtime Overview.}
-\label{fig:mscac:assembly}
+  \includegraphics[width=.6\textwidth]{./images/msf.pdf}
+  \caption{The Multi-Stencil Framework (MSF) composed of the Multi-Stencil Language (MSL), the Generic Assembly (GA), the Multi-Stencil Compiler (MSC) and which produces a specialized assembly of components.}
+  \label{fig:msf}
 \end{center}
 \end{figure}
-
-As described in Section~\ref{sect:mscac}, the compiler generates four independent parts in a component assembly in addition to a static part as represented in Figure~\ref{fig:mscac:assembly}: each rounded box represents one component instance, drawn with simple lines, or more, drawn with double lines.
-For example, the component \texttt{Data} is instantiated more than once.
-\texttt{Driver} is static and drives the whole application.
-The four other parts are generated based on the content of the different sections of the MSL program: \texttt{Computations}, from the parallelization of $\Gamma$ (done by MSC as explained in Section~\ref{sect:msc}), \emph{DDS} from $\mathcal{M}$ and $\mathcal{E}$, \emph{Data} from $\Delta$ and $\mathcal{D}$, and \emph{Time} from $T$.
-
-%, and using the kernel codes provided by the user
-
-\texttt{DDS} manages the structure and the partitioning of the mesh, a single instance is generated as MSL only supports a single mesh for now.
-The back-end presented in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}.
-\texttt{DDS} is used by \texttt{Data}, for which one instance represents one data element in the MSL program.
-\texttt{Time} is parametrized by the \texttt{iteration} value from MSL and uses the \texttt{Computations} component as many times as required.
-
-\texttt{Computations} contains the static scheduling $\Gamma_{tsp}$, or the data parallel transformation $\Gamma_{data}$, both computed by MSC.
-Both transformations, $\Gamma_{tsp}$ and $\Gamma_{data}$, are encoded using three specific components, detailed in Section~\ref{sect:control}, which manage the $P$, $S$, and $sync$ operations.
-At the leaves of $\Gamma_{tsp}$ and $\Gamma_{data}$, the computation components instances (denoted $K$), provided by the user, are connected to \texttt{Data}.
-As shown in Figure~\ref{fig:k}, a $K$ component provides a port to launch the computation and exposes a use port for each data element manipulated in the computation; a star is added on the use port to \texttt{Data} to denote it.
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (k) at (0,0) {$K$};
-   \node[provide] (p) at (-1.5,0) {};
-   \node[use] (u) at (1.5,0) {};
-   \node[right=0.2 of u] (star) {*};
-
-  \path[-]
-    (p) edge node {} (k)
-    (k) edge node {} (u);
-\end{tikzpicture}
-\end{center}
-\caption{A kernel component $K$.}
-\label{fig:k}
-\end{figure}
-
-A complete assembly is obtained by replicating the component assembly of Figure~\ref{fig:mscac:assembly} on available processors (or cores).
-Connections is done with MPI ports of \texttt{Data} components.
-
-As a result of this approach, each part of the generated code is rather independent.
-Changing the main time loop to use a convergence criteria rather than a fixed number of iterations would only require changing the \texttt{Time} component.
-Similarly, changing the approach used for scheduling would only impact the code generated in \texttt{Computations}.
-The technology used for data parallelism can also be changed by replacing the \texttt{DDS} and \texttt{Data} components, and by using the adequate interfaces in $K$.
-As a matter of fact, we have started evaluating an alternative based on Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} instead of SkelGIS.
-
-%----------------------------------------
-\subsection{Control Components and Dump}
-\label{sect:control}
-%----------------------------------------
-The series-parallel tree decomposition $\Gamma_{tsp}$ represents a static scheduling of the simulation.
-$\Gamma_{data}$, on the other hand, is an optimized merged ordered list of computations.
-Both are dumped to components by generating an assembly that exactly matches their structure.
-We introduce what we call \emph{control components} to represent all nodes of $\Gamma_{tsp}$ or $\Gamma_{data}$.
-These components can be used for any simulation, which increases code reuse between simulations.
-A control component exposes a single \emph{provide} port containing a \emph{control-only} method, without any parameter.
-Three control components have been needed: SEQ, PAR, and SYNC. They are represented in Figure~\ref{fig:ctrlcomponents}.
-
-\begin{description}
-\item[Sequence component (SEQ)] It is the direct representation of a sequence node of $\Gamma_{tsp}$ or $\Gamma_{data}$.
-This component sequentially calls an ordered list of other components.
-It exposes an ordered \emph{use-multiple} port to be connected to the components to call in sequence.
-
-\item[Parallel component (PAR)] It is the direct representation of a parallel node of $\Gamma_{tsp}$ (not $\Gamma_{data}$).
-This component simultaneously calls a set of other components.
-It exposes a \emph{use-multiple} port to be connected to the components to call in parallel.
-
-\item[Synchronization component (SYNC)] It is the direct representation to an update leaf of $\Gamma_{tsp}$ or $\Gamma_{data}$.
-This component calls the synchronization of a given data.
-It exposes a \emph{use} port to be connected to the data to be updated.
-\end{description}
-
-Then, from $\Gamma_{tsp}$ or $\Gamma_{data}$, and using the control components described above,
-a direct dump can be done to a component assembly for a data or hybrid (data
-and task) parallelization of a simulation.
-Figure~\ref{fig:tsp:assembly} displays the assembly part corresponding to
-$\Gamma_{tsp}$ of Figure~\ref{fig:tsp}. In this figure, the ports
-linked to data (use and use-multiple ports of SYNC and K) are
-represented but are not connected for readability. Moreover, each computation
-component is an instance of a specific K component using
-the identification name of the computation of an MSL instance.
-Thus, this figure represents what the component \texttt{Computations} of Figure~\ref{fig:mscac:assembly} contains if an hybrid parallelization is performed.
-On the other hand, if a data parallelization is performed, a single $SEQ$ component is used by \texttt{Time}, which invokes in the appropriate order the computation kernels and synchronizations.
-Among those kernels, a single \emph{merged} kernel would represent $c_3$ and $c_5$.
-
-\begin{figure}[t]
-\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
-   \node[component] (seq) at (0,0) {$SEQ$}; \node[provide] (p) at
-   (-1,0) {}; \node[use] (u) at (1,0) {$m$};
-
-  \path[-]
-    (p) edge node {} (seq)
-    (seq) edge node {} (u);
-\end{tikzpicture}
-\hspace{\fill}
-\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
-   \node[component] (seq) at (0,0) {$PAR$};
-   \node[provide] (p) at (-1,0) {};
-   \node[use] (u) at (1,0) {$m$};
-
-  \path[-]
-    (p) edge node {} (seq)
-    (seq) edge node {} (u);
-\end{tikzpicture}
-\hspace{\fill}
-\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
-   \node[component] (sync) at (0,0) {$SYNC$};
-   \node[provide] (p) at (-1,0) {};
-   \node[use] (u) at (1,0) {};
-
-  \path[-]
-    (p) edge node {} (sync)
-    (sync) edge node {} (u);
-\end{tikzpicture}
-\\
-a) Comp. SEQ
-\hspace{\fill}
-b) Comp. PAR
-\hspace{\fill}
-c) Comp. SYNC
-\\
-\caption{The three control components.}
-\label{fig:ctrlcomponents}
-\end{figure}
-
-\begin{figure*}[t]
-\begin{center}
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   %seq0
-   \node[component] (seq0) at (0,0) {$SEQ$};
-   \node[provide] (seq0p) [left = 0.8cm of seq0] {};
-   \node[use] (seq0u) [right = 1cm of seq0] {$m$};
-   %c0
-   \node[component] (c0) [right = 2cm of seq0u] {$K(c_0)$};
-   \node[use] (c0u) [right = 0.8cm of c0] {};
-   \node[right=0.2 of c0u] (star) {*};
-   %sync0
-   \node[component] (sync0) [below = 1cm of c0] {$SYNC$};
-   \node[use] (sync0u) [right = 1cm of sync0] {$m$};
-   %c1
-   \node[component] (c1) [below = 0.8cm of sync0] {$K(c_1)$};
-   \node[use] (c1u) [right = 0.8cm of c1] {};
-   \node[right=0.2 of c1u] (star) {*};
-   %par0
-   \node[component] (par0) [below = 0.8cm of c1] {$PAR$};
-   \node[use] (par0u) [right = 1cm of par0] {$m$};
-  %seq1
-   \node[component] (seq1) [right = 1cm of sync0u] {$SEQ$};
-   \node[use] (seq1u) [right = 1cm of seq1] {$m$};
-  %c2
-   \node[component] (c2) [right = 4cm of c0] {$K(c_2)$};
-   \node[use] (c2u) [right = 0.8cm of c2] {};
-   \node[right=0.2 of c2u] (star) {*};
-  %sync1
-   \node[component] (sync1) [below = 0.8cm of c2] {$SYNC$};
-   \node[use] (sync1u) [right = 1cm of sync1] {$m$};
-  %c4
-   \node[component] (c4) [below = 0.8cm of sync1] {$K(c_4)$};
-   \node[use] (c4u) [right = 0.8cm of c4] {};
-   \node[right=0.2 of c4u] (star) {*};
-  %c6
-   \node[component] (c6) [below = 0.8cm of c4] {$K(c_6)$};
-   \node[use] (c6u) [right = 0.8cm of c6] {};
-   \node[right=0.2 of c6u] (star) {*};
-  %seq2
-   \node[component] (seq2) [below = 3.2cm of seq1] {$SEQ$};
-   \node[use] (seq2u) [right = 1cm of seq2] {$m$};
-  %c3
-   \node[component] (c3) [right = 1cm of seq2u] {$K(c_3)$};
-   \node[use] (c3u) [right = 0.8cm of c3] {};
-   \node[right=0.2 of c3u] (star) {*};
-  %c5
-   \node[component] (c5) [below = 0.8cm of c3] {$K(c_5)$};
-   \node[use] (c5u) [right = 0.8cm of c5] {};
-   \node[right=0.2 of c5u] (star) {*};
-  %c7
-   \node[component] (c7) [below = 0.8cm of par0] {$K(c_7)$};
-   \node[use] (c7u) [right = 0.8cm of c7] {};
-   \node[right=0.2 of c7u] (star) {*};
-  %sync2
-   \node[component] (sync2) [below = 0.8cm of c7] {$SYNC$};
-   \node[use] (sync2u) [right = 1cm of sync2] {$m$};
-  %c8
-   \node[component] (c8) [below = 0.8cm of sync2] {$K(c_8)$};
-   \node[use] (c8u) [right = 0.8cm of c8] {};
-   \node[right=0.2 of c8u] (star) {*};
-
-   \path[-]
-   %seq0
-    (seq0) edge node {} (seq0u)
-    (seq0p) edge node {} (seq0)
-   %c0
-    (seq0u) edge node {} (c0)
-    (c0) edge node {} (c0u)
-    %sync0
-    (seq0u) edge node {} (sync0.west)
-    (sync0) edge node {} (sync0u)
-    %c1
-    (seq0u) edge node {} (c1)
-    (c1) edge node {} (c1u)
-   %par0
-    (seq0u) edge node {} (par0.west)
-    (par0) edge node {} (par0u)
-  %seq1
-    (par0u) edge node {} (seq1.west)
-    (seq1) edge node {} (seq1u)
-  %c2
-    (seq1u) edge node {} (c2.west)
-    (c2) edge node {} (c2u)
-  %sync0
-    (seq1u) edge node {} (sync1.west)
-    (sync1) edge node {} (sync1u)
-  %c4
-    (seq1u) edge node {} (c4.west)
-    (c4) edge node {} (c4u)
-  %c6
-    (seq1u) edge node {} (c6.west)
-    (c6) edge node {} (c6u)
-  %seq2
-    (par0u) edge node {} (seq2.west)
-    (seq2) edge node {} (seq2u)
-  %c3
-    (seq2u) edge node {} (c3)
-    (c3) edge node {} (c3u)
-  %c6
-    (seq2u) edge node {} (c5.west)
-    (c5) edge node {} (c5u)
-  %c6
-    (seq0u) edge node {} (c7.west)
-    (c7) edge node {} (c7u)
-  %sync2
-    (seq0u) edge node {} (sync2.west)
-    (sync2) edge node {} (sync2u)
-  %c8
-    (seq0u) edge [bend right] node {} (c8.west)
-    (c8) edge node {} (c8u)
-   ;
-\end{tikzpicture}
-\vspace*{.5em}
-\caption{Component assembly representing the computation part generated from $\Gamma_{tsp}$ of Figure~\ref{fig:tsp}.}
-\label{fig:tsp:assembly}
-\end{center}
-\end{figure*}
-
-%----------------------------------------

### 2017-02-15 ARTICLES/IJPP/articleR1/msl.tex
restructuration du plan + figure MSF

--- 2016-05-30 ARTICLES/IJPP/article/msl.tex
@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% unifier texte avec nouvelle version article
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%----------------------------------------
+\subsection{Generic Assembly}
+%----------------------------------------
+
+\begin{figure}[t]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component] (D) at (0,0) {$Driver$};
+   \node[provide] (Dp) at (-1,0) {};
+   \node (Ds) at (-1.5,0) {start};
+   \node[use,right=1.5cm of D] (Du1) {};
+   \node[use,below=1.75cm of D] (Du2) {};
+   \node[use,below=0.8cm of Du1] (Du3) {$m$};
+
+   \node[provide,below=0.15 of Du2] (Tp) {};
+   \node[component,below=1.6cm of Tp] (T) {$Time$};
+   \node[use,right=1cm of T] (Tu) {};
+   \node[left=0.8cm of T] (tt) {$T$};
+
+   \node[provide,right=0.15 of Tu] (Cp) {};
+   \node[component, draw=green!80!black,right=2cm of Cp] (C) {$Computations$};
+   \node[use,above=0.8cm of C] (Cu) {};
+   \node[right=0.2cm of Cu] (star) {$*$};
+   \node[right=1.5cm of C] (gamma) {$\Gamma$};
+
+   \node[provide,below right=0.2 of Du3] (Datap1) {};
+   \node[provide,above=0.15 of Cu] (Datap2) {};
+   \node[component,double, draw=red,above=0.8cm of Datap2] (Data) {$Data$};
+   \node[use,above=0.8cm of Data] (Datau) {};
+   \node[left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
+   \node[mpi,right=1cm of Data] (mpidata) {};
+   \node[above=0.2cm of mpidata] (star2) {$*$};
+
+   \node[provide,right=0.15 of Du1] (DDSp1) {};
+   \node[provide,above=0.15 of Datau] (DDSp2) {};
+   \node[component, draw=red,above=0.8cm of DDSp2] (DDS) {$DDS$};
+   \node[right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
+
+  \path[-]
+    (Dp) edge node {} (D)
+    (D) edge node {} (Du1)
+        edge node {} (Du2)
+        edge node {} (Du3)
+    (DDSp1) edge node {} (DDS)
+    (Tp) edge node {} (T)
+    (T)  edge node {} (Tu)
+    (Cp) edge node {} (C)
+    (C) edge node {} (Cu)
+    (Datap2) edge node {} (Data)
+    (Datap1) edge node {} (Data)
+    (Data) edge node {} (Datau)
+          edge node {} (mpidata)
+    (DDSp2) edge node {} (DDS);
+\end{tikzpicture}
+\vspace*{.5em}
+\caption{Generic Assembly according to the Multi-Stencil program formalism.}
+\label{fig:ga}
+\end{center}
+\end{figure}
+
+%----------------------------------------
+\subsection{The Multi-Stencil Language}
+%----------------------------------------
+
 From the formalism detailed in the previous section, the Multi-Stencil Language and its grammar can already be given.

### 2017-02-17 ARTICLES/IJPP/articleR1/article.tex
restructuration du plan + figure MSF

--- 2017-02-15 ARTICLES/IJPP/articleR1/article.tex
@@
 %
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amssymb, stmaryrd}
-\usepackage{subfig}
-\usepackage{caption}
-\usepackage{cite}
-\usepackage{enumitem}
-\usepackage{filecontents}
-\usepackage{listings}
-\usepackage{mathtools, cuted}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}
 
-\usepackage{todonotes}
-\usepackage{url}
-\usepackage{xspace}
+%----------------------------------------
+\input{header.tex}
+%----------------------------------------
 
-\newcommand{\fix}[1]{\todo[inline,color=red!80]{#1}}
-\newcommand{\cf}{\emph{cf.}\xspace}
-\newcommand{\ie}{\emph{i.e.,}\xspace}
-\newcommand{\llc}{L\textsuperscript{2}C\xspace}
+%----------------------------------------
+\begin{document}
+%----------------------------------------
 
-%\newtheorem{theorem}{Theorem}[section]
-\newtheorem{myprop}[theorem]{Proposition}
-\newtheorem{mydef}[theorem]{Definition}
-\newtheorem{myth}[theorem]{Theorem}
-\newenvironment{mydefs}[1][Definitions]{\par\medskip\noindent\textbf{#1}\nopagebreak\begin{itemize}[noitemsep,topsep=.25\topsep]}{\end{itemize}}
-
-\def\pprec{\mathrel{\scalebox{.9}[1]{$\prec$}\mkern-3mu%
-  \scalebox{.4}[1]{$\prec$}\mkern-5.5mu\scalebox{.4}[1]{$\prec$}}}
-\begin{document}
-
+%----------------------------------------
 \title{Orchestrating stencils with the Multi-Stencil Framework}
-
+%----------------------------------------
 \author{H\'el\`ene Coullon \and Julien Bigot \and Christian Perez}
-
+%----------------------------------------
 \institute{H\'el\`ene Coullon \and Christian Perez \at
@@
 }
-
+%----------------------------------------
 \maketitle
-
+%----------------------------------------
+\HC{change the title, improve it}
+\HC{is it possible to have authors on a single line ?}
 \begin{abstract}
+\HC{has to be revised according to modifications}
 As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
@@
 \end{abstract}
-
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-
 \begin{acknowledgements}
@@
 This  work  was  granted  access  to  the  HPC  resources  of TGCC under  the 
-allocations t2015067470 and x2016067617 made by GENCI.
+allocations t2015067470, x2016067617 and AP010610191 made by GENCI.
 \end{acknowledgements}
-
+%----------------------------------------
 \bibliographystyle{plain}
 \bibliography{biblio}
-
+%----------------------------------------
 \end{document}
-% end of file template.tex
+%----------------------------------------

### 2017-02-17 ARTICLES/IJPP/articleR1/header.tex
restructuration du plan + figure MSF

--- 2016-05-28 ARTICLES/IJPP/article/article.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% This is a general template file for the LaTeX package SVJour3
-% for Springer journals.          Springer Heidelberg 2010/09/16
-%
-% Copy it to a new file with a new name and use it as the basis
-% for your article. Delete % signs as needed.
-%
-% This template includes a few options for different layouts and
-% content for various journals. Please consult a previous issue of
-% your journal as needed.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% First comes an example EPS file -- just ignore it and
-% proceed on the \documentclass line
-% your LaTeX will extract the file if required
-%
-\RequirePackage{fix-cm}
-%
-%\documentclass{svjour3}                     % onecolumn (standard format)
-%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
-\documentclass[smallextended]{svjour3}       % onecolumn (second format)
-%\documentclass[twocolumn]{svjour3}          % twocolumn
-%
-\smartqed  % flush right qed marks, e.g. at end of proof
-%
 \usepackage{graphicx}
@@
 \usepackage{algpseudocode}
+\usepackage[textwidth=17mm]{todonotes}
 \usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}
 
-\usepackage{todonotes}
+\tikzset{
+    use/.style={
+    circle,draw=black,fill=black,scale=0.5,text=white
+    },
+    mpi/.style={
+    rectangle,draw=black,fill=black,scale=0.8,text=white
+    },
+    provide/.style={
+    circle,draw=black,fill=white,scale=0.5
+    },
+    component/.style={
+    rectangle,rounded corners=3pt,draw=black
+    },
+    dcomponent/.style={
+    rectangle,rounded corners=3pt,dashed,draw=black
+    },
+}
+
 \usepackage{url}
 \usepackage{xspace}
+\usepackage{color}
 
-\newcommand{\fix}[1]{\todo[inline,color=red!80]{#1}}
-\newcommand{\cf}{\emph{cf.}\xspace}
+% comments
+\newcommand{\customtodo}[4]{
+        \todo[color=#2,inline,size=\small]{
+                \ifx&#3&
+                        \textbf{#1} #4
+                \else
+                        \textbf{#1$\Rightarrow$#3} #4
+                \fi
+        }
+}
+\newcommand{\HC}[2][]{\customtodo{HC}{red!20}{#1}{#2}}
+\newcommand{\CP}[2][]{\customtodo{CP}{blue!20}{#1}{#2}}
+\newcommand{\JB}[2][]{\customtodo{JB}{green!20}{#1}{#2}}
+
 \newcommand{\ie}{\emph{i.e.,}\xspace}
@@
 \newtheorem{myprop}[theorem]{Proposition}
-% \newenvironment{proof}[1][Proof]{\begin{trivlist}
-% \item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-\newenvironment{mydef}[1][Definition]{\begin{trivlist}
-\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-\newenvironment{myth}[1][Theorem]{\begin{trivlist}
-\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+\newtheorem{mydef}[theorem]{Definition}
+\newtheorem{myth}[theorem]{Theorem}
 \newenvironment{mydefs}[1][Definitions]{\par\medskip\noindent\textbf{#1}\nopagebreak\begin{itemize}[noitemsep,topsep=.25\topsep]}{\end{itemize}}
@@
   \scalebox{.4}[1]{$\prec$}\mkern-5.5mu\scalebox{.4}[1]{$\prec$}}}
-%
-% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
-%
-% insert here the call for the packages your document requires
-%\usepackage{latexsym}
-% etc.
-%
-% please place your own definitions here and don't use \def but
-% \newcommand{}{}
-%
-% Insert the name of "your journal" with
-% \journalname{myjournal}
-%
-\begin{document}
-
-\title{The Multi-Stencil Language}
-%\subtitle{Do you have a subtitle?\\ If so, write it here}
-
-%\titlerunning{Short form of title}        % if too long for running head
-
-\author{Helene Coullon \and
-        Julien Bigot \and
-        Christian Perez
-}
-
-%\authorrunning{Short form of author list} % if too long for running head
-
-\institute{Helene Coullon \and Christian Perez \at
-              INRIA, ENS Lyon, 46 Allée d'Italie, 69007 Lyon\\
-              \email{helene.coullon@inria.fr}
-           \and
-           Julien Bigot \at
-           Maison de la Simulation, CEA, CNRS, Univ. Paris-Sud, UVSQ, Universit\'e Paris-Saclay, 91191 Gif-sur-Yvette, France\\
-           \email{julien.bigot@cea.fr}
-}
-
-%\date{Received: date / Accepted: date}
-% The correct dates will be entered by the editor
-
-\maketitle
-
-\begin{abstract}
-Insert your abstract here.
-\keywords{First keyword \and Second keyword \and More}
-\end{abstract}
-
-%------------------------------------------------------------------------------
-\section{Introduction}
-\label{sect:introduction}
-Contributions :
-\begin{itemize}
-\item Formalism of a multi-stencil program and its parallelization
-\item The Multi-stencil Language: agnostic from the type of mesh and of the choosen implementation
-\item A first implementation of the MSL compiler (static scheduling + fusion + dump to SkelGIS/components)
-\item Evaluation of the different types of parallelism introduced + compiler + fusion
-\end{itemize}
-%------------------------------------------------------------------------------
-\section{Computational model of Multi-Stencil Programs}
-\label{sect:formalism}
-\input{formalism.tex}
-%------------------------------------------------------------------------------
-\section{The Multi-Stencil Language}
-\label{sect:msl}
-\input{msl.tex}
-%------------------------------------------------------------------------------
-\section{Parallelization of Multi-Stencil Programs}
-\label{sect:parallelism}
-\input{parallelism.tex}
-%------------------------------------------------------------------------------
-\section{Static Scheduling and performance model}
-\label{sect:msp}
-\input{msp.tex}
-%------------------------------------------------------------------------------
-\section{Implementation}
-\label{sect:comp}
-\input{compiler.tex}
-%------------------------------------------------------------------------------
-\section{Evaluation}
-\label{sect:eval}
-\input{evaluation.tex}
-%------------------------------------------------------------------------------
-\section{Related work}
-\label{sect:rel}
-\begin{itemize}
-\item stencil compilers cartesian : Pochoir, PATUS for optimizations/parallelization of a single stencil kernel
-\item stencil compilers unstructured meshes : Lizst for optimizations/parallelization of a single stencil kernel
-\item Halide and Nabla for pipeline of stencil computations in image (Cartesian) = mesh dependent
-\item Reuse of external solutions as SkelGIS, GA, Lizst, Pochoir etc.
-\end{itemize}
-%------------------------------------------------------------------------------
-\section{Conclusion}
-\label{sect:concl}
-%------------------------------------------------------------------------------
-
-%\begin{acknowledgements}
-%If you'd like to thank anyone, place your comments here
-%and remove the percent signs.
-%\end{acknowledgements}
-
-\bibliographystyle{plain}
-\bibliography{biblio}
-
-\end{document}
-% end of file template.tex

### 2017-02-17 ARTICLES/IJPP/articleR1/parallelism.tex
restructuration du plan + figure MSF

--- 2016-05-30 ARTICLES/IJPP/article/parallelism.tex
@@
+\HC{has to be revised according to modifications}
+
 % Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
@@
 consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
-This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
+This step can be performed by an external graph partitionner~\cite{Pellegrini:1996:SSP:645560.658570,DBLP:conf/ieeehpcs/HeleneS13,lachat:hal-00768916} and is not adressed by this paper.
 

### 2017-02-17 ARTICLES/IJPP/articleR1/formalism.tex
restructuration du plan + figure MSF
nouvelle partie II + ajout partie IV

--- 2016-05-30 ARTICLES/IJPP/article/formalism.tex
@@
+\HC{has to be revised according to modifications + see the scatter problem mentioned by reviewers}
+
 Some approaches to numerically solve partial differential equations (PDEs) are based on direct iterative methods (\textit{e.g.} finite difference, finite volume or finite element methods).
@@
 This section gives a complete formal description of what we call a \textit{stencil program} and its computations.
-The proposed language MSL (Multi-Stencil Language) does not depend on the type of space discretization used.
-Thus, some details given in the following sections, are useful to understand the domain but do not have to be kept in mind to understand the Multi-Stencil Language.
-% As a result, at the end of each subsection a summary of what is useful to understand the language is given.
+For this formal description is extracted a Multi-Stencil Program meta-model which will be used to define MSL and GA in the next section.
+%The proposed language MSL (Multi-Stencil Language) does not depend on the type of space discretization used.
+%Thus, some details given in the following sections, are useful to understand the domain but do not have to be kept in mind to understand the Multi-Stencil Language.
 
@@
 A \textit{multi-stencil program} is defined by the octuplet 
-\begin{equation*}
+\begin{equation}
 \mathcal{MSP}(\mathcal{M},\Phi,\mathcal{D},\mathcal{N},\Delta, \mathcal{S},T,\Gamma)
-\end{equation*}
+\label{eq:msp}
+\end{equation}
 \end{mydef}
@@
 \medskip
-A stencil program has been formally defined in this section. This formalism is used in the next Section to define two parallelization techniques of a multi-stencil program.
+A stencil program has been formally defined in this section.
+This formalism is used in the next Section as a meta-model to define both a the domain specific language MSL, and the generic assembly of a multi-stencil program GA.
 
-One can note that all definitions given in this section are independent from the topology of the mesh. This property will be kept in the rest of this paper to propose the mesh-agnostic MSL language.
+One can note that all definitions given in this section are independent from the topology of the mesh.
+This property will be kept in the rest of this paper so that MSF has the capacity to be used for any kind of mesh, thus being mesh-agnostic.

### 2017-02-17 ARTICLES/IJPP/articleR1/msf.tex
restructuration du plan + figure MSF
nouvelle partie II + ajout partie IV

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% présenter le framework dans sa vision d'ensemble comme dans WOLFHPC
-% faire un joli schéma sous inkscape
-% éléments à représenter et expliquer :
-%	- MSL language
-%	- generic component assembly
-%	- MSC compiler
-%	- final component assembly
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\HC{under construction, present the framework}
+
+This section first presents a background on component models and particularly on the Low Level Components.
+This background is needed to understand the second part of the section which presents an overview of the overall Multi-Stencil Framework (MSF).
 
@@
 %----------------------------------------
-Component-based software engineering (CBSE) is a domain of software engineering~\cite{Szyperski:2002:CSB:515228} which promotes code
-re-use, separations of concerns, and thus maintainability.
-An application is made of a set of component instances.
+Component-based software engineering (CBSE) is a domain of software engineering~\cite{Szyperski:2002:CSB:515228} which improves code re-use, separations of concerns, and thus maintainability.
+An component-based application is made of a set of component instances linked together, this is also called a component \emph{assembly}.
 A component is a black box that implements an independent functionality of the application, and which interacts with its environment only through well defined interfaces: its ports.
@@
 those used either by $use$ ports for a single service instance,
-or $use-multiple$ ports for multiple service instances.
+or $use-multiple$ ports for multiple services instances.
 Services are specified as \texttt{C++} interfaces.
-\llc also offers $MPI$ ports that enable components to share MPI communicators.
-Components can also have attribute ports to be configured.
+\llc also offers $MPI$ ports that enable components to share MPI communicators. Finally, 
+components can also have attribute ports to be configured.
 %
-As illustrated in Figures~\ref{fig:ports}, a $provide$ port is
-represented by a white circle, a $use$
-port with a black circle, a $use-multiple$ port by a black circle with
-a white $m$ in it. MPI port are
-connected with a black rectangle.
-A \llc-based application is a static \emph{assembly} of components instances and connections between their ports.
+In this paper, and as illustrated in Figures~\ref{fig:ports}, a $provide$ port is represented by a white circle, a $use$ port with a black circle, a $use-multiple$ port by a black circle with a white $m$ in it.
+MPI port are connected with a black rectangle.
+A \llc-based application is a static \emph{assembly} of components instances and the connections between their ports.
 Such an assembly is described in LAD, an XML dialect.
@@
 %----------------------------------------
+The Multi-Stencil Framework helps end-users to produce high performance parallel applications for the specific case of multi-stencils.
+The multi-stencil domain will be formally defined in the next section.
+A multi-stencil program is typically a numerical simulation based on explicit numerical schemes that approximate PDEs equations with one or multiple neighborhood computations.
+
+Figure~\ref{fig:msf} presents the overview of the Multi-Stencil Framework that will be entirely detailed throughout this paper. It is composed of four different parts presented below.
+
+One can note in this figure that MSF is addressed to two different kinds of end-users: the \emph{numerician}, in other words the mathematician, and the \emph{developer}.
+Most of the time numericians do have programming knowledge, however as it is not their core domain and because of a lack of time, development is often left to engineers according to numericians needs.
+MSF has the interesting particularity to propose a clear separation of concerns between these two end-users by distinguishing the description language from the numerical code.
+
+MSF also has the interesting capability to become flexible thanks to a third user possibility to interact with the framework: the High Performance Computing (HPC) specialist.
 
@@
 \end{figure}
+
+\paragraph{\textbf{Multi-Stencil Language}}
+The Multi-Stencil Language, or MSL, is the domain specific language proposed by the framework for the numerician.
+It is a descriptive language, easy to use, without any concern onto implementation details, which fits the need of a mathematician to describe the simulation.
+The description written with MSL can be considered as an input of the framework.
+MSL is described in details in Section~\ref{sect:msl}.
+The language is built upon the meta-model described in Section~\ref{sect:formalism}.
+
+\paragraph{\textbf{Generic Assembly}}
+In addition to the language MSL and the simulation description written by the numerician, MSF needs a generic assembly of a multi-stencil program as input.
+What is called a generic assembly is a component assembly for which meta-types are represented and for which some parts need to be generated or specialized.
+It is similar to a generic pattern or a skeleton in usual programming models such as C++ or functional models.
+From this generic assembly will be built the final specialized assembly of the simulation.
+As well as MSL, this generic assembly is described in Section~\ref{sect:msl} and is built upon the meta-model described in Section~\ref{sect:formalism}.
+Moreover, an example of generic assembly is used in evaluations of Section~\ref{sect:eval}.
+\HC{Note idée : HPC specialist gives types for DDS and Data}
+
+\paragraph{\textbf{Multi-Stencil Compiler}}
+The core of the framework is the Multi-Stencil Compiler, or MSC.
+It is responsible for transforming the genric assembly into the final parallel assembly which is specific to the simulation described by the numerician.
+MSC is described in Sections~\ref{sect:parallelism} and~\ref{sect:msp}.
+
+\paragraph{\textbf{Specialized assembly}}
+Finally, the input of MSF is the component assembly generated by MSC.
+It is an instanciation and a transformation of the generic component assembly, by adding component types and by adding specific components generated by MSC.
+From this final component assembly which is specific to the simulation initially described by using MSL, the developer will finally write numerical codes.

### 2017-02-17 ARTICLES/IJPP/articleR1/msl.tex
restructuration du plan + figure MSF
nouvelle partie II + ajout partie IV

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% unifier texte avec nouvelle version article
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\HC{under revision / has to be revised according to modifications}
+
 %----------------------------------------
@@
 %----------------------------------------
+From the meta-model represented by Equation~\ref{eq:msp} the generic assembly of any multi-stencil program can be built.
+It is represented by Figure~\ref{fig:ga}.
+As indicated, the GA has five different components described below.
 
@@
    \node[provide,right=0.15 of Tu] (Cp) {};
-   \node[component, draw=green!80!black,right=2cm of Cp] (C) {$Computations$};
+   \node[component, draw=green!80!black, text=blue!80!black,right=2cm of Cp] (C) {$Computations$};
    \node[use,above=0.8cm of C] (Cu) {};
@@
    \node[use,above=0.8cm of Data] (Datau) {};
-   \node[left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
+   \node[left=1cm of Data] (delta) {$\Delta$};
    \node[mpi,right=1cm of Data] (mpidata) {};
@@
    \node[component, draw=red,above=0.8cm of DDSp2] (DDS) {$DDS$};
-   \node[right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};
+   \node[right=1.5cm of DDS] (m) {$\mathcal{M},\Phi,\mathcal{D},\mathcal{N}$};
 
@@
 
+\paragraph{\textbf{Driver}}
+This component can be compared to the main function of a usual program. It is responsible for initialization and execution of other components, just like variable initializations and function calls.
+
+\paragraph{\textbf{Time}}
+As indicated in the figure, this component is responsible for the time $T$ defined in Equation~\ref{eq:msp}. Thus it is composed of a time loop and eventually of a convergence reduction.
+
+\paragraph{\textbf{DDS}}
+This component is responsible for the mesh and its entities $\mathcal{M}$, and $\Phi$, the computation domains $\mathcal{D}$ and the set of neighborhood functions $\mathcal{N}$.
+Moreover, if a data parallelization is needed, this component should handle mesh partitioning and should provide a synchronization interface as it will be detailed in Section~\ref{sect:parallelism}.
+One can note that this component is red, this means that this component should be provided by a HPC specialist in MSF as illustrated in the previous Figure~\ref{fig:msf}.
+
+\paragraph{\textbf{Data}}
+This component is inseparable from the previous one $DDS$.
+It actually uses the DDS component to instanciate one quantity.
+Thus, it is responsible for the set of quantities $\Delta$.
+This component is also red.
+As $DDS$ it should be provided by a HPC specialist.
+Moreover, it is represented by a double line box which means that it can be instanciated multiple time will others will be instanciated once.
+Actually, this component will be instanciated as many times as the number of quantities in the simulation.
+
+\paragraph{\textbf{Computations}}
+Finally, this component is responsible for $\Gamma$, \ie the computations of the simulation.
+This component is green and blue because it will be partly generated by MSC and partly feeded by the developer.
+Actually, this component will be replaced by MSC by a sub-assembly of component for which the parallel part will be automatically generated, and for which components responsible for simulation code will be completed by the developer.
+
 %----------------------------------------
@@
 
-From the formalism detailed in the previous section, the Multi-Stencil Language and its grammar can already be given.
-This grammar is sufficient to automatically extract parallelism from the program as will be explained in the next section.
+The second element of MSF which is built upon the meta-model represented by Equation~\ref{eq:msp} is the Multi-Stencil Language and its grammar.
+This grammar is sufficient (in addition to GA) for MSC to automatically extract a parallel pattern of the simulation which is finally dumped as a specialized instanciation of GA.
 
@@
 The grammar of the Multi-Stencil Language is given in Figure~\ref{fig:grammar} and an example is provided in Figure~\ref{fig:exmsl}.
-A Multi-Stencil program is composed of eight parts.
+A Multi-Stencil program is composed of eight parts as in Equation~ref{eq:msp}.
 

### 2017-02-20 ARTICLES/IJPP/articleR1/compiler.tex
restructuration du plan + figure MSF
nouvelle partie II + ajout partie IV
modif related work + minor changes

@@
+\HC{has to be revised according to modifications}
+\HC{add a figure of the final specialized assembly with the scheduler and sync components}
+
 %--------------------
@@
 %With the increasing number of cores of current machines this has not proved to be a problem in our experience. % LN : ?
-While the version 4 of OpenMP has introduced explicit support for tasks, our implementation only requires version 3 whose fork-join model is well suited for the static scheduling introduced in Section~\ref{sect:msp}.
+Although the version 4 of OpenMP has introduced explicit support for tasks, our implementation only requires version 3 whose fork-join model is well suited for the static scheduling introduced in Section~\ref{sect:msp}.
 The use of dynamic schedulers such as provided by libgomp\footnote{\url{https://gcc.gnu.org/projects/gomp/}}, StarPU~\cite{Augonnet2011}, or XKaapi~\cite{Gautier:2013:XRS:2510661.2511383} to directly execute the DAG $\Gamma_{dep}$ will be the subject of some future work.

### 2017-02-20 ARTICLES/IJPP/articleR1/formalism.tex
modif related work + minor changes

@@
 For example, in a 2D Cartesian mesh, a group of entities where each entity is made of four vertices and four edges could form the cells.
-Another group of entities are the vertices defined as all singletons formed of a single vertex of $V$. Both groups, then, would be part of $\Phi$. This example is illustrated in Figure~\ref{fig:ex1}.
+Another group of entities are the vertices defined as all singletons formed of a single vertex of $V$.
+Both groups, then, would be part of $\Phi$.
+This example is illustrated in Figure~\ref{fig:meshbase}.
 
@@
 \item $\mathbb{V}=\Delta\cup\mathcal{S}$ is the set of \emph{variables} or \emph{quantities}.
-\item $\Delta$ are the \textit{mesh variables} that to each entity of a group and time-step associates a value $\delta: \mathcal{G}\times T\mapsto \mathcal{V}_\delta$ where $V_{\delta}$ is a value type.
+\item $\Delta$ are the \textit{mesh variables} that to each entity of a group and time-step associates a value $\delta: \mathcal{G}\times T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
 \item The group of entities a variable is mapped on is denoted $entity(\delta)=\mathcal{G}$.
@@
 \item Amongst the scalar variables is one specific variable $conv\in\mathcal{S}$, the convergence criteria whose value is $0$ except at the last step where it is 1.
-Thus, $\forall t\in \llbracket 0, T_{max-1} \rrbracket$, $conv(t)=1$, and $conv(T_{max})=1$.
+Thus, $\forall t\in \llbracket 0, T_{max-1} \rrbracket$, $conv(t)=0$, and $conv(T_{max})=1$.
 \end{mydefs}

### 2017-02-20 ARTICLES/IJPP/articleR1/introduction.tex
restructuration du plan + figure MSF
modif related work + minor changes

--- 2016-05-30 ARTICLES/IJPP/article/introduction.tex
@@
+\HC{has to be revised according to modifications}
+
 As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
-For example, the current world's top supercomputer Tianhe-2~\footnote{url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance around thirty peta floating point operations par seconds.
+For example, the current world's top supercomputer Tianhe-2~\footnote{\url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance around thirty peta floating point operations par seconds.
 However, to be able to use such a machine, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA etc., and multiple optimization techniques, such as cache or vectorization optimizations, have to be combined.

### 2017-02-20 ARTICLES/IJPP/articleR1/msl.tex
modif related work + minor changes

@@
             "mesh quantities:" listquantities
-            "scalars" listscalar
+            "scalars:" listscalar
             listloop
@@
           "computations:" listcomp
-iteration ::= num | scalar
+iteration ::= num | scalarid
 listcomp ::= comp listcomp |  comp
@@
   F[d1] = k4(D,C[nce])
-  G[d1] = k0(mu,tau,E)
+  G[d1] = k5(mu,tau,E)
   H[d2] = k6(F)

### 2017-02-20 ARTICLES/IJPP/articleR1/msp.tex
restructuration du plan + figure MSF
nouvelle partie II + ajout partie IV
modif related work + minor changes

--- 2016-05-24 ARTICLES/IJPP/article/msp.tex
@@
+\HC{has to be revised according to modifications}
+\HC{reduce and improves the part on SP graphs}
+
 In this section we detail a static scheduling of $\Gamma_{dep}$ by using minimal series-parallel directed acyclic graphs.
@@
 \subsection{Series-Parallel graph}
+\label{sect:tsp}
 
 In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of Minimal Series-Parallel DAGs (MSPD).
-Such a graph can be decomposed as a serie-parallel tree, denoted $TSP$, where each leaf is a vertex of the MSPD DAG it represents, and whose internal nodes are labelled $S$ or $P$ to indicate respectively the series or parallel composition of sub-trees.
+Such a graph can be decomposed as a serie-parallel tree, denoted $TSP$, where each leaf is a vertex of the MSPD it represents, and whose internal nodes are labelled $S$ or $P$ to indicate respectively the series or parallel composition of sub-trees.
 Such a tree can be considered as a fork-join model and as a static scheduling.
@@
 
-To remove the forbidden N-shapes from the transitive reduction of $\Gamma_{dep}=(V,E)$, we have chosen to apply an over-constraint with the relation $k_0 \prec k_3$, such that a complete bipartite graph is created for the sub-dag, and can be translated to a series-parallel decomposition, as illustrated in Figure~\ref{fig:allover}.
+To remove the forbidden N-shapes from the transitive reduction of $\Gamma_{dep}=(V,E)$, we have chosen to apply an over-constraint with the relation $k_0 \prec k_3$, such that a complete bipartite graph is created for the sub-dag, and can be translated to a series-parallel decomposition, as illustrated in Figure~\ref{fig:over}.
 
@@
 \subsection{Performance model}
+\label{sect:perfs}
 
@@
 \begin{equation}
-T_{COM1} \geq T_{COM2} + T_{TH}
+T_{COM1} \geq T_{COM2} + T_{task}
+\label{eq:hyb}
 \end{equation}
@@
 \subsection{Fusion optimization}
+\label{sect:fusion}
 
@@
 \If {parent($k_i$)==parent($k_j$)}
-\If {$type(k_i)=type(k_j)=comp$}
+\If {$type(k_i)==type(k_j)==comp$}
 \If {parent($k_i$)==$S$}

### 2017-02-20 ARTICLES/IJPP/articleR1/related.tex
restructuration du plan + figure MSF
nouvelle partie II + ajout partie IV
modif related work + minor changes

--- 2016-05-30 ARTICLES/IJPP/article/related.tex
@@
-Many domain specific languages have been proposed for stencil computations.
+% orchestration of stencils, different from stencil compilers like Pochoir and PATUS which could be used inside MSF
+First of all, many domain specific languages have been proposed for the optimization of a single stencil computation.
 Each one has its own specificities and answers to a specific stencil case or to a specific additionnal optimization.
@@
 On the other hand, PATUS~\cite{citeulike12258902} proposes to add a parallelization strategy grammar to its stencil language to perform an auto-tuning parallelization strategy.
-Moreover, Halide~\cite{Ragan-Kelley:2013:HLC:2491956.2462176}  proposes an optimization and parallelization of a pipeline of stencil codes, while ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} is specific to multigrid numerical methods etc.
+Finally, ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} is specific to multigrid numerical methods etc.
+Thus, these stencil compilers target a different scope than the Multi-Stencil Framework presented in this paper, which actually orchestrates a parallel execution of multiple stencil codes together.
+Furthermore, future works could lead to a combination of these stencil compilers with MSF to build very optimized stencil kernels $K$.
 
-The reason why each of those languages are implemented from scratch to answer its own particularities and its own optimizations is because each language is built and thought as a single block, which makes impossible or difficult code reuse from one language to another.
-In other words domain specific languages (for stencils or other domains) suffer from a lack of software engineering properties, which could increase the productivity to build a new language, by code-reuse, and also the maintainability of languages, with more separation of concerns.
-A few works propose frameworks to create DSL, which afterwards will be easier to compose and combine together~\cite{Sujeeth:2013:CRC:2524984.2524988}.
-The Mulsi-Stencil Language (MSL), presented in this paper, is itself a domain specific language for stencil-based simulations.
-It offers a way to give a mesh-agnostic description of a simulation, which could be common to different cases of simulations, and thus facilitates reuse of existing underlying data structures, optimizations or parallelization techniques.
-In other words, MSL extracts where parallelization and optimizations are needed into the simulation, by producing an empty parallel pattern, but not how to do it, which is left to existing languages (SkelGIS and OpenMP in this paper), and which is an implementation concern.
+% Stencil orchestration
+Some solutions, closer to MSF, have also been proposed to automatically orchestrate multiple stencils and local computations in a parallel manner.
+Amongs them is Halide~\cite{Ragan-Kelley:2013:HLC:2491956.2462176} that proposes an optimization and parallelization of a pipeline of stencils.
+However, Halide is limited to Cartesian meshes and is specific to images.
+Liszt~\cite{DeVito:2011:LDS:2063384.2063396} and Nabla~\cite{Camier:2015:IPP:2820083.2820107} both offer solutions for the automatic parallel orchestration of stencils applied onto any kind of mesh, from Cartesian to unstructured meshes.
+The needed mesh can be built from a set of available symbols in the grammar of each language.
+Thus, these languages generalize the definition of a mesh, as it is proposed into the MSL formalism of Section~\ref{sect:formalism}.
+Neither Liszt or Nabla, however, handle hybrid parallelism as it is proposed by MSF.
 
-Liszt~\cite{DeVito:2011:LDS:2063384.2063396} and Nabla~\cite{Camier:2015:IPP:2820083.2820107}, both offer languages for stencils applied onto any kind of mesh, from Cartesian to unstructured meshes.
-The mesh which is needed into the simulation can be built from a set of available symbols in the grammar of each language.
-Thus, those languages generalize the definition of a mesh, as it is proposed into the MSL formalism.
-However, two main differences can be noticed.
+% MSF is not a general purpose framework. It proposes a new approach to improve sep. of concerns and code-reuse in Domain Specific Languages, by combining it to component generic assemblies.
+We have made the choice in MSF to propose and use the MSL Domain Specific Language to ask the numerician a sequential description of its simulation.
+This description, is close to a dataflow representation.
+However, MSL is specific to numerical simulations and proposes a mesh abstraction adapted to numerical simulations.
+Compared to general purpose dataflow runtimes such as Legion~\cite{bauer:legion:sc:2012}, HPX~\cite{}, PFunc~\cite{}, MSL is closer to the semantic of the domain (mesh, stencils etc.), thus easier to use for non-specialists.
+Moreover, MSL is very light and only descriptive as numerical codes are left to another language and another user (the developer in Figure~\ref{fig:msf}).
+Furthermore, these three runtimes could actually be used by MSF as a backend, instead of using SkelGIS and OpenMP.
+This is due to software engineering capacities introduced by proposing a meta-model and by using component models in MSF.
 
-First, the formalism is more flexible in MSL.
-For example, a computation can be applied onto a subset of the space domain which is not possible with Liszt and Nabla.
-This functionnality is important in numerical simulations yet.
-For example, many computations performed in the simulation studied in this paper have to be applied onto a subpart of the overall space domain.
+% vision of the framework brings separation of concerns, code-reuse and flexibility
+Actually, the vision proposed by MSF is to use components to be able to increase separation of concerns and code-reuse of the overall Framework.
+%%%% sep of concerns
+Separation of concerns is illustrated in Figure~\ref{fig:msf} and all along the paper.
+The numerician is only responsible for the description of the simulation by using MSL.
+A HPC specialist can adapt or propose new components for handling the distributed data structure and quantities of the simulation.
+MSF generates from this information the parallel orchestration of computations.
+Finally, after MSF compilation (MSC), the developer of numerical codes fills computation kernels by using the chosen distributed data structure.
+In Liszt~\cite{DeVito:2011:LDS:2063384.2063396} and Nabla~\cite{Camier:2015:IPP:2820083.2820107}, there is no such separation of concerns between the numerician and the developer.
+%%%% flexibility
+Moreover, it is not possible to easily integrate a new distributed data structure in these solutions as a monolithic code is generated.
+By using components, doors to online reconfiguration of applications~\cite{Lanore:2015:RCM:2737166.2737169} are also open (even if not answered in this paper).
+%%%% code-reuse
+Finally, thanks to components, MSF improves code-reuse.
+Kernel components as well as any component (except the scheduler component which is specific to the simulation) can be reused from one simulation to another.
 
-Second, in Liszt and Nabla the description of a simulation is not splitted from its implementation concerns.
-In other words, the topology of a mesh as well as numerical codes are given at the same time than the description of the different computations to apply into the simulation, while MSL splits those two different phases.
-Thus, MSL improves separation of concerns and maintainability of a given simulation.
-Actually, a numerician could describes the different computations to perform into the simulation, while another one, later, could focuss on the implementation and the numerical code.
-This also facilitates reuse of other languages.
-
-The formalism of a Multi-Stencil program, given in this paper, is new, as far as we know.
-Finally, the parallelization techniques proposed in this paper, take place at paradigm level, with data parallelism and hybrid (data plus task) parallelism.
-Thus the parallel pattern of the simulation does not need details onto parallel architectures (distributed or shared memories, with or without accelerators).
-This makes possible a large panel of backend architectures and languages.
+% no component-based framework for stencils orchestration (FFT, MapReduce/skeletons HLCM, nothing with DSLs)
+To conclude this related work, as far as we know no component-based framework have already been proposed for stencils orchestration.

### 2017-02-21 ARTICLES/IJPP/articleR1/conclusion.tex
restructuration du plan + figure MSF
nouvelle partie II + ajout partie IV
modif related work + minor changes
modif related work minor changes

--- 2016-05-30 ARTICLES/IJPP/article/conclusion.tex
@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% unifier texte avec nouvelle version article
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 We have presented the Multi-Stencil Language (MSL), a domain specific language for stencil-based numerical simulations.
@@
 Thus using a single language, the MSL language, two different kinds of numerical simulations could be defined: one onto Cartesian meshes; and one on unstructured meshes.
+
+\HC{framework to ease HPC specialist work for proposing new meshes etc.}
+\HC{how to choose data or hybrid, change on the fly if needed}
+\HC{how to choose if a fusion is good or not, change on the fly or learning strategies onto production codes}
+\HC{In addition to this, by using components, doors to online reconfiguration of applications~\cite{Lanore:2015:RCM:2737166.2737169} are also open (even if not answered in this paper).}

### 2017-02-21 ARTICLES/IJPP/articleR1/related.tex
modif related work minor changes

@@
 First of all, many domain specific languages have been proposed for the optimization of a single stencil computation.
-Each one has its own specificities and answers to a specific stencil case or to a specific additionnal optimization.
+Each one has its own optimization specificities and answers a specific numerical method or a specific kind of mesh.
 For example, Pochoir~\cite{spaaTangCKLL11} works on cache optimization techniques for stencils applied onto Cartesian meshes.
-On the other hand, PATUS~\cite{citeulike12258902} proposes to add a parallelization strategy grammar to its stencil language to perform an auto-tuning parallelization strategy.
+On the other hand, PATUS~\cite{citeulike12258902} proposes to add a parallelization strategy grammar to its stencil language to perform an auto-tuning parallelization strategies.
 Finally, ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} is specific to multigrid numerical methods etc.
@@
 % Stencil orchestration
-Some solutions, closer to MSF, have also been proposed to automatically orchestrate multiple stencils and local computations in a parallel manner.
+Some solutions, closer to MSF, have also been proposed to automatically orchestrate multiple stencils and computations in a parallel manner.
 Amongs them is Halide~\cite{Ragan-Kelley:2013:HLC:2491956.2462176} that proposes an optimization and parallelization of a pipeline of stencils.
 However, Halide is limited to Cartesian meshes and is specific to images.
-Liszt~\cite{DeVito:2011:LDS:2063384.2063396} and Nabla~\cite{Camier:2015:IPP:2820083.2820107} both offer solutions for the automatic parallel orchestration of stencils applied onto any kind of mesh, from Cartesian to unstructured meshes.
+Liszt~\cite{DeVito:2011:LDS:2063384.2063396}, OP2~\cite{} and Nabla~\cite{Camier:2015:IPP:2820083.2820107} all offer solutions for the automatic parallel orchestration of stencils applied onto any kind of mesh, from Cartesian to unstructured meshes.
 The needed mesh can be built from a set of available symbols in the grammar of each language.
 Thus, these languages generalize the definition of a mesh, as it is proposed into the MSL formalism of Section~\ref{sect:formalism}.
-Neither Liszt or Nabla, however, handle hybrid parallelism as it is proposed by MSF.
+HOwever, neither Liszt, OP2 or Nabla handle hybrid parallelism as it is proposed by MSF.
 
 % MSF is not a general purpose framework. It proposes a new approach to improve sep. of concerns and code-reuse in Domain Specific Languages, by combining it to component generic assemblies.
-We have made the choice in MSF to propose and use the MSL Domain Specific Language to ask the numerician a sequential description of its simulation.
+MSF offers the MSL Domain Specific Language to the numerician to describe its sequential set of computations.
 This description, is close to a dataflow representation.
-However, MSL is specific to numerical simulations and proposes a mesh abstraction adapted to numerical simulations.
-Compared to general purpose dataflow runtimes such as Legion~\cite{bauer:legion:sc:2012}, HPX~\cite{}, PFunc~\cite{}, MSL is closer to the semantic of the domain (mesh, stencils etc.), thus easier to use for non-specialists.
-Moreover, MSL is very light and only descriptive as numerical codes are left to another language and another user (the developer in Figure~\ref{fig:msf}).
-Furthermore, these three runtimes could actually be used by MSF as a backend, instead of using SkelGIS and OpenMP.
-This is due to software engineering capacities introduced by proposing a meta-model and by using component models in MSF.
+However, MSL differs from general purpose dataflow languages or framework for two main reasons.
+First, MSL is specific to numerical simulations and proposes a mesh abstraction adapted to numerical simulations.
+Thus, compared to general purpose dataflow runtimes such as Legion~\cite{bauer:legion:sc:2012}, HPX~\cite{}, PFunc~\cite{}, MSL is closer to the semantic of the domain (mesh, stencils etc.) and easier to use for non-specialists.
+Second, MSL is very light and only descriptive.
+Numerical codes are left to another language and another user (the developer in Figure~\ref{fig:msf}).
+Furthermore, such dataflow runtimes could actually be used by MSF as a backend, instead of using SkelGIS and OpenMP.
 
 % vision of the framework brings separation of concerns, code-reuse and flexibility
-Actually, the vision proposed by MSF is to use components to be able to increase separation of concerns and code-reuse of the overall Framework.
+This flexibility of MSF is due to software engineering capacities introduced by proposing a meta-model and by using component models.
+Actually, MSF is designed to increase separation of concerns and code-reuse compared to existing solutions.
 %%%% sep of concerns
@@
 The numerician is only responsible for the description of the simulation by using MSL.
-A HPC specialist can adapt or propose new components for handling the distributed data structure and quantities of the simulation.
-MSF generates from this information the parallel orchestration of computations.
+A HPC specialist can propose new (or updated) components for handling the distributed data structure and quantities of the simulation.
+MSF generates from these informations the parallel orchestration of computations.
 Finally, after MSF compilation (MSC), the developer of numerical codes fills computation kernels by using the chosen distributed data structure.
-In Liszt~\cite{DeVito:2011:LDS:2063384.2063396} and Nabla~\cite{Camier:2015:IPP:2820083.2820107}, there is no such separation of concerns between the numerician and the developer.
+In Liszt, OP2 and Nabla, there is no such separation of concerns between the numerician and the developer.
 %%%% flexibility
 Moreover, it is not possible to easily integrate a new distributed data structure in these solutions as a monolithic code is generated.
-By using components, doors to online reconfiguration of applications~\cite{Lanore:2015:RCM:2737166.2737169} are also open (even if not answered in this paper).
 %%%% code-reuse

### 2017-02-22 ARTICLES/IJPP/articleR1/article.tex
modif related work + minor changes
unification article et corrections - parallelism.tex en cours

@@
 %
-
 %----------------------------------------
@@
 \begin{abstract}
-\HC{has to be revised according to modifications}
-As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
+%\HC{has to be revised according to modifications}
+%As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
 One of the big challenges of exascale is to get programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists.
@@
 %------------------------------------------------------------------------------
-\section{Static Scheduling and performance model}
-\label{sect:msp}
-\input{msp.tex}
+%\section{Static Scheduling and performance model}
+%\label{sect:msp}
+%\input{msp.tex}
 %------------------------------------------------------------------------------
@@
 \bibliographystyle{plain}
-\bibliography{biblio}
+\bibliography{article}
 %----------------------------------------

### 2017-02-22 ARTICLES/IJPP/articleR1/formalism.tex
unification article et corrections - parallelism.tex en cours

@@
-\HC{has to be revised according to modifications + see the scatter problem mentioned by reviewers}
-
 Some approaches to numerically solve partial differential equations (PDEs) are based on direct iterative methods (\textit{e.g.} finite difference, finite volume or finite element methods).
@@
 This section gives a complete formal description of what we call a \textit{stencil program} and its computations.
-For this formal description is extracted a Multi-Stencil Program meta-model which will be used to define MSL and GA in the next section.
-%The proposed language MSL (Multi-Stencil Language) does not depend on the type of space discretization used.
-%Thus, some details given in the following sections, are useful to understand the domain but do not have to be kept in mind to understand the Multi-Stencil Language.
+A Multi-Stencil Program meta-model is extracted from this formal description.
+The meta-model will be used to define MSL and GA in the next section.
 
@@
 A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) as illustrated in Figure~\ref{fig:mesh1}.
-One can notice that more than one type of mesh is also possible inside a single simulation.
+One can notice that more than one type of mesh is possible inside a single simulation.
 For example, an hybrid mesh can be defined as an unstructured mesh composed itself of a Cartesian mesh inside each of its vertices.
-% Multi-grid numerical methods is another example. %JB: pas vraiment...
-However in this paper, single mesh simulations are addressed.
+However, in this paper single mesh simulations are addressed.
 
@@
 
-For example, in a 2D Cartesian mesh, a group of entities where each entity is made of four vertices and four edges could form the cells.
+For example, in a 2D Cartesian mesh, a group of entities where each entity is made of four vertices and four edges represents the cells.
 Another group of entities are the vertices defined as all singletons formed of a single vertex of $V$.
@@
 
-\begin{mydef}[Definitions (time)]
+\begin{mydef}
 The finite sequence $T: (t_n)_{n\in\llbracket 0, T_{max} \rrbracket}$ represents the discretization of the continuous time domain $\mathcal{T}=\mathbb{R}$.
-To each discrete time-step $n\in\llbracket 0, T_{max} \rrbracket$, it associates a time value $t_n\in\mathcal{T}$.
+To each discrete time-step $n\in\llbracket 0, T_{max} \rrbracket$, a time value $t_n\in\mathcal{T}$ is associated. The time step is $t_n-t_{n-1}$.
 \end{mydef}
@@
 The time discretization can be as simple as a constant time-step with a fixed number of steps.
-It can, however, also be defined by recursion with both the time-step and the number of steps depending on the data variables (see definitions below).
+The time-step and the number of steps can also change on the fly during the simulation depending on variables values (see definitions below).
 
@@
 \item $\mathbb{V}=\Delta\cup\mathcal{S}$ is the set of \emph{variables} or \emph{quantities}.
-\item $\Delta$ are the \textit{mesh variables} that to each entity of a group and time-step associates a value $\delta: \mathcal{G}\times T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
+\item $\Delta$ are the \textit{mesh variables}.
+A mesh variable $\delta \in \Delta$ associates to each pair of entity and time-step a value $\delta: \mathcal{G}\times T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
 \item The group of entities a variable is mapped on is denoted $entity(\delta)=\mathcal{G}$.
-\item $\mathcal{S}$ are the \textit{scalar variables} that to each time-step associates a value $s: T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
-\item Amongst the scalar variables is one specific variable $conv\in\mathcal{S}$, the convergence criteria whose value is $0$ except at the last step where it is 1.
+\item $\mathcal{S}$ are the \textit{scalar variables}.
+A scalar variable $s \in \mathcal{S}$ associates to each time-step a value $s: T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
+\item Among the scalar variables is one specific variable $conv\in\mathcal{S}$, the convergence criteria, whose value is $0$ except at the last step where it is 1.
 Thus, $\forall t\in \llbracket 0, T_{max-1} \rrbracket$, $conv(t)=0$, and $conv(T_{max})=1$.
@@
 
-%For example, the simplest $conv$ is typically to have two scalars: $t$ the time, and a constant number of iterations denoted $it$.
-%At each time-step $\Delta t=1$, and $conv$ returns the boolean expression $t<it$.
-
 \medskip
 
-This section has presented the general formalism of meshes, their entities, groups of entities, time discretization, and finally quantities (mapped on mesh and scalars).
-While the details about mesh and their topologies were useful to specify the following concepts, the Multi-Stencil Language presented in this paper is mesh-agnostic.
-These details will therefore not be needed in the remaining of the paper.
+This section has presented the formalism of meshes, their associated entities, groups of entities,  and also time discretization and quantities (mapped on meshes or scalars).
 
@@
 
-%In this section four different types of kernel computations are considered: stencil kernels, boundary kernels, local kernels, and reduction kernels.
-
 \medskip
 \begin{mydefs}
-\item A computation domain $D$ is a subpart of a mesh entities group, $D \subseteq \mathcal{G} \in \Phi$.
+\item A computation domain $D$ is a subpart of a group of mesh entities, $D \subseteq \mathcal{G} \in \Phi$.
 \item The set of computation domains of a numerical simulation is denoted $\mathcal{D}$.
@@
 \item $S \in \mathcal{S}$ is the set of scalar to read,
-\item $(w,D) \in \Delta \times \mathcal{D}$ is the single data written by the kernel:
+\item $(w,D) \in \mathbb{V} \times \mathcal{D}$ is the single data written by the kernel:
 \begin{itemize}
@@
 \begin{itemize}
-  \item $r$ is a quantity read by the kernel to compute $w$,
-  \item $n : entity(w) \rightarrow entity(r)^m$ is a neighborhood function that indicates whose values of $r$ are read to compute $w$.
+  \item $r$ is a mesh variable read by the kernel to compute $w$,
+  \item $n : entity(w) \rightarrow entity(r)^m$ is a neighborhood function that indicates which values of $r$ are read to compute $w$.
 \end{itemize}
@@
 
-At each time-step, a set of computations is performed.
-During a computation kernel, it can be considered that a set of old states ($t-1$) of quantities are read ($R$), and that a new state ($t$) of a single quantity is written ($w$).
-%TODO[jb] introduire la contrainte plus tard
+In a Multi-Stencil simulation, at each time-step, a set of computations is performed.
+During a computation kernel, it can be considered that a set of old states ($t-1$) of quantities are read ($\mathcal{S}$ and $R$), and that a new state ($t$) of a single quantity is written ($w$).
 
@@
 \item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{local kernel} if $\forall (r,n) \in R$, $n = identity$.
-\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} if $w$ is a scalar.%, $D=\emptyset$, $R=\{(r,n)\}$ with $n=entity(r)$, and $comp$ is a reduction operation (usually $+$, $-$, $*$, $max$, $min$ etc.).
+\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} if $w$ is a scalar. A reduction is typically used to compute the convergence criteria of the time loop of the simulation.
 \end{itemize}
-Since we only consider explicit numerical schemes in this paper, a kernel can not write the same quantity it reads, except locally, \textit{i.e.} if $\exists (w,n) \in R\Rightarrow n=identity$
 
-%In this paper only explicit numerical schemes are considered, thus if we denote by $identity$ the identity function, we have $w \not\in R$ except if $w=(r,identity)$.
+Since we only consider explicit numerical schemes in this paper, a kernel cannot write the same quantity it reads, except locally, \textit{i.e.} if $\exists (w,n)$ where $w \in R\Rightarrow n=identity$.
 
-%In our formalism, it is forbidden for a stencil kernel to find $(r,n) \in R$ for which $r=w$ and $n \neq identity$.
-Actually, if such a property was permitted, implicit numerical schemes would be needed into the simulation, which involves linear solvers.
-Such a scheme is not a stencil and is over the scope of this paper.
-The problem does not happened for local kernels as $\forall (r,n) \in R$, $n = identity$.
-Computations for which this property is authorized are boundary kernels only.
-This type of kernel is a particular case as it commputes boundary conditions, which represents what should happened outside the space domain but still impact stencil kernels at the next time-step.
-
-% \noindent \textbf{Property}
-% A kernel for which all data read and written are applied onto a mesh of dimension $0$ is a local kernel.
-
-% \noindent \textbf{Property}
-% In a reduction kernel $k(S,R,(w,D),comp)$, $D=entity(w)$ as a single entity exists for a scalar.
-
-%\medskip
-%A reduction computation is a computation which reads a single data applied onto a mesh and returns from all its entities a single scalar (mesh with a dimension reduced to zero).
-
-A reduction is typically used to compute the convergence criteria of the time loop of the simulation. %Occasionally reductions can also be used to compute other scalars.
-
-% \medskip
-% \noindent \textbf{Property}
-% Considering a reduction kernel $k(S,R,(w,D),comp)$, $comp$ must be a binary and associative operation on the type $V$, $comp: V \times V \rightarrow V$.
+At this point, it could seem counter-intuitive to restrict the definition of a computation to a single quantity to write.
+This is actually counter-intuitive for a computer scientist who will naturally perform more than one computation into a loop to optimize cache memory use and to get a shorter code.
+The definition used above actually follows the usual "=" meaning of an equation or a numerical computation, where a single variable is written on the left of the symbol.
+We have made this choice for three reasons.
+\begin{itemize}
+\item First, we think this notation is more natural for a numerician who usually manipulate equations.
+\item Second, we call a \emph{fusion} the fact of writing more than one quantity into a single loop.
+When writing a sequential program, fusion is natural for the developer, however, writing a parallel code without knowing it (as it is by using MSF), fusion are error prone and a bad fusion could lead to an incorrect numerical result.
+\item Finally, when performing task parallelism, as it is in this paper, it is not always the good choice to perform fusions as you reduce the number of possible tasks.
+\end{itemize}
+For these reasons we have made the choice to restrict the computation definition without any fusion, and to propose an automatic loop fusion described in Section~\ref{sect:parallelism}.
+Moreover, one can note that the usual \emph{scatter} optimization of numerical codes is actually a specific case of fusion.
 
@@
 \end{mydef}
-
-% This section has presented the general formalism of what is a computation kernel and a multi-stencil program.
-% The language presented in this paper, the Multi-Stencil Language, offers a simple way to describe a numerical simulation while being able to automatically extract parallelism.
-% Its aim is to be mesh-agnostic, as previously described, but also to make a clear separation of concerns between the simulation description and its implementation.
-% As a result, and as already described, details about the topology of the mesh, which is an implementation problem, is not used into the language.
-% Moreover, in a computation kernel $k(S,R,(w,D),comp)$, $comp$ is not used by the language.
 
@@
 
-One can note that all definitions given in this section are independent from the topology of the mesh.
+One can note that the meta-definition(~\ref{eq:msp}) is independent from the topology of the mesh.
 This property will be kept in the rest of this paper so that MSF has the capacity to be used for any kind of mesh, thus being mesh-agnostic.

### 2017-02-22 ARTICLES/IJPP/articleR1/msf.tex
unification article et corrections - parallelism.tex en cours

@@
 Component-based software engineering (CBSE) is a domain of software engineering~\cite{Szyperski:2002:CSB:515228} which improves code re-use, separations of concerns, and thus maintainability.
-An component-based application is made of a set of component instances linked together, this is also called a component \emph{assembly}.
+A component-based application is made of a set of component instances linked together, this is also called a component \emph{assembly}.
 A component is a black box that implements an independent functionality of the application, and which interacts with its environment only through well defined interfaces: its ports.
@@
 that component models can achieve the needed level of performance, and
-scalability while also helping in application portability~\cite{Bernholdt01052006, bigot:inria-00388508, UCHPC2015}
+scalability while also helping in application portability~\cite{Bernholdt01052006, bigot:inria-00388508, UCHPC2015}.
 
@@
 A \llc-based application is a static \emph{assembly} of components instances and the connections between their ports.
-Such an assembly is described in LAD, an XML dialect.
+Such an assembly is described in LAD, an XML dialect, and is managed by the \llc runtime system that minimize overheads by loading simple dynamic libraries.
+One can also notice that \llc can achieve performance if the granularity of components is high enough and attentively chosen by the user.
 
@@
 The multi-stencil domain will be formally defined in the next section.
-A multi-stencil program is typically a numerical simulation based on explicit numerical schemes that approximate PDEs equations with one or multiple neighborhood computations.
+A multi-stencil program is an approximation of PDEs equations by one or multiple numerical computations, some of them using neighborhood values around an element, also called a \emph{stencil} computation.
 
 Figure~\ref{fig:msf} presents the overview of the Multi-Stencil Framework that will be entirely detailed throughout this paper. It is composed of four different parts presented below.
-
 One can note in this figure that MSF is addressed to two different kinds of end-users: the \emph{numerician}, in other words the mathematician, and the \emph{developer}.
@@
 
-MSF also has the interesting capability to become flexible thanks to a third user possibility to interact with the framework: the High Performance Computing (HPC) specialist.
+MSF also has the interesting capability to be more flexible than existing solutions thanks to a possibility for a third party to interact with the framework.
+This third party is a High Performance Computing (HPC) specialist as illustrated in Fig.~\ref{fig:msf}.
 
@@
 \begin{center}
-  \includegraphics[width=.6\textwidth]{./images/msf.pdf}
+  \includegraphics[width=.4\textwidth]{./images/msf.pdf}
   \caption{The Multi-Stencil Framework (MSF) composed of the Multi-Stencil Language (MSL), the Generic Assembly (GA), the Multi-Stencil Compiler (MSC) and which produces a specialized assembly of components.}
@@
 \paragraph{\textbf{Generic Assembly}}
-In addition to the language MSL and the simulation description written by the numerician, MSF needs a generic assembly of a multi-stencil program as input.
-What is called a generic assembly is a component assembly for which meta-types are represented and for which some parts need to be generated or specialized.
-It is similar to a generic pattern or a skeleton in usual programming models such as C++ or functional models.
-From this generic assembly will be built the final specialized assembly of the simulation.
+In addition to the language MSL and the simulation description written by the numerician, MSF needs a Generic Assembly (GA) of a multi-stencil program as input.
+What is called a GA is a component assembly for which meta-types are represented and for which some parts need to be generated or specialized.
+A GA could be compared to a generic template or a skeleton respectively in usual object programming languages such as C++ or in functional languages.
+From this generic assembly will be built the final specialized assembly of the simulation where component types will be precised, and some parts of GA will be transformed or added.
 As well as MSL, this generic assembly is described in Section~\ref{sect:msl} and is built upon the meta-model described in Section~\ref{sect:formalism}.
-Moreover, an example of generic assembly is used in evaluations of Section~\ref{sect:eval}.
-\HC{Note idée : HPC specialist gives types for DDS and Data}
 
@@
 The core of the framework is the Multi-Stencil Compiler, or MSC.
-It is responsible for transforming the genric assembly into the final parallel assembly which is specific to the simulation described by the numerician.
-MSC is described in Sections~\ref{sect:parallelism} and~\ref{sect:msp}.
+It is responsible for transforming the genric assembly into the final parallel assembly which is specific to the simulation described by the numerician with MSL.
+MSC is described in Section~\ref{sect:parallelism}.
 
 \paragraph{\textbf{Specialized assembly}}
-Finally, the input of MSF is the component assembly generated by MSC.
-It is an instanciation and a transformation of the generic component assembly, by adding component types and by adding specific components generated by MSC.
-From this final component assembly which is specific to the simulation initially described by using MSL, the developer will finally write numerical codes.
+Finally, the output of MSF is the component assembly generated by MSC.
+It is an instanciation and a transformation of the generic component assembly, by adding component types, transforming some part of the assembly, and by adding specific components generated by MSC.
+From this final component assembly which is specific to the simulation initially described with MSL, the developer will finally write numerical codes.
+This final specialized component assembly represents the parallel computations orchestration of the simulation initially described by the numerician.

### 2017-02-22 ARTICLES/IJPP/articleR1/msl.tex
unification article et corrections - parallelism.tex en cours

@@
 %----------------------------------------
-From the meta-model represented by Equation~\ref{eq:msp} the generic assembly of any multi-stencil program can be built.
+From the meta-model represented by Equation (~\ref{eq:msp}) the Generic Assembly (GA) of any multi-stencil program can be built.
 It is represented by Figure~\ref{fig:ga}.
-As indicated, the GA has five different components described below.
+As illustrated, the GA has five different components described below.
 
@@
    \node[provide,right=0.15 of Tu] (Cp) {};
-   \node[component, draw=green!80!black, text=blue!80!black,right=2cm of Cp] (C) {$Computations$};
+   \node[component, draw=blue,fill=green!80!black,text=blue!80!black,right=2cm of Cp] (C) {$Computations$};
    \node[use,above=0.8cm of C] (Cu) {};
@@
    \node[provide,above=0.15 of Cu] (Datap2) {};
-   \node[component,double, draw=red,above=0.8cm of Datap2] (Data) {$Data$};
+   \node[component,double, draw=red,fill=red!50!white,above=0.8cm of Datap2] (Data) {$Data$};
    \node[use,above=0.8cm of Data] (Datau) {};
@@
    \node[provide,above=0.15 of Datau] (DDSp2) {};
-   \node[component, draw=red,above=0.8cm of DDSp2] (DDS) {$DDS$};
+   \node[component, draw=red, fill=red!50!white,above=0.8cm of DDSp2] (DDS) {$DDS$};
    \node[right=1.5cm of DDS] (m) {$\mathcal{M},\Phi,\mathcal{D},\mathcal{N}$};
@@
 \paragraph{\textbf{Driver}}
-This component can be compared to the main function of a usual program. It is responsible for initialization and execution of other components, just like variable initializations and function calls.
+This component can be compared to the main function of a usual program.
+It is responsible for both the initialization and the execution of other components (like variable initializations and function calls).
 
 \paragraph{\textbf{Time}}
-As indicated in the figure, this component is responsible for the time $T$ defined in Equation~\ref{eq:msp}. Thus it is composed of a time loop and eventually of a convergence reduction.
+As indicated in the figure, this component is responsible for the time $T$ defined in Equation (~\ref{eq:msp}). Thus it is composed of a time loop and eventually of a convergence reduction.
 
@@
 This component is responsible for the mesh and its entities $\mathcal{M}$, and $\Phi$, the computation domains $\mathcal{D}$ and the set of neighborhood functions $\mathcal{N}$.
-Moreover, if a data parallelization is needed, this component should handle mesh partitioning and should provide a synchronization interface as it will be detailed in Section~\ref{sect:parallelism}.
+This component is generic, thus not typed.
+When GA will be instantiated and specialized by MSC, a specific type of mesh will replace this DDS meta-component.
+However, needed interfaces of such a DDS component are well known, and any component answering these interfaces can be indifferently used by MSF.
+A third party can propose new DDS components.
+In this paper, both data and task parallelism are used.
+In the case of data parallelism this component should handle mesh partitioning and should provide a synchronization interface as it will be detailed in Section~\ref{sect:parallelism}.
 One can note that this component is red, this means that this component should be provided by a HPC specialist in MSF as illustrated in the previous Figure~\ref{fig:msf}.
@@
 This component is inseparable from the previous one $DDS$.
-It actually uses the DDS component to instanciate one quantity.
-Thus, it is responsible for the set of quantities $\Delta$.
+It actually uses the DDS component to instantiate one mesh variable.
+Thus, it is responsible for the set of mesh variables $\Delta$.
 This component is also red.
 As $DDS$ it should be provided by a HPC specialist.
-Moreover, it is represented by a double line box which means that it can be instanciated multiple time will others will be instanciated once.
-Actually, this component will be instanciated as many times as the number of quantities in the simulation.
+One can note that this component is represented by a double line box.
+This means that it can be instantiated multiple time while others will be instantiated once.
+Actually, this component will be instantiated as many times as the number of quantities in the simulation.
 
@@
 This component is green and blue because it will be partly generated by MSC and partly feeded by the developer.
-Actually, this component will be replaced by MSC by a sub-assembly of component for which the parallel part will be automatically generated, and for which components responsible for simulation code will be completed by the developer.
+Actually, this component will be replaced by a sub-assembly of component, produced by MSC, for which the parallel part will be automatically generated.
+In this generated sub-assembly, components responsible for numerical codes will be completed by the developer.
+This sub-assembly generation is described in Section~\ref{sect:parallelism}.
 
@@
 
-The second element of MSF which is built upon the meta-model represented by Equation~\ref{eq:msp} is the Multi-Stencil Language and its grammar.
-This grammar is sufficient (in addition to GA) for MSC to automatically extract a parallel pattern of the simulation which is finally dumped as a specialized instanciation of GA.
+The second element of MSF which is built upon the meta-model represented by Equation (~\ref{eq:msp}) is the Multi-Stencil Language and its grammar.
+This grammar is light and descriptive only, however it is sufficient (in addition to GA) for MSC to automatically extract a parallel pattern of the simulation, which is finally dumped as a specialized instantiation of GA.
 
@@
 The grammar of the Multi-Stencil Language is given in Figure~\ref{fig:grammar} and an example is provided in Figure~\ref{fig:exmsl}.
-A Multi-Stencil program is composed of eight parts as in Equation~ref{eq:msp}.
+A Multi-Stencil program is composed of eight parts as in Equation (~\ref{eq:msp}).
 
@@
 For example \texttt{cart} in Fig.\ref{fig:exmsl},~l.1.
-Since the language is independent of the mesh topology, this is not used by the compiler but is syntactic sugar for the user.%and presents for reference only.
+The language, based on the meta-model (~\ref{eq:msp}), is independent of the mesh topology, thus this identifier is actually not used by the compiler but is syntactic sugar for the user.
 
@@
 
-\item The \texttt{independent} keyword (Fig.\ref{fig:grammar},~l.4) offers a way to declare that computation domains do not overlap, such as \texttt{d1} and \texttt{d2} in  Fig.\ref{fig:exmsl},~l.7.
+\item The \texttt{independent} keyword (Fig.\ref{fig:grammar},~l.4) offers a way to declare that computation domains do not intersect, such as \texttt{d1} and \texttt{d2} in  Fig.\ref{fig:exmsl},~l.7.
 This is used by the compiler to compute dependencies between computations.
@@
 For each shape, the source and destination group of entities (Fig.\ref{fig:grammar},~l.16) are specified.
-For example \texttt{nec} in Fig.\ref{fig:exmsl},~l.11 associates \texttt{cell} entities to each \texttt{edgex} entity.
+For example \texttt{nec} in Fig.\ref{fig:exmsl},~l.11 is a neighborhood from \texttt{edgex} to \texttt{cell}.
 
-\item The \texttt{mesh quantities} keyword (Fig.\ref{fig:grammar},~l.6) introduces identifiers for $\delta\in\Delta$, the quantities with the group of entities they are mapped on (Fig.\ref{fig:grammar},~l.16).
-For example the quantities \texttt{C} and \texttt{H} are mapped on \texttt{edgex} entities.
+\item The \texttt{mesh quantities} keyword (Fig.\ref{fig:grammar},~l.6) introduces identifiers for $\delta\in\Delta$, the mesh variables with the group of entities they are mapped on (Fig.\ref{fig:grammar},~l.16).
+For example the quantities \texttt{C} and \texttt{H} are mapped onto \texttt{edgex} entities.
 
@@
 
-\item Finally, the last part (Fig.\ref{fig:grammar},~l.8) introduces the computations loops. Each loop is made of two parts:
+\item Finally, the last part (Fig.\ref{fig:grammar},~l.8) introduces the different computation loops of the simulation. Each loop is made of two parts:
 \begin{itemize}
-\item the \texttt{time} keyword (Fig.\ref{fig:grammar},~l.22) introduces $conv$, the convergence criteria either as a number of iterations or as ${0,1}$ valued scalar (Fig.\ref{fig:grammar},~l.24).
+\item the \texttt{time} keyword (Fig.\ref{fig:grammar},~l.22) introduces $conv$, the convergence criteria either as a constant number of iterations or as a scalar variable (Fig.\ref{fig:grammar},~l.24).
 For example, 500 iterations are specified in Fig.\ref{fig:exmsl},~l.16,
@@
  \item the quantity $w$ written and its domain $D$, for example in Fig.\ref{fig:exmsl},~l.22, kernel \texttt{k4} computes the quantity \texttt{F} on domain \texttt{d1},
- \item the read scalars $S$ and mesh quantities with their associated stencil shape $R$, for example in Fig.\ref{fig:exmsl},~l.16, \texttt{k4} reads \texttt{C} with the shape \texttt{nce} and \texttt{D} with the default \textit{identity} shape, it uses no scalar.
+ \item the read scalars $S$ and mesh quantities with their associated stencil shape $R$, for example in Fig.\ref{fig:exmsl},~l.16, \texttt{k4} reads \texttt{C} with the shape \texttt{nce} and \texttt{D} with the default \textit{identity} shape, it does not read scalars.
  \end{itemize}
@@
 One can also notice that the computation to execute for each kernel is not specified.
-This is indeed not handled by MSL, which is mesh-agnostic, and which only generates a skeleton of the application (a parallel pattern of the application).
-Actually, a numerical computation is naturally mesh-dependent and is left to other languages.
-The numerical computation is specified elsewhere by referencing the identifier chosen here.
-
-% To give a better idea of a multi-stencil program, a short example is given in Figure~\ref{fig:exmsl}.
-In this example, the mesh is called \textit{cart}.
-Two groups of mesh entities are defined \textit{cell} and \textit{edgex}.
-As previously explained, it is not needed in the grammar to give details on the topology of the mesh and the group of mesh entities.
-Two computation domains are given, \textit{d1} is a subpart of the entities of the group \textit{cell}, and \textit{d2} is a subpart of the entities of the group \textit{edgex}.
-It is declared that \textit{d1} and \textit{d2} are independant.
-Three stencil shapes are given.
-The second one for example \textit{nce} returns mesh entities of the group \textit{edgex} from an entity of the group \textit{cell}.
-Eight quantities are applied onto the group \textit{cell}, and two onto the group \textit{edgex}.
-Two scalars are defined, \textit{mu} and \textit{tau}.
-The time loop will iterate 500 times.
-Finally, nine computations are declared.
-For example, the computation \textit{k8} write the quantity \textit{J} onto the computation domain \textit{d1} by reading the scalar \textit{mu} and the quantity \textit{I} with a stencil shape \textit{n1}.
-
-% The aim of the Multi-Stencil Language is to offer a way to describe a numerical simulation based on explicit numerical schemes.
-The language is mesh-agnostic as no information is asked to the user about the actual topology of the mesh.
-Moreover, information that is usefull for the implementation is also split from the MSL description.
-For this reason, the language grammar is simple.
-
-This section has introduced the grammar of MSL.
-The next section will show how parallelization can be extracted from this simple language and how an empty parallel skeleton (or pattern) of the application can be generated.
-%, introducing implementation concerns in a second phase.
+Only an identifier is given to each kernel, for example $k_4$.
+The numerical code is indeed not handled by MSL, which is mesh-agnostic, and which is used to generate a parallel orchestration of computations only.
+The numerical computation is specified after MSC compilation by the developer (Fig.~\ref{fig:msf}).

### 2017-02-22 ARTICLES/IJPP/articleR1/parallelism.tex
modif related work + minor changes
unification article et corrections - parallelism.tex en cours

--- 2016-05-30 ARTICLES/IJPP/article/parallelism.tex
@@
-% Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.
-As previously explained, in a computation $k(S,R,(w,D),comp)$, $comp$ is not handled by MSL.
+As previously explained, in a computation $k(S,R,(w,D),comp)$, $comp$ is not handled by MSF.
+Numerical codes are left in a second step to the developer by using the $DDS$ component selected during the MSF compilation.
 As a result, in the rest of this paper, and to simplify notations, we denote the same computation $k(S,R,(w,D))$.
@@
 \label{sect:dataparal}
-In a data parallelization technique, the idea is to split quantities on which the program is computed into balanced sub-parts, one for each available resource.
-The same sequential program can afterwards be applied on each sub-part simultaneously, with some additioinal synchronizations between resources to update the data not computed locally, and thus to guarantee a correct result.
+In a data parallelization technique, the idea is to split data, or quantities, on which the program is computed into balanced sub-parts, one for each available resource.
+The same sequential program can afterwards be applied on each sub-part simultaneously, with some additional synchronizations between resources to update the data not computed locally, and thus to guarantee a correct result.
 
@@
 consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
-This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper.
+This step can be performed by an external graph partitionner~\cite{Pellegrini:1996:SSP:645560.658570,DBLP:conf/ieeehpcs/HeleneS13,lachat:hal-00768916} and is addressed by the third party DDS component instantiated from GA.
 
-As entities and quantities are mapped onto the mesh, the set of groups of mesh entities and the set of quantities $\Delta$ are partitionned the same way than the mesh: $\Phi=\{\Phi_0,\dots,\Phi_{p-1}\}$, $\Delta=\{\Delta_0,\dots,\Delta_{p-1}\}$.
+As entities and quantities are mapped onto the mesh, the set of groups of mesh entities and the set of quantities $\Delta$ are partitioned the same way than the mesh: $\Phi=\{\Phi_0,\dots,\Phi_{p-1}\}$, $\Delta=\{\Delta_0,\dots,\Delta_{p-1}\}$.
 
@@
 
-Actually, a synchronization is needed by the quantity read by a stencil computation (not local), if this quantity has been modified before, which means that it has been written before.
+Actually, a synchronization is needed by the quantity read by a stencil computation (not local), if this quantity has been written before.
 This synchronization is needed because a neighborhood function $n \in \mathcal{N}$ of a stencil computation involves values computed on different resources.
@@
 \begin{mydef}
-For $n$ the number of computations in $\Gamma$ and $j<n$, if $\exists (r_j,n_j) \in R_j$ such that $n_j\neq identity$ and for all $i<j$, $k_i \not \pprec k_j$, a \textit{synchronization} is needed between $k_l$ and $k_j$, where $j<l<n$, denoted $k_l \pprec k_j$, if $w_l=r_j$.
+For $n$ the number of computations in $\Gamma$ and $j<n$, if $\exists (r_j,n_j) \in R_j$ such that $n_j\neq identity$ and such that for all $i<j$, $k_i \not \pprec k_j$, a \textit{synchronization} is needed between $k_l^{t-1}$ and $k_j^t$, where $j<l<n$, denoted $k_l^{t-1} \pprec k_j^{t}$, if $w_l=r_j$.
 The quantity to synchronize is $\{w_l\}$.
@@
 In other words, $w_i$ has to be synchronized for the neighborhood $n_j$ for all entities of $D_j$.
+\label{def:sync3}
 \end{mydef}
@@
 A reduction kernel $k_j(S_j,R_j,(w_j,D_j))$, where $w$ is a scalar, is replaced by $k^{red}_j(S_j,R_j,(w_j,D_j))$.
-%if we denote by $w^r$, $0 \leq r<p$, the local scalar $w$ computed on the resource $r$, a reduction synchronization is defined as the specific computation
-% \begin{equation*}
-% k_{j}^{sync}(S,R,(w,D),comp)
-% \end{equation*}
 \label{def:red}
 \end{mydef}
-% where, $S=\emptyset$, $R=\{(w^0,entity(w^0)) \dots (w^{p-1},entity(w^{p-1}))\}$, and $w=w_i$, $D=entity(w)=D_i$ and $comp=comp_i$.
 
@@
 In this algorithm, lines 7 to 19 apply Definition~(\ref{def:sync}), lines 20 to 29 apply Definition~(\ref{def:sync2}), and finally lines 34 and 35 apply Definition~(\ref{def:red}).
-Finally, line 39 of the algorithm is the recursive call.
+Finally, line 37 of the algorithm is the recursive call.
 
@@
 \begin{equation*}
-\Gamma = [k_0,k_1,k_2,k_3,k_4,k_0,k_6,k_7,k_8]
+\Gamma = [k_0,k_1,k_2,k_3,k_4,k_5,k_6,k_7,k_8]
 \end{equation*}
@@
 \begin{equation}
-\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_0,k_6,k_7,k_{7;8}^{sync},k_8],
+\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_5,k_6,k_7,k_{7;8}^{sync},k_8],
 \label{eq:exsync}
@@
 Such a dependency graph can directly be given to a dynamic scheduler, or can statically be scheduled.
-In this paper, we introduce task parallelism by building the dependency graph between kernels of the sequential list $\Gamma_{sync}$.
-Thus, as $\Gamma_{sync}$ takes into account data parallelism, we introduce hybrid parallelism.
+In this paper, we consider a computation kernel as a task and we introduce task parallelism by building the dependency graph between kernels of the sequential list $\Gamma_{sync}$.
+Thus, as $\Gamma_{sync}$ already takes into account data parallelism, we introduce hybrid parallelism.
 
@@
 However, a specific condition on the computation domain, due to the specific domain of multi-stencils, is introduced for the write after write case.
+One can note that the \texttt{independent} keyword of Fig.~\ref{fig:grammar} is useful in this case.
 
@@
 
-From an ordered list of computations $\Gamma_{sync}$, a directed dependency graph $\Gamma_{dep}(V,A)$ can be built finding all pairs of computations $k_i$ and $k_j$, with $i<j$, such that $k_i \prec_{raw} k_j$ or $k_i \prec_{waw} k_j$ or $k_i \prec_{war} k_j$.
+From the ordered list of computations $\Gamma_{sync}$ and from the MSL description, a directed dependency graph $\Gamma_{dep}(V,A)$ can be built finding all pairs of computations $k_i$ and $k_j$, with $i<j$, such that $k_i \prec_{raw} k_j$ or $k_i \prec_{waw} k_j$ or $k_i \prec_{war} k_j$.
 
@@
 From the synchronized ordered list of computation kernels $\Gamma_{sync}$, the dependency graph of the computations $\Gamma_{dep}(V,A)$ is obtained from the call $F_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function defined in Algorithm~\ref{alg:dep}.
-
-% \begin{equation*}
-% F_{dep}(\Gamma_{sync},j) = 
-% \begin{cases} 	\bullet (\{\},\{\}) \mbox{ if }j=|\Gamma_{sync}|\\
-% 				\bullet (k_j, \{(\overset{\frown}{k_i,k_j})\mbox{, }\forall i < j \mbox{, } k_i\prec k_j \})\\
-% 				\text{ } \qquad \cup F_{dep}(\Gamma_{sync},j+1) \mbox{ if }j<|\Gamma_{sync}|
-% \end{cases}
-% \end{equation*}
 \end{mydef}
@@
 
-% \begin{proof}
-% $\Gamma_{dep}$ is built from $\Gamma_{sync}$ which is an ordered and sequential list of computations.
-Moreover, each computation of the list $\Gamma_{sync}$ is associated to a vertex of $V$, even if the same computation is represented more than once in $\Gamma_{sync}$.
-As a result it is not possible to go back to a previous computation and to create a cycle.
-% \end{proof}
-
 As a result of the hybrid parallelization, each resource $0 \leq r \leq p-1$ perform a multi-stencil program, defined by
@@
 From $\Gamma_{sync}$ that has been built in Equation~(\ref{eq:exsync}), the dependency DAG can be built.
-For example, as $k_0$ computes $B$ and $k_1$ reads $B$, $k_0$ and $k_1$ becomes vertices of $\Gamma_{dep}$, and an arc $(\overset{\frown}{k_0,k_1})$ is added to $\Gamma_{dep}$.
+For example, as $k_4$ computes $F$ and $k_6$ reads $F$, $k_4$ and $k_6$ becomes vertices of $\Gamma_{dep}$, and an arc $(\overset{\frown}{k_4,k_6})$ is added to $\Gamma_{dep}$.
 The overall $\Gamma_{dep}$ built from the call to $F_{dep}(\Gamma_{sync},0)$ is drawn in Figure~\ref{fig:depdep}.
+By building synchronizations as defined in Definitions (\ref{def:sync}), (\ref{def:sync2}) and (\ref{def:sync3}), dependencies are respected.
+For example, $k_{0;1}^{sync}$ read and write $B$ which guarantees that $k_{0;1}^{sync}$ is performed after $k_0$ and before $k_1$.
 \begin{figure}[h!]
@@
 \end{figure}
+
+%--------------------
+\subsection{Static scheduling}
+\label{sect:tsp}
+
+In this section we detail a static scheduling of $\Gamma_{dep}$ by using minimal series-parallel directed acyclic graphs.
+Such a static scheduling may not be the most efficient one, but it offers a simple fork/join task model which make possible the design of a performance model.
+Moreover, such a scheduling offers a simple way to propose a fusion optimization.
+
+In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of Minimal Series-Parallel DAGs (MSPD).
+Such a graph can be decomposed as a serie-parallel tree, denoted $TSP$, where each leaf is a vertex of the MSPD it represents, and whose internal nodes are labelled $S$ or $P$ to indicate whether the two sub-trees form a sequence or parallel composition.
+Such a tree can be considered as a fork-join model and as a static scheduling.
+An example is given in Fig.~\ref{fig:solu}.
+
+Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or subgraph, called $N$, such that the following property is verified :
+
+\begin{myth}
+The transitive reduction of a DAG $G$ is MSPD if and only if it does not contain $N$ as an induced subgraph.
+\end{myth}
+
+As $\Gamma_{dep}$ is a DAG, by removing N-Shapes it is transformed to a MSPD.
+The intuition is illustrated in Fig.~\ref{fig:over}.
+Considering the figure without the dashed line, the sub graph forms a "N" shape.
+The fact is that this shape cannot be represented as a composition of sequences or parallel executions.
+To remove such forbidden N-shapes of $\Gamma_{dep}=(V,E)$, we have chosen to apply an over-constraint with the relation $k_0 \prec k_3$, such that a complete bipartite graph is created for the sub-dag as illustrated in Figure~\ref{fig:over}.
+By adding this arc to the DAG, it is possible to identify its execution as $sequence(parallel(k_0;k_2);parallel(k_1;k3))$ represented by the TSP tree of Fig.~\ref{fig:solu}.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (c0) at (0,0) {$k_0$};
+   \node[] (bc0) at (-1,0) {};
+   \node[] (c1) at (2,0) {$k_1$};
+   \node[] (c2) at (0,-1) {$k_2$};
+   \node[] (bc2) at (-1,-1) {};
+   \node[] (c3) at (2,-1) {$k_3$};
+
+  \path[->]
+    (bc0) edge [dotted] node {} (c0)
+    (bc2) edge [dotted] node {} (c2)
+    (c0) edge node {} (c1)
+          edge [dashed] node [swap] {} (c3)
+    (c2) edge node {} (c1)
+        edge node {} (c3);
+  \end{tikzpicture}
+\caption{Over-constraint on the forbidden $N$ shape.}
+\label{fig:over}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[] (S0) at (0,0) {$\mathcal{S}$};
+   \node[] (bS0) at (-1,-1) {};
+   \node[] (P1) at (-1,1) {$\mathcal{P}$};
+   \node[] (aP1) at (-1.5,2) {$k_0$};
+   \node[] (aaP1) at (-0.5,2) {$k_2$};
+   \node[] (P2) at (1,1) {$\mathcal{P}$};
+   \node[] (c1) at (0.5,2) {$k_1$};
+   \node[] (c3) at (1.5,2) {$k_3$};
+
+  \path[->]
+    (bS0) edge [dotted] node {} (S0)
+    (S0) edge node {} (P1)
+         edge node {} (P2)
+    (P1) edge node [swap] {} (aP1)
+          edge node [swap] {} (aaP1)
+    (P2) edge node {} (c1)
+        edge node {} (c3);
+  \end{tikzpicture}
+\caption{TSP tree of Fig.~\ref{fig:over}.}
+\label{fig:solu}
+\end{center}
+\end{figure}
+
+After these over-constraints are applied, $\Gamma_{dep}$ is MSPD.
+Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have proposed a linear algorithm to know if a DAG is MSPD and, if it is, to decompose it to its associated binary decomposition tree.
+As a result, the binary tree decomposition algorithm of Valdes \& Al can be applied on $\Gamma_{dep}$ to get the $TSP$ static scheduling of the multi-stencil program.
+
+\paragraph{\textbf{Example}} From $\Gamma_{dep}$ illustrated in Fig.~\ref{fig:depdep} the TSP tree represented in Fig.~\ref{fig:tree} can be computed.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   %\node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
+   %\node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
+
+   \node[] (s8) at (4.5,0.5) {$\mathcal{S}$};
+   \node[] (s9) at (3,1.5) {$\mathcal{S}$};
+
+   %reduction c0 c1
+   \node[] (s0) at (1.5,2.5) {$\mathcal{S}$};
+   \node[] (s1) at (1,3.5) {$\mathcal{S}$};
+   \node[] (c0) at (0.5,4.5) {$k_0$};
+   \node[] (star1) at (1.5,4.5) {$k_{0;1}^{sync}$};
+   \node[] (c1) at (2,3.5) {$k_1$};
+
+   %reduction c7 c8
+   \node[] (s2) at (7.5,1.5) {$\mathcal{S}$};
+   \node[] (s3) at (7,2.5) {$\mathcal{S}$};
+   \node[] (c7) at (6.5,3.5) {$k_7$};
+   \node[] (star8) at (7.5,3.5) {$k_{7;8}^{sync}$};
+   \node[] (c8) at (8,2.5) {$k_8$};
+
+   \node[] (p1) at (4.5,2.5) {$\mathcal{P}$};
+   %reduction c3 c5
+   \node[] (s5) at (5.5,3.5) {$\mathcal{S}$};
+   \node[] (c3) at (5,4.5) {$k_3$};
+   \node[] (c5) at (6,4.5) {$k_0$};
+   %reduction c2 c6
+   \node[] (s6) at (3.5,3.5) {$\mathcal{S}$};
+   \node[] (p0) at (2.8,4.5) {$\mathcal{P}$};
+   \node[] (c2) at (2.3,5.5) {$k_2$};
+   \node[] (star4) at (3.3,5.5) {$k_{1;4}^{sync}$};
+   \node[] (s7) at (4.2,4.5) {$\mathcal{S}$};
+   \node[] (c4) at (3.8,5.5) {$k_4$};
+   \node[] (c6) at (4.8,5.5) {$k_6$};
+
+  \path[->]
+    %(c0s) edge node {} (c8d)
+
+    (s8) edge node {} (s9)
+         edge node {} (s2)
+    (s9) edge node {} (p1)
+         edge node {} (s0)
+
+    %reduction c0 c1
+    (s0) edge node {} (s1)
+         edge node {} (c1)
+    (s1) edge node {} (c0)
+         edge node {} (star1)
+    %reduction c7 c8
+    (s2) edge node {} (s3)
+         edge node {} (c8)
+    (s3) edge node {} (c7)
+        edge node {} (star8)
+
+    (p1) edge node {} (s6)
+         edge node {} (s5)
+    %reduction c3 c5
+    (s5) edge node {} (c3)
+         edge node {} (c5)
+    %reduction c2 c6
+    (s6) edge node {} (p0)
+         edge node {} (s7)
+    %reduction c2 *4
+    (p0) edge node {} (c2)
+         edge node {} (star4)
+    (s7) edge node {} (c4)
+         edge node {} (c6);
+  \end{tikzpicture}
+\caption{Serie-Parallel tree decomposition of the example of program of Figure~\ref{fig:exmsl}}
+\label{fig:tree}
+\end{center}
+\end{figure}
+
+%--------------------
+\subsection{Fusion optimization}
+\label{sect:fusion}
+
+Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization.
+Even though the MSL language is not dedicated to produce very optimized stencil codes, but to produces the parallel pattern of the application, building the $TSP$ tree make available an easy optimization when the data parallelization technique is the only one used.
+This optimization consists in proposing a valid merge of some computation kernels inside a single space loop.
+As a result, the user can use this valid fusion of kernels or not when implementing those.
+
+Those fusions can be computed from the canonical form of the $TSP$ tree decomposition. The canonical form consists in recursively merging successive $S$ vertices or successive $P$ vertices of $TSP$.
+
+The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and where $k_{i;j}^{fus}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.
+Finally, $type(k)$ returns $comp$ if the kernel is a computation kernel, and $sync$ otherwise.
+
+\begin{algorithm}
+\caption{$F_{fus}$}
+\label{alg:fus}
+\begin{algorithmic}[1]
+\Procedure{$F_{fus}$} {$TSP(V,E)$}
+\For {$(k_i,k_j) \in V^2$}
+\If {parent($k_i$)==parent($k_j$)}
+\If {$type(k_i)==type(k_j)==comp$}
+\If {parent($k_i$)==$S$}
+\If {$D_i==D_j$}
+\State propose the fusion $k_{i;j}^{fus}$
+\EndIf
+\ElsIf {parent($k_i$)==$P$}
+\If {$D_i==D_j$ and $R_i \cap R_j \neq \emptyset$}
+\State propose the fusion $k_{i;j}^{fus}$
+\EndIf
+\EndIf
+\EndIf
+\EndIf
+\EndFor
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs for example~\cite{spaaTangCKLL11}.
+However, this fusion takes place at a different level and can bring performance improvments as it will be illustrated in Section~\ref{sect:eval}.
+This fusion algorithm relies on very simple statements:
+\begin{itemize}
+\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependant.
+As a result, what is written by the first one is read by the second one.
+Thus, at least one data is common to those computations (the one written by $k_i$).
+Thus, if the computation domains verify $D_i=D_j$, the fusion of $k_i$ and $k_j$ will decrease cache misses.
+\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependant.
+However, if the computation domains verify $D_i=D_j$, and if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}^{fus}$.
+\end{itemize}
+
+%--------------------
+\subsection{Performance model}
+\label{sect:perfs}
+
+In this subsection are introduced two performance model, one for the data parallelization technique, and one for the hybrid data and task parallelization technique, both previously explained.
+
+The performance model of a data parallelization technique is inspired from the Bulk Synchronous Parallel model.
+Actually data parallelization technique consider that each process has its own part of the domain.
+Thus the performance model reveals that the computation time is the sum of the reference sequential time divided by the number of processes, and of the time spent in communications between processes.
+Thus, for
+\begin{itemize}
+\item $T_{SEQ}$ the sequential reference time, 
+\item $P$ the total number of processes, 
+\item $T_{COM}$ the communications time, 
+\end{itemize}
+the total computation time is
+\begin{equation}
+T = \frac{T_{SEQ}}{P} + T_{COM}.
+\end{equation}
+
+Thus, when the number of processes $P$ increase in data parallelization, the performance model limit is $T_{COM}$
+\begin{equation}
+\lim\limits_{P \rightarrow +\infty} T = T_{COM}.
+\end{equation}
+
+As a result, the critical point for performance is when $T_{COM} \geq \frac{T_{SEQ}}{P}$, which happens naturally in data parallelization as $T_{COM}$ will increase with the number of processes, and $\frac{T_{SEQ}}{P}$ decrease with the number of processes.
+
+This limitation is always true, but can be delayed by different strategies.
+First, it is possible to perform communications through the network while computations are performed simultaneously.
+Second, it is possible to introduce another kind of parallelization, task parallelization.
+Thus, for the same total number of processes, only a part of them are used for data parallelization, and the rest are used for task parallelism.
+As a result, $\frac{T_{SEQ}}{P}$ will continue to decrease but $T_{COM}$ will increase later.
+This second strategy is the one studied in the following hybrid performance model.
+
+For an hybrid (data and task) parallelization technique, and for
+\begin{itemize}
+\item $P_{data}$ the number of processes used for data parallelization,
+\item $P_{task}$ the number of processes used for task parallelization,
+\item such that $P = P_{data} \times P_{task}$ is the total number of processes used,
+\item $T_{task}$ the overhead time due to task parallelization technique,
+\item and $F_{task}$ the task parallelization degree of the application,
+\end{itemize}
+the total computation time is
+\begin{equation}
+T = \frac{T_{SEQ}}{P_{data} \times F_{task}} + T_{COM} + T_{task}
+\end{equation}
+
+The time overhead due to task parallelization can be represented as the time spent to create a pool of threads and the time spent to synchronize those threads. Thus, for
+\begin{itemize}
+\item $T_{cr}$ the total time to create the pool of threads (may happened more than once), 
+\item $T_{sync}$ the total time spent to synchronize threads, 
+\end{itemize}
+the overhead is
+\begin{equation*}
+T_{task} = T_{cr} + T_{sync}.
+\end{equation*}
+
+The task parallelization degree of the application $F_{task}$ is the limitation of a task parallelization technique.
+As explained before, a task parallelization technique is based on the dependency graph of the application.
+Thus, this dependency graph must expose enough parallelism for the number of available threads.
+For this performance model we consider that
+\begin{equation*}
+F_{task} = P_{task}.
+\end{equation*}
+However, the upper bound of $F_{task}$ is constrained by the dependency graph of the application and by the time spent in each task.
+
+\medskip
+As a result when $P_{data}$ is small a data parallelization technique may be more efficient, while an hybrid parallelization could be interesting at some point to improve performance.
+The question asked here is when is it interesting to use hybrid parallelization.
+
+To answer this question lets consider the two parallelization techniques, data only and hybrid. We denote
+\begin{itemize}
+\item $P_{data1}$ the total number of processes entirely used by the data only parallelization,
+\item $P_{data2}$ the number of processes used for data parallelization in the hybrid parallelization,
+\item and $P_{task}$ the number of processes used for task parallelization in the hybrid parallelization,
+\item such that $P_{data1} = P_{data2} \times P_{task}$.
+\end{itemize}
+
+We search the point where the data parallelization is less efficient than the hybrid parallelization. Thus, 
+\begin{equation*}
+\frac{T_{SEQ}}{P_{data1}} + T_{COM1} \geq \frac{T_{SEQ}}{P_{data2} \times P_{task}} + T_{COM2} + T_{task}.
+\end{equation*}
+
+This happens when
+\begin{equation}
+T_{COM1} \geq T_{COM2} + T_{task}
+\label{eq:hyb}
+\end{equation}
+
+This performance model will be validated and will help to explain results the Section~\ref{sect:eval}.

### 2017-02-23 ARTICLES/IJPP/articleR1/evaluation.tex
restructuration du plan + figure MSF
modif related work + minor changes
corrections + evaluation en cours

--- 2016-05-31 ARTICLES/IJPP/article/evaluation.tex
@@
-This section evaluates the MSL compiler and the code it generates.
-We first evaluate the generated code in the case where only data parallelization and fusion are applied and compare this to a full SkelGIS program.
-Then, we analyze a case taking advantage of the full the hybrid parallelization combining data and task parallelism and validate the performance model proposed in Section~\ref{sect:tsp}.
+This section first presents the implementation details chosen to evaluate MSF in this paper.
+Then, the studied use case is presented.
+Once this context is clear the MSC performance is evaluated as well as both available parallelization techniques, data and task.
+Finally, the performance model proposed in Section~\ref{sect:perfs}.
+
+%-------------------------------------
+\subsection{Implementation details}
+
+The two main implementation choices to take in MSC and its specialized assembly generation concern the technologies used for data and task parallelizations.
+
+For the data-parallelization, as already detailed many times throughout the paper, a third party HPC specialist is responsible for implementing $DDS$ and $Data$ using a chosen library or external language and by following the interfaces specified by these two components.
+To evaluate MSF, we have played the role of HPC specialists and have implemented these components using SkelGIS, a C++ embedded DSL~\cite{CPE:CPE3494} that proposes a distributed Cartesian mesh as well as user API to manipulate structures while hiding their underlying distribution.
+
+For the task parallelism, we have chosen to use OpenMP~\cite{660313} to generate the code of the $Scheduler$ component.
+OpenMP targets shared-memory platforms only.
+Although the version 4 of OpenMP has introduced explicit support for dynamic task scheduling, our implementation only requires version 3 whose fork-join model is well suited for the static scheduling introduced in Section~\ref{sect:parallelism}.
+The use of dynamic schedulers, such as provided by libgomp\footnote{\url{https://gcc.gnu.org/projects/gomp/}}, StarPU~\cite{Augonnet2011}, or XKaapi~\cite{Gautier:2013:XRS:2510661.2511383},  to directly execute the DAG $\Gamma_{dep}$ will be the subject of future works.
+
+As a result, the MSL compiler generates a hybrid code which uses both SkelGIS and OpenMP.
+It also generates empty components where the developer must provide local sequential implementations of the kernels. In our case these kernels are written using SkelGIS API.
+
+%-------------------------------------
+\subsection{Use case description}
 
 All evaluations presented in this section are based on a real case study of the shallow-Water Equations as solved in the FullSWOF2D\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}~\cite{Ferrari2004,CPE:CPE3494} code from the MAPMO laboratory, University of Orl\'eans.
-We have developed a MSL version of FullSWOF2D that contains 3~mesh entities, 7~computation domains, 48~data and 98~computations (32~stencil kernels and 66~local kernels).
+In 2013, a full SkelGIS implementation of this use case has been performed by numericians and developers of the MAPMO laboratory~\cite{CPE:CPE3494,cordier2013fullswof,coullon:hal-00832660}.
+From this implementation we have kept the code of computation kernels to directly use it into $K$ components.
+Compared to a full SkelGIS implementation, where synchronizations and fusions are handled manually, MSF automatically compute where synchronizations are needed and how to perform a fusion without errors.
+To evaluate MSF on this use case we have described the FullSWOF2D simulation by using MSL.
+FullSWOF2D contains 3~mesh entities, 7~computation domains, 48~data and 98~computations (32~stencil kernels and 66~local kernels).
+Performances of the obtained implementation are compared to the full SkelGIS implementation to show that no overheads are introduced by MSF.
 
 %-------------------------------------
-\subsection{Compiler}
+\subsection{Compiler evaluation}
 
-Table~\ref{fig:exectime} illustrates the execution time of each step of the MSL compiler for the FullSWOF2D example.
+Table~\ref{fig:exectime} illustrates the execution time of each step of MSC for the FullSWOF2D example.
 This has been computed on a laptop with a dual-core Intel Core i5 1.4~GHz, and 8~GB of DDR3.
@@
 As a matter of fact, the complexity of the algorithm for N-shapes removal is $O(n^3)$.
-The replacement of the static scheduling by a direct scheduling of the DAG by dedicated tools should solve this in the future.
+If this complexity is not a problem at the moment and onto this use case it could become one for just-in-time compilation or bigger simulations.
+The replacement of the static scheduling by a dynamic scheduling using dedicated tools (such as OpenMP 4, StarPU etc.) should solve this in the future.
 
@@
  \begin{tabular}{|c|c|c|c|c|}
-  \hline
    Step & Parser & $\Gamma_{sync}$ & $\Gamma_{dep}$ & $TSP$\\
@@
    \% & 0.022 & 0.043 & 0.09 & 86.6\\
-   \hline
  \end{tabular}
@@
 %-------------------------------------
-\subsection{Data parallelism}
+\subsection{Data parallelism evaluation}
 
@@
  \begin{tabular}{|c|c|}
-   \hline
     Cluster & \textbf{TGCC Curie Thin Nodes}\\
@@
     MPI & Bullxmpi\\
-    Network & fat-tree Infiniband\\
-    \hline
  \end{tabular}
@@
 The evaluations have been performed on the Curie supercomputer (TGCC, France) described in Table~\ref{tab:TGCC}.
-Each evaluation has been performed ten times and the median is presented in results.
+Each evaluation has been performed nine times and the median is presented in results.
 
@@
 %-------------------------------------
-\subsection{Hybrid parallelism}
+\subsection{Hybrid parallelism evaluation}
 

### 2017-02-23 ARTICLES/IJPP/articleR1/msl.tex
corrections + evaluation en cours

@@
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (D) at (0,0) {$Driver$};
+   \node[component,fill=green!80!black] (D) at (0,0) {$Driver$};
    \node[provide] (Dp) at (-1,0) {};
@@
    \node[provide,below=0.15 of Du2] (Tp) {};
-   \node[component,below=1.6cm of Tp] (T) {$Time$};
+   \node[component,below=1.6cm of Tp,fill=green!80!black] (T) {$Time$};
    \node[use,right=1cm of T] (Tu) {};
@@
    \node[provide,right=0.15 of Tu] (Cp) {};
-   \node[component, draw=blue,fill=green!80!black,text=blue!80!black,right=2cm of Cp] (C) {$Computations$};
+   \node[component, draw=blue!60!white,fill=green!80!black,text=blue!80!black,line width=0.5mm,right=2cm of Cp] (C) {$Computations$};
    \node[use,above=0.8cm of C] (Cu) {};

### 2017-02-23 ARTICLES/IJPP/articleR1/parallelism.tex
corrections + evaluation en cours

@@
 For this reason, to perform the final reduction, a set of synchronizations are needed to get the final reduced scalar.
-As most parallelism libraries (MPI, OpenMP) already propose a reduction synchronization with its own optimizations, we simply choose to replace the reduction computation by itself anotated by $red$.
+As most parallelism libraries (MPI, OpenMP) already propose a reduction synchronization with its own optimizations, we simply choose to replace the reduction computation by itself annotated by $red$.
 
@@
 \begin{mydef}
-For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{read after write} dependency, denoted $k_i \prec_{raw} k_j$, if $\exists (r_j,n_j) \in R_j$ such that $w_i=r_j$.
+For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependent from $k_i$ with a \emph{read after write} dependency, denoted $k_i \prec_{raw} k_j$, if $\exists (r_j,n_j) \in R_j$ such that $w_i=r_j$.
 In this case, $k_i$ has to be computed before $k_j$.
@@
 \begin{mydef}
-For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{write after write} dependency, denoted $k_i \prec_{waw} k_j$, if $w_i = w_j$ and $D_i \cap D_j \neq \emptyset$.
+For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependent from $k_i$ with a \emph{write after write} dependency, denoted $k_i \prec_{waw} k_j$, if $w_i = w_j$ and $D_i \cap D_j \neq \emptyset$.
 In this case, $k_i$ also has to be computed before $k_j$.
@@
 \begin{mydef}
-For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependant from $k_i$ with a \emph{write after read} dependency, denoted $k_i \prec_{war} k_j$, if $\exists (r_i,n_i) \in R_i$ such that $w_j=r_i$.
+For two computations $k_i$ and $k_j$, with $i < j$, it is said that $k_j$ is dependent from $k_i$ with a \emph{write after read} dependency, denoted $k_i \prec_{war} k_j$, if $\exists (r_i,n_i) \in R_i$ such that $w_j=r_i$.
 In this case, $k_i$ also has to be computed before $k_j$ is started so that values read by $k_i$ are relevant.
@@
 In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of Minimal Series-Parallel DAGs (MSPD).
-Such a graph can be decomposed as a serie-parallel tree, denoted $TSP$, where each leaf is a vertex of the MSPD it represents, and whose internal nodes are labelled $S$ or $P$ to indicate whether the two sub-trees form a sequence or parallel composition.
+Such a graph can be decomposed as a serie-parallel tree, denoted $TSP$, where each leaf is a vertex of the MSPD it represents, and whose internal nodes are labeled $S$ or $P$ to indicate whether the two sub-trees form a sequence or parallel composition.
 Such a tree can be considered as a fork-join model and as a static scheduling.
@@
 
-Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or subgraph, called $N$, such that the following property is verified :
+Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or sub-graph, called $N$, such that the following property is verified :
 
 \begin{myth}
-The transitive reduction of a DAG $G$ is MSPD if and only if it does not contain $N$ as an induced subgraph.
+The transitive reduction of a DAG $G$ is MSPD if and only if it does not contain $N$ as an induced sub-graph.
 \end{myth}
@@
 Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization.
-Even though the MSL language is not dedicated to produce very optimized stencil codes, but to produces the parallel pattern of the application, building the $TSP$ tree make available an easy optimization when the data parallelization technique is the only one used.
+Even though the MSL language is not dedicated to produce very optimized independent stencil codes, but to produce the parallel orchestration of computations, building the $TSP$ tree make available an easy optimization when the data parallelization technique is the only one used.
 This optimization consists in proposing a valid merge of some computation kernels inside a single space loop.
-As a result, the user can use this valid fusion of kernels or not when implementing those.
+This is called a fusion.
+As previously explained in Section~\ref{sect:formalism}, MSL restrict the definition of a numerical computation by writing a single quantity at a time which is more natural for numericians, and which avoid errors in manual fusion or counter-productive fusions for task parallelization.
+MSF guarantee that proposed fusions are correct and will not cause errors in the final results of the simulation.
 
@@
 
-The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and where $k_{i;j}^{fus}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.
+The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where the $parent(k)$ function returns the parent vertex of $k$ in the tree, and where $k_{i;j}^{fus}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.
 Finally, $type(k)$ returns $comp$ if the kernel is a computation kernel, and $sync$ otherwise.
@@
 \State propose the fusion $k_{i;j}^{fus}$
+\Else
+\If {$\exists n:D_i \rightarrow D_j \in \mathcal(N)$ and $\bigcup_{\phi \in D_i} n(\phi) = D_j$}
+\State propose the fusion $k_{i;j}^{scatter}$
+\EndIf
 \EndIf
@@
 We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs for example~\cite{spaaTangCKLL11}.
-However, this fusion takes place at a different level and can bring performance improvments as it will be illustrated in Section~\ref{sect:eval}.
-This fusion algorithm relies on very simple statements:
+However, this fusion takes place at a different level and can bring performance improvements as it will be illustrated in Section~\ref{sect:eval}.
+This fusion algorithm relies on two statements:
 \begin{itemize}
-\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependant.
+\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependent.
 As a result, what is written by the first one is read by the second one.
-Thus, at least one data is common to those computations (the one written by $k_i$).
+Thus, $w_i$ the quantity written by $k_i$ is common to these computations.
 Thus, if the computation domains verify $D_i=D_j$, the fusion of $k_i$ and $k_j$ will decrease cache misses.
-\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependant.
+\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependent.
 However, if the computation domains verify $D_i=D_j$, and if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}^{fus}$.
 \end{itemize}
+These two cases are illustrated by Fig.~\ref{fig:fus1} and Fig.~\ref{fig:fus2}.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto,every text node part/.style={align=center}]
+   \node[] (S0) at (0,0) {$\mathcal{P}$};
+   \node[] (ki) at (-1,-1) {$k_i$\\ $[D_i]$};
+   \node[] (kj) at (1,-1) {$k_j$\\ $[D_j]$};
+   \node[] (fake1) at (2,-0.5) {};
+   \node[] (fake2) at (4,-0.5) {};
+   \node[] (S1) at (5,0) {$\mathcal{P}$};
+   \node[] (kij) at (5,-1) {$k_{i;j}^{fus}$\\ $[D_i]$};
+  \path[->]
+    (S0) edge node {} (ki)
+         edge node {} (kj)
+    (fake1) edge [bend left] node {$D_i=D_j$} (fake2)
+    (S1) edge node {} (kij);
+\end{tikzpicture}
+\caption{First fusion case.}
+\label{fig:fus1}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto,every text node part/.style={align=center}]
+   \node[] (S0) at (0,0) {$\mathcal{S}$};
+   \node[] (ki) at (-1,-1) {$k_i$\\ $[D_i]$};
+   \node[] (kj) at (1,-1) {$k_j$\\ $[D_j]$};
+   \node[] (fake1) at (2,-0.5) {};
+   \node[] (fake2) at (4,-0.5) {};
+   \node[] (S1) at (5,0) {$\mathcal{S}$};
+   \node[] (kij) at (5,-1) {$k_{i;j}^{fus}$\\ $[D_i]$};
+  \path[->]
+    (S0) edge node {} (ki)
+         edge node {} (kj)
+    (fake1) edge [bend left] node {$D_i=D_j$\\ $R_i \cap R_j \neq \emptyset$} (fake2)
+    (S1) edge node {} (kij);
+\end{tikzpicture}
+\caption{Second fusion case.}
+\label{fig:fus2}
+\end{center}
+\end{figure}
+
+One additional case of fusion can also happen. It is close to the first one above but is a little more difficult:
+Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependent.
+As a result, what is written by the first one is read by the second one.
+The construction of the tree also guarantee that synchronizations are not needed between these computations, otherwise a $k^{sync}$ would have been inserted between them (inherited from $\Gamma_{sync}$).
+Thus, $w_i$ the quantity written by $k_i$ is common to these computations.
+Considering the following:
+\begin{itemize}
+\item $D_i \neq D_j$, which means that loop fusion is by default not possible,
+\item $(r_j,n_j)$ is the pair read by $k_j$ for which $r_j=w_i$ and for which $n_j:D_j \rightarrow D_i^m$
+\end{itemize}
+the fusion of $k_i$ and $k_j$ is possible if and only if $\exists n:D_i \rightarrow D_j \in \mathcal(N)$ such that
+\begin{equation*}
+\bigcup_{\phi \in D_i} n(\phi) = D_j
+\end{equation*}
+This means that even if domains are different, a loop fusion is possible if an adequate neighborhood function can be found.
+One can note that this particular fusion case is equivalent to a \emph{scatter} optimization, often used when using unstructured meshes.
+One can also note that the computation $k_j$ will be written in a different manner if a scatter fusion is performed or not.
+This is not a problem by using MSF as the fusion is proposed before the developer actually write the numerical code of $k_j$.
+The developer will be notified of fusions in the output of MSF.
+This particular case is illustrated in Fig.~\ref{fig:fus3}.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto,every text node part/.style={align=center}]
+   \node[] (S0) at (0,0) {$\mathcal{S}$};
+   \node[] (ki) at (-1,-1) {$k_i$\\ $[D_i]$};
+   \node[] (kj) at (1,-1) {$k_j$\\ $[D_j]$};
+   \node[] (fake1) at (2,-0.5) {};
+   \node[] (fake2) at (4,-0.5) {};
+   \node[] (S1) at (5,0) {$\mathcal{S}$};
+   \node[] (kij) at (5,-1) {$k_{i;j}^{scatter}$\\ $[D_i]$};
+  \path[->]
+    (S0) edge node {} (ki)
+         edge node {} (kj)
+    (fake1) edge [bend left] node {$D_i \neq D_j$\\ $\exists n:D_i \rightarrow D_j$\\ $\bigcup_{\phi \in D_i} n(\phi) = D_j$} (fake2)
+    (S1) edge node {} (kij);
+\end{tikzpicture}
+\caption{Third fusion case.}
+\label{fig:fus3}
+\end{center}
+\end{figure}
+
+%--------------------
+\subsection{Overall compilation process}
+\label{sect:ovcompil}
+
+MSC takes a MSL file written following the grammar described in Section~\ref{sect:msl}, as well as the Generic Assembly of components represented in Fig.~\ref{fig:ga} as inputs, and generates a specialized component assembly that manages the parallel orchestration of the computations of the simulation.
+In this final assembly, that could be compared to a pattern or a skeleton of the simulation, the developer still has to fill the functions corresponding to the various computation kernels by using the DDS instantiation chosen into the specialized assembly.
+The overall behavior of the compiler is as follow: %presented in Algorithm~\ref{alg:compiler}.
+\begin{enumerate}
+ \item it parses the MSL input file and generates $\Gamma$, the list of computation kernels,
+ \item from $\Gamma$, it builds $\Gamma_{sync}$, the list including synchronizations for data parallelism using Algorithm $F_{sync}$ introduced in Section~\ref{sect:parallelism},
+ \item from $\Gamma_{sync}$, it builds $\Gamma_{dep}$, the DAG supporting hybrid parallelism using Algorithm $F_{dep}$ introduced in Section~\ref{sect:parallelism},
+ \item it then removes the N-Shapes from $\Gamma_{dep}$ to get a MSPD graph, and generates its serie-parallel binary tree decomposition $TSP$,
+ \item it performs the fusion of kernels in $TSP$ if required (data parallelization only),
+ \item it transforms GA to generate its output specialized component assembly.
+\end{enumerate}
+
+The last step of this compilation process is detailed below. It is composed of four steps:
+\begin{enumerate}
+ \item it instantiates $DDS$ and $Data$ components by using components implemented by a third party HPC specialist,
+ \item it generates $K$ components responsible for each computation kernel of the simulation,
+ \item it generates a new $Scheduler$ component,
+ \item it replaces $Computations$ component by a generated sub-assembly that matches $TSP$ by using $Scheduler$, $K$ and $Sync$ components.
+ \end{enumerate}
+
+% detailed each component (interfaces)
+New components have been introduced above and need to be explained.
+A $K$ component is a component into which the developer will write numerical code.
+It could represents a single computation kernel described by the numerician using MSL, or it could represents the fusion of multiple computation kernels.
+In any case the name of the generated component will use kernel identifiers used in the MSL description.
+A $K$ kernel is composed of $m$ use ports that are used to be connected to the $m$ quantities needed by the computation (\ie the numerical code).
+The component is also equipped with a provide port to be connected to the $Scheduler$ component.
+A generic $K$ component is represented in Fig.~\ref{fig:k}.
+
+A $Sync$ component is a static component (not generated) composed of a use-multiple port which is used to ask synchronizations to all quantities it is linked to ($Data$).
+The component is also equipped with a provide port to be connected to the $Scheduler$ component.
+The $Sync$ component is represented in Fig.~\ref{fig:sync}.
+
+Finally, the $Scheduler$ component is the component responsible for implementing the $TSP$ computed by MSC.
+Thus, this component represents the specific parallel orchestration of  computations.
+It is a heavy component with as much use ports as the number of $K$ components to call (\ie computations and fusions of computations).
+The component is also equipped with a provide port to be connected to the $Time$ component.
+A generic $Scheduler$ component is represented in Fig.~\ref{fig:ass}.
+
+\begin{figure}[t]
+\begin{center}
+\subfloat[][$K$\label{fig:k}]{
+\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
+   \node[component] (seq) at (0,0) {$K$}; 
+   \node[provide] (p) at (-1,0) {}; 
+   \node[use] (u) at (1,0) {}; 
+   \node[] (star) at (1,0.5) {$*$};
+  \path[-]
+    (p) edge node {} (seq)
+    (seq) edge node {} (u);
+\end{tikzpicture}
+}
+\hspace{\fill}
+\subfloat[][\label{fig:sync}$Sync$]{
+\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
+   \node[component] (seq) at (0,0) {$Sync$};
+   \node[provide] (p) at (-1,0) {};
+   \node[use] (u) at (1,0) {$m$};
+
+  \path[-]
+    (p) edge node {} (seq)
+    (seq) edge node {} (u);
+\end{tikzpicture}
+}
+\hspace{\fill}
+\subfloat[][\label{fig:sched}$Scheduler$]{
+\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
+   \node[component] (sync) at (0,0) {$Scheduler$};
+   \node[provide] (p) at (-1.5,0) {};
+   \node[use] (u) at (1.5,0) {};
+   \node[] (star) at (1.5,0.5) {$*$};
+  \path[-]
+    (p) edge node {} (sync)
+    (sync) edge node {} (u);
+\end{tikzpicture}
+}
+\caption{Specific components used to transform GA to the specialized component assembly of the simulation.}
+\label{fig:specass}
+\end{center}
+\end{figure}
+
+% draw the sepcialized assembly of the example
+To illustrate how a specialized assembly is generated, the specialized assembly of the example that has been used throughout this paper is represented by Fig.~\ref{specass}.
+As this assembly is heavy to entirely draw, only a sub part of it is represented by the figure.
+Moreover some connections are represented by numbers instead of lines to not complexity the read.
+
+\begin{figure}[t]
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+   \node[component,fill=green!80!black] (D) at (0,0) {$Driver$};
+   \node[provide] (Dp) at (-1,0) {};
+   \node (Ds) at (-1.5,0) {start};
+   \node[use,right=1.5cm of D] (Du1) {};
+   \node[use,below=1.75cm of D] (Du2) {};
+   \node[use,below=0.8cm of Du1] (Du3) {$m$};
+
+   \node[provide,below=0.15 of Du2] (Tp) {};
+   \node[component,below=1.6cm of Tp,fill=green!80!black] (T) {$Time$};
+   \node[use,right=1cm of T] (Tu) {};
+
+   \node[provide,right=0.15 of Tu] (Cp) {};
+   \node[component, fill=green!80!black,right=2cm of Cp] (C) {$Scheduler$};
+   \node[use,right=1.8 of C] (scu1) {};
+   \node[use,below=0.2 of scu1] (scu2) {};
+
+   \node[provide,below right=0.2 of Du3] (Datap1) {};
+   \node[component,draw=red,fill=red!50!white,above=0.8cm of Datap2] (Data) {$A$};
+   \node[component,draw=red,fill=red!50!white,below=0.8cm of Data] (Data2) {$B$};
+   \node[below=0.5cm of Data2] (Data3) {...};
+   \node[use,above=0.8cm of Data] (Datau) {};
+   \node[provide,right=1cm of Data] (mpidata) {};
+   \node[right=0.2 of mpidata] (u1) {1};
+   \node[provide,right=1cm of Data2] (mpidata3) {};
+   \node[right=0.2 of mpidata3] (u2) {2};
+   \node[provide,below=0.2cm of mpidata3] (mpidata2) {};
+   \node[right=0.2 of mpidata2] (u3) {3};
+
+   \node[provide,right=0.15 of Du1] (DDSp1) {};
+   \node[provide,above=0.15 of Datau] (DDSp2) {};
+   \node[component, draw=red, fill=red!50!white,above=0.8cm of DDSp2] (DDS) {cart};
+
+   \node[provide,right=0.15 of scu1] (pk0) {};
+   \node[provide,right=0.15 of scu2] (psync1) {};
+   \node[component, fill=blue!60!white,right=2cm of pk0] (k0) {$k_0$};
+   \node[use,right=0.8 of k0] (k0u) {};
+   \node[right=0.2 of k0u] (1) {1};
+   \node[use,below=0.2 of k0u] (k0u2) {};
+   \node[right=0.2 of k0u2] (1) {2};
+   \node[component, fill=blue!60!white,below=0.8cm of k0] (sync1) {$Sync(0,1)$};
+   \node[use,right=1.5 of sync1] (syncu) {};
+   \node[right=0.2 of syncu] (1) {3};
+   \node[below=0.5cm of sync1] (next) {...};
+
+  \path[-]
+    (Dp) edge node {} (D)
+    (D) edge node {} (Du1)
+        edge node {} (Du2)
+        edge node {} (Du3)
+    (DDSp1) edge node {} (DDS)
+    (Tp) edge node {} (T)
+    (T)  edge node {} (Tu)
+    (Cp) edge node {} (C)
+    (C) edge node {} (scu1)
+    	edge node {} (scu2)
+    (Datap1) edge node {} (Data.west)
+    		edge node {} (Data2.west)
+    (Data) edge node {} (Datau)
+          edge node {} (mpidata)
+     (Data2) edge node {} (mpidata3)
+     		edge node {} (mpidata2)
+    (DDSp2) edge node {} (DDS)
+    (pk0) edge node {} (k0)
+    (psync1) edge node {} (sync1)
+    (k0) edge node {} (k0u)
+    	edge node {} (k0u2)
+    (sync1) edge node {} (syncu)
+    (Data2.west) edge [bend left] node {} (Datau);
+    %(k0p.east) edge [bend right] node {} (mpidata.east)
+    %(k0p2.east) edge [bend right] node {} (mpidata2.east)
+    %(syncp.east) edge [bend right] node {} (mpidata3.east);
+\end{tikzpicture}
+\vspace*{.5em}
+\caption{Generic Assembly according to the Multi-Stencil program formalism.}
+\label{fig:specass}
+\end{center}
+\end{figure}
 
@@
 
-In this subsection are introduced two performance model, one for the data parallelization technique, and one for the hybrid data and task parallelization technique, both previously explained.
+In this subsection are introduced two performance models, one for the data parallelization technique, and one for the hybrid data and task parallelization technique, both previously explained.
 
@@
 This limitation is always true, but can be delayed by different strategies.
-First, it is possible to perform communications through the network while computations are performed simultaneously.
+First, it is possible to overlap communications by performing computations simultaneously.
 Second, it is possible to introduce another kind of parallelization, task parallelization.
@@
 \item $P_{data}$ the number of processes used for data parallelization,
-\item $P_{task}$ the number of processes used for task parallelization,
-\item such that $P = P_{data} \times P_{task}$ is the total number of processes used,
+\item $P_{task}$ the number of processes used for task parallelization, such that $P = P_{data} \times P_{task}$ is the total number of processes used,
 \item $T_{task}$ the overhead time due to task parallelization technique,
@@
 \begin{equation*}
-F_{task} = P_{task}.
+F_{task} = P_{task}, 
 \end{equation*}
-However, the upper bound of $F_{task}$ is constrained by the dependency graph of the application and by the time spent in each task.
+however, as it will be illustrated in Section~\ref{sect:eval} $F_{task}$ is more difficult to establish.
+Actually, the lower and upper bounds of $F_{task}$ are constrained by the dependency graph of the application.
 
@@
 As a result when $P_{data}$ is small a data parallelization technique may be more efficient, while an hybrid parallelization could be interesting at some point to improve performance.
-The question asked here is when is it interesting to use hybrid parallelization.
-
-To answer this question lets consider the two parallelization techniques, data only and hybrid. We denote
+The question is: when is it interesting to use hybrid parallelization ?
+This paper does not propose an intelligent system to answer this question automatically, however, it offers a way to understand how to answer the question.
+To answer this question lets consider the two parallelization techniques, data only and hybrid.
+We denote
 \begin{itemize}
@@
 
-This performance model will be validated and will help to explain results the Section~\ref{sect:eval}.
+This performance model will be validated and will help to explain results of Section~\ref{sect:eval}.

### 2017-02-23 ARTICLES/IJPP/articleR1/evaluation.tex
en cours evaluation.tex

@@
    Time (ms) & 1 & 2 & 4.2 & 3998.5\\
-   \hline
    \% & 0.022 & 0.043 & 0.09 & 86.6\\
@@
 
-In this evaluation, we disable task-parallelism (fix its degree to 1) to focus on data-parallelism.
-In the implementation evaluated in this paper, the data parallelism is handled using SkelGIS that has already been evaluated in comparison with a native MPI implementation for the FullSWOF2D example~\cite{CPE:CPE3494}.
-In addition to SkelGIS, MSL automatically inserts the required MPI synchronizations and identifies possible fusion of kernels to reduce cache misses.
+In this evaluation, we disable task-parallelism to focus on data-parallelism.
+Two versions of code are compared in this section: first a plain SkelGIS implementation of FullSWOF2D, where synchronizations and fusions are handled manually; second, the automatic synchronizations and fusions handled by MSF over SkelGIS.
+SkelGIS has already been evaluated in comparison with a native MPI implementation for the FullSWOF2D example~\cite{CPE:CPE3494}.
+For this reason, this section uses the plain SkelGIS implementation as the reference version.
+This enables to evaluate both the choices made by MSC as well as the potential overheads of using \llc~\cite{l2c} that is not used in the plain SkelGIS version.
+The evaluations have been performed on the Curie supercomputer (TGCC, France) described in Table~\ref{tab:TGCC}.
+Each evaluation has been performed nine times and the median is presented in results.
 
@@
  \begin{tabular}{|c|c|}
-    Cluster & \textbf{TGCC Curie Thin Nodes}\\
+     & TGCC Curie Thin Nodes\\
      \hline         
@@
 
-This section compares the performance of the code produced by MSL for FullSWOF2D with a plain SkelGIS program where synchronizations and fusion choices have been done by the developer without any automatic support.
-This enables to evaluate both the choices made by the compiler as well as the potential overheads of using \llc~\cite{l2c} that is not used in the plain SkelGIS version.
-The evaluations have been performed on the Curie supercomputer (TGCC, France) described in Table~\ref{tab:TGCC}.
-Each evaluation has been performed nine times and the median is presented in results.
+Figures~\ref{fig:weak1}, \ref{fig:weak2} and~\ref{fig:weak3} respectively show weak scalings for a computation domain of $400 \times 400$ cells by core, $600 \times 600$ cells by core and $800 \times 800$ cells by core from 16 cores to 16.384 cores, as resumed in Table~\ref{tab:weak}.
+
+\begin{table}[!ht]
+\begin{center}
+ \begin{tabular}{|c|c|}
+    Domain size per core & Number of iterations\\
+    \hline
+     $400 \times 400$ & $200$\\
+     $600 \times 600$ & $200$\\
+     $800 \times 800$ & $200$\\
+ \end{tabular}
+ \caption{\label{tab:weak}Weak scaling experiments of Fig.~\ref{fig:weak1}, Fig.~\ref{fig:weak2} and Fig.~\ref{fig:weak3}.}
+ \end{center}
+\end{table}
 
@@
 \begin{minipage}{.475\textwidth}
-  \includegraphics[width=\textwidth]{../results/weak_scaling/400/median_weak.pdf}
-  \caption{weak-scaling with $400 \times 400$ domain per core.}
+  \includegraphics[width=\textwidth]{../results/weak_scaling/400_200/median_weak.pdf}
+  \caption{weak-scaling with $400 \times 400$ domain per core and $200$ time iterations.}
   \label{fig:weak1}
@@
 \begin{minipage}{.475\textwidth}
-  \includegraphics[width=\textwidth]{../results/weak_scaling/800/median_weak.pdf}
-  \caption{weak-scaling with $800 \times 800$ domain per core.}
+  \includegraphics[width=\textwidth]{../results/weak_scaling/600_200/median_weak.pdf}
+  \caption{weak-scaling with $600 \times 600$ domain per core and $200$ time iterations.}
   \label{fig:weak2}
 \end{minipage}\end{figure}
-%In a weak scaling, a constant amount of work is given to each core while increasing the overall number of cores.
-%As a result, the overall domain size also increase with the number of cores.
-%A weak scaling is a good evaluation to detect overheads which are independent from the amount of computation performed by each core, such as collective communications, or overheads introduced by \llc components, for example.
-Figures~\ref{fig:weak1} and~\ref{fig:weak2} show a weak scaling for a $400 \times 400$ points by core and a $800 \times 800$ points by core from 16 cores to 16.384 cores.
-Minimum and maximum values are also shown as error bars in figures.
 
-One can notice that the MSL code produces a slightly better execution time on the $400 \times 400$ domain and a small overhead on the $800 \times 800$ domain.
-This might be explained by slightly different compiler optimization flags since \llc requires compilation as dynamic libraries with the \emph{-fpic} compilation flag while the SkelGIS version does not.
+\begin{figure}[t]
+\begin{center}
+  \includegraphics[width=.6\textwidth]{../results/weak_scaling/800_200/median_weak.pdf}
+  \caption{weak-scaling with $800 \times 800$ domain per core and $200$ time iterations.}
+  \label{fig:weak3}
+ \end{center}
+\end{figure}
+
+From these results, one can notice first that performances of MSF are very close to the reference version using plain SKelGIS.
+This is a very good result which shows first that MSC performs good synchronizations and fusions, and second that overheads introduced by \llc are limited thanks to a good granularity in GA.
+However, it seems that a slightly drop of performance happens when the domain size per core increases.
+This performance decrease is really small though, with a maximum difference between the two versions of 2.83\%.
+
+The only noticeable difference between the two versions are due to \llc which load dynamic libraries at runtime.
+Because of this particularity, components of \llc are compiled with the \emph{-fpic} compilation flag while the SkelGIS version does not.
 This flag can have slight positive or negative effect on code performance depending on the situation and might be responsible for the observed difference.
-In any case, these difference are barely noticeable and often included in the error bar of the measure and both version globally perform similarly.
-The MPI synchronizations performed by MSL perform as good as manually specified ones, and \llc does not introduce any noticeable overhead.
+
+\HC{another benchmark to show that with the static version of \llc the problem disappear ?}
 
@@
   \includegraphics[width=.6\textwidth]{../results/strong_scaling/10K_1K/median_strong.pdf}
-  \caption{Strong scaling on a $10k \times 10k$ domain.}
+  \caption{Strong scaling on a $10k \times 10k$ domain and $1000$ time iterations.}
   \label{fig:strong}
 \end{center}\end{figure}
-% The second evaluation is a strong scaling evaluation.
-% Unlike a weak scaling, a strong scaling keeps a constant overall size of problem, while the number of cores increase.
-% As a result, the sub-domain computed by each core reduces while the number of cores increases.
-% The strong scaling is a good additional evaluation to the weak scaling, showing different kinds of overheads, such as cache misses (when the size of sub-domain becomes small enough to fit into cache memory), or small constant overheads.
+
 Figure~\ref{fig:strong} shows the number of iteration per second for a 10k$\times$10k global domain size from 16 to 16.384 cores.
-The ideal strong scaling is also illustrated.
+The total number of time iterations for this benchmark is $1000$.
+In addition to the reference SkelGIS version, the ideal strong scaling is also illustrated in the figure.
 
-First, one can notice that the strong scaling evaluated for the code generated by MSL is close to the ideal speedup up to 16.384 cores, which is a very good result.
-Moreover, no overheads are introduced by MSL which shows that automatic synchronization detections and automatic fusion detections are the same one that the one written manually into the SkelGIS code of FullSWOF2D.
+First, one can notice that the strong scaling evaluated for the MSF version is close to the ideal speedup up to 16.384 cores, which is a very good result.
+Moreover, no overheads are introduced by MSF which shows that automatic synchronizations and automatic fusions are the same one that the one written manually into the plain SkelGIS version.
 Finally, no overheads are introduced by components of \llc.
@@
 
-\paragraph{\textbf{Fusion}} Finally, to evaluate the data parallelization technique automatically introduced by MSL, the fusion optimization is evaluated.
-Figure~\ref{fig:fusion} shows the number of iterations per second as a function of the number of cores, for FullSWOF2D with and without fusion optimization and onto a $500 \times 500$ domain.
-As explained in Section~\ref{sect:fusion}, the MSL fusion happens at a high level and is most of the time done naturally by a computer scientist.
-However, for a non computer scientist which write its numerical codes, an automatic proposition of such fusions makes the implementation easier.
+\paragraph{\textbf{Fusion}} Finally, we evaluate loop fusions automatically proposed by MSF from the $TSP$ tree of computation kernels.
+Figure~\ref{fig:fusion} shows the number of iterations per second as a function of the number of cores with and without fusions.
+This benchmark is performed on FullSWOF2D onto a $500 \times 500$ domain size with $200$ time iterations.
+As explained in Section~\ref{sect:fusion}, the MSF loop fusion happens at a high level and is most of the time done naturally by a computer scientist.
+However, for a non computer scientist which write its numerical codes, an automatic proposition of such fusions avoid errors, particularly for a parallel execution.
 Moreover, one can notice that the performance is clearly improved (around 40\%) by this fusion.
@@
   \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/fusVSbase.pdf}
-  \caption{Strong scaling on a 500x500 domain, with and without the fusion optimization.}
+  \caption{Strong scaling on a 500x500 domain size with $200$ time iterations, with and without fusions proposed by MSF.}
   \label{fig:fusion}
 \end{center}\end{figure}
+
+\HC{another benchmark with and without fusion fro task parallelization ?}
 
@@
 
+In this section, we introduce task parallelism to evaluate the hybrid parallelization offered by MSF. The MSF implementation evaluated in this paper relies on SkelGIS and OpenMP.
+
+The series-parallel tree decomposition $TSP$ of this simulation, extracted by MSC, is composed of 17 nodes labeled as sequence $\mathcal{S}$ and 18 nodes labeled as parallel $\mathcal{P}$.
+
+We define the \emph{level of parallelism} as the number of parallel tasks inside a fork of the fork/join model.
+
+Table~\ref{fig:freq} represents the number of time a given level of parallelism, \ie the number of tasks to perform concurrently, is observed in the final tree.
+One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1).
+Moreover, a level of 16~parallel tasks is reached two times, and five times for the fourth level.
+This means that if two threads are dedicated to task parallelism, two parts of the code at least will not take advantage of this and that no part of the code would benefit from more than 16 threads.
+The task parallelism alone is therefore clearly not enough to take advantage of even a single node of a cluster that typically supports today between 16 and 32 threads (16 for Curie).
+
 \begin{table}[th]
@@
  \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-    \hline 
    Level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
@@
    Frequency & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
-   \hline
  \end{tabular}
@@
 \end{table}
-
-In this evaluation, we introduce task parallelism to evaluate the hybrid parallelization offered by MSL that mixes data parallelization and task parallelization.
-In the implementation evaluated in this paper and in addition to SkelGIS, the hybrid parallelization relies on OpenMP.
-
-The series-parallel tree decomposition $TSP$ of this simulation, extracted by the MSL compiler, is composed of 17~sequence nodes and 18~parallel nodes.
-Table~\ref{fig:freq} represents the number of time a given level of parallelism, \ie the number of tasks to perform concurrently, is observed in the final tree.
-One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1).
-Moreover, a level of 16~parallel tasks is reached two times, and five times for the fourth level.
-This means that if two threads are dedicated to task parallelism, two parts of the code at least will not take advantage of this and that no part of the code would benefit from more than 16 threads.
-The task parallelism alone is therefore clearly not enough to take advantage of even a single node of a cluster that typically supports today between 16 and 32 threads (16 for Curie).
 

### 2017-02-24 ARTICLES/IJPP/articleR1/conclusion.tex
evaluation

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% unifier texte avec nouvelle version article
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We have presented the Multi-Stencil Language (MSL), a domain specific language for stencil-based numerical simulations.
+%We have presented the Multi-Stencil Language (MSL), a domain specific language for stencil-based numerical simulations.
 Compared to other existing stencil DSLs, MSL proposes a higher abstraction level for end-users.
@@
 
-In this paper, the formal definition of a multi-stencil program has been given.
+%In this paper, the formal definition of a multi-stencil program has been given.
 This formalism helps to understand what is defined into the light grammar of MSL, and also how parallelization can be extracted from it.
@@
 
-Experiments presented in this paper show that the back-end code produced by MSL, \ie by a mesh-agnostic language, does not introduce overheads up to 16.384 cores, and that the hybrid (data and task) parallelism automatically introduced by the language, which combines SkelGIS and OpenMP as a back-end, improves performance compared to a data parallelization only.
+E%xperiments presented in this paper show that the back-end code produced by MSL, \ie by a mesh-agnostic language, does not introduce overheads up to 16.384 cores, and that the hybrid (data and task) parallelism automatically introduced by the language, which combines SkelGIS and OpenMP as a back-end, improves performance compared to a data parallelization only.
 Thus, with the same code asked to the user, two different parallel versions of the simulation can be obtained.
 
-However, if performances are improved by the hybrid parallelization, this performance is also limited by the scheduling strategy used in this paper.
+%However, if performances are improved by the hybrid parallelization, this performance is also limited by the scheduling strategy used in this paper.
 Actually, as described in Section~\ref{sect:msp}, a static series-parallel (fork-join) scheduling is performed in this work.
@@
 
-Finally, to validate the separation of concerns introduced by MSL, a second back-end for unstructured meshes could be interesting.
+%Finally, to validate the separation of concerns introduced by MSL, a second back-end for unstructured meshes could be interesting.
 We think about PaMPA~\cite{lachat:hal-00768916} as a distributed data structure backend, instead of SkelGIS.

### 2017-02-24 ARTICLES/IJPP/articleR1/evaluation.tex
evaluation

@@
 
-We define the \emph{level of parallelism} as the number of parallel tasks inside a fork of the fork/join model.
-
-Table~\ref{fig:freq} represents the number of time a given level of parallelism, \ie the number of tasks to perform concurrently, is observed in the final tree.
-One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1).
-Moreover, a level of 16~parallel tasks is reached two times, and five times for the fourth level.
-This means that if two threads are dedicated to task parallelism, two parts of the code at least will not take advantage of this and that no part of the code would benefit from more than 16 threads.
-The task parallelism alone is therefore clearly not enough to take advantage of even a single node of a cluster that typically supports today between 16 and 32 threads (16 for Curie).
+We define the \emph{level of parallelism} as the number of parallel tasks inside one fork of the fork/join model.
+The fork/join model obtained for FullSWOF2D is composed of 18 fork phases (corresponding to $\mathcal{P}$ nodes of $TSP$).
+Table~\ref{fig:freq} represents the number of time a given level of parallelism is obtained inside fork phases.
 
@@
 
+One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1).
+Moreover, a level of 16~parallel tasks is reached two times, and five times for the fourth level.
+This means that if two cores are dedicated to task parallelism, the two sequential parts of the code will not take advantage of these two cores, and that no part of the code would benefit from more than 16 cores.
+The task parallelism, as proposed in this paper (\ie where each kernel is a task) is therefore not insufficient to take advantage of a single node of modern clusters that typically supports more than 16 cores.
+
 \begin{figure}[bh]\begin{center}
@@
 
-On the other hand, Figure~\ref{fig:limit} illustrates limitations of data parallelization technique alone.
-While the computation time decreases linearly with the number of core used, the communication behavior is much more erratic.
-Between 2 and 16 cores, the communication happen inside a single node and the time is small and nearly constant.
-There is a small oscillation that might be explained by difference between number of processors that are power of 4, where the communications are all the same, and those that are nota power of 4, where there is an imbalance.
-Starting with 32 cores, there are multiple nodes and the communication time is typically modeled as $L+S/B$ where $L$ is the latency, $S$ the data size and $B$ the bandwidth.
+On the other hand, Fig.~\ref{fig:limit} illustrates limitations of data parallelization technique alone.
+In this figure is represented the execution time (with a logarithmic scale) of FullSWOF2D while increasing the number of cores for a fix domain size of $500 \times 500$ with a total of $200$ time iterations (\ie this is a strong scaling).
+The speedup of this same benchmark is represented in blue in Fig.~\ref{fig:close}.
+One can note that the scaling is not as good is the one presented in Fig.~\ref{fig:strong}.
+The main difference between these two benchmarks is the domain size.
+In the benchmark of Fig.~\ref{fig:strong} the domain size is $10k \times 10k$ which means that using $2^8$ cores, for example, each core has to compute only a $31 \times 31$ sub-domain.
+On the other hand using $2^8$ cores in Fig.~\ref{fig:close} each core has to compute a $625 \times 625$ sub-domain.
+Fig.~\ref{fig:limit} shows why the speedup is not as good as the one with a bigger domain size.
+
+Actually, in this figure, while the computation time (in blue) decreases linearly with the number of core used, the communication behavior (in red) is much more erratic.
+Between 2 and 16 cores, communications are performed inside a single node thus the time is small and nearly constant.
+There is a small oscillation that might be explained by the partitioning differences.
+SkelGIS performs a two dimensional partitioning strategy.
+For this reason a smaller number of bytes are communicated using 2 cores than using 4, and using 8 cores than using 16 cores.
+Starting from 32 cores, each node is fully used and more than one node is used.
+From this point thus the communication time is typically modeled as $L+S/B$ where $L$ is the latency, $S$ the data size and $B$ the bandwidth.
 This explains the decrease of time from 32 to 128 cores where the data sizes communicated by each process decreases.
-The increased observed after that might be due to the fact that with the increased number of processes the fat-tree becomes deeper and the latencies increase.
+The increased observed after$128$ cores that might be due to the fact that with the increased number of processes the fat-tree becomes deeper and the latencies increase.
 
@@
 As a result, the data parallelism alone fails to provide enough parallelism to leverage the whole machine and other sources of parallelism have to be found.
-The blue curve of Figure~\ref{fig:close} shows the total execution time for the same example as in Figure~\ref{fig:limit} and as expected, the speedup bends down from 256 to 2048 cores.
+As expected, in  Figure~\ref{fig:close} the speedup bends down from 256 to 2048 cores.
+The same problem would happened in previous experiment of Fig.~\ref{fig:strong}, however as the domain size is larger, the phenomena appears with much more cores.
 
-In addition to the blue curve, Figure~\ref{fig:close} shows strong scalings for the same example but using an hybrid parallelization.
-For example, the purple curve shows the parallelization which uses 8 cores, to perform the task parallelization, for each process used for data parallelization (\ie MPI process).
+As task parallelism fails to scale from 16 cores, and as data parallelism also fails to scale when the communication cost overpass the execution time, an hybrid parallelization strategy is proposed by MSF and evaluated below.
+
+In addition to the blue curve, Fig.~\ref{fig:close} shows speedups for the same example ($500 \times 500$ domain with $200$ iterations) but using an hybrid parallelization.
+Fig.~\ref{fig:close} shows a comparison with 2, 4, 8 and 16 cores per MPI process for task parallelization.
+
+For example, the purple curve shows the parallelization which uses for each data parallelization process (\ie MPI process) 8 additional cores for task parallelization.
 As a result, for example, when using 2 machines of the TGCC cluster, with a total of 32 cores, 4 cores are used for SkelGIS MPI processes, for data parallelization, and for each one 8 cores are used for task parallelization ($4 \times 8 = 32$).
-
-As a result, and as explained in Section~\ref{sect:perfs}, data is less divided into sub-domains and the effect which is observe onto the blue curve is delayed.
-This figure shows a comparison with 2, 4, 8 and 16 cores per MPI process for task parallelization.
+This respect $P = P_{data} \times P_{task}$ presented in Section~\ref{sect:perfs}.
+As a result, and as explained in Section~\ref{sect:perfs}, quantities are less divided into sub-domains, that are responsible for communications, and the effect which is observe onto the blue curve is delayed.
 
@@
 However, reaching 16 cores, an important initial overhead appears and in addition to this, the curve bends down rapidly instead of improving the one with 8 cores for task parallelization.
-Two different phenomena happen.
+Two different phenomena happen in this case.
 
@@
 As a result, when increasing the number of OpenMP cores for task parallelization from 8 to 16 cores, an overhead is introduced by exchanges of data between memories of the two NUMA nodes.
-This phenomena is illutrated in Figure~\ref{fig:spread}.
-In this figure, a different strategy is used to bind threads onto available cores (using OpenMP).
-This strategy, called \emph{spread} (instead of \emph{close} in Figure~\ref{fig:close}), binds threads on cores in order to spread as much as possible onto resources, which means that the two NUMA nodes are directly used whatever the number of cores kept for task parallelization.
-As a result, and as shown in the figure, using 2, 4 and 8 cores an initial overhead is introduced.
+This phenomena is illutrated in Fig.~\ref{fig:spread}.
+In this figure, a different binding strategy is used.
+A binding strategy is the way the scheduler binds threads onto available cores.
+The strategy used in Fig.~\ref{fig:spread} is called \emph{spread} (instead of \emph{close} in Figure~\ref{fig:close}).
+This strategy binds threads on cores in order to spread as much as possible onto resources, which means that the two NUMA nodes are directly used whatever the number of cores used for task parallelism.
+As a result, and as shown in the figure, using 2, 4 and 8 cores an initial overhead is introduced as the one observed in Fig.~\ref{fig:close}.
+This shows that the initial overhead with 16 cores is due to NUMA effects.
 
@@
 
-The second phenomena, which happens in Figure~\ref{fig:close} with 16 cores for tasks parallelization, is due to the level of parallelization introduced by the task parallelization technique.
-Actually, as illustrated in Table~\ref{fig:freq}, only two steps of the $TSP$ static scheduling generated by the MSL compiler can take advantage of 16 cores among a total of 18 steps.
-This phenomena has been explained in Section~\ref{sect:perfs} by the variable $F_{task}$ and the fact that it is not always true that $F_{task}=P_{task}$.
-This explains why using 16 cores for task parallelization in Figure~\ref{fig:spread} is still less efficient than using 8 cores even if the two NUMA nodes are always used in this evaluation.
+The second phenomena that happens in Figure~\ref{fig:close} using 16 cores is due to the level of parallelism introduced by the task parallelization technique.
+Actually, as illustrated in Table~\ref{fig:freq}, only two forks of $TSP$ can take advantage of 16 cores among a total of 18 forks.
+This phenomena has been mentioned in Section~\ref{sect:perfs} by the variable $F_{task}$ and the fact that it is not always true that $F_{task}=P_{task}$.
+This explains why using 16 cores is less efficient than using 8 cores, even when the two NUMA nodes are always used as in Fig.~\ref{fig:spread}.
 
-Finally, to completely validate the performance model introduced in Section~\ref{sect:perfs}, and to understand when the hybrid parallelization becomes more interesting than the data parallelization, Figure~\ref{fig:tth2} represents $T_{COM1}$ and $T_{COM2}+T_{task}$ of Equation~(\ref{eq:hyb}), for the best case (\ie when 8 cores are used for task parallelization), and with a \emph{close} OpenMP bind of threads onto cores.
-Table~\ref{fig:tth} gives time details.
+Finally, to validate the performance model introduced in Section~\ref{sect:perfs}, and to understand when the hybrid parallelization becomes more interesting than the data parallelization, Figure~\ref{fig:tth2} represents $T_{COM1}$ and $T_{COM2}+T_{task}$ of Equation~(\ref{eq:hyb}), for the best case, \ie when 8 cores are used in Fig.~\ref{fig:close}.
+Figure~\ref{fig:tth2} and Table~\ref{fig:tth} presents results of these measurements.
+Results perfectly matches Fig.~\ref{fig:close} for 8 cores per MPI process.
+As a result, the hybrid parallelization becomes better with a total of 512 cores in this case.
 
@@
 \end{center}\end{figure}
-
-Figure~\ref{fig:tth2} and Table~\ref{fig:tth} perfectly matches results observed in Figure~\ref{fig:close} for 8 cores used for task parallelization per core used for data parallelization.
-As a result, the hybrid parallelization becomes better with a total of 512 cores.

### 2017-02-24 ARTICLES/IJPP/articleR1/introduction.tex
evaluation

@@
 
-As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
+%As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
 For example, the current world's top supercomputer Tianhe-2~\footnote{\url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance around thirty peta floating point operations par seconds.
@@
 
-One of the big challenges to solve to be able to use exascale computers is to propose programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}.
+%One of the big challenges to solve to be able to use exascale computers is to propose programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}.
 Actually, applications which run on supercomputers and which need such a computation power are physics, weather or genomic applications, which are not implemented by HPC specialists most of the time.
 
-One possible runtime execution model for HPC is to propose dynamic scheduling of task graphs combined to message passing models~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359} (to be able to use more than one machine).
+%One possible runtime execution model for HPC is to propose dynamic scheduling of task graphs combined to message passing models~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359} (to be able to use more than one machine).
 Those models increase HPC code portability and reach an interesting performance onto heterogeneous architectures, which is interesting to reach exascale programming.
@@
 
-However, many DSLs have been proposed for many domains yet, and, as far as we know, only a few of them have been able to reuse work already done by another language~\cite{Sujeeth:2013:CRC:2524984.2524988}.
+%However, many DSLs have been proposed for many domains yet, and, as far as we know, only a few of them have been able to reuse work already done by another language~\cite{Sujeeth:2013:CRC:2524984.2524988}.
 In other words, software engineering properties have to be integrated into DSL conception, such that a new DSL can be seen as a composition of parallelization, optimizations or even languages semantics already proposed by others.
 
-For example, to numerically solve a set of partial differential equations (PDEs), iterative methods are frequently used to approximate the exact solution through a discretized phenomena.
+%For example, to numerically solve a set of partial differential equations (PDEs), iterative methods are frequently used to approximate the exact solution through a discretized phenomena.
 A very well known and usual way to discretize PDEs is to transform them to explicit numerical schemes, also often called \emph{stencils}.
@@
 Many of them use same kind of parallelization, data structures or optimizations, however each one has been built from scratch to deal with another additionnal specific case.
-We present the Multi-Stencil Language (MSL) DSL, also for stencil-based numerical simulations.
+%We present the Multi-Stencil Language (MSL) DSL, also for stencil-based numerical simulations.
 MSL is a language with a light grammar to describe a numerical simulation without implementation details.
@@
 
-Section~\ref{sect:rel} introduces the related work on DSLs for stencils.
+%Section~\ref{sect:rel} introduces the related work on DSLs for stencils.
 Section~\ref{sect:formalism} formally explains the targetted domain and its computational model.

### 2017-03-02 ARTICLES/IJPP/articleR1/msf.tex
corrections + evaluation en cours
Section 2.0 & 2.1

@@
-\HC{under construction, present the framework}
-
 This section first presents a background on component models and particularly on the Low Level Components.
-This background is needed to understand the second part of the section which presents an overview of the overall Multi-Stencil Framework (MSF).
+This background is needed to understand the second part of the section which gives an overview of the overall Multi-Stencil Framework (MSF).
 
@@
 %----------------------------------------
-Component-based software engineering (CBSE) is a domain of software engineering~\cite{Szyperski:2002:CSB:515228} which improves code re-use, separations of concerns, and thus maintainability.
+Component-based software engineering (CBSE) is a domain of software engineering~\cite{Szyperski:2002:CSB:515228} which aims at improving code re-use, separations of concerns, and thus maintainability.
 A component-based application is made of a set of component instances linked together, this is also called a component \emph{assembly}.
 A component is a black box that implements an independent functionality of the application, and which interacts with its environment only through well defined interfaces: its ports.
-A port can, for example, specify services provided or required by the component.
+For example, a port can specify services provided or required by the component.
 With respect to high performance computing, some works have also shown
-that component models can achieve the needed level of performance, and
+that component models can achieve the needed level of performance and
 scalability while also helping in application portability~\cite{Bernholdt01052006, bigot:inria-00388508, UCHPC2015}.
@@
 %
-In this paper, and as illustrated in Figures~\ref{fig:ports}, a $provide$ port is represented by a white circle, a $use$ port with a black circle, a $use-multiple$ port by a black circle with a white $m$ in it.
+In this paper, and as illustrated in Figure~\ref{fig:ports}, a $provide$ port is represented by a white circle, a $use$ port by a black circle, a $use-multiple$ port by a black circle with a white $m$ in it.
 MPI port are connected with a black rectangle.
@@
 One can also notice that \llc can achieve performance if the granularity of components is high enough and attentively chosen by the user.
+The typical overhead of a \llc is a \texttt{C++} indirect virtual method invocation.
 

### 2017-03-02 ARTICLES/IJPP/articleR1/msf.tex
Section 2.2

@@
 
-Figure~\ref{fig:msf} presents the overview of the Multi-Stencil Framework that will be entirely detailed throughout this paper. It is composed of four different parts presented below.
-One can note in this figure that MSF is addressed to two different kinds of end-users: the \emph{numerician}, in other words the mathematician, and the \emph{developer}.
-Most of the time numericians do have programming knowledge, however as it is not their core domain and because of a lack of time, development is often left to engineers according to numericians needs.
+Figure~\ref{fig:msf} gives an overview of the Multi-Stencil Framework that is entirely detailed throughout this paper. It is composed of four distinct parts described hereafter.
+As illustrated in Figure~\ref{fig:msf}, MSF targets two different kinds of end-users: the \emph{numerician}, in other words the mathematician, and the \emph{developer}.
+Most of the time numericians do have programming knowledge, however as it is not their core domain and because of a lack of time, development is often left to engineers according to numerician needs.
 MSF has the interesting particularity to propose a clear separation of concerns between these two end-users by distinguishing the description language from the numerical code.
+\CP{pas top clair la fin de la phrase}
 
 MSF also has the interesting capability to be more flexible than existing solutions thanks to a possibility for a third party to interact with the framework.
-This third party is a High Performance Computing (HPC) specialist as illustrated in Fig.~\ref{fig:msf}.
+This third party is a High Performance Computing (HPC) specialist as displayed in Figure~\ref{fig:msf}.
 
@@
 The core of the framework is the Multi-Stencil Compiler, or MSC.
-It is responsible for transforming the genric assembly into the final parallel assembly which is specific to the simulation described by the numerician with MSL.
+It is responsible for transforming the generic assembly into the final parallel assembly which is specific to the simulation described by the numerician with MSL.
 MSC is described in Section~\ref{sect:parallelism}.
@@
 This final specialized component assembly represents the parallel computations orchestration of the simulation initially described by the numerician.
+\CP{Dire que c'est du \llc?}

### 2017-03-02 ARTICLES/IJPP/articleR1/formalism.tex
Section 3.0 & 3.1

@@
 %In other words, the PDEs are transformed to a set of numerical computations applied at each time-step on elements of the discretized space domain.
-While the computations can have various forms, we focus on three categories.
+While the computations can have various forms, we focus on three categories:
 %TODO[JB] 4 plus tard...
-\emph{Stencil} computations involve access to neighbor values only (the concept of neighborhood depending on the space discretization used).
-\emph{Local} computations depend on the computed location only (this can be seen as a stencil of size one).
+\emph{Stencil} computations involve access to neighbor values only (the concept of neighborhood depending on the space discretization used);
+\emph{Local} computations depend on the computed location only (this can be seen as a stencil of size one);
 Finally, \emph{reductions} enable to transform values mapped on the mesh to a single scalar value.
@@
 
+Let introduce some notations.
 $\Omega$ is the continuous space domain of a numerical simulation (typically $\mathbb{R}^n$).
@@
 \item A \textit{group of mesh entities} $\mathcal{G} \in \mathcal{P}(V\cup E)$ represents a set of entities of the same topology and the mesh of a group is denoted $mesh(\mathcal{G})$, \textit{i.e.} $\mathcal{G} \in \mathcal{P}(V\cup E) \Leftrightarrow mesh(\mathcal{G})=(V,E)$
+\CP{Pas clair pour moi}
 \item The \textit{set of entities groups} used in a simulation is denoted $\Phi$.
+\CP{Entities groups not defined! a propos, n'est pas set of entit*y* groups?}
 \end{mydefs}
@@
 This example is illustrated in Figure~\ref{fig:meshbase}.
+\CP{Idem, pas top clair. Il y a des implicites !}
 
@@
 The finite sequence $T: (t_n)_{n\in\llbracket 0, T_{max} \rrbracket}$ represents the discretization of the continuous time domain $\mathcal{T}=\mathbb{R}$.
-To each discrete time-step $n\in\llbracket 0, T_{max} \rrbracket$, a time value $t_n\in\mathcal{T}$ is associated. The time step is $t_n-t_{n-1}$.
+To each discrete time-step $n\in\llbracket 0, T_{max} \rrbracket$, a time value $t_n\in\mathcal{T}$ is associated. The time step for $t_n$ is equal to $t_n-t_{n-1}$.
 \end{mydef}
@@
 A mesh variable $\delta \in \Delta$ associates to each pair of entity and time-step a value $\delta: \mathcal{G}\times T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
-\item The group of entities a variable is mapped on is denoted $entity(\delta)=\mathcal{G}$.
+\item The group of entities a variable is mapped on is denoted $entity(\delta)=\mathcal{G}$.\CP{Pas clair}
 \item $\mathcal{S}$ are the \textit{scalar variables}.
@@
 Thus, $\forall t\in \llbracket 0, T_{max-1} \rrbracket$, $conv(t)=0$, and $conv(T_{max})=1$.
+\CP{Il y a des implicites : est ce un calcul jusqu'à convergence ou un calcul sur MAX time step?
+Ou veux tu modéliser les deux (suivant que MAX est connu a priori ou a postiori?}
 \end{mydefs}

### 2017-03-02 ARTICLES/IJPP/articleR1/conclusion.tex
corrections + début intro

@@
 
-E%xperiments presented in this paper show that the back-end code produced by MSL, \ie by a mesh-agnostic language, does not introduce overheads up to 16.384 cores, and that the hybrid (data and task) parallelism automatically introduced by the language, which combines SkelGIS and OpenMP as a back-end, improves performance compared to a data parallelization only.
+%Experiments presented in this paper show that the back-end code produced by MSL, \ie by a mesh-agnostic language, does not introduce overheads up to 16.384 cores, and that the hybrid (data and task) parallelism automatically introduced by the language, which combines SkelGIS and OpenMP as a back-end, improves performance compared to a data parallelization only.
 Thus, with the same code asked to the user, two different parallel versions of the simulation can be obtained.
@@
 \HC{framework to ease HPC specialist work for proposing new meshes etc.}
-\HC{how to choose data or hybrid, change on the fly if needed}
+\HC{how to choose data or hybrid, need for a network model, or learning strategies onto production codes, change on the fly if needed}
 \HC{how to choose if a fusion is good or not, change on the fly or learning strategies onto production codes}

### 2017-03-02 ARTICLES/IJPP/articleR1/introduction.tex
corrections + début intro

--- 2017-02-20 ARTICLES/IJPP/articleR1/introduction.tex
@@
 However, for non-experts end-users, general purpose languages still are difficult to use and to tune for a given application onto a given architecture.
-The current easiest (with the higher abstraction level), but still efficient, programming model for end-users is Domain Specific Language (DSL).
-Such a language is specific to the end-user domain and proposes a grammar which is easy to understand.
-The DSL compiler is able, because of the specific knowledge onto the domain targetted, to automatically apply parallelization and specific optimizations to produce a high performance back-end code.
-Thus, a DSL is able to split end-user concerns from HPC concerns which is relevant to reach exascale programming models.
+The current easiest (with the higher abstraction level), but still efficient, programming model for end-users is Domain Specific Language (DSL) and domain specific frameworks.
+Such solutions are specific to the end-user domain and propose a grammar or an API which is easy to understand.
+Moreover, the compiler is able, because of the specific knowledge onto the targeted domain, to automatically apply parallelization and specific optimization to produce a high performance back-end code.
+Thus, a domain specific solutions are able to split end-user concerns from HPC concerns which is relevant to reach exascale programming models.
 
-However, many DSLs have been proposed for many domains yet, and, as far as we know, only a few of them have been able to reuse work already done by another language~\cite{Sujeeth:2013:CRC:2524984.2524988}.
-In other words, software engineering properties have to be integrated into DSL conception, such that a new DSL can be seen as a composition of parallelization, optimizations or even languages semantics already proposed by others.
+However, many domain specific languages and frameworks have been proposed for many domains yet, and each one claims to solve a specific case that others don't.
+Furthermore instead of re-using a common basis for every language or framework of a same domain, each solution is implemented from scratch as a monolithic code.
+We claim in this paper that being able to find a common meta-model for all solutions of a given domain facilitates code re-use, and eases flexibility and maintainability of solutions.
 
@@
 Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107}, as it will be detailed in Section~\ref{sect:rel}.
+Many of them use same kind of parallelization, data structures or optimization, however each one has been built from scratch to deal with another additional specific case.
+
+We propose the Multi-Stencil Framework (MSF) that is built over a meta-model of multi-stencil simulations.
+MSF produces a parallel orchestration of a multi-stencil program without being aware of the underlying implementation choices (\eg distributed data structures, task scheduler etc.).
+Thanks to this meta-model MSF is able to easily switch from one parallelization technique to another and from one optimization to another.
+Moreover, as the meta-model is independent from implementation details, MSF can easily choose one back-end or another, thus easing code reuse of existing solutions.
+
+Contributions of the paper are: a meta-model of a multi-stencil program presented in Section~\ref{}; from this meta-model are built both a light and descriptive domain specific language MSL as well as a generic template of the application described in Section~\ref{}; the compiler MSC of the framework builds a parallel specialized version of the generic template from the specific application described with MSL; finally a performance evaluation is detailed in Section~\ref{}  and validates the performance model introduced in Section~\ref{}.
+
+%as far as we know, only a few of them have been able to reuse work already done by another language~\cite{Sujeeth:2013:CRC:2524984.2524988}.
+In other words, software engineering properties have to be integrated into DSL conception, such that a new DSL can be seen as a composition of parallelization, optimizations or even languages semantics already proposed by others.
+%For example, to numerically solve a set of partial differential equations (PDEs), iterative methods are frequently used to approximate the exact solution through a discretized phenomena.
+A very well known and usual way to discretize PDEs is to transform them to explicit numerical schemes, also often called \emph{stencils}.
+Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107}, as it will be detailed in Section~\ref{sect:rel}.
 Many of them use same kind of parallelization, data structures or optimizations, however each one has been built from scratch to deal with another additionnal specific case.
-We present the Multi-Stencil Language (MSL) DSL, also for stencil-based numerical simulations.
+%We present the Multi-Stencil Language (MSL) DSL, also for stencil-based numerical simulations.
 MSL is a language with a light grammar to describe a numerical simulation without implementation details.

### 2017-03-02 ARTICLES/IJPP/articleR1/msl.tex
corrections + début intro

--- 2017-02-22 ARTICLES/IJPP/articleR1/msl.tex
@@
-\HC{under revision / has to be revised according to modifications}
-
 %----------------------------------------
@@
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component] (D) at (0,0) {$Driver$};
+   \node[component,fill=green!50!white] (D) at (0,0) {$Driver$};
    \node[provide] (Dp) at (-1,0) {};
@@
    \node[provide,below=0.15 of Du2] (Tp) {};
-   \node[component,below=1.6cm of Tp] (T) {$Time$};
+   \node[component,below=1.6cm of Tp,fill=green!50!white] (T) {$Time$};
    \node[use,right=1cm of T] (Tu) {};
@@
    \node[provide,right=0.15 of Tu] (Cp) {};
-   \node[component, draw=blue,fill=green!80!black,text=blue!80!black,right=2cm of Cp] (C) {$Computations$};
+   \node[component, draw=blue!60!white,fill=green!50!white,text=blue!80!black,line width=0.5mm,right=2cm of Cp] (C) {$Computations$};
    \node[use,above=0.8cm of C] (Cu) {};

### 2017-03-02 ARTICLES/IJPP/articleR1/parallelism.tex
corrections + début intro

@@
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
-   \node[component,fill=green!80!black] (D) at (0,0) {$Driver$};
+   \node[component,fill=green!50!white] (D) at (0,0) {$Driver$};
    \node[provide] (Dp) at (-1,0) {};
@@
    \node[provide,below=0.15 of Du2] (Tp) {};
-   \node[component,below=1.6cm of Tp,fill=green!80!black] (T) {$Time$};
+   \node[component,below=1.6cm of Tp,fill=green!50!white] (T) {$Time$};
    \node[use,right=1cm of T] (Tu) {};
@@
    \node[provide,right=0.15 of Tu] (Cp) {};
-   \node[component, fill=green!80!black,right=2cm of Cp] (C) {$Scheduler$};
+   \node[component, fill=green!50!white,right=2cm of Cp] (C) {$Scheduler$};
    \node[use,right=1.8 of C] (scu1) {};
@@
    \node[provide,below right=0.2 of Du3] (Datap1) {};
-   \node[component,draw=red,fill=red!50!white,above=0.8cm of Datap2] (Data) {$A$};
-   \node[component,draw=red,fill=red!50!white,below=0.8cm of Data] (Data2) {$B$};
+   \node[component,draw=red,line width=0.4mm,fill=green!50!white,above=0.8cm of Datap2] (Data) {$A$};
+   \node[component,draw=red,line width=0.4mm,fill=green!50!white,below=0.8cm of Data] (Data2) {$B$};
    \node[below=0.5cm of Data2] (Data3) {...};
@@
    \node[provide,right=0.15 of scu2] (psync1) {};
-   \node[component, fill=blue!60!white,right=2cm of pk0] (k0) {$k_0$};
+   \node[component, fill=blue!50!white,right=2cm of pk0] (k0) {$k_0$};
    \node[use,right=0.8 of k0] (k0u) {};
@@
    \node[right=0.2 of k0u2] (1) {2};
-   \node[component, fill=blue!60!white,below=0.8cm of k0] (sync1) {$Sync(0,1)$};
+   \node[component, fill=green!50!white,below=0.8cm of k0] (sync1) {$Sync(0,1)$};
    \node[use,right=1.5 of sync1] (syncu) {};

### 2017-03-02 ARTICLES/IJPP/articleR1/introduction.tex
intro

@@
 For example, the current world's top supercomputer Tianhe-2~\footnote{\url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance around thirty peta floating point operations par seconds.
-However, to be able to use such a machine, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA etc., and multiple optimization techniques, such as cache or vectorization optimizations, have to be combined.
+However, to be able to use such a machine, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA etc., and multiple optimization techniques, such as cache optimization, have to be combined.
 Moreover, current architectures evolution lets think that heterogeneity and complexity in HPC will continue to grow in future.
@@
 
-However, many domain specific languages and frameworks have been proposed for many domains yet, and each one claims to solve a specific case that others don't.
-Furthermore instead of re-using a common basis for every language or framework of a same domain, each solution is implemented from scratch as a monolithic code.
-We claim in this paper that being able to find a common meta-model for all solutions of a given domain facilitates code re-use, and eases flexibility and maintainability of solutions.
+Many domain specific languages and frameworks have been proposed for many domains yet, and each one claims to solve a specific case that others don't.
+However, instead of re-using a common basis for every language or framework of a same domain, each solution is implemented from scratch as a monolithic code.
+We claim in this paper that being able to find a common meta-model to all solutions for a given domain facilitates code re-use, and eases flexibility and maintainability of solutions.
 
@@
 Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107}, as it will be detailed in Section~\ref{sect:rel}.
-Many of them use same kind of parallelization, data structures or optimization, however each one has been built from scratch to deal with another additional specific case.
+Many of them use same kind of parallelization, data structures or optimization techniques, however each one has been built from scratch.
 
@@
 Moreover, as the meta-model is independent from implementation details, MSF can easily choose one back-end or another, thus easing code reuse of existing solutions.
+To ease composition of existing solutions, MSF uses Software Engineering capabilities of component-based programming models~\cite{Szyperski:2002:CSB:515228} where applications are defined as an assembly of building blocks, or components.
 
-Contributions of the paper are: a meta-model of a multi-stencil program presented in Section~\ref{}; from this meta-model are built both a light and descriptive domain specific language MSL as well as a generic template of the application described in Section~\ref{}; the compiler MSC of the framework builds a parallel specialized version of the generic template from the specific application described with MSL; finally a performance evaluation is detailed in Section~\ref{}  and validates the performance model introduced in Section~\ref{}.
+After a short overview of the Multi-Stencil Framework given in Section~\ref{sect:msf}, contributions of the paper are presented as follows: a meta-model of a multi-stencil program presented in Section~\ref{sect:formalism}; from this meta-model are built both a light and descriptive domain specific language, namely MSL, as well as a generic component assembly of the application described in Section~\ref{sect:msl}; the compiler MSC of the framework builds a parallel specialized version of the generic assembly from the application description with MSL, MSC and its performance model are described in Section~\ref{sect:parallelism}; finally a performance evaluation is detailed in Section~\ref{sect:eval}  and validates the performance model.
 

### 2017-03-02 ARTICLES/IJPP/articleR1/evaluation.tex
new results + new structure evaluation

@@
 
-\HC{another benchmark to show that with the static version of \llc the problem disappear ?}
+%\HC{another benchmark to show that with the static version of \llc the problem disappear ?}
 
@@
 
-\paragraph{\textbf{Fusion}} Finally, we evaluate loop fusions automatically proposed by MSF from the $TSP$ tree of computation kernels.
+%\paragraph{\textbf{Fusion}} Finally, we evaluate loop fusions automatically proposed by MSF from the $TSP$ tree of computation kernels.
 Figure~\ref{fig:fusion} shows the number of iterations per second as a function of the number of cores with and without fusions.
@@
 Moreover, one can notice that the performance is clearly improved (around 40\%) by this fusion.
-
-\begin{figure}[!h]\begin{center}
-  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/fusVSbase.pdf}
-  \caption{Strong scaling on a 500x500 domain size with $200$ time iterations, with and without fusions proposed by MSF.}
-  \label{fig:fusion}
-\end{center}\end{figure}
-
-\HC{another benchmark with and without fusion fro task parallelization ?}
+%
+%\begin{figure}[!h]\begin{center}
+%  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/fusVSbase.pdf}
+%  \caption{Strong scaling on a 500x500 domain size with $200$ time iterations, with and without fusions proposed by MSF.}
+%  \label{fig:fusion}
+%\end{center}\end{figure}
+%
+%\HC{another benchmark with and without fusion fro task parallelization ?}
 
@@
 \end{center}\end{figure}
+
+\subsection{Fusion evaluation}
+
+\paragraph{\textbf{Fusion}} Finally, we evaluate loop fusions automatically proposed by MSF from the $TSP$ tree of computation kernels.
+Figure~\ref{fig:fusion} shows the number of iterations per second as a function of the number of cores with and without fusions.
+This benchmark is performed on FullSWOF2D onto a $500 \times 500$ domain size with $200$ time iterations.
+As explained in Section~\ref{sect:fusion}, the MSF loop fusion happens at a high level and is most of the time done naturally by a computer scientist.
+However, for a non computer scientist which write its numerical codes, an automatic proposition of such fusions avoid errors, particularly for a parallel execution.
+Moreover, one can notice that the performance is clearly improved (around 40\%) by this fusion.
+
+\begin{figure}[!h]\begin{center}
+  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/fusVSbase.pdf}
+  \caption{Strong scaling on a 500x500 domain size with $200$ time iterations, with and without fusions proposed by MSF.}
+  \label{fig:fusion}
+\end{center}\end{figure}
+
+\begin{figure}[ht]
+\begin{center}
+\subfloat[blabla\label{fig:fusion1}]{
+\includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout2_close_median.pdf}
+}\\
+\hspace{10pt}
+\subfloat[blabla\label{fig:fusion2}]{
+\includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout8_close_median.pdf}
+}
+\end{center}
+\caption{blabla}
+\label{fig:fusion}
+\end{figure}
+
+%\begin{figure}[!h]\begin{center}
+%  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout2_close_median.pdf}
+%  \caption{Strong scaling on a 500x500 domain size with $200$ time iterations, with and without fusions proposed by MSF.}
+%  \label{fig:fusion}
+%\end{center}\end{figure}
+%
+%\begin{figure}[!h]\begin{center}
+%  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout8_close_median.pdf}
+%  \caption{Strong scaling on a 500x500 domain size with $200$ time iterations, with and without fusions proposed by MSF.}
+%  \label{fig:fusion}
+%\end{center}\end{figure}

### 2017-03-02 ARTICLES/IJPP/results/task_scaling/500_200/max_mean.py

@@ -3,8 +3,8 @@ import glob, os, string
 # os.chdir("./base")
 # filesearch = "base_MSL_500_200_128_"
 
-os.chdir("./fus")
-filesearch = "fus_MSL_500_200_128_"
+os.chdir("./task2Fus")
+filesearch = "taskFus2MSL_500_200_1024_"
 
 # os.chdir("./task16")
 # filesearch = "task16MSL_500_200_128_"

### 2017-03-02 ARTICLES/IJPP/results/task_scaling/500_200/results_with_without_fusion.py

@@ -1,0 +1,82 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.backends.backend_pdf import PdfPages
+
+from matplotlib import rc
+matplotlib.rc('pdf', fonttype=42)
+
+cores = [16,32,64,128,256,512,1024,2048]
+
+# median 10 runs
+# without fusion close
+timesMSL = [2.35,1.05,0.54,0.3,0.2,0.19,0.24,0.49]
+timesMSLfus = [1.38,0.54,0.28,0.17,0.148,0.21,0.57,2.64]
+
+timesT2 = [2.32,1.05,0.54,0.3,0.18,0.14,0.16,0.22]
+timesT8 = [3.06,1.48,0.77,0.41,0.23,0.14,0.12,0.078]
+
+#with fusion close
+timesT2fus = [1.44,0.57,0.29,0.17,0.115,0.236,1.05,5.98]
+timesT8fus = [4.61,2.294,1.177,0.62,0.3,0.18,0.133,0.283]
+
+itsMSL = []
+itsMSLfus = []
+itsT2 = []
+itsT8 = []
+itsT2fus = []
+itsT8fus = []
+
+for i in timesMSL:
+    itsMSL.append(500/i)
+for i in timesMSLfus:
+    itsMSLfus.append(500/i)
+for i in timesT2:
+	if i!=0:
+		itsT2.append(500/i)
+	else :
+		itsT2.append(0)
+for i in timesT8:
+	if i!=0:
+		itsT8.append(500/i)
+	else :
+		itsT8.append(0)
+
+for i in timesT2fus:
+	if i!=0:
+		itsT2fus.append(500/i)
+	else :
+		itsT2fus.append(0)
+for i in timesT8fus:
+	if i!=0:
+		itsT8fus.append(500/i)
+	else :
+		itsT8fus.append(0)
+
+#relatives_its2 = []
+#for i in cores:
+#    relatives_its2.append(500/(timesMSL[0]/(i/cores[0])))
+
+matplotlib.rcParams.update({'font.size': 16})
+
+#plt.xkcd()
+
+#plt.plot(cores, relatives_its2, 'b--', label='Ideal')
+plt.loglog(cores, itsMSL, 'b-o', label='MSL data parallelization only', basex=2,basey=2)
+plt.loglog(cores, itsMSLfus, 'b-x', label='MSL data parallelization only', basex=2,basey=2)
[…]

### 2017-03-02 ARTICLES/IJPP/articleR1/evaluation.tex
new results + new structure evaluation

@@
 
-\begin{figure}[ht]
-\begin{center}
-\subfloat[blabla\label{fig:fusion1}]{
+\begin{figure}[!h]\begin{center}
   \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout2_close_median.pdf}
-}\\
-\hspace{10pt}
-\subfloat[blabla\label{fig:fusion2}]{
+  \caption{blabla}
+  \label{fig:fusion1}
+\end{center}\end{figure}
+
+\begin{figure}[!h]\begin{center}
   \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout8_close_median.pdf}
-}
-\end{center}
   \caption{blabla}
-\label{fig:fusion}
-\end{figure}
-
-%\begin{figure}[!h]\begin{center}
-%  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout2_close_median.pdf}
-%  \caption{Strong scaling on a 500x500 domain size with $200$ time iterations, with and without fusions proposed by MSF.}
-%  \label{fig:fusion}
-%\end{center}\end{figure}
-%
-%\begin{figure}[!h]\begin{center}
-%  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout8_close_median.pdf}
-%  \caption{Strong scaling on a 500x500 domain size with $200$ time iterations, with and without fusions proposed by MSF.}
-%  \label{fig:fusion}
-%\end{center}\end{figure}
+  \label{fig:fusion2}
+\end{center}\end{figure}

### 2017-03-03 ARTICLES/IJPP/articleR1/conclusion.tex
conclusion

--- 2017-02-24 ARTICLES/IJPP/articleR1/conclusion.tex
@@
-%We have presented the Multi-Stencil Language (MSL), a domain specific language for stencil-based numerical simulations.
-Compared to other existing stencil DSLs, MSL proposes a higher abstraction level for end-users.
-MSL offers a way to describe a numerical simulation independently from the type of mesh used.
-As a result, the language can be used for a larger number of simulations.
-It helps to clearly separate the description of the simulation from implementation choices, which facilitates reuse of existing languages and libraries, for example languages for data structures optimizations and parallelization strategies.
+In this paper, we have presented the Multi-Stencil Framework.
+MSF is built upon a meta-model of a multi-stencil program which has been presented in Section~\ref{sect:formalism}.
+From this meta-model has been designed, first, the generic component assembly of a multi-stencil program, and second, the domain specific language MSL that enables the description of a specific application by a numerician.
+From these entries, the MSF compiler, namely MSC, is able to generate automatically a parallel specific assembly of the simulation.
+This specific assembly represents the parallel orchestration of computations independently of implementation choices.
+Moreover, by combining a meta-model and component-based programming models MSF has the particularity to enhance separation of concerns and code-reuse of existing solutions (\eg SkelGIS and OpenMP in this paper).
+Finally, MSF has been evaluated on the real case simulation FullSWOF2D.
+Results have shown that even if component models have been used, strong and weak scalings of MSF (for its data parallelization technique) do not suffer from overheads.
+Results have also shown that the hybrid parallelization handled by MSF could increase performances when MPI communications are too heavy, which is new compared to existing solutions.
+Finally, we have presented evaluations on fusion optimization that demonstrates that fusion is not always a good choice, particularly when performing hybrid parallelism.
 
-%In this paper, the formal definition of a multi-stencil program has been given.
-This formalism helps to understand what is defined into the light grammar of MSL, and also how parallelization can be extracted from it.
-MSL produces an empty parallel pattern of the numerical simulation described.
-This pattern indicates where synchronizations are needed, but not how they are performed, which is dependent from the type of mesh and from implementation choices.
-This second step is left to other existing languages, which are chosen as a back-end of MSL.
-For example, in this paper the chosen back-end is a code using SkelGIS~\cite{CPE:CPE3494,DBLP:conf/ieeehpcs/HeleneS13,DBLP:conf/europar/CoullonL14}, a templated C++ language for distributed meshes (using MPI), and OpenMP~\cite{660313}.
+Many perspectives of future work raise from this work.
+First, even if using one distributed data structure or another is possible and facilitated by MSF, the HPC specialist still has to understand component programming models to add one.
+We think it could be interesting to also facilitate the work of the HPC specialist by proposing a semi-automated framework to add external tools to MSF.
 
-E%xperiments presented in this paper show that the back-end code produced by MSL, \ie by a mesh-agnostic language, does not introduce overheads up to 16.384 cores, and that the hybrid (data and task) parallelism automatically introduced by the language, which combines SkelGIS and OpenMP as a back-end, improves performance compared to a data parallelization only.
-Thus, with the same code asked to the user, two different parallel versions of the simulation can be obtained.
+Second, it has been shown in the performance model and evaluation results that the choice between data parallelism and hybrid parallelism in not trivial.
+It depends on the time spent in computations, the number of cores and nodes used for the execution, as well as the network behavior (Fig.~\ref{fig:limit} and~\ref{fig:close}).
+MSF, thanks to its meta-model and to the use of component models, handles the activation of one parallelism technique or another.
+Thus MSF can be considered as a performance leverage for multi-stencil programs.
+However, it would be interesting to investigate how an actor algorithm could choose the best parallelization technique for a given execution of a production multi-stencil simulation.
 
-%However, if performances are improved by the hybrid parallelization, this performance is also limited by the scheduling strategy used in this paper.
-Actually, as described in Section~\ref{sect:msp}, a static series-parallel (fork-join) scheduling is performed in this work.
-This helps to define a predictive performance model, which has been validated in experiments.
-However, getting rid of global synchronizations (join) could be an interesting improvement to get better performance, especially for irregular simulations, where tasks are not naturally balanced.
-Our immediate perspective, thus, is to use OpenMP dynamic schedulers~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359} as back-ends to study performance improvments.
-Using such dynamic schedulers also opens to more heterogeneous architectures, such as GPUs or many-cores accelerators.
+The same choice problem is raised by fusion evaluations of Section~\ref{sect:fus}.
+Actually, fusion optimization is not always relevant.
+Thus, for the same reasons than parallelization techniques, an actor algorithm could choose the best solution for a given execution.
+To design such actor algorithms different solutions should be studied, such as following an approximated behavior model, or using setup runs to use deep learning algorithms etc.
 
-%Finally, to validate the separation of concerns introduced by MSL, a second back-end for unstructured meshes could be interesting.
-We think about PaMPA~\cite{lachat:hal-00768916} as a distributed data structure backend, instead of SkelGIS.
-Thus using a single language, the MSL language, two different kinds of numerical simulations could be defined: one onto Cartesian meshes; and one on unstructured meshes.
-
-\HC{framework to ease HPC specialist work for proposing new meshes etc.}
-\HC{how to choose data or hybrid, change on the fly if needed}
-\HC{how to choose if a fusion is good or not, change on the fly or learning strategies onto production codes}
-\HC{In addition to this, by using components, doors to online reconfiguration of applications~\cite{Lanore:2015:RCM:2737166.2737169} are also open (even if not answered in this paper).}
+Finally, by using component models doors are opened to reconfiguration of applications~\cite{Lanore:2015:RCM:2737166.2737169} which means that the application could adapt its behavior and structure to external events.

### 2017-03-03 ARTICLES/IJPP/articleR1/evaluation.tex
conclusion
evaluation

@@
 
+%-------------------------------------------------
 \subsection{Fusion evaluation}
+\label{sect:fus}
+%-------------------------------------------------
 
-\paragraph{\textbf{Fusion}} Finally, we evaluate loop fusions automatically proposed by MSF from the $TSP$ tree of computation kernels.
+In this section we propose an evaluation of the fusion optimization.
+Three fusion optimizations have been formally detailed in Section~\ref{sect:fusion}.
+From the $TSP$ tree computed by MSC it is possible according to specific conditions to merge domain loops of kernels, thus optimizing the use of cache memories.
+
 Figure~\ref{fig:fusion} shows the number of iterations per second as a function of the number of cores with and without fusions.
-This benchmark is performed on FullSWOF2D onto a $500 \times 500$ domain size with $200$ time iterations.
-As explained in Section~\ref{sect:fusion}, the MSF loop fusion happens at a high level and is most of the time done naturally by a computer scientist.
-However, for a non computer scientist which write its numerical codes, an automatic proposition of such fusions avoid errors, particularly for a parallel execution.
-Moreover, one can notice that the performance is clearly improved (around 40\%) by this fusion.
+This benchmark is performed on FullSWOF2D onto a $500 \times 500$ domain size with $200$ time iterations, and by using data parallelism alone (without tasks).
+As explained in Section~\ref{sect:fusion}, the MSF loop fusion happens at a high level.
+Most of the time such fusions are done naturally by a computer scientist.
+However, for a numerician who usually writes equations more than numerical codes, an automatic detection of such fusions avoids errors, particularly for a parallel execution.
+In addition to this, more advanced fusion cases, such as a scatter are more difficult to deduce.
+In FullSWOF2D a total of  62 fusions are proposed by MSF over a total of 98 computation kernels.
+Fig.~\ref{fig:fusion} shows that the performance is clearly improved (around 40\%) by these fusions.
 
@@
 
+However, fusion optimizations are not always relevant.
+To illustrate this, we are using the same benchmark of FullSWOF2D onto a $500 \times 500$ domain size with $200$ time iterations, however we compare data parallelism and hybrid parallelism both with and without fusion.
+
+Blue curves of Fig.~\ref{fig:fusion1} represent results for data parallelism only with and without fusion.
+One can note that the best performance, as expected, is reached by the version using fusions.
+Green curves represent results by using 2 cores per MPI process dedicated for tasks, with and without fusion again.
+One can note that the best performance is also reached by the version using fusion.
+
 \begin{figure}[!h]\begin{center}
@@
 
+However, to deeper analyze this results, we propose a second evaluation presented in Fig.~\ref{fig:fusion2}.
+The blue curves are exactly the same one than in Fig.~\ref{fusion1}.
+The purple curves, on the other hand, represent results by using 8 cores per MPI process dedicated to tasks, with and without fusion.
+Interesting results appears in this figure as the hybrid version using fusions is less efficient than the one without fusions.
+As already explained, this result is due to the fact that fusions reduce the number of tasks from 98 to 36 resulting in a non optimized use of eight cores for task parallelism.
+By using only 2 cores per MPI process (in Fig.~\ref{fig:fusion1}) the 36 computation kernels were enough to feed the two cores, while it is not for eight.
+
 \begin{figure}[!h]\begin{center}
@@
 \end{center}\end{figure}
+
+As a result, if fusion optimization incurs a too large reduction of the number of tasks to feed dedicated cores, the problem observed for 16 cores in Fig.~\ref{fig:close} happens earlier which reduces obtained performances.

### 2017-03-03 ARTICLES/IJPP/articleR1/article.tex
corrections + evaluation en cours
abstract + titre

--- 2017-02-17 ARTICLES/IJPP/articleR1/article.tex
@@
 %----------------------------------------
-\title{Orchestrating stencils with the Multi-Stencil Framework}
+\title{The Multi-Stencil Framework: A Component-Based Approach to Orchestrate Stencils}
 %----------------------------------------
@@
 %----------------------------------------
-\institute{H\'el\`ene Coullon \and Christian Perez \at
-              INRIA, LIP, ENS Lyon, 46 Allée d'Italie, 69364 Lyon Cedex 07, France\\
-              \email{helene.coullon@inria.fr} \email{christian.perez@inria.fr}
+\institute{H\'el\`ene Coullon \at
+              DAPI IMT Atlantique, LS2N, Inria. Nantes, France\\
+              \email{helene.coullon@inria.fr}
 	\and
@@
            \email{julien.bigot@cea.fr}           
+           \and
+	Christian Perez \at
+              Univ. Lyon, Inria, CNRS, ENS de Lyon. Lyon, France\\
+              \email{christian.perez@inria.fr}
 }
@@
 %----------------------------------------
-\HC{change the title, improve it}
-\HC{is it possible to have authors on a single line ?}
 \begin{abstract}
-\HC{has to be revised according to modifications}
 As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
-One of the big challenges of exascale is to get programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists.
+One of the big challenges of exascale is to reach programming models that give access to high performance computing (HPC) to many scientists and not only to a few HPC specialists.
 One relevant solution to ease parallel programming for scientists is Domain Specific Language (DSL).
-However, one problem to avoid with DSLs is to not design a new DSL each time a new domain or a new problem has to be solved.
-This phenomenon happens for stencil-based numerical simulations, for which a large number of languages has been proposed without code reuse between them.
-The Multi-Stencil Language (MSL) presented in this paper is a language common to any kind of mesh used into a stencil-based numerical simulation.
-It is said that MSL is mesh-agnostic.
-Actually, from the description of a numerical simulation, MSL produces an empty parallel pattern, or skeleton, of the simulation which will be filled using other existing parallel languages and libraries.
-Thus, MSL, by finding a common language for different kinds of stencil-based simulation, facilitates code reuse.
-MSL is evaluated on a real case simulation which solves shallow-water equations.
-It is shown that MSL does not introduce overheads on data parallelism up to 16.384 cores, and that the hybrid parallelism (data and task) introduced improves performance of the simulation.
+However, one problem to avoid with DSLs is to mutualize existing codes and libraries instead of implemented from scratch.
+For example, this phenomenon occurs for stencil-based numerical simulations, for which a large number of languages has been proposed without code reuse between them.
+The Multi-Stencil Framework (MSF) presented in this paper combines a meta-model to component-based programming models to enhance code reuse and separation of concerns in the specific case of stencils.
+MSF can easily choose one parallelization technique or another as well as one optimization or another, as well as one back-end implementation or another.
+It is shown that MSF can reach same performances than a non-component-based MPI implementation over 16.384 cores.
+Finally, the performance model of the framework for hybrid parallelization is validated by evaluations.
 
-\keywords{Domain Specific Language (DSL) \and stencil \and numerical simulation \and mesh-agnostic \and data parallelism \and task parallelism \and scheduling \and MPI \and OpenMP}
+\keywords{Component programming models \and Domain Specific Language (DSL) \and Stencil \and Numerical simulation \and Data parallelism \and Task parallelism \and Scheduling \and MPI \and OpenMP}
 \end{abstract}
@@
 %------------------------------------------------------------------------------
-\section{Static Scheduling and performance model}
-\label{sect:msp}
-\input{msp.tex}
+%\section{Static Scheduling and performance model}
+%\label{sect:msp}
+%\input{msp.tex}
 %------------------------------------------------------------------------------
-\section{Implementation}
-\label{sect:comp}
-\input{compiler.tex}
+%\section{Implementation}
+%\label{sect:comp}
+%\input{compiler.tex}
 %------------------------------------------------------------------------------

### 2017-03-03 ARTICLES/IJPP/articleR1/conclusion.tex
abstract + titre

@@
 From these entries, the MSF compiler, namely MSC, is able to generate automatically a parallel specific assembly of the simulation.
-This specific assembly represents the parallel orchestration of computations independently of implementation choices.
+This specific assembly represents the parallel orchestration of computations, independently of implementation choices.
 Moreover, by combining a meta-model and component-based programming models MSF has the particularity to enhance separation of concerns and code-reuse of existing solutions (\eg SkelGIS and OpenMP in this paper).
@@
 Results have shown that even if component models have been used, strong and weak scalings of MSF (for its data parallelization technique) do not suffer from overheads.
-Results have also shown that the hybrid parallelization handled by MSF could increase performances when MPI communications are too heavy, which is new compared to existing solutions.
+Results have also shown that the hybrid parallelization handled by MSF could increase performances when MPI communications are too heavy, which is a new contribution compared to existing solutions.
 Finally, we have presented evaluations on fusion optimization that demonstrates that fusion is not always a good choice, particularly when performing hybrid parallelism.
 
-Many perspectives of future work raise from this work.
-First, even if using one distributed data structure or another is possible and facilitated by MSF, the HPC specialist still has to understand component programming models to add one.
+Many perspectives of future work raise from this paper.
+First, even if using one distributed data structure or another is possible and facilitated by MSF, the HPC specialist still has to understand component programming models and the meta-model of the framework to add new back-ends.
 We think it could be interesting to also facilitate the work of the HPC specialist by proposing a semi-automated framework to add external tools to MSF.
@@
 Thus, for the same reasons than parallelization techniques, an actor algorithm could choose the best solution for a given execution.
-To design such actor algorithms different solutions should be studied, such as following an approximated behavior model, or using setup runs to use deep learning algorithms etc.
+To design such actor algorithms different solutions should be studied, such as following an approximated behavior model, or using calibration runs to then use deep learning algorithms etc.
 

### 2017-03-03 ARTICLES/IJPP/articleR1/introduction.tex
abstract + titre

@@
-\HC{has to be revised according to modifications}
-
 As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
-For example, the current world's top supercomputer Tianhe-2~\footnote{\url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance around thirty peta floating point operations par seconds.
-However, to be able to use such a machine, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA etc., and multiple optimization techniques, such as cache optimization, have to be combined.
+For example, the current world's top supercomputer Tianhe-2~\footnote{\url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance around thirty peta floating point operations per second.
+However, to be able to use such machines, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA etc., and multiple optimization techniques, such as cache optimization, have to be combined.
 Moreover, current architectures evolution lets think that heterogeneity and complexity in HPC will continue to grow in future.
@@
 One possible runtime execution model for HPC is to propose dynamic scheduling of task graphs combined to message passing models~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359} (to be able to use more than one machine).
-Those models increase HPC code portability and reach an interesting performance onto heterogeneous architectures, which is interesting to reach exascale programming.
+These models increase HPC code portability and reach an interesting performance onto heterogeneous architectures, which is interesting to reach exascale programming.
 At a higher abstraction level, general purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} follow the direction of task graph scheduling proposed by those execution models.
@@
 Moreover, the compiler is able, because of the specific knowledge onto the targeted domain, to automatically apply parallelization and specific optimization to produce a high performance back-end code.
-Thus, a domain specific solutions are able to split end-user concerns from HPC concerns which is relevant to reach exascale programming models.
+Thus, domain specific solutions are able to distinguish end-user concerns from HPC concerns which is relevant to open HPC to non specialists.
 
@@
 
-We propose the Multi-Stencil Framework (MSF) that is built over a meta-model of multi-stencil simulations.
+We propose the Multi-Stencil Framework (MSF) that is built upon a meta-model of multi-stencil simulations.
 MSF produces a parallel orchestration of a multi-stencil program without being aware of the underlying implementation choices (\eg distributed data structures, task scheduler etc.).
@@
 Moreover, as the meta-model is independent from implementation details, MSF can easily choose one back-end or another, thus easing code reuse of existing solutions.
-To ease composition of existing solutions, MSF uses Software Engineering capabilities of component-based programming models~\cite{Szyperski:2002:CSB:515228} where applications are defined as an assembly of building blocks, or components.
+To ease composition of existing solutions, MSF uses Software Engineering capabilities of component-based programming models~\cite{Szyperski:2002:CSB:515228}, where applications are defined as an assembly of building blocks, or components.
 
-After a short overview of the Multi-Stencil Framework given in Section~\ref{sect:msf}, contributions of the paper are presented as follows: a meta-model of a multi-stencil program presented in Section~\ref{sect:formalism}; from this meta-model are built both a light and descriptive domain specific language, namely MSL, as well as a generic component assembly of the application described in Section~\ref{sect:msl}; the compiler MSC of the framework builds a parallel specialized version of the generic assembly from the application description with MSL, MSC and its performance model are described in Section~\ref{sect:parallelism}; finally a performance evaluation is detailed in Section~\ref{sect:eval}  and validates the performance model.
+After a short overview of the Multi-Stencil Framework given in Section~\ref{sect:msf}, the paper is organized as follows.
+A meta-model of a multi-stencil program is presented in Section~\ref{sect:formalism}; from this meta-model are built both a light and descriptive domain specific language, namely MSL, as well as a generic component assembly of the application described in Section~\ref{sect:msl}; the compiler MSC of the framework is described in Section~\ref{sect:parallelism}; finally a performance evaluation is detailed in Section~\ref{sect:eval} .
 

### 2017-03-03 ARTICLES/IJPP/articleR1/msl.tex
Relecture Section 4

@@
 %----------------------------------------
-From the meta-model represented by Equation (~\ref{eq:msp}) the Generic Assembly (GA) of any multi-stencil program can be built.
-It is represented by Figure~\ref{fig:ga}.
-As illustrated, the GA has five different components described below.
+From the meta-model represented by Equation~(\ref{eq:msp}) the Generic Assembly (GA) of any multi-stencil program can be built.
+As illustrated in Figure~\ref{fig:ga}, the GA has five components: {\bf \emph{Driver}, \emph{Time}, \emph{DDS}, \emph{Data}, } and {\bf \emph{Computations}}.
 
@@
 \paragraph{\textbf{Time}}
-As indicated in the figure, this component is responsible for the time $T$ defined in Equation (~\ref{eq:msp}). Thus it is composed of a time loop and eventually of a convergence reduction.
+As indicated in Figure~\ref{fig:ga}, this component is responsible for the time $T$ defined in Equation~(\ref{eq:msp}). Thus it is composed of a time loop and eventually of a convergence reduction.
 
 \paragraph{\textbf{DDS}}
-This component is responsible for the mesh and its entities $\mathcal{M}$, and $\Phi$, the computation domains $\mathcal{D}$ and the set of neighborhood functions $\mathcal{N}$.
+This component is responsible for the mesh and its entities $\mathcal{M}$ and $\Phi$, the computation domains $\mathcal{D}$, and the set of neighborhood functions $\mathcal{N}$.
 This component is generic, thus not typed.
-When GA will be instantiated and specialized by MSC, a specific type of mesh will replace this DDS meta-component.
-However, needed interfaces of such a DDS component are well known, and any component answering these interfaces can be indifferently used by MSF.
-A third party can propose new DDS components.
+When GA is instantiated and specialized by MSC, a specific type of mesh is replaced in this DDS meta-component \CP{pas clair}.
+However, needed interfaces of such a DDS component are well known, and any component providing these interfaces can be indifferently used by MSF: a third party can propose new DDS components.
 In this paper, both data and task parallelism are used.
-In the case of data parallelism this component should handle mesh partitioning and should provide a synchronization interface as it will be detailed in Section~\ref{sect:parallelism}.
-One can note that this component is red, this means that this component should be provided by a HPC specialist in MSF as illustrated in the previous Figure~\ref{fig:msf}.
+In the case of data parallelism this component needs to handle mesh partitioning and needs to provide a synchronization interface as it is detailed in Section~\ref{sect:parallelism}.
+One can note that this component is red, this means that this component should be provided by a HPC specialist in MSF as illustrated in Figure~\ref{fig:msf}.
 
@@
 One can note that this component is represented by a double line box.
-This means that it can be instantiated multiple time while others will be instantiated once.
-Actually, this component will be instantiated as many times as the number of quantities in the simulation.
+This means that it can be instantiated multiple time while others are instantiated once.
+Actually, this component is instantiated as many times as the number of quantities in the simulation.
 
@@
 Finally, this component is responsible for $\Gamma$, \ie the computations of the simulation.
-This component is green and blue because it will be partly generated by MSC and partly feeded by the developer.
-Actually, this component will be replaced by a sub-assembly of component, produced by MSC, for which the parallel part will be automatically generated.
-In this generated sub-assembly, components responsible for numerical codes will be completed by the developer.
+This component is green and blue because it is partly generated by MSC and partly feeded by the developer.
+Actually, this component is replaced by a sub-assembly of component, produced by MSC, for which the parallel part is automatically generated.
+In this generated sub-assembly, components responsible for numerical codes is completed by the developer.
 This sub-assembly generation is described in Section~\ref{sect:parallelism}.
@@
 
-The second element of MSF which is built upon the meta-model represented by Equation (~\ref{eq:msp}) is the Multi-Stencil Language and its grammar.
-This grammar is light and descriptive only, however it is sufficient (in addition to GA) for MSC to automatically extract a parallel pattern of the simulation, which is finally dumped as a specialized instantiation of GA.
+The second element of MSF which is built upon the meta-model represented by Equation~(\ref{eq:msp}) is the Multi-Stencil Language and its grammar.
+This grammar is light and descriptive only.
+However it is sufficient (in addition to GA) for MSC to automatically extract a parallel pattern of the simulation, which is finally dumped as a specialized instantiation of GA.
 
@@
 The grammar of the Multi-Stencil Language is given in Figure~\ref{fig:grammar} and an example is provided in Figure~\ref{fig:exmsl}.
-A Multi-Stencil program is composed of eight parts as in Equation (~\ref{eq:msp}).
+A Multi-Stencil program is composed of eight parts as in Equation~(\ref{eq:msp}).
 
@@
 
-\item The \texttt{mesh} keyword (Fig.\ref{fig:grammar},~l.1) introduces an identifier for $\mathcal{M}$, the single mesh of the simulation.
-For example \texttt{cart} in Fig.\ref{fig:exmsl},~l.1.
-The language, based on the meta-model (~\ref{eq:msp}), is independent of the mesh topology, thus this identifier is actually not used by the compiler but is syntactic sugar for the user.
+\item The \texttt{mesh} keyword (Fig.~\ref{fig:grammar},~l.1) introduces an identifier for $\mathcal{M}$, the single mesh of the simulation.
+For example \texttt{cart} in Fig.~\ref{fig:exmsl},~l.1.
+The language, based on the meta-model is independent of the mesh topology, thus this identifier is actually not used by the compiler but is syntactic sugar for the user.
 
-\item The \texttt{mesh entities} keyword (Fig.\ref{fig:grammar},~l.2) introduces identifiers for the groups of entities $\mathcal{G}\in\Phi$.
-For example \texttt{cell} or \texttt{edgex} in Fig.\ref{fig:exmsl},~l.2.
+\item The \texttt{mesh entities} keyword (Fig.~\ref{fig:grammar},~l.2) introduces identifiers for the groups of entities $\mathcal{G}\in\Phi$.
+For example \texttt{cell} or \texttt{edgex} in Fig.~\ref{fig:exmsl},~l.2.
 Again, this is syntactic sugar only as the compiler does not rely on the mesh topology.
 
-\item The \texttt{computation domains} keyword (Fig.\ref{fig:grammar},~l.3) introduces identifiers for the computation domains $D\in\mathcal{D}$.
-For example \texttt{d1} and \texttt{d2} in Fig.\ref{fig:exmsl},~l.4-5.
-For reference, each domain is associated to a group of entities (Fig.\ref{fig:grammar},~l.12) such as \texttt{cell} for \texttt{d1} in Fig.\ref{fig:exmsl},~l.4.
+\item The \texttt{computation domains} keyword (Fig.~\ref{fig:grammar},~l.3) introduces identifiers for the computation domains $D\in\mathcal{D}$.
+For example \texttt{d1} and \texttt{d2} in Fig.~\ref{fig:exmsl},~l.4-5.
+For reference, each domain is associated to a group of entities (Fig.~\ref{fig:grammar},~l.12) such as \texttt{cell} for \texttt{d1} in Fig.~\ref{fig:exmsl},~l.4.
 
-\item The \texttt{independent} keyword (Fig.\ref{fig:grammar},~l.4) offers a way to declare that computation domains do not intersect, such as \texttt{d1} and \texttt{d2} in  Fig.\ref{fig:exmsl},~l.7.
+\item The \texttt{independent} keyword (Fig.~\ref{fig:grammar},~l.4) offers a way to declare that computation domains do not intersect, such as \texttt{d1} and \texttt{d2} in  Fig.~\ref{fig:exmsl},~l.7.
 This is used by the compiler to compute dependencies between computations.
 
-\item The \texttt{stencil shapes} keyword (Fig.\ref{fig:grammar},~l.5) introduces identifiers for $n\in\mathcal{N}$, the stencil shapes.
-For each shape, the source and destination group of entities (Fig.\ref{fig:grammar},~l.16) are specified.
-For example \texttt{nec} in Fig.\ref{fig:exmsl},~l.11 is a neighborhood from \texttt{edgex} to \texttt{cell}.
+\item The \texttt{stencil shapes} keyword (Fig.~\ref{fig:grammar},~l.5) introduces identifiers for $n\in\mathcal{N}$, the stencil shapes.
+For each shape, the source and destination group of entities (Fig.~\ref{fig:grammar},~l.16) are specified.
+For example \texttt{nec} in Fig.~\ref{fig:exmsl},~l.11 is a neighborhood from \texttt{edgex} to \texttt{cell}.
 
-\item The \texttt{mesh quantities} keyword (Fig.\ref{fig:grammar},~l.6) introduces identifiers for $\delta\in\Delta$, the mesh variables with the group of entities they are mapped on (Fig.\ref{fig:grammar},~l.16).
+\item The \texttt{mesh quantities} keyword (Fig.~\ref{fig:grammar},~l.6) introduces identifiers for $\delta\in\Delta$, the mesh variables with the group of entities they are mapped on (Fig.~\ref{fig:grammar},~l.16).
 For example the quantities \texttt{C} and \texttt{H} are mapped onto \texttt{edgex} entities.
 
-\item The \texttt{scalars} keyword (Fig.\ref{fig:grammar},~l.7) introduces identifiers for $s\in\mathcal{S}$, the scalars. For example \texttt{mu} and \texttt{tau} in Fig.\ref{fig:exmsl},~l.15.
+\item The \texttt{scalars} keyword (Fig.~\ref{fig:grammar},~l.7) introduces identifiers for $s\in\mathcal{S}$, the scalars. For example \texttt{mu} and \texttt{tau} in Fig.~\ref{fig:exmsl},~l.15.
 
-\item Finally, the last part (Fig.\ref{fig:grammar},~l.8) introduces the different computation loops of the simulation. Each loop is made of two parts:
+\item Finally, the last part (Fig.~\ref{fig:grammar},~l.8) introduces the different computation loops of the simulation. Each loop is made of two parts:
 \begin{itemize}
-\item the \texttt{time} keyword (Fig.\ref{fig:grammar},~l.22) introduces $conv$, the convergence criteria either as a constant number of iterations or as a scalar variable (Fig.\ref{fig:grammar},~l.24).
-For example, 500 iterations are specified in Fig.\ref{fig:exmsl},~l.16,
-\item the \texttt{computations} keyword (Fig.\ref{fig:grammar},~l.23) introduces identifiers for each computation $k\in\Gamma=(S,R,(w,D),comp)$.
-Each computation (Fig.\ref{fig:grammar},~l.26) specifies:
+\item the \texttt{time} keyword (Fig.~\ref{fig:grammar},~l.22) introduces $conv$, the convergence criteria either as a constant number of iterations or as a scalar variable (Fig.~\ref{fig:grammar},~l.24).
+For example, 500 iterations are specified in Fig.~\ref{fig:exmsl},~l.16,
+\item the \texttt{computations} keyword (Fig.~\ref{fig:grammar},~l.23) introduces identifiers for each computation $k\in\Gamma=(S,R,(w,D),comp)$.
+Each computation (Fig.~\ref{fig:grammar},~l.26) specifies:
 \begin{itemize}
- \item the quantity $w$ written and its domain $D$, for example in Fig.\ref{fig:exmsl},~l.22, kernel \texttt{k4} computes the quantity \texttt{F} on domain \texttt{d1},
- \item the read scalars $S$ and mesh quantities with their associated stencil shape $R$, for example in Fig.\ref{fig:exmsl},~l.16, \texttt{k4} reads \texttt{C} with the shape \texttt{nce} and \texttt{D} with the default \textit{identity} shape, it does not read scalars.
+ \item the quantity $w$ written and its domain $D$, for example in Fig.~\ref{fig:exmsl},~l.22, kernel \texttt{k4} computes the quantity \texttt{F} on domain \texttt{d1},
+ \item the read scalars $S$ and mesh quantities with their associated stencil shape $R$, for example in Fig.~\ref{fig:exmsl},~l.16, \texttt{k4} reads \texttt{C} with the shape \texttt{nce} and \texttt{D} with the default \textit{identity} shape; it does not read scalars.
  \end{itemize}

### 2017-03-04 ARTICLES/IJPP/articleR1/introduction.tex
passe intro

@@
 As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
-For example, the current world's top supercomputer Tianhe-2~\footnote{\url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance around thirty peta floating point operations per second.
-However, to be able to use such machines, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA etc., and multiple optimization techniques, such as cache optimization, have to be combined.
-Moreover, current architectures evolution lets think that heterogeneity and complexity in HPC will continue to grow in future.
+For example, the current fastest supercomputer Tianhe-2~\footnote{\url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance of about thirty peta-flops (floating-point operations per second).
+However, to be able to use such machines, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA, etc., and multiple optimization techniques, such as cache optimization, have to be combined.
+Moreover, current architectures evolution seems to indicate that heterogeneity and complexity in HPC will continue to grow in the future.
 
-One of the big challenges to solve to be able to use exascale computers is to propose programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}.
-Actually, applications which run on supercomputers and which need such a computation power are physics, weather or genomic applications, which are not implemented by HPC specialists most of the time.
+One of the big challenges to solve to be able to use these upcoming Exascale computers is to propose programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}.
+Actually, applications that run on supercomputers and need such computation power (\textit{e.g.} physics, weather or genomic) are typically not implemented by HPC specialists but by domain scientists.
 
-One possible runtime execution model for HPC is to propose dynamic scheduling of task graphs combined to message passing models~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359} (to be able to use more than one machine).
-These models increase HPC code portability and reach an interesting performance onto heterogeneous architectures, which is interesting to reach exascale programming.
-At a higher abstraction level, general purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} follow the direction of task graph scheduling proposed by those execution models.
-However, for non-experts end-users, general purpose languages still are difficult to use and to tune for a given application onto a given architecture.
-The current easiest (with the higher abstraction level), but still efficient, programming model for end-users is Domain Specific Language (DSL) and domain specific frameworks.
-Such solutions are specific to the end-user domain and propose a grammar or an API which is easy to understand.
-Moreover, the compiler is able, because of the specific knowledge onto the targeted domain, to automatically apply parallelization and specific optimization to produce a high performance back-end code.
-Thus, domain specific solutions are able to distinguish end-user concerns from HPC concerns which is relevant to open HPC to non specialists.
+%JB: pourquoi tu parles de taches ?
+One possible runtime execution model for HPC is based on dynamic scheduling of task graphs combined with message passing~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359}.
+These models increase code portability and enable users to reach interesting performance on heterogeneous architectures.
+%JB: j'enleverais ce qui est au dessus et je commencerait tout de suite avec "OpenMP c'est compliqué"
+General purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} start to include these concepts.
+For non-expert users however, these languages are difficult to use and make it complex to tune applications for a given architecture. 
+An interesting approach that combines simplicity of use due to a high abstraction level with efficient execution consists in domain specific languages (DSL) and domain specific frameworks (DSF).
+These solutions are specific to a given domain and propose a grammar or an API which is easy to understand for specialists of this domain.
+Moreover, knowledge about the targeted domain can be embedded in the compiler that can thus automatically apply parallelization and optimizations techniques to produce high performance code.
+Domain specific solutions are therefore able to separate end-user concerns from HPC concerns which is a requirement to make HPC accessible to a wider audience.
 
-Many domain specific languages and frameworks have been proposed for many domains yet, and each one claims to solve a specific case that others don't.
-However, instead of re-using a common basis for every language or framework of a same domain, each solution is implemented from scratch as a monolithic code.
-We claim in this paper that being able to find a common meta-model to all solutions for a given domain facilitates code re-use, and eases flexibility and maintainability of solutions.
+Many domain specific languages and frameworks have been proposed that claim to handle distinct specific use-cases each.
+Each solution is however typically re-implemented from scratch.
+In this paper, we claim that the sharing of common building blocks when designing DSLs or DSFs would increases re-use, flexibility and maintainability in their implementation.
+%JB: J'ajoute cet argument, tu peux le retirer si tu trouves ça dangereux
+It would also ease the creation of approaches and applications combining multiple DSLs and DSFs.
 
-For example, to numerically solve a set of partial differential equations (PDEs), iterative methods are frequently used to approximate the exact solution through a discretized phenomena.
-A very well known and usual way to discretize PDEs is to transform them to explicit numerical schemes, also often called \emph{stencils}.
-Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107}, as it will be detailed in Section~\ref{sect:rel}.
-Many of them use same kind of parallelization, data structures or optimization techniques, however each one has been built from scratch.
+For example, some of the approaches to numerically solve partial differential equations (PDEs) lead to \emph{stencil computations} where the values associated to one point in space at a given time are computed from the values at the previous time at the exact same location together with a few neighbor locations.
+Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107} as will be detailed in Section~\ref{sect:rel}.
+Many of them use the same kind of parallelization, data structures or optimization techniques, however each one has been built from scratch.
 
@@
 Moreover, as the meta-model is independent from implementation details, MSF can easily choose one back-end or another, thus easing code reuse of existing solutions.
-To ease composition of existing solutions, MSF uses Software Engineering capabilities of component-based programming models~\cite{Szyperski:2002:CSB:515228}, where applications are defined as an assembly of building blocks, or components.
+To ease composition of existing solutions, MSF is based on component-based programming~\cite{Szyperski:2002:CSB:515228}, where applications are defined as an assembly of building blocks, or components.
 

### 2017-03-05 ARTICLES/IJPP/articleR1/msl.tex
passe Sec. 4

@@
 \vspace*{.5em}
-\caption{Generic Assembly according to the Multi-Stencil program formalism.}
+\caption{Generic Assembly according to the Multi-Stencil program formalism.
+Components circled by a double line identify those that will be instantiated multiple times at execution.
+Component colors represent the actor of Figure~\ref{fig:msf} responsible for the component implementation: green for those automatically generated by the compiler, red for those implemented by HPC specialists and blue for those implemented by the developer.}
 \label{fig:ga}
@@
 \paragraph{\textbf{Time}}
-As indicated in Figure~\ref{fig:ga}, this component is responsible for the time $T$ defined in Equation~(\ref{eq:msp}). Thus it is composed of a time loop and eventually of a convergence reduction.
+This component is responsible for the time $T$ defined in Equation~(\ref{eq:msp}). It is composed of a time loop and potentially of a convergence reduction.
 
@@
 This component is responsible for the mesh and its entities $\mathcal{M}$ and $\Phi$, the computation domains $\mathcal{D}$, and the set of neighborhood functions $\mathcal{N}$.
-This component is generic, thus not typed.
-When GA is instantiated and specialized by MSC, a specific type of mesh is replaced in this DDS meta-component \CP{pas clair}.
-However, needed interfaces of such a DDS component are well known, and any component providing these interfaces can be indifferently used by MSF: a third party can propose new DDS components.
+This component is generic; its type is a parameter of the generic assembly.
+When the generic assembly is instantiated and specialized by MSC, a component handling a specific type of mesh is used in place of the DDS placeholder.
+\CP{pas clair}
+The interfaces exposed by this component are however well defined and any component providing these interfaces can be indifferently used.
+Third party developers can therefore propose new DDS components.
 In this paper, both data and task parallelism are used.
-In the case of data parallelism this component needs to handle mesh partitioning and needs to provide a synchronization interface as it is detailed in Section~\ref{sect:parallelism}.
-One can note that this component is red, this means that this component should be provided by a HPC specialist in MSF as illustrated in Figure~\ref{fig:msf}.
+In the case of data parallelism this component handles mesh partitioning and provides a synchronization interface as detailed in Section~\ref{sect:parallelism}.
+The implementation of this component is the responsibility of HPC specialists.
 
 \paragraph{\textbf{Data}}
-This component is inseparable from the previous one $DDS$.
-It actually uses the DDS component to instantiate one mesh variable.
-Thus, it is responsible for the set of mesh variables $\Delta$.
-This component is also red.
-As $DDS$ it should be provided by a HPC specialist.
-One can note that this component is represented by a double line box.
-This means that it can be instantiated multiple time while others are instantiated once.
-Actually, this component is instantiated as many times as the number of quantities in the simulation.
+This component is responsible for the set of mesh variables $\Delta$.
+Each instance of the component uses the DDS component to handle one single mesh variable.
+It is closely related to $DDS$ and its implementation is typically provided by the same HPC specialist as $DDS$.
 
 \paragraph{\textbf{Computations}}
-Finally, this component is responsible for $\Gamma$, \ie the computations of the simulation.
-This component is green and blue because it is partly generated by MSC and partly feeded by the developer.
-Actually, this component is replaced by a sub-assembly of component, produced by MSC, for which the parallel part is automatically generated.
-In this generated sub-assembly, components responsible for numerical codes is completed by the developer.
-This sub-assembly generation is described in Section~\ref{sect:parallelism}.
+This component is responsible for $\Gamma$, \ie the computations of the simulation. 
+It is automatically replaced by a sub-assembly of components produced by MSC for which the parallel part is automatically generated and the part responsible for the numerical code is filled by the developer.
+The sub-assembly generation is described in Section~\ref{sect:parallelism}.
 
@@
 The grammar of the Multi-Stencil Language is given in Figure~\ref{fig:grammar} and an example is provided in Figure~\ref{fig:exmsl}.
-A Multi-Stencil program is composed of eight parts as in Equation~(\ref{eq:msp}).
+A Multi-Stencil program is composed of eight parts that match those of Equation~(\ref{eq:msp}).
 
@@
 The language, based on the meta-model is independent of the mesh topology, thus this identifier is actually not used by the compiler but is syntactic sugar for the user.
+\JB{J'aime pas trop l'utilisation de ``syntactic sugar'' ici...
+Syntactic sugar est generalement utilise pour un racourci de syntaxe (i.e.
+int x = myInteger au lieu de int x = myInteger.getIntVal()), ce n'est pas le cas ici}
 

### 2017-03-05 ARTICLES/IJPP/articleR1/parallelism.tex
passe Sec. 5

@@
-As previously explained, in a computation $k(S,R,(w,D),comp)$, $comp$ is not handled by MSF.
-Numerical codes are left in a second step to the developer by using the $DDS$ component selected during the MSF compilation.
-As a result, in the rest of this paper, and to simplify notations, we denote the same computation $k(S,R,(w,D))$.
+In a computation $k(S,R,(w,D),comp)$, the $comp$ part is provided by the developer after the MSC compilation phase.
+This part does therefore not have any impact on compilation concerns.
+Thus, to simplify notations in the rest of this paper, we use the shortcut notation $k(S,R,(w,D))$ instead of $k(S,R,(w,D),comp)$.
 
@@
 \label{sect:dataparal}
-In a data parallelization technique, the idea is to split data, or quantities, on which the program is computed into balanced sub-parts, one for each available resource.
-The same sequential program can afterwards be applied on each sub-part simultaneously, with some additional synchronizations between resources to update the data not computed locally, and thus to guarantee a correct result.
+In a data parallelization technique, the idea is to split data, or quantities, on which the program is computed into
+%JB: oui, certes... mais ca n'apporte rien comme info ici.
+% balanced 
+sub-domains, one for each execution resource.
+The same program is applied to each sub-domain simultaneously with some additional synchronizations to ensure coherence.
 
@@
 \end{equation*}
+consists in a partitioning of the mesh $\mathcal{M}$ in $p$ sub-meshes $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
+This step can be performed by an external graph partitioner~\cite{Pellegrini:1996:SSP:645560.658570,DBLP:conf/ieeehpcs/HeleneS13,lachat:hal-00768916} and is addressed by the third party DDS component instantiated from GA.
 
-consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
-This step can be performed by an external graph partitionner~\cite{Pellegrini:1996:SSP:645560.658570,DBLP:conf/ieeehpcs/HeleneS13,lachat:hal-00768916} and is addressed by the third party DDS component instantiated from GA.
+As entities and quantities are mapped on the mesh, the set of groups of mesh entities and the set of quantities $\Delta$ are partitioned the same way as the mesh: $\Phi=\{\Phi_0,\dots,\Phi_{p-1}\}$, $\Delta=\{\Delta_0,\dots,\Delta_{p-1}\}$.
 
-As entities and quantities are mapped onto the mesh, the set of groups of mesh entities and the set of quantities $\Delta$ are partitioned the same way than the mesh: $\Phi=\{\Phi_0,\dots,\Phi_{p-1}\}$, $\Delta=\{\Delta_0,\dots,\Delta_{p-1}\}$.
-
-The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma_{sync}$.
+The second step of the parallelization is to identify in $\Gamma$ the synchronizations required to update data.
+It leads to the construction of a new ordered list of computations $\Gamma_{sync}$.
 
@@
 
-Actually, a synchronization is needed by the quantity read by a stencil computation (not local), if this quantity has been written before.
+A synchronization is needed for the quantity read by a stencil computation (not local), if this quantity has been written before.
 This synchronization is needed because a neighborhood function $n \in \mathcal{N}$ of a stencil computation involves values computed on different resources.
@@
 \end{equation*}
+where the synchronization operation has been added.
 \end{mydef}
@@
 For this reason, to perform the final reduction, a set of synchronizations are needed to get the final reduced scalar.
-As most parallelism libraries (MPI, OpenMP) already propose a reduction synchronization with its own optimizations, we simply choose to replace the reduction computation by itself annotated by $red$.
+As most parallelism libraries (MPI, OpenMP) already propose a reduction synchronization with their own optimizations, we simply replace the reduction computation by itself annotated by $red$.
 
@@
 In this section we detail a static scheduling of $\Gamma_{dep}$ by using minimal series-parallel directed acyclic graphs.
-Such a static scheduling may not be the most efficient one, but it offers a simple fork/join task model which make possible the design of a performance model.
+Such a static scheduling may not be the most efficient one, but it offers a simple fork/join task model which makes possible the design of a performance model.
 Moreover, such a scheduling offers a simple way to propose a fusion optimization.
@@
 Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization.
-Even though the MSL language is not dedicated to produce very optimized independent stencil codes, but to produce the parallel orchestration of computations, building the $TSP$ tree make available an easy optimization when the data parallelization technique is the only one used.
+Even though the MSL language is not dedicated to produce very optimized independent stencil codes, but to produce the parallel orchestration of computations, building the $TSP$ tree makes available an easy optimization when the data parallelization technique is the only one used.
 This optimization consists in proposing a valid merge of some computation kernels inside a single space loop.
@@
 
-We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs for example~\cite{spaaTangCKLL11}.
-However, this fusion takes place at a different level and can bring performance improvements as it will be illustrated in Section~\ref{sect:eval}.
-This fusion algorithm relies on two statements:
-\begin{itemize}
-\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependent.
+We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs~\cite{spaaTangCKLL11} for example.
+However, this fusion takes place at a different level and can bring performance improvements as will be illustrated in Section~\ref{sect:eval}.
+This fusion algorithm relies on the following observations.
+
+%JB: des items sont normalement des parties de phrases, pas de point. Ici tu peux faire des paragraphes il me semble
+% \begin{itemize}
+% \item
+First, two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependent.
 As a result, what is written by the first one is read by the second one.
@@
 Thus, if the computation domains verify $D_i=D_j$, the fusion of $k_i$ and $k_j$ will decrease cache misses.
-\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependent.
+
+% \item 
+Second, two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependent.
 However, if the computation domains verify $D_i=D_j$, and if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}^{fus}$.
-\end{itemize}
+% \end{itemize}
 These two cases are illustrated by Fig.~\ref{fig:fus1} and Fig.~\ref{fig:fus2}.
@@
 
-One additional case of fusion can also happen. It is close to the first one above but is a little more difficult:
-Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependent.
-As a result, what is written by the first one is read by the second one.
-The construction of the tree also guarantee that synchronizations are not needed between these computations, otherwise a $k^{sync}$ would have been inserted between them (inherited from $\Gamma_{sync}$).
+Thirs, and similarly to the first observation, two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are data dependent and what is written by the first one is read by the second one.
+The construction of the tree also guarantees that synchronizations are not needed between these computations, otherwise a $k^{sync}$ would have been inserted between them (inherited from $\Gamma_{sync}$).
 Thus, $w_i$ the quantity written by $k_i$ is common to these computations.
@@
 MSC takes a MSL file written following the grammar described in Section~\ref{sect:msl}, as well as the Generic Assembly of components represented in Fig.~\ref{fig:ga} as inputs, and generates a specialized component assembly that manages the parallel orchestration of the computations of the simulation.
-In this final assembly, that could be compared to a pattern or a skeleton of the simulation, the developer still has to fill the functions corresponding to the various computation kernels by using the DDS instantiation chosen into the specialized assembly.
-The overall behavior of the compiler is as follow: %presented in Algorithm~\ref{alg:compiler}.
+In this final assembly, that could be compared to a pattern or a skeleton of the simulation, the developer still has to fill-in the functions corresponding to the various computation kernels by using the DDS instantiation chosen into the specialized assembly.
+The overall behavior of the compiler is as follows: %presented in Algorithm~\ref{alg:compiler}.
 \begin{enumerate}
@@
 A $K$ kernel is composed of $m$ use ports that are used to be connected to the $m$ quantities needed by the computation (\ie the numerical code).
-The component is also equipped with a provide port to be connected to the $Scheduler$ component.
+The component also exposes a provide port to be connected to the $Scheduler$ component.
 A generic $K$ component is represented in Fig.~\ref{fig:k}.
 
-A $Sync$ component is a static component (not generated) composed of a use-multiple port which is used to ask synchronizations to all quantities it is linked to ($Data$).
-The component is also equipped with a provide port to be connected to the $Scheduler$ component.
+A $Sync$ component is a static component (not generated) composed of a use-multiple port which is used to request synchronizations for all quantities it is linked to ($Data$).
+The component also exposes a provide port to be connected to the $Scheduler$ component.
 The $Sync$ component is represented in Fig.~\ref{fig:sync}.
@@
 Thus, this component represents the specific parallel orchestration of  computations.
-It is a heavy component with as much use ports as the number of $K$ components to call (\ie computations and fusions of computations).
-The component is also equipped with a provide port to be connected to the $Time$ component.
+It exposes as many use ports as there are instances of $K$ components to call (\ie computations and fusions of computations).
+The component also exposes a provide port to be connected to the $Time$ component.
 A generic $Scheduler$ component is represented in Fig.~\ref{fig:ass}.
+
+% draw the sepcialized assembly of the example
+To illustrate how a specialized assembly is generated, the specialized assembly of the example that has been used throughout this paper is represented in Fig.~\ref{fig:specass}.
 
@@
 \end{figure}
-
-% draw the sepcialized assembly of the example
-To illustrate how a specialized assembly is generated, the specialized assembly of the example that has been used throughout this paper is represented by Fig.~\ref{specass}.
-As this assembly is heavy to entirely draw, only a sub part of it is represented by the figure.
-Moreover some connections are represented by numbers instead of lines to not complexity the read.
 
@@
 \vspace*{.5em}
-\caption{Generic Assembly according to the Multi-Stencil program formalism.}
+\caption{Generic Assembly according to the Multi-Stencil program formalism.
+The complete assembly would be too complex to entirely draw and only a sub part is represented.
+Some connections are represented by numbers instead of lines in order not to make the schema more clear.
+\JB{etre plus clair sur les notations (Comme pour la Fig\ref{fig:ga}}}
 \label{fig:specass}
@@
 
-In this subsection are introduced two performance models, one for the data parallelization technique, and one for the hybrid data and task parallelization technique, both previously explained.
+In this subsection we introduce two performance models, one for the data parallelization technique, and one for the hybrid data and task parallelization technique, both previously explained.
 
-The performance model of a data parallelization technique is inspired from the Bulk Synchronous Parallel model.
-Actually data parallelization technique consider that each process has its own part of the domain.
-Thus the performance model reveals that the computation time is the sum of the reference sequential time divided by the number of processes, and of the time spent in communications between processes.
-Thus, for
+The performance model for the data parallelization technique is inspired by the Bulk Synchronous Parallel model.
+We consider that each process handles its own sub-domain that has been distributed in a perfectly balanced way.
+The performance model describes the computation time as the sum of the sequential time divided by the number of processes, and of the time spent in communications between processes. Thus, for
 \begin{itemize}
@@
 As a result, the critical point for performance is when $T_{COM} \geq \frac{T_{SEQ}}{P}$, which happens naturally in data parallelization as $T_{COM}$ will increase with the number of processes, and $\frac{T_{SEQ}}{P}$ decrease with the number of processes.
+\JB{Pour moi $T_{COM}$ est constant quand on communique entre voisins... Il diminue meme legerement si on reduit les donnees par process.}
 
 This limitation is always true, but can be delayed by different strategies.
-First, it is possible to overlap communications by performing computations simultaneously.
+First, it is possible to overlap communications and computations.
 Second, it is possible to introduce another kind of parallelization, task parallelization.
@@
 This paper does not propose an intelligent system to answer this question automatically, however, it offers a way to understand how to answer the question.
-To answer this question lets consider the two parallelization techniques, data only and hybrid.
+To answer this question let's consider the two parallelization techniques, data only and hybrid.
 We denote
@@
 
-This performance model will be validated and will help to explain results of Section~\ref{sect:eval}.
+This performance model will be validated and will help explain results of Section~\ref{sect:eval}.

### 2017-03-06 ARTICLES/IJPP/articleR1/introduction.tex
petites corrections

@@
 
-%JB: pourquoi tu parles de taches ?
+\JB{pourquoi tu parles de taches ?}
 One possible runtime execution model for HPC is based on dynamic scheduling of task graphs combined with message passing~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359}.
 These models increase code portability and enable users to reach interesting performance on heterogeneous architectures.
-%JB: j'enleverais ce qui est au dessus et je commencerait tout de suite avec "OpenMP c'est compliqué"
+\JB{j'enleverais ce qui est au dessus et je commencerait tout de suite avec "OpenMP c'est compliqué"}
 General purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} start to include these concepts.
@@
 In this paper, we claim that the sharing of common building blocks when designing DSLs or DSFs would increases re-use, flexibility and maintainability in their implementation.
-%JB: J'ajoute cet argument, tu peux le retirer si tu trouves ça dangereux
+\JB{J'ajoute cet argument, tu peux le retirer si tu trouves ça dangereux}
 It would also ease the creation of approaches and applications combining multiple DSLs and DSFs.

### 2017-03-06 ARTICLES/IJPP/articleR1/formalism.tex

@@
-Some approaches to numerically solve partial differential equations (PDEs) are based on direct iterative methods (\textit{e.g.} finite difference, finite volume or finite element methods).
-They approximate the solution through a discretized process where the continuous time and space domains are discretized and numerical computations are iteratively (time discretization) applied onto a mesh (space discretization).
-%In other words, the PDEs are transformed to a set of numerical computations applied at each time-step on elements of the discretized space domain.
-While the computations can have various forms, we focus on three categories:
-%TODO[JB] 4 plus tard...
-\emph{Stencil} computations involve access to neighbor values only (the concept of neighborhood depending on the space discretization used);
-\emph{Local} computations depend on the computed location only (this can be seen as a stencil of size one);
-Finally, \emph{reductions} enable to transform values mapped on the mesh to a single scalar value.
+%JB: on peut faire des differences/elements/volumes finis avec methodes directes ou indirectes, la difference entre les deux est est-ce que j'exprime la valeur en un point en fonction de
+% * points au pas de temps precedent seulement (direct, et du coup, souvent stencil) ou
+% * points au pas de temps courant aussi => resolution d'un systeme lineaire (indirect)
+% Concernant la discretisation, c'est la definition de resolution numérique (par opposition a analytique). Toutes les méthodes que je connais discrétisent en temps _et_ en espace...
 
-This section gives a complete formal description of what we call a \textit{stencil program} and its computations.
+The numerical solving of partial differential equations relies on the discretization of the continuous time and space domains.
+Computations are typically iteratively (time discretization) applied onto a mesh (space discretization).
+While the computations can have various forms, many direct methods can be expressed using three categories only:
+%JB: plus tard, tu parles de stencil, *boundary*, local & reduction
+\emph{stencil} computations involve access to neighbor values only (the concept of neighborhood depending on the space discretization used);
+\emph{local} computations depend on the computed location only (this can be seen as a stencil of size one);
+finally, \emph{reductions} enable to transform values mapped on the mesh to a single scalar value.
+
+This section gives a complete formal description of what we call a \textit{multi-stencil program} and its computations.
 A Multi-Stencil Program meta-model is extracted from this formal description.
@@
 
-Let introduce some notations.
+Let us introduce some notations.
 $\Omega$ is the continuous space domain of a numerical simulation (typically $\mathbb{R}^n$).
 %For example, if $n=3$, any geographic point on earth could be represented in this space in a given geographic coordinate system.
-A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ of a set of PDEs and is defined as follows.
+A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ and is defined as follows.
 
 \begin{mydef}
-A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V\subset \Omega$ is the set of vertices and $E\subseteq V^2$ the set of edges.
+A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V\subset \Omega$ is the (finite) set of vertices and $E\subseteq V^2$ the set of edges.
 The set of edges $E$ of a mesh $\mathcal{M}=(V,E)$ does not contain bridges.
@@
 \CP{Pas clair pour moi}
-\item The \textit{set of entities groups} used in a simulation is denoted $\Phi$.
+\item The \textit{set of all groups of entities} used in a simulation is denoted $\Phi$.
 \CP{Entities groups not defined! a propos, n'est pas set of entit*y* groups?}
@@
 
-For example, in a 2D Cartesian mesh, a group of entities where each entity is made of four vertices and four edges represents the cells.
-Another group of entities are the vertices defined as all singletons formed of a single vertex of $V$.
-Both groups, then, would be part of $\Phi$.
+For example, in a 2D Cartesian mesh, an entity could be a cell, made up of four vertices and four edges.
+A a group of entities could contain all the cells while another would for example contain the vertical egdes at the frontier between cells.
+Both groups would be part of $\Phi$.
 This example is illustrated in Figure~\ref{fig:meshbase}.
@@
 \begin{mydefs}[Definitions (quantity)]
-\item $\mathbb{V}=\Delta\cup\mathcal{S}$ is the set of \emph{variables} or \emph{quantities}.
 \item $\Delta$ are the \textit{mesh variables}.
-A mesh variable $\delta \in \Delta$ associates to each pair of entity and time-step a value $\delta: \mathcal{G}\times T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
+A mesh variable $\delta \in \Delta$ associates to each couple entity and time-step a value $\delta: \mathcal{G}\times T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
 \item The group of entities a variable is mapped on is denoted $entity(\delta)=\mathcal{G}$.\CP{Pas clair}
@@
 A scalar variable $s \in \mathcal{S}$ associates to each time-step a value $s: T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
+\item $\mathbb{V}=\Delta\cup\mathcal{S}$ is the set of \emph{variables} or \emph{quantities}.
 \item Among the scalar variables is one specific variable $conv\in\mathcal{S}$, the convergence criteria, whose value is $0$ except at the last step where it is 1.
 Thus, $\forall t\in \llbracket 0, T_{max-1} \rrbracket$, $conv(t)=0$, and $conv(T_{max})=1$.
-\CP{Il y a des implicites : est ce un calcul jusqu'à convergence ou un calcul sur MAX time step?
-Ou veux tu modéliser les deux (suivant que MAX est connu a priori ou a postiori?}
+This makes it possible to compute while not knowing $T_{max}$ \textit{a priori}.
+\CP{Il y a des implicites : est ce un calcul jusqu'à convergence ou un calcul sur MAX time step? Ou veux tu modéliser les deux (suivant que MAX est connu a priori ou a postiori?}
 \end{mydefs}
@@
 
-This section has presented the formalism of meshes, their associated entities, groups of entities,  and also time discretization and quantities (mapped on meshes or scalars).
+This section has presented the formalism of meshes, their associated entities, groups of entities as well as time discretization and quantities (mapped on meshes or scalars).
 
@@
 \item $S \in \mathcal{S}$ is the set of scalar to read,
-\item $(w,D) \in \mathbb{V} \times \mathcal{D}$ is the single data written by the kernel:
+\item $(w,D) \in \mathbb{V} \times \mathcal{D}$ is the single data written by the kernel:\JB{$\mathbb{V}$ contient les scalaires pour lesquels la notion de domaine de calcul n'est pas definie}
 \begin{itemize}
@@
 \begin{itemize}
-\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{stencil kernel} if $\exists (r,n) \in R$ such that $n \neq identity$ .
+\item A computation kernel $k(S,R,(w,D),comp)$ is a \emph{stencil kernel} if $\exists (r,n) \in R$ such that $n \neq identity$.
+That is, if the neighborhood it accesses is not reduced to the exact point it modifies.
 \item A \emph{boundary kernel} is a kernel $k(S,R,(w,D),comp)$ where $D$ is a specific computation domain at the border of entities, and which does not intersect with any other computation domain.
-\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{local kernel} if $\forall (r,n) \in R$, $n = identity$.
-\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} if $w$ is a scalar. A reduction is typically used to compute the convergence criteria of the time loop of the simulation.
+\item A computation kernel $k(S,R,(w,D),comp)$ is a \emph{local kernel} if $\forall (r,n) \in R$, $n = identity$.
+\item A computation kernel $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} if $w$ is a scalar.
+A reduction can for example be used to compute the convergence criteria of the time loop of the simulation.
 \end{itemize}
@@
 
-At this point, it could seem counter-intuitive to restrict the definition of a computation to a single quantity to write.
-This is actually counter-intuitive for a computer scientist who will naturally perform more than one computation into a loop to optimize cache memory use and to get a shorter code.
-The definition used above actually follows the usual "=" meaning of an equation or a numerical computation, where a single variable is written on the left of the symbol.
-We have made this choice for three reasons.
-\begin{itemize}
-\item First, we think this notation is more natural for a numerician who usually manipulate equations.
-\item Second, we call a \emph{fusion} the fact of writing more than one quantity into a single loop.
-When writing a sequential program, fusion is natural for the developer, however, writing a parallel code without knowing it (as it is by using MSF), fusion are error prone and a bad fusion could lead to an incorrect numerical result.
-\item Finally, when performing task parallelism, as it is in this paper, it is not always the good choice to perform fusions as you reduce the number of possible tasks.
-\end{itemize}
-For these reasons we have made the choice to restrict the computation definition without any fusion, and to propose an automatic loop fusion described in Section~\ref{sect:parallelism}.
-Moreover, one can note that the usual \emph{scatter} optimization of numerical codes is actually a specific case of fusion.
+It could seem counter-intuitive to restrict kernels to the computation of a single quantity.
+As a matter of fact, one often performs multiple computations in a single loop, for example for performance reasons (cache optimization, temporary computation factorization, etc.) or for readability (code length, semantically close computations, etc.).
+One can however notice that it is always possible to re-express a computation modifying $n$ quantities as $n$ computations modifying a single quantity each.
+Both approaches are therefore equivalent from an expressiveness point of view.
+
+%JB: l'argument me semble un peu specieux, en math on peut faire $(x,y,z) = ...$
+% The definition used above actually follows the usual "=" meaning of an equation or a numerical computation, where a single variable is written on the left of the symbol.
+% \item First, we think this notation is more natural for a numerician who usually manipulate equations.
+
+Modifying multiple quantities in a single loop nest does however not always improve performance.
+For example, it reduces the number of concurrent tasks available and limits the potential efficiency on parallel resources as will be shown in Section~\ref{sect:eval}.
+We therefore introduce the concept of \emph{fusion} in Section~\ref{sect:parallelism} where multiple logical kernels can be executed in a single loop nest that modifies multiple quantities.
+This transformation is much easier to implement than splitting a kernel would be, leaving more execution choices open.
+
+In addition, the modification of multiple quantities in a single loop nest can lead to subtle ordering errors when executing in parallel as will be discussed for the introduction of fusion.
+Automatically detecting kernels that can be fused instead of leaving this to the responsibility of the domain scientist avoids these potential errors.
+We have therefore chosen to restrict kernels to the computation of a single quantity as it is equivalent from an expressiveness point of view and potentially more efficient and safer.
+
+%JB: je garderais la discussion sur le scatter pour plus tard
+% Moreover, one can note that the usual \emph{scatter} optimization of numerical codes is actually a specific case of fusion.
 
@@
 \end{figure}
-For example, in Figure~\ref{fig:ex1}, assuming that the computation domain (full lines) is denoted $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined as:
+For example, in Figure~\ref{fig:ex1}, assuming that the computation domain (full lines) is denoted $dc1$ and the stencil shape described by the neighborhood function is $n1$, the stencil kernel can be defined as:
 \begin{equation*}
@@
 \medskip
-A stencil program has been formally defined in this section.
-This formalism is used in the next Section as a meta-model to define both a the domain specific language MSL, and the generic assembly of a multi-stencil program GA.
+We have formally defined stencil programs in this section.
+In the next section, we use this formalism as a meta-model to define both a the domain specific language MSL, and the generic assembly of a multi-stencil program GA.
+\JB{Plusieurs fois tu parles de meta-modele...
+C'est un peu different de la definition communement accepte plus orientee UML de meta-modele.
+Heureusement que le relecteur est pas de cette communautee :)}
 
-One can note that the meta-definition(~\ref{eq:msp}) is independent from the topology of the mesh.
+One can note that the meta-definition(\ref{eq:msp}) is independent from the topology of the mesh.
 This property will be kept in the rest of this paper so that MSF has the capacity to be used for any kind of mesh, thus being mesh-agnostic.

### 2017-03-06 ARTICLES/IJPP/articleR1/msf.tex
corrections + début intro

@@
 The multi-stencil domain will be formally defined in the next section.
-A multi-stencil program is an approximation of PDEs equations by one or multiple numerical computations, some of them using neighborhood values around an element, also called a \emph{stencil} computation.
+A multi-stencil program numerically solves PDEs using computations that can use neighborhood values around an element, also called a \emph{stencil} computation.
 
@@
 \begin{center}
-  \includegraphics[width=.4\textwidth]{./images/msf.pdf}
+  \includegraphics[width=.6\textwidth]{./images/msf.pdf}
   \caption{The Multi-Stencil Framework (MSF) composed of the Multi-Stencil Language (MSL), the Generic Assembly (GA), the Multi-Stencil Compiler (MSC) and which produces a specialized assembly of components.}
@@
 The Multi-Stencil Language, or MSL, is the domain specific language proposed by the framework for the numerician.
-It is a descriptive language, easy to use, without any concern onto implementation details, which fits the need of a mathematician to describe the simulation.
+It is a descriptive language, easy to use, without any concern about implementation details.
+It fits the need of a mathematician to describe the simulation.
 The description written with MSL can be considered as an input of the framework.
@@
 A GA could be compared to a generic template or a skeleton respectively in usual object programming languages such as C++ or in functional languages.
-From this generic assembly will be built the final specialized assembly of the simulation where component types will be precised, and some parts of GA will be transformed or added.
+From this generic assembly will be built the final specialized assembly of the simulation where component types will be specified, and some parts of GA will be transformed or added.
 As well as MSL, this generic assembly is described in Section~\ref{sect:msl} and is built upon the meta-model described in Section~\ref{sect:formalism}.

### 2017-03-06 ARTICLES/IJPP/articleR1/related.tex

@@
 % orchestration of stencils, different from stencil compilers like Pochoir and PATUS which could be used inside MSF
-First of all, many domain specific languages have been proposed for the optimization of a single stencil computation.
-Each one has its own optimization specificities and answers a specific numerical method or a specific kind of mesh.
-For example, Pochoir~\cite{spaaTangCKLL11} works on cache optimization techniques for stencils applied onto Cartesian meshes.
+Many domain specific languages have been proposed for the optimization of single stencil computations.
+Each one has its own optimization specificities and targets a specific numerical method or a specific kind of mesh.
+For example, Pochoir~\cite{spaaTangCKLL11} focuses on cache optimization techniques for stencils applied onto Cartesian meshes.
 On the other hand, PATUS~\cite{citeulike12258902} proposes to add a parallelization strategy grammar to its stencil language to perform an auto-tuning parallelization strategies.
@@
 Thus, these languages generalize the definition of a mesh, as it is proposed into the MSL formalism of Section~\ref{sect:formalism}.
-HOwever, neither Liszt, OP2 or Nabla handle hybrid parallelism as it is proposed by MSF.
+However, neither Liszt, OP2 nor Nabla handle hybrid parallelism as it is proposed by MSF.
+\JB{Manque la ref OP2}
 
@@
 Numerical codes are left to another language and another user (the developer in Figure~\ref{fig:msf}).
-Furthermore, such dataflow runtimes could actually be used by MSF as a backend, instead of using SkelGIS and OpenMP.
+Furthermore, such dataflow runtimes could actually be used by MSF as backends, instead of using SkelGIS or OpenMP.
+\JB{Manque de refs bis :)}
 
@@
 % no component-based framework for stencils orchestration (FFT, MapReduce/skeletons HLCM, nothing with DSLs)
-To conclude this related work, as far as we know no component-based framework have already been proposed for stencils orchestration.
+To conclude and as far as we know, no component-based framework has been proposed for stencils orchestration.

### 2017-03-06 ARTICLES/IJPP/articleR1/conclusion.tex
passe conclu

@@
-In this paper, we have presented the Multi-Stencil Framework.
-MSF is built upon a meta-model of a multi-stencil program which has been presented in Section~\ref{sect:formalism}.
-From this meta-model has been designed, first, the generic component assembly of a multi-stencil program, and second, the domain specific language MSL that enables the description of a specific application by a numerician.
-From these entries, the MSF compiler, namely MSC, is able to generate automatically a parallel specific assembly of the simulation.
-This specific assembly represents the parallel orchestration of computations, independently of implementation choices.
-Moreover, by combining a meta-model and component-based programming models MSF has the particularity to enhance separation of concerns and code-reuse of existing solutions (\eg SkelGIS and OpenMP in this paper).
-Finally, MSF has been evaluated on the real case simulation FullSWOF2D.
-Results have shown that even if component models have been used, strong and weak scalings of MSF (for its data parallelization technique) do not suffer from overheads.
-Results have also shown that the hybrid parallelization handled by MSF could increase performances when MPI communications are too heavy, which is a new contribution compared to existing solutions.
-Finally, we have presented evaluations on fusion optimization that demonstrates that fusion is not always a good choice, particularly when performing hybrid parallelism.
+In this paper, we have presented MSF, the multi-stencil framework.
+MSF is built upon a meta-model of a multi-stencil program which we have presented in Section~\ref{sect:formalism}.
+From this meta-model, we have designed, first, the generic component assembly of a multi-stencil program, and second, the domain specific language MSL that enables the description of a specific application by a numerician.
+From these entries, MSC, the MSF compiler, automatically generates a parallel component assembly.
+This assembly represents the parallel orchestration of computations, independently of implementation choices.
+
+By combining a meta-model and component-based programming, MSF has the particularity to enhance separation of concerns and code-reuse of existing solutions (\eg SkelGIS and OpenMP in this paper).
+MSF has been evaluated on the real case simulation FullSWOF2D.
+Results show that the MSF runtime does not induce unwanted overheads for its data parallelization technique on both strong and weak scalings.
+Results also show that the hybrid parallelization supported by MSF can increase performance when the data parallelism exposed by the simulation is not enough, which is a new contribution compared to existing solutions.
+Finally, we have evaluated the fusion optimization and shown that it often increases performance but is not always a good choice, particularly when performing hybrid parallelism.
 
@@
 
-The same choice problem is raised by fusion evaluations of Section~\ref{sect:fus}.
-Actually, fusion optimization is not always relevant.
-Thus, for the same reasons than parallelization techniques, an actor algorithm could choose the best solution for a given execution.
-To design such actor algorithms different solutions should be studied, such as following an approximated behavior model, or using calibration runs to then use deep learning algorithms etc.
+The same choice problem is raised by the fusion optimization.
+For the same reasons as for parallelization techniques, an actor algorithm could choose the best solution for a given execution.
+To design such actor algorithms different solutions should be studied, such as following an approximated behavior model, or using calibration runs to then use deep learning algorithms, etc.
 

### 2017-03-06 ARTICLES/IJPP/articleR1/article.tex
corrections abstract + intro

@@
 One relevant solution to ease parallel programming for scientists is Domain Specific Language (DSL).
-However, one problem to avoid with DSLs is to mutualize existing codes and libraries instead of implemented from scratch.
+However, one problem to avoid with DSLs is to mutualize existing codes and libraries instead of implemented each solution from scratch.
 For example, this phenomenon occurs for stencil-based numerical simulations, for which a large number of languages has been proposed without code reuse between them.
-The Multi-Stencil Framework (MSF) presented in this paper combines a meta-model to component-based programming models to enhance code reuse and separation of concerns in the specific case of stencils.
-MSF can easily choose one parallelization technique or another as well as one optimization or another, as well as one back-end implementation or another.
-It is shown that MSF can reach same performances than a non-component-based MPI implementation over 16.384 cores.
+The Multi-Stencil Framework (MSF) presented in this paper combines a new DSL to component-based programming models to enhance code reuse and separation of concerns in the specific case of stencils.
+MSF can easily choose one parallelization technique or another, one optimization or another, as well as one back-end implementation or another.
+It is shown that MSF can reach same performances than a non component-based MPI implementation over 16.384 cores.
 Finally, the performance model of the framework for hybrid parallelization is validated by evaluations.

### 2017-03-06 ARTICLES/IJPP/articleR1/introduction.tex
corrections abstract + intro

--- 2017-03-04 ARTICLES/IJPP/articleR1/introduction.tex
@@
 
-One of the big challenges to solve to be able to use these upcoming Exascale computers is to propose programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}.
+One of the big challenges to be able to use those upcoming Exascale computers is to propose programming models that give access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}.
 Actually, applications that run on supercomputers and need such computation power (\textit{e.g.} physics, weather or genomic) are typically not implemented by HPC specialists but by domain scientists.
 
-%JB: pourquoi tu parles de taches ?
-One possible runtime execution model for HPC is based on dynamic scheduling of task graphs combined with message passing~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359}.
-These models increase code portability and enable users to reach interesting performance on heterogeneous architectures.
-%JB: j'enleverais ce qui est au dessus et je commencerait tout de suite avec "OpenMP c'est compliqué"
-General purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} start to include these concepts.
-For non-expert users however, these languages are difficult to use and make it complex to tune applications for a given architecture. 
-An interesting approach that combines simplicity of use due to a high abstraction level with efficient execution consists in domain specific languages (DSL) and domain specific frameworks (DSF).
+%\JB{pourquoi tu parles de taches ?}
+%One possible runtime execution model for HPC is based on dynamic scheduling of task graphs combined with message passing~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359}.
+%These models increase code portability and enable users to reach interesting performance on heterogeneous architectures.
+%\JB{j'enleverais ce qui est au dessus et je commencerait tout de suite avec "OpenMP c'est compliqué"}
+%General purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} start to include these concepts.
+Many general purpose languages and frameworks have improved the simplicity of writing parallel codes.
+For example PGAS models~\cite{Nieplocha:2006:AAP:1125980.1125985} or task-based frameworks, such as OpenMP~\cite{660313}, Legion~\cite{bauer:legion:sc:2012} or StarPU~\cite{Augonnet2011}, partially hide intricate details of parallelism to the user.
+For non-expert users however, these languages and frameworks are still difficult to use.
+Moreover, tuning an application for a given architecture is still very complex to achieve with these solutions.
+An interesting approach that combines simplicity of use, due to a high abstraction level, with efficient execution are domain specific languages (DSL) and domain specific frameworks (DSF).
 These solutions are specific to a given domain and propose a grammar or an API which is easy to understand for specialists of this domain.
-Moreover, knowledge about the targeted domain can be embedded in the compiler that can thus automatically apply parallelization and optimizations techniques to produce high performance code.
+Moreover, knowledge about the targeted domain can be embedded in the compiler that can thus automatically apply parallelization and optimization techniques to produce high performance code.
 Domain specific solutions are therefore able to separate end-user concerns from HPC concerns which is a requirement to make HPC accessible to a wider audience.
 
-Many domain specific languages and frameworks have been proposed that claim to handle distinct specific use-cases each.
+Many domain specific languages and frameworks have been proposed. Each one claims to handle a distinct specific optimization or use case.
 Each solution is however typically re-implemented from scratch.
 In this paper, we claim that the sharing of common building blocks when designing DSLs or DSFs would increases re-use, flexibility and maintainability in their implementation.
-%JB: J'ajoute cet argument, tu peux le retirer si tu trouves ça dangereux
 It would also ease the creation of approaches and applications combining multiple DSLs and DSFs.
@@
 
-We propose the Multi-Stencil Framework (MSF) that is built upon a meta-model of multi-stencil simulations.
+We propose the Multi-Stencil Framework (MSF) that is built upon a meta-formalism of multi-stencil simulations.
 MSF produces a parallel orchestration of a multi-stencil program without being aware of the underlying implementation choices (\eg distributed data structures, task scheduler etc.).
-Thanks to this meta-model MSF is able to easily switch from one parallelization technique to another and from one optimization to another.
-Moreover, as the meta-model is independent from implementation details, MSF can easily choose one back-end or another, thus easing code reuse of existing solutions.
+Thanks to this meta-formalism MSF is able to easily switch from one parallelization technique to another and from one optimization to another.
+Moreover, as MSF is independent from implementation details, MSF can easily choose one back-end or another, thus easing code reuse of existing solutions.
 To ease composition of existing solutions, MSF is based on component-based programming~\cite{Szyperski:2002:CSB:515228}, where applications are defined as an assembly of building blocks, or components.
@@
 After a short overview of the Multi-Stencil Framework given in Section~\ref{sect:msf}, the paper is organized as follows.
-A meta-model of a multi-stencil program is presented in Section~\ref{sect:formalism}; from this meta-model are built both a light and descriptive domain specific language, namely MSL, as well as a generic component assembly of the application described in Section~\ref{sect:msl}; the compiler MSC of the framework is described in Section~\ref{sect:parallelism}; finally a performance evaluation is detailed in Section~\ref{sect:eval} .
+The meta-formalism of a multi-stencil program is presented in Section~\ref{sect:formalism}; from this formalism are built both a light and descriptive domain specific language, namely MSL, as well as a generic component assembly of the application described in Section~\ref{sect:msl}; the compiler MSC of the framework is described in Section~\ref{sect:parallelism}; finally a performance evaluation is detailed in Section~\ref{sect:eval} .
 

### 2017-03-06 ARTICLES/IJPP/articleR1/parallelism.tex
petites corrections
Relecture + questions

@@
-In a computation $k(S,R,(w,D),comp)$, the $comp$ part is provided by the developer after the MSC compilation phase.
+In a computation $k(S,R,(w,D),comp)$, the $comp$ part is provided by a developer after the MSC compilation phase.
 This part does therefore not have any impact on compilation concerns.
@@
 \label{sect:dataparal}
-In a data parallelization technique, the idea is to split data, or quantities, on which the program is computed into
+A classical data parallelim technique is to split data, or quantities, on which the program is computed into
 %JB: oui, certes... mais ca n'apporte rien comme info ici.
@@
 sub-domains, one for each execution resource.
-The same program is applied to each sub-domain simultaneously with some additional synchronizations to ensure coherence.
+The same program is applied to each sub-domain simultaneously with some synchronizations and data exchanges to ensure coherence.
 
@@
 \end{equation*}
-consists in a partitioning of the mesh $\mathcal{M}$ in $p$ sub-meshes $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
+on $p$ computing ressources consists in a partitioning of the mesh $\mathcal{M}$ in $p$ sub-meshes $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
 This step can be performed by an external graph partitioner~\cite{Pellegrini:1996:SSP:645560.658570,DBLP:conf/ieeehpcs/HeleneS13,lachat:hal-00768916} and is addressed by the third party DDS component instantiated from GA.
@@
 \label{def:sync}
+\CP{rappeler ce que sont $w_i$ et $r_j$ ?? Je les ai ajouté ci-dessous mais cela ne me semble pas suffisant. La def n'est pas self-contained.}
 \end{mydef}
 
-A synchronization is needed for the quantity read by a stencil computation (not local), if this quantity has been written before.
+A synchronization is needed for the quantity ($r_j$) read by a stencil computation (not local), if this quantity has been written before ($w_i$).
 This synchronization is needed because a neighborhood function $n \in \mathcal{N}$ of a stencil computation involves values computed on different resources.
 
-However, as a multi-stencil program is an iterative program, computations which happen after $k_j$ at the time iteration $t$ have also been computed before $k_j$ at the previous time iteration $t-1$.
+However, as a multi-stencil program is an iterative program, computations scheduled after $k_j$ at the time iteration $t$ have also been computed before it ($k_j$) at the previous time iteration $t-1$.
 For this reason another case of synchronization has to be defined.
@@
 Actually, the reduction is first applied locally on each subset of entities, on each resource.
-Thus, $p$ (number of resources) scalar values are obtained.
+Thus, $p$ (the number of resources) scalar values are obtained (one per ressource).
 For this reason, to perform the final reduction, a set of synchronizations are needed to get the final reduced scalar.
-As most parallelism libraries (MPI, OpenMP) already propose a reduction synchronization with their own optimizations, we simply replace the reduction computation by itself annotated by $red$.
+As most parallelism libraries (MPI, OpenMP) already propose a reduction synchronization with their own optimizations, we annotate reduction computation by $red$ to mark that a reduction function has to be applied.
 
@@
 One can notice that both types of synchronizations are performed by all resources.
+\CP{C'est quoi les 2 types de sync ?}
 
@@
 
-Algorithm~\ref{alg:sync} follows previous definitions to build a new ordered list which includes synchronizations.
+Algorithm~\ref{alg:sync}~\vpageref{alg:sync} follows the previous definitions to build a new ordered list which includes synchronizations.
 In this algorithm, lines 7 to 19 apply Definition~(\ref{def:sync}), lines 20 to 29 apply Definition~(\ref{def:sync2}), and finally lines 34 and 35 apply Definition~(\ref{def:red}).
-Finally, line 37 of the algorithm is the recursive call.
+Last, line 37 of the algorithm is the recursive call.
 
@@
 
- The final step of this parallelization is to run $\Gamma_{sync}$ on each resource. Thus, for each resource $0 \leq r \leq p-1$ the multi-stencil program 
+\medskip
+
+ The final step of the parallelization is to run $\Gamma_{sync}$ on each resource. Thus, for each resource $0 \leq r \leq p-1$ the multi-stencil program 
 \begin{equation}
@@
 
-\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl} gives an example of a $\mathcal{MSP}$ program. From this example, the following ordered list of computation kernels can be extracted:
+\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl}~\vpageref{fig:exmsl} gives an example of a $\mathcal{MSP}$ program.
+From this example, the following ordered list of computation kernels can be extracted:
 \begin{equation*}
@@
 \subsection{Hybrid parallelism}
-A task parallelization technique is a technique to transform a program as a dependency graph of different tasks.
-A dependency graph exhibits parallel tasks, or on the contrary sequential execution of tasks.
+A task parallelization technique aims at transforming a program into a dependency graph of tasks. A dependency graph exhibits parallel tasks, or on the contrary sequential execution of tasks.
+\CP{Phrase précédente pas claire}
+%
 Such a dependency graph can directly be given to a dynamic scheduler, or can statically be scheduled.
@@
 However, a specific condition on the computation domain, due to the specific domain of multi-stencils, is introduced for the write after write case.
-One can note that the \texttt{independent} keyword of Fig.~\ref{fig:grammar} is useful in this case.
+One can note that the \texttt{independent} keyword of Figure~\ref{fig:grammar} is useful in this case.
 
@@
 \begin{mydef}
-From the synchronized ordered list of computation kernels $\Gamma_{sync}$, the dependency graph of the computations $\Gamma_{dep}(V,A)$ is obtained from the call $F_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function defined in Algorithm~\ref{alg:dep}.
+From the synchronized ordered list of computation kernels $\Gamma_{sync}$, the dependency graph of the computations $\Gamma_{dep}(V,A)$ is obtained from the call $F_{dep}(\Gamma_{sync},0)$, where $F_{dep}$ is the recursive function defined in Algorithm~\ref{alg:dep}~\vpageref{alg:dep}.
 \end{mydef}
@@
 
-As a result of the hybrid parallelization, each resource $0 \leq r \leq p-1$ perform a multi-stencil program, defined by
+As a result of the hybrid parallelization, each resource $0 \leq r \leq p-1$ performs a multi-stencil program, defined by
 \begin{equation*}
@@
 
-\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl} gives an example of $\mathcal{MSP}$ program.
+\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl}~\vpageref{fig:exmsl} gives an example of a $\mathcal{MSP}$ program.
 From $\Gamma_{sync}$ that has been built in Equation~(\ref{eq:exsync}), the dependency DAG can be built.
@@
 Such a tree can be considered as a fork-join model and as a static scheduling.
-An example is given in Fig.~\ref{fig:solu}.
+An example is given in Figure~\ref{fig:solu}.
 
@@
 \begin{myth}
-The transitive reduction of a DAG $G$ is MSPD if and only if it does not contain $N$ as an induced sub-graph.
+The transitive reduction of a DAG $G$ is MSPD if and only if it does not contain $N$ as a sub-graph.
 \end{myth}
 
-As $\Gamma_{dep}$ is a DAG, by removing N-Shapes it is transformed to a MSPD.
-The intuition is illustrated in Fig.~\ref{fig:over}.
+As $\Gamma_{dep}$ is a DAG, all N-Shapes it contains must be removed to transform it into a MSPD.
+The intuition is illustrated in Figure~\ref{fig:over}.
 Considering the figure without the dashed line, the sub graph forms a "N" shape.
-The fact is that this shape cannot be represented as a composition of sequences or parallel executions.
+The fact is that this shape cannot be represented as a composition of sequences and parallel executions.
 To remove such forbidden N-shapes of $\Gamma_{dep}=(V,E)$, we have chosen to apply an over-constraint with the relation $k_0 \prec k_3$, such that a complete bipartite graph is created for the sub-dag as illustrated in Figure~\ref{fig:over}.
-By adding this arc to the DAG, it is possible to identify its execution as $sequence(parallel(k_0;k_2);parallel(k_1;k3))$ represented by the TSP tree of Fig.~\ref{fig:solu}.
+By adding this arc to the DAG, it is possible to identify its execution as $sequence(parallel(k_0;k_2);parallel(k_1;k3))$ represented by the TSP tree of Figure~\ref{fig:solu}.
 
@@
   \end{tikzpicture}
-\caption{TSP tree of Fig.~\ref{fig:over}.}
+\caption{TSP tree of Figure~\ref{fig:over}.}
 \label{fig:solu}
@@
 
-\paragraph{\textbf{Example}} From $\Gamma_{dep}$ illustrated in Fig.~\ref{fig:depdep} the TSP tree represented in Fig.~\ref{fig:tree} can be computed.
+\paragraph{\textbf{Example}} From $\Gamma_{dep}$ illustrated in Figure~\ref{fig:depdep}~on page~\pageref{fig:depdep}, the TSP tree represented in Figure~\ref{fig:tree}~\vpageref{fig:tree} can be computed.
 
@@
 Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization.
-Even though the MSL language is not dedicated to produce very optimized independent stencil codes, but to produce the parallel orchestration of computations, building the $TSP$ tree makes available an easy optimization when the data parallelization technique is the only one used.
+The MSL language is not dedicated to produce very optimized independent stencil codes, but to produce the parallel orchestration of computations.
+However, building the $TSP$ tree makes available an easy optimization when the data parallelization technique is the only one used.
 This optimization consists in proposing a valid merge of some computation kernels inside a single space loop.
@@
 We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs~\cite{spaaTangCKLL11} for example.
-However, this fusion takes place at a different level and can bring performance improvements as will be illustrated in Section~\ref{sect:eval}.
+However, this fusion takes place at a different level and can bring performance improvements as illustrated in Section~\ref{sect:eval}.
 This fusion algorithm relies on the following observations.
@@
 % \end{itemize}
-These two cases are illustrated by Fig.~\ref{fig:fus1} and Fig.~\ref{fig:fus2}.
+These two cases are illustrated by Figure~\ref{fig:fus1} and Figure~\ref{fig:fus2}.
 
@@
 
-Thirs, and similarly to the first observation, two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are data dependent and what is written by the first one is read by the second one.
+Third, and similarly to the first observation, two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are data dependent and what is written by the first one is read by the second one.
 The construction of the tree also guarantees that synchronizations are not needed between these computations, otherwise a $k^{sync}$ would have been inserted between them (inherited from $\Gamma_{sync}$).
@@
 The developer will be notified of fusions in the output of MSF.
-This particular case is illustrated in Fig.~\ref{fig:fus3}.
+This particular case is illustrated in Figure~\ref{fig:fus3}.
 
@@
 
-MSC takes a MSL file written following the grammar described in Section~\ref{sect:msl}, as well as the Generic Assembly of components represented in Fig.~\ref{fig:ga} as inputs, and generates a specialized component assembly that manages the parallel orchestration of the computations of the simulation.
+MSC takes as input a MSL file written following the grammar described in Section~\ref{sect:msl}, as well as the Generic Assembly of components represented in Figure~\ref{fig:ga} and a flag indicating the kind of paralelization selected (data or hybrid paralization).
+It generates a specialized component assembly that manages the parallel orchestration of the computations of the simulation.
 In this final assembly, that could be compared to a pattern or a skeleton of the simulation, the developer still has to fill-in the functions corresponding to the various computation kernels by using the DDS instantiation chosen into the specialized assembly.
@@
 % detailed each component (interfaces)
-New components have been introduced above and need to be explained.
-A $K$ component is a component into which the developer will write numerical code.
-It could represents a single computation kernel described by the numerician using MSL, or it could represents the fusion of multiple computation kernels.
-In any case the name of the generated component will use kernel identifiers used in the MSL description.
-A $K$ kernel is composed of $m$ use ports that are used to be connected to the $m$ quantities needed by the computation (\ie the numerical code).
+Let describe the newly introduced components.
+A $K$ component is a component into which a developer writes numerical code.
+It can be either a single computation kernel, or the fusion of multiple computation kernels.
+In any case the name of the generated component makes use of kernel identifiers from the MSL description.
+A $K$ kernel is composed of $m$ use ports that are connected to the $m$ quantities needed by the computation (\ie the numerical code).
 The component also exposes a provide port to be connected to the $Scheduler$ component.
-A generic $K$ component is represented in Fig.~\ref{fig:k}.
+A generic $K$ component is represented in Figure~\ref{fig:k}.
 
@@
 The component also exposes a provide port to be connected to the $Scheduler$ component.
-The $Sync$ component is represented in Fig.~\ref{fig:sync}.
+The $Sync$ component is represented in Figure~\ref{fig:sync}.
 
@@
 The component also exposes a provide port to be connected to the $Time$ component.
-A generic $Scheduler$ component is represented in Fig.~\ref{fig:ass}.
+A generic $Scheduler$ component is represented in Figure~\ref{fig:sched}.
 
 % draw the sepcialized assembly of the example
-To illustrate how a specialized assembly is generated, the specialized assembly of the example that has been used throughout this paper is represented in Fig.~\ref{fig:specass}.
+To illustrate how a specialized assembly is generated, the specialized assembly of the example used throughout this section is represented in Figure~\ref{fig:specass}.
 
@@
 
-In this subsection we introduce two performance models, one for the data parallelization technique, and one for the hybrid data and task parallelization technique, both previously explained.
+In this section we introduce a performance model for two used parallelization techniques, the data parallelization and the hybrid data and task parallelization.
 
@@
 \JB{Pour moi $T_{COM}$ est constant quand on communique entre voisins... Il diminue meme legerement si on reduit les donnees par process.}
+\CP{Pour moi $T_{COM}$ augmente avec P s'il y a au moins une op\'erations collectives (reduction/barrière) dont le cout est en log(P); Of course, s'il n'y a que des comm entre voisins, $T_{COM}$ tend vers la latence du réseau (quand il n'y a plus qu'une donnée par voisin)}
 
@@
 As a result when $P_{data}$ is small a data parallelization technique may be more efficient, while an hybrid parallelization could be interesting at some point to improve performance.
-The question is: when is it interesting to use hybrid parallelization ?
+The question is: when is it interesting to use hybrid parallelization?
 This paper does not propose an intelligent system to answer this question automatically, however, it offers a way to understand how to answer the question.
-To answer this question let's consider the two parallelization techniques, data only and hybrid.
+To answer this question let us consider the two parallelization techniques, data and hybrid.
 We denote
@@
 
-This performance model will be validated and will help explain results of Section~\ref{sect:eval}.
+This performance model is validated in Section~\ref{sect:eval} and it helps to explain some experimental results of Section~\ref{sect:eval}.

### 2017-03-06 ARTICLES/IJPP/articleR1/introduction.tex
relecture intro

@@
 As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important.
-For example, the current fastest supercomputer Tianhe-2~\footnote{\url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance of about thirty peta-flops (floating-point operations per second).
+For example, in the November 2016 TOP500 list\footnote{\url{www.top500.org}}, the fastest supercomputer Sunway TaihuLight~is composed of multi-cores processors and accelerators, and is able to reach a theoretical peak performance of about one hundred twenty five peta-flops (floating-point operations per second).
 However, to be able to use such machines, multiple programming models, such as MPI (Message Passing Interface), OpenMP, CUDA, etc., and multiple optimization techniques, such as cache optimization, have to be combined.
@@
 
-One of the big challenges to be able to use those upcoming Exascale computers is to propose programming models that give access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}.
-Actually, applications that run on supercomputers and need such computation power (\textit{e.g.} physics, weather or genomic) are typically not implemented by HPC specialists but by domain scientists.
+One of the big challenges to be able to use those supercomputers is to propose programming models that give access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}.
+Actually, applications that run on supercomputers and need such computation power (\textit{e.g.} physics, weather or genomics) are typically not implemented by HPC specialists but by domain scientists.
 
@@
 These solutions are specific to a given domain and propose a grammar or an API which is easy to understand for specialists of this domain.
-Moreover, knowledge about the targeted domain can be embedded in the compiler that can thus automatically apply parallelization and optimization techniques to produce high performance code.
+Moreover, knowledge about the targeted domain can be embedded in a DSL compiler that can thus automatically apply parallelization and optimization techniques to produce high performance code.
 Domain specific solutions are therefore able to separate end-user concerns from HPC concerns which is a requirement to make HPC accessible to a wider audience.
@@
 
-For example, some of the approaches to numerically solve partial differential equations (PDEs) lead to \emph{stencil computations} where the values associated to one point in space at a given time are computed from the values at the previous time at the exact same location together with a few neighbor locations.
+For example, some of the approaches to numerically solve partial differential equations (PDEs) lead to \emph{stencil computations} where the values associated to one point in space at a given time are computed from the values at the previous iteration at the exact same location together with a few neighbor locations.
 Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107} as will be detailed in Section~\ref{sect:rel}.
@@
 
-We propose the Multi-Stencil Framework (MSF) that is built upon a meta-formalism of multi-stencil simulations.
+In this paper, we propose the Multi-Stencil Framework (MSF) that is built upon a meta-formalism of multi-stencil simulations.
 MSF produces a parallel orchestration of a multi-stencil program without being aware of the underlying implementation choices (\eg distributed data structures, task scheduler etc.).
@@
 After a short overview of the Multi-Stencil Framework given in Section~\ref{sect:msf}, the paper is organized as follows.
-The meta-formalism of a multi-stencil program is presented in Section~\ref{sect:formalism}; from this formalism are built both a light and descriptive domain specific language, namely MSL, as well as a generic component assembly of the application described in Section~\ref{sect:msl}; the compiler MSC of the framework is described in Section~\ref{sect:parallelism}; finally a performance evaluation is detailed in Section~\ref{sect:eval} .
+The meta-formalism of a multi-stencil program is presented in Section~\ref{sect:formalism}.
+From this formalism are built both a light and descriptive domain specific language, namely MSL, as well as a generic component assembly of the application described in Section~\ref{sect:msl}.
+The compiler MSC of the framework is described in Section~\ref{sect:parallelism}.
+Finally a performance evaluation is detailed in Section~\ref{sect:eval} .
 

### 2017-03-06 ARTICLES/IJPP/articleR1/msf.tex

@@
 Most of the time numericians do have programming knowledge, however as it is not their core domain and because of a lack of time, development is often left to engineers according to numerician needs.
-MSF has the interesting particularity to propose a clear separation of concerns between these two end-users by distinguishing the description language from the numerical code.
-\CP{pas top clair la fin de la phrase}
+MSF has the interesting particularity to propose a clear separation of concerns between these two end-users by distinguishing the description of the simulation from the implementation of numerical codes.
 
@@
   \includegraphics[width=.6\textwidth]{./images/msf.pdf}
-  \caption{The Multi-Stencil Framework (MSF) composed of the Multi-Stencil Language (MSL), the Generic Assembly (GA), the Multi-Stencil Compiler (MSC) and which produces a specialized assembly of components.}
+  \caption{The Multi-Stencil Framework (MSF) is composed of the Multi-Stencil Language (MSL), the Generic Assembly (GA) and the Multi-Stencil Compiler (MSC) to produce a specialized assembly of components.
+The numerician, or mathematician uses MSL to describe its simulation.
+The developer will implement components responsible for numerical codes.
+A third party HPC specialist can interact with MSF to propose different version of HPC components.}
   \label{fig:msf}
@@
 MSL is described in details in Section~\ref{sect:msl}.
-The language is built upon the meta-model described in Section~\ref{sect:formalism}.
+The language is built upon the formalism described in Section~\ref{sect:formalism}.
 
 \paragraph{\textbf{Generic Assembly}}
-In addition to the language MSL and the simulation description written by the numerician, MSF needs a Generic Assembly (GA) of a multi-stencil program as input.
-What is called a GA is a component assembly for which meta-types are represented and for which some parts need to be generated or specialized.
-A GA could be compared to a generic template or a skeleton respectively in usual object programming languages such as C++ or in functional languages.
-From this generic assembly will be built the final specialized assembly of the simulation where component types will be specified, and some parts of GA will be transformed or added.
-As well as MSL, this generic assembly is described in Section~\ref{sect:msl} and is built upon the meta-model described in Section~\ref{sect:formalism}.
+In addition to the language MSL, used by the numerician to describe its simulation, MSF needs a Generic Assembly (GA) of a multi-stencil program as input.
+What is called a GA is a component assembly for which meta-types of components are represented and for which some parts need to be generated or specialized.
+A GA could be compared respectively to a template or a skeleton in object programming languages (such as C++) or functional languages.
+From this generic assembly will be built the final specialized assembly of the simulation where component types will be specified, and where parts of GA will be transformed.
+As well as MSL, this generic assembly is described in Section~\ref{sect:msl} and is built upon the meta-formalism described in Section~\ref{sect:formalism}.
 
@@
 Finally, the output of MSF is the component assembly generated by MSC.
-It is an instanciation and a transformation of the generic component assembly, by adding component types, transforming some part of the assembly, and by adding specific components generated by MSC.
-From this final component assembly which is specific to the simulation initially described with MSL, the developer will finally write numerical codes.
-This final specialized component assembly represents the parallel computations orchestration of the simulation initially described by the numerician.
-\CP{Dire que c'est du \llc?}
+It is an instantiation and a transformation of the generic component assembly, by adding component types, transforming some part of the assembly, and by adding specific components generated by MSC.
+From this final component assembly which is specific to the simulation initially described with MSL, the developer will finally write components associated to numerical codes, or directly re-use existing components from other simulations.
+This final specialized component assembly is a parallel orchestration of the computations of the simulation initially described by the numerician.
+Finally, the specialized assembly produced by MSF is written in \llc.

### 2017-03-06 ARTICLES/IJPP/articleR1/formalism.tex
petites corrections
corrections formalism.tex

@@
 \emph{local} computations depend on the computed location only (this can be seen as a stencil of size one);
-finally, \emph{reductions} enable to transform values mapped on the mesh to a single scalar value.
+finally, \emph{reductions} enable to transform variables mapped on the mesh to a single scalar value.
 
 This section gives a complete formal description of what we call a \textit{multi-stencil program} and its computations.
-A Multi-Stencil Program meta-model is extracted from this formal description.
-The meta-model will be used to define MSL and GA in the next section.
+This formalism is general enough to be common to any existing solution already proposed for stencil computations.
+As a result it can be considered as a meta-formalism or a meta-model of a Multi-Stencil Program.
+This meta-formalism will be used to define MSL and GA in the next section.
 
@@
 $\Omega$ is the continuous space domain of a numerical simulation (typically $\mathbb{R}^n$).
-%For example, if $n=3$, any geographic point on earth could be represented in this space in a given geographic coordinate system.
 A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ and is defined as follows.
@@
 \end{center}\end{figure}
-\begin{mydef}
-The dimension of a mesh $\mathcal{M}=(V,E)$ applied onto $\Omega=\mathbb{R}^n$ is denoted $dim(\mathcal{M})=n$.
-\end{mydef}
+%\begin{mydef}
+%The dimension of a mesh $\mathcal{M}=(V,E)$ applied onto $\Omega=\mathbb{R}^n$ is denoted $dim(\mathcal{M})=n$.
+%\end{mydef}
 A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) as illustrated in Figure~\ref{fig:mesh1}. 
-One can notice that more than one type of mesh is possible inside a single simulation.
+%One can notice that more than one type of mesh is possible inside a single simulation.
 For example, an hybrid mesh can be defined as an unstructured mesh composed itself of a Cartesian mesh inside each of its vertices.
@@
 \item An \textit{entity} $\phi$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $\phi\subset V\cup E$.
-\item A \textit{group of mesh entities} $\mathcal{G} \in \mathcal{P}(V\cup E)$ represents a set of entities of the same topology and the mesh of a group is denoted $mesh(\mathcal{G})$, \textit{i.e.} $\mathcal{G} \in \mathcal{P}(V\cup E) \Leftrightarrow mesh(\mathcal{G})=(V,E)$
-\CP{Pas clair pour moi}
-\item The \textit{set of all groups of entities} used in a simulation is denoted $\Phi$.
-\CP{Entities groups not defined! a propos, n'est pas set of entit*y* groups?}
+\item A \textit{group of mesh entities} $\mathcal{G} \in \mathcal{P}(V\cup E)$ represents a set of entities of the same topology.%and the mesh of a group is denoted $mesh(\mathcal{G})$, \textit{i.e.} $\mathcal{G} \in \mathcal{P}(V\cup E) \Leftrightarrow mesh(\mathcal{G})=(V,E)$
+%\CP{Pas clair pour moi}
+\item The \textit{set of all groups of mesh entities} used in a simulation is denoted $\Phi$.
+%\CP{Entities groups not defined! a propos, n'est pas set of entit*y* groups?}
 \end{mydefs}
@@
 For example, in a 2D Cartesian mesh, an entity could be a cell, made up of four vertices and four edges.
-A a group of entities could contain all the cells while another would for example contain the vertical egdes at the frontier between cells.
+A group of entities could contain all the cells, another would for example contain the vertical edges at the frontier between cells.
 Both groups would be part of $\Phi$.
 This example is illustrated in Figure~\ref{fig:meshbase}.
-\CP{Idem, pas top clair. Il y a des implicites !}
+%\CP{Idem, pas top clair. Il y a des implicites !}
 
@@
 The finite sequence $T: (t_n)_{n\in\llbracket 0, T_{max} \rrbracket}$ represents the discretization of the continuous time domain $\mathcal{T}=\mathbb{R}$.
-To each discrete time-step $n\in\llbracket 0, T_{max} \rrbracket$, a time value $t_n\in\mathcal{T}$ is associated. The time step for $t_n$ is equal to $t_n-t_{n-1}$.
+%To each discrete time-step $n\in\llbracket 0, T_{max} \rrbracket$, a time value $t_n\in\mathcal{T}$ is associated. The time step for $t_n$ is equal to $t_n-t_{n-1}$.
 \end{mydef}
@@
 The time discretization can be as simple as a constant time-step with a fixed number of steps.
-The time-step and the number of steps can also change on the fly during the simulation depending on variables values (see definitions below).
+The time-step and the number of steps can also change on the fly during execution. %the simulation depending on variables values (see definitions below).
 
@@
 A mesh variable $\delta \in \Delta$ associates to each couple entity and time-step a value $\delta: \mathcal{G}\times T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
-\item The group of entities a variable is mapped on is denoted $entity(\delta)=\mathcal{G}$.\CP{Pas clair}
+\item The group of entities a variable is mapped on is denoted $entity(\delta)=\mathcal{G}$.%\CP{Pas clair}
 \item $\mathcal{S}$ are the \textit{scalar variables}.
@@
 \item $\mathbb{V}=\Delta\cup\mathcal{S}$ is the set of \emph{variables} or \emph{quantities}.
-\item Among the scalar variables is one specific variable $conv\in\mathcal{S}$, the convergence criteria, whose value is $0$ except at the last step where it is 1.
-Thus, $\forall t\in \llbracket 0, T_{max-1} \rrbracket$, $conv(t)=0$, and $conv(T_{max})=1$.
+\item Among the scalar variables is one specific boolean variable $conv\in\mathcal{S}$, the convergence criteria, whose value is $0$ except at the last step where it is 1.
+This scalar can be updated on the fly according to other variables, typically by using a reduction as detailed later.%Thus, $\forall t\in \llbracket 0, T_{max-1} \rrbracket$, $conv(t)=0$, and $conv(T_{max})=1$.
 This makes it possible to compute while not knowing $T_{max}$ \textit{a priori}.
-\CP{Il y a des implicites : est ce un calcul jusqu'à convergence ou un calcul sur MAX time step? Ou veux tu modéliser les deux (suivant que MAX est connu a priori ou a postiori?}
+%\CP{Il y a des implicites : est ce un calcul jusqu'à convergence ou un calcul sur MAX time step? Ou veux tu modéliser les deux (suivant que MAX est connu a priori ou a postiori?}
 \end{mydefs}
-
-\medskip
-
-This section has presented the formalism of meshes, their associated entities, groups of entities as well as time discretization and quantities (mapped on meshes or scalars).
+%\medskip
+%
+%This section has presented the formalism of meshes, their associated entities, groups of entities as well as time discretization and quantities (mapped on meshes or scalars).
 
@@
 \item The set of computation domains of a numerical simulation is denoted $\mathcal{D}$.
-\item $\mathcal{N}$ are the neighborhood functions $n: \mathcal{G}_i \mapsto \mathcal{G}_j^m$ which for a given entity $\phi \in \mathcal{G}_i$ returns a set of $m$ entities in $\mathcal{G}_j$.
+\item $\mathcal{N}$ is the set of neighborhood functions $n: \mathcal{G}_i \mapsto \mathcal{G}_j^m$ which for a given entity $\phi \in \mathcal{G}_i$ returns a set of $m$ entities in $\mathcal{G}_j$.
 One can notice that $i = j$ is possible.
@@
 \item $S \in \mathcal{S}$ is the set of scalar to read,
-\item $(w,D) \in \mathbb{V} \times \mathcal{D}$ is the single data written by the kernel:\JB{$\mathbb{V}$ contient les scalaires pour lesquels la notion de domaine de calcul n'est pas definie}
-\begin{itemize}
-  \item $w$ is the single quantity (variable) modified by the computation kernel,
-  \item $D$ is the computation domain on which $w$ is computed, $D \subseteq entity(w)$,
-\end{itemize}
-\item $R \in \Delta \times \mathcal{N}$ is the set of tuples $(r,n)$ representing the data read where
-\begin{itemize}
-  \item $r$ is a mesh variable read by the kernel to compute $w$,
-  \item $n : entity(w) \rightarrow entity(r)^m$ is a neighborhood function that indicates which values of $r$ are read to compute $w$.
-\end{itemize}
-\item $comp$ is the numerical computation which returns a value from a set of $n$ input values, $comp: \mathcal{V}_i^n \rightarrow \mathcal{V}_j$, where $\mathcal{V}_i$ and $\mathcal{V}_j$ are value types.
-Thus, $comp$ represents the actual numerical expression which is computed by a kernel.
+\item $w \in \mathbb{V}$ is the single quantity (variable) modified by the computation kernel,
+\item $D$ is the computation domain on which $w$ is computed, $D \subseteq entity(w)$, or is $null$ if $w \in \Delta$,
+\item $R \in \Delta \times \mathcal{N}$ is the set of tuples $(r,n)$ representing the data read where $r$ is a mesh variable read by the kernel to compute $w$, and $n : entity(w) \rightarrow entity(r)^m$ is a neighborhood function that indicates which entity of $r$ are read to compute $w$.
+\item Finally, $comp$ is the numerical computation which returns a value from a set of $n$ input values, $comp: \mathcal{V}_i^n \rightarrow \mathcal{V}_j$, where $\mathcal{V}_i$ and $\mathcal{V}_j$ are value types.
+Thus, $comp$ represents the actual numerical expression  computed by a kernel.
 \end{itemize}
@@
 During a computation kernel, it can be considered that a set of old states ($t-1$) of quantities are read ($\mathcal{S}$ and $R$), and that a new state ($t$) of a single quantity is written ($w$).
-
 Such a definition of a computation kernel covers a large panel of different computations.
@@
 \item A computation kernel $k(S,R,(w,D),comp)$ is a \emph{stencil kernel} if $\exists (r,n) \in R$ such that $n \neq identity$.
-That is, if the neighborhood it accesses is not reduced to the exact point it modifies.
+%That is, if the neighborhood it accesses is not reduced to the exact point it modifies.
 \item A \emph{boundary kernel} is a kernel $k(S,R,(w,D),comp)$ where $D$ is a specific computation domain at the border of entities, and which does not intersect with any other computation domain.
@@
 
-In addition, the modification of multiple quantities in a single loop nest can lead to subtle ordering errors when executing in parallel as will be discussed for the introduction of fusion.
+In addition, the modification of multiple quantities in a single loop nest can lead to subtle ordering errors when executing in parallel as it will be discussed in Section~\ref{sect:fusion}.
 Automatically detecting kernels that can be fused instead of leaving this to the responsibility of the domain scientist avoids these potential errors.
-We have therefore chosen to restrict kernels to the computation of a single quantity as it is equivalent from an expressiveness point of view and potentially more efficient and safer.
+We have therefore chosen to restrict kernels to the computation of a single quantity.
 
-%JB: je garderais la discussion sur le scatter pour plus tard
-% Moreover, one can note that the usual \emph{scatter} optimization of numerical codes is actually a specific case of fusion.
+%\JB{je garderais la discussion sur le scatter pour plus tard}
+%\JB{Moreover, one can note that the usual \emph{scatter} optimization of numerical codes is actually a specific case of fusion.}
 
@@
 \begin{center}
-\subfloat[Mesh and two groups of mesh entities.\label{fig:meshbase}]{
+\subfloat[On the left a mesh is represented, on the right two examples of groups of mesh entities are represented: cells and vertical edges.\label{fig:meshbase}]{
 \resizebox{8cm}{!}{\includegraphics{./images/mesh.pdf}}
@@
 
-\medskip
-We have formally defined stencil programs in this section.
-In the next section, we use this formalism as a meta-model to define both a the domain specific language MSL, and the generic assembly of a multi-stencil program GA.
-\JB{Plusieurs fois tu parles de meta-modele...
+In this section, we have formally defined a stencil program.
+This formalism is mainly composed of a mesh abstraction and a simple definition of computation.
+In fact, this formalism is generic enough to be common to many existing modelizations of a stencil computation or a stencil simulation.
+Thus, the formalism summarized by Equation (\ref{eq:msp}) can be compared to a meta-model of a multi-stencil program.
+In the next section, we use this meta-model (or meta-formalism) to define both a the domain specific language MSL, and the generic assembly of a multi-stencil program GA.
+%\JB{Plusieurs fois tu parles de meta-modele...
 C'est un peu different de la definition communement accepte plus orientee UML de meta-modele.
@@
 
-One can note that the meta-definition(\ref{eq:msp}) is independent from the topology of the mesh.
+%Finally, one can note that the meta-formalism is independent from the topology of the mesh.
 This property will be kept in the rest of this paper so that MSF has the capacity to be used for any kind of mesh, thus being mesh-agnostic.

### 2017-03-06 ARTICLES/IJPP/articleR1/conclusion.tex
WIP relecture

@@
-In this paper, we have presented MSF, the multi-stencil framework.
+In this paper, we have presented MSF, a multi-stencil framework.
 MSF is built upon a meta-model of a multi-stencil program which we have presented in Section~\ref{sect:formalism}.
@@
 This assembly represents the parallel orchestration of computations, independently of implementation choices.
+Two parallelization strategies are supported: data parallelization and hybrid (data and task) parallelization.
 
@@
 Results also show that the hybrid parallelization supported by MSF can increase performance when the data parallelism exposed by the simulation is not enough, which is a new contribution compared to existing solutions.
-Finally, we have evaluated the fusion optimization and shown that it often increases performance but is not always a good choice, particularly when performing hybrid parallelism.
+Finally, we have evaluated the fusion optimization and shown that it often increases performance but it is not always a good choice, particularly when performing hybrid parallelism.
 
@@
 Second, it has been shown in the performance model and evaluation results that the choice between data parallelism and hybrid parallelism in not trivial.
-It depends on the time spent in computations, the number of cores and nodes used for the execution, as well as the network behavior (Fig.~\ref{fig:limit} and~\ref{fig:close}).
+It depends on the time spent in computations, the number of cores and nodes used for the execution, as well as the network behavior (Figure~\ref{fig:limit} and~\ref{fig:close}).
 MSF, thanks to its meta-model and to the use of component models, handles the activation of one parallelism technique or another.
 Thus MSF can be considered as a performance leverage for multi-stencil programs.
-However, it would be interesting to investigate how an actor algorithm could choose the best parallelization technique for a given execution of a production multi-stencil simulation.
+However, it would be interesting to investigate how an algorithm could choose the best parallelization technique for a given execution of a production multi-stencil simulation.
 
 The same choice problem is raised by the fusion optimization.
-For the same reasons as for parallelization techniques, an actor algorithm could choose the best solution for a given execution.
-To design such actor algorithms different solutions should be studied, such as following an approximated behavior model, or using calibration runs to then use deep learning algorithms, etc.
+For the same reasons as for parallelization techniques, an algorithm could choose the best solution for a given execution.
+To design such an algorithm different solutions should be studied, such as following an approximated behavior model, or using calibration runs to then use deep learning algorithms, etc.
 

### 2017-03-06 ARTICLES/IJPP/articleR1/formalism.tex
corrections formalism.tex + msl.tex

@@
 \end{center}
-\caption{(a) a Cartesian mesh and two kind of groups of mesh entities, (b) an example of stencil kernel on cells, (c) an example of stencil kernel on two different entities of the mesh.}
+\caption{(a) a Cartesian mesh and two kind of groups of mesh entities, (b) an example of stencil kernel on cells, (c) an example of stencil kernel on two different groups of mesh entities.}
 \label{fig:gspmsp}

### 2017-03-06 ARTICLES/IJPP/articleR1/msl.tex
corrections formalism.tex + msl.tex

@@
+From the octuplet of Equation~(\ref{eq:msp}) both the Generic Assembly (GA) of a multi-stencil program and the Multi-Stencil Language (MSL) can be built.
+GA and MSL are both are described in this section.
+
 %----------------------------------------
@@
 %----------------------------------------
-From the meta-model represented by Equation~(\ref{eq:msp}) the Generic Assembly (GA) of any multi-stencil program can be built.
 As illustrated in Figure~\ref{fig:ga}, the GA has five components: {\bf \emph{Driver}, \emph{Time}, \emph{DDS}, \emph{Data}, } and {\bf \emph{Computations}}.
+These components are generic components or abstract components.
+It means that interfaces of these components are well defined but that they are not implemented yet in GA.
+They can be compared to abstract classes and templates in \texttt{C++} for which an implementation must be given as well as specific parameters.
 
@@
 \caption{Generic Assembly according to the Multi-Stencil program formalism.
-Components circled by a double line identify those that will be instantiated multiple times at execution.
-Component colors represent the actor of Figure~\ref{fig:msf} responsible for the component implementation: green for those automatically generated by the compiler, red for those implemented by HPC specialists and blue for those implemented by the developer.}
+Components circled by a double line identify those that will be instantiated multiple times by MSC.
+Component colors represent actors of Figure~\ref{fig:msf} responsible for the component implementation: green for those automatically generated by the compiler, red for those implemented by HPC specialists and blue for those implemented by the developer.}
 \label{fig:ga}
@@
 This component can be compared to the main function of a usual program.
-It is responsible for both the initialization and the execution of other components (like variable initializations and function calls).
+It is responsible for both the initialization and the execution of other components (like variable initialization and function calls).
+This component is generated by MSC (represented in green).
 
 \paragraph{\textbf{Time}}
-This component is responsible for the time $T$ defined in Equation~(\ref{eq:msp}). It is composed of a time loop and potentially of a convergence reduction.
+This component is responsible for the time $T$ defined in Equation~(\ref{eq:msp}).
+It is composed of a time loop and potentially of a convergence reduction.
+This component is generated by MSC (represented in green).
 
 \paragraph{\textbf{DDS}}
-This component is responsible for the mesh and its entities $\mathcal{M}$ and $\Phi$, the computation domains $\mathcal{D}$, and the set of neighborhood functions $\mathcal{N}$.
-This component is generic; its type is a parameter of the generic assembly.
-When the generic assembly is instantiated and specialized by MSC, a component handling a specific type of mesh is used in place of the DDS placeholder.
-\CP{pas clair}
-The interfaces exposed by this component are however well defined and any component providing these interfaces can be indifferently used.
-Third party developers can therefore propose new DDS components.
+This component is responsible for the mesh and its entities $\mathcal{M}$ and $\Phi$, the set of computation domains $\mathcal{D}$, and the set of neighborhood functions $\mathcal{N}$.
+When the generic assembly is instantiated and specialized by MSC, an implementation of DDS is selected to handle a specific type of mesh.
+%\CP{pas clair}
+The interfaces exposed by this component are well defined and any component providing these interfaces can be indifferently used.
+A third party specialist can therefore propose new implementation of DDS.
 In this paper, both data and task parallelism are used.
-In the case of data parallelism this component handles mesh partitioning and provides a synchronization interface as detailed in Section~\ref{sect:parallelism}.
-The implementation of this component is the responsibility of HPC specialists.
+In the case of data parallelism DDS handles mesh partitioning and provides a synchronization interface as detailed in Section~\ref{sect:parallelism}.
+The implementation of this component is the responsibility of HPC specialists (represented in red).
 
@@
 Each instance of the component uses the DDS component to handle one single mesh variable.
-It is closely related to $DDS$ and its implementation is typically provided by the same HPC specialist as $DDS$.
+It is closely related to $DDS$ and its implementation is typically provided by the same HPC specialist as $DDS$ (represented in red).
 
@@
 This component is responsible for $\Gamma$, \ie the computations of the simulation. 
-It is automatically replaced by a sub-assembly of components produced by MSC for which the parallel part is automatically generated and the part responsible for the numerical code is filled by the developer.
+It is automatically replaced by a sub-assembly of components produced by MSC for which the parallel part is automatically generated.
+On the other hand, components responsible for the numerical kernels are filled by the developer.
+This is why this component is represented in green and blue in Fig.~\ref{fig:ga}.
 The sub-assembly generation is described in Section~\ref{sect:parallelism}.
@@
 For example \texttt{cart} in Fig.~\ref{fig:exmsl},~l.1.
-The language, based on the meta-model is independent of the mesh topology, thus this identifier is actually not used by the compiler but is syntactic sugar for the user.
-\JB{J'aime pas trop l'utilisation de ``syntactic sugar'' ici...
+The language, based on the meta-model is independent of the mesh topology, thus this identifier is actually not used by the compiler.% but is syntactic sugar for the user.
+%\JB{J'aime pas trop l'utilisation de ``syntactic sugar'' ici...
 Syntactic sugar est generalement utilise pour un racourci de syntaxe (i.e.
@@
 
-\item The \texttt{mesh entities} keyword (Fig.~\ref{fig:grammar},~l.2) introduces identifiers for the groups of entities $\mathcal{G}\in\Phi$.
+\item The \texttt{mesh entities} keyword (Fig.~\ref{fig:grammar},~l.2) introduces identifiers for the groups of mesh entities $\mathcal{G}\in\Phi$.
 For example \texttt{cell} or \texttt{edgex} in Fig.~\ref{fig:exmsl},~l.2.
-Again, this is syntactic sugar only as the compiler does not rely on the mesh topology.
+%Again, this is syntactic sugar only as the compiler does not rely on the mesh topology.
 
@@
 
-\item The \texttt{stencil shapes} keyword (Fig.~\ref{fig:grammar},~l.5) introduces identifiers for $n\in\mathcal{N}$, the stencil shapes.
-For each shape, the source and destination group of entities (Fig.~\ref{fig:grammar},~l.16) are specified.
+\item The \texttt{stencil shapes} keyword (Fig.~\ref{fig:grammar},~l.5) introduces identifiers for each stencil shape $n\in\mathcal{N}$.
+For each $n$, the source and destination group of entities (Fig.~\ref{fig:grammar},~l.16) are specified.
 For example \texttt{nec} in Fig.~\ref{fig:exmsl},~l.11 is a neighborhood from \texttt{edgex} to \texttt{cell}.
@@
 \item The \texttt{mesh quantities} keyword (Fig.~\ref{fig:grammar},~l.6) introduces identifiers for $\delta\in\Delta$, the mesh variables with the group of entities they are mapped on (Fig.~\ref{fig:grammar},~l.16).
-For example the quantities \texttt{C} and \texttt{H} are mapped onto \texttt{edgex} entities.
+For example the quantities \texttt{C} and \texttt{H} are mapped onto the groups of mesh entities \texttt{edgex}.
 
@@
 \begin{itemize}
-\item the \texttt{time} keyword (Fig.~\ref{fig:grammar},~l.22) introduces $conv$, the convergence criteria either as a constant number of iterations or as a scalar variable (Fig.~\ref{fig:grammar},~l.24).
+\item the \texttt{time} keyword (Fig.~\ref{fig:grammar},~l.22) introduces either a constant number of iterations or $conv$, the convergence criteria that is a scalar (Fig.~\ref{fig:grammar},~l.24).
 For example, 500 iterations are specified in Fig.~\ref{fig:exmsl},~l.16,
-\item the \texttt{computations} keyword (Fig.~\ref{fig:grammar},~l.23) introduces identifiers for each computation $k\in\Gamma=(S,R,(w,D),comp)$.
+\item the \texttt{computations} keyword (Fig.~\ref{fig:grammar},~l.23) introduces identifiers for each computation $k=(S,R,(w,D),comp) \in \Gamma$.
 Each computation (Fig.~\ref{fig:grammar},~l.26) specifies:
 \begin{itemize}
- \item the quantity $w$ written and its domain $D$, for example in Fig.~\ref{fig:exmsl},~l.22, kernel \texttt{k4} computes the quantity \texttt{F} on domain \texttt{d1},
- \item the read scalars $S$ and mesh quantities with their associated stencil shape $R$, for example in Fig.~\ref{fig:exmsl},~l.16, \texttt{k4} reads \texttt{C} with the shape \texttt{nce} and \texttt{D} with the default \textit{identity} shape; it does not read scalars.
+ \item the quantity $w$ written and its domain $D$, for example in Fig.~\ref{fig:exmsl},~l.22, kernel \texttt{k4} computes the variable \texttt{F} on domain \texttt{d1},
+ \item the read scalars $S$ and mesh variables with their associated stencil shape ($R$).
+For example in Fig.~\ref{fig:exmsl},~l.16, \texttt{k4} reads \texttt{C} with the shape \texttt{nce} and \texttt{D} with the default \textit{identity} shape; it does not read scalars.
  \end{itemize}
@@
 
-One can notice that in the example of Figure~\ref{fig:exmsl}, there are no (reduction) kernel associated to the scalars \texttt{mu} and \texttt{tau}.
+One can notice that in the example of Figure~\ref{fig:exmsl}, there are no kernel associated to the scalars \texttt{mu} and \texttt{tau} (reduction).
 In this case, those scalars are in fact constants.
@@
 Only an identifier is given to each kernel, for example $k_4$.
-The numerical code is indeed not handled by MSL, which is mesh-agnostic, and which is used to generate a parallel orchestration of computations only.
+The numerical code is indeed not handled by MSL that generates a parallel orchestration of computations only.
 The numerical computation is specified after MSC compilation by the developer (Fig.~\ref{fig:msf}).

### 2017-03-06 ARTICLES/IJPP/articleR1/evaluation.tex
relecture eval

@@
-This section first presents the implementation details chosen to evaluate MSF in this paper.
-Then, the studied use case is presented.
-Once this context is clear the MSC performance is evaluated as well as both available parallelization techniques, data and task.
-Finally, the performance model proposed in Section~\ref{sect:perfs}.
+This section first presents the implementation details chosen to evaluate MSF in this paper, and the studied use case.
+Then, the compilation time of MSC is evaluated before analyzing both available parallelization techniques, data and hybrid (data and task).
+Finally, the impact of fusioning kernel is studied.
+\CP{J'ai enlev\'e la ref au mod\`ele de perf car pas trouv\'e  dans les titres de sous-section}
+%performance model proposed in Section~\ref{sect:perfs}.
 
@@
 
-The two main implementation choices to take in MSC and its specialized assembly generation concern the technologies used for data and task parallelizations.
+The main choices to take when implementing MSC and its specialized assembly generation concern the technologies used for data and task parallelizations.
 
@@
 Although the version 4 of OpenMP has introduced explicit support for dynamic task scheduling, our implementation only requires version 3 whose fork-join model is well suited for the static scheduling introduced in Section~\ref{sect:parallelism}.
-The use of dynamic schedulers, such as provided by libgomp\footnote{\url{https://gcc.gnu.org/projects/gomp/}}, StarPU~\cite{Augonnet2011}, or XKaapi~\cite{Gautier:2013:XRS:2510661.2511383},  to directly execute the DAG $\Gamma_{dep}$ will be the subject of future works.
+The use of dynamic schedulers, such as provided by libgomp\footnote{\url{https://gcc.gnu.org/projects/gomp/}}, StarPU~\cite{Augonnet2011}, or XKaapi~\cite{Gautier:2013:XRS:2510661.2511383},  to directly execute the DAG $\Gamma_{dep}$ is left for future work.
 
 As a result, the MSL compiler generates a hybrid code which uses both SkelGIS and OpenMP.
-It also generates empty components where the developer must provide local sequential implementations of the kernels. In our case these kernels are written using SkelGIS API.
+It also generates components where the developer must provide local sequential implementations of the kernels. In our case these kernels are written using SkelGIS API.
 
@@
 %-------------------------------------
-\subsection{Compiler evaluation}
+\subsection{Multi-Stencil Compiler evaluation}
 
@@
 This has been computed on a laptop with a dual-core Intel Core i5 1.4~GHz, and 8~GB of DDR3.
+MSC has been implemented in Python.
+\CP{J'ai ajout\'e la phrase précédente. Est ce correct ? Quelle version de Python ? 2 ou 3 ? Combien de ligne de code ?}
+%
 While the overall time of 4.6 seconds remains reasonable, one can notice that the computation of the $TSP$ tree is by far the longest step.
 As a matter of fact, the complexity of the algorithm for N-shapes removal is $O(n^3)$.
-If this complexity is not a problem at the moment and onto this use case it could become one for just-in-time compilation or bigger simulations.
+If this complexity is not a problem at the moment and onto this use case it could become one for just-in-time compilation or more complex simulations.
 The replacement of the static scheduling by a dynamic scheduling using dedicated tools (such as OpenMP 4, StarPU etc.) should solve this in the future.
@@
 
-In this evaluation, we disable task-parallelism to focus on data-parallelism.
-Two versions of code are compared in this section: first a plain SkelGIS implementation of FullSWOF2D, where synchronizations and fusions are handled manually; second, the automatic synchronizations and fusions handled by MSF over SkelGIS.
+In this part, we disable task-parallelism to focus on data-parallelism.
+Two versions of the code are compared in this section: first a plain SkelGIS implementation of FullSWOF2D, where synchronizations and fusions are handled manually; second, the synchronizations and fusions are automatically handled by MSF over SkelGIS.
 SkelGIS has already been evaluated in comparison with a native MPI implementation for the FullSWOF2D example~\cite{CPE:CPE3494}.
@@
     RAM/core & 4GB\\
+    \#Nodes & 5040 \\
     Compiler [-O3] & gcc 4.9.1\\
@@
 
-Figures~\ref{fig:weak1}, \ref{fig:weak2} and~\ref{fig:weak3} respectively show weak scalings for a computation domain of $400 \times 400$ cells by core, $600 \times 600$ cells by core and $800 \times 800$ cells by core from 16 cores to 16.384 cores, as resumed in Table~\ref{tab:weak}.
+Figures~\ref{fig:weak1}, \ref{fig:weak2} and~\ref{fig:weak3} respectively show weak scalings for a computation domain of $400 \times 400$ cells by core, $600 \times 600$ cells by core and $800 \times 800$ cells by core from 16 cores to 16,384 cores, as summarized in Table~\ref{tab:weak}.
 
@@
 
-From these results, one can notice first that performances of MSF are very close to the reference version using plain SKelGIS.
-This is a very good result which shows first that MSC performs good synchronizations and fusions, and second that overheads introduced by \llc are limited thanks to a good granularity in GA.
+From these results, one can notice first that performances of MSF are very close to the reference version using plain SkelGIS.
+This is a very good result which shows first that MSC performs good synchronizations and fusions, and second that overheads introduced by \llc are limited thanks to a good component granularity in the Generic Assembly.
+
 However, it seems that a slightly drop of performance happens when the domain size per core increases.
 This performance decrease is really small though, with a maximum difference between the two versions of 2.83\%.
+\CP{Quelle figure? est ce \ref{fig:weak3} pour de petit nombre de core? Si oui, \`a mod\'erer car MSF devient meilleur quand le nombre de core augmente}
 
 The only noticeable difference between the two versions are due to \llc which load dynamic libraries at runtime.
-Because of this particularity, components of \llc are compiled with the \emph{-fpic} compilation flag while the SkelGIS version does not.
+Because of this particularity, components of \llc are compiled with the \emph{-fpic} compilation flag\footnote{\llc has been recently extended with the possibility of static linking.} while the SkelGIS version does not.
 This flag can have slight positive or negative effect on code performance depending on the situation and might be responsible for the observed difference.
@@
 
-Figure~\ref{fig:strong} shows the number of iteration per second for a 10k$\times$10k global domain size from 16 to 16.384 cores.
+Figure~\ref{fig:strong} shows the number of iteration per second for a 10k$\times$10k global domain size from 16 to 16,384 cores.
 The total number of time iterations for this benchmark is $1000$.
-In addition to the reference SkelGIS version, the ideal strong scaling is also illustrated in the figure.
+In addition to the reference SkelGIS version, the ideal strong scaling is also plotted in the figure.
 
-First, one can notice that the strong scaling evaluated for the MSF version is close to the ideal speedup up to 16.384 cores, which is a very good result.
-Moreover, no overheads are introduced by MSF which shows that automatic synchronizations and automatic fusions are the same one that the one written manually into the plain SkelGIS version.
+First, one can notice that the strong scaling evaluated for the MSF version is close to the ideal speedup up to 16,384 cores, which is a very good result.
+Moreover, no overheads are introduced by MSF which shows that automatic synchronizations and automatic fusions enable the same level of performance than the one manually written into the plain SkelGIS version.
 Finally, no overheads are introduced by components of \llc.
 A small behavior difference can be noticed with $2^9=512$ cores, however this variation is no longer observed with 1024 cores.
+\CP{Un commentaire sur le d\'ecrochage par rapport au perfect speedup pour $2^9--2^{12}$??}
 
@@
 
-In this section, we introduce task parallelism to evaluate the hybrid parallelization offered by MSF. The MSF implementation evaluated in this paper relies on SkelGIS and OpenMP.
+In this section, we add task parallelism to evaluate the hybrid parallelization offered by MSF. The MSF implementation evaluated in this paper relies on SkelGIS and OpenMP.
 
@@
 
-\begin{figure}[bh]\begin{center}
+\begin{figure}[th]\begin{center}
   \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/analytic/times.pdf}
@@
   \label{fig:limit}
+  \CP{es tu sûr que le temps soit en puissance de 2 et pas en puissance de 10? C'est strange.}
 \end{center}\end{figure}
 
-On the other hand, Fig.~\ref{fig:limit} illustrates limitations of data parallelization technique alone.
-In this figure is represented the execution time (with a logarithmic scale) of FullSWOF2D while increasing the number of cores for a fix domain size of $500 \times 500$ with a total of $200$ time iterations (\ie this is a strong scaling).
-The speedup of this same benchmark is represented in blue in Fig.~\ref{fig:close}.
-One can note that the scaling is not as good is the one presented in Fig.~\ref{fig:strong}.
+On the other hand, Figure~\ref{fig:limit} illustrates limitations of data parallelization technique alone.
+This figure displays the execution time (with a logarithmic scale) of FullSWOF2D while increasing the number of cores for a fix domain size of $500 \times 500$ with a total of $200$ time iterations (\ie this is a strong scaling).
+The speedup of this same benchmark is represented in blue in Figure~\ref{fig:close}.
+One can note that the scaling is not as good as the one presented in Figure~\ref{fig:strong}.
 The main difference between these two benchmarks is the domain size.
-In the benchmark of Fig.~\ref{fig:strong} the domain size is $10k \times 10k$ which means that using $2^8$ cores, for example, each core has to compute only a $31 \times 31$ sub-domain.
-On the other hand using $2^8$ cores in Fig.~\ref{fig:close} each core has to compute a $625 \times 625$ sub-domain.
-Fig.~\ref{fig:limit} shows why the speedup is not as good as the one with a bigger domain size.
+In the benchmark of Figure~\ref{fig:strong} the domain size is $10k \times 10k$ which means that using $2^8 = 256$ cores, for example, each core has to compute only a $31 \times 31$ sub-domain.
+On the other hand using $2^8$ cores in Figure~\ref{fig:close} each core has to compute a $625 \times 625$ sub-domain.
+Figure~\ref{fig:limit} shows why the speedup is not as good as the one with a bigger domain size.
 
@@
 
-All in all, when the number of core used increases, the computation/communication ratio becomes poorer and poorer.
+All in all, when the number of used core increases, the computation/communication ratio becomes poorer and poorer.
 As a result, the data parallelism alone fails to provide enough parallelism to leverage the whole machine and other sources of parallelism have to be found.
 As expected, in  Figure~\ref{fig:close} the speedup bends down from 256 to 2048 cores.
-The same problem would happened in previous experiment of Fig.~\ref{fig:strong}, however as the domain size is larger, the phenomena appears with much more cores.
+The same problem would happened in previous experiment of Figure~\ref{fig:strong}, however as the domain size is larger, the phenomena appears with more cores.
 
@@
 
-In addition to the blue curve, Fig.~\ref{fig:close} shows speedups for the same example ($500 \times 500$ domain with $200$ iterations) but using an hybrid parallelization.
-Fig.~\ref{fig:close} shows a comparison with 2, 4, 8 and 16 cores per MPI process for task parallelization.
+In addition to the blue curve, Figure~\ref{fig:close} shows speedups for the same example ($500 \times 500$ domain with $200$ iterations) but using an hybrid parallelization.
+Figure~\ref{fig:close} shows a comparison with 2, 4, 8 and 16 cores per MPI process for task parallelization.
 
@@
 As a result, for example, when using 2 machines of the TGCC cluster, with a total of 32 cores, 4 cores are used for SkelGIS MPI processes, for data parallelization, and for each one 8 cores are used for task parallelization ($4 \times 8 = 32$).
-This respect $P = P_{data} \times P_{task}$ presented in Section~\ref{sect:perfs}.
-As a result, and as explained in Section~\ref{sect:perfs}, quantities are less divided into sub-domains, that are responsible for communications, and the effect which is observe onto the blue curve is delayed.
+This respects $P = P_{data} \times P_{task}$ as presented in Section~\ref{sect:perfs}.
+As a result, and as explained in Section~\ref{sect:perfs}, quantities that are responsible for communications are less divided into sub-domains.
+Therefore, the effect observed with the blue curve is delayed to a higher number of cores.
 
@@
 
-First, thin nodes of the TGCC Curie are built with two NUMA nodes each of 8 cores.
-As a result, when increasing the number of OpenMP cores for task parallelization from 8 to 16 cores, an overhead is introduced by exchanges of data between memories of the two NUMA nodes.
-This phenomena is illutrated in Fig.~\ref{fig:spread}.
+First, thin nodes of the TGCC Curie are built with two NUMA socket each of 8 cores.
+As a result, when increasing the number of OpenMP cores for task parallelization from 8 to 16 cores, an overhead is introduced by exchanges of data between memories of the two NUMA sockets.
+This phenomena is illustrated in Figure~\ref{fig:spread}.
 In this figure, a different binding strategy is used.
 A binding strategy is the way the scheduler binds threads onto available cores.
-The strategy used in Fig.~\ref{fig:spread} is called \emph{spread} (instead of \emph{close} in Figure~\ref{fig:close}).
-This strategy binds threads on cores in order to spread as much as possible onto resources, which means that the two NUMA nodes are directly used whatever the number of cores used for task parallelism.
-As a result, and as shown in the figure, using 2, 4 and 8 cores an initial overhead is introduced as the one observed in Fig.~\ref{fig:close}.
+The strategy used in Figure~\ref{fig:spread} is called \emph{spread} (instead of \emph{close} in Figure~\ref{fig:close}).
+This strategy binds threads on cores in order to spread as much as possible onto resources, which means that the two NUMA sockets are used whatever the number of cores used for task parallelism.
+As a result, and as shown in the figure, using 2, 4 and 8 cores an initial overhead is introduced as the one observed in Figure~\ref{fig:close}.
 This shows that the initial overhead with 16 cores is due to NUMA effects.
@@
 This phenomena has been mentioned in Section~\ref{sect:perfs} by the variable $F_{task}$ and the fact that it is not always true that $F_{task}=P_{task}$.
-This explains why using 16 cores is less efficient than using 8 cores, even when the two NUMA nodes are always used as in Fig.~\ref{fig:spread}.
+This explains why using 16 cores is less efficient than using 8 cores, even when the two NUMA sockets are always used as in Figure~\ref{fig:spread}.
 
-Finally, to validate the performance model introduced in Section~\ref{sect:perfs}, and to understand when the hybrid parallelization becomes more interesting than the data parallelization, Figure~\ref{fig:tth2} represents $T_{COM1}$ and $T_{COM2}+T_{task}$ of Equation~(\ref{eq:hyb}), for the best case, \ie when 8 cores are used in Fig.~\ref{fig:close}.
+Finally, to validate the performance model introduced in Section~\ref{sect:perfs}, and to understand when the hybrid parallelization becomes more interesting than the data parallelization, Figure~\ref{fig:tth2} represents $T_{COM1}$ and $T_{COM2}+T_{task}$ of Equation~(\ref{eq:hyb}), for the best case, \ie when 8 cores are used in Figure~\ref{fig:close}.
 Figure~\ref{fig:tth2} and Table~\ref{fig:tth} presents results of these measurements.
-Results perfectly matches Fig.~\ref{fig:close} for 8 cores per MPI process.
-As a result, the hybrid parallelization becomes better with a total of 512 cores in this case.
+Results perfectly matches Figure~\ref{fig:close} for 8 cores per MPI process.
+As a result, the hybrid parallelization is better for 512 cores or more in this case.
 
@@
   \label{fig:tth2}
+\CP{Time en puissance de 2 ou de 10 ?}
 \end{center}\end{figure}
@@
 
-In this section we propose an evaluation of the fusion optimization.
-Three fusion optimizations have been formally detailed in Section~\ref{sect:fusion}.
+In this section we propose an evaluation of the fusion optimization. Three fusion optimizations
+\CP{les rappeler !}
+%
+have been formally detailed in Section~\ref{sect:fusion}.
 From the $TSP$ tree computed by MSC it is possible according to specific conditions to merge domain loops of kernels, thus optimizing the use of cache memories.
@@
 In FullSWOF2D a total of  62 fusions are proposed by MSF over a total of 98 computation kernels.
-Fig.~\ref{fig:fusion} shows that the performance is clearly improved (around 40\%) by these fusions.
+Figure~\ref{fig:fusion} shows that the performance is clearly improved (around 40\%) by these fusions.
 
@@
 
-Blue curves of Fig.~\ref{fig:fusion1} represent results for data parallelism only with and without fusion.
+Blue curves of Figure~\ref{fig:fusion1} represent results for data parallelism only with and without fusion.
 One can note that the best performance, as expected, is reached by the version using fusions.
@@
 One can note that the best performance is also reached by the version using fusion.
+\CP{PARAGRAPHE A ADAPTER SUIVANT LES MODIFS DE LA FIGURE: 1 ou 2 exp dans la figure?}
 
@@
   \label{fig:fusion1}
+  \CP{A FINIR !!! ERREUR DANS LEGENDES AUSSI. COULEUR PROCHES}
 \end{center}\end{figure}
 
-However, to deeper analyze this results, we propose a second evaluation presented in Fig.~\ref{fig:fusion2}.
-The blue curves are exactly the same one than in Fig.~\ref{fusion1}.
-The purple curves, on the other hand, represent results by using 8 cores per MPI process dedicated to tasks, with and without fusion.
+However, to deeper analyze this results, we propose a second evaluation presented in Figure~\ref{fig:fusion2}.
+The blue curves are exactly the same one than in Figure~\ref{fig:fusion1}.
+The purple \CP{Couleur trop subtil: peux tu prendre une autre couleur, par ex rouge} curves, on the other hand, represent results by using 8 cores per MPI process dedicated to tasks, with and without fusion.
 Interesting results appears in this figure as the hybrid version using fusions is less efficient than the one without fusions.
 As already explained, this result is due to the fact that fusions reduce the number of tasks from 98 to 36 resulting in a non optimized use of eight cores for task parallelism.
-By using only 2 cores per MPI process (in Fig.~\ref{fig:fusion1}) the 36 computation kernels were enough to feed the two cores, while it is not for eight.
+By using only 2 cores per MPI process (in Figure~\ref{fig:fusion1}) the 36 computation kernels were enough to feed the two cores, while it is not for eight.
 
@@
   \label{fig:fusion2}
+    \CP{A FINIR !!! ERREUR DANS LEGENDES AUSSI.}
 \end{center}\end{figure}
 
-As a result, if fusion optimization incurs a too large reduction of the number of tasks to feed dedicated cores, the problem observed for 16 cores in Fig.~\ref{fig:close} happens earlier which reduces obtained performances.
+As a result, if fusion optimization incurs a too large reduction of the number of tasks to feed dedicated cores, the problem observed for 16 cores in Figure~\ref{fig:close} happens earlier which reduces obtained performances.

### 2017-03-06 ARTICLES/IJPP/articleR1/related.tex
relecture RW

@@
 On the other hand, PATUS~\cite{citeulike12258902} proposes to add a parallelization strategy grammar to its stencil language to perform an auto-tuning parallelization strategies.
-Finally, ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} is specific to multigrid numerical methods etc.
+Finally, ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} is specific to multigrid numerical methods.
 Thus, these stencil compilers target a different scope than the Multi-Stencil Framework presented in this paper, which actually orchestrates a parallel execution of multiple stencil codes together.
-Furthermore, future works could lead to a combination of these stencil compilers with MSF to build very optimized stencil kernels $K$.
+Hence, an interesting future work would be to combine these stencil compilers with MSF to build very optimized stencil kernels $K$.
 
@@
 Some solutions, closer to MSF, have also been proposed to automatically orchestrate multiple stencils and computations in a parallel manner.
-Amongs them is Halide~\cite{Ragan-Kelley:2013:HLC:2491956.2462176} that proposes an optimization and parallelization of a pipeline of stencils.
+Among them is Halide~\cite{Ragan-Kelley:2013:HLC:2491956.2462176} that proposes an optimization and parallelization of a pipeline of stencils.
 However, Halide is limited to Cartesian meshes and is specific to images.
@@
 Second, MSL is very light and only descriptive.
-Numerical codes are left to another language and another user (the developer in Figure~\ref{fig:msf}).
-Furthermore, such dataflow runtimes could actually be used by MSF as backends, instead of using SkelGIS or OpenMP.
+Numerical codes are left to another language and another user (the developer in Figure~\ref{fig:msf}~\vpageref{fig:msf}).
+Furthermore, such dataflow runtimes could actually be used by MSF as back-ends, instead of using SkelGIS or OpenMP.
 \JB{Manque de refs bis :)}
@@
 % vision of the framework brings separation of concerns, code-reuse and flexibility
-This flexibility of MSF is due to software engineering capacities introduced by proposing a meta-model and by using component models.
+This flexibility of MSF is due to software engineering capacities introduced by proposing a meta-model and by using a component model.
 Actually, MSF is designed to increase separation of concerns and code-reuse compared to existing solutions.
@@
 A HPC specialist can propose new (or updated) components for handling the distributed data structure and quantities of the simulation.
-MSF generates from these informations the parallel orchestration of computations.
+MSF generates from these pieces of information the parallel orchestration of computations.
 Finally, after MSF compilation (MSC), the developer of numerical codes fills computation kernels by using the chosen distributed data structure.

### 2017-03-06 ARTICLES/IJPP/articleR1/evaluation.tex
corrections parallelism.tex + evaluation.tex

@@
 Then, the compilation time of MSC is evaluated before analyzing both available parallelization techniques, data and hybrid (data and task).
-Finally, the impact of fusioning kernel is studied.
-\CP{J'ai enlev\'e la ref au mod\`ele de perf car pas trouv\'e  dans les titres de sous-section}
+Finally, the impact of kernels fusions is studied.
+%\CP{J'ai enlev\'e la ref au mod\`ele de perf car pas trouv\'e  dans les titres de sous-section}
 %performance model proposed in Section~\ref{sect:perfs}.
@@
 
-The main choices to take when implementing MSC and its specialized assembly generation concern the technologies used for data and task parallelizations.
+The main choices to take when implementing a specialized assembly of GA concern the technologies used for data and task parallelizations, \ie implementation choices of $DDS$ and $Scheduler$ components.
 
-For the data-parallelization, as already detailed many times throughout the paper, a third party HPC specialist is responsible for implementing $DDS$ and $Data$ using a chosen library or external language and by following the interfaces specified by these two components.
+For the data-parallelization, as already detailed many times throughout the paper, a third party HPC specialist is responsible for implementing $DDS$ and $Data$ using a chosen library or external language and by following the specified interfaces of these two components.
 To evaluate MSF, we have played the role of HPC specialists and have implemented these components using SkelGIS, a C++ embedded DSL~\cite{CPE:CPE3494} that proposes a distributed Cartesian mesh as well as user API to manipulate structures while hiding their underlying distribution.
 
-For the task parallelism, we have chosen to use OpenMP~\cite{660313} to generate the code of the $Scheduler$ component.
+For task parallelism, we have chosen to use OpenMP~\cite{660313} to generate the code of the $Scheduler$ component.
 OpenMP targets shared-memory platforms only.
 Although the version 4 of OpenMP has introduced explicit support for dynamic task scheduling, our implementation only requires version 3 whose fork-join model is well suited for the static scheduling introduced in Section~\ref{sect:parallelism}.
-The use of dynamic schedulers, such as provided by libgomp\footnote{\url{https://gcc.gnu.org/projects/gomp/}}, StarPU~\cite{Augonnet2011}, or XKaapi~\cite{Gautier:2013:XRS:2510661.2511383},  to directly execute the DAG $\Gamma_{dep}$ is left for future work.
+The use of dynamic schedulers, such as provided by libgomp\footnote{\url{https://gcc.gnu.org/projects/gomp/}}, StarPU~\cite{Augonnet2011}, or XKaapi~\cite{Gautier:2013:XRS:2510661.2511383}, to directly execute the DAG $\Gamma_{dep}$ is left to future work.
 
-As a result, the MSL compiler generates a hybrid code which uses both SkelGIS and OpenMP.
-It also generates components where the developer must provide local sequential implementations of the kernels. In our case these kernels are written using SkelGIS API.
+As a result, MSC generates a hybrid code which uses both SkelGIS and OpenMP.
+It also generates the structure of $K$ components where the developer must provide local sequential implementations of the kernels using SkelGIS API.
 
@@
 FullSWOF2D contains 3~mesh entities, 7~computation domains, 48~data and 98~computations (32~stencil kernels and 66~local kernels).
-Performances of the obtained implementation are compared to the full SkelGIS implementation to show that no overheads are introduced by MSF.
+Performances of the obtained implementation are compared to the plain SkelGIS implementation to show that no overheads are introduced by MSF by using \llc.
 
@@
 This has been computed on a laptop with a dual-core Intel Core i5 1.4~GHz, and 8~GB of DDR3.
-MSC has been implemented in Python.
-\CP{J'ai ajout\'e la phrase précédente. Est ce correct ? Quelle version de Python ? 2 ou 3 ? Combien de ligne de code ?}
+MSC has been implemented in Python 2.
 %
-While the overall time of 4.6 seconds remains reasonable, one can notice that the computation of the $TSP$ tree is by far the longest step.
+While the overall time of 4.6 seconds remains reasonable for a real case study, one can notice that the computation of the $TSP$ tree is by far the longest step.
 As a matter of fact, the complexity of the algorithm for N-shapes removal is $O(n^3)$.
@@
 In this part, we disable task-parallelism to focus on data-parallelism.
-Two versions of the code are compared in this section: first a plain SkelGIS implementation of FullSWOF2D, where synchronizations and fusions are handled manually; second, the synchronizations and fusions are automatically handled by MSF over SkelGIS.
+Two versions of the code are compared in this section: first a plain SkelGIS implementation of FullSWOF2D, where synchronizations and fusions are handled manually; second, a MSF over SkelGIS version where  synchronizations and fusions are automatically handled.
 SkelGIS has already been evaluated in comparison with a native MPI implementation for the FullSWOF2D example~\cite{CPE:CPE3494}.
@@
 
-Figures~\ref{fig:weak1}, \ref{fig:weak2} and~\ref{fig:weak3} respectively show weak scalings for a computation domain of $400 \times 400$ cells by core, $600 \times 600$ cells by core and $800 \times 800$ cells by core from 16 cores to 16,384 cores, as summarized in Table~\ref{tab:weak}.
+\paragraph{\textbf{Weak scaling}}
+Figures~\ref{fig:weak1}, \ref{fig:weak2} and~\ref{fig:weak3} respectively show weak scaling experiments tha twe have conducted.
+Four computation domains are evaluated: $400 \times 400$ cells by core, $600 \times 600$ cells by core and $800 \times 800$ cells by core, from 16 to 16,384 cores, as summarized in Table~\ref{tab:weak}.
 
@@
 
-\paragraph{\textbf{Weak scaling}}
 \begin{figure}[t]
@@
 
-From these results, one can notice first that performances of MSF are very close to the reference version using plain SkelGIS.
+From these results, one can notice, first, that performances of MSF are very close to the reference version using plain SkelGIS.
 This is a very good result which shows first that MSC performs good synchronizations and fusions, and second that overheads introduced by \llc are limited thanks to a good component granularity in the Generic Assembly.
@@
 However, it seems that a slightly drop of performance happens when the domain size per core increases.
-This performance decrease is really small though, with a maximum difference between the two versions of 2.83\%.
-\CP{Quelle figure? est ce \ref{fig:weak3} pour de petit nombre de core? Si oui, \`a mod\'erer car MSF devient meilleur quand le nombre de core augmente}
+This performance decrease is really small though, with a maximum difference between the two versions of 2.83\% in Fig.~\ref{fig:weak3}.
+%\CP{Quelle figure? est ce \ref{fig:weak3} pour de petit nombre de core? Si oui, \`a mod\'erer car MSF devient meilleur quand le nombre de core augmente}
 
@@
 Because of this particularity, components of \llc are compiled with the \emph{-fpic} compilation flag\footnote{\llc has been recently extended with the possibility of static linking.} while the SkelGIS version does not.
-This flag can have slight positive or negative effect on code performance depending on the situation and might be responsible for the observed difference.
+This flag can have slight positive or negative effects on code performance depending on the situation and might be responsible for the observed difference.
 
@@
 A small behavior difference can be noticed with $2^9=512$ cores, however this variation is no longer observed with 1024 cores.
-\CP{Un commentaire sur le d\'ecrochage par rapport au perfect speedup pour $2^9--2^{12}$??}
+%We have not analyzed this behavior that we consider not fundamental.
+%\CP{Un commentaire sur le decrochage par rapport au perfect speedup pour $2^9--2^{12}$??}
 
@@
 The fork/join model obtained for FullSWOF2D is composed of 18 fork phases (corresponding to $\mathcal{P}$ nodes of $TSP$).
-Table~\ref{fig:freq} represents the number of time a given level of parallelism is obtained inside fork phases.
+Table~\ref{fig:freq} represents the number of time (denoted frequency) a given level of parallelism is obtained inside fork phases.
 
@@
  \end{tabular}
-\caption{Parallelism level (number of parallel tasks) and the number of times this level appears.}
+\caption{Parallelism level and the number of times this parallelism level appears into fork phases.}
 \label{fig:freq}
@@
 This means that if two cores are dedicated to task parallelism, the two sequential parts of the code will not take advantage of these two cores, and that no part of the code would benefit from more than 16 cores.
-The task parallelism, as proposed in this paper (\ie where each kernel is a task) is therefore not insufficient to take advantage of a single node of modern clusters that typically supports more than 16 cores.
+The task parallelism, as proposed in this paper (\ie where each kernel is a task) is therefore insufficient to take advantage of a single node of modern clusters that typically supports more than 16 cores.
 
@@
   \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/analytic/times.pdf}
-  \caption{Computation vs communication times in the data parallelization technique.}
+  \caption{Computation vs communication times for a single time iteration using the data parallelization technique.}
   \label{fig:limit}
-  \CP{es tu sûr que le temps soit en puissance de 2 et pas en puissance de 10? C'est strange.}
+  %\CP{es tu sûr que le temps soit en puissance de 2 et pas en puissance de 10? C'est strange.}
 \end{center}\end{figure}
@@
 This figure displays the execution time (with a logarithmic scale) of FullSWOF2D while increasing the number of cores for a fix domain size of $500 \times 500$ with a total of $200$ time iterations (\ie this is a strong scaling).
+One can note that times are really small.
+Actually the time represented in Fig.~\ref{fig:limit} is the time spent into a single time iteration.
 The speedup of this same benchmark is represented in blue in Figure~\ref{fig:close}.
@@
 The main difference between these two benchmarks is the domain size.
-In the benchmark of Figure~\ref{fig:strong} the domain size is $10k \times 10k$ which means that using $2^8 = 256$ cores, for example, each core has to compute only a $31 \times 31$ sub-domain.
-On the other hand using $2^8$ cores in Figure~\ref{fig:close} each core has to compute a $625 \times 625$ sub-domain.
+In the benchmark of Figure~\ref{fig:strong} the domain size is $10k \times 10k$ which means that using $2^8 = 256$ cores, for example, each core has to compute only a $625 \times 625$ sub-domain.
+On the other hand, using $2^8$ cores in Figure~\ref{fig:close} each core has to compute a $31 \times 31$ sub-domain.
 Figure~\ref{fig:limit} shows why the speedup is not as good as the one with a bigger domain size.
@@
 From this point thus the communication time is typically modeled as $L+S/B$ where $L$ is the latency, $S$ the data size and $B$ the bandwidth.
-This explains the decrease of time from 32 to 128 cores where the data sizes communicated by each process decreases.
-The increased observed after$128$ cores that might be due to the fact that with the increased number of processes the fat-tree becomes deeper and the latencies increase.
+This explains the decrease of communication time from 32 to 128 cores where the data sizes communicated by each process decreases.
+The increases observed after $128$ cores might be due to the fact that with the increased number of processes the fat-tree becomes deeper and the latencies increase.
 
@@
 
-All in all, when the number of used core increases, the computation/communication ratio becomes poorer and poorer.
+All in all, when the number of core increases, the computation/communication ratio becomes poorer and poorer.
 As a result, the data parallelism alone fails to provide enough parallelism to leverage the whole machine and other sources of parallelism have to be found.
@@
 
-As task parallelism fails to scale from 16 cores, and as data parallelism also fails to scale when the communication cost overpass the execution time, an hybrid parallelization strategy is proposed by MSF and evaluated below.
+As task parallelism fails to scale from 16 cores, and as data parallelism also fails to scale when the communication cost overpass the execution time, an hybrid parallelization strategy is proposed by MSF and is evaluated below.
 
@@
 The strategy used in Figure~\ref{fig:spread} is called \emph{spread} (instead of \emph{close} in Figure~\ref{fig:close}).
-This strategy binds threads on cores in order to spread as much as possible onto resources, which means that the two NUMA sockets are used whatever the number of cores used for task parallelism.
+This strategy binds threads on cores in order to spread as much as possible onto resources, which means that the two NUMA sockets are used whatever the number of cores used for tasks is.
 As a result, and as shown in the figure, using 2, 4 and 8 cores an initial overhead is introduced as the one observed in Figure~\ref{fig:close}.
@@
   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/tth.pdf}}
-  \caption{Execution times (seconds) of $T_{COM1}$ and $T_{COM2} + T_{task}$ for 8 cores for task parallelization. Verification of the Equation~(\ref{eq:hyb}).}
+  \caption{Execution times (seconds) for a single time iteration of $T_{COM1}$ and $T_{COM2} + T_{task}$ for 8 cores for task parallelization. Verification of the Equation~(\ref{eq:hyb}).}
   \label{fig:tth2}
-\CP{Time en puissance de 2 ou de 10 ?}
+%\CP{Time en puissance de 2 ou de 10 ?}
 \end{center}\end{figure}
@@
 
-In this section we propose an evaluation of the fusion optimization. Three fusion optimizations
-\CP{les rappeler !}
+In this section we propose an evaluation of the fusion optimization.
+From the $TSP$ tree computed by MSC it is possible, according to specific conditions, to merge domain loops of kernels, thus optimizing the use of cache memories.
+This optimization is called a fusion and has been introduced in Section\ref{sect:fusion}.
+Three fusion optimizations
+%\CP{les rappeler !}
 %
 have been formally detailed in Section~\ref{sect:fusion}.
-From the $TSP$ tree computed by MSC it is possible according to specific conditions to merge domain loops of kernels, thus optimizing the use of cache memories.
+The two first one of Figures~\ref{fig:fus1} and~\ref{fig:fus2} have been automatically detected by MSF in our case study (\ie scatter optimization has not been detected).
 
@@
 Most of the time such fusions are done naturally by a computer scientist.
-However, for a numerician who usually writes equations more than numerical codes, an automatic detection of such fusions avoids errors, particularly for a parallel execution.
+However, an automatic detection of such fusions avoids errors, particularly for a parallel execution.
 In addition to this, more advanced fusion cases, such as a scatter are more difficult to deduce.
@@
 
-Blue curves of Figure~\ref{fig:fusion1} represent results for data parallelism only with and without fusion.
+Blue curves of Figure~\ref{fig:fusion1} represent results for data parallelism with and without fusion.
 One can note that the best performance, as expected, is reached by the version using fusions.
-Green curves represent results by using 2 cores per MPI process dedicated for tasks, with and without fusion again.
+Red curves represent results by using 2 cores per MPI process dedicated to tasks, with and without fusion again.
 One can note that the best performance is also reached by the version using fusion.
-\CP{PARAGRAPHE A ADAPTER SUIVANT LES MODIFS DE LA FIGURE: 1 ou 2 exp dans la figure?}
+%\CP{PARAGRAPHE A ADAPTER SUIVANT LES MODIFS DE LA FIGURE: 1 ou 2 exp dans la figure?}
 
@@
   \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout2_close_median.pdf}
-  \caption{blabla}
+  \caption{String scaling on a 500x500 domain size with $200$ time iterations.
+Blue curves represent strong scaling for data parallelism with and without fusion.
+Red curves represent strong scaling by using 2 cores per MPI process dedicated to tasks, with and without fusion.}
   \label{fig:fusion1}
-  \CP{A FINIR !!! ERREUR DANS LEGENDES AUSSI. COULEUR PROCHES}
+  %\CP{A FINIR !!! ERREUR DANS LEGENDES AUSSI. COULEUR PROCHES}
 \end{center}\end{figure}
@@
 The blue curves are exactly the same one than in Figure~\ref{fig:fusion1}.
-The purple \CP{Couleur trop subtil: peux tu prendre une autre couleur, par ex rouge} curves, on the other hand, represent results by using 8 cores per MPI process dedicated to tasks, with and without fusion.
+The red curves, on the other hand, represent results by using 8 cores per MPI process dedicated to tasks, with and without fusion.
 Interesting results appears in this figure as the hybrid version using fusions is less efficient than the one without fusions.
@@
   \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout8_close_median.pdf}
-  \caption{blabla}
+  \caption{String scaling on a 500x500 domain size with $200$ time iterations.
+Blue curves represent strong scaling for data parallelism with and without fusion, thus are exactly the same than blue curves of Fig.~\ref{fig:fusion1}.
+Red curves represent strong scaling by using 8 cores per MPI process dedicated to tasks, with and without fusion.}
   \label{fig:fusion2}
-    \CP{A FINIR !!! ERREUR DANS LEGENDES AUSSI.}
+    %\CP{A FINIR !!! ERREUR DANS LEGENDES AUSSI.}
 \end{center}\end{figure}
@@
 As a result, if fusion optimization incurs a too large reduction of the number of tasks to feed dedicated cores, the problem observed for 16 cores in Figure~\ref{fig:close} happens earlier which reduces obtained performances.
+For this reason, MSF performs fusions only when data parallelization is used alone.
+This choice could be more intelligent but this is the subject of future work.

### 2017-03-06 ARTICLES/IJPP/articleR1/parallelism.tex
corrections parallelism.tex + evaluation.tex

--- 2017-03-05 ARTICLES/IJPP/articleR1/parallelism.tex
@@
 \medskip
-More formally, the data parallelization of a multi-stencil program 
-\begin{equation*}
-\mathcal{MSP}(\mathcal{M},\Phi,\mathcal{D},\mathcal{N},\Delta, \mathcal{S},T,\Gamma)
-\end{equation*}
+More formally, the data parallelization of a multi-stencil program of equation (\ref{eq:msp})
 consists in a partitioning of the mesh $\mathcal{M}$ in $p$ sub-meshes $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
-This step can be performed by an external graph partitioner~\cite{Pellegrini:1996:SSP:645560.658570,DBLP:conf/ieeehpcs/HeleneS13,lachat:hal-00768916} and is addressed by the third party DDS component instantiated from GA.
+This step can be performed by an external graph partitioner~\cite{Pellegrini:1996:SSP:645560.658570,DBLP:conf/ieeehpcs/HeleneS13,lachat:hal-00768916} and is handled by the DDS implementation of the third party HPC specialist.
 
@@
 
-However, as a multi-stencil program is an iterative program, computations which happen after $k_j$ at the time iteration $t$ have also been computed before $k_j$ at the previous time iteration $t-1$.
+However, as a multi-stencil program is an iterative program, computations that happen after $k_j$ at the time iteration $t$ have also been computed before $k_j$ at the previous time iteration $t-1$.
 For this reason another case of synchronization has to be defined.
@@
 
-One can notice that both types of synchronizations are performed by all resources.
-
+%One can notice that both types of synchronizations are performed by all resources.
 \begin{mydef}
@@
 
-\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl} gives an example of a $\mathcal{MSP}$ program. From this example, the following ordered list of computation kernels can be extracted:
+\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl} gives an example of a $\mathcal{MSP}$ program. From this example, the following ordered list of computation kernels is extracted:
 \begin{equation*}
@@
 
-Those definitions are known as \emph{data hazards classification}.
-However, a specific condition on the computation domain, due to the specific domain of multi-stencils, is introduced for the write after write case.
-One can note that the \texttt{independent} keyword of Fig.~\ref{fig:grammar} is useful in this case.
+These definitions are known as \emph{data hazards classification}.
+However, a specific condition on the computation domain, due the multi-stencils specific case, is introduced for the write after write case.
+One can note that the \texttt{independent} keyword of Fig.~\ref{fig:grammar} is useful in this case as the user explicitly indicates that $D_i \cap D_j=\emptyset$.
 
@@
 
-One can notice that $\Gamma_{dep}$ is the dependency graph of the computations of a multi-stencil program, but it only takes into account a single time iteration.
-A complete dependency graph of the simulation could be built.
-This is a possible extension of this work.
+One can note that $\Gamma_{dep}$ only takes into account a single time iteration. A complete dependency graph of the simulation could be built. This is a possible extension of this work.
 
@@
 
-Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or sub-graph, called $N$, such that the following property is verified :
+Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or sub-graph, called $N$, such that a DAG without this shape is MSPD.
+%the following property is verified :
+%\begin{myth}
+%The transitive reduction of a DAG $G$ is MSPD if and only if it does not contain $N$ as a sub-graph.
+%\end{myth}
 
-\begin{myth}
-The transitive reduction of a DAG $G$ is MSPD if and only if it does not contain $N$ as an induced sub-graph.
-\end{myth}
-
-As $\Gamma_{dep}$ is a DAG, by removing N-Shapes it is transformed to a MSPD.
+Thus, as $\Gamma_{dep}$ is a DAG, by removing N-Shapes it is transformed to a MSPD.
 The intuition is illustrated in Fig.~\ref{fig:over}.
@@
 \begin{center}
+\subfloat[Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
   \end{tikzpicture}
-\caption{Over-constraint on the forbidden $N$ shape.}
-\label{fig:over}
-\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-\begin{center}
+}
+\hspace{30pt}
+\subfloat[TSP tree of Fig.~\ref{fig:over}\label{fig:solu}]{
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
   \end{tikzpicture}
-\caption{TSP tree of Fig.~\ref{fig:over}.}
-\label{fig:solu}
+}
+\label{fid:nshape}
+\caption{N-Shape problem and solution}
 \end{center}
@@
 This is called a fusion.
-As previously explained in Section~\ref{sect:formalism}, MSL restrict the definition of a numerical computation by writing a single quantity at a time which is more natural for numericians, and which avoid errors in manual fusion or counter-productive fusions for task parallelization.
-MSF guarantee that proposed fusions are correct and will not cause errors in the final results of the simulation.
+As previously explained in Section~\ref{sect:formalism}, MSL restrict the definition of a numerical computation by writing a single quantity at a time which avoids errors in manual fusion or counter-productive fusions for task parallelization.
+MSF guarantees that proposed fusions are correct and will not cause errors in the final results of the simulation.
 
@@
 The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where the $parent(k)$ function returns the parent vertex of $k$ in the tree, and where $k_{i;j}^{fus}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.
-Finally, $type(k)$ returns $comp$ if the kernel is a computation kernel, and $sync$ otherwise.
+Finally, $type(k)$ returns $comp$ if the kernel is a computation kernel, and $sync$  or $red$ otherwise.
 
@@
 We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs~\cite{spaaTangCKLL11} for example.
-However, this fusion takes place at a different level and can bring performance improvements as will be illustrated in Section~\ref{sect:eval}.
+However, this fusion takes place at a different level and can bring performance improvements as  illustrated in Section~\ref{sect:eval}.
 This fusion algorithm relies on the following observations.
@@
 
-Thirs, and similarly to the first observation, two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are data dependent and what is written by the first one is read by the second one.
+Third, an additional fusion case is possible and more tricky to find.
+Similarly to the first observation, two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are data dependent and what is written by the first one is read by the second one.
 The construction of the tree also guarantees that synchronizations are not needed between these computations, otherwise a $k^{sync}$ would have been inserted between them (inherited from $\Gamma_{sync}$).
@@
 \end{itemize}
-the fusion of $k_i$ and $k_j$ is possible if and only if $\exists n:D_i \rightarrow D_j \in \mathcal(N)$ such that
+the fusion of $k_i$ and $k_j$ is possible if and only if $\exists n:D_i \rightarrow D_j \in \mathcal{N}$ such that
 \begin{equation*}
@@
 One can also note that the computation $k_j$ will be written in a different manner if a scatter fusion is performed or not.
-This is not a problem by using MSF as the fusion is proposed before the developer actually write the numerical code of $k_j$.
-The developer will be notified of fusions in the output of MSF.
 This particular case is illustrated in Fig.~\ref{fig:fus3}.
@@
 
+The developer will be notified of fusions in the output of MSC. This is not a problem by using MSF as the fusion is proposed before the developer actually write the numerical code of $k_j$.
+
 %--------------------
@@
 
-MSC takes a MSL file written following the grammar described in Section~\ref{sect:msl}, as well as the Generic Assembly of components represented in Fig.~\ref{fig:ga} as inputs, and generates a specialized component assembly that manages the parallel orchestration of the computations of the simulation.
+MSC takes a MSL file written using the grammar described in Section~\ref{sect:msl}, as well as the Generic Assembly presented in Fig.~\ref{fig:ga} as inputs, and generates a specialized component assembly that manages the parallel orchestration of the computations of the simulation.
 In this final assembly, that could be compared to a pattern or a skeleton of the simulation, the developer still has to fill-in the functions corresponding to the various computation kernels by using the DDS instantiation chosen into the specialized assembly.
@@
  \item it instantiates $DDS$ and $Data$ components by using components implemented by a third party HPC specialist,
- \item it generates $K$ components responsible for each computation kernel of the simulation,
+ \item it generates the structure of $K$ components responsible for each computation kernel of the simulation,
  \item it generates a new $Scheduler$ component,
- \item it replaces $Computations$ component by a generated sub-assembly that matches $TSP$ by using $Scheduler$, $K$ and $Sync$ components.
+ \item it replaces the $Computations$ component of GA by a generated sub-assembly that matches $TSP$ by using $Scheduler$, $K$ and $Sync$ components.
  \end{enumerate}
@@
 The component also exposes a provide port to be connected to the $Scheduler$ component.
-A generic $K$ component is represented in Fig.~\ref{fig:k}.
+Interfaces of a $K$ component are represented in Fig.~\ref{fig:k}.
 
@@
 
-Finally, the $Scheduler$ component is the component responsible for implementing the $TSP$ computed by MSC.
+Finally, the $Scheduler$ component is the component responsible for implementing the $TSP$ tree  computed by MSC.
 Thus, this component represents the specific parallel orchestration of  computations.
@@
 The component also exposes a provide port to be connected to the $Time$ component.
-A generic $Scheduler$ component is represented in Fig.~\ref{fig:ass}.
+Interfaces of a $Scheduler$ component are represented in Fig.~\ref{fig:ass}.
 
@@
    \node[use] (u) at (1,0) {}; 
-   \node[] (star) at (1,0.5) {$*$};
+   \node[] (star) at (1,0.2) {$*$};
   \path[-]
@@
    \node[use] (u) at (1.5,0) {};
-   \node[] (star) at (1.5,0.5) {$*$};
+   \node[] (star) at (1.5,0.2) {$*$};
   \path[-]
@@
    \node[provide,below right=0.2 of Du3] (Datap1) {};
-   \node[component,draw=red,line width=0.4mm,fill=green!50!white,above=0.8cm of Datap2] (Data) {$A$};
-   \node[component,draw=red,line width=0.4mm,fill=green!50!white,below=0.8cm of Data] (Data2) {$B$};
+   \node[component,fill=red!50!white,above=0.8cm of Datap2] (Data) {$A$};
+   \node[component, fill=red!50!white,below=0.8cm of Data] (Data2) {$B$};
    \node[below=0.5cm of Data2] (Data3) {...};
@@
    \node[provide,above=0.15 of Datau] (DDSp2) {};
-   \node[component, draw=red, fill=red!50!white,above=0.8cm of DDSp2] (DDS) {cart};
+   \node[component, fill=red!50!white,above=0.8cm of DDSp2] (DDS) {cart};
 
@@
     (Data2.west) edge [bend left] node {} (Datau);
-    %(k0p.east) edge [bend right] node {} (mpidata.east)
-    %(k0p2.east) edge [bend right] node {} (mpidata2.east)
-    %(syncp.east) edge [bend right] node {} (mpidata3.east);
 \end{tikzpicture}
 \vspace*{.5em}
-\caption{Generic Assembly according to the Multi-Stencil program formalism.
-The complete assembly would be too complex to entirely draw and only a sub part is represented.
-Some connections are represented by numbers instead of lines in order not to make the schema more clear.
-\JB{etre plus clair sur les notations (Comme pour la Fig\ref{fig:ga}}}
+\caption{Sub-part of the specialized assembly generated by MSC from the example of the example of Fig.\ref{fig:exmsl} used throughout the paper.
+For readability some connections are represented by numbers instead of lines.
+The entire assembly is generated by MSC, however some components are automatically generated by MSC (in green), some are written by HPC specialists (in red) and others by the developer (in blue).}
 \label{fig:specass}

### 2017-03-06 ARTICLES/IJPP/results/task_scaling/500_200/results_with_without_fusion.py

@@ -64,16 +64,16 @@ matplotlib.rcParams.update({'font.size': 16})
 #plt.plot(cores, relatives_its2, 'b--', label='Ideal')
 plt.loglog(cores, itsMSL, 'b-o', label='MSL data parallelization only', basex=2,basey=2)
 plt.loglog(cores, itsMSLfus, 'b-x', label='MSL data parallelization only', basex=2,basey=2)
-#plt.loglog(cores, itsT2, 'c-*', label='2 cores for tasks without fusion', basex=2,basey=2)
-plt.loglog(cores, itsT8, 'm-*', label='8 cores for tasks without fusion', basex=2,basey=2)
-#plt.loglog(cores, itsT2fus, 'c-+', label='2 cores for tasks with fusion', basex=2,basey=2)
-plt.loglog(cores, itsT8fus, 'm-+', label='8 cores for tasks with fusions', basex=2,basey=2)
+plt.loglog(cores, itsT2, 'r-*', label='2 cores for tasks without fusion', basex=2,basey=2)
+#plt.loglog(cores, itsT8, 'r-*', label='8 cores for tasks without fusion', basex=2,basey=2)
+plt.loglog(cores, itsT2fus, 'r-+', label='2 cores for tasks with fusion', basex=2,basey=2)
+#plt.loglog(cores, itsT8fus, 'r-+', label='8 cores for tasks with fusions', basex=2,basey=2)
 
 legend = plt.legend(loc='lower right', shadow=True,prop={'size':12})
 plt.grid(True, which="both", linestyle="dotted")
 plt.xlabel('cores')
 plt.ylabel('iterations per second')
-pp = PdfPages('withwithout8_close_median.pdf')
+pp = PdfPages('withwithout2_close_median.pdf')
 plt.savefig(pp, format='pdf', bbox_inches='tight')
 pp.close()
 

### 2017-03-06 ARTICLES/IJPP/articleR1/answer_reviewer1.txt

First of all we would like to warmly thank the reviewers for their remarks that have been very helpful to improve the paper. To take their remarks into account, we have changed the structure of the paper and  we have added details and improved sentences. Moreover we have added two experiments to clarify the results and conclusions.

In the previous submission we had tried to present the domain specific language MSL. This DSL is actually part of a larger framework, the Multi-Stencil Framework. It appears in reviews that splitting the DSL from its framework and presenting it alone is difficult to understand for readers. We have therefore reworked the paper to acknoledge that fact.
We have changed the title of the paper. We have inserted a new section after the introduction that presents an overview of the Multi-Stencil Framework (MSF) to help understanding the rest of the paper as well as its outline. We have improved the third section that presents the formalism of a multi-stencil program according to the remarks of the reviewers. Sections four and five are based on this formalism, or meta-model. In Section four, we present the MSL grammar as well as the second input of the Multi-Stencil Framework, a generic template of a multi-stencil program. In Section five, we present the overall compilation process of the framework. This part was split in two different sections in the previous submission. Furthermore, this section has been improved by adding the scatter fusion case in Section 5.4, as well as the explanation of how the back-end code is produced in Section 5.5. Section six presents the evaluation of the framework. Evaluations has been clarified and additional results have been added in Section 6.4 and 6.6. Finally, the related work has been clarified according to the Multi-Stencil Framework and not only from the DSL point of view which helps to understand the contribution compared to other solutions.

A detailed answer to each reviewer follows.

**** Reviewer 1**********************************************************

- "It would be better to swap the last two subsections, since the fusion optimization is partially built upon an intermediate data introduced for the static"
    + this swap has been done as the fusion is detailed in Section 5.4 and the performance model in Section 5.6

- "The first results show that a data-parallel version scales up to at least
2^14 processor cores in both a strong and a weak scaling experiment, which is
promising. However, the results shown in Figures 12 to 14 for a version named
"MSL data parallelization only" do not scale to more than 2^8 processors.  Based on the descriptions, both experiments should be identical, so why do the results differ that much?  I am not sure whether the discrepancy is due to an insufficient description of the experiment setup or some other major issues.  More generally stated, the experiment setup for all parts of the evaluation should be clear, and possible discrepancies between different experiments should be explained in detail."
    + The reviewer is right, we had forgotten to give a precise and clear description of this evaluation (of the old Section 8.3). This has been improved in the new section 6.5. Strong scaling results are different in figures 20 and 22 (new submission references) because, to illustrate data prallelization limitations (by using MPI), we have considerably reduced the domain size of the experiment. In Figure 20 the strong scaling was shown for a domain size of 10k x 10k, while the domain size is only 500x500 in Figure 22. As a result, when reaching 2^8 cores, for example, each core has to compute only 31x31 in Figure 22, while for 2^8 cores each core has to compute 625x625 in Figure 20. As explained and shown in the new Section 6.5, this result is due to communication overheads compared to computations. This has been clarified in the paper in Section 6.5.

- "Concerning Figures 8 and 9, for a 400*400 domain per core, the MSL code
performs better but, for an 800*800 domain, SkelGIS is better: does this trend
continue for an even larger domain per core?  Do both perform exactly
identically for a 600*600 domain size?  The explanation in the paper is
confusing: are different compiler flags used for both MSL versions, or for the
MSL + SkelGIS and the SkelGIS only version?  In the former case: why? The only
difference is the problem size."
  + It is true that results are a bit different for weak scaling using 400x400 or 800x800 domain sizes. We have conducted the same experiments again and we have added the weak scaling result for a domain size 600x600. This result confirms that the weak scaling gets worse when increasing the domain size. However, results are still very close to the version using SkelGIS only (less than 3%). As clarified in the new Section 6.4, the difference between "SkelGIS" and "MSF over SkelGIS" is the use of component models l2c, for which the flag -fpic of gcc is used to produce dynamic libraries. It is quite difficult to clearly state what are the differences due to this flag and why both code versions are not exactly behaving the same. This is probably due to some minor compiler optimizations not possible because of this option.

- "An overall problem of the paper is that the language is in parts convoluted, especially in the evaluation.  The following sentence is one example from Section 8.3: "Table 3 represents the number of time a given level of parallelism, i.e., the number of tasks to perform concurrently, is observed in the final tree."
  + the English correctness has been improved in the paper

*** Reviewer 2**********************************************************

- "The formalism developed is rigorous but doesn't seem to give any new insight into the computation"
    + The formalism of Section 3 (new submission) is the basis of the Multi-Stencil framework. It is a meta-model of any multi-stencil simulation. It means that this formalism should be common to any existing solution. MSF is based on this meta-model and this is why MSF is flexible and could combine multiple kinds of back-end solution. This has been clarified in the new submission.

- "Languages like Listz already capture and utilize the same information in almost the same way."
    + We have improved the clarity of the contribution of the paper. As indicated in the new submission, the proposed meta-model of a multi-stencil program can be compared to the mesh abstraction proposed by Listz and OP2. However, neither Listz or OP2 propose an hybrid parallelization. Finally Listz and OP2 are monolithic solutions with a limited flexibility and a limited separation of concerns. This separation of concerns eases code reuse of existing solutions as well as code-reuse from one simulation to another. We hope that the new submission is more clear and answers this remark.

- "The modeling of when to use data parallelism and when to use fork-join model is interesting but not enough details are provided as to how this exactly works."
    + It has been clarified in the paper that MSF does not propose a way to choose neither the best parallelization technique nor the best fusion choices according to the kind of run or hardware used. However, MSF and its evaluation shows that the choice is not trivial and it proposes a first performance model that could help this choice. Algorithms to automatically choose the best configuration are left for future work as described in the conclusion of the paper.

- "Further the paper seems to have been submitted in a rush without adequate editing. There are a lot of grammatical errors, spelling mistakes, even undefined references that should be fixed before any submission."
    + the English correctness has been improved in the paper.

- "My suggestion to the authors is to flush out the hybrid parallelism aspect in more detail. The huge amount of formalism developed does not contribute much to this part of the paper, and the actual interesting part gets lost in all that detail. The evaluation of this is done only for Shallow water. More examples of use of this approach would make a more compelling case for the utility of this approach across different applications."
    + The contribution of the paper has been clarified and we hope that this clarification helps the reader to understand why our formalism is important. The proposed meta-model is the basis of the Multi-Stencil framework and is the way to enhance separation of concerns and flexibility. We did not have the opportunity to perform our evaluations onto different use cases than the shallow-water equations. However, we have tried to provide a deeper analyze of results.

- "Section 4: The MSL does not support sparse-updates. Supporting scatter computation is an important in many mesh-based applications. It is often computationally cheaper to traverse a particular mesh element (like vertices in a mesh) and update all adjoining mesh elements. This restriction needs to be addressed for a true mesh-based applications.
    + The reviewer is right, we have not take into account scatter computations in the previous submission. We have added it to the new submission in Section 5.4. We consider in our model that a scatter computation is actually a particular case of fusion where the developer wants to optimize cache memories by updating adjacent elements of the one being iterated. We think that more fusion cases could be added to the work, however we think that our formalism is large enough to cover almost any case.

- "The partitioning being used is traditional mesh partitioning. Previous work [4] has shown that using a hyper-graph partitioner where nodes represent iteration points and edges represent data elements touched by each iteration gives a better partitioning by reducing the communication volume. There is no modeling of the total communication volume due to the partitioning. It seems like it is treated as a constant when really it depends on the number of edges cut during the partitioning."
    + As precised at the beginning of the Section 5 (old and new), the partitioning problem is not addressed by MSF but  by the external DDS component, SkelGIS in this paper. We are aware that hyper-graph partitioning could be more interesting in some cases, and we have cited papers. However it is true that by using a better partitioning algorithm than the one used in SkelGIS, limitations of data parallelization can be delayed.

- "Please add some description of the algorithm used. The paper should at least give an intuition of what the transformation is and why it works. For example, there should at least be a line description of what an "induced subgraph" is and why it is important in the context of this paper."
    + We have improved the description of the series-parallel graphs and their use to produce a static scheduling of tasks

- "What is task parallelization degree. Its an undefined term
How is P = P_task * P_data ? This needs to be clarified by a better description of what P_task and P_data are. The formulation using P_data1 and P-data2, etc. is completely opaque. This needs to be explained better with some running example maybe."
    + Although these terms were already defined in section "performance model" of the previous submission we have tried to improve the clarity of the paper concerning this part (Section 5.6 in the new submission)

- "This is by far the strongest section of the paper. The results are very good. I would really like to see more examples here and if the parallelization effort is explained clearly it will be easier for a reader to appreciate what the different performance evaluation means.""
    + Unfortunately we did not have the opportunity to add additional use cases in the paper. However we have added performances results to clarify our conclusions.

*** Reviewer 3 **********************************************************

- "Therefore, as a DSL for determining and orchestrating task parallelism, a major omission of the paper is that it does not compare itself against other task-parallel frameworks (Legion, HPX, PFunc, etc.), and argue why a mesh-specific abstraction is better than a general task-dependency abstraction."
    + We have added a clear statement about MSF compared to general purpose task-parallel frameworks in the related work section (Section 7 in the new submission). General purpose task-parallelism frameworks are not suited for numericians. The interesting idea behind domain specific languages and domain specific frameworks is to be closer to the domain semantic, thus easier to use by specialists of this domain. Moreover, as a specific knowledge is also known onto the domain, performances can still be interesting while the language is easier to use. Finally these solutions (domain specific or general purpose) are not mutually exclusive but take place at a different abstraction level. This is the point of MSF, to enhance the use of different general purpose back-ends. In the paper OpenMP has been used as a task-based back-end, and SkelGIS as an MPI back-end. However, Legion or StarPU could have been used too. This is the subject of future works as described in the new conclusion.

  - "The mesh abstraction is very well described, yet there is an important omission that (in its current form) would mostly limit its applicability to structured meshes: there is no support for indirectly writing/incrementing data, as Definition 4 states that the variable written must be defined on the domain over which we iterate. In unstructured mesh computations, this scatter pattern is very common."
    + The reviewer is right, we have not take into account scatter computations in the previous submission. We have added it to the new submission in Section 5.4. We consider in our model that a scatter computation is actually a particular case of fusion where the developer wants to optimize cache memories by updating adjacent elements of the one being iterated.
    => je ne pense pas que tu répondes à la question (ou alors c'est trop implicite) : est que MSF supporte les maillages irréguliers et en particulier les opérations "writing/incrementing data" ?

  - "It is also a bit tedious, that a single kernel can only write a single variable, as in many codes a single kernel will write several - granted with loop fusion this can then be recovered and fused, but splitting kernels to conform to this requirement and then merging them again is painful."
    + By asking to write a single variable into one kernel, MSF guarantees that fusions are proposed in a correct manner even if different kinds of parallelization techniques are used. Moreover, we have shown in the new Section 6.6 that fusions are not always a good choice for performances. By writing a single variable in the description MSF offers more flexibility with respect to one paralellization technique or another.

  - "The paper cites the examples of other DSLs and argues that the description of the simulation is not separated from implementation concerns, therefore making re-use more difficult. However, languages like Liszt, OP2, and OPS and others have a similar abstraction to what this paper describes (even if they don't exploit it to do task parallelsim), and what they do is separate the abstraction from the !parallel! implementation concerns, and enable the parallel execution of code without any additional work, unlike this paper, where data parallel code, the computaitonal core, and the distributed messaging still has to be implemented. This distinction isn't described very well - the proposed layering of DSLs in contrast to more monolithic approaches (where the opportunity for code re-use is less), such as the aforementioned DSLs"
    + We have improved the clarity of the contribution of the paper. As indicated in the new submission, the proposed meta-model of a multi-stencil program can be compared to the mesh abstraction proposed by Listz and OP2. However, neither Listz or OP2 propose an hybrid parallelization. Finally Listz and OP2 are monolithic solutions with a limited flexibility and a limited separation of concerns. This separation of concerns eases code reuse of existing solutions as well as code-reuse from one simulation to another. We hope that the new submission is more clear and answers this remark.

### 2017-03-06 ARTICLES/IJPP/articleR1/conclusion.tex
corrections related + conclusion

@@
 In this paper, we have presented MSF, a multi-stencil framework.
-MSF is built upon a meta-model of a multi-stencil program which we have presented in Section~\ref{sect:formalism}.
-From this meta-model, we have designed, first, the generic component assembly of a multi-stencil program, and second, the domain specific language MSL that enables the description of a specific application by a numerician.
+MSF is built upon a meta-formalism of a multi-stencil program that we have presented in Section~\ref{sect:formalism}.
+From this meta-formalism, we have designed, first, the generic component assembly of a multi-stencil program, and second, the domain specific language MSL that enables the description of a specific application by a numerician.
 From these entries, MSC, the MSF compiler, automatically generates a parallel component assembly.
@@
 
-By combining a meta-model and component-based programming, MSF has the particularity to enhance separation of concerns and code-reuse of existing solutions (\eg SkelGIS and OpenMP in this paper).
+By combining a meta-model and component-based programming, MSF has the particularity to enhance separation of concerns, as well as code-reuse and composition of existing solutions (\eg SkelGIS and OpenMP in this paper).
 MSF has been evaluated on the real case simulation FullSWOF2D.
@@
 Many perspectives of future work raise from this paper.
-First, even if using one distributed data structure or another is possible and facilitated by MSF, the HPC specialist still has to understand component programming models and the meta-model of the framework to add new back-ends.
-We think it could be interesting to also facilitate the work of the HPC specialist by proposing a semi-automated framework to add external tools to MSF.
+First, even if using one distributed data structure or another is possible and facilitated by MSF, the HPC specialist still has to understand component programming models and the meta-model of the framework to add new implementation of $DDS$ and $Data$ components.
+We think it could be interesting to also facilitate the work of the HPC specialist by proposing a semi-automated framework to add new implementations of these components to MSF.
 
@@
 It depends on the time spent in computations, the number of cores and nodes used for the execution, as well as the network behavior (Figure~\ref{fig:limit} and~\ref{fig:close}).
-MSF, thanks to its meta-model and to the use of component models, handles the activation of one parallelism technique or another.
-Thus MSF can be considered as a performance leverage for multi-stencil programs.
-However, it would be interesting to investigate how an algorithm could choose the best parallelization technique for a given execution of a production multi-stencil simulation.
+MSF, thanks to its meta-formalism and to the use of component models, easily handles the activation of one parallelism technique or another.
+Thus, MSF can be considered as a performance leverage for multi-stencil programs.
+However, it would be interesting to investigate how an algorithm could choose the best parallelization technique for a given execution of a multi-stencil simulation.
 
 The same choice problem is raised by the fusion optimization.
-For the same reasons as for parallelization techniques, an algorithm could choose the best solution for a given execution.
+For the same reasons an algorithm could choose the best solution for a given execution.
 To design such an algorithm different solutions should be studied, such as following an approximated behavior model, or using calibration runs to then use deep learning algorithms, etc.

### 2017-03-06 ARTICLES/IJPP/articleR1/related.tex
corrections related + conclusion

@@
 For example, Pochoir~\cite{spaaTangCKLL11} focuses on cache optimization techniques for stencils applied onto Cartesian meshes.
-On the other hand, PATUS~\cite{citeulike12258902} proposes to add a parallelization strategy grammar to its stencil language to perform an auto-tuning parallelization strategies.
-Finally, ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} is specific to multigrid numerical methods.
+On the other hand, PATUS~\cite{citeulike12258902} proposes to add a parallelization strategy grammar to its stencil language to perform an auto-tuning parallelization.
+ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} is specific to multigrid numerical methods.
 Thus, these stencil compilers target a different scope than the Multi-Stencil Framework presented in this paper, which actually orchestrates a parallel execution of multiple stencil codes together.
@@
 However, Halide is limited to Cartesian meshes and is specific to images.
-Liszt~\cite{DeVito:2011:LDS:2063384.2063396}, OP2~\cite{} and Nabla~\cite{Camier:2015:IPP:2820083.2820107} all offer solutions for the automatic parallel orchestration of stencils applied onto any kind of mesh, from Cartesian to unstructured meshes.
+Liszt~\cite{DeVito:2011:LDS:2063384.2063396}, OP2~\cite{Giles2011} and Nabla~\cite{Camier:2015:IPP:2820083.2820107} all offer solutions for the automatic parallel orchestration of stencils applied onto any kind of mesh, from Cartesian to unstructured meshes.
 The needed mesh can be built from a set of available symbols in the grammar of each language.
@@
 However, neither Liszt, OP2 nor Nabla handle hybrid parallelism as it is proposed by MSF.
-\JB{Manque la ref OP2}
+%\JB{Manque la ref OP2}
 
@@
 First, MSL is specific to numerical simulations and proposes a mesh abstraction adapted to numerical simulations.
-Thus, compared to general purpose dataflow runtimes such as Legion~\cite{bauer:legion:sc:2012}, HPX~\cite{}, PFunc~\cite{}, MSL is closer to the semantic of the domain (mesh, stencils etc.) and easier to use for non-specialists.
+Thus, compared to general purpose dataflow runtimes such as Legion~\cite{bauer:legion:sc:2012}, HPX~\cite{Kaiser:2014:HTB:2676870.2676883}, PFunc~\cite{Kambadur:2009:PMT:1654059.1654103}, MSL is closer to the semantic of the domain (mesh, stencils etc.) and easier to use for non-specialists.
 Second, MSL is very light and only descriptive.
@@
 Furthermore, such dataflow runtimes could actually be used by MSF as back-ends, instead of using SkelGIS or OpenMP.
-\JB{Manque de refs bis :)}
+%\JB{Manque de refs bis :)}
 
 % vision of the framework brings separation of concerns, code-reuse and flexibility
-This flexibility of MSF is due to software engineering capacities introduced by proposing a meta-model and by using a component model.
+This flexibility proposed by MSF is due to software engineering capacities introduced by proposing a meta-model and by using a component programming model.
 Actually, MSF is designed to increase separation of concerns and code-reuse compared to existing solutions.
@@
 MSF generates from these pieces of information the parallel orchestration of computations.
-Finally, after MSF compilation (MSC), the developer of numerical codes fills computation kernels by using the chosen distributed data structure.
-In Liszt, OP2 and Nabla, there is no such separation of concerns between the numerician and the developer.
+Finally, the developer of numerical codes fills computation kernels by using the chosen distributed data structure.
+In Liszt, OP2 and Nabla, for example, there is no such separation of concerns between the numerician and the developer.
 %%%% flexibility

### 2017-03-06 ARTICLES/IJPP/articleR1/parallelism.tex
corrections related + conclusion
corrections

@@
 
-\begin{figure}[h!]
+\begin{figure}[t]
+\begin{minipage}{.475\textwidth}
 \begin{center}
-\subfloat[Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
   \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
   \end{tikzpicture}
-}
-\hspace{30pt}
-\subfloat[TSP tree of Fig.~\ref{fig:over}\label{fig:solu}]{
+  \caption{Over-constraint on the forbidden $N$ shape.}
+  \label{fig:over}
+ \end{center}
+\end{minipage}
+\hfill
+\begin{minipage}{.475\textwidth}
+\begin{center}
   \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
@@
   \end{tikzpicture}
-}
-\label{fid:nshape}
-\caption{N-Shape problem and solution}
+  \caption{TSP tree of Fig.~\ref{fig:over}.}
+  \label{fig:solu}
  \end{center}
-\end{figure}
+\end{minipage}\end{figure}
+
+%\begin{figure}[h!]
+%\begin{center}
+%\subfloat[Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
+%\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%   \node[] (c0) at (0,0) {$k_0$};
+%   \node[] (bc0) at (-1,0) {};
+%   \node[] (c1) at (2,0) {$k_1$};
+%   \node[] (c2) at (0,-1) {$k_2$};
+%   \node[] (bc2) at (-1,-1) {};
+%   \node[] (c3) at (2,-1) {$k_3$};
+% 
+%  \path[->]
+%    (bc0) edge [dotted] node {} (c0)
+%    (bc2) edge [dotted] node {} (c2)
+%    (c0) edge node {} (c1)
+%          edge [dashed] node [swap] {} (c3)
+%    (c2) edge node {} (c1)
+%        edge node {} (c3);
+%  \end{tikzpicture}
+%}
+%\hspace{30pt}
+%\subfloat[TSP tree of Fig.~\ref{fig:over}\label{fig:solu}]{
+%\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
+%   \node[] (S0) at (0,0) {$\mathcal{S}$};
+%   \node[] (bS0) at (-1,-1) {};
+%   \node[] (P1) at (-1,1) {$\mathcal{P}$};
+%   \node[] (aP1) at (-1.5,2) {$k_0$};
+%   \node[] (aaP1) at (-0.5,2) {$k_2$};
+%   \node[] (P2) at (1,1) {$\mathcal{P}$};
+%   \node[] (c1) at (0.5,2) {$k_1$};
+%   \node[] (c3) at (1.5,2) {$k_3$};
+% 
+%  \path[->]
+%    (bS0) edge [dotted] node {} (S0)
+%    (S0) edge node {} (P1)
+%         edge node {} (P2)
+%    (P1) edge node [swap] {} (aP1)
+%          edge node [swap] {} (aaP1)
+%    (P2) edge node {} (c1)
+%        edge node {} (c3);
+%  \end{tikzpicture}
+%}
+%\label{fid:nshape}
+%\caption{N-Shape problem and solution}
+%\end{center}
+%\end{figure}
 
@@
 As a result, the critical point for performance is when $T_{COM} \geq \frac{T_{SEQ}}{P}$, which happens naturally in data parallelization as $T_{COM}$ will increase with the number of processes, and $\frac{T_{SEQ}}{P}$ decrease with the number of processes.
-\JB{Pour moi $T_{COM}$ est constant quand on communique entre voisins... Il diminue meme legerement si on reduit les donnees par process.}
+%\JB{Pour moi $T_{COM}$ est constant quand on communique entre voisins... Il diminue meme legerement si on reduit les donnees par process.}
 

### 2017-03-06 ARTICLES/IJPP/articleR1/answer_reviewer1.txt
corrections

@@
 It appears in reviews that splitting the DSL from its framework and presenting it alone is difficult to understand for readers.
-We have therefore reworked the paper to acknoledge that fact.
+We have therefore reworked the paper to acknowledge that fact.
 We have changed the title of the paper.
@@
 This result confirms that the weak scaling gets worse when increasing the domain size.
-However, results are still very close to the version using SkelGIS only (less than 3%).
+However, results are still very close to the version using SkelGIS only (less than 3% in the worse case).
 As clarified in the new Section 6.4, the difference between "SkelGIS" and "MSF over SkelGIS" is the use of component models l2c, for which the flag -fpic of gcc is used to produce dynamic libraries.
@@
     + The formalism of Section 3 (new submission) is the basis of the Multi-Stencil framework.
-It is a meta-model of any multi-stencil simulation.
+It can be considered as a meta-model of any multi-stencil simulation.
 It means that this formalism should be common to any existing solution.
@@
     + The contribution of the paper has been clarified and we hope that this clarification helps the reader to understand why our formalism is important.
-The proposed meta-model is the basis of the Multi-Stencil framework and is the way to enhance separation of concerns and flexibility.
+The proposed meta-formalism is the basis of the Multi-Stencil framework and is the way to enhance separation of concerns and flexibility.
 We did not have the opportunity to perform our evaluations onto different use cases than the shallow-water equations.
@@
 It seems like it is treated as a constant when really it depends on the number of edges cut during the partitioning."
-    + As precised at the beginning of the Section 5 (old and new), the partitioning problem is not addressed by MSF but  by the external DDS component, SkelGIS in this paper.
+    + As precised at the beginning of the Section 5 (old and new), the partitioning problem is not addressed by MSF but  by the external DDS component (SkelGIS in this paper).
 We are aware that hyper-graph partitioning could be more interesting in some cases, and we have cited papers.
@@
 We consider in our model that a scatter computation is actually a particular case of fusion where the developer wants to optimize cache memories by updating adjacent elements of the one being iterated.
-    => je ne pense pas que tu répondes à la question (ou alors c'est trop implicite) : est que MSF supporte les maillages irréguliers et en particulier les opérations "writing/incrementing data" ?
+Thus the new submission explains how MSF supports such scatter optimization for unstructured meshes.
 

### 2017-03-06 ARTICLES/IJPP/articleR1/article.tex
fix running title
corrections parallelism.tex + evaluation.tex
corrections

@@
 %----------------------------------------
-\title{The Multi-Stencil Framework: A Component-Based Approach to Orchestrate Stencils}
+\title{A Component-Based Framework to Orchestrate Stencils}
 %----------------------------------------
@@
 One relevant solution to ease parallel programming for scientists is Domain Specific Language (DSL).
-However, one problem to avoid with DSLs is to mutualize existing codes and libraries instead of implemented each solution from scratch.
+However, one problem to avoid with DSLs is to mutualize existing codes and libraries instead of implementing each solution from scratch.
 For example, this phenomenon occurs for stencil-based numerical simulations, for which a large number of languages has been proposed without code reuse between them.
@@
 %------------------------------------------------------------------------------
-\section{Meta-modelization of a Multi-Stencil Program}
+\section{Formalism of a Multi-Stencil Program}
 \label{sect:formalism}
@@
 \input{parallelism.tex}
-%------------------------------------------------------------------------------
-%\section{Static Scheduling and performance model}
-%\label{sect:msp}
-%\input{msp.tex}
-%------------------------------------------------------------------------------
-%\section{Implementation}
-%\label{sect:comp}
-%\input{compiler.tex}
 %------------------------------------------------------------------------------

### 2017-03-06 ARTICLES/IJPP/articleR1/evaluation.tex
corrections

@@
 \begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_spread_median.pdf}}
+  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/base_spread_median.pdf}
   \caption{Strong scaling comparisons between data parallelization and hybrid parallelization. A \emph{spread} OpenMP clause is used to bind threads onto cores.}
@@
 \begin{figure}[!h]\begin{center}
-  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/tth.pdf}}
+  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/analytic/tth.pdf}
   \caption{Execution times (seconds) for a single time iteration of $T_{COM1}$ and $T_{COM2} + T_{task}$ for 8 cores for task parallelization. Verification of the Equation~(\ref{eq:hyb}).}
@@
 \end{center}\end{figure}
-
 As a result, if fusion optimization incurs a too large reduction of the number of tasks to feed dedicated cores, the problem observed for 16 cores in Figure~\ref{fig:close} happens earlier which reduces obtained performances.

### 2017-03-06 ARTICLES/IJPP/articleR1/introduction.tex
corrections

--- 2017-03-06 ARTICLES/IJPP/articleR1/introduction.tex
@@
 For example, some of the approaches to numerically solve partial differential equations (PDEs) lead to \emph{stencil computations} where the values associated to one point in space at a given time are computed from the values at the previous time at the exact same location together with a few neighbor locations.
-Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107} as will be detailed in Section~\ref{sect:rel}.
+Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107} as detailed in Section~\ref{sect:rel}.
 Many of them use the same kind of parallelization, data structures or optimization techniques, however each one has been built from scratch.
@@
 After a short overview of the Multi-Stencil Framework given in Section~\ref{sect:msf}, the paper is organized as follows.
-The meta-formalism of a multi-stencil program is presented in Section~\ref{sect:formalism}; from this formalism are built both a light and descriptive domain specific language, namely MSL, as well as a generic component assembly of the application described in Section~\ref{sect:msl}; the compiler MSC of the framework is described in Section~\ref{sect:parallelism}; finally a performance evaluation is detailed in Section~\ref{sect:eval} .
+The meta-formalism of a multi-stencil program is presented in Section~\ref{sect:formalism}; from this formalism are built both a light and descriptive domain specific language, namely MSL, as well as a generic component assembly of the application both described in Section~\ref{sect:msl}; the compiler of the framework is described in Section~\ref{sect:parallelism}; finally a performance evaluation is detailed in Section~\ref{sect:eval} .
 

### 2017-03-07 ARTICLES/IJPP/articleR1/evaluation.tex
relecture + todo

@@
 In this section we propose an evaluation of the fusion optimization.
-From the $TSP$ tree computed by MSC it is possible, according to specific conditions, to merge domain loops of kernels, thus optimizing the use of cache memories.
+From the $TSP$ tree computed by MSC it may be possible, according to some specific conditions, to merge the domain loops of some kernels, thus optimizing the use of cache memories.
 This optimization is called a fusion and has been introduced in Section\ref{sect:fusion}.
@@
 have been formally detailed in Section~\ref{sect:fusion}.
-The two first one of Figures~\ref{fig:fus1} and~\ref{fig:fus2} have been automatically detected by MSF in our case study (\ie scatter optimization has not been detected).
+The two first ones of Figures~\ref{fig:fus1} and~\ref{fig:fus2}~\vpageref{fig:fus2} have been automatically detected by MSF in our case study (\ie scatter optimization has not been detected).
+\CP{Et la troisieme ?}
 
@@
 However, an automatic detection of such fusions avoids errors, particularly for a parallel execution.
-In addition to this, more advanced fusion cases, such as a scatter are more difficult to deduce.
+In addition to this, more advanced fusion cases, such as a scatter, are more difficult to deduce.
 In FullSWOF2D a total of  62 fusions are proposed by MSF over a total of 98 computation kernels.
@@
   \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout2_close_median.pdf}
-  \caption{String scaling on a 500x500 domain size with $200$ time iterations.
+  \caption{Strong scaling on a 500x500 domain size with $200$ time iterations.
 Blue curves represent strong scaling for data parallelism with and without fusion.
@@
   \label{fig:fusion1}
-  %\CP{A FINIR !!! ERREUR DANS LEGENDES AUSSI. COULEUR PROCHES}
+  \CP{ERREUR DANS LEGENDES !!}
 \end{center}\end{figure}
@@
   \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/withwithout8_close_median.pdf}
-  \caption{String scaling on a 500x500 domain size with $200$ time iterations.
+  \caption{Strong scaling on a 500x500 domain size with $200$ time iterations.
 Blue curves represent strong scaling for data parallelism with and without fusion, thus are exactly the same than blue curves of Fig.~\ref{fig:fusion1}.
@@
   \label{fig:fusion2}
-    %\CP{A FINIR !!! ERREUR DANS LEGENDES AUSSI.}
+    \CP{ERREUR DANS LEGENDES.}
 \end{center}\end{figure}
-As a result, if fusion optimization incurs a too large reduction of the number of tasks to feed dedicated cores, the problem observed for 16 cores in Figure~\ref{fig:close} happens earlier which reduces obtained performances.
+In conclusion, if fusion optimization incurs a too large reduction of the number of tasks to feed dedicated cores, the problem observed for 16 cores in Figure~\ref{fig:close} happens earlier which reduces performance.
 For this reason, MSF performs fusions only when data parallelization is used alone.

### 2017-03-07 ARTICLES/IJPP/articleR1/evaluation.tex
corrections figures

@@
 \begin{figure}[th]\begin{center}
-  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/analytic/times.pdf}
+  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/analytic/times2.pdf}
   \caption{Computation vs communication times for a single time iteration using the data parallelization technique.}
@@
 \begin{figure}[!h]\begin{center}
-  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/analytic/tth.pdf}
+  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/analytic/tth2.pdf}
   \caption{Execution times (seconds) for a single time iteration of $T_{COM1}$ and $T_{COM2} + T_{task}$ for 8 cores for task parallelization. Verification of the Equation~(\ref{eq:hyb}).}
@@
 From the $TSP$ tree computed by MSC it may be possible, according to some specific conditions, to merge the domain loops of some kernels, thus optimizing the use of cache memories.
-This optimization is called a fusion and has been introduced in Section\ref{sect:fusion}.
+This kind of optimization is called a fusion and has been introduced in Section\ref{sect:fusion}.
 Three fusion optimizations
@@
 have been formally detailed in Section~\ref{sect:fusion}.
-The two first ones of Figures~\ref{fig:fus1} and~\ref{fig:fus2}~\vpageref{fig:fus2} have been automatically detected by MSF in our case study (\ie scatter optimization has not been detected).
-\CP{Et la troisieme ?}
+The two first ones of Figures~\ref{fig:fus1} and~\ref{fig:fus2}~\vpageref{fig:fus2} have been automatically detected by MSF in our case study (\ie the third scatter optimization of Fig.~\ref{fig:fus3} has not been detected).
+%\CP{Et la troisieme ?}
 
@@
   \label{fig:fusion1}
-  \CP{ERREUR DANS LEGENDES !!}
+  %\CP{ERREUR DANS LEGENDES !!}
 \end{center}\end{figure}
@@
   \label{fig:fusion2}
-    \CP{ERREUR DANS LEGENDES.}
+    %\CP{ERREUR DANS LEGENDES.}
 \end{center}\end{figure}

### 2017-03-07 ARTICLES/IJPP/articleR1/evaluation.tex
minor modif

@@
 From the $TSP$ tree computed by MSC it may be possible, according to some specific conditions, to merge the domain loops of some kernels, thus optimizing the use of cache memories.
-This kind of optimization is called a fusion and has been introduced in Section\ref{sect:fusion}.
-Three fusion optimizations
+This kind of optimization is called a fusion and
+three fusion optimizations have been introduced in Section~\ref{sect:fusion}. Among 
 %\CP{les rappeler !}
 %
-have been formally detailed in Section~\ref{sect:fusion}.
-The two first ones of Figures~\ref{fig:fus1} and~\ref{fig:fus2}~\vpageref{fig:fus2} have been automatically detected by MSF in our case study (\ie the third scatter optimization of Fig.~\ref{fig:fus3} has not been detected).
+them, the two first ones (Figures~\ref{fig:fus1} and~\ref{fig:fus2}~\vpageref{fig:fus2}) have been automatically detected by MSF in this case study.% (the third scatter optimization of Fig.~\ref{fig:fus3} has not been detected).
 %\CP{Et la troisieme ?}
